### 2018-01-03 paper.tex
init commit

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\begin{document}

%% Title information
\title[Short Title]{Staged Abstract Interpreter}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  We present Staged Abstract Interpreter
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{generic programming, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Text of paper \ldots

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2018-01-03 paper.tex
references

@@
 
+\section{Related Work}
+
+\textbf{Abstract Compilation}. \citeauthor{Boucher:1996:ACN:647473.727587} present a closure generation 
+technique to optimize the monovariant flow analysis \textit{0}-CFA for programs in continuation-passing style
+\cite{Boucher:1996:ACN:647473.727587}.
+
+\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the idea closure generation 
+for an small-step abstracting abstract machine \cite{Johnson:2013:OAA:2500365.2500604}.
+
+\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
+
+Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
+
 %% Acknowledgments
@@
 %% Bibliography
-%\bibliography{bibfile}
+\bibliography{references}
 

### 2018-01-05 paper.tex
outline

--- 2018-01-03 paper.tex
@@
 
-Text of paper \ldots
+\textbf{Contribution(Proposed)}.
+
+Show that abstract compilation can be understood as an instance of staged abstract interpreter.
+
+We implement staged abstract interpreter in a big-step semantics (abstract definitional interpreter) 
+using Scala with LMS framework. Show it generates sound and high-performance analysis code with 
+various precisions and sensitivities, for examples, k-CFA, pushdown or m-CFA. 
+Generate efficient store implementation.
+
+Show the evaluation result on benchmark programs.
+
+Can we do it with small-step semantics? or staging a transition system?
+
+Modular analysis? For example, k-CFA (k > 0) is a naturally global program analysis, because 
+it need the last k calling contexts. But can we analyze programs (libraries) separately 
+which generate the specialized analysis and leave the k calling contexts as dynamic parameters, 
+and then install these contexts when we have the whole program.
+
+\section{Background}
+
+\subsection{Abstracting Abstract Machine}
+
+\subsection{Staged Interpreter}
+
+\section{Abstract Compilation, Revisited}
+
+\section{Staged Abstracting Abstract Machine}
+
+\section{Evaluation}
+
+\section{Related Work}
+
+\textbf{Abstract Compilation}. The idea in this paper was closely inspired by 
+abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
+\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
+techniques as an efficient implementation of the monovariant flow analysis 
+\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
+reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
+Specifically, they proposed two kinds of abstract compilers to achieve this. 
+The first one is to generate specialized analysis as a textual program, then 
+can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
+one is using closure as a representation of specialized analysis. 
+A closure is a function that remembers the environment where it defined. So the 
+abstract compiler generates higer-order functions on-the-fly rather than textual 
+program, then the functions can be executed immediately in a higher-order language.
+
+\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
+idea closure generation for optimizing small-step abstracting abstract machine
+\cite{Johnson:2013:OAA:2500365.2500604}.
+
+\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
+
+Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
+
+\section{Conclusion}
 

### 2018-01-05 paper.tex
outline

@@
 
-\textbf{Contribution(Proposed)}.
+\textbf{Outline(Proposed)}.
 
-Show that abstract compilation can be understood as an instance of staged abstract interpreter.
+1. Revisit abstract compilation (generates textual program and closures)
+
+Show that abstract compilation can be understood as an instance of staged 
+abstract interpreter.
+
+2. But closure is still not an efficient enough representation.
 
 We implement staged abstract interpreter in a big-step semantics (abstract definitional interpreter) 
-using Scala with LMS framework. Show it generates sound and high-performance analysis code with 
-various precisions and sensitivities, for examples, k-CFA, pushdown or m-CFA. 
-Generate efficient store implementation.
+using Scala with LMS framework.  
+Show it generates sound and high-performance low-level analysis code with 
+various precisions and sensitivities, for instances, k-CFA, pushdown or m-CFA.
 
-Show the evaluation result on benchmark programs.
+Store/set operations (join, update) is expensive. Generate efficient store implementation\cite{liang2014fast}.
 
-Can we do it with small-step semantics? or staging a transition system?
+Extension: can we do it with small-step semantics? or staging a transition system?
 
-Modular analysis? For example, k-CFA (k > 0) is a naturally global program analysis, because 
-it need the last k calling contexts. But can we analyze programs (libraries) separately 
+3. Modular analysis for free?
+
+For example, k-CFA (k > 0) is a naturally global program analysis, because 
+it need the last k calling contexts. But can we analyze programs (or libraries) separately 
 which generate the specialized analysis and leave the k calling contexts as dynamic parameters, 
 and then install these contexts when we have the whole program.
+
+For another example, programs are data for an abstract interpreter, if we have $n$ programs, then there 
+are $n$ stages. Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
+waiting for the rest $(n-m)$ programs.
+
+Not sure how much improvement of performance.
+
+4. Show the evaluation result on benchmark programs.
 
@@
 abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be executed immediately in a higher-order language.
+program, then the functions can be applied immediately in a higher-order host language.
 

### 2018-01-05 paper.tex
outline

@@
 
-For example, k-CFA (k > 0) is a naturally global program analysis, because 
+For example, k-CFA (k > 0) is naturally a kind of whole program analysis, because 
 it need the last k calling contexts. But can we analyze programs (or libraries) separately 
@@
 
-For another example, programs are data for an abstract interpreter, if we have $n$ programs, then there 
+Another perspective, programs are data for an abstract interpreter, if we have $n$ programs, then there 
 are $n$ stages. Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
@@
 
-\section{Staged Abstracting Abstract Machine}
+\section{Staged Abstract Interpreter}
 

### 2018-01-05 paper.tex
outline

@@
 
-2. But closure is still not an efficient enough representation.
+2. From a performance perspective, closure is still not an efficient enough representation.
 
@@
 
-Store/set operations (join, update) is expensive. Generate efficient store implementation\cite{liang2014fast}.
+Store/set operations (join, update) is also expensive, generating efficient store implementation\cite{liang2014fast}.
 
@@
 
-For example, k-CFA (k > 0) is naturally a kind of whole program analysis, because 
-it need the last k calling contexts. But can we analyze programs (or libraries) separately 
+Motivation: one of the challenges of modern static analysis is program usually depends on
+large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
+Can we analyze programs and libraries separately without losing precision? So that we can 
+reduce part of the overhead of repeatedly analyzing libraries for different programs.
+Similarly, some static analyzers compute summary for a function or a module, that can be reused
+later(Facebook Infer). But to my knowledge, they are mostly context-insensitive or imprecise.
+
+Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis, because 
+it need the last k calling contexts as abstraction. But can we analyze programs (libraries) separately 
 which generate the specialized analysis and leave the k calling contexts as dynamic parameters, 
@@
 
-Another perspective, programs are data for an abstract interpreter, if we have $n$ programs, then there 
-are $n$ stages. Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
+Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
+then there can be $n$ stages. 
+Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
 waiting for the rest $(n-m)$ programs.

### 2018-01-05 paper.tex
outline
update outline

@@
 
+2. Generalize the idea to staged abstract interpreter.
+We implement staged abstract interpreter within a big-step abstract semantics 
+(abstract definitional interpreter) using Scala with LMS framework.  
 Show that abstract compilation can be understood as an instance of staged 
@@
 
-2. From a performance perspective, closure is still not an efficient enough representation.
+Extension: can we do it with small-step semantics? or staging a transition system?
 
-We implement staged abstract interpreter in a big-step semantics (abstract definitional interpreter) 
-using Scala with LMS framework.  
+3. Improve performance (closure is still not an efficient enough representation).
+
 Show it generates sound and high-performance low-level analysis code with 
@@
 
-Extension: can we do it with small-step semantics? or staging a transition system?
-
-3. Modular analysis for free?
+4. Modular analysis for free?
 
@@
 Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later(Facebook Infer). But to my knowledge, they are mostly context-insensitive or imprecise.
+later (like Facebook Infer). But to my knowledge, they are mostly context-insensitive or imprecise.
 
@@
 
-4. Show the evaluation result on benchmark programs.
+5. Show the evaluation result on benchmark programs (TODO: collect benchmark programs). 
+Compare with normal abstract interpreter, 
+abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
 
@@
 
-\textbf{Abstract Compilation}. The idea in this paper was closely inspired by 
+\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
 abstract compilation \cite{Boucher:1996:ACN:647473.727587}.

### 2018-01-05 paper.tex
related work

@@
 idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}.
+\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
+to intermediate form called "abstract bytecode", which are actually higher-order functions, 
+and then executed on a abstract abstract machine.
+
+\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
+
+\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
+control flow analysis \cite{amtoft1999partial}.
 

### 2018-01-07 paper/paper.tex
update benchmarks
misc update

--- 2018-01-05 paper.tex
@@
 
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis, because 
-it need the last k calling contexts as abstraction. But can we analyze programs (libraries) separately 
-which generate the specialized analysis and leave the k calling contexts as dynamic parameters, 
+Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
+because it is interpreter and need the last k calling contexts as abstraction. 
+But can we analyze programs (libraries) separately which generate the specialized 
+analysis and leave the k calling contexts as dynamic parameters, 
 and then install these contexts when we have the whole program.
@@
 
-5. Show the evaluation result on benchmark programs (TODO: collect benchmark programs). 
+5. Show the evaluation result on benchmark programs. 
 Compare with normal abstract interpreter, 
 abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
+Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
 

### 2018-01-07 paper/paper.tex
fix typo

@@
 Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interpreter and need the last k calling contexts as abstraction. 
+because it is interprocedural and need the last k calling contexts as abstraction. 
 But can we analyze programs (libraries) separately which generate the specialized 

### 2018-01-08 paper/paper.tex
misc

--- 2018-01-07 paper/paper.tex
@@
 (abstract definitional interpreter) using Scala with LMS framework.
+(ADI in Racket: \url{https://github.com/plum-umd/ abstracting-definitional-interpreters/}).
 Show that abstract compilation can be understood as an instance of staged 
@@
 Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly context-insensitive or imprecise.
+later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
+or unsound, which both lead to imprecision.
 
 Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interpreter and need the last k calling contexts as abstraction. 
+because it is interprocedural and need the last k calling contexts to distinguish
+different call sites.
 But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the k calling contexts as dynamic parameters, 
+analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
 and then install these contexts when we have the whole program.
@@
 Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then there can be $n$ stages. 
+then maybe there can be $n$ stages. 
 Probably we can analyze first $m$ programs, and generate a residual abstract interpreter

### 2018-01-12 paper/paper.tex
misc

@@
 
-Store/set operations (join, update) is also expensive, generating efficient store implementation\cite{liang2014fast}.
+Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
 
@@
 waiting for the rest $(n-m)$ programs.
+The abstract interpreter itself might be a partial abstract interpreter.
 

### 2018-01-16 paper/paper.tex
misc

@@
 
-Extension: can we do it with small-step semantics? or staging a transition system?
-
 3. Improve performance (closure is still not an efficient enough representation).
@@
 waiting for the rest $(n-m)$ programs.
-The abstract interpreter itself might be a partial abstract interpreter.
+These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
+the abstract interpreter itself might be a partial abstract interpreter.
 
@@
 Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
+
+6. Discussion
+
+Big-step vs small-step
+
+Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
+abstract interpreter?
 

### 2018-01-27 paper/paper.tex
update structure and references

@@
 
-\textbf{Outline(Proposed)}.
+\section{Abstract Compilation, Revisited}
 
-1. Revisit abstract compilation (generates textual program and closures)
-
-2. Generalize the idea to staged abstract interpreter.
-We implement staged abstract interpreter within a big-step abstract semantics 
-(abstract definitional interpreter) using Scala with LMS framework.
-(ADI in Racket: \url{https://github.com/plum-umd/ abstracting-definitional-interpreters/}).
+\cite{Boucher:1996:ACN:647473.727587}'s work is esentially a partial evaluated
+CE abstract interpreter written in a big-step semantics for CPS. The analysis
+it can perform is monovariant. We reproduce their abstractly complied analysis
+by adding explicit staging annotation to CE abstract interpreter.
 Show that abstract compilation can be understood as an instance of staged 
@@
 
-3. Improve performance (closure is still not an efficient enough representation).
+\subsection{CPS Syntax and Concrete Interpreter}
 
-Show it generates sound and high-performance low-level analysis code with 
-various precisions and sensitivities, for instances, k-CFA, pushdown or m-CFA.
+\subsection{Abstract Semantics}
+
+\subsection{Staged Abstract CE Interpreter}
+
+\section{Staged Abstract Interpreter}
+
+Generalize the idea to staged CESK abstract machine for direct style (or ANF), 
+which has fully power for instantiating different analyses.
+
+Olivier Danvy shows the relation between small-step CEK machine and big-step
+CEK machine by applying a lightweight fix-point promotion. And it is well-known
+that a staged interpreter (big-step abstract machine) is a compiler. 
+Then a nature question is to ask what is the relation of their abstract 
+counterparts.
+
+We may show: 1) staged small-step abstract interpreter (classical AAM),
+2) staged big-step abstract interpreter (ADI), 3) staged small-step abstract
+interpreter is equivalent to staging a fix-point promoted big-step abstract 
+interpreter, 4) soundness after staging.
+
+Question: 1) Pushdown analysis can be obtained if the abstract interpreter is
+written in a definitional way. Does pushdown property still exists after 
+staging? 2) Pushdown analysis can be complicated in a small-step setting 
+\cite{vardoulakis2010cfa2, earl2012introspective, gilray2016pushdown}. 
+The Pushdown for Free (P4F for short) \cite{gilray2016pushdown} is probably the most 
+simple solution. What is the relation between a staged P4F and a staged 
+abstract defintional interpreter?
+
+(ADI in Racket: \url{https://github.com/plum-umd/ abstracting-definitional-interpreters/}).
+
+\section{Improving Performance}
+
+We may show how to generate more efficient code (rather closure).
 
@@
 
-4. Modular analysis for free?
+\section{Modular Analysis for Free?}
 
@@
 
-5. Show the evaluation result on benchmark programs. 
+\section{Evaluation}
+
+Show the evaluation result on benchmark programs.
+
 Compare with normal abstract interpreter, 
@@
 
-6. Discussion
-
-Big-step vs small-step
+\section{Discussion}
 
@@
 abstract interpreter?
-
-\section{Background}
-
-\subsection{Abstracting Abstract Machine}
-
-\subsection{Staged Interpreter}
-
-\section{Abstract Compilation, Revisited}
-
-\section{Staged Abstract Interpreter}
-
-\section{Evaluation}
 

### 2018-01-27 paper/paper.tex
misc

@@
 \cite{Boucher:1996:ACN:647473.727587}'s work is esentially a partial evaluated
-CE abstract interpreter written in a big-step semantics for CPS. The analysis
-it can perform is monovariant. We reproduce their abstractly complied analysis
+CE (\textit{Control-Environment}) abstract interpreter written in a big-step semantics for CPS.
+After the partial evluation, the \textit{Control} part of abstract interpreter
+has been eliminated. 
+The analysis it can perform is monovariant, sepcifically, 0-CFA.
+Since the target language is written a continuation-passing style, so there is
+no explict continuation component when constructing the abstract interpreter
+(like CEK machine).
+
+We reproduce their abstractly complied analysis
 by adding explicit staging annotation to CE abstract interpreter.

### 2018-01-29 paper/paper.tex
update

--- 2018-01-27 paper/paper.tex
@@
 \begin{abstract}
-  We present Staged Abstract Interpreter
+  It is well known that a staged interpreter is a compiler. A nature question is to ask that what would be for a staged abstract interpreter and how to do that.
 \end{abstract}
@@
 
-\cite{Boucher:1996:ACN:647473.727587}'s work is esentially a partial evaluated
-CE abstract interpreter written in a big-step semantics for CPS. The analysis
-it can perform is monovariant. We reproduce their abstractly complied analysis
-by adding explicit staging annotation to CE abstract interpreter.
+\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
+implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
+Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
+interpreter written in a big-step semantics for programs written in continuation-passing style.
+The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
+formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
+After the partial evaluation on input program, the \textit{control} part of abstract interpreter
+is disappeared, instead, the compiled analysis is a closure (higher-order function) and
+only takes an environment as argument,
+and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
+Since the target language is written a continuation-passing style, so there is
+no explict evaluation context component when constructing the abstract interpreter
+(rather than a CEK machine).
+
+In this section, we firstly review their work of closure generation, and reproduce abstractly complied analysis
+by adding explicit staging annotation to the CE abstract interpreter.
 Show that abstract compilation can be understood as an instance of staged 
@@
 
-\subsection{CPS Syntax and Concrete Interpreter}
+\subsection{Syntax and Concrete Semantics}
+Continuation-passing style (CPS) is a widely used intermediate representation in compilation and analysis for functional
+languages.
 
@@
 
+\subsection{Closure Generation}
+
 \subsection{Staged Abstract CE Interpreter}
+
+(Fianally, we should briefly summarize the difference between them: staging vs partial evaluation; code generation; engineering effort; performance)
 
@@
 Generalize the idea to staged CESK abstract machine for direct style (or ANF), 
-which has fully power for instantiating different analyses.
+which has fully power for instantiating different analyses. We need to show that
+it not merely reproduces abstract compilation by staged computation.
 
@@
 
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
+Store/set operations (join, update) are also expensive, generating efficient store
+implementation\cite{liang2014fast}.
 

### 2018-02-01 paper/paper.tex
add a related work
draft introduction section

@@
 \begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask that what would be for a staged abstract interpreter and how to do that.
+  It is well known that a staged interpreter is a compiler. A nature question is to ask 
+  that what would be a staged abstract interpreter and how to obtain it in a mechanized 
+  approach.
+
+  In this paper, we systematically study the family of staged abstract interpreters, including 
+  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
+  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
+  To understand their relations, we firstly study the relations of their unstaged counterparts:
+  we show how to obtain a abstracting definitional interpreter by a series of transformations
+  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
+  property of 
+  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
+  been lost in another direction.
+  (TODO: pushdown for staged abstract interpreter?)
+
+  We show staged abstract interpreters have practical applications rather than just theoretical merit.
+  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
+  the improvement of performance on benchmark programs.
+  Another application of staged abstract interpreter is to transform whole-program analysis
+  to modular analysis without losing precison. The generated code can be viewed as a 
+  partially analyzed result, which also can be composed and used latter with other modulars
+  in the program.
+
 \end{abstract}
@@
 %% comma separated list
-\keywords{generic programming, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
+\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
+\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
+They firstly stage the program code to obtain a sparse representation, then conduct
+the flow-sensitive analysis.
+
 \textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}

### 2018-02-01 paper/paper.tex
misc

@@
   It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be a staged abstract interpreter and how to obtain it in a mechanized 
+  that what would be a staged abstract interpreter and how to obtain it by a mechanized 
   approach.
@@
   from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  property of 
+  control flow property of 
   analysis been obtained when transforming from small-step AAM to ADI, as well as how it 

### 2018-04-13 paper/paper.tex
update

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall,review,anonymous]{acmart}
+\documentclass[sigplan,anonymous,review]{acmart}
+\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% Title information
-\title[Short Title]{Staged Abstract Interpreter}         %% [Short Title] is optional;
+\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Subtitle}                     %% \subtitle is optional
+\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
 \subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
 
+\section{Background}
+
+\subsection{Staged Interpreter}
+
 \section{Abstract Compilation, Revisited}
@@
 
-\subsection{Syntax and Concrete Semantics}
-Continuation-passing style (CPS) is a widely used intermediate representation in compilation and analysis for functional
-languages.
-
-\subsection{Abstract Semantics}
-
-\subsection{Closure Generation}
-
-\subsection{Staged Abstract CE Interpreter}
-
-(Fianally, we should briefly summarize the difference between them: staging vs partial evaluation; code generation; engineering effort; performance)
+(Fianally, we should briefly summarize the difference between them:
+staging vs partial evaluation; code generation; engineering effort; performance)
 
@@
 
-Generalize the idea to staged CESK abstract machine for direct style (or ANF), 
-which has fully power for instantiating different analyses. We need to show that
-it not merely reproduces abstract compilation by staged computation.
+\subsection{what not to do when writing an abstract interpreter for staging}
 
-Olivier Danvy shows the relation between small-step CEK machine and big-step
-CEK machine by applying a lightweight fix-point promotion. And it is well-known
-that a staged interpreter (big-step abstract machine) is a compiler. 
-Then a nature question is to ask what is the relation of their abstract 
-counterparts.
+\section{Heterogeneous Staging}
 
-We may show: 1) staged small-step abstract interpreter (classical AAM),
-2) staged big-step abstract interpreter (ADI), 3) staged small-step abstract
-interpreter is equivalent to staging a fix-point promoted big-step abstract 
-interpreter, 4) soundness after staging.
-
-Question: 1) Pushdown analysis can be obtained if the abstract interpreter is
-written in a definitional way. Does pushdown property still exists after 
-staging? 2) Pushdown analysis can be complicated in a small-step setting 
-\cite{vardoulakis2010cfa2, earl2012introspective, gilray2016pushdown}. 
-The Pushdown for Free (P4F for short) \cite{gilray2016pushdown} is probably the most 
-simple solution. What is the relation between a staged P4F and a staged 
-abstract defintional interpreter?
-
-(ADI in Racket: \url{https://github.com/plum-umd/ abstracting-definitional-interpreters/}).
-
-\section{Improving Performance}
+Staging to GPU?
 
@@
 
-\section{Modular Analysis for Free?}
+\section{Compositional Analysis for Free}
 

### 2018-04-19 paper/paper.tex
update

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review,anonymous]{acmart}
-\documentclass[sigplan,anonymous,review]{acmart}
+\documentclass[acmsmall,review,anonymous]{acmart}
+%\documentclass[sigplan,anonymous,review]{acmart}
 \settopmatter{printfolios=true,printccs=false,printacmref=false}
@@
 
-In this section, we firstly review their work of closure generation, and reproduce abstractly complied analysis
-by adding explicit staging annotation to the CE abstract interpreter.
-Show that abstract compilation can be understood as an instance of staged 
+In this section, we firstly review their work of closure generation,
+and reproduce their result by adding stage annotations to 0-CFA.
+Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
 abstract interpreter.
@@
 
+(Specific -> General vs General vs Specific)
+
 \section{Staged Abstract Interpreter}
 
+\subsection{Abstract Interpreter}
+
 \subsection{what not to do when writing an abstract interpreter for staging}
+
+General insights when staging a static analysis algorithm.
+
+Big-step vs small-step
+
+operational vs denotational
+
+\subsection{Adding Stage Annotations}
 
@@
 the abstract interpreter itself might be a partial abstract interpreter.
-
-Not sure how much improvement of performance.
 

### 2018-08-23 paper/paper.tex
update
reorganize

--- 2018-02-01 paper/paper.tex
@@
   It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be a staged abstract interpreter and how to obtain it by a mechanized 
+  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
   approach.
@@
 
-Not sure how much improvement of performance.
+IDEA: combining analysis with different sensitivities.
+
+TODO: evaluate the improvement of performance.
 

### 2018-08-23 paper/paper.tex

--- 2018-04-19 paper/paper.tex
@@
 
-\subsection{Staged Interpreter}
+\subsection{Interpreters}
+
+\subsection{From Interpreters to Staged Interpreters}
+
+\subsection{From Interpreters to Abstract Interpreters}
 
@@
 
-General insights when staging a static analysis algorithm.
-
 Big-step vs small-step
@@
 
-\subsection{Adding Stage Annotations}
+General insights for staging other static analysis algorithm.
 
-\section{Heterogeneous Staging}
+\subsection{Adding Stage Annotations is Easy}
+
+\section{Optimizations}
 
@@
 
-We may show how to generate more efficient code (rather closure).
+Staging inside data structures.
 
-Store/set operations (join, update) are also expensive, generating efficient store
-implementation\cite{liang2014fast}.
+Jones-optimal specialization.
+
+Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
 

### 2018-10-12 paper/paper.tex
update tex template

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+%\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
+%\documentclass[sigplan,review]{acmart}\settopmatter{printfolios=true}
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+%\documentclass[sigplan]{acmart}\settopmatter{}
 
-%% Journal information
+%% Conference information
 %% Supplied to authors by publisher for camera-ready submission;
 %% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
+\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
 \acmYear{2018}
-\acmMonth{1}
+\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
 \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
@@
 %% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
+%\citestyle{acmauthoryear}  %% For author/year citations
+%\citestyle{acmnumeric}     %% For numeric citations
+%\setcitestyle{nosort}      %% With 'acmnumeric', to disable automatic
+                            %% sorting of references within a single citation;
+                            %% e.g., \cite{Smith99,Carpenter05,Baker12}
+                            %% rendered as [14,5,2] rather than [2,5,14].
+%\setcitesyle{nocompress}   %% With 'acmnumeric', to disable automatic
+                            %% compression of sequential references within a
+                            %% single citation;
+                            %% e.g., \cite{Baker12,Baker14,Baker16}
+                            %% rendered as [2,3,4] rather than [2-4].
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
+%% Note: Authors migrating a paper from traditional SIGPLAN
+%% proceedings format to PACMPL format must update the
+%% '\documentclass' and topmatter commands above; see
+%% 'acmart-pacmpl-template.tex'.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   in the program. 
-
 \end{abstract}
@@
 %% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
+\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
 

### 2018-10-18 paper/10_intro.tex
restructured

\section{Introduction}

\subsection{Motivation}

\subsection{Contributions}

### 2018-10-18 paper/20_bg.tex
restructured

\section{Background}

\subsection{The Language}

\subsection{From Interpreter to Abstract Interpreter}

### 2018-10-18 paper/30_sai.tex
restructured

--- 2018-08-23 paper/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\begin{document}
-
-%% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
-
-  We show staged abstract interpreters have practical applications rather than just theoretical merit.
-  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-  the improvement of performance on benchmark programs.
-  Another application of staged abstract interpreter is to transform whole-program analysis
-  to modular analysis without losing precison. The generated code can be viewed as a 
-  partially analyzed result, which also can be composed and used latter with other modulars
-  in the program.
-
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
-
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
-
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
-
-(Specific -> General vs General vs Specific)
-
 \section{Staged Abstract Interpreter}
 
-\subsection{Abstract Interpreter}
+\subsection{From Abstract Interpreter to Staged Abstract Interpreter}
 
@@
 
-\subsection{Adding Stage Annotations is Easy}
+\subsection{Stage Annotations}
 
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-\section{Compositional Analysis for Free}
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
-\section{Related Work}
-
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
-
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
-
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
-\bibliography{references}
-
-%% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
-
-\end{document}
+Jones-optimal specialization?
+Futamura projection?

### 2018-10-18 paper/40_comp.tex
restructured

--- 2018-08-23 paper/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\begin{document}
-
-%% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
-
-  We show staged abstract interpreters have practical applications rather than just theoretical merit.
-  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-  the improvement of performance on benchmark programs.
-  Another application of staged abstract interpreter is to transform whole-program analysis
-  to modular analysis without losing precison. The generated code can be viewed as a 
-  partially analyzed result, which also can be composed and used latter with other modulars
-  in the program.
-
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
-
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
-
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
-
-(Specific -> General vs General vs Specific)
-
-\section{Staged Abstract Interpreter}
-
-\subsection{Abstract Interpreter}
-
-\subsection{what not to do when writing an abstract interpreter for staging}
-
-Big-step vs small-step
-
-operational vs denotational
-
-General insights for staging other static analysis algorithm.
-
-\subsection{Adding Stage Annotations is Easy}
-
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
 \section{Compositional Analysis for Free}
@@
 the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
-\section{Related Work}
-
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
-
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
-
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
-\bibliography{references}
-
-%% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
-
-\end{document}

### 2018-10-18 paper/50_cases.tex
restructured

--- 2018-08-23 paper/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+\section{Case Study}
 
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\begin{document}
-
-%% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
-
-  We show staged abstract interpreters have practical applications rather than just theoretical merit.
-  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-  the improvement of performance on benchmark programs.
-  Another application of staged abstract interpreter is to transform whole-program analysis
-  to modular analysis without losing precison. The generated code can be viewed as a 
-  partially analyzed result, which also can be composed and used latter with other modulars
-  in the program.
-
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
+\subsection{Abstract Compilation, Revisited}
 
@@
 
-\section{Staged Abstract Interpreter}
+\subsection{Control Flow Analysis}
 
-\subsection{Abstract Interpreter}
+context/path/flow-sensitive
 
-\subsection{what not to do when writing an abstract interpreter for staging}
+store-widening
 
-Big-step vs small-step
+using different $k$ for $k$-CFA
 
-operational vs denotational
+\subsection{Scale to Imperative Languages}
 
-General insights for staging other static analysis algorithm.
+Assignments, Loops, Exceptions
 
-\subsection{Adding Stage Annotations is Easy}
-
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-\section{Compositional Analysis for Free}
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
-\section{Related Work}
-
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
-
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
-
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
-\bibliography{references}
-
-%% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
-
-\end{document}
+Options: Numerical Analysis, Data-flow analysis, taint analysis

### 2018-10-18 paper/60_eval.tex
restructured

\section{Evaluation}

\subsection{Benchmarks}

\subsection{Performance}

\subsubsection{Baselines}
Small-step AAM

Big-step ADI

OAAM with AC \cite{Boucher:1996:ACN:647473.727587, Johnson:2013:OAA:2500365.2500604}

Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.

\subsection{Compositionality}

### 2018-10-18 paper/70_related.tex
restructured

--- 2018-08-23 paper/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
-
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\begin{document}
-
-%% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
-
-  We show staged abstract interpreters have practical applications rather than just theoretical merit.
-  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-  the improvement of performance on benchmark programs.
-  Another application of staged abstract interpreter is to transform whole-program analysis
-  to modular analysis without losing precison. The generated code can be viewed as a 
-  partially analyzed result, which also can be composed and used latter with other modulars
-  in the program.
-
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
-
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
-
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
-
-(Specific -> General vs General vs Specific)
-
-\section{Staged Abstract Interpreter}
-
-\subsection{Abstract Interpreter}
-
-\subsection{what not to do when writing an abstract interpreter for staging}
-
-Big-step vs small-step
-
-operational vs denotational
-
-General insights for staging other static analysis algorithm.
-
-\subsection{Adding Stage Annotations is Easy}
-
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-\section{Compositional Analysis for Free}
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
 \section{Related Work}
@@
 Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
-\bibliography{references}
-
-%% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
-
-\end{document}

### 2018-10-18 paper/abstract.tex
restructured

--- 2018-08-23 paper/paper.tex
@@
-%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}
-%\documentclass[sigplan,anonymous,review]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For double-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-%% For single-blind review submission, w/ CCS and ACM Reference
-%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
-%% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+It is well known that a staged interpreter is a compiler, which provides performance improvement
+by eliminate the overhead of traversing the abstract syntax tree.
 
-%% Journal information
-%% Supplied to authors by publisher for camera-ready submission;
-%% use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-%% Copyright information
-%% Supplied to authors (based on authors' rights management selection;
-%% see authors.acm.org) by publisher for camera-ready submission;
-%% use 'none' for review submission.
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\copyrightyear{2018}           %% If different from \acmYear
-
-%% Bibliography style
-\bibliographystyle{ACM-Reference-Format}
-%% Citation style
-%% Note: author/year citations are required for papers published as an
-%% issue of PACMPL.
-\citestyle{acmauthoryear}   %% For author/year citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Note: Authors migrating a paper from PACMPL format to traditional
-%% SIGPLAN proceedings format must update the '\documentclass' and
-%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Some recommended packages.
-\usepackage{booktabs}   %% For formal tables:
-                        %% http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
-                        %% http://ctan.org/pkg/subcaption
-
-\begin{document}
-
-%% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
-                                        %% when present, will be used in
-                                        %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
-                                        %% can be repeated if necessary;
-                                        %% contents suppressed with 'anonymous'
-
-%% Author information
-%% Contents and number of authors suppressed with 'anonymous'.
-%% Each author should be introduced by \author, followed by
-%% \authornote (optional), \orcid (optional), \affiliation, and
-%% \email.
-%% An author may have multiple affiliations and/or emails; repeat the
-%% appropriate command.
-%% Many elements are not rendered, but should be provided for metadata
-%% extraction tools.
-
-%% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
-
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-%% Abstract
-%% Note: \begin{abstract}...\end{abstract} environment must come
-%% before \maketitle command
-\begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
+In this paper, we study staged \textit{abstract} interpreters.
 
@@
 in the program.
-
-\end{abstract}
-
-%% 2012 ACM Computing Classification System (CSS) concepts
-%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
-%% End of generated code
-
-%% Keywords
-%% comma separated list
-\keywords{staged computation, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
-
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
-
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
-
-(Specific -> General vs General vs Specific)
-
-\section{Staged Abstract Interpreter}
-
-\subsection{Abstract Interpreter}
-
-\subsection{what not to do when writing an abstract interpreter for staging}
-
-Big-step vs small-step
-
-operational vs denotational
-
-General insights for staging other static analysis algorithm.
-
-\subsection{Adding Stage Annotations is Easy}
-
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-\section{Compositional Analysis for Free}
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
-\section{Related Work}
-
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
-
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
-
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
-
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-                                        %% contents suppressed with 'anonymous'
-  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-  %% acknowledge financial support and will be used by metadata
-  %% extraction tools.
-  This material is based upon work supported by the
-  \grantsponsor{GS100000001}{National Science
-    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-  conclusions or recommendations expressed in this material are those
-  of the author and do not necessarily reflect the views of the
-  National Science Foundation.
-\end{acks}
-
-%% Bibliography
-\bibliography{references}
-
-%% Appendix
-\appendix
-\section{Appendix}
-
-Text of appendix \ldots
-
-\end{document}

### 2018-10-18 paper/paper.tex
restructured

@@
 %% use defaults for review submission.
-\acmConference[PL'18]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2018}{New York, NY, USA}
-\acmYear{2018}
-\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+%\acmConference[Submitted to PLDI'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
+\acmConference[Submitted to PLDI]{}{June 2019}{Phoenix, Arizona}
+\acmYear{2019}
+%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
 \startPage{1}
@@
                                         %% header instead of Full Title.
-\titlenote{with title note}             %% \titlenote is optional;
+%\titlenote{with title note}             %% \titlenote is optional;
                                         %% can be repeated if necessary;
@@
 \subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
+%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                         %% can be repeated if necessary;
@@
 \begin{abstract}
-  It is well known that a staged interpreter is a compiler. A nature question is to ask 
-  that what would be if we can stage an abstract interpreter and how to obtain it by a mechanized 
-  approach.
-
-  In this paper, we systematically study the family of staged abstract interpreters, including 
-  \citeauthor{Boucher:1996:ACN:647473.727587}'s Abstract Compilation, staged small-step 
-  Abstracting Abstract Machines (AAM) and staged Abstracting Definitional Interpreters (ADI).
-  To understand their relations, we firstly study the relations of their unstaged counterparts:
-  we show how to obtain a abstracting definitional interpreter by a series of transformations
-  from a small-step abstracting abstract machine. A key insight we show is how the pushdown 
-  control flow property of 
-  analysis been obtained when transforming from small-step AAM to ADI, as well as how it 
-  been lost in another direction.
-  (TODO: pushdown for staged abstract interpreter?)
-
-  We show staged abstract interpreters have practical applications rather than just theoretical merit.
-  The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-  the improvement of performance on benchmark programs.
-  Another application of staged abstract interpreter is to transform whole-program analysis
-  to modular analysis without losing precison. The generated code can be viewed as a 
-  partially analyzed result, which also can be composed and used latter with other modulars
-  in the program. 
+  \input{abstract}
 \end{abstract}
@@
 
-\section{Introduction}
-
-\section{Background}
-
-\subsection{Interpreters}
-
-\subsection{From Interpreters to Staged Interpreters}
-
-\subsection{From Interpreters to Abstract Interpreters}
-
-\section{Abstract Compilation, Revisited}
-
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
-
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
-
-(Specific -> General vs General vs Specific)
-
-\section{Staged Abstract Interpreter}
-
-\subsection{Abstract Interpreter}
-
-\subsection{what not to do when writing an abstract interpreter for staging}
-
-Big-step vs small-step
-
-operational vs denotational
-
-General insights for staging other static analysis algorithm.
-
-\subsection{Adding Stage Annotations is Easy}
-
-\section{Optimizations}
-
-Staging to GPU?
-
-Staging inside data structures.
-
-Jones-optimal specialization.
-
-Store/set operations (join, update) are also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-\section{Compositional Analysis for Free}
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
-
-\section{Evaluation}
-
-Show the evaluation result on benchmark programs.
-
-Compare with normal abstract interpreter, 
-abstract compilation\cite{Boucher:1996:ACN:647473.727587}, and OAAM \cite{Johnson:2013:OAA:2500365.2500604}.
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\section{Discussion}
-
-Futamura projection for abstract interpreter. Is there a second or third Futamura projection for
-abstract interpreter?
-
-\section{Related Work}
-
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
-
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
-
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
+\input{10_intro}
+\input{20_bg}
+\input{30_sai}
+\input{40_comp}
+\input{50_cases}
+\input{60_eval}
+\input{70_related}
+\input{80_concl}
 

### 2018-10-18 paper/abstract.tex
revise abstract

@@
 It is well known that a staged interpreter is a compiler, which provides performance improvement
-by eliminate the overhead of traversing the abstract syntax tree.
+by specializing the interpreter to a given program.
 
-In this paper, we study staged \textit{abstract} interpreters.
+In this paper, we study \textit{abstract} interpreters with multi-stage programming, i.e., the 
+staged abstract interpreters. 
+By staging an abstract interpreter respect to a program, we obtain a compiled and faster analysis.
+By appying the idea to \textit{open} programs, we obtain a modular analysis that generates sound
+partial analysis results which can be composed and reused later without losing precision, though 
+the original abstract interpreter is a whole-program analysis algorithm.
 
-We show staged abstract interpreters have practical applications rather than just theoretical merit.
-The time consumed of a analysis can be decreased by staging (TODO: to what degree). We evaluate
-the improvement of performance on benchmark programs.
-Another application of staged abstract interpreter is to transform whole-program analysis
-to modular analysis without losing precison. The generated code can be viewed as a 
-partially analyzed result, which also can be composed and used latter with other modulars
-in the program.
+We show several case studies using the idea of staged abstract interpreters, including 
+\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of Shiver's 0-CFA, pushdown
+control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
+analysis on a imperative language.
+
+We also evaluate the improvement of performance on benchmark programs for a subset of Scheme.
+The result shows ...

### 2018-10-18 paper/10_intro.tex
intro

@@
 
+1. we know how to stage an interpreter, natural to ask how to stage an abstract interpreter, given 
+the structural similarity of interpreters and abstract interpreters.
+
+2. whole-program analysis is slow, how to optimize it systematically/mechanically?
+bridging the gap between designing sound analyses and implementing efficient analyses.
+
+3. analyzing modern programs with large libraries are expensive, we need to a way to create
+summaries of libraries so that we can reuse them.
+
 \subsection{Contributions}

### 2018-10-18 paper/abstract.tex
continue working on outline

@@
 staged abstract interpreters. 
-By staging an abstract interpreter respect to a program, we obtain a compiled and faster analysis.
-By appying the idea to \textit{open} programs, we obtain a modular analysis that generates sound
-partial analysis results which can be composed and reused later without losing precision, though 
-the original abstract interpreter is a whole-program analysis algorithm.
+By staging an abstract interpreter respect to a program, we obtain a specialized analysis that runs faster.
+By appying the idea to \textit{open} programs and considering the free variables as dynamic inputs, 
+we obtain a modular analysis that generates sound partial analysis results which can be composed 
+and reused later without losing precision, though the original abstract interpreter is a whole-program 
+analysis algorithm.
 

### 2018-10-19 paper/10_intro.tex
continue working on outline
draft on contribution

@@
 
-2. whole-program analysis is slow, how to optimize it systematically/mechanically?
+2. precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
 bridging the gap between designing sound analyses and implementing efficient analyses.
@@
 \subsection{Contributions}
+
+\begin{itemize}
+\item Starting from a definitional abstract interpreter for a higher-order functional language,
+  we show how to turn it into a staged abstract interpreter by adding stage annotations with 
+  Lightweight Modular Staging (LMS) framework. We later extend the abstract interpreter to support
+  imperative language features, such as states, loops and exceptions.
+\item We elaborate the idea that the staged abstract interpreter not only improves the performance 
+  but also the scalability, which are two major issues in static analysis, if we apply it on open
+  programs. TODO: more on this point.
+\item To evaluate the generality of the idea, we conduct several case studies. 
+  First we revisited Abstract Compilation (TODO: cite), which is a closure generation technique applied 
+  on 0-CFA and can be understood as an instance of staged abstract interpreter. 
+  Second, we demonstrate existing control-flow analysis techniques, such as context/flow/path-sensitive 
+  analysis and store-widening can be integrated with staged abstract interpreters.
+  Finally, we show it also can be applied to numerical analysis in an imperative language.
+  (Optional: taint/information flow analysis?)
+\item We also emprically evaluate the performance improved of staged abstract interpreter with the 
+  unstaged version and other existing optimizations on benchmarks.
+\end{itemize}

### 2018-10-19 code/src/main/scala/sai/direct/core/concrete/CESK.scala

@@ -114,7 +114,6 @@ object SmallStepCESK {
 object RefuncBigStepCESK {
   /* A refunctionalized CPS interpreter. */
   type Cont = (Value, Store) ⇒ (Value, Store)
-  case class State(e: Expr, env: Env, store: Store, k: Cont)
 
   def interp(e: Expr, ρ: Env, σ: Store, κ: Cont): (Value, Store) = e match {
     case Lit(i) ⇒ κ(NumV(i), σ)

### 2018-10-19 paper/abstract.tex
misc

@@
 
-In this paper, we study \textit{abstract} interpreters with multi-stage programming, i.e., the 
+In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
 staged abstract interpreters. 
@@
 
-We show several case studies using the idea of staged abstract interpreters, including 
+Using the idea of staged abstract interpreters, we show several case studies, including 
 \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of Shiver's 0-CFA, pushdown

### 2018-10-20 paper/10_intro.tex
misc

@@
 1. we know how to stage an interpreter, natural to ask how to stage an abstract interpreter, given 
-the structural similarity of interpreters and abstract interpreters.
+the structural similarity of interpreters and abstract interpreters. Futamura projection \cite{Futamura1999}
 
@@
   we show how to turn it into a staged abstract interpreter by adding stage annotations with 
-  Lightweight Modular Staging (LMS) framework. We later extend the abstract interpreter to support
+  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
+  We later extend the abstract interpreter to support
   imperative language features, such as states, loops and exceptions.

### 2018-10-21 paper/10_intro.tex
refactor AC 0CFA

@@
 
-1. we know how to stage an interpreter, natural to ask how to stage an abstract interpreter, given 
-the structural similarity of interpreters and abstract interpreters. Futamura projection \cite{Futamura1999}
+1. we know how to stage an interpreter, intellectually it is natural to ask what would be an staged abstract interpreter,
+and how to stage an abstract interpreter, given the structural similarity of interpreters and abstract interpreters. 
+Futamura projection \cite{Futamura1999}
 
-2. precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
+2. Multi-stage programming is widely used in high performance computation.
+Precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
 bridging the gap between designing sound analyses and implementing efficient analyses.
+
+When staging a concrete interpreter, the programmers need to distinguish static and dynamic values
+--- the given program is classified as static because it is known at compile-time, and the inputs to that program 
+are dynamic. However, when staging an abstract interpreter, this distinction not exists anymore. 
+Because an abstract interpreter treats all the inputs as some form of abstract values, which usually are 
+top elements in their abstract domains and are statically known. Then what is the point of staging?
+A surprising by-product of thinking about this question is to realize that we can apply the staged
+abstract interpreter on \textit{open} programs, and the free variables representing other parts of 
+the program (e.g., libraries) are dynamic input, then we obtain a modular analysis, mechanically.
 

### 2018-10-21 code/src/main/scala/sai/cps/zerocfa/StagedZeroCFA.scala

@@ -12,103 +12,95 @@ import sai.cps.parser._
 import sai.common._
 
 trait StagedZeroCFA extends DslExp with LamOpsExp with MapOpsExp with TupledFunctionsRecursiveExp with ImmSetOpsExp {
-  type MapT = Map[String, ImmSet[Lam]]
+  type Addr = String
+  type MapT = Map[Addr, ImmSet[Lam]]
 
-  implicit def MapTyp: Typ[MapT] = manifestTyp
+  type Ans = Rep[ImmSet[Lam]]
+  type Store = Rep[Map[Addr, ImmSet[Lam]]]
+  //TODO: The store doesn't have to be staged, the addresses are known at stage-time.
+  //      Try Store = Map[Addr, Rep[ImmSet[Lam]]]
 
-  def lift[T:Typ](lst: List[T]): Rep[List[T]] = List[T](lst.map((s) => unit[T](s)):_*)
+  implicit def StoreTyp: Typ[Store] = manifestTyp
+
+  //def lift[T:Typ](lst: List[T]): Rep[List[T]] = List[T](lst.map(lift(_)):_*)
+
   def lift[T:Typ](x: T) = unit[T](x)
 
-  def lookup(map: Rep[MapT], addr: String): Rep[ImmSet[Lam]] = {
-    map.getOrElse(addr, ImmSet[Lam]())
+  def primEval(e: Expr, σ: Store): Ans = e match {
+    case Lit(_) ⇒ ImmSet[Lam]()
+    case Var(x) ⇒ lookup(x, σ)
+    case l: Lam ⇒ ImmSet[Lam](lift(l))
   }
 
-  def lookup(map: Rep[MapT], addr: Expr): Rep[ImmSet[Lam]] = {
-    addr match {
-      case Lit(_) => ImmSet[Lam]()
-      case Var(name) => lookup(map, name)
-      case l@Lam(v, body) => ImmSet[Lam](lift(l))
-    }
+  def lookup(α: Addr, σ: Store): Ans = σ.getOrElse(α, ImmSet[Lam]())
+  def update(σ: Store, α: Rep[Addr], d: Rep[ImmSet[Lam]]): Store = {
+    val oldd = σ.getOrElse(α, ImmSet[Lam]())
+    σ ++ Map(α → (d ++ oldd))
   }
-
[…]

### 2018-10-21 code/src/main/scala/sai/cps/zerocfa/Store.scala

@@ -1,41 +1,0 @@
-package sai.cps.zerocfa
-
-import sai.cps.parser._
-
-object SetExtractor {
-  def unapplySeq[T](s: Set[T]): Option[Seq[T]] = Some(s.toSeq)
-}
-
-case class Store(map: Map[String, Set[Lam]]) {
-  case class UndefinedAddrException(addr: Expr) extends RuntimeException
-
-  def lookup(addr: Expr): Set[Lam] = {
-    addr match {
-      case Lit(_) => Set()
-      case Var(name) => lookup(name)
-      case lam@Lam(_, _) => Set(lam)
-      case _ => throw UndefinedAddrException(addr)
-    }
-  }
-
-  def lookup(addr: String): Set[Lam] = map.getOrElse(addr, Set[Lam]())
-
-  def update(addr: String, d: Set[Lam]): Store = {
-    //println(s"update $addr -> $d")
-    val oldd = map.getOrElse(addr, Set[Lam]())
-    Store(map ++ Map(addr -> (d ++ oldd)))
-  }
-
-  def update(addr: String, v: Lam): Store = update(addr, Set(v))
-
-  def update(addrs: List[String], ds: List[Set[Lam]]): Store = {
-    (addrs zip ds).foldLeft(this) {
-      case (store, (addr, v)) => store.update(addr, v)
-    }
-  }
-}
-
-object Store {
-  def bot = Set()
-  def mtStore = Store(Map())
-}

### 2018-10-21 code/src/main/scala/sai/cps/zerocfa/ZeroCFA.scala

@@ -2,64 +2,94 @@ package sai.cps.zerocfa
 
 import sai.cps.parser._
 
+object SetExtractor {
+  def unapplySeq[T](s: Set[T]): Option[Seq[T]] = Some(s.toSeq)
+}
+
+case class Store(map: Map[String, Set[Lam]]) {
+  def lookup(α: String): Set[Lam] = map.getOrElse(α, Set[Lam]())
+
+  def update(α: String, d: Set[Lam]): Store = {
+    val oldd = map.getOrElse(α, Set[Lam]())
+    Store(map ++ Map(α → (d ++ oldd)))
+  }
+
+  def update(α: String, v: Lam): Store = update(α, Set(v))
+
+  def update(αs: List[String], ds: List[Set[Lam]]): Store =
+    (αs zip ds).foldLeft (this) {
+      case (σ, (α, v)) => σ.update(α, v)
+    }
+}
+
+object Store {
+  def bot = Set()
+  def mtStore = Store(Map())
+}
+
 abstract class CFACommon {
   val debug = false
-  def analysisAbsApp(fs: Set[Lam], args: List[Expr], store: Store): Store = {
+  def primEval(e: Expr, σ: Store): Set[Lam] = e match {
+    case Lit(_) => Set()
+    case Var(x) => σ.lookup(x)
+    case l: Lam => Set(l)
+  }
+
+  def analysisAbsApp(fs: Set[Lam], args: List[Expr], σ: Store): Store = {
     if (debug) println(s"analysisAbsApp fs[$fs]")
-    if (fs.isEmpty) store
+    if (fs.isEmpty) σ
     else {
-      val newArgs = args.map(store.lookup(_))
-      val newStore = store.update(fs.head.vars, newArgs)
-      analysisAbsApp(fs.tail, args, newStore)
+      val newArgs = args.map(primEval(_, σ))
+      val σ_* = σ.update(fs.head.vars, newArgs)
+      analysisAbsApp(fs.tail, args, σ_*)
     }
   }
 }
 
 object ZeroCFA extends CFACommon {
   def analyze(prog: Expr): Store = {
-    def iter(store: Store): Store = {
-      val newStore = analysisProgram(prog, store)
-      if (store == newStore) store else iter(newStore)
+    def iter(σ: Store): Store = {
+      val σ_* = analysisProgram(prog, σ)
+      if (σ == σ_*) σ else iter(σ_*)
     }
     iter(Store.mtStore)
   }
 
[…]

### 2018-10-21 paper/10_intro.tex
misc

@@
 bridging the gap between designing sound analyses and implementing efficient analyses.
+partial evaluation on program analyzers is not a new idea, remove the interpretive overhead of repetively 
+traversing the abstract syntax tree.
+we revisit this approach from a more recent perspective by combining general abstract interpreters
+and multi-stage programming.
 
@@
 
-3. analyzing modern programs with large libraries are expensive, we need to a way to create
-summaries of libraries so that we can reuse them.
+3. analyzing modern programs with large libraries are expensive, repetitively analyze the part of library,
+modularity as in compilers, we need to a way to create summaries of libraries so that we can reuse them,
+remove redundant analyzing of libraries.
 

### 2018-10-23 paper/10_intro.tex
update

@@
 
-1. we know how to stage an interpreter, intellectually it is natural to ask what would be an staged abstract interpreter,
-and how to stage an abstract interpreter, given the structural similarity of interpreters and abstract interpreters. 
-Futamura projection \cite{Futamura1999}
+1. Futamura projection \cite{Futamura1999} shows that staging an interpreter yields a compiler.
+literature on this approach, creating DSL.
+
+Recent research uncovers a systematic approach to construct abstract interpreters from concrete interpreters.
+Given the structural similarity between abstract and concrete interpreters, intellectually it is natural to 
+wonder what would be a staged abstract interpreter, and how to stage an abstract interpreter 
+having considered their functional dissimilarity.
 
@@
 Precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
-bridging the gap between designing sound analyses and implementing efficient analyses.
+bring existing high-performance computation facilities, in the form of meta-programming, into
+semantic artifacts of abstract interpretation, i.e., abstract interpreters.
+and bridging the gap between designing sound analyses and implementing efficient analyses.
+without changing the abstract interpreter, just type
+
 partial evaluation on program analyzers is not a new idea, remove the interpretive overhead of repetively 
@@
 
-3. analyzing modern programs with large libraries are expensive, repetitively analyze the part of library,
-modularity as in compilers, we need to a way to create summaries of libraries so that we can reuse them,
-remove redundant analyzing of libraries.
+3. analyzing modern programs with large libraries are expensive. 
+why? 1. precise analysis usually required whole-program analysis; 2. to do wp analysis, need to 
+repetitively analyze the parts of library.
+As modularity in compilers, we need to a way to separately analyze programs, and create summaries of analyzed 
+part so that we can reuse them, remove redundant analyzing of libraries.
 

### 2018-10-23 paper/abstract.tex
refactor AC 0CFA
update

@@
 staged abstract interpreters. 
-By staging an abstract interpreter respect to a program, we obtain a specialized analysis that runs faster.
-By appying the idea to \textit{open} programs and considering the free variables as dynamic inputs, 
+By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
+By appying the abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
 we obtain a modular analysis that generates sound partial analysis results which can be composed 
@@
 
-We also evaluate the improvement of performance on benchmark programs for a subset of Scheme.
+We also empirically evaluate the improvement of performance on benchmark programs for a subset of Scheme.
 The result shows ...

### 2018-10-23 code/src/main/scala/sai/Main.scala

@@ -23,7 +23,7 @@ case class Timing(ts: List[Double]) {
   val perc50 = sorted_ts((sorted_ts.size / 2).toInt)
   val perc75 = sorted_ts(((sorted_ts.size / 4) * 3).toInt)
   val perc95 = sorted_ts(sorted_ts.size - (sorted_ts.size/20).toInt)
-  override def toString: String = s"Mean: ${mean}, 0/5/25/50/75/95/100: ${ub}/${perc05}/${perc25}/${perc50}/${perc75}/${perc95}/${lb}"
+  override def toString: String = s"#: ${ts.size}, Mean: ${mean}, 0/5/25/50/75/95/100: ${ub}/${perc05}/${perc25}/${perc50}/${perc75}/${perc95}/${lb}"
 }
 
 object Main {

### 2018-10-23 code/src/main/scala/sai/cps/zerocfa/StagedZeroCFA.scala

@@ -16,6 +16,10 @@ import BasicTypes._
  0CFA AC average time: 132463.2875
  0CFA Staged (with precompiled) average time: 75982.7375
  54.6% faster than unopt, 42.6% faster than AC
+ -----------------------------------------------
+ 0CFA time - #: 500, Mean: 469726.998, 0/5/25/50/75/95/100: 126457.0/129778.0/162299.0/183759.0/387371.0/1574363.0/3.8219527E7
+ 0CFA AC time - #: 500, Mean: 163236.018, 0/5/25/50/75/95/100: 104389.0/107165.0/130640.0/146222.0/163850.0/286668.0/907706.0
+ 0CFA Staged time - #: 500, Mean: 154590.39, 0/5/25/50/75/95/100: 100771.0/102149.0/107035.0/111119.0/122820.0/174272.0/1.6202133E7
  */
 
 trait StagedZeroCFA extends DslExp with LamOpsExp with MapOpsExp with TupledFunctionsRecursiveExp with SetOpsExp {
@@ -23,8 +27,6 @@ trait StagedZeroCFA extends DslExp with LamOpsExp with MapOpsExp with TupledFunc
 
   //TODO: The store doesn't have to be staged, the addresses are known at stage-time.
   //      Try Store = Map[Addr, Rep[Set[Lam]]]
-
-  //implicit def StoreTyp: Typ[Store] = manifestTyp
 
   implicit def lift[T:Typ](lst: List[T]): Rep[List[T]] = List[T](lst.map(lift(_)):_*)
   def lift[T:Typ](x: T) = unit[T](x)
@@ -78,17 +80,17 @@ trait StagedZeroCFA extends DslExp with LamOpsExp with MapOpsExp with TupledFunc
   def analysisProgram(prog: Expr, σ: RStore): RStore = analysisCall(prog, σ)
 
   def analysisCall(call: Expr, σ: RStore): RStore = call match {
-    case App(f, args) ⇒ analysisApp(f, args, analysisArg(args, σ))
+    case App(f, args) ⇒ analysisApp(f, args, analysisArgs(args, σ))
     case Letrec(bds, body) ⇒
       val σ_* = σ.update(bds.map(_.name), bds.map(b => Set(lift(b.value.asInstanceOf[Lam]))))
-      val σ_** = analysisArg(bds.map(_.value), σ_*)
+      val σ_** = analysisArgs(bds.map(_.value), σ_*)
       analysisCall(body, σ_**)
   }
 
[…]

### 2018-10-27 paper/10_intro.tex
revise on intro

@@
 
-2. Multi-stage programming is widely used in high performance computation.
-Precise whole-program analysis can be very slow, how to optimize it systematically/mechanically?
-bring existing high-performance computation facilities, in the form of meta-programming, into
-semantic artifacts of abstract interpretation, i.e., abstract interpreters.
-and bridging the gap between designing sound analyses and implementing efficient analyses.
-without changing the abstract interpreter, just type
-
-partial evaluation on program analyzers is not a new idea, remove the interpretive overhead of repetively 
-traversing the abstract syntax tree.
-we revisit this approach from a more recent perspective by combining general abstract interpreters
-and multi-stage programming.
+(TODO: more intro on MSP)
+Multi-stage programming has been widely used to improve the performance in many applications, 
+such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
+numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
+generic programming \cite{DBLP:journals/pacmpl/Yallop17}, 
+data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
+query compilation of databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
+and etc.
+Specializing static analyses by partial evaluation had been emerged in 90s \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587}, 
+and indeed it is able to effectively remove the interpretive overhead of repetively traversing the abstract syntax tree.
+However, it is still worth to revisit the idea from a modern perspective based on meta-programming for general abstract interpreters.
+Practically, precise whole-program analyses can be very slow due to the inherent complexity of the algorithm and the large scale of the analyzed programs (TODO: cite); 
+particularly, we would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
+In this paper, we show how to achieve this goal by bring existing high-performance computing facilities, in the form of generative
+programming, to abstract interpretation semantic artifacts, i.e., abstract interpreters. We use LMS framework, which enables
+multi-stage programming by just changing the types of ther analyzer program, and we thus bridge the gap between designing sound 
+program analyses and implementing efficient analyzers.
 
 When staging a concrete interpreter, the programmers need to distinguish static and dynamic values
---- the given program is classified as static because it is known at compile-time, and the inputs to that program 
-are dynamic. However, when staging an abstract interpreter, this distinction not exists anymore. 
-Because an abstract interpreter treats all the inputs as some form of abstract values, which usually are 
-top elements in their abstract domains and are statically known. Then what is the point of staging?
+--- the given program to be executed by the interpreter is classified as static because it is known 
+at compile-time, and the inputs to that program are dynamic. 
+However, when staging an abstract interpreter, this distinction not exists anymore.  
+Because the abstract interpreter instantiates all the inputs as some form of abstract values, which are usually
+top elements in their abstract domains and are statically known. 
+Then what is the point of staging if no such distinction?
 A surprising by-product of thinking about this question is to realize that we can apply the staged
@@
 the program (e.g., libraries) are dynamic input, then we obtain a modular analysis, mechanically.
+(TODO: unknown context)
 
@@
 \item Starting from a definitional abstract interpreter for a higher-order functional language,
-  we show how to turn it into a staged abstract interpreter by adding stage annotations with 
+  which the syntax and semantics are described in section \ref{bg_lang},
+  we show how to turn it into a staged abstract interpreter by adding stage annotations within
   Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
-  We later extend the abstract interpreter to support
-  imperative language features, such as states, loops and exceptions.
-\item We elaborate the idea that the staged abstract interpreter not only improves the performance 
-  but also the scalability, which are two major issues in static analysis, if we apply it on open
+  We later extend the staged abstract interpreter to support imperative language features 
+  (section \ref{cases_imp}), such as states, loops and exceptions.
+\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{performance}
+  but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
   programs. TODO: more on this point.
-\item To evaluate the generality of the idea, we conduct several case studies. 
-  First we revisited Abstract Compilation (TODO: cite), which is a closure generation technique applied 
-  on 0-CFA and can be understood as an instance of staged abstract interpreter. 
-  Second, we demonstrate existing control-flow analysis techniques, such as context/flow/path-sensitive 
-  analysis and store-widening can be integrated with staged abstract interpreters.
-  Finally, we show it also can be applied to numerical analysis in an imperative language.
+\item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
+  First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
+  technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
+  staged abstract interpreter. 
+  Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
+  analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
+  Finally, we show that in an imperative language, the idea can be applied onto an abstract interpreter with
+  an abstract domain that performs numerical analysis.
   (Optional: taint/information flow analysis?)
-\item We also emprically evaluate the performance improved of staged abstract interpreter with the
-  unstaged version and other existing optimizations on benchmarks.
+\item We also emprically evaluate the performance improved by staging an big-step abstract interpreter 
+  for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 
+  and an optimizing small-step version.
 \end{itemize}

### 2018-10-27 paper/abstract.tex
revise on intro

@@
 By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
-By appying the abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
+By appying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
 we obtain a modular analysis that generates sound partial analysis results which can be composed 

### 2018-10-28 paper/10_intro.tex
intro

@@
 
-1. Futamura projection \cite{Futamura1999} shows that staging an interpreter yields a compiler.
-literature on this approach, creating DSL.
+Futamura projection \cite{Futamura1999, futamura1971partial} reveals the close relation between 
+compilers and interpreters, 
+and the first Futamura projection specifically shows that specializing an interpreter yields a compiler. 
+Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
+multi-stage programs that specializes and generates programs; the generated program then can be executed 
+at next stage. 
+As an approach to partial evaluation, MSP requires explicit staging annotations from the implementers.
+The staging annotations controls which part should be specialized, and are considered as manual 
+binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} can be 
+considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically.
 
-Recent research uncovers a systematic approach to construct abstract interpreters from concrete interpreters.
-Given the structural similarity between abstract and concrete interpreters, intellectually it is natural to 
-wonder what would be a staged abstract interpreter, and how to stage an abstract interpreter 
-having considered their functional dissimilarity.
+Recent research uncovers a systematic approach to construct sound abstract interpreters 
+from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+Given the structural similarity between concrete and abstract interpreters, intellectually 
+it is natural to wonder what would be a staged abstract interpreter, and how to effectively 
+specializing an abstract interpreter having considered their functional dissimilarity. 
+In this paper, we study the application of the first Futamura projection on abstract interpreters, 
+and the approach to construct efficient abstract interpreters by applying multi-stage programming.
 
-(TODO: more intro on MSP)
-Multi-stage programming has been widely used to improve the performance in many applications, 
-such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
+Multi-stage programming has been widely used to improve the performance in many domains, 
+such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
+DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
 numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
@@
 data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
-query compilation of databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
+query compilation in databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
 and etc.
-Specializing static analyses by partial evaluation had been emerged in 90s \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587}, 
-and indeed it is able to effectively remove the interpretive overhead of repetively traversing the abstract syntax tree.
-However, it is still worth to revisit the idea from a modern perspective based on meta-programming for general abstract interpreters.
-Practically, precise whole-program analyses can be very slow due to the inherent complexity of the algorithm and the large scale of the analyzed programs (TODO: cite); 
-particularly, we would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
-In this paper, we show how to achieve this goal by bring existing high-performance computing facilities, in the form of generative
-programming, to abstract interpretation semantic artifacts, i.e., abstract interpreters. We use LMS framework, which enables
-multi-stage programming by just changing the types of ther analyzer program, and we thus bridge the gap between designing sound 
-program analyses and implementing efficient analyzers.
+Likewise, specializing static analyses by partial evaluation had been emerged in 90s 
+\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
+and indeed it is able to effectively remove the interpretive overhead of 
+repeatedly traversing the abstract syntax tree. However, it is still worth 
+to investigate the idea deeper from a modern perspective based on meta-programming 
+for general abstract interpreters.
+Practically, precise whole-program analyses formulated as abstract interpretation 
+can be very slow due to the inherent complexity of the algorithm and the large scale 
+of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121} \todo{other cite?};
+particularly, we would like a systematic approach to optimize program 
+analyzers and meanwhile minimally modifying the analyzer programs.
+In the present paper, we show how to achieve this goal by bring existing 
+high-performance computing facilities, in the style of multi-stage generative
+programming, to abstract interpretation semantic artifacts, i.e., 
+abstract interpreters. We show the use LMS framework, that enables 
+multi-stage programming by just changing the types of the analyzer 
+program, and we thus bridge the gap between designing sound program 
+analyses and implementing efficient program analyzers.
 
-When staging a concrete interpreter, the programmers need to distinguish static and dynamic values
---- the given program to be executed by the interpreter is classified as static because it is known 
-at compile-time, and the inputs to that program are dynamic. 
-However, when staging an abstract interpreter, this distinction not exists anymore.  
-Because the abstract interpreter instantiates all the inputs as some form of abstract values, which are usually
-top elements in their abstract domains and are statically known. 
-Then what is the point of staging if no such distinction?
-A surprising by-product of thinking about this question is to realize that we can apply the staged
-abstract interpreter on \textit{open} programs, and the free variables representing other parts of 
-the program (e.g., libraries) are dynamic input, then we obtain a modular analysis, mechanically.
-(TODO: unknown context)
+When staging a concrete interpreter, the programmers need to distinguish 
+static and dynamic values --- the given program to be executed by the 
+interpreter is classified as static because it is known at compile-time, 
+and the inputs to that program are dynamic.  However, when staging an 
+abstract interpreter, this distinction not exists anymore. Because the 
+abstract interpreter instantiates all the inputs as some form of abstract 
+values, which are usually top elements in their abstract domains and are 
+also statically known. Then what is the point of staging if no such distinction?
+A surprising by-product of thinking about this question is to realize that 
+we can apply the staged abstract interpreter on \textit{open} programs, and
+the free variables representing other parts of the program (e.g., libraries) 
+are dynamic input, therefore we obtain a modular analysis through staging, 
+mechanically. \todo{unknown context}
 
+The use of staged abstract interpreters on open programs 
+This addresses 
 3. analyzing modern programs with large libraries are expensive. 
-why? 1. precise analysis usually required whole-program analysis; 2. to do wp analysis, need to 
+why? 1. precise analysis usually required whole-program analysis; 2. to do whole-program analysis, need to 
 repetitively analyze the parts of library.
@@
 part so that we can reuse them, remove redundant analyzing of libraries.
+
+Many optimization of static analysis are ad-hoc.
 
@@
   but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
-  programs. TODO: more on this point.
+  programs. \todo{more on this point.}
 \item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
@@
   (Optional: taint/information flow analysis?)
-\item We also emprically evaluate the performance improved by staging an big-step abstract interpreter 
+\item We also empirically evaluate the performance improved by staging an big-step abstract interpreter 
   for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 

### 2018-10-28 paper/50_cases.tex
update
revise on intro
intro

@@
-\section{Case Study}
+\section{Case Study} \label{cases_study}
 
-\subsection{Abstract Compilation, Revisited}
+\subsection{Abstract Compilation a la Staging}
 
@@
 
+Advantages: (almost?) no need to modify the analysis code; staging the store data structure
+
 (Fianally, we should briefly summarize the difference between them:
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Control Flow Analysis}
 
-context/path/flow-sensitive
+\subsubsection{context/path/flow-sensitive}
 
-store-widening
+\subsubsection{Store-widening}
+
+\subsubsection{Mixed Sensitivity}
 
@@
 
-\subsection{Scale to Imperative Languages}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Numerical Analysis in Imperative Languages} \label{cases_imp}
+
+Beyond control-flow analysis, staging data structures used in the abstract interpreter, such as abstract domain.
+It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
+improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17}.
+In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.
+
+\subsubsection{Scaling to Imperative Languages}
 
@@
 
-Options: Numerical Analysis, Data-flow analysis, taint analysis
+\subsubsection{Specializing Interval Domains}
+
+Intervals, Numerical Analysis
+
+Options: Data-flow analysis, taint analysis

### 2018-10-28 paper/abstract.tex
misc

--- 2018-10-23 paper/abstract.tex
@@
 By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
-By appying the abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
+By applying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
 we obtain a modular analysis that generates sound partial analysis results which can be composed 
@@
 Using the idea of staged abstract interpreters, we show several case studies, including 
-\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of Shiver's 0-CFA, pushdown
+\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of  0-CFA, pushdown
 control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
-analysis on a imperative language.
+analysis on an imperative language.
 

### 2018-10-28 paper/10_intro.tex
intro

@@
 Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
-multi-stage programs that specializes and generates programs; the generated program then can be executed 
-at next stage. 
+generic programs that and generates specialized and efficient programs; 
+the generated program then can be executed at next stage. 
 As an approach to partial evaluation, MSP requires explicit staging annotations from the implementers.
-The staging annotations controls which part should be specialized, and are considered as manual 
-binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} can be 
-considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically.
+The staging annotations controls which part of the program should be specialized, and are considered 
+as manual binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} 
+can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically.
 
@@
 specializing an abstract interpreter having considered their functional dissimilarity. 
-In this paper, we study the application of the first Futamura projection on abstract interpreters, 
-and the approach to construct efficient abstract interpreters by applying multi-stage programming.
+This paper studies the application of the first Futamura projection on abstract interpreters, 
+and the approach to construct efficient abstract interpreters by multi-stage programming.
 
@@
 analyzers and meanwhile minimally modifying the analyzer programs.
-In the present paper, we show how to achieve this goal by bring existing 
+In the present work, we show how to achieve this goal by bring existing 
 high-performance computing facilities, in the style of multi-stage generative
 programming, to abstract interpretation semantic artifacts, i.e., 
-abstract interpreters. We show the use LMS framework, that enables 
-multi-stage programming by just changing the types of the analyzer 
-program, and we thus bridge the gap between designing sound program 
-analyses and implementing efficient program analyzers.
+abstract interpreters.
+
+The slogan of multi-stage programming is "abstraction without regret". 
+Of course, all interpreters are metalinguistic abstractions, but some interpreters are more abstract 
+than others \todo{maybe rephrase}. 
+One technical challenge is to deal with the non-determinisism, fixed-point and abstract domains in 
+abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
+should be written in a style that is easy for people to communicate the formulation, but also be friendly
+to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
+We show the use LMS framework, that enables multi-stage programming by just changing the types 
+of the analyzer program, and we thus bridge the gap between designing sound program analyses 
+and implementing efficient program analyzers.
 
@@
 are dynamic input, therefore we obtain a modular analysis through staging, 
-mechanically. \todo{unknown context}
+mechanically.
 
-The use of staged abstract interpreters on open programs 
-This addresses 
-3. analyzing modern programs with large libraries are expensive. 
-why? 1. precise analysis usually required whole-program analysis; 2. to do whole-program analysis, need to 
-repetitively analyze the parts of library.
-As modularity in compilers, we need to a way to separately analyze programs, and create summaries of analyzed 
-part so that we can reuse them, remove redundant analyzing of libraries.
+The use of staged abstract interpreters on open programs improves the scalability.
+One challenge in static analysis of modern software is that the programs are usually 
+shipped with large library code, for example, it is observed that analyzing a simple
+"Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
+Soundly and precisely analyzing these libraries sometimes are expensive and unnecessary repeatedly.
+When applying the staged abstract interpreter on libraries, we leave the unknown 
+calling context as dynamic value and generate partial analyzing result which is 
+repsented as a residul program. The partial analyzing result can be resued and composed
+with the analyzing result of application code later. \todo{we show this can be applied 
+to context/path/flow-sensitive even with different sensitivity parameters.}
+It has been observed library code requires less precision \todo{cite}.
 
-Many optimization of static analysis are ad-hoc.
+% Many optimization of static analysis are ad-hoc.
 

### 2018-10-29 paper/10_intro.tex
intro

@@
 Given the structural similarity between concrete and abstract interpreters, intellectually 
-it is natural to wonder what would be a staged abstract interpreter, and how to effectively 
-specializing an abstract interpreter having considered their functional dissimilarity. 
+it is natural to raise the question whether it is possible to stage an abstract interpreter, and how 
+to effectively specialize an abstract interpreter having considered their functional dissimilarity. 
 This paper studies the application of the first Futamura projection on abstract interpreters, 
@@
 than others \todo{maybe rephrase}. 
-One technical challenge is to deal with the non-determinisism, fixed-point and abstract domains in 
-abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
+One technical challenge is to deal with the non-determinisism, fixed-point iterations and abstract 
+domains in abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
 should be written in a style that is easy for people to communicate the formulation, but also be friendly
@@
 One challenge in static analysis of modern software is that the programs are usually 
-shipped with large library code, for example, it is observed that analyzing a simple
+shipped with large library code, for example, it has been shown that analyzing a simple
 "Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
@@
 repsented as a residul program. The partial analyzing result can be resued and composed
-with the analyzing result of application code later. \todo{we show this can be applied 
-to context/path/flow-sensitive even with different sensitivity parameters.}
-It has been observed library code requires less precision \todo{cite}.
+with the analyzing result of application code when available later. 
+It has been observed that partially applying context-sensitivity on selected portion of the program could
+improve the precision and efficiency \cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. 
+We show that staging abstract interpreters as an approach to effectively implement hybird context-sensitivity\todo{}.
 
+staged polymorphism, decide when to compile an analysis.  
 % Many optimization of static analysis are ad-hoc.
@@
   (Optional: taint/information flow analysis?)
-\item We also empirically evaluate the performance improved by staging an big-step abstract interpreter 
+\item We also empirically evaluate the speed improved by staging an big-step abstract interpreter 
   for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 

### 2018-10-29 paper/70_related.tex
intro
continue working on outline
misc

@@
 
+Abstract compilation for analyzing logic programs\cite{DBLP:conf/flops/AmatoS01}
+
 \citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
@@
 
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
+\textbf{Abstract Interpreters}.
+
+Abstract interpretation \cite{DBLP:conf/popl/CousotC77}.
+
+Abstracting Definitional Interpreter \cite{darais2017abstracting}
+
+Polyhedra, online decomposition
+
+\textbf{Modular Static Analysis}.
+
+Cousot \& Cousot \cite{DBLP:conf/cc/CousotC02}, 
+Compositional shape analysis \cite{DBLP:conf/popl/CalcagnoDOY09}
+
+\textbf{Multi-Stage Programming}.
 
 Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
+
+TODO: compare to other staging approach (e.g., MetaOCaml)

### 2018-10-29 paper/abstract.tex
intro

@@
 
-We also empirically evaluate the improvement of performance on benchmark programs for a subset of Scheme.
-The result shows ...
+We also empirically evaluate the improvement of performance on control-flow analysis of benchmark programs.
+The result shows an average speedup of Nx when staging to Scala for a monovariant analysis, and Mx for polyvariant analysis.

### 2018-10-29 paper/10_intro.tex
worked on paper

@@
 and etc.
-Likewise, specializing static analyses by partial evaluation had been emerged in 90s 
+Likewise, specializing static analyses by partial evaluation had been emerged in 1990s 
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
@@
 The slogan of multi-stage programming is "abstraction without regret". 
-Of course, all interpreters are metalinguistic abstractions, but some interpreters are more abstract 
+Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract"
 than others \todo{maybe rephrase}. 
@@
   (section \ref{cases_imp}), such as states, loops and exceptions.
-\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{performance}
+\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{efficiency}
   but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
   programs. \todo{more on this point.}
-\item To evaluate the generality and applicability of the idea, we conduct three case studies in section \ref{cases_study}. 
+\item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in section \ref{cases_study}. 
   First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 

### 2018-10-29 paper/50_cases.tex
worked on paper

@@
 implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-Their work is esentially a partial evaluated CE (\textit{Control-Environment}) abstract
-interpreter written in a big-step semantics for programs written in continuation-passing style.
-The analysis it can perform is a monovariant control flow analysis, sepcifically, Shiver's original
-formulation of 0-CFA \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
-After the partial evaluation on input program, the \textit{control} part of abstract interpreter
-is disappeared, instead, the compiled analysis is a closure (higher-order function) and
-only takes an environment as argument,
-and the overhead of traversing the abstract syntax tree of source program also has been eliminated.
-Since the target language is written a continuation-passing style, so there is
-no explict evaluation context component when constructing the abstract interpreter
-(rather than a CEK machine).
+The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
+at compile-time therefore the overhead if interpretation can be eliminated. The compiled analysis
+can be represented by text or closures (higher-order functions).
+Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
+analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
+for continuation-passing style (CPS) programs. Since the analyzed program is in CPS, the analyzer 
+is essentially a big-step CE (\textit{Control-Environment}) abstract interpreter.
+After specializing the analyzer with respect to the input program, the \textit{control} part of 
+abstract interpreter is eliminated, instead, the compiled analysis is a closure which takes an 
+environment as argument. Therefore the overhead of traversing the abstract syntax tree of input 
+program also has been eliminated.
 
-In this section, we firstly review their work of closure generation,
-and reproduce their result by adding stage annotations to 0-CFA.
-Show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be understood as an instance of staged 
-abstract interpreter.
+In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be 
+understood and implemented as an instance of staging abstract interpreters.
+We firstly revisit abstract compilation in the style of closure generation,
+and then reproduce their result by adding stage annotations to the unstaged 0-CFA algorithm.
+The generated program of our approach improves approximately the same extent of speed (roughly a factor of 2),
+but without changing a single line of the analyzer program (with the use of LMS), 
+while closure generation requires more engineering effort, such as a whole-program 
+conversion on the analyzer. Moreover, our approach is able to not only remove the interpretive
+overhead, but also specialize the data structures used in the analysis, for example, the store
+that maps variables to sets of lambda.
 
-Advantages: (almost?) no need to modify the analysis code; staging the store data structure
+\subsubsection{Closure Generation}
 
-(Fianally, we should briefly summarize the difference between them:
-staging vs partial evaluation; code generation; engineering effort; performance)
+\subsubsection{Staged 0-CFA}
 
-(Specific -> General vs General vs Specific)
+\subsubsection{Specializing the Data Structure}
 
@@
 
-\subsection{Control Flow Analysis}
+\subsection{Control-flow Analysis}
 
-\subsubsection{context/path/flow-sensitive}
+The target language we presented in section \ref{bg_lang} is essentially a higher-order functional language.
+One fundamental analysis task for functional programs is the control-flow analysis, i,e., determining
+which functions will possibly be applied at each call-site. 
+Though, in the previous section, we discussed abstraction compilation and staging on a monovariant 
+control-flow analysis for programs written in continuation-passing style; 
+in this section, based on the staged abstract interpreter we presented in section \ref{sai}, 
+we further discuss staging control-flow analysis for direct style languages in depth, 
+especially polyvariant analyses.
 
-\subsubsection{Store-widening}
+\subsubsection{Context/Path/Flow-Sensitivity}
 
@@
 using different $k$ for $k$-CFA
+
+\subsubsection{Store-widening}
 

### 2018-10-29 paper/50_cases.tex
add ref

@@
 It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
-improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17}.
+improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
 In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.

### 2018-10-29 paper/20_bg.tex
continue working on outline
revise on intro
worked on paper
misc

@@
 
-\subsection{The Language}
+\subsection{The Language} \label{bg_lang}
+
+syntax, higher-order functional, variables, lambda, app, fix, let, arith, if0
+
+concrete interpreter.
 
 \subsection{From Interpreter to Abstract Interpreter}
+
+big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800}.
+
+monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+
+\subsection{Multi-Stage Programming}
+
+general idea; LMS framework in scala. MetaOCaml?
+
+Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
+and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
+Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
+escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
+still applies with other MSP implementations.

### 2018-10-29 paper/50_cases.tex
misc

@@
 is essentially a big-step CE (\textit{Control-Environment}) abstract interpreter.
-After specializing the analyzer with respect to the input program, the \textit{control} part of 
+After specializing the analyzer with respect to the input program, the \textit{control} component of 
 abstract interpreter is eliminated, instead, the compiled analysis is a closure which takes an 

### 2018-10-29 paper/70_related.tex
misc

@@
 techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
+(\textit{0}-CFA) for programs written in continuation-passing style. 
+The key idea is to remove the interpretation overhead on traversing the syntax 
+tree by partial evaluation. Specifically, they proposed two similar kinds of abstract compilation techniques.
 The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
+can be loaded and executed by \texttt{eval} or other similar mechanisms. 
+The second one is using closure, i.e., a function that remembers its environment, 
+as a representation of specialized analysis. 
+So compiling the analysis generates higer-order functions on-the-fly with respect to the analyzed program,
+then the generated closure can be applied immediately in the higher-order host language.
 
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}. The source program is firstly compiled
-to intermediate form called "abstract bytecode", which are actually higher-order functions, 
-and then executed on a abstract abstract machine.
+\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea closure generation 
+for optimizing small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
+The analyzed program is firstly compiled to a intermediate representation called 
+"abstract bytecode", which are actually higher-order functions, and then be executed 
+on a abstract abstract machine for that IR.
 
-Abstract compilation for analyzing logic programs\cite{DBLP:conf/flops/AmatoS01}
+%\citeauthor{DBLP:conf/flops/AmatoS01} apply abstract compilation for non par-sharing and freeness analysis in logic programs \cite{DBLP:conf/flops/AmatoS01}.
 
@@
 
-\citeauthor{hardekopf2011flow} apply staging to flow-sensitive pointer analysis \cite{hardekopf2011flow}. 
+\citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. 
 They firstly stage the program code to obtain a sparse representation, then conduct
@@
 
-\textbf{Multi-Stage Programming}.
+\textbf{Meta-Programming}.
 
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
+Partial evaluation. 
+Reflexive language \cite{DBLP:conf/gpce/Asai14}
+Collapsing Towers of Interpreters \cite{Amin:2017:CTI:3177123.3158140}
 
-TODO: compare to other staging approach (e.g., MetaOCaml)
+Multi-stage programming. MetaML, MetaOCaml, Mint, LMS

### 2018-11-02 paper/abstract.tex
attempt at revising abstract

--- 2018-10-28 paper/abstract.tex
@@
-It is well known that a staged interpreter is a compiler, which provides performance improvement
-by specializing the interpreter to a given program.
+It is well known that a staged interpreter is a compiler: specializing the interpreter to a given program
+produces an equivalent program that runs faster.
+It is even more widely known that an abstract interpreter is a program analyzer: tweaking the 
+interpreter to run on a domain of abstract values produces a sound static analysis.
 
-In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
-staged abstract interpreters. 
-By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
-By applying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
+What happens when we combine these two ideas, and apply staging to an \emph{abstract} interpreter? 
+We obtain a sound static analysis, specialized for a given program, that runs faster.
+%
+More surprisingly, we show that by applying the staged abstract interpreter to \textit{open} programs and considering the free variables as dynamic inputs, 
 we obtain a modular analysis that generates sound partial analysis results which can be composed 
-and reused later without losing precision, though the original abstract interpreter is a whole-program 
+and reused later without losing precision, even though the original abstract interpreter is a whole-program 
 analysis algorithm.
 
-Using the idea of staged abstract interpreters, we show several case studies, including 
+Based on the idea of staged abstract interpreters, we show several case studies, including 
 \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of 0-CFA, pushdown
@@
 
-We also empirically evaluate the improvement of performance on benchmark programs for a subset of Scheme.
-The result shows ...
+We empirically evaluate the performance improvements on control-flow analysis of benchmark programs.
+The results show an average speedup of Nx with staging for a monovariant analysis, and Mx for 
+a polyvariant analysis.
+
+% It is well known that a staged interpreter is a compiler, which provides performance improvement
+% by specializing the interpreter to a given program.
+
+% In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
+% staged abstract interpreters. 
+% By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
+% By applying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
+% we obtain a modular analysis that generates sound partial analysis results which can be composed 
+% and reused later without losing precision, though the original abstract interpreter is a whole-program 
+% analysis algorithm.
+
+% Using the idea of staged abstract interpreters, we show several case studies, including 
+% \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of  0-CFA, pushdown
+% control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
+% analysis on an imperative language.
+
+% We also empirically evaluate the improvement of performance on control-flow analysis of benchmark programs.
+% The result shows an average speedup of Nx when staging to Scala for a monovariant analysis, and Mx for polyvariant analysis.

### 2018-11-02 paper/10_intro.tex
notes on intro

@@
 
-\subsection{Motivation}
+% \subsection{Motivation}
 
-Futamura projection \cite{Futamura1999, futamura1971partial} reveals the close relation between 
-compilers and interpreters, 
-and the first Futamura projection specifically shows that specializing an interpreter yields a compiler. 
+Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close connection between 
+compilers and interpreters. 
+The first Futamura projection specifically shows that specializing an interpreter yields a compiler. \note{TR: not true!!}
 Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
-generic programs that and generates specialized and efficient programs; 
+generic programs that and generates specialized and efficient programs;  \note{TR: connection from Futamura to MSP doesn't flow well}
 the generated program then can be executed at next stage. 
@@
 can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically. 
+\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's who to implement
+the futamura projection idea in practice''}
 
-Recent research uncovers a systematic approach to construct sound abstract interpreters 
+Recent research \note{TR: Cousot was already systematic.
+Both Futamura and Cousot are results from the 1970s -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters
 from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
@@
 query compilation in databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
-and etc.
+etc.
 Likewise, specializing static analyses by partial evaluation had been emerged in 1990s 
@@
 to investigate the idea deeper from a modern perspective based on meta-programming 
-for general abstract interpreters.
+for general abstract interpreters. \note{TR: sounds weak, what is new?}
 Practically, precise whole-program analyses formulated as abstract interpretation 
@@
 of the analyzer program, and we thus bridge the gap between designing sound program analyses 
-and implementing efficient program analyzers.
+and implementing efficient program analyzers. \note{TR: a lot of back and forth here, where is story going?}
 
@@
 and the inputs to that program are dynamic.  However, when staging an 
-abstract interpreter, this distinction not exists anymore. Because the 
+abstract interpreter, this distinction does not exist anymore. Because the 
 abstract interpreter instantiates all the inputs as some form of abstract 
 values, which are usually top elements in their abstract domains and are 
-also statically known. Then what is the point of staging if no such distinction?
+also statically known. 
+Then what is the point of staging if there is no such distinction?
+\note{TR: I don't understand this. The program
+structure is static, the abstract values are still dynamic, no? They 
+change in every iteration of the fixpoint algorithm}
 A surprising by-product of thinking about this question is to realize that 
@@
 The use of staged abstract interpreters on open programs improves the scalability. 
+\note{TR: why is it a big deal. On Closed programs we obtain a constant factor, but modular analysis has different asymptotics}
 One challenge in static analysis of modern software is that the programs are usually 
@@
 \item Starting from a definitional abstract interpreter for a higher-order functional language,
-  which the syntax and semantics are described in section \ref{bg_lang},
-  we show how to turn it into a staged abstract interpreter by adding stage annotations within
-  Lightweight Modular Staging (LMS) framework, a library-based approach to code generation. 
+  whose syntax and semantics are described in section \ref{bg_lang},
+  we show how to turn it into a staged abstract interpreter by adding stage annotations with
+  the Lightweight Modular Staging (LMS) framework. %, a library-based approach to code generation. 
   We later extend the staged abstract interpreter to support imperative language features 
   (section \ref{cases_imp}), such as states, loops and exceptions.
-\item We elaborate the idea that the staged abstract interpreter not only improves the \textit{efficiency}
+\item We demonstrate that the staged abstract interpreter not only improves the \textit{efficiency}
   but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
@@
 \item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in section \ref{cases_study}. 
-  First we revisited Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
+  First we revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
   technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
-  staged abstract interpreter. 
+  our staged abstract interpreter. 
   Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
   analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
-  Finally, we show that in an imperative language, the idea can be applied onto an abstract interpreter with
+  Finally, we show that in an imperative language, the idea can be applied to an abstract interpreter with
   an abstract domain that performs numerical analysis.
-  (Optional: taint/information flow analysis?)
-\item We also empirically evaluate the speed improved by staging an big-step abstract interpreter 
+  % (Optional: taint/information flow analysis?)
+\item We empirically evaluate the performance improvement by staging an big-step abstract interpreter 
   for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 

### 2018-11-02 paper/10_intro.tex
fix paper

@@
 compilers and interpreters. 
-The first Futamura projection specifically shows that specializing an interpreter yields a compiler. \note{TR: not true!!}
+The first Futamura projection specifically shows that specializing an interpreter with respect to the input program yields 
+an equivalent executable.
 Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
@@
   whose syntax and semantics are described in section \ref{bg_lang},
-  we show how to turn it into a staged abstract interpreter by adding stage annotations with
-  the Lightweight Modular Staging (LMS) framework. %, a library-based approach to code generation. 
+  we show how to turn it into a staged abstract interpreter (SAI) by adding stage annotations. 
+  The design and implementation of SAI using Scala and Lightweight Modular Staging (LMS) framework is 
+  described in section \ref{sai}.
   We later extend the staged abstract interpreter to support imperative language features 

### 2018-11-02 paper/20_bg.tex
fix paper

@@
+\newcommand{\TLang}{$L_\lambda$}
+
 \section{Background}
 
-\subsection{The Language} \label{bg_lang}
+In this section, we first described the language we will model and analyze, and present a unstaged concrete 
+interpreter for it. Then, starting from this unstaged concrete interpreter, we proceed in two directions:
+by abstracting it we obtain an unstaged abstract interpreter; by staging it we obtain a staged concrete
+interpreter.
 
-syntax, higher-order functional, variables, lambda, app, fix, let, arith, if0
+\subsection{The Target Language} \label{bg_lang}
 
-concrete interpreter.
+As higher-order functions had been widely introduced in mainstream programming languages \todo{cite}, in this paper, 
+we target a tiny higher-order functional language \TLang, which is based on direct-style lambda-calculus, 
+and with the support of numbers, arithmetics, recursions and conditionals.
+In section \ref{cases_imp}, we will see adding more imperative features to the language.
+Since we are mostly interested in analyzing the dynamic behavior of the program, the static semantics and
+type system will be omitted. We also assume all the input programs are well-typed.
+The abstract syntax is described below:
+
+\begin{lstlisting}
+abstract class Expr
+case class Lit(i: Int)
+case class Var(x: String)
+case class Lam(x: String, e: Expr)
+case class App(e1: Expr, e2: Expr)
+case class AOp(op: String, e1: Expr, e2: Expr)
+case class Rec(x: String, e: Expr, body: Expr)
+case class If0(cnd: Expr, thn: Expr, els: Expr)
+\end{lstlisting}
+
+We give the concrete semantics of \TLang by showing a big-step definitional interpreter for it.
+But before doing that, let's first see a general interface for interpreters, whether concrete or abstract, 
+staged or unstaged. The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Ident},
+\texttt{Addr} and etc., which will be instantiated concretely later. The abstract types \texttt{Env} and \texttt{Store}
+are required to conform their specifications, \texttt{EnvSpec} and \texttt{StoreSpec}, which define 
+the operations such as lookup and update on them, nevertheless the detail of them are omitted.  
+The store component is necessary when modeling side effects such as mutations. 
+The trait also requires the implementators provide the initial values of environment and store.
+
+%TODO: talk about alloc here?
+
+\begin{lstlisting}
+trait Semantics[R[_]] {
+  type Ident;    type Addr;   type Value;    type Ans
+  type Env <: EnvSpec[R, Ident, Addr, Env]
+  type Store <: StoreSpec[R, Addr, Value, Store]
+  val $\rho$0: Env;     val $\sigma$0: Store
+  def alloc(x: Ident, $\sigma$: Store): R[Addr]
+  type EvalFun = (Expr, Env, Store) => Ans
+  def eval(ev: EvalFun)(e: Expr, $\rho$: Env, $\sigma$: Store): Ans
+  def fix(ev: EvalFun => EvalFun): EvalFun = 
+    (e, $\rho$, $\sigma$) =>  ev(fix(ev))(e, $\rho$, $\sigma$)
+  def eval_top(e: Expr): Ans = fix(eval)(e, $\rho$0, $\sigma$0)
+}
+\end{lstlisting}
+
+The evaluator \texttt{eval} is a open-recursive function that takes one and three arguments: the first
+one \texttt{ev} represents the recursive function itself; the rest three are the AST representing program 
+will be executed, the environment that maps from identifiers to addresses, and the store that maps addresses to values.
+The interpreters that take control, environment and store components also can be obtained by 
+refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
+machine \cite{DBLP:conf/popl/FelleisenF87}.
+We provide a function \texttt{fix} to close the recursion by providing the result of \texttt{fix(ev)} 
+to argument \texttt{ev}. Finally, a default implementation for the top-level evaluation function \texttt{eval\_top} is defined.
+As we will see later, the open-recursion style is beneficial when implementing abstract interpreter and 
+staged abstract interpreter \todo{cite adi}.
+
+Readers may notice an additional notation \texttt{[R[\_]]} afterward the trait definition,
+which indicates \texttt{Semantics} takes a type parameter \texttt{R}, where \texttt{R} represents
+a higher-kinded type and will used to specify the binding-time, i.e., staged or not. 
+For the moment, \texttt{R} will be instantiated unstaged --- an identity type constructor: 
+\texttt{type NoRep[A] = A}. \todo{where talk about stage poly?}
+
+Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
+Since the interpreter is concrete, the address space is infinite thus type \texttt{Int} is used here.
+The final returned type \texttt{Ans} is a pair of \texttt{Value} and \texttt{Store}, where
+the values are either closure or numbers.
+
+\begin{lstlisting}
+trait Concrete[R[_]] extends Semantics[R] {
+  type Ident = String;         type Addr = Int
+  type Ans = (Value, Store);   abstract class Value
+  case class CloV($\lambda$: Lam, $\rho$: Env) extends Value
+  case class NumV(i: Int) extends Value
+}
+\end{lstlisting}
+
+Then we define evaluation function that dispatches differently for different expression.
+Note that an allocation function is used when we need to update the store: 
+in the concrete case, function \texttt{alloc} always returns a fresh address of the store.
+
+\begin{lstlisting}
+object NoRepConcreteEval extends Concrete[NoRep] {
+  def alloc(x: Ident, $\sigma$: Store): Addr = $\sigma$.size + 1
+  def eval(ev: EvalFun)(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = 
+    e match {
+      case Lit(i) => (NumV(i), $\sigma$)
+      case Var(x) => ($\sigma$($\rho$(x)), $\sigma$)
+      case Lam(x, e) => (CloV(Lam(x,e), $\rho$), $\sigma$)
+      case App(e1, e2) =>
+        val (CloV(Lam(x,e), $\lambda\rho$), e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
+        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
+        val $\alpha$ = alloc(x, e2$\sigma$)
+        ev(e, $\lambda\rho$ + (x -> $\alpha$), e2$\sigma$ + ($\alpha$ -> e2v))
+      case If0(cnd, thn, els) => ...
+      case AOp(op, e1, e2) => ...
+      case Rec(x, e, body) => ...
+    }
+}
+\end{lstlisting}
 
@@
 
-big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800}.
+\begin{lstlisting}
+trait Abstract[R[_]] extends Semantics[R] {
+  type Ident = String;         case class Addr(x: Ident)
+  type Ans = (Value, Store);   type Value = Set[AbsValue]
+  abstract class AbsValue
+  case class CloV($\lambda$: Lam, $\rho$: Env) extends AbsValue
+  case class NumV() extends AbsValue
+}
+\end{lstlisting}
 
-monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+Address just record the variable name, more sensitivity in section ...
 
-\subsection{Multi-Stage Programming}
+other big-step abstract interpreter, 
+big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
+monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
+arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
 
+\subsection{From Interpreter to Staged Interpreter}
+
+\subsubsection{Multi-Stage Programming}
+\todo{example, power?}
 general idea; LMS framework in scala. MetaOCaml?
@@
 still applies with other MSP implementations.
+
+\subsubsection{Staged Interpreters}

### 2018-11-02 paper/20_bg.tex
fix bug / revision

@@
 As higher-order functions had been widely introduced in mainstream programming languages \todo{cite}, in this paper, 
-we target a tiny higher-order functional language \TLang, which is based on direct-style lambda-calculus, 
-and with the support of numbers, arithmetics, recursions and conditionals.
+we target a tiny higher-order functional language \TLang, which is based on direct-style lambda-calculus
+with the support of numbers, arithmetics, recursions and conditionals.
 In section \ref{cases_imp}, we will see adding more imperative features to the language.
@@
 \begin{lstlisting}
-trait Semantics[R[_]] {
+trait Semantics {
+  type R[+T]
   type Ident;    type Addr;   type Value;    type Ans
@@
 
-Readers may notice an additional notation \texttt{[R[\_]]} afterward the trait definition,
-which indicates \texttt{Semantics} takes a type parameter \texttt{R}, where \texttt{R} represents
-a higher-kinded type and will used to specify the binding-time, i.e., staged or not. 
-For the moment, \texttt{R} will be instantiated unstaged --- an identity type constructor: 
-\texttt{type NoRep[A] = A}. \todo{where talk about stage poly?}
+Readers may notice an additional abstract type \texttt{R[+T]} is defined in the trait,
+which means \texttt{R} represents a higher-kinded type and will used to specify the binding-time, 
+i.e., staged or not. Since the LMS framework we are using allows us to specify binding-time purely
+based on types, the unstaged higher-kinded type can be simply instantiated as an identity type constructor: 
+\texttt{type NoRep[+T] = T}. \todo{where talk about stage poly?}
 
@@
 \begin{lstlisting}
-trait Concrete[R[_]] extends Semantics[R] {
+trait Concrete extends Semantics {
   type Ident = String;         type Addr = Int
@@
 \begin{lstlisting}
-object NoRepConcreteEval extends Concrete[NoRep] {
+object NoRepConcreteEval extends Concrete {
+  type R[+T] = NoRep[T]
   def alloc(x: Ident, $\sigma$: Store): Addr = $\sigma$.size + 1
@@
 \begin{lstlisting}
-trait Abstract[R[_]] extends Semantics[R] {
+trait Abstract extends Semantics {
   type Ident = String;         case class Addr(x: Ident)

### 2018-11-03 paper/10_intro.tex
bg

@@
 numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
-generic programming \cite{DBLP:journals/pacmpl/Yallop17}, 
+generic programming \cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, 
 data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
@@
 etc.
-Likewise, specializing static analyses by partial evaluation had been emerged in 1990s 
+Likewise, specializing static analyses by partial evaluation had been emerged in late 90s 
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 

### 2018-11-03 paper/20_bg.tex
bg

@@
 
-We give the concrete semantics of \TLang by showing a big-step definitional interpreter for it.
+We give the concrete semantics of \TLang by showing a big-step operational interpreter for it.
 But before doing that, let's first see a general interface for interpreters, whether concrete or abstract, 
@@
 based on types, the unstaged higher-kinded type can be simply instantiated as an identity type constructor: 
-\texttt{type NoRep[+T] = T}. \todo{where talk about stage poly?}
+\texttt{type NoRep[+T] = T}. In section \ref{stagedinterp}, \texttt{R} will be instantiated 
+as a staging annotation. \todo{where talk about stage poly?}
 
 Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
-Since the interpreter is concrete, the address space is infinite thus type \texttt{Int} is used here.
-The final returned type \texttt{Ans} is a pair of \texttt{Value} and \texttt{Store}, where
-the values are either closure or numbers.
+Since the interpreter is concrete, the address space is infinite thus type \texttt{Int} is used for \texttt{Addr}.
+The final returned type by the interpreter \texttt{Ans} is a pair of \texttt{Value} and \texttt{Store}, 
+where the values are either closure or numbers.
 
@@
 Then we define evaluation function that dispatches differently for different expression.
+As we mentioned before, the type \texttt{R} is concretize as \texttt{NoRep}.
 Note that an allocation function is used when we need to update the store: 
-in the concrete case, function \texttt{alloc} always returns a fresh address of the store.
+in the concrete case, \texttt{alloc} always returns a fresh address of the store.
 
 \begin{lstlisting}
+//TODO: maybe not show all the code
 object NoRepConcreteEval extends Concrete {
@@
         ev(e, $\lambda\rho$ + (x -> $\alpha$), e2$\sigma$ + ($\alpha$ -> e2v))
-      case If0(cnd, thn, els) => ...
-      case AOp(op, e1, e2) => ...
-      case Rec(x, e, body) => ...
+      case AOp(op, e1, e2) =>
+        val (e1v: NumV, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
+        val (e2v: NumV, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
+        (evalArith(op, e1v, e2v), e2$\sigma$)
+      case Rec(x, f, body) => 
+        val $\alpha$ = alloc(x, $\sigma$);  val $\rho$_* = $\rho$ + (x -> $\sigma$)
+        val (fv, f$\sigma$) = ev(f, $\rho$_*, $\sigma$)
+        val $\sigma$_* = f$\sigma$ + ($\alpha$ -> fv);  ev(body, $\rho$_*, $\sigma$_*)
+      case If0(cnd, thn, els) =>
+        val (cndv: NumV, cnd$\sigma$) = ev(cnd, $\rho$, $\sigma$)
+        if (cndv.i == 0) ev(thn, $\rho$, cnd$\sigma$)
+        else ev(els, $\rho$, cnd$\sigma$)
     }
@@
 
-\subsection{From Interpreter to Staged Interpreter}
+\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}
 
 \subsubsection{Multi-Stage Programming}
-\todo{example, power?}
-general idea; LMS framework in scala. MetaOCaml?
+\begin{lstlisting}
+val power5 = new DslDriver[Int, Int] {
+  def power(b: Rep[Int], x: Int): Rep[Int] =
+    if (x == 0) 1 else b * power(b, x-1)
+  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
+}
+\end{lstlisting}
+
+Talk about power example. general idea; LMS framework in scala. MetaOCaml?
 

### 2018-11-03 paper/20_bg.tex
wip...

@@
   val $\rho$0: Env;     val $\sigma$0: Store
-  def alloc(x: Ident, $\sigma$: Store): R[Addr]
+  def alloc(x: R[Ident], $\sigma$: Store): R[Addr]
   type EvalFun = (Expr, Env, Store) => Ans

### 2018-11-03 code/src/main/scala/sai/direct/core/ai/SADI2.scala

@@ -1,7 +1,7 @@
 package sai.direct.core.ai
 
 import scala.util.continuations._
-//import scala.language.implicitConversions
+import scala.language.implicitConversions
 import scala.language.higherKinds
 
 import sai.common.ai._
@@ -12,7 +12,7 @@ import sai.common._
 import scala.lms.tutorial._
 import scala.reflect.SourceContext
 import scala.lms.internal.GenericNestedCodegen
-import scala.lms.common.{SetOpsExp ⇒ _, ScalaGenSetOps ⇒ _, ListOpsExp ⇒ _, ScalaGenListOps ⇒ _, _}
+import scala.lms.common.{SetOpsExp ⇒ _, ScalaGenSetOps ⇒ _, ListOps ⇒ _, ListOpsExp ⇒ _, ScalaGenListOps ⇒ _, _}
 
 object Semantics {
   type NoRep[+T] = T
@@ -55,12 +55,47 @@ object Semantics {
 
     val ρ0: Env
     val σ0: Store
-    def alloc(x: Ident, σ: Store): R[Addr]
+    def alloc(x: R[Ident], σ: Store): R[Addr]
     def eval(ev: EvalFun)(e: Expr, ρ: Env, σ: Store): Ans
 
     type EvalFun = (Expr, Env, Store) ⇒ Ans
     def fix(ev: EvalFun => EvalFun): EvalFun = (e, ρ, σ) => ev(fix(ev))(e, ρ, σ)
     def eval_top(e: Expr): Ans = fix(eval)(e, ρ0, σ0)
+  }
+
+  trait Test extends Sem {
+    type R[+T] = NoRep[T]
+    //TODO: can this also be F-bound poly?
+    type Ident = String
+    type Addr = Int
+
+    trait EnvLike[R[+_],K,V,E] {
+      def bala(m: E, k: R[K]): R[V]
+      def +(m: E, kv: (R[K], R[V])): E
+    }
+    object EnvLike {
+      def apply[R[+_],K,V,E](implicit env: EnvLike[R,K,V,E]): EnvLike[R,K,V,E] = env
+    }
+    implicit class EnvOps[R[+_],K,V,E:({type λ[ε] = EnvLike[R,K,V,ε]})#λ](env: E) {
+      def bala(k: R[K]): R[V] = EnvLike[R,K,V,E].bala(env, k)
+    }
+
+    //trait OhMyEnv[E] extends EnvLike[NoRep,Ident,Addr,E]
+
+    type Env1 = Map[Ident, Addr]
[…]

### 2018-11-03 paper/10_intro.tex
fix typos in paper

@@
 analyzers and meanwhile minimally modifying the analyzer programs.
-In the present work, we show how to achieve this goal by bring existing 
+In the present work, we show how to achieve this goal by bringing existing 
 high-performance computing facilities, in the style of multi-stage generative
@@
 to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
-We show the use LMS framework, that enables multi-stage programming by just changing the types 
+We show the usage of the LMS framework, that enables multi-stage programming by just changing the types 
 of the analyzer program, and we thus bridge the gap between designing sound program analyses 

### 2018-11-04 paper/20_bg.tex
typo
cache type class

@@
 
-In this section, we first described the language we will model and analyze, and present a unstaged concrete 
+In this section, we first described the language we will model and analyze, and present an unstaged concrete 
 interpreter for it. Then, starting from this unstaged concrete interpreter, we proceed in two directions:
@@
 will be executed, the environment that maps from identifiers to addresses, and the store that maps addresses to values.
-The interpreters that take control, environment and store components also can be obtained by 
+This environment-and-store-passing style big-interpreter also can be obtained by 
 refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 

### 2018-11-04 code/src/main/scala/sai/direct/core/ai/SADI2.scala

@@ -29,7 +29,7 @@ object Semantics {
     def +(kv: (R[K], R[V])): R[E] = EnvSpec[R,K,V,E].+(ρ, kv)
   }
 
-  trait StoreSpec[R[_], K, V, S] {  //<: StoreSpec[R,K,V,S]] {
+  trait StoreSpec[R[+_], K, V, S] {  //<: StoreSpec[R,K,V,S]] {
     def get(s: R[S], k: R[K]): R[V]
     def getOrElse(s: R[S], k: R[K], dft: R[V]): R[V]
     def update(s: R[S], k: R[K], v: R[V]): R[S]
@@ -49,15 +49,25 @@ object Semantics {
     def ⊔(that: R[S]): R[S] = StoreSpec[R,K,V,S].⊔(σ, that)
   }
 
-  trait CacheSpec[R[_], K, V, C <: CacheSpec[R,K,V,C]] {
-    def inGet(k: R[K]): R[V]
-    def outGet(k: R[K]): R[V]
-    def inContains(k: R[K]): R[Boolean]
-    def outContains(k: R[K]): R[Boolean]
-    def outUpdate(k: R[K], v: R[V]): C
-    //def outUpdateSingle(k: K, v: R[V]): C
-    def outUpdateFromIn(k: R[K]): C
-    def ⊔(that: C): C
+  trait CacheSpec[R[+_], K, V, C] { // <: CacheSpec[R,K,V,C]
+    def inGet(c: R[C], k: R[K]): R[V]
+    def outGet(c: R[C], k: R[K]): R[V]
+    def inContains(c: R[C], k: R[K]): R[Boolean]
+    def outContains(c: R[C], k: R[K]): R[Boolean]
+    def outUpdate(c: R[C], k: R[K], v: R[V]): R[C]
+    def ⊔(c1: R[C], c2: R[C]): R[C]
+  }
+  object CacheSpec {
+    def apply[R[+_],K,V,C](implicit c: CacheSpec[R,K,V,C]): CacheSpec[R,K,V,C] = c
+  }
+  implicit class CacheOps[R[+_],K,V,C:({type λ[c] = CacheSpec[R,K,V,c]})#λ](c: R[C]) {
+    def inGet(k: R[K]): R[V] = CacheSpec[R,K,V,C].inGet(c, k)
+    def inContains(k: R[K]): R[Boolean] = CacheSpec[R,K,V,C].inContains(c, k)
[…]

### 2018-11-04 paper/20_bg.tex
misc

@@
 will be executed, the environment that maps from identifiers to addresses, and the store that maps addresses to values.
-This environment-and-store-passing style big-interpreter also can be obtained by 
+This environment-and-store-passing style big-step interpreter also can be obtained by 
 refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
@@
 
+co-inductive caching/truncated depth-first evaluation \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800, Rosendahl:AbsIntPL}
+
 \subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}

### 2018-11-04 paper/20_bg.tex
write about LMS in section 2.3.1

@@
 
+Lightweight modular staging (LMS) \todo{cite} is a library-based multi-stage programming framework 
+that enables type-safe dynamic code generation. Different from the traditional approach of using
+syntactic quasi-quotation, it solely relies on types to distinguish between binding times.
+
+In the example shown above, \texttt{power} is a staged function that takes 2 arguments where 
+\texttt{b} is declared as \texttt{Rep[Int]} meaning that \texttt{b} is a representation of type 
+\texttt{Int} whose underlying value will be available at the next stage. Meanwhile the actual 
+computation result of the \texttt{power} function  will be available at the next stage as well. 
+Thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. The generated code for
+\texttt{power5} consists of a function that takes only one argument that corresponds to its 
+representation \texttt{b} from the previous stage.
+
 Talk about power example. general idea; LMS framework in scala. MetaOCaml?

### 2018-11-04 paper/10_intro.tex
minor note

@@
 can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically. 
-\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's who to implement
+\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's how to implement
 the futamura projection idea in practice''}
 
-Recent research \note{TR: Cousot was already systematic.
-Both Futamura and Cousot are results from the 1970s -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters
+Recent research \note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have
+profited from more recent results (MSP and AAM)  -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters 
 from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
@@
 abstract interpreters.
+
+\note{TR: insert a commuting diagram here? concrete vs abstract,
+general (interpreter) vs specialized (compiler). can combine both dimensions.}
 

### 2018-11-06 paper/10_intro.tex
section refs

@@
 \item Starting from a definitional abstract interpreter for a higher-order functional language,
-  whose syntax and semantics are described in section \ref{bg_lang},
+  whose syntax and semantics are described in Section~\ref{bg_lang},
   we show how to turn it into a staged abstract interpreter (SAI) by adding stage annotations. 
   The design and implementation of SAI using Scala and Lightweight Modular Staging (LMS) framework is 
-  described in section \ref{sai}.
+  described in Section~\ref{sai}.
   We later extend the staged abstract interpreter to support imperative language features 
-  (section \ref{cases_imp}), such as states, loops and exceptions.
+  (Section~\ref{cases_imp}), such as states, loops and exceptions.
 \item We demonstrate that the staged abstract interpreter not only improves the \textit{efficiency}
@@
   programs. \todo{more on this point.}
-\item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in section \ref{cases_study}. 
+\item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in Section~\ref{cases_study}. 
   First we revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 

### 2018-11-06 paper/20_bg.tex
section refs

@@
 with the support of numbers, arithmetics, recursions and conditionals.
-In section \ref{cases_imp}, we will see adding more imperative features to the language.
+In Section~\ref{cases_imp}, we will see adding more imperative features to the language.
 Since we are mostly interested in analyzing the dynamic behavior of the program, the static semantics and
@@
 based on types, the unstaged higher-kinded type can be simply instantiated as an identity type constructor: 
-\texttt{type NoRep[+T] = T}. In section \ref{stagedinterp}, \texttt{R} will be instantiated 
+\texttt{type NoRep[+T] = T}. In Section~\ref{stagedinterp}, \texttt{R} will be instantiated 
 as a staging annotation. \todo{where talk about stage poly?}
@@
 
-Address just record the variable name, more sensitivity in section ...
+Address just record the variable name, more sensitivity in Section ...
 

### 2018-11-06 paper/50_cases.tex
section refs

@@
 
-The target language we presented in section \ref{bg_lang} is essentially a higher-order functional language.
+The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
 One fundamental analysis task for functional programs is the control-flow analysis, i,e., determining
 which functions will possibly be applied at each call-site. 
-Though, in the previous section, we discussed abstraction compilation and staging on a monovariant 
+Though, in Section~\ref{?}, we discussed abstraction compilation and staging on a monovariant 
 control-flow analysis for programs written in continuation-passing style; 
-in this section, based on the staged abstract interpreter we presented in section \ref{sai}, 
+in this section, based on the staged abstract interpreter we presented in Section~\ref{sai}, 
 we further discuss staging control-flow analysis for direct style languages in depth, 

### 2018-11-10 paper/10_intro.tex
revise intro

@@
 compilers and interpreters. 
-The first Futamura projection specifically shows that specializing an interpreter with respect to the input program yields 
-an equivalent executable.
-Multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} is a popular way to write 
-generic programs that and generates specialized and efficient programs;  \note{TR: connection from Futamura to MSP doesn't flow well}
-the generated program then can be executed at next stage. 
-As an approach to partial evaluation, MSP requires explicit staging annotations from the implementers.
-The staging annotations controls which part of the program should be specialized, and are considered 
-as manual binding-time analysis. On the other hand, partial evaluation \cite{DBLP:books/daglib/0072559} 
-can be considered as a two-stage (static or dynamic) instance of MSP but analyzing binding-time automatically. 
-\note{TR: are last two sentences relevant? perhaps rephrase in the direction of ``here's how to implement
-the futamura projection idea in practice''}
+The first Futamura projection specifically shows that specializing an interpreter with respect to the input 
+program yields an equivalent executable.
+Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to Futamura projection,
+which first identifies the binding-time of variables in the program, i.e., they can be known statically or dynamically, 
+and then specializes the program with respect to the given input of program.
+However, a precise binding-time analysis (BTA) is hard generally. As an alternative and practical approach to 
+specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 
+requires the stage annotations (i.e., binding-time annotations) from the programmers, these staging annotations 
+control which part of the input program should be specialized.
+As a simple example, we consider a power function: 
+\begin{lstlisting}
+def power(b: Int, e: Int): Int = 
+  if (e == 0) 1 else e * power(b, e-1)
+\end{lstlisting}
+If \texttt{e} is statically known, say 5, then we can specialize \texttt{power} with respect to \texttt{e = 5} and 
+generate a new specialized function as following:
+\begin{lstlisting}
+def power5(b: Int) = b * b * b * b * b
+\end{lstlisting}
 
-Recent research \note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have
-profited from more recent results (MSP and AAM)  -- build a connection?} uncovers a systematic approach to construct sound abstract interpreters 
-from concrete interpreters \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach to construct sound static analysis 
+by approximation is almost as old as Futamura projection.
+Some recent progress such as Abstracting Abstract Machines uncovers a methodology to derive sound abstract interpreters 
+from their concrete counterparts \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+\note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have
+profited from more recent results (MSP and AAM) -- build a connection?} 
 Given the structural similarity between concrete and abstract interpreters, intellectually 
@@
 and the approach to construct efficient abstract interpreters by multi-stage programming.
+
+\begin{figure}[h]
+  \begin{tikzpicture}
+  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
+  {
+    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} \\
+      \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} \\ };
+  \path[-stealth]
+    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
+    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
+    (m-2-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-2-2)
+    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
+  \end{tikzpicture}
+\end{figure}
 
@@
 and indeed it is able to effectively remove the interpretive overhead of 
-repeatedly traversing the abstract syntax tree. However, it is still worth 
-to investigate the idea deeper from a modern perspective based on meta-programming 
-for general abstract interpreters. \note{TR: sounds weak, what is new?}
-Practically, precise whole-program analyses formulated as abstract interpretation 
-can be very slow due to the inherent complexity of the algorithm and the large scale 
-of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121} \todo{other cite?};
-particularly, we would like a systematic approach to optimize program 
-analyzers and meanwhile minimally modifying the analyzer programs.
-In the present work, we show how to achieve this goal by bringing existing 
-high-performance computing facilities, in the style of multi-stage generative
-programming, to abstract interpretation semantic artifacts, i.e., 
-abstract interpreters.
+repeatedly traversing the abstract syntax tree. 
+However, despite of the previous work are mostly focus on one particular analysis,
+it is worth to investigate the idea from a modern perspective based on meta-programming,
+especially for a general abstract interpreter that models direct-style $\lambda$-calculus, imperative 
+features such as mutation, and different abstract domains.
+One technical challenge is the binding-time engineering when non-determinism, 
+fixed-point iterations, different abstract domains and staging are introduced at the same time.
+In this paper, we present an end-to-end design and implementation of staging an abstract interpreter,
+that means ont only the abstract interpreter, but also the data structure of abstract domains, 
+abstract environment and heap, and the fixpoint iteration are staged.
 
-\note{TR: insert a commuting diagram here? concrete vs abstract,
-general (interpreter) vs specialized (compiler). can combine both dimensions.}
+On the other hand, the abstract interpreter, as a semantic artifact, should be written in a 
+style that is easy for people to communicate the formulation and abstraction, but also can be implemented 
+efficiently. As the solgan of multi-stage programming said, "abstraction without regret",
+we draw connection between abstract formulation and efficient implementation \todo{wording},
+just like the connection between interpreters and compilers drawn by Futamura projection.
+Particularly, we show a systematic way of adding stage annotations to the abstract interpreter
+without changing the code of interpreter skeleton that is shared between four concrete or abstract,
+unstaged or staged interpreters. We use LMS framework for staging, which allows only use types to 
+annotate the binding-time. Therefore, the proposed approach bridges the gap between designing sound 
+static analysis and implementing efficient program analyzer.
 
-The slogan of multi-stage programming is "abstraction without regret". 
-Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract"
-than others \todo{maybe rephrase}. 
-One technical challenge is to deal with the non-determinisism, fixed-point iterations and abstract 
-domains in abstract interpreters when staging is introduced. The abstract interpreter, as a semantic artifact, 
-should be written in a style that is easy for people to communicate the formulation, but also be friendly
-to specialization and generating efficient code. \todo{We discuss the form of abstract interpreters for staging}
-We show the usage of the LMS framework, that enables multi-stage programming by just changing the types 
-of the analyzer program, and we thus bridge the gap between designing sound program analyses 
-and implementing efficient program analyzers. \note{TR: a lot of back and forth here, where is story going?}
+%Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract" than others \todo{maybe rephrase}. 
+%Particularly, we also would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
+
+From the performance-wise, precise whole-program analyses formulated as abstract 
+interpretation can be very slow due to the inherent complexity of the algorithm and the large scale 
+of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121}. \todo{where put this}
 
@@
 staged polymorphism, decide when to compile an analysis.  
+end-to-end staging: abstract interpreter, abstract domain, environment/store structure, fixpoint iteration.
 % Many optimization of static analysis are ad-hoc.

### 2018-11-10 paper/10_intro.tex
add some stuff

@@
 program yields an equivalent executable.
-Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to Futamura projection,
+Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to realize Futamura projections,
 which first identifies the binding-time of variables in the program, i.e., they can be known statically or dynamically, 
 and then specializes the program with respect to the given input of program.
-However, a precise binding-time analysis (BTA) is hard generally. As an alternative and practical approach to 
-specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 
+However, precisely analyzing binding-time for an arbitray program is hard in general. As an alternative and more practical 
+approach to specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 
 requires the stage annotations (i.e., binding-time annotations) from the programmers, these staging annotations 
@@
 by approximation is almost as old as Futamura projection.
-Some recent progress such as Abstracting Abstract Machines uncovers a methodology to derive sound abstract interpreters 
-from their concrete counterparts \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+Some recent progress such as Abstracting Abstract Machines (AAM) uncovers a methodology to derive sound abstract interpreters 
+from their concrete counterparts and has been applied to different variants of definitional interpreters and abstract machines
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
 \note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have

### 2018-11-10 paper/20_bg.tex
revise intro
add some stuff

--- 2018-11-04 paper/20_bg.tex
@@
 
-\subsection{The Target Language} \label{bg_lang}
-
-As higher-order functions had been widely introduced in mainstream programming languages \todo{cite}, in this paper, 
-we target a tiny higher-order functional language \TLang, which is based on direct-style lambda-calculus
-with the support of numbers, arithmetics, recursions and conditionals.
-In section \ref{cases_imp}, we will see adding more imperative features to the language.
-Since we are mostly interested in analyzing the dynamic behavior of the program, the static semantics and
-type system will be omitted. We also assume all the input programs are well-typed.
-The abstract syntax is described below:
-
-\begin{lstlisting}
-abstract class Expr
-case class Lit(i: Int)
-case class Var(x: String)
-case class Lam(x: String, e: Expr)
-case class App(e1: Expr, e2: Expr)
-case class AOp(op: String, e1: Expr, e2: Expr)
-case class Rec(x: String, e: Expr, body: Expr)
-case class If0(cnd: Expr, thn: Expr, els: Expr)
-\end{lstlisting}
-
-We give the concrete semantics of \TLang by showing a big-step operational interpreter for it.
-But before doing that, let's first see a general interface for interpreters, whether concrete or abstract, 
-staged or unstaged. The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Ident},
-\texttt{Addr} and etc., which will be instantiated concretely later. The abstract types \texttt{Env} and \texttt{Store}
-are required to conform their specifications, \texttt{EnvSpec} and \texttt{StoreSpec}, which define 
-the operations such as lookup and update on them, nevertheless the detail of them are omitted.  
-The store component is necessary when modeling side effects such as mutations. 
-The trait also requires the implementators provide the initial values of environment and store.
-
-%TODO: talk about alloc here?
-
-\begin{lstlisting}
-trait Semantics {
-  type R[+T]
-  type Ident;    type Addr;   type Value;    type Ans
-  type Env <: EnvSpec[R, Ident, Addr, Env]
-  type Store <: StoreSpec[R, Addr, Value, Store]
-  val $\rho$0: Env;     val $\sigma$0: Store
-  def alloc(x: R[Ident], $\sigma$: Store): R[Addr]
-  type EvalFun = (Expr, Env, Store) => Ans
-  def eval(ev: EvalFun)(e: Expr, $\rho$: Env, $\sigma$: Store): Ans
-  def fix(ev: EvalFun => EvalFun): EvalFun = 
-    (e, $\rho$, $\sigma$) =>  ev(fix(ev))(e, $\rho$, $\sigma$)
-  def eval_top(e: Expr): Ans = fix(eval)(e, $\rho$0, $\sigma$0)
-}
-\end{lstlisting}
-
-The evaluator \texttt{eval} is a open-recursive function that takes one and three arguments: the first
-one \texttt{ev} represents the recursive function itself; the rest three are the AST representing program 
-will be executed, the environment that maps from identifiers to addresses, and the store that maps addresses to values.
-This environment-and-store-passing style big-step interpreter also can be obtained by 
-refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
-machine \cite{DBLP:conf/popl/FelleisenF87}.
-We provide a function \texttt{fix} to close the recursion by providing the result of \texttt{fix(ev)} 
-to argument \texttt{ev}. Finally, a default implementation for the top-level evaluation function \texttt{eval\_top} is defined.
-As we will see later, the open-recursion style is beneficial when implementing abstract interpreter and 
-staged abstract interpreter \todo{cite adi}.
-
-Readers may notice an additional abstract type \texttt{R[+T]} is defined in the trait,
-which means \texttt{R} represents a higher-kinded type and will used to specify the binding-time, 
-i.e., staged or not. Since the LMS framework we are using allows us to specify binding-time purely
-based on types, the unstaged higher-kinded type can be simply instantiated as an identity type constructor: 
-\texttt{type NoRep[+T] = T}. In section \ref{stagedinterp}, \texttt{R} will be instantiated 
-as a staging annotation. \todo{where talk about stage poly?}
-
-Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
-Since the interpreter is concrete, the address space is infinite thus type \texttt{Int} is used for \texttt{Addr}.
-The final returned type by the interpreter \texttt{Ans} is a pair of \texttt{Value} and \texttt{Store}, 
-where the values are either closure or numbers.
-
-\begin{lstlisting}
-trait Concrete extends Semantics {
-  type Ident = String;         type Addr = Int
-  type Ans = (Value, Store);   abstract class Value
-  case class CloV($\lambda$: Lam, $\rho$: Env) extends Value
-  case class NumV(i: Int) extends Value
-}
-\end{lstlisting}
-
-Then we define evaluation function that dispatches differently for different expression.
-As we mentioned before, the type \texttt{R} is concretize as \texttt{NoRep}.
-Note that an allocation function is used when we need to update the store: 
-in the concrete case, \texttt{alloc} always returns a fresh address of the store.
-
-\begin{lstlisting}
-//TODO: maybe not show all the code
-object NoRepConcreteEval extends Concrete {
-  type R[+T] = NoRep[T]
-  def alloc(x: Ident, $\sigma$: Store): Addr = $\sigma$.size + 1
-  def eval(ev: EvalFun)(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = 
-    e match {
-      case Lit(i) => (NumV(i), $\sigma$)
-      case Var(x) => ($\sigma$($\rho$(x)), $\sigma$)
-      case Lam(x, e) => (CloV(Lam(x,e), $\rho$), $\sigma$)
-      case App(e1, e2) =>
-        val (CloV(Lam(x,e), $\lambda\rho$), e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
-        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
-        val $\alpha$ = alloc(x, e2$\sigma$)
-        ev(e, $\lambda\rho$ + (x -> $\alpha$), e2$\sigma$ + ($\alpha$ -> e2v))
-      case AOp(op, e1, e2) =>
-        val (e1v: NumV, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
-        val (e2v: NumV, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
-        (evalArith(op, e1v, e2v), e2$\sigma$)
-      case Rec(x, f, body) => 
-        val $\alpha$ = alloc(x, $\sigma$);  val $\rho$_* = $\rho$ + (x -> $\sigma$)
-        val (fv, f$\sigma$) = ev(f, $\rho$_*, $\sigma$)
-        val $\sigma$_* = f$\sigma$ + ($\alpha$ -> fv);  ev(body, $\rho$_*, $\sigma$_*)
-      case If0(cnd, thn, els) =>
-        val (cndv: NumV, cnd$\sigma$) = ev(cnd, $\rho$, $\sigma$)
-        if (cndv.i == 0) ev(thn, $\rho$, cnd$\sigma$)
-        else ev(els, $\rho$, cnd$\sigma$)
-    }
-}
-\end{lstlisting}
-
-\subsection{From Interpreter to Abstract Interpreter}
-
-\begin{lstlisting}
-trait Abstract extends Semantics {
-  type Ident = String;         case class Addr(x: Ident)
-  type Ans = (Value, Store);   type Value = Set[AbsValue]
-  abstract class AbsValue
-  case class CloV($\lambda$: Lam, $\rho$: Env) extends AbsValue
-  case class NumV() extends AbsValue
-}
-\end{lstlisting}
-
-Address just record the variable name, more sensitivity in section ...
-
-other big-step abstract interpreter, 
-big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
-monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
-arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
-
-co-inductive caching/truncated depth-first evaluation \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800, Rosendahl:AbsIntPL}
-
-\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}
-
-\subsubsection{Multi-Stage Programming}
-\begin{lstlisting}
-val power5 = new DslDriver[Int, Int] {
-  def power(b: Rep[Int], x: Int): Rep[Int] =
-    if (x == 0) 1 else b * power(b, x-1)
-  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
-}
-\end{lstlisting}
-
-Talk about power example. general idea; LMS framework in scala. MetaOCaml?
-
-Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
-and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
-Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
-escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
-still applies with other MSP implementations.
-
-\subsubsection{Staged Interpreters}
+\input{21_bg_interp}
+\input{23_bg_stage}
+\input{22_bg_abs}

### 2018-11-10 paper/21_bg_interp.tex
add some stuff

\subsection{An Unstaged Concrete Interpreter} \label{bg_lang}

As higher-order functions had been widely introduced in mainstream programming languages (such as Python and Java \todo{cite}), 
in this paper, we consider a tiny higher-order functional language \TLang, which is based on direct-style $\lambda$-calculus
with numbers, arithmetics, recursions and conditionals.
In Section~\ref{cases_imp}, we will see adding more imperative features to the language.
Since we are mostly interested in analyzing the dynamic behavior of the program, the static semantics and
type system will be omitted. We also assume the input programs are well-typed and variables are distinct.
The abstract syntax is described below:

\begin{lstlisting}
abstract class Expr
case class Lit(i: Int)
case class Var(x: String)
case class Lam(x: String, e: Expr)
case class App(e1: Expr, e2: Expr)
case class AOp(op: String, e1: Expr, e2: Expr)
case class Rec(x: String, e: Expr, body: Expr)
case class If0(cnd: Expr, thn: Expr, els: Expr)
\end{lstlisting}

We give the concrete semantics of \TLang by showing a big-step operational-style interpreter for it.
The interpreter is a recursive function that takes the program AST, environment and store, and returns 
the evaluated value and the accompanied store. The environment will be a mapping from identifiers to addresses,
and the store is a mapping from addresses to values. The store simulates the heap and is useful when modeling 
recursion and mutation.
This environment-and-store-passing style big-step interpreter also can be obtained by 
refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
machine \cite{DBLP:conf/popl/FelleisenF87}.

\textbf{Generic Interface.} Before showing the actual implementation of the concrete semantics, let's first see 
a generic interface for the interpreters, which will be used later for both unstaged/staged and concrete/abstract scenarios.
The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Addr},
\texttt{Value}, \texttt{Env}, \texttt{Store} and returned type \texttt{Ans}. 
Initial values for environments and stores, and few abstract methods that manipulate environments and stores are also declared.

\begin{lstlisting}
trait Semantics {
  type R[+_]
  type Ident = String;  type Addr;  type Value;  type Env
  type Store;  type Ans = (R[Value], R[Store])
  def get($\rho$: R[Env], x: Ident): R[Addr]
  def put($\rho$: R[Env], x: Ident, a: R[Addr]): R[Env]
  def get($\sigma$: R[Store], a: R[Addr]): R[Value]
  def put($\sigma$: R[Store], a: R[Addr], v: R[Value]): R[Store]
  def alloc($\sigma$: R[Store], x: Ident): R[Addr]
  val $\rho$0: R[Env]; val $\sigma$0: R[Store]
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: R[Env]): R[Value]
  def num(i: Lit): R[Value]
  def apply_closure(ev: EvalFun)
    (f: R[Value], arg: R[Value], $\sigma$: R[Store]): Ans
  def branch0(cnd: R[Value], thn: => Ans, els: => Ans): Ans
  def prim_eval(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
}  // to be continued
\end{lstlisting}

The actual semantics of the interpreter is also abstract now, such as lifting literal terms to values 
(\texttt{close}, \texttt{num}), applying a closure (\texttt{apply\_closure}), branching (\texttt{branch0},
and arithmetics (\texttt{prim\_eval}). But, just based on these abstract methods, a general interpreter 
can be defined:

\begin{lstlisting}
trait Semantics {  // continued
  type EvalFun = (Expr, R[Env], R[Store]) => Ans
  def eval(ev: EvalFun)(e: Expr, $\rho$: R[Env], $\sigma$: R[Store]): Ans = 
    e match {
      case Lit(i) => (num(Lit(i)), $\sigma$)
      case Var(x) => (get($\sigma$, get($\rho$, x)), $\sigma$)
      case Lam(x, e) => (close(ev)(Lam(x, e), $\rho$), $\sigma$)
      case App(e1, e2) =>
        val (e1v, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
        apply_closure(ev)(e1v, e2v, e2$\sigma$)
      case Rec(x, f, body) =>
        val $\alpha$ = alloc($\sigma$, x)
        val $\rho$_* = put($\rho$, x, $\alpha$)
        val (fv, f$\sigma$) = ev(f, $\rho$_*, $\sigma$)
        val $\sigma$_* = put(f$\sigma$, $\alpha$, fv)
        ev(body, $\rho$_*, $\sigma$_*)
      case If0(cnd, thn, els) =>
        val (cndv, cnd$\sigma$) = ev(cnd, $\rho$, $\sigma$)
        branch0(cndv, ev(thn, $\rho$, cnd$\sigma$), ev(els, $\rho$, cnd$\sigma$))
      case AOp(op, e1, e2) =>
        val (e1v, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
        (prim_eval(op, e1v, e2v), e2$\sigma$)
    }
  def fix(ev: EvalFun => EvalFun): EvalFun = 
    (e, $\rho$, $\sigma$) => ev(fix(ev))(e, $\rho$, $\sigma$)
  def eval_top(e: Expr): Ans = eval_top(e, $\rho$0, $\sigma$0)
  def eval_top(e: Expr, $\rho$: R[Env], $\sigma$: R[Store]): Ans = 
    fix(eval)(e, $\rho$, $\sigma$)
}
\end{lstlisting}

The implementation of the interpreter is straightforward.
It is worth to note that the interpreter is written in open-recursive style -- 
it can not refer to itself directly, instead, \texttt{eval} takes an additional 
parameter \texttt{ev} referring to itself; this allows us to instrument the interpreter
from outside. Accordingly, the function \texttt{close} for lifting lambda term to closure values
is also written in this style.
We define a default implementation \texttt{fix} to close the recursion by recursively use the result 
of \texttt{fix(ev)} for \texttt{ev}. 
Finally, the top-level evaluation function \texttt{eval\_top} is defined.

\textbf{Stage Polymorphism.}
Readers may notice an abstract higher-kinded type \texttt{R[+T]} is defined in the trait, that is a
type-level function and will be used to specify the binding-time, i.e., staged or not; the annotation 
\texttt{+} indicates type parameter \texttt{T} is covariant.
The LMS framework we use here allows us to specify binding-time purely based on types.
All occurrences of type \texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} 
essentially because they may potentially being dynamic when staging is introduced.
Obviously, the unstaged higher-kinded type then can be simply instantiated as an identity type function: 
\texttt{type R[+T] = T}. In Section~\ref{stagedinterp}, \texttt{R} will be instantiated 
as a staging annotation. \todo{where talk about stage poly?}

\textbf{Concrete Instantiation.}
Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
Since the interpreter is concrete, for the propose that properly models the heap, the type \texttt{Int} 
is used for addresses space \texttt{Addr}. The environment and store are just ordinary \texttt{Map} in Scala.
The values can be either a tagged number \texttt{NumV} or a closure \texttt{CloV} that contains a lambda term
and an environment.

\begin{lstlisting}
trait Concrete extends Semantics {
  type Addr = Int;   sealed trait Value
  case class CloV($\lambda$: Lam, $\rho$: Env) extends Value
  case class NumV(i: Int) extends Value
  type Env = Map[Ident, Addr];  type Store = Map[Addr, Value]
}
\end{lstlisting}

The we implement the components that define the semantics. 
As we mentioned before, the type \texttt{R} is concretize as an identity type function, thus
all types are just ordinary types.

\begin{lstlisting}
object ConcInterp extends Concrete {
  type R[+T] = T
  val $\rho$0 = Map[Ident,Addr]();  val $\sigma$0 = Map[Addr,Value]()
  def get($\rho$: Env, x: Ident): Addr = $\rho$(x)
  def put($\rho$: Env, x: Ident, a: Addr): Env = $\rho$ + (x -> a)
  def get($\sigma$: Store, a: Addr): Value = $\sigma$(a)
  def put($\sigma$: Store, a: Addr, v: Value): Store = $\sigma$ + (a -> v)
  def alloc($\sigma$: Store, x: Ident): Addr = $\sigma$.size + 1
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Env): Value = CloV($\lambda$, $\rho$)
  def num(i: Lit): Value = NumV(i.i)
  def apply_closure(ev: EvalFun)
    (f: Value, arg: Value, $\sigma$: Store): Ans = f match {
      case CloV(Lam(x, e), $\rho$) =>
        val $\alpha$ = alloc($\sigma$, x); val $\rho$_* = put($\rho$, x, $\alpha$)
        val $\sigma$_* = put($\sigma$, $\alpha$, arg); ev(e, $\rho$_*, $\sigma$_*)
    }
  def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = 
    cnd match { case NumV(i) => if (i == 0) thn else els }
  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = 
    (op, v1, v2) match {
      case ('+, NumV(a), NumV(b)) => NumV(a + b)
      ...
    }
}
\end{lstlisting}

\texttt{get} and \texttt{put} directly manipulate the map of environment and store. 
An allocation function is used when we need to update the store: 
in the concrete case, \texttt{alloc} always returns a fresh address of the store.
\texttt{close} and \texttt{num} lift syntactic literals to values.
\texttt{apply\_closure} takes the function value, argument value and the latest store,
then extends the environment and store after allocating a fresh address, and continue
evaluates the body \texttt{e} of the lambda term.

### 2018-11-10 paper/22_bg_abs.tex
add some stuff

\subsection{From Interpreter to Abstract Interpreter}

Based on the generic interface we presented, now we will describe how to instantiate an abstract interpreter.
To keep the simplicity, we intened to use simply abstract domain, and just establish a context-insensitive, 
path-insensitive, and flow-insensitive analysis in this section -- it is coarse but enough to lay a foundation for 
the staged abstract interpreter we will present later.
In Section \note{X}, we will see how to regain 
context-/path-/flow-sensitivity for control-flow analysis; in Section \note{Y}, we will see how to 
instantiate an interval abstract domain for numerical analysis.

Following the idea of AAM, store widened

\textbf{Abstract Instantiation.}
We use powerset as our abstract domain, the concrete closures directly becomes elements in the powerset, numbers
are abstracted to a single abstract number element \texttt{NumV}

\begin{lstlisting}
trait Abstract extends Semantics {
  case class Addr(x: Ident); sealed trait AbsValue
  case class CloV($\lambda$: Lam, $\rho$: Env) extends AbsValue
  case class NumV() extends AbsValue
  type Value = Set[AbsValue]
  type Env = Map[Ident, Addr]; type Store = Map[Addr, Value]
}
\end{lstlisting}

\textbf{Stage Polymorphic Lattices.}
We use type class in Scala, also works as stage polymorphic.
\begin{lstlisting}
trait Lattice[E, R[_]] {
  val bot: R[E];  val top: R[E]
  def $\sqsubseteq$(l1: R[E], l2: R[E]): R[Boolean]
  def $\sqcup$(l1: R[E], l2: R[E]): R[E]
  def $\sqcap$(l1: R[E], l2: R[E]): R[E]
}
\end{lstlisting}

For example, the abstract domain we use here, power set can be implemented as:

\begin{lstlisting}
implicit def SetLattice[T]: Lattice[Set[T], NoRep] = 
  new Lattice[Set[T], NoRep] {
    lazy val bot: Set[T] = Set[T]()
    lazy val top: Set[T] = throw new NotImplementedError()
    def $\sqsubseteq$(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf l2
    def $\sqcup$(l1: Set[T], l2: Set[T]): Set[T] = l1 union l2
    def $\sqcap$(l1: Set[T], l2: Set[T]): Set[T] = l1 intersect l2
  }
\end{lstlisting}

tuples, maps can be lifted to lattice element-wise or point-wise, the code for them are elided.

\todo{conditional, should use different store}

\begin{lstlisting}
object AbsInterp extends Abstract {
  type R[+T] = T
  val $\rho$0 = Map[Ident, Addr]();  val $\sigma$0 = Map[Addr, Value]()
  def get($\rho$: Env, x: Ident): Addr = $\rho$(x)
  def put($\rho$: Env, x: Ident, a: Addr): Env = $\rho$ + (x -> a)
  def get($\sigma$: Store, a: Addr): Value = 
    $\sigma$.getOrElse(a, Lattice[Value].bot)
  def put($\sigma$: Store, a: Addr, v: Value): Store =
    $\sigma$ + (a -> (v $\sqcup$ get($\sigma$, a)))
  def alloc($\sigma$: Store, x: Ident): Addr = Addr(x)
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Env): Value = Set(CloV($\lambda$, $\rho$))
  def num(i: Lit): Value = Set(NumV())
  def apply_closure(ev: EvalFun)(f: Value, arg: Value, $\sigma$: Store): Ans = {
    var $\sigma$0 = $\sigma$
    val vs = for (CloV(Lam(x, e), $\rho$) <- f) yield {
      val $\alpha$ = alloc($\sigma$0, x)
      val $\rho$_* = put($\rho$, x, $\alpha$)
      val $\sigma$_* = put($\sigma$0, $\alpha$, arg)
      val (v, v$\sigma$) = ev(e, $\rho$_*, $\sigma$_*)
      $\sigma$0 = v$\sigma$; v
    }
    (vs.reduce(Lattice[Value].$\sqcup$(_,_)), $\sigma$0)
  }
  def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = {
    thn $\sqcup$ els //TODO!
  }
  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = Set(NumV())
} // to be continued
\end{lstlisting}

\textbf{Fixpoint Iteration.}
The interpreter may not terminate
co-inductive caching/truncated depth-first evaluation \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800, Rosendahl:AbsIntPL}

\begin{lstlisting}
object AbsInterp extends Abstract { // continued
  case class CacheFix(evev: EvalFun => EvalFun) {
    var in = Map[Config, Ans](); var out = Map[Config, Ans]()
    def cached_ev(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = {
      val cfg: Config = (e, $\rho$, $\sigma$)
      if (out.contains(cfg)) out(cfg)
      else {
        val ans0: Ans = in.getOrElse(cfg, Lattice[(Value, Store)].bot)
        out = out + (cfg -> ans0)
        val ans1: Ans = evev(cached_ev)(e, $\rho$, $\sigma$)
        out = out + (cfg -> Lattice[(Value, Store)].$\sqcup$(ans0, ans1))
        ans1
      }
    }
    def iter(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = {
      in = out; out = Map[Config, Ans](); cached_ev(e, $\rho$, $\sigma$)
      if (in == out) out((e, $\rho$, $\sigma$)) else iter(e, $\rho$, $\sigma$)
    }
  }
  override def eval_top(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = 
    CacheFix(eval).iter(e, $\rho$, $\sigma$)
}
\end{lstlisting}

other big-step abstract interpreter, 
big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.

### 2018-11-10 paper/23_bg_stage.tex
add some stuff

\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}

\textbf{Multi-Stage Programming in LMS.}

\begin{lstlisting}
val power5 = new DslDriver[Int, Int] {
  def power(b: Rep[Int], x: Int): Rep[Int] =
    if (x == 0) 1 else b * power(b, x-1)
  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
}
\end{lstlisting}

Lightweight modular staging (LMS) \todo{cite} is a library-based multi-stage programming framework 
that enables type-safe dynamic code generation. Different from the traditional approach of using
syntactic quasi-quotation, it solely relies on types to distinguish between binding times.

In the example shown above, \texttt{power} is a staged function that takes 2 arguments where 
\texttt{b} is declared as \texttt{Rep[Int]} meaning that \texttt{b} is a representation of type 
\texttt{Int} whose underlying value will be available at the next stage. Meanwhile the actual 
computation result of the \texttt{power} function  will be available at the next stage as well. 
Thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. The generated code for
\texttt{power5} consists of a function that takes only one argument that corresponds to its 
representation \texttt{b} from the previous stage.

\textbf{Staged Concrete Interpreter.}

We share the same concrete type instantiation, i.e. trait \texttt{Concrete}, with the unstaged version,
but reimplemented the staged version for the concrete operations.

\begin{lstlisting}
trait RepConcInterpOps extends Concrete with LMSOps {
  type R[+T] = Rep[T]
  implicit def valueTyp: Typ[Value]
  val $\rho$0: Rep[Env] = Map[Ident,Addr]()
  val $\sigma$0: Rep[Store] = Map[Addr,Value]()
  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): Rep[Env] = 
    $\rho$ + (unit(x) -> a)
  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = $\sigma$(a)
  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): Rep[Store] = 
    $\sigma$ + (a -> v)
  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
    val Lam(x, e) = $\lambda$
    val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = 
      (as: Rep[(Value, Store)]) => {
        val arg = as._1; val $\sigma$ = as._2
        val $\alpha$ = alloc($\sigma$, x)
        ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, arg))
      }
    unchecked("CompiledClo(", fun(f), ",", $\lambda$, ",", $\rho$, ")")
  }
  def num(i: Lit): Rep[Value] = {
    unchecked("NumV(", i.i, ")")
  }
  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
    val i = unchecked[Int](cnd, ".asInstanceOf[NumV].i")
    (if (i == 0) thn else els).asInstanceOf[Rep[(Value,Store)]] //FIXME: Why?
  }
  def prim_eval(op: Symbol, v1: Rep[Value], v2: Rep[Value]): Rep[Value] = {
    val v1i = unchecked(v1, ".asInstanceOf[NumV].i")
    val v2i = unchecked(v2, ".asInstanceOf[NumV].i")
    unchecked("NumV(", v1i, op.toString.drop(1), v2i, ")")
  }
}
\end{lstlisting}

\textbf{Code Generation.}

code generation, runtime code

### 2018-11-10 paper/70_related.tex
fix paper
add some stuff

@@
 
-Abstracting Definitional Interpreter \cite{darais2017abstracting}
+Abstracting Definitional Interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}.
 
@@
 Multi-stage programming. MetaML, MetaOCaml, Mint, LMS
+
+Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
+and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
+Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
+escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
+still applies with other MSP implementations.

### 2018-11-11 paper/22_bg_abs.tex
work on paper

@@
 
-Based on the generic interface we presented, now we will describe how to instantiate an abstract interpreter.
-To keep the simplicity, we intened to use simply abstract domain, and just establish a context-insensitive, 
-path-insensitive, and flow-insensitive analysis in this section -- it is coarse but enough to lay a foundation for 
-the staged abstract interpreter we will present later.
-In Section \note{X}, we will see how to regain 
-context-/path-/flow-sensitivity for control-flow analysis; in Section \note{Y}, we will see how to 
-instantiate an interval abstract domain for numerical analysis.
-
-Following the idea of AAM, store widened
+Based on the generic interface we presented in Section~\ref{bg_lang}, now we describe the instantiation of an
+abstract interpreter.
+To keep the simplicity, we intend to use simply abstract domains, and just establish a 
+context-/path-/flow-insensitive, store-widened analysis in this section --- indeed, it is coarse 
+but enough to setup a foundation for the staged abstract interpreter we will present later.
+However, in Section \note{X}, we will see how to regain context-/path-/flow-sensitivity 
+for pushdown control-flow analysis; in Section \note{Y}, we will see how to instantiate 
+an interval abstract domain for numerical analysis.
 
 \textbf{Abstract Instantiation.}
-We use powerset as our abstract domain, the concrete closures directly becomes elements in the powerset, numbers
-are abstracted to a single abstract number element \texttt{NumV}
+Our instantiation roughly follows the Abstracting Abstract Machines methodology 
+\cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12} that transforms the concrete semantics 
+to abstract semantics. The store now maps addresses to sets of abstract values, meaning that an address points
+to all possible values may occur at runtime, where the abstract value is either a closure or 
+a single abstract number element \texttt{NumV}. The address space is also
+constrained to be finite, in the monovariant setting, we simply use variable names for it.
 
@@
 \textbf{Stage Polymorphic Lattices.}
-We use type class in Scala, also works as stage polymorphic.
+To effectively reuse the code between unstaged and staged part, we also make the lattice structure to
+be stage polymorphic. A lattice type class parameterizes over an element type \texttt{E} and defines
+five operations for \texttt{R[E]}, such as meet, join, and ordering relation; again, \texttt{R} 
+is the higher-kinded type indicating the binding-time.
 \begin{lstlisting}
-trait Lattice[E, R[_]] {
+trait Lattice[E, R[+_]] {
   val bot: R[E];  val top: R[E]
@@
 
-For example, the abstract domain we use here, power set can be implemented as:
+For example, here we use powerset as the abstract domain, and the unstaged lattice 
+of powerset can be easily implemented as follows:
 
@@
 
-tuples, maps can be lifted to lattice element-wise or point-wise, the code for them are elided.
+The ordering relation is to ask whether one set is a subset of the other, 
+join is to union the two set, and meet is to intersect the two sets.
+Accordingly, other structures used in the rest of the paper such as tuples 
+(products) and maps can be lifted to lattices element-wise or point-wise.
+The code for them are elided.
+
+\textbf{Operations Implementation.}
+
+The @get@ and @put@ operation on abstract stores slightly changed according to
+the abstract semantics: @get@ uses bottom element in the @Value@ lattice if the queried
+address not exists; @put@ performs the update by joining with the existing values.
+@alloc@ simply invokes the @Addr@ constructor.
+\texttt{close} and \texttt{num} lift syntactic literals to a singleton set that
+contains one \texttt{CloV} or \texttt{NumV} object.
 
 \todo{conditional, should use different store}
-
 \begin{lstlisting}
@@
   def num(i: Lit): Value = Set(NumV())
-  def apply_closure(ev: EvalFun)(f: Value, arg: Value, $\sigma$: Store): Ans = {
-    var $\sigma$0 = $\sigma$
+  def apply_closure(ev: EvalFun)
+    (f: Value, arg: Value, $\sigma$: Store): Ans = { var $\sigma$_* = $\sigma$
       val vs = for (CloV(Lam(x, e), $\rho$) <- f) yield {
-      val $\alpha$ = alloc($\sigma$0, x)
-      val $\rho$_* = put($\rho$, x, $\alpha$)
-      val $\sigma$_* = put($\sigma$0, $\alpha$, arg)
-      val (v, v$\sigma$) = ev(e, $\rho$_*, $\sigma$_*)
-      $\sigma$0 = v$\sigma$; v
+        val $\alpha$ = alloc($\sigma$_*, x)
+        val (v, v$\sigma$) = ev(e, put($\rho$, x, $\alpha$), put($\sigma$_*, $\alpha$, arg))
+        $\sigma$_* = v$\sigma$; v
       }
-    (vs.reduce(Lattice[Value].$\sqcup$(_,_)), $\sigma$0)
+    (vs.reduce(Lattice[Value].$\sqcup$), $\sigma$_*)
   }
-  def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = {
-    thn $\sqcup$ els //TODO!
-  }
+  def branch0(cnd: Value, thn: =>Ans, els: =>Ans): Ans = thn $\sqcup$ els //TODO!
   def prim_eval(op: Symbol, v1: Value, v2: Value): Value = Set(NumV())

### 2018-11-11 paper/10_intro.tex
section 2

@@
 and then specializes the program with respect to the given input of program.
-However, precisely analyzing binding-time for an arbitray program is hard in general. As an alternative and more practical 
+However, precisely analyzing binding-time for an arbitray program is hard in general. As an alternative and programmatic
 approach to specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 

### 2018-11-11 paper/21_bg_interp.tex
section 2

@@
 
-\textbf{Generic Interface.} Before showing the actual implementation of the concrete semantics, let's first see 
+\paragraph{Generic Interface.} 
+Before showing the actual implementation of the concrete semantics, let's first see 
 a generic interface for the interpreters, which will be used later for both unstaged/staged and concrete/abstract scenarios.
@@
 
-\textbf{Stage Polymorphism.}
+\paragraph{Stage Polymorphism.}
 Readers may notice an abstract higher-kinded type \texttt{R[+T]} is defined in the trait, that is a
@@
 
-\textbf{Concrete Instantiation.}
+\paragraph{Concrete Instantiation.}
 Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
@@
   def prim_eval(op: Symbol, v1: Value, v2: Value): Value = 
-    (op, v1, v2) match {
-      case ('+, NumV(a), NumV(b)) => NumV(a + b)
+    op match {
+      case '+ => v1.asInstanceOf[NumV].i + v2.asInstanceOf[NumV].i
       ...

### 2018-11-11 paper/22_bg_abs.tex
section 2

@@
 
-\textbf{Abstract Instantiation.}
+\paragraph{Abstract Instantiation.}
 Our instantiation roughly follows the Abstracting Abstract Machines methodology 
@@
 to all possible values may occur at runtime, where the abstract value is either a closure or 
-a single abstract number element \texttt{NumV}. The address space is also
+a single abstract number \texttt{NumV}, which stands for the top. The address space is also
 constrained to be finite, in the monovariant setting, we simply use variable names for it.
@@
 
-\textbf{Stage Polymorphic Lattices.}
+\paragraph{Stage Polymorphic Lattices.}
 To effectively reuse the code between unstaged and staged part, we also make the lattice structure to
@@
 
-\textbf{Operations Implementation.}
-
+\paragraph{Operations Implementation.}
 The @get@ and @put@ operation on abstract stores slightly changed according to
@@
 
-\textbf{Fixpoint Iteration.}
+\paragraph{Fixpoint Iteration.}
 The interpreter may not terminate

### 2018-11-11 paper/23_bg_stage.tex
section 2

@@
 
-\textbf{Multi-Stage Programming in LMS.}
+\paragraph{Multi-Stage Programming in LMS.}
+Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a multi-stage programming framework implemented 
+as a Scala library that enables dynamic code generation in a type-safe manner.
+Different from the syntactic approach such as MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations, 
+LMS distinguishes binding-time solely based on types.
+LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an expression will be 
+known at next stage. All operations act on a @Rep@ expression will be resulted in the generated code. However,
+we need to provide intermediate representation and code generation support for type @T@ to LMS. 
+Fortunately, the LMS framework already provides
+such support for primitive types and commonly used data structure such as arrays and maps; implementing such support
+for custom class is also straightforward. Let's go back and see how LMS can be used to specialize the power function
+we mentioned in Section 1.
 
@@
 
-Lightweight modular staging (LMS) \todo{cite} is a library-based multi-stage programming framework 
-that enables type-safe dynamic code generation. Different from the traditional approach of using
-syntactic quasi-quotation, it solely relies on types to distinguish between binding times.
+In the code shown above, \texttt{power} takes 2 arguments where \texttt{b} is declared as \texttt{Rep[Int]} type 
+meaning that \texttt{b} is a representation of \texttt{Int} but whose value will be available at the next stage. 
+Meanwhile, the result of the @power@ function will be available at the next stage as will, thus the return value 
+of \texttt{power} is also of type \texttt{Rep[Int]}. 
+Then we specialize @power@ in function @snippet@ by providing @x@ to @5@.
+The generated code for \texttt{power5} is a function that takes only one argument which corresponds to @b@,
+and body of this function multiples @b@ five times.
 
-In the example shown above, \texttt{power} is a staged function that takes 2 arguments where 
-\texttt{b} is declared as \texttt{Rep[Int]} meaning that \texttt{b} is a representation of type 
-\texttt{Int} whose underlying value will be available at the next stage. Meanwhile the actual 
-computation result of the \texttt{power} function  will be available at the next stage as well. 
-Thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. The generated code for
-\texttt{power5} consists of a function that takes only one argument that corresponds to its 
-representation \texttt{b} from the previous stage.
-
-\textbf{Staged Concrete Interpreter.}
-
-We share the same concrete type instantiation, i.e. trait \texttt{Concrete}, with the unstaged version,
-but reimplemented the staged version for the concrete operations.
+\paragraph{Staged Concrete Interpreter.} In the staging part, we share the same concrete type instantiation, 
+i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations.
+Meanwhile, @RepConcInterpOps@ also extends from @LMSOps@ which provides necessary supports for the 
+staged version of primitive types and data structures.
 
@@
   type R[+T] = Rep[T]
-  implicit def valueTyp: Typ[Value]
   val $\rho$0: Rep[Env] = Map[Ident,Addr]()
@@
   def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
+  def num(i: Lit): Rep[Value] = NumV(i)
+  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
+    //FIXME: the cast is ugly, patten match?
+    val i = cnd.asInstanceOf[Rep[NumV]].i
+    if (i == 0) thn else els
+  }
+  def prim_eval(op: Symbol, 
+    v1: Rep[Value], v2: Rep[Value]): Rep[Value] = op match {
+      case '+ => v1.asInstanceOf[Rep[NumV]].i + 
+                 v2.asInstanceOf[Rep[NumV]].i
+      ...
+    }
+  ...
+}
+\end{lstlisting}
+
+The first notable change is the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
+types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. \todo{NumV}
+The interesting thing is how we handle closures and function applications when staging is involved.
+In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, 
+however, what we desire is the a \textit{compiled} closure, i.e., the syntactic term of the body expression should be
+eliminated and specialized.
+The specialization of the interpreter with respect to the body of the lambda term proceed under the
+assumptions that the argument and latest store will be provided later. \note{to denotational style?}
+We will see this is an important observation that enables us to specialize an abstract interpreter modularly.
+In the end of @close@, we create a @CompiledClo@ object that contains a staged function, which both will be
+exhibited in the generated code.
+
+\begin{lstlisting}
 def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
   val Lam(x, e) = $\lambda$
-    val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = 
-      (as: Rep[(Value, Store)]) => {
-        val arg = as._1; val $\sigma$ = as._2
+  val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = { 
+    case (arg: Rep[Value],$\sigma$: Rep[Store]) =>
       val $\alpha$ = alloc($\sigma$, x)
@@
   }
-    unchecked("CompiledClo(", fun(f), ",", $\lambda$, ",", $\rho$, ")")
-  }
-  def num(i: Lit): Rep[Value] = {
-    unchecked("NumV(", i.i, ")")
-  }
-  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
-    val i = unchecked[Int](cnd, ".asInstanceOf[NumV].i")
-    (if (i == 0) thn else els).asInstanceOf[Rep[(Value,Store)]] //FIXME: Why?
-  }
-  def prim_eval(op: Symbol, v1: Rep[Value], v2: Rep[Value]): Rep[Value] = {
-    val v1i = unchecked(v1, ".asInstanceOf[NumV].i")
-    val v2i = unchecked(v2, ".asInstanceOf[NumV].i")
-    unchecked("NumV(", v1i, op.toString.drop(1), v2i, ")")
-  }
+  CompiledClo(fun(f))
 }
@@
 
-\textbf{Code Generation.}
+Then, @apply_closure@ takes three arguments the function value, argument value and the latest store, 
+which are both of @Rep@ type. It means basically these is nothing we can do at current stage --- 
+so we create a new node @ApplyClosure@ that contains these arguments in the intermediate representation graph,
+and in the code generation phase we can emit the code that does the application. \todo{reflectEffect}.
 
-code generation, runtime code
+\begin{lstlisting}
+case class ApplyClosure(f: Rep[Value], 
+  arg: Rep[Value], $\sigma$: Rep[Store]) extends Def[(Value, Store)]
+def apply_closure(ev: EvalFun)
+  (f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]): Ans = {
+    reflectEffect(ApplyClosure(f, arg, $\sigma$))
+  }
+\end{lstlisting}
+
+\paragraph{Code Generation.}
+When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as what it is because
+they are intended to be existed in the next stage; however, we need to treat @ApplyClosure@ a little bit 
+differently.
+
+\begin{lstlisting}
+  case ApplyClosure(f, arg, $\sigma$) => 
+    emitValDef(sym, quote(f) + 
+                    ".asInstanceOf[CompiledClo].f(" + 
+                    quote(arg) + "," + 
+                    quote($\sigma$) + ")")
+\end{lstlisting}
+
+Concretely, as shown in the code above, when an @ApplyClosure@ is matched when emitting code, we just 
+emit a function application where the function value is extracted from the @CompiledClo@
+(we assume the object program is well-typed so that we can safely cast it to @CompiledClo@), 
+and argument and store are the rest fields in @ApplyClosure@.

### 2018-11-11 paper/22_bg_abs.tex
fixpoint iter

@@
 an interval abstract domain for numerical analysis.
+
+\todo{maybe put to related work} other big-step abstract interpreter, 
+big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
+monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
+arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
 
@@
 \paragraph{Fixpoint Iteration.}
-The interpreter may not terminate
-co-inductive caching/truncated depth-first evaluation \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800, Rosendahl:AbsIntPL}
+We described the abstract semantics modularly in the last section, however, the abstract 
+interpreter may not terminate for some input program. In this section, we use a memoization
+technique to ensure its termination. This technique is also called co-inductive caching algorithm
+\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or truncated depth-first evaluation 
+\cite{Rosendahl:AbsIntPL} and widely used in different analysis.
+
+The idea is to set up two caches called @in@ and @out@, which are both mapping from the arguments of interpreter 
+(@Config@) to the returned value of interpreter (@Ans@). The @in@ contains what we already computed from the last
+iteration, the @out@ cache is what we computed after this iteration joined with the previous one.
+When starting a new iteration, the @in@ is set to be the result of last iteration, i.e., @out@; the @out@ is set
+to be empty. In the iteration, we first ask whether @out@ contains what we want, if yes then it is returned;
+otherwise, we retrieve what we have from @in@, compute the result for this time, and put the joined result back to @out@.
+Note that we also instrument the recursive call by putting @cached_ev@ to @evev@.
+After one iteration, if @in == out@, then there is no more information can be gained, thus the iteration should 
+end and we have reached the fixed point.
 
 \begin{lstlisting}
-object AbsInterp extends Abstract { // continued
 case class CacheFix(evev: EvalFun => EvalFun) {
@@
     else {
-        val ans0: Ans = in.getOrElse(cfg, Lattice[(Value, Store)].bot)
+      val ans0 = in.getOrElse(cfg, Lattice[(Value, Store)].bot)
       out = out + (cfg -> ans0)
-        val ans1: Ans = evev(cached_ev)(e, $\rho$, $\sigma$)
-        out = out + (cfg -> Lattice[(Value, Store)].$\sqcup$(ans0, ans1))
-        ans1
+      val ans1 = evev(cached_ev)(e, $\rho$, $\sigma$)
+      out = out + (cfg -> (ans0 $\sqcup$ ans1));  ans1
     }
@@
   CacheFix(eval).iter(e, $\rho$, $\sigma$)
-}
 \end{lstlisting}
 
-other big-step abstract interpreter, 
-big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
-monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
-arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
+Finally, we override the definition of @eval_top@ by instantiating @CacheFix@ with @eval@ and start 
+the first iteration.

### 2018-11-11 paper/23_bg_stage.tex
fixpoint iter

@@
 however, what we desire is the a \textit{compiled} closure, i.e., the syntactic term of the body expression should be
-eliminated and specialized.
+eliminated and specialized away.
 The specialization of the interpreter with respect to the body of the lambda term proceed under the
-assumptions that the argument and latest store will be provided later. \note{to denotational style?}
+assumptions that the argument and latest store will be provided later. \note{to denotational style?, back to HO?}
 We will see this is an important observation that enables us to specialize an abstract interpreter modularly.

### 2018-11-12 paper/10_intro.tex
section 2/3

@@
     (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
-    (m-2-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-2-2)
+    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
     (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
   \end{tikzpicture}
+  \caption{The confluence of staging and abstraction.}
 \end{figure}

### 2018-11-12 paper/21_bg_interp.tex
section 2/3

@@
 
-The actual semantics of the interpreter is also abstract now, such as lifting literal terms to values 
-(\texttt{close}, \texttt{num}), applying a closure (\texttt{apply\_closure}), branching (\texttt{branch0},
-and arithmetics (\texttt{prim\_eval}). But, just based on these abstract methods, a general interpreter 
-can be defined:
+The actual semantics and operations of the interpreter is left to be implemented later, 
+such as lifting literal terms to values (\texttt{close}, \texttt{num}), 
+applying a closure (\texttt{apply\_closure}), branching (\texttt{branch0}),
+and arithmetics (\texttt{prim\_eval}). We can choose to implement them concrete or abstract,
+unstaged or staged.
+But, just based on these abstract operations, an expressive generic interpreter can be defined:
 
@@
 
-The implementation of the interpreter is straightforward.
+The implementation of the skeleton interpreter traverses the abstract syntax tree and should be straightforward.
 It is worth to note that the interpreter is written in open-recursive style -- 

### 2018-11-12 paper/22_bg_abs.tex
section 2/3

@@
-\subsection{From Interpreter to Abstract Interpreter}
+\subsection{From Interpreter to Abstract Interpreter} \label{unstaged_abs}
 
-Based on the generic interface we presented in Section~\ref{bg_lang}, now we describe the instantiation of an
-abstract interpreter.
+Based on the generic interface we presented in Section~\ref{bg_lang}, 
+now we describe the instantiation of an abstract interpreter.
 To keep the simplicity, we intend to use simply abstract domains, and just establish a 
 context-/path-/flow-insensitive, store-widened analysis in this section --- indeed, it is coarse
-but enough to setup a foundation for the staged abstract interpreter we will present later.
-However, in Section \note{X}, we will see how to regain context-/path-/flow-sensitivity 
-for pushdown control-flow analysis; in Section \note{Y}, we will see how to instantiate 
+but simple enough to setup a foundation for the staged abstract interpreter we will present later.
+However, in Section~\ref{cfa}, we will see how to regain context-/path-/flow-sensitivity 
+for pushdown control-flow analysis; in Section~\ref{cases_imp}, we will see how to instantiate 
 an interval abstract domain for numerical analysis.
@@
 
-\paragraph{Operations Implementation.}
-The @get@ and @put@ operation on abstract stores slightly changed according to
-the abstract semantics: @get@ uses bottom element in the @Value@ lattice if the queried
-address not exists; @put@ performs the update by joining with the existing values.
+\paragraph{Abstract Semantics.}
+The operations @get@ and @put@ on abstract stores are slightly changed according to
+the abstract semantics: @get@ uses the bottom element in the @Value@ lattice if the 
+queried address not exists; @put@ performs the update by joining with the existing values.
 @alloc@ simply invokes the @Addr@ constructor.
-\texttt{close} and \texttt{num} lift syntactic literals to a singleton set that
-contains one \texttt{CloV} or \texttt{NumV} object.
+\texttt{close} and \texttt{num} lift syntactic literals to our abstract domain, i.e., 
+a singleton set that contains only one \texttt{CloV} or \texttt{NumV} object.
 
-\todo{conditional, should use different store}
+\todo{conditional, use different store?}
 \begin{lstlisting}
@@
   def apply_closure(ev: EvalFun)
-    (f: Value, arg: Value, $\sigma$: Store): Ans = { var $\sigma$_* = $\sigma$
-      val vs = for (CloV(Lam(x, e), $\rho$) <- f) yield {
+    (fs: Value, arg: Value, $\sigma$: Store): Ans = { var $\sigma$_* = $\sigma$
+      val vs = for (CloV(Lam(x, e), $\rho$) <- fs) yield {
         val $\alpha$ = alloc($\sigma$_*, x)
@@
   }
-  def branch0(cnd: Value, thn: =>Ans, els: =>Ans): Ans = thn $\sqcup$ els //TODO!
-  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = Set(NumV())
+  def branch0(cnd: Value, thn: =>Ans, els: =>Ans): Ans = 
+    thn $\sqcup$ els
+  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = 
+    Set(NumV())
 } // to be continued
 \end{lstlisting}
+
+For branching, @branch0@ simply joins the answer from two branches. For arithmetics, @prim_eval@ simply 
+returns the top abstract number element.
+It is worth to note that @apply_closure@ is where we handle the non-determinism of applications.
+The non-determinism happens because of the abstract store maps addresses to sets of abstract values;
+when dereferencing an address from the store, we need to explore all possible values to achieve a sound 
+analysis. In the case of @apply_closure@, the first argument @fs@ may contains multiple target closures,
+so we have a @for@ comprehension over all closures, apply them respectively and form a set result values @vs@;
+meanwhile, every time when we evaluate into the body expression @e@ we use the latest store $\sigma$@_*@
+which is updated iteratively. Thus it is a store-widened analysis.
 
@@
 interpreter may not terminate for some input program. In this section, we use a memoization
-technique to ensure its termination. This technique is also called co-inductive caching algorithm
+technique to ensure its termination. This technique is also widely used in the name of co-inductive caching algorithm
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or truncated depth-first evaluation 
-\cite{Rosendahl:AbsIntPL} and widely used in different analysis.
+\cite{Rosendahl:AbsIntPL}.
 

### 2018-11-12 paper/23_bg_stage.tex
section 2/3

@@
 The generated code for \texttt{power5} is a function that takes only one argument which corresponds to @b@,
-and body of this function multiples @b@ five times.
+and the body of this function multiples @b@ five times.
 
-\paragraph{Staged Concrete Interpreter.} In the staging part, we share the same concrete type instantiation, 
+\paragraph{Staged Concrete Semantics.} In the staging part, we share the same concrete type instantiation, 
 i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations.

### 2018-11-12 paper/30_sai.tex
continue working on outline
intro
worked on paper
add some stuff
section 2/3

@@
-\section{Staged Abstract Interpreter}
+\section{Staging Abstract Interpreters} \label{sai}
+
+In the previous section, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
+now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
+Unsurprisingly, the staged abstract interpreter has the same abstract semantics with the unstaged version
+we presented in Section~\ref{unstaged_abs}.
 
@@
 
-\subsection{what not to do when writing an abstract interpreter for staging}
+In this section, we present a principled approach to derive staged abstract interpreter
+from its unstaged version. One guiding principle of our approach is the code of abstract semantics
+and the code that optimizes should be separated. Thus it is also an advantage of doing staging 
+for abstract interpreters: the designer of analysis has no need to rewrite the analysis, and 
+the performance improvement comes almost for free, without any sacrifice of soundness or precision.
+
+\paragraph{Staged Lattices.} The lattice definition we introduced in Section~\ref{unstaged_abs}
+leaves space to be stage polymorphic, now we still use powerset as an example to present its 
+staged version.
+
+\begin{lstlisting}
+trait RepLattice[A] extends Lattice[A, Rep]
+implicit def RepSetLattice[T:Typ]: RepLattice[Set[T]] = 
+  new RepLattice[Set[T]] {
+    lazy val bot: Rep[Set[T]] = Set[T]()
+    lazy val top: Rep[Set[T]] = throw new NotImplementedError()
+    def $\sqsubseteq$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
+      Rep[Boolean] = l1 subsetOf l2
+    def $\sqcup$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
+      Rep[Set[T]] = l1 union l2
+    def $\sqcap$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
+      Rep[Set[T]] = l1 intersect l2
+  }
+\end{lstlisting}
+
+The type parameter @T:Typ@ of @RepLattice@ requires the element of sets
+is also can be staged, otherwise without knowing how to stage the elements in the set
+we can not stage the set as well.
+The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
+such as @union@ and @intersect@ will be mapped to a node in IR graph and emitted in 
+the generated code. Again, other structures such as maps and tuples are also implemented
+in a similar way.
+
+\paragraph{Staged Abstract Semantics.}
+
+\begin{lstlisting}
+trait RepAbsInterpOps extends Abstract with LMSOps {
+  type R[+T] = Rep[T]
+  val $\rho$0: Rep[Env] = Map[Ident, Addr]()
+  val $\sigma$0: Rep[Store] = Map[Addr, Value]()
+  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
+  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): 
+    Rep[Env] = $\rho$ + (unit(x) -> a)
+  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = 
+    $\sigma$.getOrElse(a, RepLattice[Value].bot)
+  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): 
+    Rep[Store] = $\sigma$ + (a -> RepLattice[Value].$\sqcup$(v, get($\sigma$, a)))
+  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = Addr(x)
+  def num(i: Lit): Rep[Value] = Set(NumV())
+  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
+    val Lam(x, e) = $\lambda$
+    val f: Rep[(Value, Store)]=>Rep[(Value,Store)] = 
+      (as: Rep[(Value,Store)]) => {
+        val args = as._1; val $\sigma$ = as._2; val $\alpha$ = alloc($\sigma$, x)
+        ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, args))
+      }
+    Set[AbsValue](CompiledClo(fun(f)))
+  }
+  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans =
+    thn $\sqcup$ els
+  def prim_eval(op: Symbol, 
+                v1: Rep[Value], v2: Rep[Value]): Rep[Value] = 
+    Set(NumV())
+  ...
+}
+\end{lstlisting}
+
+\paragraph{Staged Fixpoint Iteration.} staged fixed point iteration with memoization
+
+\paragraph{Specialized Data Structures.}
+
+\subsection{Discussion}
+what not to do when writing an abstract interpreter for staging?
+what kind of abstract interpreter is friendly for staging/specialization?
+
+Stage an abstract abstract machine yields another small-step abstract machine
 
@@
 
-operational vs denotational
+direct-style vs CPS
+
+Operational vs denotational
 
@@
 
-\subsection{Stage Annotations}
+\cite{10.1007/3-540-61580-6_11}
 
@@
 Futamura projection?
+
+\subsection{Lattice-theoretic Interpretation}

### 2018-11-12 paper/21_bg_interp.tex
paper

@@
 Before showing the actual implementation of the concrete semantics, let's first see 
-a generic interface for the interpreters, which will be used later for both unstaged/staged and concrete/abstract scenarios.
+a generic interface for the interpreters, which is semantics-agnostic and can be 
+"interpreted" by either concrete or abstract, unstaged or staged semantics by our choice.
 The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Addr},

### 2018-11-12 paper/23_bg_stage.tex
paper

@@
 Meanwhile, @RepConcInterpOps@ also extends from @LMSOps@ which provides necessary supports for the 
-staged version of primitive types and data structures.
+staged version of primitive types and data structures. \todo{what are static, what are dyanmic?}
 
@@
   def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
-  def num(i: Lit): Rep[Value] = NumV(i)
+  def num(i: Lit): Rep[Value] = lift(NumV(i))
   def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
@@
 The first notable change is the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
-types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. \todo{NumV}
+types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. For @num@, we use @lift@ function which is a
+built-in in LMS that lifts a current-stage value to next stage.
 The interesting thing is how we handle closures and function applications when staging is involved.
-In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, 
-however, what we desire is the a \textit{compiled} closure, i.e., the syntactic term of the body expression should be
+In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, which
+is a pair of the syntactic lambda term and the enclosing environment.
+However, what we desire is the a \textit{compiled} closure, instead of a lifted @CloV@ value like 
+what we did for @NumV@ --- in other words, the syntactic term of the lambda expression should be 
 eliminated and specialized away.

### 2018-11-12 paper/30_sai.tex
paper

@@
 
-\paragraph{Staged Lattices.} The lattice definition we introduced in Section~\ref{unstaged_abs}
-leaves space to be stage polymorphic, now we still use powerset as an example to present its 
-staged version.
+\paragraph{Staged Lattices.} In Section~\ref{unstaged_abs}, we exploited the higher-kinded type @R@
+to leave space for staging lattices, now we instantiate @R@ to @Rep@ and still use powerset as an 
+example to present its staged version.
 
@@
 
-\paragraph{Staged Abstract Semantics.}
+\paragraph{Staged Abstract Semantics.} We have seen how to obtain a staged concrete semantics
+based on types, now we take the same approach to obtain a staged abstract semantics.
+\todo{what are static, what are dynamic?}
+The basic operations largely kept same as the unstaged version, except the types are changed to @Rep@.
+Besides, when we update the environment, the identifier @x@ is known statically, but the map of environment
+has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to make it as a next-stage value.
 
@@
   def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): 
-    Rep[Env] = $\rho$ + (unit(x) -> a)
+    Rep[Env] = $\rho$ + (lift(x) -> a)
   def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = 
@@
   def num(i: Lit): Rep[Value] = Set(NumV())
-  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
-    val Lam(x, e) = $\lambda$
-    val f: Rep[(Value, Store)]=>Rep[(Value,Store)] = 
-      (as: Rep[(Value,Store)]) => {
-        val args = as._1; val $\sigma$ = as._2; val $\alpha$ = alloc($\sigma$, x)
-        ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, args))
-      }
-    Set[AbsValue](CompiledClo(fun(f)))
-  }
   def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans =
@@
 
-\paragraph{Staged Fixpoint Iteration.} staged fixed point iteration with memoization
+Once more, the way we handle closures is as same as the staged concrete interpreter:
+the recursive call to @ev@ with the body expression @e@ is compiled and specialized, 
+the wrapper function @f@ and will be a field value of @CompiledClo@ and be generated
+for the next stage. At the end, we return a singleton set.
 
-\paragraph{Specialized Data Structures.}
+\begin{lstlisting}
+def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
+  val Lam(x, e) = $\lambda$
+  val f: Rep[(Value, Store)]=>Rep[(Value,Store)] = {
+    case (args: Rep[Value], $\sigma$: Rep[Store]) =>
+      val args = as._1; val $\sigma$ = as._2; val $\alpha$ = alloc($\sigma$, x)
+      ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, args))
+    }
+  Set[AbsValue](CompiledClo(fun(f)))
+}
+def apply_closure(ev: EvalFun)
+  (f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]): Ans = {
+    reflectEffect(ApplyClosure(f, arg, $\sigma$))
+  }
+\end{lstlisting}
+
+When generating code for application, we can not directly apply the callee.
+Instead, we emit code that calls a next-stage function @apply_closures_norep@ 
+which non-deterministically applies multiple target closure with the argument
+and latest store, and finally returns the joined value and a single store.
+
+\begin{lstlisting}
+case ApplyClosures(fs, arg, $\sigma$) =>
+  emitValDef(sym, "apply_closures_norep(" + 
+                  quote(fs) + "," + quote(arg) + 
+                  "," + quote($\sigma$) + ")")
+\end{lstlisting}
+
+\paragraph{Staged Fixpoint Iteration.} 
+\todo{staged fixed point iteration with memoization}
+
+\subsection{Specialized Data Structures}
+
+Now we already obtained a working version of staged abstract interpter that is able to 
+specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
+which means any operations on a @Map@ becomes code in the next stage.
+But, as we identified before \todo{where}, the keys of environment are identifiers in the program,
+which are completely known statically. This leaves us a chance to further specialize the 
+data structures. Particularly, if we are specializing a monovariant analysis, the address space 
+is equivalent to the identifiers, then the environment component can be entirely eliminated, 
+and the store is a partially specialized map.
+\todo{maps to arrays}
 

### 2018-11-12 paper/50_cases.tex
section 2/3
paper

@@
 \section{Case Study} \label{cases_study}
+
+In Section~\ref{sai}, we have shown staging an abstract interpreter is feasible and can be 
+systematic, despite the fact that the abstract interpreter is intended to be imprecise and easy to implement. 
+In this section, we conduct several case studies to show that it is also useful and widely
+applicable to different analysis.
 
@@
 
-\subsection{Control-flow Analysis}
+\subsection{Control-flow Analysis} \label{cfa}
 
@@
 
+\subsubsection{Regain Precise Store}
+
 \subsubsection{Context/Path/Flow-Sensitivity}
@@
 using different $k$ for $k$-CFA
-
-\subsubsection{Store-widening}
 

### 2018-11-12 code/src/main/scala/sai/direct/core/ai/LC.scala

@@ -114,8 +114,10 @@ object LamCal {
     def alloc(σ: Rep[Store], x: Ident): Rep[Addr] = σ.size + 1
     def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
       val Lam(x, e) = λ
+      // TODO: runtime error when using pattern matching
+      // scala.MatchError: UnboxedTuple(List(Sym(7), Sym(8))) (of class scala.lms.common.TupledFunctionsExp$UnboxedTuple)
       //val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = {
-      //case (arg: Rep[Value], σ: Rep[Store]) =>
+      //  case (arg: Rep[Value], σ: Rep[Store]) =>
       val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = (as: Rep[(Value, Store)]) => {
         val arg = as._1; val σ = as._2;
         val α = alloc(σ, x)

### 2018-11-12 paper/70_related.tex
related work

@@
 
-Abstract interpretation \cite{DBLP:conf/popl/CousotC77}.
-
-Abstracting Definitional Interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
+As semantic artifacts, the Abstracting Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
+approach shows that abstract interpreters can be derived systematically from concrete semantic artifacts.
+The big-step abstract interpreters we presented in this paper are also inspired by the AAM framework,
+which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
 
@@
 
-\textbf{Modular Static Analysis}.
-
-Cousot \& Cousot \cite{DBLP:conf/cc/CousotC02}, 
-Compositional shape analysis \cite{DBLP:conf/popl/CalcagnoDOY09}
+\citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
+\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
 

### 2018-11-13 paper/10_intro.tex
fix typos and restructure

@@
 approach to specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 
-requires the stage annotations (i.e., binding-time annotations) from the programmers, these staging annotations 
+requires the stage annotations (i.e., binding-time annotations) from the programmers. These staging annotations 
 control which part of the input program should be specialized.
@@
 \end{lstlisting}
-If \texttt{e} is statically known, say 5, then we can specialize \texttt{power} with respect to \texttt{e = 5} and 
-generate a new specialized function as following:
+If \texttt{e} is identified as statically known by the programmer, say 5, then we can specialize \texttt{power} 
+with respect to \texttt{e = 5} and generate a new specialized function as following:
 \begin{lstlisting}
@@
 etc.
-Likewise, specializing static analyses by partial evaluation had been emerged in late 90s 
+Likewise, specializing static analyses by partial evaluation emerged in late 90s 
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
@@
 repeatedly traversing the abstract syntax tree. 
-However, despite of the previous work are mostly focus on one particular analysis,
+However, despite of the previous works mostly focused on one particular analysis,
+or required to rewrite the analyzer,
 it is worth to investigate the idea from a modern perspective based on meta-programming,
@@
 In this paper, we present an end-to-end design and implementation of staging an abstract interpreter,
-that means ont only the abstract interpreter, but also the data structure of abstract domains, 
-abstract environment and heap, and the fixpoint iteration are staged.
+that means not only the interpreter that traverses the abstract syntax tree, 
+but also the data structure of abstract domains, abstract environment and heap, 
+and the fixpoint iteration are staged.
 
@@
 the free variables representing other parts of the program (e.g., libraries) 
-are dynamic input, therefore we obtain a modular analysis through staging, 
+are dynamic inputs, therefore we obtain a modular analysis through staging, 
 mechanically.
@@
 "Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
-Soundly and precisely analyzing these libraries sometimes are expensive and unnecessary repeatedly.
+Soundly and precisely analyzing these libraries sometimes are expensive and unnecessarily repeated.
 When applying the staged abstract interpreter on libraries, we leave the unknown 
@@
   we show how to turn it into a staged abstract interpreter (SAI) by adding stage annotations. 
-  The design and implementation of SAI using Scala and Lightweight Modular Staging (LMS) framework is 
-  described in Section~\ref{sai}.
+  The design and implementation of SAI using Scala and Lightweight Modular Staging (LMS) framework 
+  are described in Section~\ref{sai}.
   We later extend the staged abstract interpreter to support imperative language features 
@@
 \item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in Section~\ref{cases_study}. 
-  First we revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
+  First, we revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
   technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 

### 2018-11-13 paper/20_bg.tex
fix typos and restructure

@@
 
-\section{Background}
+\section{Preliminary}
 
-In this section, we first described the language we will model and analyze, and present an unstaged concrete 
-interpreter for it. Then, starting from this unstaged concrete interpreter, we proceed in two directions:
-by abstracting it we obtain an unstaged abstract interpreter; by staging it we obtain a staged concrete
-interpreter.
+In this section, we first describe the abstract syntax of the language we will analyze, then present an 
+generic interface of interpreters, and instantiate an unstaged concrete interpreter.
+Then, starting from this unstaged concrete interpreter, we proceed in two directions:
+by abstracting it we obtain an unstaged abstract interpreter (Section~\ref{unstaged_abs}); 
+by staging it we obtain a staged concrete interpreter (Section~\ref{stagedinterp}).
 

### 2018-11-13 paper/21_bg_interp.tex
fix typos and restructure

--- 2018-11-12 paper/21_bg_interp.tex
@@
-\subsection{An Unstaged Concrete Interpreter} \label{bg_lang}
+\subsection{Abstract Syntax} \label{bg_lang}
 
-As higher-order functions had been widely introduced in mainstream programming languages (such as Python and Java \todo{cite}), 
+As higher-order functions had been widely introduced to mainstream programming languages (such as Python and Java \todo{cite}), 
 in this paper, we consider a tiny higher-order functional language \TLang, which is based on direct-style $\lambda$-calculus
@@
 
-\paragraph{Generic Interface.} 
+\subsection{Generic Interface} 
 Before showing the actual implementation of the concrete semantics, let's first see 
-a generic interface for the interpreters, which will be used later for both unstaged/staged and concrete/abstract scenarios.
+a generic interface for the interpreters, which is semantics-agnostic and can be 
+"interpreted" by either concrete or abstract, unstaged or staged semantics by our choice.
 The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Addr},
@@
 The implementation of the skeleton interpreter traverses the abstract syntax tree and should be straightforward.
-It is worth to note that the interpreter is written in open-recursive style -- 
+It is worth noting that the interpreter is written in open-recursive style -- 
 it can not refer to itself directly, instead, \texttt{eval} takes an additional 
@@
 is also written in this style.
-We define a default implementation \texttt{fix} to close the recursion by recursively use the result 
+We define a default implementation \texttt{fix} to close the recursion by recursively using the result 
 of \texttt{fix(ev)} for \texttt{ev}. 
@@
 All occurrences of type \texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} 
-essentially because they may potentially being dynamic when staging is introduced.
+essentially because they may potentially be dynamic when staging is introduced.
 Obviously, the unstaged higher-kinded type then can be simply instantiated as an identity type function: 
@@
 
-\paragraph{Concrete Instantiation.}
+\subsection{Concrete Instantiation}
 Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
 Since the interpreter is concrete, for the propose that properly models the heap, the type \texttt{Int} 
-is used for addresses space \texttt{Addr}. The environment and store are just ordinary \texttt{Map} in Scala.
+is used for address space \texttt{Addr}. The environment and store are just ordinary \texttt{Map} in Scala.
 The values can be either a tagged number \texttt{NumV} or a closure \texttt{CloV} that contains a lambda term
@@
 then extends the environment and store after allocating a fresh address, and continue
-evaluates the body \texttt{e} of the lambda term.
+evaluating the body \texttt{e} of the lambda term.

### 2018-11-13 paper/22_bg_abs.tex
fix typos and restructure

@@
-\subsection{From Interpreter to Abstract Interpreter} \label{unstaged_abs}
+\section{From Interpreters to Abstract Interpreters} \label{unstaged_abs}
 
@@
 
-\paragraph{Stage Polymorphic Lattices.}
+\subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
 To effectively reuse the code between unstaged and staged part, we also make the lattice structure to
@@
 
-\paragraph{Abstract Semantics.}
+\subsection{Abstract Semantics}
 The operations @get@ and @put@ on abstract stores are slightly changed according to
@@
 
-\paragraph{Fixpoint Iteration.}
+\subsection{Fixpoint Iteration}
 We described the abstract semantics modularly in the last section, however, the abstract 
@@
 
-Finally, we override the definition of @eval_top@ by instantiating @CacheFix@ with @eval@ and start 
+Finally, we override the definition of @eval_top@ by instantiating @CacheFix@ with @eval@ and starting 
 the first iteration.

### 2018-11-13 paper/23_bg_stage.tex
fix typos and restructure

@@
-\subsection{From Interpreter to Staged Interpreter} \label{stagedinterp}
+\section{From Interpreters to Staged Interpreters} \label{stagedinterp}
 
-\paragraph{Multi-Stage Programming in LMS.}
+\subsection{Multi-Stage Programming in LMS}
 Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a multi-stage programming framework implemented 
@@
 LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an expression will be 
-known at next stage. All operations act on a @Rep@ expression will be resulted in the generated code. However,
+known at next stage. All operations acting on a @Rep@ expression will be resulted in the generated code. However,
 we need to provide intermediate representation and code generation support for type @T@ to LMS. 
@@
 \begin{lstlisting}
-val power5 = new DslDriver[Int, Int] {
+.. = new DslDriver[Int, Int] {
   def power(b: Rep[Int], x: Int): Rep[Int] =
@@
 meaning that \texttt{b} is a representation of \texttt{Int} but whose value will be available at the next stage. 
-Meanwhile, the result of the @power@ function will be available at the next stage as will, thus the return value 
+Meanwhile, the result of the @power@ function will be available at the next stage as well, thus the return value 
 of \texttt{power} is also of type \texttt{Rep[Int]}. 
@@
 
-\paragraph{Staged Concrete Semantics.} In the staging part, we share the same concrete type instantiation, 
+\subsection{Staged Concrete Semantics} 
+In the staging part, we share the same concrete type instantiation, 
 i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations.
@@
 is a pair of the syntactic lambda term and the enclosing environment.
-However, what we desire is the a \textit{compiled} closure, instead of a lifted @CloV@ value like 
+However, what we desire is a \textit{compiled} closure, instead of a lifted @CloV@ value like 
 what we did for @NumV@ --- in other words, the syntactic term of the lambda expression should be 
@@
 We will see this is an important observation that enables us to specialize an abstract interpreter modularly.
-In the end of @close@, we create a @CompiledClo@ object that contains a staged function, which both will be
+In the end of @close@, we create a @CompiledClo@ object that contains a staged function, which all will be
 exhibited in the generated code.
@@
 Then, @apply_closure@ takes three arguments the function value, argument value and the latest store, 
-which are both of @Rep@ type. It means basically these is nothing we can do at current stage --- 
+which are both of @Rep@ type. It means basically there is nothing we can do at current stage --- 
 so we create a new node @ApplyClosure@ that contains these arguments in the intermediate representation graph,
@@
 \paragraph{Code Generation.}
-When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as what it is because
+When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as what they are because
 they are intended to be existed in the next stage; however, we need to treat @ApplyClosure@ a little bit 

### 2018-11-13 paper/30_sai.tex
fix typos and restructure

@@
-\section{Staging Abstract Interpreters} \label{sai}
+\section{From Abstract Interpreters to Staged Abstract Interpreters} \label{sai}
 
@@
 now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
-Unsurprisingly, the staged abstract interpreter has the same abstract semantics with the unstaged version
-we presented in Section~\ref{unstaged_abs}.
-
-\subsection{From Abstract Interpreter to Staged Abstract Interpreter}
-
-In this section, we present a principled approach to derive staged abstract interpreter
+We present a principled approach to derive staged abstract interpreter
 from its unstaged version. One guiding principle of our approach is the code of abstract semantics
@@
 the performance improvement comes almost for free, without any sacrifice of soundness or precision.
+Unsurprisingly, the staged abstract interpreter we present in this section has the same abstract 
+semantics with the unstaged version we presented in Section~\ref{unstaged_abs}.
 
-\paragraph{Staged Lattices.} In Section~\ref{unstaged_abs}, we exploited the higher-kinded type @R@
+\subsection{Staged Lattices} 
+In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@
 to leave space for staging lattices, now we instantiate @R@ to @Rep@ and still use powerset as an 
@@
 
-\paragraph{Staged Abstract Semantics.} We have seen how to obtain a staged concrete semantics
+\subsection{Staged Abstract Semantics} 
+We have seen how to obtain a staged concrete semantics
 based on types, now we take the same approach to obtain a staged abstract semantics.
@@
 
-\paragraph{Staged Fixpoint Iteration.} 
+\subsection{Staged Fixpoint Iteration.} 
 \todo{staged fixed point iteration with memoization}
@@
 Futamura projection?
-
-\subsection{Lattice-theoretic Interpretation}
+Lattice-theoretic Interpretation

### 2018-11-13 paper/30_sai.tex
fix mistake when generating functions

@@
 
+\subsection{Modular Analysis for Free}
+
+Motivation: one of the challenges of modern static analysis is program usually depends on
+large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
+Can we analyze programs and libraries separately without losing precision? So that we can 
+reduce part of the overhead of repeatedly analyzing libraries for different programs.
+Similarly, some static analyzers compute summary for a function or a module, that can be reused
+later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
+or unsound, which both lead to imprecision.
+
+Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
+because it is interprocedural and need the last k calling contexts to distinguish
+different call sites.
+But can we analyze programs (libraries) separately which generate the specialized 
+analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
+and then install these contexts when we have the whole program.
+
+Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
+then maybe there can be $n$ stages. 
+Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
+waiting for the rest $(n-m)$ programs.
+These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
+the abstract interpreter itself might be a partial abstract interpreter.
+
 \subsection{Discussion}
+Review a little bit what we have done?
 what not to do when writing an abstract interpreter for staging?
@@
 
-Big-step vs small-step
+\paragraph{Big-step vs Small-step.}
 
@@
 
+\paragraph{Soudness.}
+
 \cite{10.1007/3-540-61580-6_11}

### 2018-11-13 code/src/main/scala/sai/common/MyScalaGenTupledFunctions.scala

@@ -51,12 +51,18 @@ trait MyScalaGenTupledFunctions extends ScalaGenFunctions with GenericGenUnboxed
       val res = remap(m.typeArguments.last)
       //val targsUnboxed = targs.flatMap(t => unwrapTupleStr(t))
       val targsUnboxed = targs.map(t => unwrapTupleStr(t))
-      val size = targs.map(s => topLevelTupleSize(s)).foldLeft(0)(_ + _)
-      //println(s"targsUnboxed.length: ${targsUnboxed.length}")
-      //println(s"targsUnboxed: ${targsUnboxed.mkString(",")}")
-      val sep = if (targsUnboxed.length > 0) "," else ""
-      //"scala.Function" + (targsUnboxed.length) + "[" + targsUnboxed.mkString(",") + sep + res + "]"
-      s"scala.Function${size}[" + targsUnboxed.mkString(",") + sep + res + "]"
+      if (targsUnboxed.head == "Unit") { //FIXME: GW hack
+        val size = targs.map(s => topLevelTupleSize(s)).foldLeft(0)(_ + _)
+        val sep = if (targsUnboxed.tail.length > 0) "," else ""
+        s"scala.Function${size-1}[" + targsUnboxed.tail.mkString(",") + sep + res + "]"
+      } else {
+        val size = targs.map(s => topLevelTupleSize(s)).foldLeft(0)(_ + _)
+        //println(s"targsUnboxed.length: ${targsUnboxed.length}")
+        //println(s"targsUnboxed: ${targsUnboxed.mkString(",")}")
+        val sep = if (targsUnboxed.length > 0) "," else ""
+        //"scala.Function" + (targsUnboxed.length) + "[" + targsUnboxed.mkString(",") + sep + res + "]"
+        s"scala.Function${size}[" + targsUnboxed.mkString(",") + sep + res + "]"
+      }
 
     case _ => super.remap(m)
   }

### 2018-11-15 paper/10_intro.tex
revise intro

@@
 compilers and interpreters. 
-The first Futamura projection specifically shows that specializing an interpreter with respect to the input 
-program yields an equivalent executable.
-Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to realize Futamura projections,
-which first identifies the binding-time of variables in the program, i.e., they can be known statically or dynamically, 
-and then specializes the program with respect to the given input of program.
-However, precisely analyzing binding-time for an arbitray program is hard in general. As an alternative and programmatic
-approach to specialization and partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} 
-requires the stage annotations (i.e., binding-time annotations) from the programmers. These staging annotations 
-control which part of the input program should be specialized.
-As a simple example, we consider a power function: 
+The first Futamura projection specifically shows that specializing an interpreter with respect to 
+the input program yields an equivalent executable.
+Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to realize 
+Futamura projections, which first identifies the binding-time of variables in the program, i.e., 
+they can be known whether statically or dynamically, and then evaluates the static part, and 
+generates a residual program that relies on the dynamic part.
+However, precisely analyzing binding-time for an arbitrary program is hard in general. As an 
+alternative and programmatic approach to specialization and partial evaluation, multi-stage 
+programming (MSP for shot) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage 
+annotations (i.e., binding-time annotations) from the programmers. These staging annotations 
+identify which part of the input program should be specialized.
+As a classical example, we consider a power function: 
 \begin{lstlisting}
@@
 \end{lstlisting}
-If \texttt{e} is identified as statically known by the programmer, say 5, then we can specialize \texttt{power} 
-with respect to \texttt{e = 5} and generate a new specialized function as following:
+If \texttt{e} is identified as statically known by the programmer, say 5, then we can 
+specialize the function \texttt{power} with respect to \texttt{e = 5} and generate a new 
+specialized function that runs faster:
 \begin{lstlisting}
@@
 
-Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach to construct sound static analysis 
-by approximation is almost as old as Futamura projection.
-Some recent progress such as Abstracting Abstract Machines (AAM) uncovers a methodology to derive sound abstract interpreters 
-from their concrete counterparts and has been applied to different variants of definitional interpreters and abstract machines
+Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach to construct 
+sound static analysis by approximation is almost as old as Futamura projection.
+Some recent progress such as Abstracting Abstract Machines (AAM) uncovers a methodology to 
+derive sound abstract interpreters from their concrete counterparts and has been applied to 
+different variants of definitional interpreters and abstract machines
 \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-\note{TR: Cousot was already systematic. Both Futamura and Cousot are results from the 1970s, but have
-profited from more recent results (MSP and AAM) -- build a connection?} 
 Given the structural similarity between concrete and abstract interpreters, intellectually 
-it is natural to raise the question whether it is possible to stage an abstract interpreter, and how 
-to effectively specialize an abstract interpreter having considered their functional dissimilarity. 
-This paper studies the application of the first Futamura projection on abstract interpreters, 
-and the approach to construct efficient abstract interpreters by multi-stage programming.
+it is natural to raise the question whether it is possible to specialize an abstract interpreter, and how 
+can we effectively do that having considered their functional dissimilarity. 
+This paper studies the confluence of two old ideas --- Futamura project and abstract interpretation,
+but from the perspective of their recent realizations --- multi-stage programming and definitional
+abstract interpreters. To be specific, we present the application of the first Futamura projection
+on abstract interpreters, and the approach of constructing optimizing abstract interpreters by 
+multi-stage programming.
 
@@
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
-and indeed it is able to effectively remove the interpretive overhead of 
-repeatedly traversing the abstract syntax tree. 
-However, despite of the previous works mostly focused on one particular analysis,
-or required to rewrite the analyzer,
-it is worth to investigate the idea from a modern perspective based on meta-programming,
+and indeed it is able to effectively remove the interpretive overhead of repeatedly traversing 
+the abstract syntax tree. 
+However, despite of the previous works mostly focused on one particular analysis, or required to 
+completely rewrite the analyzer,
+it is worth to investigate the idea from a modern perspective based on generative programming,
 especially for a general abstract interpreter that models direct-style $\lambda$-calculus, imperative 
@@
 but also the data structure of abstract domains, abstract environment and heap, 
-and the fixpoint iteration are staged.
+and the fixed-point iteration are staged.
 
 On the other hand, the abstract interpreter, as a semantic artifact, should be written in a 
-style that is easy for people to communicate the formulation and abstraction, but also can be implemented 
-efficiently. As the solgan of multi-stage programming said, "abstraction without regret",
-we draw connection between abstract formulation and efficient implementation \todo{wording},
-just like the connection between interpreters and compilers drawn by Futamura projection.
-Particularly, we show a systematic way of adding stage annotations to the abstract interpreter
-without changing the code of interpreter skeleton that is shared between four concrete or abstract,
+style that is easy for people to communicate the formulation and abstraction, but also can be 
+implemented efficiently. As the slogan of multi-stage programming said, "abstraction without regret",
+we draw connection between high-level description and efficient implementation of abstract interpreters,
+just like the connection between concrete interpreters and compilers drawn by Futamura projection.
+Particularly, we show a easy but systematic way of adding stage annotations to the abstract interpreter
+without changing the code of interpreter skeleton, which is shared between four concrete or abstract,
 unstaged or staged interpreters. We use LMS framework for staging, which allows only use types to 
@@
 %Particularly, we also would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
-
-From the performance-wise, precise whole-program analyses formulated as abstract 
-interpretation can be very slow due to the inherent complexity of the algorithm and the large scale 
-of the analyzed programs \cite{toman_et_al:LIPIcs:2017:7121}. \todo{where put this}
-
+\iffalse
 When staging a concrete interpreter, the programmers need to distinguish 
@@
 Then what is the point of staging if there is no such distinction?
-\note{TR: I don't understand this. The program
-structure is static, the abstract values are still dynamic, no? They 
-change in every iteration of the fixpoint algorithm}
 A surprising by-product of thinking about this question is to realize that 
@@
 mechanically. 
+\note{TR: I don't understand this. The program structure is static, the abstract values are still dynamic, no? They change in every iteration of the fixpoint algorithm} 
+\note{TR: why is it a big deal. On Closed programs we obtain a constant factor, but modular analysis has different asymptotics}
+\fi
 
-The use of staged abstract interpreters on open programs improves the scalability. 
-\note{TR: why is it a big deal. On Closed programs we obtain a constant factor, but modular analysis has different asymptotics}
-One challenge in static analysis of modern software is that the programs are usually 
-shipped with large library code, for example, it has been shown that analyzing a simple
-"Hello World" program in Java depends on additional 3,000 classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}.
-Soundly and precisely analyzing these libraries sometimes are expensive and unnecessarily repeated.
-When applying the staged abstract interpreter on libraries, we leave the unknown 
-calling context as dynamic value and generate partial analyzing result which is 
-repsented as a residul program. The partial analyzing result can be resued and composed
-with the analyzing result of application code when available later. 
+When targeting higher-order programs, either staged concrete interpreters or staged abstract interpreters
+are able to compile a closure, i.e., specialize the call of interpreter with respect to the body
+expression of the lambda term without knowing the actual argument value. By generalizing this observation,
+we can actually specialize the abstract interpreter with any open programs, 
+which unexpectedly leads to a modular analysis and improves scalability.
+A open program contains free variables, which represent other parts of the program and will be left
+as dynamic values.
+For instances, one challenge in static analysis of modern software is that the programs are usually 
+shipped with large library code \cite{toman_et_al:LIPIcs:2017:7121}, for example, it has been shown that 
+analyzing a simple "Hello World" program in Java depends on additional 3,000 classes in the library 
+\cite{DBLP:conf/oopsla/KulkarniMZN16}.
+A precise whole-program analysis formulated as abstract interpreters can be very expensive
+due to the scale of the program and the inherent complexity of the algorithm.
+However, analyzing these libraries sometimes are unnecessarily repeated.
+When applying the staged abstract interpreter on such open programs, we leave the unknown arguments and 
+calling contexts as dynamic values and generate partial analyzing result which is 
+represented as a residual program. The partial analyzing result can be reused and composed with the 
+analyzing result of application code when available later. Therefore we mechanically obtain a modular
+analysis through staging, even though the original algorithm is formulated as a whole-program analysis.
+
 It has been observed that partially applying context-sensitivity on selected portion of the program could
@@
 
-staged polymorphism, decide when to compile an analysis.  
-end-to-end staging: abstract interpreter, abstract domain, environment/store structure, fixpoint iteration.
 % Many optimization of static analysis are ad-hoc.
@@
 \begin{itemize}
-\item Starting from a definitional abstract interpreter for a higher-order functional language,
-  whose syntax and semantics are described in Section~\ref{bg_lang},
-  we show how to turn it into a staged abstract interpreter (SAI) by adding stage annotations. 
-  The design and implementation of SAI using Scala and Lightweight Modular Staging (LMS) framework 
-  are described in Section~\ref{sai}.
-  We later extend the staged abstract interpreter to support imperative language features 
-  (Section~\ref{cases_imp}), such as states, loops and exceptions.
+  \item Starting from a generic definition of interpreter for a higher-order functional language,
+    whose syntax and semantics are described in Section~\ref{bg_lang}, we show its concrete
+    interpreter (Section~\ref{unstaged_conc}), staged concrete interpreter (Section~\ref{stagedinterp}), 
+    abstract interpreter (Section~\ref{unstaged_abs}) and staged abstract interpreter (Section~\ref{sai}).
+    Our design allows one can derive these four different kinds of interpreters by mechanically changing
+    the underlying semantics or stage annotations.
+    %We later extend the staged abstract interpreter to support imperative language features (Section~\ref{cases_imp}), such as states and loops.
   \item We demonstrate that the staged abstract interpreter not only improves the \textit{efficiency}
     but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
-  programs. \todo{more on this point.}
-\item To demonstrate the feasibility and applicability of the proposed approach, we conduct three case studies in Section~\ref{cases_study}. 
-  First, we revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587}, which is a closure generation 
-  technique applied on 0-CFA. We show that it can be understood and easily implemented as an instance of 
-  our staged abstract interpreter. 
-  Second, we demonstrate sophisticated control-flow analysis techniques, such as adaptive context/flow/path-sensitive 
-  analysis and store-widening can be integrated with staged abstract interpreters without losing any precision.
-  Finally, we show that in an imperative language, the idea can be applied to an abstract interpreter with
-  an abstract domain that performs numerical analysis.
+    programs. 
+    %\todo{more on this point.}
+  \item To demonstrate the applicability and usefulness of the proposed approach, we conduct three case studies 
+    in Section~\ref{cases_study}. 
+    \subitem (1) We revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587} in Section~\ref{cs_ac}, 
+    which was originally a closure generation technique applied on 0-CFA.
+    We show that abstract compilation can be understood and easily implemented as an instance of our staged abstract interpreter.
+    But unlike the abstract compilation, our approach does not require the implementers to rewrite the analysis.
+    \subitem (2) We demonstrate pushdown control-flow analysis techniques (Section~\ref{cfa}), such as context-sensitive 
+    (\todo{adaptive/flow})
+    analysis and more precise stores can be integrated to staged abstract interpreters without losing any precision.
+    \subitem (3) We show that in an imperative language (Section~\ref{cases_imp}), by replacing the abstract domain
+    to an interval domain (as well as staged), the staged abstract interpreter performs numerical analysis.
     % (Optional: taint/information flow analysis?)
   \item We empirically evaluate the performance improvement by staging an big-step abstract interpreter 
-  for control-flow analysis. We compare against with the unstaged version, a small-step version (unstaged), 
-  and an optimizing small-step version.
+    for control-flow analysis in Section~\ref{evaluation}. 
+    We compare our staged abstract interpreter with the unstaged version and a small-step abstract interpreter.
 \end{itemize}

### 2018-11-15 paper/23_bg_stage.tex
paper

@@
 \begin{lstlisting}
-.. = new DslDriver[Int, Int] {
+new DslDriver[Int, Int] {
   def power(b: Rep[Int], x: Int): Rep[Int] =

### 2018-11-15 paper/30_sai.tex
paper

@@
 When generating code for application, we can not directly apply the callee.
-Instead, we emit code that calls a next-stage function @apply_closures_norep@ 
-which non-deterministically applies multiple target closure with the argument
+Instead, we emit code that calls a next-stage function @apply_closures_norep@.
+Function @apply_closures_norep@ does the same job as its unstaged version,
+which non-deterministically applies multiple target closures with the argument
 and latest store, and finally returns the joined value and a single store.
+We provide the definition of @apply_closures_norep@ in the runtime supporting code.
 
@@
 
-\subsection{Staged Fixpoint Iteration.} 
-\todo{staged fixed point iteration with memoization}
+\subsection{Staged Fixpoint Iteration}
+
+Our fixed-point iteration relies two caches @in@ and @out@, but the iteration 
+can not be done at the current stage.
+Because the @in@ and @out@ are both next-stage values, thus the test of whether 
+@in@ and @out@ are equal is a generated expression to the next stage, and we can only 
+know the comparison result at the next stage. 
+In other words, we do not know how many iterations we need to reach the fixed-point.
+To achieve this, we need to stage a
+function --- @iter_aux@ is generated as a recursive function of type @Rep[Unit => (Value,Store)]@
+and will be invoked at the next stage.
+
+\begin{lstlisting}
+def iter(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
+Rep[(Value,Store)] = {
+  def iter_aux: Rep[Unit => (Value,Store)] = fun { () =>
+    in = out; out = Map[Config, (Value,Store)]()
+    cached_ev(e, $\rho$, $\sigma$)
+    if (in === out) out((unit(e), $\rho$, $\sigma$)) else iter_aux()
+  }
+  iter_aux() // generated code that invokes iter_aux()
+}
+\end{lstlisting}
+
+However, the instrumented evaluation function that uses @in@ cache and updates @out@ cache
+can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
+if it is applied on subexpressions of the analyzed program.
+
+\begin{lstlisting}
+def cached_ev(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
+Rep[(Value, Store)] = {
+  val cfg: Rep[Config] = (unit(e), $\rho$, $\sigma$)
+  if (out.contains(cfg)) { out(cfg) }
+  else {
+    val ans0 = in.getOrElse(cfg, RepLattice[(Value, Store)].bot)
+    out = out + (cfg -> ans0)
+    val ans1 = evev(cached_ev)(e, $\rho$, $\sigma$)
+    out = out + (cfg -> (ans0 $\sqcup$ ans1)); ans1
+  }
+}
+\end{lstlisting}
 
@@
 
-Now we already obtained a working version of staged abstract interpter that is able to 
+Now we already obtained an end-to-end staged abstract interpreter that is able to 
 specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,

### 2018-11-15 paper/30_sai.tex
paper

@@
 In other words, we do not know how many iterations we need to reach the fixed-point.
-To achieve this, we need to stage a
-function --- @iter_aux@ is generated as a recursive function of type @Rep[Unit => (Value,Store)]@
-and will be invoked at the next stage.
+To achieve this, we need to stage a function value --- @iter_aux@ is generated as a 
+recursive function of type @Rep[Unit => (Value,Store)]@ and will be invoked at the next stage.
 
@@
 
-\subsection{Specialized Data Structures}
+\subsection{Specialized Data Structures} \label{staged_ds}
 
@@
 \subsection{Discussion}
-Review a little bit what we have done?
-what not to do when writing an abstract interpreter for staging?
-what kind of abstract interpreter is friendly for staging/specialization?
-
-Stage an abstract abstract machine yields another small-step abstract machine
+%We have presented all of our four different interpreters as shown in Figure~\ref{confluence}.
+In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} gave several
+advice on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
+We borrow these advice to abstract interpreters and discuss decisions we made to achieve this 
+and examine some alternatives.
 
 \paragraph{Big-step vs Small-step.}
+What we implemented is a big-step, compositional \todo{or semi-compositional} abstract interpreter, 
+where the "compositional" means every recursive call of our abstract interpreter is applied with proper 
+substructures of the current syntactic parameters. 
+The compositionality ensures the specialization can be done by unfolding, as well as the termination of specialization procedure.
+Nevertheless, it is also possible to specialize a small-step, operational abstract semantics --- \citeauthor{Johnson:2013:OAA:2500365.2500604}
+showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimization 
+to Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
+However, the generated abstract bytecodes still requires another small-step abstract machine 
+to execute, which is an additional engineering efforts.
+Another alternative approach for efficient specialization is to write the abstract interpreter 
+in a big-step, monadic style \cite{DBLP:journals/pacmpl/DaraisLNH17}.
 
-direct-style vs CPS
+% direct-style vs CPS
 
-Operational vs denotational
-
-General insights for staging other static analysis algorithm.
-
-\paragraph{Soudness.}
-
-\cite{10.1007/3-540-61580-6_11}
-
-Jones-optimal specialization?
-Futamura projection?
-Lattice-theoretic Interpretation
+\paragraph{Correctness and Soundness.}
+Based on the assumption that LMS preserves the semantics during staging, and indeed it is 
+in our application, 
+we are confident that the staged abstract interpreter does the same analysis compared 
+with the unstaged one. Moreover, the optimization done by staging does not compromise
+and soundness.

### 2018-11-15 paper/50_cases.tex
revise intro
paper

--- 2018-11-06 paper/50_cases.tex
@@
 
-\subsection{Abstract Compilation a la Staging}
+In Section~\ref{sai}, we have shown staging an abstract interpreter is feasible and can be 
+systematic after correctly identifying the binding-times, despite the fact that the abstract 
+interpreter is intended to be imprecise and easy to implement. 
+In this section, we conduct several case studies to show that it is also useful and widely
+applicable to different analysis.
 
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation as a new
+\subsection{Abstract Compilation a la Staging} \label{cs_ac}
+
+\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC) as a new
 implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
 The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
-at compile-time therefore the overhead if interpretation can be eliminated. The compiled analysis
-can be represented by text or closures (higher-order functions).
+statically therefore the overhead of interpretation can be eliminated. 
+In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
+the closures can be executed immediately however the textual program need to be loaded firstly.
+
 Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
 analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
-for continuation-passing style (CPS) programs. Since the analyzed program is in CPS, the analyzer 
-is essentially a big-step CE (\textit{Control-Environment}) abstract interpreter.
-After specializing the analyzer with respect to the input program, the \textit{control} component of 
-abstract interpreter is eliminated, instead, the compiled analysis is a closure which takes an 
-environment as argument. Therefore the overhead of traversing the abstract syntax tree of input 
-program also has been eliminated.
+for continuation-passing style (CPS) programs. Since the analyzed program is written in CPS, the analyzer 
+is essentially a big-step control-environment abstract interpreter.
+Closure generation compiles the analysis as a closure which takes an environment as argument.
+And the overhead of traversing the abstract syntax tree of input program also has been eliminated.
 
@@
 understood and implemented as an instance of staging abstract interpreters.
-We firstly revisit abstract compilation in the style of closure generation,
-and then reproduce their result by adding stage annotations to the unstaged 0-CFA algorithm.
-The generated program of our approach improves approximately the same extent of speed (roughly a factor of 2),
-but without changing a single line of the analyzer program (with the use of LMS), 
-while closure generation requires more engineering effort, such as a whole-program 
-conversion on the analyzer. Moreover, our approach is able to not only remove the interpretive
-overhead, but also specialize the data structures used in the analysis, for example, the store
-that maps variables to sets of lambda.
+We firstly revisit the original implementation of abstract compilation of 0-CFA,
+and then reproduce their result by simply adding stage annotations.
+The generated program of our approach improves approximately the same extent of speed,
+but without changing a single line of the analyzer program (with the use of LMS).
+However, closure generation requires more engineering effort, such as a whole-program 
+conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
+our approach is able to not only remove the interpretive overhead, but also specialize 
+the data structures used in the analysis, for example, the environment that maps variables to sets of lambda.
+
+\begin{figure*}
+  \centering
+  \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+    \begin{lstlisting}
+type CompAnalysis = Store => Store
+def compProgram(prog: Expr): CompAnalysis = compCall(prog)
+def compCall(call: Expr): CompAnalysis = call match {
+  case Letrec(bds, body) =>
+    val C1 = compCall(body); val C2 = compArgs(bds.map(_.value))
+    ($\sigma$: Store) => C1(C2($\sigma$.update(bds.map(_.name), 
+       bds.map(b => Set(b.value.asInstanceOf[Lam])))))
+  case App(f, args) =>
+    val C1 = compApp(f, args); val C2 = compArgs(args)
+    ($\sigma$: Store) => C1(C2($\sigma$))
+}
+def compApp(f: Expr, args: List[Expr]): CompAnalysis = 
+  f match {
+    case Var(x) => ($\sigma$: Store) => 
+      analysisAbsApp($\sigma$.lookup(x), args, $\sigma$)
+    case Op(_) => compArgs(args)
+    case Lam(vars, body) =>
+      val C = compCall(body)
+      ($\sigma$: Store) => 
+        C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
+  }
+def compArgs(args: List[Expr]): CompAnalysis = args match {
+  case Nil => ($\sigma$: Store) => $\sigma$
+  case (arg@Lam(vars, body))::rest =>
+    val C1 = compCall(body); val C2 = compArgs(rest)
+    ($\sigma$: Store) => C2(C1($\sigma$))
+  case _::rest => compArgs(rest)
+}
+  \end{lstlisting}
+  \end{subfigure}
+\hfill
+  \begin{subfigure}[h]{0.49\textwidth}
+    \centering
+    \begin{lstlisting}
+def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
+  analyzeCall(prog, $\sigma$)
+def analyzeCall(call: Expr, $\sigma$: Rep[Store]): Rep[Store] = 
+  call match {
+    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
+    case Letrec(bds, body) =>
+      val $\sigma$_* = $\sigma$.update(bds.map(_.name), 
+        bds.map(b => Set(b.value.asInstanceOf[Lam])))
+      val $\sigma$_** = analyzeArgs(bds.map(_.value), $\sigma$_*)
+      analyzeCall(body, $\sigma$_**)
+  }
+def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
+  f match {
+    case Var(x) => analyzeAbsApp(args, $\sigma$(x), $\sigma$)
+    case Op(_) => analyzeArgs(args, $\sigma$)
+    case Lam(vars, body) =>
+      val $\sigma$_* = $\sigma$.update(vars, args.map(primEval(_, $\sigma$)))
+      analyzeCall(body, $\sigma$_*)
+  }
+def analyzeArgs(args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
+  args match {
+    case Nil => $\sigma$
+    case Lam(vars, body)::rest => analyzeArgs(rest, analyzeCall(body, $\sigma$))
+    case _::rest => analyzeArgs(rest, $\sigma$)
+  }
+  \end{lstlisting}
+
+  \end{subfigure}
+  \caption{Comparison of AC (left) and SAI (right). Only core code are shown.}
+  \label{compare_ac_sai}
+\end{figure*}
 
@@
 
+The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA for a CPS language consisting
+of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
+structs are decomposed to different functions, such as @analyzeCall@ and @analyzeApp@. The idea of closure
+generation is to rewrite these functions, where previously they may take both static arugments and dynamic arguments,
+but after the rewrite only the static arguments is taken. In this case, the static arguments are syntactic terms;
+the dynamic arguments are stores. After written in AC style, the functions like @compCall@ (compiled version of @analyzeCall@) 
+returns a value of type
+@CompAnalysis@, i.e., a closure that takes a store and returns a store. The result of multiple calls on such functions, for example,
+@compCall@ and @compArgs@ can be composed. The generated closure only takes stores, because the input program
+is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).
+
 \subsubsection{Staged 0-CFA}
 
-\subsubsection{Specializing the Data Structure}
+On the other side, our approach does exactly the same thing through staging: the syntactic terms are static, and 
+stores are dynamic, therefore the generated code just looks-up and updates the store.
+Figure~\ref{compare_ac_sai} (right) shows the code written with LMS. In fact, the only changes are
+the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
+at the next stage. Indeed, a amount of additional engineering efforts are required to make this happen, 
+including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
+current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
+functions of AST structs. We consider these efforts are relatively small, and does not interfere the actual analysis
+we desire.
 
@@
 which functions will possibly be applied at each call-site. 
-Though, in Section~\ref{?}, we discussed abstraction compilation and staging on a monovariant 
-control-flow analysis for programs written in continuation-passing style; 
-in this section, based on the staged abstract interpreter we presented in Section~\ref{sai}, 
-we further discuss staging control-flow analysis for direct style languages in depth, 
-especially polyvariant analyses.
+The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is already 
+a store-widened 0-CFA-like abstraction; in last section, we also reviewed AC with 0-CFA.
+In this section, based on the existing staged abstract interpreter, we further
+develop the staging techniques with control-flow analyses, including recoverying a more precise store model;
+different polyvariant analsis, and using staging as a implementation stategy for mixed sensitivity.
+
+\subsubsection{A More Precise Store Model}
 

### 2018-11-15 paper/30_sai.tex
misc

@@
 where the "compositional" means every recursive call of our abstract interpreter is applied with proper 
-substructures of the current syntactic parameters. 
+substructures of the current syntactic parameters \cite{10.1007/3-540-61580-6_11}.
 The compositionality ensures the specialization can be done by unfolding, as well as the termination of specialization procedure.
 Nevertheless, it is also possible to specialize a small-step, operational abstract semantics --- \citeauthor{Johnson:2013:OAA:2500365.2500604}
-showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimization 
-to Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
-However, the generated abstract bytecodes still requires another small-step abstract machine 
+showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimizations
+of Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
+However, the generated abstract bytecode still requires another small-step abstract machine 
 to execute, which is an additional engineering efforts.

### 2018-11-16 paper/22_bg_abs.tex
paper

@@
         val (v, v$\sigma$) = ev(e, put($\rho$, x, $\alpha$), put($\sigma$_*, $\alpha$, arg))
-        $\sigma$_* = v$\sigma$; v
+        $\sigma$_* = v$\sigma$ $\sqcup$ $\sigma$_*; v
       }

### 2018-11-16 paper/50_cases.tex
paper

@@
 which functions will possibly be applied at each call-site. 
-The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is already 
-a store-widened 0-CFA-like abstraction; in last section, we also reviewed AC with 0-CFA.
+The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 
+0-CFA-like abstraction, moreover it is also a pushdown control-flow analysis; 
+in last section, we also reviewed AC with finite-state 0-CFA.
 In this section, based on the existing staged abstract interpreter, we further
 develop the staging techniques with control-flow analyses, including recoverying a more precise store model;
-different polyvariant analsis, and using staging as a implementation stategy for mixed sensitivity.
+context-sensitive analsis, and using staging as a implementation stategy for mixed sensitivity.
 
@@
 
-\subsubsection{Context/Path/Flow-Sensitivity}
+The store-widened analysis we implemented in Section~\ref{sai} uses a single store to approximate
+the runtime store. It can be efficently computed in polynomial time together with 0-CFA-like abstraction, 
+but sometimes we may desire a 
+more precise result that distinguish the final values and stores for different closure targets. 
+To achieve this goal,
+we need to tweak our abstract interpreter and type instantiation. The answer type @Ans@ is changed to
+a set of @VS@s where a @VS@ is a pair of sets of abstract values (such as closures or abstract numbers)
+and a store.
 
-\subsubsection{Mixed Sensitivity}
+\begin{lstlisting}
+type VS = (Set[AbsValue], Store)
+type Ans = R[Set[VS]]
+\end{lstlisting}
 
-using different $k$ for $k$-CFA
+Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under staging the type @Ans@
+represents a next stage value. Then, our generic interpreter is also changed when handling function 
+applications.
+
+\begin{lstlisting}
+case App(e1, e2) =>
+  val e1ans = ev(e1, $\rho$, $\sigma$)
+  val e1vs  = choices(e1ans)
+  val e2ans = ev(e2, $\rho$, e1vs.$\sigma$)
+  val e2vs  = choices(e2ans)
+  apply_closure(ev)(e1vs.v, e2vs.v, e2vs.$\sigma$)
+\end{lstlisting}
+
+The idea is to handle the application is to explore all possible closures from @e1@ and meanwhile
+use the latest store.
+What @choices@ does is similar to McCarthy's @amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically return an
+element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side receiver, i.e., @e1vs@.
+The function @choices@ internally uses delimited control operator @shift@ to capture the continuation,
+which is the code block after its call site. 
+This allows us to perform nesting depth-first evaluation for function application but still writing 
+the program in direct-style \cite{Wei:2018:RAA:3243631.3236800}. 
+Again, we can stage this part as we did for the naive abstract interpreter.
+
+\subsubsection{Context-Sensitivity}
+
+We add k-CFA-like context-sensitivity to the analysis by introducing an abstract time stamp, 
+whose concrete instantiation is a finite list of expression that track $k$ recent calling context.
+The definition of abstract addresses is changed to a tuple of identifiers and the time it get allocated,
+meaning that this address points to some values under such calling context.
+If $k$ is 0, we obtain a monovariant analysis as demonstrated before; if $k > 0$, a higher precision
+analysis is obtained.
+
+\begin{lstlisting}
+type Time = List[Expr]
+\end{lstlisting}
+
+Every time when we call the @eval@ function, we refresh the time stamp by calling function @tick@,
+which returns a new time stamp. Here we adopt a $k$-CFA-like allocation strategy, therefore
+the @tick@ function can be implemented as appending the current expression being evaluated to the exiting
+calling context, and then taking the first $k$ elements from the list.
+
+\begin{lstlisting}
+def tick(e: Expr, $\tau$: R[Time]): R[Time] = (e :: $\tau$).take(k)
+def eval(ev: EvalFun)
+  (e: Expr, $\rho$: R[Env], $\sigma$: R[Store], $\tau$: R[Time]): Ans = {
+    val $\tau$_* = tick(e, $\tau$)
+    e match {
+      case Lit(i) => ...
+      ...
+    }
+  }
+\end{lstlisting}
+
+Accordingly, the type of return value is accompanied with the time stamp. For recursive call of @ev@,
+which would also return the latest time stamp, and that time stamp will be used for the afterward evaluation.
+
+\begin{lstlisting}
+type VST = (Value, Store, Time)
+type Ans = R[Set[VST]]
+\end{lstlisting}
+
+Using other allocation strategies to achieve different sensitivities is possible \cite{DBLP:conf/icfp/Gilray0M16} 
+and will be staged under our framework.
+
+%\subsubsection{Mixed Sensitivity}
+
+%using different $k$ for $k$-CFA
 
@@
 
-Beyond control-flow analysis, staging data structures used in the abstract interpreter, such as abstract domain.
+Now we consider in a first-order imperative language, we may care more about the data-flow because the control-flow
+is relatively easy to obtain. In this section, we show the staging of other abstract domains, particularly an 
+interval domain for numbers.
 It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
 improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
-In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.
+In this section, we first show how to support imperative language features to the abstract interpreter, 
+and present a similar idea for interval domain but show the specialization is feasible by systematically staging.
 
@@
 
-Assignments, Loops, Exceptions
+To evaluates an assignment, we first evaluate its right-hand side, and then put the value to the slot where
+the address of @v@ points to.
+For simplicity, we elect to make the value of the assignment be the value of its right-hand side.
 
-\subsubsection{Specializing Interval Domains}
+\begin{lstlisting}
+case Assign(x, e) =>
+  val (v, $\sigma$_*) = ev(e, $\rho$, $\sigma$)
+  (v, put($\sigma$_*, get($\rho$, x), v))
+\end{lstlisting}
 
-Intervals, Numerical Analysis
+without introducing new language value members.
 
-Options: Data-flow analysis, taint analysis
+\begin{lstlisting}
+case While(t, e) =>
+  val (tv, t$\sigma$) = ev(t, $\rho$, $\sigma$)
+  if (notZero(tv)) 
+\end{lstlisting}
+
+\subsubsection{Staged Interval}
+
+\begin{lstlisting}
+case class Interval(lb: Rep[Double], ub: Rep[Double]) {
+  def +(that: Interval): Interval = that match {
+    case Interval(lb_, ub_) => Interval(lb + lb_, ub + ub_)
+  }
+  def -(that: Interval): Interval = that match {
+    case Interval(lb_, ub_) => Interval(lb - lb_, ub_ - ub) 
+  }
+  ...
+}
+\end{lstlisting}

### 2018-11-16 paper/70_related.tex
paper

@@
 
-\textbf{Abstract Compilation}. The idea in this paper is closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
+\paragraph{Optimizing Static Analysis Through Specialization}
+The idea in this paper is closely inspired by abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
 \citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
@@
 tree by partial evaluation. Specifically, they proposed two similar kinds of abstract compilation techniques.
-The first one is to generate specialized analysis as a textual program, then 
+The first one is to generate specialized analysis as a textual program, which then
 can be loaded and executed by \texttt{eval} or other similar mechanisms. 
-The second one is using closure, i.e., a function that remembers its environment, 
-as a representation of specialized analysis. 
-So compiling the analysis generates higer-order functions on-the-fly with respect to the analyzed program,
-then the generated closure can be applied immediately in the higher-order host language.
+The second one is to use closures, i.e., functions that remember their environments, 
+as a representation of specialized analysis. As we show in the case study,
+compiling the analysis generates higer-order functions on-the-fly with respect to 
+the analyzed program, then the generated closure can be applied immediately in the 
+higher-order host language.
 
-\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea closure generation 
-for optimizing small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
+\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure generation 
+to optimize small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
 The analyzed program is firstly compiled to a intermediate representation called 
@@
 
-%\citeauthor{DBLP:conf/flops/AmatoS01} apply abstract compilation for non par-sharing and freeness analysis in logic programs \cite{DBLP:conf/flops/AmatoS01}.
+\citet{damian1999partial} provides a formal treatment to abstract compilation and Shiver's CFA, 
+as well as proofs to establish correctness and certified specialized analyzers \cite{damian1999partial}.
+\citeauthor{amtoft1999partial} applyies partial evaluation for constraint-based 
+control flow analysis \cite{amtoft1999partial}.
+Split the analysis to be multiple stages is also studied other than control-flow analysis,
+though the formulation may very different.
+For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer 
+analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first stage is to analyze the program code to 
+obtain a sparse representation, and then the second stage conducts the flow-sensitive analysis.
 
-\citeauthor{damian1999partial}'s work on partial evaluation for program analysis \cite{damian1999partial}.
+\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domain through online decomposition \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
+The idea is to decompose a large polyhedra into several smaller one; one abstract transformer
+usually only depends a limit number of variables, then we just need to compute the resulting polyhera
+on those smaller polyhedra that contains these related variables.
+The decomposition can be considered as a specialization of abstract domain with respect to the program
+static structure.
 
-\citeauthor{amtoft1999partial}'s work on partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-
-\citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. 
-They firstly stage the program code to obtain a sparse representation, then conduct
-the flow-sensitive analysis.
-
-\textbf{Abstract Interpreters}.
-
+\paragraph{Abstract Interpreters}
 Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
@@
 which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
-
-Polyhedra, online decomposition
-
 \citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
 \citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
+The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
 
@@
 
-Multi-stage programming. MetaML, MetaOCaml, Mint, LMS
-
+The Lightweight Modular Staging framework in Scala
 Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 

### 2018-11-16 paper/60_eval.tex
worked on paper
revise intro
Run evaluation

@@
-\section{Evaluation}
+\section{Evaluation} \label{evaluation}
 
 \subsection{Benchmarks}
+
+We have designed, implemented and evaluated a prototype abstract interpreter for Scheme.
+Our prototype includes both staged and unstaged versions of the abstract interpreter for comparison 
+purposes.
+
+We ran our prototype abstract interpreter on a suite of benchmark scheme programs that are also used 
+in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}.
 

### 2018-11-16 code/src/main/scala/sai/Evaluation.scala

@@ -10,7 +10,7 @@ object Evaluation extends AbsLamCalTrait {
   def main(args: Array[String]) = {
     val prog = boyer
     println(prog)
-//    println("staged: " + evalStaged(prog))
+    println("staged: " + evalStaged(prog))
     println("unstaged: " + evalUnstaged(prog))
   }
 }

### 2018-11-16 code/src/main/scala/sai/common/parser/SimpleSchemeTokenParser.scala

@@ -15,6 +15,7 @@ trait SchemeTokenParser extends RegexParsers {
   def RPAREN = "[\\)\\}\\]]".r
 
   def QUASIQUOTE: Parser[String] = "`" | "'"
+  def QUOTE: Parser[String] = "quote"
   def UNQUOTE: Parser[String] = ","
   def SYMBOL: Parser[String] = """[^ \t\n\(\{\[\)\}\]]+""".r
 

### 2018-11-16 code/src/main/scala/sai/direct/large/ai/AbsLC.scala

@@ -41,7 +41,7 @@ object AbsLamCal {
     type R[+T] = T
     val ρ0 = Map[Ident, Addr]()
     val σ0 = Map[Addr, Value]()
-    def get(ρ: Env, x: Ident): Addr = ρ(x)
+    def get(ρ: Env, x: Ident): Addr = ρ.getOrElse(x, Addr("__somerandomthingthatdoesntexist"))
     def put(ρ: Env, x: Ident, a: Addr): Env = ρ + (x → a)
     def get(σ: Store, a: Addr): Value = σ.getOrElse(a, Lattice[Value].bot)
     def put(σ: Store, a: Addr, v: Value): Store = {
@@ -77,13 +77,13 @@ object AbsLamCal {
 
     def prim_eval(op: String, lv: List[Value]): Value = op match {
       case op if (scala.collection.immutable.Set("+", "-", "*", "/", "%", "vector-length")(op)) => Set(IntV())
-      case op if (scala.collection.immutable.Set("eq?", ">", "<", ">=", "<=")(op)) => Set(BoolV())
-      case op if (scala.collection.immutable.Set("list", "cons", "cdr")(op)) => Set(ListV())
-      case op if ((scala.collection.immutable.Set("car", "vector-ref"))(op)) =>
+      case op if (scala.collection.immutable.Set("eq?", "null?", "pair?", "and", "or", "not", ">", "<", ">=", "<=")(op)) => Set(BoolV())
+      case op if (scala.collection.immutable.Set("list", "cons", "cdr", "cadr")(op)) => Set(ListV())
+      case op if ((scala.collection.immutable.Set("car", "caar", "caddr", "vector-ref"))(op)) =>
         Set(BoolV(), IntV(), FloatV(), CharV(), ListV(), VectorV(), SymV())
       case op if ((scala.collection.immutable.Set("read", "vector", "make-vector", "number->string"))(op)) =>
         Set(VectorV())
-      case op if (scala.collection.immutable.Set("vector-set!", "display", "write", "newline")(op)) => Set(VoidV())
+      case op if (scala.collection.immutable.Set("vector-set!", "display", "write", "newline", "set-cdr!", "error")(op)) => Set(VoidV())
     }
 
     type Config = (Expr, R[Env], R[Store])
@@ -161,7 +161,7 @@ object AbsLamCal {
     implicit def addrTyp: Typ[Addr]
[…]

### 2018-11-16 code/src/main/scala/sai/direct/large/ai/LC.scala

@@ -28,7 +28,7 @@ object LamCal {
     type Env
     type Store
     type Ans = (R[Value], R[Store])
-    val primops = Set("+", "-", "*", "/", "%", "eq?", ">", "<", ">=", "<=", "list", "cons", "car", "cdr", "vector", "make-vector", "vector-set!", "vector-ref", "vector-length", "display", "write", "newline", "read", "number->string")
+    val primops = Set("+", "-", "*", "/", "%", "eq?", "null?", "pair?", "and", "or", "not", ">", "<", ">=", "<=", "list", "cons", "car", "caar", "cadr", "caddr", "cdr", "cddr", "vector", "make-vector", "set-cdr!", "error", "vector-set!", "vector-ref", "vector-length", "display", "write", "newline", "read", "number->string")
     def get(ρ: R[Env], x: Ident): R[Addr]
     def put(ρ: R[Env], x: Ident, a: R[Addr]): R[Env]
     def get(σ: R[Store], a: R[Addr]): R[Value]

### 2018-11-16 code/src/main/scala/sai/direct/large/parser/Transform.scala

@@ -36,8 +36,20 @@ object LargeSchemeASTDesugar {
         }
     }
 
-  def desugarSequence(seq: List[Expr]): Expr =
+  def foldDefine(seq: List[Expr]): (List[String], List[Expr]) =
     seq match {
+      case Nil => (List(), List())
+      case x :: xs =>
+        val (ls, le) = foldDefine(xs)
+        x match {
+          case Define(n, v) => (n :: ls, Set_!(n, v) :: le)
+          case _ => (ls, x :: le)
+        }
+    }
+
+  def desugarSequence(seq: List[Expr]): Expr = {
+    val (ls, le) = foldDefine(seq)
+    val body = le match {
       case Nil => Void()
       case x :: Nil => apply(x)
       case x :: xs =>
@@ -48,6 +60,11 @@ object LargeSchemeASTDesugar {
             newIdentLet(apply(x)) { _ => desugarSequence(xs) }
         }
     }
+    ls.foldLeft(body) {
+      case (b, name) =>
+        App(Lam(List(name), b), List(Void()))
+    }
+  }
 
   def apply(expr: Expr): Expr = expr match {
     case App(e1, param) => App(apply(e1), param map apply)

### 2018-11-17 paper/10_intro.tex
paper
revise

@@
 However, precisely analyzing binding-time for an arbitrary program is hard in general. As an 
-alternative and programmatic approach to specialization and partial evaluation, multi-stage 
-programming (MSP for shot) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage 
+alternative and pragmatic approach to specialization and partial evaluation, multi-stage 
+programming (MSP for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage 
 annotations (i.e., binding-time annotations) from the programmers. These staging annotations 
@@
 \begin{lstlisting}
-def power(b: Int, e: Int): Int = 
+def power(b: Rep[Int], e: Int): Int = 
   if (e == 0) 1 else e * power(b, e-1)
 \end{lstlisting}
-If \texttt{e} is identified as statically known by the programmer, say 5, then we can 
-specialize the function \texttt{power} with respect to \texttt{e = 5} and generate a new 
+If the program identifies that @b@ will be known dynamically (as shown in its type @Rep[Int]@ -- a representation of @Int@) 
+and @e@ is statically known, say 5, 
+then we can specialize the function \texttt{power} with respect to \texttt{e = 5} and generate a new 
 specialized function that runs faster:
@@
 Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach to construct 
-sound static analysis by approximation is almost as old as Futamura projection.
+sound static analysis by approximation is almost as old as Futamura projections.
 Some recent progress such as Abstracting Abstract Machines (AAM) uncovers a methodology to 
@@
 it is natural to raise the question whether it is possible to specialize an abstract interpreter, and how 
-can we effectively do that having considered their functional dissimilarity. 
+can we effectively do that having considered their different functionalities.
 This paper studies the confluence of two old ideas --- Futamura project and abstract interpretation,
@@
   \caption{The confluence of staging and abstraction.}
+  \label{confluence}
 \end{figure}
@@
 Multi-stage programming has been widely used to improve the performance in many domains, 
-such as optimizing compilers and domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
+such as optimizing compilers or domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
 DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
@@
 the abstract syntax tree. 
-However, despite of the previous works mostly focused on one particular analysis, or required to 
-completely rewrite the analyzer,
-it is worth to investigate the idea from a modern perspective based on generative programming,
+However, these previous works focused mostly on one particular analysis, or required to 
+completely rewrite the analyzer. 
+Hence, it is worth to investigate the idea from a modern perspective based on generative programming,
 especially for a general abstract interpreter that models direct-style $\lambda$-calculus, imperative 
@@
 fixed-point iterations, different abstract domains and staging are introduced at the same time.
-In this paper, we present an end-to-end design and implementation of staging an abstract interpreter,
+In this paper, we present an end-to-end design and implementation of staging an abstract interpreter;
 that means not only the interpreter that traverses the abstract syntax tree, 
@@
 due to the scale of the program and the inherent complexity of the algorithm.
-However, analyzing these libraries sometimes are unnecessarily repeated.
+However, analyzing these libraries is sometimes unnecessarily repeated.
 When applying the staged abstract interpreter on such open programs, we leave the unknown arguments and 
@@
 
+This paper makes the following contributions:
+
 \begin{itemize}
@@
     but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
-    programs. 
-    %\todo{more on this point.}
+    programs. \todo{Section X}.
   \item To demonstrate the applicability and usefulness of the proposed approach, we conduct three case studies 
-    in Section~\ref{cases_study}. 
+    Section~\ref{cases_study}: 
     \subitem (1) We revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587} in Section~\ref{cs_ac}, 
@@
     \subitem (2) We demonstrate pushdown control-flow analysis techniques (Section~\ref{cfa}), such as context-sensitive 
-    (\todo{adaptive/flow})
     analysis and more precise stores can be integrated to staged abstract interpreters without losing any precision.
@@
   \item We empirically evaluate the performance improvement by staging an big-step abstract interpreter 
-    for control-flow analysis in Section~\ref{evaluation}. 
-    We compare our staged abstract interpreter with the unstaged version and a small-step abstract interpreter.
+    for control-flow analysis (Section~\ref{evaluation}). 
+    We compare our staged abstract interpreter with the unstaged version.
 \end{itemize}

### 2018-11-17 paper/20_bg.tex
revise

@@
 
-\section{Preliminary}
+\section{Preliminaries}
 
-In this section, we first describe the abstract syntax of the language we will analyze, then present an 
-generic interface of interpreters, and instantiate an unstaged concrete interpreter.
+In this section, we first describe the abstract syntax of the language we will analyze, then present a
+generic interface for interpreters, and instantiate an unstaged concrete interpreter.
 Then, starting from this unstaged concrete interpreter, we proceed in two directions:

### 2018-11-17 paper/21_bg_interp.tex
revise intro
revise

@@
 
-As higher-order functions had been widely introduced to mainstream programming languages (such as Python and Java \todo{cite}), 
-in this paper, we consider a tiny higher-order functional language \TLang, which is based on direct-style $\lambda$-calculus
+We consider a tiny higher-order functional language \TLang, which is based on direct-style $\lambda$-calculus
 with numbers, arithmetics, recursions and conditionals.
-In Section~\ref{cases_imp}, we will see adding more imperative features to the language.
-Since we are mostly interested in analyzing the dynamic behavior of the program, the static semantics and
-type system will be omitted. We also assume the input programs are well-typed and variables are distinct.
-The abstract syntax is described below:
+In Section~\ref{cases_imp}, we will add more imperative features to the language.
+Since we are mostly interested in analyzing the dynamic behavior of the program, we disguised any static semantics and
+type system. We also assume that input programs are well-typed and variables are distinct.
+The abstract syntax is as follows:
 
@@
 
-We give the concrete semantics of \TLang by showing a big-step operational-style interpreter for it.
+We give the concrete semantics of \TLang by showing a big-step definitional interpreter.
 The interpreter is a recursive function that takes the program AST, environment and store, and returns 
-the evaluated value and the accompanied store. The environment will be a mapping from identifiers to addresses,
-and the store is a mapping from addresses to values. The store simulates the heap and is useful when modeling 
-recursion and mutation.
-This environment-and-store-passing style big-step interpreter also can be obtained by 
+the evaluated value and the accompanied store. The environment is a mapping from identifiers to addresses,
+and the store is a mapping from addresses to values. 
+We use the store to model recursion and mutation in concrete semantics; it is also useful for polyvariant analysis.
+This environment-and-store-passing style big-step interpreter is standard and can also be obtained by 
 refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
@@
 \subsection{Generic Interface} 
-Before showing the actual implementation of the concrete semantics, let's first see 
+Before showing the actual implementation of the concrete semantics, let us first consider
 a generic interface for the interpreters, which is semantics-agnostic and can be 
-"interpreted" by either concrete or abstract, unstaged or staged semantics by our choice.
+"interpreted" by either concrete or abstract, unstaged or staged semantics of our choice.
 The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Addr},
 \texttt{Value}, \texttt{Env}, \texttt{Store} and returned type \texttt{Ans}. 
-Initial values for environments and stores, and few abstract methods that manipulate environments and stores are also declared.
+It also declares initial values for environments and stores, and few abstract methods that 
+manipulate environments and stores such as @put@ and @get@.
 
@@
 
-The actual semantics and operations of the interpreter is left to be implemented later, 
-such as lifting literal terms to values (\texttt{close}, \texttt{num}), 
+The actual semantics and operations of the interpreter are left to be implemented later, 
+these cinclude operations such as lifting literal terms to values (\texttt{close}, \texttt{num}), 
 applying a closure (\texttt{apply\_closure}), branching (\texttt{branch0}),
-and arithmetics (\texttt{prim\_eval}). We can choose to implement them concrete or abstract,
+and arithmetics (\texttt{prim\_eval}). We can choose to implement them in a way that is concrete or abstract,
 unstaged or staged.
@@
 
-The implementation of the skeleton interpreter traverses the abstract syntax tree and should be straightforward.
+The implementation of the skeleton interpreter traverses the abstract syntax tree and should be straightforward to understand.
 It is worth noting that the interpreter is written in open-recursive style -- 
@@
 parameter \texttt{ev} referring to itself; this allows us to instrument the interpreter
-from outside. Accordingly, the function \texttt{close} for lifting lambda term to closure values
+from the outside. Accordingly, the function \texttt{close} for lifting lambda term to closure values
 is also written in this style.
@@
 \paragraph{Stage Polymorphism.}
-Readers may notice an abstract higher-kinded type \texttt{R[+T]} is defined in the trait, that is a
-type-level function and will be used to specify the binding-time, i.e., staged or not; the annotation 
-\texttt{+} indicates type parameter \texttt{T} is covariant.
-The LMS framework we use here allows us to specify binding-time purely based on types.
+Readers may notice that an abstract higher-kinded type \texttt{R[+T]} is defined in the trait, i.e. a
+type-level function whici will be used to specify the binding-time, i.e., staged or not; the annotation 
+\texttt{+} indicates that the type parameter \texttt{T} is covariant.
+The LMS framework we use here allows us to specify binding-times purely based on types.
 All occurrences of type \texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} 
@@
 \texttt{type R[+T] = T}. In Section~\ref{stagedinterp}, \texttt{R} will be instantiated 
-as a staging annotation. \todo{where talk about stage poly?}
+as a staging annotation. \todo{Finally tagless, poly embedding}
 
-\subsection{Concrete Instantiation}
+\subsection{Concrete Instantiation} \label{unstaged_conc}
 Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
 Since the interpreter is concrete, for the propose that properly models the heap, the type \texttt{Int} 
-is used for address space \texttt{Addr}. The environment and store are just ordinary \texttt{Map} in Scala.
-The values can be either a tagged number \texttt{NumV} or a closure \texttt{CloV} that contains a lambda term
+is used for address space \texttt{Addr}. The environment and store are just ordinary \texttt{Map}s in Scala.
+A value can be either a tagged number \texttt{NumV} or a closure \texttt{CloV} that contains a lambda term
 and an environment.
@@
 
-The we implement the components that define the semantics. 
-As we mentioned before, the type \texttt{R} is concretize as an identity type function, thus
+we proceed to implement the components that define the semantics. 
+As mentioned before, type \texttt{R} is concretized as an identity type function, thus
 all types are just ordinary types.
@@
 
-\texttt{get} and \texttt{put} directly manipulate the map of environment and store. 
+Finally, \texttt{get} and \texttt{put} directly manipulate the map of environment and store. 
 An allocation function is used when we need to update the store: 

### 2018-11-17 paper/22_bg_abs.tex
revise

@@
 Based on the generic interface we presented in Section~\ref{bg_lang}, 
-now we describe the instantiation of an abstract interpreter.
-To keep the simplicity, we intend to use simply abstract domains, and just establish a 
+we now describe the instantiation of an abstract interpreter.
+To keep the presentation simple, we use simple abstract domains, and just establish a 
 context-/path-/flow-insensitive, store-widened analysis in this section --- indeed, it is coarse
 but simple enough to setup a foundation for the staged abstract interpreter we will present later.
-However, in Section~\ref{cfa}, we will see how to regain context-/path-/flow-sensitivity 
+In Section~\ref{cfa}, we will see how to regain context-/path-/flow-sensitivity 
 for pushdown control-flow analysis; in Section~\ref{cases_imp}, we will see how to instantiate 
 an interval abstract domain for numerical analysis.
-
-\todo{maybe put to related work} other big-step abstract interpreter, 
-big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
-monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
-arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
 
@@
 to abstract semantics. The store now maps addresses to sets of abstract values, meaning that an address points
-to all possible values may occur at runtime, where the abstract value is either a closure or 
-a single abstract number \texttt{NumV}, which stands for the top. The address space is also
-constrained to be finite, in the monovariant setting, we simply use variable names for it.
+to all possible values that may occur at runtime, where the abstract value is either a closure or 
+a single abstract number \texttt{NumV}, which stands for the top element of the number lattice. 
+The address space is also constrained to be finite, in the monovariant setting, we simply use variable names 
+for address.
 
@@
 \subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
-To effectively reuse the code between unstaged and staged part, we also make the lattice structure to
-be stage polymorphic. A lattice type class parameterizes over an element type \texttt{E} and defines
+To effectively reuse the code between unstaged and staged variants, we also make the lattice structure
+stage polymorphic. A lattice type class parameterizes over an element type \texttt{E} and defines
 five operations for \texttt{R[E]}, such as meet, join, and ordering relation; again, \texttt{R} 
@@
 
-For example, here we use powerset as the abstract domain, and the unstaged lattice 
-of powerset can be easily implemented as follows:
+For example, here we use powersets as the abstract domain, and the unstaged lattice 
+of powersets can be easily implemented as follows:
 
@@
 The ordering relation is to ask whether one set is a subset of the other, 
-join is to union the two set, and meet is to intersect the two sets.
+join is to union the two sets, and meet is to intersect the two sets.
 Accordingly, other structures used in the rest of the paper such as tuples 
@@
 the abstract semantics: @get@ uses the bottom element in the @Value@ lattice if the 
-queried address not exists; @put@ performs the update by joining with the existing values.
-@alloc@ simply invokes the @Addr@ constructor.
+queried address does not exist; @put@ performs the update by joining with the existing values.
+function @alloc@ simply invokes the @Addr@ constructor.
 \texttt{close} and \texttt{num} lift syntactic literals to our abstract domain, i.e., 
@@
 
-\todo{conditional, use different store?}
 \begin{lstlisting}
@@
 
-For branching, @branch0@ simply joins the answer from two branches. For arithmetics, @prim_eval@ simply 
-returns the top abstract number element.
-It is worth to note that @apply_closure@ is where we handle the non-determinism of applications.
-The non-determinism happens because of the abstract store maps addresses to sets of abstract values;
+For branching, @branch0@ joins the answers from two branches since we are doing a path-insensitive analysis. 
+For arithmetics, @prim_eval@ returns the top abstract number element.
+It is worth noting that @apply_closure@ is where we handle the non-determinism of applications.
+The non-determinism happens because the abstract store maps addresses to sets of abstract values;
 when dereferencing an address from the store, we need to explore all possible values to achieve a sound 
 analysis. In the case of @apply_closure@, the first argument @fs@ may contains multiple target closures,
-so we have a @for@ comprehension over all closures, apply them respectively and form a set result values @vs@;
-meanwhile, every time when we evaluate into the body expression @e@ we use the latest store $\sigma$@_*@
-which is updated iteratively. Thus it is a store-widened analysis.
+so we use a @for@ comprehension over all closures, apply them respectively and form a set result values @vs@;
+meanwhile, every time when evaluating the body expression @e@ we use the latest store $\sigma$@_*@
+which is updated iteratively during the loop. Thus this is a store-widened analysis.
 
@@
 interpreter may not terminate for some input program. In this section, we use a memoization
-technique to ensure its termination. This technique is also widely used in the name of co-inductive caching algorithm
+technique to ensure its termination. This technique is also widely used under the name of co-inductive caching algorithm
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or truncated depth-first evaluation 
@@
 
-The idea is to set up two caches called @in@ and @out@, which are both mapping from the arguments of interpreter 
-(@Config@) to the returned value of interpreter (@Ans@). The @in@ contains what we already computed from the last
-iteration, the @out@ cache is what we computed after this iteration joined with the previous one.
-When starting a new iteration, the @in@ is set to be the result of last iteration, i.e., @out@; the @out@ is set
-to be empty. In the iteration, we first ask whether @out@ contains what we want, if yes then it is returned;
-otherwise, we retrieve what we have from @in@, compute the result for this time, and put the joined result back to @out@.
-Note that we also instrument the recursive call by putting @cached_ev@ to @evev@.
-After one iteration, if @in == out@, then there is no more information can be gained, thus the iteration should 
+The idea is to set up two caches called @in@ and @out@, which are both mapping from the arguments of the interpreter 
+(@Config@) to the result values of the interpreter (@Ans@). The @in@ cache contains what we already have 
+computed from the last
+iteration, the @out@ cache is what we will have computed after this iteration joined with the previous one.
+When starting a new iteration, @in@ is set to be the result of the last iteration, i.e., @out@; @out@ is set
+to be empty. In the iteration, we first check whether @out@ contains what we want, if yes then it is returned;
+otherwise, we retrieve what we have from @in@, compute the result for this iteration, 
+and put the joined result back into @out@.
+Note that we also instrument the recursive call by putting @cached_ev@ into @evev@.
+After one iteration, if @in == out@, then there is no more information to be gained, thus the iteration should 
 end and we have reached the fixed point.

### 2018-11-17 paper/23_bg_stage.tex
revise

@@
 as a Scala library that enables dynamic code generation in a type-safe manner.
-Different from the syntactic approach such as MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations, 
+Different from the syntactic approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations, 
 LMS distinguishes binding-time solely based on types.
 LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an expression will be 
-known at next stage. All operations acting on a @Rep@ expression will be resulted in the generated code. However,
+known at next stage. All operations acting on a @Rep[T]@ expression will be residualized in the generated code. However,
 we need to provide intermediate representation and code generation support for type @T@ to LMS. 
@@
 
-In the code shown above, \texttt{power} takes 2 arguments where \texttt{b} is declared as \texttt{Rep[Int]} type 
+In the code shown above, \texttt{power} takes two arguments where \texttt{b} is declared as \texttt{Rep[Int]} type 
 meaning that \texttt{b} is a representation of \texttt{Int} but whose value will be available at the next stage. 
@@
 In the staging part, we share the same concrete type instantiation, 
-i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations.
-Meanwhile, @RepConcInterpOps@ also extends from @LMSOps@ which provides necessary supports for the 
-staged version of primitive types and data structures. \todo{what are static, what are dyanmic?}
+i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations
+in trait @RepConcInterpOps@.
+Meanwhile, it is also extended from @LMSOps@ which provides necessary supports for the 
+staged version of primitive types and data structures.
+To stage the interpreter, we first identifies the syntactic input is known statically, 
+and the interpreter returns a pair of values and store, which both becomes dynamic values.
+Thus the staged stores propogate the binding-time to environments, because they share the same
+address component.
 
@@
 
-The first notable change is the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
-types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. For @num@, we use @lift@ function which is a
+The first notable change is that the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
+types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. For @num@, we use the @lift@ function which is a
 built-in in LMS that lifts a current-stage value to next stage.
-The interesting thing is how we handle closures and function applications when staging is involved.
+The interesting point is how we handle closures and function applications when staging is involved.
 In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, which
@@
 eliminated and specialized away.
-The specialization of the interpreter with respect to the body of the lambda term proceed under the
-assumptions that the argument and latest store will be provided later. \note{to denotational style?, back to HO?}
-We will see this is an important observation that enables us to specialize an abstract interpreter modularly.
-In the end of @close@, we create a @CompiledClo@ object that contains a staged function, which all will be
-exhibited in the generated code.
+The specialization of the interpreter with respect to the body of the lambda term proceeds under the
+assumptions that the argument and latest store will be provided later. 
+%\note{to denotational style?, back to HO?}
+We will see this is an important observation that enables us to specialize an abstract interpreter in a modular way.
+At the end of function @close@, we create a @CompiledClo@ object that contains a staged function, which all will be
+emitted in the generated code.
 
@@
 
-Then, @apply_closure@ takes three arguments the function value, argument value and the latest store, 
-which are both of @Rep@ type. It means basically there is nothing we can do at current stage --- 
+Next, @apply_closure@ takes three arguments: the function value, argument value and the latest store, 
+which are both of @Rep@ type. This means basically there is nothing we can do at the current stage --- 
 so we create a new node @ApplyClosure@ that contains these arguments in the intermediate representation graph,
-and in the code generation phase we can emit the code that does the application. \todo{reflectEffect}.
+and in the code generation phase we can emit the code that performs the application. The additional @reflectEffect@
+is a function in LMS that handles side effects.
 
@@
 \paragraph{Code Generation.}
-When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as what they are because
-they are intended to be existed in the next stage; however, we need to treat @ApplyClosure@ a little bit 
-differently.
+When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as they are because
+they are intended exist in the next stage; however, we need to treat @ApplyClosure@ a little differently.
 
@@
 
-Concretely, as shown in the code above, when an @ApplyClosure@ is matched when emitting code, we just 
-emit a function application where the function value is extracted from the @CompiledClo@
-(we assume the object program is well-typed so that we can safely cast it to @CompiledClo@), 
-and argument and store are the rest fields in @ApplyClosure@.
+Concretely, as shown in the code above, when emitting code for @ApplyClosure@, we just 
+emit a function application where the function value is extracted from the @CompiledClo@ object
+(we assume the input program is well-typed so that we can safely cast it to @CompiledClo@).
+Argument and store are the remaining rest fields in @ApplyClosure@.

### 2018-11-17 paper/70_related.tex
revise

@@
 
+\todo{maybe put to related work} other big-step abstract interpreter, 
+big-step abstract interpreter using delimited control operators \cite{Wei:2018:RAA:3243631.3236800},
+monadic big-step abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17},
+arrow-based abstract interpreter \cite{Keidel:2018:CSP:3243631.3236767}.
+
 \textbf{Meta-Programming}.

### 2018-11-17 paper/60_eval.tex
write evaluation

--- 2018-10-18 paper/60_eval.tex
@@
-\section{Evaluation}
+\section{Evaluation} \label{evaluation}
+
+We have designed, implemented and evaluated a prototype abstract interpreter for Scheme.
+In order to compare results, our prototype includes both staged and unstaged versions of
+the abstract interpreter doing the same control flow analysis.
+
+Our evaluations have shown that the staged abstract interpreter performs well in general.
+Particularly in some interesting cases that are considered to be the worst case scenario
+in the control flow analysis, the staged version outperforms the unstaged abstract 
+interpreter by a wide margin.
+
+However, there also exists some programs where the unstaged abstract interpreter yields 
+results faster than the unstaged interpreter.
 
@@
 
+Our evaluation is based on a suite of benchmark tests. Considering that our abstract 
+interpreters are implemented in Scala which will be affected by JVM warm up times, programs
+tend to run faster after a number of runs. In order to minimize the influence of JVM, 
+we ran all experiments 10 times and take the statistical mean and median values of 
+the running times.
+
+%on a suite of benchmark scheme programs that are also used in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}. 
+We used the following benchmark programs:
+\begin{itemize}
+    \item \textbf{fib:} a program that recursively calculates the $n$-th fibonacci number in
+        exponential time.
+    \item \textbf{church:} a program introduced by \todo{cite Dimitrios Vardoulakis and Olin Shivers. CFA2: a Context-Free approach
+        to Control-Flow analysis. Logical Methods in Computer Science,
+        7(2), 2011.} to test distributivity of multiplication for numbers in Church Encoding \todo{ cite }
+\end{itemize}
+
 \subsection{Performance}
+We performed our 
+For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
 

### 2018-11-17 paper/60_eval.tex
more eval

@@
 \subsection{Performance}
-We performed our 
+All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux 4.4.0) machine 
+with 4 Intel(R) Xeon(R) Platinum 8168 CPUs running at 2.7GHz with a total of 3 TiB of RAM.
+Although we were using an environment with 96 cores and 192 threads, we used only
+one single thread for all the benchmark programs we run.
+
 For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
+\todo{Figure show data.}
 

### 2018-11-17 paper/30_sai.tex
revise

@@
 
-In the previous section, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
+In the previous sections, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
 now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
 We present a principled approach to derive staged abstract interpreter
-from its unstaged version. One guiding principle of our approach is the code of abstract semantics
-and the code that optimizes should be separated. Thus it is also an advantage of doing staging 
-for abstract interpreters: the designer of analysis has no need to rewrite the analysis, and 
+from its unstaged version. One guiding principle of our approach is that the code of the abstract semantics
+and the code that optimizes should be separated. This it is an advantage of using staging 
+for abstract interpreters: the designer of the analysis has no need to rewrite the analysis, and 
 the performance improvement comes almost for free, without any sacrifice of soundness or precision.
 Unsurprisingly, the staged abstract interpreter we present in this section has the same abstract 
-semantics with the unstaged version we presented in Section~\ref{unstaged_abs}.
+semantics as the unstaged version we presented in Section~\ref{unstaged_abs}.
 
@@
 In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@
-to leave space for staging lattices, now we instantiate @R@ to @Rep@ and still use powerset as an 
-example to present its staged version.
+to leave space for staging lattices, now we instantiate the type @R@ to @Rep@ and 
+still use powersets as an example to present its staged version.
 
@@
 
-The type parameter @T:Typ@ of @RepLattice@ requires the element of sets
-is also can be staged, otherwise without knowing how to stage the elements in the set
-we can not stage the set as well.
+The type parameter @T:Typ@ of @RepLattice@ requires that the elements of sets
+is can also be staged. Otherwise, without knowing how to stage the elements in the set,
+we can not stage the set either.
 The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
-such as @union@ and @intersect@ will be mapped to a node in IR graph and emitted in 
-the generated code. Again, other structures such as maps and tuples are also implemented
-in a similar way.
+such as @union@ and @intersect@ will be mapped to a node in the IR graph during staging and 
+eventually emitted in the generated code. 
+Again, other structures such as maps and tuples are implemented in a similar way.
 
 \subsection{Staged Abstract Semantics} 
-We have seen how to obtain a staged concrete semantics
-based on types, now we take the same approach to obtain a staged abstract semantics.
-\todo{what are static, what are dynamic?}
-The basic operations largely kept same as the unstaged version, except the types are changed to @Rep@.
-Besides, when we update the environment, the identifier @x@ is known statically, but the map of environment
-has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to make it as a next-stage value.
+We have seen how to obtain a staged concrete semantics based on types, 
+now we take the same approach to obtain a staged abstract semantics.
+The basic operations are largely kept the same as in the unstaged version, except the types are changed to @Rep@.
+Besides, when we update the environment, the identifier @x@ is known statically, but the environment map
+has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to turn it as a next-stage value.
 
@@
 
-Once more, the way we handle closures is as same as the staged concrete interpreter:
+Once more, the way we handle closures is the same as in the staged concrete interpreter:
 the recursive call to @ev@ with the body expression @e@ is compiled and specialized, 
-the wrapper function @f@ and will be a field value of @CompiledClo@ and be generated
-for the next stage. At the end, we return a singleton set.
+the wrapper function @f@ will be a field value in a @CompiledClo@ object and be generated
+for the next stage. At the end, we return a singleton set:
 
@@
 
-When generating code for application, we can not directly apply the callee.
+When generating code for an application, we can not directly apply the callee.
 Instead, we emit code that calls a next-stage function @apply_closures_norep@.
-Function @apply_closures_norep@ does the same job as its unstaged version,
-which non-deterministically applies multiple target closures with the argument
+As its unstaged counterpart, function @apply_closures_norep@ 
+non-deterministically applies multiple target closures with the argument
 and latest store, and finally returns the joined value and a single store.
@@
 
-Our fixed-point iteration relies two caches @in@ and @out@, but the iteration 
-can not be done at the current stage.
-Because the @in@ and @out@ are both next-stage values, thus the test of whether 
-@in@ and @out@ are equal is a generated expression to the next stage, and we can only 
+Our fixed-point iteration again relies on two caches @in@ and @out@, but the iteration 
+no longer be done at the current stage.
+Because the @in@ and @out@ are both next-stage values, the test of whether 
+@in@ and @out@ are equal is a generated expression in the next stage, and we can only 
 know the comparison result at the next stage. 
@@
 
-However, the instrumented evaluation function that uses @in@ cache and updates @out@ cache
+However, the instrumented evaluation function that uses the @in@ cache and updates the @out@ cache
 can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
@@
 
-Now we already obtained an end-to-end staged abstract interpreter that is able to 
+Now we have already obtained an end-to-end staged abstract interpreter that is able to 
 specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
 which means any operations on a @Map@ becomes code in the next stage.
-But, as we identified before \todo{where}, the keys of environment are identifiers in the program,
-which are completely known statically. This leaves us a chance to further specialize the 
-data structures. Particularly, if we are specializing a monovariant analysis, the address space 
+But, as we identified when introducing the generic interface, the keys of any environment maps are 
+identifiers in the program, which are completely known statically. 
+This leaves us a chance to further specialize the data structures. 
+Assume the @Map[K,V]@ is implemented as a hash map, if the keys $K$ are known, then the indices can be computed
+statically. Thus the specialized map would be an array @Array[Rep[V]]@ whose elements 
+are next-stage values; all the accesses to the array is determined during staging.
+
+Particularly, if we are specializing a monovariant analysis, the address space 
 is equivalent to the identifiers, then the environment component can be entirely eliminated, 
-and the store is a partially specialized map.
-\todo{maps to arrays}
+and the store is a specialized map as array of @Rep[Value]@ elements.
 
@@
 %We have presented all of our four different interpreters as shown in Figure~\ref{confluence}.
-In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} gave several
-advice on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
-We borrow these advice to abstract interpreters and discuss decisions we made to achieve this 
-and examine some alternatives.
+In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} provided guidelines
+on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
+We borrow these guidelines and extend them to abstract interpreters. We discuss decisions we made to 
+achieve this and examine some alternatives.
 
@@
 What we implemented is a big-step, compositional \todo{or semi-compositional} abstract interpreter, 
-where the "compositional" means every recursive call of our abstract interpreter is applied with proper 
+where "compositional" means that every recursive call of our abstract interpreter is applied to proper 
 substructures of the current syntactic parameters \cite{10.1007/3-540-61580-6_11}.
-The compositionality ensures the specialization can be done by unfolding, as well as the termination of specialization procedure.
-Nevertheless, it is also possible to specialize a small-step, operational abstract semantics --- \citeauthor{Johnson:2013:OAA:2500365.2500604}
+This compositionality ensures that specialization can be done by unfolding,
+as well as that the specialization procedure terminates.
+Nevertheless, it is also possible to specialize a small-step operational abstract semantics ---
+\citeauthor{Johnson:2013:OAA:2500365.2500604}
 showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimizations
@@
 \paragraph{Correctness and Soundness.}
-Based on the assumption that LMS preserves the semantics during staging, and indeed it is 
-in our application, 
+Based on the assumption that LMS preserves the semantics during staging,
 we are confident that the staged abstract interpreter does the same analysis compared 
 with the unstaged one. Moreover, the optimization done by staging does not compromise
-and soundness.
+any soundness.

### 2018-11-17 paper/50_cases.tex
revise

--- 2018-11-15 paper/50_cases.tex
@@
 
-In Section~\ref{sai}, we have shown staging an abstract interpreter is feasible and can be 
+In Section~\ref{sai}, we have shown that staging an abstract interpreter is feasible and can be 
 systematic after correctly identifying the binding-times, despite the fact that the abstract 
 interpreter is intended to be imprecise and easy to implement. 
-In this section, we conduct several case studies to show that it is also useful and widely
-applicable to different analysis.
+In this section, we conduct several case studies to show that this approach is also useful and
+widely applicable to different analyses.
 
@@
 implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-The idea is inspired by partial evaluation and similar with the presented paper -- the program can be known 
+The idea is inspired by partial evaluation similar to the present paper -- the program can be known 
 statically therefore the overhead of interpretation can be eliminated. 
 In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
-the closures can be executed immediately however the textual program need to be loaded firstly.
+the closures can be executed immediately however the textual program need to be compiled and loaded first.
 
@@
 Closure generation compiles the analysis as a closure which takes an environment as argument.
-And the overhead of traversing the abstract syntax tree of input program also has been eliminated.
+The overhead of traversing the abstract syntax tree of input program also has been eliminated.
 
@@
 understood and implemented as an instance of staging abstract interpreters.
-We firstly revisit the original implementation of abstract compilation of 0-CFA,
+We first revisit the original implementation of abstract compilation of 0-CFA,
 and then reproduce their result by simply adding stage annotations.
@@
 but without changing a single line of the analyzer program (with the use of LMS).
-However, closure generation requires more engineering effort, such as a whole-program 
+However, closure generation requires more engineering effort, specifically a whole-program 
 conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
@@
 of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
-structs are decomposed to different functions, such as @analyzeCall@ and @analyzeApp@. The idea of closure
-generation is to rewrite these functions, where previously they may take both static arugments and dynamic arguments,
-but after the rewrite only the static arguments is taken. In this case, the static arguments are syntactic terms;
-the dynamic arguments are stores. After written in AC style, the functions like @compCall@ (compiled version of @analyzeCall@) 
-returns a value of type
-@CompAnalysis@, i.e., a closure that takes a store and returns a store. The result of multiple calls on such functions, for example,
-@compCall@ and @compArgs@ can be composed. The generated closure only takes stores, because the input program
+constructs are decomposed into different functions, such as @analyzeCall@ and @analyzeApp@.
+The idea of closure generation is to rewrite these functions. Where previously they may take both static
+arguments and dynamic arguments, after the rewrite only the static arguments is taken.
+In this case, the static arguments are syntactic terms;
+the dynamic arguments are stores.
+After written in AC style, functions like @compCall@ (compiled version of @analyzeCall@) return a value of type
+@CompAnalysis@, i.e., a closure that takes a store and returns a store.
+The result of multiple calls on such functions, for example,
+@compCall@ and @compArgs@, can be composed.
+The generated closure only takes stores, because the input program
 is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).
@@
 the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
-at the next stage. Indeed, a amount of additional engineering efforts are required to make this happen, 
+at the next stage. Indeed, additional engineering efforts are required to make this happen, 
 including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
 current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
-functions of AST structs. We consider these efforts are relatively small, and does not interfere the actual analysis
+functions of AST structs. We consider these efforts relatively small, and they do not interfere the actual analysis
 we desire.
@@
 The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
-One fundamental analysis task for functional programs is the control-flow analysis, i,e., determining
+One fundamental analysis task for functional programs is control-flow analysis, i,e., determining
 which functions will possibly be applied at each call-site. 
-The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is already 
-a store-widened 0-CFA-like abstraction; in last section, we also reviewed AC with 0-CFA.
+The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 
+0-CFA-like abstraction, moreover it is also a pushdown control-flow analysis; 
+in the last section, we also reviewed AC with finite-state 0-CFA.
 In this section, based on the existing staged abstract interpreter, we further
-develop the staging techniques with control-flow analyses, including recoverying a more precise store model;
-different polyvariant analsis, and using staging as a implementation stategy for mixed sensitivity.
+develop the staging techniques with control-flow analyses, including recoverying a more precise store model and
+adding context-sensitivity to the analysis.
+%, and using staging as an implementation stategy for mixed sensitivity.
 
@@
 
-\subsubsection{Context/Path/Flow-Sensitivity}
+The store-widened analysis we implemented in Section~\ref{sai} uses a single store to approximate
+the runtime store. It can be efficently computed in polynomial time together with 0-CFA-like abstraction, 
+but sometimes we may desire a 
+more precise result that distinguishes the final values and stores for different closure targets. 
+To achieve this goal,
+we need to tweak our abstract interpreter and type instantiation. The answer type @Ans@ is changed to
+a set of @VS@s where a @VS@ is a pair of sets of abstract values (such as closures or abstract numbers)
+and a store.
 
-\subsubsection{Mixed Sensitivity}
+\begin{lstlisting}
+type VS = (Set[AbsValue], Store)
+type Ans = R[Set[VS]]
+\end{lstlisting}
 
-using different $k$ for $k$-CFA
+Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under staging the type @Ans@
+represents a next stage value. Then, our generic interpreter is also changed when handling function 
+applications.
+
+\begin{lstlisting}
+case App(e1, e2) =>
+  val e1ans = ev(e1, $\rho$, $\sigma$)
+  val e1vs  = choices(e1ans)
+  val e2ans = ev(e2, $\rho$, e1vs.$\sigma$)
+  val e2vs  = choices(e2ans)
+  apply_closure(ev)(e1vs.v, e2vs.v, e2vs.$\sigma$)
+\end{lstlisting}
+
+The idea to handle the application is to explore all possible closures from @e1@ and meanwhile
+use the latest store.
+What @choices@ does is similar to McCarthy's @amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
+element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side receiver, i.e., @e1vs@.
+The function @choices@ internally uses the delimited control operator @shift@ to capture the continuation,
+which is the code block after its call site. 
+This allows us to perform nesting depth-first evaluation for function application while still writing 
+the program in direct-style \cite{Wei:2018:RAA:3243631.3236800}. 
+Again, we can stage this part as we did for the naive abstract interpreter.
+
+\subsubsection{Context-Sensitivity}
+
+We add k-CFA-like context-sensitivity to the analysis by introducing an abstract timestamp, 
+whose concrete instantiation is a finite list of expressions that track $k$ recent calling contexts.
+The definition of abstract addresses is changed to a tuple of identifiers and the time it get allocated,
+meaning that this address points to some values under such calling context.
+If $k$ is 0, we obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain
+a family of analysis with increasing precision.
+
+\begin{lstlisting}
+type Time = List[Expr]
+\end{lstlisting}
+
+Every time when we call the @eval@ function, we refresh the timestamp by calling function @tick@,
+which returns a new timestamp. Here we adopt a $k$-CFA-like allocation strategy, therefore
+the @tick@ function can be implemented as appending the current expression being evaluated to the existing
+calling context, and then taking the first $k$ elements from the list.
+
+\begin{lstlisting}
+def tick(e: Expr, $\tau$: R[Time]): R[Time] = (e :: $\tau$).takewhile
+def eval(ev: EvalFun)
+  (e: Expr, $\rho$: R[Env], $\sigma$: R[Store], $\tau$: R[Time]): Ans = {
+    val $\tau$_* = tick(e, $\tau$)
+    e match {
+      case Lit(i) => ...
+      ...
+    }
+  }
+\end{lstlisting}
+
+Accordingly, the type of return value is accompanied by the timestamp. For a recursive call of @ev@,
+which would also return the latest timestamp, and that timestamp will be used for the evaluation afterward.
+
+\begin{lstlisting}
+type VST = (Value, Store, Time)
+type Ans = R[Set[VST]]
+\end{lstlisting}
+
+Using other allocation strategies to achieve different sensitivities is also possible \cite{DBLP:conf/icfp/Gilray0M16} 
+and can be staged under our framework.
+
+%\subsubsection{Mixed Sensitivity}
+
+%using different $k$ for $k$-CFA
 
@@
 
-Beyond control-flow analysis, staging data structures used in the abstract interpreter, such as abstract domain.
+Now we consider in a first-order imperative language, we may care more about the data-flow because the control-flow
+is relatively easy to obtain. In this section, we show the staging of other abstract domains, particularly an 
+interval domain for numbers.
 It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
 improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
-In this section, we present a similar idea for interval abstract domain using an \textit{offline} fashion by staging.
+In this section, we first show how to support imperative language features in the generic abstract interpreter.
+Then we present a similar idea for the interval domain and show that the specialization is feasible by staging systematically.
 
@@
 
-Assignments, Loops, Exceptions
+To evaluates an assignment, we first evaluate its right-hand side, and then put the value
+into the slot where the address of @v@ points to.
+For simplicity, we elect to make the value of the assignment be an @void()@ value.
 
-\subsubsection{Specializing Interval Domains}
+\begin{lstlisting}
+case Assign(x, e) =>
+  val (v, $\sigma$_*) = ev(e, $\rho$, $\sigma$)
+  (void(), put($\sigma$_*, get($\rho$, x), v))
+\end{lstlisting}
 
-Intervals, Numerical Analysis
+To evaluates a @while@ loop statement, we evaluate the condition first.
+Then similar to how we treat @branch0@, we have a generic @branch@ function but works on boolean
+values. For the @true@ branch, we recursively call @ev@ on a newly constructed expression
+@Seq(e, While(t, e))@ meaning that first evaluates @e@ and the repeat the loop.
+Otherwise, for the other branch, we simply return a void value and current store.
 
-Options: Data-flow analysis, taint analysis
+\begin{lstlisting}
+case While(t, e) =>
+  val (tv, t$\sigma$) = ev(t, $\rho$, $\sigma$)
+  branch(tv, ev(Seq(e, While(t, e)), $\rho$, t$\sigma$), (void(), t$\sigma$))
+\end{lstlisting}
+
+\subsubsection{Staged Interval}
+
+\begin{lstlisting}
+case class Interval(lb: Rep[Double], ub: Rep[Double]) {
+  def +(that: Interval): Interval = that match {
+    case Interval(lb_, ub_) => Interval(lb + lb_, ub + ub_)
+  }
+  def -(that: Interval): Interval = that match {
+    case Interval(lb_, ub_) => Interval(lb - lb_, ub_ - ub) 
+  }
+  ...
+}
+\end{lstlisting}

### 2018-11-17 paper/70_related.tex
revise

--- 2018-11-16 paper/70_related.tex
@@
 The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}. 
+\citet{Keidel:2018:CSP:3243631.3236767} demonstrate a similar implementation of definitional abstract interpreter using arrows.
 
-\textbf{Meta-Programming}.
-
-Partial evaluation. 
-Reflexive language \cite{DBLP:conf/gpce/Asai14}
-Collapsing Towers of Interpreters \cite{Amin:2017:CTI:3177123.3158140}
-
-The Lightweight Modular Staging framework in Scala
+\paragraph{Meta-Programming}
+The Lightweight Modular Staging framework relies on types as stage annotations, and the staging
+procedure is modular.
 Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
@@
 still applies with other MSP implementations.
+Partial evaluation as an automatic technique is studied comprehensively
+\cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
+Futamura projections reveals close relations between interpreters and compilers \cite{Futamura1999, futamura1971partial}.
+\citet{DBLP:conf/gpce/Asai14} studies the compilation of reflective language using MSP and MetaOCaml.
+\citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of concrete interpreters by MSP and LMS framework.

### 2018-11-17 paper/80_concl.tex
revise

\section{Conclusion}

In this paper, we show the feasibility of using multi-staging programming as a systematic approach 
to optimize abstract interpreters, while minimally modifying the abstract interpreter and keeping
the analysis sound.
We presented an end-to-end design and implementation
of an staged abstract interpreter for a small higher-order stateful language, which 
models the core part of modern programming languages.
The design that uses abstracting definitional interpreters captures the commonality between
abstract interpreters and concrete interpreters; the implementation that uses staging polymorphism
captures the commonality between unstage programs and staged programs.

### 2018-11-17 paper/60_eval.tex
revise

--- 2018-11-17 paper/60_eval.tex
@@
 
-We have designed, implemented and evaluated a prototype abstract interpreter for Scheme.
-In order to compare results, our prototype includes both staged and unstaged versions of
-the abstract interpreter doing the same control flow analysis.
+To evaluate the performance can be improved through staging, we use the abstract
+interpreter demonstrated in the paper to analyze some Scheme programs. We implement
+a desugaring transformation for a large subset of scheme that transforms into the
+small language we used in the paper.
+We use the same generic interface with a 0-CFA-like abstraction, and the unstage abstract
+interpreter forms the baseline.
 
-Our evaluations have shown that the staged abstract interpreter performs well in general.
-Particularly in some interesting cases that are considered to be the worst case scenario
-in the control flow analysis, the staged version outperforms the unstaged abstract 
-interpreter by a wide margin.
-
-However, there also exists some programs where the unstaged abstract interpreter yields 
-results faster than the unstaged interpreter.
+All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux kernel 4.4.0)
+machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM.
+Although the machine has 96 cores and 192 threads in total,
+the abstract interpreters only use one thread to run all the benchmark programs.
 
@@
 \subsection{Performance}
-We performed our 
+
 For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
+\todo{Figure show data.}
 
-\subsubsection{Baselines}
-Small-step AAM
+Our evaluations have shown that the staged abstract interpreter performs well in general.
+Particularly in some interesting cases that are considered to be the worst case scenario
+in the control-flow analysis, the staged version outperforms the unstaged abstract 
+interpreter by a wide margin.
 
-Big-step ADI
-
-OAAM with AC \cite{Boucher:1996:ACN:647473.727587, Johnson:2013:OAA:2500365.2500604}
-
-Benchmarks and other reference implementations: \url{https://github.com/ilyasergey/reachability}.
-
-\subsection{Compositionality}
+However, there also exists some programs where the unstaged abstract interpreter yields 
+results faster than the unstaged interpreter.

### 2018-11-17 paper/10_intro.tex
revise

@@
 
+\iffalse
 It has been observed that partially applying context-sensitivity on selected portion of the program could
@@
 We show that staging abstract interpreters as an approach to effectively implement hybird context-sensitivity\todo{}.
+\fi
 
@@
     %We later extend the staged abstract interpreter to support imperative language features (Section~\ref{cases_imp}), such as states and loops.
-  \item We demonstrate that the staged abstract interpreter not only improves the \textit{efficiency}
-    but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
-    programs. \todo{Section X}.
+  \item As shown in Section~\ref{sai}, we demonstrate that if we apply the abstract interpreters on open 
+    programs, it not only improves the \textit{efficiency} but also the \textit{scalability}, 
+    which are two major issues in static analysis.
   \item To demonstrate the applicability and usefulness of the proposed approach, we conduct three case studies 

### 2018-11-17 paper/30_sai.tex
revise

@@
 
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later (like Facebook Infer). But to my knowledge, they are mostly too conservative (context-insensitive) 
-or unsound, which both lead to imprecision.
+One of the challenges of modern static analysis is program usually depends on
+large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. 
+Can we analyze programs and libraries separately and reuse the result without losing precision? 
+Then we can reduce part of the overhead of repeatedly analyzing libraries for different programs.
+Indeed, some static analyzers compute summary for a function or a module, which can be reused
+later, however they are mostly too conservative or unsound, which both lead to imprecision.
 
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis,
-because it is interprocedural and need the last k calling contexts to distinguish
-different call sites.
-But can we analyze programs (libraries) separately which generate the specialized 
-analysis and leave the unavailable programs (for the moment) as dynamic parameters, 
-and then install these contexts when we have the whole program.
+The specialization of abstract interpreter provides a chance to obtain such partial analysis 
+result in a mechanized way, but still keeps the analysis sound.
+As we see when compiling the closures, we can specialize the abstract interpreter
+with respect the body expression of the lambda term without knowing the actual argument.
+The programs with some unknown variables are open programs, which is exactly the case 
+if we want to analyze programs in a modular way.
 
+For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a whole-program analysis,
+because it is inter-procedural and needs the last $k$ calling contexts to distinguish
+different call sites, where the calling contexts are dynamic values.
+However, it is possible to analyze programs (libraries) separately through specializing an abstract interpreter 
+that generates the analysis as the next-stage program and leave the unavailable programs and calling contexts 
+as dynamic parameters, and then install these contexts when we have the whole program.
+
+\iffalse
 Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
@@
 the abstract interpreter itself might be a partial abstract interpreter.
+\fi
 
@@
 \paragraph{Big-step vs Small-step.}
-What we implemented is a big-step, compositional \todo{or semi-compositional} abstract interpreter, 
+What we implemented is a big-step, compositional abstract interpreter, 
 where "compositional" means that every recursive call of our abstract interpreter is applied to proper 

### 2018-11-17 paper/abstract.tex
abstract

@@
 %
-More surprisingly, we show that by applying the staged abstract interpreter to \textit{open} programs and considering the free variables as dynamic inputs, 
+More surprisingly, we show that by applying the staged abstract interpreter to \textit{open} 
+programs and considering the free variables as dynamic inputs, 
 we obtain a modular analysis that generates sound partial analysis results which can be composed 
@@
 \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of 0-CFA, pushdown
-control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
+control-flow analysis with context-sensitivity and precise stores, and a numerical
 analysis on an imperative language.
@@
 We empirically evaluate the performance improvements on control-flow analysis of benchmark programs.
-The results show an average speedup of Nx with staging for a monovariant analysis, and Mx for 
-a polyvariant analysis.
+The results show speedups up to 2.3x with staging on a monovariant analysis.
 

### 2018-11-17 paper/50_cases.tex
interval
revise paper

@@
   }
-def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): Rep[Store] = 
-  f match {
+def analyzeApp(f: Expr, args: List[Expr], $\sigma$: Rep[Store]): 
+  Rep[Store] = f match {
     case Var(x) => analyzeAbsApp(args, $\sigma$(x), $\sigma$)
@@
     case Nil => $\sigma$
-    case Lam(vars, body)::rest => analyzeArgs(rest, analyzeCall(body, $\sigma$))
+    case Lam(vars, body)::rest => 
+      analyzeArgs(rest, analyzeCall(body, $\sigma$))
     case _::rest => analyzeArgs(rest, $\sigma$)
@@
 
+Interval domain is a relative simple domain which consists of two numeric fields, an upper bound 
+and a lower bound. Here we annotate the upper bound and lower bound fields to be of type @Rep[Double]@.
+The operations such as @+@ of two intervals produce a new
+
 \begin{lstlisting}
@@
   }
-  def -(that: Interval): Interval = that match {
-    case Interval(lb_, ub_) => Interval(lb - lb_, ub_ - ub) 
-  }
   ...

### 2018-11-17 paper/60_eval.tex
revise paper

@@
 To evaluate the performance can be improved through staging, we use the abstract
-interpreter demonstrated in the paper to analyze some Scheme programs. We implement
+interpreter demonstrated in the paper to analyze some Scheme programs. We first implement
 a desugaring transformation for a large subset of scheme that transforms into the
@@
 the abstract interpreters only use one thread to run all the benchmark programs.
+
+Our evaluations have shown that the staged abstract interpreter performs well in some
+interesting cases that are considered to be the worst case scenario in the control flow
+analysis, the staged version outperforms the unstaged abstract interpreter by a wide
+margin (up to 3x performance gain).
 
@@
         7(2), 2011.} to test distributivity of multiplication for numbers in Church Encoding \todo{ cite }
+    \item \textbf{kcfa3/kcfa-worst-case-n:} Benchmark programs that are supposed to
+      be tough cases for CFA.
+    \item \textbf{fermat:} Fermat and Solovay-Strassen primality testing in Scheme by
+      Matthew Might.
+    \item \textbf{rsa:} The RSA public key encryption algorithm in Scheme.
 \end{itemize}
@@
 
-Our evaluations have shown that the staged abstract interpreter performs well in general.
-Particularly in some interesting cases that are considered to be the worst case scenario
-in the control-flow analysis, the staged version outperforms the unstaged abstract 
-interpreter by a wide margin.
+\begin{figure}[h]
+\begin{tabular}{@{}llll@{}}
+\toprule
+    Program            & unstaged   & staged     & $\Delta$ of median \\ \midrule
+    fib                & 1.020      & 2.046      & -50.14\%          \\
+    rsa                & 35.928     & 145.306    & -75.27\%          \\
+    kcfa3              & 2.934      & 5.671      & -48.27\%          \\
+    church             & 884.800    & 545.736    & +62.13\%          \\
+    fermat             & 15.673     & 68.790     & -77.22\%          \\
+    kcfa-worst-16      & 320.689    & 301.564    & +6.34\%           \\
+    kcfa-worst-32      & 3,685.615  & 1,854.461  & +98.74\%          \\
+    kcfa-worst-64      & 49,422.666 & 14,849.725 & +232.82\%         \\
+    \bottomrule
+\end{tabular}
+\caption{Evaluation result.} \label{evaluation_result}
+\end{figure}
+
+Figure~\ref{evaluation_result} shows the benchmark evaluation results. The "unstaged" and "staged" columns
+show the median time to finish the abstract interpretation (in milliseconds). The "$\Delta_{median}$"
+column shows the performance change from the unstaged abstract interpreter to the staged version.\\
 
@@
 results faster than the unstaged interpreter.
+From the data shown above, we observed that 
+However, in general, the unstaged abstract interpreter yields results faster
+than the unstaged interpreter.

### 2018-11-17 paper/60_eval.tex
evaluation

@@
 
-Our evaluations have shown that the staged abstract interpreter performs well in some
-interesting cases that are considered to be the worst case scenario in the control flow
-analysis, the staged version outperforms the unstaged abstract interpreter by a wide
-margin (up to 3x performance gain).
+Our evaluations shows that the staged abstract interpreter performs well in some
+cases that are considered to be the worst scenarios for traditional $k$-CFA.
+The staged version outperforms the unstaged abstract interpreter by a significant
+margin (up to ~230\% performance gain) on these difficult ones.
 
 \subsection{Benchmarks}
+The benchmark programs we used are collected from previous papers \cite{Johnson:2013:OAA:2500365.2500604,
+ashley:practical, DBLP:journals/corr/abs-1102-3676} and exisiting artifacts \footnote{https://github.com/ilyasergey/reachability}.
 
-Our evaluation is based on a suite of benchmark tests. Considering that our abstract 
-interpreters are implemented in Scala which will be affected by JVM warm up times, programs
-tend to run faster after a number of runs. In order to minimize the influence of JVM, 
-we ran all experiments 10 times and take the statistical mean and median values of 
-the running times.
-
-%on a suite of benchmark scheme programs that are also used in \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical}. 
 We used the following benchmark programs:
 \begin{itemize}
-    \item \textbf{fib:} a program that recursively calculates the $n$-th fibonacci number in
-        exponential time.
-    \item \textbf{church:} a program introduced by \todo{cite Dimitrios Vardoulakis and Olin Shivers. CFA2: a Context-Free approach
-        to Control-Flow analysis. Logical Methods in Computer Science,
-        7(2), 2011.} to test distributivity of multiplication for numbers in Church Encoding \todo{ cite }
-    \item \textbf{kcfa3/kcfa-worst-case-n:} Benchmark programs that are supposed to
-      be tough cases for CFA.
-    \item \textbf{fermat:} Fermat and Solovay-Strassen primality testing in Scheme by
-      Matthew Might.
-    \item \textbf{rsa:} The RSA public key encryption algorithm in Scheme.
+    \item \textbf{fib}: Calculates the $n$-th fibonacci number.
+    \item \textbf{rsa}: The RSA public key encryption algorithm.
+    \item \textbf{kcfa3}: A difficult benchmark for $k$-CFA.
+    \item \textbf{church}: Church numerals with additions and mutiplications.
+    \item \textbf{fermat}: Fermat and Solovay-Strassen primality testing.
+    \item \textbf{kcfa-worst-case-}\textit{n}: 
+      Benchmark programs that are supposed to be tough cases for $k$-CFA; 
+      the number $n$ indicates the depth of nesting lambda terms.
 \end{itemize}
@@
 
-For each benchmark program, we ran our prototype abstract interpreters, both unstaged and staged.
-\todo{Figure show data.}
+Considering that our abstract interpreters are implemented in Scala which will be 
+affected by the JIT warm up times, we run all experiments 10 times and reprot
+the statistical median values of the running times.
 
@@
 Figure~\ref{evaluation_result} shows the benchmark evaluation results. The "unstaged" and "staged" columns
-show the median time to finish the abstract interpretation (in milliseconds). The "$\Delta_{median}$"
-column shows the performance change from the unstaged abstract interpreter to the staged version.\\
+show the median time to finish the analysis (in milliseconds). The "$\Delta$ of median"
+column shows the performance change from the unstaged abstract interpreter to the staged version.
 
-However, there also exists some programs where the unstaged abstract interpreter yields 
-results faster than the unstaged interpreter.
-From the data shown above, we observed that 
-However, in general, the unstaged abstract interpreter yields results faster
-than the unstaged interpreter.
+As we can see from the table, the staged version significantly outperforms the unstaged especially 
+for the difficult ones.
+However, for some benchmarks, we observe degraded performance. 
+We conjecture this happens due to the large size of the generated code is unoptimized -- 
+as what we test here is a simple implementation almost identical to what we described in the previous sections,
+we believe the performance still can be improved if more engineering effort is taken, 
+such as optimizing code generation and avoiding code duplication.

### 2018-11-17 paper/50_cases.tex
revise

@@
 The idea is inspired by partial evaluation similar to the present paper -- the program can be known 
-statically therefore the overhead of interpretation can be eliminated. 
+statically, the overhead of interpretation can be eliminated. 
 In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
-the closures can be executed immediately however the textual program need to be compiled and loaded first.
+though the closures can be executed immediately, the textual programs need to be compiled and loaded first.
 
@@
 is essentially a big-step control-environment abstract interpreter.
-Closure generation compiles the analysis as a closure which takes an environment as argument.
+Closure generation compiles the analysis as a closure taking an environment argument.
 The overhead of traversing the abstract syntax tree of input program also has been eliminated.

### 2018-11-17 paper/60_eval.tex
revise

@@
 
-Our evaluations shows that the staged abstract interpreter performs well in some
+Our evaluations show that the staged abstract interpreter performs well in some
 cases that are considered to be the worst scenarios for traditional $k$-CFA.

### 2018-11-18 paper/60_eval.tex
fix typo

@@
 Considering that our abstract interpreters are implemented in Scala which will be 
-affected by the JIT warm up times, we run all experiments 10 times and reprot
+affected by the JIT warm up times, we run all experiments 10 times and report
 the statistical median values of the running times.

### 2018-11-20 paper/paper.tex
intro
notes on intro
fix paper
revise intro
work on paper
fixpoint iter
delete empty appendix
preprint for website

@@
+\documentclass[sigplan,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+% \documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage[scaled=0.83]{beramono}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{xcolor,colortbl}
+\usepackage{url}
+\usepackage{listings}
+\usepackage{paralist}
+%\usepackage[compact]{titlesec}
+\usepackage[font={small}]{caption}
+\usepackage{wrapfig}
+\usepackage{enumitem}
+\usepackage{multicol}
+\usepackage{flushend}
+\usepackage{bcprules}
+
+\usepackage{tikz}
+\usetikzlibrary{matrix}
+
+\input{macros}
+
+\newcommand{\updownarrows}{\mathbin\uparrow\hspace{-.5em}\downarrow}
+\newcommand{\downuparrows}{\mathbin\downarrow\hspace{-.5em}\uparrow}
+
 \begin{document}
@@
 %% Title information
-\title[Short Title]{Staged Abstract Interpreters}         %% [Short Title] is optional;
+\title{Staged Abstract Interpreters}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
                                         %% contents suppressed with 'anonymous'
-\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
+%\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
+\subtitle{Fast and Modular Whole-Program Analysis via Meta-Programming}                     %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
                                         %% contents suppressed with 'anonymous'
+
+\author
+[Guannan Wei, Yuxuan Chen, Tiark Rompf]
+{
+\vspace{-2ex}
+Guannan Wei, Yuxuan Chen, Tiark Rompf\\
+Purdue University
+\vspace{-0.5ex}
+}
 
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
-}
-\email{first1.last1@inst1.edu}          %% \email is recommended
+% \author{First1 Last1}
+% \authornote{with author1 note}          %% \authornote is optional;
+%                                         %% can be repeated if necessary
+% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+% \affiliation{
+%   \position{Position1}
+%   \department{Department1}              %% \department is recommended
+%   \institution{Institution1}            %% \institution is required
+%   \streetaddress{Street1 Address1}
+%   \city{City1}
+%   \state{State1}
+%   \postcode{Post-Code1}
+%   \country{Country1}                    %% \country is recommended
+% }
+% \email{first1.last1@inst1.edu}          %% \email is recommended
 
-%% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
-                                        %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
-}
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
+% %% Author with two affiliations and emails.
+% \author{First2 Last2}
+% \authornote{with author2 note}          %% \authornote is optional;
+%                                         %% can be repeated if necessary
+% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+% \affiliation{
+%   \position{Position2a}
+%   \department{Department2a}             %% \department is recommended
+%   \institution{Institution2a}           %% \institution is required
+%   \streetaddress{Street2a Address2a}
+%   \city{City2a}
+%   \state{State2a}
+%   \postcode{Post-Code2a}
+%   \country{Country2a}                   %% \country is recommended
+% }
+% \email{first2.last2@inst2a.com}         %% \email is recommended
+% \affiliation{
+%   \position{Position2b}
+%   \department{Department2b}             %% \department is recommended
+%   \institution{Institution2b}           %% \institution is required
+%   \streetaddress{Street3b Address2b}
+%   \city{City2b}
+%   \state{State2b}
+%   \postcode{Post-Code2b}
+%   \country{Country2b}                   %% \country is recommended
+% }
+% \email{first2.last2@inst2b.org}         %% \email is recommended
+
+\lstMakeShortInline[keywordstyle=,%
+                    flexiblecolumns=false,%
+                    %basewidth={0.56em, 0.52em},%
+                    mathescape=false,%
+                    basicstyle=\tt]@
 
@@
 %% comma separated list
-\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
 
@@
 \maketitle
+
+\renewcommand\thefootnotecopyrightpermission{}
+\footnotetextcopyrightpermission{
+Preprint, November 2018.\\ Copyright held by the authors.}
+\fancyhead[RO,LE]{Preprint, November 2018}
 
@@
 
-%% Appendix
-\appendix
-\section{Appendix}
+% %% Appendix
+% \appendix
+% \section{Appendix}
 
-Text of appendix \ldots
+% Text of appendix \ldots
 

### 2019-02-04 paper/10_intro.tex
reorganize and address typos

--- 2018-11-17 paper/10_intro.tex
@@
 
-% \subsection{Motivation}
+Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close
+connection between compilers and interpreters. The first Futamura projection
+specifically shows that specializing an interpreter with respect to the input
+program yields an equivalent executable. Partial evaluation
+\cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
+Futamura projections, which first identifies the binding-time of variables in
+the program, i.e., they can be known whether statically or dynamically, and then
+evaluates the static part, and generates a residual program that relies on the
+dynamic part. However, precisely analyzing binding-time for an arbitrary program
+is hard in general. As an alternative and pragmatic approach to specialization
+and partial evaluation, multi-stage programming (MSP for short)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage annotations
+(i.e., binding-time annotations) from the programmers. These staging annotations
+identify which part of the input program should be specialized. As a classical
+example, we consider a power function:
 
-Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close connection between 
-compilers and interpreters. 
-The first Futamura projection specifically shows that specializing an interpreter with respect to 
-the input program yields an equivalent executable.
-Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed approach to realize 
-Futamura projections, which first identifies the binding-time of variables in the program, i.e., 
-they can be known whether statically or dynamically, and then evaluates the static part, and 
-generates a residual program that relies on the dynamic part.
-However, precisely analyzing binding-time for an arbitrary program is hard in general. As an 
-alternative and pragmatic approach to specialization and partial evaluation, multi-stage 
-programming (MSP for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage 
-annotations (i.e., binding-time annotations) from the programmers. These staging annotations 
-identify which part of the input program should be specialized.
-As a classical example, we consider a power function: 
 \begin{lstlisting}
@@
 \end{lstlisting}
-If the program identifies that @b@ will be known dynamically (as shown in its type @Rep[Int]@ -- a representation of @Int@) 
-and @e@ is statically known, say 5, 
-then we can specialize the function \texttt{power} with respect to \texttt{e = 5} and generate a new 
-specialized function that runs faster:
+
+If the program identifies that @b@ will be known dynamically (as shown in its
+type @Rep[Int]@ -- a representation of @Int@) and @e@ is statically known, say
+5, then we can specialize the function \texttt{power} with respect to \texttt{e
+= 5} and generate a new specialized function that runs faster:
+
 \begin{lstlisting}
@@
 
-Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach to construct 
-sound static analysis by approximation is almost as old as Futamura projections.
-Some recent progress such as Abstracting Abstract Machines (AAM) uncovers a methodology to 
-derive sound abstract interpreters from their concrete counterparts and has been applied to 
-different variants of definitional interpreters and abstract machines
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-Given the structural similarity between concrete and abstract interpreters, intellectually 
-it is natural to raise the question whether it is possible to specialize an abstract interpreter, and how 
-can we effectively do that having considered their different functionalities.
-This paper studies the confluence of two old ideas --- Futamura project and abstract interpretation,
-but from the perspective of their recent realizations --- multi-stage programming and definitional
-abstract interpreters. To be specific, we present the application of the first Futamura projection
-on abstract interpreters, and the approach of constructing optimizing abstract interpreters by 
-multi-stage programming.
+Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach
+to construct sound static analysis by approximation is almost as old as Futamura
+projections. Some recent progress such as Abstracting Abstract Machines (AAM)
+uncovers a methodology to derive sound abstract interpreters from their concrete
+counterparts and has been applied to different variants of definitional
+interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
+DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}. Given the structural
+similarity between concrete and abstract interpreters, intellectually it is
+natural to raise the question whether it is possible to specialize an abstract
+interpreter, and how can we effectively do that having considered their
+different functionalities. This paper studies the confluence of two old ideas
+--- Futamura projection and abstract interpretation, but from the perspective of
+their recent realizations --- multi-stage programming and definitional abstract
+interpreters. To be specific, we present the application of the first Futamura
+projection on abstract interpreters, and the approach of constructing optimizing
+abstract interpreters by multi-stage programming.
 
@@
   \end{tikzpicture}
-  \caption{The confluence of staging and abstraction.}
+  \caption{The confluence of staging and approximation.}
   \label{confluence}
@@
 
-Multi-stage programming has been widely used to improve the performance in many domains, 
-such as optimizing compilers or domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13, DBLP:journals/jfp/CaretteKS09},
-numerical computation \cite{PGL-038, DBLP:conf/pepm/AktemurKKS13}, 
-generic programming \cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, 
-data processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14, DBLP:conf/popl/KiselyovBPS17}, 
-query compilation in databases \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
-etc.
+Multi-stage programming has been widely used to improve the performance in many
+domains, such as optimizing compilers or domain-specific languages
+\cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
+DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
+DBLP:journals/jfp/CaretteKS09}, numerical computation \cite{PGL-038,
+DBLP:conf/pepm/AktemurKKS13}, generic programming
+\cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, data
+processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
+DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
+\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
 Likewise, specializing static analyses by partial evaluation emerged in late 90s
-\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical}, 
-and indeed it is able to effectively remove the interpretive overhead of repeatedly traversing 
-the abstract syntax tree. 
-However, these previous works focused mostly on one particular analysis, or required to 
-completely rewrite the analyzer. 
-Hence, it is worth to investigate the idea from a modern perspective based on generative programming,
-especially for a general abstract interpreter that models direct-style $\lambda$-calculus, imperative 
-features such as mutation, and different abstract domains.
-One technical challenge is the binding-time engineering when non-determinism, 
-fixed-point iterations, different abstract domains and staging are introduced at the same time.
-In this paper, we present an end-to-end design and implementation of staging an abstract interpreter;
-that means not only the interpreter that traverses the abstract syntax tree, 
-but also the data structure of abstract domains, abstract environment and heap, 
-and the fixed-point iteration are staged.
+\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
+ashley:practical}, and indeed it is able to effectively remove the interpretive
+overhead of repeatedly traversing the abstract syntax tree. However, these
+previous works focused mostly on one particular analysis, or required to
+completely rewrite the analyzer. Hence, it is worth to investigate the idea from
+a modern perspective based on generative programming, especially for a general
+abstract interpreter that models direct-style $\lambda$-calculus, imperative
+features such as mutation, and different abstract domains. One technical
+challenge is the binding-time engineering when non-determinism, fixed-point
+iterations, different abstract domains and staging are introduced at the same
+time. In this paper, we present an end-to-end design and implementation of
+staging an abstract interpreter; that means not only the interpreter that
+traverses the abstract syntax tree, but also the data structure of abstract
+domains, abstract environment and heap, and the fixed-point iteration are
+staged.
 
-On the other hand, the abstract interpreter, as a semantic artifact, should be written in a 
-style that is easy for people to communicate the formulation and abstraction, but also can be 
-implemented efficiently. As the slogan of multi-stage programming said, "abstraction without regret",
-we draw connection between high-level description and efficient implementation of abstract interpreters,
-just like the connection between concrete interpreters and compilers drawn by Futamura projection.
-Particularly, we show a easy but systematic way of adding stage annotations to the abstract interpreter
-without changing the code of interpreter skeleton, which is shared between four concrete or abstract,
-unstaged or staged interpreters. We use LMS framework for staging, which allows only use types to 
-annotate the binding-time. Therefore, the proposed approach bridges the gap between designing sound 
-static analysis and implementing efficient program analyzer.
+\todo{rewrite this paragraph}
+On the other hand, the abstract interpreter, as a semantic artifact, should be
+written in a style that is easy for people to communicate the formulation and
+abstraction, but also can be implemented efficiently. As the slogan of
+multi-stage programming said, "abstraction without regret", we draw connection
+between high-level description and efficient implementation of abstract
+interpreters, just like the connection between concrete interpreters and
+compilers drawn by Futamura projection. Particularly, we show an easy but
+systematic way of adding stage annotations to the abstract interpreter without
+changing the code of the interpreter skeleton, which is shared between four concrete
+or abstract, unstaged or staged interpreters. We use the LMS framework for staging,
+which allows us only use types to annotate the binding-time. Therefore, the
+proposed approach bridges the gap between designing sound static analysis and
+implementing efficient program analyzer.
 
-%Of course, all interpreters are metalinguistic abstractions, but some interpreters are more "abstract" than others \todo{maybe rephrase}. 
-%Particularly, we also would like a systematic approach to optimize program analyzers and meanwhile minimally modifying the analyzer programs.
+%Of course, all interpreters are metalinguistic abstractions, but some
+%interpreters are more "abstract" than others \todo{maybe rephrase}.
+%Particularly, we also would like a systematic approach to optimize program
+%analyzers and meanwhile minimally modifying the analyzer programs.
+
 \iffalse
-When staging a concrete interpreter, the programmers need to distinguish 
-static and dynamic values --- the given program to be executed by the 
-interpreter is classified as static because it is known at compile-time, 
-and the inputs to that program are dynamic.  However, when staging an 
-abstract interpreter, this distinction does not exist anymore. Because the 
-abstract interpreter instantiates all the inputs as some form of abstract 
-values, which are usually top elements in their abstract domains and are 
-also statically known. 
-Then what is the point of staging if there is no such distinction?
-A surprising by-product of thinking about this question is to realize that 
-we can apply the staged abstract interpreter on \textit{open} programs, and
-the free variables representing other parts of the program (e.g., libraries) 
-are dynamic inputs, therefore we obtain a modular analysis through staging, 
-mechanically. 
-\note{TR: I don't understand this. The program structure is static, the abstract values are still dynamic, no? They change in every iteration of the fixpoint algorithm} 
-\note{TR: why is it a big deal. On Closed programs we obtain a constant factor, but modular analysis has different asymptotics}
+When staging a concrete interpreter, the programmers need to distinguish static
+and dynamic values --- the given program to be executed by the interpreter is
+classified as static because it is known at compile-time, and the inputs to that
+program are dynamic. However, when staging an abstract interpreter, this
+distinction does not exist anymore. Because the abstract interpreter
+instantiates all the inputs as some form of abstract values, which are usually
+top elements in their abstract domains and are also statically known. Then what
+is the point of staging if there is no such distinction? A surprising by-product
+of thinking about this question is to realize that we can apply the staged
+abstract interpreter on \textit{open} programs, and the free variables
+representing other parts of the program (e.g., libraries) are dynamic inputs,
+therefore we obtain a modular analysis through staging, mechanically.
+\note{TR: I don't understand this. The program structure is static, the abstract values
+are still dynamic, no? They change in every iteration of the fixpoint algorithm}
+\note{TR: why is it a big deal. On Closed programs we obtain a constant factor,
+but modular analysis has different asymptotics}
 \fi
 
-When targeting higher-order programs, either staged concrete interpreters or staged abstract interpreters
-are able to compile a closure, i.e., specialize the call of interpreter with respect to the body
-expression of the lambda term without knowing the actual argument value. By generalizing this observation,
-we can actually specialize the abstract interpreter with any open programs, 
-which unexpectedly leads to a modular analysis and improves scalability.
-A open program contains free variables, which represent other parts of the program and will be left
-as dynamic values.
-For instances, one challenge in static analysis of modern software is that the programs are usually 
-shipped with large library code \cite{toman_et_al:LIPIcs:2017:7121}, for example, it has been shown that 
-analyzing a simple "Hello World" program in Java depends on additional 3,000 classes in the library 
-\cite{DBLP:conf/oopsla/KulkarniMZN16}.
-A precise whole-program analysis formulated as abstract interpreters can be very expensive
+When targeting higher-order programs, either staged concrete interpreters or
+staged abstract interpreters are able to compile a closure, i.e., specialize the
+call of interpreter with respect to the body expression of the lambda term
+without knowing the actual argument value. By generalizing this observation, we
+can actually specialize the abstract interpreter with any open programs, which
+unexpectedly leads to a modular analysis and improves scalability. An open
+program contains free variables, which represent other parts of the program and
+will be left as dynamic values. For instance, one challenge in static analysis
+of modern software is that the programs are usually shipped with large library
+code \cite{toman_et_al:LIPIcs:2017:7121}, for example, it has been shown that
+analyzing a simple "Hello World" program in Java depends on additional 3,000
+classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}. A precise
+whole-program analysis formulated as abstract interpreters can be very expensive
 due to the scale of the program and the inherent complexity of the algorithm.
-However, analyzing these libraries is sometimes unnecessarily repeated.
-When applying the staged abstract interpreter on such open programs, we leave the unknown arguments and 
-calling contexts as dynamic values and generate partial analyzing result which is 
-represented as a residual program. The partial analyzing result can be reused and composed with the 
-analyzing result of application code when available later. Therefore we mechanically obtain a modular
-analysis through staging, even though the original algorithm is formulated as a whole-program analysis.
+However, analyzing these libraries is sometimes unnecessarily repeated. When
+applying the staged abstract interpreter on such open programs, we leave the
+unknown arguments and calling contexts as dynamic values and generate partial
+analyzing result which is represented as a residual program. The partial
+analyzing result can be reused and composed with the analyzing result of
+application code when available later. Therefore we mechanically obtain a
+modular analysis through staging, even though the original algorithm is
+formulated as a whole-program analysis.
 
-It has been observed that partially applying context-sensitivity on selected portion of the program could
-improve the precision and efficiency \cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}.
-We show that staging abstract interpreters as an approach to effectively implement hybird context-sensitivity\todo{}.
-
-% Many optimization of static analysis are ad-hoc.
+\iffalse
+It has been observed that partially applying context-sensitivity on selected
+portion of the program could improve the precision and efficiency
+\cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. We show that staging
+abstract interpreters as an approach to effectively implement hybird
+context-sensitivity\todo{}.
+\fi
 
@@
 \begin{itemize}
-  \item Starting from a generic definition of interpreter for a higher-order functional language,
-    whose syntax and semantics are described in Section~\ref{bg_lang}, we show its concrete
-    interpreter (Section~\ref{unstaged_conc}), staged concrete interpreter (Section~\ref{stagedinterp}), 
-    abstract interpreter (Section~\ref{unstaged_abs}) and staged abstract interpreter (Section~\ref{sai}).
-    Our design allows one can derive these four different kinds of interpreters by mechanically changing
-    the underlying semantics or stage annotations.
-    %We later extend the staged abstract interpreter to support imperative language features (Section~\ref{cases_imp}), such as states and loops.
-  \item We demonstrate that the staged abstract interpreter not only improves the \textit{efficiency}
-    but also the \textit{scalability}, which are two major issues in static analysis, if we apply it on open
-    programs. \todo{Section X}.
-  \item To demonstrate the applicability and usefulness of the proposed approach, we conduct three case studies 
-    Section~\ref{cases_study}: 
+  \item Starting from a generic definition of an interpreter for a higher-order
+    functional language, whose syntax and semantics are described in
+    Section~\ref{bg_lang}, we show its concrete interpreter
+    (Section~\ref{unstaged_conc}), staged concrete interpreter
+    (Section~\ref{stagedinterp}), abstract interpreter (Section~\ref{unstaged_abs})
+    and staged abstract interpreter (Section~\ref{sai}). Our design allows to
+    derive these four different kinds of interpreters by mechanically changing the
+    underlying semantics or stage annotations.
+    %We later extend the staged abstract interpreter to support imperative
+    %language features (Section~\ref{cases_imp}), such as states and loops.
+
+  \item As shown in Section~\ref{sai}, we demonstrate that if we apply the
+    abstract interpreters on open programs, it not only improves the
+    \textit{efficiency} but also the \textit{scalability}, which are two major
+    issues in static analysis.
+
+  \item To demonstrate the applicability and usefulness of the proposed
+    approach, we conduct three case studies in Section~\ref{cases_study}:
+
     \subitem (1) We revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587} in Section~\ref{cs_ac}, 
     which was originally a closure generation technique applied on 0-CFA.
-    We show that abstract compilation can be understood and easily implemented as an instance of our staged abstract interpreter.
-    But unlike the abstract compilation, our approach does not require the implementers to rewrite the analysis.
-    \subitem (2) We demonstrate pushdown control-flow analysis techniques (Section~\ref{cfa}), such as context-sensitive 
-    analysis and more precise stores can be integrated to staged abstract interpreters without losing any precision.
-    \subitem (3) We show that in an imperative language (Section~\ref{cases_imp}), by replacing the abstract domain
-    to an interval domain (as well as staged), the staged abstract interpreter performs numerical analysis.
+    We show that abstract compilation can be understood and easily implemented
+    as an instance of our staged abstract interpreter. But unlike abstract
+    compilation, our approach does not require the implementers to rewrite the
+    analysis.
+
+    \subitem (2) We demonstrate pushdown control-flow analysis techniques
+    (Section~\ref{cfa}), such as context-sensitive analysis and more precise
+    stores can be integrated to staged abstract interpreters without losing any
+    precision.
+
+    \subitem (3) We show that in an imperative language
+    (Section~\ref{cases_imp}), by using an interval domain as abstract domain
+    (as well as staged), the staged abstract interpreter performs numerical 
+    analysis.
+
     % (Optional: taint/information flow analysis?)
-  \item We empirically evaluate the performance improvement by staging an big-step abstract interpreter 
-    for control-flow analysis (Section~\ref{evaluation}). 
-    We compare our staged abstract interpreter with the unstaged version.
+  \item We empirically evaluate the performance improvement by staging a
+    big-step abstract interpreter for control-flow analysis
+    (Section~\ref{evaluation}). We compare our staged abstract interpreter with the
+    unstaged version.
 \end{itemize}

### 2019-02-04 paper/21_bg_interp.tex
revise
reorganize and address typos

@@
 
-We consider a tiny higher-order functional language \TLang, which is based on direct-style $\lambda$-calculus
-with numbers, arithmetics, recursions and conditionals.
-In Section~\ref{cases_imp}, we will add more imperative features to the language.
-Since we are mostly interested in analyzing the dynamic behavior of the program, we disguised any static semantics and
-type system. We also assume that input programs are well-typed and variables are distinct.
-The abstract syntax is as follows:
+We consider a tiny higher-order functional language \TLang, which is based on
+direct-style $\lambda$-calculus with numbers, arithmetics, recursions and
+conditionals. In Section~\ref{cases_imp}, we will add more imperative features
+to the language. Since we are mostly interested in analyzing the dynamic
+behavior of the program, we disguised any static semantics and type system. We
+also assume that input programs are well-typed and variables are distinct. The
+abstract syntax is as follows:
 
@@
 
-We give the concrete semantics of \TLang by showing a big-step definitional interpreter.
-The interpreter is a recursive function that takes the program AST, environment and store, and returns 
-the evaluated value and the accompanied store. The environment is a mapping from identifiers to addresses,
-and the store is a mapping from addresses to values. 
-We use the store to model recursion and mutation in concrete semantics; it is also useful for polyvariant analysis.
-This environment-and-store-passing style big-step interpreter is standard and can also be obtained by 
-refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03, Wei:2018:RAA:3243631.3236800} a small-step CESK 
-machine \cite{DBLP:conf/popl/FelleisenF87}.
+We give the concrete semantics of \TLang by showing a big-step definitional
+interpreter. The interpreter is a recursive function that takes the program AST,
+environment and store, and returns the evaluated value and the accompanied
+store. The environment is a mapping from identifiers to addresses, and the store
+is a mapping from addresses to values. We use the store to model recursion and
+mutation in concrete semantics; it is also useful for polyvariant analysis. This
+environment-and-store-passing style big-step interpreter is standard and can
+also be obtained by refunctionalizing \cite{DBLP:conf/ppdp/AgerBDM03,
+Wei:2018:RAA:3243631.3236800} a small-step CESK machine
+\cite{DBLP:conf/popl/FelleisenF87}.
 
 \subsection{Generic Interface}
-Before showing the actual implementation of the concrete semantics, let us first consider
-a generic interface for the interpreters, which is semantics-agnostic and can be 
-"interpreted" by either concrete or abstract, unstaged or staged semantics of our choice.
-The trait \texttt{Semantics} first declares several abstract type members, such as \texttt{Addr},
-\texttt{Value}, \texttt{Env}, \texttt{Store} and returned type \texttt{Ans}. 
-It also declares initial values for environments and stores, and few abstract methods that 
-manipulate environments and stores such as @put@ and @get@.
+
+Before showing the actual implementation of the concrete semantics, let us first
+consider a generic interface for the interpreters, which is semantics-agnostic
+and can be "interpreted" by either concrete or abstract, unstaged or staged
+semantics of our choice. The trait \texttt{Semantics} first declares several
+abstract type members, such as \texttt{Addr}, \texttt{Value}, \texttt{Env},
+\texttt{Store} and returned type \texttt{Ans}. It also declares initial values
+for environments and stores, and few abstract methods that manipulate
+environments and stores such as @put@ and @get@.
 
@@
 
-The actual semantics and operations of the interpreter are left to be implemented later, 
-these cinclude operations such as lifting literal terms to values (\texttt{close}, \texttt{num}), 
-applying a closure (\texttt{apply\_closure}), branching (\texttt{branch0}),
-and arithmetics (\texttt{prim\_eval}). We can choose to implement them in a way that is concrete or abstract,
-unstaged or staged.
-But, just based on these abstract operations, an expressive generic interpreter can be defined:
+The actual semantics and operations of the interpreter are left to be
+implemented later, these include operations such as lifting literal terms to
+values (\texttt{close}, \texttt{num}), applying a closure
+(\texttt{apply\_closure}), branching (\texttt{branch0}), and arithmetics
+(\texttt{prim\_eval}). We can choose to implement them in a way that is concrete
+or abstract, unstaged or staged. But, just based on these abstract operations,
+an expressive generic interpreter can be defined:
 
@@
 
-The implementation of the skeleton interpreter traverses the abstract syntax tree and should be straightforward to understand.
-It is worth noting that the interpreter is written in open-recursive style -- 
-it can not refer to itself directly, instead, \texttt{eval} takes an additional 
-parameter \texttt{ev} referring to itself; this allows us to instrument the interpreter
-from the outside. Accordingly, the function \texttt{close} for lifting lambda term to closure values
-is also written in this style.
-We define a default implementation \texttt{fix} to close the recursion by recursively using the result 
-of \texttt{fix(ev)} for \texttt{ev}. 
-Finally, the top-level evaluation function \texttt{eval\_top} is defined.
+The implementation of the skeleton interpreter traverses the abstract syntax
+tree and should be straightforward to understand. It is worth noting that the
+interpreter is written in open-recursive style -- it can not refer to itself
+directly, instead, \texttt{eval} takes an additional parameter \texttt{ev}
+referring to itself; this allows us to instrument the interpreter from the
+outside. Accordingly, the function \texttt{close} for lifting lambda term to
+closure values is also written in this style. We define a default implementation
+\texttt{fix} to close the recursion by recursively using the result of
+\texttt{fix(ev)} for \texttt{ev}. Finally, the top-level evaluation function
+\texttt{eval\_top} is defined.
 
-\paragraph{Stage Polymorphism.}
-Readers may notice that an abstract higher-kinded type \texttt{R[+T]} is defined in the trait, i.e. a
-type-level function whici will be used to specify the binding-time, i.e., staged or not; the annotation 
-\texttt{+} indicates that the type parameter \texttt{T} is covariant.
-The LMS framework we use here allows us to specify binding-times purely based on types.
-All occurrences of type \texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} 
-essentially because they may potentially be dynamic when staging is introduced.
-Obviously, the unstaged higher-kinded type then can be simply instantiated as an identity type function: 
-\texttt{type R[+T] = T}. In Section~\ref{stagedinterp}, \texttt{R} will be instantiated 
-as a staging annotation. \todo{Finally tagless, poly embedding}
+\paragraph{Stage Polymorphism}
+
+Readers may notice that an abstract higher-kinded type \texttt{R[+T]} is defined
+in the trait, i.e. a type-level function which will be used to specify the
+binding-time, i.e., staged or not; the annotation \texttt{+} indicates that the
+type parameter \texttt{T} is covariant. The LMS framework we use here allows us
+to specify binding-times purely based on types. All occurrences of type
+\texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} essentially
+because they may potentially be dynamic when staging is introduced. Obviously,
+the unstaged higher-kinded type then can be simply instantiated as an identity
+type function: \texttt{type R[+T] = T}. In Section~\ref{stagedinterp},
+\texttt{R} will be instantiated as a staging annotation.
+
+\todo{Finally tagless, poly embedding}
 
 \subsection{Concrete Instantiation} \label{unstaged_conc}
-Now we can instantiate a concrete interpreter. We first concretize the abstract types as follows.
-Since the interpreter is concrete, for the propose that properly models the heap, the type \texttt{Int} 
-is used for address space \texttt{Addr}. The environment and store are just ordinary \texttt{Map}s in Scala.
-A value can be either a tagged number \texttt{NumV} or a closure \texttt{CloV} that contains a lambda term
-and an environment.
+
+Now we can instantiate a concrete interpreter. We first concretize the abstract
+types as follows. Since the interpreter is concrete, for the purpose that
+properly models the heap, the type \texttt{Int} is used for address space
+\texttt{Addr}. The environment and store are just ordinary \texttt{Map}s in
+Scala. A value can be either a tagged number \texttt{NumV} or a closure
+\texttt{CloV} that contains a lambda term and an environment.
 
@@
 
-we proceed to implement the components that define the semantics. 
-As mentioned before, type \texttt{R} is concretized as an identity type function, thus
-all types are just ordinary types.
+We proceed to implement the components that define the semantics. As mentioned
+before, type \texttt{R} is concretized as an identity type function, thus all
+types are just ordinary types.
 
@@
 
-Finally, \texttt{get} and \texttt{put} directly manipulate the map of environment and store. 
-An allocation function is used when we need to update the store: 
-in the concrete case, \texttt{alloc} always returns a fresh address of the store.
-\texttt{close} and \texttt{num} lift syntactic literals to values.
-\texttt{apply\_closure} takes the function value, argument value and the latest store,
-then extends the environment and store after allocating a fresh address, and continue
-evaluating the body \texttt{e} of the lambda term.
+Finally, \texttt{get} and \texttt{put} directly manipulate the map of
+environment and store. An allocation function is used when we need to update the
+store: in the concrete case, \texttt{alloc} always returns a fresh address of
+the store. \texttt{close} and \texttt{num} lift syntactic literals to values.
+\texttt{apply\_closure} takes the function value, argument value and the latest
+store, then extends the environment and store after allocating a fresh address,
+and continue evaluating the body \texttt{e} of the lambda term.

### 2019-02-04 paper/23_bg_stage.tex
reorganize and address typos

@@
 \subsection{Multi-Stage Programming in LMS}
-Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a multi-stage programming framework implemented 
-as a Scala library that enables dynamic code generation in a type-safe manner.
-Different from the syntactic approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations, 
-LMS distinguishes binding-time solely based on types.
-LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an expression will be 
-known at next stage. All operations acting on a @Rep[T]@ expression will be residualized in the generated code. However,
-we need to provide intermediate representation and code generation support for type @T@ to LMS. 
-Fortunately, the LMS framework already provides
-such support for primitive types and commonly used data structure such as arrays and maps; implementing such support
-for custom class is also straightforward. Let's go back and see how LMS can be used to specialize the power function
-we mentioned in Section 1.
+
+Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
+multi-stage programming framework implemented as a Scala library that enables
+dynamic code generation in a type-safe manner. Different from the syntactic
+approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations,
+LMS distinguishes binding-time solely based on types. LMS provides a type
+constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an
+expression will be known at next stage. All operations acting on a @Rep[T]@
+expression will be residualized in the generated code. However, we need to
+provide intermediate representation and code generation support for type @T@ to
+LMS. Fortunately, the LMS framework already provides such support for primitive
+types and commonly used data structure such as arrays and maps; implementing
+such support for custom class is also straightforward. Let's go back and see how
+LMS can be used to specialize the power function we mentioned in Section 1.
 
@@
 
-In the code shown above, \texttt{power} takes two arguments where \texttt{b} is declared as \texttt{Rep[Int]} type 
-meaning that \texttt{b} is a representation of \texttt{Int} but whose value will be available at the next stage. 
-Meanwhile, the result of the @power@ function will be available at the next stage as well, thus the return value 
-of \texttt{power} is also of type \texttt{Rep[Int]}. 
-Then we specialize @power@ in function @snippet@ by providing @x@ to @5@.
-The generated code for \texttt{power5} is a function that takes only one argument which corresponds to @b@,
-and the body of this function multiples @b@ five times.
+In the code shown above, \texttt{power} takes two arguments where \texttt{b} is
+declared as \texttt{Rep[Int]} type meaning that \texttt{b} is a representation
+of \texttt{Int} but whose value will be available at the next stage. Meanwhile,
+the result of the @power@ function will be available at the next stage as well,
+thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. Then
+we specialize @power@ in function @snippet@ by providing @x@ to @5@. The
+generated code for \texttt{power5} is a function that takes only one argument
+which corresponds to @b@, and the body of this function multiplies @b@ five
+times.
 
 \subsection{Staged Concrete Semantics}
-In the staging part, we share the same concrete type instantiation, 
-i.e. trait \texttt{Concrete}, but we reimplement the staged version for the concrete operations
-in trait @RepConcInterpOps@.
-Meanwhile, it is also extended from @LMSOps@ which provides necessary supports for the 
-staged version of primitive types and data structures.
-To stage the interpreter, we first identifies the syntactic input is known statically, 
-and the interpreter returns a pair of values and store, which both becomes dynamic values.
-Thus the staged stores propogate the binding-time to environments, because they share the same
-address component.
+
+In the staging part, we share the same concrete type instantiation, i.e., trait
+\texttt{Concrete}, but we reimplement the staged version for the concrete
+operations in trait @RepConcInterpOps@. Meanwhile, it is also extended from
+@LMSOps@ which provides necessary support for the staged version of primitive
+types and data structures. To stage the interpreter, we first identify the
+syntactic input that is known statically, and the interpreter returns a pair of
+values and store, which both becomes dynamic values. Thus the staged stores
+propagate the binding-time to environments, because they share the same address
+component.
 
@@
 
-The first notable change is that the abstract type member @R@ is assigned to be @Rep@, and accordingly the affected 
-types such as @Env@ and @Store@ become @Rep[Env]@ and @Rep[Store]@. For @num@, we use the @lift@ function which is a
-built-in in LMS that lifts a current-stage value to next stage.
-The interesting point is how we handle closures and function applications when staging is involved.
-In the unstaged version, function @close@ is used to lift a literal lambda term to a closure, which
-is a pair of the syntactic lambda term and the enclosing environment.
-However, what we desire is a \textit{compiled} closure, instead of a lifted @CloV@ value like 
-what we did for @NumV@ --- in other words, the syntactic term of the lambda expression should be 
-eliminated and specialized away.
-The specialization of the interpreter with respect to the body of the lambda term proceeds under the
-assumptions that the argument and latest store will be provided later. 
+The first notable change is that the abstract type member @R@ is assigned to be
+@Rep@, and accordingly the affected types such as @Env@ and @Store@ become
+@Rep[Env]@ and @Rep[Store]@. For @num@, we use the @lift@ function which is a
+built-in in LMS that lifts a current-stage value to next stage. The interesting
+point is how we handle closures and function applications when staging is
+involved. In the unstaged version, function @close@ is used to lift a literal
+lambda term to a closure, which is a pair of the syntactic lambda term and the
+enclosing environment. However, what we desire is a \textit{compiled} closure,
+instead of a lifted @CloV@ value like what we did for @NumV@ --- in other words,
+the syntactic term of the lambda expression should be eliminated and specialized
+away. The specialization of the interpreter with respect to the body of the
+lambda term proceeds under the assumptions that the argument and latest store
+will be provided later.
 %\note{to denotational style?, back to HO?}
-We will see this is an important observation that enables us to specialize an abstract interpreter in a modular way.
-At the end of function @close@, we create a @CompiledClo@ object that contains a staged function, which all will be
+We will see this is an important observation that enables us to specialize an
+abstract interpreter in a modular way. At the end of function @close@, we create
+a @CompiledClo@ object that contains a staged function, which all will be
 emitted in the generated code.
@@
 
-Next, @apply_closure@ takes three arguments: the function value, argument value and the latest store, 
-which are both of @Rep@ type. This means basically there is nothing we can do at the current stage --- 
-so we create a new node @ApplyClosure@ that contains these arguments in the intermediate representation graph,
-and in the code generation phase we can emit the code that performs the application. The additional @reflectEffect@
-is a function in LMS that handles side effects.
+Next, @apply_closure@ takes three arguments: the function value, argument value
+and the latest store, which are all of @Rep@ type. This means basically there
+is nothing we can do at the current stage --- so we create a new node
+@ApplyClosure@ that contains these arguments in the intermediate representation
+graph, and in the code generation phase we can emit the code that performs the
+application. The additional @reflectEffect@ is a function in LMS that handles
+side effects.
 
@@
 
-\paragraph{Code Generation.}
-When generating code for the next stage, values like @CompiledClo@ and @NumV@ are kept as they are because
-they are intended exist in the next stage; however, we need to treat @ApplyClosure@ a little differently.
+\paragraph{Code Generation}
+
+When generating code for the next stage, values like @CompiledClo@ and @NumV@
+are kept as they are because they are intended exist in the next stage; however,
+we need to treat @ApplyClosure@ a little differently.
 
@@
 
-Concretely, as shown in the code above, when emitting code for @ApplyClosure@, we just 
-emit a function application where the function value is extracted from the @CompiledClo@ object
-(we assume the input program is well-typed so that we can safely cast it to @CompiledClo@).
-Argument and store are the remaining rest fields in @ApplyClosure@.
+Concretely, as shown in the code above, when emitting code for @ApplyClosure@,
+we just emit a function application where the function value is extracted from
+the @CompiledClo@ object (we assume the input program is well-typed so that we
+can safely cast it to @CompiledClo@). Argument and a store are the remaining rest
+fields in @ApplyClosure@.

### 2019-02-04 paper/30_sai.tex
reorganize and address typos

@@
 
-In the previous sections, we have seen an unstaged abstract interpreter and a staged concrete interpreter, 
-now we begin describing the implementation of their confluence -- a staged abstract interpreter. 
-We present a principled approach to derive staged abstract interpreter
-from its unstaged version. One guiding principle of our approach is that the code of the abstract semantics
-and the code that optimizes should be separated. This it is an advantage of using staging 
-for abstract interpreters: the designer of the analysis has no need to rewrite the analysis, and 
-the performance improvement comes almost for free, without any sacrifice of soundness or precision.
-Unsurprisingly, the staged abstract interpreter we present in this section has the same abstract 
-semantics as the unstaged version we presented in Section~\ref{unstaged_abs}.
+In the previous sections, we have seen an unstaged abstract interpreter and a
+staged concrete interpreter, now we begin describing the implementation of their
+confluence -- a staged abstract interpreter. We present a principled approach to
+derive staged abstract interpreter from its unstaged version. One guiding
+principle of our approach is that the code of the abstract semantics and the
+code that optimizes should be separated. This it is an advantage of using
+staging for abstract interpreters: the designer of the analysis has no need to
+rewrite the analysis, and the performance improvement comes almost for free,
+without any sacrifice of soundness or precision. Unsurprisingly, the staged
+abstract interpreter we present in this section has the same abstract semantics
+as the unstaged version we presented in Section~\ref{unstaged_abs}.
 
 \subsection{Staged Lattices}
-In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@
-to leave space for staging lattices, now we instantiate the type @R@ to @Rep@ and 
+
+In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
+leave space for staging lattices, now we instantiate the type @R@ to @Rep@ and
 still use powersets as an example to present its staged version.
@@
 The type parameter @T:Typ@ of @RepLattice@ requires that the elements of sets
-is can also be staged. Otherwise, without knowing how to stage the elements in the set,
-we can not stage the set either.
-The methods defined operate on type @Rep[Set[T]]@, thus the underlying implementation
-such as @union@ and @intersect@ will be mapped to a node in the IR graph during staging and 
-eventually emitted in the generated code. 
-Again, other structures such as maps and tuples are implemented in a similar way.
+can also be staged. Otherwise, without knowing how to stage the elements in the
+set, we can not stage the set either. The methods defined operate on type
+@Rep[Set[T]]@, thus the underlying implementation such as @union@ and
+@intersect@ will be mapped to a node in the IR graph during staging and
+eventually emitted in the generated code. Again, other structures such as maps
+and tuples are implemented in a similar way.
 
 \subsection{Staged Abstract Semantics}
-We have seen how to obtain a staged concrete semantics based on types, 
-now we take the same approach to obtain a staged abstract semantics.
-The basic operations are largely kept the same as in the unstaged version, except the types are changed to @Rep@.
-Besides, when we update the environment, the identifier @x@ is known statically, but the environment map
-has type @Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to turn it as a next-stage value.
+
+We have seen how to obtain a staged concrete semantics based on types, now we
+take the same approach to obtain a staged abstract semantics. The basic
+operations are largely kept the same as in the unstaged version, except the
+types are changed to @Rep@. Besides, when we update the environment, the
+identifier @x@ is known statically, but the environment map has type
+@Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to turn it as a next-stage
+value.
 
@@
 
-Once more, the way we handle closures is the same as in the staged concrete interpreter:
-the recursive call to @ev@ with the body expression @e@ is compiled and specialized, 
-the wrapper function @f@ will be a field value in a @CompiledClo@ object and be generated
-for the next stage. At the end, we return a singleton set:
+Once more, the way we handle closures is the same as in the staged concrete
+interpreter: the recursive call to @ev@ with the body expression @e@ is compiled
+and specialized, the wrapper function @f@ will be a field value in a
+@CompiledClo@ object and be generated for the next stage. At the end, we return
+a singleton set:
 
@@
 As its unstaged counterpart, function @apply_closures_norep@
-non-deterministically applies multiple target closures with the argument
-and latest store, and finally returns the joined value and a single store.
-We provide the definition of @apply_closures_norep@ in the runtime supporting code.
+non-deterministically applies multiple target closures with the argument and
+latest store, and finally returns the joined value and a single store. We
+provide the definition of @apply_closures_norep@ in the runtime supporting code.
 
@@
 
-Our fixed-point iteration again relies on two caches @in@ and @out@, but the iteration 
-no longer be done at the current stage.
-Because the @in@ and @out@ are both next-stage values, the test of whether 
-@in@ and @out@ are equal is a generated expression in the next stage, and we can only 
-know the comparison result at the next stage. 
-In other words, we do not know how many iterations we need to reach the fixed-point.
-To achieve this, we need to stage a function value --- @iter_aux@ is generated as a 
-recursive function of type @Rep[Unit => (Value,Store)]@ and will be invoked at the next stage.
+Our fixed-point iteration again relies on two caches @in@ and @out@, but the
+iteration no longer be done at the current stage. Because the @in@ and @out@ are
+both next-stage values, the test of whether @in@ and @out@ are equal is a
+generated expression in the next stage, and we can only know the comparison
+result at the next stage. In other words, we do not know how many iterations we
+need to reach the fixed-point. To achieve this, we need to stage a function
+value --- @iter_aux@ is generated as a recursive function of type @Rep[Unit => (Value,Store)]@ 
+and will be invoked at the next stage.
 
@@
 
-However, the instrumented evaluation function that uses the @in@ cache and updates the @out@ cache
-can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
-if it is applied on subexpressions of the analyzed program.
+However, the instrumented evaluation function that uses the @in@ cache and
+updates the @out@ cache can be completely eliminated by staging. Each recursive
+call to @cached_ev@ will also be specialized if it is applied on subexpressions
+of the analyzed program.
 
@@
 
-Now we have already obtained an end-to-end staged abstract interpreter that is able to 
-specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
-which means any operations on a @Map@ becomes code in the next stage.
-But, as we identified when introducing the generic interface, the keys of any environment maps are 
-identifiers in the program, which are completely known statically. 
-This leaves us a chance to further specialize the data structures. 
-Assume the @Map[K,V]@ is implemented as a hash map, if the keys $K$ are known, then the indices can be computed
-statically. Thus the specialized map would be an array @Array[Rep[V]]@ whose elements 
-are next-stage values; all the accesses to the array is determined during staging.
+Now we have already obtained an end-to-end staged abstract interpreter that is
+able to specialize an analysis. However, we treat the data structures such as
+@Map@s as black-boxes, which means any operations on a @Map@ become code in the
+next stage. But, as we identified when introducing the generic interface, the
+keys of any environment maps are identifiers in the program, which are
+completely known statically. This leaves us a chance to further specialize the
+data structures. Assume the @Map[K,V]@ is implemented as a hash map, if the keys
+$K$ are known, then the indices can be computed statically. Thus the specialized
+map would be an array @Array[Rep[V]]@ whose elements are next-stage values; all
+the accesses to the array is determined during staging.
 
 Particularly, if we are specializing a monovariant analysis, the address space
-is equivalent to the identifiers, then the environment component can be entirely eliminated, 
-and the store is a specialized map as array of @Rep[Value]@ elements.
+is equivalent to the identifiers, then the environment component can be entirely
+eliminated, and the store is a specialized map as array of @Rep[Value]@
+elements.
 
@@
 One of the challenges of modern static analysis is program usually depends on
-large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately and reuse the result without losing precision? 
-Then we can reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Indeed, some static analyzers compute summary for a function or a module, which can be reused
-later, however they are mostly too conservative or unsound, which both lead to imprecision.
+large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. Can we analyze
+programs and libraries separately and reuse the result without losing precision?
+Then we can reduce part of the overhead of repeatedly analyzing libraries for
+different programs. Indeed, some static analyzers compute summary for a function
+or a module, which can be reused later, however they are mostly too conservative
+or unsound, which both lead to imprecision.
 
-The specialization of abstract interpreter provides a chance to obtain such partial analysis 
-result in a mechanized way, but still keeps the analysis sound.
-As we see when compiling the closures, we can specialize the abstract interpreter
-with respect the body expression of the lambda term without knowing the actual argument.
-The programs with some unknown variables are open programs, which is exactly the case 
-if we want to analyze programs in a modular way.
+The specialization of abstract interpreter provides a chance to obtain such
+partial analysis result in a mechanized way, but still keeps the analysis sound.
+As we see when compiling the closures, we can specialize the abstract
+interpreter with respect the body expression of the lambda term without knowing
+the actual argument. The programs with some unknown variables are open programs,
+which is exactly the case if we want to analyze programs in a modular way.
 
-For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a whole-program analysis,
-because it is inter-procedural and needs the last $k$ calling contexts to distinguish
-different call sites, where the calling contexts are dynamic values.
-However, it is possible to analyze programs (libraries) separately through specializing an abstract interpreter 
-that generates the analysis as the next-stage program and leave the unavailable programs and calling contexts 
-as dynamic parameters, and then install these contexts when we have the whole program.
+For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a
+whole-program analysis, because it is inter-procedural and needs the last $k$
+calling contexts to distinguish different call sites, where the calling contexts
+are dynamic values. However, it is possible to analyze programs (libraries)
+separately through specializing an abstract interpreter that generates the
+analysis as the next-stage program and leave the unavailable programs and
+calling contexts as dynamic parameters, and then install these contexts when we
+have the whole program.
 
 \iffalse
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then maybe there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-These $(n-m)$ programs might be (abstract) arguments for the first $n$ programs, and
-the abstract interpreter itself might be a partial abstract interpreter.
+Another perspective: programs are data for an abstract interpreter, so if we
+have $n$ programs, then maybe there can be $n$ stages. Probably we can analyze
+first $m$ programs, and generate a residual abstract interpreter waiting for the
+rest $(n-m)$ programs. These $(n-m)$ programs might be (abstract) arguments for
+the first $n$ programs, and the abstract interpreter itself might be a partial
+abstract interpreter.
 \fi
@@
 \subsection{Discussion}
-%We have presented all of our four different interpreters as shown in Figure~\ref{confluence}.
-In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11} provided guidelines
-on what to do and not to do when specializing a concrete interpreter \cite{10.1007/3-540-61580-6_11}. 
-We borrow these guidelines and extend them to abstract interpreters. We discuss decisions we made to 
-achieve this and examine some alternatives.
 
-\paragraph{Big-step vs Small-step.}
-What we implemented is a big-step, compositional abstract interpreter, 
-where "compositional" means that every recursive call of our abstract interpreter is applied to proper 
-substructures of the current syntactic parameters \cite{10.1007/3-540-61580-6_11}.
-This compositionality ensures that specialization can be done by unfolding,
-as well as that the specialization procedure terminates.
-Nevertheless, it is also possible to specialize a small-step operational abstract semantics ---
-\citeauthor{Johnson:2013:OAA:2500365.2500604}
-showed this in abstract compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their optimizations
-of Abstract Abstract Machines \cite{Johnson:2013:OAA:2500365.2500604}. 
-However, the generated abstract bytecode still requires another small-step abstract machine 
-to execute, which is an additional engineering efforts.
-Another alternative approach for efficient specialization is to write the abstract interpreter 
-in a big-step, monadic style \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11}
+provided guidelines on what to do and not to do when specializing a concrete
+interpreter \cite{10.1007/3-540-61580-6_11}. We borrow these guidelines and
+extend them to abstract interpreters. We discuss decisions we made to achieve
+this and examine some alternatives.
+
+\paragraph{Big-step vs Small-step}
+
+What we implemented is a big-step, compositional abstract interpreter, where
+"compositional" means that every recursive call of our abstract interpreter is
+applied to proper substructures of the current syntactic parameters
+\cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
+specialization can be done by unfolding, as well as that the specialization
+procedure terminates. Nevertheless, it is also possible to specialize a
+small-step operational abstract semantics ---
+\citeauthor{Johnson:2013:OAA:2500365.2500604} showed this in abstract
+compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their
+optimizations of Abstract Abstract Machines
+\cite{Johnson:2013:OAA:2500365.2500604}. However, the generated abstract
+bytecode still requires another small-step abstract machine to execute, which is
+an additional engineering efforts. Another alternative approach for efficient
+specialization is to write the abstract interpreter in a big-step, monadic style
+\cite{DBLP:journals/pacmpl/DaraisLNH17}.
 
@@
 
-\paragraph{Correctness and Soundness.}
-Based on the assumption that LMS preserves the semantics during staging,
-we are confident that the staged abstract interpreter does the same analysis compared 
-with the unstaged one. Moreover, the optimization done by staging does not compromise
-any soundness.
+\paragraph{Correctness and Soundness}
+
+Based on the assumption that LMS preserves the semantics during staging, we are
+confident that the staged abstract interpreter does the same analysis compared
+with the unstaged one. Moreover, the optimization done by staging does not
+compromise any soundness.

### 2019-02-04 paper/50_cases.tex
reorganize and address typos

@@
 
-In Section~\ref{sai}, we have shown that staging an abstract interpreter is feasible and can be 
-systematic after correctly identifying the binding-times, despite the fact that the abstract 
-interpreter is intended to be imprecise and easy to implement. 
-In this section, we conduct several case studies to show that this approach is also useful and
-widely applicable to different analyses.
+In Section~\ref{sai}, we have shown that staging an abstract interpreter is
+feasible and can be systematic after correctly identifying the binding-times,
+despite the fact that the abstract interpreter is intended to be imprecise and
+easy to implement. In this section, we conduct several case studies to show that
+this approach is also useful and widely applicable to different analyses.
 
@@
 
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC) as a new
-implementation technique for abstract interpretation based static analysis \cite{Boucher:1996:ACN:647473.727587}.
-The idea is inspired by partial evaluation similar to the present paper -- the program can be known 
-statically, the overhead of interpretation can be eliminated. 
-In AC, the compiled analysis can be represented by either text or closures (higher-order functions);
-though the closures can be executed immediately, the textual programs need to be compiled and loaded first.
+\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC)
+as a new implementation technique for abstract interpretation based static
+analysis \cite{Boucher:1996:ACN:647473.727587}. The idea is inspired by partial
+evaluation similar to the present paper -- the program can be known statically,
+the overhead of interpretation can be eliminated. In AC, the compiled analysis
+can be represented by either text or closures (higher-order functions); though
+the closures can be executed immediately, the textual programs need to be
+compiled and loaded first.
 
-Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a monovariant control-flow 
-analysis \cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007} 
-for continuation-passing style (CPS) programs. Since the analyzed program is written in CPS, the analyzer 
-is essentially a big-step control-environment abstract interpreter.
-Closure generation compiles the analysis as a closure taking an environment argument.
-The overhead of traversing the abstract syntax tree of input program also has been eliminated.
+Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a
+monovariant control-flow analysis \cite{Shivers:1991:SSC:115865.115884,
+Shivers:1988:CFA:53990.54007} for continuation-passing style (CPS) programs.
+Since the analyzed program is written in CPS, the analyzer is essentially a
+big-step control-environment abstract interpreter. Closure generation compiles
+the analysis as a closure taking an environment argument. The overhead of
+traversing the abstract syntax tree of input program also has been eliminated.
 
-In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation can be 
-understood and implemented as an instance of staging abstract interpreters.
-We first revisit the original implementation of abstract compilation of 0-CFA,
-and then reproduce their result by simply adding stage annotations.
-The generated program of our approach improves approximately the same extent of speed,
-but without changing a single line of the analyzer program (with the use of LMS).
-However, closure generation requires more engineering effort, specifically a whole-program 
-conversion on the analyzer. Moreover, as shown in Section~\ref{staged_ds}, 
-our approach is able to not only remove the interpretive overhead, but also specialize 
-the data structures used in the analysis, for example, the environment that maps variables to sets of lambda.
+In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s
+abstract compilation can be understood and implemented as an instance of staging
+abstract interpreters. We first revisit the original implementation of abstract
+compilation of 0-CFA, and then reproduce their result by simply adding stage
+annotations. The generated program of our approach improves approximately the
+same extent of speed, but without changing a single line of the analyzer program
+(with the use of LMS). However, closure generation requires more engineering
+effort, specifically a whole-program conversion on the analyzer. Moreover, as
+shown in Section~\ref{staged_ds}, our approach is able to not only remove the
+interpretive overhead, but also specialize the data structures used in the
+analysis, for example, the environment that maps variables to sets of lambda.
 
@@
 
-The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA for a CPS language consisting
-of lambda terms, applications, @letrec@ and primtive operators. The analyses for different syntactic 
-constructs are decomposed into different functions, such as @analyzeCall@ and @analyzeApp@.
-The idea of closure generation is to rewrite these functions. Where previously they may take both static
-arguments and dynamic arguments, after the rewrite only the static arguments is taken.
-In this case, the static arguments are syntactic terms;
-the dynamic arguments are stores.
-After written in AC style, functions like @compCall@ (compiled version of @analyzeCall@) return a value of type
-@CompAnalysis@, i.e., a closure that takes a store and returns a store.
-The result of multiple calls on such functions, for example,
-@compCall@ and @compArgs@, can be composed.
-The generated closure only takes stores, because the input program
-is specialized into the closure. The code is shown in Figure~\ref{compare_ac_sai} (left).
+The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
+for a CPS language consisting of lambda terms, applications, @letrec@ and
+primtive operators. The analyses for different syntactic constructs are
+decomposed into different functions, such as @analyzeCall@ and @analyzeApp@. The
+idea of closure generation is to rewrite these functions. Where previously they
+may take both static arguments and dynamic arguments, after the rewrite only the
+static arguments is taken. In this case, the static arguments are syntactic
+terms; the dynamic arguments are stores. After written in AC style, functions
+like @compCall@ (compiled version of @analyzeCall@) return a value of type
+@CompAnalysis@, i.e., a closure that takes a store and returns a store. The
+result of multiple calls on such functions, for example, @compCall@ and
+@compArgs@, can be composed. The generated closure only takes stores, because
+the input program is specialized into the closure. The code is shown in
+Figure~\ref{compare_ac_sai} (left).
 
@@
 
-On the other side, our approach does exactly the same thing through staging: the syntactic terms are static, and 
-stores are dynamic, therefore the generated code just looks-up and updates the store.
-Figure~\ref{compare_ac_sai} (right) shows the code written with LMS. In fact, the only changes are
-the type of @Store@ is replaced with @Rep[Store]@ indicating that the values of type @Store@  will be known
-at the next stage. Indeed, additional engineering efforts are required to make this happen, 
-including: the staged version @Map@ which is already included in LMS; implicit @lift@ function that transform a
-current-stage constant value to next stage; next-stage representation of the syntactic terms, i.e., proper @toString@
-functions of AST structs. We consider these efforts relatively small, and they do not interfere the actual analysis
-we desire.
+On the other side, our approach does exactly the same thing through staging: the
+syntactic terms are static, and stores are dynamic, therefore the generated code
+just looks-up and updates the store. Figure~\ref{compare_ac_sai} (right) shows
+the code written with LMS. In fact, the only changes are the type of @Store@ is
+replaced with @Rep[Store]@ indicating that the values of type @Store@ will be
+known at the next stage. Indeed, additional engineering efforts are required to
+make this happen, including: the staged version @Map@ which is already included
+in LMS; implicit @lift@ function that transform a current-stage constant value
+to next stage; next-stage representation of the syntactic terms, i.e., proper
+@toString@ functions of AST structs. We consider these efforts relatively small,
+and they do not interfere the actual analysis we desire.
 
@@
 
-The target language we presented in Section~\ref{bg_lang} is essentially a higher-order functional language.
-One fundamental analysis task for functional programs is control-flow analysis, i,e., determining
-which functions will possibly be applied at each call-site. 
-The abstract interpreter we used in Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 
-0-CFA-like abstraction, moreover it is also a pushdown control-flow analysis; 
-in the last section, we also reviewed AC with finite-state 0-CFA.
-In this section, based on the existing staged abstract interpreter, we further
-develop the staging techniques with control-flow analyses, including recoverying a more precise store model and
-adding context-sensitivity to the analysis.
-%, and using staging as an implementation stategy for mixed sensitivity.
+The target language we presented in Section~\ref{bg_lang} is essentially a
+higher-order functional language. One fundamental analysis task for functional
+programs is control-flow analysis, i,e., determining which functions will
+possibly be applied at each call-site. The abstract interpreter we used in
+Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 0-CFA-like
+abstraction, moreover it is also a pushdown control-flow analysis; in the last
+section, we also reviewed AC with finite-state 0-CFA. In this section, based on
+the existing staged abstract interpreter, we further develop the staging
+techniques with control-flow analyses, including recoverying a more precise
+store model and adding context-sensitivity to the analysis.
 
@@
 
-The store-widened analysis we implemented in Section~\ref{sai} uses a single store to approximate
-the runtime store. It can be efficently computed in polynomial time together with 0-CFA-like abstraction, 
-but sometimes we may desire a 
-more precise result that distinguishes the final values and stores for different closure targets. 
-To achieve this goal,
-we need to tweak our abstract interpreter and type instantiation. The answer type @Ans@ is changed to
-a set of @VS@s where a @VS@ is a pair of sets of abstract values (such as closures or abstract numbers)
-and a store.
+The store-widened analysis we implemented in Section~\ref{sai} uses a single
+store to approximate the runtime store. It can be efficently computed in
+polynomial time together with 0-CFA-like abstraction, but sometimes we may
+desire a more precise result that distinguishes the final values and stores for
+different closure targets. To achieve this goal, we need to tweak our abstract
+interpreter and type instantiation. The answer type @Ans@ is changed to a set of
+@VS@s where a @VS@ is a pair of sets of abstract values (such as closures or
+abstract numbers) and a store.
 
@@
 
-Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under staging the type @Ans@
-represents a next stage value. Then, our generic interpreter is also changed when handling function 
-applications.
+Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under
+staging the type @Ans@ represents a next stage value. Then, our generic
+interpreter is also changed when handling function applications.
 
@@
 
-The idea to handle the application is to explore all possible closures from @e1@ and meanwhile
-use the latest store.
-What @choices@ does is similar to McCarthy's @amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
-element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side receiver, i.e., @e1vs@.
-The function @choices@ internally uses the delimited control operator @shift@ to capture the continuation,
-which is the code block after its call site. 
-This allows us to perform nesting depth-first evaluation for function application while still writing 
-the program in direct-style \cite{Wei:2018:RAA:3243631.3236800}. 
-Again, we can stage this part as we did for the naive abstract interpreter.
+The idea to handle the application is to explore all possible closures from @e1@
+and meanwhile use the latest store. What @choices@ does is similar to McCarthy's
+@amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
+element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side
+receiver, i.e., @e1vs@. The function @choices@ internally uses the delimited
+control operator @shift@ to capture the continuation, which is the code block
+after its call site. This allows us to perform nesting depth-first evaluation
+for function application while still writing the program in direct-style
+\cite{Wei:2018:RAA:3243631.3236800}. Again, we can stage this part as we did for
+the naive abstract interpreter.
 
@@
 
-We add k-CFA-like context-sensitivity to the analysis by introducing an abstract timestamp, 
-whose concrete instantiation is a finite list of expressions that track $k$ recent calling contexts.
-The definition of abstract addresses is changed to a tuple of identifiers and the time it get allocated,
-meaning that this address points to some values under such calling context.
-If $k$ is 0, we obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain
-a family of analysis with increasing precision.
+We add k-CFA-like context-sensitivity to the analysis by introducing an abstract
+timestamp, whose concrete instantiation is a finite list of expressions that
+track $k$ recent calling contexts. The definition of abstract addresses is
+changed to a tuple of identifiers and the time it get allocated, meaning that
+this address points to some values under such calling context. If $k$ is 0, we
+obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain a
+family of analysis with increasing precision.
 
@@
 
-Every time when we call the @eval@ function, we refresh the timestamp by calling function @tick@,
-which returns a new timestamp. Here we adopt a $k$-CFA-like allocation strategy, therefore
-the @tick@ function can be implemented as appending the current expression being evaluated to the existing
-calling context, and then taking the first $k$ elements from the list.
+Every time when we call the @eval@ function, we refresh the timestamp by calling
+function @tick@, which returns a new timestamp. Here we adopt a $k$-CFA-like
+allocation strategy, therefore the @tick@ function can be implemented as
+appending the current expression being evaluated to the existing calling
+context, and then taking the first $k$ elements from the list.
 
@@
 
-Accordingly, the type of return value is accompanied by the timestamp. For a recursive call of @ev@,
-which would also return the latest timestamp, and that timestamp will be used for the evaluation afterward.
+Accordingly, the type of return value is accompanied by the timestamp. For a
+recursive call of @ev@, which would also return the latest timestamp, and that
+timestamp will be used for the evaluation afterward.
 
@@
 
-Using other allocation strategies to achieve different sensitivities is also possible \cite{DBLP:conf/icfp/Gilray0M16} 
-and can be staged under our framework.
+Using other allocation strategies to achieve different sensitivities is also
+possible \cite{DBLP:conf/icfp/Gilray0M16} and can be staged under our framework.
 
@@
 
-Now we consider in a first-order imperative language, we may care more about the data-flow because the control-flow
-is relatively easy to obtain. In this section, we show the staging of other abstract domains, particularly an 
-interval domain for numbers.
-It has been shown that specializing abstract domains with respect to the structure of analyzed program significantly 
-improves the performance: a recent example is online decomposition of polyhedra \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
-In this section, we first show how to support imperative language features in the generic abstract interpreter.
-Then we present a similar idea for the interval domain and show that the specialization is feasible by staging systematically.
+Now we consider in a first-order imperative language, we may care more about the
+data-flow because the control-flow is relatively easy to obtain. In this
+section, we show the staging of other abstract domains, particularly an interval
+domain for numbers. It has been shown that specializing abstract domains with
+respect to the structure of analyzed program significantly improves the
+performance: a recent example is online decomposition of polyhedra
+\cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}. In this
+section, we first show how to support imperative language features in the
+generic abstract interpreter. Then we present a similar idea for the interval
+domain and show that the specialization is feasible by staging systematically.
 
@@
 
-To evaluates an assignment, we first evaluate its right-hand side, and then put the value
-into the slot where the address of @v@ points to.
-For simplicity, we elect to make the value of the assignment be an @void()@ value.
+To evaluates an assignment, we first evaluate its right-hand side, and then put
+the value into the slot where the address of @v@ points to. For simplicity, we
+elect to make the value of the assignment be an @void()@ value.
 
@@
 
-To evaluates a @while@ loop statement, we evaluate the condition first.
-Then similar to how we treat @branch0@, we have a generic @branch@ function but works on boolean
-values. For the @true@ branch, we recursively call @ev@ on a newly constructed expression
-@Seq(e, While(t, e))@ meaning that first evaluates @e@ and the repeat the loop.
-Otherwise, for the other branch, we simply return a void value and current store.
+To evaluates a @while@ loop statement, we evaluate the condition first. Then
+similar to how we treat @branch0@, we have a generic @branch@ function but works
+on boolean values. For the @true@ branch, we recursively call @ev@ on a newly
+constructed expression @Seq(e, While(t, e))@ meaning that first evaluates @e@
+and the repeat the loop. Otherwise, for the other branch, we simply return a
+void value and current store.
 
@@
 
-Interval domain is a relative simple domain which consists of two numeric fields, an upper bound 
-and a lower bound. Here we annotate the upper bound and lower bound fields to be of type @Rep[Double]@.
-The operations such as @+@ of two intervals produce a new
+Interval domain is a relative simple domain which consists of two numeric
+fields, an upper bound and a lower bound. Here we annotate the upper bound and
+lower bound fields to be of type @Rep[Double]@. The operations such as @+@ of
+two intervals produce a new
 

### 2019-02-04 paper/60_eval.tex
reorganize and address typos

@@
 
-To evaluate the performance can be improved through staging, we use the abstract
-interpreter demonstrated in the paper to analyze some Scheme programs. We first implement
-a desugaring transformation for a large subset of scheme that transforms into the
-small language we used in the paper.
-We use the same generic interface with a 0-CFA-like abstraction, and the unstage abstract
-interpreter forms the baseline.
+To evaluate whether the performance can be improved through staging, we use the
+abstract interpreter demonstrated in the paper to analyze some Scheme programs.
+We first implement a desugaring transformation for a large subset of Scheme that
+transforms into the small language we used in the paper. We use the same generic
+interface with a 0-CFA-like abstraction, and the unstage abstract interpreter
+forms the baseline.
 
-All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux kernel 4.4.0)
-machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM.
-Although the machine has 96 cores and 192 threads in total,
-the abstract interpreters only use one thread to run all the benchmark programs.
+All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux
+kernel 4.4.0) machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of
+RAM. Although the machine has 96 cores and 192 threads in total, the abstract
+interpreters only use one thread to run all the benchmark programs.
 
 Our evaluations show that the staged abstract interpreter performs well in some
-cases that are considered to be the worst scenarios for traditional $k$-CFA.
-The staged version outperforms the unstaged abstract interpreter by a significant
+cases that are considered to be the worst scenarios for traditional $k$-CFA. The
+staged version outperforms the unstaged abstract interpreter by a significant
 margin (up to ~230\% performance gain) on these difficult ones.
@@
 \subsection{Benchmarks}
-The benchmark programs we used are collected from previous papers \cite{Johnson:2013:OAA:2500365.2500604,
-ashley:practical, DBLP:journals/corr/abs-1102-3676} and exisiting artifacts \footnote{https://github.com/ilyasergey/reachability}.
+
+The benchmark programs we used are collected from previous papers
+\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
+DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
+\footnote{https://github.com/ilyasergey/reachability}.
 
 We used the following benchmark programs:
+
 \begin{itemize}
@@
 
-Figure~\ref{evaluation_result} shows the benchmark evaluation results. The "unstaged" and "staged" columns
-show the median time to finish the analysis (in milliseconds). The "$\Delta$ of median"
-column shows the performance change from the unstaged abstract interpreter to the staged version.
+Figure~\ref{evaluation_result} shows the benchmark evaluation results. The
+"unstaged" and "staged" columns show the median time to finish the analysis (in
+milliseconds). The "$\Delta$ of median" column shows the performance change from
+the unstaged abstract interpreter to the staged version.
 
-As we can see from the table, the staged version significantly outperforms the unstaged especially 
-for the difficult ones.
-However, for some benchmarks, we observe degraded performance. 
-We conjecture this happens due to the large size of the generated code is unoptimized -- 
-as what we test here is a simple implementation almost identical to what we described in the previous sections,
-we believe the performance still can be improved if more engineering effort is taken, 
-such as optimizing code generation and avoiding code duplication.
+As we can see from the table, the staged version significantly outperforms the
+unstaged especially for the difficult ones. However, for some benchmarks, we
+observe degraded performance. We conjecture this happens due to the fact that
+the large size of the generated code is unoptimized -- as what we test here is a
+simple implementation almost identical to what we described in the previous
+sections, we believe the performance still can be improved if more engineering
+effort is taken, such as optimizing code generation and avoiding code
+duplication.

### 2019-02-04 paper/70_related.tex
reorganize and address typos

@@
 \paragraph{Optimizing Static Analysis Through Specialization}
-The idea in this paper is closely inspired by abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
+The idea in this paper is closely inspired by abstract compilation
+\cite{Boucher:1996:ACN:647473.727587}.
 \citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
 techniques as an efficient implementation of the monovariant flow analysis
-(\textit{0}-CFA) for programs written in continuation-passing style. 
-The key idea is to remove the interpretation overhead on traversing the syntax 
-tree by partial evaluation. Specifically, they proposed two similar kinds of abstract compilation techniques.
-The first one is to generate specialized analysis as a textual program, which then
-can be loaded and executed by \texttt{eval} or other similar mechanisms. 
-The second one is to use closures, i.e., functions that remember their environments, 
-as a representation of specialized analysis. As we show in the case study,
-compiling the analysis generates higer-order functions on-the-fly with respect to 
-the analyzed program, then the generated closure can be applied immediately in the 
-higher-order host language.
+(\textit{0}-CFA) for programs written in continuation-passing style. The key
+idea is to remove the interpretation overhead on traversing the syntax tree by
+partial evaluation. Specifically, they proposed two similar kinds of abstract
+compilation techniques. The first one is to generate specialized analysis as a
+textual program, which then can be loaded and executed by \texttt{eval} or other
+similar mechanisms. The second one is to use closures, i.e., functions that
+remember their environments, as a representation of specialized analysis. As we
+show in the case study, compiling the analysis generates higer-order functions
+on-the-fly with respect to the analyzed program, then the generated closure can
+be applied immediately in the higher-order host language.
 
-\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure generation 
-to optimize small-step abstract interpreter in state-transition style \cite{Johnson:2013:OAA:2500365.2500604}.
-The analyzed program is firstly compiled to a intermediate representation called 
-"abstract bytecode", which are actually higher-order functions, and then be executed 
-on a abstract abstract machine for that IR.
+\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
+generation to optimize small-step abstract interpreter in state-transition style
+\cite{Johnson:2013:OAA:2500365.2500604}. The analyzed program is firstly
+compiled to a intermediate representation called "abstract bytecode", which are
+actually higher-order functions, and then be executed on a abstract abstract
+machine for that IR.
 
-\citet{damian1999partial} provides a formal treatment to abstract compilation and Shiver's CFA, 
-as well as proofs to establish correctness and certified specialized analyzers \cite{damian1999partial}.
-\citeauthor{amtoft1999partial} applyies partial evaluation for constraint-based 
-control flow analysis \cite{amtoft1999partial}.
-Split the analysis to be multiple stages is also studied other than control-flow analysis,
-though the formulation may very different.
-For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to flow-sensitive pointer 
-analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first stage is to analyze the program code to 
-obtain a sparse representation, and then the second stage conducts the flow-sensitive analysis.
+\citet{damian1999partial} provides a formal treatment to abstract compilation
+and Shiver's CFA, as well as proofs to establish correctness and certified
+specialized analyzers \cite{damian1999partial}. \citeauthor{amtoft1999partial}
+applyies partial evaluation for constraint-based control flow analysis
+\cite{amtoft1999partial}. Split the analysis to be multiple stages is also
+studied other than control-flow analysis, though the formulation may very
+different. For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to
+flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first
+stage is to analyze the program code to obtain a sparse representation, and then
+the second stage conducts the flow-sensitive analysis.
 
-\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domain through online decomposition \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}.
-The idea is to decompose a large polyhedra into several smaller one; one abstract transformer
-usually only depends a limit number of variables, then we just need to compute the resulting polyhera
-on those smaller polyhedra that contains these related variables.
-The decomposition can be considered as a specialization of abstract domain with respect to the program
-static structure.
+\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domain
+through online decomposition \cite{DBLP:conf/popl/SinghPV17,
+Singh:2017:PCD:3177123.3158143}. The idea is to decompose a large polyhedra into
+several smaller one; one abstract transformer usually only depends a limit
+number of variables, then we just need to compute the resulting polyhera on
+those smaller polyhedra that contains these related variables. The decomposition
+can be considered as a specialization of abstract domain with respect to the
+program static structure.
 
-\paragraph{Abstract Interpreters}
-Abstract interpretation was proposed as a semantic-based approach to construct sound static analysis by approximation \cite{DBLP:conf/popl/CousotC77}.
-As semantic artifacts, the Abstracting Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
-approach shows that abstract interpreters can be derived systematically from concrete semantic artifacts.
-The big-step abstract interpreters we presented in this paper are also inspired by the AAM framework,
-which are later adopted for big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}.
-\citet{DBLP:conf/cc/CousotC02} also proposed an abstract interpretation framework for modular analysis.
-\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by using bi-abduction.
-The big-step abstract interpreter we presented in this paper is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800}. 
-\citet{Keidel:2018:CSP:3243631.3236767} demonstrate a similar implementation of definitional abstract interpreter using arrows.
+\paragraph{Abstract Interpreters} Abstract interpretation was proposed as a
+semantic-based approach to construct sound static analysis by approximation
+\cite{DBLP:conf/popl/CousotC77}. As semantic artifacts, the Abstracting Abstract
+Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} approach
+shows that abstract interpreters can be derived systematically from concrete
+semantic artifacts. The big-step abstract interpreters we presented in this
+paper are also inspired by the AAM framework, which are later adopted for
+big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17,
+Wei:2018:RAA:3243631.3236800}. \citet{DBLP:conf/cc/CousotC02} also proposed an
+abstract interpretation framework for modular analysis.
+\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by
+using bi-abduction. The big-step abstract interpreter we presented in this paper
+is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17,
+Wei:2018:RAA:3243631.3236800}. \citet{Keidel:2018:CSP:3243631.3236767}
+demonstrate a similar implementation of definitional abstract interpreter using
+arrows.
 
-\paragraph{Meta-Programming}
-The Lightweight Modular Staging framework relies on types as stage annotations, and the staging
-procedure is modular.
-Other notable implementations of MSP exist in ML family, for examples, MetaML \cite{DBLP:conf/pepm/TahaS97} 
-and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
-Compared with the LMS approach in Scala, MetaML/MetaOCaml use term-level annotations such as brackets, 
-escape, and run. Notwithstanding, we use LMS in this paper, the idea of staging an abstract interpreter
-still applies with other MSP implementations.
-Partial evaluation as an automatic technique is studied comprehensively
-\cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
-Futamura projections reveals close relations between interpreters and compilers \cite{Futamura1999, futamura1971partial}.
-\citet{DBLP:conf/gpce/Asai14} studies the compilation of reflective language using MSP and MetaOCaml.
-\citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of concrete interpreters by MSP and LMS framework.
+\paragraph{Meta-Programming} The Lightweight Modular Staging framework relies on
+types as stage annotations, and the staging procedure is modular. Other notable
+implementations of MSP exist in ML family, for examples, MetaML
+\cite{DBLP:conf/pepm/TahaS97} and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
+DBLP:conf/flops/Kiselyov14}. Compared with the LMS approach in Scala,
+MetaML/MetaOCaml use term-level annotations such as brackets, escape, and run.
+Notwithstanding, we use LMS in this paper, the idea of staging an abstract
+interpreter still applies with other MSP implementations. Partial evaluation as
+an automatic technique is studied comprehensively
+\cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}. Futamura projections
+reveals close relations between interpreters and compilers \cite{Futamura1999,
+futamura1971partial}. \citet{DBLP:conf/gpce/Asai14} studies the compilation of
+reflective language using MSP and MetaOCaml.
+\citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of
+concrete interpreters by MSP and LMS framework.

### 2019-02-11 paper/10_intro.tex
revise

@@
 the program, i.e., they can be known whether statically or dynamically, and then
-evaluates the static part, and generates a residual program that relies on the
-dynamic part. However, precisely analyzing binding-time for an arbitrary program
-is hard in general. As an alternative and pragmatic approach to specialization
-and partial evaluation, multi-stage programming (MSP for short)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage annotations
-(i.e., binding-time annotations) from the programmers. These staging annotations
-identify which part of the input program should be specialized. As a classical
-example, we consider a power function:
+evaluates the static part, and finally generates a residual program that solely
+relies on the dynamic part. However, given an arbitrary program, precisely
+analyzing its binding-time is hard in general. As an alternative and pragmatic
+approach to specialization and partial evaluation, multi-stage programming (MSP
+for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
+annotations (i.e., binding-time annotations) to be explict from the programmers.
+These staging annotations identify which part of the inputs should be
+specialized. As a classical example, we consider the power function:
 
@@
 
-If the program identifies that @b@ will be known dynamically (as shown in its
-type @Rep[Int]@ -- a representation of @Int@) and @e@ is statically known, say
-5, then we can specialize the function \texttt{power} with respect to \texttt{e
-= 5} and generate a new specialized function that runs faster:
+If the programmer identifies that @b@ will be known in the future stage (as shown
+in its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known in the
+current stage, say 5, then we can specialize the function @power@ with
+respect to @e = 5@ and generate a specialized function that runs faster:
 
@@
 
-Abstract interpretation \cite{DBLP:conf/popl/CousotC77} as a semantic approach
-to construct sound static analysis by approximation is almost as old as Futamura
-projections. Some recent progress such as Abstracting Abstract Machines (AAM)
+Roughly at the same time when Futamura projections was visioned in 1970s,
+\citeauthor{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
+lattice-based semantic approach to construct sound static analysis, by
+approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}.
+Some recent progress such as Abstracting Abstract Machines (AAM)
 uncovers a methodology to derive sound abstract interpreters from their concrete
@@
 
+\note{
+  Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},
+  deep embedding, four different interpretations.
+}
+
 \subsection{Contributions}

### 2019-02-11 paper/paper.tex
reorganize and address typos
revert to review mode
revise

@@
-\documentclass[sigplan,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[sigplan,nonacm]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-% \documentclass[sigplan,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 
-%% Conference information
+%% Journal information
 %% Supplied to authors by publisher for camera-ready submission;
 %% use defaults for review submission.
-%\acmConference[Submitted to PLDI'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
-\acmConference[Submitted to PLDI]{}{June 2019}{Phoenix, Arizona}
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
 \acmYear{2019}
-%\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
-%\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
 \startPage{1}
@@
 %% Citation style
-%\citestyle{acmauthoryear}  %% For author/year citations
+\citestyle{acmauthoryear}  %% For author/year citations
 %\citestyle{acmnumeric}     %% For numeric citations
@@
 
+\iffalse
+
 \renewcommand\thefootnotecopyrightpermission{}
@@
 \fancyhead[RO,LE]{Preprint, November 2018}
+
+\fi
 

### 2019-02-12 paper/10_intro.tex
code & revision

@@
 for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
-annotations (i.e., binding-time annotations) to be explict from the programmers.
+annotations (i.e., binding-time annotations) to be explicit from the programmers.
 These staging annotations identify which part of the inputs should be
-specialized. As a classical example, we consider the power function:
+specialized. As a classical example, we consider the power function to introduce
+the idea of MSP:
 
@@
 If the programmer identifies that @b@ will be known in the future stage (as shown
-in its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known in the
+on its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known in the
 current stage, say 5, then we can specialize the function @power@ with
@@
 approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}.
-Some recent progress such as Abstracting Abstract Machines (AAM)
-uncovers a methodology to derive sound abstract interpreters from their concrete
-counterparts and has been applied to different variants of definitional
+Recent progress such as Abstracting Abstract Machines (AAM) uncovers a
+methodology to derive sound abstract interpreters from their concrete
+counterparts, and it has been applied to different variants of definitional
 interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
-DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}. Given the structural
-similarity between concrete and abstract interpreters, intellectually it is
-natural to raise the question whether it is possible to specialize an abstract
-interpreter, and how can we effectively do that having considered their
-different functionalities. This paper studies the confluence of two old ideas
---- Futamura projection and abstract interpretation, but from the perspective of
-their recent realizations --- multi-stage programming and definitional abstract
-interpreters. To be specific, we present the application of the first Futamura
-projection on abstract interpreters, and the approach of constructing optimizing
-abstract interpreters by multi-stage programming.
+  DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+Given the structural similarity between concrete and abstract interpreters,
+intellectually it is natural to raise the question whether it is possible to
+specialize an abstract interpreter, and how can we effectively do that having
+considered their different functionalities.
+
+This paper studies the confluence of two old ideas --- Futamura projection and
+abstract interpretation, but from the perspective of their recent realizations
+--- multi-stage programming and definitional abstract interpreters.
+To be specific, we present the application of the first Futamura projection on
+definitional abstract interpreters, which intersects as staged abstract interpreters.
+By borrowing the ideas from \note{monadic interpreters} and embedded
+domain-specific languages, we develop an approach to fulfill four different
+semantics (Figure~\ref{confluence}) in a unified framework.
+
+% FIXME: and the approach of constructing optimizing abstract interpreters by multi-stage programming.
 
@@
 
+\note{besides the intellectual merit of the confluence diagram, practical
+problem: can we speed up a static analyzer without pervasive changes of the
+abstract semantics, even without changing a single line of the abstract
+interpreter?}
+
+\todo{Move some into related work}
 Multi-stage programming has been widely used to improve the performance in many

### 2019-02-12 paper/21_bg_interp.tex
code & revision

@@
 abstract class Expr
-case class Lit(i: Int)
-case class Var(x: String)
-case class Lam(x: String, e: Expr)
-case class App(e1: Expr, e2: Expr)
-case class AOp(op: String, e1: Expr, e2: Expr)
-case class Rec(x: String, e: Expr, body: Expr)
-case class If0(cnd: Expr, thn: Expr, els: Expr)
+case class Lit(i: Int)                         // numbers
+case class Var(x: String)                      // variables
+case class Lam(x: String, e: Expr)             // lambda abstractions
+case class App(e1: Expr, e2: Expr)             // applications
+case class AOp(op: String, e1: Expr, e2: Expr) // primitive operations
+case class Rec(x: String, rhs: Expr, e: Expr)  // recursive bindings
+case class If0(e1: Expr, e2: Expr, e3: Expr)   // conditionals
 \end{lstlisting}

### 2019-02-12 paper/10_intro.tex
revise intro

@@
 
+\todo{example of interpreter -> compiler, mix}
+
 Roughly at the same time when Futamura projections was visioned in 1970s,
@@
 lattice-based semantic approach to construct sound static analysis, by
-approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}.
-Recent progress such as Abstracting Abstract Machines (AAM) uncovers a
-methodology to derive sound abstract interpreters from their concrete
-counterparts, and it has been applied to different variants of definitional
-interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
+approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}. However,
+constructing sound abstract interpreters was considered abstruse and complicated
+for a long time.
+Recent progress such as Abstracting Abstract Machines (AAM)
+uncovers a systematic method to derive sound abstract interpreters from their
+concrete counterparts, where the soundness can be easily established by
+examining the transformation of semantics.
+The AAM approach also has been been applied to different variants of
+definitional interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
 DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-Given the structural similarity between concrete and abstract interpreters,
-intellectually it is natural to raise the question whether it is possible to
-specialize an abstract interpreter, and how can we effectively do that having
-considered their different functionalities.
+\todo{CEK example?}
 
-This paper studies the confluence of two old ideas --- Futamura projection and
+Now we have seen two orthogonal transformations of interpreters: by
+speclization, we can refactor interpreters to compilers; by abstraction, we can
+refactor interpreters to static analyzers. As an intellectual quest, it is
+natural to ask, starting from an interpreter, can we derive sound abstract
+interpreters but it works like a compiler and generates fast analysis code?
+
+%artifacts of concrete semantics can be refactored to abstract interpreters.
+In this paper, we study the confluence of two old ideas --- Futamura projection and
 abstract interpretation, but from the perspective of their recent realizations
@@
 domain-specific languages, we develop an approach to fulfill four different
-semantics (Figure~\ref{confluence}) in a unified framework.
+semantics (Figure~\ref{confluence}) in a unified framework. \todo{explain four sem}
 
@@
 
-\note{besides the intellectual merit of the confluence diagram, practical
-problem: can we speed up a static analyzer without pervasive changes of the
-abstract semantics, even without changing a single line of the abstract
-interpreter?}
+Besides the intellectual merit of the confluence diagram, this paper contributes
+to a practical problem in static analysis: can we speed up a static analyzer
+without any intrusive changes and soundness compromises?
 

### 2019-02-12 paper/10_intro.tex
code & revision

@@
 refactor interpreters to static analyzers. As an intellectual quest, it is
-natural to ask, starting from an interpreter, can we derive sound abstract
+natural to ask --- starting from an interpreter, can we derive sound abstract
 interpreters but it works like a compiler and generates fast analysis code?
@@
 without any intrusive changes and soundness compromises?
+\todo{specializing analysis wrt program is not new; first staging; we argue that
+staging is less intrusive compared with AC (ie, closure generation), in
+fact,LMS-based staging only requires type annotations }
 

### 2019-02-13 paper/10_intro.tex
code & revision

@@
 \begin{lstlisting}
-def power5(b: Int) = b * b * b * b * b
+def power5(b: Int) = b * b * b * b * b * 1
 \end{lstlisting}
@@
 
-%artifacts of concrete semantics can be refactored to abstract interpreters.
-In this paper, we study the confluence of two old ideas --- Futamura projection and
-abstract interpretation, but from the perspective of their recent realizations
---- multi-stage programming and definitional abstract interpreters.
+In this paper, we study the confluence of two old ideas --- Futamura projection
+and abstract interpretation, but from the perspective of their recent
+realizations --- multi-stage programming and definitional abstract interpreters.
 To be specific, we present the application of the first Futamura projection on
-definitional abstract interpreters, which intersects as staged abstract interpreters.
-By borrowing the ideas from \note{monadic interpreters} and embedded
-domain-specific languages, we develop an approach to fulfill four different
-semantics (Figure~\ref{confluence}) in a unified framework. \todo{explain four sem}
+definitional abstract interpreters, which intersects as staged abstract
+interpreters. By borrowing the ideas from \note{monadic interpreters}
+\cite{DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
+Sergey:2013:MAI:2491956.2491979} and embedded domain-specific languages
+\cite{DBLP:conf/snapl/RompfBLSJAOSKDK15, DBLP:journals/jfp/CaretteKS09,
+DBLP:conf/icfp/GibbonsW14}, we develop an approach to fulfill four different
+semantics (Figure~\ref{confluence}) in a unified framework. The four semantics
+share one same generic interpreter and some abstract types; by instantiating the
+abstract types differently, the interpreter behaves as a concrete interpreter,
+an abstracter interperter, or their staged versions, respectively.
 

### 2019-02-18 code/src/main/scala/sai/common/IfThenElse.scala
fix a code dup bug

@@ -25,12 +25,8 @@ trait MyScalaGenIfThenElse extends ScalaGenEffect with BaseGenIfThenElse {
       stream.println("}")
 
       stream.println("val " + if_sym + " = if (" + quote(c) + ") {")
-      //emitBlock(a)
-      //stream.println(quote(getBlockResult(a)))
       stream.println(then_sym + "()")
       stream.println("} else {")
-      //emitBlock(b)
-      //stream.println(quote(getBlockResult(b)))
       stream.println(else_sym + "()")
       stream.println("}")
     case _ => super.emitNode(sym, rhs)

### 2019-02-18 code/src/main/scala/sai/direct/core/ai/AbsLC.scala

@@ -36,9 +36,7 @@ object AbsLamCal {
     def get(ρ: Env, x: Ident): Addr = ρ(x)
     def put(ρ: Env, x: Ident, a: Addr): Env = ρ + (x → a)
     def get(σ: Store, a: Addr): Value = σ.getOrElse(a, Lattice[Value].bot)
-    def put(σ: Store, a: Addr, v: Value): Store = {
-      val oldv = get(σ, a); σ + (a → (v ⊔ oldv))
-    }
+    def put(σ: Store, a: Addr, v: Value): Store = { val oldv = get(σ, a); σ + (a → (v ⊔ oldv)) }
     def alloc(σ: Store, x: Ident): Addr = Addr(x)
     def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
     def num(i: Lit): Value = Set(NumV())
@@ -53,7 +51,9 @@ object AbsLamCal {
       }
       (vs.reduce(Lattice[Value].⊔), σ0)
     }
-    def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = { thn ⊔ els }
+    def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = {
+      thn ⊔ els
+    }
     def prim_eval(op: Symbol, v1: Value, v2: Value): Value = Set(NumV())
 
     type Config = (Expr, R[Env], R[Store])
@@ -149,14 +149,13 @@ object AbsLamCal {
         val α = alloc(σ, x)
         ev(e, put(ρ, x, α), put(σ, α, args))
       }
-      //unchecked[Value](s"Set[AbsValue](CompiledClo(${fun(f)}, $λ, $ρ))")
       unchecked[Value]("Set[AbsValue](CompiledClo(", fun(f), ",", λ, ",", ρ, "))")
     }
     def num(i: Lit): Rep[Value] = unchecked[Value]("Set[AbsValue](NumV())")
     def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
       val thnans = thn
       val elsans = els
-      (RepLattice[Value].⊔(thn._1, els._1), RepLattice[Store].⊔(thn._2, els._2))
+      (RepLattice[Value].⊔(thnans._1, elsans._1), RepLattice[Store].⊔(thnans._2, elsans._2))
     }
[…]

### 2019-02-18 code/src/main/scala/sai/direct/core/parser/AST.scala

@@ -19,6 +19,7 @@ case class Bind(x: String, e: Expr) {
 }
 case class Let(x: String, e: Expr, body: Expr) extends Expr {
   def toApp: App = App(Lam(x, body), e)
+  override def toString = "Let(\"" + x + "\"," + e + "," + body + ")"
 }
 case class Lrc(bds: List[Bind], body: Expr) extends Expr {
   require(bds.nonEmpty)

### 2019-02-18 code/src/main/scala/sai/Evaluation4.scala
evaluation

@@ -1,60 +1,0 @@
-package sai
-
-import scala.io.Source
-
-import sai.utils._
-import sai.direct.large.ai._
-import sai.evaluation.TestPrograms._
-import sai.direct.large.parser._
-
-object Evaluation extends AbsLamCalTrait {
-
-  def main(args: Array[String]) {
-    val progs = List(
-      fib,
-      euclid,
-      euclid_imp
-    )
-
-    progs foreach { compare(_) }
-  }
-
-  def run[R](n: Int, block: => R): Timing = {
-    /* warm up*/
-    //for (i <- 1 to n) block
-    var tsum: List[Double] = Nil
-    for (i <- 1 to n) {
-      val (result, t) = Utils.time(block)
-      println(s"iteration #$i took $t")
-      tsum = t::tsum
-    }
-    val dropN = (0.0 * n).toInt
-    val dropTs = tsum.sorted.drop(dropN).take(n - dropN*2)
-    Timing(dropTs)
-  }
-
-  def compare(e: Expr) {
-    /*
-    val ex4fold = new SnippetEx4Fold()
-    val t5 = run(n, { ex4fold() })
-    println(s"0CFA Staged (fold) average time: $t5")
-    */
-    val t1 =  run(10, { evalUnstaged(e) })
-    println(s"Unstaged time - ${t1}")
-
-    val code = specialize(e)
-    code.precompile
-    println("finished precompile")
-    // val code_str = code.code
-    // val writer = new java.io.PrintWriter(new java.io.File("CodeGen.scala.ignore"))
-    // writer.write(code_str)
-    // writer.close()
-    println("start running")
-
-    val t2 = run(10, { code.eval(()) })
-
-    println("-------------------------------------------")
-    println(s"Unstaged time - ${t1}")
-    println(s"Staged time - ${t2}")
-  }
-}

### 2019-02-18 code/src/main/scala/sai/Main.scala

@@ -13,19 +13,6 @@ object GenerateCode {
   }
 }
 
-case class Timing(ts: List[Double]) {
-  val mean: Double = ts.sum/ts.size
-  val sorted_ts: List[Double] = ts.sorted
-  val ub: Double = sorted_ts.head
-  val lb: Double = sorted_ts.last
-  val perc05 = sorted_ts((sorted_ts.size / 20).toInt)
-  val perc25 = sorted_ts((sorted_ts.size / 4).toInt)
-  val perc50 = sorted_ts((sorted_ts.size / 2).toInt)
-  val perc75 = sorted_ts(((sorted_ts.size / 4) * 3).toInt)
-  val perc95 = sorted_ts(sorted_ts.size - (sorted_ts.size/20).toInt - 1)
-  override def toString: String = s"#: ${ts.size}, Mean: ${mean}, Median: ${perc50}"
-}
-
 object Main {
   def run[R](n: Int, block: => R): Timing = {
     /* warm up*/

### 2019-02-18 code/src/main/scala/sai/direct/large/ai/AbsLC.scala

@@ -41,11 +41,10 @@ object AbsLamCal {
     type R[+T] = T
     val ρ0 = Map[Ident, Addr]()
     val σ0 = Map[Addr, Value]()
-    def get(ρ: Env, x: Ident): Addr = ρ.getOrElse(x, Addr("__somerandomthingthatdoesntexist"))
+    def get(ρ: Env, x: Ident): Addr = ρ(x) //ρ.getOrElse(x, Addr("__somerandomthingthatdoesntexist"))
     def put(ρ: Env, x: Ident, a: Addr): Env = ρ + (x → a)
     def get(σ: Store, a: Addr): Value = σ.getOrElse(a, Lattice[Value].bot)
     def put(σ: Store, a: Addr, v: Value): Store = {
-      //val oldv = get(σ, a); σ + (a → Lattice[Value].⊔(v, oldv))
       val oldv = get(σ, a); σ + (a → (v ⊔ oldv))
     }
     def alloc(σ: Store, x: Ident): Addr = Addr(x)
@@ -77,7 +76,7 @@ object AbsLamCal {
 
     def prim_eval(op: String, lv: List[Value]): Value = op match {
       case op if (scala.collection.immutable.Set("+", "-", "*", "/", "%", "vector-length")(op)) => Set(IntV())
-      case op if (scala.collection.immutable.Set("eq?", "null?", "pair?", "and", "or", "not", ">", "<", ">=", "<=")(op)) => Set(BoolV())
+      case op if (scala.collection.immutable.Set("eq?", "null?", "pair?", "and", "or", "not", ">", "<", ">=", "<=", "=")(op)) => Set(BoolV())
       case op if (scala.collection.immutable.Set("list", "cons", "cdr", "cadr")(op)) => Set(ListV())
       case op if ((scala.collection.immutable.Set("car", "caar", "caddr", "vector-ref"))(op)) =>
         Set(BoolV(), IntV(), FloatV(), CharV(), ListV(), VectorV(), SymV())
@@ -159,9 +158,10 @@ object AbsLamCal {
 
     implicit def absValueTyp: Typ[AbsValue]
     implicit def addrTyp: Typ[Addr]
+
     val ρ0: Rep[Env] = Map[Ident, Addr]()
     val σ0: Rep[Store] = Map[Addr, Value]()
[…]

### 2019-02-18 code/src/main/scala/sai/direct/large/ai/LC.scala

@@ -28,7 +28,11 @@ object LamCal {
     type Env
     type Store
     type Ans = (R[Value], R[Store])
-    val primops = Set("+", "-", "*", "/", "%", "eq?", "null?", "pair?", "and", "or", "not", ">", "<", ">=", "<=", "list", "cons", "car", "caar", "cadr", "caddr", "cdr", "cddr", "vector", "make-vector", "set-cdr!", "error", "vector-set!", "vector-ref", "vector-length", "display", "write", "newline", "read", "number->string")
+    val primops = Set("+", "-", "*", "/", "%", "eq?", "null?", "pair?", "and", 
+      "or", "not", ">", "<", ">=", "<=", "=", "list", "cons", "car", "caar", "cadr", 
+      "caddr", "cdr", "cddr", "vector", "make-vector", "set-cdr!", "error", 
+      "vector-set!", "vector-ref", "vector-length", "display", "write", "newline", 
+      "read", "number->string")
     def get(ρ: R[Env], x: Ident): R[Addr]
     def put(ρ: R[Env], x: Ident, a: R[Addr]): R[Env]
     def get(σ: R[Store], a: R[Addr]): R[Value]

### 2019-02-19 paper/10_intro.tex
misc

@@
 staging is less intrusive compared with AC (ie, closure generation), in
-fact,LMS-based staging only requires type annotations }
+fact,LMS-based staging only requires type annotations, static analysis trade-off
+between precision and efficiency, we argue that mp is an approach to improve the
+latter one while fix the former one}
 

### 2019-02-19 paper/70_related.tex
revise
some moands
misc

@@
 \citet{damian1999partial} provides a formal treatment to abstract compilation
-and Shiver's CFA, as well as proofs to establish correctness and certified
-specialized analyzers \cite{damian1999partial}. \citeauthor{amtoft1999partial}
-applyies partial evaluation for constraint-based control flow analysis
-\cite{amtoft1999partial}. Split the analysis to be multiple stages is also
-studied other than control-flow analysis, though the formulation may very
+and Shiver's CFA, as well as proofs to establish correctness of the certified
+specialized analyzer \cite{damian1999partial}. \citeauthor{amtoft1999partial}
+applied partial evaluation for constraint-based control flow analysis
+\cite{amtoft1999partial}. Splitting an analysis into multiple stages is also
+studied for analysis other than for control-flow, though the formulation may very
 different. For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to
@@
 stage is to analyze the program code to obtain a sparse representation, and then
-the second stage conducts the flow-sensitive analysis.
+the second stage conducts the flow-sensitive analysis based on the first one.
 
-\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domain
+\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domains
 through online decomposition \cite{DBLP:conf/popl/SinghPV17,
@@
 number of variables, then we just need to compute the resulting polyhera on
-those smaller polyhedra that contains these related variables. The decomposition
+those smaller polyhedras which contain these related variables. The decomposition
 can be considered as a specialization of abstract domain with respect to the
-program static structure.
+static program structure.
 
@@
 big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17,
-Wei:2018:RAA:3243631.3236800}. \citet{DBLP:conf/cc/CousotC02} also proposed an
+Wei:2018:RAA:3243631.3236800}. \cite{DBLP:conf/cc/CousotC02} also proposed an
 abstract interpretation framework for modular analysis.
@@
 concrete interpreters by MSP and LMS framework.
+
+\cite{10.1007/11561347_18}

### 2019-02-19 code/src/main/scala/sai/direct/large/ai/AbsLC.scala

@@ -220,7 +220,7 @@ object AbsLamCal {
       var out = Map[Config, (Value,Store)]()
 
       def cached_ev(e: Expr, ρ: Rep[Env], σ: Rep[Store]): Rep[(Value, Store)] = {
-        println(s"calling cachev_ev $e")
+        //println(s"calling cachev_ev $e")
         val cfg: Rep[Config] = (unit(e), ρ, σ)
         if (out.contains(cfg)) {
           out(cfg)

### 2019-02-19 paper/10_intro.tex
intro

--- 2019-02-13 paper/10_intro.tex
@@
 
-\todo{example of interpreter -> compiler, mix}
+Now consider that we have an interpreter @eval@ for some language as following:
+
+\begin{lstlisting}
+def eval(e: Expr)(arg: Input): Value = ...
+\end{lstlisting}
+
+where $e$ is the program of type @Expr@ and @arg@ is the input to program $e$,
+and @eval@ produces a value.
+Semantically, the interpreter satisfies $ \texttt{eval}(e)(arg) = [\![ e ]\!] arg$ for all
+programs and inputs. Given a program $e_0$, by applying the first Futamura
+projection, we may obtain a specialized interpreter
+$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$, and by definition of the interpreter, 
+$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
+argument @arg@ to an environment (as well as a companion heap object, if for the
+interest of using assginments) that contains values mapped from free variables
+in $e$, the interpreter is a standard environment/store-passing interpreter,
+where the specialization through MSP is also applicable.
 
@@
 refactor interpreters to static analyzers. As an intellectual quest, it is
-natural to ask --- starting from an interpreter, can we derive sound abstract
-interpreters but it works like a compiler and generates fast analysis code?
+natural to ask --- starting from an interpreter, \textbf{can we derive sound abstract
+interpreters but it works like a compiler and generates fast analysis code?}
+Similar to the concrete interpreter, for an abstract interpreter
+$\widehat{\texttt{eval}}$ that takes a program, an abstract environment and
+returns abstract values, shown as the following:
+
+\begin{lstlisting}
+def $\widehat{\texttt{eval}}$(e: Expr)($\rho$: $\widehat{\texttt{Env}}$): $\widehat{\texttt{Value}}$ = ...
+\end{lstlisting}
+
+We would like to have a specialized abstract interpreter
+$\widehat{\texttt{eval}}_{\texttt{e0}}$ given a program $e_0$ statically, but
+leave the abstract environment $\rho$ dynamic.
+
+\begin{figure}[h]
+  \begin{tikzpicture}
+  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
+  {
+    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} \\
+      \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} \\ };
+  \path[-stealth]
+    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
+    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
+    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
+    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
+  \end{tikzpicture}
+  \caption{The confluence of specialization and approximation.}
+  \label{confluence}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-  \begin{tikzpicture}
-  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
-  {
-    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} \\
-      \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} \\ };
-  \path[-stealth]
-    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
-    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
-    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
-    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
-  \end{tikzpicture}
-  \caption{The confluence of staging and approximation.}
-  \label{confluence}
-\end{figure}
+Besides the intellectual merit of the confluence diagram, this paper contributes
+to a practical problem in static analysis: \textbf{can we speed up a static analyzer
+without any intrusive changes and soundness compromises?} Static analysis is
+known as a trade-off game between precision and efficiency, we argue that
+applying meta-programming (particularly, multi-stage programming) with abstract
+interpreters is an effective approach to improve the latter one while fix the
+former one. Specialization removes the overhead of abstractions introduced by
+high-level programming, but not changes the meaning of the program -- in this
+case, the abstract semantics implemented by the analyzer.
+Indeed, specializing analysis with respect to a program by partial evaluation
+is not new \cite{damian1999partial, amtoft1999partial,
+  Boucher:1996:ACN:647473.727587, ashley:practical}, despite the fact that these
+work are ad-hoc on a particular analysis or requiring significant changes to
+the analysis. For example, abstract compilation \cite{Boucher:1996:ACN:647473.727587}
+requires the whole analyzer to be rewritten as closure generation from. In this
+paper, we show a systematic approach to meta-programming achieves the
+performance goal with less intrusive changes; in fact, a case study
+(Section~\ref{cs_ac}) that compared with abstract compilation shows, by
+utilizing type-directed binding-time annotations, the analyzer program does not
+need any changes.
 
-Besides the intellectual merit of the confluence diagram, this paper contributes
-to a practical problem in static analysis: can we speed up a static analyzer
-without any intrusive changes and soundness compromises?
-\todo{specializing analysis wrt program is not new; first staging; we argue that
-staging is less intrusive compared with AC (ie, closure generation), in
-fact,LMS-based staging only requires type annotations }
+\todo{modular analysis}
 
-\todo{Move some into related work}
-Multi-stage programming has been widely used to improve the performance in many
-domains, such as optimizing compilers or domain-specific languages
-\cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
-DBLP:journals/jfp/CaretteKS09}, numerical computation \cite{PGL-038,
-DBLP:conf/pepm/AktemurKKS13}, generic programming
-\cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, data
-processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
-DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
-\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
+\iffalse
 Likewise, specializing static analyses by partial evaluation emerged in late 90s
-\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
-ashley:practical}, and indeed it is able to effectively remove the interpretive
+, and indeed it is able to effectively remove the interpretive
 overhead of repeatedly traversing the abstract syntax tree. However, these
@@
 
-\iffalse
 When staging a concrete interpreter, the programmers need to distinguish static

### 2019-02-19 paper/70_related.tex
intro

@@
 
+Multi-stage programming has been widely used to improve the performance in many
+domains, such as optimizing compilers or domain-specific languages
+\cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
+DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
+DBLP:journals/jfp/CaretteKS09}, numerical computation \cite{PGL-038,
+DBLP:conf/pepm/AktemurKKS13}, generic programming
+\cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, data
+processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
+DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
+\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
+
 \cite{10.1007/11561347_18}

### 2019-02-19 paper/10_intro.tex
intro

@@
 refactor interpreters to static analyzers. As an intellectual quest, it is
-natural to ask --- starting from an interpreter, \textbf{can we derive sound abstract
-interpreters but it works like a compiler and generates fast analysis code?}
-Similar to the concrete interpreter, for an abstract interpreter
+natural to ask --- starting from an interpreter, \textbf{can we derive a sound
+abstract interpreter but it works like a compiler and generates fast analysis
+code?} Similar to the concrete interpreter, for an abstract interpreter
 $\widehat{\texttt{eval}}$ that takes a program, an abstract environment and
@@
 
-We would like to have a specialized abstract interpreter
-$\widehat{\texttt{eval}}_{\texttt{e0}}$ given a program $e_0$ statically, but
-leave the abstract environment $\rho$ dynamic.
+Given a program $e_0$ statically, We would like to have a specialized abstract
+interpreter $\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
+\widehat{\texttt{Value}}$, but leave the abstract environment $\rho$ dynamic.
 
@@
 Besides the intellectual merit of the confluence diagram, this paper contributes
-to a practical problem in static analysis: \textbf{can we speed up a static analyzer
-without any intrusive changes and soundness compromises?} Static analysis is
-known as a trade-off game between precision and efficiency, we argue that
-applying meta-programming (particularly, multi-stage programming) with abstract
-interpreters is an effective approach to improve the latter one while fix the
-former one. Specialization removes the overhead of abstractions introduced by
-high-level programming, but not changes the meaning of the program -- in this
-case, the abstract semantics implemented by the analyzer.
-Indeed, specializing analysis with respect to a program by partial evaluation
-is not new \cite{damian1999partial, amtoft1999partial,
-  Boucher:1996:ACN:647473.727587, ashley:practical}, despite the fact that these
-work are ad-hoc on a particular analysis or requiring significant changes to
-the analysis. For example, abstract compilation \cite{Boucher:1996:ACN:647473.727587}
-requires the whole analyzer to be rewritten as closure generation from. In this
-paper, we show a systematic approach to meta-programming achieves the
-performance goal with less intrusive changes; in fact, a case study
-(Section~\ref{cs_ac}) that compared with abstract compilation shows, by
-utilizing type-directed binding-time annotations, the analyzer program does not
-need any changes.
+to a practical problem in static analysis: \textbf{can we speed up a static
+analyzer without any intrusive changes and soundness compromises?} Static
+analysis is known as a trade-off game between precision and efficiency, we argue
+that applying meta-programming (particularly, multi-stage programming) with
+abstract interpreters is an effective approach to improve the latter one while
+fix the former one. Specialization removes the overhead of abstractions
+introduced by high-level programming, but not changes the meaning of the program
+-- in this case, the abstract semantics implemented by the analyzer. Indeed,
+specializing analysis with respect to a program by partial evaluation
+\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
+ashley:practical} is not a new idea, despite the fact that most of the previous
+work are ad-hoc on a particular analysis or requiring significant changes to the
+analysis. For example, abstract compilation
+\cite{Boucher:1996:ACN:647473.727587} requires the whole analyzer to be
+rewritten as closure generation from. In this paper, we show a systematic
+approach to meta-programming achieves the performance goal with less intrusive
+changes; in fact, a case study (Section~\ref{cs_ac}) comparing with abstract
+compilation shows, by utilizing type-based stage annotations, the analyzer
+program does not need any changes.
 
-\todo{modular analysis}
+\todo{modular analysis: a whole-program abstract
+  interpreter can be run on different components of program separately, we
+  achieve this by meta-programming mechnically}
 
@@
 
-\subsection{Contributions}
-
-This paper makes the following contributions:
+\paragraph{Contributions and Outline}
 
 \begin{itemize}
-  \item Starting from a generic definition of an interpreter for a higher-order
-    functional language, whose syntax and semantics are described in
-    Section~\ref{bg_lang}, we show its concrete interpreter
-    (Section~\ref{unstaged_conc}), staged concrete interpreter
-    (Section~\ref{stagedinterp}), abstract interpreter (Section~\ref{unstaged_abs})
-    and staged abstract interpreter (Section~\ref{sai}). Our design allows to
-    derive these four different kinds of interpreters by mechanically changing the
-    underlying semantics or stage annotations.
-    %We later extend the staged abstract interpreter to support imperative
-    %language features (Section~\ref{cases_imp}), such as states and loops.
+  \item We present a unified implementation framework that naturally extends the
+    first Futamura projection to abstract interpreters. Starting from a generic
+    interface of the interpreter for a tiny functional language
+    (Section~\ref{generic_if}), we first show the instantiation of concrete
+    interpretation in Section~\ref{unstaged_conc}. Then we lift it to the staged
+    version by adding stage annotations to the semantic operations
+    (Section~\ref{stagedinterp}); and to the abstract interpreter by applying
+    abstraction to the environment, store, and values. Finally, we show the
+    combination of the two transformations, dubbed \textit{staged abstract
+    interpreters}, can be easily obtained (Section~\ref{sai}).
 
-  \item As shown in Section~\ref{sai}, we demonstrate that if we apply the
-    abstract interpreters on open programs, it not only improves the
-    \textit{efficiency} but also the \textit{scalability}, which are two major
-    issues in static analysis.
+  \item \todo{modular analysis} As shown in Section~\ref{sai}, we demonstrate
+    that if we apply the abstract interpreters on open programs, it not only
+    improves the \textit{efficiency} but also the \textit{scalability}, which are
+    two major issues in static analysis.
 
-  \item To demonstrate the applicability and usefulness of the proposed
-    approach, we conduct three case studies in Section~\ref{cases_study}:
+  \item In Section~\ref{cases_study}, we conduct three case studies to
+    demonstrate the applicability and usefulness of the proposed approach to
+    specialize analyses.
+    (1) We first revisit the Abstract Compilation (AC)
+    \cite{Boucher:1996:ACN:647473.727587} technique in Section~\ref{cs_ac}.
+    Our approach and AC are able to achieve the same goal, however, as we show,
+    one of the advantage of staging through types is it does not require
+    whole-analyzer refactoring.
+    (2) We extend the staged abstract interpreter shown in Section~\ref{sai} to
+    two variants of control-flow analysis: context-sensitive pushdown
+    control-flow analysis and \todo{precise-store?}.
+    (3) In Section~\ref{cases_imp}, we plug a \note{staged numerical domain}
+    \todo{widening} into the staged abstract interpreter, demonstrating that a
+    numerical analysis on imperative languages can be also implemented and
+    speeded-up in this manner.
 
-    \subitem (1) We revisit Abstract Compilation \cite{Boucher:1996:ACN:647473.727587} in Section~\ref{cs_ac}, 
-    which was originally a closure generation technique applied on 0-CFA.
-    We show that abstract compilation can be understood and easily implemented
-    as an instance of our staged abstract interpreter. But unlike abstract
-    compilation, our approach does not require the implementers to rewrite the
-    analysis.
+  \item Based on the instantiation of control-flow analysis, we empirically
+    evaluate the performance improvement by staging and specialization
+    (Section~\ref{evaluation}). We compare the both context-insensitive and
+    \note{context-sensitive} analysis.
 
-    \subitem (2) We demonstrate pushdown control-flow analysis techniques
-    (Section~\ref{cfa}), such as context-sensitive analysis and more precise
-    stores can be integrated to staged abstract interpreters without losing any
-    precision.
-
-    \subitem (3) We show that in an imperative language
-    (Section~\ref{cases_imp}), by using an interval domain as abstract domain
-    (as well as staged), the staged abstract interpreter performs numerical 
-    analysis.
-
-    % (Optional: taint/information flow analysis?)
-  \item We empirically evaluate the performance improvement by staging a
-    big-step abstract interpreter for control-flow analysis
-    (Section~\ref{evaluation}). We compare our staged abstract interpreter with the
-    unstaged version.
 \end{itemize}

### 2019-02-20 paper/10_intro.tex
revision

@@
 semantics (Figure~\ref{confluence}) in a unified framework. The four semantics
-share one same generic interpreter and some abstract types; by instantiating the
-abstract types differently, the interpreter behaves as a concrete interpreter,
-an abstracter interperter, or their staged versions, respectively.
-
-% FIXME: and the approach of constructing optimizing abstract interpreters by multi-stage programming.
+share one same generic interpreter. However, by instantiating the interpreter
+differently, it can behave as a concrete interpreter, an abstracter interperter,
+or their staged versions, respectively. The key idea to enable this flexibility
+is to use abstract type members \note{or type parameters} abstract over concrete/abstract components
+(e.g., concrete values or abstract values) of the interpreter, as well as the
+binding-time (e.g., static or dyanmic).
 

### 2019-02-20 paper/20_bg.tex
revision

--- 2018-11-10 paper/20_bg.tex
@@
 
-\section{Background}
+\section{Preliminaries}
 
-In this section, we first described the language we will model and analyze, and present an unstaged concrete 
-interpreter for it. Then, starting from this unstaged concrete interpreter, we proceed in two directions:
-by abstracting it we obtain an unstaged abstract interpreter; by staging it we obtain a staged concrete
-interpreter.
+In this section, we first describe the abstract syntax of the language we will
+analyze, then present the generic interface shared between the four different
+interpreters. We first instantiate the interpreter to the concrete one; then we
+proceed in two directions: by staging it we obtain a staged concrete interpreter
+(Section~\ref{stagedinterp}), by abstracting it we obtain an unstaged abstract
+interpreter (Section~\ref{unstaged_abs});
 

### 2019-02-20 paper/21_bg_interp.tex
intro
revision

@@
 
-We consider a tiny higher-order functional language \TLang, which is based on
-direct-style $\lambda$-calculus with numbers, arithmetics, recursions and
-conditionals. In Section~\ref{cases_imp}, we will add more imperative features
-to the language. Since we are mostly interested in analyzing the dynamic
-behavior of the program, we disguised any static semantics and type system. We
-also assume that input programs are well-typed and variables are distinct. The
-abstract syntax is as follows:
+We consider a call-by-value $\lambda$-calculus in direct-style, but extended
+with numbers, arithmetics, recursions and conditionals. Other effectful features
+such as assignments also can be supported easily; in Section~\ref{cases_imp}, we
+will add more imperative features to the language. Since we are mostly
+interested in analyzing the dynamic behaviors of the program, we disguised any
+static semantics and type system. We also assume that input programs are
+well-typed and all variables are distinct. The abstract syntax is shown as
+follows:
 
@@
 
-\subsection{Generic Interface}
+\subsection{Generic Interface} \label{generic_if}
 

### 2019-02-20 paper/21_bg_interp.tex
misc

@@
 abstract class Expr
-case class Lit(i: Int)                         // numbers
-case class Var(x: String)                      // variables
-case class Lam(x: String, e: Expr)             // lambda abstractions
-case class App(e1: Expr, e2: Expr)             // applications
-case class AOp(op: String, e1: Expr, e2: Expr) // primitive operations
-case class Rec(x: String, rhs: Expr, e: Expr)  // recursive bindings
-case class If0(e1: Expr, e2: Expr, e3: Expr)   // conditionals
+case class Lit(i: Int) extends Expr                         // numbers
+case class Var(x: String) extends Expr                      // variables
+case class Lam(x: String, e: Expr) extends Expr             // abstractions
+case class App(e1: Expr, e2: Expr) extends Expr             // applications
+case class If0(e1: Expr, e2: Expr, e3: Expr) extends Expr   // conditionals
+case class Rec(x: String, rhs: Expr, e: Expr) extends Expr  // recursions
+case class AOp(op: String, e1: Expr, e2: Expr) extends Expr // primitives
 \end{lstlisting}

### 2019-02-20 paper/10_intro.tex
misc

@@
 These staging annotations identify which part of the inputs should be
-specialized. As a classical example, we consider the power function to introduce
+specialized. As a classical example, we use the power function to introduce
 the idea of MSP:
@@
 \begin{lstlisting}
-def power(b: Rep[Int], e: Int): Int = 
+def power(b: Rep[Int], e: Int): Rep[Int] = 
   if (e == 0) 1 else e * power(b, e-1)
@@
 DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-\todo{CEK example?}
+% \todo{CEK example?}
 

### 2019-02-20 code/CodeGen.scala

@@ -1,1323 +1,0 @@
-import sai.direct.large.parser._
-import sai.common.ai.Lattices._
-object RTSupport {
-  case class Addr(x: String)
-  trait AbsValue
-  case class IntV() extends AbsValue
-  case class FloatV() extends AbsValue
-  case class CharV() extends AbsValue
-  case class BoolV() extends AbsValue
-  case class ListV() extends AbsValue
-  case class VectorV() extends AbsValue
-  case class VoidV() extends AbsValue
-  case class SymV() extends AbsValue
-  type Value = Set[AbsValue]
-  case class CompiledClo(f: (List[Value], Map[Addr,Value]) => (Value, Map[Addr,Value]), λ: Int, ρ: Int) extends AbsValue {
-    def canEqual(a: Any) = a.isInstanceOf[CompiledClo]
-    override def equals(that: Any): Boolean = that match {
-      case that: CompiledClo => that.canEqual(this) && this.λ == that.λ && this.ρ == that.ρ //this.hashCode == that.hashCode
-      case _ => false
-    }
-    override def hashCode: Int = {
-      val prime = 31
-      var result = 1
-      result = prime * result + λ //.hashCode
-      result = prime * result + ρ //.hashCode
-      result
-    }
-  }
-  def apply_closures_norep(f: Value, args: List[Value], σ: Map[Addr,Value]) = {
-    val (σ0, vs) = f.foldLeft((σ, Set[Value]())) {
-      case ((σ0, vs), CompiledClo(fun, _, _)) =>
-        val (v, vσ) = fun(args, σ0)
-        (vσ ⊔ σ0, vs + v)
-      case ((σ0, vs), _) =>
-        (σ0, vs)
-    }
-    (vs.foldLeft(Lattice[Value].bot)(Lattice[Value].⊔), σ0)
-  }
-}
-import RTSupport._
-
-/*****************************************
- Emitting Generated Code                  
- *******************************************/
-class Snippet extends ((Unit)=>(Unit)) {
-  def apply(x999:Unit): Unit = {
-    var x1002 = null.asInstanceOf[scala.Function0[Tuple2[scala.collection.immutable.Set[AbsValue],scala.collection.immutable.Map[Addr, scala.collection.immutable.Set[AbsValue]]]]]
[…]

### 2019-02-20 code/src/main/scala/sai/direct/core/ai/AbsLC.scala

@@ -189,6 +189,8 @@ object AbsLamCal {
     }
 
     override def eval_top(e: Expr, ρ: R[Env], σ: R[Store]): Ans = CacheFix(eval).iter(e, ρ, σ)
+
+    def eval_top_nocache(e: Expr): Ans = fix(eval)(e, ρ0, σ0)
   }
 
 
@@ -281,7 +283,8 @@ object AbstractLC {
     def specialize(p: Expr): DslDriver[Unit, Unit] =
       new RepAbsInterpDriver {
         def snippet(unit: Rep[Unit]): Rep[Unit] = {
-          val (v, s) = eval_top(p)
+          //val (v, s) = eval_top(p)
+          val (v, s) = eval_top_nocache(p)
           println(v); println(s)
         }
       }
@@ -298,9 +301,9 @@ object AbstractLC {
 
     //println(AbsInterp.eval_top(id4))
     //println(AbsInterp.eval_top(omega))
-    println(AbsInterp.eval_top(fact5))
+    //println(AbsInterp.eval_top(fact5))
 
-    val code = specialize(fact5)
+    val code = specialize(omega)
     println(code.code)
     code.eval(())
 

### 2019-02-21 paper/10_intro.tex
add lms support

@@
 interpreters. By borrowing the ideas from \note{monadic interpreters}
-\cite{DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
+\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
 Sergey:2013:MAI:2491956.2491979} and embedded domain-specific languages
@@
 analyzer without any intrusive changes and soundness compromises?} Static
-analysis is known as a trade-off game between precision and efficiency, we argue
-that applying meta-programming (particularly, multi-stage programming) with
+analysis is known as a trade-off game between \textit{precision} and \textit{efficiency}, we argue
+that applying meta-programming (multi-stage programming, in particular) with
 abstract interpreters is an effective approach to improve the latter one while
-fix the former one. Specialization removes the overhead of abstractions
+the former one is untouched. Specialization removes the overhead of abstractions
 introduced by high-level programming, but not changes the meaning of the program
@@
 \begin{itemize}
-  \item We present a unified implementation framework that naturally extends the
+  \item We progressively present a unifying construction that naturally extends the
     first Futamura projection to abstract interpreters. Starting from a generic
@@
     combination of the two transformations, dubbed \textit{staged abstract
-    interpreters}, can be easily obtained (Section~\ref{sai}).
+    interpreters}, can be easily derived (Section~\ref{sai}).
 

### 2019-02-21 virtmacro-code/src/main/scala/sai/lms/DslApi.scala

@@ -1,0 +1,325 @@
+package sai
+package lms
+
+import org.scala_lang.virtualized.virtualize
+import org.scala_lang.virtualized.SourceContext
+
+import scala.virtualization.lms.common._
+
+import java.nio._
+import java.nio.file._
+import java.io._
+
+/** This file is adapted from LMS-tutorial.
+  * https://github.com/scala-lms/tutorials/blob/master/src/test/scala/lms/tutorial/dslapi.scala
+  */
+
+// should this be added to LMS?
+trait UtilOps extends Base { this: Dsl =>
+  def infix_HashCode[T:Manifest](o: Rep[T])(implicit pos: SourceContext): Rep[Long]
+  def infix_HashCode(o: Rep[String], len: Rep[Int])(implicit v: Overloaded1, pos: SourceContext): Rep[Long]
+}
+
+trait UtilOpsExp extends UtilOps with BaseExp { this: DslExp =>
+  case class ObjHashCode[T: Manifest](o: Rep[T])(implicit pos: SourceContext) extends Def[Long] { def m = manifest[T] }
+  case class StrSubHashCode(o: Rep[String], len: Rep[Int])(implicit pos: SourceContext) extends Def[Long]
+  def infix_HashCode[T:Manifest](o: Rep[T])(implicit pos: SourceContext) = ObjHashCode(o)
+  def infix_HashCode(o: Rep[String], len: Rep[Int])(implicit v: Overloaded1, pos: SourceContext) = StrSubHashCode(o,len)
+
+  override def mirror[A:Manifest](e: Def[A], f: Transformer)(implicit pos: SourceContext): Exp[A] = (e match {
+    case e@ObjHashCode(a) => infix_HashCode(f(a))(e.m,pos)
+    case e@StrSubHashCode(o,len) => infix_HashCode(f(o),f(len))
+    case _ => super.mirror(e,f)
+  }).asInstanceOf[Exp[A]]
+}
+
+trait ScalaGenUtilOps extends ScalaGenBase {
+  val IR: UtilOpsExp
+  import IR._
+
+  override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
+    case ObjHashCode(o) => emitValDef(sym, src"$o.##")
+    case _ => super.emitNode(sym, rhs)
+  }
+}
+
+trait CGenUtilOps extends CGenBase {
+  val IR: UtilOpsExp
+  import IR._
+
[…]

### 2019-02-22 paper/10_intro.tex
paper

@@
 If the programmer identifies that @b@ will be known in the future stage (as shown
-on its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known in the
+on its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known at the
 current stage, say 5, then we can specialize the function @power@ with
-respect to @e = 5@ and generate a specialized function that runs faster:
+respect to @e = 5@, and generate a specialized function where the overhead of
+recursion is eliminated. The following is the generated code:
 
 \begin{lstlisting}
-def power5(b: Int) = b * b * b * b * b * 1
+def power5(b: Int) = {
+  val x1 = x0 * x0; val x2 = x0 * x1; val x3 = x0 * x2; val x4 = x0 * x3
+  x4
+}
 \end{lstlisting}
 
-Now consider that we have an interpreter @eval@ for some language as following:
+Now to concretely illustrate the idea of Futamura projection, consider that we
+have an interpreter @eval@ of some language as following:
 
@@
 definitional abstract interpreters, which intersects as staged abstract
-interpreters. By borrowing the ideas from \note{monadic interpreters}
-\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17,
-Sergey:2013:MAI:2491956.2491979} and embedded domain-specific languages
-\cite{DBLP:conf/snapl/RompfBLSJAOSKDK15, DBLP:journals/jfp/CaretteKS09,
-DBLP:conf/icfp/GibbonsW14}, we develop an approach to fulfill four different
-semantics (Figure~\ref{confluence}) in a unified framework. The four semantics
-share one same generic interpreter. However, by instantiating the interpreter
-differently, it can behave as a concrete interpreter, an abstracter interperter,
-or their staged versions, respectively. The key idea to enable this flexibility
-is to use abstract type members \note{or type parameters} abstract over concrete/abstract components
-(e.g., concrete values or abstract values) of the interpreter, as well as the
-binding-time (e.g., static or dyanmic).
+interpreters. By borrowing the ideas from monadic interpreters
+\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95,
+DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979} and embedded
+domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
+DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14}, we develop an
+approach to fulfill four different semantics (Figure~\ref{confluence}) in a
+unified framework. The four semantics share one same generic interpreter.
+However, by instantiating the interpreter differently, it can behave as a
+concrete interpreter, an abstracter interperter, or their code-generator
+versions, respectively. The key idea that enables this flexibility is to use
+abstract type members abstract over concrete/abstract components (e.g., concrete
+values or abstract values), as well as the binding-time of them (e.g., static or
+dyanmic). By staging, the overhead caused by the monadic layers is eliminated in
+the generated code.
 
-Besides the intellectual merit of the confluence diagram, this paper contributes
-to a practical problem in static analysis: \textbf{can we speed up a static
-analyzer without any intrusive changes and soundness compromises?} Static
-analysis is known as a trade-off game between \textit{precision} and \textit{efficiency}, we argue
-that applying meta-programming (multi-stage programming, in particular) with
-abstract interpreters is an effective approach to improve the latter one while
-the former one is untouched. Specialization removes the overhead of abstractions
-introduced by high-level programming, but not changes the meaning of the program
--- in this case, the abstract semantics implemented by the analyzer. Indeed,
-specializing analysis with respect to a program by partial evaluation
-\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
-ashley:practical} is not a new idea, despite the fact that most of the previous
-work are ad-hoc on a particular analysis or requiring significant changes to the
-analysis. For example, abstract compilation
+Besides the intellectual merit shown in the confluence diagram, this paper
+contributes to a practical problem in static analysis: \textbf{can we speed up a
+static analyzer without any intrusive changes and soundness compromises?} Static
+analysis is known as a trade-off game between \textit{precision} and
+\textit{efficiency}, we argue that applying meta-programming (multi-stage
+programming, in particular) with abstract interpreters is an effective approach
+to improve the latter one while the former one is untouched. Specialization
+removes the overhead of abstractions introduced by high-level programming, but
+not changes the meaning of the program -- in this case, the abstract semantics
+implemented by the analyzer. Indeed, specializing analysis with respect to a
+program by partial evaluation \cite{damian1999partial, amtoft1999partial,
+Boucher:1996:ACN:647473.727587, ashley:practical} is not a new idea, despite the
+fact that most of the previous work are ad-hoc on a particular analysis or
+requiring significant changes to the analysis. For example, abstract compilation
 \cite{Boucher:1996:ACN:647473.727587} requires the whole analyzer to be

### 2019-02-22 paper/70_related.tex
misc
paper

@@
 shows that abstract interpreters can be derived systematically from concrete
-semantic artifacts. The big-step abstract interpreters we presented in this
-paper are also inspired by the AAM framework, which are later adopted for
-big-step abstract interpreters \cite{DBLP:journals/pacmpl/DaraisLNH17,
-Wei:2018:RAA:3243631.3236800}. \cite{DBLP:conf/cc/CousotC02} also proposed an
+semantic artifacts.
+Using monads to construct abstract interpreters is explored by
+\citet{Sergey:2013:MAI:2491956.2491979} and
+\citet{DBLP:journals/pacmpl/DaraisLNH17}.
+Similar to monadic abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
+reconstruct big-step abstract interpreters with delimited continuations;
+\citet{Keidel:2018:CSP:3243631.3236767} present an arrow-based abstract
+interpreter that makes proof of soundness can be constructed compositionally.
+\cite{DBLP:conf/cc/CousotC02} proposed an
 abstract interpretation framework for modular analysis.
 \citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by
-using bi-abduction. The big-step abstract interpreter we presented in this paper
-is inspired by \cite{DBLP:journals/pacmpl/DaraisLNH17,
-Wei:2018:RAA:3243631.3236800}. \citet{Keidel:2018:CSP:3243631.3236767}
-demonstrate a similar implementation of definitional abstract interpreter using
-arrows.
+using bi-abduction.
 
@@
 
-\cite{10.1007/11561347_18}
+MSP with functors and monads \cite{10.1007/11561347_18}
+
+staged monadic interpreters \cite{DBLP:conf/dsl/SheardBP99}

### 2019-02-22 paper/10_intro.tex
intro

@@
 Now we have seen two orthogonal transformations of interpreters: by
-speclization, we can refactor interpreters to compilers; by abstraction, we can
-refactor interpreters to static analyzers. As an intellectual quest, it is
-natural to ask --- starting from an interpreter, \textbf{can we derive a sound
-abstract interpreter but it works like a compiler and generates fast analysis
-code?} Similar to the concrete interpreter, for an abstract interpreter
-$\widehat{\texttt{eval}}$ that takes a program, an abstract environment and
-returns abstract values, shown as the following:
-
-\begin{lstlisting}
-def $\widehat{\texttt{eval}}$(e: Expr)($\rho$: $\widehat{\texttt{Env}}$): $\widehat{\texttt{Value}}$ = ...
-\end{lstlisting}
-
-Given a program $e_0$ statically, We would like to have a specialized abstract
-interpreter $\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
-\widehat{\texttt{Value}}$, but leave the abstract environment $\rho$ dynamic.
+speclization, we can refactor interpreters to code generators; by abstraction,
+we can refactor interpreters to static analyzers. As an intellectual quest, it
+is natural to ask --- starting from an interpreter, \textbf{can we derive a
+sound abstract interpreter but it works like a code generator which produces fast
+analysis code?} Conceptually, it is similar to the concrete setting: for an
+abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
+\widehat{\texttt{Env}} \to \widehat{\texttt{Value}}$, which takes a program, an abstract
+environment and returns abstract values, we would like to specialize it with
+repsect to a program $e_0$ and produce an code generator
+$\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
+\widehat{\texttt{Value}}$, such that $
+\widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
+\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$, but the staged one leaves the
+abstract environment $\widehat{\rho}$ dynamic.
 
@@
   \begin{tikzpicture}
-  \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em]
+  \matrix (m) [matrix of math nodes,row sep=3em,column sep=3em,minimum width=2em]
   {

### 2019-02-22 paper/10_intro.tex
typo

@@
 Now we have seen two orthogonal transformations of interpreters: by
-speclization, we can refactor interpreters to code generators; by abstraction,
+specliazation, we can refactor interpreters to code generators; by approximation,
 we can refactor interpreters to static analyzers. As an intellectual quest, it

### 2019-02-22 virtmacro-code/src/main/scala/sai/Gen.scala
migrate to scalaz

@@ -1,0 +1,54 @@
+package sai
+import PCFLang._
+sealed trait Value
+case class IntV(i: Int) extends Value
+case class CompiledClo(f: (Map[Int,Value], Value) => (Map[Int,Value], Value), λ: Lam, ρ: Map[String,Int]) extends Value
+
+/*****************************************
+ Emitting Generated Code
+ *******************************************/
+class Snippet extends ((Unit)=>(Unit)) {
+  def apply(x2:Unit): Unit = {
+    val x0 = Map[java.lang.String, Int]()
+    val x1 = Map[Int, Value]()
+    val x3 = x1.size
+    val x4 = x3 + 1
+    val x5 = x0 + ("fact" -> x4)
+    val x6 = {(x7:scala.collection.immutable.Map[Int, Value],x8:Value) =>
+      val x9 = x7
+      val x10 = x8
+      val x11 = x9.size
+      val x12 = x11 + 1
+      val x14 = x9 + (x12 -> x10)
+      val x13 = x5 + ("n" -> x12)
+      val x15 = x13("n")
+      val x16 = x14(x15)
+      val x17 = x16.asInstanceOf[IntV].i
+      val x20 = IntV(1)
+      val x21 = x16.asInstanceOf[IntV].i
+      val x22 = x20.asInstanceOf[IntV].i
+      val x23 = IntV(x21-x22)
+      val x18 = x13("fact")
+      val x19 = x14(x18)
+      val x24 = x19.asInstanceOf[CompiledClo].f(x14, x23)
+      val x27 = x16.asInstanceOf[IntV].i
+      val x26 = x24._2
+      val x28 = x26.asInstanceOf[IntV].i
+      val x29 = IntV(x27*x28)
+      val x25 = x24._1
+      val x30 = new Tuple2[scala.collection.immutable.Map[Int, Value],Value](x25,x29)
+      x30: Tuple2[scala.collection.immutable.Map[Int, Value],Value]
+    }
+    val x32 = CompiledClo(x6,Lam("n",If0(Var("n"),Lit(1),Aop('*,Var("n"),App(Var("fact"),Aop('-,Var("n"),Lit(1)))))),x5)
+    val x36 = IntV(5)
+    val x33 = x1 + (x4 -> x32)
+    val x34 = x5("fact")
+    val x35 = x33(x34)
+    val x37 = x35.asInstanceOf[CompiledClo].f(x33, x36)
+    val x38 = x37._1
[…]

### 2019-02-22 virtmacro-code/src/main/scala/sai/LMSSupport.scala

@@ -1,0 +1,25 @@
+package sai
+
+import scala.virtualization.lms.internal.GenericNestedCodegen
+import scala.virtualization.lms.common.{
+  SetOps => _, SetOpsExp => _, ScalaGenSetOps => _,
+  ListOps => _, ListOpsExp => _, ScalaGenListOps => _,
+  _}
+import org.scala_lang.virtualized.virtualize
+import org.scala_lang.virtualized.SourceContext
+
+import sai.lms._
+
+trait SAIDsl extends Dsl
+    with MapOps
+    with SetOps
+    with TupleOps
+    with UncheckedOps
+    with TupledFunctions
+
+trait SAIOpsExp extends DslExp
+    with MapOpsExp
+    with SetOpsExp
+    with TupleOpsExp
+    with UncheckedOpsExp
+    with TupledFunctionsRecursiveExp

### 2019-02-22 virtmacro-code/src/main/scala/sai/Lang.scala

@@ -1,0 +1,43 @@
+package sai
+
+object PCFLang {
+  sealed trait Expr
+  sealed trait AExpr extends Expr
+  sealed trait CExpr extends Expr
+
+  case class Lit(i: Int) extends AExpr
+  case class Var(x: String) extends AExpr {
+    override def toString: String = "Var(\"" + x + "\")"
+  }
+  case class Lam(x: String, e: Expr) extends AExpr {
+    override def toString: String = "Lam(\"" + x + "\"," + e + ")"
+  }
+  case class Amb(a: AExpr, b: AExpr) extends AExpr
+
+  case class App(e1: Expr, e2: Expr) extends CExpr
+  case class Let(x: String, rhs: Expr, body: Expr) extends CExpr {
+    override def toString = "Let(\"" + x + "\"," + rhs + "," + body + ")"
+  }
+  case class Rec(x: String, rhs: Expr, body: Expr) extends CExpr {
+    override def toString: String = "Rec(\"" + x + "\"," + rhs + "," + body + ")"
+  }
+  case class If0(e1: Expr, e2: Expr, e3: Expr) extends CExpr
+  case class Aop(op: Symbol, e1: Expr, e2: Expr) extends CExpr
+
+  object Values {
+    sealed trait Value
+    case class IntV(i: Int) extends Value
+    case class CloV[Env](lam: Lam, e: Env) extends Value
+  }
+
+  val p1 = Let("x", Lit(1),
+               Let("y", Lit(2),
+                   Let("f", Lam("z", Aop('+, Var("z"), Var("y"))),
+                       App(Var("f"), Var("x")))))
+
+  val fact = Lam("n",
+                 If0(Var("n"),
+                     Lit(1),
+                     Aop('*, Var("n"), App(Var("fact"), Aop('-, Var("n"), Lit(1))))))
+  val fact5 = Rec("fact", fact, App(Var("fact"), Lit(5)))
+}

### 2019-02-22 virtmacro-code/src/main/scala/sai/examples/MainCats.scala

@@ -1,9 +1,20 @@
 package sai
+package examples
 
 import scala.language.implicitConversions
 
-import scalaz._
-import Scalaz._
+import cats._
+import cats.{Id, ~>}
+import cats.implicits._
+import cats.data.State
+import cats.data.Nested
+import cats.data.IndexedStateT
+import cats.data.Kleisli
+import cats.free.Free
+import cats.free.Free.liftF
+import cats.arrow.FunctionK
+import cats.syntax.applicative._
+import cats.instances.list._
 
 import scala.virtualization.lms.internal.GenericNestedCodegen
 import scala.virtualization.lms.common.{
@@ -13,10 +24,18 @@ import scala.virtualization.lms.common.{
 import org.scala_lang.virtualized.virtualize
 import org.scala_lang.virtualized.SourceContext
 
+import sai._
 import sai.lms._
 import sai.examples._
 
-object EnvInterpreter1 {
+object Misc {
+  type Reader[A, B] = Kleisli[Id, A, B]
+  object Reader {
+    def apply[A, B](f: A => B): Reader[A, B] = Kleisli[Id, A, B](f)
+  }
+}
+
+object EnvInterpreter {
   /* An environment interpreter using Reader Monad */
   import PCFLang._
   import PCFLang.Values._
@@ -27,10 +46,14 @@ object EnvInterpreter1 {
   val ReaderT = Kleisli
   import ReaderT._
 
-  type AnsT[T] = ReaderT[Id, Env, T]
-  type Ans = AnsT[Value]
+  type Ans = ReaderT[Id, Env, Value]
 
-  def num(i: Int): Ans = IntV(i).asInstanceOf[Value].pure[AnsT]
+  def ap_clo(fun: Value, arg: Value): Ans = fun match {
+    case CloV(Lam(x, e), ρ: Env) => eval(e).local[Env](_ => ρ + (x → arg))
+  }
+
+  def branch0(test: Value, thn: Expr, els: Expr): Ans =
+    if (test == IntV(0)) eval(thn) else eval(els)
 
   def prim(op: Symbol, v1: Value, v2: Value): Value = (op, v1, v2) match {
     case ('+, IntV(x), IntV(y)) => IntV(x + y)
[…]

### 2019-02-24 paper/10_intro.tex
misc
typo, monadic ai

--- 2019-02-22 paper/10_intro.tex
@@
 def power5(b: Int) = {
-  val x1 = x0 * x0; val x2 = x0 * x1; val x3 = x0 * x2; val x4 = x0 * x3
+  val x1 = b * b; val x2 = b * x1; val x3 = b * x2; val x4 = b * x3
   x4
@@
 argument @arg@ to an environment (as well as a companion heap object, if for the
-interest of using assginments) that contains values mapped from free variables
+interest of using assignments) that contains values mapped from free variables
 in $e$, the interpreter is a standard environment/store-passing interpreter,
@@
 lattice-based semantic approach to construct sound static analysis, by
-approximation of fixpoints \cite{DBLP:conf/popl/CousotC77}. However,
+approximation of fix-points \cite{DBLP:conf/popl/CousotC77}. However,
 constructing sound abstract interpreters was considered abstruse and complicated
@@
 Now we have seen two orthogonal transformations of interpreters: by
-speclization, we can refactor interpreters to code generators; by abstraction,
+specialization, we can refactor interpreters to code generators; by approximation,
 we can refactor interpreters to static analyzers. As an intellectual quest, it
@@
 environment and returns abstract values, we would like to specialize it with
-repsect to a program $e_0$ and produce an code generator
+respect to a program $e_0$ and produce a compiled analysis
 $\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
@@
 \widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
-\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$, but the staged one leaves the
-abstract environment $\widehat{\rho}$ dynamic.
+\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$, but the compiled one leaves the
+abstract environment $\widehat{\rho}$ as remaining input.
 
@@
 However, by instantiating the interpreter differently, it can behave as a
-concrete interpreter, an abstracter interperter, or their code-generator
+concrete interpreter, an abstracter interpreter, or their code-generator
 versions, respectively. The key idea that enables this flexibility is to use
@@
 values or abstract values), as well as the binding-time of them (e.g., static or
-dyanmic). By staging, the overhead caused by the monadic layers is eliminated in
+dynamic). By staging, the overhead caused by the monadic layers is eliminated in
 the generated code.

### 2019-02-24 virtmacro-code/src/main/scala/sai/Main.scala

@@ -27,10 +27,10 @@ object EnvInterpreter {
   val ReaderT = Kleisli
   import ReaderT._
 
-  type AnsT[T] = ReaderT[Id, Env, T]
-  type Ans = AnsT[Value]
+  type AnsM[T] = ReaderT[Id, Env, T]
+  type Ans = AnsM[Value]
 
-  def num(i: Int): Ans = IntV(i).asInstanceOf[Value].pure[AnsT]
+  def num(i: Int): Ans = IntV(i).asInstanceOf[Value].pure[AnsM]
 
   def prim(op: Symbol, v1: Value, v2: Value): Value = (op, v1, v2) match {
     case ('+, IntV(x), IntV(y)) => IntV(x + y)
@@ -86,24 +86,24 @@ object EnvStoreInterpreter {
 
   type ReaderT[F[_], B] = Kleisli[F, Env, B]
   val ReaderT = Kleisli
-
-  type StateT[T] = IndexedStateT[Id, Store, Store, T]
+  type StateT[F[_], B] = IndexedStateT[F, Store, Store, B]
   val StateT = IndexedStateT
 
-  type AnsT[T] = ReaderT[StateT, T]
-  type Ans = AnsT[Value]
+  type StateM[T] = StateT[Id, T]
+  type AnsM[T] = ReaderT[StateM, T]
+  type Ans = AnsM[Value]
 
-  def ask_env: AnsT[Env] = ReaderT.ask[StateT, Env]
-  def ext_env(xv: (String, Addr)): AnsT[Env] = for { ρ <- ask_env } yield ρ + xv
+  def ask_env: AnsM[Env] = ReaderT.ask[StateM, Env]
+  def ext_env(xv: (String, Addr)): AnsM[Env] = for { ρ <- ask_env } yield ρ + xv
 
-  def get_store: AnsT[Store] = MonadTrans[ReaderT].liftM[StateT, Store](State.get)
-  def alloc(x: String): AnsT[Addr] = for { σ <- get_store } yield σ.size + 1
+  def get_store: AnsM[Store] = MonadTrans[ReaderT].liftM[StateM, Store](State.get)
+  def alloc(x: String): AnsM[Addr] = for { σ <- get_store } yield σ.size + 1
 
-  def update_store(av: (Addr, Value)): AnsT[Unit] =
-    MonadTrans[ReaderT].liftM[StateT, Unit](State.modify(σ => σ + av))
[…]

### 2019-02-24 virtmacro-code/src/main/scala/sai/examples/NDMonad.scala

@@ -21,12 +21,72 @@ object NDTest {
 
     val b: Ans = for {
       s <- StateT.stateTMonadState[Store, NondetT].get
-      x <- MonadTrans[StateT].liftM[NondetT, Int](ListT.fromList[Id, Int](List(4,5,6)))
-      y <- MonadTrans[StateT].liftM[NondetT, Int](ListT.fromList[Id, Int](List(1,2,3)))
+      x <- MonadTrans[StateT].liftM[NondetT, Int](ListT.fromList[Id, Int](List(4,5)))
+      y <- MonadTrans[StateT].liftM[NondetT, Int](ListT.fromList[Id, Int](List(1,2)))
       _ <- StateT.stateTMonadState[Store, NondetT].modify(s => s + (x -> y))
     } yield x * y
 
-    println(a.run(Map()))
+    val c: Ans = for {
+      v <- StateT.stateTMonadPlus[Store, NondetT].plus(a, b)
+    } yield v
+
+    println(c.run(Map[Int, Int](-1 -> -1, -2 -> -2)))
+  }
+
+  def test2() = {
+    type Store = Map[Int, Int]
+    type Cache = Map[String, Set[String]]
+
+    type OutCacheT[F[_], B] = IndexedStateT[F, Cache, Cache, B]
+    type InCacheT[F[_], B] = Kleisli[F, Cache, B]
+    type NondetT[F[_], B] = ListT[F, B]
+    type StateT[F[_], B] = IndexedStateT[F, Store, Store, B]
+
+    type OutCacheM[T] = OutCacheT[Id, T]
+    type InOutCacheM[T] = InCacheT[OutCacheM, T]
+    type NondetM[T] = NondetT[InOutCacheM, T]
+    type AnsM[T] = StateT[NondetM, T]
+
+    type Ans = AnsM[Int] // StateT[NondetT[InCacheT[OutCacheT[Id, ?], ?], ?], Int]
+
+    def ask_incache: AnsM[Cache] = MonadTrans[StateT].liftM[NondetM, Cache](
+      MonadTrans[NondetT].liftM[InOutCacheM, Cache](
+        Kleisli.ask[OutCacheM, Cache]
+      ))
+    def get_outcache: AnsM[Cache] =
+      MonadTrans[StateT].liftM[NondetM, Cache](
+        MonadTrans[NondetT].liftM[InOutCacheM, Cache](
+          MonadTrans[InCacheT].liftM[OutCacheM, Cache](
[…]

### 2019-02-26 paper/10_intro.tex
RepListT wip

@@
 for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
-annotations (i.e., binding-time annotations) to be explicit from the programmers.
+annotations (i.e., binding-time annotations) to be explicit from the
+programmers. The stage annotations can be either syntactic (as in
+MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \todo{cite}.
 These staging annotations identify which part of the inputs should be
-specialized. As a classical example, we use the power function to introduce
-the idea of MSP:
+specialized. As a classical example, we use the power function and type-based
+annotations to introduce the idea of MSP:
 
@@
 respect to @e = 5@, and generate a specialized function where the overhead of
-recursion is eliminated. The following is the generated code:
+recursion is eliminated. Conceptually, the generated code looks like the following:
 
 \begin{lstlisting}
-def power5(b: Int) = {
-  val x1 = b * b; val x2 = b * x1; val x3 = b * x2; val x4 = b * x3
-  x4
-}
+def power5(b: Int) = b * b * b * b * b
 \end{lstlisting}
@@
   DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-% \todo{CEK example?}
+
+\paragraph{Futamura Projection of Abstract Interpreters}
 
@@
 domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14}, we develop an
+DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14, Hofer:2008:PED:1449913.1449935}, we develop an
 approach to fulfill four different semantics (Figure~\ref{confluence}) in a
@@
 the generated code.
+
+\paragraph{Performant via Specialization}
 
@@
 program does not need any changes.
+
+\paragraph{Modularity via Meta-Programming}
 

### 2019-02-26 virtmacro-code/src/main/scala/sai/LMSSupport.scala

@@ -13,6 +13,7 @@ import sai.lms._
 trait SAIDsl extends Dsl
     with MapOps
     with SetOps
+    with ListOps
     with TupleOps
     with UncheckedOps
     with TupledFunctions
@@ -20,6 +21,7 @@ trait SAIDsl extends Dsl
 trait SAIOpsExp extends DslExp
     with MapOpsExp
     with SetOpsExp
+    with ListOpsExp
     with TupleOpsExp
     with UncheckedOpsExp
     with TupledFunctionsRecursiveExp

### 2019-02-26 paper/20_bg.tex
paper
migrate to scalaz

@@
 analyze, then present the generic interface shared between the four different
-interpreters. We first instantiate the interpreter to the concrete one; then we
+interpreters. We choose to use Scala to demonstrate the idea, but the idea of
+utilizing monads and multi-stages also can be easily constructed in other functional
+languages, such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
+  DBLP:conf/flops/Kiselyov14} or Template Haskell \cite{Sheard:2002:TMH:636517.636528}.
+We first instantiate the interpreter to the concrete one; then we
 proceed in two directions: by staging it we obtain a staged concrete interpreter
-(Section~\ref{stagedinterp}), by abstracting it we obtain an unstaged abstract
-interpreter (Section~\ref{unstaged_abs});
+(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
+interpreter (Section~\ref{unstaged_abs}).
 
 \input{21_bg_interp}
-\input{23_bg_stage}
-\input{22_bg_abs}
+\input{22_bg_stage}
+\input{23_bg_abs}

### 2019-02-26 paper/21_bg_interp.tex
paper

@@
 
-We consider a call-by-value $\lambda$-calculus in direct-style, but extended
+We consider a call-by-value $\lambda$-calculus in direct-style, extended
 with numbers, arithmetics, recursions and conditionals. Other effectful features
-such as assignments also can be supported easily; in Section~\ref{cases_imp}, we
+such as assignments can also be supported readily. In Section~\ref{cases_imp}, we
 will add more imperative features to the language. Since we are mostly
@@
 
-We give the concrete semantics of \TLang by showing a big-step definitional
+We will give the concrete semantics using a big-step definitional
 interpreter. The interpreter is a recursive function that takes the program AST,

### 2019-02-26 paper/22_bg_stage.tex
paper

--- 2019-02-04 paper/23_bg_stage.tex
@@
 dynamic code generation in a type-safe manner. Different from the syntactic
-approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations,
-LMS distinguishes binding-time solely based on types. LMS provides a type
-constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an
-expression will be known at next stage. All operations acting on a @Rep[T]@
-expression will be residualized in the generated code. However, we need to
-provide intermediate representation and code generation support for type @T@ to
-LMS. Fortunately, the LMS framework already provides such support for primitive
-types and commonly used data structure such as arrays and maps; implementing
-such support for custom class is also straightforward. Let's go back and see how
-LMS can be used to specialize the power function we mentioned in Section 1.
+approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14,
+DBLP:conf/gpce/CalcagnoTHL03} that uses quotations, LMS distinguishes
+binding-time solely based on types. LMS provides a type constructor @Rep[T]@
+where @T@ can be an arbitrary type, indicating an expression will be known at
+next stage. All operations acting on a @Rep[T]@ expression will be residualized
+in the generated code. However, we need to provide intermediate representation
+and code generation support for type @T@ to LMS. Fortunately, the LMS framework
+already provides such support for primitive types and commonly used data
+structure such as arrays and maps; implementing such support for custom class is
+also straightforward. Let's go back and see how LMS can be used to specialize
+the power function we mentioned in Section 1.
 

### 2019-02-27 paper/10_intro.tex
update

@@
 def power(b: Rep[Int], e: Int): Rep[Int] = 
-  if (e == 0) 1 else e * power(b, e-1)
+  if (e == 0) 1 else b * power(b, e - 1)
 \end{lstlisting}
@@
 \begin{lstlisting}
-def power5(b: Int) = b * b * b * b * b
+def power5(b: Int): Int = b * b * b * b * b
 \end{lstlisting}
@@
 we can refactor interpreters to static analyzers. As an intellectual quest, it
-is natural to ask --- starting from an interpreter, \textbf{can we derive a
-sound abstract interpreter but it works like a code generator which produces fast
-analysis code?} Conceptually, it is similar to the concrete setting: for an
+is natural to ask --- starting from an interpreter, \textit{can we derive a
+sound abstract interpreter while it produces fast code that does the analysis?}
+Conceptually, it is similar to the concrete setting: for an
 abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
@@
 Besides the intellectual merit shown in the confluence diagram, this paper
-contributes to a practical problem in static analysis: \textbf{can we speed up a
+contributes to a practical problem in static analysis: \textit{can we speed up a
 static analyzer without any intrusive changes and soundness compromises?} Static

### 2019-02-27 paper/21_bg_interp.tex
update

@@
 
+The abstract syntax we present in fact can be seen as a deep embedding of the
+language -- we use datatypes to represent programs. This design choice makes us
+easily add different interpretations over the AST; with the inheritance and
+overriding mechanism in Scala, we may also add new language construct and reuse
+the existing interpretation.
+
+\iffalse
 We will give the concrete semantics using a big-step definitional
@@
 \cite{DBLP:conf/popl/FelleisenF87}.
+\fi
+
+\subsection{Monads in Scala}
+
+In Scala, thanks for its nature that combines functional programming and
+object-oriented programming, we may use traits and objects to implement monads.
+A trait poses monadic operations as abstract methods that objects extending it
+should implement. An object encapsulates the data and implements monadic methods
+that operate on these data. The monadic @bind@ operation is usually called
+@flatMap@ in Scala. For example, the following trait @Monad@ declares the
+@return@ and @flatMap@ operation:
+
+\begin{lstlisting}
+  trait Monad[F[_]] {
+    def return[A](a: => A): F[A]
+    def flatMap[A,B](fa: F[A])(f: A => F[B]): F[B]
+  }
+\end{lstlisting}
+
+An object may encapsulate the data of type @F[A]@ and only provide
+@flatMap(f: A => F[B]): F[B]@. This allows us to utilize the for-comprehensions in
+Scala, which provide special syntactic support for moandic @flatMap@.
+For example, a list object of element type @A@ is a monad and it implements the
+methods @flatMap[B](f: A => List[B]): List[B]@ and @map[B](f: A => B): List[B]@.
+To compute the Cartesian product of two lists of numbers, we may directly call
+@flatMap@ and pass an anonymous function that internally calls the monadic
+@return@ operation (the @List@ constructor in this case):
+
+\begin{lstlisting}
+  val xs = List(1, 2); val ys = List(4, 5)
+  xs.flatMap(x => ys.flatMap(y => List((x, y)))) // List((1,4), (1,5), (2,4), (2,5))
+\end{lstlisting}
+
+Or, we can use for-comprehensions that computes the same value:
+
+\begin{lstlisting}
+  for {x <- xs; y <- ys} yield (x, y) // List((1,4), (1,5), (2,4), (2,5))
+\end{lstlisting}
+
+The Scala compiler will translate the above for-comprehension expression into
+an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}:
+
+\begin{lstlisting}
+  xs.flatMap { case x => ys.map { case y => (x, y) } }
+\end{lstlisting}
+
+In the rest of the paper, we will use for-comprehension syntax for monadic
+operations, the implementation of monads and monad transformers essentially
+borrows the ground-truth from Haskell.
 
 \subsection{Generic Interface} \label{generic_if}
+
+Monads provide an elegant and extensible way to build
+interpreters\cite{DBLP:conf/popl/LiangHJ95, Steele:1994:BIC:174675.178068,
+Wadler:1992:EFP:143165.143169}, and recently the approach has been extended to
+abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979,
+DBLP:journals/pacmpl/DaraisLNH17}. In section, with the multi-stage programming
+and abstract interpretation in mind, we use moandic operations to construct an
+generic interface in the style of big-step definitional interpreter.
 

### 2019-02-27 code/src/main/scala/sai/direct/core/ai/AbsLC.scala

@@ -298,11 +298,12 @@ object AbstractLC {
                        AOp('*, Var("n"), App(Var("fact"), AOp('-, Var("n"), Lit(1))))))
     val fact5 = Rec("fact", fact, App(Var("fact"), Lit(5)))
     val omega = App(lam, lam)
-
+  
+    /*
     println(AbsInterp.eval_top(id4))
     println(AbsInterp.eval_top(omega))
     println(AbsInterp.eval_top(fact5))
-
+    */
     //val code = specialize(omega)
     //println(code.code)
     //code.eval(())
@@ -313,8 +314,8 @@ object AbstractLC {
                              Lit(2),
                              Lit(3)),
                      Lit(4)))
-    //val code = specialize(p1)
-    //println(code.code)
+    val code = specialize(fact5)
+    println(code.code)
     //code.eval(())
   }
 }

### 2019-02-27 code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -12,7 +12,7 @@ object Evaluation extends AbsLamCalTrait {
 
   def main(args: Array[String]) {
     val progs = List(
-      boyer
+      //boyer
       //euclid,
       //euclid_imp
       //kcfa32

### 2019-02-27 virtmacro-code/src/main/scala/sai/examples/NDMonad.scala

@@ -4,6 +4,14 @@ package examples
 import scalaz._
 import Scalaz._
 
+import scala.virtualization.lms.internal.GenericNestedCodegen
+import scala.virtualization.lms.common.{
+  SetOps => _, SetOpsExp => _, ScalaGenSetOps => _,
+  ListOps => _, ListOpsExp => _, ScalaGenListOps => _,
+  _}
+import org.scala_lang.virtualized.virtualize
+import org.scala_lang.virtualized.SourceContext
+
 object NDTest {
   def test() = {
     type Store = Map[Int, Int]
@@ -96,5 +104,49 @@ object NDTest {
     val result: (Cache, List[(Store, Int)]) = c.run(Map()).run(Map()).run(Map())
     println(result)
 
+  }
+}
+
+object RepListTExample {
+  trait ExampleOps extends SAIDsl with RepListTransfomer {
+    type Store0 = Map[Int, Int]
+    type Cache0 = Map[String, Set[String]]
+
+    type Store = Rep[Store0]
+    type Cache = Rep[Cache0]
+
+    type OutCacheT[F[_], B] = StateT[F, Cache, B]
+    type NondetT[F[_], B] = RepListT[F, B]
+    type StoreT[F[_], B] = StateT[F, Store, B]
+
+    type OutCacheM[T] = OutCacheT[Id, T]
+    type NondetM[T] = NondetT[OutCacheM, T]
+    type AnsM[T] = StoreT[NondetM, T]
+    type Ans = AnsM[Int]
+
+    def RepListLiftM[G[_], A: Manifest](a: G[Rep[A]])(implicit G: Monad[G]): RepListT[G, A] =
+      RepListT(G.map(a)(entry => List(entry)))
+
+    def get_outcache: AnsM[Cache] = {
+      val fa = StateT.stateTMonadState[Cache, Id].get
+      val ga = RepListT[OutCacheM, Cache0](fa.map((entry: Cache) => List(entry)))
+      val f: Store => RepListT[OutCacheM, (Store0, Cache0)] =
+        (s: Store) => ga.map((a: Cache) => {
+                               val res: Rep[(Store0, Cache0)] = (s, a)
+                               res
+                             })
+      StateT[NondetM, Store, Cache](f)
+      // other monads uses Rep type, add pass into NondetT, where only takes non-Rep type.
+      // TODO: rewrite StateT to use Rep type internally
[…]

### 2019-02-28 paper/10_intro.tex
misc

@@
 \begin{itemize}
-  \item We progressively present a unifying construction that naturally extends the
-    first Futamura projection to abstract interpreters. Starting from a generic
-    interface of the interpreter for a tiny functional language
+  \item We progressively present a unifying construction that naturally extends
+    the first Futamura projection to abstract interpreters. Starting from a generic
+    interface of the interpreter for a small functional language
     (Section~\ref{generic_if}), we first show the instantiation of concrete
-    interpretation in Section~\ref{unstaged_conc}. Then we lift it to the staged
-    version by adding stage annotations to the semantic operations
-    (Section~\ref{stagedinterp}); and to the abstract interpreter by applying
-    abstraction to the environment, store, and values. Finally, we show the
-    combination of the two transformations, dubbed \textit{staged abstract
-    interpreters}, can be easily derived (Section~\ref{sai}).
+    interpretation (Section~\ref{unstaged_conc}). Then we lift it to the staged
+    version by replacing the binding-time type (Section~\ref{stagedinterp}); and to
+    the abstract interpreter by applying abstraction to the environment, store, and
+    values (Section~\ref{unstaged_abs}). Finally, we show the combination of the two
+    transformations, dubbed \textit{staged abstract interpreters}, can be easily
+    derived (Section~\ref{sai}).
 
-  \item \todo{modular analysis} As shown in Section~\ref{sai}, we demonstrate
-    that if we apply the abstract interpreters on open programs, it not only
-    improves the \textit{efficiency} but also the \textit{scalability}, which are
-    two major issues in static analysis.
+  \item We demonstrate that if we apply the abstract interpreter on components
+    of the program separately, it not only improves the \textit{efficiency} but
+    also the \textit{scalability} (Section~\ref{sai} \todo{}), which are two
+    major issues in static analysis.
 
-  \item In Section~\ref{cases_study}, we conduct three case studies to
-    demonstrate the applicability and usefulness of the proposed approach to
-    specialize analyses.
+  \item We conduct two case studies to show the applicability and usefulness
+    of applying meta-programming to static analysis (Section~\ref{cases_study}).
     (1) We first revisit the Abstract Compilation (AC)
-    \cite{Boucher:1996:ACN:647473.727587} technique in Section~\ref{cs_ac}.
+    \cite{Boucher:1996:ACN:647473.727587} technique (Section~\ref{cs_ac}).
     Our approach and AC are able to achieve the same goal, however, as we show,
@@
     whole-analyzer refactoring.
-    (2) We extend the staged abstract interpreter shown in Section~\ref{sai} to
-    two variants of control-flow analysis: context-sensitive pushdown
-    control-flow analysis and \todo{precise-store?}.
-    (3) In Section~\ref{cases_imp}, we plug a \note{staged numerical domain}
-    \todo{widening} into the staged abstract interpreter, demonstrating that a
-    numerical analysis on imperative languages can be also implemented and
-    speeded-up in this manner.
+    (2) We extend the staged abstract interpreter to two variants of
+    control-flow analysis: context-sensitive pushdown control-flow analysis and
+    \todo{precise-store?} (Section~\ref{sai}).
 

### 2019-02-28 code/src/main/scala/sai/direct/core/ai/Monadic.scala

@@ -333,43 +333,45 @@ object Monadics {
     def mod(f: S => S): State[S, Unit] = State(s => ((), f(s)))
   }
 
-  case class StateT[M[_]: Monad, S, A](run: S => M[(A, S)])
+  case class StateT[M[_]: Monad, S, A](run: S => M[(A, S)]) {
+    def flatMap[B](f: A => StateT[M,S,B]): StateT[M,S,B] =
+      StateT(s => run(s).flatMap { case (a, s1) => f(a).run(s1) })
+    def map[B](f: A => B): StateT[M, S, B] =
+      flatMap(a => StateT.pure[M, S, B](f(a)))
+  }
+
   object StateT {
+    def pure[M[_]: Monad, S, A](a: A): StateT[M,S,A] = StateT(s => Monad[M].pure((a, s)))
     def apply[M[_]: Monad, S, A](implicit m: StateT[M, S, A]): StateT[M, S, A] = m
     implicit def apply[M[_]: Monad, S]: Monad[StateT[M, S, ?]] = StateTMonad[M, S]
   }
+
   def StateTMonad[M[_]: Monad, S] = new MonadState[StateT[M, S, ?], S] {
-    def pure[A](a: A): StateT[M,S,A] = StateT(s => Monad[M].pure((a, s)))
-    def flatMap[A, B](sa: StateT[M,S,A])(f: A => StateT[M,S,B]): StateT[M,S,B] =
-      StateT(s => sa.run(s).flatMap { case (a, s1) => f(a).run(s1) })
     def get: StateT[M,S,S] = StateT(s => Monad[M].pure((s, s)))
     def put(s: S): StateT[M,S,Unit] = StateT(s => Monad[M].pure(((), s)))
     def mod(f: S => S): StateT[M,S,Unit] = StateT(s => Monad[M].pure(((), f(s))))
   }
 
-  /*
   case class ReaderStateInterp() extends ReaderSemantics {
     type Store = Int
     type M[T] = ReaderT[StateT[Id, Store, ?], Env, T]
     implicit val s = StateTMonad[Id, Store]
     implicit val m = ReaderTMonad[StateT[Id, Store, ?], Env]
-    import m._
-    import s._
     override def interp(e: Term): M[Value] = e match {
+      /*
       case Tick() => for {
-        _ <- mod(_ + 1)
-        t <- get
+        _ <- s.mod(_ + 1)
+        t <- s.get
         i <- num(t)
       } yield i
+      */
       case Fetch() => for {
-        t <- get
[…]

### 2019-03-07 paper/10_intro.tex
update

@@
 Now we have seen two orthogonal transformations of interpreters: by
-specialization, we can refactor interpreters to code generators; by approximation,
-we can refactor interpreters to static analyzers. As an intellectual quest, it
-is natural to ask --- starting from an interpreter, \textit{can we derive a
-sound abstract interpreter while it produces fast code that does the analysis?}
+specialization, we can refactor interpreters to code generators; by
+approximation, we can refactor interpreters to static analyzers. As an
+intellectual quest, we should naturally ask -- starting from an interpreter,
+\textit{can we derive a sound abstract interpreter while it produces fast code
+that does the analysis?}
 Conceptually, it is similar to the concrete setting: for an
@@
 domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14, Hofer:2008:PED:1449913.1449935}, we develop an
-approach to fulfill four different semantics (Figure~\ref{confluence}) in a
-unified framework. The four semantics share one same generic interpreter.
-However, by instantiating the interpreter differently, it can behave as a
-concrete interpreter, an abstracter interpreter, or their code-generator
-versions, respectively. The key idea that enables this flexibility is to use
-abstract type members abstract over concrete/abstract components (e.g., concrete
-values or abstract values), as well as the binding-time of them (e.g., static or
-dynamic). By staging, the overhead caused by the monadic layers is eliminated in
-the generated code.
+DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14,
+Hofer:2008:PED:1449913.1449935}, we develop an approach to fulfill four
+different semantics (Figure~\ref{confluence}) in a unified framework. The four
+semantics share one same generic interpreter interface. Later, by instantiating
+the interpreter differently, it can behave as a concrete interpreter, an
+abstracter interpreter, or their code-generator versions, respectively. The key
+idea that enables this flexibility is to use abstract type members abstract over
+concrete/abstract components (e.g., concrete values or abstract values), as well
+as the binding-time of them (e.g., static or dynamic). By staging, the overhead
+caused by the monadic layers is eliminated in the generated code.
 
-\paragraph{Performant via Specialization}
+\paragraph{Performance via Specialization}
 
@@
 to improve the latter one while the former one is untouched. Specialization
-removes the overhead of abstractions introduced by high-level programming, but
-not changes the meaning of the program -- in this case, the abstract semantics
-implemented by the analyzer. Indeed, specializing analysis with respect to a
-program by partial evaluation \cite{damian1999partial, amtoft1999partial,
-Boucher:1996:ACN:647473.727587, ashley:practical} is not a new idea, despite the
-fact that most of the previous work are ad-hoc on a particular analysis or
-requiring significant changes to the analysis. For example, abstract compilation
-\cite{Boucher:1996:ACN:647473.727587} requires the whole analyzer to be
-rewritten as closure generation from. In this paper, we show a systematic
-approach to meta-programming achieves the performance goal with less intrusive
-changes; in fact, a case study (Section~\ref{cs_ac}) comparing with abstract
-compilation shows, by utilizing type-based stage annotations, the analyzer
-program does not need any changes.
+removes the overhead of abstractions introduced by high-level programming (e.g.,
+monads), but not loses any advantages of this programming style (e.g.,
+extensibility), nor changes the meaning of the program -- in this case, the
+abstract semantics that does the analysis. Indeed, specializing analysis
+with respect to a program by partial evaluation \cite{damian1999partial,
+amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical} is not a
+new idea, despite the fact that most of the previous work are ad-hoc on a
+particular analysis or requiring significant changes to the analysis. For
+example, abstract compilation \cite{Boucher:1996:ACN:647473.727587} requires the
+whole analyzer to be rewritten as closure generation from. In this paper, we
+show a systematic approach to meta-programming that achieves the performance
+goal with less intrusive changes; in fact, a case study (Section~\ref{cs_ac})
+comparing with abstract compilation shows, by utilizing type-based stage
+annotations, the analyzer program does not need any changes.
 
@@
 
-\todo{modular analysis: a whole-program abstract
-  interpreter can be run on different components of program separately, we
-  achieve this by meta-programming mechnically}
+Our second contribution is to enable modular analysis based on a whole-program
+analyzer by meta-programming. Modern software are shipped with large libraries
+\cite{toman_et_al:LIPIcs:2017:7121}, and analyzing programs with large libraries
+is usually expensive or unnecessarily repeated for whole-program analyzing
+algorithms. For example, it has been shown that analyzing a simple ``Hello
+World'' program in Java depends on additional 3,000 classes in the library
+\cite{DBLP:conf/oopsla/KulkarniMZN16}. Meta-programming gives us a chance to
+tweak a whole-program analyzing algorithm into an analyzer that can work on
+different components of program separately. Still, the idea is to specialize the
+abstract interpreter used for whole-program analysis, but on a subset of the
+program, for examples, several generic functions for list manipulation. Such
+specialization works on a lambda expression and generates code, which is partial
+analysis results and can be completed by running with the necessary arguments
+(for example, an environment that closes the lambda term) at next stages. Multiple
+partial analysis results also can be reused later and composed together as they
+are just normal programs at the next stage. Through staging, we mechanically
+obtain a modular analysis, even though the original algorithm is formulated as a
+whole-program analysis.
 
@@
 but modular analysis has different asymptotics}
-\fi
 
@@
 
-\iffalse
 It has been observed that partially applying context-sensitivity on selected
@@
   Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},
-  deep embedding, four different interpretations.
 }

### 2019-03-07 paper/21_bg_interp.tex
update

@@
 object-oriented programming, we may use traits and objects to implement monads.
-A trait poses monadic operations as abstract methods that objects extending it
-should implement. An object encapsulates the data and implements monadic methods
-that operate on these data. The monadic @bind@ operation is usually called
-@flatMap@ in Scala. For example, the following trait @Monad@ declares the
-@return@ and @flatMap@ operation:
+For example, the following trait @Monad@ declares the @pure@ and @flatMap@
+operation. The trait itself takes a higher-kinded type @F[_]: * -> *@
+which takes a type and returns a type. The method @pure@ promotes values of type
+@A@ to values of type @F[A]@. The monadic @bind@ operation is usually called
+@flatMap@ in Scala, which takes a monad-encapsulated value of type @F[A]@, a function of
+@A => F[B]@ and returns values of type @F[B]@.
 
@@
   trait Monad[F[_]] {
-    def return[A](a: => A): F[A]
+    def pure[A](a: A): F[A]
     def flatMap[A,B](fa: F[A])(f: A => F[B]): F[B]
@@
 
-An object may encapsulate the data of type @F[A]@ and only provide
-@flatMap(f: A => F[B]): F[B]@. This allows us to utilize the for-comprehensions in
-Scala, which provide special syntactic support for moandic @flatMap@.
-For example, a list object of element type @A@ is a monad and it implements the
-methods @flatMap[B](f: A => List[B]): List[B]@ and @map[B](f: A => B): List[B]@.
-To compute the Cartesian product of two lists of numbers, we may directly call
-@flatMap@ and pass an anonymous function that internally calls the monadic
-@return@ operation (the @List@ constructor in this case):
+An object implementing the trait @Monad@ may also encapsulates the data of type
+@F[A]@ internally, in this case, it only needs to provide method
+@flatMap(f: A => F[B]): F[B]@. This allows us to utilize the for-comprehensions
+in Scala, simliar to Haskell's do-notation, which provides special syntactic
+support for moandic @flatMap@. For example, a list object of element type @A@ is
+a monad and it implements the methods @flatMap[B](f: A => List[B]): List[B]@ and
+@map[B](f: A => B): List[B]@. To compute the Cartesian product of two lists of
+numbers, we may directly call @flatMap@ and pass an anonymous function that
+internally calls the monadic @pure@ operation (the @List@ constructor in this
+case):
 
@@
 
-In the rest of the paper, we will use for-comprehension syntax for monadic
-operations, the implementation of monads and monad transformers essentially
-borrows the ground-truth from Haskell.
+In the rest of the paper, we use Scala's for-comprehension syntax and monad
+transformers @ReaderT@, @StateT@, @ListT@ to write our interpreters. The
+implementation of monads and monad transformers essentially borrows the
+ground-truth from Haskell.
 

### 2019-03-08 paper/10_intro.tex
misc

@@
 World'' program in Java depends on additional 3,000 classes in the library
-\cite{DBLP:conf/oopsla/KulkarniMZN16}. Meta-programming gives us a chance to
-tweak a whole-program analyzing algorithm into an analyzer that can work on
-different components of program separately. Still, the idea is to specialize the
-abstract interpreter used for whole-program analysis, but on a subset of the
-program, for examples, several generic functions for list manipulation. Such
-specialization works on a lambda expression and generates code, which is partial
-analysis results and can be completed by running with the necessary arguments
-(for example, an environment that closes the lambda term) at next stages. Multiple
+\cite{DBLP:conf/oopsla/KulkarniMZN16}. A natural idea is to analyze libraries
+and generate summaries, later we can plugin the necessary information into the
+summaries and finish the analysis; therefore such summaries should be modular
+and reusable. Meta-programming gives us a chance to readily tweak a
+whole-program analyzing algorithm into an analyzer that works modularly and
+produces summaries. Still, the idea is to specialize the abstract interpreter
+that was used for whole-program analysis, but on a subset of the program, for
+examples, several generic functions for list manipulation. Such specialization
+works on lambda expressions and generates code, which is partial analysis
+results and can be completed by running with the necessary arguments (for
+example, an environment that closes the lambda term) at next stages. Multiple
 partial analysis results also can be reused later and composed together as they
@@
 
-\note{
-  Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},
-}
+%\note{Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},}
 

### 2019-03-08 virtmacro-code/src/main/scala/sai/lms/LMSSupport.scala

@@ -1,0 +1,26 @@
+package sai
+package lms
+
+import scala.virtualization.lms.internal.GenericNestedCodegen
+import scala.virtualization.lms.common.{
+  SetOps => _, SetOpsExp => _, ScalaGenSetOps => _,
+  ListOps => _, ListOpsExp => _, ScalaGenListOps => _,
+  _}
+import org.scala_lang.virtualized.virtualize
+import org.scala_lang.virtualized.SourceContext
+
+trait SAIDsl extends Dsl
+    with MapOps
+    with SetOps
+    with ListOps
+    with TupleOps
+    with UncheckedOps
+    with TupledFunctions
+
+trait SAIOpsExp extends DslExp
+    with MapOpsExp
+    with SetOpsExp
+    with ListOpsExp
+    with TupleOpsExp
+    with UncheckedOpsExp
+    with TupledFunctionsRecursiveExp

### 2019-03-08 virtmacro-code/src/main/scala/sai/monads/RepMonads.scala

@@ -172,6 +172,8 @@ trait SAIMonads { self: SAIDsl =>
 
   def emit_foldMap[M[_]: Monad, A: Manifest, B: Manifest](xs: Rep[List[A]], f: Rep[A => ListT[M, B]]): Rep[M[List[B]]]
 
+
+
   case class ListT[M[_]: Monad, A: Manifest](run: M[List[A]]) {
     import ListT._
 
@@ -191,17 +193,18 @@ trait SAIMonads { self: SAIDsl =>
                 /*
                 list.foldLeft(ListT.empty[M, B]) {
                   case (acc: ListT[M, B], a: Rep[A]) => acc ++ f(a)
-                }
-                */
+                 }*/
                 list.foldLeft(List[B]()) {
                   case (acc: Rep[List[B]], a: Rep[A]) =>
                     val fa: ListT[M, B] = f(a)
                     val listb: M[List[B]] = Monad[M].map(fa.run) { case falist: Rep[List[B]] => acc ++ falist }
+                    Comonad[M].extract(listb)
                     ??? // expected Rep[List[B]], a comonad operation `extract: w a -> a`!
                 }
               Monad[M].pure[List[B]](merge)
             })
 
+    /*
     @virtualize
     def flatMapCps[M[_]: Monad, B: Manifest](list: Rep[List[A]], acc: ListT[M, B], f: Rep[A] => ListT[M, B], k: Rep[List[B]] => Rep[List[B]]): Rep[List[B]] = {
       if (list.isEmpty) {
@@ -212,6 +215,7 @@ trait SAIMonads { self: SAIDsl =>
         flatMapCps(list.tl, nacc, f, k)
       }
     }
+     */
 
     def map[B: Manifest](f: Rep[A] => Rep[B]): ListT[M, B] =
       ListT(Monad[M].flatMap(run) { list => Monad[M].pure(list.map(f)) })

### 2019-03-15 paper/10_intro.tex
some revision

@@
 \begin{lstlisting}
-def power(b: Rep[Int], e: Int): Rep[Int] = 
-  if (e == 0) 1 else b * power(b, e - 1)
+def power(b: Rep[Int], e: Int): Rep[Int] = if (e == 0) 1 else b * power(b, e - 1)
 \end{lstlisting}
@@
 projection, we may obtain a specialized interpreter
-$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$, and by definition of the interpreter, 
+$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$ -- the so-called \textit{equivalent executable}. By definition of the interpreter, 
 $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
 argument @arg@ to an environment (as well as a companion heap object, if for the
-interest of using assignments) that contains values mapped from free variables
-in $e$, the interpreter is a standard environment/store-passing interpreter,
+interest of using assignments) that maps free variables in $e$ to some values, 
+the interpreter is a standard environment/store-passing interpreter,
 where the specialization through MSP is also applicable.
@@
 definitional abstract interpreters, which intersects as staged abstract
-interpreters. By borrowing the ideas from monadic interpreters
+interpreters. We develop a unified approach to fulfill the four different
+semantics, as shown in Figure~\ref{confluence}. Starting with a generic but
+parametric interpreter shared by the four semantics, by simply instantiating
+the interpreter differently, we can automatically derive a concrete interpreter,
+or an abstract interpreter, or their code-generator versions, respectively.
+
+\iffalse
+embedded domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
+DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14,
+Hofer:2008:PED:1449913.1449935},
+The key idea that enables this flexibility is to use abstract type members first
+abstract over concrete/abstract components (e.g., concrete values or abstract
+values), then as well as the binding-time of them (e.g., static or dynamic). By
+staging, the overhead caused by the monadic layers is eliminated in the
+generated code.
+\fi
+
+\paragraph{Abstraction, Abstraction and Abstraction}
+To enable such flexibility between different semantics, we extensively use
+abstract type members over multiple abstractions, which can be later instantiated.
+
+\begin{itemize}
+\item Monadic abstractions, which encapsulate the following two abstractions.
+  The interpreter itself is written in monadic form
   \cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95,
-DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979} and embedded
-domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14,
-Hofer:2008:PED:1449913.1449935}, we develop an approach to fulfill four
-different semantics (Figure~\ref{confluence}) in a unified framework. The four
-semantics share one same generic interpreter interface. Later, by instantiating
-the interpreter differently, it can behave as a concrete interpreter, an
-abstracter interpreter, or their code-generator versions, respectively. The key
-idea that enables this flexibility is to use abstract type members abstract over
-concrete/abstract components (e.g., concrete values or abstract values), as well
-as the binding-time of them (e.g., static or dynamic). By staging, the overhead
-caused by the monadic layers is eliminated in the generated code.
+  DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}, by plugging
+  different monads, the underlying semantics of the interpreter is changed.
+  \todo{mention that direct-style also works?}
+\item Env-store-value abstractions, which control whether the interpreter
+  behaves as a concrete interpreter or an abstract interpreter.
+  The environments, stores and values declared in the generic interpreter are
+  abstract types; we also define several abstract methods that manipulate
+  environments or stores.
+  As shown by the AAM approach \cite{DBLP:journals/jfp/HornM12,
+  DBLP:conf/icfp/HornM10}, if we widen the values to abstract domains (such as
+  intervals) and refactor the store to a mapping from finite addresses to sets of
+  values, we can derive a computable and sound abstract interpreter.
+\item Binding-time abstractions, which control whether the interpreter produces
+  values or residual programs. The binding-times differentiate whether the
+  data is known statically or dynamically. Unsurprisingly, such binding-time
+  information can be encoded into higher-kinded types \todo{cite FTPE, stage poly}.
+\end{itemize}
+
+\paragraph{Abstraction without Regret}
+
+We show that these well-known abstractions can be combined smoothly, and more
+importantly the price is almost free. Through staging, the major overhead of
+abstractions -- from the monadic layers -- is completely eliminated in the
+generated code. However, we do not lose any advantages of this high-level
+programming style (e.g., the extensibility); the meaning of the program also
+remains the same, particularly, the abstract semantics that does the analysis
+when instantiated as an abstract interpreter.
 
@@
 programming, in particular) with abstract interpreters is an effective approach
-to improve the latter one while the former one is untouched. Specialization
-removes the overhead of abstractions introduced by high-level programming (e.g.,
-monads), but not loses any advantages of this programming style (e.g.,
-extensibility), nor changes the meaning of the program -- in this case, the
-abstract semantics that does the analysis. Indeed, specializing analysis
-with respect to a program by partial evaluation \cite{damian1999partial,
-amtoft1999partial, Boucher:1996:ACN:647473.727587, ashley:practical} is not a
-new idea, despite the fact that most of the previous work are ad-hoc on a
-particular analysis or requiring significant changes to the analysis. For
-example, abstract compilation \cite{Boucher:1996:ACN:647473.727587} requires the
-whole analyzer to be rewritten as closure generation from. In this paper, we
-show a systematic approach to meta-programming that achieves the performance
-goal with less intrusive changes; in fact, a case study (Section~\ref{cs_ac})
-comparing with abstract compilation shows, by utilizing type-based stage
-annotations, the analyzer program does not need any changes.
+to improve the latter one while the former one is untouched.
+In the literature, specializing analysis with respect to a program by partial
+evaluation \cite{damian1999partial, amtoft1999partial,
+Boucher:1996:ACN:647473.727587, ashley:practical} is not a new, despite the fact
+that most of the previous work are ad-hoc on a particular analysis or requiring
+significant changes to the analyzer. For example, abstract compilation
+\cite{Boucher:1996:ACN:647473.727587} requires the whole analyzer to be
+rewritten as closure generation from. In this paper, we advocate a systematic
+approach to meta-programming that achieves the performance goal with less
+intrusive changes; in fact, a case study (Section~\ref{cs_ac}) comparing with
+abstract compilation shows, by utilizing type-based stage annotations, the
+analyzer program does not need any changes.
 
@@
 
-Our second contribution is to enable modular analysis based on a whole-program
+Our next contribution is to enable modular analysis based on a whole-program
 analyzer by meta-programming. Modern software are shipped with large libraries
@@
 that was used for whole-program analysis, but on a subset of the program, for
-examples, several generic functions for list manipulation. Such specialization
+examples, a set of generic functions for list manipulation. Such specialization
 works on lambda expressions and generates code, which is partial analysis
@@
 whole-program analysis.
+
+\paragraph{The paper is organized as follows:}
+
+\begin{itemize}
+  \item We progressively present a unifying construction that naturally extends
+    the first Futamura projection to abstract interpreters. Starting from a generic
+    interface of the interpreter for a small functional language
+    (Section~\ref{generic_if}), we first show the instantiation of concrete
+    interpretation (Section~\ref{unstaged_conc}). Then we lift it to the staged
+    version by replacing the binding-time types (Section~\ref{stagedinterp}); and to
+    the abstract interpreter by applying abstraction to the environment, store, and
+    values (Section~\ref{unstaged_abs}). Finally, we show the combination of the two
+    transformations, dubbed \textit{staged abstract interpreters}, can be easily
+    derived (Section~\ref{sai}).
+
+  \item We demonstrate that if we apply the abstract interpreter on components
+    of the program separately, it not only improves the \textit{efficiency} but
+    also the \textit{scalability} (Section~\ref{sai} \todo{}), which are two
+    major issues in static analysis.
+
+  \item We conduct two case studies to show the applicability and usefulness
+    of applying meta-programming to static analysis (Section~\ref{cases_study}).
+    (1) We first revisit the Abstract Compilation (AC)
+    \cite{Boucher:1996:ACN:647473.727587} technique (Section~\ref{cs_ac}).
+    Our approach and AC are able to achieve the same goal, however, as we show,
+    one of the advantage of staging through types is it does not require
+    whole-analyzer refactoring.
+    (2) We extend the staged abstract interpreter to two variants of
+    control-flow analysis: context-sensitive pushdown control-flow analysis and
+    \todo{precise-store?} (Section~\ref{sai}).
+
+  \item Based on the instantiation of control-flow analysis, we empirically
+    evaluate the performance improvement by staging and specialization
+    (Section~\ref{evaluation}). We compare the both context-insensitive and
+    \note{context-sensitive} analysis.
+
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 %\note{Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},}
-
-\paragraph{Contributions and Outline}
-
-\begin{itemize}
-  \item We progressively present a unifying construction that naturally extends
-    the first Futamura projection to abstract interpreters. Starting from a generic
-    interface of the interpreter for a small functional language
-    (Section~\ref{generic_if}), we first show the instantiation of concrete
-    interpretation (Section~\ref{unstaged_conc}). Then we lift it to the staged
-    version by replacing the binding-time type (Section~\ref{stagedinterp}); and to
-    the abstract interpreter by applying abstraction to the environment, store, and
-    values (Section~\ref{unstaged_abs}). Finally, we show the combination of the two
-    transformations, dubbed \textit{staged abstract interpreters}, can be easily
-    derived (Section~\ref{sai}).
-
-  \item We demonstrate that if we apply the abstract interpreter on components
-    of the program separately, it not only improves the \textit{efficiency} but
-    also the \textit{scalability} (Section~\ref{sai} \todo{}), which are two
-    major issues in static analysis.
-
-  \item We conduct two case studies to show the applicability and usefulness
-    of applying meta-programming to static analysis (Section~\ref{cases_study}).
-    (1) We first revisit the Abstract Compilation (AC)
-    \cite{Boucher:1996:ACN:647473.727587} technique (Section~\ref{cs_ac}).
-    Our approach and AC are able to achieve the same goal, however, as we show,
-    one of the advantage of staging through types is it does not require
-    whole-analyzer refactoring.
-    (2) We extend the staged abstract interpreter to two variants of
-    control-flow analysis: context-sensitive pushdown control-flow analysis and
-    \todo{precise-store?} (Section~\ref{sai}).
-
-  \item Based on the instantiation of control-flow analysis, we empirically
-    evaluate the performance improvement by staging and specialization
-    (Section~\ref{evaluation}). We compare the both context-insensitive and
-    \note{context-sensitive} analysis.
-
-\end{itemize}

### 2019-03-15 paper/20_bg.tex
some revision

@@
 languages, such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
-  DBLP:conf/flops/Kiselyov14} or Template Haskell \cite{Sheard:2002:TMH:636517.636528}.
+  DBLP:conf/flops/Kiselyov14} or Template Haskell \cite{Sheard:2002:TMH:636517.636528} \todo{cite the recent one}.
 We first instantiate the interpreter to the concrete one; then we

### 2019-03-15 paper/21_bg_interp.tex
some revision

@@
 For example, the following trait @Monad@ declares the @pure@ and @flatMap@
-operation. The trait itself takes a higher-kinded type @F[_]: * -> *@
+operation. The trait itself takes a higher-kinded type @F[_]: * -> *@ as argument,
 which takes a type and returns a type. The method @pure@ promotes values of type
-@A@ to values of type @F[A]@. The monadic @bind@ operation is usually called
+@A@ to values of type @F[A]@ \footnote{We elect to use \texttt{pure} as the name, since \texttt{return} is a keyword in Scala and \texttt{unit} is a built-in function in LMS.}.
+The monadic @bind@ operation is usually called
 @flatMap@ in Scala, which takes a monad-encapsulated value of type @F[A]@, a function of
@@
 
-\paragraph{Stage Polymorphism}
+\paragraph{Binding-time Abstraction}
+%\paragraph{Stage Polymorphism}
+% Binding-time abstraction via abstract types
 

### 2019-03-18 paper/10_intro.tex
misc

@@
 Now to concretely illustrate the idea of Futamura projection, consider that we
-have an interpreter @eval@ of some language as following:
+have an operational interpreter @eval@ of some language as following:
 
@@
 projection, we may obtain a specialized interpreter
-$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$ -- the so-called \textit{equivalent executable}. By definition of the interpreter, 
+$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$ -- the so-called \textit{equivalent executable}. 
+By definition of the interpreter, 
 $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
@@
 
-\paragraph{Abstraction, Abstraction and Abstraction}
+\paragraph{Abstraction, Abstraction, and Abstraction}
 To enable such flexibility between different semantics, we extensively use

### 2019-03-21 paper/10_intro.tex
misc

@@
 \paragraph{Abstraction, Abstraction, and Abstraction}
-To enable such flexibility between different semantics, we extensively use
-abstract type members over multiple abstractions, which can be later instantiated.
+To enable such flexibility between different semantics, we extensively use generic 
+programming and abstract type members over multiple abstractions, which can be 
+later instantiated.
 

### 2019-03-21 virtmacro-code/src/main/scala/sai/Main.scala

@@ -184,7 +184,7 @@ trait StagedCESOps extends SAIDsl with SAIMonads {
 
   sealed trait Value
   //case class IntV(i: Int) extends Value
-  //case class CloV[Env](lam: Lam, e: Env) extends Value
+  case class CloV[Env](lam: Lam, e: Env) extends Value
 
   type Ident = String
 

### 2019-03-21 paper/10_intro.tex
replace Id with IdM

--- 2019-03-18 paper/10_intro.tex
@@
 \paragraph{Abstraction, Abstraction, and Abstraction}
-To enable such flexibility between different semantics, we extensively use
-abstract type members over multiple abstractions, which can be later instantiated.
+To enable such flexibility between different semantics, we combine generic 
+programming and generative programming, and extensively use abstract type members 
+over multiple abstractions, which can be later instantiated.
 
@@
 \item Binding-time abstractions, which control whether the interpreter produces
-  values or residual programs. The binding-times differentiate whether the
-  data is known statically or dynamically. Unsurprisingly, such binding-time
-  information can be encoded into higher-kinded types \todo{cite FTPE, stage poly}.
+  values or generates code. The binding-times differentiate whether data are 
+  known statically or dynamically. Unsurprisingly, such binding-time
+  information can be encoded into higher-kinded types 
+    \cite{DBLP:journals/jfp/CaretteKS09, Ofenbeck:2017:SGP:3136040.3136060}.
 \end{itemize}

### 2019-03-21 virtmacro-code/src/main/scala/sai/Main.scala

@@ -19,15 +19,15 @@ object EnvInterpreter {
   import PCFLang.Values._
   import ReaderT._
   import StateT._
-  import IdMonadInstance._
+  import IdM._
 
   type Ident = String
   type Env = Map[Ident, Value]
 
-  type AnsM[T] = ReaderT[Id, Env, T]
+  type AnsM[T] = ReaderT[IdM, Env, T]
   type Ans = AnsM[Value]
 
-  def num(i: Int): Ans = ReaderTMonad[Id, Env].pure[Value](IntV(i))
+  def num(i: Int): Ans = ReaderTMonad[IdM, Env].pure[Value](IntV(i))
 
   def prim(op: Symbol, v1: Value, v2: Value): Value = (op, v1, v2) match {
     case ('+, IntV(x), IntV(y)) => IntV(x + y)
@@ -38,7 +38,7 @@ object EnvInterpreter {
 
   def ap_clo(fun: Value, arg: Value): Ans = fun match {
     case CloV(Lam(x, e), ρ: Env) =>
-      ReaderTMonad[Id, Env].local(eval(e))(_ => ρ + (x → arg))
+      ReaderTMonad[IdM, Env].local(eval(e))(_ => ρ + (x → arg))
   }
 
   def br0(test: Value, thn: => Ans, els: => Ans): Ans =
@@ -47,10 +47,10 @@ object EnvInterpreter {
   def eval(e: Expr): Ans = e match {
     case Lit(i) => num(i)
     case Var(x) => for {
-      ρ <- ReaderTMonad[Id, Env].ask
+      ρ <- ReaderTMonad[IdM, Env].ask
     } yield ρ(x)
     case Lam(x, e) => for {
-      ρ <- ReaderTMonad[Id, Env].ask
+      ρ <- ReaderTMonad[IdM, Env].ask
     } yield CloV(Lam(x, e), ρ)
     case App(e1, e2) => for {
       v1 <- eval(e1)
@@ -59,8 +59,8 @@ object EnvInterpreter {
     } yield rt
     case Let(x, rhs, e) => for {
       v <- eval(rhs)
-      ρ <- ReaderTMonad[Id, Env].ask
-      rt <- ReaderTMonad[Id, Env].local(eval(e))(_ => ρ + (x → v))
+      ρ <- ReaderTMonad[IdM, Env].ask
+      rt <- ReaderTMonad[IdM, Env].local(eval(e))(_ => ρ + (x → v))
     } yield rt
[…]

### 2019-03-21 virtmacro-code/src/main/scala/sai/monads/MonadOps.scala

@@ -1,14 +1,15 @@
 package sai
 package monads
 
-import sai.lattices._
-import sai.lattices.Lattices._
-
 object NoRep {
-  type NoRep[+T] = T
+  type NoRep[T] = T
 }
+
+import NoRep._
 
 trait RMonadOps[R[_], M[_], A] {
   def map[B](f: R[A] => R[B])(implicit mB: Manifest[B] = null): M[B]
   def flatMap[B](f: R[A] => M[B])(implicit mB: Manifest[B] = null): M[B]
 }
+
+trait MonadOps[M[_], A] extends RMonadOps[NoRep, M, A]

### 2019-03-21 virtmacro-code/src/main/scala/sai/monads/Monads.scala

@@ -20,7 +20,7 @@ trait RMonad[R[_], M[_]] {
          (implicit mA: Manifest[A] = null, mB: Manifest[B] = null): M[B] = flatMap(ma)(a => pure(f(a)))
   def filter[A](ma: M[A])(f: R[A] => R[Boolean])(implicit mA: Manifest[A] = null): M[A]
 }
- */
+*/
 
 trait Monad[M[_]] extends RMonad[NoRep.NoRep, M]
 
@@ -39,6 +39,7 @@ object MonadPlus {
 
 /////////////////////////////////////////////////
 
+@deprecated("Use IdM", "")
 object IdMonadInstance {
   type Id[T] = T
   implicit val IdMonad: Monad[Id] = new Monad[Id] {
@@ -51,6 +52,27 @@ object IdMonadInstance {
     def mzero[A: Lattice]: A = Lattice[A].bot
     def mplus[A: Lattice](a: A, b: A) = Lattice[A].⊔(a, b)
   }
+}
+
+object IdM {
+  def apply[A](implicit m: IdM[A]): IdM[A] = m
+
+  implicit val IdMonadInstance: Monad[IdM] = new Monad[IdM] {
+    def pure[A](a: A): IdM[A] = IdM(a)
+    def flatMap[A, B](ma: IdM[A])(f: A => IdM[B]): IdM[B] = ma.flatMap(f)
+    def filter[A](ma: IdM[A])(f: A => Boolean): IdM[A] = throw new Exception("Not supported")
+  }
+
+  implicit val IdMonadPlus: MonadPlus[IdM] = new MonadPlus[IdM] {
+    def mzero[A: Lattice]: IdM[A] = IdM(Lattice[A].bot)
+    def mplus[A: Lattice](a: IdM[A], b: IdM[A]): IdM[A] = IdM(a.run ⊔ b.run)
+  }
+}
+
+case class IdM[A](run: A) extends MonadOps[IdM, A] {
+  def apply: A = run
+  def flatMap[B](f: A => IdM[B])(implicit mB: Manifest[B] = null): IdM[B] = f(run)
+  def map[B](f: A => B)(implicit mB: Manifest[B] = null): IdM[B] = IdM(f(run))
 }
 
 /////////////////////////////////////////////////

### 2019-03-23 paper/21_bg_interp.tex
[paper] monadic sec

--- 2019-02-27 paper/21_bg_interp.tex
@@
 object-oriented programming, we may use traits and objects to implement monads.
-A trait poses monadic operations as abstract methods that objects extending it
-should implement. An object encapsulates the data and implements monadic methods
-that operate on these data. The monadic @bind@ operation is usually called
-@flatMap@ in Scala. For example, the following trait @Monad@ declares the
-@return@ and @flatMap@ operation:
+For example, the trait @Monad@ (Figure \ref{fig:monad}) declares the @pure@ and @flatMap@
+operation. The trait itself takes a higher-kinded type @M[_]: * -> *@ as
+argument, which takes a type and returns a type. The method @pure@ promotes
+values of type @A@ to values of type @M[A]@ \footnote{We elect to use
+\texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
+\texttt{unit} is a built-in function in LMS.}. The monadic @bind@ operation is
+usually called @flatMap@ in Scala, which takes a monad-encapsulated value of
+type @M[A]@, a function of @A => M[B]@ and returns values of type @M[B]@.
 
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[b]{0.55\textwidth}
     \begin{lstlisting}
-  trait Monad[F[_]] {
-    def return[A](a: => A): F[A]
-    def flatMap[A,B](fa: F[A])(f: A => F[B]): F[B]
+  trait Monad[M[_]] {                                  
+    def pure[A](a: A): M[A]                            
+    def flatMap[A,B](fa: M[A])(f: A => M[B]): M[B]     
   }                                                    
     \end{lstlisting}
+    \caption{trait \texttt{Monad}} \label{fig:monad}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.4\textwidth}
+    \begin{lstlisting}
+trait MonadOps[M[_], A] {
+  def flatMap[B](f: A => M[B]): M[B]
+  def map[B](f: A => B): M[B]
+}
+    \end{lstlisting}
+    \caption{trait \texttt{MonadOps}} \label{fig:monadops}
+  \end{subfigure}
+\end{figure}
 
-An object may encapsulate the data of type @F[A]@ and only provide
-@flatMap(f: A => F[B]): F[B]@. This allows us to utilize the for-comprehensions in
-Scala, which provide special syntactic support for moandic @flatMap@.
-For example, a list object of element type @A@ is a monad and it implements the
-methods @flatMap[B](f: A => List[B]): List[B]@ and @map[B](f: A => B): List[B]@.
-To compute the Cartesian product of two lists of numbers, we may directly call
-@flatMap@ and pass an anonymous function that internally calls the monadic
-@return@ operation (the @List@ constructor in this case):
+Similar to Haskell's @do@-notation, Scala provides special syntactic support for
+monadic operations through @for@-comprehension.
+For example, a object of @List[A]@, where @A@ is the element type. Then to
+compute the the Cartesian product of two lists of numbers, we can use Scala's
+@for@-comprehension syntax.
 
@@
   val xs = List(1, 2); val ys = List(4, 5)
-  xs.flatMap(x => ys.flatMap(y => List((x, y)))) // List((1,4), (1,5), (2,4), (2,5))
-\end{lstlisting}
-
-Or, we can use for-comprehensions that computes the same value:
-
-\begin{lstlisting}
   for {x <- xs; y <- ys} yield (x, y) // List((1,4), (1,5), (2,4), (2,5))
@@
 
-The Scala compiler will translate the above for-comprehension expression into
-an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}:
+The Scala compiler will translate the above @for@-comprehension expression into
+an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
+in the @for@-comprehension is translated into a @map@, where the expression of
+@yield@ becomes the body expression of that @map@ application. The foregoing
+bindings are all translated into calls of @flatMap@.
 
@@
 
-In the rest of the paper, we will use for-comprehension syntax for monadic
-operations, the implementation of monads and monad transformers essentially
-borrows the ground-truth from Haskell.
+Note that here the monadic object @List[_]@ encapsulates the data internally.
+Therefore it only exposes simplified version of @flatMap@ and @map@, where the
+monad @M[A]@ is not introduced as a function argument. The trait @MonadOps@
+(Figure \ref{fig:monadops}) defines the simplified version of monadic
+operations. The conversion between @Monad@ and @MonadOps@ can be done by using
+implicit design pattern.
+In the rest of the paper, we use Scala's for-comprehension syntax and monad
+transformers such as @ReaderT@, @StateT@, and @ListT@ to write our interpreters.
+The implementation of monads and monad transformers essentially borrows the
+ground-truth from Haskell.
 

### 2019-03-23 paper/21_bg_interp.tex
revise

@@
 
-In Scala, thanks for its nature that combines functional programming and
-object-oriented programming, we may use traits and objects to implement monads.
-For example, the trait @Monad@ (Figure \ref{fig:monad}) declares the @pure@ and @flatMap@
-operation. The trait itself takes a higher-kinded type @M[_]: * -> *@ as
-argument, which takes a type and returns a type. The method @pure@ promotes
-values of type @A@ to values of type @M[A]@ \footnote{We elect to use
-\texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
-\texttt{unit} is a built-in function in LMS.}. The monadic @bind@ operation is
+A monad is a type constructor @M[_]: * -> *@ with two operations, often called @return@ and @bind@.
+Informally, @return@ wraps a value into the monad @M@, and @bind@ unwraps the monadic value and 
+transforms it to a new monadic value.
+Pragmatically in Scala, we define a monad using trait @Monad@ (Figure \ref{fig:monad}),
+where it declares the @pure@ \footnote{We elect to use \texttt{pure} as the name, 
+since \texttt{return} is a keyword in Scala and \texttt{unit} is a built-in function in LMS.}
+and @flatMap@ operation. The trait itself takes the monad type @M[_]@ as
+argument, which is a higher-kinded type that takes a type and returns a type. 
+The method @pure@ promotes values of type @A@ to values of type @M[A]@.
+The monadic @bind@ operation is
 usually called @flatMap@ in Scala, which takes a monad-encapsulated value of
@@
 trait MonadOps[M[_], A] {
+  def map[B](f: A => B): M[B]
   def flatMap[B](f: A => M[B]): M[B]
-  def map[B](f: A => B): M[B]
 }
@@
 monadic operations through @for@-comprehension.
-For example, a object of @List[A]@, where @A@ is the element type. Then to
-compute the the Cartesian product of two lists of numbers, we can use Scala's
+For example, an object of @List[A]@ is an instance of @List@ monad, where @A@ is the element type. 
+Then to compute the the Cartesian product of two lists of numbers, we can use Scala's
 @for@-comprehension syntax.
@@
   val xs = List(1, 2); val ys = List(4, 5)
-  for {x <- xs; y <- ys} yield (x, y) // List((1,4), (1,5), (2,4), (2,5))
+  for { x <- xs; y <- ys } yield (x, y) // List((1,4), (1,5), (2,4), (2,5))
 \end{lstlisting}
@@
 @yield@ becomes the body expression of that @map@ application. The foregoing
-bindings are all translated into calls of @flatMap@.
+bindings in the comprehension are all translated into calls of @flatMap@.
 
@@
 Note that here the monadic object @List[_]@ encapsulates the data internally.
-Therefore it only exposes simplified version of @flatMap@ and @map@, where the
+Therefore it only exposes the simplified version of @flatMap@, where the
 monad @M[A]@ is not introduced as a function argument. The trait @MonadOps@
 (Figure \ref{fig:monadops}) defines the simplified version of monadic
-operations. The conversion between @Monad@ and @MonadOps@ can be done by using
-implicit design pattern.
-In the rest of the paper, we use Scala's for-comprehension syntax and monad
+operations that are necessary for @for@-comprehension. 
+The conversion between @Monad@ and @MonadOps@ can be done by using implicit design pattern.
+In the rest of the paper, we use Scala's @for@-comprehension syntax and monad
 transformers such as @ReaderT@, @StateT@, and @ListT@ to write our interpreters.

### 2019-03-23 virtmacro-code/src/main/scala/sai/MainGeneric.scala

@@ -467,9 +467,9 @@ trait AbstractSemantics extends AbstractComponents {
   def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0).run(cache0)(cache0).run
 }
 
-trait StagedAbstractSemantics extends AbstractComponents with RepMonads with SAIDsl {
+trait StagedAbstractSemantics extends AbstractComponents with RepMonads with RepLattices with SAIDsl {
   type R[T] = Rep[T]
-  type AnsM[T] = ReaderT[StateT[ListT[ReaderT[StateT[IdM, Cache, ?], Cache, ?], ?], Store, ?], Env, T]
+  type AnsM[T] = ReaderT[StateT[ListReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
 }
 
 ////////////////////////////////////////////////

### 2019-03-23 paper/10_intro.tex
revise

@@
 
-\paragraph{Abstraction, Abstraction, and Abstraction}
+\paragraph{Abstraction without Regret}
+
 To enable such flexibility between different semantics, we combine generic 
@@
 \item Monadic abstractions, which encapsulate the following two abstractions.
-  The interpreter itself is written in monadic form
-  \cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95,
-  DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}, by plugging
-  different monads, the underlying semantics of the interpreter is changed.
+  The interpreter itself is written in monadic form, and by plugging different monads, 
+  the underlying semantics of the interpreter is changed.
   \todo{mention that direct-style also works?}
+  %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
 \item Env-store-value abstractions, which control whether the interpreter
   behaves as a concrete interpreter or an abstract interpreter.
-  The environments, stores and values declared in the generic interpreter are
-  abstract types; we also define several abstract methods that manipulate
-  environments or stores.
+  The environments, stores and values are initially declared as abstract types in the generic interpreter.
   As shown by the AAM approach \cite{DBLP:journals/jfp/HornM12,
@@
 \end{itemize}
-
-\paragraph{Abstraction without Regret}
 
@@
 
-\paragraph{The paper is organized as follows:}
+\paragraph{Contributions} The contribution of the paper is twofold:
+1) Intellectually, we present a unifying construction that naturally extends the first 
+   Futamura projection to abstract interpreters. 
+2) Pragmatically, staging abstract interpreters is an approach to improve performance 
+   while untouch the soundness. We study the applicability and evaluate the performance 
+   of this approach on various analyses.
+
+The paper is organized as follows:
 
 \begin{itemize}
-  \item We progressively present a unifying construction that naturally extends
-    the first Futamura projection to abstract interpreters. Starting from a generic
-    interface of the interpreter for a small functional language
+  \item Starting from a generic interface of the interpreter for a functional language
     (Section~\ref{generic_if}), we first show the instantiation of concrete
@@
     values (Section~\ref{unstaged_abs}). Finally, we show the combination of the two
-    transformations, dubbed \textit{staged abstract interpreters}, can be easily
+    transformations, dubbed \textit{staged abstract interpreters}, can be readily
     derived (Section~\ref{sai}).
@@
     control-flow analysis: context-sensitive pushdown control-flow analysis and
-    \todo{precise-store?} (Section~\ref{sai}).
+    \todo{store-widened?} (Section~\ref{sai}).
 

### 2019-03-23 paper/20_bg.tex
revise

--- 2019-02-20 paper/20_bg.tex
@@
 In this section, we first describe the abstract syntax of the language we will
-analyze, then present the generic interface shared between the four different
-interpreters. We first instantiate the interpreter to the concrete one; then we
-proceed in two directions: by staging it we obtain a staged concrete interpreter
-(Section~\ref{stagedinterp}), by abstracting it we obtain an unstaged abstract
-interpreter (Section~\ref{unstaged_abs});
+analyze, then present the generic interpreter shared among the four different
+semantics, after which, we instantiate the interpreter to the concrete one. Then we
+will proceed in two directions: by staging it we obtain a staged concrete interpreter
+(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
+interpreter (Section~\ref{unstaged_abs}).
+
+We choose to use Scala and monad to demonstrate the idea, but the approach
+is not restricted to our choice. One can use imperative style and other MSP languages 
+(such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} 
+or Template Haskell \cite{Sheard:2002:TMH:636517.636528} \todo{cite the recent one})
+to construct such staged abstract interpreters.
 
 \input{21_bg_interp}
-\input{23_bg_stage}
-\input{22_bg_abs}
+\input{22_bg_stage}
+\input{23_bg_abs}

### 2019-03-23 paper/21_bg_interp.tex
revise

@@
 
-Monads provide an elegant and extensible way to build
-interpreters\cite{DBLP:conf/popl/LiangHJ95, Steele:1994:BIC:174675.178068,
-Wadler:1992:EFP:143165.143169}, and recently the approach has been extended to
-abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979,
-DBLP:journals/pacmpl/DaraisLNH17}. In section, with the multi-stage programming
-and abstract interpretation in mind, we use moandic operations to construct an
-generic interface in the style of big-step definitional interpreter.
+Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
+takes a monad as argument and produces another monad. By using monad transformers, 
+we can combine multiple monads into a single one.
+Constructing extensible interpreters using monad transformers was first proposed by \citet{DBLP:conf/popl/LiangHJ95},
+and later has been applied to abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17}.
+In section, with the multi-stage programming and monad transformers in mind
+while leave them as abstract type members, we present the generic interface of
+a big-step definitional interpreter.
+
+\todo{}
 

### 2019-03-23 paper/20_prelim.tex
reorganize paper structure

--- 2019-03-23 paper/21_bg_interp.tex
@@
+\newcommand{\TLang}{$L_\lambda$}
+
+\section{Preliminaries}
+
+In this section, we first describe the abstract syntax of the language we will
+analyze, then present the generic interpreter shared among the four different
+semantics, after which, we instantiate the interpreter to the concrete one. Then we
+will proceed in two directions: by staging it we obtain a staged concrete interpreter
+(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
+interpreter (Section~\ref{unstaged_abs}).
+
+We choose to use Scala and monad to demonstrate the idea, but the approach
+is not restricted to our choice. One can use imperative style and other MSP languages 
+(such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} 
+or Template Haskell \cite{Sheard:2002:TMH:636517.636528} \todo{cite the recent one})
+to construct such staged abstract interpreters.
+
 \subsection{Abstract Syntax} \label{bg_lang}

### 2019-03-23 paper/40_absinterp.tex
reorganize paper structure

--- 2018-11-17 paper/22_bg_abs.tex
@@
 
-Based on the generic interface we presented in Section~\ref{bg_lang}, 
-we now describe the instantiation of an abstract interpreter.
-To keep the presentation simple, we use simple abstract domains, and just establish a 
-context-/path-/flow-insensitive, store-widened analysis in this section --- indeed, it is coarse
-but simple enough to setup a foundation for the staged abstract interpreter we will present later.
-In Section~\ref{cfa}, we will see how to regain context-/path-/flow-sensitivity 
-for pushdown control-flow analysis; in Section~\ref{cases_imp}, we will see how to instantiate 
-an interval abstract domain for numerical analysis.
+Based on the generic interface we presented in Section~\ref{bg_lang}, we now
+describe the instantiation of an abstract interpreter. To keep the presentation
+simple, we use simple abstract domains, and just establish a
+context-/path-/flow-insensitive, store-widened analysis in this section ---
+indeed, it is coarse but simple enough to setup a foundation for the staged
+abstract interpreter we will present later. In Section~\ref{cfa}, we will see
+how to regain context-/path-/flow-sensitivity \todo{path/flow} for pushdown
+control-flow analysis; in Section~\ref{cases_imp}, we will see how to
+instantiate an interval abstract domain for numerical analysis.
 
-\paragraph{Abstract Instantiation.}
+\paragraph{Abstract Instantiation}
+
 Our instantiation roughly follows the Abstracting Abstract Machines methodology
-\cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12} that transforms the concrete semantics 
-to abstract semantics. The store now maps addresses to sets of abstract values, meaning that an address points
-to all possible values that may occur at runtime, where the abstract value is either a closure or 
-a single abstract number \texttt{NumV}, which stands for the top element of the number lattice. 
-The address space is also constrained to be finite, in the monovariant setting, we simply use variable names 
-for address.
+\cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12} that transforms the
+concrete semantics to abstract semantics. The store now maps addresses to sets
+of abstract values, meaning that an address points to all possible values that
+may occur at runtime, where the abstract value is either a closure or a single
+abstract number \texttt{NumV}, which stands for the top element of the number
+lattice. The address space is also constrained to be finite, in the monovariant
+setting, we simply use variable names for address.
 
@@
 \subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
-To effectively reuse the code between unstaged and staged variants, we also make the lattice structure
-stage polymorphic. A lattice type class parameterizes over an element type \texttt{E} and defines
-five operations for \texttt{R[E]}, such as meet, join, and ordering relation; again, \texttt{R} 
-is the higher-kinded type indicating the binding-time.
+
+To effectively reuse the code between unstaged and staged variants, we also make
+the lattice structure stage polymorphic. A lattice type class parameterizes over
+an element type \texttt{E} and defines five operations for \texttt{R[E]}, such
+as meet, join, and ordering relation; again, \texttt{R} is the higher-kinded
+type indicating the binding-time.
+
 \begin{lstlisting}
@@
 
-For example, here we use powersets as the abstract domain, and the unstaged lattice 
-of powersets can be easily implemented as follows:
+For example, here we use powersets as the abstract domain, and the unstaged
+lattice of powersets can be easily implemented as follows:
 
@@
 \subsection{Abstract Semantics}
-The operations @get@ and @put@ on abstract stores are slightly changed according to
-the abstract semantics: @get@ uses the bottom element in the @Value@ lattice if the 
-queried address does not exist; @put@ performs the update by joining with the existing values.
-function @alloc@ simply invokes the @Addr@ constructor.
-\texttt{close} and \texttt{num} lift syntactic literals to our abstract domain, i.e., 
-a singleton set that contains only one \texttt{CloV} or \texttt{NumV} object.
+
+The operations @get@ and @put@ on abstract stores are slightly changed according
+to the abstract semantics: @get@ uses the bottom element in the @Value@ lattice
+if the queried address does not exist; @put@ performs the update by joining with
+the existing values. function @alloc@ simply invokes the @Addr@ constructor.
+\texttt{close} and \texttt{num} lift syntactic literals to our abstract domain,
+i.e., a singleton set that contains only one \texttt{CloV} or \texttt{NumV}
+object.
 
@@
 
-For branching, @branch0@ joins the answers from two branches since we are doing a path-insensitive analysis. 
-For arithmetics, @prim_eval@ returns the top abstract number element.
-It is worth noting that @apply_closure@ is where we handle the non-determinism of applications.
-The non-determinism happens because the abstract store maps addresses to sets of abstract values;
-when dereferencing an address from the store, we need to explore all possible values to achieve a sound 
-analysis. In the case of @apply_closure@, the first argument @fs@ may contains multiple target closures,
-so we use a @for@ comprehension over all closures, apply them respectively and form a set result values @vs@;
-meanwhile, every time when evaluating the body expression @e@ we use the latest store $\sigma$@_*@
-which is updated iteratively during the loop. Thus this is a store-widened analysis.
+For branching, @branch0@ joins the answers from two branches since we are doing
+a path-insensitive analysis. For arithmetics, @prim_eval@ returns the top
+abstract number element. It is worth noting that @apply_closure@ is where we
+handle the non-determinism of applications. The non-determinism happens because
+the abstract store maps addresses to sets of abstract values; when dereferencing
+an address from the store, we need to explore all possible values to achieve a
+sound analysis. In the case of @apply_closure@, the first argument @fs@ may
+contain multiple target closures, so we use a @for@ comprehension over all
+closures, apply them respectively and form a set result values @vs@; meanwhile,
+every time when evaluating the body expression @e@ we use the latest store
+$\sigma$@_*@ which is updated iteratively during the loop. Thus this is a
+store-widened analysis.
 
 \subsection{Fixpoint Iteration}
-We described the abstract semantics modularly in the last section, however, the abstract 
-interpreter may not terminate for some input program. In this section, we use a memoization
-technique to ensure its termination. This technique is also widely used under the name of co-inductive caching algorithm
-\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or truncated depth-first evaluation 
-\cite{Rosendahl:AbsIntPL}.
 
-The idea is to set up two caches called @in@ and @out@, which are both mapping from the arguments of the interpreter 
-(@Config@) to the result values of the interpreter (@Ans@). The @in@ cache contains what we already have 
-computed from the last
-iteration, the @out@ cache is what we will have computed after this iteration joined with the previous one.
-When starting a new iteration, @in@ is set to be the result of the last iteration, i.e., @out@; @out@ is set
-to be empty. In the iteration, we first check whether @out@ contains what we want, if yes then it is returned;
-otherwise, we retrieve what we have from @in@, compute the result for this iteration, 
-and put the joined result back into @out@.
-Note that we also instrument the recursive call by putting @cached_ev@ into @evev@.
-After one iteration, if @in == out@, then there is no more information to be gained, thus the iteration should 
-end and we have reached the fixed point.
+We described the abstract semantics modularly in the last section, however, the
+abstract interpreter may not terminate for some input program. In this section,
+we use a memoization technique to ensure its termination. This technique is also
+widely used under the name of co-inductive caching algorithm
+\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or
+truncated depth-first evaluation \cite{Rosendahl:AbsIntPL}.
+
+The idea is to set up two caches called @in@ and @out@, which are both mapping
+from the arguments of the interpreter (@Config@) to the result values of the
+interpreter (@Ans@). The @in@ cache contains what we already have computed from
+the last iteration, the @out@ cache is what we will have computed after this
+iteration joined with the previous one. When starting a new iteration, @in@ is
+set to be the result of the last iteration, i.e., @out@; @out@ is set to be
+empty. In the iteration, we first check whether @out@ contains what we want, if
+yes then it is returned; otherwise, we retrieve what we have from @in@, compute
+the result for this iteration, and put the joined result back into @out@. Note
+that we also instrument the recursive call by putting @cached_ev@ into @evev@.
+After one iteration, if @in == out@, then there is no more information to be
+gained, thus the iteration should end and we have reached the fixed point.
 
@@
 
-Finally, we override the definition of @eval_top@ by instantiating @CacheFix@ with @eval@ and starting 
-the first iteration.
+Finally, we override the definition of @eval_top@ by instantiating @CacheFix@
+with @eval@ and starting the first iteration.

### 2019-03-23 paper/70_eval.tex
reorganize paper structure

--- 2019-02-04 paper/60_eval.tex
@@
 \begin{itemize}
-    \item \textbf{fib}: Calculates the $n$-th fibonacci number.
+  %\item \textbf{fib}: Calculates the $n$-th fibonacci number.
   \item \textbf{rsa}: The RSA public key encryption algorithm.
@@
 
+\todo{remove fib; report LOC; mean}
+
 \begin{figure}[h]
@@
     Program            & unstaged   & staged     & $\Delta$ of median \\ \midrule
-    fib                & 1.020      & 2.046      & -50.14\%          \\
+  % fib                & 1.020      & 2.046      & -50.14\%          \\
     rsa                & 35.928     & 145.306    & -75.27\%          \\

### 2019-03-23 paper/20_prelim.tex
reference

@@
 
-\subsection{Generic Interface} \label{generic_if}
+\subsection{Generic Interpreter} \label{generic_if}
 
@@
 Constructing extensible interpreters using monad transformers was first proposed by \citet{DBLP:conf/popl/LiangHJ95},
-and later has been applied to abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17}.
+and later has been applied to abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
 In section, with the multi-stage programming and monad transformers in mind

### 2019-03-23 paper/20_prelim.tex
section 2

@@
 
-\todo{}
+\paragraph{Basic type definitions}
+
+\begin{lstlisting}
+trait Semantics {
+  type Ident = String
+  type Addr
+  type Value
+  type Env = Map[Ident, Addr]
+  type Store = Map[Addr, Value]
+  ...
+}
+\end{lstlisting}
+
+The definitions in the rest of this section are enclosed in trait @Semantics@.
+
+\paragraph{Binding-time abstraction} As mentioned before, the binding-time is declared
+as a higher-kinded type @R[_]@.
+
+\begin{lstlisting}
+  type R[_]
+\end{lstlisting}
+
+\paragraph{Moandic operations}
+
+\begin{lstlisting}
+  type MonadOps[R[_], M[_], A] = {
+    def map[B](f: R[A] => R[B]): M[B]
+    def flatMap[B](f: R[A] => M[B]): M[B]
+  }
+  type AnsM[T] <: MonadOps[R, AnsM, T]
+  type Ans = AnsM[Value]
+
+  // Environment operations
+  def ask_env: AnsM[Env]
+  def ext_env(ans: Ans)(xα: (String, R[Addr])): Ans
+
+  // Store operations
+  def get_store: AnsM[Store]
+  def put_store(σ: R[Store]): AnsM[Unit]
+  def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
+\end{lstlisting}
+
+\paragraph{Primitive operations}
+
+\begin{lstlisting}
+  // allocate addresses
+  def alloc(x: String): AnsM[Addr]
+  def alloc(σ: R[Store], x: String): R[Addr]
+
+  // Primitive operations
+  def num(i: Int): Ans
+  def get(σ: R[Store], ρ: R[Env], x: String): R[Value]
+  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
+  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
+  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
+  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
+\end{lstlisting}
+
+\paragraph{The interpreter}
+
+\begin{lstlisting}
+  def eval(ev: Expr => Ans)(e: Expr): Ans = e match {
+    case Lit(i) => num(i)
+    case Var(x) => for {
+      ρ <- ask_env
+      σ <- get_store
+    } yield get(σ, ρ, x)
+    case Lam(x, e) => for {
+      ρ <- ask_env
+    } yield close(ev)(Lam(x, e), ρ)
+    case App(e1, e2) => for {
+      v1 <- ev(e1)
+      v2 <- ev(e2)
+      rt <- ap_clo(ev)(v1, v2)
+    } yield rt
+    case Let(x, rhs, e) => for {
+      v <- ev(rhs)
+      α <- alloc(x)
+      _ <- set_store(α → v)
+      rt <- ext_env(ev(e))(x → α)
+    } yield rt
+    case If0(e1, e2, e3) => for {
+      cnd <- ev(e1)
+      rt <- br0(cnd, ev(e2), ev(e3))
+    } yield rt
+    case Aop(op, e1, e2) => for {
+      v1 <- ev(e1)
+      v2 <- ev(e2)
+    } yield arith(op, v1, v2)
+    case Rec(x, rhs, e) => for {
+      α <- alloc(x)
+      v <- ext_env(ev(rhs))(x → α)
+      _ <- set_store(α → v)
+      rt <- ext_env(ev(e))(x → α)
+    } yield rt
+  }
+\end{lstlisting}
+
+\begin{lstlisting}
+  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans
+  type Result;    def run(e: Expr): Result
+\end{lstlisting}
 

### 2019-03-23 virtmacro-code/src/main/scala/sai/MainGeneric.scala

@@ -18,40 +18,40 @@ import sai.lattices.Lattices._
 import PCFLang._
 
 trait Semantics {
-  type R[_]
+  // Basic type definitions
   type Ident = String
   type Addr
   type Value
   type Env = Map[Ident, Addr]
   type Store = Map[Addr, Value]
 
-  implicit def mValue: Manifest[Value]
+  // Binding-time polymorphism
+  type R[_]
 
+  // Monadic interface
   type MonadOps[R[_], M[_], A] = {
     def map[B: Manifest](f: R[A] => R[B]): M[B]
     def flatMap[B: Manifest](f: R[A] => M[B]): M[B]
   }
   type AnsM[T] <: MonadOps[R, AnsM, T]
-
   type Ans = AnsM[Value]
-  type Result
-
-  type EvalFun = Expr => Ans
 
   // Environment operations
   def ask_env: AnsM[Env]
   def ext_env(ans: Ans)(xα: (String, R[Addr])): Ans
 
+  // Store operations
+  def get_store: AnsM[Store]
+  def put_store(σ: R[Store]): AnsM[Unit]
+  def set_store(αv: (R[Addr], R[Value])): AnsM[Unit]
+
   // allocate addresses
   def alloc(x: String): AnsM[Addr]
   def alloc(σ: R[Store], x: String): R[Addr]
 
-  // Store operations
-  def get_store: AnsM[Store]
-  def put_store(σ: R[Store]): AnsM[Unit]
-  def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
-
   // Primitive operations
+  type EvalFun = Expr => Ans
+
   def num(i: Int): Ans
   def get(σ: R[Store], ρ: R[Env], x: String): R[Value]
   def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
@@ -59,9 +59,9 @@ trait Semantics {
   def close(ev: EvalFun)(λ: Lam, ρ: R[Env]): R[Value]
   def ap_clo(ev: EvalFun)(fun: R[Value], arg: R[Value]): Ans
 
-  val ρ0: R[Env]
-  val σ0: R[Store]
+  // Fixpoint wrapper and top-level interface
   def fix(ev: EvalFun => EvalFun): EvalFun
+  type Result
   def run(e: Expr): Result
 
[…]

### 2019-03-24 paper/20_prelim.tex
sec 2

--- 2019-03-23 paper/20_prelim.tex
@@
   case class Rec(x: String, rhs: Expr, e: Expr) extends Expr  // recursions
-  case class AOp(op: String, e1: Expr, e2: Expr) extends Expr // primitives
+  case class Aop(op: String, e1: Expr, e2: Expr) extends Expr // arithmetics
 \end{lstlisting}
@@
 
-\subsection{Monads in Scala}
+\subsection{Monads in Scala} \label{monadscala}
 
-A monad is a type constructor @M[_]: * -> *@ with two operations, often called @return@ and @bind@.
-Informally, @return@ wraps a value into the monad @M@, and @bind@ unwraps the monadic value and 
-transforms it to a new monadic value.
-Pragmatically in Scala, we define a monad using trait @Monad@ (Figure \ref{fig:monad}),
-where it declares the @pure@ \footnote{We elect to use \texttt{pure} as the name, 
-since \texttt{return} is a keyword in Scala and \texttt{unit} is a built-in function in LMS.}
-and @flatMap@ operation. The trait itself takes the monad type @M[_]@ as
-argument, which is a higher-kinded type that takes a type and returns a type. 
-The method @pure@ promotes values of type @A@ to values of type @M[A]@.
-The monadic @bind@ operation is
-usually called @flatMap@ in Scala, which takes a monad-encapsulated value of
-type @M[A]@, a function of @A => M[B]@ and returns values of type @M[B]@.
+A monad is a type constructor @M[_]: * -> *@ with two operations, often called
+@return@ and @bind@. Informally, @return@ wraps a value into the monad @M@, and
+@bind@ unwraps the monadic value and transforms it to a new monadic value.
+Pragmatically in Scala, we define a monad type class using trait @Monad@ (Figure
+\ref{fig:monad}), where it declares the @pure@ \footnote{We elect to use
+\texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
+\texttt{unit} is a built-in function in LMS.} and @flatMap@ operation. The trait
+itself takes the monad type @M[_]@ as argument, which is a higher-kinded type
+that takes a type and returns a type. The method @pure@ promotes values of type
+@A@ to values of type @M[A]@. The monadic @bind@ operation is usually called
+@flatMap@ in Scala, which takes a monad-encapsulated value of type @M[A]@, a
+function of @A => M[B]@ and returns values of type @M[B]@.
 
@@
     def pure[A](a: A): M[A]                            
-    def flatMap[A,B](fa: M[A])(f: A => M[B]): M[B]     
+    def flatMap[A,B](ma: M[A])(f: A => M[B]): M[B]     
   }                                                    
@@
 Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
-takes a monad as argument and produces another monad. By using monad transformers, 
-we can combine multiple monads into a single one.
-Constructing extensible interpreters using monad transformers was first proposed by \citet{DBLP:conf/popl/LiangHJ95},
-and later has been applied to abstract interpreters \cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
-In section, with the multi-stage programming and monad transformers in mind
-while leave them as abstract type members, we present the generic interface of
-a big-step definitional interpreter.
+takes a monad as argument and produces another monad. By using monad
+transformers, we can combine multiple monads into a single one. Constructing
+extensible interpreters using monad transformers was first proposed by
+\citet{DBLP:conf/popl/LiangHJ95}, and later has been applied to abstract
+interpreters \cite{Sergey:2013:MAI:2491956.2491979,
+DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}. In section,
+with the multi-stage programming and monad transformers in mind while leave them
+as abstract type members, we present the generic interface of a big-step
+definitional interpreter.
 
-\todo{}
-
-Before showing the actual implementation of the concrete semantics, let us first
-consider a generic interface for the interpreters, which is semantics-agnostic
-and can be "interpreted" by either concrete or abstract, unstaged or staged
-semantics of our choice. The trait \texttt{Semantics} first declares several
-abstract type members, such as \texttt{Addr}, \texttt{Value}, \texttt{Env},
-\texttt{Store} and returned type \texttt{Ans}. It also declares initial values
-for environments and stores, and few abstract methods that manipulate
-environments and stores such as @put@ and @get@.
+\paragraph{Basic type definitions} We start with some basic type definitions.
+The identifiers in the program are represented by strings. The two default
+components in the interpreter are environments @Env@ and stores @Store@, i.e.,
+mappings from identifiers to addressess and mappings from addresses to values,
+respectively. But the @Addr@ and @Value@ are just declared as abstracrt types.
 
@@
   trait Semantics {
-  type R[+_]
-  type Ident = String;  type Addr;  type Value;  type Env
-  type Store;  type Ans = (R[Value], R[Store])
-  def get($\rho$: R[Env], x: Ident): R[Addr]
-  def put($\rho$: R[Env], x: Ident, a: R[Addr]): R[Env]
-  def get($\sigma$: R[Store], a: R[Addr]): R[Value]
-  def put($\sigma$: R[Store], a: R[Addr], v: R[Value]): R[Store]
-  def alloc($\sigma$: R[Store], x: Ident): R[Addr]
-  val $\rho$0: R[Env]; val $\sigma$0: R[Store]
-  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: R[Env]): R[Value]
-  def num(i: Lit): R[Value]
-  def apply_closure(ev: EvalFun)
-    (f: R[Value], arg: R[Value], $\sigma$: R[Store]): Ans
-  def branch0(cnd: R[Value], thn: => Ans, els: => Ans): Ans
-  def prim_eval(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
-}  // to be continued
-\end{lstlisting}
-
-The actual semantics and operations of the interpreter are left to be
-implemented later, these include operations such as lifting literal terms to
-values (\texttt{close}, \texttt{num}), applying a closure
-(\texttt{apply\_closure}), branching (\texttt{branch0}), and arithmetics
-(\texttt{prim\_eval}). We can choose to implement them in a way that is concrete
-or abstract, unstaged or staged. But, just based on these abstract operations,
-an expressive generic interpreter can be defined:
-
-\begin{lstlisting}
-trait Semantics {  // continued
-  type EvalFun = (Expr, R[Env], R[Store]) => Ans
-  def eval(ev: EvalFun)(e: Expr, $\rho$: R[Env], $\sigma$: R[Store]): Ans = 
-    e match {
-      case Lit(i) => (num(Lit(i)), $\sigma$)
-      case Var(x) => (get($\sigma$, get($\rho$, x)), $\sigma$)
-      case Lam(x, e) => (close(ev)(Lam(x, e), $\rho$), $\sigma$)
-      case App(e1, e2) =>
-        val (e1v, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
-        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
-        apply_closure(ev)(e1v, e2v, e2$\sigma$)
-      case Rec(x, f, body) =>
-        val $\alpha$ = alloc($\sigma$, x)
-        val $\rho$_* = put($\rho$, x, $\alpha$)
-        val (fv, f$\sigma$) = ev(f, $\rho$_*, $\sigma$)
-        val $\sigma$_* = put(f$\sigma$, $\alpha$, fv)
-        ev(body, $\rho$_*, $\sigma$_*)
-      case If0(cnd, thn, els) =>
-        val (cndv, cnd$\sigma$) = ev(cnd, $\rho$, $\sigma$)
-        branch0(cndv, ev(thn, $\rho$, cnd$\sigma$), ev(els, $\rho$, cnd$\sigma$))
-      case AOp(op, e1, e2) =>
-        val (e1v, e1$\sigma$) = ev(e1, $\rho$, $\sigma$)
-        val (e2v, e2$\sigma$) = ev(e2, $\rho$, e1$\sigma$)
-        (prim_eval(op, e1v, e2v), e2$\sigma$)
-    }
-  def fix(ev: EvalFun => EvalFun): EvalFun = 
-    (e, $\rho$, $\sigma$) => ev(fix(ev))(e, $\rho$, $\sigma$)
-  def eval_top(e: Expr): Ans = eval_top(e, $\rho$0, $\sigma$0)
-  def eval_top(e: Expr, $\rho$: R[Env], $\sigma$: R[Store]): Ans = 
-    fix(eval)(e, $\rho$, $\sigma$)
+    type Ident = String; type Addr; type Value
+    type Env = Map[Ident, Addr]; type Store = Map[Addr, Value]
+    type R[_] // Binding-time as a higher-kinded type
+    ... // The definitions in the rest of this section are enclosed in trait Semantics.
   }
@@
 
-The implementation of the skeleton interpreter traverses the abstract syntax
-tree and should be straightforward to understand. It is worth noting that the
-interpreter is written in open-recursive style -- it can not refer to itself
-directly, instead, \texttt{eval} takes an additional parameter \texttt{ev}
-referring to itself; this allows us to instrument the interpreter from the
-outside. Accordingly, the function \texttt{close} for lifting lambda term to
-closure values is also written in this style. We define a default implementation
-\texttt{fix} to close the recursion by recursively using the result of
-\texttt{fix(ev)} for \texttt{ev}. Finally, the top-level evaluation function
-\texttt{eval\_top} is defined.
+\paragraph{Binding-time abstraction} As mentioned before, the binding-time is
+declared as a higher-kinded type @R[_]@. If we instantiate @R@ as an identity
+type (i.e., @type R[T] = T@), then the generic interpreter will execute the program.
+In Section \ref{stagedinterp}, we will instantiate @R@ using LMS's built-in
+next-stage type annotation @Rep@, which makes the interpreter act as a compiler.
 
-\paragraph{Binding-time Abstraction}
-%\paragraph{Stage Polymorphism}
-% Binding-time abstraction via abstract types
+\paragraph{Moandic operations} We define the return type of the interpreter as
+@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
+As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
+syntax, certain constraints have to be posed onto the type @AnsM@. Here, we use a
+structural type @MonadOps@ to require @AnsM@ to at least implement @map@ and
+@flatMap@. It is worth noting that @MonadOps@ takes another type parameter
+@R[_]@ as binding-time; accordingly the outer @R@ defined in the trait is passed
+to @MonadOps@. Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
+encapsulated by the monad, but not the monad type @M@ itself. When acting as
+compilers, we will also replace the monads to the ones that work on staged
+values.
 
-Readers may notice that an abstract higher-kinded type \texttt{R[+T]} is defined
-in the trait, i.e. a type-level function which will be used to specify the
-binding-time, i.e., staged or not; the annotation \texttt{+} indicates that the
-type parameter \texttt{T} is covariant. The LMS framework we use here allows us
-to specify binding-times purely based on types. All occurrences of type
-\texttt{Env}, \texttt{Store}, etc. are wrapped with \texttt{R} essentially
-because they may potentially be dynamic when staging is introduced. Obviously,
-the unstaged higher-kinded type then can be simply instantiated as an identity
-type function: \texttt{type R[+T] = T}. In Section~\ref{stagedinterp},
-\texttt{R} will be instantiated as a staging annotation.
+We also defines several methods to obtain and modify the environment and store.
+These methods either returns monadic values with environment or store inside, or
+simply a @Unit@ value for effects. For example, @ext_env@ takes a value @ans@ of
+type @Ans@ and a pair of binding that maps @x@ to $\alpha$, and installs this
+binding to the environment when evaluating @ans@.
 
-\todo{Finally tagless, poly embedding}
+\vspace{-1em}
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+    \begin{lstlisting}
+  type MonadOps[R[_], M[_], A] = {
+    def map[B](f: R[A] => R[B]): M[B]
+    def flatMap[B](f: R[A] => M[B]): M[B]
+  }
 
-\subsection{Concrete Instantiation} \label{unstaged_conc}
+  type AnsM[T] <: MonadOps[R, AnsM, T]
+  type Ans = AnsM[Value]
+    \end{lstlisting}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.55\textwidth}
+    \begin{lstlisting}
+// Environment operations
+def ask_env: AnsM[Env]
+def ext_env(ans: Ans)(xα: (Ident, R[Addr])): Ans
+// Store operations
+def get_store: AnsM[Store]
+def put_store(σ: R[Store]): AnsM[Unit]
+def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
+    \end{lstlisting}
+  \end{subfigure}
+\end{figure}
+\vspace{-1em}
+
+\paragraph{Primitive operations} Next we define several primitive operations.
+Two versions of @alloc@ are declared. The first one takes a store and an
+identifier, and produces a fresh address of non-monadic type @R[Addr]@. Since
+the freshness of the address may depend on the store, which might be a
+next-stage value as indicated by its type, the type of addresses is also wrapped
+by @R[_]@. The other one would simply wrap the address with our monadic type
+@AnsM[_]@.
+\begin{lstlisting}
+  def alloc(σ: R[Store], x: Ident): R[Addr];  def alloc(x: Ident): AnsM[Addr]
+\end{lstlisting}
+
+Other primitive operations provide basic functionality for the interpreter.
+The method @num@ and @close@ deal with primitive values, which lift literal
+terms (e.g., lambdas) to our value representation (e.g., closures).
+The method @get@ simply retrieves the value mapped from the identifier @x@ in
+the environment and store. Conditionals and arithmetics is handled by @br0@
+and @arith@, respectively. The methods @ap_clo@ takes a function value and an
+arugment value and then does the application.
+
+%TODO: close/ap_clo takes an ev
+
+\begin{lstlisting}
+  def num(i: Int): Ans
+  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
+  def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
+  def br0(test: R[Value], thn: => Ans, els: => Ans): Ans
+  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
+  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
+\end{lstlisting}
+
+\paragraph{The interpreter} Now we can define the semantics-agnostic interpreter
+in monadic form. The essential idea is to traverse the abstract syntax tree
+while maintaining the effects such as reader and state. 
+It is worth noting that the interpreter is written in open-recursive style -- it
+can not refer to itself directly, instead, @eval@ takes an additional parameter
+@ev@ of type @Expr => Ans@ to refer to itself. Accordingly, the method @close@
+that lifts lambda terms to closures and @ap_clo@ that applies functions also
+takes an extra @ev@.
+
+\begin{lstlisting}
+  def eval(ev: Expr => Ans)(e: Expr): Ans = e match {
+    case Lit(i) => num(i)                      case Let(x, rhs, e) => for {
+    case Var(x) => for {                         v <- ev(rhs)
+      ρ <- ask_env                               α <- alloc(x)
+      σ <- get_store                             _ <- set_store(α → v)
+     } yield get(σ, ρ, x)                        rt <- ext_env(ev(e))(x → α)
+    case Lam(x, e) => for {                    } yield rt
+      ρ <- ask_env                             case Aop(op, e1, e2) => for {
+    } yield close(ev)(Lam(x, e), ρ)              v1 <- ev(e1)                                               
+    case App(e1, e2) => for {                    v2 <- ev(e2)
+      v1 <- ev(e1)                             } yield arith(op, v1, v2)
+      v2 <- ev(e2)                             case Rec(x, rhs, e) => for {
+      rt <- ap_clo(ev)(v1, v2)                   α <- alloc(x)
+    } yield rt                                   v <- ext_env(ev(rhs))(x → α)
+    case If0(e1, e2, e3) => for {                _ <- set_store(α → v)
+      cnd <- ev(e1)                              rt <- ext_env(ev(e))(x → α)
+      rt <- br0(cnd, ev(e2), ev(e3))           } yield rt                    
+    } yield rt
+  }
+\end{lstlisting}
+
+Since the interpreter is written in open-recursive style, we declare an abstract
+combinator @fix@ used to close the recursion. For concrete-interpretation instantiation, it
+works like the Y combinator; for abstract-interpretation instantiation, we will
+instrument the interpreter by defining a memorized version of @fix@ to ensure
+termination. Finally, a top-level wrapper @run@ is declared; the return type
+@Result@ depends on what kind of monad we will be using, thus is also an abstract type.
+
+\begin{lstlisting}
+  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans
+  type Result; def run(e: Expr): Result
+\end{lstlisting}
+
+%==========================================================================
+
+\subsection{A Concrete Interpreter} \label{unstaged_conc}
 

### 2019-03-24 paper/20_prelim.tex
misc

@@
 The abstract syntax we present in fact can be seen as a deep embedding of the
-language -- we use datatypes to represent programs. This design choice makes us
+language -- we use data-types to represent programs. This design choice makes us
 easily add different interpretations over the AST; with the inheritance and
@@
 components in the interpreter are environments @Env@ and stores @Store@, i.e.,
-mappings from identifiers to addressess and mappings from addresses to values,
-respectively. But the @Addr@ and @Value@ are just declared as abstracrt types.
+mappings from identifiers to addresses and mappings from addresses to values,
+respectively. But the @Addr@ and @Value@ are just declared as abstract types.
 
@@
 
-\paragraph{Moandic operations} We define the return type of the interpreter as
-@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
+\paragraph{Monadic operations} We define the return type of the interpreter as
+@Ans@, which is an arbitrary monad type @AnsM[_]@ wrapping the type @Value@. 
 As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
@@
 and @arith@, respectively. The methods @ap_clo@ takes a function value and an
-arugment value and then does the application.
+argument value and then does the application.
 

### 2019-03-25 paper/20_prelim.tex
section 2

@@
 
-\paragraph{Basic type definitions} We start with some basic type definitions.
-The identifiers in the program are represented by strings. The two default
-components in the interpreter are environments @Env@ and stores @Store@, i.e.,
-mappings from identifiers to addresses and mappings from addresses to values,
-respectively. But the @Addr@ and @Value@ are just declared as abstract types.
+\paragraph{Basic Types} We start with some basic type definitions used in the
+interpreter. The identifiers in the program are represented by strings. The two
+default components in the interpreter are environments @Env@ and stores @Store@,
+i.e., mappings from identifiers to addresses and mappings from addresses to
+values, respectively. The @Env@ models accessible variables, and @Store@ models
+the persistent heap through the program runtime. But the @Addr@ and @Value@ are
+just declared as abstract types.
 
@@
 
-\paragraph{Binding-time abstraction} As mentioned before, the binding-time is
-declared as a higher-kinded type @R[_]@. If we instantiate @R@ as an identity
+\paragraph{Binding-time Abstraction} As mentioned before, the binding-time is
+declared as a higher-kinded type @R[_]@. If we simply instantiate @R@ as an identity
 type (i.e., @type R[T] = T@), then the generic interpreter will execute the program.
@@
 
-\paragraph{Monadic operations} We define the return type of the interpreter as
+\paragraph{Monadic Operations} We define the return type of the interpreter as
 @Ans@, which is an arbitrary monad type @AnsM[_]@ wrapping the type @Value@. 
@@
 
-\paragraph{Primitive operations} Next we define several primitive operations.
+\paragraph{Primitive Operations} Next we define several primitive operations.
 Two versions of @alloc@ are declared. The first one takes a store and an
@@
 and @arith@, respectively. The methods @ap_clo@ takes a function value and an
-argument value and then does the application.
+argument value and then does the application. Note that the @Env@, @Store@, and
+@Value@ are all wrapped by @R[_]@ since they will be known as next-stage value
+when acting as compilers.
 
@@
 
-\paragraph{The interpreter} Now we can define the semantics-agnostic interpreter
+\paragraph{The Interpreter} Now we can define the semantics-agnostic interpreter
 in monadic form. The essential idea is to traverse the abstract syntax tree
@@
 
-Now we can instantiate a concrete interpreter. We first concretize the abstract
-types as follows. Since the interpreter is concrete, for the purpose that
-properly models the heap, the type \texttt{Int} is used for address space
-\texttt{Addr}. The environment and store are just ordinary \texttt{Map}s in
-Scala. A value can be either a tagged number \texttt{NumV} or a closure
-\texttt{CloV} that contains a lambda term and an environment.
+\paragraph{Concrete Components}
+Now we can instantiate the interpreter as a standard environment-store
+interpreter. Such interpreter can be obtained by refunctionalization of CESK
+machines \cite{Felleisen:1987:CAH:41625.41654, DBLP:conf/ppdp/AgerBDM03}.
+The two types we need to concretize are @Addr@ and @Value@, afterward, the @Env@
+and @Store@ is concretized automatically. To assure the freshness, we use @Int@
+for the address space @Addr@. A value can be either a tagged number
+\texttt{IntV}, or a closure \texttt{CloV} that contains a lambda term and an
+environment. The final return value of the interpreter is a pair of values and
+stores, where the @Value@ and @Store@ are obviously next-stage objects. We also
+provide a standard fixed-point combinator to close the open-recursive function @ev@.
 
 \begin{lstlisting}
-trait Concrete extends Semantics {
-  type Addr = Int;   sealed trait Value
-  case class CloV($\lambda$: Lam, $\rho$: Env) extends Value
-  case class NumV(i: Int) extends Value
-  type Env = Map[Ident, Addr];  type Store = Map[Addr, Value]
+  trait ConcreteComponents extends Semantics {
+    type Addr = Int
+    sealed trait Value
+    case class IntV(i: Int) extends Value
+    case class CloV(λ: Lam, e: Env) extends Value
+    type Result = (R[Value], R[Store])
+    def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans = e => ev(fix(ev))(e)
+  }
+\end{lstlisting}
+%\vspace{-1em}
+
+\paragraph{Unstaged Concrete Monads}
+% The data manipulated by the monad are staged value. But the monad is known at
+% current stage.
+The environment and store can be modeled by reader effect and state effect,
+i.e., a Reader monad and a State monad. We combine them using monad transformers
+@ReaderT@ and @StateT@.
+In other word, we instantiate the @AnsM@ moand as a stack of @ReaderT@ and @StateT@
+transformers, where the @ReaderT@ is parameterized by type @Env@, @StateT@ is
+parameterized by type @Store@, and the inner-most monad @IdM@ is just an identity
+monad.
+
+\begin{lstlisting}
+  trait ConcreteSemantics extends ConcreteComponents {
+    type R[T] = T
+    type AnsM[T] = ReaderT[StateT[IdM, Store, ?], Env, T]
+    ...
   }
@@
 
-We proceed to implement the components that define the semantics. As mentioned
-before, type \texttt{R} is concretized as an identity type function, thus all
-types are just ordinary types.
+Here we sketch the basic idea of unstaged @ReaderT@ and @StateT@. Readers may
+refer to \todo{cite, or appendix?} for more detail and implementation of monad
+transformers.
+A @ReaderT@ monad transformer encapsulates the computation @R => M[A]@, where
+@R@ is the reader type, @M[_]@ is a monad type. The @ReaderT@ monad returns the
+transformed value of type @M[A]@. Similarly, a @StateT@ monad encapsulates the
+computation @S => M[(A, S)]@, where @S@ is the state type, @M[_]@ is a monad
+type. The @StateT@ monad returns the transformed value of type @M[(A, S)]@,
+where the new state is accompanied.
+Note that the binding-time type @R@ is an identity type, thus the monads are
+operated on unstaged data. We can also see this from the signature of @flatMap@:
+the function @f@ takes an unstaged value of type @A@ and produces a monadic
+value.
 
 \begin{lstlisting}
-object ConcInterp extends Concrete {
-  type R[+T] = T
-  val $\rho$0 = Map[Ident,Addr]();  val $\sigma$0 = Map[Addr,Value]()
-  def get($\rho$: Env, x: Ident): Addr = $\rho$(x)
-  def put($\rho$: Env, x: Ident, a: Addr): Env = $\rho$ + (x -> a)
-  def get($\sigma$: Store, a: Addr): Value = $\sigma$(a)
-  def put($\sigma$: Store, a: Addr, v: Value): Store = $\sigma$ + (a -> v)
-  def alloc($\sigma$: Store, x: Ident): Addr = $\sigma$.size + 1
-  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Env): Value = CloV($\lambda$, $\rho$)
-  def num(i: Lit): Value = NumV(i.i)
-  def apply_closure(ev: EvalFun)
-    (f: Value, arg: Value, $\sigma$: Store): Ans = f match {
-      case CloV(Lam(x, e), $\rho$) =>
-        val $\alpha$ = alloc($\sigma$, x); val $\rho$_* = put($\rho$, x, $\alpha$)
-        val $\sigma$_* = put($\sigma$, $\alpha$, arg); ev(e, $\rho$_*, $\sigma$_*)
-    }
-  def branch0(cnd: Value, thn: => Ans, els: => Ans): Ans = 
-    cnd match { case NumV(i) => if (i == 0) thn else els }
-  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = 
-    op match {
-      case '+ => v1.asInstanceOf[NumV].i + v2.asInstanceOf[NumV].i
-      ...
-    }
+  case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {
+    def flatMap[B](f: A => ReaderT[M, R, B]): ReaderT[M, R, B] = ... }
+  case class StateT[M[_]: Monad, S, A](run: S => M[(A, S)]) {
+    def flatMap[B](f: A => StateT[M, S, B]): StateT[M, S, B] = ... }
+\end{lstlisting}
+
+Then, the monadic operations that manipulate the environment and store are
+simply constructing the proper monads and lifting it to the top-level of our
+moand stack. To modify the store, for example, we may have a @StateT@ monad that
+transforms the current store $\sigma$ to @σ + αv@, which results in
+@StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to @ReaderT@,
+i.e., @AnsM[Unit]@.
+\begin{lstlisting}
+  def set_store(αv: (Addr, Value)): AnsM[Unit] = liftM(StateTMonad.mod(σ => σ + αv))
+\end{lstlisting}
+
+\paragraph{Primitive Operations}
+Other primitive operations can be implemented straightforwardly, for example,
+@alloc@ simply takes the size of current store and produce the successor number.
+We elide most of them, but describe a little bit on how do we handle functions
+and applications as they will be very different when staging and abstract
+interpretation is involved. The method @close@ takes a lambda term and an
+environment, then create a closure value @CloV@.
+
+\begin{lstlisting}
+  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = CloV(λ, ρ)
+\end{lstlisting}
+
+To apply a function, the method @ap_clo@ takes a function value and an arugment
+value, extracts the lambda term and environment enclosed in the function value,
+and evaluates the body expression of that lambda under the new environment and store.
+
+\begin{lstlisting}
+  def ap_clo(ev: EvalFun)(f: Value, arg: Value): Ans = f match {
+    case CloV(Lam(x, e), ρ: Env) => for {
+      α <- alloc(x)
+      _ <- set_store(α → arg)
+      rt <- ext_env(ev(e))(x → α)
+    } yield rt
   }
@@
 
-Finally, \texttt{get} and \texttt{put} directly manipulate the map of
-environment and store. An allocation function is used when we need to update the
-store: in the concrete case, \texttt{alloc} always returns a fresh address of
-the store. \texttt{close} and \texttt{num} lift syntactic literals to values.
-\texttt{apply\_closure} takes the function value, argument value and the latest
-store, then extends the environment and store after allocating a fresh address,
-and continue evaluating the body \texttt{e} of the lambda term.
+With other primitive operations implementated, we can implement the top-level
+@run@ method, where $\rho_0$ and $\sigma_0$ are initial empty environment and
+store.
+
+\begin{lstlisting}
+  def run(e: Expr): Result = fix(eval)(e)(ρ$_0$)(σ$_0$)
+\end{lstlisting}

### 2019-03-25 virtmacro-code/src/main/scala/sai/MainGeneric.scala

@@ -29,6 +29,7 @@ trait Semantics {
   type R[_]
 
   // Monadic interface
+  // TODO: why not a trait?
   type MonadOps[R[_], M[_], A] = {
     def map[B: Manifest](f: R[A] => R[B]): M[B]
     def flatMap[B: Manifest](f: R[A] => M[B]): M[B]

### 2019-03-25 virtmacro-code/src/main/scala/sai/monads/Monads.scala

@@ -26,6 +26,11 @@ trait Monad[M[_]] extends RMonad[NoRep.NoRep, M]
 
 object Monad {
   def apply[M[_]](implicit m: Monad[M]): Monad[M] = m
+
+  trait RMonadOps[R[_], M[_], A] {
+    def flatMap[B: Manifest](f: R[A] => M[B]): M[B]
+    def map[B: Manifest](f: R[A] => R[B]): M[B]
+  }
 }
 
 trait MonadPlus[M[_]] {
@@ -109,6 +114,10 @@ object ReaderT {
 
   def liftM[G[_]: Monad, R, A](ga: G[A]): ReaderT[G, R, A] =
     ReaderT(r => ga)
+}
+
+trait SPReaderT[BT[_], M[_], R, A] extends Monad.RMonadOps[BT, SPReaderT[BT, M, R, ?], A] {
+  val run: BT[R] => M[A]
 }
 
 case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {

### 2019-03-25 virtmacro-code/src/main/scala/sai/monads/RepMonads.scala

@@ -108,7 +108,7 @@ trait RepMonads extends RepLattices { self: SAIDsl =>
       ReaderT(r => ga)
   }
 
-  case class ReaderT[M[_]: Monad, R: Manifest, A: Manifest](run: Rep[R] => M[A]) {
+  case class ReaderT[M[_]: Monad, R: Manifest, A: Manifest](run: Rep[R] => M[A]) /*extends SPReaderT[Rep, M, R, A, ReaderT]*/ {
     import ReaderT._
     def apply(r: Rep[R]): M[A] = run(r)
     def flatMap[B: Manifest](f: Rep[A] => ReaderT[M, R, B]): ReaderT[M, R, B] =

### 2019-03-25 paper/20_prelim.tex
tweak code

@@
 
-\paragraph{Concrete Components}
 Now we can instantiate the interpreter as a standard environment-store
@@
 machines \cite{Felleisen:1987:CAH:41625.41654, DBLP:conf/ppdp/AgerBDM03}.
+
+\paragraph{Concrete Components}
 The two types we need to concretize are @Addr@ and @Value@, afterward, the @Env@

### 2019-03-25 paper/60_cases.tex
tweak code

--- 2019-02-04 paper/50_cases.tex
@@
     \centering
-    \begin{lstlisting}
+    \begin{lstlisting}[
+      language=Scala,%
+      mathescape=true,%
+      %  columns=[c]fixed,%
+      %  basewidth={0.5em, 0.40em},%
+      aboveskip=2pt,%\smallskipamount,
+      belowskip=1pt,%\negsmallskipamount,
+      lineskip=-1pt,
+      basewidth={0.6em, 0.45em},%
+      %  backgroundcolor=\color{listingbg},
+      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
+      keywordstyle=\keywordstyle,
+      commentstyle=\commentstyle,
+      stringstyle=\stringstyle,
+    ]
 type CompAnalysis = Store => Store
@@
       val C = compCall(body)
-      ($\sigma$: Store) => 
-        C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
+      ($\sigma$: Store) => C($\sigma$.update(vars, args.map(primEval(_, $\sigma$))))
   }
@@
     \centering
-    \begin{lstlisting}
+    \begin{lstlisting}[
+      language=Scala,%
+      mathescape=true,%
+      %  columns=[c]fixed,%
+      %  basewidth={0.5em, 0.40em},%
+      aboveskip=2pt,%\smallskipamount,
+      belowskip=1pt,%\negsmallskipamount,
+      lineskip=-1pt,
+      basewidth={0.6em, 0.45em},%
+      %  backgroundcolor=\color{listingbg},
+      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
+      keywordstyle=\keywordstyle,
+      commentstyle=\commentstyle,
+      stringstyle=\stringstyle,
+    ]
 def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 

### 2019-03-25 paper/20_prelim.tex
section 3

@@
 
-\section{Preliminaries}
+\section{Preliminaries} \label{prelim}
 
@@
 Now we can instantiate the interpreter as a standard environment-store
-interpreter. Such interpreter can be obtained by refunctionalization of CESK
-machines \cite{Felleisen:1987:CAH:41625.41654, DBLP:conf/ppdp/AgerBDM03}.
+interpreter. Such interpreter also can be obtained by refunctionalization of
+CESK machines \cite{Felleisen:1987:CAH:41625.41654, DBLP:conf/ppdp/AgerBDM03}.
 
@@
 
-\paragraph{Unstaged Concrete Monads}
-% The data manipulated by the monad are staged value. But the monad is known at
-% current stage.
+\paragraph{Unstaged Monads}
 The environment and store can be modeled by reader effect and state effect,
@@
 In other word, we instantiate the @AnsM@ moand as a stack of @ReaderT@ and @StateT@
-transformers, where the @ReaderT@ is parameterized by type @Env@, @StateT@ is
-parameterized by type @Store@, and the inner-most monad @IdM@ is just an identity
-monad.
+transformers\footnote{The question mark syntax is a kind projector
+  \cite{kindprojector}, thus \texttt{StateT[IdM,Store,?]} is equivalent to \newline
+  \texttt{(\{type M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is
+parameterized by type @Env@, @StateT@ is parameterized by type @Store@, and the
+inner-most monad @IdM@ is just an identity monad.
 
@@
 value, extracts the lambda term and environment enclosed in the function value,
-and evaluates the body expression of that lambda under the new environment and store.
+and evaluates the body expression of that lambda term under the new environment and store.
 

### 2019-03-25 paper/30_stageconc.tex
section 3

--- 2019-02-04 paper/23_bg_stage.tex
@@
 
-\subsection{Multi-Stage Programming in LMS}
+In this section, based on the generic interpreter and concrete components we
+presented in Section \ref{prelim}, we show how to obtain a staged concrete
+interpreters by changing the monads and refactoring several primitive
+operations. We begin by briefly introducing the Lightweight Modular Staging
+framework in Scala.
+
+\subsection{Multi-Stage Programming with LMS}
 
@@
 multi-stage programming framework implemented as a Scala library that enables
-dynamic code generation in a type-safe manner. Different from the syntactic
-approach in MetaOCaml \cite{DBLP:conf/flops/Kiselyov14} that uses quotations,
-LMS distinguishes binding-time solely based on types. LMS provides a type
-constructor @Rep[T]@ where @T@ can be an arbitrary type, indicating an
-expression will be known at next stage. All operations acting on a @Rep[T]@
-expression will be residualized in the generated code. However, we need to
-provide intermediate representation and code generation support for type @T@ to
-LMS. Fortunately, the LMS framework already provides such support for primitive
-types and commonly used data structure such as arrays and maps; implementing
-such support for custom class is also straightforward. Let's go back and see how
-LMS can be used to specialize the power function we mentioned in Section 1.
+dynamic code generation in a type-safe manner. Different from the approach of
+MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} that
+uses syntactic quotation and escape, LMS distinguishes binding-time solely based
+on types. LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary
+type, indicating a value of @T@ will be known at next stage. All operations
+acting a @Rep[T]@ expression will be residualized as generated code.
+
+The LMS framework provides staging support for primitive data types such as
+@Int@ and @Double@, and commonly-used data structures such as arrays and maps.
+The idea is to construct intermediate representations (IR) for the next-stage
+program at current stage. Such conversion from expressions of type @Rep[_]@ to
+their IR is done by using implicit design pattern. As we will see later,
+implementing the staging support for custom classes is also straightforward.
+
+For example, let us go back and see how LMS can be used to specialize the
+@power@ function from Section ~\ref{intro}.
 
@@
 new DslDriver[Int, Int] {
-  def power(b: Rep[Int], x: Int): Rep[Int] =
-    if (x == 0) 1 else b * power(b, x-1)
+  def power(b: Rep[Int], x: Int): Rep[Int] = if (x == 0) 1 else b * power(b, x-1)
   def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
@@
 
-In the code shown above, \texttt{power} takes two arguments where \texttt{b} is
-declared as \texttt{Rep[Int]} type meaning that \texttt{b} is a representation
-of \texttt{Int} but whose value will be available at the next stage. Meanwhile,
-the result of the @power@ function will be available at the next stage as well,
-thus the return value of \texttt{power} is also of type \texttt{Rep[Int]}. Then
-we specialize @power@ in function @snippet@ by providing @x@ to @5@. The
-generated code for \texttt{power5} is a function that takes only one argument
-which corresponds to @b@, and the body of this function multiplies @b@ five
-times.
+\todo{Show some IR}
 
 \subsection{Staged Concrete Semantics}
+
+To implement the staged concrete interpreter, we reuse the exactly same concrete
+components from Section \ref{unstaged_conc}, but we instantiate the binding-time
+type @R[T]@ as @Rep[T]@.
+
+\paragraph{Staged Monads} We also need to use monads that operate on staged
+values. Note that the monadic objects such as @StateT@ or the internal value
+@R[S] => M[(A,S)]@ are not staged, only the data @S@ or @A@ are staged.
+This is why we can completely eliminate the overhead caused by the monadic layers.
+Here we use the prefix name @Rep@ to differentitate unstaged monads and staged
+monads, such as @RepReaderT@. So the @AnsM[_]@ is instantiated as the same monad
+stack before, but with @Rep@ versions.
+
+\begin{lstlisting}
+  type R[T] = Rep[T]
+  type AnsM[T] = RepReaderT[RepStateT[RepIdM, Store, ?], Env, T]
+\end{lstlisting}
+
+We use @RepStateT@ as an example to give a sketch of how the @Rep@ versions of monad
+transformers look like. The generic type @M[_]@ is a @RepMonad@ and used as the
+inner monad, which means the pair @(A,S)@ is also staged when appears inside of monad
+@M[_]@. The @flatMap@ is similar to the unstaged one, but the function @f@ takes
+a @Rep[A]@ value. Again, @f@ is not a next-stage function, but a current-stage
+function that transforms a next-stage value of @Rep[A]@ to @StateT[M,S,B]@.
+\todo{highlight the changes in the code}
+
+\begin{lstlisting}
+  case class RepStateT[M[_]: RepMonad, S, A](run: Rep[S] => M[(A, S)]) {
+    def flatMap[B](f: Rep[A] => StateT[M, S, B]): StateT[M, S, B] = ...
+  }
+\end{lstlisting}
+
+Readers may notice that the conversion between unstaged monads and monads is 
+simply changing the unstaged data to staged data, which in fact can be achieved without
+changing the implementation of monads. This is true so far, but as will see in
+the abstract interpreter when introducing the nondeterminism monad (@ListT@), as
+the whole list will be a next-stage value instead of merely the elements in the list.
+For this reason, we explicitly distinguish and present the two version of
+monads, as a more consistent approach.
+
+\paragraph{Primitive Operations} Most of the primitive operations can be easily
+translated to staged version -- we just need to let them works on staged data.
+
+\begin{lstlisting}
+  def emit_ap_clo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]): Rep[(Value, Store)]
+  def emit_compiled_clo(fun: (Rep[Value], Rep[Store]) => Rep[(Value, Store)],
+                        λ: Lam, ρ: Rep[Env]): Rep[Value]
+
+  def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
+    val Lam(x, e) = λ
+    val f: (Rep[Value], Rep[Store]) => Rep[(Value, Store)] = {
+      case (v: Rep[Value], σ: Rep[Store]) =>
+        val α = alloc(σ, x)
+        ev(e)(ρ + (unit(x) → α))(σ + (α → v))
+    }; emit_compiled_clo(f, λ, ρ)
+  }
+  def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = for {
+    σ <- get_store
+    val res: Rep[(Value, Store)] = emit_ap_clo(fun, arg, σ)
+    _ <- put_store(res._2)
+  } yield res._1
+\end{lstlisting}
+
+\paragraph{Code Generation}
+
+\begin{lstlisting}
+  case class IRApClo(f: Exp[Value], arg: Exp[Value], σ: Exp[Store]) extends Def[(Value, Store)]
+  override def emit_ap_clo(fun: Exp[Value], arg: Exp[Value], σ: Exp[Store]): Exp[(Value, Store)] =
+    IRApClo(fun, arg, σ)
+\end{lstlisting}
+
+\begin{lstlisting}
+  override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
+    case IRApClo(f, arg, σ) =>
+      emitValDef(sym, s"$\textdollar${quote(f)}.asInstanceOf[CompiledClo].f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
+    ...
+  }
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 component.
-
-\begin{lstlisting}
-trait RepConcInterpOps extends Concrete with LMSOps {
-  type R[+T] = Rep[T]
-  val $\rho$0: Rep[Env] = Map[Ident,Addr]()
-  val $\sigma$0: Rep[Store] = Map[Addr,Value]()
-  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
-  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): Rep[Env] = 
-    $\rho$ + (unit(x) -> a)
-  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = $\sigma$(a)
-  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): Rep[Store] = 
-    $\sigma$ + (a -> v)
-  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = $\sigma$.size + 1
-  def num(i: Lit): Rep[Value] = lift(NumV(i))
-  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans = {
-    //FIXME: the cast is ugly, patten match?
-    val i = cnd.asInstanceOf[Rep[NumV]].i
-    if (i == 0) thn else els
-  }
-  def prim_eval(op: Symbol, 
-    v1: Rep[Value], v2: Rep[Value]): Rep[Value] = op match {
-      case '+ => v1.asInstanceOf[Rep[NumV]].i + 
-                 v2.asInstanceOf[Rep[NumV]].i
-      ...
-    }
-  ...
-}
-\end{lstlisting}
 
@@
 will be provided later.
-%\note{to denotational style?, back to HO?}
+
 We will see this is an important observation that enables us to specialize an
@@
 \begin{lstlisting}
-case class ApplyClosure(f: Rep[Value], 
-  arg: Rep[Value], $\sigma$: Rep[Store]) extends Def[(Value, Store)]
+case class ApplyClosure(f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]) extends Def[(Value, Store)]
 def apply_closure(ev: EvalFun)
@@
 
-\paragraph{Code Generation}
-
 When generating code for the next stage, values like @CompiledClo@ and @NumV@
@@
 we need to treat @ApplyClosure@ a little differently.
-
-\begin{lstlisting}
-  case ApplyClosure(f, arg, $\sigma$) => 
-    emitValDef(sym, quote(f) + 
-                    ".asInstanceOf[CompiledClo].f(" + 
-                    quote(arg) + "," + 
-                    quote($\sigma$) + ")")
-\end{lstlisting}
 

### 2019-03-25 paper/30_stageconc.tex
misc

@@
 simply changing the unstaged data to staged data, which in fact can be achieved without
-changing the implementation of monads. This is true so far, but as will see in
-the abstract interpreter when introducing the nondeterminism monad (@ListT@), as
+changing the implementation of monads. This is true so far, but as we will see
+when introducing the nondeterminism monad (@ListT@) for abstract interpreters, as
 the whole list will be a next-stage value instead of merely the elements in the list.

### 2019-03-25 paper/30_stageconc.tex
section 3

@@
 translated to staged version -- we just need to let them works on staged data.
+As we mentioned before, we elaborate how to handle functions and applications in
+detail. One of the desired goal is to eliminate the monadic layers through
+staging, this becomes tricky for clsoures, because we need to actually compile
+the lambda term instead of generate a @CloV@ value. To make this happen, when
+closing the lambda term with the environment, we need to \textit{collpase} the
+monads to normal values. The following function @close@ illustrates the idea:
+@f@ is a current stage function but takes two next-stage values @v@ and
+$\sigma$, inside of which, we eagerly collapse the monads to values by
+@ev(e)(ρ_*)(σ_*)@, which has type @Rep[(Value, Store)]@.
+Then we generate an intermediate representation of the compiled closure using
+@emit_compiled_clo@. The intermediate representation stands for a next-stage
+value, and contains the compiled function, which will be generated as a
+next-stage function.
 
 \begin{lstlisting}
-  def emit_ap_clo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]): Rep[(Value, Store)]
-  def emit_compiled_clo(fun: (Rep[Value], Rep[Store]) => Rep[(Value, Store)],
-                        λ: Lam, ρ: Rep[Env]): Rep[Value]
-
+  def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[(Value, Store)]): Rep[Value]
   def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
@@
       case (v: Rep[Value], σ: Rep[Store]) =>
-        val α = alloc(σ, x)
-        ev(e)(ρ + (unit(x) → α))(σ + (α → v))
-    }; emit_compiled_clo(f, λ, ρ)
+        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ + (α + v)
+        ev(e)(ρ_*)(σ_*)
+    }; emit_compiled_clo(f)
   }
-  def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = for {
+\end{lstlisting}
+
+When applying a function in @ap_clo@, we generate an intermediate representation
+for closure application through @emit_ap_clo@, assuming that the @rator@ is already an
+IR of compiled closure, @rand@ is an arbitrary stage value, and $\sigma$ is the
+latest store. @emit_ap_clo@ represents a next-stage value of type
+@Rep[(Value, Store)]@, which is a collapsed monadic value from next stage, now
+we need to reify the value and store back into the monad stack through
+@put_store@ and @yield@.
+
+\begin{lstlisting}
+  def emit_ap_clo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]): Rep[(Value, Store)]
+  def ap_clo(ev: EvalFun)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
     σ <- get_store
-    val res: Rep[(Value, Store)] = emit_ap_clo(fun, arg, σ)
+    val res: Rep[(Value, Store)] = emit_ap_clo(rator, rand, σ)
     _ <- put_store(res._2)
@@
 
+The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces IR
+node in LMS. We define these IR node as follow: \note{Def, quote are omitted}
+
 \begin{lstlisting}
-  case class IRApClo(f: Exp[Value], arg: Exp[Value], σ: Exp[Store]) extends Def[(Value, Store)]
-  override def emit_ap_clo(fun: Exp[Value], arg: Exp[Value], σ: Exp[Store]): Exp[(Value, Store)] =
-    IRApClo(fun, arg, σ)
+  case class IRCompiledClo(f: Rep[((Value,Store)) => (Value, Store)]) extends Rep[Value]
+  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Rep[(Value, Store)]
 \end{lstlisting}
+
+After constructing the IR graph, LMS provides an @emitNode@ to generate code for
+each kind of IR node. To generate code for @IRApClo@, we know that @rator@ is a
+compiled closure @IRCompiledClo@, which contains a next-stage function @f@, so
+we just need to generate code that invokes @f@ with @rand@ and @σ@ as arguments
+provided. Similarly, we directly generate a next-stage value @CompiledClo(f)@
+for @IRCompiledClo@.
 
@@
   override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
-    case IRApClo(f, arg, σ) =>
-      emitValDef(sym, s"$\textdollar${quote(f)}.asInstanceOf[CompiledClo].f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
+    case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar$rand, $\textdollar$σ)")
+    case IRCompiledClo(f) => emitValDef(sym, s"CompiledClo($\textdollar$f)")
     ...
@@
 \end{lstlisting}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-In the staging part, we share the same concrete type instantiation, i.e., trait
-\texttt{Concrete}, but we reimplement the staged version for the concrete
-operations in trait @RepConcInterpOps@. Meanwhile, it is also extended from
-@LMSOps@ which provides necessary support for the staged version of primitive
-types and data structures. To stage the interpreter, we first identify the
-syntactic input that is known statically, and the interpreter returns a pair of
-values and store, which both becomes dynamic values. Thus the staged stores
-propagate the binding-time to environments, because they share the same address
-component.
-
-The first notable change is that the abstract type member @R@ is assigned to be
-@Rep@, and accordingly the affected types such as @Env@ and @Store@ become
-@Rep[Env]@ and @Rep[Store]@. For @num@, we use the @lift@ function which is a
-built-in in LMS that lifts a current-stage value to next stage. The interesting
-point is how we handle closures and function applications when staging is
-involved. In the unstaged version, function @close@ is used to lift a literal
-lambda term to a closure, which is a pair of the syntactic lambda term and the
-enclosing environment. However, what we desire is a \textit{compiled} closure,
-instead of a lifted @CloV@ value like what we did for @NumV@ --- in other words,
-the syntactic term of the lambda expression should be eliminated and specialized
-away. The specialization of the interpreter with respect to the body of the
-lambda term proceeds under the assumptions that the argument and latest store
-will be provided later.
-
-We will see this is an important observation that enables us to specialize an
-abstract interpreter in a modular way. At the end of function @close@, we create
-a @CompiledClo@ object that contains a staged function, which all will be
-emitted in the generated code.
-
-\begin{lstlisting}
-def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
-  val Lam(x, e) = $\lambda$
-  val f: Rep[(Value,Store)]=>Rep[(Value,Store)] = { 
-    case (arg: Rep[Value],$\sigma$: Rep[Store]) =>
-      val $\alpha$ = alloc($\sigma$, x)
-      ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, arg)) 
-  }
-  CompiledClo(fun(f))
-}
-\end{lstlisting}
-
-Next, @apply_closure@ takes three arguments: the function value, argument value
-and the latest store, which are all of @Rep@ type. This means basically there
-is nothing we can do at the current stage --- so we create a new node
-@ApplyClosure@ that contains these arguments in the intermediate representation
-graph, and in the code generation phase we can emit the code that performs the
-application. The additional @reflectEffect@ is a function in LMS that handles
-side effects.
-
-\begin{lstlisting}
-case class ApplyClosure(f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]) extends Def[(Value, Store)]
-def apply_closure(ev: EvalFun)
-  (f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]): Ans = {
-    reflectEffect(ApplyClosure(f, arg, $\sigma$))
-  }
-\end{lstlisting}
-
-When generating code for the next stage, values like @CompiledClo@ and @NumV@
-are kept as they are because they are intended exist in the next stage; however,
-we need to treat @ApplyClosure@ a little differently.
-
-Concretely, as shown in the code above, when emitting code for @ApplyClosure@,
-we just emit a function application where the function value is extracted from
-the @CompiledClo@ object (we assume the input program is well-typed so that we
-can safely cast it to @CompiledClo@). Argument and a store are the remaining rest
-fields in @ApplyClosure@.

### 2019-03-25 paper/50_stageabs.tex
section 3

--- 2019-02-04 paper/30_sai.tex
@@
 
-\subsection{Specialized Data Structures} \label{staged_ds}
-
-Now we have already obtained an end-to-end staged abstract interpreter that is
-able to specialize an analysis. However, we treat the data structures such as
-@Map@s as black-boxes, which means any operations on a @Map@ become code in the
-next stage. But, as we identified when introducing the generic interface, the
-keys of any environment maps are identifiers in the program, which are
-completely known statically. This leaves us a chance to further specialize the
-data structures. Assume the @Map[K,V]@ is implemented as a hash map, if the keys
-$K$ are known, then the indices can be computed statically. Thus the specialized
-map would be an array @Array[Rep[V]]@ whose elements are next-stage values; all
-the accesses to the array is determined during staging.
-
-Particularly, if we are specializing a monovariant analysis, the address space
-is equivalent to the identifiers, then the environment component can be entirely
-eliminated, and the store is a specialized map as array of @Rep[Value]@
-elements.
-
 \subsection{Modular Analysis for Free}
@@
 
+\subsection{Optimizations} \label{staged_ds}
+
+Our staging schema works well in theory, but would suffer from code explosion,
+and possibly runtime GC overhead when analyzing (specializing) large programs.
+We present several optimizations that largely mitigate these issues. Still,
+these optimizations do not need to change the generic interpreter.
+
+\paragraph{Specialized Data Structures}
+
+Now we have already obtained an end-to-end staged abstract interpreter that is
+able to specialize an analysis. However, we treat the data structures such as
+@Map@s as black-boxes, which means any operations on a @Map@ become code in the
+next stage. But, as we identified when introducing the generic interface, the
+keys of any environment maps are identifiers in the program, which are
+completely known statically. This leaves us a chance to further specialize the
+data structures. Assume the @Map[K,V]@ is implemented as a hash map, if the keys
+$K$ are known, then the indices can be computed statically. Thus the specialized
+map would be an array @Array[Rep[V]]@ whose elements are next-stage values; all
+the accesses to the array is determined during staging.
+
+Particularly, if we are specializing a monovariant analysis, the address space
+is equivalent to the identifiers, then the environment component can be entirely
+eliminated, and the store is a specialized map as array of @Rep[Value]@
+elements.
+
+\paragraph{Selective Caching} It is observed that the two-fold caching is
+wrapped for recursive call of our abstract interpreter. But this is even more
+expensive when generating code for atomic expressions such as literal numbers or
+variables. \todo{caching prevents nontermination}
+
+\paragraph{Partially-static Data}
+
+For example, to fold a singleton list, e.g., @List(x).foldLeft(init)(f)@ where
+@x@ and @init@ are staged values, a naive code generator would faithfully
+apply @foldLeft@ to the list with an anonymous function. But we can also utilize
+the algebraic property of @foldLeft@ to generate cheaper code:
+@List(x).foldLeft(init)(f) = f(init, x)@.
+Since the function @f@ is known at current stage, we completely eliminate the
+fold operation and function application. We apply several rewritings enabled by
+partial-static data structures, such as @List@ and @Map@, which greatly reduce
+the size of residual programs.
+
+\paragraph{Heterogeneous Staging} The generated program is in A-Normal form,
+
 \subsection{Discussion}
 
-In the literature of partial evaluation, \citeauthor{10.1007/3-540-61580-6_11}
-provided guidelines on what to do and not to do when specializing a concrete
-interpreter \cite{10.1007/3-540-61580-6_11}. We borrow these guidelines and
-extend them to abstract interpreters. We discuss decisions we made to achieve
-this and examine some alternatives.
+We have gradually presented the confluence of specialization and abstraction of
+concrete interpreters. In this section, we discuss several decisions we made to
+achieve this and examine some alternatives.
+
+\paragraph{Are Monads Necessary?} No. One can always inline the monads and
+obtain an abstract interpreter in continuation-passing style, or simply use
+explicit side-effects such mutation to implement the artifact with the same
+abstract semantics. In either cases, we can still apply the staging schema to
+abstract interpreter. As an evidence, we provide an staged abstract interpreter
+written in direct-style in the accompanying artifact.
 
@@
 
-What we implemented is a big-step, compositional abstract interpreter, where
-"compositional" means that every recursive call of our abstract interpreter is
-applied to proper substructures of the current syntactic parameters
-\cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
-specialization can be done by unfolding, as well as that the specialization
-procedure terminates. Nevertheless, it is also possible to specialize a
-small-step operational abstract semantics ---
-\citeauthor{Johnson:2013:OAA:2500365.2500604} showed this in abstract
-compilation \cite{Boucher:1996:ACN:647473.727587} style as one of their
-optimizations of Abstract Abstract Machines
-\cite{Johnson:2013:OAA:2500365.2500604}. However, the generated abstract
+What we implemented is a big-step, compositional abstract interpreter in monadic
+style, where "compositional" means that every recursive call of our abstract
+interpreter is applied to proper substructures of the current syntactic
+parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
+specialization can be done by unfolding, as well as the termination of
+specialization procedure. It is also possible to specialize small-step
+operational abstract semantics through abstract compilation
+\cite{Boucher:1996:ACN:647473.727587} --
+\citet{Johnson:2013:OAA:2500365.2500604} implemented it for
+optimizating Abstract Abstract Machines. However, the generated abstract
 bytecode still requires another small-step abstract machine to execute, which is
-an additional engineering efforts. Another alternative approach for efficient
-specialization is to write the abstract interpreter in a big-step, monadic style
-\cite{DBLP:journals/pacmpl/DaraisLNH17}.
+an additional engineering efforts.
 
@@
 
-Based on the assumption that LMS preserves the semantics during staging, we are
-confident that the staged abstract interpreter does the same analysis compared
-with the unstaged one. Moreover, the optimization done by staging does not
-compromise any soundness.
+As one of the merit of our approach, the staging does not compromise any
+soundness, because we only change the underlying monads and data structures to
+the staged version, not any part of the abstract semantics. Based on the
+assumption that LMS preserves the equivalence during staging, we are confident
+that the staged abstract interpreter does the same analysis as the unstaged one.

### 2019-03-25 virtmacro-code/src/main/scala/sai/MainGeneric.scala

@@ -646,7 +646,7 @@ trait StagedAbstractSemantics extends AbstractComponents with RepMonads with Rep
   def run_nonsel(e: Expr): (Rep[List[(Value, Store)]], Rep[Cache]) = fix_nonsel(eval)(e)(ρ0)(σ0)(cache0)(cache0)
 }
 
-trait StagedAbstractSemanticsExp extends StagedAbstractSemantics with SAIOpsExp {
+trait StagedAbstractSemanticsExp extends StagedAbstractSemantics with SAIOpsExp with AbsEnvExpOpt {
   //TODO: when evaluating, change lam/rho to hash code
   case class IRCompiledClo(f: (Exp[Value], Exp[Store], Exp[Cache], Exp[Cache]) => Exp[(List[(Value, Store)], Cache)],
                            rf: Exp[((Value, Store, Cache, Cache)) => (List[(Value, Store)], Cache)],

### 2019-03-26 paper/40_absinterp.tex
section 4

@@
 
-Based on the generic interface we presented in Section~\ref{bg_lang}, we now
-describe the instantiation of an abstract interpreter. To keep the presentation
-simple, we use simple abstract domains, and just establish a
-context-/path-/flow-insensitive, store-widened analysis in this section ---
-indeed, it is coarse but simple enough to setup a foundation for the staged
-abstract interpreter we will present later. In Section~\ref{cfa}, we will see
-how to regain context-/path-/flow-sensitivity \todo{path/flow} for pushdown
-control-flow analysis; in Section~\ref{cases_imp}, we will see how to
-instantiate an interval abstract domain for numerical analysis.
+After seeing both unstaged and staged concrete interpreters, now we trun our
+focus on abstract interpreters under the same framework. We will first present a
+lattice framework with binding-time annotations, but for the purpose of clearly
+illustrating the idea, we just use simple abstract domains such as power set
+lattice or point-wise map lattice. Then we show the abstract components such as
+abstract environments and abstract stores.
+The abstract interpreter we present in this section is similar to
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s, it is also
+context/path/flow-insensitive by our choice.  In Section \ref{cfa}, we will
+instantiate a context-sensitive analysis;
+\citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
+flow-sensitive by varying the monads, which is also applicable to our unstaged or
+staged abstract interpreter. \todo{numerical/imp?}
 
-\paragraph{Abstract Instantiation}
+\subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
+
+A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
+\bot, \sqcup, \sqcap \rangle$, where $\sqsubseteq : S \to S \to Bool$ is the
+ordering relation, $\sqcup: S \to S \to S$ is the join (least upper bound)
+operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
+operator. The trait @SPLattice@ declares the type class for lattices.
+Similar to @MonadOps@ in Section \ref{generic_if}, we introduce an additional
+higher-kinded type @R[_]@ to annotate the binding-times.
+
+\begin{lstlisting}
+  trait SPLattice[S, R[_]] {
+    val $\top$: R[S];  val $\bot$: R[S]
+    def $\sqsubseteq$(l1: R[S], l2: R[S]): R[Boolean]
+    def $\sqcup$(l1: R[S], l2: R[S]): R[S]
+    def $\sqcap$(l1: R[S], l2: R[S]): R[S]
+  }
+  trait Lattice[S] extends SPLattice[S, NoRep]
+\end{lstlisting}
+
+In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
+unstaged lattice type class. An simple instance of lattice we used is the power
+set abstract domain, where the ordering is subset relation, join is the union
+operation, and meet is the intersect operation. The power set lattice can be
+implemented as follows:
+
+\begin{lstlisting}
+  def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
+    lazy val $\bot$: Set[T] = Set[T]()
+    lazy val $\top$: Set[T] = throw new NotImplementedError("No representation for top")
+    def $\sqsubseteq$(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf l2
+    def $\sqcup$(l1: Set[T], l2: Set[T]): Set[T] = l1 union l2
+    def $\sqcap$(l1: Set[T], l2: Set[T]): Set[T] = l1 intersect l2
+  }
+\end{lstlisting}
+
+Other lattices used in the rest of the paper, such as products and maps, can be
+implemented similarly, or lifted from existing lattices element-wise or point-wise.
+
+\subsection{Abstract Semantics}
+
+\paragraph{Abstract Components}
 
@@
 
-\subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
+\paragraph{Monads for Abstract Interpretation} nondetT
 
-To effectively reuse the code between unstaged and staged variants, we also make
-the lattice structure stage polymorphic. A lattice type class parameterizes over
-an element type \texttt{E} and defines five operations for \texttt{R[E]}, such
-as meet, join, and ordering relation; again, \texttt{R} is the higher-kinded
-type indicating the binding-time.
-
-\begin{lstlisting}
-trait Lattice[E, R[+_]] {
-  val bot: R[E];  val top: R[E]
-  def $\sqsubseteq$(l1: R[E], l2: R[E]): R[Boolean]
-  def $\sqcup$(l1: R[E], l2: R[E]): R[E]
-  def $\sqcap$(l1: R[E], l2: R[E]): R[E]
-}
-\end{lstlisting}
-
-For example, here we use powersets as the abstract domain, and the unstaged
-lattice of powersets can be easily implemented as follows:
-
-\begin{lstlisting}
-implicit def SetLattice[T]: Lattice[Set[T], NoRep] = 
-  new Lattice[Set[T], NoRep] {
-    lazy val bot: Set[T] = Set[T]()
-    lazy val top: Set[T] = throw new NotImplementedError()
-    def $\sqsubseteq$(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf l2
-    def $\sqcup$(l1: Set[T], l2: Set[T]): Set[T] = l1 union l2
-    def $\sqcap$(l1: Set[T], l2: Set[T]): Set[T] = l1 intersect l2
-  }
-\end{lstlisting}
-
-The ordering relation is to ask whether one set is a subset of the other, 
-join is to union the two sets, and meet is to intersect the two sets.
-Accordingly, other structures used in the rest of the paper such as tuples 
-(products) and maps can be lifted to lattices element-wise or point-wise.
-The code for them are elided.
-
-\subsection{Abstract Semantics}
+\paragraph{Primitive Operations}
 
@@
 
-\subsection{Fixpoint Iteration}
+\paragraph{Caching and Fixpoint Iteration}
 

### 2019-03-26 paper/20_prelim.tex
paper writing

@@
 \paragraph{The Interpreter} Now we can define the semantics-agnostic interpreter
-in monadic form. The essential idea is to traverse the abstract syntax tree
-while maintaining the effects such as reader and state. 
+in monadic form, shown in Figure \ref{fig:shared_int}.
+The essential idea is to traverse the abstract syntax tree while maintaining the
+effects such as reader and state.
 It is worth noting that the interpreter is written in open-recursive style -- it
@@
 
+\vspace{-1em}
+\begin{figure}[h!]
+  \centering
   \begin{lstlisting}
@@
   \end{lstlisting}
+\caption{The generic interpreter,
+  shared by the unstaged/staged + concrete/abstract interpreter.}
+\label{fig:shared_int}
+\end{figure}
+\vspace{-1em}
 

### 2019-03-26 paper/30_stageconc.tex
paper writing

@@
 changing the implementation of monads. This is true so far, but as we will see
-when introducing the nondeterminism monad (@ListT@) for abstract interpreters, as
-the whole list will be a next-stage value instead of merely the elements in the list.
+it not that easy when introducing the nondeterminism monad (@ListT@) for
+abstract interpreters, as the whole list will be a next-stage value instead of
+merely the elements in the list.
 For this reason, we explicitly distinguish and present the two version of

### 2019-03-26 paper/40_absinterp.tex
paper writing

@@
 operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
-operator. The trait @SPLattice@ declares the type class for lattices.
-Similar to @MonadOps@ in Section \ref{generic_if}, we introduce an additional
-higher-kinded type @R[_]@ to annotate the binding-times.
+operator. The trait @SPLattice@ defines the type class of stage polymorphic
+lattices. Similar to @MonadOps@ in Section \ref{generic_if}, we introduce an
+additional higher-kinded type @R[_]@ to annotate the binding-times.
 
@@
 unstaged lattice type class. An simple instance of lattice we used is the power
-set abstract domain, where the ordering is subset relation, join is the union
-operation, and meet is the intersect operation. The power set lattice can be
+set abstract domain, where the ordering is subset relation, the join is union
+operation, and the meet is intersect operation. The power set lattice can be
 implemented as follows:
@@
 
-Our instantiation roughly follows the Abstracting Abstract Machines methodology
-\cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12} that transforms the
-concrete semantics to abstract semantics. The store now maps addresses to sets
-of abstract values, meaning that an address points to all possible values that
-may occur at runtime, where the abstract value is either a closure or a single
-abstract number \texttt{NumV}, which stands for the top element of the number
-lattice. The address space is also constrained to be finite, in the monovariant
-setting, we simply use variable names for address.
+% \cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12}
+
+We follow the \textit{abstracting definitional interpreters} (ADIs) approach
+\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
+an abstract interpreter. We first widen the concrete values to @AbsValue@: numbers
+are lifted to a singleton object @IntTop@ representing the set of all numbers;
+closures remain the same. Then the type @Value@ is sets of @AbsValue@ (we abuse
+the type names a little bit, as the generic interface defined the abstract type
+@Value@).
+The address space is constrained to be finited to ensure the computatbility of
+analysis -- for the simplest monovariant analysis, we use variable names for
+addresses. After setteling the @Value@ and @Addr@ type, the environments and
+stores are automatically lifted to abstract environments and abstract stores,
+i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
+It is importatnt that the abstract store now maps addresses to sets of abstract
+values, meaning an address may points to an over-approximated set of values
+occurred at runtime.
 
 \begin{lstlisting}
-trait Abstract extends Semantics {
-  case class Addr(x: Ident);  sealed trait AbsValue
-  case class CloV($\lambda$: Lam, $\rho$: Env) extends AbsValue
-  case class NumV() extends AbsValue
-  type Value = Set[AbsValue]
-  type Env = Map[Ident, Addr];  type Store = Map[Addr, Value]
+  trait AbstractComponents extends Semantics {
+    sealed trait AbsValue
+    case object IntTop extends AbsValue
+    case class CloV(lam: Lam, env: Env) extends AbsValue
+    type Value = Set[AbsValue]; case class Addr(x: String)
+    ...
   }
+\end{lstlisting}
+
+A cache is a mapping from configurations to sets of value-store pairs, where the
+configuration is a triple of current expression being evaluated, environment and store.
+
+\begin{lstlisting}
+  type Config = (Expr, Env, Store); type Cache = Map[Config, Set[(Value, Store)]]
+  type Result = (R[List[(Value, Store)]], R[Cache])
 \end{lstlisting}
@@
 \begin{lstlisting}
-object AbsInterp extends Abstract {
-  type R[+T] = T
-  val $\rho$0 = Map[Ident, Addr]();  val $\sigma$0 = Map[Addr, Value]()
-  def get($\rho$: Env, x: Ident): Addr = $\rho$(x)
-  def put($\rho$: Env, x: Ident, a: Addr): Env = $\rho$ + (x -> a)
-  def get($\sigma$: Store, a: Addr): Value = 
-    $\sigma$.getOrElse(a, Lattice[Value].bot)
-  def put($\sigma$: Store, a: Addr, v: Value): Store =
-    $\sigma$ + (a -> (v $\sqcup$ get($\sigma$, a)))
-  def alloc($\sigma$: Store, x: Ident): Addr = Addr(x)
-  def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Env): Value = Set(CloV($\lambda$, $\rho$))
-  def num(i: Lit): Value = Set(NumV())
-  def apply_closure(ev: EvalFun)
-    (fs: Value, arg: Value, $\sigma$: Store): Ans = { var $\sigma$_* = $\sigma$
-      val vs = for (CloV(Lam(x, e), $\rho$) <- fs) yield {
-        val $\alpha$ = alloc($\sigma$_*, x)
-        val (v, v$\sigma$) = ev(e, put($\rho$, x, $\alpha$), put($\sigma$_*, $\alpha$, arg))
-        $\sigma$_* = v$\sigma$ $\sqcup$ $\sigma$_*; v
-      }
-    (vs.reduce(Lattice[Value].$\sqcup$), $\sigma$_*)
-  }
-  def branch0(cnd: Value, thn: =>Ans, els: =>Ans): Ans = 
-    thn $\sqcup$ els
-  def prim_eval(op: Symbol, v1: Value, v2: Value): Value = 
-    Set(NumV())
-} // to be continued
 \end{lstlisting}
@@
 \paragraph{Caching and Fixpoint Iteration}
+
+\begin{lstlisting}
+  def ask_in_cache: AnsM[Cache]
+  def get_out_cache: AnsM[Cache]
+  def put_out_cache(out: R[Cache]): AnsM[Unit]
+  def update_out_cache(cfg: R[Config], vs: R[(Value, Store)]): AnsM[Unit]
+\end{lstlisting}
 
@@
 
-\begin{lstlisting}
-case class CacheFix(evev: EvalFun => EvalFun) {
-  var in = Map[Config, Ans](); var out = Map[Config, Ans]()
-  def cached_ev(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = {
-    val cfg: Config = (e, $\rho$, $\sigma$)
-    if (out.contains(cfg)) out(cfg)
-    else {
-      val ans0 = in.getOrElse(cfg, Lattice[(Value, Store)].bot)
-      out = out + (cfg -> ans0)
-      val ans1 = evev(cached_ev)(e, $\rho$, $\sigma$)
-      out = out + (cfg -> (ans0 $\sqcup$ ans1));  ans1
-    }
-  }
-  def iter(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = {
-    in = out; out = Map[Config, Ans](); cached_ev(e, $\rho$, $\sigma$)
-    if (in == out) out((e, $\rho$, $\sigma$)) else iter(e, $\rho$, $\sigma$)
-  }
-}
-override def eval_top(e: Expr, $\rho$: Env, $\sigma$: Store): Ans = 
-  CacheFix(eval).iter(e, $\rho$, $\sigma$)
-\end{lstlisting}
-
 Finally, we override the definition of @eval_top@ by instantiating @CacheFix@

### 2019-03-26 paper/10_intro.tex
section 3
section 4

@@
-\section{Introduction}
+\section{Introduction} \label{intro}
 
@@
 Futamura projections, which first identifies the binding-time of variables in
-the program, i.e., they can be known whether statically or dynamically, and then
+the program, i.e., they can be known whether statically or dynamically, then
 evaluates the static part, and finally generates a residual program that solely
 relies on the dynamic part. However, given an arbitrary program, precisely
-analyzing its binding-time is hard in general. As an alternative and pragmatic
+analyzing its binding-times is hard in general. As an alternative and pragmatic
 approach to specialization and partial evaluation, multi-stage programming (MSP
@@
 Now to concretely illustrate the idea of Futamura projection, consider that we
-have an operational interpreter @eval@ of some language as following:
+have an operational interpreter @eval@ of some language:
 
@@
 $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
-argument @arg@ to an environment (as well as a companion heap object, if for the
-interest of using assignments) that maps free variables in $e$ to some values, 
-the interpreter is a standard environment/store-passing interpreter,
+argument @arg@ to an environment (as well as an accompanying heap object, if for
+the interest of using assignments) that maps free variables in $e$ to some
+values, the interpreter is a standard environment-passing interpreter,
 where the specialization through MSP is also applicable.
@@
 Roughly at the same time when Futamura projections was visioned in 1970s,
-\citeauthor{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
+\citet{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
 lattice-based semantic approach to construct sound static analysis, by
-approximation of fix-points \cite{DBLP:conf/popl/CousotC77}. However,
-constructing sound abstract interpreters was considered abstruse and complicated
-for a long time.
+approximation of fix-points. However, constructing sound abstract interpreters
+was considered abstruse and complicated for a long time.
 Recent progress such as Abstracting Abstract Machines (AAM)
-uncovers a systematic method to derive sound abstract interpreters from their
-concrete counterparts, where the soundness can be easily established by
-examining the transformation of semantics.
-The AAM approach also has been been applied to different variants of
-definitional interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
+\cite{DBLP:conf/icfp/HornM10} uncovers a systematic method to derive sound
+abstract interpreters from their concrete counterparts, where the soundness can
+be easily established by examining the transformation of semantics. The AAM
+approach also has been been applied to different variants of definitional
+interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
 DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
@@
 approximation, we can refactor interpreters to static analyzers. As an
-intellectual quest, we should naturally ask -- starting from an interpreter,
-\textit{can we derive a sound abstract interpreter while it produces fast code
-that does the analysis?}
+intellectual quest, we should naturally ask -- starting from an abstract
+interpreter, \textit{can we derive a sound abstract interpreter while it
+produces fast code that does the analysis?}
 Conceptually, it is similar to the concrete setting: for an
@@
 environment and returns abstract values, we would like to specialize it with
-respect to a program $e_0$ and produce a compiled analysis
+respect to some program $e_0$ and produce a compiled analysis
 $\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
@@
 
+\vspace{-1em}
 \begin{figure}[h]
@@
   {
-    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \end{smallmatrix} \\
-      \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} & \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \end{smallmatrix} \\ };
+    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_abs}) \end{smallmatrix} &
+    \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{sai}) \end{smallmatrix} \\
+    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{prelim}) \end{smallmatrix} &
+    \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{stagedinterp}) \end{smallmatrix} \\ };
   \path[-stealth]
@@
 \end{figure}
+\vspace{-1em}
 
@@
 abstractions -- from the monadic layers -- is completely eliminated in the
-generated code. However, we do not lose any advantages of this high-level
-programming style (e.g., the extensibility); the meaning of the program also
+generated code. \todo{put some numbers} However, we do not lose any advantages of this high-level
+programming style (e.g., modularity and extensibility); the meaning of the program also
 remains the same, particularly, the abstract semantics that does the analysis
@@
 to improve the latter one while the former one is untouched.
-In the literature, specializing analysis with respect to a program by partial
-evaluation \cite{damian1999partial, amtoft1999partial,
-Boucher:1996:ACN:647473.727587, ashley:practical} is not a new, despite the fact
-that most of the previous work are ad-hoc on a particular analysis or requiring
-significant changes to the analyzer. For example, abstract compilation
-\cite{Boucher:1996:ACN:647473.727587} requires the whole analyzer to be
-rewritten as closure generation from. In this paper, we advocate a systematic
+In the literature, specializing analysis with respect to the input program
+\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
+ashley:practical} is not a new, despite the fact that most of the previous work
+are ad-hoc on a particular analysis or requiring significant changes to the
+analyzer. For example, abstract compilation is one of the techniques
+\cite{Boucher:1996:ACN:647473.727587}. But it requires the whole analyzer to be
+rewritten to the closure generation from. In this paper, we advocate a systematic
 approach to meta-programming that achieves the performance goal with less
-intrusive changes; in fact, a case study (Section~\ref{cs_ac}) comparing with
-abstract compilation shows, by utilizing type-based stage annotations, the
-analyzer program does not need any changes.
+intrusive changes; in fact, a case study (Section~\ref{cs_ac}) comparing our
+approach with abstract compilation shows that, by utilizing type-based stage
+annotations, the analyzer program does not need any changes.
 
-\paragraph{Modularity via Meta-Programming}
+\paragraph{Modularity via Meta-Programming} \todo{compress this part}
 
@@
 \end{itemize}
+
+\todo{mention LMS somewhere}
 

### 2019-03-26 paper/30_stageconc.tex
section 4

--- 2019-03-25 paper/30_stageconc.tex
@@
 changing the implementation of monads. This is true so far, but as we will see
-when introducing the nondeterminism monad (@ListT@) for abstract interpreters, as
-the whole list will be a next-stage value instead of merely the elements in the list.
+it is not that easy when introducing the nondeterminism monad (@ListT@) for
+abstract interpreters, as the whole list will be a next-stage value instead of
+merely the elements in the list.
 For this reason, we explicitly distinguish and present the two version of
@@
 staging, this becomes tricky for clsoures, because we need to actually compile
-the lambda term instead of generate a @CloV@ value. To make this happen, when
-closing the lambda term with the environment, we need to \textit{collpase} the
-monads to normal values. The following function @close@ illustrates the idea:
+the lambda term instead of generating a @CloV@ value. To make this happen, when
+closing the lambda term with the environment, we need to staging into the body
+expression by \textit{collpasing} the @AnsM@ monads to normal values.
+The following function @close@ illustrates the idea:
 @f@ is a current stage function but takes two next-stage values @v@ and
@@
 The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces IR
-node in LMS. We define these IR node as follow: \note{Def, quote are omitted}
+node in LMS. We define these IR node as follow: \note{Exp, Def, quote are omitted}
 

### 2019-03-26 paper/40_absinterp.tex
section 4

@@
 \citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
-flow-sensitive by varying the monads, which is also applicable to our unstaged or
-staged abstract interpreter. \todo{numerical/imp?}
+flow-sensitivity by varying the monads, which is also applicable to our unstaged
+or staged abstract interpreter.
 
@@
   trait SPLattice[S, R[_]] {
-    val $\top$: R[S];  val $\bot$: R[S]
-    def $\sqsubseteq$(l1: R[S], l2: R[S]): R[Boolean]
-    def $\sqcup$(l1: R[S], l2: R[S]): R[S]
-    def $\sqcap$(l1: R[S], l2: R[S]): R[S]
+    val ⊤: R[S];  val ⊥: R[S]
+    def ⊑(l1: R[S], l2: R[S]): R[Boolean]
+    def ⊔(l1: R[S], l2: R[S]): R[S]
+    def ⊓(l1: R[S], l2: R[S]): R[S]
   }
@@
   def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
-    lazy val $\bot$: Set[T] = Set[T]()
-    lazy val $\top$: Set[T] = throw new NotImplementedError("No representation for top")
-    def $\sqsubseteq$(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf l2
-    def $\sqcup$(l1: Set[T], l2: Set[T]): Set[T] = l1 union l2
-    def $\sqcap$(l1: Set[T], l2: Set[T]): Set[T] = l1 intersect l2
+    lazy val ⊤: Set[T] = Set[T]()
+    lazy val ⊥: Set[T] = throw new NotImplementedError("No representation for top")
+    def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
+    def ⊔(l1: Set[T], l2: Set[T]): Set[T]  = l1 union     l2
+    def ⊓(l1: Set[T], l2: Set[T]): Set[T]  = l1 intersect l2
   }
@@
 
-A cache is a mapping from configurations to sets of value-store pairs, where the
-configuration is a triple of current expression being evaluated, environment and store.
+To prevent the abstract interpreter from falling into nontermination, the ADI
+approach uses a two-cache mechanism. Here we first provides the type
+definitions, and later we will describe the algorithm in detail. A @Cache@ is a
+mapping from configurations @Config@ to sets of value-store pairs, where the
+configuration is a triple of current expression being evaluated, environment and
+store.
 
@@
   type Config = (Expr, Env, Store); type Cache = Map[Config, Set[(Value, Store)]]
+\end{lstlisting}
+
+As of now, the store is a mapping from addresses to sets of abstract values.
+Therefore, when applying a function, for instance @f(a)@, we may retrieve a set
+of closures by dereferencing the address of @f@. We need to apply all of the
+target closures, where the nondeterminism is introduced. Another nondeterminism
+happens when evaluating conditionals, we have to evaluate both branches if the
+condition can not be determined. In other words, we will have multiple computing
+results.
+
+\begin{lstlisting}
   type Result = (R[List[(Value, Store)]], R[Cache])
@@
 
-\paragraph{Monads for Abstract Interpretation} nondetT
+This oberservation leads us to the type definition of @Result@: in concrete setting,
+the @Result@ is a pair of @Value@ and @Store@, now we use a list of such pairs
+@List[(Value, Store)]@ to represent multiple results. The results is also linked
+with the resulting cache used in this iteration.
+Note that the abstract binding-time annotation is wrapped to the two objects in
+the @Result@ pair, but not the whole pair. Because we statically know the result
+is a tuple of two elements, but we don't know how many value-store pairs we
+would have, neither the content of the cache.
 
-\paragraph{Primitive Operations}
+\paragraph{Monads for Abstract Interpretation} Now we incorporate the
+nondeterminism and caching into the monad stack. For concrete evaluation, we
+already have a reader effect for environments and a state effect for stores. We
+add three more effects based on it: the nondeterminism effect is represented by
+the @ListT[M[_], A]@ monad transformer, inside of which, we have a @ReaderT@ for
+a @in@ cache, and a @StateT@ for an @out@ cache.
+\todo{highlight where changed}
+
+\begin{lstlisting}
+  trait AbstractSemantics extends AbstractComponents {
+    type R[T] = T
+    type AnsM[T] = ReaderT[StateT[ListT[ReaderT[StateT[IdM, Cache, ?], Cache, ?], ?], Store, ?], Env, T]
+    ...
+  }
+\end{lstlisting}
+
+The sketch implementation of @ListT@ is shown below. The value @run@ it encapsulates
+is of type @M[List[A]]@, where @M[_]@ is the inner monad, and @A@ the element
+type of the list.
+
+\begin{lstlisting}
+  case class ListT[M[_]: Monad, A](run: M[List[A]]) { ... }
+\end{lstlisting}
+
+\paragraph{Primitive Operations} \todo{wip}
 
@@
 
-\paragraph{Caching and Fixpoint Iteration}
+\paragraph{Caching and Fixpoint Iteration} Caching
 
 \begin{lstlisting}
-  def ask_in_cache: AnsM[Cache]
-  def get_out_cache: AnsM[Cache]
+  def ask_in_cache: AnsM[Cache]; def get_out_cache: AnsM[Cache]
   def put_out_cache(out: R[Cache]): AnsM[Unit]
-  def update_out_cache(cfg: R[Config], vs: R[(Value, Store)]): AnsM[Unit]
+  def set_out_cache(cfg: R[Config], vs: R[(Value, Store)]): AnsM[Unit]
 \end{lstlisting}

### 2019-03-26 paper/50_stageabs.tex
section 4

@@
 
-\subsection{Modular Analysis for Free}
-
-One of the challenges of modern static analysis is program usually depends on
-large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. Can we analyze
-programs and libraries separately and reuse the result without losing precision?
-Then we can reduce part of the overhead of repeatedly analyzing libraries for
-different programs. Indeed, some static analyzers compute summary for a function
-or a module, which can be reused later, however they are mostly too conservative
-or unsound, which both lead to imprecision.
-
-The specialization of abstract interpreter provides a chance to obtain such
-partial analysis result in a mechanized way, but still keeps the analysis sound.
-As we see when compiling the closures, we can specialize the abstract
-interpreter with respect the body expression of the lambda term without knowing
-the actual argument. The programs with some unknown variables are open programs,
-which is exactly the case if we want to analyze programs in a modular way.
-
-For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a
-whole-program analysis, because it is inter-procedural and needs the last $k$
-calling contexts to distinguish different call sites, where the calling contexts
-are dynamic values. However, it is possible to analyze programs (libraries)
-separately through specializing an abstract interpreter that generates the
-analysis as the next-stage program and leave the unavailable programs and
-calling contexts as dynamic parameters, and then install these contexts when we
-have the whole program.
-
-\iffalse
-Another perspective: programs are data for an abstract interpreter, so if we
-have $n$ programs, then maybe there can be $n$ stages. Probably we can analyze
-first $m$ programs, and generate a residual abstract interpreter waiting for the
-rest $(n-m)$ programs. These $(n-m)$ programs might be (abstract) arguments for
-the first $n$ programs, and the abstract interpreter itself might be a partial
-abstract interpreter.
-\fi
-
 \subsection{Optimizations} \label{staged_ds}
@@
 We present several optimizations that largely mitigate these issues. Still,
-these optimizations do not need to change the generic interpreter.
+implementing these optimizations do not need to change the generic interpreter.
 
@@
 abstract semantics. In either cases, we can still apply the staging schema to
-abstract interpreter. As an evidence, we provide an staged abstract interpreter
-written in direct-style in the accompanying artifact.
+the abstract interpreter. As an evidence, we provide an staged abstract
+interpreter written in direct-style in the accompanying artifact.
 
@@
 What we implemented is a big-step, compositional abstract interpreter in monadic
-style, where "compositional" means that every recursive call of our abstract
+style, where \textit{compositional} means that every recursive call of our abstract
 interpreter is applied to proper substructures of the current syntactic
@@
 the staged version, not any part of the abstract semantics. Based on the
-assumption that LMS preserves the equivalence during staging, we are confident
-that the staged abstract interpreter does the same analysis as the unstaged one.
+assumption that MSP system and staged data structure preserve the equivalence
+during staging, we are confident that the staged abstract interpreter does the
+same analysis as the unstaged one.

### 2019-03-26 paper/60_cases.tex
section 4

@@
-\section{Case Study} \label{cases_study}
+\section{Case Studies} \label{cases_study}
 
@@
 
-%\subsubsection{Mixed Sensitivity}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%using different $k$ for $k$-CFA
+\subsection{Modular Analysis for Free}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% \subsubsection{Mixed Sensitivity}
+% using different k?
+
+One of the challenges of modern static analysis is program usually depends on
+large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. Can we analyze
+programs and libraries separately and reuse the result without losing precision?
+Then we can reduce part of the overhead of repeatedly analyzing libraries for
+different programs. Indeed, some static analyzers compute summary for a function
+or a module, which can be reused later, however they are mostly too conservative
+or unsound, which both lead to imprecision.
+
+The specialization of abstract interpreter provides a chance to obtain such
+partial analysis result in a mechanized way, but still keeps the analysis sound.
+As we see when compiling the closures, we can specialize the abstract
+interpreter with respect the body expression of the lambda term without knowing
+the actual argument. The programs with some unknown variables are open programs,
+which is exactly the case if we want to analyze programs in a modular way.
+
+For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a
+whole-program analysis, because it is inter-procedural and needs the last $k$
+calling contexts to distinguish different call sites, where the calling contexts
+are dynamic values. However, it is possible to analyze programs (libraries)
+separately through specializing an abstract interpreter that generates the
+analysis as the next-stage program and leave the unavailable programs and
+calling contexts as dynamic parameters, and then install these contexts when we
+have the whole program.
+
+\iffalse
+Another perspective: programs are data for an abstract interpreter, so if we
+have $n$ programs, then maybe there can be $n$ stages. Probably we can analyze
+first $m$ programs, and generate a residual abstract interpreter waiting for the
+rest $(n-m)$ programs. These $(n-m)$ programs might be (abstract) arguments for
+the first $n$ programs, and the abstract interpreter itself might be a partial
+abstract interpreter.
+\fi
 

### 2019-03-26 paper/20_prelim.tex
section 4

@@
   def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
-  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
+  def ap_clo(ev: Expr => Ans)(rator: R[Value], rand: R[Value]): Ans
 \end{lstlisting}
@@
 \begin{lstlisting}
-  def ap_clo(ev: EvalFun)(f: Value, arg: Value): Ans = f match {
+  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = rator match {
     case CloV(Lam(x, e), ρ: Env) => for {
       α <- alloc(x)
-      _ <- set_store(α → arg)
+      _ <- set_store(α → rand)
       rt <- ext_env(ev(e))(x → α)

### 2019-03-26 paper/30_stageconc.tex
section 4

@@
 
-Readers may notice that the conversion between unstaged monads and monads is 
-simply changing the unstaged data to staged data, which in fact can be achieved without
-changing the implementation of monads. This is true so far, but as we will see
-it is not that easy when introducing the nondeterminism monad (@ListT@) for
-abstract interpreters, as the whole list will be a next-stage value instead of
-merely the elements in the list.
-For this reason, we explicitly distinguish and present the two version of
-monads, as a more consistent approach.
+Readers may notice that the conversion between unstaged monads and staged monads
+is simply changing the unstaged data to staged data, which in fact can be
+achieved without modifying the implementation of monads. This is true so far, but
+as we will see, it is not that easy when introducing the nondeterminism monad
+(@ListT@) for abstract interpreters, as the whole list will be a next-stage
+value instead of merely the elements in the list. For this reason, we explicitly
+distinguish and present the two versions of monads, as a more consistent
+approach.
 
 \paragraph{Primitive Operations} Most of the primitive operations can be easily
-translated to staged version -- we just need to let them works on staged data.
+translated to their staged versions -- we just need to let them works on staged data.
 As we mentioned before, we elaborate how to handle functions and applications in
 detail. One of the desired goal is to eliminate the monadic layers through
-staging, this becomes tricky for clsoures, because we need to actually compile
-the lambda term instead of generating a @CloV@ value. To make this happen, when
-closing the lambda term with the environment, we need to staging into the body
-expression by \textit{collpasing} the @AnsM@ monads to normal values.
-The following function @close@ illustrates the idea:
-@f@ is a current stage function but takes two next-stage values @v@ and
-$\sigma$, inside of which, we eagerly collapse the monads to values by
-@ev(e)(ρ_*)(σ_*)@, which has type @Rep[(Value, Store)]@.
-Then we generate an intermediate representation of the compiled closure using
-@emit_compiled_clo@. The intermediate representation stands for a next-stage
-value, and contains the compiled function, which will be generated as a
-next-stage function.
+staging, this becomes a little bit tricky for clsoures, because we need actually
+compiling the lambda term instead of generating a @CloV@ value, but the @ev(e)@
+returns a monad and prevents the specialization. To make this happen, when
+closing a lambda term, we need to staging into the body expression by
+\textit{collpasing} the @AnsM@ monads to normal values. The
+following function @close@ illustrates the idea: @f@ is a current stage function
+but takes two next-stage values @v@ and $\sigma$, inside of which, we eagerly
+collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
+@Rep[(Value, Store)]@. Then we generate an intermediate representation of the compiled
+closure using @emit_compiled_clo@. The intermediate representation stands for a
+next-stage value, and contains the compiled function, which will be generated as
+a next-stage function.
 

### 2019-03-26 paper/40_absinterp.tex
section 4

@@
 focus on abstract interpreters under the same framework. We will first present a
-lattice framework with binding-time annotations, but for the purpose of clearly
-illustrating the idea, we just use simple abstract domains such as power set
-lattice or point-wise map lattice. Then we show the abstract components such as
-abstract environments and abstract stores.
+lattice framework with binding-time abstraction, but for the purpose of clearly
+illustrating the idea, we will just use simple abstract domains such as power set
+lattice or point-wise map lattice. Then we show the abstract components, namely
+abstract values, abstract environments and abstract stores.
 The abstract interpreter we present in this section is similar to
@@
 context/path/flow-insensitive by our choice.  In Section \ref{cfa}, we will
-instantiate a context-sensitive analysis;
+instantiate a commonly-used context-sensitive analysis;
 \citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
@@
 operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
-operator. The trait @SPLattice@ defines the type class of stage polymorphic
-lattices. Similar to @MonadOps@ in Section \ref{generic_if}, we introduce an
-additional higher-kinded type @R[_]@ to annotate the binding-times.
+operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines the type
+class of stage polymorphic lattices. Similar to @MonadOps@ in Section
+\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
+annotate the binding-times, where @R[_]@ wraps @S@ and @Boolean@ in these
+operations.
 
-\begin{lstlisting}
+\begin{figure}[h!]
+  \centering
+  \begin{subfigure}[b]{0.45\textwidth}
+  \begin{lstlisting}[style=small]
 trait SPLattice[S, R[_]] {
@@
   \end{lstlisting}
-
-In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
-unstaged lattice type class. An simple instance of lattice we used is the power
-set abstract domain, where the ordering is subset relation, the join is union
-operation, and the meet is intersect operation. The power set lattice can be
-implemented as follows:
-
-\begin{lstlisting}
+  \caption{trait \texttt{SPLattice} and \texttt{Lattice}} \label{fig:splattice}
+  \end{subfigure}
+  ~
+  \begin{subfigure}[b]{0.6\textwidth}
+\begin{lstlisting}[style=small]
 def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
-    lazy val ⊤: Set[T] = Set[T]()
-    lazy val ⊥: Set[T] = throw new NotImplementedError("No representation for top")
+  lazy val ⊥: Set[T] = Set[T]()
+  lazy val ⊤: Set[T] = error("No representation for ⊤")
   def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
@@
 \end{lstlisting}
+  \caption{The power set instance for lattice} \label{fig:powerset}
+\end{subfigure}
+\end{figure}
 
+In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
+unstaged lattice type class. An simple instance of lattice we used is the power
+set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
+operation, and the meet is intersect operation.
 Other lattices used in the rest of the paper, such as products and maps, can be
-implemented similarly, or lifted from existing lattices element-wise or point-wise.
+implemented similarly, or lifted from existing lattices element-wise or
+point-wise.
 
@@
 It is importatnt that the abstract store now maps addresses to sets of abstract
-values, meaning an address may points to an over-approximated set of values
+values, meaning an address may point to an over-approximated set of values
 occurred at runtime.
@@
 To prevent the abstract interpreter from falling into nontermination, the ADI
-approach uses a two-cache mechanism. Here we first provides the type
+approach uses a two-cache mechanism. Here we first provide the type
 definitions, and later we will describe the algorithm in detail. A @Cache@ is a
@@
 Therefore, when applying a function, for instance @f(a)@, we may retrieve a set
-of closures by dereferencing the address of @f@. We need to apply all of the
-target closures, where the nondeterminism is introduced. Another nondeterminism
-happens when evaluating conditionals, we have to evaluate both branches if the
-condition can not be determined. In other words, we will have multiple computing
-results.
+of closures by dereferencing the address of @f@. We need to nondeterministically
+apply all of the target closures.
+Another nondeterminism happens when evaluating conditionals, we have to evaluate
+both branches if the condition can not be determined. In other words, we will
+have multiple computing results.
 
@@
 already have a reader effect for environments and a state effect for stores. We
-add three more effects based on it: the nondeterminism effect is represented by
-the @ListT[M[_], A]@ monad transformer, inside of which, we have a @ReaderT@ for
-a @in@ cache, and a @StateT@ for an @out@ cache.
-\todo{highlight where changed}
+further add three more effects based on it: the nondeterminism effect is
+represented by the @ListT[M[_], A]@ monad transformer, inside of which, we have
+a @ReaderT@ for a @in@ cache, and a @StateT@ for an @out@ cache. \todo{highlight
+where changed}
 
@@
 
-The sketch implementation of @ListT@ is shown below. The value @run@ it encapsulates
-is of type @M[List[A]]@, where @M[_]@ is the inner monad, and @A@ the element
-type of the list.
+The sketch implementation of @ListT@ transformer is shown below. The value @run@
+it encapsulates is of type @M[List[A]]@, where @M[_]@ is the inner monad, and
+@A@ the element type of the list. The @flatMap@ method first unwraps the inner
+monadic value @M[List[A]]@ and obtain a value of @List[A]@, then applies @f@ to
+each element in that list, and finally folds the results back to a single
+@List[M, B]@.
 
 \begin{lstlisting}
-  case class ListT[M[_]: Monad, A](run: M[List[A]]) { ... }
+  case class ListT[M[_]: Monad, A](run: M[List[A]]) {
+    def flatMap[B](f: A => ListT[M, B]): ListT[M, B] = ...
+  }
 \end{lstlisting}
 
-\paragraph{Primitive Operations} \todo{wip}
-
-The operations @get@ and @put@ on abstract stores are slightly changed according
-to the abstract semantics: @get@ uses the bottom element in the @Value@ lattice
-if the queried address does not exist; @put@ performs the update by joining with
-the existing values. function @alloc@ simply invokes the @Addr@ constructor.
-\texttt{close} and \texttt{num} lift syntactic literals to our abstract domain,
-i.e., a singleton set that contains only one \texttt{CloV} or \texttt{NumV}
-object.
+\paragraph{Primitive Operations} Some of the primitive operations are changed
+according to the new monad stack. A notable change is that the store update
+operation joins the value with the existing values, for the same address. Under
+abstract interpretation, we may have to explore the results from both branches
+of a conditional expression for soundness. In @br0@, we combine the results
+using @mplus@ from @MonadPlus@, which requires the values are join-semilattices
+(in our case, @List@s and @Map@s).
 
 \begin{lstlisting}
+  def set_store(αv: (Addr, Value)): AnsM[Unit] = liftM(StateTMonad.mod(σ => σ ⊔ Map(αv)))
+  def br0(test: Value, thn: => Ans, els: => Ans): Ans = ReaderTMonadPlus.mplus(thn, els)
 \end{lstlisting}
 
-For branching, @branch0@ joins the answers from two branches since we are doing
-a path-insensitive analysis. For arithmetics, @prim_eval@ returns the top
-abstract number element. It is worth noting that @apply_closure@ is where we
-handle the non-determinism of applications. The non-determinism happens because
-the abstract store maps addresses to sets of abstract values; when dereferencing
-an address from the store, we need to explore all possible values to achieve a
-sound analysis. In the case of @apply_closure@, the first argument @fs@ may
-contain multiple target closures, so we use a @for@ comprehension over all
-closures, apply them respectively and form a set result values @vs@; meanwhile,
-every time when evaluating the body expression @e@ we use the latest store
-$\sigma$@_*@ which is updated iteratively during the loop. Thus this is a
-store-widened analysis.
+The value representation of lambda terms is stil @CloV@, but we lift it to a
+singleton set.
 
-\paragraph{Caching and Fixpoint Iteration} Caching
+\begin{lstlisting}
+  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
+\end{lstlisting}
+
+For function applications, @ap_clo@ loooks roughly the same as the concrete
+interpreter. The difference is now the function @rator@ is a set of closures, so
+we can not directly destruct it by pattern matching. Instead, we use an
+auxiliary function @lift_nd@ that takes a list and lifts the elements into the
+monad stack. Then we can still implement @ap_clo@ in monadic style, where the
+@CloV@ comes from the @ListT@ and the nondeterminism is naturally handled.
+
+\begin{lstlisting}
+  def lift_nd[T](vs: List[T]): AnsM[T]
+  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = for {
+    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](f.toList)
+    α <- alloc(x)
+    _ <- set_store(α → rator)
+    rt <- ext_env(ev(e))(x → α)
+  } yield rt
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\paragraph{Caching and Fixpoint Iteration}
+
+As we mentioned ealier, the ADI approach uses a two-cache mechanism to
+compute the least fixed-point and prevent nontermination, which is also called a
+\textit{co-inductive} caching algorithm or \textit{truncated depth-first
+evaluation} \cite{Rosendahl:AbsIntPL}. It has been widely used in abstract
+interpretation or fixed point computation
+\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
+Rosendahl:AbsIntPL}. Since the @in@ and @out@ cache are wrapped in the monad
+stack, we first defined several methods to obtain or update the @in@ and @out@
+cache through the monad stack.
 
@@
 
-We described the abstract semantics modularly in the last section, however, the
-abstract interpreter may not terminate for some input program. In this section,
-we use a memoization technique to ensure its termination. This technique is also
-widely used under the name of co-inductive caching algorithm
-\cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800} or
-truncated depth-first evaluation \cite{Rosendahl:AbsIntPL}.
+Figure \ref{fig:coind_cache} shows the co-inductive caching algorithm that
+closes the open-recursive @eval@.
+The idea is to first set up two caches called @in@ and @out@ both of type
+@Cache@. Then similar to Kleene's iteration algorithm, we compute the analysis
+result iteratively. The @in@ cache contains what we already have computed from
+the last iteration, and the @out@ cache is what we would have computed after this
+iteration joined with the contents from @in@ cache.
+When starting a new iteration, @in@ is set to be the result of the last
+iteration, i.e., @out@; and @out@ is assigned to be empty.
 
-The idea is to set up two caches called @in@ and @out@, which are both mapping
-from the arguments of the interpreter (@Config@) to the result values of the
-interpreter (@Ans@). The @in@ cache contains what we already have computed from
-the last iteration, the @out@ cache is what we will have computed after this
-iteration joined with the previous one. When starting a new iteration, @in@ is
-set to be the result of the last iteration, i.e., @out@; @out@ is set to be
-empty. In the iteration, we first check whether @out@ contains what we want, if
-yes then it is returned; otherwise, we retrieve what we have from @in@, compute
-the result for this iteration, and put the joined result back into @out@. Note
-that we also instrument the recursive call by putting @cached_ev@ into @evev@.
-After one iteration, if @in == out@, then there is no more information to be
-gained, thus the iteration should end and we have reached the fixed point.
+During the iteration, we first check whether @out@ contains the configuration
+@cfg@, representing the current desired computation, if
+yes then result is directly returned through the monads.
 
-Finally, we override the definition of @eval_top@ by instantiating @CacheFix@
-with @eval@ and starting the first iteration.
+Otherwise, we first retrieve the results from @in@ (@⊥@ if not contained in
+@in@), and put it into the @out@ cache.
+Then we invoke @ev@ to compute the result for this iteration.
+After the evaluation, the store @σ@ may have been changed. So we update the
+@out@ cache with the new result, i.e., @(v, σ)@. Similar to store update, cache
+update is also point-wise join.
+Note that for the recursive call of @ev@, we also instrument it by puting
+@fix(ev)@ as the first argument to @ev@.
+
+The iteration terminates when the resulting @out@ cache is equivalent to the
+input @in@ cache, which means there is no more analysis result can be gained,
+thus the iteration should end and we have reached the least fixed point.
+
+\begin{figure}[t!]
+  \centering
+\begin{lstlisting}
+  def fix(ev: EvalFun => EvalFun): EvalFun = e => for {
+    ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache; val cfg = (e, ρ, σ)
+    rt <- if (out.contains(cfg)) for {
+            (v, s) <- lift_nd[(Value, Store)](out(cfg).toList)
+             _ <- put_store(s)
+          } yield v
+          else for {
+            _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
+            v <- ev(fix(ev))(e)
+            σ <- get_store
+            _ <- update_out_cache(cfg, (v, σ))
+          } yield v
+  } yield rt
+\end{lstlisting}
+\vspace{-1em}
+\caption{The unstaged co-inductive caching algorithm.}
+\label{fig:coind_cache}
+\end{figure}
+
+\begin{lstlisting}
+  def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0)(cache0)(cache0).run
+\end{lstlisting}
+
+Finally, we override the top-level @run@ method by running the monad with
+the empty abstract environment $\rho_0$, initial store $\sigma_0$, and initial
+caches @cache@$_0$.

### 2019-03-26 paper/60_cases.tex
section 4

@@
     \centering
-    \begin{lstlisting}[
-      language=Scala,%
-      mathescape=true,%
-      %  columns=[c]fixed,%
-      %  basewidth={0.5em, 0.40em},%
-      aboveskip=2pt,%\smallskipamount,
-      belowskip=1pt,%\negsmallskipamount,
-      lineskip=-1pt,
-      basewidth={0.6em, 0.45em},%
-      %  backgroundcolor=\color{listingbg},
-      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
-      keywordstyle=\keywordstyle,
-      commentstyle=\commentstyle,
-      stringstyle=\stringstyle,
-    ]
+    \begin{lstlisting}[style=extrasmall]
 type CompAnalysis = Store => Store
@@
     \centering
-    \begin{lstlisting}[
-      language=Scala,%
-      mathescape=true,%
-      %  columns=[c]fixed,%
-      %  basewidth={0.5em, 0.40em},%
-      aboveskip=2pt,%\smallskipamount,
-      belowskip=1pt,%\negsmallskipamount,
-      lineskip=-1pt,
-      basewidth={0.6em, 0.45em},%
-      %  backgroundcolor=\color{listingbg},
-      basicstyle=\fontsize{6}{8}\selectfont\ttfamily,
-      keywordstyle=\keywordstyle,
-      commentstyle=\commentstyle,
-      stringstyle=\stringstyle,
-    ]
+    \begin{lstlisting}[style=extrasmall]
 def analyzeProgram(prog: Expr, $\sigma$: Rep[Store]): Rep[Store] = 

### 2019-03-27 paper/10_intro.tex
mostly section 5

@@
 \cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
-Futamura projections, which first identifies the binding-time of variables in
-the program, i.e., they can be known whether statically or dynamically, then
+Futamura projections: it identifies the binding-time of variables in
+the program (i.e., they can be known statically or dynamically),
 evaluates the static part, and finally generates a residual program that solely
@@
 approach to specialization and partial evaluation, multi-stage programming (MSP
-for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the stage
+for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
 annotations (i.e., binding-time annotations) to be explicit from the
-programmers. The stage annotations can be either syntactic (as in
+programmers. The staging annotations can be either syntactic (as in
 MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \todo{cite}.
@@
 $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
-argument @arg@ to an environment (as well as an accompanying heap object, if for
-the interest of using assignments) that maps free variables in $e$ to some
+argument @arg@ to an environment (as well as an accompanying heap object to
+support mutable variables) that maps free variables in $e$ to some
 values, the interpreter is a standard environment-passing interpreter,
@@
 
-Roughly at the same time when Futamura projections was visioned in 1970s,
+Roughly at the same time when Futamura projections was envisioned in 1970s,
 \citet{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
@@
 interpreter, \textit{can we derive a sound abstract interpreter while it
-produces fast code that does the analysis?}
+produces optimized code that does the analysis?}
 Conceptually, it is similar to the concrete setting: for an
@@
 programming and generative programming, and extensively use abstract type members 
-over multiple abstractions, which can be later instantiated.
+over the following abstractions, which can be later instantiated.
 

### 2019-03-27 paper/30_stageconc.tex
mostly section 5

@@
       case (v: Rep[Value], σ: Rep[Store]) =>
-        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ + (α + v)
+        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ + (α → v)
         ev(e)(ρ_*)(σ_*)

### 2019-03-27 paper/50_stageabs.tex
section 4
mostly section 5

@@
 staged concrete interpreter, now we begin describing the implementation of their
-confluence -- a staged abstract interpreter. We present a principled approach to
-derive staged abstract interpreter from its unstaged version. One guiding
-principle of our approach is that the code of the abstract semantics and the
-code that optimizes should be separated. This it is an advantage of using
-staging for abstract interpreters: the designer of the analysis has no need to
-rewrite the analysis, and the performance improvement comes almost for free,
-without any sacrifice of soundness or precision. Unsurprisingly, the staged
-abstract interpreter we present in this section has the same abstract semantics
-as the unstaged version we presented in Section~\ref{unstaged_abs}.
+confluence -- a staged abstract interpreter.
+Unsurprisingly, the staged abstract interpreter in this section has the same
+abstract semantics with the unstaged version from Section~\ref{unstaged_abs}.
+The approach from unstaged to staged abstract interpreter is modular,
+principled, and does not sacrifice soundness or precision. The designer of the
+analyzer has no need to rewrite the analysis. We first present the staged
+lattices and staged monads, and then the staged version of primitive operations,
+especially @close@, @ap_clo@ and @fix@, at last we discuss several optimizations
+and our design.
 
@@
 
-In Section~\ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
-leave space for staging lattices, now we instantiate the type @R@ to @Rep@ and
-still use powersets as an example to present its staged version.
+In Section \ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
+achieve stage polymorphism. Now we instantiate the type @R@ to @Rep@ and
+still use power sets as an example to present its staged version.
 
 \begin{lstlisting}
-trait RepLattice[A] extends Lattice[A, Rep]
-implicit def RepSetLattice[T:Typ]: RepLattice[Set[T]] = 
-  new RepLattice[Set[T]] {
-    lazy val bot: Rep[Set[T]] = Set[T]()
-    lazy val top: Rep[Set[T]] = throw new NotImplementedError()
-    def $\sqsubseteq$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
-      Rep[Boolean] = l1 subsetOf l2
-    def $\sqcup$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
-      Rep[Set[T]] = l1 union l2
-    def $\sqcap$(l1: Rep[Set[T]], l2: Rep[Set[T]]): 
-      Rep[Set[T]] = l1 intersect l2
+  trait RepLattice[S] extends Lattice[S, Rep]
+  def RepSetLattice[T]: RepLattice[Set[T]] = new RepLattice[Set[T]] {
+    lazy val ⊥: Rep[Set[T]] = Set[T]()
+    lazy val ⊤: Set[T] = error("No representation for ⊤")
+    def ⊑(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Boolean] = l1 subsetOf  l2
+    def ⊔(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Set[T]]  = l1 union     l2
+    def ⊓(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Set[T]]  = l1 intersect l2
   }
@@
 
-The type parameter @T:Typ@ of @RepLattice@ requires that the elements of sets
-can also be staged. Otherwise, without knowing how to stage the elements in the
-set, we can not stage the set either. The methods defined operate on type
-@Rep[Set[T]]@, thus the underlying implementation such as @union@ and
-@intersect@ will be mapped to a node in the IR graph during staging and
-eventually emitted in the generated code. Again, other structures such as maps
-and tuples are implemented in a similar way.
+As we can see, the @RepSetLattice@ is an instance of @RepLattice@, which is also
+an abstraction of the manipulations on the foundamental data structure. For
+staged versions, these operations eventually are delegated to the @subsetOf@,
+@union@ and @intersect@ operation on @Rep[Set[T]]@. We even don't have to change
+the implementation code, but just the types – from @Set[T]@ to @Rep[Set[T]]@. In
+the code generation part, these operations emit their corresponding next-stage
+code. Again, other lattice structures such as products and maps are implemented
+in a similar way.
 
@@
 
-We have seen how to obtain a staged concrete semantics based on types, now we
-take the same approach to obtain a staged abstract semantics. The basic
-operations are largely kept the same as in the unstaged version, except the
-types are changed to @Rep@. Besides, when we update the environment, the
-identifier @x@ is known statically, but the environment map has type
-@Rep[Map[Ident,Addr]]@, so we apply @lift@ to @x@ to turn it as a next-stage
-value.
+Now we implement the staged abstract semantics, where @R@ is instantiated as
+@Rep@, and the abstract components are shared from the unstaged version.
+
+\paragraph{Staged Monads for Abstract Interpretation}
+We use the same moand stack structure as in the unstaged abstract interpreter,
+but replacing \textit{all} the transformers to their staged version. The
+following code shows this change. For the readability, we squeeze the three
+inner transformers into a single monad @RepListReaderStateM[A, B, C]@, where @A@
+and @B@ are both @Cache@ when used in @AnsM@. Now the result type is a pair of
+two staged value: @Rep[List[(Value, Store)]]@ and @Rep[Cache]@.
 
 \begin{lstlisting}
-trait RepAbsInterpOps extends Abstract with LMSOps {
-  type R[+T] = Rep[T]
-  val $\rho$0: Rep[Env] = Map[Ident, Addr]()
-  val $\sigma$0: Rep[Store] = Map[Addr, Value]()
-  def get($\rho$: Rep[Env], x: Ident): Rep[Addr] = $\rho$(x)
-  def put($\rho$: Rep[Env], x: Ident, a: Rep[Addr]): 
-    Rep[Env] = $\rho$ + (lift(x) -> a)
-  def get($\sigma$: Rep[Store], a: Rep[Addr]): Rep[Value] = 
-    $\sigma$.getOrElse(a, RepLattice[Value].bot)
-  def put($\sigma$: Rep[Store], a: Rep[Addr], v: Rep[Value]): 
-    Rep[Store] = $\sigma$ + (a -> RepLattice[Value].$\sqcup$(v, get($\sigma$, a)))
-  def alloc($\sigma$: Rep[Store], x: Ident): Rep[Addr] = Addr(x)
-  def num(i: Lit): Rep[Value] = Set(NumV())
-  def branch0(cnd: Rep[Value], thn: => Ans, els: => Ans): Ans =
-    thn $\sqcup$ els
-  def prim_eval(op: Symbol, 
-                v1: Rep[Value], v2: Rep[Value]): Rep[Value] = 
-    Set(NumV())
+  trait StagedAbstractSemantics extends AbstractComponents {
+    type R[T] = Rep[T]
+    type RepListReaderStateM[A, B, C] = RepListT[RepReaderT[RepStateT[RepIdM, A, ?], B, ?], C]
+    type AnsM[T] = ReaderT[StateT[RepListReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
+    type Result = (Rep[List[(Value, Store)]], Rep[Cache])
     ...
@@
 \end{lstlisting}
+
+Since our monad stack is uniformly using staged data, so the staged @ListT@ now
+stores a staged value of type @Rep[List[A]]@, inside of a staged monad @M@.
+\todo{staged ListT implementation}
+
+\begin{lstlisting}
+  case class ListT[M[_]: RepMonad, A](run: M[List[A]]) {
+    def flatMap[B: Manifest](f: Rep[A] => ListT[M, B]): ListT[M, B] = ...
+  }
+\end{lstlisting}
+
+\paragraph{Primitive Operations}
 
@@
 \begin{lstlisting}
-def close(ev: EvalFun)($\lambda$: Lam, $\rho$: Rep[Env]): Rep[Value] = {
-  val Lam(x, e) = $\lambda$
-  val f: Rep[(Value, Store)]=>Rep[(Value,Store)] = {
-    case (args: Rep[Value], $\sigma$: Rep[Store]) =>
-      val args = as._1; val $\sigma$ = as._2; val $\alpha$ = alloc($\sigma$, x)
-      ev(e, put($\rho$, x, $\alpha$), put($\sigma$, $\alpha$, args))
+  def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
+    val Lam(x, e) = λ
+    val f: (Rep[Value],Rep[Store],Rep[Cache],Rep[Cache]) => Rep[(List[(Value,Store)],Cache)] = {
+      case (arg, σ, in, out) =>
+        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ ⊔ Map(α → arg)
+        ev(e)(ρ_*)(σ_*)(in)(out)
     }
-  Set[AbsValue](CompiledClo(fun(f)))
-}
-def apply_closure(ev: EvalFun)
-  (f: Rep[Value], arg: Rep[Value], $\sigma$: Rep[Store]): Ans = {
-    reflectEffect(ApplyClosure(f, arg, $\sigma$))
+    Set[AbsValue](emit_compiled_clo(f, λ, ρ))
   }
@@
 
-When generating code for an application, we can not directly apply the callee.
-Instead, we emit code that calls a next-stage function @apply_closures_norep@.
-As its unstaged counterpart, function @apply_closures_norep@
-non-deterministically applies multiple target closures with the argument and
-latest store, and finally returns the joined value and a single store. We
-provide the definition of @apply_closures_norep@ in the runtime supporting code.
-
 \begin{lstlisting}
-case ApplyClosures(fs, arg, $\sigma$) =>
-  emitValDef(sym, "apply_closures_norep(" + 
-                  quote(fs) + "," + quote(arg) + 
-                  "," + quote($\sigma$) + ")")
+  def lift_nd[T](vs: Rep[List[T]]): AnsM[T]
+  def ap_clo(ev: EvalFun)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
+    clo <- lift_nd[AbsValue](rator.toList)
+    σ <- get_store; in <- ask_in_cache; out <- get_out_cache
+    val res: Rep[(List[(Value, Store)], Cache)] = emit_ap_clo(clo, rand, σ, in, out)
+    _ <- put_out_cache(res._2)
+    vs <- lift_nd[(Value, Store)](res._1)
+    _ <- put_store(vs._2)
+  } yield vs._1
 \end{lstlisting}
 
-\subsection{Staged Fixpoint Iteration}
+\paragraph{Staged Caching and Fixpoint Iteration}
 
@@
 
+\begin{figure}[t!]
+  \centering
 \begin{lstlisting}
-def iter(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
-Rep[(Value,Store)] = {
-  def iter_aux: Rep[Unit => (Value,Store)] = fun { () =>
-    in = out; out = Map[Config, (Value,Store)]()
-    cached_ev(e, $\rho$, $\sigma$)
-    if (in === out) out((unit(e), $\rho$, $\sigma$)) else iter_aux()
-  }
-  iter_aux() // generated code that invokes iter_aux()
-}
+  def fix(ev: EvalFun => EvalFun): EvalFun = e => for {
+    ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache
+    val cfg: Rep[Config] = (unit(e), ρ, σ)
+    val res: Rep[(List[(Value, Store)], Cache)] =
+      if (out.contains(cfg)) (out(cfg).toList, out) // a next-stage if expression
+      else { val m: Ans = for {
+               _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
+               v <- ev(fix(ev))(e)
+               σ <- get_store
+               _ <- update_out_cache(cfg, (v, σ))
+             } yield v
+             m(ρ)(σ)(in)(out) }
+    _ <- put_out_cache(res._2); vs <- lift_nd(res._1); _ <- put_store(vs._2)
+  } yield vs._1
 \end{lstlisting}
+\vspace{-1em}
+\caption{The staged co-inductive caching algorithm.}
+\label{fig:staged_coind_cache}
+\end{figure}
 
@@
 \begin{lstlisting}
-def cached_ev(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
-Rep[(Value, Store)] = {
-  val cfg: Rep[Config] = (unit(e), $\rho$, $\sigma$)
-  if (out.contains(cfg)) { out(cfg) }
-  else {
-    val ans0 = in.getOrElse(cfg, RepLattice[(Value, Store)].bot)
-    out = out + (cfg -> ans0)
-    val ans1 = evev(cached_ev)(e, $\rho$, $\sigma$)
-    out = out + (cfg -> (ans0 $\sqcup$ ans1)); ans1
-  }
-}
 \end{lstlisting}
+
+\paragraph{Code Generation} Code generation
 
@@
 
-Our staging schema works well in theory, but would suffer from code explosion,
-and possibly runtime GC overhead when analyzing (specializing) large programs.
-We present several optimizations that largely mitigate these issues. Still,
-implementing these optimizations do not need to change the generic interpreter.
+Our staging schema works well in theory, but in practice would suffer from code
+explosion, and possibly runtime GC overhead when analyzing (specializing) large
+programs. In this section, we present several optimizations that largely
+mitigate these issues. Still, implementing these optimizations do not need to
+change the generic interpreter.
 
-\paragraph{Specialized Data Structures}
+\paragraph{Specialized Data Structures} \todo{need revise}
 
@@
 
-\paragraph{Selective Caching} It is observed that the two-fold caching is
-wrapped for recursive call of our abstract interpreter. But this is even more
-expensive when generating code for atomic expressions such as literal numbers or
-variables. \todo{caching prevents nontermination}
+\paragraph{Selective Caching} It is observed that the two-fold co-inductive
+caching is used for every recursive call of our abstract interpreter. But this
+is not necessary and even very expensive when generating code for atomic
+expressions such as literals or variables, as they always terminate. Borrowing
+the idea from the partition in ANF \cite{Flanagan:1993:ECC:155090.155113}, we
+may use a selective caching algorithm:
+
+\begin{lstlisting}
+  def fix_select: EvalFun = e => e match {
+    case Lit(_) | Var(_) | Lam(_, _) => eval(fix_select)(e)
+    case _ => fix_cache(e)
+  }
+\end{lstlisting}
 
@@
 
-For example, to fold a singleton list, e.g., @List(x).foldLeft(init)(f)@ where
-@x@ and @init@ are staged values, a naive code generator would faithfully
-apply @foldLeft@ to the list with an anonymous function. But we can also utilize
-the algebraic property of @foldLeft@ to generate cheaper code:
-@List(x).foldLeft(init)(f) = f(init, x)@.
-Since the function @f@ is known at current stage, we completely eliminate the
-fold operation and function application. We apply several rewritings enabled by
-partial-static data structures, such as @List@ and @Map@, which greatly reduce
+Our treatment to binding-times is coarse-grained: @Expr@s is static, the rest of
+the world are all dynamic. But this is not always true, since the static data
+has to be used somewhere with the dyanmic operation. Partial-static data is a
+way to improve binding-times and optimize the generated code.
+
+For example, to fold a singleton list (often appears in @ListT@), e.g.,
+@List(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
+generator would faithfully apply @foldLeft@ to the list with an anonymous
+function. But we can also utilize the algebraic property of @foldLeft@ to
+generate cheaper code: @List(x).foldLeft(init)(f) = f(init, x)@. Since the
+function @f@ is known at current stage, we completely eliminate the fold
+operation and function application. We apply several rewritings enabled by
+partial-static data structures, such as @List@ and @Map@, which greatly reduces
 the size of residual programs.
 
-\paragraph{Heterogeneous Staging} The generated program is in A-Normal form,
+\paragraph{Heterogeneous Staging} The generated program is in A-Normal form. \todo{TODO}
 
@@
 parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
-specialization can be done by unfolding, as well as the termination of
-specialization procedure. It is also possible to specialize small-step
+specialization can be done by unfolding, as well as guarantees the termination
+of specialization procedure. It is also possible to specialize small-step
 operational abstract semantics through abstract compilation
-\cite{Boucher:1996:ACN:647473.727587} --
+\cite{Boucher:1996:ACN:647473.727587} -- as
 \citet{Johnson:2013:OAA:2500365.2500604} implemented it for

### 2019-03-27 paper/40_absinterp.tex
section 5

@@
   def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = for {
-    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](f.toList)
+    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](rator.toList)
     α <- alloc(x)

### 2019-03-27 paper/50_stageabs.tex
section 5

@@
 
-\paragraph{Primitive Operations}
-
-Once more, the way we handle closures is the same as in the staged concrete
-interpreter: the recursive call to @ev@ with the body expression @e@ is compiled
-and specialized, the wrapper function @f@ will be a field value in a
-@CompiledClo@ object and be generated for the next stage. At the end, we return
-a singleton set:
+\paragraph{Primitive Operations} We keep our road map that focuses on functions
+and applications. The @close@ method now is a rough mixing of the staged
+concrete and unstaged abstract version: we have a current-stage function @f@,
+that takes four next-stage values, @in@ and @out@ additionally; inside of @f@,
+we collapse the @Ans@ monad to values of type @Result@ by providing the desired
+arguments, i.e., the new environment, new store, and caches. The collapse
+happens at current-stage, so the call of @ev@ is unfolded after staging. At
+last, we generate a singleton set containing the compiled closure
+@emit_compiled_clo(f)@, represented by an IR node in LMS.
 
 \begin{lstlisting}
+  def emit_compiled_clo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
+                           => Rep[(List[(Value,Store)], Cache)]): Rep[AbsValue]
   def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
@@
         ev(e)(ρ_*)(σ_*)(in)(out)
-    }
-    Set[AbsValue](emit_compiled_clo(f, λ, ρ))
+    }; Set[AbsValue](emit_compiled_clo(f))
   }
@@
 
+The @ap_clo@ method is also similar: we use the staged version of @lift_nd@ to
+lift the set of closures into the monad stack, and generate a next-stage value
+using @emit_ap_clo@, which represets the result of application, and finally we
+put the @out@ cache, store and values back into the monads. The @emit_ap_clo@
+method takes the target closure @clo@, argument @rand@, store and caches.
+
 \begin{lstlisting}
   def lift_nd[T](vs: Rep[List[T]]): AnsM[T]
+  def emit_ap_clo(rator: Rep[AbsValue], rand: Rep[Value], σ: Rep[Store],
+                  in: Rep[Cache], out: Rep[Cache]): Rep[(List[(Value, Store)], Cache)]
   def ap_clo(ev: EvalFun)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
@@
 
-Our fixed-point iteration again relies on two caches @in@ and @out@, but the
-iteration no longer be done at the current stage. Because the @in@ and @out@ are
-both next-stage values, the test of whether @in@ and @out@ are equal is a
-generated expression in the next stage, and we can only know the comparison
-result at the next stage. In other words, we do not know how many iterations we
-need to reach the fixed-point. To achieve this, we need to stage a function
-value --- @iter_aux@ is generated as a recursive function of type @Rep[Unit => (Value,Store)]@ 
-and will be invoked at the next stage.
+Our fixed-point iteration again relies on two caches @in@ and @out@, which are
+both staged values now. Therefore, the query of whether the @out@ cache contains
+the current configuration produces a next-stage Boolean value, i.e.,
+@Rep[Boolean]@. Consequently, the branching can not be determined statically --
+we need to generate code for @if@. Figure \ref{fig:staged_coind_cache} shows the
+staged version of @fix@.
 
-\begin{figure}[t!]
+\vspace{-1em}
+\begin{figure}[h!]
   \centering
@@
 \end{figure}
+\vspace{-1em}
 
-However, the instrumented evaluation function that uses the @in@ cache and
-updates the @out@ cache can be completely eliminated by staging. Each recursive
-call to @cached_ev@ will also be specialized if it is applied on subexpressions
-of the analyzed program.
+The @res@ variable is a next-stage result, which consists of a next-stage @if@
+expression. The true branch simply returns a pair of queried result form the
+@out@ cache and the @out@ cache. The else branch constructs a monad @m@ first,
+which evaluates @e@ under the new @out@ cache. After which, we collapse the
+monad @m@ with desired arguments. Finally, we obtain a placeholder @res@ since
+it is a next-stage value, and put the content of @res@ back into the moand
+stack.
+
+\paragraph{Code Generation} The IR nodes for compiled closures and
+applications are defined as following:
 
 \begin{lstlisting}
+  case class IRCompiledClo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
+                           => Rep[(List[(Value, Store)], Cache)]) extends Rep[AbsValue]
+  case class IRApClo(clo: Rep[AbsValue], arg: Rep[Value], σ: Rep[Store],
+                     in: Rep[Cache], out: Rep[Cache]) extends Rep[(List[(Value, Store)], Cache)]
 \end{lstlisting}
 
-\paragraph{Code Generation} Code generation
+The code generator for these two IR nodes is similar to staged concrete
+interpreter, which generates a next-stage @CompiledClo@ object and a next-stage
+function call to @f@, respectively.
 
@@
 Our staging schema works well in theory, but in practice would suffer from code
-explosion, and possibly runtime GC overhead when analyzing (specializing) large
-programs. In this section, we present several optimizations that largely
-mitigate these issues. Still, implementing these optimizations do not need to
-change the generic interpreter.
+explosion, and possibly runtime GC overhead (at the next stage) when analyzing
+(specializing) large programs. In this section, we present several optimizations
+that largely mitigate these issues. Still, implementing these optimizations do
+not need to change the generic interpreter.
 
-\paragraph{Specialized Data Structures} \todo{need revise}
+\paragraph{Specialized Data Structures}
 
-Now we have already obtained an end-to-end staged abstract interpreter that is
-able to specialize an analysis. However, we treat the data structures such as
-@Map@s as black-boxes, which means any operations on a @Map@ become code in the
-next stage. But, as we identified when introducing the generic interface, the
-keys of any environment maps are identifiers in the program, which are
-completely known statically. This leaves us a chance to further specialize the
-data structures. Assume the @Map[K,V]@ is implemented as a hash map, if the keys
-$K$ are known, then the indices can be computed statically. Thus the specialized
-map would be an array @Array[Rep[V]]@ whose elements are next-stage values; all
-the accesses to the array is determined during staging.
+Although every components except @Expr@s are staged, we treat the data
+structures such as @Map@s as black-boxes, which means the operations on a @Map@
+directly become code in the next stage, but we don't inspect any further inside.
+As we identified when introducing the generic interface, the keys of @Env@s are
+identifiers (i.e., strings) in the program, which are completely known
+statically. This leaves us a chance to further specialize the data structures.
+For example, let's assume that the @Map[K, V]@ is implemented as a hash map. If
+the keys of type @K@ are all known statically, then the indices also can be
+computed statically. Thus the specialized map would be an array of type
+@Array[Rep[V]]@, whose elements are next-stage values; the size of the array is
+also known statically. An access to the map is translated to an access to the
+array with pre-determined index during staging.
 
 Particularly, if we are specializing a monovariant analysis, the address space
-is equivalent to the identifiers, then the environment component can be entirely
-eliminated, and the store is a specialized map as array of @Rep[Value]@
+is equivalent to the set of identifiers in the program. Then the accesses to the
+environment can be completely determined statically, and generates the addresses
+directly. After which, the stores can be specialized as arrays of @Rep[Value]@
 elements.
@@
 
-\paragraph{Heterogeneous Staging} The generated program is in A-Normal form. \todo{TODO}
+%\paragraph{Heterogeneous Staging} The generated program is in A-Normal form. \todo{TODO}
 
@@
 achieve this and examine some alternatives.
+
+\paragraph{What the generated code look like?} In the generated code, all the
+primtive operations (such as @eval@, @fix@, @ap_clo@, etc.) and moandic
+operations (such as @flatMap@ and @map@) are eliminated. The residual program
+consists of statements and expressions that purely manipulate the environment,
+store, and two caches, whose underlying representions are all @Rep[Map[K,V]]@. We
+also have several operations on tuples and lists, which are residualized from
+the use of monads internally.
 
@@
 
-% direct-style vs CPS
-
 \paragraph{Correctness and Soundness}

### 2019-03-27 paper/50_stageabs.tex
misc

@@
 As we can see, the @RepSetLattice@ is an instance of @RepLattice@, which is also
-an abstraction of the manipulations on the foundamental data structure. For
-staged versions, these operations eventually are delegated to the @subsetOf@,
-@union@ and @intersect@ operation on @Rep[Set[T]]@. We even don't have to change
-the implementation code, but just the types – from @Set[T]@ to @Rep[Set[T]]@. In
-the code generation part, these operations emit their corresponding next-stage
-code. Again, other lattice structures such as products and maps are implemented
-in a similar way.
+an abstraction of the underlying data structure. The lattice operations
+eventually are delegated to @subsetOf@, @union@ and @intersect@ on
+@Rep[Set[T]]@. We even don't have to change the implementation code, but just
+the types – from @Set[T]@ to @Rep[Set[T]]@. LMS library provides an implicit
+conversion lifting to next-stage values. In the code generation part, these
+operations emit their corresponding next-stage code. Again, other lattice
+structures such as products and maps are implemented in a similar way.
 

### 2019-03-27 paper/20_prelim.tex
mostly section 5
few typos

@@
 We consider a call-by-value $\lambda$-calculus in direct-style, extended
-with numbers, arithmetics, recursions and conditionals. Other effectful features
+with numbers, arithmetic, recursions and conditionals. Other effectful features
 such as assignments can also be supported readily. In Section~\ref{cases_imp}, we
@@
   case class Rec(x: String, rhs: Expr, e: Expr) extends Expr  // recursions
-  case class Aop(op: String, e1: Expr, e2: Expr) extends Expr // arithmetics
+  case class Aop(op: String, e1: Expr, e2: Expr) extends Expr // arithmetic
 \end{lstlisting}
@@
 The method @get@ simply retrieves the value mapped from the identifier @x@ in
-the environment and store. Conditionals and arithmetics is handled by @br0@
+the environment and store. Conditionals and arithmetic is handled by @br0@
 and @arith@, respectively. The methods @ap_clo@ takes a function value and an
@@
 @ReaderT@ and @StateT@.
-In other word, we instantiate the @AnsM@ moand as a stack of @ReaderT@ and @StateT@
+In other word, we instantiate the @AnsM@ monad as a stack of @ReaderT@ and @StateT@
 transformers\footnote{The question mark syntax is a kind projector
@@
 \begin{lstlisting}
-  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = CloV(λ, ρ)
+  def close(ev: Expr => Ans)(λ: Lam, ρ: Env): Value = CloV(λ, ρ)
 \end{lstlisting}
@@
 \begin{lstlisting}
-  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = rator match {
+  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = rator match {
     case CloV(Lam(x, e), ρ: Env) => for {
@@
 
-With other primitive operations implementated, we can implement the top-level
+With other primitive operations implemented, we can implement the top-level
 @run@ method, where $\rho_0$ and $\sigma_0$ are initial empty environment and

### 2019-03-27 paper/30_stageconc.tex
few typos

@@
   def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[(Value, Store)]): Rep[Value]
-  def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
+  def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(x, e) = λ
@@
   def emit_ap_clo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]): Rep[(Value, Store)]
-  def ap_clo(ev: EvalFun)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
+  def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
     σ <- get_store

### 2019-03-27 paper/40_absinterp.tex
few typos

--- 2019-03-26 paper/40_absinterp.tex
@@
 
-After seeing both unstaged and staged concrete interpreters, now we trun our
+After seeing both unstaged and staged concrete interpreters, now we turn our
 focus on abstract interpreters under the same framework. We will first present a
@@
 @Value@).
-The address space is constrained to be finited to ensure the computatbility of
+The address space is constrained to be finite to ensure the computatbility of
 analysis -- for the simplest monovariant analysis, we use variable names for
-addresses. After setteling the @Value@ and @Addr@ type, the environments and
+addresses. After settling the @Value@ and @Addr@ type, the environments and
 stores are automatically lifted to abstract environments and abstract stores,
 i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
-It is importatnt that the abstract store now maps addresses to sets of abstract
+It is important that the abstract store now maps addresses to sets of abstract
 values, meaning an address may point to an over-approximated set of values
@@
 
-To prevent the abstract interpreter from falling into nontermination, the ADI
+To prevent the abstract interpreter from falling into non-termination, the ADI
 approach uses a two-cache mechanism. Here we first provide the type
@@
 
-This oberservation leads us to the type definition of @Result@: in concrete setting,
+This observation leads us to the type definition of @Result@: in concrete setting,
 the @Result@ is a pair of @Value@ and @Store@, now we use a list of such pairs
@@
 
-The value representation of lambda terms is stil @CloV@, but we lift it to a
+The value representation of lambda terms is still @CloV@, but we lift it to a
 singleton set.
@@
 \begin{lstlisting}
-  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
+  def close(ev: Expr => Ans)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
 \end{lstlisting}
 
-For function applications, @ap_clo@ loooks roughly the same as the concrete
+For function applications, @ap_clo@ looks roughly the same as the concrete
 interpreter. The difference is now the function @rator@ is a set of closures, so
@@
   def lift_nd[T](vs: List[T]): AnsM[T]
-  def ap_clo(ev: EvalFun)(rator: Value, rand: Value): Ans = for {
-    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](f.toList)
+  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
+    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](rator.toList)
     α <- alloc(x)
@@
 
-As we mentioned ealier, the ADI approach uses a two-cache mechanism to
-compute the least fixed-point and prevent nontermination, which is also called a
+As we mentioned earlier, the ADI approach uses a two-cache mechanism to
+compute the least fixed-point and prevent non-termination, which is also called a
 \textit{co-inductive} caching algorithm or \textit{truncated depth-first
@@
 update is also point-wise join.
-Note that for the recursive call of @ev@, we also instrument it by puting
+Note that for the recursive call of @ev@, we also instrument it by putting
 @fix(ev)@ as the first argument to @ev@.
@@
 \begin{lstlisting}
-  def fix(ev: EvalFun => EvalFun): EvalFun = e => for {
+  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans = e => for {
     ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache; val cfg = (e, ρ, σ)

### 2019-03-27 paper/50_stageabs.tex
few typos

@@
 \paragraph{Staged Monads for Abstract Interpretation}
-We use the same moand stack structure as in the unstaged abstract interpreter,
+We use the same monad stack structure as in the unstaged abstract interpreter,
 but replacing \textit{all} the transformers to their staged version. The
@@
     type RepListReaderStateM[A, B, C] = RepListT[RepReaderT[RepStateT[RepIdM, A, ?], B, ?], C]
-    type AnsM[T] = ReaderT[StateT[RepListReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
+    type AnsM[T] = RepReaderT[RepStateT[RepListReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
     type Result = (Rep[List[(Value, Store)]], Rep[Cache])
@@
                            => Rep[(List[(Value,Store)], Cache)]): Rep[AbsValue]
-  def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
+  def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(x, e) = λ
@@
 lift the set of closures into the monad stack, and generate a next-stage value
-using @emit_ap_clo@, which represets the result of application, and finally we
+using @emit_ap_clo@, which represents the result of application, and finally we
 put the @out@ cache, store and values back into the monads. The @emit_ap_clo@
@@
                   in: Rep[Cache], out: Rep[Cache]): Rep[(List[(Value, Store)], Cache)]
-  def ap_clo(ev: EvalFun)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
+  def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
     clo <- lift_nd[AbsValue](rator.toList)
@@
 \begin{lstlisting}
-  def fix(ev: EvalFun => EvalFun): EvalFun = e => for {
+  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans = e => for {
     ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache
@@
 monad @m@ with desired arguments. Finally, we obtain a placeholder @res@ since
-it is a next-stage value, and put the content of @res@ back into the moand
+it is a next-stage value, and put the content of @res@ back into the monad
 stack.
@@
 \begin{lstlisting}
-  def fix_select: EvalFun = e => e match {
+  def fix_select: Expr => Ans = e => e match {
     case Lit(_) | Var(_) | Lam(_, _) => eval(fix_select)(e)
@@
 the world are all dynamic. But this is not always true, since the static data
-has to be used somewhere with the dyanmic operation. Partial-static data is a
+has to be used somewhere with the dynamic operation. Partial-static data is a
 way to improve binding-times and optimize the generated code.
@@
 \paragraph{What the generated code look like?} In the generated code, all the
-primtive operations (such as @eval@, @fix@, @ap_clo@, etc.) and moandic
+primitive operations (such as @eval@, @fix@, @ap_clo@, etc.) and monadic
 operations (such as @flatMap@ and @map@) are eliminated. The residual program
 consists of statements and expressions that purely manipulate the environment,
-store, and two caches, whose underlying representions are all @Rep[Map[K,V]]@. We
+store, and two caches, whose underlying representations are all @Rep[Map[K,V]]@. We
 also have several operations on tuples and lists, which are residualized from
@@
 \citet{Johnson:2013:OAA:2500365.2500604} implemented it for
-optimizating Abstract Abstract Machines. However, the generated abstract
-bytecode still requires another small-step abstract machine to execute, which is
+optimizing Abstract Abstract Machines. However, the generated abstract
+byte-code still requires another small-step abstract machine to execute, which is
 an additional engineering efforts.

### 2019-03-27 paper/10_intro.tex
few typos
misc and reorganize

@@
     \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{sai}) \end{smallmatrix} \\
-    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{prelim}) \end{smallmatrix} &
+    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_conc}) \end{smallmatrix} &
     \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{stagedinterp}) \end{smallmatrix} \\ };
@@
 
-The paper is organized as follows:
+The paper is organized as follows: \todo{restate this part}
 

### 2019-03-27 paper/40_absinterp.tex
misc and reorganize

@@
   def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
+    α <- alloc(x)
     CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](rator.toList)
-    α <- alloc(x)
     _ <- set_store(α → rator)

### 2019-03-27 paper/50_stageabs.tex
misc and reorganize

@@
   def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
-    clo <- lift_nd[AbsValue](rator.toList)
-    σ <- get_store; in <- ask_in_cache; out <- get_out_cache
+    σ <- get_store;     clo <- lift_nd[AbsValue](rator.toList)
+    in <- ask_in_cache; out <- get_out_cache
     val res: Rep[(List[(Value, Store)], Cache)] = emit_ap_clo(clo, rand, σ, in, out)
@@
 
-Our fixed-point iteration again relies on two caches @in@ and @out@, which are
+Our fixed-point iteration again relies on the two caches @in@ and @out@, which are
 both staged values now. Therefore, the query of whether the @out@ cache contains

### 2019-03-27 paper/60_cases.tex
flag
evaluation & paper

@@
 
-In Section~\ref{sai}, we have shown that staging an abstract interpreter is
-feasible and can be systematic after correctly identifying the binding-times,
-despite the fact that the abstract interpreter is intended to be imprecise and
-easy to implement. In this section, we conduct several case studies to show that
-this approach is also useful and widely applicable to different analyses.
+In Section~\ref{sai}, we use a toy language to show that staging an abstract
+interpreter is feasible and can be systematically derived from staged concrete
+interpreters and unstaged abstract interpreter.
+In this section, we conduct several case studies to further show that this
+methodology is practically useful and widely applicable to diverse analyses.
 
-\subsection{Abstract Compilation a la Staging} \label{cs_ac}
+\subsection{Abstract Compilation \`a la Staging} \label{cs_ac}
 
-\citeauthor{Boucher:1996:ACN:647473.727587} introduced abstract compilation (AC)
-as a new implementation technique for abstract interpretation based static
-analysis \cite{Boucher:1996:ACN:647473.727587}. The idea is inspired by partial
-evaluation similar to the present paper -- the program can be known statically,
-the overhead of interpretation can be eliminated. In AC, the compiled analysis
-can be represented by either text or closures (higher-order functions); though
-the closures can be executed immediately, the textual programs need to be
-compiled and loaded first.
+We first revisit a similar technique called \textit{abstract compilation} (AC).
+AC is introduced by \citet{Boucher:1996:ACN:647473.727587} as a new
+implementation technique for abstract interpretation based static analysis.
+The idea is inspired by partial evaluation similar to the present paper --
+the program is known statically, and the interpretive overhead can be
+eliminated. In AC, the compiled analysis can be represented by either text or
+closures (higher-order functions). The minute difference is that the closures
+can be executed immediately, while the textual programs need to be compiled and
+loaded first.
 
-Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} show how to compile a
+Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} showed how to compile a
 monovariant control-flow analysis \cite{Shivers:1991:SSC:115865.115884,
 Shivers:1988:CFA:53990.54007} for continuation-passing style (CPS) programs.
-Since the analyzed program is written in CPS, the analyzer is essentially a
-big-step control-environment abstract interpreter. Closure generation compiles
-the analysis as a closure taking an environment argument. The overhead of
-traversing the abstract syntax tree of input program also has been eliminated.
+Since the analyzed program is in CPS form, the analyzer is a big-step
+control-environment abstract interpreter. The abstract compilation transforms
+the analysis as a closure just taking an environment argument. Therefore the
+overhead of traversing the AST of input program has been eliminated.
 
-In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s
-abstract compilation can be understood and implemented as an instance of staging
-abstract interpreters. We first revisit the original implementation of abstract
-compilation of 0-CFA, and then reproduce their result by simply adding stage
-annotations. The generated program of our approach improves approximately the
-same extent of speed, but without changing a single line of the analyzer program
-(with the use of LMS). However, closure generation requires more engineering
-effort, specifically a whole-program conversion on the analyzer. Moreover, as
-shown in Section~\ref{staged_ds}, our approach is able to not only remove the
-interpretive overhead, but also specialize the data structures used in the
-analysis, for example, the environment that maps variables to sets of lambda.
+In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s AC
+(closure generation as example) can be understood and implemented as an instance
+of staging abstract interpreters. We first revisit the original implementation
+of abstract compilation of 0-CFA, and then reproduce their result by simply
+adding stage annotations.
+The generated program of our approach improves approximately the same extent of
+speed, but without changing a single line of the analyzer program using
+type-based staging annotations. However, AC requires more engineering effort,
+i.e., rewriting the whole analyzer into the closure generation form. Moreover,
+as shown in Section~\ref{staged_ds}, our approach is able to not only remove the
+interpretive overhead, but also enabling several optimizations such specialized
+data structures and equivalent rewritings.
 
@@
 
-\subsubsection{Closure Generation}
+\paragraph{Original AC} \todo{WIP}
 
@@
 
-\subsubsection{Staged 0-CFA}
+\paragraph{AC through Staging}
 

### 2019-03-27 oopsla19-code/src/main/scala/sai/evaluation/parser/AST.scala

@@ -1,4 +1,4 @@
-package sai.direct.large.parser
+package sai.evaluation.parser
 
 /* Author: Yuxuan Chen, Guannan Wei */
 
@@ -25,6 +25,7 @@ case class Bind(x: String, e: Expr) {
   override def toString: String = "Bind(\"" + x + "\" " + e + ")"
   def toSet: Set_! = Set_!(x, e)
 }
+
 case class Let(bds: List[Bind], body: Expr) extends Expr {
   def toApp: App = {
     val args = bds map { case Bind(name, _) => name }
@@ -32,6 +33,7 @@ case class Let(bds: List[Bind], body: Expr) extends Expr {
     App(Lam(args, body), vals)
   }
 }
+
 case class Lrc(bds: List[Bind], body: Expr) extends Expr {
   def toApp: App =
     Let(bds.map { case Bind(x, _) => Bind(x, Void()) },
@@ -79,15 +81,5 @@ object SExpPrinter {
   }
 
   def apply(e: Expr): Unit = println(exprToString(e))
-
-  def main(args: Array[String]) = {
-    assert(SExpPrinter.exprToString(CharLit('a')) == "#\\a")
-
-    assert(SExpPrinter.exprToString(App(Lam(List("x", "y"),
-      App(Var("+"), List(Var("x"), Var("y")))), List(IntLit(1), IntLit(3)))) ==
-      "((lambda (x y) (+ x y)) 1 3)")
-
-    assert(SExpPrinter.exprToString(If(BoolLit(false), Var("a"), Lam(List("t"), Var("t")))) ==
-      "(if #f a (lambda (t) t))")
-  }
 }
+

### 2019-03-27 oopsla19-code/src/main/scala/sai/evaluation/parser/Read.scala

@@ -1,0 +1,13 @@
+package sai.evaluation.parser
+
+trait Read[A] {
+  def read(s: String): Option[A]
+}
+
+object Read {
+  def apply[A](implicit e: Read[A]): Read[A] = e
+
+  implicit class ReadOps(s: String) {
+    def read[A: Read]: Option[A] = Read[A].read(s)
+  }
+}

### 2019-03-27 oopsla19-code/src/main/scala/sai/evaluation/parser/SimpleSchemeTokenParser.scala

@@ -1,0 +1,57 @@
+package sai.evaluation.parser
+
+import scala.util.parsing.combinator._
+
+/** Syntax reference:
+  * TSPL 4th: https://www.scheme.com/tspl4/grammar.html
+  * Boucher and Feeley (1996): http://www.iro.umontreal.ca/~feeley/papers/BoucherFeeleyCC96.pdf
+  */
+
+trait SchemeTokenParser extends RegexParsers {
+  override def skipWhitespace = true
+  override protected val whiteSpace = """(\s|;.*)+""".r
+
+  def LPAREN = "[\\(\\{\\[]".r
+  def RPAREN = "[\\)\\}\\]]".r
+
+  def QUASIQUOTE: Parser[String] = "`" | "'"
+  def QUOTE: Parser[String] = "quote"
+  def UNQUOTE: Parser[String] = ","
+  def SYMBOL: Parser[String] = """[^ \t\n\(\{\[\)\}\]]+""".r
+
+  def LISTLPAREN: Parser[String] = "'("
+  def VECLPAREN: Parser[String] = "#("
+
+  def keyword = LAMBDA | DEF | VOID | BEGIN | SET | IF | IF0 | COND | ELSE | CASE | LET | LETSTAR | LETREC
+  def LAMBDA: Parser[String] = "lambda"
+  def IDENT: Parser[String] = """[a-zA-Z!$%*/:<=>?~_^.+\-@][a-zA-Z0-9!$%*/:<=>?~_^.+\-@]*""".r
+
+  def DEF: Parser[String] = "define\\s".r
+  def VOID: Parser[String] = "void"
+  def BEGIN: Parser[String] = "begin"
+  def SET: Parser[String] = "set!"
+
+  def IF: Parser[String] = "if"
+  def IF0: Parser[String] = "if0"
+  def COND: Parser[String] = "cond"
+  def RARROW: Parser[String] = "=>"
+  def ELSE: Parser[String] = "else"
+  def CASE: Parser[String] = "case"
+
+  def LET: Parser[String] = "let"
+  def REC: Parser[String] = "rec"
+  def LETSTAR: Parser[String] = "let*"
+  def LETREC: Parser[String] = "letrec"
+
+  def TRUE: Parser[Boolean] = ("#t"|"#T") ^^ { _ => true }
+  def FALSE: Parser[Boolean] = ("#f"|"#F") ^^ { _ => false }
[…]

### 2019-03-27 oopsla19-code/src/main/scala/sai/evaluation/utils/TestTrait.scala

@@ -1,0 +1,62 @@
+package sai.evaluation.utils
+
+import java.lang._
+
+trait TestTrait {
+  protected var testNameFilter = ""
+  private var tests = 0
+  private var passed = 0
+  private var failed = 0
+  private var fatal = 0
+  private var omitted = 0
+  def test(name: String)(f: => Unit) = {
+    if (testNameFilter == "" || testNameFilter == name) {
+      val dots = "." * (32 - name.length)
+      print(s"""Running test "${name}"$dots""")
+      tests += 1
+      try {
+        f
+        passed += 1
+        println("\u001B[92m✔\u001B[39m")
+      } catch {
+        case e: AssertionError =>
+          failed += 1
+          println("\u001B[91m✗ Assertion Failed\u001B[39m")
+        case e: Throwable =>
+          fatal += 1
+          println("\u001B[91m✗ Exception Thrown\u001B[39m\u001B[91m")
+          e.printStackTrace()
+          print("\u001B[39m")
+      }
+    } else {
+      omitted += 1
+    }
+  }
+
+  def doNothing(f: => Unit) = Unit
+
+  def testOmit(name: String) =
+    if (name == testNameFilter) {
+      test(name) _
+    } else {
+      omitted += 1
+      doNothing _
+    }
+
+  def runtest(filter: String = "") = {
+    testNameFilter = filter
+    tests = 0
+    passed = 0
+    failed = 0
+    fatal = 0
+    testall()
+    println("Test Summary:")
+    println(s"\u001B[94mTESTS:\u001B[39m $tests")
+    println(s"\u001B[92mPASSED:\u001B[39m $passed")
+    println(s"\u001B[91mFAILED:\u001B[39m $failed")
+    println(s"\u001B[93mFATAL:\u001B[39m $fatal")
+    println(s"\u001B[90mOMITTED:\u001B[39m $omitted")
+  }
+
+  def testall(): Unit
+}

### 2019-03-27 oopsla19-code/src/main/scala/sai/evaluation/utils/Utils.scala

@@ -1,0 +1,26 @@
+package sai.evaluation.utils
+
+object Utils {
+  def time[R](block: => R): (R, Double) = {
+    val t0 = System.nanoTime()
+    val result = block    // call-by-name
+    val t1 = System.nanoTime()
+    val t = (t1 - t0) //1000000.0
+    //println("Elapsed time: " + t + "ms")
+    (result, t)
+  }
+}
+
+case class Timing(ts: List[Double]) {
+  val mean: Double = ts.sum/ts.size
+  val sorted_ts: List[Double] = ts.sorted
+  val ub: Double = sorted_ts.head
+  val lb: Double = sorted_ts.last
+  val perc05 = sorted_ts((sorted_ts.size / 20).toInt)
+  val perc25 = sorted_ts((sorted_ts.size / 4).toInt)
+  val perc50 = sorted_ts((sorted_ts.size / 2).toInt)
+  val perc75 = sorted_ts(((sorted_ts.size / 4) * 3).toInt)
+  val perc95 = sorted_ts(sorted_ts.size - (sorted_ts.size/20).toInt - 1)
+  def toSec(t: Double): Double = t / 1000000.0
+  override def toString: String = s"#: ${ts.size}, Mean: ${toSec(mean)}, Median: ${toSec(perc50)}"
+}

### 2019-03-28 paper/60_cases.tex
evaluation

@@
 implementation technique for abstract interpretation based static analysis.
-The idea is inspired by partial evaluation similar to the present paper --
-the program is known statically, and the interpretive overhead can be
+Similar to the present paper, the idea is inspired by partial evaluation:
+the program is known statically and the interpretive overhead can be
 eliminated. In AC, the compiled analysis can be represented by either text or
@@
 can be executed immediately, while the textual programs need to be compiled and
-loaded first.
+loaded.
 

### 2019-03-28 oopsla19-code/src/main/scala/sai/evaluation/Analyzer.scala

@@ -1,0 +1,137 @@
+package sai.evaluation
+
+import scala.virtualization.lms.internal.GenericNestedCodegen
+import scala.virtualization.lms.common.{
+  SetOps => _, SetOpsExp => _, ScalaGenSetOps => _,
+  ListOps => _, ListOpsExp => _, ScalaGenListOps => _,
+  _}
+import org.scala_lang.virtualized.virtualize
+import org.scala_lang.virtualized.SourceContext
+
+import sai.lms._
+import sai.monads._
+import sai.lattices._
+import sai.lattices.Lattices._
+
+import sai.evaluation.parser._
+
+trait SchemeAnalyzer {
+  type MonadOps[R[_], M[_], A] = {
+    def map[B: Manifest](f: R[A] => R[B]): M[B]
+    def flatMap[B: Manifest](f: R[A] => M[B]): M[B]
+  }
+  type Ident = String
+  type R[_]
+  type Addr
+  type Value
+  type Env = Map[Ident, Addr]
+  type Store = Map[Addr, Value]
+  type AnsM[T] <: MonadOps[R, AnsM, T]
+  type Ans = AnsM[Value]
+  type Result
+  type EvalFun = Expr => Ans
+
+  // Environment operations
+  def ask_env: AnsM[Env]
+  def ext_env(ans: Ans)(xα: (String, R[Addr])): Ans
+
+  // Store operations
+  def get_store: AnsM[Store]
+  def put_store(σ: R[Store]): AnsM[Unit]
+  def set_store(αv: (R[Addr], R[Value])): AnsM[Unit]
+
+  // allocate addresses
+  def alloc(x: String): AnsM[Addr]
+  def alloc(σ: R[Store], x: String): R[Addr]
+
+  // Primitive operations
+  def void: Ans
+  def literal(i: Any): Ans
+  def get(ρ: R[Env], x: String): R[Addr]
+  def get(σ: R[Store], ρ: R[Env], x: String): R[Value]
+  def br(test: R[Value], thn: => Ans, els: => Ans): Ans
+  def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
+  def close(ev: EvalFun)(λ: Lam, ρ: R[Env]): R[Value]
+  def ap_clo(ev: EvalFun)(fun: R[Value], arg: R[List[Value]]): Ans
+
+  // Fixpoint wrapper and top-level interface
[…]

### 2019-03-28 oopsla19-code/src/main/scala/sai/monads/Monads.scala

@@ -27,9 +27,21 @@ trait Monad[M[_]] extends RMonad[NoRep.NoRep, M]
 object Monad {
   def apply[M[_]](implicit m: Monad[M]): Monad[M] = m
 
-  trait RMonadOps[R[_], M[_], A] {
-    def flatMap[B: Manifest](f: R[A] => M[B]): M[B]
-    def map[B: Manifest](f: R[A] => R[B]): M[B]
+  // TODO: test
+  def mapM[M[_]: Monad, A, B](xs: List[A])(f: A => M[B]): M[List[B]] = xs match {
+    case Nil => Monad[M].pure(List.empty[B])
+    case x::xs => Monad[M].flatMap(f(x)) { b =>
+      Monad[M].flatMap(mapM(xs)(f)) { bs =>
+        Monad[M].pure(b::bs)
+      }
+    }
+  }
+
+  // Returns the last result
+  // TODO: test
+  def forM[M[_]: Monad, A, B](xs: List[A])(f: A => M[B]): M[B] = xs match {
+    case x::Nil => f(x)
+    case x::xs => Monad[M].flatMap(f(x)) { _ => forM(xs)(f) }
   }
 }
 
@@ -114,10 +126,6 @@ object ReaderT {
 
   def liftM[G[_]: Monad, R, A](ga: G[A]): ReaderT[G, R, A] =
     ReaderT(r => ga)
-}
-
-trait SPReaderT[BT[_], M[_], R, A] extends Monad.RMonadOps[BT, SPReaderT[BT, M, R, ?], A] {
-  val run: BT[R] => M[A]
 }
 
 case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {

### 2019-03-28 oopsla19-code/src/main/scala/sai/monads/RepMonads.scala

@@ -25,6 +25,21 @@ trait RepMonads extends RepLattices { self: SAIDsl =>
 
   object Monad {
     def apply[M[_]](implicit m: Monad[M]): Monad[M] = m
+
+    // A doesn't have to be a Rep type
+    def mapM[M[_]: Monad, A, B: Manifest](xs: List[A])(f: A => M[B]): M[List[B]] = xs match {
+      case Nil => Monad[M].pure(List[B]())
+      case x::xs => Monad[M].flatMap(f(x)) { b =>
+        Monad[M].flatMap(mapM(xs)(f)) { bs =>
+          Monad[M].pure(b::bs)
+        }
+      }
+    }
+
+    def forM[M[_]: Monad, A, B: Manifest](xs: List[A])(f: A => M[B]): M[B] = xs match {
+      case x::Nil => f(x)
+      case x::xs => Monad[M].flatMap(f(x)) { _ => forM(xs)(f) }
+    }
   }
 
   /////////////////////////////////////////////////
@@ -217,13 +232,12 @@ trait RepMonads extends RepLattices { self: SAIDsl =>
 
     def flatMap[B: Manifest](f: Rep[A] => ListT[M, B]): ListT[M, B] =
       ListT(Monad[M].flatMap(run) { case list: Rep[List[A]] =>
-              val merge: Rep[List[B]] =
-                list.foldLeft(List[B]()) {
-                  case (acc: Rep[List[B]], a: Rep[A]) =>
-                    val fa: M[List[B]] = f(a).run
-                    val listb: M[List[B]] = Monad[M].map(fa) { case falist: Rep[List[B]] => acc ++ falist }
+              val merge: Rep[List[B]] = list.foldLeft(List[B]()) {
+                case (acc: Rep[List[B]], a: Rep[A]) =>
+                  val fa: M[List[B]] = f(a).run
+                  val listb: M[List[B]] = Monad[M].map(fa) { case falist: Rep[List[B]] => acc ++ falist }
                     ??? // expected Rep[List[B]], a comonad operation `extract: w a -> a`!
-                }
+              }
               Monad[M].pure[List[B]](merge)
             })
 

### 2019-03-28 paper/20_prelim.tex
misc and reorganize
fix a mistake

@@
 These methods either returns monadic values with environment or store inside, or
-simply a @Unit@ value for effects. For example, @ext_env@ takes a value @ans@ of
-type @Ans@ and a pair of binding that maps @x@ to $\alpha$, and installs this
-binding to the environment when evaluating @ans@.
+simply a @Unit@ value for effects. For example, @local_env@ takes a value @ans@ of
+type @Ans@ and a new environment which will be installed when evaluating @ans@.
 
@@
 def ask_env: AnsM[Env]
-def ext_env(ans: Ans)(xα: (Ident, R[Addr])): Ans
+def local_env(ans: Ans)(ρ: R[Env]): Ans
 // Store operations
@@
               case Var(x) => for {                       v <- ev(rhs)
-                ρ <- ask_env                             α <- alloc(x)
-                σ <- get_store                           _ <- set_store(α → v)
-              } yield get(σ, ρ, x)                      rt <- ext_env(ev(e))(x → α)
-              case Lam(x, e) => for {                  } yield rt
-                ρ <- ask_env                           case Aop(op, e1, e2) => for {
-              } yield close(ev)(Lam(x, e), ρ)            v1 <- ev(e1)                                               
-              case App(e1, e2) => for {                  v2 <- ev(e2)
-                v1 <- ev(e1)                           } yield arith(op, v1, v2)
-                v2 <- ev(e2)                           case Rec(x, rhs, e) => for {
-                rt <- ap_clo(ev)(v1, v2)                 α <- alloc(x)
-              } yield rt                                 v <- ext_env(ev(rhs))(x → α)
-              case If0(e1, e2, e3) => for {              _ <- set_store(α → v)
-                cnd <- ev(e1)                            rt <- ext_env(ev(e))(x → α)
-                rt <- br0(cnd, ev(e2), ev(e3))         } yield rt                    
+                ρ <- ask_env                             ρ <- ask_env
+                σ <- get_store                           α <- alloc(x)
+              } yield get(σ, ρ, x)                       _ <- set_store(α → v)
+              case Lam(x, e) => for {                   rt <- local_env(ev(e))(ρ + (x → α))
+                ρ <- ask_env                          } yield rt
+              } yield close(ev)(Lam(x, e), ρ)         case Aop(op, e1, e2) => for {
+              case App(e1, e2) => for {                 v1 <- ev(e1)                                               
+                v1 <- ev(e1)                            v2 <- ev(e2)
+                v2 <- ev(e2)                          } yield arith(op, v1, v2)
+                rt <- ap_clo(ev)(v1, v2)              case Rec(x, rhs, e) => for {
+              } yield rt                                α <- alloc(x)
+              case If0(e1, e2, e3) => for {             ρ <- ask_env
+                cnd <- ev(e1)                           v <- local_env(ev(rhs))(ρ + (x → α))
+                rt <- br0(cnd, ev(e2), ev(e3))          _ <- set_store(α → v)
+              } yield rt                                rt <- local_env(ev(e))(ρ + (x → α))
                                                       } yield rt                    
@@
 
-\subsection{A Concrete Interpreter} \label{unstaged_conc}
+\section{A Concrete Interpreter} \label{unstaged_conc}
 
@@
       _ <- set_store(α → rand)
-      rt <- ext_env(ev(e))(x → α)
+      rt <- local_env(ev(e))(ρ + (x → α))
     } yield rt

### 2019-03-28 paper/40_absinterp.tex
fix a mistake

@@
   def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
+    CloV(Lam(x, e), ρ) <- lift_nd[AbsValue](rator.toList)
     α <- alloc(x)
-    CloV(Lam(x, e), ρ: Env) <- lift_nd[AbsValue](rator.toList)
     _ <- set_store(α → rator)
-    rt <- ext_env(ev(e))(x → α)
+    rt <- local_env(ev(e))(ρ + (x → α))
   } yield rt

### 2019-03-28 oopsla19-code/src/main/scala/sai/MainGeneric.scala

@@ -40,6 +40,7 @@ trait Semantics {
   // Environment operations
   def ask_env: AnsM[Env]
   def ext_env(ans: Ans)(xα: (String, R[Addr])): Ans
+  def local_env(ans: Ans)(ρ: R[Env]): Ans
 
   // Store operations
   def get_store: AnsM[Store]
@@ -134,7 +135,11 @@ trait ConcreteSemantics extends ConcreteComponents {
 
   // Environment operations
   def ask_env: AnsM[Env] = ReaderTMonad[StoreM, Env].ask
-  def ext_env(ans: Ans)(xα: (String, Addr)): Ans = ReaderTMonad[StoreM, Env].local(ans)(ρ => ρ + xα)
+  def ext_env(ans: Ans)(xα: (String, Addr)): Ans = for {
+    ρ <- ask_env
+    v <- local_env(ans)(ρ + xα)
+  } yield v
+  def local_env(ans: Ans)(ρ: Env): Ans = ReaderTMonad[StoreM, Env].local(ans)(_ => ρ)
 
   // Allocating addresses
   def alloc(σ: Store, x: String) = σ.size + 1
@@ -155,7 +160,7 @@ trait ConcreteSemantics extends ConcreteComponents {
     case CloV(Lam(x, e), ρ: Env) => for {
       α <- alloc(x)
       _ <- set_store(α → arg)
-      rt <- ext_env(ev(e))(x → α)
+      rt <- local_env(ev(e))(ρ + (x → α))
     } yield rt
   }
   def br0(test: Value, thn: => Ans, els: => Ans): Ans =
@@ -194,7 +199,9 @@ trait StagedConcreteSemantics extends SAIDsl with ConcreteComponents with RepMon
 
   // Environment operations
   def ask_env: AnsM[Env] = ReaderTMonad[StoreM, Env].ask
-  def ext_env(ans: Ans)(xα: (String, Rep[Addr])): Ans = ReaderTMonad[StoreM, Env].local(ans)(ρ => ρ + (unit(xα._1) → xα._2))
+  def ext_env(ans: Ans)(xα: (String, Rep[Addr])): Ans =
+    ReaderTMonad[StoreM, Env].local(ans)(ρ => ρ + (unit(xα._1) → xα._2))
+  def local_env(ans: Ans)(ρ: Rep[Env]): Ans = ReaderTMonad[StoreM, Env].local(ans)(_ => ρ)
 
[…]

### 2019-03-28 oopsla19-code/src/main/scala/sai/evaluation/Analyzer.scala

@@ -185,8 +185,13 @@ trait UnstagedSchemeAnalyzer extends SchemeAnalyzer {
   def br(test: Value, thn: => Ans, els: => Ans): Ans =
     ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(thn, els)
   def arith(op: Symbol, v1: Value, v2: Value): Value = Set(IntV, FloatV)
-  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = ???
-  def ap_clo(ev: EvalFun)(fun: Value, arg: List[Value]): Ans = ???
+  def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
+  def ap_clo(ev: EvalFun)(fun: Value, arg: List[Value]): Ans = for {
+    CloV(Lam(x, e), ρ: Env) <- lift_nd(fun.toList)
+    αs <- mapM(x)(alloc)
+    //_ <- set_store(x → arg(0))
+    //v <- local_env(ev(e))(ρ + (x → α))
+  } yield ???
 
   def primtives(v: AbsValue, args: List[Value]): Value = ???
 
@@ -251,6 +256,7 @@ trait UnstagedSchemeAnalyzer extends SchemeAnalyzer {
   val σ0: Store = Map()
   val cache0: Cache = Map()
 
+  type Result = (List[(Value, Store)], Cache)
   def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0).run(cache0)(cache0).run
 
   def mValue: Manifest[Value] = manifest[Value]

### 2019-03-29 paper/40_absinterp.tex
evaluation

@@
 operation, and the meet is intersect operation.
-Other lattices used in the rest of the paper, such as products and maps, can be
+Other lattices used in the remainder of the paper, such as products and maps, can be
 implemented similarly, or lifted from existing lattices element-wise or

### 2019-03-29 paper/70_eval.tex
section 4
evaluation

@@
-\section{Evaluation} \label{evaluation}
+\section{Empirical Evaluation} \label{evaluation}
 
-To evaluate whether the performance can be improved through staging, we use the
-abstract interpreter demonstrated in the paper to analyze some Scheme programs.
-We first implement a desugaring transformation for a large subset of Scheme that
-transforms into the small language we used in the paper. We use the same generic
-interface with a 0-CFA-like abstraction, and the unstage abstract interpreter
-forms the baseline.
+To evaluate the performance can be improved by our staging abstract interpreters
+approach, we scale our toy interpreter to a large subset of the Scheme language
+and evaluate the analyzing time against with the unstaged versions. The result
+shows an average \todo{X}\% running time improvement, as a strong evidence of
+our \textit{abstraction without regret} approach to construct abstract
+interpreters.
 
-All of our evaluation benchmarks were performed on an Ubuntu 16.04 LTS (Linux
-kernel 4.4.0) machine with 4 Intel Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of
-RAM. Although the machine has 96 cores and 192 threads in total, the abstract
-interpreters only use one thread to run all the benchmark programs.
+We implement the abstract interpreters for control-flow analysis, i.e., a
+value-flow analysis focus on lambda terms in functional languages. A frontend
+desugars Scheme to a small core language that make the analyzer easy to
+implement. Macros and @call/cc@ is not handled. As described in Section ???, we
+implement a monovariant 0-CFA-like analysis, a context-sensitive $1$-CFA-like
+analysis, both their versions with and without store-widening.
 
-Our evaluations show that the staged abstract interpreter performs well in some
-cases that are considered to be the worst scenarios for traditional $k$-CFA. The
-staged version outperforms the unstaged abstract interpreter by a significant
-margin (up to ~230\% performance gain) on these difficult ones.
+\paragraph{Setting}
 
-\subsection{Benchmarks}
+We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04LTS
+(kernel 4.4.0). All of our evaluation were performed on a machine with 4 Intel
+Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
+cores and 192 threads in total, the abstract interpreters only use one thread.
+To minimize the effect of warming-up from the JVM JIT compiler, all the
+experiments are executed for 20 times and we report the statistical median value
+of the running times (in seconds).
 
+\paragraph{Benchmarks} \todo{Rewrite}
 The benchmark programs we used are collected from previous papers
@@
 
-\subsection{Performance}
-
-Considering that our abstract interpreters are implemented in Scala which will be 
-affected by the JIT warm up times, we run all experiments 10 times and report
-the statistical median values of the running times.
-
-\todo{remove fib; report LOC; mean}
+\paragraph{Result}
 
 \begin{figure}[h]
-\begin{tabular}{@{}llll@{}}
+\tiny
+\begin{tabular}{@{}lllll@{}}
 \toprule
-    Program            & unstaged   & staged     & $\Delta$ of median \\ \midrule
-  % fib                & 1.020      & 2.046      & -50.14\%          \\
-    rsa                & 35.928     & 145.306    & -75.27\%          \\
-    kcfa3              & 2.934      & 5.671      & -48.27\%          \\
-    church             & 884.800    & 545.736    & +62.13\%          \\
-    fermat             & 15.673     & 68.790     & -77.22\%          \\
-    kcfa-worst-16      & 320.689    & 301.564    & +6.34\%           \\
-    kcfa-worst-32      & 3,685.615  & 1,854.461  & +98.74\%          \\
-    kcfa-worst-64      & 49,422.666 & 14,849.725 & +232.82\%         \\
+    Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
+  % fib              &      & 1.020      & 2.046      & -50.14\%          \\
+    @rsa@            &      &            &            & \%          \\
+    @kcfa3@          &      &            &            & \%          \\
+    @church@         &      &            &            & \%          \\
+    @fermat@         &      &            &            & \%          \\
+    @kcfa-worst-16@  &  198 & 0.352      & 0.028      & +1157.14\%        \\
+    @kcfa-worst-32@  &  390 & 4.234      & 0.289      & +1365.05\%          \\
+    @kcfa-worst-64@  &  774 & 61.475     & 4.425      & +1289.27\%         \\
     \bottomrule
 \end{tabular}
-\caption{Evaluation result.} \label{evaluation_result}
+\caption{Evaluation result for 0CFA without store-widening.} \label{evaluation_result}
 \end{figure}
 
-Figure~\ref{evaluation_result} shows the benchmark evaluation results. The
-"unstaged" and "staged" columns show the median time to finish the analysis (in
-milliseconds). The "$\Delta$ of median" column shows the performance change from
-the unstaged abstract interpreter to the staged version.
+Figure~\ref{evaluation_result} shows the evaluation results. The
+\textit{unstaged} and \textit{staged} columns show the median time to finish the
+analysis (in seconds). The "$\Delta$" column shows the median runtime change
+from the unstaged to the staged version. \todo{0cfa, 1cfa, store-widen}
 
+\iffalse
 As we can see from the table, the staged version significantly outperforms the
@@
 duplication.
+\fi

### 2019-03-29 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -12,21 +12,22 @@ import scala.io.Source
 import sai.lms._
 import sai.evaluation.utils._
 import sai.evaluation.parser._
+import sai.evaluation.parser.ASTUtils._
 import sai.evaluation.TestPrograms._
 
 object Evaluation {
 
   def main(args: Array[String]) {
     val progs = List[(Expr, String)](
-      //boyer
-      //euclid,
-      //euclid_imp
-      //kcfa32
+      //(boyer, "boyer"),
+      //(euclid, "euclid"),
+      //(euclid_imp, "euclid_imp"),
       //(fact5, "fact5")
       //(kcfa16, "kcfa16"),
       //(kcfa32, "kcfa32"),
       //(kcfa64, "kcfa64"),
-      (kcfa256, "kcfa256")
+      //(kcfa256, "kcfa256"),
+      (church, "church")
     )
     progs foreach { case (e, id) => compare(e, id) }
   }
@@ -71,8 +72,9 @@ object Evaluation {
   val output = "CodeGen.out"
 
   def compare(e: Expr, id: String) {
+    println(s"Running $id, AST size: ${size(e)}")
+    /*
     val N = 20
-
     val t1 =  run(N, { evalUnstaged(e) })
     println(s"[$id] [unstaged] - ${t1}s")
   
@@ -83,9 +85,6 @@ object Evaluation {
 
     val t2 = run(N, { code.eval(()) })
     println(s"[$id] [staged] - ${t2}s")
-
-    //println("Result:")
-    //println(s"  Unstaged time - ${t1}")
-    //println(s"  Staged time - ${t2}")
+    */
   }
 }

### 2019-03-29 oopsla19-code/src/main/scala/sai/evaluation/GenericAnalyzer.scala

@@ -128,6 +128,5 @@ trait AbstractComponents extends SchemeAnalyzer {
 
   type Value = Set[AbsValue]
   type Ctx = R[List[Expr]]
-
 }
 

### 2019-03-29 oopsla19-code/src/main/scala/sai/evaluation/parser/AST.scala

@@ -63,7 +63,7 @@ case class Set_!(x: String, e: Expr) extends Expr with AtomExpr {
 case class Begin(es: List[Expr]) extends Expr
 case class Define(x: String, e: Expr) extends Expr
 
-object SExpPrinter {
+object ASTUtils {
   def exprToString(e: Expr): String = e match {
     case Var(x) => x
     case Void() => "(void)"
@@ -80,6 +80,22 @@ object SExpPrinter {
     case Lam(params, body) => "(lambda (" + params.mkString(" ") + ") " + exprToString(body) + ")"
   }
 
-  def apply(e: Expr): Unit = println(exprToString(e))
+  def prettyPrint(e: Expr): Unit = println(exprToString(e))
+
+  def size(e: Expr): Int = e match {
+    case Var(x) => 1
+    case Void() => 1
+    case Sym(x) => 1
+    case CharLit(x)   => 1
+    case IntLit(x)    => 1
+    case FloatLit(x)  => 1
+    case BoolLit(x)   => 1
+    case Set_!(x, e)  => 1 + size(e)
+    case Define(x, e) => 1 + size(e)
+    case App(x, l)    => 1 + size(x) + l.foldLeft(0)(_ + size(_))
+    case Begin(es)    => 1 + es.foldLeft(0)(_ + size(_))
+    case If(c, t, e)  => 1 + size(c) + size(t) + size(e)
+    case Lam(params, body) => 1 + size(body)
+  }
 }
 

### 2019-03-29 oopsla19-code/src/main/scala/sai/evaluation/parser/SchemeParser.scala

@@ -4,7 +4,9 @@ package sai.evaluation.parser
 
 import scala.util.parsing.combinator._
 import scala.io.Source
+
 import sai.evaluation.utils.TestTrait
+import sai.evaluation.parser.ASTUtils._
 
 trait SchemeParserTrait extends SchemeTokenParser {
   def id[T](x: T) = x
@@ -149,34 +151,34 @@ object SchemeParser extends SchemeParserTrait {
 object SchemeParserTest extends TestTrait {
 
   def testPrettyPrinter = {
-    assert(SExpPrinter.exprToString(CharLit('a')) == "#\\a")
-    assert(SExpPrinter.exprToString(App(Lam(List("x", "y"),
+    assert(prettyPrint(CharLit('a')) == "#\\a")
+    assert(prettyPrint(App(Lam(List("x", "y"),
       App(Var("+"), List(Var("x"), Var("y")))), List(IntLit(1), IntLit(3)))) ==
       "((lambda (x y) (+ x y)) 1 3)")
-    assert(SExpPrinter.exprToString(If(BoolLit(false), Var("a"), Lam(List("t"), Var("t")))) ==
+    assert(prettyPrint(If(BoolLit(false), Var("a"), Lam(List("t"), Var("t")))) ==
       "(if #f a (lambda (t) t))")
   }
 
   def testDesugar = {
     assert(SchemeASTDesugar(IntLit(1)) == IntLit(1))
-    SExpPrinter(SchemeASTDesugar(
+    prettyPrint(SchemeASTDesugar(
       Begin(List(Define("x", IntLit(2)), Set_!("x", IntLit(3)), Var("x")))))
-    SExpPrinter(SchemeASTDesugar(
+    prettyPrint(SchemeASTDesugar(
       Cond(List(
         CondBr(
           App(Var("positive?"),List(IntLit(-5))),
           App(Var("error"),List())),
         CondBr(App(Var("zero?"),List(IntLit(-5))),App(Var("error"),List())),
         CondBr(App(Var("positive?"),List(IntLit(5))),Sym("here"))))))
-    SExpPrinter(SchemeASTDesugar(
+    prettyPrint(SchemeASTDesugar(
       Case(IntLit(3), List(
         CaseBranch(List(IntLit(3), IntLit(4), IntLit(5)), BoolLit(true)),
[…]

### 2019-03-29 paper/70_eval.tex
related work

@@
 
-We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04LTS
-(kernel 4.4.0). All of our evaluation were performed on a machine with 4 Intel
+We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
+(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
 Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
-cores and 192 threads in total, the abstract interpreters only use one thread.
+cores (192 threads) in total, the abstract interpreters are single-threaded.
 To minimize the effect of warming-up from the JVM JIT compiler, all the

### 2019-03-29 paper/80_related.tex
related work

--- 2019-02-22 paper/70_related.tex
@@
 \paragraph{Optimizing Static Analysis Through Specialization}
-The idea in this paper is closely inspired by abstract compilation
-\cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis
-(\textit{0}-CFA) for programs written in continuation-passing style. The key
-idea is to remove the interpretation overhead on traversing the syntax tree by
-partial evaluation. Specifically, they proposed two similar kinds of abstract
-compilation techniques. The first one is to generate specialized analysis as a
-textual program, which then can be loaded and executed by \texttt{eval} or other
-similar mechanisms. The second one is to use closures, i.e., functions that
-remember their environments, as a representation of specialized analysis. As we
-show in the case study, compiling the analysis generates higer-order functions
-on-the-fly with respect to the analyzed program, then the generated closure can
-be applied immediately in the higher-order host language.
-
-\citeauthor{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
-generation to optimize small-step abstract interpreter in state-transition style
-\cite{Johnson:2013:OAA:2500365.2500604}. The analyzed program is firstly
-compiled to a intermediate representation called "abstract bytecode", which are
-actually higher-order functions, and then be executed on a abstract abstract
-machine for that IR.
-
-\citet{damian1999partial} provides a formal treatment to abstract compilation
-and Shiver's CFA, as well as proofs to establish correctness of the certified
-specialized analyzer \cite{damian1999partial}. \citeauthor{amtoft1999partial}
-applied partial evaluation for constraint-based control flow analysis
-\cite{amtoft1999partial}. Splitting an analysis into multiple stages is also
-studied for analysis other than for control-flow, though the formulation may very
-different. For example, \citeauthor{DBLP:conf/cgo/HardekopfL11} apply staging to
-flow-sensitive pointer analysis \cite{DBLP:conf/cgo/HardekopfL11}. The first
-stage is to analyze the program code to obtain a sparse representation, and then
-the second stage conducts the flow-sensitive analysis based on the first one.
-
-\citeauthor{DBLP:conf/popl/SinghPV17} optimizes polyhedra abstract domains
-through online decomposition \cite{DBLP:conf/popl/SinghPV17,
-Singh:2017:PCD:3177123.3158143}. The idea is to decompose a large polyhedra into
-several smaller one; one abstract transformer usually only depends a limit
-number of variables, then we just need to compute the resulting polyhera on
-those smaller polyhedras which contain these related variables. The decomposition
-can be considered as a specialization of abstract domain with respect to the
-static program structure.
+The underlying idea in this paper is closely related to abstract compilation (AC)
+\cite{Boucher:1996:ACN:647473.727587}: removing the interpretation overhead on
+traversing the syntax tree by specialization. Specifically, the residual program
+of AC can be either textual or closures. However, as we studied in Section
+\ref{cs_ac}, with the perspective from high-performance DSLs and monadic
+interpreters, we make it more general, extensible and easy.
+\citet{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
+generation to optimize small-step abstract interpreter in state-transition
+style. The analyzed program is firstly compiled to a IR called "abstract
+bytecode", which are actually higher-order functions and will be executed later
+on a abstract abstract machine for that IR. \citet{damian1999partial} provides a
+formal treatment to abstract compilation and Shiver's CFA, as well as proofs to
+establish correctness of the certified specialized analyzer.
+\citet{amtoft1999partial} applied partial evaluation for constraint-based
+control flow analysis. Splitting an analysis into multiple stages is also
+studied for analysis other than for control-flow, though the formulation may
+very different. For example, \citet{DBLP:conf/cgo/HardekopfL11} apply staging to
+flow-sensitive pointer analysis. The first stage is to analyze the program code
+to obtain a sparse representation, and then the second stage conducts the
+flow-sensitive analysis based on the first one.
+The abstract domains can be also specialized w.r.t. the static program
+structure, e.g., decomposing polyhedras \cite{DBLP:conf/popl/SinghPV17,
+Singh:2017:PCD:3177123.3158143} to smaller ones such that the abstract
+transformers can compute using less time and space.
 
@@
 staged monadic interpreters \cite{DBLP:conf/dsl/SheardBP99}
+
+compiling monads \cite{danvy1991compiling}

### 2019-03-29 paper/60_cases.tex
tweak code

@@
 }
-def compApp(f: Expr, args: List[Expr]): CompAnalysis = 
-  f match {
+def compApp(f: Expr, args: List[Expr]): CompAnalysis = f match {
   case Var(x) => ($\sigma$: Store) => 
@@
   call match {
-    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
     case Letrec(bds, body) =>
@@
       analyzeCall(body, $\sigma$_**)
+    case App(f, args) => analyzeApp(f, args, analyzeArgs(args, $\sigma$))
   }

### 2019-03-30 paper/50_stageabs.tex
misc

@@
 
-\paragraph{What the generated code look like?} In the generated code, all the
+\paragraph{What does the generated code look like?} In the generated code, all the
 primitive operations (such as @eval@, @fix@, @ap_clo@, etc.) and monadic

### 2019-03-30 paper/70_eval.tex
misc

@@
 desugars Scheme to a small core language that make the analyzer easy to
-implement. Macros and @call/cc@ is not handled. As described in Section ???, we
-implement a monovariant 0-CFA-like analysis, a context-sensitive $1$-CFA-like
-analysis, both their versions with and without store-widening.
+implement. Macros and @call/cc@ is not handled. As described in Section
+\ref{cfa}, we implement a monovariant 0-CFA-like analysis, a context-sensitive
+$1$-CFA-like analysis, both their versions with and without store-widening.
 
@@
 DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
-\footnote{https://github.com/ilyasergey/reachability}.
-
+\footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
 We used the following benchmark programs:
@@
     Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
-  % fib              &      & 1.020      & 2.046      & -50.14\%          \\
     @rsa@            &      &            &            & \%          \\

### 2019-03-30 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -19,17 +19,32 @@ object Evaluation {
 
   def main(args: Array[String]) {
     val progs = List[(Expr, String)](
-      //(boyer, "boyer"),
       //(euclid, "euclid"),
       //(euclid_imp, "euclid_imp"),
-      //(fact5, "fact5")
+      //(fact5, "fact5"),
       //(kcfa16, "kcfa16"),
       //(kcfa32, "kcfa32"),
       //(kcfa64, "kcfa64"),
       //(kcfa256, "kcfa256"),
       (church, "church")
+      //(fermat, "fermat"),
+      //(rsa, "rsa"),
+      //(blur, "blur"), // 56
+      //(fib, "fib"), // 36
+      //(sat, "sat"), // 73
+      //(boyer, "boyer"),
+      //(earley, "earley"),
+      //(mbrotZ, "mbrotZ"),
+      //(dynamic, "dynamic"),
+      //(graphs, "graphs"),
+      //(lattice, "lattice"),
+      //(matrix, "matrix"),
+      //(nbody, "nbody"),
+      //(nucleic, "nucleic"),
     )
-    progs foreach { case (e, id) => compare(e, id) }
+    //progs foreach { case (e, id) => compare(e, id) }
+    //progs foreach { case (e, id) => println(s"$id - ${size(e)}") }
+    progs foreach { case (e, id) => evalUnstaged(e) }
   }
 
   def run[R](n: Int, block: => R): Timing = {
@@ -48,7 +63,8 @@ object Evaluation {
 
   def evalUnstaged(e: Expr): Unit = {
     val res = UnstagedSchemeAnalyzer.run(e)
-    //println(s"Number of values: ${res._1.size}")
+    println(s"Values: ${res._1}")
+    println(s"Number of values: ${res._1.size}")
     //println(s"Size of cache: ${res._2.size}")
   }
 
@@ -73,7 +89,6 @@ object Evaluation {
 
   def compare(e: Expr, id: String) {
     println(s"Running $id, AST size: ${size(e)}")
-    /*
     val N = 20
     val t1 =  run(N, { evalUnstaged(e) })
     println(s"[$id] [unstaged] - ${t1}s")
@@ -85,6 +100,5 @@ object Evaluation {
 
     val t2 = run(N, { code.eval(()) })
     println(s"[$id] [staged] - ${t2}s")
-    */
   }
 }

### 2019-03-30 oopsla19-code/src/main/scala/sai/evaluation/TestPrograms.scala

@@ -87,7 +87,7 @@ object TestPrograms {
   def boyer = getAST(Source.fromFile("benchmarks/toplas98/boyer.sch").mkString)
 
   // Dynamic -- Fritz's dynamic type inferencer, set up to run on itself
-  def dynamic = getAST(Source.fromFile("benchmarks/toplas98/boyer.sch").mkString)
+  def dynamic = getAST(Source.fromFile("benchmarks/toplas98/dynamic.sch").mkString)
 
   // graphs
   def graphs = getAST(Source.fromFile("benchmarks/toplas98/graphs.sch").mkString)

### 2019-03-30 oopsla19-code/src/main/scala/sai/evaluation/parser/Transform.scala

@@ -6,7 +6,7 @@ object SchemeASTDesugar {
   var lastIdent = 0
 
   def newIdentLet(e: Expr)(f: Expr => Expr): App = {
-    val prefix = "__$"
+    val prefix = "$"
     val id = lastIdent
     lastIdent += 1
     val newIdent: String = prefix + id

### 2019-03-30 oopsla19-code/src/main/scala/sai/lms/ListOps.scala

@@ -194,12 +194,14 @@ trait ListOpsExpOpt extends ListOpsExp with TupleOpsExp {
     case (xs1, Def(ListNew(Seq(), _))) => xs1
     case _ => super.list_concat(xs1, xs2)
   }
+
   override def list_foldLeft[A: Manifest, B: Manifest](s: Exp[List[A]], z: Exp[B], f: (Exp[B], Exp[A]) => Exp[B])(implicit pos: SourceContext): Exp[B] = s match {
     case Def(ListNew(xs, _)) if xs.size == 0 => z
     case Def(ListNew(xs, _)) if xs.size == 1 => f(z, xs(0))
     case Def(ListNew(xs, _)) => xs.foldLeft(z)(f)
     case _ => super.list_foldLeft(s, z, f)
   }
+
   override def list_foldLeftPair[A: Manifest, B: Manifest, C: Manifest](s: Exp[List[A]], z: Exp[(B, C)], f: ((Exp[B], Exp[C]), Exp[A]) => Exp[(B,C)])(implicit pos: SourceContext): Exp[(B,C)] = s match {
     case Def(ListNew(xs, _)) if xs.size == 0 => z
     case Def(ListNew(xs, _)) if xs.size == 1 => f(z, xs(0))

### 2019-04-02 paper/40_absinterp.tex
misc

@@
     rt <- if (out.contains(cfg)) for {
-            (v, s) <- lift_nd[(Value, Store)](out(cfg).toList)
-             _ <- put_store(s)
+            (v, σ) <- lift_nd[(Value, Store)](out(cfg).toList)
+             _ <- put_store(σ)
           } yield v

### 2019-04-02 paper/abstract.tex
misc

@@
-It is well known that a staged interpreter is a compiler: specializing the interpreter to a given program
-produces an equivalent program that runs faster.
-It is even more widely known that an abstract interpreter is a program analyzer: tweaking the 
-interpreter to run on a domain of abstract values produces a sound static analysis.
+It is well known that a staged interpreter is a compiler: specializing the
+interpreter to a given program produces an equivalent program that runs faster,
+which known as the first Futamura projection. It is even more widely known that an
+abstract interpreter is a program analyzer: tweaking the interpreter to run on a
+domain of abstract values produces a sound static analysis. What happens when we
+combine these two ideas, and apply staging to an \emph{abstract} interpreter?
 
-What happens when we combine these two ideas, and apply staging to an \emph{abstract} interpreter? 
-We obtain a sound static analysis, specialized for a given program, that runs faster.
-%
-More surprisingly, we show that by applying the staged abstract interpreter to \textit{open} 
-programs and considering the free variables as dynamic inputs, 
-we obtain a modular analysis that generates sound partial analysis results which can be composed 
-and reused later without losing precision, even though the original abstract interpreter is a whole-program 
-analysis algorithm.
+In this paper, we present a unifying framework that naturally extends the first
+Futamura projection of concrete interpreters to abstract interpreters. Our
+approach derives a sound staged abstract interpreter based on a
+semantic-agnostic interpreter with type-level binding-time abstraction and
+monadic abstraction. By using different instantiations of these abstractions,
+the generic interpreter can flexibly behave in four modes: unstaged concrete
+interpreter, staged concrete interpreter, unstaged abstract interpreter, or
+staged abstract interpreter.
 
-Based on the idea of staged abstract interpreters, we show several case studies, including 
-\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of 0-CFA, pushdown
-control-flow analysis with context-sensitivity and precise stores, and a numerical
-analysis on an imperative language.
+As an example of abstraction without regret, we show that staging abstract
+interpreters is practical and useful to optimize static analysis while requiring
+least engineering efforts and not compromising any soundness. We conduct three
+case studies, including a comparison with
+\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
+the applicability on various control-flow analyses, and the use to modular
+analysis. We also empirically evaluate the performance improved by staging. The
+overhead of monads is eliminated in the generated code, and the experiment shows
+ an average speedup of \note{10}x with staging for a value-flow analysis.
 
-We empirically evaluate the performance improvements on control-flow analysis of benchmark programs.
-The results show speedups up to 2.3x with staging on a monovariant analysis.
+\iffalse
+We obtain a sound static analysis, specialized for
+a given program, that runs faster. More surprisingly, we show that by applying
+the staged abstract interpreter to \textit{open} programs and considering the
+free variables as dynamic inputs, we obtain a modular analysis that generates
+sound partial analysis results which can be composed and reused later without
+losing precision, even though the original abstract interpreter is a
+whole-program analysis algorithm.
 
-% It is well known that a staged interpreter is a compiler, which provides performance improvement
-% by specializing the interpreter to a given program.
+Based on the idea of staged abstract interpreters, we show several case studies,
+including \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of
+0-CFA, pushdown control-flow analysis with context-sensitivity and precise
+stores, and a numerical analysis on an imperative language.
 
-% In this paper, we study \textit{abstract} interpreters combined with multi-stage programming, i.e., the 
-% staged abstract interpreters. 
-% By staging the abstract interpreter with respect to a program, we obtain a specialized analysis that runs faster.
-% By applying the staged abstract interpreter with \textit{open} programs and considering the free variables as dynamic inputs, 
-% we obtain a modular analysis that generates sound partial analysis results which can be composed 
-% and reused later without losing precision, though the original abstract interpreter is a whole-program 
-% analysis algorithm.
+We empirically evaluate the performance improvements on control-flow analysis of
+benchmark programs. The results show speedups up to 2.3x with staging on a
+monovariant analysis.
+\fi
 
-% Using the idea of staged abstract interpreters, we show several case studies, including 
-% \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of  0-CFA, pushdown
-% control-flow analysis with context/path/flow-sensitivity and store-widening, and a numerical
-% analysis on an imperative language.
-
-% We also empirically evaluate the improvement of performance on control-flow analysis of benchmark programs.
-% The result shows an average speedup of Nx when staging to Scala for a monovariant analysis, and Mx for polyvariant analysis.
+% It is well known that a staged interpreter is a compiler, which provides
+% performance improvement by specializing the interpreter to a given program. In
+% this paper, we study \textit{abstract} interpreters combined with multi-stage
+% programming, i.e., the staged abstract interpreters. By staging the abstract
+% interpreter with respect to a program, we obtain a specialized analysis that
+% runs faster. By applying the staged abstract interpreter with \textit{open}
+% programs and considering the free variables as dynamic inputs, we obtain a
+% modular analysis that generates sound partial analysis results which can be
+% composed and reused later without losing precision, though the original
+% abstract interpreter is a whole-program analysis algorithm. Using the idea of
+% staged abstract interpreters, we show several case studies, including
+% \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation of 0-CFA,
+% pushdown control-flow analysis with context/path/flow-sensitivity and
+% store-widening, and a numerical analysis on an imperative language. We also
+% empirically evaluate the improvement of performance on control-flow analysis
+% of benchmark programs. The result shows an average speedup of Nx when staging
+% to Scala for a monovariant analysis, and Mx for polyvariant analysis.

### 2019-04-02 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -44,7 +44,16 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
   def put_store(σ: Store): AnsM[Unit] =
     ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].put(σ))
   def set_store(αv: (Addr, Value)): AnsM[Unit] =
-    ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].mod(σ => σ ⊔ Map(αv)))
+    ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].mod(σ => {
+                                                                                         val news = σ ⊔ Map(αv)
+                                                                                         val oldc = σ.getOrElse(αv._1, Set())
+                                                                                         val newc = news(αv._1)
+                                                                                         if (oldc != newc) {
+                                                                                           println(s"growing: ${αv._1} oldsize: ${oldc.size} newsize: ${newc.size}")
+                                                                                         }
+
+                                                                                         news
+                                                                                       }))
 
   // allocate addresses
   def alloc(σ: Store, x: String): Addr = ZCFAAddr(x)
@@ -66,17 +75,25 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
   }
   def get(ρ: Env, x: String): Addr = ρ(x)
   def get(σ: Store, ρ: Env, x: String): Value = σ(ρ(x))
-  def br(test: Value, thn: => Ans, els: => Ans): Ans = for {
-    v1 <- thn
-    v2 <- els
-  } yield v1 ++ v2
-    //ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(thn, els)
+  def br(ev: EvalFun)(test: Value, thn: Expr, els: Expr): Ans =
+    //ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(ev(thn), ev(els)) // they use different store and cache
+    for { //Note: they use the same store and cache
+      ρ <- ask_env
[…]

### 2019-04-02 oopsla19-code/src/main/scala/sai/monads/Monads.scala

@@ -27,7 +27,6 @@ trait Monad[M[_]] extends RMonad[NoRep.NoRep, M]
 object Monad {
   def apply[M[_]](implicit m: Monad[M]): Monad[M] = m
 
-  // TODO: test
   def mapM[M[_]: Monad, A, B](xs: List[A])(f: A => M[B])(implicit mB: Manifest[B] = null): M[List[B]] = xs match {
     case Nil => Monad[M].pure(List.empty[B])
     case x::xs => Monad[M].flatMap(f(x)) { b =>
@@ -90,8 +89,6 @@ case class IdM[A](run: A) {
   def apply: A = run
   def flatMap[B](f: A => IdM[B])(implicit mB: Manifest[B] = null): IdM[B] = f(run)
   def map[B](f: A => B)(implicit mB: Manifest[B] = null): IdM[B] = IdM(f(run))
-  //def flatMap[B](f: A => IdM[B]): IdM[B] = f(run)
-  //def map[B](f: A => B): IdM[B] = IdM(f(run))
 }
 
 /////////////////////////////////////////////////
@@ -131,8 +128,6 @@ object ReaderT {
 case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {
   import ReaderT._
   def apply(r: R): M[A] = run(r)
-  //def flatMap[B](f: A => ReaderT[M, R, B]): ReaderT[M, R, B] = ReaderT(r => Monad[M].flatMap(run(r))(a => f(a).run(r)))
-  //def map[B](f: A => B): ReaderT[M, R, B] = ReaderT(r => Monad[M].map(run(r))(f))
 
   def flatMap[B](f: A => ReaderT[M, R, B])(implicit mB: Manifest[B] = null): ReaderT[M, R, B] =
     ReaderT(r => Monad[M].flatMap(run(r))(a => f(a).run(r)))

### 2019-04-03 paper/abstract.tex
wip

@@
 analysis. We also empirically evaluate the performance improved by staging. The
-overhead of monads is eliminated in the generated code, and the experiment shows
- an average speedup of \note{10}x with staging for a value-flow analysis.
+overhead of the abstraction layers is eliminated in the generated code, and the 
+experiment shows an average speedup of \note{10}x with staging for a value-flow analysis.
 

### 2019-04-03 oopsla19-code/src/main/scala/sai/evaluation/GenericAnalyzer.scala

@@ -50,7 +50,7 @@ trait SchemeAnalyzer {
   def get(ρ: R[Env], x: String): R[Addr]
   def get(σ: R[Store], ρ: R[Env], x: String): R[Value]
   //def br(test: R[Value], thn: => Ans, els: => Ans): Ans
-  def br(ev: EvalFun)(test: R[Value], thn: Expr, els: Expr): Ans
+  def br(ev: EvalFun)(test: Expr, thn: Expr, els: Expr): Ans
   def close(ev: EvalFun)(λ: Lam, ρ: R[Env]): R[Value]
   def ap_clo(ev: EvalFun)(fun: R[Value], arg: R[List[Value]]): Ans
   def primtives(v: R[Value], args: List[R[Value]]): R[Value]
@@ -96,12 +96,20 @@ trait SchemeAnalyzer {
       v2 <- mapM(args)(ev)
       rt <- ap_clo(ev)(v1, v2)
     } yield rt
+    case Begin(Define(x, rhs)::rest) => for {
+      α <- alloc(x)
+      v <- ext_env(ev(rhs))(x → α)
+      _ <- set_store(α → v)
+      rv <- ext_env(ev(Begin(rest)))(x → α)
+    } yield rv
     case Begin(es) => forM(es)(ev)
     case If(cnd, thn, els) => for {
       t <- ev(cnd)
-      rt <- br(ev)(t, thn, els)
+      rt <- br(ev)(cnd, thn, els)
     } yield rt
   }
+
+  def foldVss(vss: R[List[Value]]): R[Value]
 }
 
 trait AbstractComponents extends SchemeAnalyzer {

### 2019-04-03 oopsla19-code/src/main/scala/sai/evaluation/StagedAnalyzer.scala

@@ -63,6 +63,8 @@ trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with Rep
   def set_store(αv: (Rep[Addr], Rep[Value])): AnsM[Unit] =
     ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].mod(σ => σ ⊔ Map(αv)))
 
+  def foldVss(vss: Rep[List[Value]]): Rep[Value] = ???
+
   def void: Ans = ReaderTMonad[StoreNdInOutCacheM, Env].pure[Value](Set[AbsValue]())
   def literal(i: Any): Ans = {
     val v: AbsValue = i match {
@@ -78,7 +80,8 @@ trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with Rep
   }
   def get(ρ: Rep[Env], x: String): Rep[Addr] = ρ(x)
   def get(σ: Rep[Store], ρ: Rep[Env], x: String): Rep[Value] = σ(ρ(x))
-  def br(ev: EvalFun)(test: Rep[Value], thn: Expr, els: Expr): Ans =
+  //def br(ev: EvalFun)(test: Rep[Value], thn: Expr, els: Expr): Ans =
+  def br(ev: EvalFun)(test: Expr, thn: Expr, els: Expr): Ans =
     ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(ev(thn), ev(els))
   def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(params, e) = λ

### 2019-04-03 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -10,7 +10,8 @@ import sai.evaluation.parser._
 object UnstagedSchemeAnalyzer extends AbstractComponents {
   import ReaderT._
   import StateT._
-  import ListT._
+  //import ListT._
+  import SetT._
   import IdM._
 
   type R[T] = T
@@ -20,7 +21,8 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
   // Transformers
   type EnvT[F[_], B] = ReaderT[F, Env, B]
   type StoreT[F[_], B] = StateT[F, Store, B]
-  type NondetT[F[_], B] = ListT[F, B]
+  //type NondetT[F[_], B] = ListT[F, B]
+  type NondetT[F[_], B] = SetT[F, B]
   type InCacheT[F[_], B] = ReaderT[F, Cache, B]
   type OutCacheT[F[_], B] = StateT[F, Cache, B]
   // Monads
@@ -48,12 +50,11 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
                                                                                          val news = σ ⊔ Map(αv)
                                                                                          val oldc = σ.getOrElse(αv._1, Set())
                                                                                          val newc = news(αv._1)
-                                                                                         if (oldc.size == 0) {
-                                                                                           println(s"growing: ${αv}")
+                                                                                         if (oldc != newc) {
+                                                                                           //println(s"growing: ${αv._1} oldsize: ${oldc.size} newsize: ${newc.size}")
+                                                                                           //println(s"DIFF ${newc -- oldc}")
                                                                                          }
-                                                                                         //  println(s"growing: ${αv._1} oldsize: ${oldc.size} newsize: ${newc.size}")
[…]

### 2019-04-03 oopsla19-code/src/main/scala/sai/monads/Monads.scala

@@ -27,6 +27,7 @@ trait Monad[M[_]] extends RMonad[NoRep.NoRep, M]
 object Monad {
   def apply[M[_]](implicit m: Monad[M]): Monad[M] = m
 
+  // TODO: test
   def mapM[M[_]: Monad, A, B](xs: List[A])(f: A => M[B])(implicit mB: Manifest[B] = null): M[List[B]] = xs match {
     case Nil => Monad[M].pure(List.empty[B])
     case x::xs => Monad[M].flatMap(f(x)) { b =>
@@ -89,6 +90,8 @@ case class IdM[A](run: A) {
   def apply: A = run
   def flatMap[B](f: A => IdM[B])(implicit mB: Manifest[B] = null): IdM[B] = f(run)
   def map[B](f: A => B)(implicit mB: Manifest[B] = null): IdM[B] = IdM(f(run))
+  //def flatMap[B](f: A => IdM[B]): IdM[B] = f(run)
+  //def map[B](f: A => B): IdM[B] = IdM(f(run))
 }
 
 /////////////////////////////////////////////////
@@ -128,6 +131,8 @@ object ReaderT {
 case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {
   import ReaderT._
   def apply(r: R): M[A] = run(r)
+  //def flatMap[B](f: A => ReaderT[M, R, B]): ReaderT[M, R, B] = ReaderT(r => Monad[M].flatMap(run(r))(a => f(a).run(r)))
+  //def map[B](f: A => B): ReaderT[M, R, B] = ReaderT(r => Monad[M].map(run(r))(f))
 
   def flatMap[B](f: A => ReaderT[M, R, B])(implicit mB: Manifest[B] = null): ReaderT[M, R, B] =
     ReaderT(r => Monad[M].flatMap(run(r))(a => f(a).run(r)))
@@ -234,4 +239,53 @@ case class ListT[M[_]: Monad, A](run: M[List[A]]) {
   def filter(f: A => Boolean): ListT[M, A] =
     ListT(Monad[M].map(run) { list => list.filter(f) })
   def withFilter(f: A => Boolean): ListT[M, A] = filter(f)
+}
+
+/////////////////////////////////////////////////////
+
+object SetT {
+  def apply[M[_]: Monad, A](implicit m: SetT[M, A]): SetT[M, A] = m
+
[…]

### 2019-04-04 paper/10_intro.tex
paper revision

@@
 programmers. The staging annotations can be either syntactic (as in
-MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \todo{cite}.
+MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \cite{DBLP:conf/gpce/RompfO10}.
 These staging annotations identify which part of the inputs should be
@@
 In this paper, we study the confluence of two old ideas --- Futamura projection
-and abstract interpretation, but from the perspective of their recent
+and abstract interpretation, but from the perspective of their very recent
 realizations --- multi-stage programming and definitional abstract interpreters.
@@
 To enable such flexibility between different semantics, we combine generic
-programming and generative programming, and extensively use abstract type members 
-over the following abstractions, which can be later instantiated.
+programming and generative programming, and extensively use abstract type
+members over the following abstractions, which can be later instantiated.
 
@@
   the underlying semantics of the interpreter is changed.
-  \todo{mention that direct-style also works?}
+  %\todo{mention that direct-style also works?}
   %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
@@
 abstractions -- from the monadic layers -- is completely eliminated in the
-generated code. \todo{put some numbers} However, we do not lose any advantages of this high-level
-programming style (e.g., modularity and extensibility); the meaning of the program also
-remains the same, particularly, the abstract semantics that does the analysis
-when instantiated as an abstract interpreter.
+generated code. \todo{put some numbers} However, we do not lose any advantages
+of this high-level programming style (e.g., modularity and extensibility); the
+meaning of the program also remains the same, particularly, the abstract
+semantics that does the analysis when instantiated as an abstract interpreter.
 
@@
    Futamura projection to abstract interpreters.
-2) Pragmatically, staging abstract interpreters is an approach to improve performance 
-   while untouch the soundness. We study the applicability and evaluate the performance 
-   of this approach on various analyses.
+2) Pragmatically, we show that staging abstract interpreters is an approach to
+  improve performance while untouch the soundness by case studies and emprical
+  evaluation. The paper is organized as follows:
 
-The paper is organized as follows: \todo{restate this part}
-
-\begin{itemize}
-  \item Starting from a generic interface of the interpreter for a functional language
+\begin{itemize}[leftmargin=2em]
+\item Starting from a generic interface of the interpreter for a small higher-order language
   (Section~\ref{generic_if}), we first show the instantiation of concrete
-    interpretation (Section~\ref{unstaged_conc}). Then we lift it to the staged
-    version by replacing the binding-time types (Section~\ref{stagedinterp}); and to
-    the abstract interpreter by applying abstraction to the environment, store, and
-    values (Section~\ref{unstaged_abs}). Finally, we show the combination of the two
-    transformations, dubbed \textit{staged abstract interpreters}, can be readily
-    derived (Section~\ref{sai}).
-
-  \item We demonstrate that if we apply the abstract interpreter on components
-    of the program separately, it not only improves the \textit{efficiency} but
-    also the \textit{scalability} (Section~\ref{sai} \todo{}), which are two
-    major issues in static analysis.
-
-  \item We conduct two case studies to show the applicability and usefulness
-    of applying meta-programming to static analysis (Section~\ref{cases_study}).
-    (1) We first revisit the Abstract Compilation (AC)
-    \cite{Boucher:1996:ACN:647473.727587} technique (Section~\ref{cs_ac}).
-    Our approach and AC are able to achieve the same goal, however, as we show,
-    one of the advantage of staging through types is it does not require
-    whole-analyzer refactoring.
-    (2) We extend the staged abstract interpreter to two variants of
-    control-flow analysis: context-sensitive pushdown control-flow analysis and
-    \todo{store-widened?} (Section~\ref{sai}).
-
-  \item Based on the instantiation of control-flow analysis, we empirically
-    evaluate the performance improvement by staging and specialization
-    (Section~\ref{evaluation}). We compare the both context-insensitive and
-    \note{context-sensitive} analysis.
-
+  interpretation (Section~\ref{unstaged_conc}), and then lift it to the staged
+  version by replacing the binding-time types (Section~\ref{stagedinterp}).
+\item We discuss the unstaged abstract interpreter under the same framework by
+  replacing the environment, store and values (Section~\ref{unstaged_abs}).
+  Finally, we show the combination of approximation and specialization, dubbed
+  \textit{staged abstract interpreters}, can be readily derived (Section~\ref{sai}).
+\item We conduct three case studies on the applicability of our approach
+  (Section~\ref{cases_study}).
+  \begin{enumerate}
+  \item We first revisit the Abstract Compilation (AC)
+    \cite{Boucher:1996:ACN:647473.727587} (Section~\ref{cs_ac}), showing that we can
+    achieve the same optimization with least engineering effort.
+  \item We extend the staged abstract interpreter to various control-flow
+    analysis (Section \ref{cfa}), including context-sensitivity, store-widening, and abstract
+    \note{garbage collection}, showing that our approach is widely applicable.
+  \item We apply our approach to modular analysis (Section \ref{modular}),
+    i.e., analyzing the components of program separately and composing them later,
+    showing a novel application of generative programming with program analysis.
+  \end{enumerate}
+\item We empirically evaluate the performance improvement by staging on
+    control-flow analysis (Section~\ref{evaluation}). We compare the both
+    context-insensitive, \note{context-sensitive, and store-widening} analysis.
 \end{itemize}

### 2019-04-04 paper/20_prelim.tex
misc
paper revision

@@
 analyze, then present the generic interpreter shared among the four different
-semantics, after which, we instantiate the interpreter to the concrete one. Then we
+semantics, after which, we instantiate the interpreter to the concrete one.
+It is worth noting that we choose to use Scala and monad to demonstrate the
+idea, but the approach is not restricted to our choice. One can use
+imperative/direct style and other MSP languages (e.g., MetaOCaml
+\cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} and Template
+Haskell \cite{Sheard:2002:TMH:636517.636528} to construct such staged abstract
+interpreters.
+
+\iffalse
+Then we
 will proceed in two directions: by staging it we obtain a staged concrete interpreter
@@
 interpreter (Section~\ref{unstaged_abs}).
-
-We choose to use Scala and monad to demonstrate the idea, but the approach
-is not restricted to our choice. One can use imperative style and other MSP languages 
-(such as MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} 
-or Template Haskell \cite{Sheard:2002:TMH:636517.636528} \todo{cite the recent one})
-to construct such staged abstract interpreters.
+\fi
 
@@
 
-\vspace{-1em}
+%\vspace{-1em}
 \begin{figure}[h!]
@@
 \end{figure}
-\vspace{-1em}
+%\vspace{-1em}
 

### 2019-04-04 paper/40_absinterp.tex
paper revision

@@
 \begin{lstlisting}
-  type Result = (R[List[(Value, Store)]], R[Cache])
+  type Result = (R[Set[(Value, Store)]], R[Cache])
 \end{lstlisting}
@@
 This observation leads us to the type definition of @Result@: in concrete setting,
-the @Result@ is a pair of @Value@ and @Store@, now we use a list of such pairs
-@List[(Value, Store)]@ to represent multiple results. The results is also linked
+the @Result@ is a pair of @Value@ and @Store@, now we use a set of such pairs
+@Set[(Value, Store)]@ to represent multiple results. The results is also linked
 with the resulting cache used in this iteration.
@@
 further add three more effects based on it: the nondeterminism effect is
-represented by the @ListT[M[_], A]@ monad transformer, inside of which, we have
+represented by the @SetT[M[_], A]@ monad transformer, inside of which, we have
 a @ReaderT@ for a @in@ cache, and a @StateT@ for an @out@ cache. \todo{highlight
@@
     type R[T] = T
-    type AnsM[T] = ReaderT[StateT[ListT[ReaderT[StateT[IdM, Cache, ?], Cache, ?], ?], Store, ?], Env, T]
+    type AnsM[T] = ReaderT[StateT[SetT[ReaderT[StateT[IdM, Cache, ?], Cache, ?], ?], Store, ?], Env, T]
     ...
@@
 
-The sketch implementation of @ListT@ transformer is shown below. The value @run@
-it encapsulates is of type @M[List[A]]@, where @M[_]@ is the inner monad, and
+The sketch implementation of @SetT@ transformer is shown below. The value @run@
+it encapsulates is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and
 @A@ the element type of the list. The @flatMap@ method first unwraps the inner
-monadic value @M[List[A]]@ and obtain a value of @List[A]@, then applies @f@ to
+monadic value @M[SetT[A]]@ and obtain a value of @Set[A]@, then applies @f@ to
 each element in that list, and finally folds the results back to a single
-@List[M, B]@.
+@SetT[M, B]@.
 
 \begin{lstlisting}
-  case class ListT[M[_]: Monad, A](run: M[List[A]]) {
-    def flatMap[B](f: A => ListT[M, B]): ListT[M, B] = ...
+  case class SetT[M[_]: Monad, A](run: M[Set[A]]) {
+    def flatMap[B](f: A => SetT[M, B]): SetT[M, B] = ...
   }
@@
 using @mplus@ from @MonadPlus@, which requires the values are join-semilattices
-(in our case, @List@s and @Map@s).
+(in our case, @Set@s and @Map@s).
 
@@
 we can not directly destruct it by pattern matching. Instead, we use an
-auxiliary function @lift_nd@ that takes a list and lifts the elements into the
+auxiliary function @lift_nd@ that takes a set and lifts the elements into the
 monad stack. Then we can still implement @ap_clo@ in monadic style, where the
-@CloV@ comes from the @ListT@ and the nondeterminism is naturally handled.
+@CloV@ comes from the @SetT@ and the nondeterminism is naturally handled.
 
 \begin{lstlisting}
-  def lift_nd[T](vs: List[T]): AnsM[T]
+  def lift_nd[T](vs: Set[T]): AnsM[T]
   def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
-    CloV(Lam(x, e), ρ) <- lift_nd[AbsValue](rator.toList)
+    CloV(Lam(x, e), ρ) <- lift_nd[AbsValue](rator)
     α <- alloc(x)
@@
     rt <- if (out.contains(cfg)) for {
-            (v, σ) <- lift_nd[(Value, Store)](out(cfg).toList)
+            (v, σ) <- lift_nd[(Value, Store)](out(cfg))
              _ <- put_store(σ)

### 2019-04-04 paper/50_stageabs.tex
paper revision

@@
 following code shows this change. For the readability, we squeeze the three
-inner transformers into a single monad @RepListReaderStateM[A, B, C]@, where @A@
+inner transformers into a single monad @RepSetReaderStateM[A, B, C]@, where @A@
 and @B@ are both @Cache@ when used in @AnsM@. Now the result type is a pair of
-two staged value: @Rep[List[(Value, Store)]]@ and @Rep[Cache]@.
+two staged value: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
 
@@
     type R[T] = Rep[T]
-    type RepListReaderStateM[A, B, C] = RepListT[RepReaderT[RepStateT[RepIdM, A, ?], B, ?], C]
-    type AnsM[T] = RepReaderT[RepStateT[RepListReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
-    type Result = (Rep[List[(Value, Store)]], Rep[Cache])
+    type RepSetReaderStateM[A, B, C] = RepSetT[RepReaderT[RepStateT[RepIdM, A, ?], B, ?], C]
+    type AnsM[T] = RepReaderT[RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
+    type Result = (Rep[Set[(Value, Store)]], Rep[Cache])
     ...
@@
 
-Since our monad stack is uniformly using staged data, so the staged @ListT@ now
-stores a staged value of type @Rep[List[A]]@, inside of a staged monad @M@.
-\todo{staged ListT implementation}
+Since our monad stack is uniformly using staged data, so the staged @SetT@ now
+stores a staged value of type @Rep[Set[A]]@, inside of a staged monad @M@.
+\todo{staged SetT implementation}
 
 \begin{lstlisting}
-  case class ListT[M[_]: RepMonad, A](run: M[List[A]]) {
-    def flatMap[B: Manifest](f: Rep[A] => ListT[M, B]): ListT[M, B] = ...
+  case class SetT[M[_]: RepMonad, A](run: M[Set[A]]) {
+    def flatMap[B: Manifest](f: Rep[A] => SetT[M, B]): SetT[M, B] = ...
   }
@@
   def emit_compiled_clo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
-                           => Rep[(List[(Value,Store)], Cache)]): Rep[AbsValue]
+                           => Rep[(Set[(Value,Store)], Cache)]): Rep[AbsValue]
   def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(x, e) = λ
-    val f: (Rep[Value],Rep[Store],Rep[Cache],Rep[Cache]) => Rep[(List[(Value,Store)],Cache)] = {
+    val f: (Rep[Value],Rep[Store],Rep[Cache],Rep[Cache]) => Rep[(Set[(Value,Store)],Cache)] = {
       case (arg, σ, in, out) =>
@@
 \begin{lstlisting}
-  def lift_nd[T](vs: Rep[List[T]]): AnsM[T]
+  def lift_nd[T](vs: Rep[Set[T]]): AnsM[T]
   def emit_ap_clo(rator: Rep[AbsValue], rand: Rep[Value], σ: Rep[Store],
-                  in: Rep[Cache], out: Rep[Cache]): Rep[(List[(Value, Store)], Cache)]
+                  in: Rep[Cache], out: Rep[Cache]): Rep[(Set[(Value, Store)], Cache)]
   def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
-    σ <- get_store;     clo <- lift_nd[AbsValue](rator.toList)
+    σ <- get_store;     clo <- lift_nd[AbsValue](rator)
     in <- ask_in_cache; out <- get_out_cache
-    val res: Rep[(List[(Value, Store)], Cache)] = emit_ap_clo(clo, rand, σ, in, out)
+    val res: Rep[(Set[(Value, Store)], Cache)] = emit_ap_clo(clo, rand, σ, in, out)
     _ <- put_out_cache(res._2)
@@
     val cfg: Rep[Config] = (unit(e), ρ, σ)
-    val res: Rep[(List[(Value, Store)], Cache)] =
-      if (out.contains(cfg)) (out(cfg).toList, out) // a next-stage if expression
+    val res: Rep[(Set[(Value, Store)], Cache)] =
+      if (out.contains(cfg)) (out(cfg), out) // a next-stage if expression
       else { val m: Ans = for {
@@
   case class IRCompiledClo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
-                           => Rep[(List[(Value, Store)], Cache)]) extends Rep[AbsValue]
+                           => Rep[(Set[(Value, Store)], Cache)]) extends Rep[AbsValue]
   case class IRApClo(clo: Rep[AbsValue], arg: Rep[Value], σ: Rep[Store],
-                     in: Rep[Cache], out: Rep[Cache]) extends Rep[(List[(Value, Store)], Cache)]
+                     in: Rep[Cache], out: Rep[Cache]) extends Rep[(Set[(Value, Store)], Cache)]
 \end{lstlisting}
@@
 
-For example, to fold a singleton list (often appears in @ListT@), e.g.,
-@List(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully apply @foldLeft@ to the list with an anonymous
+For example, to fold a singleton set (often appears in @SetT@), e.g.,
+@Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
+generator would faithfully apply @foldLeft@ to the set with an anonymous
 function. But we can also utilize the algebraic property of @foldLeft@ to
-generate cheaper code: @List(x).foldLeft(init)(f) = f(init, x)@. Since the
+generate cheaper code: @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
 function @f@ is known at current stage, we completely eliminate the fold
 operation and function application. We apply several rewritings enabled by
-partial-static data structures, such as @List@ and @Map@, which greatly reduces
+partial-static data structures, such as @Set@ and @Map@, which greatly reduces
 the size of residual programs.

### 2019-04-04 paper/60_cases.tex
paper revision

@@
 
-\paragraph{Original AC} \todo{WIP}
-
+\paragraph{Original AC}
 The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
-for a CPS language consisting of lambda terms, applications, @letrec@ and
-primtive operators. The analyses for different syntactic constructs are
-decomposed into different functions, such as @analyzeCall@ and @analyzeApp@. The
-idea of closure generation is to rewrite these functions. Where previously they
-may take both static arguments and dynamic arguments, after the rewrite only the
-static arguments is taken. In this case, the static arguments are syntactic
-terms; the dynamic arguments are stores. After written in AC style, functions
-like @compCall@ (compiled version of @analyzeCall@) return a value of type
-@CompAnalysis@, i.e., a closure that takes a store and returns a store. The
-result of multiple calls on such functions, for example, @compCall@ and
-@compArgs@, can be composed. The generated closure only takes stores, because
-the input program is specialized into the closure. The code is shown in
-Figure~\ref{compare_ac_sai} (left).
+for a small CPS language consisting of abstractions, applications, recursion and
+primtive operators. Figure \ref{compare_ac_sai} (left) shows the AC
+implementation: different syntactic constructs are handled in different functions
+(e.g., @compCall@, @compApp@, etc.).
+Before refactoring into AC, these functions would take the store as a
+regular argument; now every function returns a value of type @CompAnalysis@,
+which is a function (closure) value that takes and returns stores. After the
+first run of analysis, we have traversed the AST and obtained a single closure
+takes a store and returns a store. Therefore, the ASTs are static terms, and the
+stores are dynamic terms. Now we can invoke the compiled closure with an initial
+store to complete the analysis.
 
@@
 On the other side, our approach does exactly the same thing through staging: the
-syntactic terms are static, and stores are dynamic, therefore the generated code
-just looks-up and updates the store. Figure~\ref{compare_ac_sai} (right) shows
-the code written with LMS. In fact, the only changes are the type of @Store@ is
-replaced with @Rep[Store]@ indicating that the values of type @Store@ will be
-known at the next stage. Indeed, additional engineering efforts are required to
-make this happen, including: the staged version @Map@ which is already included
-in LMS; implicit @lift@ function that transform a current-stage constant value
-to next stage; next-stage representation of the syntactic terms, i.e., proper
-@toString@ functions of AST structs. We consider these efforts relatively small,
-and they do not interfere the actual analysis we desire.
+AST are static, and stores are dynamic, therefore we annotate the type @Store@
+with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
+changes of our appraoch using LMS framework are the types of @Store@ to
+@Rep[Store]@, since they will be known at the next stage. The generated code
+will just looks-up and update the store. The functions at current stage are
+eliminated in the residual program, therefore we do not generate higher-order
+functions, which provides additional performance improvement. Nevertheless,
+the support from LMS framework is necessary, such as staging support for @Map@,
+we consider these efforts are reusable and complete orthogonal to the
+implementation of analysis.
 
@@
 
-The target language we presented in Section~\ref{bg_lang} is essentially a
-higher-order functional language. One fundamental analysis task for functional
-programs is control-flow analysis, i,e., determining which functions will
-possibly be applied at each call-site. The abstract interpreter we used in
-Section~\ref{unstaged_abs} and Section~\ref{sai} is a store-widened 0-CFA-like
-abstraction, moreover it is also a pushdown control-flow analysis; in the last
-section, we also reviewed AC with finite-state 0-CFA. In this section, based on
-the existing staged abstract interpreter, we further develop the staging
-techniques with control-flow analyses, including recoverying a more precise
-store model and adding context-sensitivity to the analysis.
+In this section, to demonstrate the SAI approach is widely applicable, we extend
+and refactor the staged abstract interpreter we presented in Section \ref{sai}
+to various control-flow analysis (CFA). We first discuss calling-context
+sensitive analysis for staged analyzers; then we will see how to implement
+store-widening with staging, which is a common optimization in CFA; finally, we
+integrate abstract garbage collection with staging, \todo{where the addresses to
+  be collected can be computed at the staging time}.
+
+\subsubsection{Context-sensitivity}
+
+\subsubsection{Store-widened Analysis}
+
+\subsubsection{Abstract Garbage Collection at Staging Time}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
 
@@
 
-\subsection{Modular Analysis for Free}
+\subsection{Modular Analysis for Free} \label{modular}
 

### 2019-04-04 oopsla19-code/src/main/scala/sai/evaluation/GenericAnalyzer.scala
paper

@@ -53,7 +53,7 @@ trait SchemeAnalyzer {
   def br(ev: EvalFun)(test: Expr, thn: Expr, els: Expr): Ans
   def close(ev: EvalFun)(λ: Lam, ρ: R[Env]): R[Value]
   def ap_clo(ev: EvalFun)(fun: R[Value], arg: R[List[Value]]): Ans
-  def primtives(v: R[Value], args: List[R[Value]]): R[Value]
+  def primitives(v: R[Value], args: List[R[Value]]): R[Value]
 
   // Fixpoint wrapper and top-level interface
   def fix(ev: EvalFun => EvalFun): EvalFun

### 2019-04-04 oopsla19-code/src/main/scala/sai/evaluation/StagedAnalyzer.scala

@@ -82,11 +82,6 @@ trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with Rep
   def get(ρ: Rep[Env], x: String): Rep[Addr] = ρ(x)
   def get(σ: Rep[Store], ρ: Rep[Env], x: String): Rep[Value] = σ(ρ(x))
   def br(ev: EvalFun)(test: Expr, thn: Expr, els: Expr): Ans =
-    /*for {
-     v1 <- ev(thn)
-     v2 <- ev(els)
-     } yield v1 ++ v2
-     */
     ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(ev(thn), ev(els))
   def close(ev: EvalFun)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(params, e) = λ
@@ -110,7 +105,7 @@ trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with Rep
               lift_nd[(Value, Store)](res._1).flatMap { vs =>
                 put_store(vs._2).map { _ => vs._1 }
               } } } } } } }
-  def primtives(v: Rep[Value], args: List[Rep[Value]]): Rep[Value] = ???
+  def primitives(v: Rep[Value], args: List[Rep[Value]]): Rep[Value] = ???
 
   // auxiliary function that lifts values
   def lift_nd[T: Manifest](vs: Rep[Set[T]]): AnsM[T] =

### 2019-04-04 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -46,23 +46,10 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
   def put_store(σ: Store): AnsM[Unit] =
     ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].put(σ))
   def set_store(αv: (Addr, Value)): AnsM[Unit] =
-    ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].mod(σ => {
-                                                                                         val news = σ ⊔ Map(αv)
-                                                                                         /*
-                                                                                         val oldc = σ.getOrElse(αv._1, Set())
-                                                                                         val newc = news(αv._1)
-                                                                                         if (oldc != newc) {
-                                                                                           //println(s"growing: ${αv._1} oldsize: ${oldc.size} newsize: ${newc.size}")
-                                                                                           //println(s"DIFF ${newc -- oldc}")
-                                                                                         }
-                                                                                         */
-                                                                                         news
-                                                                                       }))
+    ReaderT.liftM[StoreNdInOutCacheM, Env, Unit](StateTMonad[NdInOutCacheM, Store].mod(σ => { σ ⊔ Map(αv) }))
 
   // allocate addresses
-  def alloc(σ: Store, x: String): Addr = {
-    ZCFAAddr(x)
-  }
+  def alloc(σ: Store, x: String): Addr = { ZCFAAddr(x) }
   def alloc(x: String): AnsM[Addr] = for { σ <- get_store } yield alloc(σ, x)
 
   // Primitive operations
@@ -107,7 +94,7 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
     */
   } yield v
 
-  def primtives(v: Value, args: List[Value]): Value = ???
+  def primitives(v: Value, args: List[Value]): Value = ???
 
   def foldVss(vss: List[Value]): Value = vss.foldLeft(Lattice[Value].bot)(_ ⊔ _)
 

### 2019-04-05 paper/10_intro.tex
paper

@@
 analyzing its binding-times is hard in general. As an alternative and pragmatic
-approach to specialization and partial evaluation, multi-stage programming (MSP
-for short) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
+approach to specialization and partial evaluation, multi-stage programming (MSP) 
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
 annotations (i.e., binding-time annotations) to be explicit from the
@@
 \begin{itemize}
-\item Monadic abstractions, which encapsulate the following two abstractions.
-  The interpreter itself is written in monadic form, and by plugging different monads, 
-  the underlying semantics of the interpreter is changed.
-  %\todo{mention that direct-style also works?}
-  %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
 \item Env-store-value abstractions, which control whether the interpreter
@@
     \cite{DBLP:journals/jfp/CaretteKS09, Ofenbeck:2017:SGP:3136040.3136060}.
+\item Monadic abstractions, which encapsulate the above two abstractions.
+  The interpreter itself is written in monadic form, and by plugging different monads, 
+  the underlying semantics of the interpreter is changed.
+  %\todo{mention that direct-style also works?}
+  %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
 \end{itemize}
@@
 We show that these well-known abstractions can be combined smoothly, and more
-importantly the price is almost free. Through staging, the major overhead of
+importantly almost free. Through staging, the major overhead of
 abstractions -- from the monadic layers -- is completely eliminated in the
@@
 of this high-level programming style (e.g., modularity and extensibility); the
-meaning of the program also remains the same, particularly, the abstract
-semantics that does the analysis when instantiated as an abstract interpreter.
+meaning of the program also remains the same, particularly, the soundness of 
+static analysis after staging.
 
@@
 programming, in particular) with abstract interpreters is an effective approach
-to improve the latter one while the former one is untouched.
+to improve the latter while the former is untouched.
 In the literature, specializing analysis with respect to the input program
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
-ashley:practical} is not a new, despite the fact that most of the previous work
+ashley:practical} is not new, despite the fact that most of the previous work
 are ad-hoc on a particular analysis or requiring significant changes to the
@@
 \cite{Boucher:1996:ACN:647473.727587}. But it requires the whole analyzer to be
-rewritten to the closure generation from. In this paper, we advocate a systematic
+rewritten to the closure generation form. In this paper, we advocate a systematic
 approach to meta-programming that achieves the performance goal with less
@@
 
-\paragraph{Contributions} The contribution of the paper is twofold:
-1) Intellectually, we present a unifying construction that naturally extends the first 
-   Futamura projection to abstract interpreters.
-2) Pragmatically, we show that staging abstract interpreters is an approach to
-  improve performance while untouch the soundness by case studies and emprical
-  evaluation. The paper is organized as follows:
-
+\paragraph{Contributions} Briefly, the contribution of the paper is twofold:
+\begin{enumerate}[leftmargin=2em]
+  \item Intellectually, through multi-stage programming, we present a unified construction 
+    that naturally extends the first Futamura projection to abstract interpreters.
+ \item Pragmatically, by case studies and emprical evaluation, we show that staging 
+   abstract interpreters is useful to improve performance while untouch 
+   the soundness. 
+\end{enumerate}
+\paragraph{Organization} The paper is organized as follows:
 \begin{itemize}[leftmargin=2em]
 \item Starting from a generic interface of the interpreter for a small higher-order language
-  (Section~\ref{generic_if}), we first show the instantiation of concrete
+  (Section~\ref{prelim}), we first show the instantiation of concrete
   interpretation (Section~\ref{unstaged_conc}), and then lift it to the staged
@@
     \cite{Boucher:1996:ACN:647473.727587} (Section~\ref{cs_ac}), showing that we can
-    achieve the same optimization with least engineering effort.
+    achieve the same optimization with much less engineering effort.
   \item We extend the staged abstract interpreter to various control-flow

### 2019-04-05 paper/20_prelim.tex
paper

@@
 
-In this section, we first describe the abstract syntax of the language we will
-analyze, then present the generic interpreter shared among the four different
+In this section, we first describe the abstract syntax of the language for our interpreter, 
+then present the generic interpreter shared among the four different
 semantics, after which, we instantiate the interpreter to the concrete one.
@@
 will add more imperative features to the language. Since we are mostly
-interested in analyzing the dynamic behaviors of the program, we disguised any
+interested in analyzing the dynamic behaviors of the program, we disguise any
 static semantics and type system. We also assume that input programs are
@@
 The abstract syntax we present in fact can be seen as a deep embedding of the
-language -- we use data-types to represent programs. This design choice makes us
-easily add different interpretations over the AST; with the inheritance and
+language -- we use data-types to represent programs. This design choice allows us to
+easily use different interpretations over the AST; with the inheritance and
 overriding mechanism in Scala, we may also add new language construct and reuse
@@
 Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
-takes a monad as argument and produces another monad. By using monad
+take a monad as argument and produces another monad. By using monad
 transformers, we can combine multiple monads into a single one. Constructing
@@
 interpreters \cite{Sergey:2013:MAI:2491956.2491979,
-DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}. In section,
-with the multi-stage programming and monad transformers in mind while leave them
+DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}. In this section,
+with the multi-stage programming and monad transformers in mind while leaving them
 as abstract type members, we present the generic interface of a big-step
@@
 We also defines several methods to obtain and modify the environment and store.
-These methods either returns monadic values with environment or store inside, or
-simply a @Unit@ value for effects. For example, @local_env@ takes a value @ans@ of
-type @Ans@ and a new environment which will be installed when evaluating @ans@.
+These methods return monadic values of type @AnsM[_]@, which encapsulate the 
+environment or store, or simply a @Unit@ value for effects.
+For example, @local_env@ takes a value @ans@ of type @Ans@ and an environment 
+which will be installed when evaluating @ans@.
 
@@
 Here we sketch the basic idea of unstaged @ReaderT@ and @StateT@. Readers may
-refer to \todo{cite, or appendix?} for more detail and implementation of monad
-transformers.
+refer to \cite{DBLP:conf/popl/LiangHJ95, Chiusano:2014:FPS:2688794} for more detail 
+and implementation of monad transformers.
 A @ReaderT@ monad transformer encapsulates the computation @R => M[A]@, where

### 2019-04-05 paper/70_eval.tex
misc
paper

@@
 
-We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
+We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
 (kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
@@
 \toprule
-    Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
-    @rsa@            &      &            &            & \%          \\
-    @kcfa3@          &      &            &            & \%          \\
-    @church@         &      &            &            & \%          \\
-    @fermat@         &      &            &            & \%          \\
-    @kcfa-worst-16@  &  198 & 0.352      & 0.028      & +1157.14\%        \\
-    @kcfa-worst-32@  &  390 & 4.234      & 0.289      & +1365.05\%          \\
-    @kcfa-worst-64@  &  774 & 61.475     & 4.425      & +1289.27\%         \\
+    program          &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
+    @church@         & 120  & 71.811     & 23.990     & 299.34\%          \\
+    @rsa@            & 451  & 0.308      & 0.064      & 459.15\%          \\
+    @latice@         & 609  & 31.874     & 8.816      & 358.69\%          \\
+    @fermat@         & 310  & 0.043      & 0.006      & 716.67\%          \\
+    @mbrotZ@         & 331  & 1.112      & 0.214      & 519.63\%          \\
+    @kcfa-worst-16@  & 182  & 0.083      & 0.013      & 638.46\%          \\
+    @kcfa-worst-32@  & 358  & 0.452      & 0.044      & 1027.27\%         \\
+    @kcfa-worst-64@  & 774  & 3.329      & 0.236      & 1410.59\%         \\
     \bottomrule

### 2019-04-05 paper/abstract.tex
paper

@@
 interpreter to a given program produces an equivalent program that runs faster,
-which known as the first Futamura projection. It is even more widely known that an
+which is known as the first Futamura projection. It is even more widely known that an
 abstract interpreter is a program analyzer: tweaking the interpreter to run on a

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -26,8 +26,9 @@ object Evaluation {
       //(fermat, "fermat"),
       //(rsa, "rsa"),
       //(mbrotZ, "mbrotZ"),
+      ////(lattice, "lattice"),
       /********************/
-      (lattice, "lattice"),
+      (matrix, "matrix"),
       //(letloop, "letloop")
       //(euclid, "euclid"),
       //(euclid_imp, "euclid_imp"),
@@ -43,22 +44,19 @@ object Evaluation {
       //(earley, "earley"),
       //(dynamic, "dynamic"),
       //(graphs, "graphs"),
-      //(matrix, "matrix"),
       //(nbody, "nbody"),
       //(nucleic, "nucleic"),
     )
-    progs foreach { case (e, id) => compare(e, id) }
-    /*
+    //progs foreach { case (e, id) => compare(e, id) }
     progs foreach { case (e, id) => 
       println(s"$id - ${size(e)}")
       println(e)
       println(ASTUtils.exprToString(e))
       println(ASTUtils.free(e).map("\"" + _ + "\""))
       println(ASTUtils.free(e).map("\"" + _ + "\" =>\n"))
-      val t1 =  run(1, { evalUnstaged(e) })
-      println(s"[$id] [unstaged] - ${t1}s")
+      //val t1 =  run(1, { evalUnstaged(e) })
+      //println(s"[$id] [unstaged] - ${t1}s")
     }
-    */
     /*
     progs foreach { case (e, id) =>
       val code = specialize(e)
@@ -114,11 +112,9 @@ object Evaluation {
 
   def compare(e: Expr, id: String) {
     println(s"Running $id, AST size: ${size(e)}")
-    val N = 1
-    /*
+    val N = 20
     val t1 =  run(N, { evalUnstaged(e) })
     println(s"[$id] [unstaged] - ${t1}s")
-    */
     val code = specialize(e)
     code.precompile
     val outfile = output(id)

### 2019-04-05 paper/60_cases.tex
evaluation

@@
 We first revisit a similar technique called \textit{abstract compilation} (AC).
-AC is introduced by \citet{Boucher:1996:ACN:647473.727587} as a new
+AC was introduced by \citet{Boucher:1996:ACN:647473.727587} as an
 implementation technique for abstract interpretation based static analysis.
@@
 changes of our appraoch using LMS framework are the types of @Store@ to
-@Rep[Store]@, since they will be known at the next stage. The generated code
+@Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
+The generated code
 will just looks-up and update the store. The functions at current stage are

### 2019-04-05 paper/70_eval.tex
evaluation

--- 2019-03-30 paper/70_eval.tex
@@
 \paragraph{Setting}
-
-We use Scala 2.12.4 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
+We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
 (kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
@@
 
-\paragraph{Benchmarks} \todo{Rewrite}
-The benchmark programs we used are collected from previous papers
+\paragraph{Benchmarks}
+The benchmark programs we used in the experiment are collected from several previous papers
 \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
@@
 \footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
+Some of the benchmarks are small programs designed for experiments, such as the @kcfa-worst-@$n$ series, 
+which are intended to challenge $k$-CFA; while some other benchmarks are from real-world applications, 
+for examples, the RSA public key encryption algorithm @rsa@ and matrix computation @matrix@.
+In the Table \ref{?}, we report the number of AST node after parsing and desugaring 
+(excluding comments) as a proper measurement of program size.
+
+\iffalse
 We used the following benchmark programs:
-
 \begin{itemize}
@@
 \end{itemize}
+\fi
 
@@
 \toprule
-    Program          &\#AST & unstaged   & staged     & $\Delta$ \\ \midrule
-    @rsa@            &      &            &            & \%          \\
-    @kcfa3@          &      &            &            & \%          \\
-    @church@         &      &            &            & \%          \\
-    @fermat@         &      &            &            & \%          \\
-    @kcfa-worst-16@  &  198 & 0.352      & 0.028      & +1157.14\%        \\
-    @kcfa-worst-32@  &  390 & 4.234      & 0.289      & +1365.05\%          \\
-    @kcfa-worst-64@  &  774 & 61.475     & 4.425      & +1289.27\%         \\
+    program          &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
+    @church@         & 120  & 71.811     & 23.990     & 299.34\%          \\
+    @rsa@            & 451  & 0.308      & 0.064      & 459.15\%          \\
+    @lattice@        & 609  & 31.874     & 8.816      & 358.69\%          \\
+    @fermat@         & 310  & 0.043      & 0.006      & 716.67\%          \\
+    @mbrotZ@         & 331  & 1.112      & 0.214      & 519.63\%          \\
+    @kcfa-worst-16@  & 182  & 0.083      & 0.013      & 638.46\%          \\
+    @kcfa-worst-32@  & 358  & 0.452      & 0.044      & 1027.27\%         \\
+    @kcfa-worst-64@  & 774  & 3.329      & 0.236      & 1410.59\%         \\
     \bottomrule

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -54,8 +54,8 @@ object Evaluation {
       println(ASTUtils.exprToString(e))
       println(ASTUtils.free(e).map("\"" + _ + "\""))
       println(ASTUtils.free(e).map("\"" + _ + "\" =>\n"))
-      //val t1 =  run(1, { evalUnstaged(e) })
-      //println(s"[$id] [unstaged] - ${t1}s")
+      val t1 =  run(1, { evalUnstaged(e) })
+      println(s"[$id] [unstaged] - ${t1}s")
     }
     /*
     progs foreach { case (e, id) =>

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/TestPrograms.scala

@@ -6,7 +6,8 @@ import sai.evaluation.parser._
 object TestPrograms {
   def getAST(prog: String) = {
     SchemeParser(prog) match {
-      case Some(expr) => SchemeASTDesugar(expr)
+      case Some(expr) => 
+        ASTUtils.alpha(SchemeASTDesugar(expr), Map())._1
     }
   }
 

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -147,6 +147,23 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
       , "append" -> Set(ListVTop)
       , "else" -> Set(BoolV)
       , "list" -> Set(ListVTop)
+      // matrix
+      , "child" -> Set(SymV)
+      , "now" -> Set(SymV)
+      , "puke" -> Set(SymV)
+      , "brother" -> Set(SymV) //matrix symbols
+      , "caar" -> Set(IntV, FloatV, CharV, BoolV)
+      , "for-each" -> Set()
+      , "map" -> Set(ListVTop)
+      , "expt" -> Set(IntV)
+      , "even?" -> Set(BoolV)
+      , "length" -> Set(IntV)
+      , "reverse" -> Set(ListVTop)
+      , "cadr" -> Set(ListVTop)
+      , "vector-ref" -> Set(IntV, FloatV, CharV, BoolV)
+      , "cddr" -> Set(IntV, FloatV, CharV, BoolV)
+      , "zero?" -> Set(BoolV)
+      ,
   )
 
   def primitives(ev: EvalFun)(x: String, args: List[Expr]): Ans = {
@@ -206,10 +223,11 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
     val cfg = (e, ρ, σ)
     in <- ask_in_cache
     out <- get_out_cache
+    val _ = println(s"Eval out: ${out.size}")
     rt <- if (out.contains(cfg)) {
       //val _ = println(s"MISS ${ASTUtils.exprToString(e)}, ρ: ${ρ.size}, σ: ${σ.size}, out: ${out.size}")
       for {
-        _ <- put_out_cache(out)
+        //_ <- put_out_cache(out)
         (v, s) <- lift_nd[(Value, Store)](out(cfg))
         _ <- put_store(s)
       } yield v

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/parser/AST.scala

@@ -125,5 +125,49 @@ object ASTUtils {
     }
     fv -- defVars(e)
   }
-}
 
+  var count = 1
+  def fresh(pre: String = "var"): String = {
+    val name = pre + count.toString
+    count = count + 1
+    name
+  }
+  def alpha(e: Expr, map: Map[String, String]): (Expr, Map[String, String]) = {
+    e match {
+      case Var(x) =>
+        (if (map.contains(x)) Var(map(x)) else {
+           //println(s"$x not exist!")
+           e
+         }, map)
+      case Void() => (e, map)
+      case Sym(x) => (e, map)
+      case CharLit(x)   => (e, map)
+      case IntLit(x)    => (e, map)
+      case FloatLit(x)  => (e, map)
+      case BoolLit(x)   => (e, map)
+      case Set_!(x, e)  =>
+        val (ne, nm) = alpha(e, map)
+        (Set_!(map(x), ne), map)
+      case Define(x, e) =>
+        val name = if (map.contains(x)) map(x) else fresh()
+        val nm = map + (x -> name)
+        val (ne, nnm) = alpha(e, nm)
+        (Define(name, ne), nm)
+      case App(x, l)    => (App(alpha(x, map)._1, l.map(alpha(_, map)._1)), map)
+      case Begin(es)    =>
+        val defmap = (es.filter(_.isInstanceOf[Define]).map { case Define(x, _) => (x, fresh()) }).toMap
+        val (res, mm) = es.foldLeft((List[Expr](), defmap ++ map)) {
+          case ((nes, m), e) =>
+            val (ne, nm) = alpha(e, m)
+            (nes ++ List(ne), nm)
+        }
+        (Begin(res), map)
+      case If(c, t, e)  => (If(alpha(c, map)._1, alpha(t, map)._1, alpha(e, map)._1), map)
+      case Lam(params, body) =>
+        val nparams = params.map(x => (x, fresh("var")))
+        val nmap = map ++ nparams.toMap
+        val (nbody, _) = alpha(body, nmap)
+        (Lam(nparams.map(_._2), nbody), map)
+    }
+  }
+}

### 2019-04-05 paper/70_eval.tex
evaluation

@@
     program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
+    @rsa@               & 451  & 0.308      & 0.064      & 459.15\%          \\
     @church@            & 120  & 71.811     & 23.990     & 299.34\%          \\
-    @rsa@            & 451  & 0.308      & 0.064      & 459.15\%          \\
-    @lattice@        & 609  & 31.874     & 8.816      & 358.69\%          \\
     @fermat@            & 310  & 0.043      & 0.006      & 716.67\%          \\
     @mbrotZ@            & 331  & 1.112      & 0.214      & 519.63\%          \\
+    @lattice@           & 609  & 31.874     & 8.816      & 358.69\%          \\
     @kcfa-worst-16@     & 182  & 0.083      & 0.013      & 638.46\%          \\
@@
     @kcfa-worst-64@     & 774  & 3.329      & 0.236      & 1410.59\%         \\
+    @solovay-strassen@  & 523  & 5.94       & 1.189      & 428.43\%          \\
     \bottomrule

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -26,9 +26,15 @@ object Evaluation {
       //(fermat, "fermat"),
       //(rsa, "rsa"),
       //(mbrotZ, "mbrotZ"),
-      ////(lattice, "lattice"),
+      //(lattice, "lattice"),
+      //(blur, "blur"), // 56
+      //(sat, "sat"), // 73
+      //(solovay, "solovay")
+      //(metacirc, "meta")
+      (regex, "regex"),
+      //(scheme2java, "scheme2java")
       /********************/
-      (matrix, "matrix"),
+      //(matrix, "matrix"),
       //(letloop, "letloop")
       //(euclid, "euclid"),
       //(euclid_imp, "euclid_imp"),
@@ -36,10 +42,7 @@ object Evaluation {
       //(kcfa3, "kcfa3"),
       //(kcfa256, "kcfa256"),
       //(omega, "omega")
-      //(church_exp, "church_exp")
-      //(blur, "blur"), // 56
       //(fib, "fib"), // 36
-      //(sat, "sat"), // 73
       //(boyer, "boyer"),
       //(earley, "earley"),
       //(dynamic, "dynamic"),
@@ -115,6 +118,7 @@ object Evaluation {
     val N = 20
     val t1 =  run(N, { evalUnstaged(e) })
     println(s"[$id] [unstaged] - ${t1}s")
+    
     val code = specialize(e)
     code.precompile
     val outfile = output(id)

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/StagedAnalyzer.scala

@@ -164,6 +164,23 @@ trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with Rep
     , "append" -> Set[AbsValue](unit(ListVTop))
     , "else" -> Set[AbsValue](unit(BoolV))
     , "list" -> Set[AbsValue](unit(ListVTop))
+    , "%" -> Set[AbsValue](unit(IntV))
+    , "brother" -> Set[AbsValue](unit(SymV)) //matrix symbols
+    , "caar" -> Set[AbsValue](unit(IntV), unit(FloatV), unit(CharV), unit(BoolV))
+    , "for-each" -> Set[AbsValue]()
+    , "map" -> Set[AbsValue](unit(ListVTop))
+    , "expt" -> Set[AbsValue](unit(IntV))
+    , "even?" -> Set[AbsValue](unit(BoolV))
+    , "length" -> Set[AbsValue](unit(IntV))
+    , "reverse" -> Set[AbsValue](unit(ListVTop))
+    , "cadr" -> Set[AbsValue](unit(ListVTop))
+    , "vector-ref" -> Set[AbsValue](unit(IntV), unit(FloatV), unit(CharV), unit(BoolV))
+    , "cddr" -> Set[AbsValue](unit(IntV), unit(FloatV), unit(CharV), unit(BoolV))
+    , "zero?" -> Set[AbsValue](unit(BoolV))
+    , "symbol?" -> Set[AbsValue](unit(BoolV))
+    , "equal?" ->  Set[AbsValue](unit(BoolV))
+    , "pair?" ->   Set[AbsValue](unit(BoolV))
+    , "char?" ->   Set[AbsValue](unit(BoolV))
   )
 
   def primitives(ev: EvalFun)(x: String, args: List[Expr]): Ans = {

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/TestPrograms.scala

@@ -7,7 +7,8 @@ object TestPrograms {
   def getAST(prog: String) = {
     SchemeParser(prog) match {
       case Some(expr) => 
-        ASTUtils.alpha(SchemeASTDesugar(expr), Map())._1
+        SchemeASTDesugar(expr)
+        //ASTUtils.alpha(SchemeASTDesugar(expr), Map())._1
     }
   }
 
@@ -132,4 +133,11 @@ object TestPrograms {
   // splay -- old match
   def splay = getAST(Source.fromFile("benchmarks/toplas98/splay.scm").mkString)
 
+  def regex = getAST(Source.fromFile("benchmarks/regex-derivative.scm").mkString)
+
+  def metacirc = getAST(Source.fromFile("benchmarks/meta-circ.scm").mkString)
+
+  def solovay = getAST(Source.fromFile("benchmarks/solovay-strassen.scm").mkString)
+
+  def scheme2java = getAST(Source.fromFile("benchmarks/scheme2java.scm").mkString)
 }

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -163,7 +163,11 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
       , "vector-ref" -> Set(IntV, FloatV, CharV, BoolV)
       , "cddr" -> Set(IntV, FloatV, CharV, BoolV)
       , "zero?" -> Set(BoolV)
-      ,
+          , "%" -> Set(BoolV)
+          , "symbol?" -> Set(BoolV)
+          , "equal?" -> Set(BoolV)
+          , "pair?" -> Set(BoolV)
+          , "char?" -> Set(BoolV)
   )
 
   def primitives(ev: EvalFun)(x: String, args: List[Expr]): Ans = {
@@ -223,7 +227,7 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
     val cfg = (e, ρ, σ)
     in <- ask_in_cache
     out <- get_out_cache
-    val _ = println(s"Eval out: ${out.size}")
+    //val _ = println(s"Eval out: ${out.size}")
     rt <- if (out.contains(cfg)) {
       //val _ = println(s"MISS ${ASTUtils.exprToString(e)}, ρ: ${ρ.size}, σ: ${σ.size}, out: ${out.size}")
       for {

### 2019-04-05 paper/10_intro.tex
paper

@@
     that naturally extends the first Futamura projection to abstract interpreters.
- \item Pragmatically, by case studies and emprical evaluation, we show that staging 
+ \item Pragmatically, by case studies and empirical evaluation, we show that staging 
    abstract interpreters is useful to improve performance while untouch 

### 2019-04-05 paper/70_eval.tex
paper

@@
 and evaluate the analyzing time against with the unstaged versions. The result
-shows an average \todo{X}\% running time improvement, as a strong evidence of
-our \textit{abstraction without regret} approach to construct abstract
-interpreters.
+shows that the staged versions are averagely 6 times as fast of the unstaged
+version, as a strong evidence of our \textit{abstraction without regret}
+approach.
 
+\paragraph{Implementation and Evaluation Environment}
 We implement the abstract interpreters for control-flow analysis, i.e., a
@@
 desugars Scheme to a small core language that make the analyzer easy to
-implement. Macros and @call/cc@ is not handled. As described in Section
-\ref{cfa}, we implement a monovariant 0-CFA-like analysis, a context-sensitive
-$1$-CFA-like analysis, both their versions with and without store-widening.
+implement. As described in Section \ref{cfa}, we implement a monovariant
+0-CFA-like analysis, \todo{a context-sensitive $1$-CFA-like analysis, both their
+  versions with and without store-widening}. We also implement several
+optimizations mentioned in Section \ref{staged_ds} (less than ~30 LOC in total).
 
-\paragraph{Setting}
 We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
@@
 \paragraph{Benchmarks}
-The benchmark programs we used in the experiment are collected from several previous papers
-\cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
+The benchmark programs we used in the experiment are collected from several
+previous papers \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
 DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
 \footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
-Some of the benchmarks are small programs designed for experiments, such as the @kcfa-worst-@$n$ series, 
-which are intended to challenge $k$-CFA; while some other benchmarks are from real-world applications, 
-for examples, the RSA public key encryption algorithm @rsa@ and matrix computation @matrix@.
-In the Table \ref{?}, we report the number of AST node after parsing and desugaring 
-(excluding comments) as a proper measurement of program size.
+Some of the benchmarks are small programs designed for experiments, such as the
+@kcfa-worst-@$n$ series, which are intended to challenge $k$-CFA; while some
+other benchmarks are from real-world applications, for examples, the RSA public
+key encryption algorithm @rsa@. In the Figure \ref{evaluation_result}, we report
+the number of AST node after parsing and desugaring (excluding comments) as a
+proper measurement of program size.
 
@@
 
-Figure~\ref{evaluation_result} shows the evaluation results. The
+Figure~\ref{evaluation_result} shows the evaluation result, comparing the
+performance improved by staging on 0CFA without store-widening. The
 \textit{unstaged} and \textit{staged} columns show the median time to finish the
-analysis (in seconds). The "$\Delta$" column shows the median runtime change
-from the unstaged to the staged version. \todo{0cfa, 1cfa, store-widen}
+analysis (in seconds). The last column $\frac{\text{unstaged}}{\text{staged}}$
+shows the improvement from staging. \todo{0cfa, 1cfa, store-widen}
+As we can see from the table, the staged version significantly outperforms the
+unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
+We observe the most improvement of 14x times faster than the unstaged version on
+@kcfa-worst-64@. On average, the staging produces code that run 6x times faster
+than the unstaged analyzer.
+
+We also observe that much overhead of unstaged analyzers are from the monadic
+layers. Thus our evaluation result provides an abstraction without regret
+approach to construct optimizing abstract interpreters -- user may write an
+high-level modular and composable abstract interpreter, and then derive a staged
+version in a principled way that runs significantly faster.
 

### 2019-04-05 paper/80_related.tex
paper

@@
 \citet{Sergey:2013:MAI:2491956.2491979} and
-\citet{DBLP:journals/pacmpl/DaraisLNH17}.
+\citet{DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
 Similar to monadic abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
 reconstruct big-step abstract interpreters with delimited continuations;
-\citet{Keidel:2018:CSP:3243631.3236767} present an arrow-based abstract
+\citet{Keidel:2018:CSP:3243631.3236767} present a modular arrow-based abstract
 interpreter that makes proof of soundness can be constructed compositionally.
-\cite{DBLP:conf/cc/CousotC02} proposed an
-abstract interpretation framework for modular analysis.
-\citet{DBLP:conf/popl/CalcagnoDOY09} developed compositional shape analysis by
-using bi-abduction.
+\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
+modular analysis.
 
-\paragraph{Meta-Programming} The Lightweight Modular Staging framework relies on
-types as stage annotations, and the staging procedure is modular. Other notable
-implementations of MSP exist in ML family, for examples, MetaML
-\cite{DBLP:conf/pepm/TahaS97} and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
-DBLP:conf/flops/Kiselyov14}. Compared with the LMS approach in Scala,
-MetaML/MetaOCaml use term-level annotations such as brackets, escape, and run.
-Notwithstanding, we use LMS in this paper, the idea of staging an abstract
-interpreter still applies with other MSP implementations. Partial evaluation as
-an automatic technique is studied comprehensively
-\cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}. Futamura projections
-reveals close relations between interpreters and compilers \cite{Futamura1999,
-futamura1971partial}. \citet{DBLP:conf/gpce/Asai14} studies the compilation of
-reflective language using MSP and MetaOCaml.
+\paragraph{Meta-Programming} In this paper, we use the Lightweight Modular
+Staging framework \cite{DBLP:conf/gpce/RompfO10}, which relies on types as stage
+annotations. Other notable implementations of MSP exist in ML family, e.g.,
+MetaML \cite{DBLP:conf/pepm/TahaS97} and MetaOCaml
+\cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
+The idea of staging an abstract interpreter we presented in this paper is also
+applicable to other MSP system. Partial evaluation as an automatic technique is
+studied comprehensively \cite{10.1007/3-540-61580-6_11,
+DBLP:books/daglib/0072559}. Futamura projections reveals close relations between
+interpreters and compilers, which was originally proposed by
+\citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial}, and later
+reprinted \cite{Futamura1999}.
 \citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of
-concrete interpreters by MSP and LMS framework.
+concrete interpreters by MSP -- it would be also interesting to explore multiple
+layers of abstract interpreters.
+\citet{10.1007/11561347_18} discussed combining
+multi-stage programming with functors and monads.
+Similar to the idea in this paper, specializing monadic interpreters was
+explored by \citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}.
 
@@
 \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
-
-MSP with functors and monads \cite{10.1007/11561347_18}
-
-staged monadic interpreters \cite{DBLP:conf/dsl/SheardBP99}
-
-compiling monads \cite{danvy1991compiling}

### 2019-04-05 paper/90_conc.tex
paper

\section{Conclusion}

In this paper, we first present the implementaion of first Futamura projection
on abstract interpreters. To realize it, we design a generic interpreter that
abstracts over both binding-time and value domains; by instantiating it
differently, the staged abstract interpreter can be readily derived. We also
conduct case studies and empirical evaluation, demonstrating a wide
applicability of our approach and evaluating the performance improved by staging.

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/TestPrograms.scala

@@ -7,8 +7,8 @@ object TestPrograms {
   def getAST(prog: String) = {
     SchemeParser(prog) match {
       case Some(expr) => 
-        SchemeASTDesugar(expr)
-        //ASTUtils.alpha(SchemeASTDesugar(expr), Map())._1
+        //SchemeASTDesugar(expr)
+        ASTUtils.alpha(SchemeASTDesugar(expr), Map())._1
     }
   }
 

### 2019-04-05 paper/70_eval.tex
typo

@@
 We implement the abstract interpreters for control-flow analysis, i.e., a
-value-flow analysis focus on lambda terms in functional languages. A frontend
+value-flow analysis focus on lambda terms in functional languages. A front-end
 desugars Scheme to a small core language that make the analyzer easy to
@@
 previous papers \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
-DBLP:journals/corr/abs-1102-3676} and exisiting artifacts
+DBLP:journals/corr/abs-1102-3676} and existing artifacts
 \footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
@@
 layers. Thus our evaluation result provides an abstraction without regret
-approach to construct optimizing abstract interpreters -- user may write an
+approach to construct optimizing abstract interpreters -- user may write a
 high-level modular and composable abstract interpreter, and then derive a staged

### 2019-04-05 paper/10_intro.tex
paper

@@
 abstractions -- from the monadic layers -- is completely eliminated in the
-generated code. \todo{put some numbers} However, we do not lose any advantages
-of this high-level programming style (e.g., modularity and extensibility); the
-meaning of the program also remains the same, particularly, the soundness of 
-static analysis after staging.
+generated code. We observe at most 14x times faster than the unstage abstract
+interpreter.
+However, we do not lose any advantages of this high-level programming style
+(e.g., modularity and extensibility); the meaning of the program also remains
+the same, particularly, the soundness of static analysis after staging.
 
-\paragraph{Performance via Specialization}
-
+%\paragraph{Performance via Specialization}
 Besides the intellectual merit shown in the confluence diagram, this paper
@@
 rewritten to the closure generation form. In this paper, we advocate a systematic
-approach to meta-programming that achieves the performance goal with less
-intrusive changes; in fact, a case study (Section~\ref{cs_ac}) comparing our
-approach with abstract compilation shows that, by utilizing type-based stage
-annotations, the analyzer program does not need any changes.
+approach of meta-programming that achieves the performance goal with less
+intrusive changes and engineering efforts.
 
-\paragraph{Modularity via Meta-Programming} \todo{compress this part}
+\paragraph{Applications}
 
+Further more, we show various applications of staging an abstract interpreter.
+(1) We compare our approach with abstract compilation, and show that we can
+achieve the same optimization by utilizing type-based stage annotations,
+meanwhile the analyzer program does not need any changes. (2) \todo{} (3)
+Staging abstract interpreter enables modular analysis and provides a mechanized
+way to create library summaries. We notice that we may apply the staged abstract
+interpreter separately on the different part of the analyzed program, the
+generated code is reusable and composable. Therefore, we can turn a
+whole-program analyzer into a analyzer that runs modularly.
+
+\iffalse
 Our next contribution is to enable modular analysis based on a whole-program
@@
 whole-program analysis.
+\fi
 

### 2019-04-05 paper/40_absinterp.tex
paper

@@
 implemented similarly, or lifted from existing lattices element-wise or
-point-wise.
+point-wise. Naturally, abstract domains such intervals also can be implemented
+in a stage polymorphic way.
 
@@
 The sketch implementation of @SetT@ transformer is shown below. The value @run@
-it encapsulates is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and
-@A@ the element type of the list. The @flatMap@ method first unwraps the inner
+encapsulated by the monad is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and
+@A@ is the element type of the list. The @flatMap@ method first unwraps the inner
 monadic value @M[SetT[A]]@ and obtain a value of @Set[A]@, then applies @f@ to

### 2019-04-05 paper/60_cases.tex
paper

@@
 
-\subsubsection{Store-widened Analysis}
-
-\subsubsection{Abstract Garbage Collection at Staging Time}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
-
-\subsubsection{A More Precise Store Model}
-
-The store-widened analysis we implemented in Section~\ref{sai} uses a single
-store to approximate the runtime store. It can be efficently computed in
-polynomial time together with 0-CFA-like abstraction, but sometimes we may
-desire a more precise result that distinguishes the final values and stores for
-different closure targets. To achieve this goal, we need to tweak our abstract
-interpreter and type instantiation. The answer type @Ans@ is changed to a set of
-@VS@s where a @VS@ is a pair of sets of abstract values (such as closures or
-abstract numbers) and a store.
+The abstract addresses we introduced in Section \ref{sai} is simply identical to
+the variable names, which is a context-insensitive schema.
+Through tweaking the address allocation strategy, we can achieve various
+context-sensitive analysis \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we
+demonstrate a calling-context-sensitive analysis, i.e., $k$-CFA-like analysis
+\cite{DBLP:journals/jfp/HornM12}, in our staged abstract interpreter.
 
 \begin{lstlisting}
-type VS = (Set[AbsValue], Store)
-type Ans = R[Set[VS]]
+  type Time = List[Expr]
+  case class KCFAAddr(x: Ident, time: Time) extends Addr
 \end{lstlisting}
 
-Note that the type @Ans@ uses our stage polymorphic type @R@, meaning that under
-staging the type @Ans@ represents a next stage value. Then, our generic
-interpreter is also changed when handling function applications.
-
-\begin{lstlisting}
-case App(e1, e2) =>
-  val e1ans = ev(e1, $\rho$, $\sigma$)
-  val e1vs  = choices(e1ans)
-  val e2ans = ev(e2, $\rho$, e1vs.$\sigma$)
-  val e2vs  = choices(e2ans)
-  apply_closure(ev)(e1vs.v, e2vs.v, e2vs.$\sigma$)
-\end{lstlisting}
-
-The idea to handle the application is to explore all possible closures from @e1@
-and meanwhile use the latest store. What @choices@ does is similar to McCarthy's
-@amb@ operator ~\cite{MCCARTHY196333}: it non-deterministically returns an
-element of type @VS@ from its argument, e.g., @e1ans@, to its right-hand side
-receiver, i.e., @e1vs@. The function @choices@ internally uses the delimited
-control operator @shift@ to capture the continuation, which is the code block
-after its call site. This allows us to perform nesting depth-first evaluation
-for function application while still writing the program in direct-style
-\cite{Wei:2018:RAA:3243631.3236800}. Again, we can stage this part as we did for
-the naive abstract interpreter.
-
-\subsubsection{Context-Sensitivity}
-
-We add k-CFA-like context-sensitivity to the analysis by introducing an abstract
-timestamp, whose concrete instantiation is a finite list of expressions that
-track $k$ recent calling contexts. The definition of abstract addresses is
-changed to a tuple of identifiers and the time it get allocated, meaning that
+The timestamp is essentially a finite list of expressions that tracks the $k$
+recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
+changed to contain a identifier and the time when it get allocated, meaning that
 this address points to some values under such calling context. If $k$ is 0, we
@@
 \begin{lstlisting}
-type Time = List[Expr]
+  type AnsM[T] = RepReaderT[RepStateT[
+                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //Add Time into a RepStateT
+                 Store, ?], Env, T]
+  type Result = (Rep[Set[(Value, Store, Time)]], Rep[Cache])
+\end{lstlisting}
+
+We then integrate the timestamp into our moand stack by adding a staged
+@StateT@ monad. The result type is passingly get updated to contain a set of
+tuples, where the time is added.
+
+\begin{lstlisting}
+  def tick(e: Expr): AnsM[Unit] = for {
+    τ <- get_time; u <- liftM(StateTMonad.put((e::τ).take($k$)))
+  } yield u
+  def alloc(x: String): AnsM[Addr] = for {
+    τ <- get_time
+  } yield KCFAAddr(x, τ)
 \end{lstlisting}
@@
 Every time when we call the @eval@ function, we refresh the timestamp by calling
-function @tick@, which returns a new timestamp. Here we adopt a $k$-CFA-like
-allocation strategy, therefore the @tick@ function can be implemented as
-appending the current expression being evaluated to the existing calling
-context, and then taking the first $k$ elements from the list.
+the @tick@ function, which update the timestamp in the state monad. The @tick@
+function is implemented as prepending the current expression being evaluated @e@
+to the existing calling context, and then taking the first $k$ elements from the list.
+The type @Config@ used in caching is also changed to included the timestamp.
+
+\subsubsection{Store-widened Analysis}
+
+An commonly used technique to improve the running time in control-flow analysis
+is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
+DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
+implemented in monadic abstract interpreter by swaping the @StateT@ (for store)
+and the @SetT@ (for nondeterminism). In this section, we show this optimization
+is orthogonal to staging and we just need to swap the same \textit{staged}
+version of monads. After the adjustment, we have the following monad stack:
 
 \begin{lstlisting}
-def tick(e: Expr, $\tau$: R[Time]): R[Time] = (e :: $\tau$).takewhile
-def eval(ev: EvalFun)
-  (e: Expr, $\rho$: R[Env], $\sigma$: R[Store], $\tau$: R[Time]): Ans = {
-    val $\tau$_* = tick(e, $\tau$)
-    e match {
-      case Lit(i) => ...
-      ...
-    }
-  }
+  type AnsM[T] = RepReaderT[RepSetT[RepStateT[
+                   RepReaderT[RepStateT[IdM, Cache, ?], Cache, ?], Store, ?], ?], Env, T]
 \end{lstlisting}
 
-Accordingly, the type of return value is accompanied by the timestamp. For a
-recursive call of @ev@, which would also return the latest timestamp, and that
-timestamp will be used for the evaluation afterward.
-
-\begin{lstlisting}
-type VST = (Value, Store, Time)
-type Ans = R[Set[VST]]
-\end{lstlisting}
-
-Using other allocation strategies to achieve different sensitivities is also
-possible \cite{DBLP:conf/icfp/Gilray0M16} and can be staged under our framework.
+\subsubsection{Abstract Garbage Collection}
 
@@
 abstract interpreter.
-\fi
 
@@
 \end{lstlisting}
+
+\fi

### 2019-04-05 paper/70_eval.tex
paper

@@
 \begin{figure}[h]
-\tiny
+\footnotesize
 \begin{tabular}{@{}lllll@{}}
@@
     program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
+    @fib@               & 32   & 0.013      & 0.002      & 650.00\%          \\
     @rsa@               & 451  & 0.308      & 0.064      & 459.15\%          \\

### 2019-04-05 paper/80_related.tex
paper

@@
 structure, e.g., decomposing polyhedras \cite{DBLP:conf/popl/SinghPV17,
-Singh:2017:PCD:3177123.3158143} to smaller ones such that the abstract
-transformers can compute using less time and space.
+Singh:2017:PCD:3177123.3158143} to smaller ones depending on the variables
+involved such that the abstract transformers can compute using less time and
+space.
 
@@
 shows that abstract interpreters can be derived systematically from concrete
-semantic artifacts.
+semantic artifacts. The AAM approach is closely related to control-flow analysis
+for higher-order languages \cite{Midtgaard:2012:CAF:2187671.2187672}.
 Using monads to construct abstract interpreters is explored by
@@
 \citet{DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
-Similar to monadic abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
+Similar to definitional abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
 reconstruct big-step abstract interpreters with delimited continuations;

### 2019-04-05 paper/90_conc.tex
paper

@@
 abstracts over both binding-time and value domains; by instantiating it
-differently, the staged abstract interpreter can be readily derived. We also
-conduct case studies and empirical evaluation, demonstrating a wide
-applicability of our approach and evaluating the performance improved by staging.
+differently, the staged abstract interpreter can be readily derived.
+The meta-programming approach brings new insight and implementation techniques
+to optimize abstract interpretation based static analysis. The optimization can
+be realized in a compeltely orthogonal way to the abstract semantics and actual
+implementation code. We empirical evaluate the performance improved by our approach.

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -31,8 +31,9 @@ object Evaluation {
       //(sat, "sat"), // 73
       //(solovay, "solovay")
       //(metacirc, "meta")
-      (regex, "regex"),
+      //(regex, "regex"),
       //(scheme2java, "scheme2java")
+      (fib, "fib"),
       /********************/
       //(matrix, "matrix"),
       //(letloop, "letloop")
@@ -42,7 +43,6 @@ object Evaluation {
       //(kcfa3, "kcfa3"),
       //(kcfa256, "kcfa256"),
       //(omega, "omega")
-      //(fib, "fib"), // 36
       //(boyer, "boyer"),
       //(earley, "earley"),
       //(dynamic, "dynamic"),
@@ -50,7 +50,8 @@ object Evaluation {
       //(nbody, "nbody"),
       //(nucleic, "nucleic"),
     )
-    //progs foreach { case (e, id) => compare(e, id) }
+    progs foreach { case (e, id) => compare(e, id) }
+    /*
     progs foreach { case (e, id) => 
       println(s"$id - ${size(e)}")
       println(e)
@@ -60,6 +61,7 @@ object Evaluation {
       val t1 =  run(1, { evalUnstaged(e) })
       println(s"[$id] [unstaged] - ${t1}s")
     }
+    */
     /*
     progs foreach { case (e, id) =>
       val code = specialize(e)

### 2019-04-05 paper/60_cases.tex
paper

@@
 to various control-flow analysis (CFA). We first discuss calling-context
-sensitive analysis for staged analyzers; then we will see how to implement
-store-widening with staging, which is a common optimization in CFA; finally, we
-integrate abstract garbage collection with staging, \todo{where the addresses to
-  be collected can be computed at the staging time}.
+sensitive analysis for staged analyzers by tweaking allocation strategy; then we
+will see how to implement store-widening with staged monads, which is a common
+optimization in CFA; finally, we integrate abstract garbage collection with
+staging.
 
@@
 
-% \subsubsection{Mixed Sensitivity}
 % using different k?
@@
 One of the challenges of modern static analysis is program usually depends on
-large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. Can we analyze
-programs and libraries separately and reuse the result without losing precision?
-Then we can reduce part of the overhead of repeatedly analyzing libraries for
-different programs. Indeed, some static analyzers compute summary for a function
-or a module, which can be reused later, however they are mostly too conservative
-or unsound, which both lead to imprecision.
+large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. If we can analyze
+programs and libraries separately and reuse the result without losing precision,
+then we can reduce the overhead of repeatedly analyzing libraries.
+Some static analyzers compute a summary for a function or a module, which
+can be reused later, however they are mostly too conservative or unsound, which
+both lead to imprecision. In this section, we show meta-programming can be used
+to create sound summaries in the code generation from, such generated code can
+be resued later. The summary contains information that not depend on the
+function argument.
 
-The specialization of abstract interpreter provides a chance to obtain such
-partial analysis result in a mechanized way, but still keeps the analysis sound.
-As we see when compiling the closures, we can specialize the abstract
-interpreter with respect the body expression of the lambda term without knowing
-the actual argument. The programs with some unknown variables are open programs,
-which is exactly the case if we want to analyze programs in a modular way.
+As we see when compiling a higher-order function (closure), we specialize the
+abstract interpreter to the body expression of that lambda term under some
+environment, without knowing the actual arguments to the function. We may reuse
+the infrastratures and refactor the @close@ function mentioned in Section
+\ref{sai} to achieve modular analysis. The @close@ function takes a lambda term
+@f@ and an environment, produces a compiled closure, which can be freely used at
+the next stage.
 
-For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a
-whole-program analysis, because it is inter-procedural and needs the last $k$
-calling contexts to distinguish different call sites, where the calling contexts
-are dynamic values. However, it is possible to analyze programs (libraries)
-separately through specializing an abstract interpreter that generates the
-analysis as the next-stage program and leave the unavailable programs and
-calling contexts as dynamic parameters, and then install these contexts when we
-have the whole program.
+\begin{lstlisting}
+  def close(ev: Eval => Ans)(f: Expr, ρ: Rep[Env]): Rep[CompiledClo] = { ...}
+\end{lstlisting}
+
+\paragraph{Partially-Static Context} Under a polyvariant analysis (like
+$k$-CFA), such summary generated by specializing static analyzer even could do
+more for calling context by partially-static data optimization. For example,
+consider that we have a library consists of three functions @f@, @g@ and @h@,
+where @f@ calls @g@ and @h@ internally and returns @y@ fianlly (we extend the
+language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
+would like to modularlly specialize these functions, starting from @f@.
+
+\begin{lstlisting}
+  def f(a, b, c) = val x = g(a, b); val y = h(x, c);  y
+\end{lstlisting}
+
+Since $k$ is bounded to a fixed number, after the analysis running $k$ steps
+inside of @f@, the calling context will purely depend on the static structure of
+our module being analyzed.
 

### 2019-04-05 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala

@@ -19,21 +19,21 @@ object Evaluation {
 
   def main(args: Array[String]) {
     val progs = List[(Expr, String)](
-      //(kcfa16, "kcfa16"),
-      //(kcfa32, "kcfa32"),
-      //(kcfa64, "kcfa64"),
-      //(church, "church")
-      //(fermat, "fermat"),
-      //(rsa, "rsa"),
-      //(mbrotZ, "mbrotZ"),
-      //(lattice, "lattice"),
+      (fib, "fib"),
+      (rsa, "rsa"),
+      (church, "church"),
+      (fermat, "fermat"),
+      (mbrotZ, "mbrotZ"),
+      (lattice, "lattice"),
+      (kcfa16, "kcfa16"),
+      (kcfa32, "kcfa32"),
+      (kcfa64, "kcfa64"),
+      (solovay, "solovay")
       //(blur, "blur"), // 56
       //(sat, "sat"), // 73
-      //(solovay, "solovay")
       //(metacirc, "meta")
       //(regex, "regex"),
       //(scheme2java, "scheme2java")
-      (fib, "fib"),
       /********************/
       //(matrix, "matrix"),
       //(letloop, "letloop")

### 2019-04-05 paper/10_intro.tex
paper

@@
 We show that these well-known abstractions can be combined smoothly, and more
-importantly almost free. Through staging, the major overhead of
-abstractions -- from the monadic layers -- is completely eliminated in the
-generated code. We observe at most 14x times faster than the unstage abstract
-interpreter.
-However, we do not lose any advantages of this high-level programming style
-(e.g., modularity and extensibility); the meaning of the program also remains
-the same, particularly, the soundness of static analysis after staging.
+importantly almost free. We implement the four interpreters using Scala and the
+Lightweight Modular Staging \cite{DBLP:conf/gpce/RompfO10} framework. Through
+staging, the major overhead of abstractions -- from the monadic layers -- is
+completely eliminated in the generated code. We observe at most 14x times faster
+than the unstage abstract interpreter. However, we do not lose any advantages of
+this high-level programming style (e.g., modularity and extensibility); the
+meaning of the program also remains the same, particularly, the soundness of
+static analysis after staging.
 
@@
     analysis (Section \ref{cfa}), including context-sensitivity, store-widening, and abstract
-    \note{garbage collection}, showing that our approach is widely applicable.
+    garbage collection, showing that our approach is widely applicable.
   \item We apply our approach to modular analysis (Section \ref{modular}),
@@
     control-flow analysis (Section~\ref{evaluation}). We compare the both
-    context-insensitive, \note{context-sensitive, and store-widening} analysis.
+    context-insensitive and store-widening analysis.
 \end{itemize}
-
-\todo{mention LMS somewhere}
 

### 2019-04-05 paper/20_prelim.tex
paper

@@
 with numbers, arithmetic, recursions and conditionals. Other effectful features
-such as assignments can also be supported readily. In Section~\ref{cases_imp}, we
-will add more imperative features to the language. Since we are mostly
+such as assignments can also be supported readily.
+%In Section~\ref{cases_imp}, we will add more imperative features to the
+language. Since we are mostly
 interested in analyzing the dynamic behaviors of the program, we disguise any
@@
 The abstract syntax we present in fact can be seen as a deep embedding of the
-language -- we use data-types to represent programs. This design choice allows us to
+language -- we use algebraic data-types to represent programs. This design choice allows us to
 easily use different interpretations over the AST; with the inheritance and

### 2019-04-05 paper/30_stageconc.tex
paper

@@
 collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
-@Rep[(Value, Store)]@. Then we generate an intermediate representation of the compiled
-closure using @emit_compiled_clo@. The intermediate representation stands for a
-next-stage value, and contains the compiled function, which will be generated as
-a next-stage function.
+@Rep[(Value, Store)]@. Then we generate the next-stage value of the compiled
+closure using @emit_compiled_clo@, containing the compiled closure.
 
@@
 
-When applying a function in @ap_clo@, we generate an intermediate representation
+When applying a function in @ap_clo@, we also generate the next-stage value
 for closure application through @emit_ap_clo@, assuming that the @rator@ is already an
 IR of compiled closure, @rand@ is an arbitrary stage value, and $\sigma$ is the
-latest store. @emit_ap_clo@ represents a next-stage value of type
-@Rep[(Value, Store)]@, which is a collapsed monadic value from next stage, now
-we need to reify the value and store back into the monad stack through
-@put_store@ and @yield@.
+latest store. @emit_ap_clo@ represents the result value of application, which has type
+@Rep[(Value, Store)]@. Note that @emit_ap_clo@ is a ordinary value from the
+future, and there is no monads in the future, so we need to reify the value and
+store back into the monad stack through @put_store@ and @yield@.
 
@@
 
-The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces IR
-node in LMS. We define these IR node as follow: \note{Exp, Def, quote are omitted}
+The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces
+intermediate representations for the denoted operations in LMS. These IR are
+wrappers of the underlying staged data. We define these IR nodes using
+@case class@ extending from @Def[T]@, meaning that they are next-staged definitions of
+type @T@.
 
 \begin{lstlisting}
-  case class IRCompiledClo(f: Rep[((Value,Store)) => (Value, Store)]) extends Rep[Value]
-  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Rep[(Value, Store)]
+  case class IRCompiledClo(f: Rep[((Value,Store)) => (Value, Store)], λ: Lam, ρ: Rep[Env]) extends Def[Value]
+  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[(Value, Store)]
 \end{lstlisting}
@@
     case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar$rand, $\textdollar$σ)")
-    case IRCompiledClo(f) => emitValDef(sym, s"CompiledClo($\textdollar$f)")
+    case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar$f, $\textdollar$λ, $\textdollar$ρ)")
     ...

### 2019-04-05 paper/50_stageabs.tex
paper

@@
 Since our monad stack is uniformly using staged data, so the staged @SetT@ now
-stores a staged value of type @Rep[Set[A]]@, inside of a staged monad @M@.
-\todo{staged SetT implementation}
+stores a staged value of type @Rep[Set[A]]@, inside of anohter staged monad @M@.
 
@@
 lift the set of closures into the monad stack, and generate a next-stage value
-using @emit_ap_clo@, which represents the result of application, and finally we
-put the @out@ cache, store and values back into the monads. The @emit_ap_clo@
-method takes the target closure @clo@, argument @rand@, store and caches.
+using @emit_ap_clo@, which represents the result of application from future.
+Finally we put the future @out@ cache, store and values back into the
+current-stage monads. The @emit_ap_clo@ method takes the target closure @clo@,
+argument @rand@, store and caches.
 
@@
   case class IRCompiledClo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
-                           => Rep[(Set[(Value, Store)], Cache)]) extends Rep[AbsValue]
+                           => Rep[(Set[(Value, Store)], Cache)], λ: Lam, ρ: Rep[Env]) extends Def[AbsValue]
   case class IRApClo(clo: Rep[AbsValue], arg: Rep[Value], σ: Rep[Store],
-                     in: Rep[Cache], out: Rep[Cache]) extends Rep[(Set[(Value, Store)], Cache)]
+                     in: Rep[Cache], out: Rep[Cache]) extends Def[(Set[(Value, Store)], Cache)]
 \end{lstlisting}

### 2019-04-05 paper/60_cases.tex
paper

@@
     τ <- get_time
-  } yield KCFAAddr(x, τ)
+  } yield unit(KCFAAddr(x, τ)) // we unit to turn a current-stage constant to next-stage value
 \end{lstlisting}

### 2019-04-05 paper/70_eval.tex
paper

--- 2019-04-05 paper/70_eval.tex
@@
     program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
-    @rsa@               & 451  & 0.308      & 0.064      & 459.15\%          \\
+    @fib@               & 32   & 0.013      & 0.002      & 650.00\%          \\
+    @rsa@               & 451  & 0.042      & 0.232      & 552.38\%          \\
     @church@            & 120  & 71.811     & 23.990     & 299.34\%          \\

### 2019-04-05 paper/abstract.tex
paper

@@
 overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of \note{10}x with staging for a value-flow analysis.
+experiment shows an average speedup of 6.6x with staging for control-flow analysis.
 

### 2019-04-06 paper/10_intro.tex
Grammar Intro

@@
 
-Roughly at the same time when Futamura projections was envisioned in 1970s,
+Roughly at the same time when Futamura projections were envisioned in the 1970s,
 \citet{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
@@
 be easily established by examining the transformation of semantics. The AAM
-approach also has been been applied to different variants of definitional
+approach also has been applied to different variants of definitional
 interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
@@
   behaves as a concrete interpreter or an abstract interpreter.
-  The environments, stores and values are initially declared as abstract types in the generic interpreter.
+  The environments, stores, and values are initially declared as abstract types in the generic interpreter.
   As shown by the AAM approach \cite{DBLP:journals/jfp/HornM12,
@@
 \cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
-ashley:practical} is not new, despite the fact that most of the previous work
+ashley:practical} is not new, despite the fact that most of the previous works
 are ad-hoc on a particular analysis or requiring significant changes to the
@@
 
-Further more, we show various applications of staging an abstract interpreter.
+Furthermore, we show various applications of staging an abstract interpreter.
 (1) We compare our approach with abstract compilation, and show that we can
 achieve the same optimization by utilizing type-based stage annotations,
-meanwhile the analyzer program does not need any changes. (2) \todo{} (3)
+meanwhile, the analyzer program does not need any changes. (2) \todo{} (3)
 Staging abstract interpreter enables modular analysis and provides a mechanized
@@
 generated code is reusable and composable. Therefore, we can turn a
-whole-program analyzer into a analyzer that runs modularly.
+whole-program analyzer into an analyzer that runs modularly.
 
@@
 \cite{DBLP:conf/oopsla/KulkarniMZN16}. A natural idea is to analyze libraries
-and generate summaries, later we can plugin the necessary information into the
+and generate summaries, later we can plug in the necessary information into the
 summaries and finish the analysis; therefore such summaries should be modular
@@
 produces summaries. Still, the idea is to specialize the abstract interpreter
-that was used for whole-program analysis, but on a subset of the program, for
+that was used for whole-program analysis but on a subset of the program, for
 examples, a set of generic functions for list manipulation. Such specialization
@@
  \item Pragmatically, by case studies and empirical evaluation, we show that staging 
-   abstract interpreters is useful to improve performance while untouch 
+   abstract interpreters is useful to improve performance while \note{not a word} untouch 
    the soundness. 
@@
   \item We apply our approach to modular analysis (Section \ref{modular}),
-    i.e., analyzing the components of program separately and composing them later,
+    i.e., analyzing the components of the program separately and composing them later,
     showing a novel application of generative programming with program analysis.
@@
 \iffalse
-Likewise, specializing static analyses by partial evaluation emerged in late 90s
+Likewise, specializing static analyses by partial evaluation emerged in the late 90s
 , and indeed it is able to effectively remove the interpretive
@@
 challenge is the binding-time engineering when non-determinism, fixed-point
-iterations, different abstract domains and staging are introduced at the same
+iterations, different abstract domains, and staging are introduced at the same
 time. In this paper, we present an end-to-end design and implementation of
@@
 abstraction, but also can be implemented efficiently. As the slogan of
-multi-stage programming said, "abstraction without regret", we draw connection
+multi-stage programming said, "abstraction without regret", we draw a connection
 between high-level description and efficient implementation of abstract
 interpreters, just like the connection between concrete interpreters and
-compilers drawn by Futamura projection. Particularly, we show an easy but
+compilers drawn by the Futamura projections. Particularly, we show an easy but
 systematic way of adding stage annotations to the abstract interpreter without
@@
 are still dynamic, no? They change in every iteration of the fixpoint algorithm}
-\note{TR: why is it a big deal. On Closed programs we obtain a constant factor,
+\note{TR: why is it a big deal. On Closed programs, we obtain a constant factor,
 but modular analysis has different asymptotics}
@@
 It has been observed that partially applying context-sensitivity on selected
-portion of the program could improve the precision and efficiency
+portions of the program could improve the precision and efficiency
 \cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. We show that staging
-abstract interpreters as an approach to effectively implement hybird
+abstract interpreters as an approach to effectively implement hybrid
 context-sensitivity\todo{}.

### 2019-04-06 paper/60_cases.tex
paper

--- 2019-04-05 paper/60_cases.tex
@@
   type AnsM[T] = RepReaderT[RepStateT[
-                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //Add Time into a RepStateT
+                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //add Time into a RepStateT
                  Store, ?], Env, T]
@@
     τ <- get_time
-  } yield KCFAAddr(x, τ)
+  } yield unit(KCFAAddr(x, τ)) // we use unit to turn a current-stage constant to next-stage value
 \end{lstlisting}

### 2019-04-06 paper/70_eval.tex
paper

--- 2019-04-05 paper/70_eval.tex
@@
     program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
-    @rsa@               & 451  & 0.308      & 0.064      & 459.15\%          \\
-    @church@            & 120  & 71.811     & 23.990     & 299.34\%          \\
-    @fermat@            & 310  & 0.043      & 0.006      & 716.67\%          \\
-    @mbrotZ@            & 331  & 1.112      & 0.214      & 519.63\%          \\
-    @lattice@           & 609  & 31.874     & 8.816      & 358.69\%          \\
-    @kcfa-worst-16@     & 182  & 0.083      & 0.013      & 638.46\%          \\
-    @kcfa-worst-32@     & 358  & 0.452      & 0.044      & 1027.27\%         \\
-    @kcfa-worst-64@     & 774  & 3.329      & 0.236      & 1410.59\%         \\
-    @solovay-strassen@  & 523  & 5.94       & 1.189      & 428.43\%          \\
+    @fib@               & 32   & 0.013      & 0.002      & 650.00\%          \\
+    @rsa@               & 451  & 0.232      & 0.042      & 552.38\%          \\
+    @church@            & 120  & 67.342     & 18.917     & 355.99\%          \\
+    @fermat@            & 310  & 0.024      & 0.004      & 600.00\%          \\
+    @mbrotZ@            & 331  & 0.763      & 0.127      & 600.79\%          \\
+    @lattice@           & 609  & 31.874     & 8.816      & 361.59\%          \\
+    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 750.00\%          \\
+    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 1290.91\%         \\
+    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 1689.92\%         \\
+    @solovay-strassen@  & 523  & 5.094      & 1.189      & 428.43\%          \\
     \bottomrule
@@
 unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
-We observe the most improvement of 14x times faster than the unstaged version on
-@kcfa-worst-64@. On average, the staging produces code that run 6x times faster
+We observe the most improvement of 17x times faster than the unstaged version on
+@kcfa-worst-64@. On average, the staging produces code that run 7x times faster
 than the unstaged analyzer.

### 2019-04-06 paper/abstract.tex
paper

@@
 overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of 6.6x with staging for control-flow analysis.
+experiment shows an average speedup of 7x with staging for control-flow analysis.
 

### 2019-04-06 paper/20_prelim.tex
Grammar Prelim

@@
 We consider a call-by-value $\lambda$-calculus in direct-style, extended
-with numbers, arithmetic, recursions and conditionals. Other effectful features
+with numbers, arithmetic, recursions, and conditionals. Other effectful features
 such as assignments can also be supported readily.
@@
 interpreter. The interpreter is a recursive function that takes the program AST,
-environment and store, and returns the evaluated value and the accompanied
+environment, and store, and returns the evaluated value and the accompanying
 store. The environment is a mapping from identifiers to addresses, and the store
@@
 @return@ and @bind@. Informally, @return@ wraps a value into the monad @M@, and
-@bind@ unwraps the monadic value and transforms it to a new monadic value.
+@bind@ unwraps the monadic value and transforms it into a new monadic value.
 Pragmatically in Scala, we define a monad type class using trait @Monad@ (Figure
@@
 For example, an object of @List[A]@ is an instance of @List@ monad, where @A@ is the element type. 
-Then to compute the the Cartesian product of two lists of numbers, we can use Scala's
+Then to compute the Cartesian product of two lists of numbers, we can use Scala's
 @for@-comprehension syntax.
@@
 operations that are necessary for @for@-comprehension. 
-The conversion between @Monad@ and @MonadOps@ can be done by using implicit design pattern.
+The conversion between @Monad@ and @MonadOps@ can be done by using the implicit design pattern.
 In the rest of the paper, we use Scala's @for@-comprehension syntax and monad
@@
 As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
-syntax, certain constraints have to be posed onto the type @AnsM@. Here, we use a
+syntax, certain constraints have to be added on the type @AnsM@. Here, we use a
 structural type @MonadOps@ to require @AnsM@ to at least implement @map@ and
@@
 
-We also defines several methods to obtain and modify the environment and store.
+We also define several methods to obtain and modify the environment and store.
 These methods return monadic values of type @AnsM[_]@, which encapsulate the 
@@
 Two versions of @alloc@ are declared. The first one takes a store and an
-identifier, and produces a fresh address of non-monadic type @R[Addr]@. Since
+identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
 the freshness of the address may depend on the store, which might be a
@@
 
-Since the interpreter is written in open-recursive style, we declare an abstract
+Since the interpreter is written in an open-recursive style, we declare an abstract
 combinator @fix@ used to close the recursion. For concrete-interpretation instantiation, it
@@
 termination. Finally, a top-level wrapper @run@ is declared; the return type
-@Result@ depends on what kind of monad we will be using, thus is also an abstract type.
+@Result@ depends on what kind of monad we will be using, this is also an abstract type.
 
@@
 @ReaderT@ and @StateT@.
-In other word, we instantiate the @AnsM@ monad as a stack of @ReaderT@ and @StateT@
+In other words, we instantiate the @AnsM@ monad as a stack of @ReaderT@ and @StateT@
 transformers\footnote{The question mark syntax is a kind projector
@@
 simply constructing the proper monads and lifting it to the top-level of our
-moand stack. To modify the store, for example, we may have a @StateT@ monad that
+monad stack. To modify the store, for example, we may have a @StateT@ monad that
 transforms the current store $\sigma$ to @σ + αv@, which results in
@@
 Other primitive operations can be implemented straightforwardly, for example,
-@alloc@ simply takes the size of current store and produce the successor number.
-We elide most of them, but describe a little bit on how do we handle functions
+@alloc@ simply takes the size of the current store and produce the successor number.
+We elide most of them but describe a little bit on how do we handle functions
 and applications as they will be very different when staging and abstract
@@
 
-To apply a function, the method @ap_clo@ takes a function value and an arugment
-value, extracts the lambda term and environment enclosed in the function value,
+To apply a function, the method @ap_clo@ takes a function value and an argument
+value, and extracts the lambda term and environment enclosed in the function value,
 and evaluates the body expression of that lambda term under the new environment and store.
@@
 With other primitive operations implemented, we can implement the top-level
-@run@ method, where $\rho_0$ and $\sigma_0$ are initial empty environment and
+@run@ method, where $\rho_0$ and $\sigma_0$ are the initial empty environment and
 store.

### 2019-04-06 paper/10_intro.tex
typo

@@
  \item Pragmatically, by case studies and empirical evaluation, we show that staging 
-   abstract interpreters is useful to improve performance while \note{not a word} untouch 
+   abstract interpreters is useful to improve performance while untouching
    the soundness. 

### 2019-04-06 paper/30_stageconc.tex
Grammar stagecontext

@@
 on types. LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary
-type, indicating a value of @T@ will be known at next stage. All operations
+type, indicating a value of @T@ will be known at the next stage. All operations
 acting a @Rep[T]@ expression will be residualized as generated code.
@@
 @Int@ and @Double@, and commonly-used data structures such as arrays and maps.
-The idea is to construct intermediate representations (IR) for the next-stage
-program at current stage. Such conversion from expressions of type @Rep[_]@ to
-their IR is done by using implicit design pattern. As we will see later,
+The idea is to construct intermediate representations (IR) for the next stage
+program at the current stage. Such conversion from expressions of type @Rep[_]@ to
+their IR is done by using the implicit design pattern. As we will see later,
 implementing the staging support for custom classes is also straightforward.
@@
 This is why we can completely eliminate the overhead caused by the monadic layers.
-Here we use the prefix name @Rep@ to differentitate unstaged monads and staged
+Here we use the prefix name @Rep@ to differentiate unstaged monads and staged
 monads, such as @RepReaderT@. So the @AnsM[_]@ is instantiated as the same monad
@@
 transformers look like. The generic type @M[_]@ is a @RepMonad@ and used as the
-inner monad, which means the pair @(A,S)@ is also staged when appears inside of monad
+inner monad, which means the pair @(A,S)@ is also staged when appears inside of the monad
 @M[_]@. The @flatMap@ is similar to the unstaged one, but the function @f@ takes
@@
 translated to their staged versions -- we just need to let them works on staged data.
-As we mentioned before, we elaborate how to handle functions and applications in
-detail. One of the desired goal is to eliminate the monadic layers through
-staging, this becomes a little bit tricky for clsoures, because we need actually
-compiling the lambda term instead of generating a @CloV@ value, but the @ev(e)@
+As we mentioned before, we elaborate on how to handle functions and applications in
+detail. One of the desired goals is to eliminate the monadic layers through
+staging, this becomes a little bit tricky for closures because we actually need 
+ to compile the lambda term instead of generating a @CloV@ value, but the @ev(e)@
 returns a monad and prevents the specialization. To make this happen, when
-closing a lambda term, we need to staging into the body expression by
+closing a lambda term, we need to specialize the body expression by
 \textit{collpasing} the @AnsM@ monads to normal values. The
@@
 IR of compiled closure, @rand@ is an arbitrary stage value, and $\sigma$ is the
-latest store. @emit_ap_clo@ represents the result value of application, which has type
-@Rep[(Value, Store)]@. Note that @emit_ap_clo@ is a ordinary value from the
-future, and there is no monads in the future, so we need to reify the value and
+latest store. @emit_ap_clo@ represents the result value of the application, which has type
+@Rep[(Value, Store)]@. Note that @emit_ap_clo@ is an ordinary value from the
+future, and there are no monads in the future, so we need to reify the value and
 store back into the monad stack through @put_store@ and @yield@.
@@
 The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces
-intermediate representations for the denoted operations in LMS. These IR are
+intermediate representations for the denoted operations in LMS. These IRs are
 wrappers of the underlying staged data. We define these IR nodes using

### 2019-04-06 paper/40_absinterp.tex
Grammar absinterpreter

@@
 lattice or point-wise map lattice. Then we show the abstract components, namely
-abstract values, abstract environments and abstract stores.
+abstract values, abstract environments, and abstract stores.
 The abstract interpreter we present in this section is similar to
@@
 In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
-unstaged lattice type class. An simple instance of lattice we used is the power
+unstaged lattice type class. A simple instance of lattice we used is the power
 set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
-operation, and the meet is intersect operation.
+operation, and the meet is the intersect operation.
 Other lattices used in the remainder of the paper, such as products and maps, can be
-implemented similarly, or lifted from existing lattices element-wise or
+implemented similarly or lifted from existing lattices element-wise or
 point-wise. Naturally, abstract domains such intervals also can be implemented
@@
 @Value@).
-The address space is constrained to be finite to ensure the computatbility of
+The address space is constrained to be finite to ensure the computability of
 analysis -- for the simplest monovariant analysis, we use variable names for
@@
 As of now, the store is a mapping from addresses to sets of abstract values.
-Therefore, when applying a function, for instance @f(a)@, we may retrieve a set
+Therefore, when applying a function, for instance, @f(a)@, we may retrieve a set
 of closures by dereferencing the address of @f@. We need to nondeterministically
@@
 
-This observation leads us to the type definition of @Result@: in concrete setting,
+This observation leads us to the type definition of @Result@: in a concrete setting,
 the @Result@ is a pair of @Value@ and @Store@, now we use a set of such pairs
-@Set[(Value, Store)]@ to represent multiple results. The results is also linked
+@Set[(Value, Store)]@ to represent multiple results. The results are also linked
 with the resulting cache used in this iteration.
@@
 represented by the @SetT[M[_], A]@ monad transformer, inside of which, we have
-a @ReaderT@ for a @in@ cache, and a @StateT@ for an @out@ cache. \todo{highlight
+a @ReaderT@ for a @in@ cache and a @StateT@ for an @out@ cache. \todo{highlight
 where changed}
@@
 
-The sketch implementation of @SetT@ transformer is shown below. The value @run@
+The sketch implementation of the @SetT@ transformer is shown below. The value @run@
 encapsulated by the monad is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and
@@
 using @mplus@ from @MonadPlus@, which requires the values are join-semilattices
-(in our case, @Set@s and @Map@s).
+(in our case @Set@s and @Map@s).
 
@@
 the last iteration, and the @out@ cache is what we would have computed after this
-iteration joined with the contents from @in@ cache.
+iteration joined with the contents from the @in@ cache.
 When starting a new iteration, @in@ is set to be the result of the last
@@
 @cfg@, representing the current desired computation, if
-yes then result is directly returned through the monads.
+yes, then the result is directly returned through the monads.
 

### 2019-04-06 paper/50_stageabs.tex
Grammar stageabs

@@
 lattices and staged monads, and then the staged version of primitive operations,
-especially @close@, @ap_clo@ and @fix@, at last we discuss several optimizations
+especially @close@, @ap_clo@ and @fix@, at last, we discuss several optimizations
 and our design.
@@
 Since our monad stack is uniformly using staged data, so the staged @SetT@ now
-stores a staged value of type @Rep[Set[A]]@, inside of anohter staged monad @M@.
+stores a staged value of type @Rep[Set[A]]@, inside of another staged monad @M@.
 
@@
 using @emit_ap_clo@, which represents the result of application from future.
-Finally we put the future @out@ cache, store and values back into the
+Finally, we put the future @out@ cache, store, and values back into the
 current-stage monads. The @emit_ap_clo@ method takes the target closure @clo@,
-argument @rand@, store and caches.
+argument @rand@, store, and caches.
 
@@
 the current configuration produces a next-stage Boolean value, i.e.,
-@Rep[Boolean]@. Consequently, the branching can not be determined statically --
+@Rep[Boolean]@. Consequently, the branching cannot be determined statically --
 we need to generate code for @if@. Figure \ref{fig:staged_coind_cache} shows the
@@
 \paragraph{Code Generation} The IR nodes for compiled closures and
-applications are defined as following:
+applications are defined as follows:
 
@@
 
-Although every components except @Expr@s are staged, we treat the data
+Although every component except @Expr@s are staged, we treat the data
 structures such as @Map@s as black-boxes, which means the operations on a @Map@
@@
 @Array[Rep[V]]@, whose elements are next-stage values; the size of the array is
-also known statically. An access to the map is translated to an access to the
+also known statically. An access to the map is translated into an access to the
 array with pre-determined index during staging.
@@
 is equivalent to the set of identifiers in the program. Then the accesses to the
-environment can be completely determined statically, and generates the addresses
+environment can be completely determined statically and generates the addresses
 directly. After which, the stores can be specialized as arrays of @Rep[Value]@
@@
 Our treatment to binding-times is coarse-grained: @Expr@s is static, the rest of
-the world are all dynamic. But this is not always true, since the static data
+the world are all dynamic. But this is not always true since the static data
 has to be used somewhere with the dynamic operation. Partial-static data is a
@@
 generate cheaper code: @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
-function @f@ is known at current stage, we completely eliminate the fold
+function @f@ is known at the current stage, we completely eliminate the fold
 operation and function application. We apply several rewritings enabled by
@@
 explicit side-effects such mutation to implement the artifact with the same
-abstract semantics. In either cases, we can still apply the staging schema to
-the abstract interpreter. As an evidence, we provide an staged abstract
+abstract semantics. In either case, we can still apply the staging schema to
+the abstract interpreter. As evidence, we provide a staged abstract
 interpreter written in direct-style in the accompanying artifact.
@@
 byte-code still requires another small-step abstract machine to execute, which is
-an additional engineering efforts.
+an additional engineering effort.
 
@@
 
-As one of the merit of our approach, the staging does not compromise any
+As one of the merits of our approach, the staging does not compromise any
 soundness, because we only change the underlying monads and data structures to

### 2019-04-06 paper/70_eval.tex
Grammar eval

@@
 
-To evaluate the performance can be improved by our staging abstract interpreters
+To evaluate how the performance can be improved by our staged abstract interpreters
 approach, we scale our toy interpreter to a large subset of the Scheme language
@@
 shows that the staged versions are averagely 6 times as fast of the unstaged
-version, as a strong evidence of our \textit{abstraction without regret}
+version, as strong evidence of our \textit{abstraction without regret}
 approach.
@@
 value-flow analysis focus on lambda terms in functional languages. A front-end
-desugars Scheme to a small core language that make the analyzer easy to
+desugars Scheme to a small core language that makes the analyzer easy to
 implement. As described in Section \ref{cfa}, we implement a monovariant
@@
 We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
-(kernel 4.4.0) machine. All of our evaluation were performed on a machine with 4 Intel
+(kernel 4.4.0) machine. All of our evaluations were performed on a machine with 4 Intel
 Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
@@
 We observe the most improvement of 17x times faster than the unstaged version on
-@kcfa-worst-64@. On average, the staging produces code that run 7x times faster
+@kcfa-worst-64@. On average, the staging produces code that runs 7x times faster
 than the unstaged analyzer.
@@
 layers. Thus our evaluation result provides an abstraction without regret
-approach to construct optimizing abstract interpreters -- user may write a
+approach to construct optimizing abstract interpreters -- the user may write a
 high-level modular and composable abstract interpreter, and then derive a staged

### 2019-04-06 paper/60_cases.tex
paper

@@
 optimization in CFA; finally, we integrate abstract garbage collection with
-staging.
+staging. All of them are in fact orthogonal to staging.
 
@@
 
+Abstract garbage collection is a technique to reclaim unreachable addresses in
+the store while doing the abstract interpretation
+\cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
+shows that, for big-step monadic abstract interpreter, we may add a @Reader@
+monad to track a set of root addresses and compute the reachable addresses every
+time when we obtain a value.
+
+\begin{lstlisting}
+  def ask_root: AnsM[Set[Addr]]
+  def eval(e: Expr): AnsM
+\end{lstlisting}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
-would like to modularlly specialize these functions, starting from @f@.
+would like to modularlly specialize these functions, starting from @f@ with an
+initial environment containing @f@, @g@ and @h@.
 

### 2019-04-06 paper/80_related.tex
paper
Grammar related

@@
 of AC can be either textual or closures. However, as we studied in Section
-\ref{cs_ac}, with the perspective from high-performance DSLs and monadic
+\ref{cs_ac}, with the perspective from generative programming and monadic
 interpreters, we make it more general, extensible and easy.
@@
 generation to optimize small-step abstract interpreter in state-transition
-style. The analyzed program is firstly compiled to a IR called "abstract
+style. The analyzed program is firstly compiled to an IR called "abstract
 bytecode", which are actually higher-order functions and will be executed later
-on a abstract abstract machine for that IR. \citet{damian1999partial} provides a
+on an abstract abstract machine for that IR. \citet{damian1999partial} provides a
 formal treatment to abstract compilation and Shiver's CFA, as well as proofs to
-establish correctness of the certified specialized analyzer.
+establish the correctness of the certified specialized analyzer.
 \citet{amtoft1999partial} applied partial evaluation for constraint-based
@@
 \paragraph{Abstract Interpreters} Abstract interpretation was proposed as a
-semantic-based approach to construct sound static analysis by approximation
+semantic-based approach to build sound static analysis by approximation
 \cite{DBLP:conf/popl/CousotC77}. As semantic artifacts, the Abstracting Abstract
@@
 The idea of staging an abstract interpreter we presented in this paper is also
-applicable to other MSP system. Partial evaluation as an automatic technique is
+applicable to other MSP systems. Partial evaluation as an automatic technique is
 studied comprehensively \cite{10.1007/3-540-61580-6_11,

### 2019-04-06 paper/90_conc.tex
Grammar conclusion

@@
 
-In this paper, we first present the implementaion of first Futamura projection
+In this paper, we first present the implementation of first Futamura projection
 on abstract interpreters. To realize it, we design a generic interpreter that
@@
 to optimize abstract interpretation based static analysis. The optimization can
-be realized in a compeltely orthogonal way to the abstract semantics and actual
+be realized in a completely orthogonal way to the abstract semantics and actual
 implementation code. We empirical evaluate the performance improved by our approach.

### 2019-04-06 paper/60_cases.tex
paper

@@
 
-An commonly used technique to improve the running time in control-flow analysis
+A commonly used technique to improve the running time in control-flow analysis
 is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
@@
 monad to track a set of root addresses and compute the reachable addresses every
-time when we obtain a value.
+time when we obtain a value. Here we adopt it with staged monads:
 
 \begin{lstlisting}
-  def ask_root: AnsM[Set[Addr]]
-  def eval(e: Expr): AnsM
+  type AnsM[T] = RepReaderT[RepReaderT[
+                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?],
+                 Set[Addr], ?], Env, T] //add a reader monad for Set[Addr]
 \end{lstlisting}
+
+The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the @fix@
+function with caching (from Section \ref{sai}) to @fix_gc@, but the caching part
+is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
+current root addresses from the moand stack; @root@ compute the reachable
+addresses from a value; and @gc@ cleans up the store according to the reachable
+addresses and returns a new store. For the generic interpreter, we also need to
+compute the root addresses for compound expressions, readers may refer to
+\cite{DBLP:journals/pacmpl/DaraisLNH17}.
+
+\begin{figure}[h!]
+\begin{lstlisting}
+  def fix_gc(ev: (Expr => Ans) => (Expr => Ans))(e: Expr): Ans = for {
+    ψ <- ask_roots ... // omit code for retrieving ρ, σ, in and out
+    val res: Rep[(Set[(Value,Store)], Cache)] = ... // omit code for checking caches
+    _ <- put_out_cache(res._2)
+    vs <- lift_nd(res._1)
+    σ <- gc(ψ ⊔ root(vs._2)) // perform abstract garbage collection
+    _ <- put_store(σ)
+  } yield v
+\end{lstlisting}
+\vspace{-1em}
+\caption{\texttt{fix\_gc} for abstract garbage collection} \label{fixgc}
+\end{figure}
+
+%def root(v: Rep[Value]): Set[Addr]  def ask_root: AnsM[Set[Addr]]  def gc(ra: Rep[Set[Addr]]): AnsM[Store]
 

### 2019-04-06 paper/60_cases.tex
paper

@@
 
-In this section, we show that \citeauthor{Boucher:1996:ACN:647473.727587}'s AC
-(closure generation as example) can be understood and implemented as an instance
-of staging abstract interpreters. We first revisit the original implementation
-of abstract compilation of 0-CFA, and then reproduce their result by simply
-adding stage annotations.
+In this section, we show that the AC (closure generation as example) can be
+understood and implemented as an instance of staging abstract interpreters. We
+first revisit the original implementation of abstract compilation of 0-CFA, and
+then reproduce their result by simply adding stage annotations.
 The generated program of our approach improves approximately the same extent of

### 2019-04-06 oopsla19-code/src/main/scala/sai/evaluation/UnstagedAnalyzer.scala

@@ -77,12 +77,6 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
   } yield v1 ++ v2
      */
   ReaderTMonadPlus[StoreNdInOutCacheM, Env].mplus(ev(thn), ev(els)) // they use different store and cache?
-  /*
-    for {
-      exp <- lift_nd[Expr](Set(thn, els)) //Note: use same store
-      v <- ev(exp)
-    } yield v
-  */
   def close(ev: EvalFun)(λ: Lam, ρ: Env): Value = Set(CloV(λ, ρ))
   def ap_clo(ev: EvalFun)(fun: Value, args: List[Value]): Ans = for {
     CloV(Lam(params, e), ρ: Env) <- lift_nd(fun)
@@ -163,11 +157,11 @@ object UnstagedSchemeAnalyzer extends AbstractComponents {
       , "vector-ref" -> Set(IntV, FloatV, CharV, BoolV)
       , "cddr" -> Set(IntV, FloatV, CharV, BoolV)
       , "zero?" -> Set(BoolV)
-          , "%" -> Set(BoolV)
-          , "symbol?" -> Set(BoolV)
-          , "equal?" -> Set(BoolV)
-          , "pair?" -> Set(BoolV)
-          , "char?" -> Set(BoolV)
+      , "%" -> Set(BoolV)
+      , "symbol?" -> Set(BoolV)
+      , "equal?" -> Set(BoolV)
+      , "pair?" -> Set(BoolV)
+      , "char?" -> Set(BoolV)
   )
 
   def primitives(ev: EvalFun)(x: String, args: List[Expr]): Ans = {

### 2019-04-06 paper/10_intro.tex
paper

@@
 staging, the major overhead of abstractions -- from the monadic layers -- is
-completely eliminated in the generated code. We observe at most 14x times faster
+completely eliminated in the generated code. We observe up to 17x times faster
 than the unstage abstract interpreter. However, we do not lose any advantages of
@@
 achieve the same optimization by utilizing type-based stage annotations,
-meanwhile, the analyzer program does not need any changes. (2) \todo{} (3)
-Staging abstract interpreter enables modular analysis and provides a mechanized
+meanwhile, the analyzer program does not need any changes.
+(2) We extend the basic staged abstract interpreter to diverse control-flow analysis, including store-widening analysis, context-sensitive analysis and abstract garbage collection.
+(3) Staging abstract interpreter enables modular analysis and provides a mechanized
 way to create library summaries. We notice that we may apply the staged abstract

### 2019-04-06 paper/20_prelim.tex
paper

@@
 such as assignments can also be supported readily.
-%In Section~\ref{cases_imp}, we will add more imperative features to the
-language. Since we are mostly
-interested in analyzing the dynamic behaviors of the program, we disguise any
-static semantics and type system. We also assume that input programs are
-well-typed and all variables are distinct. The abstract syntax is shown as
-follows:
+%In Section~\ref{cases_imp}, we will add more imperative features to the language.
+Since we are mostly interested in analyzing the dynamic behaviors of the
+program, we disguise any static semantics and type system. We also assume that
+input programs are well-typed and all variables are distinct. The abstract
+syntax is shown as follows:
 
@@
 Other primitive operations can be implemented straightforwardly, for example,
-@alloc@ simply takes the size of the current store and produce the successor number.
+@alloc@ simply takes the size of the current store and produces the successor number.
 We elide most of them but describe a little bit on how do we handle functions

### 2019-04-06 paper/30_stageconc.tex
paper

@@
 presented in Section \ref{prelim}, we show how to obtain a staged concrete
-interpreters by changing the monads and refactoring several primitive
+interpreter by changing the monads and refactoring several primitive
 operations. We begin by briefly introducing the Lightweight Modular Staging
@@
 type, indicating a value of @T@ will be known at the next stage. All operations
-acting a @Rep[T]@ expression will be residualized as generated code.
+acting on @Rep[T]@ expressions will be residualized as generated code.
 
-The LMS framework provides staging support for primitive data types such as
-@Int@ and @Double@, and commonly-used data structures such as arrays and maps.
-The idea is to construct intermediate representations (IR) for the next stage
-program at the current stage. Such conversion from expressions of type @Rep[_]@ to
-their IR is done by using the implicit design pattern. As we will see later,
-implementing the staging support for custom classes is also straightforward.
-
-For example, let us go back and see how LMS can be used to specialize the
-@power@ function from Section ~\ref{intro}.
+For the example we introduced in Section \ref{intro}, we have identified @b@ will 
+be known later. To specialize it we write a function @snippet@ that provides a 
+concrete value @x@ for @x@. After this, we may instantiate an @DslDriver@ object 
+which provide the specialized program.
 
@@
 
-\todo{Show some IR}
+The LMS framework provides staging support for primitive data types such as
+@Int@ and @Double@, and commonly-used data structures such as arrays and maps.
+The idea is to construct intermediate representations (IR) for the next stage
+program at the current stage. Such conversion from expressions of type @Rep[_]@ to
+their IR is done by using the implicit design pattern. As we will see later,
+implementing the staging support for custom classes is also straightforward.
 
@@
 function that transforms a next-stage value of @Rep[A]@ to @StateT[M,S,B]@.
-\todo{highlight the changes in the code}
+%\todo{highlight the changes in the code}
 

### 2019-04-06 paper/abstract.tex
paper

@@
 \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
-the applicability on various control-flow analyses, and the use to modular
+the application of various control-flow analyses, and the use to modular
 analysis. We also empirically evaluate the performance improved by staging. The

### 2019-04-06 paper/30_stageconc.tex
misc

@@
 be known later. To specialize it we write a function @snippet@ that provides a 
-concrete value @x@ for @x@. After this, we may instantiate an @DslDriver@ object 
+concrete value for @x@. After this, we may instantiate an @DslDriver@ object 
 which provide the specialized program.
@@
     def power(b: Rep[Int], x: Int): Rep[Int] = if (x == 0) 1 else b * power(b, x-1)
-    def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
+    def snippet(b: Rep[Int]): Rep[Int] = power(b, 5)
   }

### 2019-04-06 paper/70_eval.tex
wip evaluation

@@
 experiments are executed for 20 times and we report the statistical median value
-of the running times (in seconds).
+of the running times. We set a 5 minutes timeout.
 
@@
 \footnotesize
-\begin{tabular}{@{}lllll@{}}
+\begin{tabular}{@{}ll|lll|lll@{}}
 \toprule
-    program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ \\ \midrule
-    @fib@               & 32   & 0.013      & 0.002      & 650.00\%          \\
-    @rsa@               & 451  & 0.232      & 0.042      & 552.38\%          \\
-    @church@            & 120  & 67.342     & 18.917     & 355.99\%          \\
-    @fermat@            & 310  & 0.024      & 0.004      & 600.00\%          \\
-    @mbrotZ@            & 331  & 0.763      & 0.127      & 600.79\%          \\
-    @lattice@           & 609  & 31.874     & 8.816      & 361.59\%          \\
-    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 750.00\%          \\
-    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 1290.91\%         \\
-    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 1689.92\%         \\
-    @solovay-strassen@  & 523  & 5.094      & 1.189      & 428.43\%          \\
+    program             &\#AST & unstaged   & staged     & $\frac{\text{unstaged}}{\text{staged}}$ & unstaged   & staged    & $\frac{\text{unstaged}}{\text{staged}}$  \\ 
+    \midrule
+                        &      & \multicolumn{3}{c}{w/o store-widening} &  \multicolumn{3}{c}{w/ store-widening}  \\
+    \midrule
+    @fib@               & 32   & 0.013      & 0.002      & 650.00\%  & & &        \\
+    @rsa@               & 451  & 0.232      & 0.042      & 552.38\%  & & &        \\
+    @church@            & 120  & 67.342     & 18.917     & 355.99\%  & & &        \\
+    @fermat@            & 310  & 0.024      & 0.004      & 600.00\%  & & &        \\
+    @mbrotZ@            & 331  & 0.763      & 0.127      & 600.79\%  & & &        \\
+    @lattice@           & 609  & 31.874     & 8.816      & 361.59\%  & & &       \\
+    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 750.00\%  & & &      \\
+    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 1290.91\% & & &      \\
+    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 1689.92\% & & &      \\
+    @solovay-strassen@  & 523  & 5.094      & 1.189      & 428.43\%  & & &       \\
+    @regex@             & 550  & -          & -          & -         & & &       \\
+    @matrix@            & 1891 & -          & -          & -         & & &       \\
     \bottomrule
 \end{tabular}
-\caption{Evaluation result for 0CFA without store-widening.} \label{evaluation_result}
+\caption{Evaluation result for monovariant control-flow analysis.} \label{evaluation_result}
 \end{figure}
@@
 analysis (in seconds). The last column $\frac{\text{unstaged}}{\text{staged}}$
-shows the improvement from staging. \todo{0cfa, 1cfa, store-widen}
+shows the improvement from staging. The dash - in some cells indicates timeout.
+\todo{0cfa, 1cfa, store-widen}
 As we can see from the table, the staged version significantly outperforms the

### 2019-04-06 oopsla19-code/src/main/scala/sai/evaluation/SAIRuntime.scala

@@ -32,3 +32,33 @@ object SAIRuntime {
   type Config = (Int, Env, Store)
   type Cache = Map[Config, Set[(Value, Store)]]
 }
+
+object SWSAIRuntime {
+  trait AbsValue
+  case object IntV extends AbsValue
+  case object SymV extends AbsValue
+  case object BoolV extends AbsValue
+  case object VoidV extends AbsValue
+  case object FloatV extends AbsValue
+  case object CharV extends AbsValue
+  case object EofV extends AbsValue
+  case object InputPortV extends AbsValue
+  case object OutputPortV extends AbsValue
+  case object MtListV extends AbsValue
+  case object VectorVTop extends AbsValue
+  case object ListVTop extends AbsValue
+  case class PrimOpV(op: String) extends AbsValue
+  case class ConsV(a: Addr, b: Addr) extends AbsValue
+  case class VectorV(vs: List[Addr]) extends AbsValue
+  case class CompiledClo(f: (List[Value], Store, Cache, Cache) => (Set[(Value, Store)], Cache), λ: Int, ρ: Env) extends AbsValue
+
+  trait Addr
+  case class ZCFAAddr(x: String) extends Addr
+  case class OCfaAddr(x: String, ctx: Expr) extends Addr
+
+  type Value = Set[AbsValue]
+  type Env = Map[String, Addr] 
+  type Store = Map[Addr, Value]
+  type Config = (Int, Env)
+  type Cache = Map[Config, Set[Value]]
+}

### 2019-04-06 oopsla19-code/src/main/scala/sai/lms/SetOps.scala

@@ -8,7 +8,7 @@ import org.scala_lang.virtualized.SourceContext
 
 /* LMS support for immutable Sets. */
 
-trait SetOps extends Base with Variables with ListOps {
+trait SetOps extends Base with Variables with ListOps with TupleOps with TupledFunctions {
   //implicit def setTyp[T:Manifest]: Manifest[Set[T]]
 
   object Set {
@@ -31,6 +31,7 @@ trait SetOps extends Base with Variables with ListOps {
     def toList()(implicit pos: SourceContext) = set_toList(s)
     def map[B:Manifest](f: Rep[A] => Rep[B]) = set_map(s, f)
     def filter(f: Rep[A] => Rep[Boolean]) = set_filter(s, f)
+    def foldLeftPairPair[B:Manifest,C:Manifest,D:Manifest](z: Rep[((B,C),D)])(f: (((Rep[B], Rep[C]),Rep[D]), Rep[A]) => Rep[((B,C),D)]) = set_foldLeftPairPair(s, z, f)
   }
 
   def set_new[A:Manifest](xs: Seq[Rep[A]])(implicit pos: SourceContext): Rep[Set[A]]
@@ -44,11 +45,13 @@ trait SetOps extends Base with Variables with ListOps {
   def set_foldLeft[A:Manifest,B:Manifest](s: Rep[Set[A]], z: Rep[B], f: (Rep[B], Rep[A]) => Rep[B])(implicit pos: SourceContext): Rep[B]
   def set_toList[A:Manifest](s: Rep[Set[A]])(implicit pos: SourceContext): Rep[List[A]]
   def set_foldLeftPair[A:Manifest,B:Manifest,C:Manifest](s: Rep[Set[A]], z: Rep[(B,C)], f: ((Rep[B], Rep[C]), Rep[A]) => Rep[(B,C)])(implicit pos: SourceContext): Rep[(B,C)]
+  def set_foldLeftPairPair[A:Manifest,B:Manifest,C:Manifest,D:Manifest](s: Rep[Set[A]], z: Rep[((B,C),D)],
+                                                                        f: (((Rep[B], Rep[C]), Rep[D]), Rep[A]) => Rep[((B,C), D)])(implicit pos: SourceContext): Rep[((B,C), D)]
   def set_map[A:Manifest,B:Manifest](s: Rep[Set[A]], f: Rep[A] => Rep[B])(implicit pos: SourceContext): Rep[Set[B]]
[…]

### 2019-04-06 oopsla19-code/src/main/scala/sai/monads/RepMonads.scala

@@ -392,5 +392,85 @@ trait RepMonads extends RepLattices { self: SAIDsl =>
     def withFilter(f: Rep[A] => Rep[Boolean]): SetReaderStateM[R, S, A] = filter(f)
   }
 
+  /////////////////////////////////////////////
+  object SetStateReaderStateM {
+    type SSRS[R,S1,S2,A] = SetStateReaderStateM[R, S1, S2, A]
+    def apply[R:Manifest,S1:Manifest,S2:Manifest,A:Manifest](implicit m: SetStateReaderStateM[R,S1,S2,A]): SetStateReaderStateM[R,S1,S2,A] = m
+
+    implicit def SetStateReaderStateMonad[R: Manifest, S1: Manifest, S2: Manifest] = new Monad[SetStateReaderStateM[R,S1,S2,?]] {
+        def flatMap[A:Manifest, B:Manifest](ma: SSRS[R,S1,S2,A])(f: Rep[A] => SSRS[R,S1,S2,B]) = ma.flatMap(f)
+        def pure[A: Manifest](a: Rep[A]): SSRS[R,S1,S2,A] = SetStateReaderStateM(r=>s1=>s2=> ((Set[A](a),s1),s2))
+        def filter[A: Manifest](ma: SSRS[R,S1,S2,A])(f: Rep[A] => Rep[Boolean]): SSRS[R,S1,S2,A] = ma.filter(f)
+
+        def put1(s1: => Rep[S1]): SSRS[R,S1,S2,Unit] = SetStateReaderStateM(r => _ => s2 => ((Set[Unit](()), s1), s2))
+        def put2(s2: => Rep[S2]): SSRS[R,S1,S2,Unit] = SetStateReaderStateM(r => s1 => _ => ((Set[Unit](()), s1), s2))
+        def get1: SSRS[R,S1,S2,S1] = SetStateReaderStateM(r => s1 => s2 => ((Set[S1](s1), s1), s2))
+        def get2: SSRS[R,S1,S2,S2] = SetStateReaderStateM(r => s1 => s2 => ((Set[S2](s2), s1), s2))
+
+        def mod1(f: Rep[S1] => Rep[S1]): SSRS[R,S1,S2,Unit] = SetStateReaderStateM(r => s1 => s2 => ((Set[Unit](()), f(s1)), s2))
[…]

### 2019-04-06 oopsla19-code/src/main/scala/sai/evaluation/Evaluation.scala
evaluation

@@ -52,7 +52,8 @@ object Evaluation {
       //(nbody, "nbody"),
       //(nucleic, "nucleic"),
     )
-    progs foreach { case (e, id) => compare(e, id) }
+    //progs foreach { case (e, id) => compare(e, id) }
+    progs foreach { case (e, id) => compareSW(e, id) }
     /*
     progs foreach { case (e, id) => 
       println(s"$id - ${size(e)}")
@@ -109,6 +110,17 @@ object Evaluation {
     }
   }
 
+  def specializeSW(e: Expr): DslDriver[Unit, Unit] = new SWStagedSchemeAnalyzerDriver {
+    @virtualize
+    def snippet(u: Rep[Unit]): Rep[Unit] = {
+      val res = run(e)
+      println(s"Number of values:" + res._1)
+      //println(s"Size of cache:" + res._2.size)
+      //println(res._1)
+      //println(res._2.size)
+    }
+  }
+
   def writeTo(filename: String, content: String): Unit = {
     val writer = new java.io.PrintWriter(new java.io.File(filename))
     writer.write(content)
@@ -117,15 +129,28 @@ object Evaluation {
 
   val output = "CodeGen.out"
   def output(id: String): String = s"CodeGen_$id.out"
+  val N = 20
 
-  def compare(e: Expr, id: String) {
+  def compareSW(e: Expr, id: String) {
     println(s"Running $id, AST size: ${size(e)}")
-    val N = 20
-    //val t1 =  run(N, { evalUnstaged(e) })
     val t1 = run(N, { evalUnstagedSW(e) })
     println(s"[$id] [unstaged] - ${t1}s")
     
-    /*
+    val code = specializeSW(e)
+    code.precompile
+    val outfile = output(id)
+    println(s"[$id] [staged] Finished precompile, writing code to ${outfile}")
+    writeTo(outfile, code.code)
+
+    val t2 = run(N, { code.eval(()) })
+    println(s"[$id] [staged] - ${t2}s")
+  }
+
+  def compare(e: Expr, id: String) {
+    println(s"Running $id, AST size: ${size(e)}")
+    val t1 =  run(N, { evalUnstaged(e) })
+    println(s"[$id] [unstaged] - ${t1}s")
+    
     val code = specialize(e)
     code.precompile
     val outfile = output(id)
@@ -134,6 +159,5 @@ object Evaluation {
 
     val t2 = run(N, { code.eval(()) })
     println(s"[$id] [staged] - ${t2}s")
-    */
   }
 }

### 2019-04-06 oopsla19-code/src/main/scala/sai/evaluation/GenericAnalyzer.scala

@@ -132,8 +132,6 @@ trait SchemeAnalyzer {
       rt <- br(ev)(cnd, thn, els)
     } yield rt
   }
-
-  def foldVss(vss: R[List[Value]]): R[Value]
 }
 
 trait AbstractComponents extends SchemeAnalyzer {

### 2019-04-06 oopsla19-code/src/main/scala/sai/evaluation/SWUnstagedAnalyzer.scala

@@ -1,0 +1,269 @@
+package sai.evaluation
+
+import sai.lms._
+import sai.monads._
+import sai.lattices._
+import sai.lattices.Lattices._
+import sai.evaluation.parser._
+
+// Unstaged 0CFA for Scheme, with store widening
+object SWUnstagedSchemeAnalyzer extends AbstractComponents {
+  import ReaderT._
+  import StateT._
+  import SetT._
+  import IdM._
+
+  type R[T] = T
+  type Config = (Expr, Env)
+  type Cache = Map[Config, Set[Value]]
+
+  // Transformers
+  type EnvT[F[_], B] = ReaderT[F, Env, B]
+  type StoreT[F[_], B] = StateT[F, Store, B]
+  //type NondetT[F[_], B] = ListT[F, B]
+  type NondetT[F[_], B] = SetT[F, B]
+  type InCacheT[F[_], B] = ReaderT[F, Cache, B]
+  type OutCacheT[F[_], B] = StateT[F, Cache, B]
+  // Monads
+  type OutCacheM[T] = OutCacheT[IdM, T]
+  type InOutCacheM[T] = InCacheT[OutCacheM, T]
+  type StoreInOutCacheM[T] = StoreT[InOutCacheM, T]
+  type NdStoreInOutCacheM[T] = NondetT[StoreInOutCacheM, T]
+  type AnsM[T] = EnvT[NdStoreInOutCacheM, T]
+
+  def mapM[A, B](xs: List[A])(f: A => AnsM[B])(implicit mB: Manifest[B]): AnsM[List[B]] = Monad.mapM(xs)(f)
+  def forM[A, B](xs: List[A])(f: A => AnsM[B])(implicit mB: Manifest[B]): AnsM[B] = Monad.forM(xs)(f)
+
+  // Environment operations
+  def ask_env: AnsM[Env] = ReaderTMonad[NdStoreInOutCacheM, Env].ask
+  def ext_env(ans: Ans)(xα: (String, Addr)): Ans = ReaderTMonad[NdStoreInOutCacheM, Env].local(ans)(ρ => ρ + xα)
+  def local_env(ans: Ans)(ρ: Env): Ans = ReaderTMonad[NdStoreInOutCacheM, Env].local(ans)(_ => ρ)
+
+  // Store operations
+  def get_store: AnsM[Store] =
+    ReaderT.liftM[NdStoreInOutCacheM, Env, Store](
+      SetT.liftM[StoreInOutCacheM, Store](
[…]

### 2019-04-06 paper/10_intro.tex
some number and typos

@@
 \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
-annotations (i.e., binding-time annotations) to be explicit from the
+annotations (i.e., binding-time annotations) to be given explicit from the
 programmers. The staging annotations can be either syntactic (as in
@@
 Now to concretely illustrate the idea of Futamura projection, consider that we
-have an operational interpreter @eval@ of some language:
+have an interpreter @eval@ of some language:
 
@@
 
-Now we have seen two orthogonal transformations of interpreters: by
+Now that we have seen two orthogonal transformations of interpreters: by
 specialization, we can refactor interpreters to code generators; by
@@
 We show that these well-known abstractions can be combined smoothly, and more
-importantly almost free. We implement the four interpreters using Scala and the
+importantly for free. We implement the four interpreters using Scala and the
 Lightweight Modular Staging \cite{DBLP:conf/gpce/RompfO10} framework. Through
 staging, the major overhead of abstractions -- from the monadic layers -- is
-completely eliminated in the generated code. We observe up to 17x times faster
-than the unstage abstract interpreter. However, we do not lose any advantages of
-this high-level programming style (e.g., modularity and extensibility); the
-meaning of the program also remains the same, particularly, the soundness of
-static analysis after staging.
+completely eliminated in the generated code. We observe that the staged version is 
+11x times as fast as the unstage abstract interpreter. 
+However, we do not lose any advantages of this high-level programming style 
+(e.g., modularity and extensibility); the meaning of the program also remains 
+the same, particularly, the soundness of static analysis after staging.
 

### 2019-04-06 paper/70_eval.tex
evaluation
some number and typos

@@
 and evaluate the analyzing time against with the unstaged versions. The result
-shows that the staged versions are averagely 6 times as fast of the unstaged
+shows that the staged versions are averagely 11x times as fast as the unstaged
 version, as strong evidence of our \textit{abstraction without regret}
@@
 desugars Scheme to a small core language that makes the analyzer easy to
-implement. As described in Section \ref{cfa}, we implement a monovariant
-0-CFA-like analysis, \todo{a context-sensitive $1$-CFA-like analysis, both their
-  versions with and without store-widening}. We also implement several
-optimizations mentioned in Section \ref{staged_ds} (less than ~30 LOC in total).
+implement. As described in Section \ref{cfa}, we implement two monovariant
+0-CFA-like analysis, one is equipped with store-widening, the other does not.
+We also implement several optimizations mentioned in Section \ref{staged_ds} 
+(less than ~30 LOC in total).
 
@@
     \midrule
-                        &      & \multicolumn{3}{c}{w/o store-widening} &  \multicolumn{3}{c}{w/ store-widening}  \\
+                        &      & \multicolumn{3}{c}{w/o store-widening (sec.)}  &  \multicolumn{3}{c}{w/ store-widening (ms)}\\
     \midrule
-    @fib@               & 32   & 0.013      & 0.002      & 650.00\%  & & &        \\
-    @rsa@               & 451  & 0.232      & 0.042      & 552.38\%  & & &        \\
-    @church@            & 120  & 67.342     & 18.917     & 355.99\%  & & &        \\
-    @fermat@            & 310  & 0.024      & 0.004      & 600.00\%  & & &        \\
-    @mbrotZ@            & 331  & 0.763      & 0.127      & 600.79\%  & & &        \\
-    @lattice@           & 609  & 31.874     & 8.816      & 361.59\%  & & &       \\
-    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 750.00\%  & & &      \\
-    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 1290.91\% & & &      \\
-    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 1689.92\% & & &      \\
-    @solovay-strassen@  & 523  & 5.094      & 1.189      & 428.43\%  & & &       \\
-    @regex@             & 550  & -          & -          & -         & & &       \\
-    @matrix@            & 1891 & -          & -          & -         & & &       \\
+    @fib@               & 32   & 0.013      & 0.002      & 6.50x      & 1.537 & 0.067  &  22.87x       \\
+    @rsa@               & 451  & 0.232      & 0.042      & 5.52x      & 15.113& 1.449 &  10.43x       \\
+    @church@            & 120  & 67.342     & 18.917     & 3.56x      & 2.962 & 0.291 &  10.19x       \\
+    @fermat@            & 310  & 0.024      & 0.004      & 6.00x      & 7.624 & 0.622 &  12.26x       \\
+    @mbrotZ@            & 331  & 0.763      & 0.127      & 6.01x      & 12.229& 0.729 &  16.77x       \\
+    @lattice@           & 609  & 31.874     & 8.816      & 3.62x      & 17.943& 0.909 &  19.75x      \\
+    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 7.50x      & 4.658 & 0.785 &  5.93x     \\
+    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 12.91x     & 10.005& 1.680 &  5.95x     \\
+    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 16.90x     & 22.999& 3.872 &  5.94x     \\
+    @solovay-strassen@  & 523  & 5.094      & 1.189      & 4.28x      & 19.118& 0.654 &  29.25x      \\
+    @regex@             & 550  & -          & -          & -          & 7.133 & 0.624 &  11.43x      \\
+    @matrix@            & 1891 & -          & -          & -          & 86.176& 4.592 &  18.76x      \\
     \bottomrule
@@
 Figure~\ref{evaluation_result} shows the evaluation result, comparing the
-performance improved by staging on 0CFA without store-widening. The
+performance improved by staging on 0CFA without and with store-widening. The
 \textit{unstaged} and \textit{staged} columns show the median time to finish the
-analysis (in seconds). The last column $\frac{\text{unstaged}}{\text{staged}}$
+analysis. The column $\frac{\text{unstaged}}{\text{staged}}$
 shows the improvement from staging. The dash - in some cells indicates timeout.
-\todo{0cfa, 1cfa, store-widen}
 As we can see from the table, the staged version significantly outperforms the
 unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
-We observe the most improvement of 17x times faster than the unstaged version on
-@kcfa-worst-64@. On average, the staging produces code that runs 7x times faster
+We observe the most improvement of 29x times faster than the unstaged version on
+@solovay-strassen@. On average, the staging produces code that runs 11x times faster
 than the unstaged analyzer.

### 2019-04-06 paper/abstract.tex
some number and typos

@@
 overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of 7x with staging for control-flow analysis.
+experiment shows an average speedup of 11x with staging for control-flow analysis.
 

### 2019-04-06 paper/10_intro.tex
tweak figure
revise intro

@@
 \cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
-annotations (i.e., binding-time annotations) to be given explicit from the
+annotations (i.e., binding-time annotations) to be given explicitly from the
 programmers. The staging annotations can be either syntactic (as in
@@
 
-where $e$ is the program of type @Expr@ and @arg@ is the input to program $e$,
+where $e$ is a program of type @Expr@ and @arg@ is the input to $e$,
 and @eval@ produces a value.
@@
 programs and inputs. Given a program $e_0$, by applying the first Futamura
-projection, we may obtain a specialized interpreter
-$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$ -- the so-called \textit{equivalent executable}. 
+projection, we can obtain a specialized interpreter
+$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$, which is the so-called \textit{equivalent executable}. 
 By definition of the interpreter, 
 $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
-argument @arg@ to an environment (as well as an accompanying heap object to
-support mutable variables) that maps free variables in $e$ to some
+argument @arg@ to an environment that maps free variables in $e$ to some
 values, the interpreter is a standard environment-passing interpreter,
@@
 DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
-
-\paragraph{Futamura Projection of Abstract Interpreters}
-
-Now that we have seen two orthogonal transformations of interpreters: by
-specialization, we can refactor interpreters to code generators; by
-approximation, we can refactor interpreters to static analyzers. As an
-intellectual quest, we should naturally ask -- starting from an abstract
-interpreter, \textit{can we derive a sound abstract interpreter while it
-produces optimized code that does the analysis?}
-Conceptually, it is similar to the concrete setting: for an
-abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
-\widehat{\texttt{Env}} \to \widehat{\texttt{Value}}$, which takes a program, an abstract
-environment and returns abstract values, we would like to specialize it with
-respect to some program $e_0$ and produce a compiled analysis
-$\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
-\widehat{\texttt{Value}}$, such that $
-\widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
-\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$, but the compiled one leaves the
-abstract environment $\widehat{\rho}$ as remaining input.
 
@@
 
+\paragraph{Futamura Projection of Abstract Interpreters}
+
+Now that we have seen two orthogonal transformations of interpreters: by
+specialization, we can refactor interpreters to code generators; by
+approximation, we can refactor interpreters to static analyzers. As an
+intellectual quest, we should naturally ask -- starting from an abstract
+interpreter, \textit{can we derive a sound abstract interpreter while it
+produces optimized code that does the analysis?}
+Conceptually, it is similar to the concrete setting: given an
+abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
+\widehat{\texttt{Env}} \to \widehat{\texttt{Value}}$, which takes a program, an abstract
+environment and returns abstract values, we would like to specialize it with
+respect to some program $e_0$ and produce a compiled analysis
+$\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
+\widehat{\texttt{Value}}$, where the abstract environment $\widehat{\rho}$ is kept as input.
+The specialized analysis satisfies $\widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
+\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$.
+
 In this paper, we study the confluence of two old ideas --- Futamura projection
@@
 the interpreter differently, we can automatically derive a concrete interpreter,
-or an abstract interpreter, or their code-generator versions, respectively.
+or an abstract interpreter, or their staged versions that generate code, respectively.
 
@@
 \begin{itemize}
-\item Env-store-value abstractions, which control whether the interpreter
-  behaves as a concrete interpreter or an abstract interpreter.
+\item Env-store-value abstractions, which define the domains that the interpreter will be running with.
   The environments, stores, and values are initially declared as abstract types in the generic interpreter.
@@
   The interpreter itself is written in monadic form, and by plugging different monads, 
-  the underlying semantics of the interpreter is changed.
+  the underlying semantics of the interpreter is changed. We also use monad transformers to combine multiple monadic effects.
   %\todo{mention that direct-style also works?}
@@
 completely eliminated in the generated code. We observe that the staged version is 
-11x times as fast as the unstage abstract interpreter. 
-However, we do not lose any advantages of this high-level programming style 
-(e.g., modularity and extensibility); the meaning of the program also remains 
+on average 11x times as fast as the unstaged abstract interpreter. 
+However, we do not lose any advantages of this high-level programming style,
+e.g., modularity and extensibility. The meaning of generated programs also remain
 the same, particularly, the soundness of static analysis after staging.
@@
 rewritten to the closure generation form. In this paper, we advocate a systematic
-approach of meta-programming that achieves the performance goal with less
+approach of generative programming that achieves the performance goal with less
 intrusive changes and engineering efforts.
@@
 achieve the same optimization by utilizing type-based stage annotations,
-meanwhile, the analyzer program does not need any changes.
-(2) We extend the basic staged abstract interpreter to diverse control-flow analysis, including store-widening analysis, context-sensitive analysis and abstract garbage collection.
+meanwhile, the analyzer program does not need any changes. (2) We extend the 
+basic staged abstract interpreter to diverse control-flow analysis, 
+including a store-widened analysis, a context-sensitive analysis and abstract garbage collection.
 (3) Staging abstract interpreter enables modular analysis and provides a mechanized
 way to create library summaries. We notice that we may apply the staged abstract
-interpreter separately on the different part of the analyzed program, the
+interpreter separately on the different part of the analyzed program, and the
 generated code is reusable and composable. Therefore, we can turn a
@@
   version by replacing the binding-time types (Section~\ref{stagedinterp}).
-\item We discuss the unstaged abstract interpreter under the same framework by
+\item We present the unstaged abstract interpreter under the same framework by
   replacing the environment, store and values (Section~\ref{unstaged_abs}).

### 2019-04-06 paper/abstract.tex
revise intro

--- 2019-04-06 paper/abstract.tex
@@
 \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
-the application of various control-flow analyses, and the use to modular
+the application on various control-flow analyses, and the use to modular
 analysis. We also empirically evaluate the performance improved by staging. The
 overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of 7x with staging for control-flow analysis.
+experiment shows an average speedup of 11x times with staging for control-flow analysis.
 

### 2019-04-06 paper/10_intro.tex
misc

@@
  \item Pragmatically, by case studies and empirical evaluation, we show that staging 
-   abstract interpreters is useful to improve performance while untouching
+   abstract interpreters is useful to improve performance while not touching
    the soundness. 

### 2019-04-06 paper/70_eval.tex
eval

@@
 Figure~\ref{evaluation_result} shows the evaluation result, comparing the
-performance improved by staging on 0CFA without and with store-widening. The
+performance improved by staging on two monovariant CFA (without and with store-widening). The
 \textit{unstaged} and \textit{staged} columns show the median time to finish the
@@
 As we can see from the table, the staged version significantly outperforms the
-unstaged version on all benchmarks, especially for the @kcfa-worst-@$n$ series.
+unstaged version on all benchmarks. %, especially for the @kcfa-worst-@$n$ series.
 We observe the most improvement of 29x times faster than the unstaged version on

### 2019-06-14 paper/50_stageabs.tex
revision note

@@
 
+\iffalse
+Revision: Solving Practical Challenges.
+theoretically, all the things should work nicely. Unfolding the interpreter over the AST.
+But the generated code is blowed up. For example .... This should not affect the correctness, 
+but poses burden on the MSP system (ie LMS) and the next stage compiler/runtime (ie, scalac and JVM).
+1) LMS becomes slower since a large IR graph is contructed during the staging.
+2) Scalac becomes slower when reading a such large source code.
+JVM has certain limitation on the size of a single method.
+
+TODO: can we formulate selective caching as a partially-static data law.
+TODO: lambda lifting for if
+\fi
+
 Our staging schema works well in theory, but in practice would suffer from code
@@
 expressions such as literals or variables, as they always terminate. Borrowing
-the idea from the partition in ANF \cite{Flanagan:1993:ECC:155090.155113}, we
-may use a selective caching algorithm:
+the idea from the partition of A-Normal Form \cite{Flanagan:1993:ECC:155090.155113},
+we may use a selective caching algorithm:
 

### 2019-06-15 paper/10_intro.tex
update

@@
 \section{Introduction} \label{intro}
+
+\iffalse
+  Revision: 
+  also motivate from the perspective that ADI is slow,
+  analyze the reason: 1) interpretive overhead, 2) the use of monadic layers,
+  3) the intrinsic complexity of abs int of HO language (ie CFA particularly).
+  Then introduce the abstraction without regret.
+  contribution: derive staged monad transformers.
+\fi
 
@@
 annotations to introduce the idea of MSP:
+
+\rev{this power is based on the induction of natural number e}
 
@@
 
-\paragraph{Futamura Projection of Abstract Interpreters}
+\paragraph{Intellectual Motivation: Futamura Projection of Abstract Interpreters}
 
@@
 
-\paragraph{Abstraction without Regret}
+\paragraph{Practical Motivation: Abstraction without Regret}
+\rev{Abstraction without performance regret; Performance without soundness regret}
 

### 2019-06-15 oopsla19-code/src/main/scala/sai/MainAbs.scala

@@ -268,6 +268,7 @@ trait StagedAbsInterpreterOps extends SAIDsl with RepMonads with RepLattices {
   }
   //FIXME: check v1 && v2 contains Int
   def prim(op: Symbol, v1: Rep[Value], v2: Rep[Value]): Rep[Value] = Set[AbsValue](emit_inttop)
+
   def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = {
     lift_nd[AbsValue](fun.toList).flatMap { clo =>
       ask_in_cache.flatMap { in =>

### 2019-06-15 oopsla19-code/src/main/scala/sai/evaluation/StagedAnalyzer.scala

@@ -18,7 +18,6 @@ import sai.evaluation.parser._
 trait StagedSchemeAnalyzerOps extends AbstractComponents with RepMonads with RepLattices with SAIDsl {
   import ReaderT._
   import StateT._
-  //import ListReaderStateM._
   import SetReaderStateM._
 
   //type Config = (Expr, Env, Store)

### 2019-06-29 paper/10_intro.tex
misc

@@
   Revision: 
+  source of overhead: 1) interpreration, dispatch, recursion.
+  2) modern ai uses effect systems to construcut abs interp, in a purely functional way.
+  which has certain benefits, but introduce additional abstraction layers may lead to 
+  performance overhead. Such as MAAM, ADI, Semantic
+
   also motivate from the perspective that ADI is slow,

### 2019-06-29 paper/10_intro.tex
revise intro

--- 2019-04-06 paper/10_intro.tex
@@
 
-Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close
-connection between compilers and interpreters. The first Futamura projection
-specifically shows that specializing an interpreter with respect to the input
-program yields an equivalent executable. Partial evaluation
-\cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
-Futamura projections: it identifies the binding-time of variables in
-the program (i.e., they can be known statically or dynamically),
-evaluates the static part, and finally generates a residual program that solely
-relies on the dynamic part. However, given an arbitrary program, precisely
-analyzing its binding-times is hard in general. As an alternative and pragmatic
-approach to specialization and partial evaluation, multi-stage programming (MSP) 
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
-annotations (i.e., binding-time annotations) to be given explicitly from the
-programmers. The staging annotations can be either syntactic (as in
-MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \cite{DBLP:conf/gpce/RompfO10}.
-These staging annotations identify which part of the inputs should be
-specialized. As a classical example, we use the power function and type-based
-annotations to introduce the idea of MSP:
+Statically analyzing semantic properties of a program is a widely known
+undecidable class of problem. Abstract interpretation as a lattice-based
+approach to construct sound static analysis was proposed by
+\citet{DBLP:conf/popl/CousotC77}. With the Galois connection, one can
+approximate to the program runtime behaviors by computing the fixed points on
+the abstract domain. Despite the tremendous theoretical development of abstract
+interpretation over the years, on the side of pragmatics, constructing
+artifacts and analyzers that perform sound abstract interpretation was 
+considered abstruse and complicated in practice for a long time.
 
-\begin{lstlisting}
-  def power(b: Rep[Int], e: Int): Rep[Int] = if (e == 0) 1 else b * power(b, e - 1)
-\end{lstlisting}
+In recent years, we observe a rich progress on how we can construct abstract
+interpreters from systematic principles, instead of ad-hoc engineerings. A
+notable one is the Abstracting Abstract Machine (AAM) methodology
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}, which uncovers an
+approach to derive sound abstract interpreters from their concrete counterparts,
+where the soundness can be easily established by examining the transformation of
+semantic artifacts. For example, a CEK machine
+\cite{DBLP:conf/popl/FelleisenF87} for concrete execution can be easily
+refactored to an effective 0-CFA control-flow analysis
+\cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672} by first
+tweaking the environment dereference as a nondeterministic choice, then
+allocating continuations in the environment, and constraining the addresses
+space to be finite. This systematic abstraction approach can be tailored to
+different analyses \todo{cite} and and has been applied to multiple variants of
+small-step abstract machines \cite{DBLP:journals/jfp/HornM12,
+DBLP:conf/icfp/HornM10, Sergey:2013:MAI:2491956.2491979} and big-step
+definitional interpreters \cite{Wei:2018:RAA:3243631.3236800,
+DBLP:journals/pacmpl/DaraisLNH17, Keidel:2018:CSP:3243631.3236767}.
 
-If the programmer identifies that @b@ will be known in the future stage (as shown
-on its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known at the
-current stage, say 5, then we can specialize the function @power@ with
-respect to @e = 5@, and generate a specialized function where the overhead of
-recursion is eliminated. Conceptually, the generated code looks like the following:
+More pragmatically, as an implementation strategy, several purely functional
+programming approaches to build abstract interpreters were emerged, for
+examples, using monads or arrows. The pure approach provides much benefits: 1)
+the abstract interpretation artifacts are compositional and can be constructed
+modularly, 2) one can easily prove the soundness of analysis, and 3) one can
+easily reason the correctness of implementation by leveraging equational
+reasoning. For example, monadic abstract interpreters \todo{two level semantics,
+arrow, Galois transformers}
 
-\begin{lstlisting}
-  def power5(b: Int): Int = b * b * b * b * b
-\end{lstlisting}
+However, besides the intrinsic complexity of static analysis, there are
+additional abstraction penalties with these high-level programming approaches.
+First, similar to concrete interpreters, the abstract interpreter analyzes the
+program by traversing the abstract syntax tree, which poses the interpretive
+overhead, e.g., pattern matching on the AST and recursive calls on the sub
+expressions. Those kind of overhead can be negligible if the abstract
+interpreter only runs on the program once, but also can be accumulated
+significantly if it runs repeatedly, for example, on libraries. Second, the
+abstract interpreter written in pure languages usually extensively uses effect
+systems to model the behaviors of abstract interpretatation, such as
+nondeterminism. For example, \citet{DBLP:journals/pacmpl/DaraisLNH17} and
+\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract interpreters use
+monad transformers; the very recent \citet{Githubsemantic}'s Semantic framework
+uses extensible effects. The effect systems may introduce additional performance
+overhead.
 
-Now to concretely illustrate the idea of Futamura projection, consider that we
-have an interpreter @eval@ of some language:
+In this paper, we propose an abstraction-without-regret approach to eliminate
+those performance penalties for abstract interpreters written in purely
+functional programming style. In short, our approach borrows ideas from program
+specialization and high-performance DSLs: 1) Using multi-stage programming at
+type level (the Lightweight Modular Staging framework
+\cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
+specialize the abstract interpreter with respect to the input program and then
+generate efficient low-level code that does the actual analysis. 2) We use the
+tagless-final approach to embed the target language interpreter in a high-level
+host language (Scala, in this paper), allowing the user to implement different
+semantics modularly, including the abstract semantics, without changing the main
+generic interpreter. This design choice allows user to derive staged abstract
+interpreters without intrusive changes to the unstaged one. Therefore, our
+approach is no regret in the sense of both performance and least engineering
+efforts to achieve such performance. We elaborate these two main ideas in detail
+as follow.
 
-\begin{lstlisting}
-  def eval(e: Expr)(arg: Input): Value = ...
-\end{lstlisting}
+\paragraph{Futumura Projection of Abstract Interpreters}
 
-where $e$ is a program of type @Expr@ and @arg@ is the input to $e$,
-and @eval@ produces a value.
-Semantically, the interpreter satisfies $ \texttt{eval}(e)(arg) = [\![ e ]\!] arg$ for all
-programs and inputs. Given a program $e_0$, by applying the first Futamura
-projection, we can obtain a specialized interpreter
-$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$, which is the so-called \textit{equivalent executable}. 
-By definition of the interpreter, 
-$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
-argument @arg@ to an environment that maps free variables in $e$ to some
-values, the interpreter is a standard environment-passing interpreter,
-where the specialization through MSP is also applicable.
+specialization of interpreters/abstract interpreters
 
-Roughly at the same time when Futamura projections were envisioned in the 1970s,
-\citet{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
-lattice-based semantic approach to construct sound static analysis, by
-approximation of fix-points. However, constructing sound abstract interpreters
-was considered abstruse and complicated for a long time.
-Recent progress such as Abstracting Abstract Machines (AAM)
-\cite{DBLP:conf/icfp/HornM10} uncovers a systematic method to derive sound
-abstract interpreters from their concrete counterparts, where the soundness can
-be easily established by examining the transformation of semantics. The AAM
-approach also has been applied to different variants of definitional
-interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
-DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.
+\paragraph{Generic Interpreters}
 
-\vspace{-1em}
-\begin{figure}[h]
-  \begin{tikzpicture}
-  \matrix (m) [matrix of math nodes,row sep=3em,column sep=3em,minimum width=2em]
-  {
-    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_abs}) \end{smallmatrix} &
-    \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{sai}) \end{smallmatrix} \\
-    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_conc}) \end{smallmatrix} &
-    \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{stagedinterp}) \end{smallmatrix} \\ };
-  \path[-stealth]
-    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
-    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
-    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
-    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
-  \end{tikzpicture}
-  \caption{The confluence of specialization and approximation.}
-  \label{confluence}
-\end{figure}
-\vspace{-1em}
-
-\paragraph{Futamura Projection of Abstract Interpreters}
-
-Now that we have seen two orthogonal transformations of interpreters: by
-specialization, we can refactor interpreters to code generators; by
-approximation, we can refactor interpreters to static analyzers. As an
-intellectual quest, we should naturally ask -- starting from an abstract
-interpreter, \textit{can we derive a sound abstract interpreter while it
-produces optimized code that does the analysis?}
-Conceptually, it is similar to the concrete setting: given an
-abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
-\widehat{\texttt{Env}} \to \widehat{\texttt{Value}}$, which takes a program, an abstract
-environment and returns abstract values, we would like to specialize it with
-respect to some program $e_0$ and produce a compiled analysis
-$\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
-\widehat{\texttt{Value}}$, where the abstract environment $\widehat{\rho}$ is kept as input.
-The specialized analysis satisfies $\widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
-\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$.
-
-In this paper, we study the confluence of two old ideas --- Futamura projection
-and abstract interpretation, but from the perspective of their very recent
-realizations --- multi-stage programming and definitional abstract interpreters.
-To be specific, we present the application of the first Futamura projection on
-definitional abstract interpreters, which intersects as staged abstract
-interpreters. We develop a unified approach to fulfill the four different
-semantics, as shown in Figure~\ref{confluence}. Starting with a generic but
-parametric interpreter shared by the four semantics, by simply instantiating
-the interpreter differently, we can automatically derive a concrete interpreter,
-or an abstract interpreter, or their staged versions that generate code, respectively.
+specializing interpreters is not new, Futumura projections, four different tasks
+for language designers and engineers, writing definitional interpreters for the
+language (concrete semantics close to deno/op semantics), compilers for the
+language (translate to another language, low-level), writing static analyzer for
+the language (abstract semantics property we want to analyze), and a performant
+static analyzer (ultimate goal of this paper). shared a single generic
+interpreter
 
 \iffalse
-embedded domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
-DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14,
-Hofer:2008:PED:1449913.1449935},
-The key idea that enables this flexibility is to use abstract type members first
-abstract over concrete/abstract components (e.g., concrete values or abstract
-values), then as well as the binding-time of them (e.g., static or dynamic). By
-staging, the overhead caused by the monadic layers is eliminated in the
-generated code.
+On the other side, static analysis is a tradeoff between performance and
+precision: higher precision usually leads to longer running time.
+
+4. Existing method to improve the performance is adhoc, engineering heavy, require to rewrite the optimized version, therefore harder to reason about the correctness
+6. program analyzers are also meta-programs, they manipulate other programs as data objects
 \fi
-
-\paragraph{Abstraction without Regret}
-
-To enable such flexibility between different semantics, we combine generic
-programming and generative programming, and extensively use abstract type
-members over the following abstractions, which can be later instantiated.
-
-\begin{itemize}
-\item Env-store-value abstractions, which define the domains that the interpreter will be running with.
-  The environments, stores, and values are initially declared as abstract types in the generic interpreter.
-  As shown by the AAM approach \cite{DBLP:journals/jfp/HornM12,
-  DBLP:conf/icfp/HornM10}, if we widen the values to abstract domains (such as
-  intervals) and refactor the store to a mapping from finite addresses to sets of
-  values, we can derive a computable and sound abstract interpreter.
-\item Binding-time abstractions, which control whether the interpreter produces
-  values or generates code. The binding-times differentiate whether data are 
-  known statically or dynamically. Unsurprisingly, such binding-time
-  information can be encoded into higher-kinded types 
-    \cite{DBLP:journals/jfp/CaretteKS09, Ofenbeck:2017:SGP:3136040.3136060}.
-\item Monadic abstractions, which encapsulate the above two abstractions.
-  The interpreter itself is written in monadic form, and by plugging different monads, 
-  the underlying semantics of the interpreter is changed. We also use monad transformers to combine multiple monadic effects.
-  %\todo{mention that direct-style also works?}
-  %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
-\end{itemize}
-
-We show that these well-known abstractions can be combined smoothly, and more
-importantly for free. We implement the four interpreters using Scala and the
-Lightweight Modular Staging \cite{DBLP:conf/gpce/RompfO10} framework. Through
-staging, the major overhead of abstractions -- from the monadic layers -- is
-completely eliminated in the generated code. We observe that the staged version is 
-on average 11x times as fast as the unstaged abstract interpreter. 
-However, we do not lose any advantages of this high-level programming style,
-e.g., modularity and extensibility. The meaning of generated programs also remain
-the same, particularly, the soundness of static analysis after staging.
-
-%\paragraph{Performance via Specialization}
-Besides the intellectual merit shown in the confluence diagram, this paper
-contributes to a practical problem in static analysis: \textit{can we speed up a
-static analyzer without any intrusive changes and soundness compromises?} Static
-analysis is known as a trade-off game between \textit{precision} and
-\textit{efficiency}, we argue that applying meta-programming (multi-stage
-programming, in particular) with abstract interpreters is an effective approach
-to improve the latter while the former is untouched.
-In the literature, specializing analysis with respect to the input program
-\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
-ashley:practical} is not new, despite the fact that most of the previous works
-are ad-hoc on a particular analysis or requiring significant changes to the
-analyzer. For example, abstract compilation is one of the techniques
-\cite{Boucher:1996:ACN:647473.727587}. But it requires the whole analyzer to be
-rewritten to the closure generation form. In this paper, we advocate a systematic
-approach of generative programming that achieves the performance goal with less
-intrusive changes and engineering efforts.
-
-\paragraph{Applications}
-
-Furthermore, we show various applications of staging an abstract interpreter.
-(1) We compare our approach with abstract compilation, and show that we can
-achieve the same optimization by utilizing type-based stage annotations,
-meanwhile, the analyzer program does not need any changes. (2) We extend the 
-basic staged abstract interpreter to diverse control-flow analysis, 
-including a store-widened analysis, a context-sensitive analysis and abstract garbage collection.
-(3) Staging abstract interpreter enables modular analysis and provides a mechanized
-way to create library summaries. We notice that we may apply the staged abstract
-interpreter separately on the different part of the analyzed program, and the
-generated code is reusable and composable. Therefore, we can turn a
-whole-program analyzer into an analyzer that runs modularly.
-
-\iffalse
-Our next contribution is to enable modular analysis based on a whole-program
-analyzer by meta-programming. Modern software are shipped with large libraries
-\cite{toman_et_al:LIPIcs:2017:7121}, and analyzing programs with large libraries
-is usually expensive or unnecessarily repeated for whole-program analyzing
-algorithms. For example, it has been shown that analyzing a simple ``Hello
-World'' program in Java depends on additional 3,000 classes in the library
-\cite{DBLP:conf/oopsla/KulkarniMZN16}. A natural idea is to analyze libraries
-and generate summaries, later we can plug in the necessary information into the
-summaries and finish the analysis; therefore such summaries should be modular
-and reusable. Meta-programming gives us a chance to readily tweak a
-whole-program analyzing algorithm into an analyzer that works modularly and
-produces summaries. Still, the idea is to specialize the abstract interpreter
-that was used for whole-program analysis but on a subset of the program, for
-examples, a set of generic functions for list manipulation. Such specialization
-works on lambda expressions and generates code, which is partial analysis
-results and can be completed by running with the necessary arguments (for
-example, an environment that closes the lambda term) at next stages. Multiple
-partial analysis results also can be reused later and composed together as they
-are just normal programs at the next stage. Through staging, we mechanically
-obtain a modular analysis, even though the original algorithm is formulated as a
-whole-program analysis.
-\fi
-
-\paragraph{Contributions} Briefly, the contribution of the paper is twofold:
-\begin{enumerate}[leftmargin=2em]
-  \item Intellectually, through multi-stage programming, we present a unified construction 
-    that naturally extends the first Futamura projection to abstract interpreters.
- \item Pragmatically, by case studies and empirical evaluation, we show that staging 
-   abstract interpreters is useful to improve performance while not touching
-   the soundness. 
-\end{enumerate}
-\paragraph{Organization} The paper is organized as follows:
-\begin{itemize}[leftmargin=2em]
-\item Starting from a generic interface of the interpreter for a small higher-order language
-  (Section~\ref{prelim}), we first show the instantiation of concrete
-  interpretation (Section~\ref{unstaged_conc}), and then lift it to the staged
-  version by replacing the binding-time types (Section~\ref{stagedinterp}).
-\item We present the unstaged abstract interpreter under the same framework by
-  replacing the environment, store and values (Section~\ref{unstaged_abs}).
-  Finally, we show the combination of approximation and specialization, dubbed
-  \textit{staged abstract interpreters}, can be readily derived (Section~\ref{sai}).
-\item We conduct three case studies on the applicability of our approach
-  (Section~\ref{cases_study}).
-  \begin{enumerate}
-  \item We first revisit the Abstract Compilation (AC)
-    \cite{Boucher:1996:ACN:647473.727587} (Section~\ref{cs_ac}), showing that we can
-    achieve the same optimization with much less engineering effort.
-  \item We extend the staged abstract interpreter to various control-flow
-    analysis (Section \ref{cfa}), including context-sensitivity, store-widening, and abstract
-    garbage collection, showing that our approach is widely applicable.
-  \item We apply our approach to modular analysis (Section \ref{modular}),
-    i.e., analyzing the components of the program separately and composing them later,
-    showing a novel application of generative programming with program analysis.
-  \end{enumerate}
-\item We empirically evaluate the performance improvement by staging on
-    control-flow analysis (Section~\ref{evaluation}). We compare the both
-    context-insensitive and store-widening analysis.
-\end{itemize}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\iffalse
-Likewise, specializing static analyses by partial evaluation emerged in the late 90s
-, and indeed it is able to effectively remove the interpretive
-overhead of repeatedly traversing the abstract syntax tree. However, these
-previous works focused mostly on one particular analysis, or required to
-completely rewrite the analyzer. Hence, it is worth to investigate the idea from
-a modern perspective based on generative programming, especially for a general
-abstract interpreter that models direct-style $\lambda$-calculus, imperative
-features such as mutation, and different abstract domains. One technical
-challenge is the binding-time engineering when non-determinism, fixed-point
-iterations, different abstract domains, and staging are introduced at the same
-time. In this paper, we present an end-to-end design and implementation of
-staging an abstract interpreter; that means not only the interpreter that
-traverses the abstract syntax tree, but also the data structure of abstract
-domains, abstract environment and heap, and the fixed-point iteration are
-staged.
-
-\todo{rewrite this paragraph}
-On the other hand, the abstract interpreter, as a semantic artifact, should be
-written in a style that is easy for people to communicate the formulation and
-abstraction, but also can be implemented efficiently. As the slogan of
-multi-stage programming said, "abstraction without regret", we draw a connection
-between high-level description and efficient implementation of abstract
-interpreters, just like the connection between concrete interpreters and
-compilers drawn by the Futamura projections. Particularly, we show an easy but
-systematic way of adding stage annotations to the abstract interpreter without
-changing the code of the interpreter skeleton, which is shared between four concrete
-or abstract, unstaged or staged interpreters. We use the LMS framework for staging,
-which allows us only use types to annotate the binding-time. Therefore, the
-proposed approach bridges the gap between designing sound static analysis and
-implementing efficient program analyzer.
-
-%Of course, all interpreters are metalinguistic abstractions, but some
-%interpreters are more "abstract" than others \todo{maybe rephrase}.
-%Particularly, we also would like a systematic approach to optimize program
-%analyzers and meanwhile minimally modifying the analyzer programs.
-
-When staging a concrete interpreter, the programmers need to distinguish static
-and dynamic values --- the given program to be executed by the interpreter is
-classified as static because it is known at compile-time, and the inputs to that
-program are dynamic. However, when staging an abstract interpreter, this
-distinction does not exist anymore. Because the abstract interpreter
-instantiates all the inputs as some form of abstract values, which are usually
-top elements in their abstract domains and are also statically known. Then what
-is the point of staging if there is no such distinction? A surprising by-product
-of thinking about this question is to realize that we can apply the staged
-abstract interpreter on \textit{open} programs, and the free variables
-representing other parts of the program (e.g., libraries) are dynamic inputs,
-therefore we obtain a modular analysis through staging, mechanically.
-\note{TR: I don't understand this. The program structure is static, the abstract values
-are still dynamic, no? They change in every iteration of the fixpoint algorithm}
-\note{TR: why is it a big deal. On Closed programs, we obtain a constant factor,
-but modular analysis has different asymptotics}
-
-When targeting higher-order programs, either staged concrete interpreters or
-staged abstract interpreters are able to compile a closure, i.e., specialize the
-call of interpreter with respect to the body expression of the lambda term
-without knowing the actual argument value. By generalizing this observation, we
-can actually specialize the abstract interpreter with any open programs, which
-unexpectedly leads to a modular analysis and improves scalability. An open
-program contains free variables, which represent other parts of the program and
-will be left as dynamic values. For instance, one challenge in static analysis
-of modern software is that the programs are usually shipped with large library
-code \cite{toman_et_al:LIPIcs:2017:7121}, for example, it has been shown that
-analyzing a simple "Hello World" program in Java depends on additional 3,000
-classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}. A precise
-whole-program analysis formulated as abstract interpreters can be very expensive
-due to the scale of the program and the inherent complexity of the algorithm.
-However, analyzing these libraries is sometimes unnecessarily repeated. When
-applying the staged abstract interpreter on such open programs, we leave the
-unknown arguments and calling contexts as dynamic values and generate partial
-analyzing result which is represented as a residual program. The partial
-analyzing result can be reused and composed with the analyzing result of
-application code when available later. Therefore we mechanically obtain a
-modular analysis through staging, even though the original algorithm is
-formulated as a whole-program analysis.
-
-It has been observed that partially applying context-sensitivity on selected
-portions of the program could improve the precision and efficiency
-\cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. We show that staging
-abstract interpreters as an approach to effectively implement hybrid
-context-sensitivity\todo{}.
-\fi
-
-%\note{Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},}

### 2019-06-30 paper/10_intro.tex
intro

@@
 
-Statically analyzing semantic properties of a program is a widely known
-undecidable class of problem. Abstract interpretation as a lattice-based
-approach to construct sound static analysis was proposed by
-\citet{DBLP:conf/popl/CousotC77}. With the Galois connection, one can
-approximate to the program runtime behaviors by computing the fixed points on
-the abstract domain. Despite the tremendous theoretical development of abstract
-interpretation over the years, on the side of pragmatics, constructing
-artifacts and analyzers that perform sound abstract interpretation was 
-considered abstruse and complicated in practice for a long time.
+Statically analyzing semantic properties of a program is a widely
+known undecidable class of problem. Abstract interpretation as a
+lattice-based approach to construct sound static analysis was proposed
+by \citet{DBLP:conf/popl/CousotC77}. With the Galois connection, one
+can approximate to the program runtime behaviors by computing the
+fixed points on the abstract domain. Despite the tremendous
+theoretical development of abstract interpretation over the years, on
+the side of pragmatics, constructing artifacts and analyzers that
+perform sound abstract interpretation was considered abstruse and
+complicated in practice for a long time.
 
-In recent years, we observe a rich progress on how we can construct abstract
-interpreters from systematic principles, instead of ad-hoc engineerings. A
-notable one is the Abstracting Abstract Machine (AAM) methodology
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}, which uncovers an
-approach to derive sound abstract interpreters from their concrete counterparts,
-where the soundness can be easily established by examining the transformation of
-semantic artifacts. For example, a CEK machine
-\cite{DBLP:conf/popl/FelleisenF87} for concrete execution can be easily
-refactored to an effective 0-CFA control-flow analysis
-\cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672} by first
-tweaking the environment dereference as a nondeterministic choice, then
-allocating continuations in the environment, and constraining the addresses
-space to be finite. This systematic abstraction approach can be tailored to
-different analyses \todo{cite} and and has been applied to multiple variants of
-small-step abstract machines \cite{DBLP:journals/jfp/HornM12,
-DBLP:conf/icfp/HornM10, Sergey:2013:MAI:2491956.2491979} and big-step
-definitional interpreters \cite{Wei:2018:RAA:3243631.3236800,
+In recent years, we observe a rich progress on how we can construct
+abstract interpreters from systematic principles, instead of ad-hoc
+engineerings. A notable one is the Abstracting Abstract Machine (AAM)
+methodology \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10},
+which uncovers an approach to derive sound abstract interpreters from
+their concrete counterparts, where the soundness can be easily
+established by examining the transformation of semantic artifacts. For
+example, a CEK machine \cite{DBLP:conf/popl/FelleisenF87} for concrete
+execution can be easily refactored to an effective 0-CFA control-flow
+analysis \cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672}
+by first tweaking the environment dereference as a nondeterministic choice,
+then allocating continuations in the environment, and constraining the
+addresses space to be finite. This systematic abstraction approach can
+be tailored to different analyses \todo{cite}
+\cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16} and and
+has been applied to multiple variants of small-step abstract machines
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
+  Sergey:2013:MAI:2491956.2491979} and big-step definitional
+interpreters \cite{Wei:2018:RAA:3243631.3236800,
   DBLP:journals/pacmpl/DaraisLNH17, Keidel:2018:CSP:3243631.3236767}.
 
-More pragmatically, as an implementation strategy, several purely functional
-programming approaches to build abstract interpreters were emerged, for
-examples, using monads or arrows. The pure approach provides much benefits: 1)
-the abstract interpretation artifacts are compositional and can be constructed
-modularly, 2) one can easily prove the soundness of analysis, and 3) one can
-easily reason the correctness of implementation by leveraging equational
-reasoning. For example, monadic abstract interpreters \todo{two level semantics,
-arrow, Galois transformers}
+More pragmatically, as an implementation strategy, several purely
+functional programming approaches to build abstract interpreters were
+emerged, for examples, using monads or arrows. The pure approach
+provides much benefits: 1) the abstract interpretation artifacts are
+compositional and can be constructed modularly, e.g., by using
+monad transformers. 2) The soundness of analysis can be proved more easily, by
+mechnization \cite{Darais:2016:CGC:2951913.2951934} or paper-based
+proof \cite{Keidel:2018:CSP:3243631.3236767}.  3) The correctness of
+implementations can be reasoned by the programmers more condidently,
+e.g., by leveraging equational reasoning.
 
@@
 \citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract interpreters use
-monad transformers; the very recent \citet{Githubsemantic}'s Semantic framework
-uses extensible effects. The effect systems may introduce additional performance
-overhead.
+monad transformers; the very recent \citet{Githubsemantic}'s \textsc{Semantic} framework
+uses extensible effects. With certain merits and elegance of the implementation,
+the effect systems may also introduce additional performance overhead.
 
-In this paper, we propose an abstraction-without-regret approach to eliminate
-those performance penalties for abstract interpreters written in purely
-functional programming style. In short, our approach borrows ideas from program
-specialization and high-performance DSLs: 1) Using multi-stage programming at
-type level (the Lightweight Modular Staging framework
-\cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
-specialize the abstract interpreter with respect to the input program and then
-generate efficient low-level code that does the actual analysis. 2) We use the
-tagless-final approach to embed the target language interpreter in a high-level
-host language (Scala, in this paper), allowing the user to implement different
-semantics modularly, including the abstract semantics, without changing the main
-generic interpreter. This design choice allows user to derive staged abstract
-interpreters without intrusive changes to the unstaged one. Therefore, our
-approach is no regret in the sense of both performance and least engineering
-efforts to achieve such performance. We elaborate these two main ideas in detail
-as follow.
+In this paper, we propose an abstraction-without-regret approach to
+eliminate those performance penalties for abstract interpreters
+written in purely functional programming style. In short, our approach
+borrows ideas from program specialization and high-performance DSLs,
+and applies them to a particular kind of meta-programs, abstract
+interpreters.
+1) Using multi-stage programming (the Lightweight Modular Staging
+framework \cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
+specialize the abstract interpreter with respect to an input program
+and then generate efficient low-level code that does the actual
+analysis. The specialization can also eliminate the effect layers in the generated code.
+2) We use a tagless-final approach to embed the target
+language interpreter in a high-level host language (Scala), allowing
+the user to implement different semantics modularly, including the
+abstract semantics and staging, without changing the main generic
+interpreter. Together with the multi-stage programming and type-level
+annotations, this design allows user to derive staged abstract
+interpreters without intrusive changes to the unstaged one. Therefore,
+our approach is no regret in the sense of both performance and least
+engineering efforts to achieve such performance. We elaborate these
+two main ideas in detail as follow.
 
 \paragraph{Futumura Projection of Abstract Interpreters}
+
+The idea of
 
@@
 
+\paragraph{Applications}
+
+\paragraph{Contribution}
+
+\paragraph{Oragnization}
+
 \iffalse

### 2019-06-30 paper/10_intro.tex
misc

@@
 annotations, this design allows user to derive staged abstract
-interpreters without intrusive changes to the unstaged one. Therefore,
-our approach is no regret in the sense of both performance and least
+interpreters without intrusive changes to the unstaged one, thus the soundness is preserved. 
+Therefore, our approach is no regret in the sense of both performance and least
 engineering efforts to achieve such performance. We elaborate these

### 2019-07-01 paper/10_intro.tex
revise intro

@@
 
-Statically analyzing semantic properties of a program is a widely
-known undecidable class of problem. Abstract interpretation as a
-lattice-based approach to construct sound static analysis was proposed
-by \citet{DBLP:conf/popl/CousotC77}. With the Galois connection, one
+Statically analyzing semantic properties of a program is a
+widely-known undecidable problem. Abstract interpretation as a
+lattice-based approach to sound static analyses was proposed
+by \citet{DBLP:conf/popl/CousotC77}. With the Galois connections, one
 can approximate to the program runtime behaviors by computing the
@@
 the side of pragmatics, constructing artifacts and analyzers that
-perform sound abstract interpretation was considered abstruse and
-complicated in practice for a long time.
+perform sound abstract interpretation for modern, high-level and
+expressive languages was considered abstruse and complicated for a
+long time.
 
-In recent years, we observe a rich progress on how we can construct
-abstract interpreters from systematic principles, instead of ad-hoc
-engineerings. A notable one is the Abstracting Abstract Machine (AAM)
-methodology \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10},
-which uncovers an approach to derive sound abstract interpreters from
-their concrete counterparts, where the soundness can be easily
-established by examining the transformation of semantic artifacts. For
-example, a CEK machine \cite{DBLP:conf/popl/FelleisenF87} for concrete
-execution can be easily refactored to an effective 0-CFA control-flow
-analysis \cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672}
-by first tweaking the environment dereference as a nondeterministic choice,
-then allocating continuations in the environment, and constraining the
-addresses space to be finite. This systematic abstraction approach can
-be tailored to different analyses \todo{cite}
-\cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16} and and
-has been applied to multiple variants of small-step abstract machines
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
-  Sergey:2013:MAI:2491956.2491979} and big-step definitional
-interpreters \cite{Wei:2018:RAA:3243631.3236800,
+In recent years, we observe a rich progress on the methodologies for
+constructing abstract interpreters from systematic principles, instead
+of ad-hoc engineerings. A notable one is the Abstracting Abstract
+Machine (AAM) methodology \cite{DBLP:journals/jfp/HornM12,
+  DBLP:conf/icfp/HornM10}, which uncovers an approach to derive sound
+abstract interpreters from their concrete counterparts, i.e., abstract
+machines, where the soundness can be easily established by examining
+the transformation of semantic artifacts. For example, a CEK machine
+\cite{DBLP:conf/popl/FelleisenF87} for concrete execution can be
+readily refactored to an effective $0$-CFA control-flow analysis
+\cite{Shivers:1988:CFA:53990.54007,
+  Midtgaard:2012:CAF:2187671.2187672} by first tweaking the
+environment dereference as a nondeterministic choice, then allocating
+continuations in the environment, and constraining the addresses space
+to be finite. This systematic abstraction approach can be tailored to
+various language features (such as states, first-class controls,
+exceptions and concurrency) and different sensitivity analyses
+\cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16,
+  Darais:2015:GTM:2814270.2814308}, and has been applied to multiple
+small-step abstract machines \cite{DBLP:journals/jfp/HornM12,
+  DBLP:conf/icfp/HornM10, Sergey:2013:MAI:2491956.2491979} and
+big-step definitional interpreters \cite{Wei:2018:RAA:3243631.3236800,
   DBLP:journals/pacmpl/DaraisLNH17, Keidel:2018:CSP:3243631.3236767}.
 
-More pragmatically, as an implementation strategy, several purely
-functional programming approaches to build abstract interpreters were
-emerged, for examples, using monads or arrows. The pure approach
-provides much benefits: 1) the abstract interpretation artifacts are
-compositional and can be constructed modularly, e.g., by using
-monad transformers. 2) The soundness of analysis can be proved more easily, by
-mechnization \cite{Darais:2016:CGC:2951913.2951934} or paper-based
-proof \cite{Keidel:2018:CSP:3243631.3236767}.  3) The correctness of
+Based on the idea of AAM, and more pragmatically, as an implementation
+strategy, several purely functional programming approaches to build
+abstract interpreters were emerged, for examples, using monads or
+arrows. The pure approach provides much benefits: 1) the abstract
+interpretation artifacts are compositional and can be constructed
+modularly, e.g., by using monad transformers. 2) The soundness of
+analysis can be proved more easily, by mechnization
+\cite{Darais:2016:CGC:2951913.2951934} or paper-based proof
+\cite{Keidel:2018:CSP:3243631.3236767}.  3) The correctness of
 implementations can be reasoned by the programmers more condidently,
@@
 In this paper, we propose an abstraction-without-regret approach to
-eliminate those performance penalties for abstract interpreters
-written in purely functional programming style. In short, our approach
-borrows ideas from program specialization and high-performance DSLs,
-and applies them to a particular kind of meta-programs, abstract
-interpreters.
-1) Using multi-stage programming (the Lightweight Modular Staging
-framework \cite{DBLP:conf/gpce/RompfO10}, in this paper), we can
-specialize the abstract interpreter with respect to an input program
-and then generate efficient low-level code that does the actual
-analysis. The specialization can also eliminate the effect layers in the generated code.
-2) We use a tagless-final approach to embed the target
-language interpreter in a high-level host language (Scala), allowing
-the user to implement different semantics modularly, including the
-abstract semantics and staging, without changing the main generic
-interpreter. Together with the multi-stage programming and type-level
+eliminate those performance penalties for abstract interpreters, while
+still keeping these benefits from purely functional programming.  In
+short, our approach borrows ideas from program specialization and
+embedded DSLs, and applies them to this particular kind of
+meta-programs, abstract interpreters.  1) Using multi-stage
+programming , we can specialize the
+abstract interpreter with respect to an input program and then
+generate efficient low-level code that does the actual analysis. The
+result of specialization is reusable, and the effect layers have been
+eleiminated in the generated code. 2) We use a tagless-final approach
+to embed the target language interpreter in a high-level host language,
+which defines a generic interpreter and allows user to
+implement different semantics modularly, including the abstract
+semantics and staging, without changing the generic one.
+Together with the multi-stage programming and type-level
 annotations, this design allows user to derive staged abstract
@@
 
-\paragraph{Futumura Projection of Abstract Interpreters}
+\paragraph{Futamura Projection of Abstract Interpreters}
 
-The idea of
+The idea of specializing interpreters can be traced to Futamura
+projections \cite{Futamura1999, futamura1971partial}.
+The first Futamura projection specifically shows that
+specializing an interpreter w.r.t. the input program yields an
+equivalent executable.  For instance, we have an interpreter @eval@ of
+some language:
+\begin{lstlisting}
+  def eval(e: Expr)(arg: Input): Value
+\end{lstlisting}
+Given a program $e_0$ of type @Expr@, by applying the first Futamura
+projection, we can obtain a specialized interpreter
+$\texttt{eval}_{\texttt{e0}}$ : @Input@ $\to$ @Value@, which is the so-called
+\textit{equivalent executable}. By definition of the interpreter,
+$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
 
-specialization of interpreters/abstract interpreters
+Partial evaluation \cite{DBLP:books/daglib/0072559} was the first
+proposed approach to realize Futamura projections: it discovers the
+binding-time (static or dynamic) of variables, evaluates the static
+part, and residualizes a program that solely relies on the dynamic
+part. However, it is hard to precisely analyze binding-times given an
+arbitrary program. As an alternative approach to specialization and
+partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97}
+lets the programmer to explicitly control and annotate the
+binding-times in the program, and the MSP system will specialize the
+program. The staging annotations can be either syntactic (e.g.,
+MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging framework
+\cite{DBLP:conf/gpce/RompfO10} in Scala).
 
-\paragraph{Generic Interpreters}
+Our proposed framework adopts the type-based multi-stage programming
+and implements the first Futamura projection of a big-step abstract
+interpreter for a stateful, higher-order core language.  We use
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
+interpreter that uses monad transformers as the unstaged baseline. By
+deriving staged monads that can be used to do code generation, we may
+obtain a staged abstract interpreter that fulfills the specializtion.
+After staging, the generated code is specialized to the input program,
+and monadic operations are all inlined.
 
-specializing interpreters is not new, Futumura projections, four different tasks
-for language designers and engineers, writing definitional interpreters for the
-language (concrete semantics close to deno/op semantics), compilers for the
-language (translate to another language, low-level), writing static analyzer for
-the language (abstract semantics property we want to analyze), and a performant
-static analyzer (ultimate goal of this paper). shared a single generic
-interpreter
+\input{fig_confluence.tex}
 
-\paragraph{Applications}
+\paragraph{Generic Interpreter and Reinterpretation}
 
-\paragraph{Contribution}
+The central basis of our approach is a semantic-agnostic interpreter, which can be
+instantiated by different interpretations or compilations (Figure~\ref{confluence}).
+The different semantics in Figure~\ref{confluence} correspond to four different
+artifacts involved in software language engineering: definitional interpreters
+which are defined by the denotational or operational semantics, compilers that
+translate the program into another language, analyzers that statically compute the
+runtime properties of programs, and optimizing analyzers that achieve the same goal
+of analyzers but in the way of compilers. We observe that these four
+tasks may look like very differently at first glance, but in fact are
+all firmly rooted in the concrete semantics of the language.
 
-\paragraph{Oragnization}
+To achieve the flexibility of reinterpreting the shared interpreter by
+different semantics, our framework abstracts the primitive operations
+from these semantics, and extensively uses abstract type members over
+the binding-times, the components of the interpreter, and the returned
+monadic type of the interpreter. The binding-time abstraction is
+represented by a higher-kinded type, and controls whether the
+interpreter produces values or generates code. The components of the
+interpreter, such as the environment and values, determines the domain
+that the interpreter will be running with. Following the AAM approach,
+by applying a series of abstractions to these components, we can
+derive computable and sound abstract interpreters. Finally, the
+monadic abstraction encapsulates both the binding-times and
+components, which appears as the returned type and defines the
+behaviors and datatype the interpreter operatres on.
+
+\paragraph{Applications and Evaluations}
+We evaluate the idea of staging an abstract interpreter through
+multiple case studies and emipical evaluation on performance.
+1) We compare our approach with abstract compilation
+\cite{Boucher:1996:ACN:647473.727587}, and show that we can achieve
+the same optimization by utilizing type-based stage annotations;
+meanwhile, the analyzer program does not need any changes, thereby
+requres little engineering efforts.
+2) We extend the basic staged abstract interpreter to diverse flow
+analysis, including a store-widened analysis, a context-sensitive
+analysis and abstract garbage collection.
+3) We notice that staging an abstract interpreter enables modularly
+compiling an analysis to programs. The compiled analysis is reusable,
+therefore it provides a mechnized way to create library summaries by
+reusing a whole-program analyzer.
+4) We also emipically evaluate the performance improved by staging,
+showing an average 11x speed-up on flow-analysis.
+
+\paragraph{Contributions} Briefly, the contribution of the paper is as follows:
+\begin{itemize}[leftmargin=2em]
+  \item To our best knowledge, we present the first one to use
+    multi-stage programming to specialize a general abstract interpreter,
+    and our approach does not touch the soundness of analyses.
+  \item Intellectually, our framework naturally extends the first
+    Futamura projection to abstract interpreters, showing a
+    well-grounded approach to optimize static analyses via
+    metaprogramming.
+  \item Pragmatically, we show that staging abstract interpreters is
+    useful to improve performance and scalability of analyses by case
+    studies and emiprical evaluation.
+\end{itemize}
+
+\paragraph{Organization} The paper is organized as follows:
+\begin{itemize}[leftmargin=2em]
+  \item We begin by introducing our core target language and reviewing
+    monads in Scala, and then presenting the generic interpreter (Section~\ref{prelim}).
+    After which, we review its instantiations of concrete interpretation
+    (Section~\ref{unstaged_conc}) and staged concrete interpretation
+    (Section~\ref{stagedinterp}).
+  \item We present the unstaged abstract interpreter under the same framework by
+    replacing the environment, store and values (Section~\ref{unstaged_abs}).
+    After, we show the combination of approximation and specialization, dubbed
+    \textit{staged abstract interpreters}, can be readily derived (Section~\ref{sai}).
+  \item We conduct three case studies (Section~\ref{cases_study}), which demonstrate that
+    our approach requires less engineering efforts, is applicable to various analysis,
+    and enables modular analysis.
+  \item We empirically evaluate the performance improvement by staging on
+    control-flow analysis (Section~\ref{evaluation}). We compare the both
+    context-insensitive and store-widening analysis.
+\end{itemize}
 

### 2019-07-01 paper/fig_confluence.tex
revise intro

\vspace{-1em}
\begin{figure}[h]
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=2.5em,column sep=2.5em,minimum width=2em]
  {
    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_abs}) \end{smallmatrix} &
    \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{sai}) \end{smallmatrix} \\
    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_conc}) \end{smallmatrix} &
    \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{stagedinterp}) \end{smallmatrix} \\ };
  \path[-stealth]
    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
  \end{tikzpicture}
  \vspace{-0.5em}
  \caption{The confluence of specialization and approximation.}
  \label{confluence}
\end{figure}
\vspace{-1em}

### 2019-07-01 paper/paper.tex
fix a code dup bug
reorganize paper structure
sec 2
paper
revise intro

@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[sigplan]{acmart}\settopmatter{}
+%\documentclass[acmsmall]{acmart}\settopmatter{}
 
@@
 \usepackage{amssymb}
+%\usepackage{MnSymbol}
 \usepackage{xcolor,colortbl}
@@
 %% Title information
-\title{Staged Abstract Interpreters}         %% [Short Title] is optional;
+%\title{Staged Abstract Interpreters}         %% [Short Title] is optional;
+\title{Staging Abstract Interpreters}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
+\iffalse
 \author
@@
 }
+\fi
 
@@
 %% Author with single affiliation.
-% \author{First1 Last1}
-% \authornote{with author1 note}          %% \authornote is optional;
-%                                         %% can be repeated if necessary
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-% \affiliation{
-%   \position{Position1}
-%   \department{Department1}              %% \department is recommended
-%   \institution{Institution1}            %% \institution is required
-%   \streetaddress{Street1 Address1}
-%   \city{City1}
-%   \state{State1}
-%   \postcode{Post-Code1}
-%   \country{Country1}                    %% \country is recommended
-% }
-% \email{first1.last1@inst1.edu}          %% \email is recommended
+\author{First1 Last1}
+\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position1}
+  \department{Department1}              %% \department is recommended
+  \institution{Institution1}            %% \institution is required
+  \streetaddress{Street1 Address1}
+  \city{City1}
+  \state{State1}
+  \postcode{Post-Code1}
+  \country{Country1}                    %% \country is recommended
+}
+\email{first1.last1@inst1.edu}          %% \email is recommended
 
-% %% Author with two affiliations and emails.
-% \author{First2 Last2}
-% \authornote{with author2 note}          %% \authornote is optional;
-%                                         %% can be repeated if necessary
-% \orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-% \affiliation{
-%   \position{Position2a}
-%   \department{Department2a}             %% \department is recommended
-%   \institution{Institution2a}           %% \institution is required
-%   \streetaddress{Street2a Address2a}
-%   \city{City2a}
-%   \state{State2a}
-%   \postcode{Post-Code2a}
-%   \country{Country2a}                   %% \country is recommended
-% }
-% \email{first2.last2@inst2a.com}         %% \email is recommended
-% \affiliation{
-%   \position{Position2b}
-%   \department{Department2b}             %% \department is recommended
-%   \institution{Institution2b}           %% \institution is required
-%   \streetaddress{Street3b Address2b}
-%   \city{City2b}
-%   \state{State2b}
-%   \postcode{Post-Code2b}
-%   \country{Country2b}                   %% \country is recommended
-% }
-% \email{first2.last2@inst2b.org}         %% \email is recommended
+%% Author with two affiliations and emails.
+\author{First2 Last2}
+\authornote{with author2 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  \position{Position2a}
+  \department{Department2a}             %% \department is recommended
+  \institution{Institution2a}           %% \institution is required
+  \streetaddress{Street2a Address2a}
+  \city{City2a}
+  \state{State2a}
+  \postcode{Post-Code2a}
+  \country{Country2a}                   %% \country is recommended
+}
+\email{first2.last2@inst2a.com}         %% \email is recommended
+\affiliation{
+  \position{Position2b}
+  \department{Department2b}             %% \department is recommended
+  \institution{Institution2b}           %% \institution is required
+  \streetaddress{Street3b Address2b}
+  \city{City2b}
+  \state{State2b}
+  \postcode{Post-Code2b}
+  \country{Country2b}                   %% \country is recommended
+}
+\email{first2.last2@inst2b.org}         %% \email is recommended
 
@@
 \input{10_intro}
-\input{20_bg}
-\input{30_sai}
-\input{40_comp}
-\input{50_cases}
-\input{60_eval}
-\input{70_related}
-\input{80_concl}
+\input{20_prelim}
+\input{30_stageconc}
+\input{40_absinterp}
+\input{50_stageabs}
+\input{60_cases}
+\input{70_eval}
+\input{80_related}
+\input{90_conc}
 
 %% Acknowledgments
+% Ack Fei, Greg, James, Qianchuan, Guanhong
 \begin{acks}                            %% acks environment is optional

### 2019-07-01 paper/10_intro.tex
misc
typo

@@
 constructing abstract interpreters from systematic principles, instead
-of ad-hoc engineerings. A notable one is the Abstracting Abstract
+of ad-hoc engineering. A notable one is the Abstracting Abstract
 Machine (AAM) methodology \cite{DBLP:journals/jfp/HornM12,
@@
 modularly, e.g., by using monad transformers. 2) The soundness of
-analysis can be proved more easily, by mechnization
+analysis can be proved more easily, by either mechanization
 \cite{Darais:2016:CGC:2951913.2951934} or paper-based proof
 \cite{Keidel:2018:CSP:3243631.3236767}.  3) The correctness of
-implementations can be reasoned by the programmers more condidently,
+implementations can be reasoned by the programmers more confidently,
 e.g., by leveraging equational reasoning.
@@
 program by traversing the abstract syntax tree, which poses the interpretive
-overhead, e.g., pattern matching on the AST and recursive calls on the sub
+overhead, e.g., pattern matching on the ASTs and recursive calls on the sub
 expressions. Those kind of overhead can be negligible if the abstract
@@
 abstract interpreter written in pure languages usually extensively uses effect
-systems to model the behaviors of abstract interpretatation, such as
+systems to model the behaviors of abstract interpretation, such as
 nondeterminism. For example, \citet{DBLP:journals/pacmpl/DaraisLNH17} and
@@
 result of specialization is reusable, and the effect layers have been
-eleiminated in the generated code. 2) We use a tagless-final approach
+eliminated in the generated code. 2) We use a tagless-final approach
 to embed the target language interpreter in a high-level host language,
@@
 binding-time (static or dynamic) of variables, evaluates the static
-part, and residualizes a program that solely relies on the dynamic
+part, and generates a residualized program that solely relies on the dynamic
 part. However, it is hard to precisely analyze binding-times given an
@@
 deriving staged monads that can be used to do code generation, we may
-obtain a staged abstract interpreter that fulfills the specializtion.
+obtain a staged abstract interpreter that fulfills the specialization.
 After staging, the generated code is specialized to the input program,
-and monadic operations are all inlined.
+and all monadic operations are all inlined.
 
@@
 components, which appears as the returned type and defines the
-behaviors and datatype the interpreter operatres on.
+behaviors and datatype the interpreter operates on.
 
@@
 We evaluate the idea of staging an abstract interpreter through
-multiple case studies and emipical evaluation on performance.
+multiple case studies and empirical evaluation on performance.
 1) We compare our approach with abstract compilation
@@
 meanwhile, the analyzer program does not need any changes, thereby
-requres little engineering efforts.
+requires little engineering efforts.
 2) We extend the basic staged abstract interpreter to diverse flow
@@
 compiling an analysis to programs. The compiled analysis is reusable,
-therefore it provides a mechnized way to create library summaries by
+therefore it provides a mechanized way to create library summaries by
 reusing a whole-program analyzer.
-4) We also emipically evaluate the performance improved by staging,
+4) We also empirically evaluate the performance improved by staging,
 showing an average 11x speed-up on flow-analysis.
@@
     useful to improve performance and scalability of analyses by case
-    studies and emiprical evaluation.
+    studies and empirical evaluation.
 \end{itemize}

### 2019-07-07 paper/10_intro.tex
paper

@@
 nondeterminism. For example, \citet{DBLP:journals/pacmpl/DaraisLNH17} and
-\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract interpreters use
-monad transformers; the very recent \citet{Githubsemantic}'s \textsc{Semantic} framework
+\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract interpreters;
+the very recent \citet{Githubsemantic}'s \textsc{Semantic} framework
 uses extensible effects. With certain merits and elegance of the implementation,
@@
 After staging, the generated code is specialized to the input program,
-and all monadic operations are all inlined.
+and all monadic operations are all inlined and compiled down to low-level
+Scala code.
 

### 2019-07-07 paper/70_eval.tex
paper

@@
     @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 12.91x     & 10.005& 1.680 &  5.95x        \\
-    @kcfa-worst-64@     & 774  & 2.18       & 0.129      & 16.90x     & 22.999& 3.872 &  5.94x     \\
+    @kcfa-worst-64@     & 710  & 2.18       & 0.129      & 16.90x     & 22.999& 3.872 &  5.94x        \\
     @solovay-strassen@  & 523  & 5.094      & 1.189      & 4.28x      & 19.118& 0.654 &  29.25x       \\
     @regex@             & 550  & -          & -          & -          & 7.133 & 0.624 &  11.43x       \\
-    @matrix@            & 1891 & -          & -          & -          & 86.176& 4.592 &  18.76x      \\
+    @matrix@            & 1732 & -          & -          & -          & 86.176& 4.592 &  18.76x       \\
     \bottomrule
@@
 As we can see from the table, the staged version significantly outperforms the
-unstaged version on all benchmarks. %, especially for the @kcfa-worst-@$n$ series.
+unstaged version on all benchmarks.
 We observe the most improvement of 29x times faster than the unstaged version on
@@
 version in a principled way that runs significantly faster.
-
-\iffalse
-As we can see from the table, the staged version significantly outperforms the
-unstaged especially for the difficult ones. However, for some benchmarks, we
-observe degraded performance. We conjecture this happens due to the fact that
-the large size of the generated code is unoptimized -- as what we test here is a
-simple implementation almost identical to what we described in the previous
-sections, we believe the performance still can be improved if more engineering
-effort is taken, such as optimizing code generation and avoiding code
-duplication.
-\fi

### 2019-07-08 paper/70_eval.tex
update numbers

@@
 and evaluate the analyzing time against with the unstaged versions. The result
-shows that the staged versions are averagely 11x times as fast as the unstaged
+shows that the staged versions are averagely 10+ times as fast as the unstaged
 version, as strong evidence of our \textit{abstraction without regret}
@@
 0-CFA-like analysis, one is equipped with store-widening, the other does not.
-We also implement several optimizations mentioned in Section \ref{staged_ds} 
-(less than ~30 LOC in total).
+We also implement several optimizations mentioned in Section \ref{staged_ds}.
 
-We use Scala 2.12.7 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
+We use Scala 2.12.8 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
 (kernel 4.4.0) machine. All of our evaluations were performed on a machine with 4 Intel
@@
 
-\iffalse
-We used the following benchmark programs:
-\begin{itemize}
-  %\item \textbf{fib}: Calculates the $n$-th fibonacci number.
-  \item \textbf{rsa}: The RSA public key encryption algorithm.
-  \item \textbf{kcfa3}: A difficult benchmark for $k$-CFA.
-  \item \textbf{church}: Church numerals with additions and mutiplications.
-  \item \textbf{fermat}: Fermat and Solovay-Strassen primality testing.
-  \item \textbf{kcfa-worst-case-}\textit{n}: 
-    Benchmark programs that are supposed to be tough cases for $k$-CFA; 
-    the number $n$ indicates the depth of nesting lambda terms.
-\end{itemize}
-\fi
-
 \paragraph{Result}
@@
     \midrule
-                        &      & \multicolumn{3}{c}{w/o store-widening (sec.)}  &  \multicolumn{3}{c}{w/ store-widening (ms)}\\
+                        &      & \multicolumn{3}{c}{w/o store-widening}  &  \multicolumn{3}{c}{w/ store-widening}\\
     \midrule
-    @fib@               & 32   & 0.013      & 0.002      & 6.50x      & 1.537 & 0.067  &  22.87x      \\
-    @rsa@               & 451  & 0.232      & 0.042      & 5.52x      & 15.113& 1.449 &  10.43x       \\
-    @church@            & 120  & 67.342     & 18.917     & 3.56x      & 2.962 & 0.291 &  10.19x       \\
-    @fermat@            & 310  & 0.024      & 0.004      & 6.00x      & 7.624 & 0.622 &  12.26x       \\
-    @mbrotZ@            & 331  & 0.763      & 0.127      & 6.01x      & 12.229& 0.729 &  16.77x       \\
-    @lattice@           & 609  & 31.874     & 8.816      & 3.62x      & 17.943& 0.909 &  19.75x       \\
-    @kcfa-worst-16@     & 182  & 0.045      & 0.006      & 7.50x      & 4.658 & 0.785 &  5.93x        \\
-    @kcfa-worst-32@     & 358  & 0.284      & 0.022      & 12.91x     & 10.005& 1.680 &  5.95x        \\
-    @kcfa-worst-64@     & 710  & 2.18       & 0.129      & 16.90x     & 22.999& 3.872 &  5.94x        \\
-    @solovay-strassen@  & 523  & 5.094      & 1.189      & 4.28x      & 19.118& 0.654 &  29.25x       \\
-    @regex@             & 550  & -          & -          & -          & 7.133 & 0.624 &  11.43x       \\
-    @matrix@            & 1732 & -          & -          & -          & 86.176& 4.592 &  18.76x       \\
+    @fib@               & 32   & 3.288 ms   & 0.154 ms   & 21.33x      & 1.434 ms   & 0.098 ms  &  14.62x       \\
+    @rsa@               & 451  & 238.171 ms & 23.333 ms  & 10.20x      & 11.977 ms  & 1.197 ms  &  10.00x       \\
+    @church@            & 120  & 61.001 s   & 4.277 s    & 14.26x      & 2.338 ms   & 0.534 ms  &  4.37x        \\
+    @fermat@            & 310  & 23.540 ms  & 2.885 ms   & 8.05x       & 7.146 ms   & 0.915 ms  &  7.81x        \\
+    @mbrotZ@            & 331  & 665.456 ms & 66.070 ms  & 10.07x      & 11.008 ms  & 1.476 ms  &  7.45x        \\
+    @lattice@           & 609  & 29.230 s   & 2.627 s    & 11.12x      & 16.432 ms  & 2.427 ms  &  6.76x        \\
+    @kcfa-worst-16@     & 182  & 44.431 ms  & 3.211 ms   & 13.83x      & 4.425 ms   & 0.850 ms  &  5.20x        \\
+    @kcfa-worst-32@     & 358  & 284.268 ms & 9.065 ms   & 31.35x      & 10.109 ms  & 1.661 ms  &  6.08x        \\
+    @kcfa-worst-64@     & 710  & 2.057 s    & 0.029 s    & 70.23x      & 23.269 ms  & 3.312 ms  &  7.02x        \\
+    @solovay-strassen@  & 523  & 5.078 s    & 0.766 s    & 6.62x       & 18.757 ms  & 3.142 ms  &  5.96x       \\
+    @regex@             & 550  & -          & -          & -           & 6.803 ms   & 1.088 ms  &  6.24x       \\
+    @matrix@            & 1732 & -          & -          & -           & 85.611 ms  & 9.297 ms  &  9.20x       \\
     \bottomrule
@@
 unstaged version on all benchmarks.
-We observe the most improvement of 29x times faster than the unstaged version on
-@solovay-strassen@. On average, the staging produces code that runs 11x times faster
+On average, the staging produces code that runs 17x times faster
 than the unstaged analyzer.

### 2019-07-08 paper/abstract.tex
update numbers

@@
 overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of 11x times with staging for control-flow analysis.
+experiment shows an order of magnitude speedup with staging for control-flow analysis.
 

### 2019-07-08 paper/paper.tex
author info
update numbers

@@
 %% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% Author with single affiliation.
-\author{First1 Last1}
-\authornote{with author1 note}          %% \authornote is optional;
+\author{Guannan Wei}
+%\authornote{with author1 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
 \affiliation{
-  \position{Position1}
-  \department{Department1}              %% \department is recommended
-  \institution{Institution1}            %% \institution is required
-  \streetaddress{Street1 Address1}
-  \city{City1}
-  \state{State1}
-  \postcode{Post-Code1}
-  \country{Country1}                    %% \country is recommended
+  %\position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Purdue University}            %% \institution is required
+  %\streetaddress{305 N. University Street}
+  %\city{West Lafayette}
+  %\state{IN}
+  %\postcode{47906}
+  \country{USA}                    %% \country is recommended
 }
-\email{first1.last1@inst1.edu}          %% \email is recommended
+\email{guannanwei@purdue.edu}          %% \email is recommended
 
 %% Author with two affiliations and emails.
-\author{First2 Last2}
-\authornote{with author2 note}          %% \authornote is optional;
+\author{Yuxuan Chen}
+%\authornote{with author2 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
 \affiliation{
-  \position{Position2a}
-  \department{Department2a}             %% \department is recommended
-  \institution{Institution2a}           %% \institution is required
-  \streetaddress{Street2a Address2a}
-  \city{City2a}
-  \state{State2a}
-  \postcode{Post-Code2a}
-  \country{Country2a}                   %% \country is recommended
+  %\position{Position2a}
+  \department{Department of Computer Science}             %% \department is recommended
+  \institution{Purdue University}           %% \institution is required
+  %\streetaddress{305 N. University Street}
+  %\city{West Lafayette}
+  %\state{IN}
+  %\postcode{47906}
+  \country{USA}                    %% \country is recommended
 }
+
 \email{first2.last2@inst2a.com}         %% \email is recommended
 \affiliation{
-  \position{Position2b}
-  \department{Department2b}             %% \department is recommended
-  \institution{Institution2b}           %% \institution is required
-  \streetaddress{Street3b Address2b}
-  \city{City2b}
-  \state{State2b}
-  \postcode{Post-Code2b}
-  \country{Country2b}                   %% \country is recommended
+  %\position{Position2b}
+  %\department{Department2b}             %% \department is recommended
+  \institution{Facebook}           %% \institution is required
+  %\streetaddress{Street3b Address2b}
+  %\city{City2b}
+  %\state{State2b}
+  %\postcode{Post-Code2b}
+  \country{USA}                   %% \country is recommended
 }
 \email{first2.last2@inst2b.org}         %% \email is recommended
+
+\author{Tiark Rompf}
+%\authornote{with author1 note}          %% \authornote is optional;
+                                        %% can be repeated if necessary
+%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
+\affiliation{
+  %\position{Position1}
+  \department{Department of Computer Science}              %% \department is recommended
+  \institution{Purdue University}            %% \institution is required
+  %\streetaddress{305 N. University Street}
+  %\city{West Lafayette}
+  %\state{IN}
+  %\postcode{47906}
+  \country{USA}                    %% \country is recommended
+}
+\email{tiark@purdue.edu}          %% \email is recommended
 

### 2019-08-11 paper/10_intro.tex
update numbers
revise intro

@@
 lattice-based approach to sound static analyses was proposed
-by \citet{DBLP:conf/popl/CousotC77}. With the Galois connections, one
-can approximate to the program runtime behaviors by computing the
-fixed points on the abstract domain. Despite the tremendous
-theoretical development of abstract interpretation over the years, on
-the side of pragmatics, constructing artifacts and analyzers that
-perform sound abstract interpretation for modern, high-level and
-expressive languages was considered abstruse and complicated for a
-long time.
+by \citet{DBLP:conf/popl/CousotC77}. Equipped with Galois connections,
+the analyzer can obtain the program runtime behaviors approximately
+by computing the fixed points on the abstract domain. Despite the tremendous
+theoretical development of abstract interpretation over the years,
+constructing artifacts and analyzers that perform sound abstract
+interpretation for modern and expressive languages was considered
+abstruse and complicated for a long time.
 
@@
 Machine (AAM) methodology \cite{DBLP:journals/jfp/HornM12,
-  DBLP:conf/icfp/HornM10}, which uncovers an approach to derive sound
-abstract interpreters from their concrete counterparts, i.e., abstract
-machines, where the soundness can be easily established by examining
-the transformation of semantic artifacts. For example, a CEK machine
+  DBLP:conf/icfp/HornM10}. It uncovers an approach to derive sound
+abstract interpreters from their concrete counterparts, for example,
+abstract machines, where the soundness can be easily established by examining
+the transformation of semantic artifacts. For example, the CEK machine
 \cite{DBLP:conf/popl/FelleisenF87} for concrete execution can be
 readily refactored to an effective $0$-CFA control-flow analysis
-\cite{Shivers:1988:CFA:53990.54007,
-  Midtgaard:2012:CAF:2187671.2187672} by first tweaking the
-environment dereference as a nondeterministic choice, then allocating
-continuations in the environment, and constraining the addresses space
-to be finite. This systematic abstraction approach can be tailored to
-various language features (such as states, first-class controls,
-exceptions and concurrency) and different sensitivity analyses
+\cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672}:
+first we tweak the environment dereference as a nondeterministic choice,
+such that the environment may contain mutiple possible values for a varaiable,
+then allocating continuations in the environment to exploit return-flow information,
+and also constrain the addresses space to be finite.
+This systematic abstraction approach can be tailored to
+different language features (such as states, first-class controls,
+exceptions and concurrency) and sensitivity analyses
 \cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16,
-  Darais:2015:GTM:2814270.2814308}, and has been applied to multiple
+  Darais:2015:GTM:2814270.2814308}. It also has been applied to various
 small-step abstract machines \cite{DBLP:journals/jfp/HornM12,
@@
 
-Based on the idea of AAM, and more pragmatically, as an implementation
-strategy, several purely functional programming approaches to build
-abstract interpreters were emerged, for examples, using monads or
-arrows. The pure approach provides much benefits: 1) the abstract
-interpretation artifacts are compositional and can be constructed
-modularly, e.g., by using monad transformers. 2) The soundness of
-analysis can be proved more easily, by either mechanization
-\cite{Darais:2016:CGC:2951913.2951934} or paper-based proof
-\cite{Keidel:2018:CSP:3243631.3236767}.  3) The correctness of
-implementations can be reasoned by the programmers more confidently,
-e.g., by leveraging equational reasoning.
+Based on the idea of abstracting abstract machines, more pragmatically,
+several implementation strategies utilizing purely functional programming
+to build abstract interpreters were emerged. Such techniques include monads,
+monad transformers \cite{DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979},
+arrows \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects \cite{Githubsemantic} and etc.
+The pure approaches provide certain benefits. The abstract interpretation
+artifacts can be built compositionally and modularly, e.g., by using monad
+transformers. Therefore, the soundness of analysis can be proved with less efforts,
+either by mechanized \cite{Darais:2016:CGC:2951913.2951934} or paper-based proof
+\cite{Keidel:2018:CSP:3243631.3236767}. Also, referential transparent and
+equational reasoning allow programmers to reason the correctness of their
+implementations more confidently.
 
 However, besides the intrinsic complexity of static analysis, there are
-additional abstraction penalties with these high-level programming approaches.
+additional abstraction penalties with these high-level implementation approaches.
 First, similar to concrete interpreters, the abstract interpreter analyzes the
-program by traversing the abstract syntax tree, which poses the interpretive
-overhead, e.g., pattern matching on the ASTs and recursive calls on the sub
+program by traversing the abstract syntax tree, which poses an interpretive
+overhead, such as the pattern matching on the ASTs and recursive calls on the sub
 expressions. Those kind of overhead can be negligible if the abstract
-interpreter only runs on the program once, but also can be accumulated
-significantly if it runs repeatedly, for example, on libraries. Second, the
-abstract interpreter written in pure languages usually extensively uses effect
-systems to model the behaviors of abstract interpretation, such as
-nondeterminism. For example, \citet{DBLP:journals/pacmpl/DaraisLNH17} and
-\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract interpreters;
-the very recent \citet{Githubsemantic}'s \textsc{Semantic} framework
-uses extensible effects. With certain merits and elegance of the implementation,
-the effect systems may also introduce additional performance overhead.
+interpreter only runs on the program for one time, but also can be accumulated
+significantly if it runs repeatedly, for example, on libraries.
+Second, the abstract interpreter written in pure languages usually extensively
+uses effect systems to implement the semantics of abstract interpretation.
+For example, the abstract interpreter that returns all possible runtime values
+is a form of nondeterminism, where nondeterminism monads can help.
+Although such pure approaches have its own merits and elegance,
+compared with imperative stateful implementations, they are significantly slower.
 
 In this paper, we propose an abstraction-without-regret approach to
-eliminate those performance penalties for abstract interpreters, while
-still keeping these benefits from purely functional programming.  In
-short, our approach borrows ideas from program specialization and
-embedded DSLs, and applies them to this particular kind of
-meta-programs, abstract interpreters.  1) Using multi-stage
-programming , we can specialize the
-abstract interpreter with respect to an input program and then
-generate efficient low-level code that does the actual analysis. The
-result of specialization is reusable, and the effect layers have been
-eliminated in the generated code. 2) We use a tagless-final approach
-to embed the target language interpreter in a high-level host language,
-which defines a generic interpreter and allows user to
-implement different semantics modularly, including the abstract
-semantics and staging, without changing the generic one.
-Together with the multi-stage programming and type-level
-annotations, this design allows user to derive staged abstract
-interpreters without intrusive changes to the unstaged one, thus the soundness is preserved. 
-Therefore, our approach is no regret in the sense of both performance and least
-engineering efforts to achieve such performance. We elaborate these
-two main ideas in detail as follow.
+eliminate those performance penalties for abstract interpreters, meanwhile
+still keeping the benefits come from purely functional programming.
+In short, our approach applies ideas and techniques from program specialization
+and embedded DSLs to abstract interpreters. 1) Using multi-stage
+programming , we can specialize the abstract interpreter with respect to an
+input program and then generate efficient low-level code that does the actual analysis.
+The result of specialization is reusable, and the effect layers have been
+eliminated in the generated code. 2) Inspired by the tagless-final interpreters,
+we define a generic interpreter that abstracts over binding-time and
+different semantics, which allows user to implement different semantics
+modularly, including the staged abstract interpretation semantics.
+Together with the multi-stage programming and type-level annotations of binding-times,
+we can derive the staged abstract interpreters without intrusive changes to
+the corresponding unstaged one, thus the soundness is untouched.
+In this sense, our approach is no regret of both performance and engineering
+effort. We elaborate these two main ideas in detail as follow.
 
@@
 specializing an interpreter w.r.t. the input program yields an
-equivalent executable.  For instance, we have an interpreter @eval@ of
+equivalent executable. For instance, @eval@ is an interpreter for
 some language:
@@
 \end{lstlisting}
-Given a program $e_0$ of type @Expr@, by applying the first Futamura
-projection, we can obtain a specialized interpreter
-$\texttt{eval}_{\texttt{e0}}$ : @Input@ $\to$ @Value@, which is the so-called
-\textit{equivalent executable}. By definition of the interpreter,
-$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
+Given a program $e_0$ of type @Expr@, by applying the specialization,
+we can obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ :
+@Input@ $\to$ @Value@, which is the so-called \textit{equivalent executable}.
+By definition of the interpreter, they produce the same result when applied
+with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
 
 Partial evaluation \cite{DBLP:books/daglib/0072559} was the first
-proposed approach to realize Futamura projections: it discovers the
-binding-time (static or dynamic) of variables, evaluates the static
-part, and generates a residualized program that solely relies on the dynamic
-part. However, it is hard to precisely analyze binding-times given an
-arbitrary program. As an alternative approach to specialization and
-partial evaluation, multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97}
+proposed approach to realize Futamura projections: it first discovers
+the binding-times (\textit{static} or \textit{dynamic}) of variables,
+then evaluates the static part of the program, and generates a residualized
+program that solely relies on the dynamic part. However, it is hard to
+precisely analyze binding-times given an arbitrary program.
+As an alternative approach to specialization, multi-stage programming (MSP)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97}
 lets the programmer to explicitly control and annotate the
-binding-times in the program, and the MSP system will specialize the
-program. The staging annotations can be either syntactic (e.g.,
-MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging framework
-\cite{DBLP:conf/gpce/RompfO10} in Scala).
+binding-times in the program, then the MSP system will check whether these annotations
+are consistent and specialize the program using that information.
+The staging annotations can be either syntactic (e.g., quote and quasiquote
+in MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging
+framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
 
-Our proposed framework adopts the type-based multi-stage programming
-and implements the first Futamura projection of a big-step abstract
-interpreter for a stateful, higher-order core language.  We use
-\citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
+Our proposed framework adopts the type-based multi-stage programming from
+the Lightweight Modular Staging framework and implements the Futamura
+projection of a big-step abstract interpreter for a small stateful,
+higher-order language.
+We use \citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
 interpreter that uses monad transformers as the unstaged baseline. By
@@
 4) We also empirically evaluate the performance improved by staging,
-showing an average 11x speed-up on flow-analysis.
+showing an order of magnitude speed-up on flow-analysis.
 

### 2019-08-11 paper/abstract.tex
revise intro

--- 2019-04-06 paper/abstract.tex
@@
 It is well known that a staged interpreter is a compiler: specializing the
-interpreter to a given program produces an equivalent program that runs faster,
-which is known as the first Futamura projection. It is even more widely known that an
-abstract interpreter is a program analyzer: tweaking the interpreter to run on a
-domain of abstract values produces a sound static analysis. What happens when we
-combine these two ideas, and apply staging to an \emph{abstract} interpreter?
+interpreter to a given program produces an equivalent executable that runs faster.
+This connection is known as the first Futamura projection.
+It is even more widely known that an abstract interpreter is a program analyzer:
+tweaking the interpreter to run on an abstract domain produces a sound static
+analysis. What happens when we combine these two ideas, and apply staging to
+an \emph{abstract} interpreter?
 
 In this paper, we present a unifying framework that naturally extends the first
-Futamura projection of concrete interpreters to abstract interpreters. Our
+Futamura projection from concrete interpreters to abstract interpreters. Our
 approach derives a sound staged abstract interpreter based on a
@@
 staged abstract interpreter.
-
-As an example of abstraction without regret, we show that staging abstract
-interpreters is practical and useful to optimize static analysis while requiring
-least engineering efforts and not compromising any soundness. We conduct three
-case studies, including a comparison with
-\citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
-the application of various control-flow analyses, and the use to modular
-analysis. We also empirically evaluate the performance improved by staging. The
-overhead of the abstraction layers is eliminated in the generated code, and the 
-experiment shows an average speedup of 11x with staging for control-flow analysis.
+As an example of \emph{abstraction without regret}, the overhead of these
+abstraction layers is eliminated in the generated code after staging.
+We show that staging abstract interpreters is practical and useful to
+optimize static analysis while requiring less engineering efforts and not
+compromising soundness. We conduct three case studies, including a comparison
+with \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
+applications on various control-flow analyses, and a demonstration shows that
+it can be used for modular analysis.
+We also empirically evaluate the running time improved by staging.
+The experiment shows an order of magnitude speedup with staging for
+control-flow analyses.
 

### 2019-08-11 paper/10_intro.tex
revise intro

@@
 
-The central basis of our approach is a semantic-agnostic interpreter, which can be
-instantiated by different interpretations or compilations (Figure~\ref{confluence}).
-The different semantics in Figure~\ref{confluence} correspond to four different
-artifacts involved in software language engineering: definitional interpreters
-which are defined by the denotational or operational semantics, compilers that
-translate the program into another language, analyzers that statically compute the
-runtime properties of programs, and optimizing analyzers that achieve the same goal
-of analyzers but in the way of compilers. We observe that these four
-tasks may look like very differently at first glance, but in fact are
-all firmly rooted in the concrete semantics of the language.
+Program specialization and abstract interpretation are two orthogonal concepts.
+To implement the confluence of them, we first construct a generic interpreter that
+is agnostic to both binding-times and value domains used in the semantics.
+Later, the generic interpreter can be instantiated from these two dimensions (Figure~\ref{confluence}):
+\begin{itemize}
+\item With a flat binding-time and concrete domains, it is a ordinary definitional interpreter based
+  on big-step operational semantics;
+\item With two-level binding-times and concrete domain, it is a compiler that translate to program
+  intro antoher language;
+\item With a flat binding-time and abstract domains, it is an definitional abstract interpreter
+  \cite{DBLP:journals/pacmpl/DaraisLNH17} that statically computes runtime properties;
+\item With two-level binding-times and abstract domains, it is an optimizing program analyzer but
+  works in the fashion of compilation.
+\end{itemize}
 
-To achieve the flexibility of reinterpreting the shared interpreter by
-different semantics, our framework abstracts the primitive operations
-from these semantics, and extensively uses abstract type members over
-the binding-times, the components of the interpreter, and the returned
-monadic type of the interpreter. The binding-time abstraction is
-represented by a higher-kinded type, and controls whether the
-interpreter produces values or generates code. The components of the
-interpreter, such as the environment and values, determines the domain
-that the interpreter will be running with. Following the AAM approach,
-by applying a series of abstractions to these components, we can
-derive computable and sound abstract interpreters. Finally, the
-monadic abstraction encapsulates both the binding-times and
-components, which appears as the returned type and defines the
-behaviors and datatype the interpreter operates on.
+Although the four artifacts may look like very differently at first glance,
+but in fact are all firmly rooted in the concrete semantics of the language.
+This observation provides a way to abstract over the interpreter and achieve
+the flexibility of reinterpreting the shared interpreter. We adapt the monadic
+interpreter schema, which already have been applied to abstract interpreters,
+and introduce the binding-time abstraction into the generic interpreter.
+The generic interpreter returns a value of monadic type, which can be varied
+by different semantics. The domain of the interpreter and the effects such as read,
+state and nondeterminism can be both wrapped into this monadic type.
+The binding-time abstraction is represented by a higher-kinded type,
+and controls whether the interpreter produces values directly or generates code.
+It is worth to mention that the binding-time type is also instrumented into
+the monadic type, so that we will distinguish normal monads and staged monads.
 
@@
 We evaluate the idea of staging an abstract interpreter through
-multiple case studies and empirical evaluation on performance.
+case studies and an empirical evaluation on performance.
 1) We compare our approach with abstract compilation
-\cite{Boucher:1996:ACN:647473.727587}, and show that we can achieve
-the same optimization by utilizing type-based stage annotations;
-meanwhile, the analyzer program does not need any changes, thereby
-requires little engineering efforts.
-2) We extend the basic staged abstract interpreter to diverse flow
-analysis, including a store-widened analysis, a context-sensitive
+\cite{Boucher:1996:ACN:647473.727587}, a implementation technique for
+control-flow analyses, and show that by utilizing type-based stage
+annotations we can achieve the same optimization, and meanwhile,
+the analyzer does not need to change, thereby requires much less
+engineering efforts.
+2) We extend the basic staged abstract interpreter to different flow
+analyses, including a store-widened analysis, a context-sensitive
 analysis and abstract garbage collection.
 3) We notice that staging an abstract interpreter enables modularly
-compiling an analysis to programs. The compiled analysis is reusable,
-therefore it provides a mechanized way to create library summaries by
-reusing a whole-program analyzer.
+compiling an analysis to programs. Here we borrow the concept of modular
+analysis, and show that the compiled analysis is reusable.
+Therefore the approach provides a modular way to create optimized analysis
+code by mechanized reusing a whole-program analyzer.
 4) We also empirically evaluate the performance improved by staging,
-showing an order of magnitude speed-up on flow-analysis.
+showing an order of magnitude speed-up on flow-analysis \todo{explicit note the number}.
 

### 2019-08-12 paper/10_intro.tex
revision

@@
     metaprogramming.
-  \item Pragmatically, we show that staging abstract interpreters is
+  \item Practically, we show that staging an abstract interpreter is
     useful to improve performance and scalability of analyses by case
-    studies and empirical evaluation.
+    studies and an empirical evaluation.
 \end{itemize}
@@
 \begin{itemize}[leftmargin=2em]
-  \item We begin by introducing our core target language and reviewing
+  \item We begin by introducing our target language and reviewing
     monads in Scala, and then presenting the generic interpreter (Section~\ref{prelim}).
@@
   \item We present the unstaged abstract interpreter under the same framework by
-    replacing the environment, store and values (Section~\ref{unstaged_abs}).
+    replacing the environment, store and values to their abstract counterparts (Section~\ref{unstaged_abs}).
     After, we show the combination of approximation and specialization, dubbed
@@
     our approach requires less engineering efforts, is applicable to various analysis,
-    and enables modular analysis.
+    and enables compiling analysis modularly.
   \item We empirically evaluate the performance improvement by staging on

### 2019-08-12 paper/20_prelim.tex
intro
revision

@@
 idea, but the approach is not restricted to our choice. One can use
-imperative/direct style and other MSP languages (e.g., MetaOCaml
+imperative or direct style in other MSP languages (e.g., MetaOCaml
 \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} and Template
@@
 interpreters.
-
-\iffalse
-Then we
-will proceed in two directions: by staging it we obtain a staged concrete interpreter
-(Section~\ref{stagedinterp}); by abstracting it we obtain an unstaged abstract
-interpreter (Section~\ref{unstaged_abs}).
-\fi
 
@@
 The abstract syntax we present in fact can be seen as a deep embedding of the
-language -- we use algebraic data-types to represent programs. This design choice allows us to
-easily use different interpretations over the AST; with the inheritance and
-overriding mechanism in Scala, we may also add new language construct and reuse
-the existing interpretation.
+language -- we use data-types to represent programs. This design choice allows us
+to easily use different interpretations over the AST; with the inheritance and
+overriding mechanism in Scala, we may also add new language constructs and reuse
+existing interpretations \todo{cite Bruno?}.
 

### 2019-08-12 paper/50_stageabs.tex
revision

@@
 
-\subsection{Discussion}
+\section{Discussion}
 
 We have gradually presented the confluence of specialization and abstraction of
-concrete interpreters. In this section, we discuss several decisions we made to
-achieve this and examine some alternatives.
+concrete interpreters. In this section, we review and summarize our recipe
+to achieve the staged abstract interpreter, discuss the correctness issue
+and different design choices.
 
-\paragraph{What does the generated code look like?} In the generated code, all the
+\subsection{Summarizing the Approach}
+
+\todo{TODO}
+
+\paragraph{What has been eliminated?} In the generated code, all the
 primitive operations (such as @eval@, @fix@, @ap_clo@, etc.) and monadic
@@
 the use of monads internally.
+
+\todo{may be show an small example?}
+
+\subsection{Correctness}
+
+\todo{TODO}
+
+As one of the merits of our approach, the staging does not compromise any
+soundness, because we only change the underlying monads and data structures to
+the staged version, not any part of the abstract semantics. Based on the
+assumption that MSP system and staged data structure preserve the equivalence
+during staging, we are confident that the staged abstract interpreter does the
+same analysis as the unstaged one.
+
+\subsection{Design Choices}
 
@@
 an additional engineering effort.
-
-\paragraph{Correctness and Soundness}
-
-As one of the merits of our approach, the staging does not compromise any
-soundness, because we only change the underlying monads and data structures to
-the staged version, not any part of the abstract semantics. Based on the
-assumption that MSP system and staged data structure preserve the equivalence
-during staging, we are confident that the staged abstract interpreter does the
-same analysis as the unstaged one.

### 2019-08-12 paper/80_related.tex
update
revision

@@
 \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis.
+modular analysis. We borrow the notation of modular analysis, \todo{but intentional properties
+  are different, compare with summary}
 
-\paragraph{Meta-Programming} In this paper, we use the Lightweight Modular
-Staging framework \cite{DBLP:conf/gpce/RompfO10}, which relies on types as stage
-annotations. Other notable implementations of MSP exist in ML family, e.g.,
+\paragraph{Two-level Semantics}
+
+\paragraph{Partial Evaluation and Multi-stage Programming}
+Partial evaluation as an automatic technique for program specialization
+was studied comprehensively by \citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
+In this paper, we use mutli-stage programming as an approach to
+program specialization. The Lightweight Modular Staging framework \cite{DBLP:conf/gpce/RompfO10}
+we used in the paper relies on type-level stage annotations.
+Other notable implementations of MSP exist in ML family, e.g.,
 MetaML \cite{DBLP:conf/pepm/TahaS97} and MetaOCaml
@@
 The idea of staging an abstract interpreter we presented in this paper is also
-applicable to other MSP systems. Partial evaluation as an automatic technique is
-studied comprehensively \cite{10.1007/3-540-61580-6_11,
-DBLP:books/daglib/0072559}. Futamura projections reveals close relations between
-interpreters and compilers, which was originally proposed by
-\citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial}, and later
-reprinted \cite{Futamura1999}.
-\citeauthor{Amin:2017:CTI:3177123.3158140} shows how to collapse a tower of
-concrete interpreters by MSP -- it would be also interesting to explore multiple
-layers of abstract interpreters.
-\citet{10.1007/11561347_18} discussed combining
-multi-stage programming with functors and monads.
-Similar to the idea in this paper, specializing monadic interpreters was
-explored by \citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}.
-
+applicable to other MSP systems.
 Multi-stage programming has been widely used to improve the performance in many
@@
 \cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
+
+As an source of inspiration of this paper, Futamura projections reveals a
+hierarchy and close relation between interpreters and compilers, which was
+originally proposed by \citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial},
+and later reprinted \cite{Futamura1999}.
+\citeauthor{Amin:2017:CTI:3177123.3158140} consider a tower of concrete
+interpreters and how to collapse them by using MSP -- it would be
+interesting to explore this idea for multiple layers of abstract interpreters
+\cite{Cousot:2019:AAI:3302515.3290355, Giacobazzi:2015:APA:2676726.2676987}.
+\citet{10.1007/11561347_18} discussed combining
+multi-stage programming with functors and monads. But they didn't \todo{monad transfomers?}
+Similar to the idea in this paper, specializing monadic interpreters was
+explored by \citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}. \todo{what's the difference?}
+
+\paragraph{Control-flow Analysis}

### 2019-08-13 paper/70_eval.tex
revision

@@
 We implement the abstract interpreters for control-flow analysis, i.e., a
-value-flow analysis focus on lambda terms in functional languages. A front-end
+value-flow analysis that computes which lambda terms can be called at each
+call-sites in functional languages. With other suitable abstract domains, the
+abstract interpreter also can be extended to other analyses. A front-end
 desugars Scheme to a small core language that makes the analyzer easy to
@@
 0-CFA-like analysis, one is equipped with store-widening, the other does not.
-We also implement several optimizations mentioned in Section \ref{staged_ds}.
+\todo{why store widening?} Our prototype implementation currently generates
+Scala code; the generated Scala code will be compiled and executed also on JVM.
+it is possible to also generate C/C++ code in the future. In the experiments, we
+implement several optimizations mentioned in Section \ref{staged_ds},
+specifically, the selective caching and rewriting rules exploiting
+partially-static data. These optimizations are useful to reduce the size of
+generated code.
 
-We use Scala 2.12.8 and Oracle JVM 1.8.0-191, running on an Ubuntu 16.04 LTS
+We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for running
+  JVM is set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
+  -XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}}, running on an Ubuntu 16.04 LTS
 (kernel 4.4.0) machine. All of our evaluations were performed on a machine with 4 Intel
@@
 cores (192 threads) in total, the abstract interpreters are single-threaded.
-To minimize the effect of warming-up from the JVM JIT compiler, all the
+To minimize the effect of warming-up from the HotSpot JIT compiler, all the
 experiments are executed for 20 times and we report the statistical median value
@@
 version in a principled way that runs significantly faster.
+
+\todo{analyze some interesting examples}

### 2019-08-13 paper/80_related.tex
revision

@@
 semantic-based approach to build sound static analysis by approximation
-\cite{DBLP:conf/popl/CousotC77}. As semantic artifacts, the Abstracting Abstract
-Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} approach
-shows that abstract interpreters can be derived systematically from concrete
+\cite{DBLP:conf/popl/CousotC77}. As to build semantic artifacts, the Abstracting
+Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
+approach shows that abstract interpreters can be derived systematically from concrete
 semantic artifacts. The AAM approach is closely related to control-flow analysis
@@
 \citet{DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
+The unstaged abstract interpreter in this paper follows the abstracting definitional
+interpreters approach from \citet{DBLP:journals/pacmpl/DaraisLNH17}.
 Similar to definitional abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
@@
 
-\paragraph{Two-level Semantics}
+\paragraph{Two-level Semantics} The idea of reinterpreting the semantics as
+abstract interpretation can be traced to Nielson's two-level semantics
+\cite{NIELSON1989117}; using two-level semantics for code generation was also
+explored by \citet{NIELSON198859}. \citet{Sergey:2013:MAI:2491956.2491979}'s
+work of monadic abstract interpreters is also closely related to the two-level
+semantics: the use of a generic interface with monads and then reinterpret it by
+different semantics is already two-level. Instead of focusing on semantics, this
+paper shows how a staged analyzer can be built and used to increase efficieny of
+static analysis, we augment the monadic abstract interpreter from another
+dimension, using code generation to produce efficient low-level code.
 
@@
 
-\paragraph{Control-flow Analysis}
+\paragraph{Control-flow Analysis} useful applications, why important

### 2019-08-13 paper/50_stageabs.tex
revision

@@
 
-Our staging schema works well in theory, but in practice would suffer from code
-explosion, and possibly runtime GC overhead (at the next stage) when analyzing
-(specializing) large programs. In this section, we present several optimizations
-that largely mitigate these issues. Still, implementing these optimizations do
-not need to change the generic interpreter.
+Our staging schema works by unfolding the interpreter over the input AST. In
+practice, it would suffer from code explosion when analyzing (specializing) large
+programs, which increases the compile time.
+If we generate next stage programs running on JVM, such large generated program
+would also cause runtime GC overhead. In this section, we present several
+optimizations that largely mitigate these issues. For all that, implementing
+these optimizations do not need to change the generic interpreter.
 
@@
 
-\todo{TODO}
+\todo{1. construct generic monadic interpreter that abstract over binding-time,
+  primtive operations, return type.
+  2. implement unstaged abstract interpreter with monads, ie. using approriate
+  monad stack
+  3. if we want to generate code, we just need to change the monad stack.
+  the data wrapped are staged.
+  we show readerT monad, stateT monad, fused monad with nondeterminism. }
 
@@
 
-\todo{TODO}
+As the central concern of static analysis, the soundness is important for
+every user who would like to use it. Here we briefly summarize that how our
+approach preserve the corretness and soundness of the analysis, under the
+assumption that the unstaged one is sound. It is worth noting that the rationale
+listed here are based on empirical evidence, a rigorous proof of soundness
+perservation of the staged abstract interpreters is still an open question.
 
-As one of the merits of our approach, the staging does not compromise any
-soundness, because we only change the underlying monads and data structures to
-the staged version, not any part of the abstract semantics. Based on the
-assumption that MSP system and staged data structure preserve the equivalence
-during staging, we are confident that the staged abstract interpreter does the
-same analysis as the unstaged one.
+\begin{itemize}
+  \item We usually assume the underlying MSP system achieves cross-stage
+    persistent, i.e., the generated program does not change the observational
+    behaviours of the original program. In the case of this paper, we use the
+    LMS framework, which 1) relies on the type checking facilities of Scala, and
+    2) generates code in A-Normal form that preserves the
+    execution order within a stage \todo{cite WF16}. However, in general, it is
+    possible to subvert these guarantees by using unsafe features, such as
+    casts.
+  \item As shown in the previous section, the generic interpreter is untouched
+    and shared by the four artifacts. We only instantiate binding-time
+    annotations and the monadic types. This allows the programmer to easily
+    check the correctness of implementation of staged monads modularlly.
+  \item We construct the staged monads and staged data structures in a
+    correct-by-construct way that directly corresponds to their unstaged
+    version. For instance, the staged data structure we used here are simply
+    a black-box that wraps the Scala's library data structures (in the next
+    stage).
+  \item In our experiments, the staged abstract interpreter produces the same
+    result with the unstaged one on all benchmark programs we tested.
+\end{itemize}
 

### 2019-08-13 paper/80_related.tex
revision

@@
 \citet{amtoft1999partial} applied partial evaluation for constraint-based
-control flow analysis. Splitting an analysis into multiple stages is also
-studied for analysis other than for control-flow, though the formulation may
-very different. For example, \citet{DBLP:conf/cgo/HardekopfL11} apply staging to
+control flow analysis.
+
+Splitting an analysis into multiple stages is also studied for analysis other
+than abstract-interpretation-based control-flow analysis, though the formulation may
+very different. \citet{DBLP:conf/cgo/HardekopfL11} apply staging to a
 flow-sensitive pointer analysis. The first stage is to analyze the program code
 to obtain a sparse representation, and then the second stage conducts the
-flow-sensitive analysis based on the first one.
-The abstract domains can be also specialized w.r.t. the static program
-structure, e.g., decomposing polyhedras \cite{DBLP:conf/popl/SinghPV17,
-Singh:2017:PCD:3177123.3158143} to smaller ones depending on the variables
-involved such that the abstract transformers can compute using less time and
-space.
+flow-sensitive analysis by refining the first one. Another large class of static
+analysis are implemented by generating constraints firstly and then let a solver
+discharge these proof obligations or find the fixed-point. Examples include SAT-
+and SMT-based program analysis\todo{cite}, Datalog-based points-to
+analysis\todo{cite}, and \todo{}. Similar to the idea in this paper, such
+analyses can be splitted to staged.
+One of the future work is to explore that whether these analyses can be
+implemented and optimized from metaprogramming. Souffle, one of the
+state-of-the-art Datalog engine that is commonly used in points-to analysis,
+uses the idea of Futamura projection for efficent implemetation. But it only
+specializes Datalog rules to the running engine, instead of a full pipeline from
+program AST to generating constraints and solving constraints.
 
-\paragraph{Abstract Interpreters} Abstract interpretation was proposed as a
+Recent work shows that the numerical abstract domains can be also specialized
+w.r.t. the static program structure. For example, decomposing polyhedras
+\cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to smaller ones
+depending on the variables involved in a statement can significantly improve the
+running time and space of abstract transformers.
+
+\paragraph{Construction of Abstract Interpreters} Abstract interpretation was proposed as a
 semantic-based approach to build sound static analysis by approximation
@@
 interpreter that makes proof of soundness can be constructed compositionally.
+
 \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for

### 2019-08-14 paper/10_intro.tex
revision

@@
 widely-known undecidable problem. Abstract interpretation as a
-lattice-based approach to sound static analyses was proposed
-by \citet{DBLP:conf/popl/CousotC77}. Equipped with Galois connections,
-the analyzer can obtain the program runtime behaviors approximately
-by computing the fixed points on the abstract domain. Despite the tremendous
-theoretical development of abstract interpretation over the years,
-constructing artifacts and analyzers that perform sound abstract
-interpretation for modern and expressive languages was considered
-abstruse and complicated for a long time.
+lattice-based approach to sound static analyses was proposed by
+\citet{DBLP:conf/popl/CousotC77}. Equipped with Galois connections,
+the analyzer can obtain the program runtime behaviors approximately by
+computing the fixed points on the abstract domain. Despite the
+tremendous theoretical development of abstract interpretation over the
+years, constructing artifacts and analyzers that perform sound
+abstract interpretation for modern and expressive languages was
+considered abstruse and complicated for a long time.
 
@@
 abstract interpreters from their concrete counterparts, for example,
-abstract machines, where the soundness can be easily established by examining
-the transformation of semantic artifacts. For example, the CEK machine
-\cite{DBLP:conf/popl/FelleisenF87} for concrete execution can be
-readily refactored to an effective $0$-CFA control-flow analysis
-\cite{Shivers:1988:CFA:53990.54007, Midtgaard:2012:CAF:2187671.2187672}:
-first we tweak the environment dereference as a nondeterministic choice,
-such that the environment may contain mutiple possible values for a varaiable,
-then allocating continuations in the environment to exploit return-flow information,
-and also constrain the addresses space to be finite.
-This systematic abstraction approach can be tailored to
-different language features (such as states, first-class controls,
-exceptions and concurrency) and sensitivity analyses
-\cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16,
-  Darais:2015:GTM:2814270.2814308}. It also has been applied to various
-small-step abstract machines \cite{DBLP:journals/jfp/HornM12,
-  DBLP:conf/icfp/HornM10, Sergey:2013:MAI:2491956.2491979} and
-big-step definitional interpreters \cite{Wei:2018:RAA:3243631.3236800,
+abstract machines, where the soundness can be easily established by
+examining the transformation of semantic artifacts. For example, the
+CEK machine \cite{DBLP:conf/popl/FelleisenF87} for concrete execution
+can be readily refactored to an effective $0$-CFA control-flow
+analysis \cite{Shivers:1988:CFA:53990.54007,
+  Midtgaard:2012:CAF:2187671.2187672}: first we tweak the environment
+dereference as a nondeterministic choice, such that the environment
+may contain mutiple possible values for a varaiable, then allocating
+continuations in the environment to exploit return-flow information,
+and also constrain the addresses space to be finite.  This systematic
+abstraction approach can be tailored to different language features
+(such as states, first-class controls, exceptions and concurrency) and
+sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
+  DBLP:conf/popl/GilrayL0MH16, Darais:2015:GTM:2814270.2814308}. It
+also has been applied to various small-step abstract machines
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
+  Sergey:2013:MAI:2491956.2491979} and big-step definitional
+interpreters \cite{Wei:2018:RAA:3243631.3236800,
   DBLP:journals/pacmpl/DaraisLNH17, Keidel:2018:CSP:3243631.3236767}.
 
-Based on the idea of abstracting abstract machines, more pragmatically,
-several implementation strategies utilizing purely functional programming
-to build abstract interpreters were emerged. Such techniques include monads,
-monad transformers \cite{DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979},
-arrows \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects \cite{Githubsemantic} and etc.
-The pure approaches provide certain benefits. The abstract interpretation
-artifacts can be built compositionally and modularly, e.g., by using monad
-transformers. Therefore, the soundness of analysis can be proved with less efforts,
-either by mechanized \cite{Darais:2016:CGC:2951913.2951934} or paper-based proof
-\cite{Keidel:2018:CSP:3243631.3236767}. Also, referential transparent and
-equational reasoning allow programmers to reason the correctness of their
-implementations more confidently.
+Based on the idea of abstracting abstract machines, more
+pragmatically, several implementation strategies utilizing purely
+functional programming to build abstract interpreters were
+emerged. Such techniques include monads, monad transformers
+\cite{DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979},
+arrows \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
+\cite{Githubsemantic} and etc.  The pure approaches provide certain
+benefits. The abstract interpretation artifacts can be built
+compositionally and modularly, e.g., by using monad
+transformers. Therefore, the soundness of analysis can be proved with
+less efforts, either by mechanized \cite{Darais:2016:CGC:2951913.2951934}
+or paper-based proof \cite{Keidel:2018:CSP:3243631.3236767}. Also,
+referential transparent and equational reasoning allow programmers to
+reason the correctness of their implementations more confidently.
 
-However, besides the intrinsic complexity of static analysis, there are
-additional abstraction penalties with these high-level implementation approaches.
-First, similar to concrete interpreters, the abstract interpreter analyzes the
-program by traversing the abstract syntax tree, which poses an interpretive
-overhead, such as the pattern matching on the ASTs and recursive calls on the sub
+However, besides the intrinsic complexity of static analysis, there
+are additional abstraction penalties with these high-level
+implementation approaches.  First, similar to concrete interpreters,
+the abstract interpreter analyzes the program by traversing the
+abstract syntax tree, which poses an interpretive overhead, such as
+the pattern matching on the ASTs and recursive calls on the sub
 expressions. Those kind of overhead can be negligible if the abstract
-interpreter only runs on the program for one time, but also can be accumulated
-significantly if it runs repeatedly, for example, on libraries.
-Second, the abstract interpreter written in pure languages usually extensively
-uses effect systems to implement the semantics of abstract interpretation.
-For example, the abstract interpreter that returns all possible runtime values
-is a form of nondeterminism, where nondeterminism monads can help.
-Although such pure approaches have its own merits and elegance,
-compared with imperative stateful implementations, they are significantly slower.
+interpreter only runs on the program for one time, but also can be
+accumulated significantly if it runs repeatedly, for example, on
+libraries.  Second, the abstract interpreter written in pure languages
+usually extensively uses effect systems to implement the semantics of
+abstract interpretation.  For example, the abstract interpreter that
+returns all possible runtime values is a form of nondeterminism, where
+nondeterminism monads can help.  Although such pure approaches have
+its own merits and elegance, compared with imperative stateful
+implementations, they are significantly slower.
 
 In this paper, we propose an abstraction-without-regret approach to
-eliminate those performance penalties for abstract interpreters, meanwhile
-still keeping the benefits come from purely functional programming.
-In short, our approach applies ideas and techniques from program specialization
-and embedded DSLs to abstract interpreters. 1) Using multi-stage
-programming , we can specialize the abstract interpreter with respect to an
-input program and then generate efficient low-level code that does the actual analysis.
-The result of specialization is reusable, and the effect layers have been
-eliminated in the generated code. 2) Inspired by the tagless-final interpreters,
+eliminate those performance penalties for abstract interpreters,
+meanwhile still keeping the benefits come from purely functional
+programming.  In short, our approach applies ideas and techniques from
+program specialization and embedded DSLs to abstract interpreters. 1)
+Using multi-stage programming , we can specialize the abstract
+interpreter with respect to an input program and then generate
+efficient low-level code that does the actual analysis.  The result of
+specialization is reusable, and the effect layers have been eliminated
+in the generated code. 2) Inspired by the tagless-final interpreters,
 we define a generic interpreter that abstracts over binding-time and
-different semantics, which allows user to implement different semantics
-modularly, including the staged abstract interpretation semantics.
-Together with the multi-stage programming and type-level annotations of binding-times,
-we can derive the staged abstract interpreters without intrusive changes to
-the corresponding unstaged one, thus the soundness is untouched.
-In this sense, our approach is no regret of both performance and engineering
-effort. We elaborate these two main ideas in detail as follow.
+different semantics, which allows user to implement different
+semantics modularly, including the staged abstract interpretation
+semantics.  Together with the multi-stage programming and type-level
+annotations of binding-times, we can derive the staged abstract
+interpreters without intrusive changes to the corresponding unstaged
+one, thus the soundness is untouched.  In this sense, our approach is
+no regret of both performance and engineering effort. We elaborate
+these two main ideas in detail as follow.
 
@@
 The idea of specializing interpreters can be traced to Futamura
-projections \cite{Futamura1999, futamura1971partial}.
-The first Futamura projection specifically shows that
-specializing an interpreter w.r.t. the input program yields an
-equivalent executable. For instance, @eval@ is an interpreter for
-some language:
+projections \cite{Futamura1999, futamura1971partial}.  The first
+Futamura projection specifically shows that specializing an
+interpreter w.r.t. the input program yields an equivalent
+executable. For instance, @eval@ is an interpreter for some language:
 \begin{lstlisting}
@@
 Given a program $e_0$ of type @Expr@, by applying the specialization,
-we can obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ :
-@Input@ $\to$ @Value@, which is the so-called \textit{equivalent executable}.
-By definition of the interpreter, they produce the same result when applied
-with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
+we can obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$
+: @Input@ $\to$ @Value@, which is the so-called \textit{equivalent
+  executable}.  By definition of the interpreter, they produce the
+same result when applied with the argument
+$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
 
@@
 the binding-times (\textit{static} or \textit{dynamic}) of variables,
-then evaluates the static part of the program, and generates a residualized
-program that solely relies on the dynamic part. However, it is hard to
-precisely analyze binding-times given an arbitrary program.
-As an alternative approach to specialization, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97}
-lets the programmer to explicitly control and annotate the
-binding-times in the program, then the MSP system will check whether these annotations
-are consistent and specialize the program using that information.
-The staging annotations can be either syntactic (e.g., quote and quasiquote
-in MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging
-framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
+then evaluates the static part of the program, and generates a
+residualized program that solely relies on the dynamic part. However,
+it is hard to precisely analyze binding-times given an arbitrary
+program.  As an alternative approach to specialization, multi-stage
+programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets
+the programmer to explicitly control and annotate the binding-times in
+the program, then the MSP system will check whether these annotations
+are consistent and specialize the program using that information.  The
+staging annotations can be either syntactic (e.g., quote and
+quasiquote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
+Modular Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
 
-Our proposed framework adopts the type-based multi-stage programming from
-the Lightweight Modular Staging framework and implements the Futamura
-projection of a big-step abstract interpreter for a small stateful,
-higher-order language.
-We use \citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
+Our proposed framework adopts the type-based multi-stage programming
+from the Lightweight Modular Staging framework and implements the
+Futamura projection of a big-step abstract interpreter for a small
+stateful, higher-order language.  We use
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
 interpreter that uses monad transformers as the unstaged baseline. By
@@
 After staging, the generated code is specialized to the input program,
-and all monadic operations are all inlined and compiled down to low-level
-Scala code.
+and all monadic operations are all inlined and compiled down to
+low-level Scala code.
 
@@
 
-Program specialization and abstract interpretation are two orthogonal concepts.
-To implement the confluence of them, we first construct a generic interpreter that
-is agnostic to both binding-times and value domains used in the semantics.
-Later, the generic interpreter can be instantiated from these two dimensions (Figure~\ref{confluence}):
+Program specialization and abstract interpretation are two orthogonal
+concepts.  To implement the confluence of them, we first construct a
+generic interpreter that is agnostic to both binding-times and value
+domains used in the semantics.  Later, the generic interpreter can be
+instantiated from these two dimensions (Figure~\ref{confluence}):
 \begin{itemize}
-\item With a flat binding-time and concrete domains, it is a ordinary definitional interpreter based
-  on big-step operational semantics;
-\item With two-level binding-times and concrete domain, it is a compiler that translate to program
-  intro antoher language;
-\item With a flat binding-time and abstract domains, it is an definitional abstract interpreter
-  \cite{DBLP:journals/pacmpl/DaraisLNH17} that statically computes runtime properties;
-\item With two-level binding-times and abstract domains, it is an optimizing program analyzer but
-  works in the fashion of compilation.
+\item With a flat binding-time and concrete domains, it is a ordinary
+  definitional interpreter based on big-step operational semantics;
+\item With two-level binding-times and concrete domain, it is a
+  compiler that translate to program intro antoher language;
+\item With a flat binding-time and abstract domains, it is an
+  definitional abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}
+  that statically computes runtime properties;
+\item With two-level binding-times and abstract domains, it is an optimizing
+  program analyzer but works in the fashion of compilation.
 \end{itemize}
 
-Although the four artifacts may look like very differently at first glance,
-but in fact are all firmly rooted in the concrete semantics of the language.
-This observation provides a way to abstract over the interpreter and achieve
-the flexibility of reinterpreting the shared interpreter. We adapt the monadic
-interpreter schema, which already have been applied to abstract interpreters,
-and introduce the binding-time abstraction into the generic interpreter.
-The generic interpreter returns a value of monadic type, which can be varied
-by different semantics. The domain of the interpreter and the effects such as read,
-state and nondeterminism can be both wrapped into this monadic type.
-The binding-time abstraction is represented by a higher-kinded type,
-and controls whether the interpreter produces values directly or generates code.
-It is worth to mention that the binding-time type is also instrumented into
-the monadic type, so that we will distinguish normal monads and staged monads.
+Although the four artifacts may look like very differently at first
+glance, but in fact are all firmly rooted in the concrete semantics of
+the language.  This observation provides a way to abstract over the
+interpreter and achieve the flexibility of reinterpreting the shared
+interpreter. We adapt the monadic interpreter schema, which already
+have been applied to abstract interpreters, and introduce the
+binding-time abstraction into the generic interpreter.  The generic
+interpreter returns a value of monadic type, which can be varied by
+different semantics. The domain of the interpreter and the effects
+such as read, state and nondeterminism can be both wrapped into this
+monadic type.  The binding-time abstraction is represented by a
+higher-kinded type, and controls whether the interpreter produces
+values directly or generates code.  It is worth to mention that the
+binding-time type is also instrumented into the monadic type, so that
+we will distinguish normal monads and staged monads.
 
@@
 3) We notice that staging an abstract interpreter enables modularly
-compiling an analysis to programs. Here we borrow the concept of modular
-analysis, and show that the compiled analysis is reusable.
-Therefore the approach provides a modular way to create optimized analysis
-code by mechanized reusing a whole-program analyzer.
+compiling an analysis to programs. Here we borrow the concept of
+modular analysis, and show that the compiled analysis is reusable.
+Therefore the approach provides a modular way to create optimized
+analysis code by mechanized reusing a whole-program analyzer.
 4) We also empirically evaluate the performance improved by staging,
-showing an order of magnitude speed-up on flow-analysis \todo{explicit note the number}.
+showing an order of magnitude speed-up on flow-analysis \todo{explicit
+  note the number}.
 
@@
   \item We begin by introducing our target language and reviewing
-    monads in Scala, and then presenting the generic interpreter (Section~\ref{prelim}).
-    After which, we review its instantiations of concrete interpretation
-    (Section~\ref{unstaged_conc}) and staged concrete interpretation
-    (Section~\ref{stagedinterp}).
-  \item We present the unstaged abstract interpreter under the same framework by
-    replacing the environment, store and values to their abstract counterparts (Section~\ref{unstaged_abs}).
-    After, we show the combination of approximation and specialization, dubbed
-    \textit{staged abstract interpreters}, can be readily derived (Section~\ref{sai}).
-  \item We conduct three case studies (Section~\ref{cases_study}), which demonstrate that
-    our approach requires less engineering efforts, is applicable to various analysis,
-    and enables compiling analysis modularly.
-  \item We empirically evaluate the performance improvement by staging on
-    control-flow analysis (Section~\ref{evaluation}). We compare the both
-    context-insensitive and store-widening analysis.
+    monads in Scala, and then presenting the generic interpreter
+    (Section~\ref{prelim}).  After which, we review its instantiations
+    of concrete interpretation (Section~\ref{unstaged_conc}) and
+    staged concrete interpretation (Section~\ref{stagedinterp}).
+  \item We present the unstaged abstract interpreter under the same
+    framework by replacing the environment, store and values to their
+    abstract counterparts (Section~\ref{unstaged_abs}). Then we
+    show the combination of approximation and specialization, dubbed
+    \textit{staged abstract interpreters}, can be readily derived
+    (Section~\ref{sai}). We also summarize the approach and discuss
+    the soundness issue after showing the four artifacts.
+  \item We conduct three case studies (Section~\ref{cases_study}) and
+    an empirical evaluation on performance improved (Section~\ref{evaluation}).
+    Finally we discuss related work and conclude.
 \end{itemize}

### 2019-08-14 paper/20_prelim.tex
revision

@@
 
-In this section, we first describe the abstract syntax of the language for our interpreter, 
-then present the generic interpreter shared among the four different
+In this section, we first describe the abstract syntax of the language for our interpreters. 
+Then present the generic interpreter shared among the four different
 semantics, after which, we instantiate the interpreter to the concrete one.
-It is worth noting that we choose to use Scala and monad to demonstrate the
-idea, but the approach is not restricted to our choice. One can use
-imperative or direct style in other MSP languages (e.g., MetaOCaml
-\cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} and Template
-Haskell \cite{Sheard:2002:TMH:636517.636528} to construct such staged abstract
-interpreters.
+It is worth noting that we choose to use Scala and monadic style to
+demonstrate the idea, but the approach is not restricted to our
+choice. One can use imperative or direct style in other MSP languages
+(e.g., MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}
+and Template Haskell \cite{Sheard:2002:TMH:636517.636528} to construct
+such staged abstract interpreters.
 
@@
 
-Note that here the monadic object @List[_]@ encapsulates the data internally.
-Therefore it only exposes the simplified version of @flatMap@, where the
-monad @M[A]@ is not introduced as a function argument. The trait @MonadOps@
-(Figure \ref{fig:monadops}) defines the simplified version of monadic
-operations that are necessary for @for@-comprehension. 
-The conversion between @Monad@ and @MonadOps@ can be done by using the implicit design pattern.
-In the rest of the paper, we use Scala's @for@-comprehension syntax and monad
-transformers such as @ReaderT@, @StateT@, and @ListT@ to write our interpreters.
-The implementation of monads and monad transformers essentially borrows the
-ground-truth from Haskell.
+Note that here the monadic object @List[_]@ encapsulates the data
+internally.  Therefore it only exposes the simplified version of
+@flatMap@, where the monad @M[A]@ is not introduced as a function
+argument. The trait @MonadOps@ (Figure \ref{fig:monadops}) defines the
+simplified version of monadic operations that are necessary for
+@for@-comprehension. The conversion between @Monad@ and @MonadOps@
+can be done by using the implicit design pattern.
+In the rest of the paper, we use Scala's @for@-comprehension syntax
+and a few of monads and monad transformers such as @ReaderT@,
+@StateT@, and @ListT@ to write our interpreters. The implementation
+of monads and monad transformers essentially borrows the ground-truth
+from Haskell.
 
@@
 transformers, we can combine multiple monads into a single one. Constructing
-extensible interpreters using monad transformers was first proposed by
+extensible interpreters using monad transformers was first explored by
 \citet{DBLP:conf/popl/LiangHJ95}, and later has been applied to abstract
 interpreters \cite{Sergey:2013:MAI:2491956.2491979,
-DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}. In this section,
-with the multi-stage programming and monad transformers in mind while leaving them
-as abstract type members, we present the generic interface of a big-step
-definitional interpreter.
+  DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
+In this section, we present the generic inferface in the style of big-step
+definitional interpreter. The key idea is to keep the binding-time type and
+returned monadic type abstract, so that they can be instantiated differently.
+We also need to abstract the primitive operations on those types.
 
 \paragraph{Basic Types} We start with some basic type definitions used in the
-interpreter. The identifiers in the program are represented by strings. The two
-default components in the interpreter are environments @Env@ and stores @Store@,
-i.e., mappings from identifiers to addresses and mappings from addresses to
-values, respectively. The @Env@ models accessible variables, and @Store@ models
-the persistent heap through the program runtime. But the @Addr@ and @Value@ are
-just declared as abstract types.
+interpreter. The identifiers in the program are represented by
+strings. To represent states, two required components for the
+interpreter are environments @Env@ and stores @Store@. @Env@ maps from
+identifiers to addresses, and @Store@ maps from addresses to values,
+respectively. The @Env@ captures bound variables in scope, and @Store@
+models a persistent heap through the program runtime.  At this moment,
+the domains of addresses and values are still abstract, i.e., they are
+declared as abstract types.
 
@@
 \paragraph{Binding-time Abstraction} As mentioned before, the binding-time is
-declared as a higher-kinded type @R[_]@. If we simply instantiate @R@ as an identity
-type (i.e., @type R[T] = T@), then the generic interpreter will execute the program.
-In Section \ref{stagedinterp}, we will instantiate @R@ using LMS's built-in
-next-stage type annotation @Rep@, which makes the interpreter act as a compiler.
+declared as a higher-kinded type @R[_]@. We are free to use @R@ to annotate
+other datatypes in the interpreter; it is also injected into the monadic
+type @MonadOps@.
+If we simply instantiate @R@ an identity type (i.e., @type R[T] = T@),
+then the generic interpreter is a standard definitional interpreter
+that will execute the program.  In Section \ref{stagedinterp}, we will
+instantiate @R@ using LMS's built-in next-stage type annotation @Rep@,
+which makes the interpreter act as a compiler.
 
 \paragraph{Monadic Operations} We define the return type of the interpreter as
-@Ans@, which is an arbitrary monad type @AnsM[_]@ wrapping the type @Value@. 
+@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
 As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
-syntax, certain constraints have to be added on the type @AnsM@. Here, we use a
-structural type @MonadOps@ to require @AnsM@ to at least implement @map@ and
+syntax, certain operations have to be added on the type @AnsM@. Here, we use a
+structural type @MonadOps@ to require that type @AnsM@ at least implements @map@ and
 @flatMap@. It is worth noting that @MonadOps@ takes another type parameter
-@R[_]@ as binding-time; accordingly the outer @R@ defined in the trait is passed
-to @MonadOps@. Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
-encapsulated by the monad, but not the monad type @M@ itself. When acting as
-compilers, we will also replace the monads to the ones that work on staged
-values.
+@R[_]@ as binding-time that annotates data; accordingly the type @R@
+defined in the outer scope is passed to @MonadOps@. Inside of @MonadOps@,
+@R[_]@ wraps the data types @A@ and @B@ that are encapsulated by the
+monad, but not the monad type @M@ itself. When acting as compilers, we
+will also replace the monads to the ones that work on staged data values.
+\todo{it is important that R only annotates data, not the monad}
 
-We also define several methods to obtain and modify the environment and store.
-These methods return monadic values of type @AnsM[_]@, which encapsulate the 
-environment or store, or simply a @Unit@ value for effects.
-For example, @local_env@ takes a value @ans@ of type @Ans@ and an environment 
-which will be installed when evaluating @ans@.
+We also declare several operations to mannipulate environments and
+stores. These methods return monadic values of type @AnsM[_]@, which
+may encapsulate over the environment or store, or simply a @Unit@
+value for effects. For examples, @local_env@ installs a new environment
+@ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
+addresses and values (potentially staged) and updates the store accordingly.
 
@@
 the freshness of the address may depend on the store, which might be a
-next-stage value as indicated by its type, the type of addresses is also wrapped
-by @R[_]@. The other one would simply wrap the address with our monadic type
-@AnsM[_]@.
+next-stage value as indicated by its type, the type of addresses is therefore
+wrapped by @R[_]@. The other @alloc@ is simply the monadic version of the addresses.
+
 \begin{lstlisting}
@@
 
-Other primitive operations provide basic functionality for the interpreter.
-The method @num@ and @close@ deal with primitive values, which lift literal
-terms (e.g., lambdas) to our value representation (e.g., closures).
-The method @get@ simply retrieves the value mapped from the identifier @x@ in
-the environment and store. Conditionals and arithmetic is handled by @br0@
-and @arith@, respectively. The methods @ap_clo@ takes a function value and an
-argument value and then does the application. Note that the @Env@, @Store@, and
-@Value@ are all wrapped by @R[_]@ since they will be known as next-stage value
-when acting as compilers.
-
-%TODO: close/ap_clo takes an ev
+Other primitive operations provide basic functionality for the
+interpreter.  The method @num@ and @close@ deal with primitive values,
+which lift literal terms (e.g., lambdas) to our value representation
+(e.g., closures).  The method @get@ retrieves the value of an
+identifier @x@, through the environment and store. Conditionals and
+arithmetic is handled by @br0@ and @arith@, respectively. The method
+@ap_clo@ is used for function applications: it takes a function value
+and an argument value. Note that the @Env@, @Store@, and @Value@ are
+all annotated by @R[_]@ since they can potentially be next-stage
+values when acting as compilers.
 
@@
 
-\paragraph{The Interpreter} Now we can define the semantics-agnostic interpreter
-in monadic form, shown in Figure \ref{fig:shared_int}.
-The essential idea is to traverse the abstract syntax tree while maintaining the
-effects such as reader and state.
-It is worth noting that the interpreter is written in open-recursive style -- it
-can not refer to itself directly, instead, @eval@ takes an additional parameter
-@ev@ of type @Expr => Ans@ to refer to itself. Accordingly, the method @close@
-that lifts lambda terms to closures and @ap_clo@ that applies functions also
-takes an extra @ev@.
-
-%\vspace{-1em}
 \begin{figure}[h!]
@@
 \end{figure}
+
+\paragraph{The Interpreter} Now we construct the semantics-agnostic interpreter
+in monadic form, shown in Figure \ref{fig:shared_int}. The basic idea
+of a generic interpretation is to traverse the abstract syntax tree
+while maintaining the effects such as reader and state.  Note that the
+interpreter is written in open-recursive style -- it can not refer to
+itself directly, instead, @eval@ takes an additional parameter @ev@ of
+type @Expr => Ans@ that refers to itself. Consequently, the method
+@close@ that lifts lambda terms to closures and method @ap_clo@ that
+applies functions also takes an extra @ev@, because evaluation may
+happen inside of them.
+To close the open-recursion, we use a fixed-point opertor @fix@.
+For concrete-interpretation instantiation, it works like the Y
+combinator; for abstract-interpretation instantiation, it instruments
+the interpreter by memorizing its input and output, which will ensure
+the termination of abstract interpretation.
+Finally, a top-level wrapper @run@ is declared; the return type
+@Result@ depends on what kind of monad we will be using, this is also
+an abstract type.
+
 %\vspace{-1em}
-
-Since the interpreter is written in an open-recursive style, we declare an abstract
-combinator @fix@ used to close the recursion. For concrete-interpretation instantiation, it
-works like the Y combinator; for abstract-interpretation instantiation, we will
-instrument the interpreter by defining a memorized version of @fix@ to ensure
-termination. Finally, a top-level wrapper @run@ is declared; the return type
-@Result@ depends on what kind of monad we will be using, this is also an abstract type.
 

### 2019-08-14 paper/50_stageabs.tex
revision

@@
     LMS framework, which 1) relies on the type checking facilities of Scala, and
-    2) generates code in A-Normal form that preserves the
-    execution order within a stage \todo{cite WF16}. However, in general, it is
-    possible to subvert these guarantees by using unsafe features, such as
-    casts.
+    2) generates code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113}
+    that preserves the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
+    However, in general, it is possible to subvert these guarantees by
+    using unsafe features, such as casts.
   \item As shown in the previous section, the generic interpreter is untouched

### 2019-08-14 paper/80_related.tex
revision

@@
 \paragraph{Optimizing Static Analysis Through Specialization}
-The underlying idea in this paper is closely related to abstract compilation (AC)
-\cite{Boucher:1996:ACN:647473.727587}: removing the interpretation overhead on
-traversing the syntax tree by specialization. Specifically, the residual program
-of AC can be either textual or closures. However, as we studied in Section
-\ref{cs_ac}, with the perspective from generative programming and monadic
+The underlying idea in this paper is closely related to abstract
+compilation (AC) \cite{Boucher:1996:ACN:647473.727587}: removing the
+interpretation overhead on traversing the syntax tree by
+specialization. Specifically, the residual program of AC can be either
+textual or closures. However, as we studied in Section \ref{cs_ac},
+with the perspective from generative programming and monadic
 interpreters, we make it more general, extensible and easy.
 \citet{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
-generation to optimize small-step abstract interpreter in state-transition
-style. The analyzed program is firstly compiled to an IR called "abstract
-bytecode", which are actually higher-order functions and will be executed later
-on an abstract abstract machine for that IR. \citet{damian1999partial} provides a
-formal treatment to abstract compilation and Shiver's CFA, as well as proofs to
-establish the correctness of the certified specialized analyzer.
-\citet{amtoft1999partial} applied partial evaluation for constraint-based
-control flow analysis.
+generation to optimize small-step abstract interpreter in
+state-transition style. The analyzed program is firstly compiled to an
+IR called "abstract bytecode", which are actually higher-order
+functions and will be executed later on an abstract abstract machine
+for that IR. \citet{damian1999partial} provides a formal treatment to
+abstract compilation and Shiver's CFA, as well as proofs to establish
+the correctness of the certified specialized analyzer.
+\citet{amtoft1999partial} applied partial evaluation for
+constraint-based control flow analysis.
 
-Splitting an analysis into multiple stages is also studied for analysis other
-than abstract-interpretation-based control-flow analysis, though the formulation may
-very different. \citet{DBLP:conf/cgo/HardekopfL11} apply staging to a
-flow-sensitive pointer analysis. The first stage is to analyze the program code
-to obtain a sparse representation, and then the second stage conducts the
-flow-sensitive analysis by refining the first one. Another large class of static
-analysis are implemented by generating constraints firstly and then let a solver
-discharge these proof obligations or find the fixed-point. Examples include SAT-
-and SMT-based program analysis\todo{cite}, Datalog-based points-to
-analysis\todo{cite}, and \todo{}. Similar to the idea in this paper, such
-analyses can be splitted to staged.
-One of the future work is to explore that whether these analyses can be
-implemented and optimized from metaprogramming. Souffle, one of the
-state-of-the-art Datalog engine that is commonly used in points-to analysis,
-uses the idea of Futamura projection for efficent implemetation. But it only
-specializes Datalog rules to the running engine, instead of a full pipeline from
-program AST to generating constraints and solving constraints.
+Splitting an analysis into multiple stages is also studied for
+analysis other than abstract-interpretation-based control-flow
+analysis, though the formulation may very
+different. \citet{DBLP:conf/cgo/HardekopfL11} apply staging to a
+flow-sensitive pointer analysis. The first stage is to analyze the
+program code to obtain a sparse representation, and then the second
+stage conducts the flow-sensitive analysis by refining the first
+one. Another large body of static analyses are implemented by
+generating constraints firstly and then let a solver discharge these
+proof obligations or find the fixed-point. Notable examples include
+SAT- and SMT-based program analysis
+\cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to
+analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
+Similar to the idea in this paper, such analyses can be splitted to
+staged.  One of the future work is to explore that whether these
+analyses can be implemented and optimized from
+metaprogramming. Souffl{\'e}, one of the state-of-the-art Datalog
+engine that is commonly used in points-to analysis
+\cite{Antoniadis:2017:PDS:3088515.3088522}, uses the idea of Futamura
+projection for efficent implemetation \cite{10.1007/978-3-319-41540-6_23}.
+But it only specializes Datalog rules to the running engine, instead
+of a full pipeline from program AST to generating constraints and
+solving constraints.
 
-Recent work shows that the numerical abstract domains can be also specialized
-w.r.t. the static program structure. For example, decomposing polyhedras
-\cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to smaller ones
-depending on the variables involved in a statement can significantly improve the
-running time and space of abstract transformers.
+Recent work shows that the numerical abstract domains can be
+specialized w.r.t. the static program structure, although not in the
+way of using staging. For example, decomposing polyhedras
+\cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to
+smaller ones that depend on the variables involved in a statement can
+significantly improve the running time and space of abstract
+transformers.
 
-\paragraph{Construction of Abstract Interpreters} Abstract interpretation was proposed as a
-semantic-based approach to build sound static analysis by approximation
-\cite{DBLP:conf/popl/CousotC77}. As to build semantic artifacts, the Abstracting
-Abstract Machines (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
-approach shows that abstract interpreters can be derived systematically from concrete
-semantic artifacts. The AAM approach is closely related to control-flow analysis
-for higher-order languages \cite{Midtgaard:2012:CAF:2187671.2187672}.
-Using monads to construct abstract interpreters is explored by
+\paragraph{Construction of Abstract Interpreters} Abstract interpretation
+was proposed as a semantic-based approach to build sound static analysis by
+approximation \cite{DBLP:conf/popl/CousotC77}. As to build semantic
+artifacts, the Abstracting Abstract Machines (AAM)
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} approach
+shows that abstract interpreters can be derived systematically from
+concrete semantic artifacts. The AAM approach is closely related to
+control-flow analysis for higher-order languages
+\cite{Midtgaard:2012:CAF:2187671.2187672}.  Using monads to construct
+abstract interpreters is explored by
 \citet{Sergey:2013:MAI:2491956.2491979} and
-\citet{DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
-The unstaged abstract interpreter in this paper follows the abstracting definitional
-interpreters approach from \citet{DBLP:journals/pacmpl/DaraisLNH17}.
-Similar to definitional abstract interpreters, \citet{Wei:2018:RAA:3243631.3236800}
-reconstruct big-step abstract interpreters with delimited continuations;
-\citet{Keidel:2018:CSP:3243631.3236767} present a modular arrow-based abstract
-interpreter that makes proof of soundness can be constructed compositionally.
+\citet{DBLP:journals/pacmpl/DaraisLNH17,
+  Darais:2015:GTM:2814270.2814308}.  The unstaged abstract interpreter
+in this paper follows the abstracting definitional interpreters
+approach from \citet{DBLP:journals/pacmpl/DaraisLNH17}.  Similar to
+definitional abstract interpreters,
+\citet{Wei:2018:RAA:3243631.3236800} reconstruct big-step abstract
+interpreters with delimited continuations;
+\citet{Keidel:2018:CSP:3243631.3236767} present a modular arrow-based
+abstract interpreter that makes proof of soundness can be constructed
+compositionally.
 
-\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis. We borrow the notation of modular analysis, \todo{but intentional properties
-  are different, compare with summary}
+\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation
+framework for modular analysis. We borrow the notation of modular
+analysis, \todo{but intentional properties are different, compare with
+  summary}
 
@@
 abstract interpretation can be traced to Nielson's two-level semantics
-\cite{NIELSON1989117}; using two-level semantics for code generation was also
-explored by \citet{NIELSON198859}. \citet{Sergey:2013:MAI:2491956.2491979}'s
-work of monadic abstract interpreters is also closely related to the two-level
-semantics: the use of a generic interface with monads and then reinterpret it by
-different semantics is already two-level. Instead of focusing on semantics, this
-paper shows how a staged analyzer can be built and used to increase efficieny of
-static analysis, we augment the monadic abstract interpreter from another
-dimension, using code generation to produce efficient low-level code.
+\cite{NIELSON1989117}; using two-level semantics for code generation
+was also explored by \citet{NIELSON198859}.
+\citet{Sergey:2013:MAI:2491956.2491979}'s work of monadic abstract
+interpreters is also closely related to the two-level semantics: the
+use of a generic interface with monads and then reinterpret it by
+different semantics is already two-level. Instead of focusing on
+semantics, this paper shows how a staged analyzer can be built and
+used to increase efficieny of static analysis, we augment the monadic
+abstract interpreter from another dimension, using code generation to
+produce efficient low-level code. Hence the presented work can be
+considered as an two-dimension application of the two-level semantics.
 
 \paragraph{Partial Evaluation and Multi-stage Programming}
-Partial evaluation as an automatic technique for program specialization
-was studied comprehensively by \citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
-In this paper, we use mutli-stage programming as an approach to
-program specialization. The Lightweight Modular Staging framework \cite{DBLP:conf/gpce/RompfO10}
-we used in the paper relies on type-level stage annotations.
-Other notable implementations of MSP exist in ML family, e.g.,
-MetaML \cite{DBLP:conf/pepm/TahaS97} and MetaOCaml
-\cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}.
-The idea of staging an abstract interpreter we presented in this paper is also
-applicable to other MSP systems.
-Multi-stage programming has been widely used to improve the performance in many
-domains, such as optimizing compilers or domain-specific languages
-\cite{DBLP:conf/pldi/RompfSBLCO14, DBLP:conf/snapl/RompfBLSJAOSKDK15,
+Partial evaluation as an automatic technique for program
+specialization was studied comprehensively by
+\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.  In this
+paper, we use mutli-stage programming as an approach to program
+specialization. The Lightweight Modular Staging framework
+\cite{DBLP:conf/gpce/RompfO10} we used in the paper relies on
+type-level stage annotations.  Other notable implementations of MSP
+exist in ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97} and
+MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
+  DBLP:conf/flops/Kiselyov14}.  The idea of staging an abstract
+interpreter we presented in this paper is also applicable to other MSP
+systems.  Multi-stage programming has been widely used to improve the
+performance in many domains, such as optimizing compilers or
+domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14,
+  DBLP:conf/snapl/RompfBLSJAOSKDK15,
   DBLP:journals/tecs/SujeethBLRCOO14, DBLP:conf/gpce/SujeethGBLROO13,
@@
   DBLP:conf/pepm/AktemurKKS13}, generic programming
-\cite{DBLP:journals/pacmpl/Yallop17, Ofenbeck:2017:SGP:3136040.3136060}, data
-processing \cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
+\cite{DBLP:journals/pacmpl/Yallop17,
+  Ofenbeck:2017:SGP:3136040.3136060}, data processing
+\cite{DBLP:conf/oopsla/JonnalageddaCSRO14,
   DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
-\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18}, etc.
+\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
+etc.
 
-As an source of inspiration of this paper, Futamura projections reveals a
-hierarchy and close relation between interpreters and compilers, which was
-originally proposed by \citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial},
+As an source of inspiration of this paper, Futamura projections
+reveals a hierarchy and close relation between interpreters and
+compilers, which was originally proposed by
+\citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial},
 and later reprinted \cite{Futamura1999}.
-\citeauthor{Amin:2017:CTI:3177123.3158140} consider a tower of concrete
-interpreters and how to collapse them by using MSP -- it would be
-interesting to explore this idea for multiple layers of abstract interpreters
-\cite{Cousot:2019:AAI:3302515.3290355, Giacobazzi:2015:APA:2676726.2676987}.
-\citet{10.1007/11561347_18} discussed combining
-multi-stage programming with functors and monads. But they didn't \todo{monad transfomers?}
-Similar to the idea in this paper, specializing monadic interpreters was
-explored by \citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}. \todo{what's the difference?}
+\citeauthor{Amin:2017:CTI:3177123.3158140} consider a tower of
+concrete interpreters and how to collapse them by using MSP -- it
+would be interesting to explore this idea for multiple layers of
+abstract interpreters \cite{Cousot:2019:AAI:3302515.3290355,
+  Giacobazzi:2015:APA:2676726.2676987}.  \citet{10.1007/11561347_18}
+discussed combining multi-stage programming with functors and
+monads. But they didn't \todo{monad transfomers?}  Similar to the idea
+in this paper, specializing monadic interpreters was explored by
+\citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}. \todo{what's the
+  difference?}
 

### 2019-08-14 paper/20_prelim.tex
update

@@
 \paragraph{Primitive Operations} Next we define several primitive operations.
-Two versions of @alloc@ are declared. The first one takes a store and an
+First, two versions of @alloc@ are declared. The first one takes a store and an
 identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
@@
 combinator; for abstract-interpretation instantiation, it instruments
-the interpreter by memorizing its input and output, which will ensure
+the interpreter by memoizing its input and output, which will ensure
 the termination of abstract interpretation.
@@
 
-Now we can instantiate the interpreter as a standard environment-store
-interpreter. Such interpreter also can be obtained by refunctionalization of
-CESK machines \cite{Felleisen:1987:CAH:41625.41654, DBLP:conf/ppdp/AgerBDM03}.
+As the first step in our roadmap, we instantiate the interpreter concretely in
+this section. The result of such instantiation is a standard definitional
+interpreter with environments and stores, it also can be obtained by
+refunctionalization of CESK machines \cite{Felleisen:1987:CAH:41625.41654,
+DBLP:conf/ppdp/AgerBDM03}. We first present the concrete components, i.e., the
+value domain, then show the monad stack for concrete interpretation, finally
+sketch the how the primitive operations are implemented.
 
 \paragraph{Concrete Components}
-The two types we need to concretize are @Addr@ and @Value@, afterward, the @Env@
-and @Store@ is concretized automatically. To assure the freshness, we use @Int@
-for the address space @Addr@. A value can be either a tagged number
-\texttt{IntV}, or a closure \texttt{CloV} that contains a lambda term and an
-environment. The final return value of the interpreter is a pair of values and
-stores, where the @Value@ and @Store@ are obviously next-stage objects. We also
-provide a standard fixed-point combinator to close the open-recursive function @ev@.
+The two types we need to concretize are addresses @Addr@ and values @Value@. The
+type @Env@ and @Store@ is derived automatically. To assure the freshness of
+address allocation, we use @Int@ and always return successor of the size of
+current store. A value can be either a tagged number \texttt{IntV}, or a closure
+\texttt{CloV} that contains a lambda term and an environment. The final grounded
+result of the interpreter is a pair of values and stores, where the @Value@ and
+@Store@ are next-stage objects, as annotated by @R@. We also define a standard
+fixed-point combinator to close the open-recursive function @ev@.
 
@@
 \end{lstlisting}
-%\vspace{-1em}
 
 \paragraph{Unstaged Monads}
-The environment and store can be modeled by reader effect and state effect,
-i.e., a Reader monad and a State monad. We combine them using monad transformers
-@ReaderT@ and @StateT@.
-In other words, we instantiate the @AnsM@ monad as a stack of @ReaderT@ and @StateT@
-transformers\footnote{The question mark syntax is a kind projector
-  \cite{kindprojector}, thus \texttt{StateT[IdM,Store,?]} is equivalent to \newline
-  \texttt{(\{type M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is
-parameterized by type @Env@, @StateT@ is parameterized by type @Store@, and the
-inner-most monad @IdM@ is just an identity monad.
+For the purpose of concrete interpretation, the monad needs to model read
+effect and state effect, which correspond to environment and store,
+respectively. Following the monad transformer approach
+\cite{DBLP:conf/popl/LiangHJ95}, we use a @ReaderT@ and @StateT@ monad
+transformer to compose the monad stack. In other words, the type @AnsM@ is
+instantiated by layering @ReaderT@ and @StateT@ transformers\footnote{The
+question mark syntax is a kind projector \cite{kindprojector}, thus
+\texttt{StateT[IdM,Store,?]} is equivalent to \newline \texttt{(\{type
+M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is parameterized by type
+@Env@, @StateT@ is parameterized by type @Store@, and the inner-most monad @IdM@
+is simply an identity monad.
 
@@
     type R[T] = T
-    type AnsM[T] = ReaderT[StateT[IdM, Store, ?], Env, T]
-    ...
-  }
+    type AnsM[T] = ReaderT[StateT[IdM, Store, ?], Env, T] // the monad stack
+    ... }
 \end{lstlisting}
 
-Here we sketch the basic idea of unstaged @ReaderT@ and @StateT@. Readers may
-refer to \cite{DBLP:conf/popl/LiangHJ95, Chiusano:2014:FPS:2688794} for more detail 
-and implementation of monad transformers.
-A @ReaderT@ monad transformer encapsulates the computation @R => M[A]@, where
-@R@ is the reader type, @M[_]@ is a monad type. The @ReaderT@ monad returns the
-transformed value of type @M[A]@. Similarly, a @StateT@ monad encapsulates the
-computation @S => M[(A, S)]@, where @S@ is the state type, @M[_]@ is a monad
-type. The @StateT@ monad returns the transformed value of type @M[(A, S)]@,
-where the new state is accompanied.
-Note that the binding-time type @R@ is an identity type, thus the monads are
-operated on unstaged data. We can also see this from the signature of @flatMap@:
-the function @f@ takes an unstaged value of type @A@ and produces a monadic
-value.
+Here we sketch the basic idea of @ReaderT@ and @StateT@. Readers may
+refer to \cite{DBLP:conf/popl/LiangHJ95, Chiusano:2014:FPS:2688794} for more detail.
+A @ReaderT@ monad transformer encapsulates computation @R => M[A]@, where
+@R@ is the reader type, and @M[_]@ is the inner monad type.
+Given a @R@, @ReaderT@ monad produecs a transformed value of type @M[A]@.
+Similarly, a @StateT@ monad encapsulates computation @S => M[(A, S)]@, where
+@S@ is the state type, and @M[_]@ is the inner monad type.
+Given a @S@, @StateT@ monad produces a transformed value of type @M[(A, S)]@,
+where the new state (type @S@) is accompanied with the value (type @A@).
+Note that for the moment, the binding-time type @R@ is an identity type, thus
+these monads operate on unstaged data. We can also see this from the
+signature of @flatMap@: the argument function @f@ takes an unstaged value of type @A@ and
+produces a monadic value. In the following code, we elide operations other than @flatMap@.
 
@@
   case class ReaderT[M[_]: Monad, R, A](run: R => M[A]) {
-    def flatMap[B](f: A => ReaderT[M, R, B]): ReaderT[M, R, B] = ... }
+    def flatMap[B](f: A => ReaderT[M, R, B]): ReaderT[M, R, B] =
+      ReaderT(r => Monad[M].flatMap(run(r))(a => f(a).run(r))); ... }
   case class StateT[M[_]: Monad, S, A](run: S => M[(A, S)]) {
-    def flatMap[B](f: A => StateT[M, S, B]): StateT[M, S, B] = ... }
+    def flatMap[B](f: A => StateT[M, S, B]): StateT[M, S, B] =
+      StateT(s => Monad[M].flatMap(run(s)) { case (a, s1) => f(a).run(s1) }); ... }
 \end{lstlisting}
 
-Then, the monadic operations that manipulate the environment and store are
-simply constructing the proper monads and lifting it to the top-level of our
-monad stack. To modify the store, for example, we may have a @StateT@ monad that
-transforms the current store $\sigma$ to @σ + αv@, which results in
-@StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to @ReaderT@,
-i.e., @AnsM[Unit]@.
+After defining the monad stack, the operations that manipulate the environment
+and store can be defined by constructing the proper monad and lifting it to the
+top-level of our monad stack. To modify the store, for example, we can construct a
+@StateT@ value that transforms the current store $\sigma$ to @σ + αv@, which
+results in @StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to the
+top-level @ReaderT@ type, i.e., @AnsM[Unit]@.
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations}
-Other primitive operations can be implemented straightforwardly, for example,
-@alloc@ simply takes the size of the current store and produces the successor number.
-We elide most of them but describe a little bit on how do we handle functions
-and applications as they will be very different when staging and abstract
-interpretation is involved. The method @close@ takes a lambda term and an
-environment, then create a closure value @CloV@.
+Other primitive operations operates on the value domain can be implemented
+straightforwardly. We elide most of them but describe a bit on how we handle
+functions and applications. Because lambda terms are data but also part of the
+control flow, although the way we treat them is simple now, but will be very
+different when staging and abstract interpretation is involved. The method
+@close@ takes a lambda term and an environment, and produces the
+defunctionalized representation of closures.
 
@@
 
-To apply a function, the method @ap_clo@ takes a function value and an argument
-value, and extracts the lambda term and environment enclosed in the function value,
-and evaluates the body expression of that lambda term under the new environment and store.
+To apply a function, @ap_clo@ takes a function value @rator@ and an argument
+@rand@, extracts the lambda term and environment enclosed in @rator@, allocates
+the addresses and updates the environment and store accordingly. Finally, it
+evaluates the body expression of the lambda term under the new environment and
+store.
 
@@
 
-With other primitive operations implemented, we can implement the top-level
-@run@ method, where $\rho_0$ and $\sigma_0$ are the initial empty environment and
-store.
+The top-level @run@ method invokes the fixed-pointer operator directly,
+given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$.
 

### 2019-08-14 paper/30_stageconc.tex
update

@@
 In this section, based on the generic interpreter and concrete components we
-presented in Section \ref{prelim}, we show how to obtain a staged concrete
-interpreter by changing the monads and refactoring several primitive
-operations. We begin by briefly introducing the Lightweight Modular Staging
-framework in Scala.
+presented in Section \ref{prelim}, we show how to stage the concrete interpreter
+by changing the monad type and refactoring several primitive operations. We
+begin by briefly introducing the Lightweight Modular Staging framework in Scala,
+and then replay the same step in the unstaged counterpart. At last, we briefly
+describe code generation.
 

### 2019-08-14 paper/30_stageconc.tex
revision

@@
 begin by briefly introducing the Lightweight Modular Staging framework in Scala,
-and then replay the same step in the unstaged counterpart. At last, we briefly
+and then replay the same steps from the unstaged counterpart. At last, we briefly
 describe code generation.
@@
 dynamic code generation in a type-safe manner. Different from the approach of
-MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} that
-uses syntactic quotation and escape, LMS distinguishes binding-time solely based
-on types. LMS provides a type constructor @Rep[T]@ where @T@ can be an arbitrary
-type, indicating a value of @T@ will be known at the next stage. All operations
-acting on @Rep[T]@ expressions will be residualized as generated code.
+MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} that
+uses syntactic quotations and quasiquotations, LMS distinguishes binding-time
+based on types. LMS provides a type constructor @Rep[T]@ where @T@ can be
+an arbitrary type, indicating a value of @T@ will be known at the next stage.
+All operations acting on @Rep[T]@ expressions will be residualized as generated
+code.
 
-For the example we introduced in Section \ref{intro}, we have identified @b@ will 
-be known later. To specialize it we write a function @snippet@ that provides a 
-concrete value for @x@. After this, we may instantiate an @DslDriver@ object 
-which provide the specialized program.
+A classic example for introducing multi-stage programming is the power function
+that computes $b^x$, which is usually implemented as a recursive function:
+
+\begin{lstlisting}
+  def power(b: Int, x: Int): Int = if (x == 0) 1 else b * power(b, x - 1)
+\end{lstlisting}
+
+If @x@ is a value know at current stage, we may specialize the power function to
+some value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
+first adding the @Rep@ type to the variable known at next stage. In this case,
+@b : Rep[Int]@ is know later, and @x : Int@ is know currently, as shown in the
+below code recipe.
+The way we use this @power@ function is to create a @DslDriver@ and override the
+@snippet@ method give the currently know value @x@ (5 in this example).
 
@@
     def power(b: Rep[Int], x: Int): Rep[Int] = if (x == 0) 1 else b * power(b, x-1)
-    def snippet(b: Rep[Int]): Rep[Int] = power(b, 5)
+    def snippet(b: Rep[Int]): Rep[Int] = power(b, 5) // specialize the power to b^5
   }
@@
 The LMS framework provides staging support for primitive data types such as
-@Int@ and @Double@, and commonly-used data structures such as arrays and maps.
-The idea is to construct intermediate representations (IR) for the next stage
-program at the current stage. Such conversion from expressions of type @Rep[_]@ to
-their IR is done by using the implicit design pattern. As we will see later,
-implementing the staging support for custom classes is also straightforward.
+@Int@ and @Double@, and commonly-used data structures such as lists and maps.
+The idea behind the framework is to construct a sea-of-node intermediate
+representations (IR) for the next stage program at the current stage
+\cite{DBLP:conf/birthday/Rompf16}. For convinence, the conversion from
+expressions of type @Rep[_]@ to their IR is done by using the implicit design
+pattern. As we will see later, implementing staging and code generation
+support for user-defined classes is also straightforward.
 
@@
 
-To implement the staged concrete interpreter, we reuse the exactly same concrete
-components from Section \ref{unstaged_conc}, but we instantiate the binding-time
-type @R[T]@ as @Rep[T]@.
+To implement the staged concrete interpreter, we replay the steps from
+instantiating unstaged concrete interpreter in Section \ref{unstaged_conc}.
+But now we use the @Rep@ type to annotate value domains, environments and
+stores, and redefine the staged version of monads and primitive operations.
 
-\paragraph{Staged Monads} We also need to use monads that operate on staged
-values. Note that the monadic objects such as @StateT@ or the internal value
-@R[S] => M[(A,S)]@ are not staged, only the data @S@ or @A@ are staged.
-This is why we can completely eliminate the overhead caused by the monadic layers.
-Here we use the prefix name @Rep@ to differentiate unstaged monads and staged
-monads, such as @RepReaderT@. So the @AnsM[_]@ is instantiated as the same monad
-stack before, but with @Rep@ versions.
+\paragraph{Staged Monads}
+We use the same structure of monad stack in the unstaged interpreter: a reader
+monad with a state monad. But now the monads operate on staged values, for
+brivity, we call them \textit{staged monads}. In the code, we also use a @Rep@
+prefix on the constructors and types to differentiate them. But it is important
+to note that objects of a case class @ReaderT@/@StateT@ are not staged, the
+monadic computation like @R => M[A]@ are also not staged. Instead, the internal
+data that these monads operate on are staged. The following code snippet shows
+the idea (we use light gray to highlight what have been changed to staged
+version except naming):
+
+\begin{lstlisting}[escapechar=!]
+  case class RepReaderT[M[_]: RepMonad, R, A](run: !\hl{Rep[R]}! => M[A]) {
+    def flatMap[B](f: !\hl{Rep[A]}! => ReaderT[M, R, B]): RepReaderT[M, R, B] =
+      RepReaderT(r => RepMonad[M].flatMap(run(r))(a => f(a).run(r))); ... }
+  case class RepStateT[M[_]: RepMonad, S, A](run: !\hl{Rep[S]}! => M[(A, S)]) {
+    def flatMap[B](f: !\hl{Rep[A]}! => StateT[M, S, B]): RepStateT[M, S, B] =
+      RepStateT(s => RepMonad[M].flatMap(run(s))(as => f(as._1).run(as._2))); ... }
+\end{lstlisting}
+
+The function @f@ passed to @flatMap@ is also a current-stage-known value, so
+that all the invocation of @flatMap@s can be completed before generating code.
+The fact that we only staged the data but not the monadic computation or monadic
+values, is the reason that we can peel of the monad stack in the generated code.
+Now the @AnsM[_]@ is instantiated as the same structure as before, but using the
+@Rep@ versions.
 
@@
 \end{lstlisting}
+
+TODOTODO====
 
@@
 %\todo{highlight the changes in the code}
-
-\begin{lstlisting}
-  case class RepStateT[M[_]: RepMonad, S, A](run: Rep[S] => M[(A, S)]) {
-    def flatMap[B](f: Rep[A] => StateT[M, S, B]): StateT[M, S, B] = ...
-  }
-\end{lstlisting}
 
@@
 
-\paragraph{Code Generation}
+\subsection{A Little Bit Code Generation}
 
@@
 wrappers of the underlying staged data. We define these IR nodes using
-@case class@ extending from @Def[T]@, meaning that they are next-staged definitions of
-type @T@.
+@case class@ extending from @Def[T]@, meaning that they are next-staged definitions of type @T@.
 
 \begin{lstlisting}
-  case class IRCompiledClo(f: Rep[((Value,Store)) => (Value, Store)], λ: Lam, ρ: Rep[Env]) extends Def[Value]
-  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[(Value, Store)]
+  type ValSt = (Value, Store)
+  case class IRCompiledClo(f: Rep[(ValSt) => ValSt], λ: Lam, ρ: Rep[Env]) extends Def[Value]
+  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[ValSt]
 \end{lstlisting}

### 2019-08-15 paper/20_prelim.tex
misc
update

@@
 @for@-comprehension syntax.
-
 \begin{lstlisting}
@@
 bindings in the comprehension are all translated into calls of @flatMap@.
-
 \begin{lstlisting}
@@
 and a few of monads and monad transformers such as @ReaderT@,
-@StateT@, and @ListT@ to write our interpreters. The implementation
+@StateT@, and @ListT/SetT@ to write our interpreters. The implementation
 of monads and monad transformers essentially borrows the ground-truth
@@
 declared as abstract types.
-
 \begin{lstlisting}
@@
 wrapped by @R[_]@. The other @alloc@ is simply the monadic version of the addresses.
-
 \begin{lstlisting}
@@
 values when acting as compilers.
-
 \begin{lstlisting}
@@
 an abstract type.
-
-%\vspace{-1em}
-
 \begin{lstlisting}
@@
 fixed-point combinator to close the open-recursive function @ev@.
-
 \begin{lstlisting}
@@
 is simply an identity monad.
-
 \begin{lstlisting}
@@
 produces a monadic value. In the following code, we elide operations other than @flatMap@.
-
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations}
-Other primitive operations operates on the value domain can be implemented
+Other primitive operations over the value domain can be implemented
 straightforwardly. We elide most of them but describe a bit on how we handle
@@
 defunctionalized representation of closures.
-
 \begin{lstlisting}
@@
 store.
-
 \begin{lstlisting}
@@
 given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$.
-
 \begin{lstlisting}

### 2019-08-15 paper/30_stageconc.tex
update

@@
 that computes $b^x$, which is usually implemented as a recursive function:
-
 \begin{lstlisting}
@@
 @snippet@ method give the currently know value @x@ (5 in this example).
-
 \begin{lstlisting}
@@
 monadic computation like @R => M[A]@ are also not staged. Instead, the internal
-data that these monads operate on are staged. The following code snippet shows
-the idea (we use light gray to highlight what have been changed to staged
-version except naming):
-
+data that these monads operate on are staged, i.e., @R@ and @A@ in the
+reader monad, @S@ and @A@ in the state monad. The following code
+snippet shows the idea. We use \hl{light gray} to highlight where @Rep@ type is added:
 \begin{lstlisting}[escapechar=!]
@@
 that all the invocation of @flatMap@s can be completed before generating code.
-The fact that we only staged the data but not the monadic computation or monadic
+The fact that we only stage data but not monadic computation or monadic
 values, is the reason that we can peel of the monad stack in the generated code.
 Now the @AnsM[_]@ is instantiated as the same structure as before, but using the
-@Rep@ versions.
-
+@Rep@ versions:
 \begin{lstlisting}
@@
 
-TODOTODO====
-
-We use @RepStateT@ as an example to give a sketch of how the @Rep@ versions of monad
-transformers look like. The generic type @M[_]@ is a @RepMonad@ and used as the
-inner monad, which means the pair @(A,S)@ is also staged when appears inside of the monad
-@M[_]@. The @flatMap@ is similar to the unstaged one, but the function @f@ takes
-a @Rep[A]@ value. Again, @f@ is not a next-stage function, but a current-stage
-function that transforms a next-stage value of @Rep[A]@ to @StateT[M,S,B]@.
-%\todo{highlight the changes in the code}
-
 Readers may notice that the conversion between unstaged monads and staged monads
-is simply changing the unstaged data to staged data, which in fact can be
+is merely changing the type of unstaged data to staged data, which in fact can be
 achieved without modifying the implementation of monads. This is true so far, but
-as we will see, it is not that easy when introducing the nondeterminism monad
-(@ListT@) for abstract interpreters, as the whole list will be a next-stage
-value instead of merely the elements in the list. For this reason, we explicitly
-distinguish and present the two versions of monads, as a more consistent
-approach.
+as we will see, it is not straightfoward when for nondeterminism monad
+(@SetT@) in abstract interpreters. Because not only the elements in the set are staged,
+but the whole set is also staged, and we have no knowledge about how many elements
+in the set. In this section, we explicitly distinguish the two versions of monadic interfaces,
+later we will instantiated staged set monad using a different strategy.
 
 \paragraph{Primitive Operations} Most of the primitive operations can be easily
-translated to their staged versions -- we just need to let them works on staged data.
-As we mentioned before, we elaborate on how to handle functions and applications in
-detail. One of the desired goals is to eliminate the monadic layers through
-staging, this becomes a little bit tricky for closures because we actually need 
- to compile the lambda term instead of generating a @CloV@ value, but the @ev(e)@
-returns a monad and prevents the specialization. To make this happen, when
-closing a lambda term, we need to specialize the body expression by
-\textit{collpasing} the @AnsM@ monads to normal values. The
-following function @close@ illustrates the idea: @f@ is a current stage function
-but takes two next-stage values @v@ and $\sigma$, inside of which, we eagerly
-collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
-@Rep[(Value, Store)]@. Then we generate the next-stage value of the compiled
-closure using @emit_compiled_clo@, containing the compiled closure.
+translated to their staged versions -- we just need to change the types.
+As we mentioned before, we will take more care on how functions and
+applications are handled. The problem here is what should we do for
+lambda terms when staging? We cannot create a next-stage
+defunctionalized closure for it, because it means we still need to
+access the interpreter at the next stage. The desired goal is to eliminate the
+interpretation overhead staging, so the right thing is compiling the lambda term
+and its environment to a next-stage Scala function. To do this, we need to
+recursively call the interpreter function, i.e., the @ev@ argument passed to @close@.
+The following code implements the idea:
 
@@
 \end{lstlisting}
+
+We have created a current function @f@ that takes a next-stage value
+and store, and returns staged value. Then we use @emit_complied_clo@
+to delay @f@ to a next Scala function, represented by the type
+@Rep[Value]@. Inside of @f@, we
+
+To make this happen, when
+closing a lambda term, we need to specialize the body expression by
+\textit{collpasing} the @AnsM@ monads to normal values. The
+following function @close@ illustrates the idea: @f@ is a current stage function
+but takes two next-stage values @v@ and $\sigma$, inside of which, we eagerly
+collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
+@Rep[(Value, Store)]@. Then we generate the next-stage value of the compiled
+closure using @emit_compiled_clo@, containing the compiled closure.
 

### 2019-08-15 paper/30_stageconc.tex
revision

@@
 Readers may notice that the conversion between unstaged monads and staged monads
-is merely changing the type of unstaged data to staged data, which in fact can be
-achieved without modifying the implementation of monads. This is true so far, but
-as we will see, it is not straightfoward when for nondeterminism monad
-(@SetT@) in abstract interpreters. Because not only the elements in the set are staged,
-but the whole set is also staged, and we have no knowledge about how many elements
-in the set. In this section, we explicitly distinguish the two versions of monadic interfaces,
-later we will instantiated staged set monad using a different strategy.
+is merely changing the type of unstaged data to staged data, which in fact can
+be achieved without modifying the implementation of monads. This is true so far,
+but as we will see, it is not straightfoward when for nondeterminism monad
+(@SetT@) in abstract interpreters. Because not only the elements in the set are
+staged, but the whole set is also staged, and we have no knowledge about how
+many elements in the set. In this section, we explicitly distinguish the two
+versions of monadic interfaces, later we will instantiated staged set monad
+using a different strategy.
 
@@
 \begin{lstlisting}
-  def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[(Value, Store)]): Rep[Value]
+  type ValSt = (Value, Store)
+  def emit_compiled_clo(f: (Rep[Value], Rep[Store]) => Rep[ValSt], λ: Lam, ρ: Exp[Env]): Rep[Value]
   def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
     val Lam(x, e) = λ
-    val f: (Rep[Value], Rep[Store]) => Rep[(Value, Store)] = {
-      case (v: Rep[Value], σ: Rep[Store]) =>
-        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ + (α → v)
-        ev(e)(ρ_*)(σ_*)
-    }; emit_compiled_clo(f)
+    val f: (Rep[Value], Rep[Store]) => Rep[ValSt] = { case (v: Rep[Value], σ: Rep[Store]) =>
+        val α = alloc(σ, x)
+        ev(e)(ρ + (unit(x) → α))(σ + (α → v))
+    }; emit_compiled_clo(f, λ, ρ)
   }
@@
 
-We have created a current function @f@ that takes a next-stage value
-and store, and returns staged value. Then we use @emit_complied_clo@
-to delay @f@ to a next Scala function, represented by the type
-@Rep[Value]@. Inside of @f@, we
+We have created a current-stage function @f@ that takes a next-stage value and
+store, and returns staged value. Then we use @emit_complied_clo@ to delay @f@ to
+a next-stage Scala function, represented by the type @Rep[Value]@. Inside of
+@f@, we can access the evaluator via @ev@, but @ev@ produces a monadic value of
+type @AnsM@, which can only exist at the current stage. To connect the
+current-stage monadic values and future-stage grounded values, the evaluator
+@ev@ is invoked on the body expression @e@ to be specialized w.r.t. @e@, and
+then we \textit{collpase} the @AnsM@ monad to grounded values by providing the
+prepared environment and store.
 
-To make this happen, when
-closing a lambda term, we need to specialize the body expression by
-\textit{collpasing} the @AnsM@ monads to normal values. The
-following function @close@ illustrates the idea: @f@ is a current stage function
-but takes two next-stage values @v@ and $\sigma$, inside of which, we eagerly
-collapse the monads to values by @ev(e)(ρ_*)(σ_*)@, which has type
-@Rep[(Value, Store)]@. Then we generate the next-stage value of the compiled
-closure using @emit_compiled_clo@, containing the compiled closure.
-
-When applying a function in @ap_clo@, we also generate the next-stage value
-for closure application through @emit_ap_clo@, assuming that the @rator@ is already an
-IR of compiled closure, @rand@ is an arbitrary stage value, and $\sigma$ is the
-latest store. @emit_ap_clo@ represents the result value of the application, which has type
-@Rep[(Value, Store)]@. Note that @emit_ap_clo@ is an ordinary value from the
-future, and there are no monads in the future, so we need to reify the value and
-store back into the monad stack through @put_store@ and @yield@.
+For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
+and @arg@, but what we can do on these values of @Rep@ type? In fact, @fun@ does
+not have an intensional application operation we can use directly; but we know
+that it is a staged value produced by the @close@ function, which represents a
+next-stage Scala function. Therefore we can just generate a next-stage function
+application, also in Scala. Since @ap_clo@ returns a monadic value of type
+@Ans@, we can still use the monadic operation @get_store@ to obtain the latest
+store object. Then we use @emit_ap_clo@ to generate a current-stage
+representation of the future-stage application result, which is a pair of values
+and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
+replace the store and reify the result to the current-stage monadic value.
 
 \begin{lstlisting}
-  def emit_ap_clo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]): Rep[(Value, Store)]
-  def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
+  def emit_ap_clo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]): Rep[ValSt]
+  def ap_clo(ev: EvalFun)(fun: Rep[Value], arg: Rep[Value]): Ans = for {
     σ  <- get_store
-    val res: Rep[(Value, Store)] = emit_ap_clo(rator, rand, σ)
-    _ <- put_store(res._2)
-  } yield res._1
+    vs <- lift[ValSt](emit_ap_clo(fun, arg, σ))
+    _  <- put_store(vs._2)
+  } yield vs._1
 \end{lstlisting}
+
+Compared with the unstaged version, we can observe a shift of using the
+evaluation function @ev@ from application-time to value-representation-time: in
+the unstaged interpreter, the invoke of @ev@ happens in @ap_clo@; while in the
+staged interpreter, we eagerly apply @ev@ when denotating the lambda term to our
+value domain.
 
@@
 
-The method @emit_compiled_clo@ and @emit_ap_clo@ mentioned above produces
-intermediate representations for the denoted operations in LMS. These IRs are
-wrappers of the underlying staged data. We define these IR nodes using
-@case class@ extending from @Def[T]@, meaning that they are next-staged definitions of type @T@.
+In the previous section, we merely show the type of @emit_compiled_clo@ and
+@emit_ap_clo@ without a concrete implemtation. In this section, we brifely
+discuss the IR node generated by them and sketch code generation.
+We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
+extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
+representing next-stage definitions of type @T@.
 
 \begin{lstlisting}
-  type ValSt = (Value, Store)
   case class IRCompiledClo(f: Rep[(ValSt) => ValSt], λ: Lam, ρ: Rep[Env]) extends Def[Value]
@@
 
-After constructing the IR graph, LMS provides an @emitNode@ to generate code for
-each kind of IR node. To generate code for @IRApClo@, we know that @rator@ is a
-compiled closure @IRCompiledClo@, which contains a next-stage function @f@, so
-we just need to generate code that invokes @f@ with @rand@ and @σ@ as arguments
-provided. Similarly, we directly generate a next-stage value @CompiledClo(f)@
-for @IRCompiledClo@.
+The IR nodes are manipulated by the LMS passes and finally generated to
+next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
+programmers to control what is generated for each kind of node. Here we have
+matched @IRCompiledClo@ and @IRApClo@ and will generate definitions for them.
+For compiled closures, we put the function @f@ and its companying syntactic term
+and environment to its next-stage value representation @CompiledClo@. For
+applications, we have the knowledge that @fun@ is a @CompiledClo@ at the stage,
+and it has a callable field @f@, so we generate a next-stage function
+application @f(arg, σ)@. The @quote@s are invocations to generate code for the
+other next-stage values.
 
@@
 override def emitNode(sym: Sym[Any], rhs: Def[Any]) = rhs match {
-    case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar$rand, $\textdollar$σ)")
-    case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar$f, $\textdollar$λ, $\textdollar$ρ)")
+  case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
+  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
   ...

### 2019-08-15 paper/40_absinterp.tex
paper
revision

@@
 After seeing both unstaged and staged concrete interpreters, now we turn our
-focus on abstract interpreters under the same framework. We will first present a
-lattice framework with binding-time abstraction, but for the purpose of clearly
-illustrating the idea, we will just use simple abstract domains such as power set
-lattice or point-wise map lattice. Then we show the abstract components, namely
-abstract values, abstract environments, and abstract stores.
-The abstract interpreter we present in this section is similar to
-\citet{DBLP:journals/pacmpl/DaraisLNH17}'s, it is also
-context/path/flow-insensitive by our choice.  In Section \ref{cfa}, we will
-instantiate a commonly-used context-sensitive analysis;
+focus on abstract interpreters under the same framework. We first present a
+lattice framework with binding-time abstraction, as well as simple abstract
+domains such as pwoer set lattice.
+Then we show the abstract components, namely abstract values, abstract
+environments, and abstract stores. The abstract interpreter we construct in this
+section is similar to \citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For
+simplicity, it is also context/path/flow-insensitive by our choice. In Section
+\ref{cfa}, we will instantiate a commonly-used context-sensitive analysis;
 \citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
@@
 
+Abstract interpreters run on abstract domains, such abstract domains usually can
+be presented by a lattice.
 A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
-\bot, \sqcup, \sqcap \rangle$, where $\sqsubseteq : S \to S \to Bool$ is the
+\bot, \sqcup, \sqcap \rangle$, where $\top$ is the top element, $\bot$ is the
+bottom element, $\sqsubseteq : S \to S \to Bool$ is the
 ordering relation, $\sqcup: S \to S \to S$ is the join (least upper bound)
 operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
-operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines the type
+operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines a type
 class of stage polymorphic lattices. Similar to @MonadOps@ in Section
 \ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
-annotate the binding-times, where @R[_]@ wraps @S@ and @Boolean@ in these
-operations.
+annotate the binding-times, where @R[_]@ wraps datatype @S@ and @Boolean@ in
+these operations.
 
@@
 def SetLattice[T]: Lattice[Set[T]] = new Lattice[Set[T]] {
+  lazy val ⊤: Set[T] = error("No representation for ⊤")
   lazy val ⊥: Set[T] = Set[T]()
-  lazy val ⊤: Set[T] = error("No representation for ⊤")
   def ⊑(l1: Set[T], l2: Set[T]): Boolean = l1 subsetOf  l2
@@
 
-In this section, we simply let @R[T] = T@ (i.e., @NoRep@) to instantiate the
-unstaged lattice type class. A simple instance of lattice we used is the power
-set abstract domain (shown in Figure \ref{fig:powerset}), where the ordering is subset relation, the join is union
-operation, and the meet is the intersect operation.
-Other lattices used in the remainder of the paper, such as products and maps, can be
+In this section, we simply let the binding-time type @R[T] = T@ (i.e., @NoRep@)
+to instantiate an unstaged lattice type class. An example of lattice we used in
+the rest of the paper is the power set abstract domain (shown in Figure
+\ref{fig:powerset}), where the ordering is subset relation, the join is union
+operation, and the meet is the intersect operation. The bottom element of a
+power set is empty set, and we don't have a top representation for power set.
+Other lattices used in the the paper, such as products and maps, can be
 implemented similarly or lifted from existing lattices element-wise or
-point-wise. Naturally, abstract domains such intervals also can be implemented
-in a stage polymorphic way.
+point-wise. Non-relational abstract domains such intervals also can be
+implemented in a stage polymorphic way.
 
@@
 
+We follow the \textit{abstracting definitional interpreters} (ADI) approach
+\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
+an abstract interpreter.
+% \cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12}
+
 \paragraph{Abstract Components}
 
-% \cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12}
-
-We follow the \textit{abstracting definitional interpreters} (ADIs) approach
-\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
-an abstract interpreter. We first widen the concrete values to @AbsValue@: numbers
-are lifted to a singleton object @IntTop@ representing the set of all numbers;
-closures remain the same. Then the type @Value@ is sets of @AbsValue@ (we abuse
-the type names a little bit, as the generic interface defined the abstract type
-@Value@).
+We first widen the concrete values to @AbsValue@: numbers are lifted to a
+singleton object @IntTop@ representing the set of all numbers; closures remain
+the same. Then the type @Value@ is a set of @AbsValue@ (we abuse the type names
+a bit, as the generic interface already defined the abstract type @Value@).
 The address space is constrained to be finite to ensure the computability of
 analysis -- for the simplest monovariant analysis, we use variable names for
-addresses. After settling the @Value@ and @Addr@ type, the environments and
+addresses. After defining the @Value@ and @Addr@ type, the environments and
 stores are automatically lifted to abstract environments and abstract stores,
@@
 It is important that the abstract store now maps addresses to sets of abstract
-values, meaning an address may point to an over-approximated set of values
-occurred at runtime.
+values, indicating that an address may point to an over-approximated set of
+values occurred at runtime.
 
@@
 
-To prevent the abstract interpreter from falling into non-termination, the ADI
-approach uses a two-cache mechanism. Here we first provide the type
-definitions, and later we will describe the algorithm in detail. A @Cache@ is a
-mapping from configurations @Config@ to sets of value-store pairs, where the
-configuration is a triple of current expression being evaluated, environment and
-store.
+To prevent the abstract interpreter from falling into non-termination when
+computing the fixed-point, the ADI approach uses a co-inductive mechanism
+consisting of two caches that remember the input and output of functions.
+Here we first provide the necessary type definitions, and later we will describe
+the algorithm in detail. A @Cache@ is a mapping from configurations @Config@ to
+sets of value-store pairs, where the configuration is a triple of current
+expression being evaluated, environment and store. Intuitively, a cache memoizes
+the corresponding values and stores for a given program configuration.
 
@@
 
-As of now, the store is a mapping from addresses to sets of abstract values.
-Therefore, when applying a function, for instance, @f(a)@, we may retrieve a set
-of closures by dereferencing the address of @f@. We need to nondeterministically
-apply all of the target closures.
-Another nondeterminism happens when evaluating conditionals, we have to evaluate
-both branches if the condition can not be determined. In other words, we will
-have multiple computing results.
+Now, the store is a mapping from addresses to sets of abstract values. This can
+be justied by the approximation nature of abstract interpretation and
+nondeterminism in the abstract interpreter.
+First, when analyzing a conditional experssion, we may not have enough
+information on which branch will be taken, thus a sound treatment is to
+nondeterministically explore both branches, i.e., compute the join of two path.
+Second, when applying a function, for instance, @f(a)@, the possible lambda
+terms of @f@ is computed approximately. As a result, we may retrieve a set of
+closures when dereferencing the address of @f@ and need to apply all of them.
+
+\paragraph{Monads for Abstract Interpretation} Compared with the concrete
+interpreter that uses reader and state effects, the abstract interpreter
+further introduces a nondeterminism effect and another reader and state effect
+for the two caches. The nondeterminism effect is represented by the
+@Set[M[_], A]@ monad transformer, where @M@ is the inner monad being
+transformed, and @A@ is the type of elements in the set. We use a @ReaderT@ for
+one cache that will not be changed during one fixed-point iteration, and use a
+@StateT@ for another cache that will be constantly updated during analysis.
+\citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permuatation of the
+moand stack for abstract interpretation. The following @AnsM@ type shows the
+monad stack (we use \hl{light gray} to highlight what have been changed from
+the monad stack of concrete interpretation):
+
+\begin{lstlisting}[escapechar=!]
+trait AbstractSemantics extends AbstractComponents {
+  type R[T] = T
+  type AnsM[T] = ReaderT[StateT[!\hl{SetT}![!\hl{ReaderT}![!\hl{StateT}![IdM, !\hl{Cache}!, ?], !\hl{Cache}!, ?], ?], Store, ?], Env, T]
+  ...
+}
+\end{lstlisting}
+
+Similar to the concrete scenario, we sketch the implementation of @flatMap@ for
+@SetT@ transformer and omit the rest. The value @run@ encapsulated by the monad
+is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and @A@ is the element
+type of the set. The @flatMap@ method takes a function that transforms each
+element of type @A@ from @M[Set[A]]@ into a @SetT[M, B]@ value, and then folds
+all the transformed values into a single monadic value of type @Set[M, B]@.
+
+\begin{lstlisting}
+  case class SetT[M[_]: Monad, A](run: M[Set[A]]) {
+    def flatMap[B](f: A => SetT[M, B]): SetT[M, B] =
+      SetT(Monad[M].flatMap(run) { (s: Set[A]) =>
+        s.foldLeft(SetT.empty[M,B])((acc, a) => acc ++ f(a)).run
+      }); ... }
+\end{lstlisting}
+
+Such monad stack scheme leads to a different type of result type. In the concrete
+setting, we only have a reader monad (for environment) and a state monad (for
+store), and the reader effect is not persistent to the final result. Together
+with the value produced by the interpreter, the final result type is a pair of
+@Value@ and @Store@. In the abstract setting, we have a @SetT@ inside of the
+environment monad and store monad, which becomes the container type of the pair
+of values and stores, i.e., @Set[(Value, Store)]@. Also note that the reader and
+state monad for caches live internally of the nondeterminism monad, therefore
+the cache type of state monad becomes adjacent to the set type.
 
@@
 
-This observation leads us to the type definition of @Result@: in a concrete setting,
-the @Result@ is a pair of @Value@ and @Store@, now we use a set of such pairs
-@Set[(Value, Store)]@ to represent multiple results. The results are also linked
-with the resulting cache used in this iteration.
-Note that the abstract binding-time annotation is wrapped to the two objects in
-the @Result@ pair, but not the whole pair. Because we statically know the result
-is a tuple of two elements, but we don't know how many value-store pairs we
-would have, neither the content of the cache.
-
-\paragraph{Monads for Abstract Interpretation} Now we incorporate the
-nondeterminism and caching into the monad stack. For concrete evaluation, we
-already have a reader effect for environments and a state effect for stores. We
-further add three more effects based on it: the nondeterminism effect is
-represented by the @SetT[M[_], A]@ monad transformer, inside of which, we have
-a @ReaderT@ for a @in@ cache and a @StateT@ for an @out@ cache. \todo{highlight
-where changed}
-
-\begin{lstlisting}
-  trait AbstractSemantics extends AbstractComponents {
-    type R[T] = T
-    type AnsM[T] = ReaderT[StateT[SetT[ReaderT[StateT[IdM, Cache, ?], Cache, ?], ?], Store, ?], Env, T]
-    ...
-  }
-\end{lstlisting}
-
-The sketch implementation of the @SetT@ transformer is shown below. The value @run@
-encapsulated by the monad is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and
-@A@ is the element type of the list. The @flatMap@ method first unwraps the inner
-monadic value @M[SetT[A]]@ and obtain a value of @Set[A]@, then applies @f@ to
-each element in that list, and finally folds the results back to a single
-@SetT[M, B]@.
-
-\begin{lstlisting}
-  case class SetT[M[_]: Monad, A](run: M[Set[A]]) {
-    def flatMap[B](f: A => SetT[M, B]): SetT[M, B] = ...
-  }
-\end{lstlisting}
-
-\paragraph{Primitive Operations} Some of the primitive operations are changed
-according to the new monad stack. A notable change is that the store update
-operation joins the value with the existing values, for the same address. Under
-abstract interpretation, we may have to explore the results from both branches
-of a conditional expression for soundness. In @br0@, we combine the results
-using @mplus@ from @MonadPlus@, which requires the values are join-semilattices
-(in our case @Set@s and @Map@s).
+\paragraph{Primitive Operations} The primitive operations are changed to adapt
+the new monad stack scheme and value domains. A notable change is that the store
+update operation joins the value with the existing values, for the same address.
+As what mentioned before, we may have to explore both branches for conditionals:
+in @br0@ method, we combine the results using @mplus@ from @MonadPlus@, which
+requires the value domains are join-semilattices (in our case @Set@s and
+@Map@s).
 
@@
 
-The value representation of lambda terms is still @CloV@, but we lift it to a
-singleton set.
+The value representation of lambda terms is still a defunctionalized closure
+@CloV@, but we lift it to a singleton set domain.
 
@@
 
-For function applications, @ap_clo@ looks roughly the same as the concrete
-interpreter. The difference is now the function @rator@ is a set of closures, so
-we can not directly destruct it by pattern matching. Instead, we use an
-auxiliary function @lift_nd@ that takes a set and lifts the elements into the
-monad stack. Then we can still implement @ap_clo@ in monadic style, where the
-@CloV@ comes from the @SetT@ and the nondeterminism is naturally handled.
+For function applications, @ap_clo@ looks the same as the concrete interpreter.
+The difference is now the first argument @rator@ is a set of closures (remember
+that @Value@ is now an alias of @Set[AbsValue]@), so we can not directly
+destruct it by pattern matching. Instead, we use an auxiliary function @lift_nd@
+that takes a set and lifts the elements into the monad stack. Then we can still
+implement @ap_clo@ in monadic style, where the closures come from the @SetT@ and
+thus the nondeterminism can be naturally handled. The light gray code shows what
+is added from the concrete @ap_clo@.
 
-\begin{lstlisting}
+\begin{lstlisting}[escapechar=!]
   def lift_nd[T](vs: Set[T]): AnsM[T]
   def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
-    CloV(Lam(x, e), ρ) <- lift_nd[AbsValue](rator)
+    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](rator)}!
     α <- alloc(x)
@@
 \textit{co-inductive} caching algorithm or \textit{truncated depth-first
-evaluation} \cite{Rosendahl:AbsIntPL}. It has been widely used in abstract
-interpretation or fixed point computation
+evaluation} \cite{Rosendahl:AbsIntPL}. It has been used in other abstract
+interpretater or fixed point computation
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
-Rosendahl:AbsIntPL}. Since the @in@ and @out@ cache are wrapped in the monad
-stack, we first defined several methods to obtain or update the @in@ and @out@
-cache through the monad stack.
+  Rosendahl:AbsIntPL}. The idea is to use a @in@ cache and a @out@ cache during
+the depth-first evaluation, where the @in@ cache stores the result from last
+iteration, and @out@ cache is used for updates in current iteration. In the next
+iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
+is plugged-in to the @out@ slot. Once the @out@ cache does not contain anything
+new than the @in@ cache, the fixed-point is reached.
+
+In out monad stack, the @in@ and @out@ cache modeled by a reader effect and
+state effect, respectively. We first defined several methods to manipulate
+the two caches through the monad stack; the implementations are elided.
 
@@
 
-Figure \ref{fig:coind_cache} shows the co-inductive caching algorithm that
-closes the open-recursive @eval@.
-The idea is to first set up two caches called @in@ and @out@ both of type
-@Cache@. Then similar to Kleene's iteration algorithm, we compute the analysis
-result iteratively. The @in@ cache contains what we already have computed from
-the last iteration, and the @out@ cache is what we would have computed after this
-iteration joined with the contents from the @in@ cache.
-When starting a new iteration, @in@ is set to be the result of the last
-iteration, i.e., @out@; and @out@ is assigned to be empty.
-
+The co-inductive caching algorithm is implemented as a instrumentation over
+@eval@ (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
+Initially, the two caches are both empty.
 During the iteration, we first check whether @out@ contains the configuration
-@cfg@, representing the current desired computation, if
-yes, then the result is directly returned through the monads.
-
+@cfg@, representing the current desired computation. If it does, then the result
+is directly returned through the monads.
 Otherwise, we first retrieve the results from @in@ (@⊥@ if not contained in
-@in@), and put it into the @out@ cache.
-Then we invoke @ev@ to compute the result for this iteration.
-After the evaluation, the store @σ@ may have been changed. So we update the
-@out@ cache with the new result, i.e., @(v, σ)@. Similar to store update, cache
-update is also point-wise join.
-Note that for the recursive call of @ev@, we also instrument it by putting
-@fix(ev)@ as the first argument to @ev@.
+@in@), and join it into the @out@ cache.
+Then we invoke @ev@ to compute the result for this iteration, where @ev@ takes
+@fix(ev)@ as a self reference.
+After the evaluation, the store @σ@ may have been changed. We update the
+@out@ cache with the new result, i.e., a pair of value and store @(v, σ)@.
+Similar to store update, cache update is also a point-wise join.
 
 The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which means there is no more analysis result can be gained,
-thus the iteration should end and we have reached the least fixed point.
+input @in@ cache, which indicates that there is no more fact can be gained, thus
+the iteration should end and we have reached the least fixed point.
 
-\begin{figure}[t!]
+\begin{figure}[h!]
   \centering
@@
     ρ <- ask_env; σ <- get_store; in <- ask_in_cache; out <- get_out_cache; val cfg = (e, ρ, σ)
-    rt <- if (out.contains(cfg)) for {
+    rt <- if (out.contains(cfg)) for { // ask if out already contains the desired result
             (v, σ) <- lift_nd[(Value, Store)](out(cfg))
@@
 
-Finally, we override the top-level @run@ method by running the monad with
-the empty abstract environment $\rho_0$, initial store $\sigma_0$, and initial
-caches @cache@$_0$.
+Finally, we override the top-level @run@ method by running the monadic value
+with the empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
+and initial caches @cache@$_0$.

### 2019-08-15 paper/50_stageabs.tex
revision

@@
 The approach from unstaged to staged abstract interpreter is modular,
-principled, and does not sacrifice soundness or precision. The designer of the
-analyzer has no need to rewrite the analysis. We first present the staged
-lattices and staged monads, and then the staged version of primitive operations,
-especially @close@, @ap_clo@ and @fix@, at last, we discuss several optimizations
-and our design.
+and does not sacrifice soundness or precision. The designer of the
+analyzer therefore has no need to rewrite the analysis. We first present the
+staged lattices and staged monads, and then the staged version of primitive
+operations, especially @close@, @ap_clo@ and @fix@. At last, we discuss several
+optimizations and summarize our approach.
 
@@
 
-\begin{lstlisting}
+\begin{lstlisting}[escapechar=!]
   trait RepLattice[S] extends Lattice[S, Rep]
   def RepSetLattice[T]: RepLattice[Set[T]] = new RepLattice[Set[T]] {
-    lazy val ⊥: Rep[Set[T]] = Set[T]()
-    lazy val ⊤: Set[T] = error("No representation for ⊤")
-    def ⊑(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Boolean] = l1 subsetOf  l2
-    def ⊔(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Set[T]]  = l1 union     l2
-    def ⊓(l1: Rep[Set[T]], l2: Rep[Set[T]]): Rep[Set[T]]  = l1 intersect l2
+    lazy val ⊥: !\hl{Rep[Set[T]]}! = Set[T]()
+    lazy val ⊤: !\hl{Rep[Set[T]]}! = error("No representation for ⊤")
+    def ⊑(l1: !\hl{Rep[Set[T]]}!, l2: !\hl{Rep[Set[T]]}!): !\hl{Rep[Boolean]}! = l1 subsetOf  l2
+    def ⊔(l1: !\hl{Rep[Set[T]]}!, l2: !\hl{Rep[Set[T]]}!): !\hl{Rep[Set[T]]}!  = l1 union     l2
+    def ⊓(l1: !\hl{Rep[Set[T]]}!, l2: !\hl{Rep[Set[T]]}!): !\hl{Rep[Set[T]]}!  = l1 intersect l2
   }
@@
 
-As we can see, the @RepSetLattice@ is an instance of @RepLattice@, which is also
-an abstraction of the underlying data structure. The lattice operations
-eventually are delegated to @subsetOf@, @union@ and @intersect@ on
-@Rep[Set[T]]@. We even don't have to change the implementation code, but just
-the types – from @Set[T]@ to @Rep[Set[T]]@. LMS library provides an implicit
-conversion lifting to next-stage values. In the code generation part, these
+The trait @RepLattice@ is obtained by specializing @Lattice@ with @Rep@.
+The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
+eventually are delegated to operations on a staged set data type, including
+@subsetOf@, @union@ and @intersect@. We even do not have to change the
+implementation code, but just the types -- from @Set[T]@ to @Rep[Set[T]]@, shown
+in light gray code. The LMS library provides an implicit conversion that lifts a
+current-stage set value to next-stage values. In the code generation part, these
 operations emit their corresponding next-stage code. Again, other lattice
@@
 
-Now we implement the staged abstract semantics, where @R@ is instantiated as
-@Rep@, and the abstract components are shared from the unstaged version.
+Now we can implement the staged abstract semantics, where @R@ is instantiated as
+@Rep@, and the abstract components are reused from the unstaged version.
 
 \paragraph{Staged Monads for Abstract Interpretation}
-We use the same monad stack structure as in the unstaged abstract interpreter,
+We use the same monad stack structure from the unstaged abstract interpreter,
 but replacing \textit{all} the transformers to their staged version. The
-following code shows this change. For the readability, we squeeze the three
-inner transformers into a single monad @RepSetReaderStateM[A, B, C]@, where @A@
-and @B@ are both @Cache@ when used in @AnsM@. Now the result type is a pair of
-two staged value: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
+following code shows this change. For readability, we squeeze the three inner
+transformers (@IdM@ omitted) into a single monad @RepSetReaderStateM[A, B, C]@,
+where @A@ is the type parameter for reader effect and @B@ for the state effects.
+In our monad stack scheme, they are both instantiated with @Cache@. Now the
+result type is a pair of two staged value: @Rep[Set[(Value, Store)]]@ and
+@Rep[Cache]@.
 
@@
     type R[T] = Rep[T]
-    type RepSetReaderStateM[A, B, C] = RepSetT[RepReaderT[RepStateT[RepIdM, A, ?], B, ?], C]
     type AnsM[T] = RepReaderT[RepStateT[RepSetReaderStateM[Cache, Cache, ?], Store, ?], Env, T]
@@
 
-Since our monad stack is uniformly using staged data, so the staged @SetT@ now
+\iffalse
+Since our monad stack is uniformly using staged data, the staged @SetT@ now
 stores a staged value of type @Rep[Set[A]]@, inside of another staged monad @M@.
@@
 \end{lstlisting}
+\fi
 
-\paragraph{Primitive Operations} We keep our road map that focuses on functions
-and applications. The @close@ method now is a rough mixing of the staged
-concrete and unstaged abstract version: we have a current-stage function @f@,
-that takes four next-stage values, @in@ and @out@ additionally; inside of @f@,
-we collapse the @Ans@ monad to values of type @Result@ by providing the desired
-arguments, i.e., the new environment, new store, and caches. The collapse
-happens at current-stage, so the call of @ev@ is unfolded after staging. At
-last, we generate a singleton set containing the compiled closure
+\paragraph{Primitive Operations} When deriving the staged concrete interpreter
+from the unstaged counterpart, we notice that recursive calls of @ev@ on the
+lambda body are shifted from @ap_clo@ to @close@. In other word, we need to
+eagerly specialize the interpreter and generate code every time when we see a
+$\lambda$ term, instead of lazily when we actually need to apply a function.
+With no surprise, it is similar in the staged abstract interpreter,
+additionally, we need to handle nondeterminism. So we keep our track and focus
+on discussing closure representations and function applications for the staged
+version.
+
+The @close@ method now is a mixing of the \textit{staged} concrete version and
+unstaged \textit{abstract} version: we build a current-stage function @f@, which
+takes four next-stage values, including the argument and latest store, and
+@in/out@ caches additionally. Inside of @f@, we collapse the @Ans@ monad to
+grounded values of type @Result@ by providing the desired arguments, i.e., the
+new environment, new store, and caches. The collapsing of monad happens at
+current-stage, so the call of @ev@ is unfolded through staging. Finally, we
+generate a singleton set containing the compiled next-stage closure
 @emit_compiled_clo(f)@, represented by an IR node in LMS.
@@
   def emit_compiled_clo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
-                           => Rep[(Set[(Value,Store)], Cache)]): Rep[AbsValue]
+                           => Rep[(Set[(Value,Store)], Cache)], λ: Lam, ρ: Exp[Env]): Rep[AbsValue]
   def close(ev: Expr => Ans)(λ: Lam, ρ: Rep[Env]): Rep[Value] = {
@@
       case (arg, σ, in, out) =>
-        val α = alloc(σ, x); val ρ_* = ρ + (unit(x) → α); val σ_* = σ ⊔ Map(α → arg)
-        ev(e)(ρ_*)(σ_*)(in)(out)
-    }; Set[AbsValue](emit_compiled_clo(f))
+        val α = alloc(σ, x)
+        ev(e)(ρ + (unit(x) → α))(σ ⊔ Map(α → arg))(in)(out)
+    }; Set[AbsValue](emit_compiled_clo(f, λ, ρ))
   }
@@
 The @ap_clo@ method is also similar: we use the staged version of @lift_nd@ to
-lift the set of closures into the monad stack, and generate a next-stage value
-using @emit_ap_clo@, which represents the result of application from future.
-Finally, we put the future @out@ cache, store, and values back into the
-current-stage monads. The @emit_ap_clo@ method takes the target closure @clo@,
-argument @rand@, store, and caches.
+lift the next-stage set of closures into the monad stack, and generate a
+next-stage value for each closure @clo@ in the set, by using @emit_ap_clo@. The
+method @emit_ap_clo@ generates the current-stage representation of the
+future-stage result of application.
+Finally, we put the future @out@ cache, store, and values back into
+the current-stage monads.
 
 \begin{lstlisting}
-  def lift_nd[T](vs: Rep[Set[T]]): AnsM[T]
   def emit_ap_clo(rator: Rep[AbsValue], rand: Rep[Value], σ: Rep[Store],
-                  in: Rep[Cache], out: Rep[Cache]): Rep[(Set[(Value, Store)], Cache)]
+                  in: Rep[Cache], out: Rep[Cache]): Rep[(Set[ValSt], Cache)]
   def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
@@
     in  <- ask_in_cache;   out <- get_out_cache
-    val res: Rep[(Set[(Value, Store)], Cache)] = emit_ap_clo(clo, rand, σ, in, out)
+    res <- lift_nd[(Set[ValSt], Cache)](Set(emit_ap_clo(clo, arg, σ, in, out)))
     _   <- put_out_cache(res._2)
-    vs <- lift_nd[(Value, Store)](res._1)
+    vs  <- lift_nd[ValSt](res._1)
     _   <- put_store(vs._2)
@@
 
-\paragraph{Staged Caching and Fixpoint Iteration}
+% Monad is the bridge that connects future-stage values
 
-Our fixed-point iteration again relies on the two caches @in@ and @out@, which are
-both staged values now. Therefore, the query of whether the @out@ cache contains
-the current configuration produces a next-stage Boolean value, i.e.,
-@Rep[Boolean]@. Consequently, the branching cannot be determined statically --
-we need to generate code for @if@. Figure \ref{fig:staged_coind_cache} shows the
-staged version of @fix@.
-
-\vspace{-1em}
 \begin{figure}[h!]
@@
 \begin{lstlisting}
-  def fix(ev: (Expr => Ans) => (Expr => Ans)): Expr => Ans = e => for {
+  def fix_cache(e: Expr): Ans = for {
     ρ   <- ask_env;  σ <- get_store;  in <- ask_in_cache;  out <- get_out_cache
-    val cfg: Rep[Config] = (unit(e), ρ, σ)
-    val res: Rep[(Set[(Value, Store)], Cache)] =
-      if (out.contains(cfg)) (out(cfg), out) // a next-stage if expression
+    cfg <- lift_nd[Config](Set((unit(e), ρ, σ)))
+    res <- lift_nd[(Set[ValSt], Cache)](Set(
+      // a next-stage value of type Rep[(Set[ValSt], Cache)] generated by if
+      if (out.contains(cfg)) (repMapToMapOps(out).apply(cfg).toList, out)
       else { val m: Ans = for {
                _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
-               v <- ev(fix(ev))(e)
+               v <- eval(fix_cache)(e)
                σ <- get_store
@@
              } yield v
-             m(ρ)(σ)(in)(out) }
+             m(ρ)(σ)(in)(out) }))
     _ <- put_out_cache(res._2);  vs <- lift_nd(res._1);  _ <- put_store(vs._2)
@@
 \end{lstlisting}
-\vspace{-1em}
 \caption{The staged co-inductive caching algorithm.}
@@
 \end{figure}
-\vspace{-1em}
 
-The @res@ variable is a next-stage result, which consists of a next-stage @if@
-expression. The true branch simply returns a pair of queried result form the
-@out@ cache and the @out@ cache. The else branch constructs a monad @m@ first,
-which evaluates @e@ under the new @out@ cache. After which, we collapse the
-monad @m@ with desired arguments. Finally, we obtain a placeholder @res@ since
-it is a next-stage value, and put the content of @res@ back into the monad
-stack.
+\paragraph{Staged Caching and Fixpoint Iteration} Our fixed-point iteration
+again relies on the two caches @in@ and @out@, which are both staged maps now.
+Therefore, the query of whether the @out@ cache contains the current
+configuration will produce a next-stage Boolean value, i.e., of type
+@Rep[Boolean]@. Consequently, the branching condition cannot be determined
+statically -- we need to generate code for @if@. Figure
+\ref{fig:staged_coind_cache} shows the staged version of @fix@. The
+monadic-binded variable @res@ is a next-stage result, consisting of a next-stage
+@if@ expression. The true branch simply returns a pair of queried result form
+the @out@ cache and the @out@ cache itself. The else branch constructs a monadic
+value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@ cache.
+After which, we use the similar technique that collapses the monad @m@ to
+grounded values by providing its desired environment and etc. Finally, we have a
+current-stage representation of the future-stage values @res@, and put the
+content of @res@ back into the monad stack.
 
-\paragraph{Code Generation} The IR nodes for compiled closures and
-applications are defined as follows:
+\subsection{A Little Bit Code Generation, Again}
+The code generation for compiled closures and function applications are similar
+to the staged concrete interpreter. We have two the IR node as case classes, they
+take additional caches as arguments. We also elide how to generate next-stage Scala
+code for them.
 
@@
 case class IRCompiledClo(f: (Rep[Value], Rep[Store], Rep[Cache], Rep[Cache])
-                           => Rep[(Set[(Value, Store)], Cache)], λ: Lam, ρ: Rep[Env]) extends Def[AbsValue]
+                          => Rep[(Set[ValSt], Cache)], λ: Lam, ρ: Rep[Env]) extends Def[AbsValue]
 case class IRApClo(clo: Rep[AbsValue], arg: Rep[Value], σ: Rep[Store],
@@
 \end{lstlisting}
-
-The code generator for these two IR nodes is similar to staged concrete
-interpreter, which generates a next-stage @CompiledClo@ object and a next-stage
-function call to @f@, respectively.
 
@@
 Our staging schema works by unfolding the interpreter over the input AST. In
-practice, it would suffer from code explosion when analyzing (specializing) large
-programs, which increases the compile time.
-If we generate next stage programs running on JVM, such large generated program
-would also cause runtime GC overhead. In this section, we present several
-optimizations that largely mitigate these issues. For all that, implementing
-these optimizations do not need to change the generic interpreter.
+practice, however, it would suffer from code explosion when analyzing
+(specializing) large programs, which increases the compile time. If we generate
+next stage programs running on JVM, such large generated program would also
+cause runtime GC overhead. In this section, we present several optimizations
+that largely mitigate these issues. For all that, implementing these
+optimizations do not need to change the generic interpreter.
 
@@
 is not necessary and even very expensive when generating code for atomic
-expressions such as literals or variables, as they always terminate. Borrowing
-the idea from the partition of A-Normal Form \cite{Flanagan:1993:ECC:155090.155113},
-we may use a selective caching algorithm:
+expressions such as literals or variables, because they always terminate.
+Borrowing the idea from the partition of expressions used in A-Normal Form
+\cite{Flanagan:1993:ECC:155090.155113}, we may use a selective caching
+algorithm that does not generate caching for atomic expressions:
 
@@
 
-\todo{1. construct generic monadic interpreter that abstract over binding-time,
-  primtive operations, return type.
-  2. implement unstaged abstract interpreter with monads, ie. using approriate
-  monad stack
-  3. if we want to generate code, we just need to change the monad stack.
-  the data wrapped are staged.
-  we show readerT monad, stateT monad, fused monad with nondeterminism. }
+We summarize our approach to staging an abstract interpreter as follows:
+
+\begin{itemize}
+  \item Constructing a generic interpreter that abstracts over binding-time,
+    value domains and primitive operations. In this paper, the generic interpreter
+    is implemented in monadic style, therefore the semantics can be encapsulated
+    into monads.
+  \item By using an approriate moand stack, we can implement a modular, unstaged
+    abstract interpreter. This step has been explored in the previous literature.
+  \item Replacing the monad stack to staged monad stack, such monads operate on
+    staged data types, i.e., next-stage values. We also need to reimplement several
+    primitive operations that on staged values, modularly.
+\end{itemize}
+
+Monadic interperters are known to be able to decouple the interpretation process
+and the underlying semantics. The key insight in this paper is, by making the
+monadic interpreter binding-time polymorphic, it can be extended for generating
+efficient code in a staging framework. The underying semantics and staging are
+two orthogonal dimensions. It is important to note that the computation
+encapsulated by the monads are not staged, only data such as sets and maps are
+staged. All the monadic computation, i.e., functions passed to the monadic bind
+operation @flatMap@ are statically known. This is why we can eliminate the
+monadic layer and its overhead.
 
@@
 
-\todo{may be show an small example?}
+%\todo{may be show an small example?}
 
@@
 As the central concern of static analysis, the soundness is important for
-every user who would like to use it. Here we briefly summarize that how our
+every user who would like to use it. Here we briefly justify that how our
 approach preserve the corretness and soundness of the analysis, under the

### 2019-08-15 paper/60_cases.tex
some number and typos
revision

@@
   case Var(x) => ($\sigma$: Store) => 
-    analysisAbsApp($\sigma$.lookup(x), args, $\sigma$)
+    analyzeAbsApp(args, $\sigma$(x), $\sigma$)
   case Op(_) => compArgs(args)
@@
 obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain a
-family of analysis with increasing precision.
+family of analysis with increasing precision. Note that when $k = 0$, the
+address space is statically determined by the set of variables appeared in the
+program, and we may exploit it to optimize the generated code. But when $k > 0$,
+the address generation happens at the next stage, i.e., analysis time.
 
@@
   type AnsM[T] = RepReaderT[RepStateT[
-                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //add Time into a RepStateT
+                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //add Time with a RepStateT
                  Store, ?], Env, T]
@@
 \end{lstlisting}
-\vspace{-1em}
+\vspace{-2em}
 \caption{\texttt{fix\_gc} for abstract garbage collection} \label{fixgc}

### 2019-08-15 paper/80_related.tex
revision

@@
 
-\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation
-framework for modular analysis. We borrow the notation of modular
-analysis, \todo{but intentional properties are different, compare with
-  summary}
+\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
+modular analysis. We borrow the notation of modular analysis for our approach.
+But the staged abstract interpreter does not generate complete summaries for
+modules, instead, it is able to specialize the analysis modularly. Similar, AC
+is considered as an example of this kind of mdoular analysis \cite{DBLP:conf/cc/CousotC02}.
+
+\paragraph{Control-flow Analysis} Our case studies and evaluation focus on
+control-flow anslysis for functional languages
+\cite{Shivers:1991:SSC:115865.115884, Midtgaard:2012:CAF:2187671.2187672}. In
+general, $k$-CFA (where $k > 0$) is considered impractical to use as a compiler
+pass, due to its EXPTIME lower bound \cite{VanHorn:2008:DKC:1411204.1411243}.
+However, several variants of 0CFA were developed and shown to be useful in
+practice and program optimization \cite{Adams:2011:FTR:2048066.2048105,
+Bergstrom:2014:PEH:2628136.2628153, ashley:practical, Reppy:2006:TCA:1159876.1159888}.
 
@@
   difference?}
-
-\paragraph{Control-flow Analysis} useful applications, why important

### 2019-08-15 paper/90_conc.tex
revision

--- 2019-04-05 paper/90_conc.tex
@@
 
-In this paper, we first present the implementaion of first Futamura projection
-on abstract interpreters. To realize it, we design a generic interpreter that
-abstracts over both binding-time and value domains; by instantiating it
-differently, the staged abstract interpreter can be readily derived. We also
-conduct case studies and empirical evaluation, demonstrating a wide
-applicability of our approach and evaluating the performance improved by staging.
+In this paper, we present a simple and elegant framework that extends the first
+Futamura projection to abstract interpreters.
+To realize it, we design a generic interpreter that abstracts over both
+binding-time and value domains; by instantiating it with staged monads for
+abstract interpretation, the staged abstract interpreter can be readily derived.
+The meta-programming approach brings new insight and implementation techniques
+to optimize static analyzers. The optimization can be realized in a completely
+orthogonal way to the interpreter and abstract semantics.
+In our empirical evaluation, staging improves an order of magnitude of
+performance in terms of running time.

### 2019-08-15 paper/10_intro.tex
revision
Proofreading intro

@@
 
-In recent years, we observe a rich progress on the methodologies for
-constructing abstract interpreters from systematic principles, instead
-of ad-hoc engineering. A notable one is the Abstracting Abstract
+In recent years, we observe rich progress on the methodologies for
+constructing abstract interpreters from systematic principles rather than
+ad-hoc engineering. Of particular note is the Abstracting Abstract
 Machine (AAM) methodology \cite{DBLP:journals/jfp/HornM12,
-  DBLP:conf/icfp/HornM10}. It uncovers an approach to derive sound
+  DBLP:conf/icfp/HornM10}. This methodology uncovers an approach to derive sound
 abstract interpreters from their concrete counterparts, for example,
@@
 analysis \cite{Shivers:1988:CFA:53990.54007,
-  Midtgaard:2012:CAF:2187671.2187672}: first we tweak the environment
+  Midtgaard:2012:CAF:2187671.2187672}: first, tweak the environment
 dereference as a nondeterministic choice, such that the environment
-may contain mutiple possible values for a varaiable, then allocating
+may contain mutiple possible values for a varaiable, then allocate
 continuations in the environment to exploit return-flow information,
@@
 abstraction approach can be tailored to different language features
-(such as states, first-class controls, exceptions and concurrency) and
+(such as states, first-class controls, exceptions, and concurrency) and
 sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
   DBLP:conf/popl/GilrayL0MH16, Darais:2015:GTM:2814270.2814308}. It
-also has been applied to various small-step abstract machines
+has also been applied to various small-step abstract machines
 \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
@@
 pragmatically, several implementation strategies utilizing purely
-functional programming to build abstract interpreters were
+functional programming to build abstract interpreters have
 emerged. Such techniques include monads, monad transformers
@@
 arrows \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Githubsemantic} and etc.  The pure approaches provide certain
-benefits. The abstract interpretation artifacts can be built
+\cite{Githubsemantic}, etc.  The pure approaches provide certain
+benefits, including that the abstract interpretation artifacts can be built
 compositionally and modularly, e.g., by using monad
-transformers. Therefore, the soundness of analysis can be proved with
-less efforts, either by mechanized \cite{Darais:2016:CGC:2951913.2951934}
-or paper-based proof \cite{Keidel:2018:CSP:3243631.3236767}. Also,
+transformers. Therefore, the soundness of analysis can be proven with
+less effort, either by mechanized \cite{Darais:2016:CGC:2951913.2951934}
+or paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}. Also,
 referential transparent and equational reasoning allow programmers to
@@
 the abstract interpreter analyzes the program by traversing the
-abstract syntax tree, which poses an interpretive overhead, such as
-the pattern matching on the ASTs and recursive calls on the sub
-expressions. Those kind of overhead can be negligible if the abstract
-interpreter only runs on the program for one time, but also can be
-accumulated significantly if it runs repeatedly, for example, on
-libraries.  Second, the abstract interpreter written in pure languages
+abstract syntax tree (AST), which poses an interpretive overhead, such as
+the pattern matching on the ASTs and recursive calls on the subexpressions.
+Such overhead can be negligible if the abstract
+interpreter only runs on the program a single time, but can also
+accumulate significantly if run repeatedly (for example, on
+libraries).  Second, the abstract interpreter written in pure languages
 usually extensively uses effect systems to implement the semantics of
@@
 nondeterminism monads can help.  Although such pure approaches have
-its own merits and elegance, compared with imperative stateful
+their own merits and elegance, compared with imperative stateful
 implementations, they are significantly slower.
@@
 In this paper, we propose an abstraction-without-regret approach to
-eliminate those performance penalties for abstract interpreters,
-meanwhile still keeping the benefits come from purely functional
+eliminating performance penalties for abstract interpreters,
+while still keeping the benefits inherent in purely functional
 programming.  In short, our approach applies ideas and techniques from
@@
 4) We also empirically evaluate the performance improved by staging,
-showing an order of magnitude speed-up on flow-analysis \todo{explicit
-  note the number}.
+showing an order of magnitude speed-up on flow-analysis.
 

### 2019-08-15 paper/50_stageabs.tex
Titles and spaces

@@
 
-\subsection{A Little Bit Code Generation, Again}
+\subsection{A Little Bit of Code Generation, Again}
 The code generation for compiled closures and function applications are similar

### 2019-08-15 paper/20_prelim.tex
revision
update

@@
 overriding mechanism in Scala, we may also add new language constructs and reuse
-existing interpretations \todo{cite Bruno?}.
+existing interpretations.
+
+%\todo{cite Bruno?}.
 
@@
 will also replace the monads to the ones that work on staged data values.
-\todo{it is important that R only annotates data, not the monad}
 

### 2019-08-15 paper/50_stageabs.tex
update

@@
 still use power sets as an example to present its staged version.
-
 \begin{lstlisting}[escapechar=!]
@@
 @Rep[Cache]@.
-
 \begin{lstlisting}
@@
 stores a staged value of type @Rep[Set[A]]@, inside of another staged monad @M@.
-
 \begin{lstlisting}
@@
 @emit_compiled_clo(f)@, represented by an IR node in LMS.
-
 \begin{lstlisting}
@@
 the current-stage monads.
-
 \begin{lstlisting}
@@
 code for them.
-
 \begin{lstlisting}
@@
 algorithm that does not generate caching for atomic expressions:
-
 \begin{lstlisting}
@@
     check the correctness of implementation of staged monads modularlly.
-  \item We construct the staged monads and staged data structures in a
-    correct-by-construct way that directly corresponds to their unstaged
+  \item We build the staged monads and staged data structures in a
+    correct-by-construction way that directly corresponds to their unstaged
     version. For instance, the staged data structure we used here are simply
@@
 
-\paragraph{Are Monads Necessary?} No. One can always inline the monads and
-obtain an abstract interpreter in continuation-passing style, or simply use
-explicit side-effects such mutation to implement the artifact with the same
-abstract semantics. In either case, we can still apply the staging schema to
-the abstract interpreter. As evidence, we provide a staged abstract
-interpreter written in direct-style in the accompanying artifact.
+\paragraph{Monadic-style vs Direct-style} We use monadic interpreter through the paper,
+but it is not necessary for staging. One can inline the monadic
+operations and obtain an abstract interpreter in continuation-passing
+style, or even translate back to direct-style that may use explicit
+side-effects such as mutations. In either case, we can still apply the
+staging idea to the abstract interpreter and remove the interpretation
+overhead. Although, monads make the staged abstract interpreter can be
+implemented in a modular and extensible way.
 

### 2019-08-15 paper/80_related.tex
update

@@
   Giacobazzi:2015:APA:2676726.2676987}.  \citet{10.1007/11561347_18}
-discussed combining multi-stage programming with functors and
-monads. But they didn't \todo{monad transfomers?}  Similar to the idea
-in this paper, specializing monadic interpreters was explored by
-\citet{DBLP:conf/dsl/SheardBP99, danvy1991compiling}. \todo{what's the
-  difference?}
+discussed a let-insertion, memoizing monads for code generation, using
+Gaussian Elimination in MetaOCaml as example. In this work,
+let-insertion is handled by the ANF transformation in LMS, and the
+memoizing monad here just combines a reader monad and state monad
+operating on staged data.
+Similar to the idea in this paper, specializing monadic interpreters
+was explored by \citet{danvy1991compiling, DBLP:conf/dsl/SheardBP99},
+but different from the performance motivation here, the generated code
+is in monadic form.

### 2019-08-15 oopsla19-code/src/main/scala/sai/evaluation/TestPrograms.scala

@@ -38,6 +38,14 @@ object TestPrograms {
     (fib 10)
   """)
 
+  def map = getAST("""
+    (define (map xs f)
+      (if (null? xs)
+          (list)
+          (cons (f (car xs)) (map (cdr xs) f))))
+    (map (list 1 2 3) (lambda (x) (+ x 1)))
+    """)
+
   def id4 = App(Lam(List("x"), App(App(Var("x"), List(Var("x"))), List(Var("x")))), List(Lam(List("y"), Var("y"))))
   def oneplusone = App(Var("+"), List(IntLit(1), IntLit(1)))
   def fact5 = getAST("(define (fact n) (if (eq? n 0) 1 (* n (fact (- n 1))))) (fact 5)")

### 2019-08-15 paper/50_stageabs.tex
Proofread 7

--- 2019-08-15 paper/50_stageabs.tex
@@
 concrete interpreters. In this section, we review and summarize our recipe
-to achieve the staged abstract interpreter, discuss the correctness issue
+to achieve the staged abstract interpreter, discuss correctness issues,
 and different design choices.
@@
   \item Constructing a generic interpreter that abstracts over binding-time,
-    value domains and primitive operations. In this paper, the generic interpreter
-    is implemented in monadic style, therefore the semantics can be encapsulated
+    value domains, and primitive operations. In this paper, the generic interpreter
+    is implemented in monadic style; therefore, the semantics can be encapsulated
     into monads.
-  \item By using an approriate moand stack, we can implement a modular, unstaged
+  \item By using an appropriate monad stack, we can implement a modular, unstaged,
     abstract interpreter. This step has been explored in the previous literature.
@@
     staged data types, i.e., next-stage values. We also need to reimplement several
-    primitive operations that on staged values, modularly.
+    primitive operations on staged values, in a modular fashion.
 \end{itemize}
 
-Monadic interperters are known to be able to decouple the interpretation process
-and the underlying semantics. The key insight in this paper is, by making the
+Monadic interpreters are known to be able to decouple the interpretation process
+and the underlying semantics. The key insight in this paper is that by making the
 monadic interpreter binding-time polymorphic, it can be extended for generating
-efficient code in a staging framework. The underying semantics and staging are
+efficient code in a staging framework. The underlying semantics and staging are
 two orthogonal dimensions. It is important to note that the computation
-encapsulated by the monads are not staged, only data such as sets and maps are
+encapsulated by the monads are not staged; only data (such as sets and maps) are
 staged. All the monadic computation, i.e., functions passed to the monadic bind
-operation @flatMap@ are statically known. This is why we can eliminate the
-monadic layer and its overhead.
+operation @flatMap@, are statically known. This is why we can eliminate the
+monadic layer and its associated overhead.
 
@@
 
-%\todo{may be show an small example?}
+%\todo{maybe show a small example?}
 
@@
 
-As the central concern of static analysis, the soundness is important for
-every user who would like to use it. Here we briefly justify that how our
-approach preserve the corretness and soundness of the analysis, under the
+Soundness is the central concern of static analysis, and as such, is vital
+for prospective users. We now briefly examine how our
+approach preserves the correctness and soundness of the analysis, under the
 assumption that the unstaged one is sound. It is worth noting that the rationale
-listed here are based on empirical evidence, a rigorous proof of soundness
-perservation of the staged abstract interpreters is still an open question.
+listed here is based on empirical evidence; a rigorous proof of soundness
+preservation of the staged abstract interpreters remains an open question.
 
@@
   \item We usually assume the underlying MSP system achieves cross-stage
-    persistent, i.e., the generated program does not change the observational
-    behaviours of the original program. In the case of this paper, we use the
+    persistence, i.e., the generated program does not change the observational
+    behaviors of the original program. In the case of this paper, we use the
     LMS framework, which 1) relies on the type checking facilities of Scala, and
-    2) generates code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113}
+    2) generates code in A-Normal Form \cite{Flanagan:1993:ECC:155090.155113}
     that preserves the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
@@
     annotations and the monadic types. This allows the programmer to easily
-    check the correctness of implementation of staged monads modularlly.
+    check the correctness of implementation of staged monads modularly.
   \item We construct the staged monads and staged data structures in a
     correct-by-construct way that directly corresponds to their unstaged
-    version. For instance, the staged data structure we used here are simply
-    a black-box that wraps the Scala's library data structures (in the next
+    version. For instance, the staged data structures we used here are simply
+    black boxes that wrap the Scala's library data structures (in the next
     stage).
@@
 obtain an abstract interpreter in continuation-passing style, or simply use
-explicit side-effects such mutation to implement the artifact with the same
+explicit side-effects such as mutation to implement the artifact with the same
 abstract semantics. In either case, we can still apply the staging schema to
@@
 
-What we implemented is a big-step, compositional abstract interpreter in monadic
+We implemented a big-step, compositional abstract interpreter in monadic
 style, where \textit{compositional} means that every recursive call of our abstract
@@
 specialization can be done by unfolding, as well as guarantees the termination
-of specialization procedure. It is also possible to specialize small-step
+of the specialization procedure. It is also possible to specialize small-step
 operational abstract semantics through abstract compilation
@@
 optimizing Abstract Abstract Machines. However, the generated abstract
-byte-code still requires another small-step abstract machine to execute, which is
+bytecode still requires another small-step abstract machine to execute, which is
 an additional engineering effort.

### 2019-08-15 paper/60_cases.tex

@@
 \cite{DBLP:journals/jfp/HornM12}, in our staged abstract interpreter.
-
 \begin{lstlisting}
@@
 the address generation happens at the next stage, i.e., analysis time.
-
 \begin{lstlisting}
@@
 tuples, where the time is added.
-
 \begin{lstlisting}
@@
 version of monads. After the adjustment, we have the following monad stack:
-
 \begin{lstlisting}
@@
 time when we obtain a value. Here we adopt it with staged monads:
-
 \begin{lstlisting}
@@
 
-\subsection{Modular Analysis for Free} \label{modular}
+\subsection{Modular Staged Analysis for Free} \label{modular}
 
@@
 
-One of the challenges of modern static analysis is program usually depends on
-large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. If we can analyze
-programs and libraries separately and reuse the result without losing precision,
-then we can reduce the overhead of repeatedly analyzing libraries.
-Some static analyzers compute a summary for a function or a module, which
-can be reused later, however they are mostly too conservative or unsound, which
-both lead to imprecision. In this section, we show meta-programming can be used
-to create sound summaries in the code generation from, such generated code can
-be resued later. The summary contains information that not depend on the
-function argument.
+One of the challenges of modern static analysis is that program
+usually depends on large libraries programs
+\cite{toman_et_al:LIPIcs:2017:7121}. If we are able to analyze
+programs and libraries separately and reuse the result without losing
+precision, then we can reduce the overhead of repeatedly analyzing
+libraries. Some static analyzers compute a summary for a function or a
+module, which can be reused later. However, when lacking the context
+information, those analyses can be too conservative or unsound, which
+both lead to imprecision. In this section, we show that
+meta-programming can provide a trade-off between summary-based modular
+analysis and slow whole-program analysis, by ultilizing an existing
+whole-program analyzer.
 
-As we see when compiling a higher-order function (closure), we specialize the
-abstract interpreter to the body expression of that lambda term under some
-environment, without knowing the actual arguments to the function. We may reuse
-the infrastratures and refactor the @close@ function mentioned in Section
-\ref{sai} to achieve modular analysis. The @close@ function takes a lambda term
-@f@ and an environment, produces a compiled closure, which can be freely used at
-the next stage.
-
+As we have already seen from Section \ref{sai}, and recall the
+primitive operation @close@ that denotes a $\lambda$ term to next-stage
+Scala function.  The specializing of staged abstract interpreter to
+$\lambda$ terms is independent to other parts of the analyzed program.
 \begin{lstlisting}
-  def close(ev: Eval => Ans)(f: Expr, ρ: Rep[Env]): Rep[CompiledClo] = { ...}
+  def close(ev: Eval => Ans)(λ: Expr, ρ: Rep[Env]): Rep[Value]
 \end{lstlisting}
 
+Consider there is a library that contains a list map function @map@, 
+\begin{lstlisting}
+  (define (map xs f)
+    (if (null? xs) '() (cons (f (car xs)) (map (cdr xs) f))))
+\end{lstlisting}
+
+The specialization (0CFA and without store-widening) of @close@ on
+@map@ generates the following next-stage Scala code, that is, the
+function @x39@:
+\begin{lstlisting}
+  val x39 = {(x40: List[Set[AbsValue]],          // argument value list
+              x41: Map[Addr, Set[AbsValue]],     // store
+              x42: Map[Config, Set[ValSt]],      // in cache
+              x43: Map[Config, Set[ValSt]]) =>   // out cache
+    val x50 = List[Addr](ZCFAAddr("xs"), ZCFAAddr("f")) // addresses of arguments
+    val x51 = x50.zip(x40)                              // addresses paired with their values
+    val x61 = x51.foldLeft (x41) { case (x52, x53) =>   // prepare a new store that joins
+      val x54 = x53._1;  val x55 = x53._2               //   the arguments and their values
+        val x58 = x52.getOrElse(x54, x57)               //   into the latest store, intially x41
+        val x59 = x58.union(x55)
+        x52 + (x54 -> x59)
+    } ...
+  }
+\end{lstlisting}
+
+We borrow the notation traditional modular analysis and realize that
+such generated function is a reusable artifact which can be used
+modularly.  The generated functions already eliminate the
+interpretation overhead and monadic overhead, therefore can be
+consider as partial summaries. Although, we still need to run the
+generated function at next stage to obtain the analysis result.  At
+next stage, we can provide different abstract arugments, store and
+caches, depending on the context where it is called, thus the analysis
+is modular.
+
+Compared with traditional summary-based analysis, our partial
+summaries are represented by executable functions that are
+parameterized to the context information. In our approach, the
+transition from a whole-program analyzer to modular staged analyzer is
+mechanized and for free from staging. We do not need to design any
+intermediate representation or data structures of summaries.  In some
+scenario, traditional summary-based analyses are able to carry more
+information within a mdoule, e.g., the summary of their internal
+submodules. The staged approach presented here may still need to run
+through the fixed-point for submodules.
+
+\iffalse
 \paragraph{Partially-Static Context} Under a polyvariant analysis (like
@@
 initial environment containing @f@, @g@ and @h@.
-
 \begin{lstlisting}
@@
 our module being analyzed.
+\fi
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 elect to make the value of the assignment be an @void()@ value.
-
 \begin{lstlisting}
@@
 void value and current store.
-
 \begin{lstlisting}
@@
 two intervals produce a new
-
 \begin{lstlisting}

### 2019-08-15 paper/10_intro.tex
typos

@@
 dereference as a nondeterministic choice, such that the environment
-may contain mutiple possible values for a varaiable, then allocate
+may contain multiple possible values for a variable, then allocate
 continuations in the environment to exploit return-flow information,
@@
 abstract syntax tree (AST), which poses an interpretive overhead, such as
-the pattern matching on the ASTs and recursive calls on the subexpressions.
+the pattern matching on the ASTs and recursive calls on the sub-expressions.
 Such overhead can be negligible if the abstract
@@
 staging annotations can be either syntactic (e.g., quote and
-quasiquote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
+quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
 Modular Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
@@
 \item With two-level binding-times and concrete domain, it is a
-  compiler that translate to program intro antoher language;
+  compiler that translate to program intro another language;
 \item With a flat binding-time and abstract domains, it is an
@@
     well-grounded approach to optimize static analyses via
-    metaprogramming.
+    meta-programming.
   \item Practically, we show that staging an abstract interpreter is
@@
 \iffalse
-On the other side, static analysis is a tradeoff between performance and
+On the other side, static analysis is a trade-off between performance and
 precision: higher precision usually leads to longer running time.

### 2019-08-15 paper/20_prelim.tex
typos

@@
   DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
-In this section, we present the generic inferface in the style of big-step
+In this section, we present the generic interface in the style of big-step
 definitional interpreter. The key idea is to keep the binding-time type and
@@
 respectively. The @Env@ captures bound variables in scope, and @Store@
-models a persistent heap through the program runtime.  At this moment,
+models a persistent heap through the program run-time.  At this moment,
 the domains of addresses and values are still abstract, i.e., they are
@@
 declared as a higher-kinded type @R[_]@. We are free to use @R@ to annotate
-other datatypes in the interpreter; it is also injected into the monadic
+other data-types in the interpreter; it is also injected into the monadic
 type @MonadOps@.
@@
 
-We also declare several operations to mannipulate environments and
+We also declare several operations to manipulate environments and
 stores. These methods return monadic values of type @AnsM[_]@, which
@@
 happen inside of them.
-To close the open-recursion, we use a fixed-point opertor @fix@.
+To close the open-recursion, we use a fixed-point operator @fix@.
 For concrete-interpretation instantiation, it works like the Y
@@
 @R@ is the reader type, and @M[_]@ is the inner monad type.
-Given a @R@, @ReaderT@ monad produecs a transformed value of type @M[A]@.
+Given a @R@, @ReaderT@ monad produces a transformed value of type @M[A]@.
 Similarly, a @StateT@ monad encapsulates computation @S => M[(A, S)]@, where

### 2019-08-15 paper/30_stageconc.tex
Titles and spaces
typos

@@
 MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} that
-uses syntactic quotations and quasiquotations, LMS distinguishes binding-time
+uses syntactic quotations and quasi-quotations, LMS distinguishes binding-time
 based on types. LMS provides a type constructor @Rep[T]@ where @T@ can be
@@
 representations (IR) for the next stage program at the current stage
-\cite{DBLP:conf/birthday/Rompf16}. For convinence, the conversion from
+\cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
 expressions of type @Rep[_]@ to their IR is done by using the implicit design
@@
 monad with a state monad. But now the monads operate on staged values, for
-brivity, we call them \textit{staged monads}. In the code, we also use a @Rep@
+brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
 prefix on the constructors and types to differentiate them. But it is important
@@
 be achieved without modifying the implementation of monads. This is true so far,
-but as we will see, it is not straightfoward when for nondeterminism monad
+but as we will see, it is not straightforward when for nondeterminism monad
 (@SetT@) in abstract interpreters. Because not only the elements in the set are
@@
 The following code implements the idea:
-
 \begin{lstlisting}
@@
 @ev@ is invoked on the body expression @e@ to be specialized w.r.t. @e@, and
-then we \textit{collpase} the @AnsM@ monad to grounded values by providing the
+then we \textit{collapse} the @AnsM@ monad to grounded values by providing the
 prepared environment and store.
@@
 replace the store and reify the result to the current-stage monadic value.
-
 \begin{lstlisting}
@@
 the unstaged interpreter, the invoke of @ev@ happens in @ap_clo@; while in the
-staged interpreter, we eagerly apply @ev@ when denotating the lambda term to our
+staged interpreter, we eagerly apply @ev@ when denoting the lambda term to our
 value domain.
 
-\subsection{A Little Bit Code Generation}
+\subsection{A Little Bit of Code Generation}
 
 In the previous section, we merely show the type of @emit_compiled_clo@ and
-@emit_ap_clo@ without a concrete implemtation. In this section, we brifely
+@emit_ap_clo@ without a concrete implementation. In this section, we briefly
 discuss the IR node generated by them and sketch code generation.
@@
 representing next-stage definitions of type @T@.
-
 \begin{lstlisting}
@@
 matched @IRCompiledClo@ and @IRApClo@ and will generate definitions for them.
-For compiled closures, we put the function @f@ and its companying syntactic term
+For compiled closures, we put the function @f@ and its accompanying syntactic term
 and environment to its next-stage value representation @CompiledClo@. For
@@
 other next-stage values.
-
 \begin{lstlisting}

### 2019-08-15 paper/40_absinterp.tex
typos

@@
 lattice framework with binding-time abstraction, as well as simple abstract
-domains such as pwoer set lattice.
+domains such as power set lattice.
 Then we show the abstract components, namely abstract values, abstract
@@
 values, indicating that an address may point to an over-approximated set of
-values occurred at runtime.
-
+values occurred at run-time.
 \begin{lstlisting}
@@
 the corresponding values and stores for a given program configuration.
-
 \begin{lstlisting}
@@
 Now, the store is a mapping from addresses to sets of abstract values. This can
-be justied by the approximation nature of abstract interpretation and
+be justified by the approximation nature of abstract interpretation and
 nondeterminism in the abstract interpreter.
-First, when analyzing a conditional experssion, we may not have enough
+First, when analyzing a conditional expression, we may not have enough
 information on which branch will be taken, thus a sound treatment is to
@@
 @StateT@ for another cache that will be constantly updated during analysis.
-\citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permuatation of the
-moand stack for abstract interpretation. The following @AnsM@ type shows the
+\citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permutation of the
+monad stack for abstract interpretation. The following @AnsM@ type shows the
 monad stack (we use \hl{light gray} to highlight what have been changed from
 the monad stack of concrete interpretation):
-
 \begin{lstlisting}[escapechar=!]
@@
 all the transformed values into a single monadic value of type @Set[M, B]@.
-
 \begin{lstlisting}
@@
 the cache type of state monad becomes adjacent to the set type.
-
 \begin{lstlisting}
@@
 @Map@s).
-
 \begin{lstlisting}
@@
 @CloV@, but we lift it to a singleton set domain.
-
 \begin{lstlisting}
@@
 is added from the concrete @ap_clo@.
-
 \begin{lstlisting}[escapechar=!]
@@
 evaluation} \cite{Rosendahl:AbsIntPL}. It has been used in other abstract
-interpretater or fixed point computation
+interpreter or fixed point computation
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
@@
 the two caches through the monad stack; the implementations are elided.
-
 \begin{lstlisting}

### 2019-08-15 paper/50_stageabs.tex
typos

@@
 assumption that the unstaged one is sound. It is worth noting that the rationale
-listed here is based on empirical evidence; a rigorous proof of soundness
+listed here is based on empirical evidence; a rigorous proof for soundness
 preservation of the staged abstract interpreters remains an open question.
@@
     check the correctness of implementation of staged monads modularly.
-  \item We construct the staged monads and staged data structures in a
+  \item We build the staged monads and staged data structures in a
     correct-by-construct way that directly corresponds to their unstaged

### 2019-08-15 paper/60_cases.tex
typos

@@
 for a small CPS language consisting of abstractions, applications, recursion and
-primtive operators. Figure \ref{compare_ac_sai} (left) shows the AC
+primitive operators. Figure \ref{compare_ac_sai} (left) shows the AC
 implementation: different syntactic constructs are handled in different functions
@@
 with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
-changes of our appraoch using LMS framework are the types of @Store@ to
+changes of our approach using LMS framework are the types of @Store@ to
 @Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
@@
 
-We then integrate the timestamp into our moand stack by adding a staged
+We then integrate the timestamp into our monad stack by adding a staged
 @StateT@ monad. The result type is passingly get updated to contain a set of
@@
 DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
-implemented in monadic abstract interpreter by swaping the @StateT@ (for store)
+implemented in monadic abstract interpreter by swapping the @StateT@ (for store)
 and the @SetT@ (for nondeterminism). In this section, we show this optimization
@@
 is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
-current root addresses from the moand stack; @root@ compute the reachable
+current root addresses from the monad stack; @root@ compute the reachable
 addresses from a value; and @gc@ cleans up the store according to the reachable
@@
 meta-programming can provide a trade-off between summary-based modular
-analysis and slow whole-program analysis, by ultilizing an existing
+analysis and slow whole-program analysis, by utilizing an existing
 whole-program analyzer.
@@
       val x54 = x53._1;  val x55 = x53._2               //   the arguments and their values
-        val x58 = x52.getOrElse(x54, x57)               //   into the latest store, intially x41
+        val x58 = x52.getOrElse(x54, x57)               //   into the latest store, initially x41
         val x59 = x58.union(x55)
@@
 generated function at next stage to obtain the analysis result.  At
-next stage, we can provide different abstract arugments, store and
+next stage, we can provide different abstract arguments, store and
 caches, depending on the context where it is called, thus the analysis
@@
 scenario, traditional summary-based analyses are able to carry more
-information within a mdoule, e.g., the summary of their internal
-submodules. The staged approach presented here may still need to run
-through the fixed-point for submodules.
+information within a module, e.g., the summary of their internal
+sub-modules. The staged approach presented here may still need to run
+through the fixed-point for sub-modules.
 
@@
 consider that we have a library consists of three functions @f@, @g@ and @h@,
-where @f@ calls @g@ and @h@ internally and returns @y@ fianlly (we extend the
+where @f@ calls @g@ and @h@ internally and returns @y@ finally (we extend the
 language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
-would like to modularlly specialize these functions, starting from @f@ with an
+would like to modularly specialize these functions, starting from @f@ with an
 initial environment containing @f@, @g@ and @h@.
@@
 respect to the structure of analyzed program significantly improves the
-performance: a recent example is online decomposition of polyhedra
+performance: a recent example is online decomposition of polyhedron
 \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143}. In this
@@
 similar to how we treat @branch0@, we have a generic @branch@ function but works
-on boolean values. For the @true@ branch, we recursively call @ev@ on a newly
+on Boolean values. For the @true@ branch, we recursively call @ev@ on a newly
 constructed expression @Seq(e, While(t, e))@ meaning that first evaluates @e@

### 2019-08-15 paper/70_eval.tex
revision
typos

@@
 
-We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for running
-  JVM is set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
-  -XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}}, running on an Ubuntu 16.04 LTS
-(kernel 4.4.0) machine. All of our evaluations were performed on a machine with 4 Intel
-Xeon Platinum 8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96
-cores (192 threads) in total, the abstract interpreters are single-threaded.
-To minimize the effect of warming-up from the HotSpot JIT compiler, all the
-experiments are executed for 20 times and we report the statistical median value
-of the running times. We set a 5 minutes timeout.
+We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for
+running JVM is set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
+-XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
+running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All of our
+evaluations were performed on a machine with 4 Intel Xeon Platinum
+8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96 cores
+(192 threads) in total, the abstract interpreters are single-threaded.
+To minimize the effect of warming-up from the HotSpot JIT compiler,
+all the experiments are executed for 20 times and we report the
+statistical median value of the running times. We set a 5 minutes
+timeout.
 
@@
 DBLP:journals/corr/abs-1102-3676} and existing artifacts
-\footnote{https://github.com/ilyasergey/reachability} for control-flow analysis.
-Some of the benchmarks are small programs designed for experiments, such as the
-@kcfa-worst-@$n$ series, which are intended to challenge $k$-CFA; while some
-other benchmarks are from real-world applications, for examples, the RSA public
-key encryption algorithm @rsa@. In the Figure \ref{evaluation_result}, we report
-the number of AST node after parsing and desugaring (excluding comments) as a
-proper measurement of program size.
+\footnote{https://github.com/ilyasergey/reachability} for control-flow
+analysis.  Some of the benchmarks are small programs designed for
+experiments, such as the @kcfa-worst-@$n$ series, which are intended
+to challenge $k$-CFA; while some other benchmarks are from real-world
+applications, for examples, the RSA public key encryption algorithm
+@rsa@. In the Figure \ref{evaluation_result}, we report the number of
+AST node after parsing and desugaring (excluding comments) as a proper
+measurement of program size.
 
@@
 
-Figure~\ref{evaluation_result} shows the evaluation result, comparing the
-performance improved by staging on two monovariant CFA (without and with store-widening). The
-\textit{unstaged} and \textit{staged} columns show the median time to finish the
-analysis. The column $\frac{\text{unstaged}}{\text{staged}}$
-shows the improvement from staging. The dash - in some cells indicates timeout.
-As we can see from the table, the staged version significantly outperforms the
-unstaged version on all benchmarks.
-On average, the staging produces code that runs 17x times faster
-than the unstaged analyzer.
-
-We also observe that much overhead of unstaged analyzers are from the monadic
-layers. Thus our evaluation result provides an abstraction without regret
-approach to construct optimizing abstract interpreters -- the user may write a
-high-level modular and composable abstract interpreter, and then derive a staged
-version in a principled way that runs significantly faster.
+Figure~\ref{evaluation_result} shows the evaluation result, comparing
+the performance improved by staging on two monovariant CFA (without
+and with store-widening). The \textit{unstaged} and \textit{staged}
+columns show the median time to finish the analysis. The column
+$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement from
+staging. The dash - in some cells indicates timeout.  As we can see
+from the table, the staged version significantly outperforms the
+unstaged version on all benchmarks.  On average, the staging produces
+code that runs 17x times faster than the unstaged analyzer.
 
 \todo{analyze some interesting examples}
+
+Our evaluation result provides an abstraction without regret
+approach to construct optimizing abstract interpreters -- the user may write a
+high-level modular and compositional abstract interpreter, and then derive a
+staged version in a principled way that runs significantly faster.

### 2019-08-15 paper/80_related.tex
typos

@@
 analyses can be implemented and optimized from
-metaprogramming. Souffl{\'e}, one of the state-of-the-art Datalog
+meta-programming. Souffl{\'e}, one of the state-of-the-art Datalog
 engine that is commonly used in points-to analysis
 \cite{Antoniadis:2017:PDS:3088515.3088522}, uses the idea of Futamura
-projection for efficent implemetation \cite{10.1007/978-3-319-41540-6_23}.
+projection for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
 But it only specializes Datalog rules to the running engine, instead
@@
 specialized w.r.t. the static program structure, although not in the
-way of using staging. For example, decomposing polyhedras
+way of using staging. For example, decomposing polyhedrons
 \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to
@@
 \paragraph{Control-flow Analysis} Our case studies and evaluation focus on
-control-flow anslysis for functional languages
+control-flow analysis for functional languages
 \cite{Shivers:1991:SSC:115865.115884, Midtgaard:2012:CAF:2187671.2187672}. In
@@
 semantics, this paper shows how a staged analyzer can be built and
-used to increase efficieny of static analysis, we augment the monadic
+used to increase efficiency of static analysis, we augment the monadic
 abstract interpreter from another dimension, using code generation to

### 2019-08-15 paper/50_stageabs.tex
update

@@
 
-\paragraph{Are Monads Necessary?} No. One can always inline the monads and
-obtain an abstract interpreter in continuation-passing style, or simply use
-explicit side-effects such as mutation to implement the artifact with the same
-abstract semantics. In either case, we can still apply the staging schema to
-the abstract interpreter. As evidence, we provide a staged abstract
-interpreter written in direct-style in the accompanying artifact.
+\paragraph{Monadic-style vs Direct-style} We use monadic interpreter through the paper,
+but it is not necessary for staging. One can inline the monadic
+operations and obtain an abstract interpreter in continuation-passing
+style, or even translate back to direct-style that may use explicit
+side-effects such as mutations. In either case, we can still apply the
+staging idea to the abstract interpreter and remove the interpretation
+overhead. Although, monads make the staged abstract interpreter can be
+implemented in a modular and extensible way.
 

### 2019-08-15 paper/10_intro.tex
at sign

@@
 interpreter w.r.t. the input program yields an equivalent
-executable. For instance, @eval@ is an interpreter for some language:
+executable. For instance, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} is an interpreter for some language:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Given a program $e_0$ of type @Expr@, by applying the specialization,
+Given a program $e_0$ of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr}, by applying the specialization,
 we can obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$
-: @Input@ $\to$ @Value@, which is the so-called \textit{equivalent
+: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Input} $\to$ \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value}, which is the so-called \textit{equivalent
   executable}.  By definition of the interpreter, they produce the

### 2019-08-15 paper/20_prelim.tex
at sign

@@
 
-A monad is a type constructor @M[_]: * -> *@ with two operations, often called
-@return@ and @bind@. Informally, @return@ wraps a value into the monad @M@, and
-@bind@ unwraps the monadic value and transforms it into a new monadic value.
-Pragmatically in Scala, we define a monad type class using trait @Monad@ (Figure
-\ref{fig:monad}), where it declares the @pure@ \footnote{We elect to use
+A monad is a type constructor \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[_]: * -> *} with two operations, often called
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{return} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{bind}.
+Informally, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{return} wraps a value into the monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M}, and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{bind} unwraps the monadic value and transforms it into a new monadic value.
+Pragmatically in Scala, we define a monad type class using trait \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Monad} (Figure
+\ref{fig:monad}), where it declares the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{pure} \footnote{We elect to use
 \texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
-\texttt{unit} is a built-in function in LMS.} and @flatMap@ operation. The trait
-itself takes the monad type @M[_]@ as argument, which is a higher-kinded type
-that takes a type and returns a type. The method @pure@ promotes values of type
-@A@ to values of type @M[A]@. The monadic @bind@ operation is usually called
-@flatMap@ in Scala, which takes a monad-encapsulated value of type @M[A]@, a
-function of @A => M[B]@ and returns values of type @M[B]@.
+\texttt{unit} is a built-in function in LMS.} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} operation. The trait
+itself takes the monad type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[_]} as argument, which is a higher-kinded type
+that takes a type and returns a type. The method \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{pure} promotes values of type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} to values of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[A]}.
+The monadic \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{bind} operation is usually called
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} in Scala, which takes a monad-encapsulated value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[A]}, a
+function of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A => M[B]} and returns values of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[B]}.
 
@@
 
-Similar to Haskell's @do@-notation, Scala provides special syntactic support for
-monadic operations through @for@-comprehension.
-For example, an object of @List[A]@ is an instance of @List@ monad, where @A@ is the element type. 
+Similar to Haskell's \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{do}-notation, Scala provides special syntactic support for
+monadic operations through \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension.
+For example, an object of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{List[A]} is an instance of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{List} monad, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} is the element type.
 Then to compute the Cartesian product of two lists of numbers, we can use Scala's
-@for@-comprehension syntax.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension syntax.
 \begin{lstlisting}
@@
 
-The Scala compiler will translate the above @for@-comprehension expression into
-an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
-in the @for@-comprehension is translated into a @map@, where the expression of
-@yield@ becomes the body expression of that @map@ application. The foregoing
-bindings in the comprehension are all translated into calls of @flatMap@.
+The Scala compiler will translate the above \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension expression into
+an equivalent one using \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map} ~\cite{scala_spec}.
+The last binding
+in the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension is translated into a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}, where the expression of
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{yield} becomes the body expression of that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map} application.
+The foregoing
+bindings in the comprehension are all translated into calls of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}.
 \begin{lstlisting}
@@
 
-Note that here the monadic object @List[_]@ encapsulates the data
+Note that here the monadic object \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{List[_]} encapsulates the data
 internally.  Therefore it only exposes the simplified version of
-@flatMap@, where the monad @M[A]@ is not introduced as a function
-argument. The trait @MonadOps@ (Figure \ref{fig:monadops}) defines the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}, where the monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[A]} is not introduced as a function
+argument. The trait \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps} (Figure \ref{fig:monadops}) defines the
 simplified version of monadic operations that are necessary for
-@for@-comprehension. The conversion between @Monad@ and @MonadOps@
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension.
+The conversion between \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Monad} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps}
 can be done by using the implicit design pattern.
-In the rest of the paper, we use Scala's @for@-comprehension syntax
-and a few of monads and monad transformers such as @ReaderT@,
-@StateT@, and @ListT/SetT@ to write our interpreters. The implementation
+In the rest of the paper, we use Scala's \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension syntax
+and a few of monads and monad transformers such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT},
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT}, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ListT/SetT} to write our interpreters.
+The implementation
 of monads and monad transformers essentially borrows the ground-truth
@@
 
-Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
+Monad transformers are type constructors of kind \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{(* -> *) -> (* -> *)}, which
 take a monad as argument and produces another monad. By using monad
@@
 strings. To represent states, two required components for the
-interpreter are environments @Env@ and stores @Store@. @Env@ maps from
-identifiers to addresses, and @Store@ maps from addresses to values,
-respectively. The @Env@ captures bound variables in scope, and @Store@
+interpreter are environments \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env} and stores \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env} maps from
+identifiers to addresses, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store} maps from addresses to values,
+respectively.
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env} captures bound variables in scope, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}
 models a persistent heap through the program run-time.  At this moment,
@@
 \paragraph{Binding-time Abstraction} As mentioned before, the binding-time is
-declared as a higher-kinded type @R[_]@. We are free to use @R@ to annotate
+declared as a higher-kinded type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]}.
+We are free to use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} to annotate
 other data-types in the interpreter; it is also injected into the monadic
-type @MonadOps@.
-If we simply instantiate @R@ an identity type (i.e., @type R[T] = T@),
+type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps}.
+If we simply instantiate \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} an identity type (i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{type R[T] = T}),
 then the generic interpreter is a standard definitional interpreter
 that will execute the program.  In Section \ref{stagedinterp}, we will
-instantiate @R@ using LMS's built-in next-stage type annotation @Rep@,
+instantiate \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} using LMS's built-in next-stage type annotation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep},
 which makes the interpreter act as a compiler.
@@
 \paragraph{Monadic Operations} We define the return type of the interpreter as
-@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
-As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
-syntax, certain operations have to be added on the type @AnsM@. Here, we use a
-structural type @MonadOps@ to require that type @AnsM@ at least implements @map@ and
-@flatMap@. It is worth noting that @MonadOps@ takes another type parameter
-@R[_]@ as binding-time that annotates data; accordingly the type @R@
-defined in the outer scope is passed to @MonadOps@. Inside of @MonadOps@,
-@R[_]@ wraps the data types @A@ and @B@ that are encapsulated by the
-monad, but not the monad type @M@ itself. When acting as compilers, we
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Ans}, which is a monad type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM[_]} wrapping the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value}.
+As mentioned in Section ~\ref{monadscala}, to use the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{for}-comprehension
+syntax, certain operations have to be added on the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM}. Here, we use a
+structural type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps} to require that type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM} at least implements \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map} and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}.
+It is worth noting that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps} takes another type parameter
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]} as binding-time that annotates data; accordingly the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R}
+defined in the outer scope is passed to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps}.
+Inside of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps},
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]} wraps the data types \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{B} that are encapsulated by the
+monad, but not the monad type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M} itself. When acting as compilers, we
 will also replace the monads to the ones that work on staged data values.
@@
 We also declare several operations to manipulate environments and
-stores. These methods return monadic values of type @AnsM[_]@, which
-may encapsulate over the environment or store, or simply a @Unit@
-value for effects. For examples, @local_env@ installs a new environment
-@ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
+stores. These methods return monadic values of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM[_]}, which
+may encapsulate over the environment or store, or simply a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Unit}
+value for effects. For examples, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{local_env} installs a new environment
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ρ} when evaluating the monadic value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ans}; \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{set_store} takes a pair of
 addresses and values (potentially staged) and updates the store accordingly.
@@
 \paragraph{Primitive Operations} Next we define several primitive operations.
-First, two versions of @alloc@ are declared. The first one takes a store and an
-identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
+First, two versions of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{alloc} are declared. The first one takes a store and an
+identifier and produces a fresh address of non-monadic type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[Addr]}. Since
 the freshness of the address may depend on the store, which might be a
 next-stage value as indicated by its type, the type of addresses is therefore
-wrapped by @R[_]@. The other @alloc@ is simply the monadic version of the addresses.
+wrapped by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]}.
+The other \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{alloc} is simply the monadic version of the addresses.
 \begin{lstlisting}
@@
 Other primitive operations provide basic functionality for the
-interpreter.  The method @num@ and @close@ deal with primitive values,
+interpreter.
+The method \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{num} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} deal with primitive values,
 which lift literal terms (e.g., lambdas) to our value representation
-(e.g., closures).  The method @get@ retrieves the value of an
-identifier @x@, through the environment and store. Conditionals and
-arithmetic is handled by @br0@ and @arith@, respectively. The method
-@ap_clo@ is used for function applications: it takes a function value
-and an argument value. Note that the @Env@, @Store@, and @Value@ are
-all annotated by @R[_]@ since they can potentially be next-stage
+(e.g., closures).  The method \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{get} retrieves the value of an
+identifier \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x}, through the environment and store. Conditionals and
+arithmetic is handled by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{br0} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{arith}, respectively.
+The method
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} is used for function applications: it takes a function value
+and an argument value.
+Note that the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} are
+all annotated by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]} since they can potentially be next-stage
 values when acting as compilers.
@@
 interpreter is written in open-recursive style -- it can not refer to
-itself directly, instead, @eval@ takes an additional parameter @ev@ of
-type @Expr => Ans@ that refers to itself. Consequently, the method
-@close@ that lifts lambda terms to closures and method @ap_clo@ that
-applies functions also takes an extra @ev@, because evaluation may
+itself directly, instead, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} takes an additional parameter \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} of
+type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr => Ans} that refers to itself. Consequently, the method
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} that lifts lambda terms to closures and method \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} that
+applies functions also takes an extra \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev}, because evaluation may
 happen inside of them.
-To close the open-recursion, we use a fixed-point operator @fix@.
+To close the open-recursion, we use a fixed-point operator \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}.
 For concrete-interpretation instantiation, it works like the Y
@@
 the termination of abstract interpretation.
-Finally, a top-level wrapper @run@ is declared; the return type
-@Result@ depends on what kind of monad we will be using, this is also
+Finally, a top-level wrapper \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{run} is declared; the return type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Result} depends on what kind of monad we will be using, this is also
 an abstract type.
@@
 \paragraph{Concrete Components}
-The two types we need to concretize are addresses @Addr@ and values @Value@. The
-type @Env@ and @Store@ is derived automatically. To assure the freshness of
-address allocation, we use @Int@ and always return successor of the size of
+The two types we need to concretize are addresses \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Addr} and values \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value}.
+The
+type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store} is derived automatically.
+To assure the freshness of
+address allocation, we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Int} and always return successor of the size of
 current store. A value can be either a tagged number \texttt{IntV}, or a closure
 \texttt{CloV} that contains a lambda term and an environment. The final grounded
-result of the interpreter is a pair of values and stores, where the @Value@ and
-@Store@ are next-stage objects, as annotated by @R@. We also define a standard
-fixed-point combinator to close the open-recursive function @ev@.
+result of the interpreter is a pair of values and stores, where the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store} are next-stage objects, as annotated by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R}.
+We also define a standard
+fixed-point combinator to close the open-recursive function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev}.
 \begin{lstlisting}
@@
 respectively. Following the monad transformer approach
-\cite{DBLP:conf/popl/LiangHJ95}, we use a @ReaderT@ and @StateT@ monad
-transformer to compose the monad stack. In other words, the type @AnsM@ is
-instantiated by layering @ReaderT@ and @StateT@ transformers\footnote{The
+\cite{DBLP:conf/popl/LiangHJ95}, we use a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} monad
+transformer to compose the monad stack. In other words, the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM} is
+instantiated by layering \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} transformers\footnote{The
 question mark syntax is a kind projector \cite{kindprojector}, thus
 \texttt{StateT[IdM,Store,?]} is equivalent to \newline \texttt{(\{type
-M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is parameterized by type
-@Env@, @StateT@ is parameterized by type @Store@, and the inner-most monad @IdM@
+M[T]=StateT[IdM,Store,T]\})\#M}}, where the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} is parameterized by type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} is parameterized by type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}, and the inner-most monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IdM}
 is simply an identity monad.
@@
 
-Here we sketch the basic idea of @ReaderT@ and @StateT@. Readers may
+Here we sketch the basic idea of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT}.
+Readers may
 refer to \cite{DBLP:conf/popl/LiangHJ95, Chiusano:2014:FPS:2688794} for more detail.
-A @ReaderT@ monad transformer encapsulates computation @R => M[A]@, where
-@R@ is the reader type, and @M[_]@ is the inner monad type.
-Given a @R@, @ReaderT@ monad produces a transformed value of type @M[A]@.
-Similarly, a @StateT@ monad encapsulates computation @S => M[(A, S)]@, where
-@S@ is the state type, and @M[_]@ is the inner monad type.
-Given a @S@, @StateT@ monad produces a transformed value of type @M[(A, S)]@,
-where the new state (type @S@) is accompanied with the value (type @A@).
-Note that for the moment, the binding-time type @R@ is an identity type, thus
+A \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} monad transformer encapsulates computation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R => M[A]}, where
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} is the reader type, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[_]} is the inner monad type.
+Given a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} monad produces a transformed value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[A]}.
+Similarly, a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} monad encapsulates computation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S => M[(A, S)]}, where
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S} is the state type, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[_]} is the inner monad type.
+Given a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} monad produces a transformed value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[(A, S)]},
+where the new state (type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S}) is accompanied with the value (type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A}).
+Note that for the moment, the binding-time type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} is an identity type, thus
 these monads operate on unstaged data. We can also see this from the
-signature of @flatMap@: the argument function @f@ takes an unstaged value of type @A@ and
-produces a monadic value. In the following code, we elide operations other than @flatMap@.
+signature of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}: the argument function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} takes an unstaged value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} and
+produces a monadic value. In the following code, we elide operations other than \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}.
 \begin{lstlisting}
@@
 top-level of our monad stack. To modify the store, for example, we can construct a
-@StateT@ value that transforms the current store $\sigma$ to @σ + αv@, which
-results in @StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to the
-top-level @ReaderT@ type, i.e., @AnsM[Unit]@.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} value that transforms the current store $\sigma$ to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{σ + αv}, which
+results in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT[IdM, Store, Unit]}, and then lift this \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} value to the
+top-level \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} type, i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM[Unit]}.
 \begin{lstlisting}
@@
 different when staging and abstract interpretation is involved. The method
-@close@ takes a lambda term and an environment, and produces the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} takes a lambda term and an environment, and produces the
 defunctionalized representation of closures.
@@
 
-To apply a function, @ap_clo@ takes a function value @rator@ and an argument
-@rand@, extracts the lambda term and environment enclosed in @rator@, allocates
+To apply a function, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} takes a function value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rator} and an argument
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rand}, extracts the lambda term and environment enclosed in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rator}, allocates
 the addresses and updates the environment and store accordingly. Finally, it
@@
 
-The top-level @run@ method invokes the fixed-pointer operator directly,
-given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$.
+The top-level \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{run} method invokes the fixed-pointer operator directly,
+given the expression \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}, initial environment $\rho_0$ and store $\sigma_0$.
 \begin{lstlisting}

### 2019-08-15 paper/30_stageconc.tex
at sign

@@
 uses syntactic quotations and quasi-quotations, LMS distinguishes binding-time
-based on types. LMS provides a type constructor @Rep[T]@ where @T@ can be
-an arbitrary type, indicating a value of @T@ will be known at the next stage.
-All operations acting on @Rep[T]@ expressions will be residualized as generated
+based on types.
+LMS provides a type constructor \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[T]} where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T} can be
+an arbitrary type, indicating a value of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T} will be known at the next stage.
+All operations acting on \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[T]} expressions will be residualized as generated
 code.
@@
 
-If @x@ is a value know at current stage, we may specialize the power function to
-some value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
-first adding the @Rep@ type to the variable known at next stage. In this case,
-@b : Rep[Int]@ is know later, and @x : Int@ is know currently, as shown in the
+If \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} is a value know at current stage, we may specialize the power function to
+some value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} -- by unrolling the recursive calls. In LMS, this is fulfilled by
+first adding the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type to the variable known at next stage. In this case,
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{b : Rep[Int]} is know later, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x : Int} is know currently, as shown in the
 below code recipe.
-The way we use this @power@ function is to create a @DslDriver@ and override the
-@snippet@ method give the currently know value @x@ (5 in this example).
+The way we use this \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{power} function is to create a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{DslDriver} and override the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{snippet} method give the currently know value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} (5 in this example).
 \begin{lstlisting}
@@
 The LMS framework provides staging support for primitive data types such as
-@Int@ and @Double@, and commonly-used data structures such as lists and maps.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Int} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Double}, and commonly-used data structures such as lists and maps.
 The idea behind the framework is to construct a sea-of-node intermediate
@@
 \cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
-expressions of type @Rep[_]@ to their IR is done by using the implicit design
+expressions of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[_]} to their IR is done by using the implicit design
 pattern. As we will see later, implementing staging and code generation
@@
 instantiating unstaged concrete interpreter in Section \ref{unstaged_conc}.
-But now we use the @Rep@ type to annotate value domains, environments and
+But now we use the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type to annotate value domains, environments and
 stores, and redefine the staged version of monads and primitive operations.
@@
 monad with a state monad. But now the monads operate on staged values, for
-brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
+brevity, we call them \textit{staged monads}. In the code, we also use a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep}
 prefix on the constructors and types to differentiate them. But it is important
-to note that objects of a case class @ReaderT@/@StateT@ are not staged, the
-monadic computation like @R => M[A]@ are also not staged. Instead, the internal
-data that these monads operate on are staged, i.e., @R@ and @A@ in the
-reader monad, @S@ and @A@ in the state monad. The following code
-snippet shows the idea. We use \hl{light gray} to highlight where @Rep@ type is added:
+to note that objects of a case class \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT}/\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} are not staged, the
+monadic computation like \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R => M[A]} are also not staged. Instead, the internal
+data that these monads operate on are staged, i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} in the
+reader monad, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} in the state monad.
+The following code
+snippet shows the idea. We use \hl{light gray} to highlight where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type is added:
 \begin{lstlisting}[escapechar=!]
@@
 
-The function @f@ passed to @flatMap@ is also a current-stage-known value, so
-that all the invocation of @flatMap@s can be completed before generating code.
+The function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} passed to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} is also a current-stage-known value, so
+that all the invocation of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}s can be completed before generating code.
 The fact that we only stage data but not monadic computation or monadic
 values, is the reason that we can peel of the monad stack in the generated code.
-Now the @AnsM[_]@ is instantiated as the same structure as before, but using the
-@Rep@ versions:
+Now the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM[_]} is instantiated as the same structure as before, but using the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} versions:
 \begin{lstlisting}
@@
 but as we will see, it is not straightforward when for nondeterminism monad
-(@SetT@) in abstract interpreters. Because not only the elements in the set are
+(\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT}) in abstract interpreters. Because not only the elements in the set are
 staged, but the whole set is also staged, and we have no knowledge about how
@@
 and its environment to a next-stage Scala function. To do this, we need to
-recursively call the interpreter function, i.e., the @ev@ argument passed to @close@.
+recursively call the interpreter function, i.e., the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} argument passed to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close}.
 The following code implements the idea:
@@
 
-We have created a current-stage function @f@ that takes a next-stage value and
-store, and returns staged value. Then we use @emit_complied_clo@ to delay @f@ to
-a next-stage Scala function, represented by the type @Rep[Value]@. Inside of
-@f@, we can access the evaluator via @ev@, but @ev@ produces a monadic value of
-type @AnsM@, which can only exist at the current stage. To connect the
+We have created a current-stage function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} that takes a next-stage value and
+store, and returns staged value.
+Then we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_complied_clo} to delay \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} to
+a next-stage Scala function, represented by the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Value]}. Inside of
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, we can access the evaluator via \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev}, but \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} produces a monadic value of
+type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM}, which can only exist at the current stage. To connect the
 current-stage monadic values and future-stage grounded values, the evaluator
-@ev@ is invoked on the body expression @e@ to be specialized w.r.t. @e@, and
-then we \textit{collapse} the @AnsM@ monad to grounded values by providing the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} is invoked on the body expression \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e} to be specialized w.r.t.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}, and
+then we \textit{collapse} the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM} monad to grounded values by providing the
 prepared environment and store.
 
-For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
-and @arg@, but what we can do on these values of @Rep@ type? In fact, @fun@ does
+For function applications \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}, we have two \textit{next-stage} values \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun}
+and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{arg}, but what we can do on these values of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type?
+In fact, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} does
 not have an intensional application operation we can use directly; but we know
-that it is a staged value produced by the @close@ function, which represents a
+that it is a staged value produced by the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} function, which represents a
 next-stage Scala function. Therefore we can just generate a next-stage function
-application, also in Scala. Since @ap_clo@ returns a monadic value of type
-@Ans@, we can still use the monadic operation @get_store@ to obtain the latest
-store object. Then we use @emit_ap_clo@ to generate a current-stage
+application, also in Scala. Since \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} returns a monadic value of type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Ans}, we can still use the monadic operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{get_store} to obtain the latest
+store object. Then we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo} to generate a current-stage
 representation of the future-stage application result, which is a pair of values
-and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
+and stores \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ValSt}.
+After getting that result \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{vs}, we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{put_store} to
 replace the store and reify the result to the current-stage monadic value.
@@
 Compared with the unstaged version, we can observe a shift of using the
-evaluation function @ev@ from application-time to value-representation-time: in
-the unstaged interpreter, the invoke of @ev@ happens in @ap_clo@; while in the
-staged interpreter, we eagerly apply @ev@ when denoting the lambda term to our
+evaluation function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} from application-time to value-representation-time: in
+the unstaged interpreter, the invoke of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} happens in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}; while in the
+staged interpreter, we eagerly apply \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} when denoting the lambda term to our
 value domain.
@@
 
-In the previous section, we merely show the type of @emit_compiled_clo@ and
-@emit_ap_clo@ without a concrete implementation. In this section, we briefly
+In the previous section, we merely show the type of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_compiled_clo} and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo} without a concrete implementation. In this section, we briefly
 discuss the IR node generated by them and sketch code generation.
-We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
-extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
-representing next-stage definitions of type @T@.
+We define the IR nodes \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} using \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{case class}
+extending from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Def[T]}.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Def[T]} is a built-in type in the LMS framework,
+representing next-stage definitions of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T}.
 \begin{lstlisting}
@@
 The IR nodes are manipulated by the LMS passes and finally generated to
-next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
+next-stage Scala code. To generate code, LMS provides an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emitNode} method for
 programmers to control what is generated for each kind of node. Here we have
-matched @IRCompiledClo@ and @IRApClo@ and will generate definitions for them.
-For compiled closures, we put the function @f@ and its accompanying syntactic term
-and environment to its next-stage value representation @CompiledClo@. For
-applications, we have the knowledge that @fun@ is a @CompiledClo@ at the stage,
-and it has a callable field @f@, so we generate a next-stage function
-application @f(arg, σ)@. The @quote@s are invocations to generate code for the
+matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions for them.
+For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term
+and environment to its next-stage value representation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo}. For
+applications, we have the knowledge that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} is a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo} at the stage,
+and it has a callable field \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, so we generate a next-stage function
+application \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f(arg, σ)}.
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{quote}s are invocations to generate code for the
 other next-stage values.

### 2019-08-15 paper/40_absinterp.tex
at sign

@@
 operator, and $\sqcap: S \to S \to S$ is the meet (greatest lower bound)
-operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines a type
-class of stage polymorphic lattices. Similar to @MonadOps@ in Section
-\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
-annotate the binding-times, where @R[_]@ wraps datatype @S@ and @Boolean@ in
+operator. The trait \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SPLattice} (Figure \ref{fig:splattice}) defines a type
+class of stage polymorphic lattices. Similar to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadOps} in Section
+\ref{generic_if}, we introduce an additional higher-kinded type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]} to
+annotate the binding-times, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[_]} wraps datatype \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Boolean} in
 these operations.
@@
 
-In this section, we simply let the binding-time type @R[T] = T@ (i.e., @NoRep@)
+In this section, we simply let the binding-time type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R[T] = T} (i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{NoRep})
 to instantiate an unstaged lattice type class. An example of lattice we used in
@@
 
-We first widen the concrete values to @AbsValue@: numbers are lifted to a
-singleton object @IntTop@ representing the set of all numbers; closures remain
-the same. Then the type @Value@ is a set of @AbsValue@ (we abuse the type names
-a bit, as the generic interface already defined the abstract type @Value@).
+We first widen the concrete values to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AbsValue}: numbers are lifted to a
+singleton object \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IntTop} representing the set of all numbers; closures remain
+the same.
+Then the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} is a set of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AbsValue} (we abuse the type names
+a bit, as the generic interface already defined the abstract type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value}).
 The address space is constrained to be finite to ensure the computability of
 analysis -- for the simplest monovariant analysis, we use variable names for
-addresses. After defining the @Value@ and @Addr@ type, the environments and
+addresses.
+After defining the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Addr} type, the environments and
 stores are automatically lifted to abstract environments and abstract stores,
-i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
+i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map[Ident, Addr]} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map[Addr, Set[AbsValue]]} respectively.
 It is important that the abstract store now maps addresses to sets of abstract
@@
 Here we first provide the necessary type definitions, and later we will describe
-the algorithm in detail. A @Cache@ is a mapping from configurations @Config@ to
+the algorithm in detail.
+A \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Cache} is a mapping from configurations \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Config} to
 sets of value-store pairs, where the configuration is a triple of current
@@
 nondeterministically explore both branches, i.e., compute the join of two path.
-Second, when applying a function, for instance, @f(a)@, the possible lambda
-terms of @f@ is computed approximately. As a result, we may retrieve a set of
-closures when dereferencing the address of @f@ and need to apply all of them.
+Second, when applying a function, for instance, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f(a)}, the possible lambda
+terms of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} is computed approximately. As a result, we may retrieve a set of
+closures when dereferencing the address of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and need to apply all of them.
 
@@
 for the two caches. The nondeterminism effect is represented by the
-@Set[M[_], A]@ monad transformer, where @M@ is the inner monad being
-transformed, and @A@ is the type of elements in the set. We use a @ReaderT@ for
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set[M[_], A]} monad transformer, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M} is the inner monad being
+transformed, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} is the type of elements in the set.
+We use a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT} for
 one cache that will not be changed during one fixed-point iteration, and use a
-@StateT@ for another cache that will be constantly updated during analysis.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} for another cache that will be constantly updated during analysis.
 \citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permutation of the
-monad stack for abstract interpretation. The following @AnsM@ type shows the
+monad stack for abstract interpretation. The following \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM} type shows the
 monad stack (we use \hl{light gray} to highlight what have been changed from
@@
 
-Similar to the concrete scenario, we sketch the implementation of @flatMap@ for
-@SetT@ transformer and omit the rest. The value @run@ encapsulated by the monad
-is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and @A@ is the element
-type of the set. The @flatMap@ method takes a function that transforms each
-element of type @A@ from @M[Set[A]]@ into a @SetT[M, B]@ value, and then folds
-all the transformed values into a single monadic value of type @Set[M, B]@.
+Similar to the concrete scenario, we sketch the implementation of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} for
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} transformer and omit the rest.
+The value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{run} encapsulated by the monad
+is of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[SetT[A]]}, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[_]} is the inner monad, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} is the element
+type of the set. The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} method takes a function that transforms each
+element of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M[Set[A]]} into a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT[M, B]} value, and then folds
+all the transformed values into a single monadic value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set[M, B]}.
 \begin{lstlisting}
@@
 with the value produced by the interpreter, the final result type is a pair of
-@Value@ and @Store@. In the abstract setting, we have a @SetT@ inside of the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}.
+In the abstract setting, we have a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} inside of the
 environment monad and store monad, which becomes the container type of the pair
-of values and stores, i.e., @Set[(Value, Store)]@. Also note that the reader and
+of values and stores, i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set[(Value, Store)]}. Also note that the reader and
 state monad for caches live internally of the nondeterminism monad, therefore
@@
 As what mentioned before, we may have to explore both branches for conditionals:
-in @br0@ method, we combine the results using @mplus@ from @MonadPlus@, which
-requires the value domains are join-semilattices (in our case @Set@s and
-@Map@s).
+in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{br0} method, we combine the results using \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{mplus} from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{MonadPlus}, which
+requires the value domains are join-semilattices (in our case \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set}s and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map}s).
 \begin{lstlisting}
@@
 The value representation of lambda terms is still a defunctionalized closure
-@CloV@, but we lift it to a singleton set domain.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CloV}, but we lift it to a singleton set domain.
 \begin{lstlisting}
@@
 
-For function applications, @ap_clo@ looks the same as the concrete interpreter.
-The difference is now the first argument @rator@ is a set of closures (remember
-that @Value@ is now an alias of @Set[AbsValue]@), so we can not directly
-destruct it by pattern matching. Instead, we use an auxiliary function @lift_nd@
+For function applications, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} looks the same as the concrete interpreter.
+The difference is now the first argument \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rator} is a set of closures (remember
+that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value} is now an alias of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set[AbsValue]}), so we can not directly
+destruct it by pattern matching. Instead, we use an auxiliary function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{lift_nd}
 that takes a set and lifts the elements into the monad stack. Then we can still
-implement @ap_clo@ in monadic style, where the closures come from the @SetT@ and
+implement \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} in monadic style, where the closures come from the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} and
 thus the nondeterminism can be naturally handled. The light gray code shows what
-is added from the concrete @ap_clo@.
+is added from the concrete \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}.
 \begin{lstlisting}[escapechar=!]
@@
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
-  Rosendahl:AbsIntPL}. The idea is to use a @in@ cache and a @out@ cache during
-the depth-first evaluation, where the @in@ cache stores the result from last
-iteration, and @out@ cache is used for updates in current iteration. In the next
-iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
-is plugged-in to the @out@ slot. Once the @out@ cache does not contain anything
-new than the @in@ cache, the fixed-point is reached.
+  Rosendahl:AbsIntPL}.
+The idea is to use a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} cache and a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache during
+the depth-first evaluation, where the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} cache stores the result from last
+iteration, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache is used for updates in current iteration. In the next
+iteration, the last \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache will be served as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} cache, and an empty cache
+is plugged-in to the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} slot.
+Once the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache does not contain anything
+new than the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} cache, the fixed-point is reached.
 
-In out monad stack, the @in@ and @out@ cache modeled by a reader effect and
+In out monad stack, the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache modeled by a reader effect and
 state effect, respectively. We first defined several methods to manipulate
@@
 The co-inductive caching algorithm is implemented as a instrumentation over
-@eval@ (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
 Initially, the two caches are both empty.
-During the iteration, we first check whether @out@ contains the configuration
-@cfg@, representing the current desired computation. If it does, then the result
+During the iteration, we first check whether \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} contains the configuration
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{cfg}, representing the current desired computation. If it does, then the result
 is directly returned through the monads.
-Otherwise, we first retrieve the results from @in@ (@⊥@ if not contained in
-@in@), and join it into the @out@ cache.
-Then we invoke @ev@ to compute the result for this iteration, where @ev@ takes
-@fix(ev)@ as a self reference.
-After the evaluation, the store @σ@ may have been changed. We update the
-@out@ cache with the new result, i.e., a pair of value and store @(v, σ)@.
+Otherwise, we first retrieve the results from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} (\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{⊥} if not contained in
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in}), and join it into the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache.
+Then we invoke \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} to compute the result for this iteration, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} takes
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix(ev)} as a self reference.
+After the evaluation, the store \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{σ} may have been changed. We update the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache with the new result, i.e., a pair of value and store \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{(v, σ)}.
 Similar to store update, cache update is also a point-wise join.
 
-The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which indicates that there is no more fact can be gained, thus
+The iteration terminates when the resulting \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache is equivalent to the
+input \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} cache, which indicates that there is no more fact can be gained, thus
 the iteration should end and we have reached the least fixed point.
@@
 
-Finally, we override the top-level @run@ method by running the monadic value
+Finally, we override the top-level \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{run} method by running the monadic value
 with the empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
-and initial caches @cache@$_0$.
+and initial caches \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{cache}$_0$.

### 2019-08-15 paper/50_stageabs.tex
at sign

@@
 staged lattices and staged monads, and then the staged version of primitive
-operations, especially @close@, @ap_clo@ and @fix@. At last, we discuss several
+operations, especially \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}.
+At last, we discuss several
 optimizations and summarize our approach.
@@
 
-In Section \ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
-achieve stage polymorphism. Now we instantiate the type @R@ to @Rep@ and
+In Section \ref{stagedpoly_lat}, we exploited the higher-kinded type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} to
+achieve stage polymorphism.
+Now we instantiate the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} and
 still use power sets as an example to present its staged version.
@@
 
-The trait @RepLattice@ is obtained by specializing @Lattice@ with @Rep@.
-The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
+The trait \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{RepLattice} is obtained by specializing \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Lattice} with \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep}.
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{RepSetLattice} is an instance of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{RepLattice}, where the lattice operations
 eventually are delegated to operations on a staged set data type, including
-@subsetOf@, @union@ and @intersect@. We even do not have to change the
-implementation code, but just the types -- from @Set[T]@ to @Rep[Set[T]]@, shown
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{subsetOf}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{union} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{intersect}.
+We even do not have to change the
+implementation code, but just the types -- from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set[T]} to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Set[T]]}, shown
 in light gray code. The LMS library provides an implicit conversion that lifts a
@@
 
-Now we can implement the staged abstract semantics, where @R@ is instantiated as
-@Rep@, and the abstract components are reused from the unstaged version.
+Now we can implement the staged abstract semantics, where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} is instantiated as
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep}, and the abstract components are reused from the unstaged version.
 
@@
 following code shows this change. For readability, we squeeze the three inner
-transformers (@IdM@ omitted) into a single monad @RepSetReaderStateM[A, B, C]@,
-where @A@ is the type parameter for reader effect and @B@ for the state effects.
-In our monad stack scheme, they are both instantiated with @Cache@. Now the
-result type is a pair of two staged value: @Rep[Set[(Value, Store)]]@ and
-@Rep[Cache]@.
+transformers (\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IdM} omitted) into a single monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{RepSetReaderStateM[A, B, C]},
+where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} is the type parameter for reader effect and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{B} for the state effects.
+In our monad stack scheme, they are both instantiated with \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Cache}. Now the
+result type is a pair of two staged value: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Set[(Value, Store)]]} and
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Cache]}.
 \begin{lstlisting}
@@
 \iffalse
-Since our monad stack is uniformly using staged data, the staged @SetT@ now
-stores a staged value of type @Rep[Set[A]]@, inside of another staged monad @M@.
+Since our monad stack is uniformly using staged data, the staged \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} now
+stores a staged value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Set[A]]}, inside of another staged monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{M}.
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations} When deriving the staged concrete interpreter
-from the unstaged counterpart, we notice that recursive calls of @ev@ on the
-lambda body are shifted from @ap_clo@ to @close@. In other word, we need to
+from the unstaged counterpart, we notice that recursive calls of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} on the
+lambda body are shifted from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close}.
+In other word, we need to
 eagerly specialize the interpreter and generate code every time when we see a
@@
 
-The @close@ method now is a mixing of the \textit{staged} concrete version and
-unstaged \textit{abstract} version: we build a current-stage function @f@, which
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} method now is a mixing of the \textit{staged} concrete version and
+unstaged \textit{abstract} version: we build a current-stage function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, which
 takes four next-stage values, including the argument and latest store, and
-@in/out@ caches additionally. Inside of @f@, we collapse the @Ans@ monad to
-grounded values of type @Result@ by providing the desired arguments, i.e., the
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in/out} caches additionally.
+Inside of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, we collapse the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Ans} monad to
+grounded values of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Result} by providing the desired arguments, i.e., the
 new environment, new store, and caches. The collapsing of monad happens at
-current-stage, so the call of @ev@ is unfolded through staging. Finally, we
+current-stage, so the call of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} is unfolded through staging. Finally, we
 generate a singleton set containing the compiled next-stage closure
-@emit_compiled_clo(f)@, represented by an IR node in LMS.
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_compiled_clo(f)}, represented by an IR node in LMS.
 \begin{lstlisting}
@@
 
-The @ap_clo@ method is also similar: we use the staged version of @lift_nd@ to
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} method is also similar: we use the staged version of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{lift_nd} to
 lift the next-stage set of closures into the monad stack, and generate a
-next-stage value for each closure @clo@ in the set, by using @emit_ap_clo@. The
-method @emit_ap_clo@ generates the current-stage representation of the
+next-stage value for each closure \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{clo} in the set, by using \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo}.
+The
+method \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo} generates the current-stage representation of the
 future-stage result of application.
-Finally, we put the future @out@ cache, store, and values back into
+Finally, we put the future \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache, store, and values back into
 the current-stage monads.
@@
 \paragraph{Staged Caching and Fixpoint Iteration} Our fixed-point iteration
-again relies on the two caches @in@ and @out@, which are both staged maps now.
-Therefore, the query of whether the @out@ cache contains the current
+again relies on the two caches \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out}, which are both staged maps now.
+Therefore, the query of whether the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache contains the current
 configuration will produce a next-stage Boolean value, i.e., of type
-@Rep[Boolean]@. Consequently, the branching condition cannot be determined
-statically -- we need to generate code for @if@. Figure
-\ref{fig:staged_coind_cache} shows the staged version of @fix@. The
-monadic-binded variable @res@ is a next-stage result, consisting of a next-stage
-@if@ expression. The true branch simply returns a pair of queried result form
-the @out@ cache and the @out@ cache itself. The else branch constructs a monadic
-value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@ cache.
-After which, we use the similar technique that collapses the monad @m@ to
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Boolean]}. Consequently, the branching condition cannot be determined
+statically -- we need to generate code for \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{if}. Figure
+\ref{fig:staged_coind_cache} shows the staged version of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}. The
+monadic-binded variable \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{res} is a next-stage result, consisting of a next-stage
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{if} expression. The true branch simply returns a pair of queried result form
+the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache and the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache itself.
+The else branch constructs a monadic
+value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{m} of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Ans} first, which evaluates \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e} under the new \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache.
+After which, we use the similar technique that collapses the monad \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{m} to
 grounded values by providing its desired environment and etc. Finally, we have a
-current-stage representation of the future-stage values @res@, and put the
-content of @res@ back into the monad stack.
+current-stage representation of the future-stage values \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{res}, and put the
+content of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{res} back into the monad stack.
 
@@
 
-Although every component except @Expr@s are staged, we treat the data
-structures such as @Map@s as black-boxes, which means the operations on a @Map@
+Although every component except \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr}s are staged, we treat the data
+structures such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map}s as black-boxes, which means the operations on a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map}
 directly become code in the next stage, but we don't inspect any further inside.
-As we identified when introducing the generic interface, the keys of @Env@s are
+As we identified when introducing the generic interface, the keys of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Env}s are
 identifiers (i.e., strings) in the program, which are completely known
 statically. This leaves us a chance to further specialize the data structures.
-For example, let's assume that the @Map[K, V]@ is implemented as a hash map. If
-the keys of type @K@ are all known statically, then the indices also can be
+For example, let's assume that the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map[K, V]} is implemented as a hash map. If
+the keys of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{K} are all known statically, then the indices also can be
 computed statically. Thus the specialized map would be an array of type
-@Array[Rep[V]]@, whose elements are next-stage values; the size of the array is
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Array[Rep[V]]}, whose elements are next-stage values; the size of the array is
 also known statically. An access to the map is translated into an access to the
@@
 environment can be completely determined statically and generates the addresses
-directly. After which, the stores can be specialized as arrays of @Rep[Value]@
+directly. After which, the stores can be specialized as arrays of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Value]}
 elements.
@@
 
-Our treatment to binding-times is coarse-grained: @Expr@s is static, the rest of
+Our treatment to binding-times is coarse-grained: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr}s is static, the rest of
 the world are all dynamic. But this is not always true since the static data
@@
 
-For example, to fold a singleton set (often appears in @SetT@), e.g.,
-@Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully apply @foldLeft@ to the set with an anonymous
-function. But we can also utilize the algebraic property of @foldLeft@ to
-generate cheaper code: @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
-function @f@ is known at the current stage, we completely eliminate the fold
+For example, to fold a singleton set (often appears in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT}), e.g.,
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set(x).foldLeft(init)(f)} where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{init} are staged values, a naive code
+generator would faithfully apply \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{foldLeft} to the set with an anonymous
+function. But we can also utilize the algebraic property of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{foldLeft} to
+generate cheaper code: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set(x).foldLeft(init)(f) = f(init, x)}. Since the
+function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} is known at the current stage, we completely eliminate the fold
 operation and function application. We apply several rewritings enabled by
-partial-static data structures, such as @Set@ and @Map@, which greatly reduces
+partial-static data structures, such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Set} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map}, which greatly reduces
 the size of residual programs.
@@
 staged. All the monadic computation, i.e., functions passed to the monadic bind
-operation @flatMap@, are statically known. This is why we can eliminate the
+operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}, are statically known. This is why we can eliminate the
 monadic layer and its associated overhead.
@@
 \paragraph{What has been eliminated?} In the generated code, all the
-primitive operations (such as @eval@, @fix@, @ap_clo@, etc.) and monadic
-operations (such as @flatMap@ and @map@) are eliminated. The residual program
+primitive operations (such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}, etc.) and monadic
+operations (such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}) are eliminated.
+The residual program
 consists of statements and expressions that purely manipulate the environment,
-store, and two caches, whose underlying representations are all @Rep[Map[K,V]]@. We
+store, and two caches, whose underlying representations are all \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Map[K,V]]}. We
 also have several operations on tuples and lists, which are residualized from

### 2019-08-15 paper/60_cases.tex
at sign

@@
 implementation: different syntactic constructs are handled in different functions
-(e.g., @compCall@, @compApp@, etc.).
+(e.g., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{compCall}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{compApp}, etc.).
 Before refactoring into AC, these functions would take the store as a
-regular argument; now every function returns a value of type @CompAnalysis@,
+regular argument; now every function returns a value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompAnalysis},
 which is a function (closure) value that takes and returns stores. After the
@@
 On the other side, our approach does exactly the same thing through staging: the
-AST are static, and stores are dynamic, therefore we annotate the type @Store@
-with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
-changes of our approach using LMS framework are the types of @Store@ to
-@Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
+AST are static, and stores are dynamic, therefore we annotate the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}
+with \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
+changes of our approach using LMS framework are the types of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store} to
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Store]}, since they will be known at the next stage; other parts do not need to change. 
 The generated code
@@
 functions, which provides additional performance improvement. Nevertheless,
-the support from LMS framework is necessary, such as staging support for @Map@,
+the support from LMS framework is necessary, such as staging support for \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map},
 we consider these efforts are reusable and complete orthogonal to the
@@
 The timestamp is essentially a finite list of expressions that tracks the $k$
-recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
+recent calling contexts. The definition of abstract addresses \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{KCFAAddr} is also
 changed to contain a identifier and the time when it get allocated, meaning that
@@
 We then integrate the timestamp into our monad stack by adding a staged
-@StateT@ monad. The result type is passingly get updated to contain a set of
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} monad. The result type is passingly get updated to contain a set of
 tuples, where the time is added.
@@
 
-Every time when we call the @eval@ function, we refresh the timestamp by calling
-the @tick@ function, which update the timestamp in the state monad. The @tick@
-function is implemented as prepending the current expression being evaluated @e@
+Every time when we call the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} function, we refresh the timestamp by calling
+the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick} function, which update the timestamp in the state monad.
+The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick}
+function is implemented as prepending the current expression being evaluated \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}
 to the existing calling context, and then taking the first $k$ elements from the list.
-The type @Config@ used in caching is also changed to included the timestamp.
+The type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Config} used in caching is also changed to included the timestamp.
 
@@
 DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
-implemented in monadic abstract interpreter by swapping the @StateT@ (for store)
-and the @SetT@ (for nondeterminism). In this section, we show this optimization
+implemented in monadic abstract interpreter by swapping the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} (for store)
+and the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} (for nondeterminism). In this section, we show this optimization
 is orthogonal to staging and we just need to swap the same \textit{staged}
@@
 \cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
-shows that, for big-step monadic abstract interpreter, we may add a @Reader@
+shows that, for big-step monadic abstract interpreter, we may add a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Reader}
 monad to track a set of root addresses and compute the reachable addresses every
@@
 
-The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the @fix@
-function with caching (from Section \ref{sai}) to @fix_gc@, but the caching part
-is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
-current root addresses from the monad stack; @root@ compute the reachable
-addresses from a value; and @gc@ cleans up the store according to the reachable
+The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}
+function with caching (from Section \ref{sai}) to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix_gc}, but the caching part
+is omitted. Several auxiliary functions are used: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ask_root} retrieves the
+current root addresses from the monad stack; \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{root} compute the reachable
+addresses from a value; and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{gc} cleans up the store according to the reachable
 addresses and returns a new store. For the generic interpreter, we also need to
@@
 As we have already seen from Section \ref{sai}, and recall the
-primitive operation @close@ that denotes a $\lambda$ term to next-stage
+primitive operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} that denotes a $\lambda$ term to next-stage
 Scala function.  The specializing of staged abstract interpreter to
@@
 
-Consider there is a library that contains a list map function @map@, 
+Consider there is a library that contains a list map function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}, 
 \begin{lstlisting}
@@
 
-The specialization (0CFA and without store-widening) of @close@ on
-@map@ generates the following next-stage Scala code, that is, the
-function @x39@:
+The specialization (0CFA and without store-widening) of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} on
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map} generates the following next-stage Scala code, that is, the
+function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x39}:
 \begin{lstlisting}
@@
 more for calling context by partially-static data optimization. For example,
-consider that we have a library consists of three functions @f@, @g@ and @h@,
-where @f@ calls @g@ and @h@ internally and returns @y@ finally (we extend the
-language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
-would like to modularly specialize these functions, starting from @f@ with an
-initial environment containing @f@, @g@ and @h@.
+consider that we have a library consists of three functions \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h},
+where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} calls \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h} internally and returns \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{y} finally (we extend the
+language to have multiple arguments).
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h} do not depend on \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}.
+Now we
+would like to modularly specialize these functions, starting from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} with an
+initial environment containing \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h}.
 \begin{lstlisting}
@@
 Since $k$ is bounded to a fixed number, after the analysis running $k$ steps
-inside of @f@, the calling context will purely depend on the static structure of
+inside of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, the calling context will purely depend on the static structure of
 our module being analyzed.
@@
 To evaluates an assignment, we first evaluate its right-hand side, and then put
-the value into the slot where the address of @v@ points to. For simplicity, we
-elect to make the value of the assignment be an @void()@ value.
+the value into the slot where the address of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{v} points to. For simplicity, we
+elect to make the value of the assignment be an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{void()} value.
 \begin{lstlisting}
@@
 
-To evaluates a @while@ loop statement, we evaluate the condition first. Then
-similar to how we treat @branch0@, we have a generic @branch@ function but works
-on Boolean values. For the @true@ branch, we recursively call @ev@ on a newly
-constructed expression @Seq(e, While(t, e))@ meaning that first evaluates @e@
+To evaluates a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{while} loop statement, we evaluate the condition first. Then
+similar to how we treat \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{branch0}, we have a generic \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{branch} function but works
+on Boolean values.
+For the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{true} branch, we recursively call \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} on a newly
+constructed expression \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Seq(e, While(t, e))} meaning that first evaluates \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}
 and the repeat the loop. Otherwise, for the other branch, we simply return a
@@
 fields, an upper bound and a lower bound. Here we annotate the upper bound and
-lower bound fields to be of type @Rep[Double]@. The operations such as @+@ of
+lower bound fields to be of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Double]}.
+The operations such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{+} of
 two intervals produce a new

### 2019-08-15 paper/70_eval.tex
at sign

@@
 analysis.  Some of the benchmarks are small programs designed for
-experiments, such as the @kcfa-worst-@$n$ series, which are intended
+experiments, such as the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{kcfa-worst-}$n$ series, which are intended
 to challenge $k$-CFA; while some other benchmarks are from real-world
 applications, for examples, the RSA public key encryption algorithm
-@rsa@. In the Figure \ref{evaluation_result}, we report the number of
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rsa}. In the Figure \ref{evaluation_result}, we report the number of
 AST node after parsing and desugaring (excluding comments) as a proper

### 2019-08-15 paper/80_related.tex
Related work

@@
 compilation (AC) \cite{Boucher:1996:ACN:647473.727587}: removing the
-interpretation overhead on traversing the syntax tree by
+interpretative overhead on traversing the syntax tree by
 specialization. Specifically, the residual program of AC can be either
-textual or closures. However, as we studied in Section \ref{cs_ac},
-with the perspective from generative programming and monadic
-interpreters, we make it more general, extensible and easy.
+text or closures. However, as we studied in Section \ref{cs_ac},
+with the perspective of generative programming and monadic
+interpreters, we make it more general and extensible, and easier.
 \citet{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
 generation to optimize small-step abstract interpreter in
-state-transition style. The analyzed program is firstly compiled to an
-IR called "abstract bytecode", which are actually higher-order
-functions and will be executed later on an abstract abstract machine
+state-transition style. The analyzed program is first compiled to an
+IR called "abstract bytecode", which is actually a series of higher-order
+functions which will be executed later on an abstract abstract machine
 for that IR. \citet{damian1999partial} provides a formal treatment to
@@
 
-Splitting an analysis into multiple stages is also studied for
-analysis other than abstract-interpretation-based control-flow
-analysis, though the formulation may very
+Splitting an analysis into multiple stages has also been studied for
+analysis other than abstract-interpretation-based control flow
+analysis, though the formulation is quite
 different. \citet{DBLP:conf/cgo/HardekopfL11} apply staging to a
@@
 analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
-Similar to the idea in this paper, such analyses can be splitted to
-staged.  One of the future work is to explore that whether these
+Similar to the idea in this paper, such analyses can be split to
+staged.  One piece of future work is to explore whether these
 analyses can be implemented and optimized from
 meta-programming. Souffl{\'e}, one of the state-of-the-art Datalog
-engine that is commonly used in points-to analysis
+engines that is commonly used in points-to analysis
 \cite{Antoniadis:2017:PDS:3088515.3088522}, uses the idea of Futamura
-projection for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
-But it only specializes Datalog rules to the running engine, instead
-of a full pipeline from program AST to generating constraints and
+projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
+However, it only specializes Datalog rules to the running engine, instead
+of a full pipeline from program AST to generating and
 solving constraints.
@@
 Recent work shows that the numerical abstract domains can be
-specialized w.r.t. the static program structure, although not in the
-way of using staging. For example, decomposing polyhedrons
+specialized w.r.t. the static program structure, although not by
+way of staging. For example, decomposing polyhedrons
 \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to
@@
 \paragraph{Construction of Abstract Interpreters} Abstract interpretation
-was proposed as a semantic-based approach to build sound static analysis by
-approximation \cite{DBLP:conf/popl/CousotC77}. As to build semantic
+was proposed as a semantics-based approach to build sound static analysis by
+approximation \cite{DBLP:conf/popl/CousotC77}. As for building semantic
 artifacts, the Abstracting Abstract Machines (AAM)
@@
 definitional abstract interpreters,
-\citet{Wei:2018:RAA:3243631.3236800} reconstruct big-step abstract
-interpreters with delimited continuations;
-\citet{Keidel:2018:CSP:3243631.3236767} present a modular arrow-based
-abstract interpreter that makes proof of soundness can be constructed
+the reconstruction of big-step abstract interpreters with delimited
+continuations shown by \citet{Wei:2018:RAA:3243631.3236800},
+\citet{Keidel:2018:CSP:3243631.3236767} present a modular, arrow-based
+abstract interpreter that allows proof of soundness to be constructed
 compositionally.
@@
 modular analysis. We borrow the notation of modular analysis for our approach.
-But the staged abstract interpreter does not generate complete summaries for
-modules, instead, it is able to specialize the analysis modularly. Similar, AC
-is considered as an example of this kind of mdoular analysis \cite{DBLP:conf/cc/CousotC02}.
+However the staged abstract interpreter does not generate complete summaries for
+modules; instead, it is able to specialize the analysis modularly. Similar, AC
+is considered as an example of this kind of modular analysis \cite{DBLP:conf/cc/CousotC02}.
 
@@
 interpreters is also closely related to the two-level semantics: the
-use of a generic interface with monads and then reinterpret it by
+use of a generic interface with monads and then reinterpreting it with
 different semantics is already two-level. Instead of focusing on
 semantics, this paper shows how a staged analyzer can be built and
-used to increase efficiency of static analysis, we augment the monadic
+used to increase efficiency of static analysis. We augment the monadic
 abstract interpreter from another dimension, using code generation to
 produce efficient low-level code. Hence the presented work can be
-considered as an two-dimension application of the two-level semantics.
+considered as an two-dimensional application of the two-level semantics.
 
@@
 type-level stage annotations.  Other notable implementations of MSP
-exist in ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97} and
+exist in the ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97} and
 MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
@@
 
-As an source of inspiration of this paper, Futamura projections
-reveals a hierarchy and close relation between interpreters and
+As a source of inspiration of this paper, the Futamura projections
+reveal a hierarchy and close relationship between interpreters and
 compilers, which was originally proposed by
-\citeauthor{futamura1971partial} in 1970s\cite{futamura1971partial},
+\citeauthor{futamura1971partial} in the 1970s\cite{futamura1971partial},
 and later reprinted \cite{Futamura1999}.
-\citeauthor{Amin:2017:CTI:3177123.3158140} consider a tower of
+\citeauthor{Amin:2017:CTI:3177123.3158140} considers a tower of
 concrete interpreters and how to collapse them by using MSP -- it
@@
 was explored by \citet{danvy1991compiling, DBLP:conf/dsl/SheardBP99},
-but different from the performance motivation here, the generated code
+but is different from the performance motivation here: the generated code
 is in monadic form.

### 2019-08-15 paper/10_intro.tex
update intro

--- 2019-08-15 paper/10_intro.tex
@@
 
-Statically analyzing semantic properties of a program is a
-widely-known undecidable problem. Abstract interpretation as a
-lattice-based approach to sound static analyses was proposed by
-\citet{DBLP:conf/popl/CousotC77}. Equipped with Galois connections,
-the analyzer can obtain the program runtime behaviors approximately by
-computing the fixed points on the abstract domain. Despite the
-tremendous theoretical development of abstract interpretation over the
-years, constructing artifacts and analyzers that perform sound
-abstract interpretation for modern and expressive languages was
-considered abstruse and complicated for a long time.
+Abstract interpretation as a lattice-based approach to sound static
+analyses was proposed by \citet{DBLP:conf/popl/CousotC77}. Equipped
+with Galois connections \cite{CousotCousot79-1}, the analyzer can
+soundly approximate the program runtime behaviors by computing fixed
+points on the abstract domain. Despite the tremendous theoretical
+development of abstract interpretation over the years, constructing
+artifacts and analyzers that perform sound abstract interpretation for
+modern and expressive languages was considered complicated for a long
+time.
 
-In recent years, we observe rich progress on the methodologies for
-constructing abstract interpreters from systematic principles rather than
-ad-hoc engineering. Of particular note is the Abstracting Abstract
-Machine (AAM) methodology \cite{DBLP:journals/jfp/HornM12,
-  DBLP:conf/icfp/HornM10}. This methodology uncovers an approach to derive sound
-abstract interpreters from their concrete counterparts, for example,
-abstract machines, where the soundness can be easily established by
-examining the transformation of semantic artifacts. For example, the
-CEK machine \cite{DBLP:conf/popl/FelleisenF87} for concrete execution
-can be readily refactored to an effective $0$-CFA control-flow
-analysis \cite{Shivers:1988:CFA:53990.54007,
-  Midtgaard:2012:CAF:2187671.2187672}: first, tweak the environment
-dereference as a nondeterministic choice, such that the environment
-may contain multiple possible values for a variable, then allocate
-continuations in the environment to exploit return-flow information,
-and also constrain the addresses space to be finite.  This systematic
-abstraction approach can be tailored to different language features
-(such as states, first-class controls, exceptions, and concurrency) and
-sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
+Recent progress on methodologies such as the Abstracting Abstract
+Machine (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
+uncovers an approach to construct abstract interpreters from
+systematic principles rather than ad-hoc engineering.  By deriving
+semantics artifacts for abstract interpretation from theire concrete
+counterparts, for example, abstract machines, the soundness can be
+easily established by examining the transformation of semantic
+artifacts.
+This systematic abstraction approach can be tailored to different
+language features (such as state, first-class control, exception, and
+concurrency) and sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
   DBLP:conf/popl/GilrayL0MH16, Darais:2015:GTM:2814270.2814308}. It
@@
 emerged. Such techniques include monads, monad transformers
-\cite{DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979},
-arrows \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
+\cite{DBLP:journals/pacmpl/DaraisLNH17,
+  Sergey:2013:MAI:2491956.2491979}, arrows
+\cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
 \cite{Githubsemantic}, etc. The pure approaches provide certain
-benefits, including that the abstract interpretation artifacts can be built
-compositionally and modularly, e.g., by using monad
+benefits, including that the abstract interpretation artifacts can be
+built compositionally and modularly, e.g., by using monad
 transformers. Therefore, the soundness of analysis can be proven with
-less effort, either by mechanized \cite{Darais:2016:CGC:2951913.2951934}
-or paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}. Also,
-referential transparent and equational reasoning allow programmers to
-reason the correctness of their implementations more confidently.
+less effort, by either mechanized proof
+\cite{Darais:2016:CGC:2951913.2951934} or paper-based proofs
+\cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
+transparency and equational reasoning allow programmers to reason the
+correctness of their implementations more confidently.
 
@@
 the abstract interpreter analyzes the program by traversing the
-abstract syntax tree (AST), which poses an interpretive overhead, such as
-the pattern matching on the ASTs and recursive calls on the sub-expressions.
-Such overhead can be negligible if the abstract
-interpreter only runs on the program a single time, but can also
-accumulate significantly if run repeatedly (for example, on
-libraries).  Second, the abstract interpreter written in pure languages
-usually extensively uses effect systems to implement the semantics of
-abstract interpretation.  For example, the abstract interpreter that
-returns all possible runtime values is a form of nondeterminism, where
-nondeterminism monads can help.  Although such pure approaches have
-their own merits and elegance, compared with imperative stateful
-implementations, they are significantly slower.
+abstract syntax tree (AST), which poses an interpretive overhead, such
+as the pattern matching on the ASTs and recursive calls on the
+sub-expressions.
+Such overhead can be negligible if the abstract interpreter only runs
+on the program a single time, but can also accumulate significantly if
+run repeatedly (for example, on libraries).  Second, it is common that
+abstract interpreters written in pure languages extensively use effect
+systems to implement the semantics of abstract interpretation. For
+example, the abstract interpreter that returns all possible runtime
+values is a form of nondeterminism, where nondeterminism monads can
+help.  Although such pure approaches have their own merits, compared
+with imperative stateful implementations, they are significantly
+slower.
 
-In this paper, we propose an abstraction-without-regret approach to
-eliminating performance penalties for abstract interpreters,
-while still keeping the benefits inherent in purely functional
-programming.  In short, our approach applies ideas and techniques from
-program specialization and embedded DSLs to abstract interpreters. 1)
-Using multi-stage programming , we can specialize the abstract
+Roughly at the same when abstract interpretation was proposed in
+1970s, \citet{futamura1971partial, Futamura1999} revealed a colse
+connection between interpreters and compilers through a hierarchy of
+specializations, namely, Futamura Projections. The first Futamura
+projection shows that specializing an interpreter to the input program
+removes interpretation overhead and yields an equivalent executable
+that runs faster.  The challenge here is can we apply the same idea to
+general abstract interpreters to make it run faster?
+To achieve this, we need to handle a non-standard semantics that
+extensively uses nondeterminism, operates on abstract domains, and
+computes fixed-point iteratively with a cache.
+
+In this paper, we show how to naturally extend the first Futamura
+projection to abstract interpreters. We present an
+abstraction-without-regret framework to eliminating performance
+penalties for abstract interpreters, while keeping the benefits
+inherent in purely functional programming. In short, our approach
+borrows ideas and techniques from multi-stage programming and embedded
+DSLs to abstract interpreters.  1) To remove the overhead from
+interpretation and effect layers, we specialize the abstract
 interpreter with respect to an input program and then generate
-efficient low-level code that does the actual analysis.  The result of
-specialization is reusable, and the effect layers have been eliminated
-in the generated code. 2) Inspired by the tagless-final interpreters,
-we define a generic interpreter that abstracts over binding-time and
-different semantics, which allows user to implement different
-semantics modularly, including the staged abstract interpretation
-semantics.  Together with the multi-stage programming and type-level
-annotations of binding-times, we can derive the staged abstract
-interpreters without intrusive changes to the corresponding unstaged
-one, thus the soundness is untouched.  In this sense, our approach is
-no regret of both performance and engineering effort. We elaborate
-these two main ideas in detail as follow.
+efficient low-level code that does the actual analysis.  2) Inspired
+by the tagless-final interpreters
+\cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic
+interpreter that abstracts over binding-time and different semantics,
+which allows user to implement different semantics modularly,
+including the staged abstract interpretation semantics. As the result,
+we can derive the staged abstract interpreters without intrusive
+changes to its corresponding unstaged one and preserve the soundness
+of analysis. In this sense, our approach is no regret of both
+performance and engineering effort. We elaborate these two main ideas
+in detail as follows.
+
+%The result of specialization is reusable, and the effect layers have
+%been eliminated in the generated code.
 
@@
 
-The idea of specializing interpreters can be traced to Futamura
-projections \cite{Futamura1999, futamura1971partial}.  The first
-Futamura projection specifically shows that specializing an
+The first Futamura projection specifically shows that specializing an
 interpreter w.r.t. the input program yields an equivalent
-executable. For instance, @eval@ is an interpreter for some language:
+executable. For instance,
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval}
+is an interpreter for some language:
 \begin{lstlisting}
@@
 \end{lstlisting}
-Given a program $e_0$ of type @Expr@, by applying the specialization,
-we can obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$
-: @Input@ $\to$ @Value@, which is the so-called \textit{equivalent
-  executable}.  By definition of the interpreter, they produce the
-same result when applied with the argument
-$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
+Given a program $e_0$ of type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr},
+by applying the specialization, we can obtain a specialized
+interpreter $\texttt{eval}_{\texttt{e0}}$ :
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Input}
+$\to$
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value},
+which is the so-called \textit{equivalent executable}.
+
+% By definition of the interpreter, they produce the same result when applied with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
 
@@
 it is hard to precisely analyze binding-times given an arbitrary
-program.  As an alternative approach to specialization, multi-stage
-programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets
-the programmer to explicitly control and annotate the binding-times in
-the program, then the MSP system will check whether these annotations
+program. As an alternative approach, multi-stage programming (MSP)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly
+programming over a two-level semantics \cite{NIELSON198859},
+by manually annotating the binding-times of variables in the program.
+Then the MSP system will check whether these annotations
 are consistent and specialize the program using that information. The
@@
 
-Our proposed framework adopts the type-based multi-stage programming
-from the Lightweight Modular Staging framework and implements the
-Futamura projection of a big-step abstract interpreter for a small
-stateful, higher-order language.  We use
-\citet{DBLP:journals/pacmpl/DaraisLNH17}'s abstract definitional
-interpreter that uses monad transformers as the unstaged baseline. By
-deriving staged monads that can be used to do code generation, we may
-obtain a staged abstract interpreter that fulfills the specialization.
-After staging, the generated code is specialized to the input program,
-and all monadic operations are all inlined and compiled down to
-low-level Scala code.
+Similar to the two-level semantics \cite{NIELSON1989117}, we found
+monads provide a good abstraction to hide the detail of abstract
+interpretation semantics, and allows us to fit code generation into
+the semantics. We use \citet{DBLP:journals/pacmpl/DaraisLNH17}'s
+abstract definitional interpreter that uses monad transformers as the
+unstaged baseline.  Our proposed framework adopts the type-based
+multi-stage programming from the Lightweight Modular Staging framework
+and implements the Futamura projection of a big-step abstract
+interpreter for a small stateful, higher-order language. By deriving
+staged monads that can be used to do code generation, we obtain a
+staged abstract interpreter that fulfills the specialization. After
+staging, the generated code is specialized to the input program, and
+all monadic operations are all inlined and compiled down to low-level
+Scala code.
 
@@
 interpreter and achieve the flexibility of reinterpreting the shared
-interpreter. We adapt the monadic interpreter schema, which already
-have been applied to abstract interpreters, and introduce the
+interpreter. Generic interpreters using monads have already been
+applied to abstract interpreters, we further introduce the
 binding-time abstraction into the generic interpreter.  The generic
@@
 values directly or generates code. It is worth to mention that the
-binding-time type is also instrumented into the monadic type, so that
-we will distinguish normal monads and staged monads.
+binding-time type is also injected into the monadic type, so that we
+will distinguish normal monads and staged monads.
 

### 2019-08-15 paper/40_absinterp.tex
update intro

@@
 Then we show the abstract components, namely abstract values, abstract
-environments, and abstract stores. The abstract interpreter we construct in this
+environments, and abstract stores.
+The abstract interpreter we construct in this
 section is similar to \citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For
@@
 or staged abstract interpreter.
+
+\iffalse
+For example, the CEK machine \cite{DBLP:conf/popl/FelleisenF87} for
+concrete execution can be readily refactored to an effective $0$-CFA
+control-flow analysis \cite{Shivers:1988:CFA:53990.54007,
+  Midtgaard:2012:CAF:2187671.2187672}: first, tweak the environment
+dereference as a nondeterministic choice, such that the environment
+may contain multiple possible values for a variable, then allocate
+continuations in the environment to exploit return-flow information,
+and also constrain the address space to be finite.
+\fi
 
@@
 
-In out monad stack, the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache modeled by a reader effect and
+In our monad stack, the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{in} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{out} cache modeled by a reader effect and
 state effect, respectively. We first defined several methods to manipulate

### 2019-08-16 paper/10_intro.tex
misc
Proofread intro

@@
 Machine (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
-uncovers an approach to construct abstract interpreters from
-systematic principles rather than ad-hoc engineering.  By deriving
-semantics artifacts for abstract interpretation from theire concrete
-counterparts, for example, abstract machines, the soundness can be
+uncovers an approach to constructing abstract interpreters from
+systematic principles, rather than ad-hoc engineering.  By deriving
+semantics artifacts for abstract interpretation from their concrete
+counterparts (for example, abstract machines), the soundness can be
 easily established by examining the transformation of semantic
@@
 transformers. Therefore, the soundness of analysis can be proven with
-less effort, by either mechanized proof
+less effort, whether by mechanized proofs
 \cite{Darais:2016:CGC:2951913.2951934} or paper-based proofs
@@
 the abstract interpreter analyzes the program by traversing the
-abstract syntax tree (AST), which poses an interpretive overhead, such
+abstract syntax tree (AST), which incurs an interpretive overhead, such
 as the pattern matching on the ASTs and recursive calls on the
-sub-expressions.
+subexpressions.
 Such overhead can be negligible if the abstract interpreter only runs
-on the program a single time, but can also accumulate significantly if
+on the program a single time, but can accumulate significantly if
 run repeatedly (for example, on libraries).  Second, it is common that
@@
 Roughly at the same when abstract interpretation was proposed in
-1970s, \citet{futamura1971partial, Futamura1999} revealed a colse
+1970s, \citet{futamura1971partial, Futamura1999} revealed a close
 connection between interpreters and compilers through a hierarchy of
@@
 removes interpretation overhead and yields an equivalent executable
-that runs faster.  The challenge here is can we apply the same idea to
-general abstract interpreters to make it run faster?
-To achieve this, we need to handle a non-standard semantics that
+that runs faster.  The challenge here: can we apply the same idea to
+general abstract interpreters to make hem run faster?
+To achieve this, we need to handle a nonstandard semantics that
 extensively uses nondeterminism, operates on abstract domains, and
@@
 
-In this paper, we show how to naturally extend the first Futamura
-projection to abstract interpreters. We present an
-abstraction-without-regret framework to eliminating performance
-penalties for abstract interpreters, while keeping the benefits
-inherent in purely functional programming. In short, our approach
-borrows ideas and techniques from multi-stage programming and embedded
-DSLs to abstract interpreters.  1) To remove the overhead from
-interpretation and effect layers, we specialize the abstract
-interpreter with respect to an input program and then generate
-efficient low-level code that does the actual analysis.  2) Inspired
-by the tagless-final interpreters
-\cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic
-interpreter that abstracts over binding-time and different semantics,
-which allows user to implement different semantics modularly,
-including the staged abstract interpretation semantics. As the result,
-we can derive the staged abstract interpreters without intrusive
-changes to its corresponding unstaged one and preserve the soundness
-of analysis. In this sense, our approach is no regret of both
-performance and engineering effort. We elaborate these two main ideas
-in detail as follows.
+In this paper, we show that the first Futamura projection can be naturally
+extended to abstract interpreters. We present an abstraction-without-regret
+framework to eliminate performance penalties for abstract interpreters, while
+keeping the benefits inherent in purely functional programming. In short, our
+approach borrows ideas and techniques from multi-stage programming and embedded
+DSLs to abstract interpreters:  1) To remove the overhead from interpretation
+and effect layers, we specialize the abstract interpreter with respect to an
+input program and then generate efficient low-level code that does the actual
+analysis.  2) Inspired by the tagless-final interpreters
+\cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
+abstracts over binding-time and different semantics, which allows user to
+implement different semantics modularly, including the staged abstract
+interpretation semantics. As a result, we can derive the staged abstract
+interpreters without intrusive changes to its corresponding unstaged
+counterpart, and also preserve the soundness of analysis. In this sense, our
+approach allows for no regret for both performance and engineering effort. We
+elaborate these two main ideas in detail, as follows.
 
@@
 
-Partial evaluation \cite{DBLP:books/daglib/0072559} was the first
-proposed approach to realize Futamura projections: it first discovers
-the binding-times (\textit{static} or \textit{dynamic}) of variables,
-then evaluates the static part of the program, and generates a
-residualized program that solely relies on the dynamic part. However,
-it is hard to precisely analyze binding-times given an arbitrary
-program. As an alternative approach, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly
-programming over a two-level semantics \cite{NIELSON198859},
-by manually annotating the binding-times of variables in the program.
-Then the MSP system will check whether these annotations
-are consistent and specialize the program using that information. The
-staging annotations can be either syntactic (e.g., quote and
-quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
-Modular Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
+Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed
+approach to realize Futamura projections: it first discovers the binding-times
+(\textit{static} or \textit{dynamic}) of variables, then evaluates the static
+part of the program, and generates a residualized program that solely relies on
+the dynamic part. However, it is hard to precisely analyze binding-times given
+an arbitrary program. As an alternative approach, multi-stage programming (MSP)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly program over
+a two-level semantics \cite{NIELSON198859}, by manually annotating the
+binding-times of variables in the program. Then, the MSP system will check
+whether these annotations are consistent, and specialize the program using that
+information. The staging annotations can be either syntactic (e.g., quote and
+quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular
+Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
 
-Similar to the two-level semantics \cite{NIELSON1989117}, we found
+Similar to two-level semantics \cite{NIELSON1989117}, we found
 monads provide a good abstraction to hide the detail of abstract
@@
 unstaged baseline.  Our proposed framework adopts the type-based
-multi-stage programming from the Lightweight Modular Staging framework
+multi-stage programming from the Lightweight Modular Staging framework,
 and implements the Futamura projection of a big-step abstract
-interpreter for a small stateful, higher-order language. By deriving
+interpreter for a small, stateful, higher-order language. By deriving
 staged monads that can be used to do code generation, we obtain a
@@
 staging, the generated code is specialized to the input program, and
-all monadic operations are all inlined and compiled down to low-level
+all monadic operations are inlined and compiled down to low-level
 Scala code.
@@
 \begin{itemize}
-\item With a flat binding-time and concrete domains, it is a ordinary
+\item With a flat binding-time and concrete domains, it is an ordinary
   definitional interpreter based on big-step operational semantics;
-\item With two-level binding-times and concrete domain, it is a
-  compiler that translate to program intro another language;
-\item With a flat binding-time and abstract domains, it is an
+\item With two-level binding-times and concrete domains, it is a
+  compiler that translates a program into another language;
+\item With a flat binding-time and abstract domains, it is a
   definitional abstract interpreter \cite{DBLP:journals/pacmpl/DaraisLNH17}
@@
 \item With two-level binding-times and abstract domains, it is an optimizing
-  program analyzer but works in the fashion of compilation.
+  program analyzer, but works in the fashion of compilation.
 \end{itemize}
 
-Although the four artifacts may look like very differently at first
-glance, but in fact are all firmly rooted in the concrete semantics of
+Although the four artifacts may look like dissimilar at first
+glance, they in fact are all firmly rooted in the concrete semantics of
 the language.  This observation provides a way to abstract over the
@@
 interpreter. Generic interpreters using monads have already been
-applied to abstract interpreters, we further introduce the
+applied to abstract interpreters; we further introduce the
 binding-time abstraction into the generic interpreter.  The generic
@@
 different semantics. The domain of the interpreter and the effects
-such as read, state and nondeterminism can be both wrapped into this
+such as read, state, and nondeterminism can all be wrapped into this
 monadic type.  The binding-time abstraction is represented by a
 higher-kinded type, and controls whether the interpreter produces
-values directly or generates code. It is worth to mention that the
+values directly or generates code. It is worth mentioning that the
 binding-time type is also injected into the monadic type, so that we
@@
 1) We compare our approach with abstract compilation
-\cite{Boucher:1996:ACN:647473.727587}, a implementation technique for
+\cite{Boucher:1996:ACN:647473.727587}, an implementation technique for
 control-flow analyses, and show that by utilizing type-based stage
-annotations we can achieve the same optimization, and meanwhile,
-the analyzer does not need to change, thereby requires much less
-engineering efforts.
+annotations we can achieve the same optimizations. Meanwhile,
+the analyzer does not need to change, thereby requiring significantly less
+engineering effort.
 2) We extend the basic staged abstract interpreter to different flow
 analyses, including a store-widened analysis, a context-sensitive
-analysis and abstract garbage collection.
-3) We notice that staging an abstract interpreter enables modularly
-compiling an analysis to programs. Here we borrow the concept of
+analysis, and abstract garbage collection.
+3) We show that staging an abstract interpreter enables modular
+compilation of an analysis to programs. Here we borrow the concept of
 modular analysis, and show that the compiled analysis is reusable.
-Therefore the approach provides a modular way to create optimized
-analysis code by mechanized reusing a whole-program analyzer.
-4) We also empirically evaluate the performance improved by staging,
-showing an order of magnitude speed-up on flow-analysis.
+Therefore, the approach provides a modular way to create optimized
+analysis code by mechanized reuse of a whole-program analyzer.
+4) We also empirically evaluate the performance improvements gained by staging,
+showing an order of magnitude speedup on flow-analysis.
 
-\paragraph{Contributions} Briefly, the contribution of the paper is as follows:
+\paragraph{Contributions} Briefly, the contributions of the paper are as follows:
 \begin{itemize}[leftmargin=2em]
-  \item To our best knowledge, we present the first one to use
-    multi-stage programming to specialize a general abstract interpreter,
-    and our approach does not touch the soundness of analyses.
+  \item To our best knowledge, we present the first
+    multi-stage programming system to specialize a general abstract interpreter,
+    without our approach affecting the soundness of analyses.
   \item Intellectually, our framework naturally extends the first
     Futamura projection to abstract interpreters, showing a
-    well-grounded approach to optimize static analyses via
+    well-grounded approach to optimizing static analyses via
     meta-programming.
@@
     monads in Scala, and then presenting the generic interpreter
-    (Section~\ref{prelim}).  After which, we review its instantiations
+    (Section~\ref{prelim}), after which we review instantiations
     of concrete interpretation (Section~\ref{unstaged_conc}) and
@@
   \item We present the unstaged abstract interpreter under the same
-    framework by replacing the environment, store and values to their
-    abstract counterparts (Section~\ref{unstaged_abs}). Then we
-    show the combination of approximation and specialization, dubbed
+    framework by replacing the environment, store, and values to their
+    abstract counterparts (Section~\ref{unstaged_abs}). We then
+    show that the combination of approximation and specialization, dubbed
     \textit{staged abstract interpreters}, can be readily derived
     (Section~\ref{sai}). We also summarize the approach and discuss
-    the soundness issue after showing the four artifacts.
+    soundness properties after showing the four artifacts.
   \item We conduct three case studies (Section~\ref{cases_study}) and
-    an empirical evaluation on performance improved (Section~\ref{evaluation}).
+    an empirical evaluation on performance improvements (Section~\ref{evaluation}).
     Finally we discuss related work and conclude.

### 2019-08-16 paper/abstract.tex
Grammar

@@
-It is well known that a staged interpreter is a compiler: specializing the
+It is well known that a staged interpreter is a compiler: specializing and
 interpreter to a given program produces an equivalent executable that runs faster.
@@
 It is even more widely known that an abstract interpreter is a program analyzer:
-tweaking the interpreter to run on an abstract domain produces a sound static
+tweaking an interpreter to run on an abstract domain produces a sound static
 analysis. What happens when we combine these two ideas, and apply staging to

### 2019-08-16 paper/10_intro.tex
Grammar
update

@@
 with Galois connections \cite{CousotCousot79-1}, the analyzer can
-soundly approximate the program runtime behaviors by computing fixed
-points on the abstract domain. Despite the tremendous theoretical
+soundly approximate the program runtime behavior by computing fixed
+points on an abstract domain. Despite the tremendous theoretical
 development of abstract interpretation over the years, constructing
@@
 
-Recent progress on methodologies such as the Abstracting Abstract
+Recent progress on methodologies such as Abstracting Abstract
 Machine (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
@@
 semantics artifacts for abstract interpretation from their concrete
-counterparts (for example, abstract machines), the soundness can be
+counterparts (for example, abstract machines), soundness can be
 easily established by examining the transformation of semantic
@@
 pragmatically, several implementation strategies utilizing purely
-functional programming to build abstract interpreters have
+functional programming techniques to build abstract interpreters have
 emerged. Such techniques include monads, monad transformers
@@
 \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Githubsemantic}, etc. The pure approaches provide certain
+\cite{Kiselyov:2015:FMM:2804302.2804319}, etc. The pure approaches provide certain
 benefits, including that the abstract interpretation artifacts can be
@@
 \cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
-transparency and equational reasoning allow programmers to reason the
-correctness of their implementations more confidently.
+transparency and equational reasoning allow programmers to
+confidently reason about the correctness of their implementations.
 
 However, besides the intrinsic complexity of static analysis, there
-are additional abstraction penalties with these high-level
+are additional abstraction penalties from these high-level
 implementation approaches. First, similar to concrete interpreters,
-the abstract interpreter analyzes the program by traversing the
-abstract syntax tree (AST), which incurs an interpretive overhead, such
-as the pattern matching on the ASTs and recursive calls on the
+an abstract interpreter analyzes the program by traversing the
+abstract syntax tree (AST), which incurs interpretive overhead, such
+as pattern matching on the ASTs and recursive calls on the
 subexpressions.
@@
 systems to implement the semantics of abstract interpretation. For
-example, the abstract interpreter that returns all possible runtime
+example, the collecting interpreter that returns all possible runtime
 values is a form of nondeterminism, where nondeterminism monads can

### 2019-08-16 paper/abstract.tex
update

@@
-It is well known that a staged interpreter is a compiler: specializing and
+It is well known that a staged interpreter is a compiler: specializing an
 interpreter to a given program produces an equivalent executable that runs faster.
@@
 tweaking an interpreter to run on an abstract domain produces a sound static
-analysis. What happens when we combine these two ideas, and apply staging to
+analysis. What happens when we combine these two ideas, and apply specialization to
 an \emph{abstract} interpreter?
@@
 We show that staging abstract interpreters is practical and useful to
-optimize static analysis while requiring less engineering efforts and not
+optimize static analysis while requiring less engineering efforts without 
 compromising soundness. We conduct three case studies, including a comparison
 with \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
-applications on various control-flow analyses, and a demonstration shows that
-it can be used for modular analysis.
-We also empirically evaluate the running time improved by staging.
+applications on various control-flow analyses, and a demonstration that
+the approach can be used to compile an analysis to the program in a modular
+fashion.  We also empirically evaluate the running time improved by staging.
 The experiment shows an order of magnitude speedup with staging for

### 2019-08-16 paper/10_intro.tex
Moar proofreading

@@
 
-Abstract interpretation as a lattice-based approach to sound static
-analyses was proposed by \citet{DBLP:conf/popl/CousotC77}. Equipped
-with Galois connections \cite{CousotCousot79-1}, the analyzer can
-soundly approximate the program runtime behavior by computing fixed
-points on an abstract domain. Despite the tremendous theoretical
-development of abstract interpretation over the years, constructing
-artifacts and analyzers that perform sound abstract interpretation for
-modern and expressive languages was considered complicated for a long
-time.
+Abstract interpretation as a lattice-based approach to sound static analyses was
+proposed by \citet{DBLP:conf/popl/CousotC77}. Based on the notion of Galois
+connections \cite{CousotCousot79-1}, the analyzer can soundly approximate the
+concrete program runtime behavior by computing fixed points on an abstract
+domain. Despite the tremendous theoretical development of abstract
+interpretation over the years, constructing artifacts and analyzers that perform
+sound abstract interpretation for modern and expressive languages was considered
+complicated for a long time.
 
@@
 \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Kiselyov:2015:FMM:2804302.2804319}, etc. The pure approaches provide certain
+\cite{Githubsemantic}, etc. The pure approaches provide certain
 benefits, including that the abstract interpretation artifacts can be
@@
 \cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
-transparency and equational reasoning allow programmers to
+transparency and equational reasoning allow programmers to more
 confidently reason about the correctness of their implementations.
@@
 However, besides the intrinsic complexity of static analysis, there
-are additional abstraction penalties from these high-level
+are additional abstraction penalties with these high-level
 implementation approaches. First, similar to concrete interpreters,
 an abstract interpreter analyzes the program by traversing the
-abstract syntax tree (AST), which incurs interpretive overhead, such
-as pattern matching on the ASTs and recursive calls on the
+abstract syntax tree (AST), which incurs an interpretive overhead, such
+as the pattern matching on the ASTs and recursive calls on the
 subexpressions.
@@
 systems to implement the semantics of abstract interpretation. For
-example, the collecting interpreter that returns all possible runtime
+example, the abstract interpreter that returns all possible runtime
 values is a form of nondeterminism, where nondeterminism monads can

### 2019-08-16 paper/abstract.tex
Moar proofreading

@@
 tweaking an interpreter to run on an abstract domain produces a sound static
-analysis. What happens when we combine these two ideas, and apply specialization to
+analysis. What happens when we combine these two ideas, and apply staging to
 an \emph{abstract} interpreter?
@@
 interpreter, staged concrete interpreter, unstaged abstract interpreter, or
-staged abstract interpreter.
-As an example of \emph{abstraction without regret}, the overhead of these
-abstraction layers is eliminated in the generated code after staging.
-We show that staging abstract interpreters is practical and useful to
-optimize static analysis while requiring less engineering efforts without 
-compromising soundness. We conduct three case studies, including a comparison
-with \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract compilation,
-applications on various control-flow analyses, and a demonstration that
-the approach can be used to compile an analysis to the program in a modular
-fashion.  We also empirically evaluate the running time improved by staging.
-The experiment shows an order of magnitude speedup with staging for
-control-flow analyses.
+staged abstract interpreter. As an example of \emph{abstraction without regret},
+the overhead of these abstraction layers is eliminated in the generated code
+after staging. We show that staging abstract interpreters is practical and
+useful to optimize static analysis while requiring less engineering effort,
+without compromising soundness. We conduct three case studies, including a
+comparison with \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract
+compilation, applications to various control-flow analyses, and a demonstration
+that can be used for modular analysis. We also empirically evaluate the effect
+of staging on the execution time. The experiment shows an order of magnitude
+speedup with staging for control-flow analyses.
 

### 2019-08-16 paper/10_intro.tex

--- 2019-08-16 paper/10_intro.tex
@@
 abstract syntax tree (AST), which incurs interpretive overhead, such
-as pattern matching on the ASTs and recursive calls on the
+as the pattern matching on the ASTs and recursive calls on the
 subexpressions.

### 2019-08-16 paper/abstract.tex

@@
 after staging. We show that staging abstract interpreters is practical and
-useful to optimize static analysis while requiring less engineering effort,
+useful to optimize static analysis, while requiring less engineering effort,
 without compromising soundness. We conduct three case studies, including a

### 2019-08-16 paper/10_intro.tex
Fixing 62

@@
 
-However, besides the intrinsic complexity of static analysis, there
-are additional abstraction penalties from these high-level
-implementation approaches. First, similar to concrete interpreters,
-an abstract interpreter analyzes the program by traversing the
-abstract syntax tree (AST), which incurs interpretive overhead, such
-as the pattern matching on the ASTs and recursive calls on the
-subexpressions.
-Such overhead can be negligible if the abstract interpreter only runs
-on the program a single time, but can accumulate significantly if
-run repeatedly (for example, on libraries).  Second, it is common that
-abstract interpreters written in pure languages extensively use effect
-systems to implement the semantics of abstract interpretation. For
-example, the collecting interpreter that returns all possible runtime
-values is a form of nondeterminism, where nondeterminism monads can
-help.  Although such pure approaches have their own merits, compared
-with imperative stateful implementations, they are significantly
-slower.
+However, besides the intrinsic complexity of static analysis, there are
+additional abstraction penalties from these high-level implementation
+approaches. First, similar to concrete interpreters, an abstract interpreter
+analyzes the program by traversing the abstract syntax tree (AST), which incurs
+interpretive overhead, such as the pattern matching on the ASTs and recursive
+calls on the subexpressions. Such overhead can be negligible if the abstract
+interpreter only runs on the program a single time, but can accumulate
+significantly if run repeatedly (for example, on libraries).  Second, it is
+common that abstract interpreters written in pure languages extensively use
+effect systems to implement the semantics of abstract interpretation. For
+example, the collecting interpreter that returns all possible runtime values is
+nondeterministic.  Although such pure approaches have their own merits, compared
+with imperative, stateful, implementations, they are significantly slower. In
+order to reconcile these approaches, one must use a programming technique which
+allows for high-level, functional abstractions during implementation, but which
+bypasses the incurred overhead at runtime.
 
-Roughly at the same when abstract interpretation was proposed in
+Roughly at the same time when abstract interpretation was proposed in
 1970s, \citet{futamura1971partial, Futamura1999} revealed a close

### 2019-08-16 paper/10_intro.tex
Mention 2-level semantics earlier
specialization comment

@@
 part of the program, and generates a residualized program that solely relies on
-the dynamic part. However, it is hard to precisely analyze binding-times given
-an arbitrary program. As an alternative approach, multi-stage programming (MSP)
+the dynamic part. This can be viewed as a \textit{two-level semantics}:
+different parts of the original program will be executed as different programs,
+possibly in an environment with different semantics.
+However, it is hard to precisely analyze binding-times given an arbitrary
+program. As an alternative approach, multi-stage programming (MSP)
 \cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly program over
@@
 staged monads that can be used to do code generation, we obtain a
-staged abstract interpreter that fulfills the specialization. After
+staged abstract interpreter which is fully specialized. After
 staging, the generated code is specialized to the input program, and

### 2019-08-16 paper/10_intro.tex
update and spaces

--- 2019-08-16 paper/10_intro.tex
@@
 
-Recent progress on methodologies such as Abstracting Abstract
-Machine (AAM) \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10}
-uncovers an approach to constructing abstract interpreters from
-systematic principles, rather than ad-hoc engineering.  By deriving
+Recent progress on methodologies such as Abstracting Abstract Machine (AAM)
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} uncovers an
+operational approach to constructing abstract interpreters.  By deriving
 semantics artifacts for abstract interpretation from their concrete
-counterparts (for example, abstract machines), soundness can be
-easily established by examining the transformation of semantic
-artifacts.
+counterparts (for example, abstract machines), soundness can be easily
+established by examining the transformation of semantic artifacts.
 This systematic abstraction approach can be tailored to different
@@
 \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Kiselyov:2015:FMM:2804302.2804319}, etc. The pure approaches provide certain
-benefits, including that the abstract interpretation artifacts can be
-built compositionally and modularly, e.g., by using monad
-transformers. Therefore, the soundness of analysis can be proven with
-less effort, whether by mechanized proofs
-\cite{Darais:2016:CGC:2951913.2951934} or paper-based proofs
-\cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
-transparency and equational reasoning allow programmers to
-confidently reason about the correctness of their implementations.
+\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. The pure
+approaches provide certain benefits, including that the abstract interpretation
+artifacts can be built compositionally and modularly, e.g., by using monad
+transformers. Therefore, the soundness of analysis can be proven with less
+effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
+paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
+transparency and equational reasoning allow programmers to confidently reason
+about the correctness of their implementations.
 
@@
 common that abstract interpreters written in pure languages extensively use
-effect systems to implement the semantics of abstract interpretation. For
-example, the collecting interpreter that returns all possible runtime values is
-nondeterministic.  Although such pure approaches have their own merits, compared
-with imperative, stateful, implementations, they are significantly slower. In
-order to reconcile these approaches, one must use a programming technique which
-allows for high-level, functional abstractions during implementation, but which
+functional encodings of side-effects to implement the semantics of abstract interpretation. 
+For example, the collecting interpreter that returns all possible runtime values is
+nondeterministic, which can be modeled by a list monad.  Although such pure
+approaches have their own merits, compared with imperative, stateful,
+implementations, they are significantly slower.
+In order to reconcile these approaches, one must use a programming technique
+which allows for high-level abstractions during implementation, but which
 bypasses the incurred overhead at runtime.
@@
 removes interpretation overhead and yields an equivalent executable
-that runs faster.  The challenge here: can we apply the same idea to
-general abstract interpreters to make hem run faster?
-To achieve this, we need to handle a nonstandard semantics that
-extensively uses nondeterminism, operates on abstract domains, and
-computes fixed-point iteratively with a cache.
+that runs faster. The idea of Futamura projection has been applied to 
+other similar performance problems and shown success, examples including
+query compilers\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
+\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275} and etc.
+The challenge here: can we apply the same idea to general abstract interpreters
+to make hem run faster? To achieve this, we need to specialize a nonstandard
+semantics that extensively uses nondeterminism, operates on abstract domains,
+and computes fixed-point iteratively with a cache. The solution is not simply obivious.
 
@@
 framework to eliminate performance penalties for abstract interpreters, while
-keeping the benefits inherent in purely functional programming. In short, our
-approach borrows ideas and techniques from multi-stage programming and embedded
-DSLs to abstract interpreters:  1) To remove the overhead from interpretation
-and effect layers, we specialize the abstract interpreter with respect to an
-input program and then generate efficient low-level code that does the actual
-analysis.  2) Inspired by the tagless-final interpreters
+keeping the implementation as close to its conceptual model.
+In short, our approach borrows ideas and techniques from multi-stage
+programming and embedded DSLs to abstract interpreters:  1) To remove the
+overhead from interpretation and effect layers, we specialize the abstract
+interpreter via staging and then generate efficient low-level code that does
+the actual analysis. 2) Inspired by the tagless-final interpreters
 \cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
-abstracts over binding-time and different semantics, which allows user to
-implement different semantics modularly, including the staged abstract
-interpretation semantics. As a result, we can derive the staged abstract
-interpreters without intrusive changes to its corresponding unstaged
-counterpart, and also preserve the soundness of analysis. In this sense, our
-approach allows for no regret for both performance and engineering effort. We
-elaborate these two main ideas in detail, as follows.
+abstracts over binding-time and different semantics, which allows the staged
+abstract interpreter can be derived from its unstaged counterpart.
+As a result, the derived staged abstract interpreters has no intrusive changes
+to the underlying abstract semantics, therefore preserves the soundness.
+In this sense, our approach allows for no regret for both performance and
+engineering effort. We elaborate these two main ideas in detail, as follows.
 
@@
 executable. For instance,
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval}
-is an interpreter for some language:
-\begin{lstlisting}
-  def eval(e: Expr)(arg: Input): Value
-\end{lstlisting}
-Given a program $e_0$ of type
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval : Expr -> Input -> Value}
+is an interpreter for some language. Given a program $e_0$ of type
 \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr},
@@
 interpreter $\texttt{eval}_{\texttt{e0}}$ :
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Input}
-$\to$
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Value},
+\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Input -> Value},
 which is the so-called \textit{equivalent executable}.
@@
 
-Partial evaluation \cite{DBLP:books/daglib/0072559} was the first proposed
-approach to realize Futamura projections: it first discovers the binding-times
-(\textit{static} or \textit{dynamic}) of variables, then evaluates the static
-part of the program, and generates a residualized program that solely relies on
-the dynamic part. However, it is hard to precisely analyze binding-times given
-an arbitrary program. As an alternative approach, multi-stage programming (MSP)
+The first proposed approach to realize Futamura projections is partial
+evaluation \cite{DBLP:books/daglib/0072559}, which relies on an automatic
+binding-time analysis (\textit{static} or \textit{dynamic}).  The underlying
+representation can be viewed as \textit{two-level semantics}
+\cite{NIELSON1989117, NIELSON198859} that distinguishes compile-time
+computation and run-time computation.
+However, it is hard to precisely analyze binding-times given an arbitrary
+program. As an alternative approach, multi-stage programming (MSP)
 \cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly program over
-a two-level semantics \cite{NIELSON198859}, by manually annotating the
-binding-times of variables in the program. Then, the MSP system will check
-whether these annotations are consistent, and specialize the program using that
-information. The staging annotations can be either syntactic (e.g., quote and
-quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular
-Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
+a two-level syntax \cite{Nielson:1992:TFL:130665}, by manually
+annotating the binding-times of variables in the program. Then, the MSP system
+will check whether these annotations are consistent, and specialize the program
+using that information. The staging annotations can be either syntactic (e.g.,
+quote and quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
+Modular Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
 
@@
 abstract definitional interpreter that uses monad transformers as the
-unstaged baseline.  Our proposed framework adopts the type-based
-multi-stage programming from the Lightweight Modular Staging framework,
-and implements the Futamura projection of a big-step abstract
-interpreter for a small, stateful, higher-order language. By deriving
-staged monads that can be used to do code generation, we obtain a
-staged abstract interpreter that fulfills the specialization. After
-staging, the generated code is specialized to the input program, and
-all monadic operations are inlined and compiled down to low-level
-Scala code.
+unstaged baseline. However, the monads do not have a clear stage distinction.
+We further introduce binding-time and make the monads binding-time polymorphic.
+Our proposed framework adopts the type-based multi-stage programming from the
+Lightweight Modular Staging framework, and implements the Futamura projection
+of a big-step abstract interpreter for a small, stateful, higher-order
+language. By deriving staged monads that can be used to do code generation, we
+obtain a staged abstract interpreter that can be used for specialization. After
+staging, the generated code is specialized to the input program, and all
+monadic operations are inlined and compiled down to low-level Scala code.
 
@@
 
-Although the four artifacts may look like dissimilar at first
-glance, they in fact are all firmly rooted in the concrete semantics of
-the language.  This observation provides a way to abstract over the
-interpreter and achieve the flexibility of reinterpreting the shared
-interpreter. Generic interpreters using monads have already been
-applied to abstract interpreters; we further introduce the
-binding-time abstraction into the generic interpreter.  The generic
-interpreter returns a value of monadic type, which can be varied by
-different semantics. The domain of the interpreter and the effects
-such as read, state, and nondeterminism can all be wrapped into this
-monadic type.  The binding-time abstraction is represented by a
-higher-kinded type, and controls whether the interpreter produces
-values directly or generates code. It is worth mentioning that the
-binding-time type is also injected into the monadic type, so that we
+Although the four artifacts may look like dissimilar at first glance, they in
+fact are all firmly rooted in the concrete semantics of the language.  This
+observation provides a way to abstract over the interpreter and achieve the
+flexibility of reinterpreting the shared interpreter.  The generic interpreter
+returns a value of monadic type, which can be varied by different semantics.
+The domain of the interpreter and the effects such as read, state, and
+nondeterminism can all be wrapped into this monadic type.  The binding-time
+abstraction is represented by a higher-kinded type, and controls whether the
+interpreter produces values directly or generates code. It is worth mentioning
+that the binding-time type is also injected into the monadic type, so that we
 will distinguish normal monads and staged monads.

### 2019-08-16 paper/50_stageabs.tex
update and spaces

@@
     res <- lift_nd[(Set[ValSt], Cache)](Set(emit_ap_clo(clo, arg, σ, in, out)))
-    _   <- put_out_cache(res._2)
-    vs  <- lift_nd[ValSt](res._1)
-    _   <- put_store(vs._2)
+    _   <- put_out_cache(res._2);  vs <- lift_nd[ValSt](res._1);  _ <- put_store(vs._2)
   } yield vs._1
 \end{lstlisting}
-
-% Monad is the bridge that connects future-stage values
 
@@
 
-%\paragraph{Heterogeneous Staging} The generated program is in A-Normal form. \todo{TODO}
-
 \section{Discussion}
@@
 We have gradually presented the confluence of specialization and abstraction of
-concrete interpreters. In this section, we review and summarize our recipe
-to achieve the staged abstract interpreter, discuss correctness issues,
-and different design choices.
+concrete interpreters from an \textit{operational} perspective.
+In this section, we review and summarize our recipe to achieve the staged
+abstract interpreter, discuss correctness issues, and different design choices.
 
@@
 the use of monads internally.
-
-%\todo{maybe show a small example?}
 

### 2019-08-16 paper/60_cases.tex
update and spaces

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\vspace{-2em}
 \subsection{Control-flow Analysis} \label{cfa}
@@
 
+\vspace{-1em}
 \begin{figure}[h!]
@@
     val res: Rep[(Set[(Value,Store)], Cache)] = ...  // omit code for checking caches
-    _ <- put_out_cache(res._2)
-    vs <- lift_nd(res._1)
-    σ <- gc(ψ ⊔ root(vs._2)) // perform abstract garbage collection
-    _ <- put_store(σ)
+    _ <- put_out_cache(res._2);  vs <- lift_nd(res._1)
+    σ <- gc(ψ ⊔ root(vs._2));     _ <- put_store(σ)  // gc performs abstract garbage collection
   } yield v
 \end{lstlisting}
-\vspace{-2em}
 \caption{\texttt{fix\_gc} for abstract garbage collection} \label{fixgc}
@@
 
+\vspace{-2em}
 \subsection{Modular Staged Analysis for Free} \label{modular}
-
-% using different k?
 
@@
 
-Consider there is a library that contains a list map function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}, 
+Consider there is a library that contains a list map function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}:
 \begin{lstlisting}
-  (define (map xs f)
-    (if (null? xs) '() (cons (f (car xs)) (map (cdr xs) f))))
+  (define (map xs f) (if (null? xs) '() (cons (f (car xs)) (map (cdr xs) f))))
 \end{lstlisting}
@@
         val x58 = x52.getOrElse(x54, x57)               //   into the latest store, initially x41
-        val x59 = x58.union(x55)
-        x52 + (x54 -> x59)
-    } ...
-  }
+        val x59 = x58.union(x55);  x52 + (x54 -> x59) } 
+    ... }
 \end{lstlisting}
 
-We borrow the notation traditional modular analysis and realize that
-such generated function is a reusable artifact which can be used
-modularly.  The generated functions already eliminate the
-interpretation overhead and monadic overhead, therefore can be
-consider as partial summaries. Although, we still need to run the
-generated function at next stage to obtain the analysis result.  At
-next stage, we can provide different abstract arguments, store and
-caches, depending on the context where it is called, thus the analysis
-is modular.
+We borrow the notation traditional modular analysis
+\cite{DBLP:conf/cc/CousotC02} and realize that such generated function is a
+reusable artifact which can be used modularly.  The generated functions already
+eliminate the interpretation overhead and monadic overhead, therefore can be
+consider as partial summaries. Although, we still need to run the generated
+function at next stage to obtain the analysis result.  At next stage, we can
+provide different abstract arguments, store and caches, depending on the
+context where it is called, thus the analysis is modular.
 

### 2019-08-16 paper/70_eval.tex
update and spaces

@@
 desugars Scheme to a small core language that makes the analyzer easy to
-implement. As described in Section \ref{cfa}, we implement two monovariant
-0-CFA-like analysis, one is equipped with store-widening, the other does not.
-\todo{why store widening?} Our prototype implementation currently generates
-Scala code; the generated Scala code will be compiled and executed also on JVM.
-it is possible to also generate C/C++ code in the future. In the experiments, we
-implement several optimizations mentioned in Section \ref{staged_ds},
-specifically, the selective caching and rewriting rules exploiting
+implement. We implement two common monovariant 0-CFA-like analysis, one is
+equipped with store-widening (0CFA-SW for short), the other does not (0CFA for
+short). The store-widened analysis is expected to be faster than the one
+without.  Our prototype implementation currently generates Scala code; the
+generated Scala code will be compiled and executed on JVM. it is possible to
+generate C/C++ code in the future. In the experiments, we implement several
+optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
+caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps) exploiting
 partially-static data. These optimizations are useful to reduce the size of
@@
     @kcfa-worst-32@     & 358  & 284.268 ms & 9.065 ms   & 31.35x      & 10.109 ms  & 1.661 ms  &  6.08x        \\
-    @kcfa-worst-64@     & 710  & 2.057 s    & 0.029 s    & 70.23x      & 23.269 ms  & 3.312 ms  &  7.02x        \\
+    @kcfa-worst-64@     & 710  & 2.165 s    & 0.0425 s   & 50.85x      & 23.269 ms  & 3.312 ms  &  7.02x        \\
     @solovay-strassen@  & 523  & 5.078 s    & 0.766 s    & 6.62x       & 18.757 ms  & 3.142 ms  &  5.96x       \\
@@
 Figure~\ref{evaluation_result} shows the evaluation result, comparing
-the performance improved by staging on two monovariant CFA (without
-and with store-widening). The \textit{unstaged} and \textit{staged}
-columns show the median time to finish the analysis. The column
-$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement from
-staging. The dash - in some cells indicates timeout.  As we can see
-from the table, the staged version significantly outperforms the
-unstaged version on all benchmarks.  On average, the staging produces
-code that runs 17x times faster than the unstaged analyzer.
+the performance improved by staging on the two monovariant CFAs 
+The \textit{unstaged} and \textit{staged} columns show the median time to
+finish the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows
+the improvement from staging. The dash - in some cells indicates timeout.  As
+we can see from the table, the staged version significantly outperforms the
+unstaged version on all benchmarks.
+For 0CFA w/o store-widening, we observe an average speed-up of ~17x times.  For
+0CFA w/ store-widening, the average speed-up is ~7x times.  The @kcfa-worst-64@
+benchmark has the most speed-up among all the tests. Although it has a
+comparablely large size of AST, it has a quite simple structure which only
+involves booleans, functions and applications. We conjecture that the simple
+domain and structure allow the generated code to be efficiently optimized by
+the JVM. But on the corresponding 0CFA-SW, we do not observe such great
+speed-up. Due to the complexity of generated code and next-stage running
+platform (JVM), we leave the detailed empirical analysis of which static
+analysis can benefit more from specialization as a future work.
 
-\todo{analyze some interesting examples}
-
-Our evaluation result provides an abstraction without regret
-approach to construct optimizing abstract interpreters -- the user may write a
-high-level modular and compositional abstract interpreter, and then derive a
-staged version in a principled way that runs significantly faster.
+%Our evaluation result provides an abstraction without regret
+%approach to construct optimizing abstract interpreters -- the user may write a
+%high-level modular and compositional abstract interpreter, and then derive a
+%staged version in a principled way that runs significantly faster.

### 2019-08-16 paper/80_related.tex
update and spaces

@@
 abstract interpreter that allows proof of soundness to be constructed
-compositionally.
-
-\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
+compositionally. \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
 modular analysis. We borrow the notation of modular analysis for our approach.
@@
 However, several variants of 0CFA were developed and shown to be useful in
-practice and program optimization \cite{Adams:2011:FTR:2048066.2048105,
+practice \cite{Adams:2011:FTR:2048066.2048105,
 Bergstrom:2014:PEH:2628136.2628153, ashley:practical, Reppy:2006:TCA:1159876.1159888}.

### 2019-08-16 paper/30_stageconc.tex
update

@@
 programmers to control what is generated for each kind of node. Here we have
-matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions for them.
+matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions via @emitValDef@ for them.
 For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term

### 2019-08-16 paper/60_cases.tex
update

@@
 \cite{DBLP:conf/cc/CousotC02} and realize that such generated function is a
-reusable artifact which can be used modularly.  The generated functions already
-eliminate the interpretation overhead and monadic overhead, therefore can be
-consider as partial summaries. Although, we still need to run the generated
-function at next stage to obtain the analysis result.  At next stage, we can
-provide different abstract arguments, store and caches, depending on the
-context where it is called, thus the analysis is modular.
+reusable artifact which can be used modularly, i.e., independent to its context. 
+The generated functions already eliminate the interpretation overhead and
+monadic overhead, therefore can be consider as partial summaries. Although, we
+still need to run the generated function at next stage to obtain the analysis
+result.  At next stage, we can provide different abstract arguments, store and
+caches, depending on the context where it is called.
 
@@
 transition from a whole-program analyzer to modular staged analyzer is
-mechanized and for free from staging. We do not need to design any
+mechanized and for free via staging. We do not need to design any
 intermediate representation or data structures of summaries.  In some

### 2019-08-16 paper/70_eval.tex
update

@@
 
-Figure~\ref{evaluation_result} shows the evaluation result, comparing
-the performance improved by staging on the two monovariant CFAs 
-The \textit{unstaged} and \textit{staged} columns show the median time to
-finish the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows
-the improvement from staging. The dash - in some cells indicates timeout.  As
-we can see from the table, the staged version significantly outperforms the
-unstaged version on all benchmarks.
-For 0CFA w/o store-widening, we observe an average speed-up of ~17x times.  For
-0CFA w/ store-widening, the average speed-up is ~7x times.  The @kcfa-worst-64@
-benchmark has the most speed-up among all the tests. Although it has a
-comparablely large size of AST, it has a quite simple structure which only
-involves booleans, functions and applications. We conjecture that the simple
-domain and structure allow the generated code to be efficiently optimized by
-the JVM. But on the corresponding 0CFA-SW, we do not observe such great
-speed-up. Due to the complexity of generated code and next-stage running
-platform (JVM), we leave the detailed empirical analysis of which static
-analysis can benefit more from specialization as a future work.
+Figure~\ref{evaluation_result} shows the evaluation result, comparing the
+performance improved by staging on the two monovariant CFAs The
+\textit{unstaged} and \textit{staged} columns show the median time to finish
+the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows the
+improvement. The dash - in some cells indicates timeout.  As
+expected, the staged version significantly outperforms the unstaged version on
+all benchmarks. We observe that simple and homogeneous domain used in the
+benchmark usualy leads to better speed-ups.  For 0CFA w/o store-widening, we
+observe an average speed-up of ~17x times; for 0CFA w/ store-widening, the
+average speed-up is ~7x times.  The @kcfa-worst-64@ benchmark has the most
+speed-up among all the tests. Although it has a comparablely large size of AST,
+it has a quite simple structure which only involves booleans, functions and
+applications. We conjecture that the simple domain and structure allow the
+generated code to be efficiently optimized by the JVM. But on the corresponding
+0CFA-SW, we do not observe such great speed-up. Due to the complexity of
+generated code and next-stage running platform (JVM), we leave the detailed
+empirical analysis of which static analysis can benefit more from
+specialization as future work.
 

### 2019-08-16 paper/20_prelim.tex
update and spaces
typo

@@
 identifier \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x}, through the environment and store. Conditionals and
-arithmetic is handled by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{br0} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{arith}, respectively.
+arithmetic are handled by \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{br0} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{arith}, respectively.
 The method

### 2019-08-16 paper/60_cases.tex
typo

@@
 which is a function (closure) value that takes and returns stores. After the
-first run of analysis, we have traversed the AST and obtained a single closure
+first run of analysis, we have traversed the AST and obtained a single closure that
 takes a store and returns a store. Therefore, the ASTs are static terms, and the
@@
 and refactor the staged abstract interpreter we presented in Section \ref{sai}
-to various control-flow analysis (CFA). We first discuss calling-context
+to various control-flow analyses (CFA). We first discuss calling-context
 sensitive analysis for staged analyzers by tweaking allocation strategy; then we
@@
 
-The abstract addresses we introduced in Section \ref{sai} is simply identical to
+The abstract addresses we introduced in Section \ref{sai} are simply identical to
 the variable names, which is a context-insensitive schema.
 Through tweaking the address allocation strategy, we can achieve various
-context-sensitive analysis \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we
+context-sensitive analyses \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we
 demonstrate a calling-context-sensitive analysis, i.e., $k$-CFA-like analysis
@@
 recent calling contexts. The definition of abstract addresses \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{KCFAAddr} is also
-changed to contain a identifier and the time when it get allocated, meaning that
+changed to contain an identifier and the time when it get allocated, meaning that
 this address points to some values under such calling context. If $k$ is 0, we
@@
 Every time when we call the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} function, we refresh the timestamp by calling
-the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick} function, which update the timestamp in the state monad.
+the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick} function, which updates the timestamp in the state monad.
 The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick}

### 2019-08-16 paper/30_stageconc.tex
tweak for latexdiff

--- 2019-08-15 paper/30_stageconc.tex
@@
 programmers to control what is generated for each kind of node. Here we have
-matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions for them.
+matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions via \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emitValDef} for them.
 For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term

### 2019-08-31 paper/70_eval.tex
tweak for latexdiff
misc update

@@
 (192 threads) in total, the abstract interpreters are single-threaded.
-To minimize the effect of warming-up from the HotSpot JIT compiler,
+To minimize the effect of warm-up from the HotSpot JIT compiler,
 all the experiments are executed for 20 times and we report the
@@
 expected, the staged version significantly outperforms the unstaged version on
-all benchmarks. We observe that simple and homogeneous domain used in the
-benchmark usualy leads to better speed-ups.  For 0CFA w/o store-widening, we
+all benchmarks. We observe that simple and homogeneous value domains used in the
+benchmark usually leads to better speed-ups.  For 0CFA w/o store-widening, we
 observe an average speed-up of ~17x times; for 0CFA w/ store-widening, the
-average speed-up is ~7x times.  The @kcfa-worst-64@ benchmark has the most
+average speed-up is ~7x times.  The \texttt{kcfa-worst-64} benchmark has the most
 speed-up among all the tests. Although it has a comparablely large size of AST,

### 2019-08-31 paper/60_cases.tex
typo

@@
 
-As we have already seen from Section \ref{sai}, and recall the
-primitive operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} that denotes a $\lambda$ term to next-stage
-Scala function.  The specializing of staged abstract interpreter to
+As we have already seen from Section \ref{sai}, and recall that the
+primitive operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} can denote a $\lambda$ term to a next-stage Scala function.
+The specializing of the staged abstract interpreter to
 $\lambda$ terms is independent to other parts of the analyzed program.

### 2019-09-01 paper/70_eval.tex
typos

@@
 
-\paragraph{Implementation and Evaluation Environment}
-We implement the abstract interpreters for control-flow analysis, i.e., a
-value-flow analysis that computes which lambda terms can be called at each
-call-sites in functional languages. With other suitable abstract domains, the
-abstract interpreter also can be extended to other analyses. A front-end
-desugars Scheme to a small core language that makes the analyzer easy to
-implement. We implement two common monovariant 0-CFA-like analysis, one is
-equipped with store-widening (0CFA-SW for short), the other does not (0CFA for
-short). The store-widened analysis is expected to be faster than the one
-without.  Our prototype implementation currently generates Scala code; the
-generated Scala code will be compiled and executed on JVM. it is possible to
-generate C/C++ code in the future. In the experiments, we implement several
-optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
-caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps) exploiting
+\paragraph{Implementation and Evaluation Environment} We implement the abstract
+interpreters for control-flow analysis, i.e., a value-flow analysis that
+computes which lambda terms can be called at each call-sites in functional
+languages. With other suitable abstract domains, the abstract interpreter also
+can be extended to other analyses. A front-end desugars Scheme to a small core
+language that makes the analyzer easy to implement. We implement two common
+monovariant 0-CFA-like analysis, one is equipped with store-widening (0CFA-SW
+for short), the other does not (0CFA for short). The store-widened analysis is
+expected to be faster than the one without.  Our prototype implementation
+currently generates Scala code; the generated Scala code will be compiled and
+executed on JVM. It is possible to generate C/C++ code in the future. In the
+experiments, we implement several optimizations mentioned in Section
+\ref{staged_ds}, specifically, the selective caching scheme and IR-level
+rewriting rules (for pairs, lists, sets, and maps) that exploit
 partially-static data. These optimizations are useful to reduce the size of
@@
 average speed-up is ~7x times.  The \texttt{kcfa-worst-64} benchmark has the most
-speed-up among all the tests. Although it has a comparablely large size of AST,
+speed-up among all the tests. Although it has a comparably large size of AST,
 it has a quite simple structure which only involves booleans, functions and

### 2019-09-03 paper/10_intro.tex
typos
misc

@@
 approaches. First, similar to concrete interpreters, an abstract interpreter
-analyzes the program by traversing the abstract syntax tree (AST), which incurs
+analyzes the program by traversing the abstract syntax tree (AST). The traversing incurs
 interpretive overhead, such as the pattern matching on the ASTs and recursive
-calls on the subexpressions. Such overhead can be negligible if the abstract
+calls on subexpressions. Such overhead can be negligible if the abstract
 interpreter only runs on the program a single time, but can accumulate
-significantly if run repeatedly (for example, on libraries).  Second, it is
+significantly if it runs repeatedly (for example, on libraries).  Second, it is
 common that abstract interpreters written in pure languages extensively use
 functional encodings of side-effects to implement the semantics of abstract interpretation. 
-For example, the collecting interpreter that returns all possible runtime values is
-nondeterministic, which can be modeled by a list monad.  Although such pure
-approaches have their own merits, compared with imperative, stateful,
-implementations, they are significantly slower.
-In order to reconcile these approaches, one must use a programming technique
-which allows for high-level abstractions during implementation, but which
-bypasses the incurred overhead at runtime.
+For example, the list monad can be used to model nondeterminism from collecting
+interpreters.
+%For example, a collecting interpreter that returns all possible runtime values is
+%nondeterministic, which can be modeled by a list monad.  
+Although such pure approaches have their own merits, compared with imperative,
+stateful implementations, they are significantly slower.  In order to
+reconcile these approaches, one must use a programming technique which allows
+for high-level abstractions during implementation, but which bypasses the
+incurred overhead at runtime.
 
 Roughly at the same time when abstract interpretation was proposed in
-1970s, \citet{futamura1971partial, Futamura1999} revealed a close
+1970s, \citet{futamura1971partial} revealed a close
 connection between interpreters and compilers through a hierarchy of
@@
 \url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275} and etc.
-The challenge here: can we apply the same idea to general abstract interpreters
-to make hem run faster? To achieve this, we need to specialize a nonstandard
+The challenge here: can we apply the same idea to a general abstract interpreter
+to make it run faster? To achieve this, we need to specialize a nonstandard
 semantics that extensively uses nondeterminism, operates on abstract domains,
-and computes fixed-point iteratively with a cache. The solution is not simply obivious.
+and computes fixed-point iteratively over a cache. The solution is not simply obvious.
 
@@
 extended to abstract interpreters. We present an abstraction-without-regret
-framework to eliminate performance penalties for abstract interpreters, while
-keeping the implementation as close to its conceptual model.
+framework that eliminates performance penalties for monadic abstract
+interpreters, while keeping the implementation as close to their conceptual
+model.
 In short, our approach borrows ideas and techniques from multi-stage
@@
 \cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
-abstracts over binding-time and different semantics, which allows the staged
+abstracts over binding-times and different semantics, which allows the staged
 abstract interpreter can be derived from its unstaged counterpart.
@@
 binding-time analysis (\textit{static} or \textit{dynamic}).  The underlying
-representation can be viewed as \textit{two-level semantics}
+representation can be viewed as a \textit{two-level semantics}
 \cite{NIELSON1989117, NIELSON198859} that distinguishes compile-time

### 2019-09-03 paper/paper.tex
revise intro
revision
update
misc update
ack

@@
 %\usepackage{MnSymbol}
+\usepackage{color}
 \usepackage{xcolor,colortbl}
@@
 
+\definecolor{light-gray}{gray}{0.85}
+
 \input{macros}
@@
 %\title{Staged Abstract Interpreters}         %% [Short Title] is optional;
-\title{Staging Abstract Interpreters}         %% [Short Title] is optional;
+\title{Staged Abstract Interpreters}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 \author{Guannan Wei}
+\author{Yuxuan Chen}
+\author{Tiark Rompf}
 %\authornote{with author1 note}          %% \authornote is optional;
@@
 \email{guannanwei@purdue.edu}          %% \email is recommended
+\email{chen1797@purdue.edu}
+\email{tiark@purdue.edu}          %% \email is recommended
 
 %% Author with two affiliations and emails.
-\author{Yuxuan Chen}
 %\authornote{with author2 note}          %% \authornote is optional;
                                         %% can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  %\position{Position2a}
-  \department{Department of Computer Science}             %% \department is recommended
-  \institution{Purdue University}           %% \institution is required
-  %\streetaddress{305 N. University Street}
-  %\city{West Lafayette}
-  %\state{IN}
-  %\postcode{47906}
-  \country{USA}                    %% \country is recommended
-}
-
-\email{first2.last2@inst2a.com}         %% \email is recommended
-\affiliation{
-  %\position{Position2b}
-  %\department{Department2b}             %% \department is recommended
-  \institution{Facebook}           %% \institution is required
-  %\streetaddress{Street3b Address2b}
-  %\city{City2b}
-  %\state{State2b}
-  %\postcode{Post-Code2b}
-  \country{USA}                   %% \country is recommended
-}
-\email{first2.last2@inst2b.org}         %% \email is recommended
-
-\author{Tiark Rompf}
 %\authornote{with author1 note}          %% \authornote is optional;
@@
 %\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
-\affiliation{
-  %\position{Position1}
-  \department{Department of Computer Science}              %% \department is recommended
-  \institution{Purdue University}            %% \institution is required
-  %\streetaddress{305 N. University Street}
-  %\city{West Lafayette}
-  %\state{IN}
-  %\postcode{47906}
-  \country{USA}                    %% \country is recommended
-}
-\email{tiark@purdue.edu}          %% \email is recommended
 
@@
 %% Acknowledgments
-% Ack Fei, Greg, James, Qianchuan, Guanhong
 \begin{acks}                            %% acks environment is optional
@@
   %% extraction tools.
+  This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE
+  award DE-SC0018050, as well as gifts from Google, Facebook, and VMware.
+
+  \iffalse
   This material is based upon work supported by the
@@
   National Science Foundation.
+  \fi
 \end{acks}

### 2019-09-03 paper/10_intro.tex
revert to as sign; typo

@@
 executable. For instance,
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval : Expr -> Input -> Value}
+@eval : Expr -> Input -> Value@
 is an interpreter for some language. Given a program $e_0$ of type
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Expr},
+@Expr@,
 by applying the specialization, we can obtain a specialized
 interpreter $\texttt{eval}_{\texttt{e0}}$ :
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Input -> Value},
+@Input -> Value@,
 which is the so-called \textit{equivalent executable}.

### 2019-09-03 paper/30_stageconc.tex
revert to as sign; typo

--- 2019-08-15 paper/30_stageconc.tex
@@
 uses syntactic quotations and quasi-quotations, LMS distinguishes binding-time
-based on types.
-LMS provides a type constructor \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[T]} where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T} can be
-an arbitrary type, indicating a value of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T} will be known at the next stage.
-All operations acting on \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[T]} expressions will be residualized as generated
+based on types. LMS provides a type constructor @Rep[T]@ where @T@ can be
+an arbitrary type, indicating a value of @T@ will be known at the next stage.
+All operations acting on @Rep[T]@ expressions will be residualized as generated
 code.
@@
 
-If \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} is a value know at current stage, we may specialize the power function to
-some value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} -- by unrolling the recursive calls. In LMS, this is fulfilled by
-first adding the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type to the variable known at next stage. In this case,
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{b : Rep[Int]} is know later, and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x : Int} is know currently, as shown in the
+If @x@ is a value know at current stage, we may specialize the power function to
+some value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
+first adding the @Rep@ type to the variable known at next stage. In this case,
+@b : Rep[Int]@ is know later, and @x : Int@ is know currently, as shown in the
 below code recipe.
-The way we use this \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{power} function is to create a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{DslDriver} and override the
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{snippet} method give the currently know value \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x} (5 in this example).
+The way we use this @power@ function is to create a @DslDriver@ and override the
+@snippet@ method give the currently know value @x@ (5 in this example).
 \begin{lstlisting}
@@
 The LMS framework provides staging support for primitive data types such as
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Int} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Double}, and commonly-used data structures such as lists and maps.
+@Int@ and @Double@, and commonly-used data structures such as lists and maps.
 The idea behind the framework is to construct a sea-of-node intermediate
@@
 \cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
-expressions of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[_]} to their IR is done by using the implicit design
+expressions of type @Rep[_]@ to their IR is done by using the implicit design
 pattern. As we will see later, implementing staging and code generation
@@
 instantiating unstaged concrete interpreter in Section \ref{unstaged_conc}.
-But now we use the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type to annotate value domains, environments and
+But now we use the @Rep@ type to annotate value domains, environments and
 stores, and redefine the staged version of monads and primitive operations.
@@
 monad with a state monad. But now the monads operate on staged values, for
-brevity, we call them \textit{staged monads}. In the code, we also use a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep}
+brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
 prefix on the constructors and types to differentiate them. But it is important
-to note that objects of a case class \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ReaderT}/\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} are not staged, the
-monadic computation like \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R => M[A]} are also not staged. Instead, the internal
-data that these monads operate on are staged, i.e., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{R} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} in the
-reader monad, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{S} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{A} in the state monad.
-The following code
-snippet shows the idea. We use \hl{light gray} to highlight where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type is added:
+to note that objects of a case class @ReaderT@/@StateT@ are not staged, the
+monadic computation like @R => M[A]@ are also not staged. Instead, the internal
+data that these monads operate on are staged, i.e., @R@ and @A@ in the
+reader monad, @S@ and @A@ in the state monad. The following code
+snippet shows the idea. We use \hl{light gray} to highlight where @Rep@ type is added:
 \begin{lstlisting}[escapechar=!]
@@
 
-The function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} passed to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap} is also a current-stage-known value, so
-that all the invocation of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{flatMap}s can be completed before generating code.
+The function @f@ passed to @flatMap@ is also a current-stage-known value, so
+that all the invocation of @flatMap@s can be completed before generating code.
 The fact that we only stage data but not monadic computation or monadic
 values, is the reason that we can peel of the monad stack in the generated code.
-Now the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM[_]} is instantiated as the same structure as before, but using the
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} versions:
+Now the @AnsM[_]@ is instantiated as the same structure as before, but using the
+@Rep@ versions:
 \begin{lstlisting}
@@
 but as we will see, it is not straightforward when for nondeterminism monad
-(\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT}) in abstract interpreters. Because not only the elements in the set are
+(@SetT@) in abstract interpreters. Because not only the elements in the set are
 staged, but the whole set is also staged, and we have no knowledge about how
@@
 and its environment to a next-stage Scala function. To do this, we need to
-recursively call the interpreter function, i.e., the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} argument passed to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close}.
+recursively call the interpreter function, i.e., the @ev@ argument passed to @close@.
 The following code implements the idea:
@@
 
-We have created a current-stage function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} that takes a next-stage value and
-store, and returns staged value.
-Then we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_complied_clo} to delay \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} to
-a next-stage Scala function, represented by the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Value]}. Inside of
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, we can access the evaluator via \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev}, but \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} produces a monadic value of
-type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM}, which can only exist at the current stage. To connect the
+We have created a current-stage function @f@ that takes a next-stage value and
+store, and returns staged value. Then we use @emit_complied_clo@ to delay @f@ to
+a next-stage Scala function, represented by the type @Rep[Value]@. Inside of
+@f@, we can access the evaluator via @ev@, but @ev@ produces a monadic value of
+type @AnsM@, which can only exist at the current stage. To connect the
 current-stage monadic values and future-stage grounded values, the evaluator
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} is invoked on the body expression \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e} to be specialized w.r.t.
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}, and
-then we \textit{collapse} the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{AnsM} monad to grounded values by providing the
+@ev@ is invoked on the body expression @e@ to be specialized w.r.t. @e@, and
+then we \textit{collapse} the @AnsM@ monad to grounded values by providing the
 prepared environment and store.
 
-For function applications \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}, we have two \textit{next-stage} values \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun}
-and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{arg}, but what we can do on these values of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} type?
-In fact, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} does
+For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
+and @arg@, but what we can do on these values of @Rep@ type? In fact, @fun@ does
 not have an intensional application operation we can use directly; but we know
-that it is a staged value produced by the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} function, which represents a
+that it is a staged value produced by the @close@ function, which represents a
 next-stage Scala function. Therefore we can just generate a next-stage function
-application, also in Scala. Since \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo} returns a monadic value of type
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Ans}, we can still use the monadic operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{get_store} to obtain the latest
-store object. Then we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo} to generate a current-stage
+application, also in Scala. Since @ap_clo@ returns a monadic value of type
+@Ans@, we can still use the monadic operation @get_store@ to obtain the latest
+store object. Then we use @emit_ap_clo@ to generate a current-stage
 representation of the future-stage application result, which is a pair of values
-and stores \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ValSt}.
-After getting that result \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{vs}, we use \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{put_store} to
+and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
 replace the store and reify the result to the current-stage monadic value.
@@
 Compared with the unstaged version, we can observe a shift of using the
-evaluation function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} from application-time to value-representation-time: in
-the unstaged interpreter, the invoke of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} happens in \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ap_clo}; while in the
-staged interpreter, we eagerly apply \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} when denoting the lambda term to our
+evaluation function @ev@ from application-time to value-representation-time: in
+the unstaged interpreter, the invoke of @ev@ happens in @ap_clo@; while in the
+staged interpreter, we eagerly apply @ev@ when denoting the lambda term to our
 value domain.
@@
 
-In the previous section, we merely show the type of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_compiled_clo} and
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emit_ap_clo} without a concrete implementation. In this section, we briefly
+In the previous section, we merely show the type of @emit_compiled_clo@ and
+@emit_ap_clo@ without a concrete implementation. In this section, we briefly
 discuss the IR node generated by them and sketch code generation.
-We define the IR nodes \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} using \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{case class}
-extending from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Def[T]}.
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Def[T]} is a built-in type in the LMS framework,
-representing next-stage definitions of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{T}.
+We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
+extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
+representing next-stage definitions of type @T@.
 \begin{lstlisting}
@@
 The IR nodes are manipulated by the LMS passes and finally generated to
-next-stage Scala code. To generate code, LMS provides an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{emitNode} method for
+next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
 programmers to control what is generated for each kind of node. Here we have
-matched \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRCompiledClo} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{IRApClo} and will generate definitions for them.
-For compiled closures, we put the function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} and its accompanying syntactic term
-and environment to its next-stage value representation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo}. For
-applications, we have the knowledge that \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fun} is a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompiledClo} at the stage,
-and it has a callable field \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, so we generate a next-stage function
-application \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f(arg, σ)}.
-The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{quote}s are invocations to generate code for the
+matched @IRCompiledClo@ and @IRApClo@ and will generate definitions via @emitValDef@ for them.
+For compiled closures, we put the function @f@ and its accompanying syntactic term
+and environment to its next-stage value representation @CompiledClo@. For
+applications, we have the knowledge that @fun@ is a @CompiledClo@ at the stage,
+and it has a callable field @f@, so we generate a next-stage function
+application @f(arg, σ)@. The @quote@s are invocations to generate code for the
 other next-stage values.

### 2019-09-03 paper/40_absinterp.tex
revert to as sign; typo

--- 2019-08-15 paper/40_absinterp.tex
@@
 
-In out monad stack, the @in@ and @out@ cache modeled by a reader effect and
+In our monad stack, the @in@ and @out@ cache modeled by a reader effect and
 state effect, respectively. We first defined several methods to manipulate

### 2019-09-03 paper/60_cases.tex
revert to as sign; typo

@@
 implementation: different syntactic constructs are handled in different functions
-(e.g., \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{compCall}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{compApp}, etc.).
+(e.g., @compCall@, @compApp@, etc.).
 Before refactoring into AC, these functions would take the store as a
-regular argument; now every function returns a value of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{CompAnalysis},
+regular argument; now every function returns a value of type @CompAnalysis@,
 which is a function (closure) value that takes and returns stores. After the
@@
 On the other side, our approach does exactly the same thing through staging: the
-AST are static, and stores are dynamic, therefore we annotate the type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store}
-with \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep} (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
-changes of our approach using LMS framework are the types of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Store} to
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Store]}, since they will be known at the next stage; other parts do not need to change. 
+AST are static, and stores are dynamic, therefore we annotate the type @Store@
+with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
+changes of our approach using LMS framework are the types of @Store@ to
+@Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
 The generated code
@@
 functions, which provides additional performance improvement. Nevertheless,
-the support from LMS framework is necessary, such as staging support for \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Map},
+the support from LMS framework is necessary, such as staging support for @Map@,
 we consider these efforts are reusable and complete orthogonal to the
@@
 The timestamp is essentially a finite list of expressions that tracks the $k$
-recent calling contexts. The definition of abstract addresses \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{KCFAAddr} is also
+recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
 changed to contain an identifier and the time when it get allocated, meaning that
@@
 We then integrate the timestamp into our monad stack by adding a staged
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} monad. The result type is passingly get updated to contain a set of
+@StateT@ monad. The result type is passingly get updated to contain a set of
 tuples, where the time is added.
@@
 
-Every time when we call the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{eval} function, we refresh the timestamp by calling
-the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick} function, which updates the timestamp in the state monad.
-The \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{tick}
-function is implemented as prepending the current expression being evaluated \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}
+Every time when we call the @eval@ function, we refresh the timestamp by calling
+the @tick@ function, which updates the timestamp in the state monad. The @tick@
+function is implemented as prepending the current expression being evaluated @e@
 to the existing calling context, and then taking the first $k$ elements from the list.
-The type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Config} used in caching is also changed to included the timestamp.
+The type @Config@ used in caching is also changed to included the timestamp.
 
@@
 DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
-implemented in monadic abstract interpreter by swapping the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{StateT} (for store)
-and the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{SetT} (for nondeterminism). In this section, we show this optimization
+implemented in monadic abstract interpreter by swapping the @StateT@ (for store)
+and the @SetT@ (for nondeterminism). In this section, we show this optimization
 is orthogonal to staging and we just need to swap the same \textit{staged}
@@
 \cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
-shows that, for big-step monadic abstract interpreter, we may add a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Reader}
+shows that, for big-step monadic abstract interpreter, we may add a @Reader@
 monad to track a set of root addresses and compute the reachable addresses every
@@
 
-The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix}
-function with caching (from Section \ref{sai}) to \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{fix_gc}, but the caching part
-is omitted. Several auxiliary functions are used: \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ask_root} retrieves the
-current root addresses from the monad stack; \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{root} compute the reachable
-addresses from a value; and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{gc} cleans up the store according to the reachable
+The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the @fix@
+function with caching (from Section \ref{sai}) to @fix_gc@, but the caching part
+is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
+current root addresses from the monad stack; @root@ compute the reachable
+addresses from a value; and @gc@ cleans up the store according to the reachable
 addresses and returns a new store. For the generic interpreter, we also need to
@@
 As we have already seen from Section \ref{sai}, and recall that the
-primitive operation \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} can denote a $\lambda$ term to a next-stage Scala function.
-The specializing of the staged abstract interpreter to
+primitive operation @close@ can denote a $\lambda$ term to a next-stage Scala function.  The specializing of the staged abstract interpreter to
 $\lambda$ terms is independent to other parts of the analyzed program.
@@
 
-Consider there is a library that contains a list map function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map}:
+Consider there is a library that contains a list map function @map@:
 \begin{lstlisting}
@@
 
-The specialization (0CFA and without store-widening) of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{close} on
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{map} generates the following next-stage Scala code, that is, the
-function \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{x39}:
+The specialization (0CFA and without store-widening) of @close@ on
+@map@ generates the following next-stage Scala code, that is, the
+function @x39@:
 \begin{lstlisting}
@@
 more for calling context by partially-static data optimization. For example,
-consider that we have a library consists of three functions \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h},
-where \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} calls \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h} internally and returns \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{y} finally (we extend the
-language to have multiple arguments).
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h} do not depend on \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}.
-Now we
-would like to modularly specialize these functions, starting from \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f} with an
-initial environment containing \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{g} and \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{h}.
+consider that we have a library consists of three functions @f@, @g@ and @h@,
+where @f@ calls @g@ and @h@ internally and returns @y@ finally (we extend the
+language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
+would like to modularly specialize these functions, starting from @f@ with an
+initial environment containing @f@, @g@ and @h@.
 \begin{lstlisting}
@@
 Since $k$ is bounded to a fixed number, after the analysis running $k$ steps
-inside of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{f}, the calling context will purely depend on the static structure of
+inside of @f@, the calling context will purely depend on the static structure of
 our module being analyzed.
@@
 To evaluates an assignment, we first evaluate its right-hand side, and then put
-the value into the slot where the address of \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{v} points to. For simplicity, we
-elect to make the value of the assignment be an \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{void()} value.
+the value into the slot where the address of @v@ points to. For simplicity, we
+elect to make the value of the assignment be an @void()@ value.
 \begin{lstlisting}
@@
 
-To evaluates a \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{while} loop statement, we evaluate the condition first. Then
-similar to how we treat \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{branch0}, we have a generic \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{branch} function but works
-on Boolean values.
-For the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{true} branch, we recursively call \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{ev} on a newly
-constructed expression \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Seq(e, While(t, e))} meaning that first evaluates \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{e}
+To evaluates a @while@ loop statement, we evaluate the condition first. Then
+similar to how we treat @branch0@, we have a generic @branch@ function but works
+on Boolean values. For the @true@ branch, we recursively call @ev@ on a newly
+constructed expression @Seq(e, While(t, e))@ meaning that first evaluates @e@
 and the repeat the loop. Otherwise, for the other branch, we simply return a
@@
 fields, an upper bound and a lower bound. Here we annotate the upper bound and
-lower bound fields to be of type \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{Rep[Double]}.
-The operations such as \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{+} of
+lower bound fields to be of type @Rep[Double]@. The operations such as @+@ of
 two intervals produce a new

### 2019-09-03 paper/70_eval.tex
revert to as sign; typo

@@
 analysis.  Some of the benchmarks are small programs designed for
-experiments, such as the \lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{kcfa-worst-}$n$ series, which are intended
+experiments, such as the @kcfa-worst-@$n$ series, which are intended
 to challenge $k$-CFA; while some other benchmarks are from real-world
 applications, for examples, the RSA public key encryption algorithm
-\lstinline[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\tt]{rsa}. In the Figure \ref{evaluation_result}, we report the number of
+@rsa@. In the Figure \ref{evaluation_result}, we report the number of
 AST node after parsing and desugaring (excluding comments) as a proper

### 2019-09-03 paper/80_related.tex
revert to as sign; typo

@@
 
-\paragraph{Optimizing Static Analysis Through Specialization}
-The underlying idea in this paper is closely related to abstract
-compilation (AC) \cite{Boucher:1996:ACN:647473.727587}: removing the
-interpretative overhead on traversing the syntax tree by
-specialization. Specifically, the residual program of AC can be either
-text or closures. However, as we studied in Section \ref{cs_ac},
-with the perspective of generative programming and monadic
-interpreters, we make it more general and extensible, and easier.
-\citet{Johnson:2013:OAA:2500365.2500604} adapt the idea of closure
-generation to optimize small-step abstract interpreter in
-state-transition style. The analyzed program is first compiled to an
-IR called "abstract bytecode", which is actually a series of higher-order
-functions which will be executed later on an abstract abstract machine
-for that IR. \citet{damian1999partial} provides a formal treatment to
-abstract compilation and Shiver's CFA, as well as proofs to establish
-the correctness of the certified specialized analyzer.
-\citet{amtoft1999partial} applied partial evaluation for
-constraint-based control flow analysis.
+\paragraph{Optimizing Static Analysis Through Specialization} The underlying
+idea in this paper is closely related to abstract compilation (AC)
+\cite{Boucher:1996:ACN:647473.727587}: removing the interpretative overhead of
+traversing the syntax tree by specialization. The residual programs of AC can
+be either textual (using quasi-quotations) or closures (using higher-order
+functions).  Our approach is more flexible between representations and requires
+less engineering efforts to refactor the analyzer (Section \ref{cs_ac}).
+\citet{Johnson:2013:OAA:2500365.2500604} adapted closure generation to optimize
+small-step abstract interpreter in state-transition style. The program
+is first compiled to an "abstract bytecode" IR. The IR consists of higher-order
+functions, which will be executed later on an abstract abstract machine for
+that IR.  \citet{damian1999partial} provided a formal treatment to abstract
+compilation and control-flow analysis, as well as proofs to establish the
+correctness of a certified specialized analyzer.  \citet{amtoft1999partial}
+applied partial evaluation for constraint-based control-flow analysis.
 
-Splitting an analysis into multiple stages has also been studied for
-analysis other than abstract-interpretation-based control flow
-analysis, though the formulation is quite
-different. \citet{DBLP:conf/cgo/HardekopfL11} apply staging to a
+Splitting an analysis into multiple stages has also been studied for other
+forms of static analyses. 
+\citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
 flow-sensitive pointer analysis. The first stage is to analyze the
-program code to obtain a sparse representation, and then the second
+programs to obtain a sparse representation, and then the second
 stage conducts the flow-sensitive analysis by refining the first
 one. Another large body of static analyses are implemented by
-generating constraints firstly and then let a solver discharge these
+generating constraints firstly and then use a solver to discharge the
 proof obligations or find the fixed-point. Notable examples include
@@
 analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
-Similar to the idea in this paper, such analyses can be split to
-staged.  One piece of future work is to explore whether these
-analyses can be implemented and optimized from
-meta-programming. Souffl{\'e}, one of the state-of-the-art Datalog
-engines that is commonly used in points-to analysis
-\cite{Antoniadis:2017:PDS:3088515.3088522}, uses the idea of Futamura
+Such analyses can be considered having stage distinctions.  One piece of future
+work is to explore whether they can be implemented and optimized from
+meta-programming. 
+Souffl{\'e}, one of the state-of-the-art Datalog engines commonly used in
+points-to analysis \cite{Antoniadis:2017:PDS:3088515.3088522}, uses Futamura
 projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
-However, it only specializes Datalog rules to the running engine, instead
-of a full pipeline from program AST to generating and
-solving constraints.
+However, it only specializes Datalog rules to the running engine, instead of a
+full pipeline from a program AST to a specialized solver.
 
-Recent work shows that the numerical abstract domains can be
+Recent work shows that numerical abstract domains can be
 specialized w.r.t. the static program structure, although not by
@@
 \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to
-smaller ones that depend on the variables involved in a statement can
-significantly improve the running time and space of abstract
+smaller ones depending on the variables involved in a statement can
+significantly improve the running time and space consumption of abstract
 transformers.
 
-\paragraph{Construction of Abstract Interpreters} Abstract interpretation
-was proposed as a semantics-based approach to build sound static analysis by
-approximation \cite{DBLP:conf/popl/CousotC77}. As for building semantic
-artifacts, the Abstracting Abstract Machines (AAM)
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} approach
-shows that abstract interpreters can be derived systematically from
-concrete semantic artifacts. The AAM approach is closely related to
-control-flow analysis for higher-order languages
-\cite{Midtgaard:2012:CAF:2187671.2187672}.  Using monads to construct
-abstract interpreters is explored by
-\citet{Sergey:2013:MAI:2491956.2491979} and
-\citet{DBLP:journals/pacmpl/DaraisLNH17,
-  Darais:2015:GTM:2814270.2814308}.  The unstaged abstract interpreter
-in this paper follows the abstracting definitional interpreters
-approach from \citet{DBLP:journals/pacmpl/DaraisLNH17}.  Similar to
-definitional abstract interpreters,
-the reconstruction of big-step abstract interpreters with delimited
-continuations shown by \citet{Wei:2018:RAA:3243631.3236800},
-\citet{Keidel:2018:CSP:3243631.3236767} present a modular, arrow-based
-abstract interpreter that allows proof of soundness to be constructed
-compositionally. \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis. We borrow the notation of modular analysis for our approach.
-However the staged abstract interpreter does not generate complete summaries for
-modules; instead, it is able to specialize the analysis modularly. Similar, AC
-is considered as an example of this kind of modular analysis \cite{DBLP:conf/cc/CousotC02}.
+\paragraph{Construction of Abstract Interpreters} Abstract interpretation is a
+semantics-based approach to build sound static analysis by approximation
+\cite{DBLP:conf/popl/CousotC77, Cousot98-5}. 
+As for building semantic artifacts, the Abstracting Abstract Machines (AAM)
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} approach shows that
+abstract interpreters can be derived systematically from concrete semantic
+artifacts based on operational semantics. 
+The AAM approach is closely related to control-flow analysis for higher-order languages
+\cite{Midtgaard:2012:CAF:2187671.2187672, Shivers:1991:SSC:115865.115884}. 
+Using monads to construct abstract interpreters was explored by
+\citet{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17,
+Darais:2015:GTM:2814270.2814308}.
+The unstaged abstract interpreter in this paper follows the abstracting
+definitional interpreters (ADI) approach from
+\citet{DBLP:journals/pacmpl/DaraisLNH17}.  Similar to ADI, reconstructing
+big-step abstract interpreters with delimited continuations was shown by
+\citet{Wei:2018:RAA:3243631.3236800}; \citet{Keidel:2018:CSP:3243631.3236767}
+presented a modular, arrow-based abstract interpreter that allows soundness
+proof to be constructed compositionally. 
+\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
+modular analysis. We borrow the notation of modular analysis to our approach.
+However, the staged abstract interpreter does not generate complete summaries for
+modules; instead, it is able to specialize the analysis modularly. Similarly, AC
+is considered as an example of this kind of modular analyses \cite{DBLP:conf/cc/CousotC02}.
 
@@
 \cite{Shivers:1991:SSC:115865.115884, Midtgaard:2012:CAF:2187671.2187672}. In
-general, $k$-CFA (where $k > 0$) is considered impractical to use as a compiler
-pass, due to its EXPTIME lower bound \cite{VanHorn:2008:DKC:1411204.1411243}.
-However, several variants of 0CFA were developed and shown to be useful in
-practice \cite{Adams:2011:FTR:2048066.2048105,
-Bergstrom:2014:PEH:2628136.2628153, ashley:practical, Reppy:2006:TCA:1159876.1159888}.
+general, $k$-CFA (where $k > 0$) is considered impractical to be used as a
+compiler pass, due to its EXPTIME lower bound
+\cite{VanHorn:2008:DKC:1411204.1411243}.  However, several variants of 0CFA
+were developed and shown to be useful in practice
+\cite{Adams:2011:FTR:2048066.2048105, Bergstrom:2014:PEH:2628136.2628153,
+ashley:practical, Reppy:2006:TCA:1159876.1159888}.
 
@@
 was also explored by \citet{NIELSON198859}.
-\citet{Sergey:2013:MAI:2491956.2491979}'s work of monadic abstract
-interpreters is also closely related to the two-level semantics: the
+\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract
+interpreters is closely related to the two-level semantics: the
 use of a generic interface with monads and then reinterpreting it with
@@
 produce efficient low-level code. Hence the presented work can be
-considered as an two-dimensional application of the two-level semantics.
+considered as a two-dimensional application of the two-level semantics.
 
 \paragraph{Partial Evaluation and Multi-stage Programming}
-Partial evaluation as an automatic technique for program
-specialization was studied comprehensively by
-\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.  In this
-paper, we use mutli-stage programming as an approach to program
-specialization. The Lightweight Modular Staging framework
+Partial evaluation is an automatic technique for program
+specialization and was studied comprehensively by
+\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}. 
+In this paper, we use mutli-stage programming to specialize programs.
+The Lightweight Modular Staging framework
 \cite{DBLP:conf/gpce/RompfO10} we used in the paper relies on
@@
 MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
-  DBLP:conf/flops/Kiselyov14}.  The idea of staging an abstract
-interpreter we presented in this paper is also applicable to other MSP
-systems.  Multi-stage programming has been widely used to improve the
-performance in many domains, such as optimizing compilers or
+  DBLP:conf/flops/Kiselyov14}.  
+The idea of staging an abstract interpreter we presented here 
+is also applicable to other MSP systems.  
+Multi-stage programming has been widely used to improve the
+performance in many other domains, such as optimizing compilers or
 domain-specific languages \cite{DBLP:conf/pldi/RompfSBLCO14,
@@
 
-As a source of inspiration of this paper, the Futamura projections
-reveal a hierarchy and close relationship between interpreters and
-compilers, which was originally proposed by
-\citeauthor{futamura1971partial} in the 1970s\cite{futamura1971partial},
-and later reprinted \cite{Futamura1999}.
-\citeauthor{Amin:2017:CTI:3177123.3158140} considers a tower of
-concrete interpreters and how to collapse them by using MSP -- it
-would be interesting to explore this idea for multiple layers of
-abstract interpreters \cite{Cousot:2019:AAI:3302515.3290355,
-  Giacobazzi:2015:APA:2676726.2676987}.  \citet{10.1007/11561347_18}
-discussed a let-insertion, memoizing monads for code generation, using
-Gaussian Elimination in MetaOCaml as example. In this work,
-let-insertion is handled by the ANF transformation in LMS, and the
-memoizing monad here just combines a reader monad and state monad
-operating on staged data.
-Similar to the idea in this paper, specializing monadic interpreters
-was explored by \citet{danvy1991compiling, DBLP:conf/dsl/SheardBP99},
-but is different from the performance motivation here: the generated code
-is in monadic form.
+As a source of inspiration of this paper, the Futamura projections reveal a
+close relation between interpreters and compilers, which was
+originally proposed by \citet{futamura1971partial}, and later reprinted \cite{Futamura1999}.
+\citeauthor{Amin:2017:CTI:3177123.3158140} considered a tower of concrete
+interpreters and showed how to collapse them by using MSP -- it would be interesting
+to explore this idea for multiple abstract interpreters
+\cite{Cousot:2019:AAI:3302515.3290355, Giacobazzi:2015:APA:2676726.2676987}.
+\citet{10.1007/11561347_18} discussed a monad with let-insertion and
+memoization for code generation, using Gaussian Elimination in MetaOCaml as
+example. In this work, let-insertion is handled by the ANF transformation in
+LMS, and the memoizing monad is just a combination of a reader monad and a state monad
+operating on staged data.  Similar to the idea in this paper, specializing
+monadic interpreters was explored by \citet{danvy1991compiling,
+DBLP:conf/dsl/SheardBP99}. But they are different from the performance motivation
+here: the generated code is in monadic form.

### 2019-09-03 paper/90_conc.tex
misc update
revert to as sign; typo

@@
 To realize it, we design a generic interpreter that abstracts over both
-binding-time and value domains; by instantiating it with staged monads for
+binding-times and value domains; by instantiating it with staged monads for
 abstract interpretation, the staged abstract interpreter can be readily derived.
-The meta-programming approach brings new insight and implementation techniques
-to optimize static analyzers. The optimization can be realized in a completely
-orthogonal way to the interpreter and abstract semantics.
-In our empirical evaluation, staging improves an order of magnitude of
-performance in terms of running time.
+The meta-programming approach brings new insights and implementation techniques
+to optimize static analyzers. Optimizations can be achieved orthogonally to 
+the interpreter and abstract semantics.  In our empirical evaluation, staging
+improves performance by an order of magnitude in terms of running time.

### 2019-09-03 paper/abstract.tex
revert to as sign; typo

@@
 It is even more widely known that an abstract interpreter is a program analyzer:
-tweaking an interpreter to run on an abstract domain produces a sound static
+tweaking an interpreter to run on abstract domains produces a sound static
 analysis. What happens when we combine these two ideas, and apply specialization to

### 2019-09-03 paper/10_intro.tex
typos

@@
 reconcile these approaches, one must use a programming technique which allows
-for high-level abstractions during implementation, but which bypasses the
+for high-level abstractions during implementation, but bypasses the
 incurred overhead at runtime.
@@
 removes interpretation overhead and yields an equivalent executable
-that runs faster. The idea of Futamura projection has been applied to 
-other similar performance problems and shown success, examples including
+that runs faster. The Futamura projections has been successfully applied to 
+other similar performance problems, examples including
 query compilers\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
@@
 The first Futamura projection specifically shows that specializing an
-interpreter w.r.t. the input program yields an equivalent
-executable. For instance,
-@eval : Expr -> Input -> Value@
-is an interpreter for some language. Given a program $e_0$ of type
-@Expr@,
-by applying the specialization, we can obtain a specialized
-interpreter $\texttt{eval}_{\texttt{e0}}$ :
-@Input -> Value@,
-which is the so-called \textit{equivalent executable}.
+interpreter w.r.t. the input program yields an equivalent executable. For
+instance, @eval : Expr -> Input -> Value@ is an interpreter for some language.
+Given a program $e_0$ of type @Expr@, by applying the specialization, we can
+obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ : 
+@Input -> Value@, which is the so-called \textit{equivalent executable}.
 
@@
 However, it is hard to precisely analyze binding-times given an arbitrary
-program. As an alternative approach, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} lets users explicitly program over
+program. As alternative approach, multi-stage programming (MSP)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users explicitly program over
 a two-level syntax \cite{Nielson:1992:TFL:130665}, by manually
@@
 
-Similar to two-level semantics \cite{NIELSON1989117}, we found
-monads provide a good abstraction to hide the detail of abstract
-interpretation semantics, and allows us to fit code generation into
-the semantics. We use \citet{DBLP:journals/pacmpl/DaraisLNH17}'s
-abstract definitional interpreter that uses monad transformers as the
-unstaged baseline. However, the monads do not have a clear stage distinction.
-We further introduce binding-time and make the monads binding-time polymorphic.
+To apply Futamura projections on abstract interpreters, we consider
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s big-step monadic abstract
+interpreter as the unstaged baseline.  Similar to the two-level semantics
+\cite{NIELSON1989117}, monads provide a good abstraction to hide the detail of
+abstract interpretation semantics. However, the monads do not have a clear
+stage distinction.
+We further introduce binding-times and make the monads binding-time polymorphic.
 Our proposed framework adopts the type-based multi-stage programming from the
@@
 obtain a staged abstract interpreter that can be used for specialization. After
-staging, the generated code is specialized to the input program, and all
+staging, the generated code is specialized to the input program, and all the
 monadic operations are inlined and compiled down to low-level Scala code.
@@
 
-Although the four artifacts may look like dissimilar at first glance, they in
+Although the four artifacts may look dissimilar at first glance, they in
 fact are all firmly rooted in the concrete semantics of the language.  This
@@
 
-\paragraph{Applications and Evaluations}
+\paragraph{Applications and Evaluation}
 We evaluate the idea of staging an abstract interpreter through
@@
 \cite{Boucher:1996:ACN:647473.727587}, an implementation technique for
-control-flow analyses, and show that by utilizing type-based stage
-annotations we can achieve the same optimizations. Meanwhile,
-the analyzer does not need to change, thereby requiring significantly less
-engineering effort.
+control-flow analyses. We show that by utilizing type-based stage
+annotations, we can achieve the same optimizations. Meanwhile,
+the analyzer does not need to change, except the stage annotations, 
+thereby requiring significantly less engineering effort.
 2) We extend the basic staged abstract interpreter to different flow
@@
 analysis code by mechanized reuse of a whole-program analyzer.
-4) We also empirically evaluate the performance improvements gained by staging,
-showing an order of magnitude speedup on flow-analysis.
+4) We empirically evaluate the performance improvements gained by staging,
+showing an order of magnitude speedup on flow analyses.
 
@@
 \begin{itemize}[leftmargin=2em]
-  \item To our best knowledge, we present the first
-    multi-stage programming system to specialize a general abstract interpreter,
-    without our approach affecting the soundness of analyses.
-  \item Intellectually, our framework naturally extends the first
-    Futamura projection to abstract interpreters, showing a
-    well-grounded approach to optimizing static analyses via
-    meta-programming.
-  \item Practically, we show that staging an abstract interpreter is
-    useful to improve performance and scalability of analyses by case
-    studies and an empirical evaluation.
+  \item Intellectually, we present a framework that naturally extends the first
+    Futamura projection to definitional abstract interpreters, showing a
+    well-grounded approach to optimizing static analyses via meta-programming.
+  \item Practically, we show that staging an abstract interpreter is useful to
+    improve performance and scalability of analyses by case studies and an
+    empirical evaluation.
 \end{itemize}

### 2019-09-03 paper/60_cases.tex
typos

@@
 
-We borrow the notation traditional modular analysis
-\cite{DBLP:conf/cc/CousotC02} and realize that such generated function is a
+We borrow the notation from traditional modular analysis
+\cite{DBLP:conf/cc/CousotC02}: such generated function is a
 reusable artifact which can be used modularly, i.e., independent to its context. 

### 2019-09-04 paper/20_prelim.tex
revert to as sign; typo
revision

--- 2019-08-15 paper/20_prelim.tex
@@
 In this section, we first describe the abstract syntax of the language for our interpreters. 
-Then present the generic interpreter shared among the four different
-semantics, after which, we instantiate the interpreter to the concrete one.
+Then we present the generic interpreter shared among the four different
+semantics. After which, we instantiate the interpreter to the concrete one.
 It is worth noting that we choose to use Scala and monadic style to
-demonstrate the idea, but the approach is not restricted to our
+demonstrate the idea, but the approach is not restricted to this
 choice. One can use imperative or direct style in other MSP languages
@@
 such as assignments can also be supported readily.
-%In Section~\ref{cases_imp}, we will add more imperative features to the language.
 Since we are mostly interested in analyzing the dynamic behaviors of the
-program, we disguise any static semantics and type system. We also assume that
+program, we will disguise the static semantics. We also assume that
 input programs are well-typed and all variables are distinct. The abstract
 syntax is shown as follows:
-
 \begin{lstlisting}
@@
 \begin{lstlisting}
-  val xs = List(1, 2); val ys = List(4, 5)
-  for { x <- xs; y <- ys } yield (x, y) // List((1,4), (1,5), (2,4), (2,5))
+  val xs = List(1, 2); val ys = List(3, 4)
+  for { x <- xs; y <- ys } yield (x, y)   // List((1,3), (1,4), (2,3), (2,4))
 \end{lstlisting}
@@
 identifier @x@, through the environment and store. Conditionals and
-arithmetic is handled by @br0@ and @arith@, respectively. The method
+arithmetic are handled by @br0@ and @arith@, respectively. The method
 @ap_clo@ is used for function applications: it takes a function value

### 2019-09-04 paper/20_prelim.tex
revision

@@
 
-Similar to Haskell's @do@-notation, Scala provides special syntactic support for
-monadic operations through @for@-comprehension.
-For example, an object of @List[A]@ is an instance of @List@ monad, where @A@ is the element type. 
-Then to compute the Cartesian product of two lists of numbers, we can use Scala's
-@for@-comprehension syntax.
+Similar to Haskell's @do@-notation, Scala provides special syntactic support
+for monadic operations through @for@-comprehension.  For example, an object of
+@List[A]@ is a @List@ monad, where @A@ is the element type.
+To compute the Cartesian product of two lists of numbers, we can use Scala's
+@for@-comprehension syntax:
 \begin{lstlisting}
@@
 in the @for@-comprehension is translated into a @map@, where the expression of
-@yield@ becomes the body expression of that @map@ application. The foregoing
-bindings in the comprehension are all translated into calls of @flatMap@.
+@yield@ becomes the body expression inside of that @map@ operation. The foregoing
+bindings in the comprehension are all translated into calls of @flatMap@:
 \begin{lstlisting}
@@
 
-Note that here the monadic object @List[_]@ encapsulates the data
-internally.  Therefore it only exposes the simplified version of
-@flatMap@, where the monad @M[A]@ is not introduced as a function
-argument. The trait @MonadOps@ (Figure \ref{fig:monadops}) defines the
-simplified version of monadic operations that are necessary for
-@for@-comprehension. The conversion between @Monad@ and @MonadOps@
-can be done by using the implicit design pattern.
-In the rest of the paper, we use Scala's @for@-comprehension syntax
-and a few of monads and monad transformers such as @ReaderT@,
-@StateT@, and @ListT/SetT@ to write our interpreters. The implementation
-of monads and monad transformers essentially borrows the ground-truth
-from Haskell.
+Note that here the monadic object @List[_]@ encapsulates the data internally.
+Therefore it only exposes the simplified version of @flatMap@, where the monad
+@M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
+\ref{fig:monadops}) defines the simplified version of monadic operations that
+are necessary for @for@-comprehension. The conversion between @Monad@ and
+@MonadOps@ can be done by using implicit design pattern.
+In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
+of monads and monad transformers such as @ReaderT@, @StateT@, and @SetT@ to
+write our interpreters.
+Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
+take a monad as argument and produces another monad. By using monad
+transformers, we can combine multiple monads into a single one.  The
+implementation of monads and monad transformers essentially borrows the
+ground-truth from Haskell.
 
@@
 
-Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
-take a monad as argument and produces another monad. By using monad
-transformers, we can combine multiple monads into a single one. Constructing
-extensible interpreters using monad transformers was first explored by
-\citet{DBLP:conf/popl/LiangHJ95}, and later has been applied to abstract
-interpreters \cite{Sergey:2013:MAI:2491956.2491979,
-  DBLP:journals/pacmpl/DaraisLNH17, Darais:2015:GTM:2814270.2814308}.
 In this section, we present the generic interface in the style of big-step
 definitional interpreter. The key idea is to keep the binding-time type and
-returned monadic type abstract, so that they can be instantiated differently.
-We also need to abstract the primitive operations on those types.
+returned monadic type both abstract, so that they can be instantiated
+differently; we also need to abstract the primitive operations on those types.
+In later sections, we will instantiate the monadic type to perform concrete
+interpretation \citet{DBLP:conf/popl/LiangHJ95} or abstract interpretation
+\cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17}.
 
-\paragraph{Basic Types} We start with some basic type definitions used in the
-interpreter. The identifiers in the program are represented by
-strings. To represent states, two required components for the
-interpreter are environments @Env@ and stores @Store@. @Env@ maps from
-identifiers to addresses, and @Store@ maps from addresses to values,
-respectively. The @Env@ captures bound variables in scope, and @Store@
-models a persistent heap through the program run-time.  At this moment,
-the domains of addresses and values are still abstract, i.e., they are
-declared as abstract types.
+\paragraph{Basic Types} 
+We start with some basic type definitions that are used in the interpreter. The
+identifiers in the program are represented by strings. To represent states, two
+required components for the interpreter are environments @Env@ and stores
+@Store@. @Env@ maps from identifiers to addresses, and @Store@ maps from
+addresses to values, respectively. The @Env@ captures bound variables in the
+scope of current control, and @Store@ models a persistent heap through the
+program run-time.  At this moment, the domains of addresses and values are
+still abstract, i.e., they are declared as abstract types.
 \begin{lstlisting}
@@
 
-\paragraph{Binding-time Abstraction} As mentioned before, the binding-time is
-declared as a higher-kinded type @R[_]@. We are free to use @R@ to annotate
-other data-types in the interpreter; it is also injected into the monadic
-type @MonadOps@.
-If we simply instantiate @R@ an identity type (i.e., @type R[T] = T@),
-then the generic interpreter is a standard definitional interpreter
-that will execute the program.  In Section \ref{stagedinterp}, we will
-instantiate @R@ using LMS's built-in next-stage type annotation @Rep@,
-which makes the interpreter act as a compiler.
+\paragraph{Binding-time Abstraction}
+The binding-time is declared as a higher-kinded type @R[_]@. 
+Now we are able to use @R@ to annotate other data-types in the interpreter;
+it is also injected into the monadic type @MonadOps@.  If we simply instantiate
+@R@ an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
+standard definitional interpreter that will execute the program.  In Section
+\ref{stagedinterp}, we will instantiate @R@ using LMS's built-in next-stage
+type annotation @Rep@, which makes the interpreter act as a compiler.
 
 \paragraph{Monadic Operations} We define the return type of the interpreter as
-@Ans@, which is a monad type @AnsM[_]@ wrapping the type @Value@. 
+@Ans@, which is a monadic type @AnsM[_]@ wrapping the type @Value@. 
 As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
 syntax, certain operations have to be added on the type @AnsM@. Here, we use a
-structural type @MonadOps@ to require that type @AnsM@ at least implements @map@ and
-@flatMap@. It is worth noting that @MonadOps@ takes another type parameter
-@R[_]@ as binding-time that annotates data; accordingly the type @R@
-defined in the outer scope is passed to @MonadOps@. Inside of @MonadOps@,
-@R[_]@ wraps the data types @A@ and @B@ that are encapsulated by the
-monad, but not the monad type @M@ itself. When acting as compilers, we
-will also replace the monads to the ones that work on staged data values.
+structural type @MonadOps@ to require that type @AnsM@ should at least
+implement @map@ and @flatMap@. It is worth noting that @MonadOps@ takes another
+type parameter @R[_]@ as binding-time, i.e., @MonadOps[R, AnsM, T]@.
+Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
+encapsulated by the monad, but not the monad type @M@ itself. When acting as
+compilers, we will also replace the monads to the ones that work on staged data
+values.
 
@@
 stores. These methods return monadic values of type @AnsM[_]@, which
-may encapsulate over the environment or store, or simply a @Unit@
+may parameterized over the environment or store type, or simply a @Unit@
 value for effects. For examples, @local_env@ installs a new environment
@@
 
-\paragraph{Primitive Operations} Next we define several primitive operations.
+\paragraph{Primitive Operations} Next we define a few primitive operations.
 First, two versions of @alloc@ are declared. The first one takes a store and an
@@
 the freshness of the address may depend on the store, which might be a
-next-stage value as indicated by its type, the type of addresses is therefore
-wrapped by @R[_]@. The other @alloc@ is simply the monadic version of the addresses.
+next-stage value as indicated by its type, the type of addresses is
+consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
+of the addresses, therefore can be used in monadic computations.
+An auxiliary method @get@ retrieves the value of an identifier @x@, through the
+environment and store. 
 \begin{lstlisting}
   def alloc(σ: R[Store], x: Ident): R[Addr];  def alloc(x: Ident): AnsM[Addr]
+  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
 \end{lstlisting}
 
-Other primitive operations provide basic functionality for the
-interpreter.  The method @num@ and @close@ deal with primitive values,
-which lift literal terms (e.g., lambdas) to our value representation
-(e.g., closures).  The method @get@ retrieves the value of an
-identifier @x@, through the environment and store. Conditionals and
-arithmetic are handled by @br0@ and @arith@, respectively. The method
-@ap_clo@ is used for function applications: it takes a function value
-and an argument value. Note that the @Env@, @Store@, and @Value@ are
-all annotated by @R[_]@ since they can potentially be next-stage
-values when acting as compilers.
+Other primitive operations in the interpreter handles language constructs.  The
+method @num@ and @close@ deal with primitive values, which lift literal terms
+(e.g., lambdas) to our value representation (e.g., closures).  
+Conditionals and arithmetic are handled by @br0@ and @arith@, respectively. The
+method @ap_clo@ is used for applying functions, which takes a function value
+and an argument value. Note that the @Env@, @Store@, and @Value@ are all
+annotated by @R[_]@ because they can potentially be next-stage values when the
+interpreter acts as a compiler.
 \begin{lstlisting}
   def num(i: Int): Ans
-  def get(σ: R[Store], ρ: R[Env], x: Ident): R[Value]
   def close(ev: Expr => Ans)(λ: Lam, ρ: R[Env]): R[Value]
@@
 
-\paragraph{The Interpreter} Now we construct the semantics-agnostic interpreter
+\paragraph{The Interpreter} Now we can construct the semantics-agnostic interpreter
 in monadic form, shown in Figure \ref{fig:shared_int}. The basic idea
 of a generic interpretation is to traverse the abstract syntax tree
-while maintaining the effects such as reader and state.  Note that the
-interpreter is written in open-recursive style -- it can not refer to
-itself directly, instead, @eval@ takes an additional parameter @ev@ of
-type @Expr => Ans@ that refers to itself. Consequently, the method
-@close@ that lifts lambda terms to closures and method @ap_clo@ that
-applies functions also takes an extra @ev@, because evaluation may
-happen inside of them.
+while maintaining the effects such as environment and state updates.  
+Note that the interpreter is written in open-recursive style -- it cannot
+refer to itself directly, instead, @eval@ takes an additional parameter @ev@ of
+type @Expr => Ans@ that refers to itself. Consequently, the method @close@ that
+lifts lambda terms to closures and method @ap_clo@ that applies functions also
+takes an extra @ev@, because evaluations may happen inside of them.
 To close the open-recursion, we use a fixed-point operator @fix@.
-For concrete-interpretation instantiation, it works like the Y
-combinator; for abstract-interpretation instantiation, it instruments
-the interpreter by memoizing its input and output, which will ensure
-the termination of abstract interpretation.
-Finally, a top-level wrapper @run@ is declared; the return type
-@Result@ depends on what kind of monad we will be using, this is also
-an abstract type.
+For concrete-interpretation instantiation, @fix@ works like the Y combinator;
+for abstract-interpretation instantiation, it will instrument the interpreter
+by memoizing @ev@'s inputs and outputs, which ensures the termination of
+abstract interpretation.
+Finally, a top-level wrapper @run@ is declared; the return type @Result@
+depends on what kind of monads is being used, therefore this is also an
+abstract type.
 \begin{lstlisting}

### 2019-09-04 paper/70_eval.tex
revision

@@
 
-To evaluate how the performance can be improved by our staged abstract interpreters
+To evaluate how performance can be improved by our staged abstract interpreters
 approach, we scale our toy interpreter to a large subset of the Scheme language
@@
 shows that the staged versions are averagely 10+ times as fast as the unstaged
-version, as strong evidence of our \textit{abstraction without regret}
-approach.
+version, as evidence of our \textit{abstraction without regret} approach.
 
@@
 interpreters for control-flow analysis, i.e., a value-flow analysis that
-computes which lambda terms can be called at each call-sites in functional
-languages. With other suitable abstract domains, the abstract interpreter also
-can be extended to other analyses. A front-end desugars Scheme to a small core
-language that makes the analyzer easy to implement. We implement two common
-monovariant 0-CFA-like analysis, one is equipped with store-widening (0CFA-SW
-for short), the other does not (0CFA for short). The store-widened analysis is
-expected to be faster than the one without.  Our prototype implementation
-currently generates Scala code; the generated Scala code will be compiled and
-executed on JVM. It is possible to generate C/C++ code in the future. In the
-experiments, we implement several optimizations mentioned in Section
-\ref{staged_ds}, specifically, the selective caching scheme and IR-level
-rewriting rules (for pairs, lists, sets, and maps) that exploit
-partially-static data. These optimizations are useful to reduce the size of
-generated code.
+computes which lambda terms can be called at each call-site. 
+With other suitable abstract domains, the abstract interpreter
+can also be extended to other analyses. 
+A front-end desugars Scheme to a small core language that makes the analyzer
+easy to implement. We implement two common monovariant 0-CFA-like analysis, one
+is equipped with store-widening (0CFA-SW for short), the other does not (0CFA
+for short). Usually, the store-widened analysis is expected to be faster than the one
+without.  Our prototype implementation currently generates Scala code; the
+generated Scala code will be compiled and executed on JVM. It is possible to
+generate C/C++ code in the future. In the experiments, we implement several
+optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
+caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps)
+that exploit partially-static data. These optimizations are useful to reduce
+the size of generated code.
 
@@
 \paragraph{Benchmarks}
-The benchmark programs we used in the experiment are collected from several
+The benchmark programs we used in the experiment are collected from
 previous papers \cite{Johnson:2013:OAA:2500365.2500604, ashley:practical,
@@
 applications, for examples, the RSA public key encryption algorithm
-@rsa@. In the Figure \ref{evaluation_result}, we report the number of
-AST node after parsing and desugaring (excluding comments) as a proper
+@rsa@. In Figure \ref{evaluation_result}, we report the number of
+AST nodes after parsing and desugaring (excluding comments) as a proper
 measurement of program size.
@@
 \vspace{-1em}
-\begin{figure}[h]
-\footnotesize
+\begin{figure}[h] \footnotesize
 \begin{tabular}{@{}ll|lll|lll@{}}
@@
 Figure~\ref{evaluation_result} shows the evaluation result, comparing the
-performance improved by staging on the two monovariant CFAs The
-\textit{unstaged} and \textit{staged} columns show the median time to finish
-the analysis. The column $\frac{\text{unstaged}}{\text{staged}}$ shows the
-improvement. The dash - in some cells indicates timeout.  As
-expected, the staged version significantly outperforms the unstaged version on
-all benchmarks. We observe that simple and homogeneous value domains used in the
-benchmark usually leads to better speed-ups.  For 0CFA w/o store-widening, we
-observe an average speed-up of ~17x times; for 0CFA w/ store-widening, the
-average speed-up is ~7x times.  The \texttt{kcfa-worst-64} benchmark has the most
-speed-up among all the tests. Although it has a comparably large size of AST,
-it has a quite simple structure which only involves booleans, functions and
-applications. We conjecture that the simple domain and structure allow the
-generated code to be efficiently optimized by the JVM. But on the corresponding
-0CFA-SW, we do not observe such great speed-up. Due to the complexity of
-generated code and next-stage running platform (JVM), we leave the detailed
-empirical analysis of which static analysis can benefit more from
-specialization as future work.
+performance improved by staging on the two monovariant CFAs. The
+\textit{unstaged} and \textit{staged} columns show the median time (excluding
+pre-compilation) to finish the analysis. The column
+$\frac{\text{unstaged}}{\text{staged}}$ shows the improvement. The dash @-@ in
+some cells indicates timeout.  As expected, the staged version significantly
+outperforms the unstaged version on all benchmarks. 
+%We observe that simple value domains used in the benchmark usually leads to
+%better speed-ups, for example, the @kcfa-worst@ series only uses lambdas and
+%booleans.
+For 0CFA w/o store-widening, we observe an average speed-up of ~17x times; for
+0CFA w/ store-widening, the average speed-up is ~7x times.
+\texttt{kcfa-worst-64} has the most speed-up among all the tests.
+Although it has a comparably large size of AST, it has a quite simple structure
+which only involves booleans, lambda terms and applications. We conjecture that
+the simple domain and structure allow the generated code to be efficiently
+optimized by the JVM. But on the corresponding 0CFA-SW, we do not observe such
+great speed-up. Due to the complexity of generated code and next-stage running
+platform (JVM), we leave the detailed empirical analysis of which static
+analysis can benefit more from the specialization as future work.
 

### 2019-09-04 paper/30_stageconc.tex
revision

@@
 In this section, based on the generic interpreter and concrete components we
-presented in Section \ref{prelim}, we show how to stage the concrete interpreter
-by changing the monad type and refactoring several primitive operations. We
-begin by briefly introducing the Lightweight Modular Staging framework in Scala,
-and then replay the same steps from the unstaged counterpart. At last, we briefly
-describe code generation.
+presented in Section \ref{prelim}, we show that how to stage the concrete
+interpreter by changing the monad type and refactoring primitive
+operations. We begin by briefly introducing the Lightweight Modular Staging
+framework in Scala, and then replay the same steps from the unstaged
+counterpart. Finally, we briefly describe code generation.
 
@@
 Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
-multi-stage programming framework implemented as a Scala library that enables
-dynamic code generation in a type-safe manner. Different from the approach of
-MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} that
-uses syntactic quotations and quasi-quotations, LMS distinguishes binding-time
-based on types. LMS provides a type constructor @Rep[T]@ where @T@ can be
-an arbitrary type, indicating a value of @T@ will be known at the next stage.
-All operations acting on @Rep[T]@ expressions will be residualized as generated
+multi-stage programming framework implemented as a Scala library. LMS enables
+runtime code generation in a type-safe manner. Different from the approach of
+MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} which
+use syntactic quotations, LMS distinguishes binding-time based on types. 
+LMS provides a type constructor @Rep[T]@ where @T@ can be
+an arbitrary type, indicating a value of type @T@ will be known at future stage.
+Operations acting on @Rep[T]@ expressions will be residualized as generated
 code.
@@
 A classic example for introducing multi-stage programming is the power function
-that computes $b^x$, which is usually implemented as a recursive function:
+that computes $b^x$. Usually it is implemented as a recursive function:
 \begin{lstlisting}
@@
 
-If @x@ is a value know at current stage, we may specialize the power function to
-some value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
-first adding the @Rep@ type to the variable known at next stage. In this case,
-@b : Rep[Int]@ is know later, and @x : Int@ is know currently, as shown in the
+If @x@ is a value known at the current stage, we may specialize the power function to
+the value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
+first adding the @Rep@ type annotation to variables known at next stage. In this case,
+@b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
 below code recipe.
-The way we use this @power@ function is to create a @DslDriver@ and override the
-@snippet@ method give the currently know value @x@ (5 in this example).
+The way we use this staged @power@ function is to create a @DslDriver@ and override the
+@snippet@ method that gives the currently know value @x@ (5 in this example).
 \begin{lstlisting}
@@
 The idea behind the framework is to construct a sea-of-node intermediate
-representations (IR) for the next stage program at the current stage
+representations (IR) for the next-stage program at the current stage
 \cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
-expressions of type @Rep[_]@ to their IR is done by using the implicit design
+expressions of type @Rep[_]@ to their IR is done by using implicit design
 pattern. As we will see later, implementing staging and code generation
@@
 
-To implement the staged concrete interpreter, we replay the steps from
-instantiating unstaged concrete interpreter in Section \ref{unstaged_conc}.
-But now we use the @Rep@ type to annotate value domains, environments and
+To implement the staged concrete interpreter, we replay the steps from the
+instantiation of unstaged concrete interpreter in Section \ref{unstaged_conc}.
+But, now we use the @Rep@ type to annotate the value domains, environments and
 stores, and redefine the staged version of monads and primitive operations.
@@
 We use the same structure of monad stack in the unstaged interpreter: a reader
-monad with a state monad. But now the monads operate on staged values, for
-brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
+monad with a state monad. The difference here is, now the monads operate on staged values. 
+For brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
 prefix on the constructors and types to differentiate them. But it is important
-to note that objects of a case class @ReaderT@/@StateT@ are not staged, the
+to note that instances of the case class @ReaderT@/@StateT@ are not staged, also the
 monadic computation like @R => M[A]@ are also not staged. Instead, the internal
@@
 
-The function @f@ passed to @flatMap@ is also a current-stage-known value, so
-that all the invocation of @flatMap@s can be completed before generating code.
+The function @f@ passed to @flatMap@ is also a function value known at current stage.
+Therefore all invocations of @flatMap@s can be reduced before generating code.
 The fact that we only stage data but not monadic computation or monadic
 values, is the reason that we can peel of the monad stack in the generated code.
-Now the @AnsM[_]@ is instantiated as the same structure as before, but using the
+Now type @AnsM[_]@ is instantiated as the same structure as before, but using the
 @Rep@ versions:
@@
 be achieved without modifying the implementation of monads. This is true so far,
-but as we will see, it is not straightforward when for nondeterminism monad
+but as we will see, it is not straightforward for nondeterminism monad
 (@SetT@) in abstract interpreters. Because not only the elements in the set are
@@
 many elements in the set. In this section, we explicitly distinguish the two
-versions of monadic interfaces, later we will instantiated staged set monad
-using a different strategy.
+versions of monadic interfaces, later we will instantiated the staged set monad
+by manually fusing part of the monad stack.
 
@@
 As we mentioned before, we will take more care on how functions and
-applications are handled. The problem here is what should we do for
-lambda terms when staging? We cannot create a next-stage
-defunctionalized closure for it, because it means we still need to
-access the interpreter at the next stage. The desired goal is to eliminate the
-interpretation overhead staging, so the right thing is compiling the lambda term
-and its environment to a next-stage Scala function. To do this, we need to
-recursively call the interpreter function, i.e., the @ev@ argument passed to @close@.
-The following code implements the idea:
+applications are handled.  One problem here is what should we do for
+$\lambda$ terms when staging? We cannot directly create a next-stage
+defunctionalized closure for it, because that means we still need to
+interpret over these closure at the next stage.  The desired goal is to
+eliminate the interpretation overhead by staging, so the right thing is to
+compile the lambda term with its environment to a next-stage Scala function. To
+do this, we need to recursively call the interpreter function at current stage,
+i.e., the @ev@ argument passed to @close@.  The following code implements the
+idea:
 \begin{lstlisting}
@@
 
-We have created a current-stage function @f@ that takes a next-stage value and
-store, and returns staged value. Then we use @emit_complied_clo@ to delay @f@ to
-a next-stage Scala function, represented by the type @Rep[Value]@. Inside of
-@f@, we can access the evaluator via @ev@, but @ev@ produces a monadic value of
-type @AnsM@, which can only exist at the current stage. To connect the
-current-stage monadic values and future-stage grounded values, the evaluator
-@ev@ is invoked on the body expression @e@ to be specialized w.r.t. @e@, and
-then we \textit{collapse} the @AnsM@ monad to grounded values by providing the
-prepared environment and store.
+In @close@, we first create a current-stage function @f@ that takes a
+next-stage value and store, and returns staged values. Then we use
+@emit_complied_clo@ to denote @f@ to a next-stage Scala function, represented
+by the type @Rep[Value]@. Inside of @f@, we can access the evaluator via @ev@.
+However, @ev@ produces a monadic value of type @AnsM@, which can only exist at
+the current stage. To connect the current-stage monadic values and future-stage
+grounded values, the evaluator @ev@ is invoked on the body expression @e@ to be
+specialized w.r.t. @e@, and then we \textit{collapse} the monadic value of
+@AnsM@ to grounded values by providing the prepared environment and store.
 
 For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
-and @arg@, but what we can do on these values of @Rep@ type? In fact, @fun@ does
-not have an intensional application operation we can use directly; but we know
-that it is a staged value produced by the @close@ function, which represents a
+and @arg@. But what we can do on the values of @Rep@ type is limited. In fact, @fun@ does
+not have an intensional application operation we can use directly. We only know
+that it is a staged value produced by @close@, and represents a
 next-stage Scala function. Therefore we can just generate a next-stage function
-application, also in Scala. Since @ap_clo@ returns a monadic value of type
+application, also in Scala. Since @ap_clo@ produces a monadic value of type
 @Ans@, we can still use the monadic operation @get_store@ to obtain the latest
-store object. Then we use @emit_ap_clo@ to generate a current-stage
+store object. Then @emit_ap_clo@ is used to generate a current-stage
 representation of the future-stage application result, which is a pair of values
 and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
-replace the store and reify the result to the current-stage monadic value.
+update the store and reify the result into the current-stage monadic value.
 \begin{lstlisting}
@@
 evaluation function @ev@ from application-time to value-representation-time: in
-the unstaged interpreter, the invoke of @ev@ happens in @ap_clo@; while in the
-staged interpreter, we eagerly apply @ev@ when denoting the lambda term to our
-value domain.
+the unstaged interpreter, invocations of @ev@ happens in @ap_clo@; while in the
+staged interpreter, we eagerly apply @ev@ when denoting lambda terms to our
+next-stage value domain.
 
@@
 
-In the previous section, we merely show the type of @emit_compiled_clo@ and
-@emit_ap_clo@ without a concrete implementation. In this section, we briefly
+In the previous section, we merely show the types of @emit_compiled_clo@ and
+@emit_ap_clo@ without concrete implementations. In this section, we briefly
 discuss the IR node generated by them and sketch code generation.
 We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
-extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
-representing next-stage definitions of type @T@.
+extending from @Def[T]@. @Def[T]@ is a built-in type from the LMS framework,
+representing next-stage value definitions of type @T@.
 \begin{lstlisting}
@@
 next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
-programmers to control what is generated for each kind of node. Here we have
-matched @IRCompiledClo@ and @IRApClo@ and will generate definitions via @emitValDef@ for them.
-For compiled closures, we put the function @f@ and its accompanying syntactic term
-and environment to its next-stage value representation @CompiledClo@. For
-applications, we have the knowledge that @fun@ is a @CompiledClo@ at the stage,
-and it has a callable field @f@, so we generate a next-stage function
-application @f(arg, σ)@. The @quote@s are invocations to generate code for the
-other next-stage values.
+programmers to control what is generated for each kind of nodes. Here we match
+@IRCompiledClo@ and @IRApClo@ and generate definitions via @emitValDef@ for
+them.  For compiled closures, we put the function @f@ and its accompanying
+syntactic term and environment to its next-stage value representation
+@CompiledClo@. For applications, we have the knowledge indicating that @fun@ is
+a @CompiledClo@ at the stage, and it has a callable field @f@. Hence, we
+generate a next-stage function application @f(arg, σ)@. The invocations of
+@quote@s are used to generate code for the other next-stage values.
 \begin{lstlisting}

### 2019-09-04 paper/10_intro.tex
Proofreading intro

@@
 approaches. First, similar to concrete interpreters, an abstract interpreter
-analyzes the program by traversing the abstract syntax tree (AST). The traversing incurs
+analyzes the program by traversing the abstract syntax tree (AST). The traversal incurs
 interpretive overhead, such as the pattern matching on the ASTs and recursive
@@
 projection shows that specializing an interpreter to the input program
-removes interpretation overhead and yields an equivalent executable
-that runs faster. The Futamura projections has been successfully applied to 
+removes interpretative overhead and yields an equivalent executable
+that runs faster. The Futamura projections have been successfully applied to
 other similar performance problems, examples including
 query compilers\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
-\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275} and etc.
+\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275}, and etc.
 The challenge here: can we apply the same idea to a general abstract interpreter
@@
 semantics that extensively uses nondeterminism, operates on abstract domains,
-and computes fixed-point iteratively over a cache. The solution is not simply obvious.
+and computes fixed-points iteratively over a cache. The solution is not simply obvious.
 
@@
 interpreters, while keeping the implementation as close to their conceptual
-model.
+model as possible.
 In short, our approach borrows ideas and techniques from multi-stage
@@
 abstracts over binding-times and different semantics, which allows the staged
-abstract interpreter can be derived from its unstaged counterpart.
+abstract interpreter to be derived from its unstaged counterpart.
 As a result, the derived staged abstract interpreters has no intrusive changes
-to the underlying abstract semantics, therefore preserves the soundness.
+to the underlying abstract semantics, thereby preserving the soundness.
 In this sense, our approach allows for no regret for both performance and
@@
 However, it is hard to precisely analyze binding-times given an arbitrary
-program. As alternative approach, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users explicitly program over
+program. As an alternative approach, multi-stage programming (MSP)
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users to explicitly program over
 a two-level syntax \cite{Nielson:1992:TFL:130665}, by manually
@@
 concepts.  To implement the confluence of them, we first construct a
-generic interpreter that is agnostic to both binding-times and value
+generic interpreter that is agnostic to both binding times and value
 domains used in the semantics.  Later, the generic interpreter can be
@@
 annotations, we can achieve the same optimizations. Meanwhile,
-the analyzer does not need to change, except the stage annotations, 
+the analyzer does not need to change, except the addition of stage annotations,
 thereby requiring significantly less engineering effort.

### 2019-09-04 paper/20_prelim.tex

@@
 this section. The result of such instantiation is a standard definitional
-interpreter with environments and stores, it also can be obtained by
+interpreter with environments and stores. It can also be obtained by the
 refunctionalization of CESK machines \cite{Felleisen:1987:CAH:41625.41654,
 DBLP:conf/ppdp/AgerBDM03}. We first present the concrete components, i.e., the
-value domain, then show the monad stack for concrete interpretation, finally
-sketch the how the primitive operations are implemented.
+value domains, then show the monad stack for concrete interpretation, finally
+sketch how the primitive operations are implemented.
 
@@
 type @Env@ and @Store@ is derived automatically. To assure the freshness of
-address allocation, we use @Int@ and always return successor of the size of
-current store. A value can be either a tagged number \texttt{IntV}, or a closure
+address allocations, we use type @Int@ and always return a number greater
+than the size of current store.
+A value can be either a tagged number \texttt{IntV}, or a closure
 \texttt{CloV} that contains a lambda term and an environment. The final grounded
 result of the interpreter is a pair of values and stores, where the @Value@ and
-@Store@ are next-stage objects, as annotated by @R@. We also define a standard
-fixed-point combinator to close the open-recursive function @ev@.
+@Store@ can be next-stage objects, as annotated by the binding-time @R@. We
+also define a standard fixed-point combinator to close the open-recursive
+function @ev@.
 \begin{lstlisting}
@@
 respectively. Following the monad transformer approach
-\cite{DBLP:conf/popl/LiangHJ95}, we use a @ReaderT@ and @StateT@ monad
-transformer to compose the monad stack. In other words, the type @AnsM@ is
+\cite{DBLP:conf/popl/LiangHJ95}, a @ReaderT@ and @StateT@ monad
+transformer are used to compose the monad stack. In other words, the type @AnsM@ is
 instantiated by layering @ReaderT@ and @StateT@ transformers\footnote{The
@@
 M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is parameterized by type
-@Env@, @StateT@ is parameterized by type @Store@, and the inner-most monad @IdM@
+@Env@, and @StateT@ is parameterized by type @Store@, and the inner-most monad @IdM@
 is simply an identity monad.
@@
 A @ReaderT@ monad transformer encapsulates computation @R => M[A]@, where
-@R@ is the reader type, and @M[_]@ is the inner monad type.
-Given a @R@, @ReaderT@ monad produces a transformed value of type @M[A]@.
+@R@ is the reader type, and @M[_]@ is the inner monadic type.
+Given a value of @R@, a @ReaderT@ monad produces a transformed value of type @M[A]@.
 Similarly, a @StateT@ monad encapsulates computation @S => M[(A, S)]@, where
 @S@ is the state type, and @M[_]@ is the inner monad type.
-Given a @S@, @StateT@ monad produces a transformed value of type @M[(A, S)]@,
-where the new state (type @S@) is accompanied with the value (type @A@).
+Given a value of @S@, a @StateT@ monad produces a transformed value of type @M[(A, S)]@,
+where the new state (of type @S@) is accompanied with the value (of type @A@).
 Note that for the moment, the binding-time type @R@ is an identity type, thus
@@
 
-After defining the monad stack, the operations that manipulate the environment
-and store can be defined by constructing the proper monad and lifting it to the
-top-level of our monad stack. To modify the store, for example, we can construct a
+After defining the monad stack, the operations manipulating environments
+and stores can be defined by constructing the proper monadic value and lifting it to the
+top-level of our monad stack. To modify the store, for example, we construct a
 @StateT@ value that transforms the current store $\sigma$ to @σ + αv@, which
-results in @StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to the
+results in a @StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to the
 top-level @ReaderT@ type, i.e., @AnsM[Unit]@.
@@
 straightforwardly. We elide most of them but describe a bit on how we handle
-functions and applications. Because lambda terms are data but also part of the
-control flow, although the way we treat them is simple now, but will be very
-different when staging and abstract interpretation is involved. The method
-@close@ takes a lambda term and an environment, and produces the
-defunctionalized representation of closures.
+functions and applications. This is because lambda terms are data but are also
+part of the control flow. Although the way we treat them is simple now, but it
+will be very different when staging and abstract interpretation is involved.
+At the moment, the method @close@ takes a lambda term and an environment, and
+produces the defunctionalized representation of closures.
 \begin{lstlisting}
@@
 To apply a function, @ap_clo@ takes a function value @rator@ and an argument
-@rand@, extracts the lambda term and environment enclosed in @rator@, allocates
-the addresses and updates the environment and store accordingly. Finally, it
-evaluates the body expression of the lambda term under the new environment and
-store.
+@rand@. It extracts the lambda term and environment enclosed in @rator@, and
+allocates the addresses and updates the environment and store accordingly.
+Finally, body expression of the lambda term is evaluated under the new
+environment and store.
 \begin{lstlisting}

### 2019-09-04 paper/50_stageabs.tex
Proofread Section 7

@@
 In this section, we review and summarize our recipe to achieve the staged
-abstract interpreter, discuss correctness issues, and different design choices.
+abstract interpreter and discuss correctness issues and different design choices.
 
@@
 \begin{itemize}
-  \item Constructing a generic interpreter that abstracts over binding-time,
+  \item Constructing a generic interpreter that abstracts over binding times,
     value domains, and primitive operations. In this paper, the generic interpreter
@@
     abstract interpreter. This step has been explored in the previous literature.
-  \item Replacing the monad stack to staged monad stack, such monads operate on
+  \item Replacing the monad stack to a staged monad stack; such monads operate on
     staged data types, i.e., next-stage values. We also need to reimplement several
@@
 two orthogonal dimensions. It is important to note that the computation
-encapsulated by the monads are not staged; only data (such as sets and maps) are
+encapsulated by the monads are not staged: only data (such as sets and maps) are
 staged. All the monadic computation, i.e., functions passed to the monadic bind
@@
 
-\paragraph{Monadic-style vs Direct-style} We use monadic interpreter through the paper,
+\paragraph{Monadic-style vs Direct-style} We use a monadic interpreter throughout the paper,
 but it is not necessary for staging. One can inline the monadic
 operations and obtain an abstract interpreter in continuation-passing
-style, or even translate back to direct-style that may use explicit
+style, or even translate back to a direct-style that may use explicit
 side-effects such as mutations. In either case, we can still apply the
 staging idea to the abstract interpreter and remove the interpretation
-overhead. Although, monads make the staged abstract interpreter can be
+overhead. However, monads allow the staged abstract interpreter to be
 implemented in a modular and extensible way.

### 2019-09-04 paper/10_intro.tex
revision

@@
     soundness properties after showing the four artifacts.
-  \item We conduct three case studies (Section~\ref{cases_study}) and
+  \item We summarize our approach and discuss correctness issues (Section~\ref{discussion}).
+    After, we conduct three case studies (Section~\ref{cases_study}) and
     an empirical evaluation on performance improvements (Section~\ref{evaluation}).

### 2019-09-04 paper/40_absinterp.tex
revision

@@
 
-After seeing both unstaged and staged concrete interpreters, now we turn our
+After seeing the unstaged and staged concrete interpreter, now we turn our
 focus on abstract interpreters under the same framework. We first present a
-lattice framework with binding-time abstraction, as well as simple abstract
-domains such as power set lattice.
+lattice representation with binding-time abstraction, as well as simple
+abstract domains such as power set lattice.
 Then we show the abstract components, namely abstract values, abstract
@@
 simplicity, it is also context/path/flow-insensitive by our choice. In Section
-\ref{cfa}, we will instantiate a commonly-used context-sensitive analysis;
+\ref{cfa}, we will instantiate it to a commonly-used context-sensitive analysis;
 \citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
@@
 
-\iffalse
-For example, the CEK machine \cite{DBLP:conf/popl/FelleisenF87} for
-concrete execution can be readily refactored to an effective $0$-CFA
-control-flow analysis \cite{Shivers:1988:CFA:53990.54007,
-  Midtgaard:2012:CAF:2187671.2187672}: first, tweak the environment
-dereference as a nondeterministic choice, such that the environment
-may contain multiple possible values for a variable, then allocate
-continuations in the environment to exploit return-flow information,
-and also constrain the address space to be finite.
-\fi
-
 \subsection{Stage Polymorphic Lattices} \label{stagedpoly_lat}
 
-Abstract interpreters run on abstract domains, such abstract domains usually can
-be presented by a lattice.
+Abstract interpreters operate on abstract domains, such abstract domains
+usually can be formulated by complete lattices.
 A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
@@
 the rest of the paper is the power set abstract domain (shown in Figure
-\ref{fig:powerset}), where the ordering is subset relation, the join is union
+\ref{fig:powerset}): the ordering is subset relation, the join is the union
 operation, and the meet is the intersect operation. The bottom element of a
 power set is empty set, and we don't have a top representation for power set.
-Other lattices used in the the paper, such as products and maps, can be
+Other lattices used in the paper, such as products and maps, can be
 implemented similarly or lifted from existing lattices element-wise or
-point-wise. Non-relational abstract domains such intervals also can be
+point-wise. Non-relational abstract domains such intervals can also be
 implemented in a stage polymorphic way.
@@
 stores are automatically lifted to abstract environments and abstract stores,
-i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@ respectively.
-It is important that the abstract store now maps addresses to sets of abstract
-values, indicating that an address may point to an over-approximated set of
-values occurred at run-time.
+i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@, respectively.
 \begin{lstlisting}
@@
 
+It is important that the abstract store now maps addresses to sets of abstract
+values, indicating that an address may point to an over-approximated set of
+values occurred at run-time.  This can be justified by the approximation nature
+of abstract interpretation and nondeterminism in the abstract interpreter.
+First, when analyzing a conditional expression, we may not have enough
+information on which branch will be taken, thus a sound treatment is to
+nondeterministically explore both branches, i.e., compute the join of two path.
+Second, when applying a function, for instance, @f(a)@, the possible lambda
+terms of @f@ is computed approximately. As a result, we may retrieve a set of
+closures when dereferencing the address of @f@ and need to apply all of them.
+
 To prevent the abstract interpreter from falling into non-termination when
@@
 consisting of two caches that remember the input and output of functions.
-Here we first provide the necessary type definitions, and later we will describe
+Here, we first provide the necessary type definitions, and later we will describe
 the algorithm in detail. A @Cache@ is a mapping from configurations @Config@ to
@@
 
-Now, the store is a mapping from addresses to sets of abstract values. This can
-be justified by the approximation nature of abstract interpretation and
-nondeterminism in the abstract interpreter.
-First, when analyzing a conditional expression, we may not have enough
-information on which branch will be taken, thus a sound treatment is to
-nondeterministically explore both branches, i.e., compute the join of two path.
-Second, when applying a function, for instance, @f(a)@, the possible lambda
-terms of @f@ is computed approximately. As a result, we may retrieve a set of
-closures when dereferencing the address of @f@ and need to apply all of them.
-
 \paragraph{Monads for Abstract Interpretation} Compared with the concrete
@@
 one cache that will not be changed during one fixed-point iteration, and use a
-@StateT@ for another cache that will be constantly updated during analysis.
+@StateT@ for another cache that will be constantly updated during the analysis.
 \citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permutation of the
@@
 
-Such monad stack scheme leads to a different type of result type. In the concrete
+Such monad stack scheme leads to a different type of grounded result. In the concrete
 setting, we only have a reader monad (for environment) and a state monad (for
@@
 @Value@ and @Store@. In the abstract setting, we have a @SetT@ inside of the
-environment monad and store monad, which becomes the container type of the pair
+environment monad and store monad, which becomes the container type of pairs
 of values and stores, i.e., @Set[(Value, Store)]@. Also note that the reader and
-state monad for caches live internally of the nondeterminism monad, therefore
+state monad for caches inhibit internally of the nondeterminism monad, therefore
 the cache type of state monad becomes adjacent to the set type.

### 2019-09-04 paper/40_absinterp.tex
revision

@@
 
-\paragraph{Primitive Operations} The primitive operations are changed to adapt
+\paragraph{Primitive Operations} The primitive operations are changed according to
 the new monad stack scheme and value domains. A notable change is that the store
-update operation joins the value with the existing values, for the same address.
-As what mentioned before, we may have to explore both branches for conditionals:
+update operation joins the new value with existing values, for the same address.
+As we mentioned before, we may have to explore both branches for conditionals:
 in @br0@ method, we combine the results using @mplus@ from @MonadPlus@, which
@@
 For function applications, @ap_clo@ looks the same as the concrete interpreter.
-The difference is now the first argument @rator@ is a set of closures (remember
-that @Value@ is now an alias of @Set[AbsValue]@), so we can not directly
+The difference is that now the first argument @rator@ is a set of closures (remember
+that @Value@ is an alias of @Set[AbsValue]@), so we can not directly
 destruct it by pattern matching. Instead, we use an auxiliary function @lift_nd@
@@
 the depth-first evaluation, where the @in@ cache stores the result from last
-iteration, and @out@ cache is used for updates in current iteration. In the next
+iteration, and @out@ cache is used for updating in current iteration. In the next
 iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
@@
 
-In our monad stack, the @in@ and @out@ cache modeled by a reader effect and
-state effect, respectively. We first defined several methods to manipulate
-the two caches through the monad stack; the implementations are elided.
+In our monad stack, the @in@ and @out@ cache are modeled by a reader effect and
+state effect, respectively. We first define several methods to manipulate
+the two caches through the monad stack (the implementations are elided).
 \begin{lstlisting}
@@
 The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which indicates that there is no more fact can be gained, thus
-the iteration should end and we have reached the least fixed point.
+input @in@ cache, which indicates that there is no more fact can be discovered,
+therefore the iteration should end and we have reached the least fixed point.
 
@@
 Finally, we override the top-level @run@ method by running the monadic value
-with the empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
+with an empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
 and initial caches @cache@$_0$.

### 2019-09-04 paper/50_stageabs.tex
revision

@@
 In the previous sections, we have seen an unstaged abstract interpreter and a
-staged concrete interpreter, now we begin describing the implementation of their
-confluence -- a staged abstract interpreter.
-Unsurprisingly, the staged abstract interpreter in this section has the same
-abstract semantics with the unstaged version from Section~\ref{unstaged_abs}.
-The approach from unstaged to staged abstract interpreter is modular,
-and does not sacrifice soundness or precision. The designer of the
-analyzer therefore has no need to rewrite the analysis. We first present the
-staged lattices and staged monads, and then the staged version of primitive
-operations, especially @close@, @ap_clo@ and @fix@. At last, we discuss several
-optimizations and summarize our approach.
+staged concrete interpreter, now we begin describing the implementation of
+their confluence -- a staged abstract interpreter.  Unsurprisingly, the staged
+abstract interpreter in this section has the same abstract semantics with the
+unstaged version from Section~\ref{unstaged_abs}.  The approach that refactors
+the unstaged one to the staged abstract interpreter is modular, and does not
+sacrifice soundness or precision. The designer of the analyzer therefore has no
+need to rewrite the analysis. We first present the staged lattices and staged
+monads, and then the staged version of primitive operations, especially
+@close@, @ap_clo@ and @fix@. At last, we discuss several optimizations.
 
@@
 achieve stage polymorphism. Now we instantiate the type @R@ to @Rep@ and
-still use power sets as an example to present its staged version.
+still use power-set as an example to describe its staged version.
 \begin{lstlisting}[escapechar=!]
@@
 The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
-eventually are delegated to operations on a staged set data type, including
+are eventually delegated to operations on a staged set data type, such as
 @subsetOf@, @union@ and @intersect@. We even do not have to change the
-implementation code, but just the types -- from @Set[T]@ to @Rep[Set[T]]@, shown
-in light gray code. The LMS library provides an implicit conversion that lifts a
-current-stage set value to next-stage values. In the code generation part, these
+implementation code, except the types -- from @Set[T]@ to @Rep[Set[T]]@, as shown
+in light gray code. The LMS library provides implicit conversions that lift
+current-stage values to next-stage values. In the code generation part, these
 operations emit their corresponding next-stage code. Again, other lattice
@@
 Now we can implement the staged abstract semantics, where @R@ is instantiated as
-@Rep@, and the abstract components are reused from the unstaged version.
+@Rep@, and the types of abstract components are reused from the unstaged version.
 
@@
 but replacing \textit{all} the transformers to their staged version. The
-following code shows this change. For readability, we squeeze the three inner
+following code shows this change. We manually squeeze the three inner
 transformers (@IdM@ omitted) into a single monad @RepSetReaderStateM[A, B, C]@,
-where @A@ is the type parameter for reader effect and @B@ for the state effects.
-In our monad stack scheme, they are both instantiated with @Cache@. Now the
-result type is a pair of two staged value: @Rep[Set[(Value, Store)]]@ and
+where @A@ is the type parameter for the reader effect and @B@ for the state effect.
+In our monad stack scheme, they are both instantiated with type @Cache@. Now the
+grounded result type is a pair of two staged value: @Rep[Set[(Value, Store)]]@ and
 @Rep[Cache]@.
@@
 \paragraph{Primitive Operations} When deriving the staged concrete interpreter
-from the unstaged counterpart, we notice that recursive calls of @ev@ on the
-lambda body are shifted from @ap_clo@ to @close@. In other word, we need to
+from the unstaged counterpart, we notice that the recursive calls of @ev@ on the
+lambda body are shifted from @ap_clo@ to @close@. In other words, we need to
 eagerly specialize the interpreter and generate code every time when we see a
-$\lambda$ term, instead of lazily when we actually need to apply a function.
-With no surprise, it is similar in the staged abstract interpreter,
-additionally, we need to handle nondeterminism. So we keep our track and focus
-on discussing closure representations and function applications for the staged
-version.
+$\lambda$ term, instead of lazily when applications happen.
+With no surprise, it is similar in the staged abstract interpreter;
+additionally, we have to handle nondeterminism occurred by over-approximation
+of runtime behaviors.
+In the rest of this part, we keep our roadmap and focus on discussing closure
+representations and function applications for the staged version.
 
@@
 next stage programs running on JVM, such large generated program would also
-cause runtime GC overhead. In this section, we present several optimizations
-that largely mitigate these issues. For all that, implementing these
-optimizations do not need to change the generic interpreter.
+potentially cause runtime GC overhead. In this section, we present several
+optimizations that largely mitigate these issues. For all that, implementing
+these optimizations do not need to change the generic interpreter.
 
@@
 
-\section{Discussion}
+\section{Discussion} \label{discussion}
 

### 2019-09-04 paper/60_cases.tex
revision

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\vspace{-2em}
 \subsection{Control-flow Analysis} \label{cfa}
 
-In this section, to demonstrate the SAI approach is widely applicable, we extend
-and refactor the staged abstract interpreter we presented in Section \ref{sai}
-to various control-flow analyses (CFA). We first discuss calling-context
-sensitive analysis for staged analyzers by tweaking allocation strategy; then we
-will see how to implement store-widening with staged monads, which is a common
-optimization in CFA; finally, we integrate abstract garbage collection with
-staging. All of them are in fact orthogonal to staging.
+In this section, to demonstrate the SAI approach is widely applicable, we
+extend and refactor the staged abstract interpreter we presented in Section
+\ref{sai} to various control-flow analyses (CFA). We first discuss
+calling-context sensitive analysis for staged analyzers by tweaking the
+allocation strategy. Then we describe how to implement store-widening with
+staged monads, which is a common optimization in CFA. Finally, we integrate
+abstract garbage collection with staging. All of them are in fact orthogonal to
+staging.
 
@@
 
-The abstract addresses we introduced in Section \ref{sai} are simply identical to
-the variable names, which is a context-insensitive schema.
-Through tweaking the address allocation strategy, we can achieve various
-context-sensitive analyses \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we
-demonstrate a calling-context-sensitive analysis, i.e., $k$-CFA-like analysis
-\cite{DBLP:journals/jfp/HornM12}, in our staged abstract interpreter.
+The abstract addresses we introduced in Section \ref{sai} are simply identical
+to the variable names, which is a context-insensitive schema.  Through tweaking
+the address allocation strategy, we can achieve various context-sensitive
+analyses \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we demonstrate a
+calling-context-sensitive analysis, i.e., $k$-CFA-like analysis
+\cite{DBLP:journals/jfp/HornM12}, can be implemented our staged abstract
+interpreter.
 \begin{lstlisting}
@@
 recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
-changed to contain an identifier and the time when it get allocated, meaning that
+changed to include the identifier and the time when it get allocated, meaning that
 this address points to some values under such calling context. If $k$ is 0, we
@@
 We then integrate the timestamp into our monad stack by adding a staged
-@StateT@ monad. The result type is passingly get updated to contain a set of
+@StateT@ monad. The result type is accordingly updated to contain a set of
 tuples, where the time is added.
@@
 
-Every time when we call the @eval@ function, we refresh the timestamp by calling
-the @tick@ function, which updates the timestamp in the state monad. The @tick@
-function is implemented as prepending the current expression being evaluated @e@
-to the existing calling context, and then taking the first $k$ elements from the list.
-The type @Config@ used in caching is also changed to included the timestamp.
+Every time when we call the @eval@ function, we refresh the timestamp by
+calling the @tick@ function, which updates the timestamp in the state monad.
+The @tick@ function is implemented as prepending the current control expression
+being evaluated (@e@) to the existing calling context, and then taking the
+first $k$ elements from the list.  The type @Config@ used in caching is also
+changed to include timestamps.
 
@@
 and the @SetT@ (for nondeterminism). In this section, we show this optimization
-is orthogonal to staging and we just need to swap the same \textit{staged}
+is orthogonal to staging and we just need to similarly swap the same \textit{staged}
 version of monads. After the adjustment, we have the following monad stack:
@@
 monad to track a set of root addresses and compute the reachable addresses every
-time when we obtain a value. Here we adopt it with staged monads:
+time when we obtain a value. Here, we adopt the idea with staged monads:
 \begin{lstlisting}
@@
 
-The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We modify the @fix@
-function with caching (from Section \ref{sai}) to @fix_gc@, but the caching part
-is omitted. Several auxiliary functions are used: @ask_root@ retrieves the
-current root addresses from the monad stack; @root@ compute the reachable
-addresses from a value; and @gc@ cleans up the store according to the reachable
-addresses and returns a new store. For the generic interpreter, we also need to
-compute the root addresses for compound expressions, readers may refer to
+The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We
+update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@,
+but the caching part is omitted in the presented code. Several auxiliary
+functions are used: @ask_root@ retrieves the current root addresses from the
+monad stack; @root@ computes the reachable addresses from a value; and @gc@
+performs garbage collection according to the reachable addresses and returns a new
+store. For the generic interpreter, we also need to compute the root addresses
+for compound expressions, readers may refer to
 \cite{DBLP:journals/pacmpl/DaraisLNH17}.

### 2019-09-04 paper/60_cases.tex
revision

@@
 
-One of the challenges of modern static analysis is that program
-usually depends on large libraries programs
-\cite{toman_et_al:LIPIcs:2017:7121}. If we are able to analyze
-programs and libraries separately and reuse the result without losing
-precision, then we can reduce the overhead of repeatedly analyzing
-libraries. Some static analyzers compute a summary for a function or a
-module, which can be reused later. However, when lacking the context
-information, those analyses can be too conservative or unsound, which
-both lead to imprecision. In this section, we show that
-meta-programming can provide a trade-off between summary-based modular
-analysis and slow whole-program analysis, by utilizing an existing
-whole-program analyzer.
+One of the challenges of modern static analysis is that programs usually depend
+on large libraries and analyzing these libraries is time consuming
+\cite{toman_et_al:LIPIcs:2017:7121}.
+If we are able to analyze programs and libraries separately and reuse the
+result without losing precision, then we can reduce the overhead of repeatedly
+analyzing libraries. Some static analyzers compute a summary for a function or
+a module, which can be reused later. However, when lacking of the context
+information, those analyses can be too conservative or unsound, which both lead
+to imprecision. In this section, we show that meta-programming can provide a
+trade-off between summary-based modular analysis and slow whole-program
+analysis, by utilizing an existing whole-program analyzer.
 
-As we have already seen from Section \ref{sai}, and recall that the
-primitive operation @close@ can denote a $\lambda$ term to a next-stage Scala function.  The specializing of the staged abstract interpreter to
-$\lambda$ terms is independent to other parts of the analyzed program.
+As we have already seen from Section \ref{sai}, and recall that the primitive
+operation @close@ can denote a $\lambda$ term to a next-stage Scala function.
+The specializing of the staged abstract interpreter to a $\lambda$ term is
+independent to other parts of the analyzed program.
 \begin{lstlisting}
@@
 
-Consider there is a library that contains a list map function @map@:
+We use Scheme as an example, and consider there is a library that contains a
+list map function @map@:
 \begin{lstlisting}
@@
 
-The specialization (0CFA and without store-widening) of @close@ on
+The specialization (0CFA and without store-widening) of @close@ to
 @map@ generates the following next-stage Scala code, that is, the
@@
 We borrow the notation from traditional modular analysis
-\cite{DBLP:conf/cc/CousotC02}: such generated function is a
-reusable artifact which can be used modularly, i.e., independent to its context. 
+\cite{DBLP:conf/cc/CousotC02}: such generated function is a modularly reusable
+artifact, i.e., independent to its context. 
 The generated functions already eliminate the interpretation overhead and

### 2019-09-04 paper/paper.tex
revert to as sign; typo
typos
revision
ccs

@@
 %\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review]{acmart}\settopmatter{}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% use defaults for review submission.
+%\startPage{1}
+
+\setcopyright{rightsretained}
 \acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
 \acmYear{2019} 
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
+\acmVolume{3} 
+\acmNumber{OOPSLA} 
+\acmArticle{126} 
+\acmMonth{10} 
+\acmPrice{}\acmDOI{10.1145/3360552}
 
@@
 %% use 'none' for review submission.
-\setcopyright{none}
+%\setcopyright{none}
 %\setcopyright{acmcopyright}
@@
 <concept>
-<concept_id>10011007.10011006.10011008</concept_id>
-<concept_desc>Software and its engineering~General programming languages</concept_desc>
+<concept_id>10011007.10010940.10010992.10010998.10011000</concept_id>
+<concept_desc>Software and its engineering~Automated static analysis</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-<concept_id>10003456.10003457.10003521.10003525</concept_id>
-<concept_desc>Social and professional topics~History of programming languages</concept_desc>
+<concept_id>10011007.10011006.10011008.10011009.10011012</concept_id>
+<concept_desc>Software and its engineering~Functional languages</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011041.10010943</concept_id>
+<concept_desc>Software and its engineering~Interpreters</concept_desc>
 <concept_significance>300</concept_significance>
@@
 
-\ccsdesc[500]{Software and its engineering~General programming languages}
-\ccsdesc[300]{Social and professional topics~History of programming languages}
+\ccsdesc[500]{Software and its engineering~Automated static analysis}
+\ccsdesc[300]{Software and its engineering~Functional languages}
+\ccsdesc[300]{Software and its engineering~Interpreters}
 %% End of generated code
@@
   %% extraction tools.
+  We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
+  Guanhong Tao, Gr\`egory Essertel, and Fei Wang for their feedback on the early
+  drafts. We thank David Van Horn and David Darais for their encouragement on
+  this work.
   This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE
@@
 \end{document}
+
+%s/@\([^@]*\)@/\\lstinline\[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\\tt\]{\1}/g
+
+%s/\\lstinline\[keywordstyle=,flexiblecolumns=false,mathescape=false,basicstyle=\\tt\]{\([^}]*\)}/@\1@/g

### 2019-09-04 paper/40_absinterp.tex
space

@@
     rt <- if (out.contains(cfg)) for { // ask if out already contains the desired result
-            (v, σ) <- lift_nd[(Value, Store)](out(cfg))
-             _ <- put_store(σ)
+            (v, σ) <- lift_nd[(Value, Store)](out(cfg));  _ <- put_store(σ)
           } yield v
@@
             _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
-            v <- ev(fix(ev))(e)
-            σ <- get_store
-            _ <- update_out_cache(cfg, (v, σ))
+            v <- ev(fix(ev))(e);  σ <- get_store;  _ <- update_out_cache(cfg, (v, σ))
           } yield v
@@
 \label{fig:coind_cache}
+\vspace{-1em}
 \end{figure}

### 2019-09-04 paper/50_stageabs.tex
misc
space

@@
 
-\begin{figure}[h!]
-  \centering
-\begin{lstlisting}
-  def fix_cache(e: Expr): Ans = for {
-    ρ   <- ask_env;  σ <- get_store;  in <- ask_in_cache;  out <- get_out_cache
-    cfg <- lift_nd[Config](Set((unit(e), ρ, σ)))
-    res <- lift_nd[(Set[ValSt], Cache)](Set(
-      // a next-stage value of type Rep[(Set[ValSt], Cache)] generated by if
-      if (out.contains(cfg)) (repMapToMapOps(out).apply(cfg).toList, out)
-      else { val m: Ans = for {
-               _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
-               v <- eval(fix_cache)(e)
-               σ <- get_store
-               _ <- update_out_cache(cfg, (v, σ))
-             } yield v
-             m(ρ)(σ)(in)(out) }))
-    _ <- put_out_cache(res._2);  vs <- lift_nd(res._1);  _ <- put_store(vs._2)
-  } yield vs._1
-\end{lstlisting}
-\caption{The staged co-inductive caching algorithm.}
-\label{fig:staged_coind_cache}
-\end{figure}
-
 \paragraph{Staged Caching and Fixpoint Iteration} Our fixed-point iteration
@@
 content of @res@ back into the monad stack.
+
+\begin{figure}[h!]
+\centering
+\begin{lstlisting}
+  def fix_cache(e: Expr): Ans = for {
+    ρ   <- ask_env;  σ <- get_store;  in <- ask_in_cache;  out <- get_out_cache
+    cfg <- lift_nd[Config](Set((unit(e), ρ, σ)))
+    res <- lift_nd[(Set[ValSt], Cache)](Set(
+      // a next-stage value of type Rep[(Set[ValSt], Cache)] generated by if
+      if (out.contains(cfg)) (repMapToMapOps(out).apply(cfg).toList, out)
+      else { val m: Ans = for {
+               _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
+               v <- eval(fix_cache)(e);  σ <- get_store;  _ <- update_out_cache(cfg, (v, σ))
+             } yield v
+             m(ρ)(σ)(in)(out) }))
+    _ <- put_out_cache(res._2);  vs <- lift_nd(res._1);  _ <- put_store(vs._2)
+  } yield vs._1
+\end{lstlisting}
+\caption{The staged co-inductive caching algorithm.}
+\label{fig:staged_coind_cache}
+\vspace{-2em}
+\end{figure}
 
@@
 statically. This leaves us a chance to further specialize the data structures.
-For example, let's assume that the @Map[K, V]@ is implemented as a hash map. If
+For example, let us assume that the @Map[K, V]@ is implemented as a hash map. If
 the keys of type @K@ are all known statically, then the indices also can be

### 2019-09-04 paper/abstract.tex
space

@@
 without compromising soundness. We conduct three case studies, including a
-comparison with \citeauthor{Boucher:1996:ACN:647473.727587}'s abstract
-compilation, applications to various control-flow analyses, and a demonstration
-that can be used for modular analysis. We also empirically evaluate the effect
-of staging on the execution time. The experiment shows an order of magnitude
-speedup with staging for control-flow analyses.
+comparison with Boucher and Feeley's abstract compilation, applications to
+various control-flow analyses, and a demonstration that can be used for modular
+analysis. We also empirically evaluate the effect of staging on the execution
+time. The experiment shows an order of magnitude speedup with staging for
+control-flow analyses.
 

### 2019-09-05 paper/60_cases.tex
ccs
Case Studies

@@
 
-In Section~\ref{sai}, we use a toy language to show that staging an abstract
+In Section~\ref{sai}, we used a toy language to show that staging an abstract
 interpreter is feasible and can be systematically derived from staged concrete
-interpreters and unstaged abstract interpreter.
+interpreters and unstaged abstract interpreters.
 In this section, we conduct several case studies to further show that this
@@
 AC was introduced by \citet{Boucher:1996:ACN:647473.727587} as an
-implementation technique for abstract interpretation based static analysis.
+implementation technique for abstract interpretation-based static analysis.
 Similar to the present paper, the idea is inspired by partial evaluation:
@@
 then reproduce their result by simply adding stage annotations.
-The generated program of our approach improves approximately the same extent of
+The program generated by our approach provides approximately the same improvement with regards to
 speed, but without changing a single line of the analyzer program using
 type-based staging annotations. However, AC requires more engineering effort,
-i.e., rewriting the whole analyzer into the closure generation form. Moreover,
+i.e., rewriting the whole analyzer into the closure-generation form. Moreover,
 as shown in Section~\ref{staged_ds}, our approach is able to not only remove the
-interpretive overhead, but also enabling several optimizations such specialized
+interpretive overhead, but also enable several optimizations such as specialized
 data structures and equivalent rewritings.
@@
 The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
-for a small CPS language consisting of abstractions, applications, recursion and
+for a small CPS language consisting of abstractions, applications, recursion, and
 primitive operators. Figure \ref{compare_ac_sai} (left) shows the AC
@@
 
-On the other side, our approach does exactly the same thing through staging: the
-AST are static, and stores are dynamic, therefore we annotate the type @Store@
+On the other side, our approach works equivalently through staging: the
+ASTs are static, whereas stores are dynamic; therefore, we annotate the type @Store@
 with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
-changes of our approach using LMS framework are the types of @Store@ to
-@Rep[Store]@, since they will be known at the next stage; other parts do not need to change. 
+changes of our approach using the LMS framework are the types of @Store@ to
+@Rep[Store]@, since they will be known at the next stage: no other parts need to change.
 The generated code
-will just looks-up and update the store. The functions at current stage are
-eliminated in the residual program, therefore we do not generate higher-order
+will consist entirely of looks-up and updating the store. The functions at the current stage are
+eliminated in the residual program; therefore, we do not generate higher-order
 functions, which provides additional performance improvement. Nevertheless,
-the support from LMS framework is necessary, such as staging support for @Map@,
-we consider these efforts are reusable and complete orthogonal to the
-implementation of analysis.
+the support from the LMS framework is necessary, such as staging support for @Map@.
+We consider these efforts reusable and completely orthogonal to the
+implementation of the analysis.
 
@@
 
-In this section, to demonstrate the SAI approach is widely applicable, we
+In this section, to demonstrate that the SAI approach is widely applicable, we
 extend and refactor the staged abstract interpreter we presented in Section
@@
 
-The timestamp is essentially a finite list of expressions that tracks the $k$
+The timestamp is essentially a finite list of expressions that tracks the $k$-most
 recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
-changed to include the identifier and the time when it get allocated, meaning that
+changed to include an identifier and the time when it gets allocated, meaning that
 this address points to some values under such calling context. If $k$ is 0, we
 obtain a monovariant analysis as demonstrated before; if $k > 0$, we obtain a
-family of analysis with increasing precision. Note that when $k = 0$, the
+family of analyses with increasing precision. Note that when $k = 0$, the
 address space is statically determined by the set of variables appeared in the
@@
 
-Every time when we call the @eval@ function, we refresh the timestamp by
+Every time we call the @eval@ function, we refresh the timestamp by
 calling the @tick@ function, which updates the timestamp in the state monad.
@@
 
-A commonly used technique to improve the running time in control-flow analysis
+A commonly-used technique to improve the running time in control-flow analysis
 is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
@@
 \cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
-shows that, for big-step monadic abstract interpreter, we may add a @Reader@
+showed that, for big-step monadic abstract interpreters, we may add a @Reader@
 monad to track a set of root addresses and compute the reachable addresses every
@@
 
-\vspace{-1em}
 \begin{figure}[h!]
@@
 
-\vspace{-2em}
 \subsection{Modular Staged Analysis for Free} \label{modular}
@@
 One of the challenges of modern static analysis is that programs usually depend
-on large libraries and analyzing these libraries is time consuming
+on large libraries, and analyzing these libraries is time-consuming
 \cite{toman_et_al:LIPIcs:2017:7121}.
 If we are able to analyze programs and libraries separately and reuse the
-result without losing precision, then we can reduce the overhead of repeatedly
+results without losing precision, we can reduce the overhead of repeatedly
 analyzing libraries. Some static analyzers compute a summary for a function or
@@
 trade-off between summary-based modular analysis and slow whole-program
-analysis, by utilizing an existing whole-program analyzer.
+analysis by utilizing an existing whole-program analyzer.
 
-As we have already seen from Section \ref{sai}, and recall that the primitive
+As we have already seen from Section \ref{sai}, the primitive
 operation @close@ can denote a $\lambda$ term to a next-stage Scala function.
@@
 We borrow the notation from traditional modular analysis
-\cite{DBLP:conf/cc/CousotC02}: such generated function is a modularly reusable
+\cite{DBLP:conf/cc/CousotC02}: such a generated function is a modularly reusable
 artifact, i.e., independent to its context.
 The generated functions already eliminate the interpretation overhead and
-monadic overhead, therefore can be consider as partial summaries. Although, we
+monadic overhead, and can therefore be considered as partial summaries. However, we
 still need to run the generated function at next stage to obtain the analysis
-result.  At next stage, we can provide different abstract arguments, store and
+result.  At the next stage, we can provide different abstract arguments, stores, and
 caches, depending on the context where it is called.
@@
 parameterized to the context information. In our approach, the
-transition from a whole-program analyzer to modular staged analyzer is
-mechanized and for free via staging. We do not need to design any
+transition from a whole-program analyzer to a modular-staged analyzer is
+mechanized and for-free via staging. We do not need to design any
 intermediate representation or data structures of summaries.  In some
-scenario, traditional summary-based analyses are able to carry more
+scenarios, traditional summary-based analyses are able to carry more
 information within a module, e.g., the summary of their internal
-sub-modules. The staged approach presented here may still need to run
-through the fixed-point for sub-modules.
+submodules. The staged approach presented here may still need to run
+through the fixed-point for submodules.
 

### 2019-09-05 paper/80_related.tex
Case Studies

@@
 one. Another large body of static analyses are implemented by
-generating constraints firstly and then use a solver to discharge the
+first generating constraints and then using a solver to discharge the
 proof obligations or find the fixed-point. Notable examples include

### 2019-09-05 paper/70_eval.tex
Proofreading eval

@@
 approach, we scale our toy interpreter to a large subset of the Scheme language
-and evaluate the analyzing time against with the unstaged versions. The result
-shows that the staged versions are averagely 10+ times as fast as the unstaged
+and evaluate the analysis time against the unstaged versions. The result
+shows that the staged versions are, on average, 10+ times as fast as the unstaged
 version, as evidence of our \textit{abstraction without regret} approach.
@@
 A front-end desugars Scheme to a small core language that makes the analyzer
-easy to implement. We implement two common monovariant 0-CFA-like analysis, one
-is equipped with store-widening (0CFA-SW for short), the other does not (0CFA
-for short). Usually, the store-widened analysis is expected to be faster than the one
-without.  Our prototype implementation currently generates Scala code; the
-generated Scala code will be compiled and executed on JVM. It is possible to
+easy to implement. We implement two common monovariant 0-CFA-like analyses: one
+is equipped with store-widening (0CFA-SW for short), the other is not (0CFA
+for short). It is traditionally expected that the store-widened analysis be faster than the
+alternative.  Our prototype implementation currently generates Scala code; the
+generated Scala code will be compiled and executed on the JVM, though it is possible to
 generate C/C++ code in the future. In the experiments, we implement several
@@
 We use Scala 2.12.8 and Oracle JVM 1.8.0-191 \footnote{The options for
-running JVM is set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
+running the JVM are set to the following: \texttt{-Xms2G -Xmx8G -Xss1024M
 -XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
@@
 evaluations were performed on a machine with 4 Intel Xeon Platinum
-8168 CPU at 2.7GHz and 3 TiB of RAM. Although the machine has 96 cores
+8168 CPU at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores
 (192 threads) in total, the abstract interpreters are single-threaded.
@@
 all the experiments are executed for 20 times and we report the
-statistical median value of the running times. We set a 5 minutes
+statistical median value of the running times. We set a 5 minute
 timeout.
@@
 to challenge $k$-CFA; while some other benchmarks are from real-world
-applications, for examples, the RSA public key encryption algorithm
+applications, for example, the RSA public key encryption algorithm
 @rsa@. In Figure \ref{evaluation_result}, we report the number of
@@
 %booleans.
-For 0CFA w/o store-widening, we observe an average speed-up of ~17x times; for
-0CFA w/ store-widening, the average speed-up is ~7x times.
-\texttt{kcfa-worst-64} has the most speed-up among all the tests.
-Although it has a comparably large size of AST, it has a quite simple structure
-which only involves booleans, lambda terms and applications. We conjecture that
-the simple domain and structure allow the generated code to be efficiently
-optimized by the JVM. But on the corresponding 0CFA-SW, we do not observe such
-great speed-up. Due to the complexity of generated code and next-stage running
-platform (JVM), we leave the detailed empirical analysis of which static
-analysis can benefit more from the specialization as future work.
+For 0CFA without store-widening, we observe an average speedup of ~17x times;
+for 0CFA with store-widening, the average speedup is ~7x times.
+\texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
+has a comparably large size of AST, it has a rather simple structure which only
+involves booleans, lambda terms, and applications. We conjecture that the simple
+domain and structure allow the generated code to be efficiently optimized by the
+JVM. However, on the corresponding 0CFA-SW, we do not observe such a significant
+speedup. Due to the complexity of generated code and next-stage running platform
+(JVM), we leave the detailed empirical analysis of which static analysis can
+benefit more from the specialization as future work.
 

### 2019-09-05 paper/80_related.tex
Finish proofreading

@@
 small-step abstract interpreter in state-transition style. The program
-is first compiled to an "abstract bytecode" IR. The IR consists of higher-order
+is first compiled to an ``abstract bytecode'' IR. The IR consists of higher-order
 functions, which will be executed later on an abstract abstract machine for
@@
 stage conducts the flow-sensitive analysis by refining the first
-one. Another large body of static analyses are implemented by
+one. Another large body of static analyses is implemented by
 first generating constraints and then using a solver to discharge the
@@
 \citet{10.1007/11561347_18} discussed a monad with let-insertion and
-memoization for code generation, using Gaussian Elimination in MetaOCaml as
+memoization for code generation, using Gaussian Elimination in MetaOCaml as an
 example. In this work, let-insertion is handled by the ANF transformation in
@@
 monadic interpreters was explored by \citet{danvy1991compiling,
-DBLP:conf/dsl/SheardBP99}. But they are different from the performance motivation
+DBLP:conf/dsl/SheardBP99}. However, they are different from the performance motivation
 here: the generated code is in monadic form.

### 2019-09-05 paper/90_conc.tex
typos
Finish proofreading

@@
 
-In this paper, we present a simple and elegant framework that extends the first
-Futamura projection to abstract interpreters.
+We present a simple and elegant framework that extends the first Futamura
+projection to abstract interpreters.
 To realize it, we design a generic interpreter that abstracts over both
-binding-times and value domains; by instantiating it with staged monads for
+binding times and value domains; by instantiating it with staged monads for
 abstract interpretation, the staged abstract interpreter can be readily derived.

### 2019-09-05 paper/60_cases.tex
margin

@@
 \begin{lstlisting}
-  type AnsM[T] = RepReaderT[RepStateT[
-                   RepStateT[RepSetReaderStateM[Cache, Cache, ?], Time, ?], //add Time with a RepStateT
-                 Store, ?], Env, T]
+  type AnsM[T] = RepReaderT[RepStateT[RepStateT[ // add Time with a RepStateT
+                  RepSetReaderStateM[Cache, Cache, ?], Time, ?], Store, ?], Env, T]
   type Result = (Rep[Set[(Value, Store, Time)]], Rep[Cache])

### 2019-09-05 paper/paper.tex
misc

@@
   We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
-  Guanhong Tao, Gr\`egory Essertel, and Fei Wang for their feedback on the early
+  Guanhong Tao, Gr\'egory Essertel, and Fei Wang for their feedback on the early
   drafts. We thank David Van Horn and David Darais for their encouragement on

### 2019-09-05 paper/10_intro.tex
space
intro

@@
 Abstract interpretation as a lattice-based approach to sound static analyses was
-proposed by \citet{DBLP:conf/popl/CousotC77}. Based on the notion of Galois
-connections \cite{CousotCousot79-1}, the analyzer can soundly approximate the
+proposed by \citet{DBLP:conf/popl/CousotC77, CousotCousot79-1}. Based on the notion of Galois
+connections, the analyzer can soundly approximate the
 concrete program runtime behavior by computing fixed points on an abstract
@@
 
-Based on the idea of abstracting abstract machines, more
-pragmatically, several implementation strategies utilizing purely
-functional programming techniques to build abstract interpreters have
-emerged. Such techniques include monads, monad transformers
-\cite{DBLP:journals/pacmpl/DaraisLNH17,
+%Based on the idea of abstracting abstract machines, more
+%pragmatically, several implementation strategies utilizing purely
+%functional programming techniques to build abstract interpreters have
+%emerged.
+
+Based on the idea of AAM, more pragmatically, several implementation strategies
+in high-level functional programming languages have emerged.  Such techniques
+include monads and monad transformers \cite{DBLP:journals/pacmpl/DaraisLNH17,
 Sergey:2013:MAI:2491956.2491979}, arrows
@@
 \cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. The pure
-approaches provide certain benefits, including that the abstract interpretation
-artifacts can be built compositionally and modularly, e.g., by using monad
-transformers. Therefore, the soundness of analysis can be proven with less
+approaches provide certain benefits, for examples, the abstract interpretation
+artifacts can be built in a compositional and modular fashion, and referential
+transparency allow programmers to confidently reason about the correctness of
+their implementations.  The soundness of analysis can also be proven with less
 effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
-paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.  Also, referential
-transparency and equational reasoning allow programmers to confidently reason
-about the correctness of their implementations.
+paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.
 
@@
 
+\input{fig_confluence.tex}
+
 \paragraph{Futamura Projection of Abstract Interpreters}
@@
 Given a program $e_0$ of type @Expr@, by applying the specialization, we can
-obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ :
-@Input -> Value@, which is the so-called \textit{equivalent executable}.
+obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ : @Input -> Value@,
+which is the so-called \textit{equivalent executable}.
 
@@
 monadic operations are inlined and compiled down to low-level Scala code.
-
-\input{fig_confluence.tex}
 

### 2019-09-05 paper/10_intro.tex
intro

@@
 approaches. First, similar to concrete interpreters, an abstract interpreter
-analyzes the program by traversing the abstract syntax tree (AST). The traversal incurs
+analyzes the program by traversing the AST. The traversal incurs
 interpretive overhead, such as the pattern matching on the ASTs and recursive
@@
 The challenge here: can we apply the same idea to a general abstract interpreter
-to make it run faster? To achieve this, we need to specialize a nonstandard
-semantics that extensively uses nondeterminism, operates on abstract domains,
-and computes fixed-points iteratively over a cache. The solution is not simply obvious.
+to make it run faster? The solution is not obvious, as we need to specialize
+a nonstandard semantics that extensively uses nondeterminism, operates on
+abstract domains, and computes fixed-points iteratively over a cache.
 
@@
 instance, @eval : Expr -> Input -> Value@ is an interpreter for some language.
-Given a program $e_0$ of type @Expr@, by applying the specialization, we can
+Given a program $e_0$@ : Expr@, by applying the specialization, we can
 obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ : @Input -> Value@,
@@
 a two-level syntax \cite{Nielson:1992:TFL:130665}, by manually
-annotating the binding-times of variables in the program. Then, the MSP system
+annotating the binding-times of variables. Then, the MSP system
 will check whether these annotations are consistent, and specialize the program
@@
 Lightweight Modular Staging framework, and implements the Futamura projection
-of a big-step abstract interpreter for a small, stateful, higher-order
+of a abstract interpreter for a stateful higher-order
 language. By deriving staged monads that can be used to do code generation, we
@@
 returns a value of monadic type, which can be varied by different semantics.
-The domain of the interpreter and the effects such as read, state, and
-nondeterminism can all be wrapped into this monadic type.  The binding-time
-abstraction is represented by a higher-kinded type, and controls whether the
-interpreter produces values directly or generates code. It is worth mentioning
-that the binding-time type is also injected into the monadic type, so that we
-will distinguish normal monads and staged monads.
+The value domains of the interpreter and the effects such as state and nondeterminism
+can all be wrapped into this monadic type.  The binding-time abstraction is
+represented by a higher-kinded type, and injected into the monadic type.  The
+binding-times will control whether the interpreter produces values directly or
+generates code.
+
+%It is worth mentioning
+%that the binding-time type is also injected into the monadic type, so that we
+%will distinguish normal monads and staged monads.
 
@@
 4) We empirically evaluate the performance improvements gained by staging,
-showing an order of magnitude speedup on flow analyses.
+showing an order of magnitude speedup.
 

### 2019-09-06 paper/10_intro.tex
revision

@@
 connections, the analyzer can soundly approximate the
-concrete program runtime behavior by computing fixed points on an abstract
+concrete program behavior at runtime by computing fixed points on an abstract
 domain. Despite the tremendous theoretical development of abstract
@@
 This systematic abstraction approach can be tailored to different
-language features (such as state, first-class control, exception, and
+language features (such as state, first-class control, exceptions, and
 concurrency) and sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
@@
 \cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. The pure
-approaches provide certain benefits, for examples, the abstract interpretation
+\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. These pure
+approaches provide certain benefits, for example, the abstract interpretation
 artifacts can be built in a compositional and modular fashion, and referential
-transparency allow programmers to confidently reason about the correctness of
-their implementations.  The soundness of analysis can also be proven with less
+transparency allows programmers to confidently reason about the correctness of
+their implementations.  The soundness of an analysis can also be proven with less
 effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
@@
 approaches. First, similar to concrete interpreters, an abstract interpreter
-analyzes the program by traversing the AST. The traversal incurs
-interpretive overhead, such as the pattern matching on the ASTs and recursive
-calls on subexpressions. Such overhead can be negligible if the abstract
-interpreter only runs on the program a single time, but can accumulate
-significantly if it runs repeatedly (for example, on libraries).  Second, it is
-common that abstract interpreters written in pure languages extensively use
-functional encodings of side-effects to implement the semantics of abstract interpretation.
-For example, the list monad can be used to model nondeterminism from collecting
-interpreters.
+analyzes a program by traversing its AST. This traversal incurs
+interpretive overhead, such as pattern matching on the AST nodes and recursive
+calls on subexpressions. If it is a single-time traversal, such overhead can be
+negligible. But usually, abstract interpreters iteratively analyze and traverse
+the AST multiple times to reach fixed-points, therefore the accumulated
+overhead is even magnified. 
+%Repeatedly analyzing library programs also poses significant overhead.
+%Such overhead can be negligible if the abstract
+%interpreter only runs on the program a single time, but can accumulate
+%significantly if it runs repeatedly (for example, on libraries).  
+Second, it is common that abstract interpreters written in pure languages
+extensively use functional encodings of side-effects to implement the semantics
+of abstract interpretation.  For example, the list monad can be used to model
+nondeterminism from collecting interpreters.
 %For example, a collecting interpreter that returns all possible runtime values is
@@
 
-Roughly at the same time when abstract interpretation was proposed in
-1970s, \citet{futamura1971partial} revealed a close
+Roughly at the same time when abstract interpretation was proposed in the
+1970s, \citet{futamura1971partial} observed a close
 connection between interpreters and compilers through a hierarchy of
-specializations, namely, Futamura Projections. The first Futamura
-projection shows that specializing an interpreter to the input program
+specializations, since then known as Futamura Projections. The first Futamura
+projection describes that specializing an interpreter to the input program
 removes interpretative overhead and yields an equivalent executable
@@
 other similar performance problems, examples including
-query compilers\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
-\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275}, and etc.
+query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
+\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
 The challenge here: can we apply the same idea to a general abstract interpreter
@@
 In short, our approach borrows ideas and techniques from multi-stage
-programming and embedded DSLs to abstract interpreters:  1) To remove the
-overhead from interpretation and effect layers, we specialize the abstract
-interpreter via staging and then generate efficient low-level code that does
-the actual analysis. 2) Inspired by the tagless-final interpreters
+programming and embedded DSLs, and apply them to abstract interpreters:  1) To
+remove the overhead from interpretation and effect layers, we specialize the
+abstract interpreter via staging and then generate efficient low-level code
+that does the actual analysis. 2) Inspired by the tagless-final interpreters
 \cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
 abstracts over binding-times and different semantics, which allows the staged
-abstract interpreter to be derived from its unstaged counterpart.
-As a result, the derived staged abstract interpreters has no intrusive changes
-to the underlying abstract semantics, thereby preserving the soundness.
-In this sense, our approach allows for no regret for both performance and
-engineering effort. We elaborate these two main ideas in detail, as follows.
+abstract interpreter to be derived from its unstaged counterpart.  As a result,
+the derived staged abstract interpreters has no intrusive changes to the
+underlying abstract semantics, thereby preserving soundness.  In this
+sense, our approach allows for no regret for both performance and engineering
+effort. We elaborate these two main ideas in detail, as follows.
 
@@
 interpreter w.r.t. the input program yields an equivalent executable. For
-instance, @eval : Expr -> Input -> Value@ is an interpreter for some language.
+instance, assume that @eval : Expr --> Input --> Value@ is an interpreter for some language.
 Given a program $e_0$@ : Expr@, by applying the specialization, we can
-obtain a specialized interpreter $\texttt{eval}_{\texttt{e0}}$ : @Input -> Value@,
-which is the so-called \textit{equivalent executable}.
-
+obtain a specialized program $\texttt{eval}_{\texttt{e0}}$ : @Input --> Value@.
+%which is the so-called \textit{equivalent executable}.
 % By definition of the interpreter, they produce the same result when applied with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
-
 The first proposed approach to realize Futamura projections is partial
@@
 \cite{NIELSON1989117}, monads provide a good abstraction to hide the detail of
-abstract interpretation semantics. However, the monads do not have a clear
+abstract interpretation semantics. However, monads do not have a clear
 stage distinction.
@@
 Lightweight Modular Staging framework, and implements the Futamura projection
-of a abstract interpreter for a stateful higher-order
-language. By deriving staged monads that can be used to do code generation, we
-obtain a staged abstract interpreter that can be used for specialization. After
+of an abstract interpreter for a stateful higher-order
+language. By deriving staged monads that can be used to perform code generation, we
+obtain a staged abstract interpreter that can be used for specialization. Through
 staging, the generated code is specialized to the input program, and all the
@@
 represented by a higher-kinded type, and injected into the monadic type.  The
-binding-times will control whether the interpreter produces values directly or
+binding-times control whether the interpreter produces values directly or
 generates code.
@@
 We evaluate the idea of staging an abstract interpreter through
-case studies and an empirical evaluation on performance.
+case studies and an empirical performance evaluation.
 1) We compare our approach with abstract compilation
@@
     soundness properties after showing the four artifacts.
-  \item We summarize our approach and discuss correctness issues (Section~\ref{discussion}).
-    After, we conduct three case studies (Section~\ref{cases_study}) and
+  \item We summarize our approach and discuss correctness (Section~\ref{discussion}).
+    We conduct three case studies (Section~\ref{cases_study}) and
     an empirical evaluation on performance improvements (Section~\ref{evaluation}).

### 2019-09-06 paper/20_prelim.tex
revision

@@
 
-A monad is a type constructor @M[_]: * -> *@ with two operations, often called
+A monad is a type constructor @M[_]: * --> *@ with two operations, often called
 @return@ and @bind@. Informally, @return@ wraps a value into the monad @M@, and
@@
 write our interpreters.
-Monad transformers are type constructors of kind @(* -> *) -> (* -> *)@, which
+Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
 take a monad as argument and produces another monad. By using monad

### 2019-09-06 paper/40_absinterp.tex
revision

@@
 
-The co-inductive caching algorithm is implemented as a instrumentation over
+The co-inductive caching algorithm is implemented as an instrumentation over
 @eval@ (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
@@
 @out@ cache with the new result, i.e., a pair of value and store @(v, σ)@.
-Similar to store update, cache update is also a point-wise join.
+Similar to store updates, the cache update is also a point-wise join.
 

### 2019-09-06 paper/80_related.tex
revision

@@
 \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis. We borrow the notation of modular analysis to our approach.
+modular analysis, where we borrow the notation of modular analysis.
 However, the staged abstract interpreter does not generate complete summaries for

### 2019-09-06 paper/20_prelim.tex
typo

@@
 (e.g., MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}
-and Template Haskell \cite{Sheard:2002:TMH:636517.636528} to construct
+and Template Haskell \cite{Sheard:2002:TMH:636517.636528}) to construct
 such staged abstract interpreters.
@@
 We consider a call-by-value $\lambda$-calculus in direct-style, extended
-with numbers, arithmetic, recursions, and conditionals. Other effectful features
+with numbers, arithmetic, recursions, and conditionals. Other effects
 such as assignments can also be supported readily.
-Since we are mostly interested in analyzing the dynamic behaviors of the
-program, we will disguise the static semantics. We also assume that
+Since we are mostly interested in analyzing dynamic behaviors of the
+program, we disguise the static semantics. We also assume that
 input programs are well-typed and all variables are distinct. The abstract
@@
 
-The abstract syntax we present in fact can be seen as a deep embedding of the
+The abstract syntax we present can be seen as a deep embedding of the
 language -- we use data-types to represent programs. This design choice allows us
-to easily use different interpretations over the AST; with the inheritance and
+to use different interpretations over the AST; with the inheritance and
 overriding mechanism in Scala, we may also add new language constructs and reuse
@@
 A monad is a type constructor @M[_]: * --> *@ with two operations, often called
-@return@ and @bind@. Informally, @return@ wraps a value into the monad @M@, and
-@bind@ unwraps the monadic value and transforms it into a new monadic value.
-Pragmatically in Scala, we define a monad type class using trait @Monad@ (Figure
+@return@ and @bind@. Informally, @return@ wraps a value into a monad @M@, and
+@bind@ unwraps a monadic value and transforms it into a new monadic value.
+Pragmatically in Scala, we define the monad type class using trait @Monad@ (Figure
 \ref{fig:monad}), where it declares the @pure@ \footnote{We elect to use
@@
 \texttt{unit} is a built-in function in LMS.} and @flatMap@ operation. The trait
-itself takes the monad type @M[_]@ as argument, which is a higher-kinded type
+itself takes the monad type @M[_]@ as an argument, which is a higher-kinded type
 that takes a type and returns a type. The method @pure@ promotes values of type
@@
 
-The Scala compiler will translate the above @for@-comprehension expression into
+The Scala compiler translates the above @for@-comprehension expression into
 an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
-in the @for@-comprehension is translated into a @map@, where the expression of
+in @for@-comprehensions is translated into a @map@, where the expression of
 @yield@ becomes the body expression inside of that @map@ operation. The foregoing
@@
 
-Note that here the monadic object @List[_]@ encapsulates the data internally.
-Therefore it only exposes the simplified version of @flatMap@, where the monad
-@M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
+Note that here, the monadic object @List[_]@ encapsulates the data internally.
+Therefore it only exposes the simplified version of @flatMap@, where the monadic
+value of @M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
 \ref{fig:monadops}) defines the simplified version of monadic operations that
 are necessary for @for@-comprehension. The conversion between @Monad@ and
-@MonadOps@ can be done by using implicit design pattern.
+@MonadOps@ can be done by using the implicit design pattern.
 In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
@@
 Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
-take a monad as argument and produces another monad. By using monad
+take a monad as an argument and produces another monad. By using monad
 transformers, we can combine multiple monads into a single one.  The
@@
 
-In this section, we present the generic interface in the style of big-step
+In this section, we present the generic interface in the style of a big-step
 definitional interpreter. The key idea is to keep the binding-time type and
@@
 \paragraph{Binding-time Abstraction}
-The binding-time is declared as a higher-kinded type @R[_]@. 
-Now we are able to use @R@ to annotate other data-types in the interpreter;
-it is also injected into the monadic type @MonadOps@.  If we simply instantiate
-@R@ an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
+The binding-time type is declared as a higher-kinded type @R[_]@. 
+Now we can use @R@ to annotate other data-types in the interpreter;
+it is also injected into the monadic type @MonadOps@.  If we instantiate
+@R@ with an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
 standard definitional interpreter that will execute the program.  In Section
-\ref{stagedinterp}, we will instantiate @R@ using LMS's built-in next-stage
+\ref{stagedinterp}, we instantiate @R@ using LMS's built-in next-stage
 type annotation @Rep@, which makes the interpreter act as a compiler.
@@
 implement @map@ and @flatMap@. It is worth noting that @MonadOps@ takes another
-type parameter @R[_]@ as binding-time, i.e., @MonadOps[R, AnsM, T]@.
-Inside of @MonadOps@, @R[_]@ wraps the data types @A@ and @B@ that are
-encapsulated by the monad, but not the monad type @M@ itself. When acting as
-compilers, we will also replace the monads to the ones that work on staged data
-values.
+type parameter @R[_]@ as the binding-time, i.e., @MonadOps[R, AnsM, T]@.
+Inside of @MonadOps@, @R[_]@ annotates the data types @A@ and @B@ that are
+encapsulated by the monad, but not the monad type @M@ itself. When the generic
+interpreter acts as a compiler, we will replace the monads to the ones that
+work on staged data values.
 
@@
 stores. These methods return monadic values of type @AnsM[_]@, which
-may parameterized over the environment or store type, or simply a @Unit@
-value for effects. For examples, @local_env@ installs a new environment
+may be parameterized over the environment or store type, or merely a @Unit@
+value for performing effects. For examples, @local_env@ installs a new environment
 @ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
@@
 
-\paragraph{Primitive Operations} Next we define a few primitive operations.
+\paragraph{Primitive Operations} Next, we define a few primitive operations.
 First, two versions of @alloc@ are declared. The first one takes a store and an
@@
 
-Other primitive operations in the interpreter handles language constructs.  The
+Other primitive operations in the interpreter handle the language constructs.  The
 method @num@ and @close@ deal with primitive values, which lift literal terms
@@
 Note that the interpreter is written in open-recursive style -- it cannot
-refer to itself directly, instead, @eval@ takes an additional parameter @ev@ of
+refer to itself directly. Instead, @eval@ takes an additional parameter @ev@ of
 type @Expr => Ans@ that refers to itself. Consequently, the method @close@ that
@@
 abstract interpretation.
-Finally, a top-level wrapper @run@ is declared; the return type @Result@
-depends on what kind of monads is being used, therefore this is also an
+Finally, a top-level wrapper @run@ is declared. The return type @Result@
+depends on what kind of monads is being used; therefore this is also an
 abstract type.
@@
 address allocations, we use type @Int@ and always return a number greater
-than the size of current store.
+than the size of the current store.
 A value can be either a tagged number \texttt{IntV}, or a closure
@@
 \paragraph{Unstaged Monads}
-For the purpose of concrete interpretation, the monad needs to model read
-effect and state effect, which correspond to environment and store,
-respectively. Following the monad transformer approach
-\cite{DBLP:conf/popl/LiangHJ95}, a @ReaderT@ and @StateT@ monad
-transformer are used to compose the monad stack. In other words, the type @AnsM@ is
-instantiated by layering @ReaderT@ and @StateT@ transformers\footnote{The
-question mark syntax is a kind projector \cite{kindprojector}, thus
-\texttt{StateT[IdM,Store,?]} is equivalent to \newline \texttt{(\{type
-M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is parameterized by type
-@Env@, and @StateT@ is parameterized by type @Store@, and the inner-most monad @IdM@
-is simply an identity monad.
+For concrete interpretation, the monad needs to model reader and state
+effects, which correspond to the environment and the store, respectively. Following
+the monad transformer approach \cite{DBLP:conf/popl/LiangHJ95}, a @ReaderT@ and
+@StateT@ monad transformer are used to compose the monad stack. In other words,
+the type @AnsM@ is instantiated by layering @ReaderT@ and @StateT@
+transformers\footnote{The question mark syntax is a kind projector
+\cite{kindprojector}, thus \texttt{StateT[IdM,Store,?]} is equivalent to
+\newline \texttt{(\{type M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@
+is parameterized by the type @Env@, and @StateT@ is parameterized by the type @Store@,
+and the inner-most monad @IdM@ is merely an identity monad.
 \begin{lstlisting}
@@
 where the new state (of type @S@) is accompanied with the value (of type @A@).
-Note that for the moment, the binding-time type @R@ is an identity type, thus
+Note that for the moment, the binding-time type @R@ is an identity type; thus,
 these monads operate on unstaged data. We can also see this from the
@@
 straightforwardly. We elide most of them but describe a bit on how we handle
-functions and applications. This is because lambda terms are data but are also
-part of the control flow. Although the way we treat them is simple now, but it
+functions and applications. The reason is that lambda terms are data but are also
+part of the control flow. Although the way we treat them is simple now, it
 will be very different when staging and abstract interpretation is involved.
@@
 
-To apply a function, @ap_clo@ takes a function value @rator@ and an argument
+When applying a function, @ap_clo@ takes a function value @rator@ and an argument
 @rand@. It extracts the lambda term and environment enclosed in @rator@, and
@@
 
-The top-level @run@ method invokes the fixed-pointer operator directly,
-given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$.
+Given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$,
+the top-level @run@ method invokes the fixed-point operator directly,
 \begin{lstlisting}

### 2019-09-06 paper/30_stageconc.tex
typo

@@
 In this section, based on the generic interpreter and concrete components we
-presented in Section \ref{prelim}, we show that how to stage the concrete
+presented in Section \ref{prelim}, we show how to stage the concrete
 interpreter by changing the monad type and refactoring primitive
@@
 framework in Scala, and then replay the same steps from the unstaged
-counterpart. Finally, we briefly describe code generation.
+counterpart. Finally, we briefly describe the code generation.
 
@@
 LMS provides a type constructor @Rep[T]@ where @T@ can be
-an arbitrary type, indicating a value of type @T@ will be known at future stage.
+an arbitrary type, indicating a value of type @T@ will be known at the future stage.
 Operations acting on @Rep[T]@ expressions will be residualized as generated
@@
 A classic example for introducing multi-stage programming is the power function
-that computes $b^x$. Usually it is implemented as a recursive function:
+that computes $b^x$. Usually, it is implemented as a recursive function:
 \begin{lstlisting}
@@
 the value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
-first adding the @Rep@ type annotation to variables known at next stage. In this case,
+first adding the @Rep@ type annotation to variables known at the next stage. In this case,
 @b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
@@
 The idea behind the framework is to construct a sea-of-node intermediate
-representations (IR) for the next-stage program at the current stage
+representation (IR) for the next-stage program at the current stage
 \cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
-expressions of type @Rep[_]@ to their IR is done by using implicit design
+expressions of type @Rep[_]@ to their IR is done by using the implicit design
 pattern. As we will see later, implementing staging and code generation
@@
 To implement the staged concrete interpreter, we replay the steps from the
-instantiation of unstaged concrete interpreter in Section \ref{unstaged_conc}.
-But, now we use the @Rep@ type to annotate the value domains, environments and
+instantiation of the unstaged concrete interpreter in Section \ref{unstaged_conc}.
+However, now we use the @Rep@ type to annotate the value domains, environments and
 stores, and redefine the staged version of monads and primitive operations.
@@
 We use the same structure of monad stack in the unstaged interpreter: a reader
-monad with a state monad. The difference here is, now the monads operate on staged values. 
-For brevity, we call them \textit{staged monads}. In the code, we also use a @Rep@
-prefix on the constructors and types to differentiate them. But it is important
-to note that instances of the case class @ReaderT@/@StateT@ are not staged, also the
-monadic computation like @R => M[A]@ are also not staged. Instead, the internal
-data that these monads operate on are staged, i.e., @R@ and @A@ in the
-reader monad, @S@ and @A@ in the state monad. The following code
-snippet shows the idea. We use \hl{light gray} to highlight where @Rep@ type is added:
+monad with a state monad. The difference here is, now the monads operate on
+staged values.  For brevity, we call them \textit{staged monads}. In the code,
+we also use a @Rep@ prefix on the constructors and types to differentiate them.
+But it is important to note that instances of the case class @ReaderT@/@StateT@
+are not staged; also the monadic computation, such as functions of @R => M[A]@,
+are not staged. Instead, the internal data that these monads operate on are
+staged, i.e., @R@ and @A@ in the reader monad, @S@ and @A@ in the state monad.
+The following code snippet shows the idea. We use \hl{light gray} to highlight
+where @Rep@ type is added:
 \begin{lstlisting}[escapechar=!]
@@
 
-The function @f@ passed to @flatMap@ is also a function value known at current stage.
+The function @f@ passed to @flatMap@ is also a function value known at the current stage.
 Therefore all invocations of @flatMap@s can be reduced before generating code.
 The fact that we only stage data but not monadic computation or monadic
-values, is the reason that we can peel of the monad stack in the generated code.
-Now type @AnsM[_]@ is instantiated as the same structure as before, but using the
+values is the reason that we can peel of the monad stack in the generated code.
+Now the type @AnsM[_]@ is instantiated as the same structure as before, but using the
 @Rep@ versions:
@@
 but as we will see, it is not straightforward for nondeterminism monad
-(@SetT@) in abstract interpreters. Because not only the elements in the set are
+(@SetT@) in the abstract interpreter. Because not only the elements in the set are
 staged, but the whole set is also staged, and we have no knowledge about how
 many elements in the set. In this section, we explicitly distinguish the two
-versions of monadic interfaces, later we will instantiated the staged set monad
-by manually fusing part of the monad stack.
+versions of monadic interfaces; later we will instantiate the staged set monad
+by manually fusing a fragment of the monad stack.
 

### 2019-09-06 paper/20_prelim.tex
typo

@@
 such as assignments can also be supported readily.
-Since we are mostly interested in analyzing dynamic behaviors of the
-program, we disguise the static semantics. We also assume that
+Since we are mostly interested in analyzing dynamic behavior of
+programs, we elide the static semantics. We also assume that
 input programs are well-typed and all variables are distinct. The abstract
-syntax is shown as follows:
+syntax is as follows:
 \begin{lstlisting}
@@
 to use different interpretations over the AST; with the inheritance and
-overriding mechanism in Scala, we may also add new language constructs and reuse
+overriding mechanisms in Scala, we may also add new language constructs and reuse
 existing interpretations.
@@
 Pragmatically in Scala, we define the monad type class using trait @Monad@ (Figure
-\ref{fig:monad}), where it declares the @pure@ \footnote{We elect to use
+\ref{fig:monad}), which it declares the @pure@ \footnote{We elect to use
 \texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
 \texttt{unit} is a built-in function in LMS.} and @flatMap@ operation. The trait
-itself takes the monad type @M[_]@ as an argument, which is a higher-kinded type
-that takes a type and returns a type. The method @pure@ promotes values of type
-@A@ to values of type @M[A]@. The monadic @bind@ operation is usually called
-@flatMap@ in Scala, which takes a monad-encapsulated value of type @M[A]@, a
-function of @A => M[B]@ and returns values of type @M[B]@.
+itself takes the monad type constructor @M[_]@ as an argument, which, as a
+higher-kinded type, takes a type and returns a type. The method @pure@
+promotes values of type @A@ to values of type @M[A]@. The monadic @bind@
+operation is usually called @flatMap@ in Scala. It takes a
+monad-encapsulated value of type @M[A]@, a function of @A => M[B]@ and returns
+values of type @M[B]@.
 
@@
 for monadic operations through @for@-comprehension.  For example, an object of
-@List[A]@ is a @List@ monad, where @A@ is the element type.
+@List[A]@ is an instance of the @List@ monad, with element type @A@.
 To compute the Cartesian product of two lists of numbers, we can use Scala's
@@
 an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
-in @for@-comprehensions is translated into a @map@, where the expression of
-@yield@ becomes the body expression inside of that @map@ operation. The foregoing
-bindings in the comprehension are all translated into calls of @flatMap@:
+in @for@-comprehensions is translated into a @map@, where the argument of
+@yield@ becomes the body expression inside of that @map@ operation. The 
+bindings before the last one are all translated into calls of @flatMap@:
 \begin{lstlisting}
@@
 are necessary for @for@-comprehension. The conversion between @Monad@ and
-@MonadOps@ can be done by using the implicit design pattern.
+@MonadOps@ can be achieved by using the implicit design pattern.
 In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
@@
 transformers, we can combine multiple monads into a single one.  The
-implementation of monads and monad transformers essentially borrows the
-ground-truth from Haskell.
+implementation of monads and monad transformers follows the one from Haskell.
 
@@
 required components for the interpreter are environments @Env@ and stores
-@Store@. @Env@ maps from identifiers to addresses, and @Store@ maps from
+@Store@. @Env@ maps identifiers to addresses, and @Store@ maps 
 addresses to values, respectively. The @Env@ captures bound variables in the
@@
 program run-time.  At this moment, the domains of addresses and values are
-still abstract, i.e., they are declared as abstract types.
+still abstract, and hence, they are declared as abstract types.
 \begin{lstlisting}
@@
 it is also injected into the monadic type @MonadOps@.  If we instantiate
-@R@ with an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
+@R@ as an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
 standard definitional interpreter that will execute the program.  In Section
@@
 
-\paragraph{Monadic Operations} We define the return type of the interpreter as
-@Ans@, which is a monadic type @AnsM[_]@ wrapping the type @Value@. 
+\paragraph{Monadic Operations} We define the return type of the interpreter
+@Ans@, as a monadic type @AnsM[_]@ wrapping the type @Value@. 
 As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
@@
 encapsulated by the monad, but not the monad type @M@ itself. When the generic
-interpreter acts as a compiler, we will replace the monads to the ones that
+interpreter acts as a compiler, we will replace the monads with the ones that
 work on staged data values.
@@
 may be parameterized over the environment or store type, or merely a @Unit@
-value for performing effects. For examples, @local_env@ installs a new environment
+value for performing effects. For example, @local_env@ installs a new environment
 @ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
@@
 \end{figure}
-%\vspace{-1em}
+\vspace{-1em}
 
 \paragraph{Primitive Operations} Next, we define a few primitive operations.
-First, two versions of @alloc@ are declared. The first one takes a store and an
+First, we delcare two versions of @alloc@. The first one takes a store and an
 identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
@@
 consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
-of the addresses, therefore can be used in monadic computations.
+of the addresses, and therefore can be used in monadic computations.
 An auxiliary method @get@ retrieves the value of an identifier @x@, through the
@@
 Other primitive operations in the interpreter handle the language constructs.  The
-method @num@ and @close@ deal with primitive values, which lift literal terms
+methods @num@ and @close@ deal with primitive values, which lift literal terms
 (e.g., lambdas) to our value representation (e.g., closures).  
@@
   def arith(op: Symbol, v1: R[Value], v2: R[Value]): R[Value]
-  def ap_clo(ev: Expr => Ans)(rator: R[Value], rand: R[Value]): Ans
+  def ap_clo(ev: Expr => Ans)(fun: R[Value], arg: R[Value]): Ans
 \end{lstlisting}
@@
 \end{figure}
-\vspace{-1em}
 
@@
 takes an extra @ev@, because evaluations may happen inside of them.
-To close the open-recursion, we use a fixed-point operator @fix@.
+To close the open recursion, we use a fixed-point operator @fix@.
 For concrete-interpretation instantiation, @fix@ works like the Y combinator;
@@
 abstract interpretation.
-Finally, a top-level wrapper @run@ is declared. The return type @Result@
-depends on what kind of monads is being used; therefore this is also an
-abstract type.
+Finally, there is a top-level wrapper @run@. The return type @Result@
+depends on the kind of monads being used; and therefore is also left abstract.
 \begin{lstlisting}
@@
 
-As the first step in our roadmap, we instantiate the interpreter concretely in
-this section. The result of such instantiation is a standard definitional
-interpreter with environments and stores. It can also be obtained by the
-refunctionalization of CESK machines \cite{Felleisen:1987:CAH:41625.41654,
+As the first step in our roadmap, we instantiate the generic interpreter for
+concrete execution in this section. The result is a standard definitional
+interpreter with environments and stores. It can also be obtained by
+refunctionalizing a standard CESK machine \cite{Felleisen:1987:CAH:41625.41654,
 DBLP:conf/ppdp/AgerBDM03}. We first present the concrete components, i.e., the
@@
 The two types we need to concretize are addresses @Addr@ and values @Value@. The
-type @Env@ and @Store@ is derived automatically. To assure the freshness of
+types @Env@ and @Store@ are derived automatically. To ensure the freshness of
 address allocations, we use type @Int@ and always return a number greater
@@
 effects, which correspond to the environment and the store, respectively. Following
-the monad transformer approach \cite{DBLP:conf/popl/LiangHJ95}, a @ReaderT@ and
-@StateT@ monad transformer are used to compose the monad stack. In other words,
+the monad transformer approach \cite{DBLP:conf/popl/LiangHJ95}, @ReaderT@ and
+@StateT@ monad transformers are used to compose the monad stack. In other words,
 the type @AnsM@ is instantiated by layering @ReaderT@ and @StateT@
@@
 functions and applications. The reason is that lambda terms are data but are also
-part of the control flow. Although the way we treat them is simple now, it
-will be very different when staging and abstract interpretation is involved.
+part of the control flow. The way we treat them is simple now, but
+will become more involved when staging the abstract interpreter.
 At the moment, the method @close@ takes a lambda term and an environment, and
@@
 
-When applying a function, @ap_clo@ takes a function value @rator@ and an argument
-@rand@. It extracts the lambda term and environment enclosed in @rator@, and
+When applying a function, @ap_clo@ takes a function value @fun@ and an argument
+@arg@. It extracts the lambda term and environment enclosed in @fun@, and
 allocates the addresses and updates the environment and store accordingly.
@@
 \begin{lstlisting}
-  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = rator match {
+  def ap_clo(ev: Expr => Ans)(fun: Value, arg: Value): Ans = fun match {
     case CloV(Lam(x, e), ρ: Env) => for {
       α <- alloc(x)
-      _ <- set_store(α → rand)
+      _ <- set_store(α → arg)
       rt <- local_env(ev(e))(ρ + (x → α))

### 2019-09-06 paper/30_stageconc.tex
typo

@@
 @b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
-below code recipe.
+below code snippet.
 The way we use this staged @power@ function is to create a @DslDriver@ and override the
-@snippet@ method that gives the currently know value @x@ (5 in this example).
+@snippet@ method that supplies 5 as the currently know value for @x@.
 \begin{lstlisting}
@@
 @Int@ and @Double@, and commonly-used data structures such as lists and maps.
-The idea behind the framework is to construct a sea-of-node intermediate
-representation (IR) for the next-stage program at the current stage
-\cite{DBLP:conf/birthday/Rompf16}. For convenience, the conversion from
-expressions of type @Rep[_]@ to their IR is done by using the implicit design
-pattern. As we will see later, implementing staging and code generation
-support for user-defined classes is also straightforward.
+The framework internally constructs a sea-of-nodes intermediate representation (IR)
+for the next-stage program when executing the current stage \cite{DBLP:conf/birthday/Rompf16}.
+For convenience, the conversion from expressions of type @Rep[_]@ to their IR
+is done by using the implicit design pattern. As we will see later,
+implementing staging and code generation support for user-defined classes is
+also straightforward.
 
@@
 \paragraph{Staged Monads}
-We use the same structure of monad stack in the unstaged interpreter: a reader
-monad with a state monad. The difference here is, now the monads operate on
+We use the same monad stack structure as in the unstaged interpreter: a reader
+monad with a state monad. The difference here is that, now the monads operate on
 staged values.  For brevity, we call them \textit{staged monads}. In the code,
@@
 But it is important to note that instances of the case class @ReaderT@/@StateT@
-are not staged; also the monadic computation, such as functions of @R => M[A]@,
-are not staged. Instead, the internal data that these monads operate on are
+are not staged, and that the monadic computation, such as functions of @R => M[A]@,
+are also not staged. Instead, the internal data that these monads operate on are
 staged, i.e., @R@ and @A@ in the reader monad, @S@ and @A@ in the state monad.
 The following code snippet shows the idea. We use \hl{light gray} to highlight
-where @Rep@ type is added:
+where @Rep@ types are added:
 \begin{lstlisting}[escapechar=!]
@@
 The function @f@ passed to @flatMap@ is also a function value known at the current stage.
-Therefore all invocations of @flatMap@s can be reduced before generating code.
+Therefore all invocations of @flatMap@ can be reduced before generating code.
 The fact that we only stage data but not monadic computation or monadic
@@
 be achieved without modifying the implementation of monads. This is true so far,
-but as we will see, it is not straightforward for nondeterminism monad
-(@SetT@) in the abstract interpreter. Because not only the elements in the set are
+but as we will see, it is not as straightforward for the nondeterminism monad
+(@SetT@) in the abstract interpreter, because not only the elements in the set are
 staged, but the whole set is also staged, and we have no knowledge about how
 many elements in the set. In this section, we explicitly distinguish the two
-versions of monadic interfaces; later we will instantiate the staged set monad
+versions of monadic interfaces. Later, we will instantiate the staged set monad
 by manually fusing a fragment of the monad stack.
@@
 compile the lambda term with its environment to a next-stage Scala function. To
-do this, we need to recursively call the interpreter function at current stage,
+do this, we need to recursively call the interpreter function at the current stage,
 i.e., the @ev@ argument passed to @close@.  The following code implements the

### 2019-09-06 paper/30_stageconc.tex
revision

@@
 @emit_complied_clo@ to denote @f@ to a next-stage Scala function, represented
-by the type @Rep[Value]@. Inside of @f@, we can access the evaluator via @ev@.
+by the type @Rep[Value]@. Inside of function @f@, we can access the evaluator via @ev@.
 However, @ev@ produces a monadic value of type @AnsM@, which can only exist at
-the current stage. To connect the current-stage monadic values and future-stage
-grounded values, the evaluator @ev@ is invoked on the body expression @e@ to be
-specialized w.r.t. @e@, and then we \textit{collapse} the monadic value of
-@AnsM@ to grounded values by providing the prepared environment and store.
+the current stage. To connect current-stage monadic values and future-stage
+grounded values, the evaluator @ev@ is supplied with not only the body
+expression @e@ to be specialized, but also the prepared environment and store.
+By doing so, we \textit{collapse} the monadic value of @AnsM@ to grounded
+values within staging; therefore in the future stage, there is no monads.
 
 For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
-and @arg@. But what we can do on the values of @Rep@ type is limited. In fact, @fun@ does
-not have an intensional application operation we can use directly. We only know
-that it is a staged value produced by @close@, and represents a
-next-stage Scala function. Therefore we can just generate a next-stage function
-application, also in Scala. Since @ap_clo@ produces a monadic value of type
-@Ans@, we can still use the monadic operation @get_store@ to obtain the latest
-store object. Then @emit_ap_clo@ is used to generate a current-stage
-representation of the future-stage application result, which is a pair of values
-and stores @ValSt@. After getting that result @vs@, we use @put_store@ to
+and @arg@. But what we can do with these values of type @Rep@ is limited. 
+In fact, @fun@ does not have an intensional application operation we can use
+directly. We only know that it is a staged value produced by @close@, and
+it represents a next-stage Scala function. Therefore, we can only generate a
+next-stage function application, also in Scala. @ap_clo@ produces a monadic
+value of type @Ans@, so that we can still use the monadic operation @get_store@
+to obtain the latest store. Then @emit_ap_clo@ takes the function and argument
+values, and the latest store to generate a current-stage representation of the
+future-stage application result, which is of type @Rep[ValSt]@, a pair of
+values and stores.  After obtaining the result @vs@, we use @put_store@ to
 update the store and reify the result into the current-stage monadic value.
@@
 
-Compared with the unstaged version, we can observe a shift of using the
-evaluation function @ev@ from application-time to value-representation-time: in
-the unstaged interpreter, invocations of @ev@ happens in @ap_clo@; while in the
-staged interpreter, we eagerly apply @ev@ when denoting lambda terms to our
-next-stage value domain.
+Compared with the unstaged version, we can observe a shift of where we use the
+evaluation function @ev@: in the unstaged interpreter, the invocation of @ev@
+happens in @ap_clo@, i.e., at the application time; while in the staged
+interpreter, we eagerly apply @ev@ when denoting the lambda terms to our
+next-stage value domains, i.e., at the value-representation time.
 
@@
 
-In the previous section, we merely show the types of @emit_compiled_clo@ and
+In the previous section, we showed the types of @emit_compiled_clo@ and
 @emit_ap_clo@ without concrete implementations. In this section, we briefly
-discuss the IR node generated by them and sketch code generation.
+discuss the IR nodes generated by them and sketch the code generation.
 We define the IR nodes @IRCompiledClo@ and @IRApClo@ using @case class@
-extending from @Def[T]@. @Def[T]@ is a built-in type from the LMS framework,
+extending from @Def[T]@. @Def[T]@ is a built-in type in the LMS framework,
 representing next-stage value definitions of type @T@.
@@
 
-The IR nodes are manipulated by the LMS passes and finally generated to
+The IR nodes are manipulated by the LMS passes and finally generated as
 next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
-programmers to control what is generated for each kind of nodes. Here we match
-@IRCompiledClo@ and @IRApClo@ and generate definitions via @emitValDef@ for
-them.  For compiled closures, we put the function @f@ and its accompanying
-syntactic term and environment to its next-stage value representation
-@CompiledClo@. For applications, we have the knowledge indicating that @fun@ is
-a @CompiledClo@ at the stage, and it has a callable field @f@. Hence, we
-generate a next-stage function application @f(arg, σ)@. The invocations of
-@quote@s are used to generate code for the other next-stage values.
+programmers to control over what code is generated for each kind of nodes. Here,
+we match @IRCompiledClo@ and @IRApClo@, and generate definitions via
+@emitValDef@ for them.  For compiled closures, we put the function @f@ and its
+accompanying syntactic term and environment to their next-stage value
+representation @CompiledClo@. For applications, we know the fact that @fun@ is
+an instance of @CompiledClo@ at the next stage, and it has a callable field
+@f@. Hence, we generate a next-stage function application @f(arg, σ)@.  The
+invocations of @quote@ are used to generate proper names for the other
+next-stage values.
 \begin{lstlisting}

### 2019-09-06 paper/40_absinterp.tex
revision

--- 2019-09-04 paper/40_absinterp.tex
@@
 After seeing the unstaged and staged concrete interpreter, now we turn our
-focus on abstract interpreters under the same framework. We first present a
+focus to abstract interpreters under the same framework. We first present a
 lattice representation with binding-time abstraction, as well as simple
-abstract domains such as power set lattice.
-Then we show the abstract components, namely abstract values, abstract
-environments, and abstract stores.
-The abstract interpreter we construct in this
-section is similar to \citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For
-simplicity, it is also context/path/flow-insensitive by our choice. In Section
-\ref{cfa}, we will instantiate it to a commonly-used context-sensitive analysis;
-\citet{Darais:2015:GTM:2814270.2814308} shows how to achieve path- and
-flow-sensitivity by varying the monads, which is also applicable to our unstaged
-or staged abstract interpreter.
+abstract domains, such as power set lattice.  Then we show the abstract
+components, namely abstract values, abstract environments, and abstract stores.
+The abstract interpreter we construct in this section is similar to
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s. For simplicity, it is
+context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
+instantiate it to a commonly-used context-sensitive analysis;
+\citet{Darais:2015:GTM:2814270.2814308} show how to achieve path- and
+flow-sensitivity by varying the monads, which is also applicable to our
+unstaged or staged abstract interpreter.
 
@@
 
-Abstract interpreters operate on abstract domains, such abstract domains
-usually can be formulated by complete lattices.
-A complete lattice over set $S$ is a tuple of $\langle S, \sqsubseteq, \top,
+The abstract domains operated by abstract interpreters usually can be
+formulated by complete lattices.
+A complete lattice over set $S$ is a tuple $\langle S, \sqsubseteq, \top,
 \bot, \sqcup, \sqcap \rangle$, where $\top$ is the top element, $\bot$ is the
@@
 \ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
-annotate the binding-times, where @R[_]@ wraps datatype @S@ and @Boolean@ in
-these operations.
+annotate the binding-times, where @R[_]@ wraps the data type @S@ and @Boolean@
+in these operations.
 
@@
 
-In this section, we simply let the binding-time type @R[T] = T@ (i.e., @NoRep@)
-to instantiate an unstaged lattice type class. An example of lattice we used in
-the rest of the paper is the power set abstract domain (shown in Figure
-\ref{fig:powerset}): the ordering is subset relation, the join is the union
-operation, and the meet is the intersect operation. The bottom element of a
-power set is empty set, and we don't have a top representation for power set.
-Other lattices used in the paper, such as products and maps, can be
-implemented similarly or lifted from existing lattices element-wise or
-point-wise. Non-relational abstract domains such intervals can also be
-implemented in a stage polymorphic way.
+In this section, we simply instantiate @SPLattice@ with the flat binding-time type
+@R[T] = T@ (i.e., @NoRep@).
+An example of lattices we used in the rest of the paper is the power set
+abstract domain (shown in Figure \ref{fig:powerset}): the ordering is the subset
+relation, the join is the union operation, and the meet is the intersect
+operation. The bottom element of a power set is empty set, and we do not have a
+representation of top element for power set.  Other lattices used in the paper, such as
+products and maps, can be implemented similarly or by lifting the existing
+lattices element-wise or point-wise. Non-relational numerical abstract domains
+such intervals can also be implemented in a stage polymorphic way.
 
@@
 
-We follow the \textit{abstracting definitional interpreters} (ADI) approach
-\cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete interpreter to
-an abstract interpreter.
-% \cite{DBLP:conf/icfp/HornM10, DBLP:journals/jfp/HornM12}
+In this section, we follow the \textit{abstracting definitional interpreters}
+(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete
+interpreter to an abstract interpreter.
 
@@
 
-We first widen the concrete values to @AbsValue@: numbers are lifted to a
-singleton object @IntTop@ representing the set of all numbers; closures remain
-the same. Then the type @Value@ is a set of @AbsValue@ (we abuse the type names
-a bit, as the generic interface already defined the abstract type @Value@).
-The address space is constrained to be finite to ensure the computability of
-analysis -- for the simplest monovariant analysis, we use variable names for
-addresses. After defining the @Value@ and @Addr@ type, the environments and
-stores are automatically lifted to abstract environments and abstract stores,
-i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@, respectively.
+We first widen the concrete values to @AbsValue@. There are two variants of
+@AbsValue@: numbers are lifted to a singleton object @IntTop@, which represents
+the set of all numbers; closures remain the same. Then the type @Value@ is
+defined as a set of @AbsValue@.  The address space is constrained to be finite
+to ensure the computability of analysis -- for a simple monovariant analysis,
+we directly use variable names for addresses. After defining the @Value@ and
+@Addr@ type, the types of environments and stores are automatically lifted to
+abstract environments and abstract stores, i.e., @Map[Ident, Addr]@ and
+@Map[Addr, Set[AbsValue]]@, respectively.  
 \begin{lstlisting}
@@
 
-It is important that the abstract store now maps addresses to sets of abstract
+It is important that the abstract stores now map addresses to sets of abstract
 values, indicating that an address may point to an over-approximated set of
-values occurred at run-time.  This can be justified by the approximation nature
-of abstract interpretation and nondeterminism in the abstract interpreter.
-First, when analyzing a conditional expression, we may not have enough
-information on which branch will be taken, thus a sound treatment is to
-nondeterministically explore both branches, i.e., compute the join of two path.
-Second, when applying a function, for instance, @f(a)@, the possible lambda
-terms of @f@ is computed approximately. As a result, we may retrieve a set of
-closures when dereferencing the address of @f@ and need to apply all of them.
+runtime values.  This can be justified by the approximation nature of abstract
+interpretation and nondeterminism in the abstract interpreter.  First, when
+analyzing a conditional expression, we may not have enough information to
+decide which branch will be taken, thus a sound treatment is to explore the both
+branches. Also, at some later point, we need to compute the join of two path.
+Another nondeterminism comes from function applications, for instance, @f(a)@,
+the possible callees ($\lambda$ terms) of @f@ is computed approximately.  As a
+result, we may retrieve a set of closures when dereferencing the address of @f@
+and we need to apply all of them.
 
-To prevent the abstract interpreter from falling into non-termination when
-computing the fixed-point, the ADI approach uses a co-inductive mechanism
-consisting of two caches that remember the input and output of functions.
-Here, we first provide the necessary type definitions, and later we will describe
-the algorithm in detail. A @Cache@ is a mapping from configurations @Config@ to
-sets of value-store pairs, where the configuration is a triple of current
+To ensure the abstract interpreter terminates on all programs when computing
+the fixed-point, the ADI approach uses a co-inductive mechanism consisting of
+two caches that remember the input and output of the @eval@ function.
+Here, we first provide the necessary type definitions, and describe
+the algorithm later. A @Cache@ is a mapping from configurations @Config@ to
+sets of value-store pairs, where the configuration is a triple of the current
 expression being evaluated, environment and store. Intuitively, a cache memoizes
-the corresponding values and stores for a given program configuration.
+the result values and stores for a given program configuration.
 \begin{lstlisting}
@@
 
-\paragraph{Monads for Abstract Interpretation} Compared with the concrete
-interpreter that uses reader and state effects, the abstract interpreter
-further introduces a nondeterminism effect and another reader and state effect
-for the two caches. The nondeterminism effect is represented by the
-@Set[M[_], A]@ monad transformer, where @M@ is the inner monad being
-transformed, and @A@ is the type of elements in the set. We use a @ReaderT@ for
-one cache that will not be changed during one fixed-point iteration, and use a
-@StateT@ for another cache that will be constantly updated during the analysis.
-\citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permutation of the
-monad stack for abstract interpretation. The following @AnsM@ type shows the
-monad stack (we use \hl{light gray} to highlight what have been changed from
-the monad stack of concrete interpretation):
+\paragraph{Monads for Abstract Interpretation}
+Compared with the concrete interpreter that uses reader and state effects, the
+abstract interpreter further introduces the nondeterminism effect and another
+reader and state effect for the two caches. The nondeterminism effect is
+represented by the @Set[M[_], A]@ monad transformer, where @M@ is the inner
+monad type being transformed, and @A@ is the type of elements in the set. We use a
+@ReaderT@ for one cache that is not changed during one fixed-point
+iteration, and use a @StateT@ for another cache that will be constantly updated
+during the analysis.  \citet{DBLP:journals/pacmpl/DaraisLNH17} discuss
+different permutations of the monad stack for abstract interpretation. The
+following @AnsM@ type shows the monad stack for abstract interpretation (we use
+\hl{light gray} to highlight what have been changed from the monad stack of
+concrete interpretation):
 \begin{lstlisting}[escapechar=!]
@@
 
-Similar to the concrete scenario, we sketch the implementation of @flatMap@ for
-@SetT@ transformer and omit the rest. The value @run@ encapsulated by the monad
+Similar to the concrete scenario, we sketch the @flatMap@ implementation of the
+@SetT@ transformer and omit the rest. The field @run@ encapsulated by the monad
 is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and @A@ is the element
 type of the set. The @flatMap@ method takes a function that transforms each
-element of type @A@ from @M[Set[A]]@ into a @SetT[M, B]@ value, and then folds
+element of type @A@ into a @SetT[M, B]@ value, and then folds
 all the transformed values into a single monadic value of type @Set[M, B]@.
@@
       SetT(Monad[M].flatMap(run) { (s: Set[A]) =>
-        s.foldLeft(SetT.empty[M,B])((acc, a) => acc ++ f(a)).run
+        s.foldLeft(SetT.empty[M, B])((acc, a) => acc ++ f(a)).run
       }); ... }
@@
 Such monad stack scheme leads to a different type of grounded result. In the concrete
-setting, we only have a reader monad (for environment) and a state monad (for
-store), and the reader effect is not persistent to the final result. Together
+setting, we only have a reader monad (for environments) and a state monad (for
+stores). The reader effect is not persistent to the final result. Together
 with the value produced by the interpreter, the final result type is a pair of
-@Value@ and @Store@. In the abstract setting, we have a @SetT@ inside of the
-environment monad and store monad, which becomes the container type of pairs
-of values and stores, i.e., @Set[(Value, Store)]@. Also note that the reader and
-state monad for caches inhibit internally of the nondeterminism monad, therefore
-the cache type of state monad becomes adjacent to the set type.
+@Value@ and @Store@. However, under abstract interpretation, we have a @SetT@
+inside of the environment monad and store monad. The @Set@ type becomes the
+container type of the pairs of values and stores, i.e., @Set[(Value, Store)]@.
+Also, note that the reader and state monad for the caches inhibits internally
+of the nondeterminism monad; therefore, the final result type is paired
+with a cache type from the state monad.
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations} The primitive operations are changed according to
-the new monad stack scheme and value domains. A notable change is that the store
-update operation joins the new value with existing values, for the same address.
-As we mentioned before, we may have to explore both branches for conditionals:
-in @br0@ method, we combine the results using @mplus@ from @MonadPlus@, which
-requires the value domains are join-semilattices (in our case @Set@s and
-@Map@s).
+the new monad stack scheme and value domains. One of the notable changes is
+that the store update operator merges the new values with existing values.
+As we mentioned before, sometimes we have to explore the both branches for
+conditionals: in method @br0@, we combine the results using the @mplus@
+operation from @MonadPlus@, which requires the value domains are
+join-semilattices (in our case, @Set@ and @Map@).
 \begin{lstlisting}
@@
 The value representation of lambda terms is still a defunctionalized closure
-@CloV@, but we lift it to a singleton set domain.
+@CloV@, but we lift it to a singleton set.
 \begin{lstlisting}
@@
 For function applications, @ap_clo@ looks the same as the concrete interpreter.
-The difference is that now the first argument @rator@ is a set of closures (remember
-that @Value@ is an alias of @Set[AbsValue]@), so we can not directly
-destruct it by pattern matching. Instead, we use an auxiliary function @lift_nd@
-that takes a set and lifts the elements into the monad stack. Then we can still
-implement @ap_clo@ in monadic style, where the closures come from the @SetT@ and
-thus the nondeterminism can be naturally handled. The light gray code shows what
-is added from the concrete @ap_clo@.
+The difference is that now the first argument @funs@ is a set of closures (remember
+that type @Value@ is an alias of @Set[AbsValue]@), so we can not directly
+destruct it to syntactic @Lam@ by pattern matching. Instead, we use an
+auxiliary function @lift_nd@ that takes a set and lifts the elements in the set
+into the monad stack. Then we can straightforwardly implement the @ap_clo@ in
+monadic style, where the closures come from the @SetT@ and thus the
+nondeterminism can be naturally handled. The light gray code shows what is
+added from the concrete
+@ap_clo@.
 \begin{lstlisting}[escapechar=!]
   def lift_nd[T](vs: Set[T]): AnsM[T]
-  def ap_clo(ev: Expr => Ans)(rator: Value, rand: Value): Ans = for {
-    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](rator)}!
+  def ap_clo(ev: Expr => Ans)(funs: Value, rand: Value): Ans = for {
+    !\hl{CloV(Lam(x, e), $\rho$) <- lift\_nd[AbsValue](funs)}!
     α <- alloc(x)
-    _ <- set_store(α → rator)
+    _ <- set_store(α → funs)
     rt <- local_env(ev(e))(ρ + (x → α))
@@
 As we mentioned earlier, the ADI approach uses a two-cache mechanism to
-compute the least fixed-point and prevent non-termination, which is also called a
-\textit{co-inductive} caching algorithm or \textit{truncated depth-first
-evaluation} \cite{Rosendahl:AbsIntPL}. It has been used in other abstract
-interpreter or fixed point computation
+compute the least fixed-point and prevent non-termination.
+The caching algorithm is also called \textit{co-inductive} caching or
+\textit{truncated depth-first evaluation} \cite{Rosendahl:AbsIntPL}. It has
+been used in other abstract interpreter or fixed point computation
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
   Rosendahl:AbsIntPL}. The idea is to use a @in@ cache and a @out@ cache during
-the depth-first evaluation, where the @in@ cache stores the result from last
-iteration, and @out@ cache is used for updating in current iteration. In the next
+the depth-first evaluation. The @in@ cache stores the result from the last
+iteration, and @out@ cache is used for updating in the current iteration. In the next
 iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
@@
 
-In our monad stack, the @in@ and @out@ cache are modeled by a reader effect and
-state effect, respectively. We first define several methods to manipulate
+In our monad stack, the @in@ and @out@ caches are modeled by reader effects and
+state effects, respectively. We first define several methods to manipulate
 the two caches through the monad stack (the implementations are elided).
@@
 
-The co-inductive caching algorithm is implemented as a instrumentation over
-@eval@ (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
-Initially, the two caches are both empty.
+The co-inductive caching algorithm is implemented as an instrumentation over the
+@eval@ function (Figure \ref{fig:coind_cache}), and it also closes the open recursion.
+The instrumentation works as follows. Initially, the two caches are both empty.
 During the iteration, we first check whether @out@ contains the configuration
 @cfg@, representing the current desired computation. If it does, then the result
-is directly returned through the monads.
-Otherwise, we first retrieve the results from @in@ (@⊥@ if not contained in
-@in@), and join it into the @out@ cache.
-Then we invoke @ev@ to compute the result for this iteration, where @ev@ takes
+is directly returned through the monad stack.
+Otherwise, we first retrieve the result from @in@ (@⊥@ if @in@ not contains
+@cfg@), and update the result into the @out@ cache in the fashion of join.
+Then, we invoke @ev@ to compute the result for this iteration, where @ev@ takes
 @fix(ev)@ as a self reference.
-After the evaluation, the store @σ@ may have been changed. We update the
-@out@ cache with the new result, i.e., a pair of value and store @(v, σ)@.
-Similar to store update, cache update is also a point-wise join.
+After the evaluation, the store @σ@ may have been changed. Therefore, 
+we get the latest store and construct the result @(v, $\sigma$)@; the
+result will be used to update the out cache.
 
 The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which indicates that there is no more fact can be discovered,
-therefore the iteration should end and we have reached the least fixed point.
+input @in@ cache, which indicates that there is no more fact can be discovered.
+Therefore, the iteration should end, and we have reached the least fixed point.
 

### 2019-09-06 paper/50_stageabs.tex
revision
typo

@@
 In the previous sections, we have seen an unstaged abstract interpreter and a
-staged concrete interpreter, now we begin describing the implementation of
+staged concrete interpreter. Now we begin describing the implementation of
 their confluence -- a staged abstract interpreter.  Unsurprisingly, the staged
-abstract interpreter in this section has the same abstract semantics with the
-unstaged version from Section~\ref{unstaged_abs}.  The approach that refactors
+abstract interpreter in this section uses the same abstract semantics from the
+unstaged version in Section~\ref{unstaged_abs}.  The approach we refactor
 the unstaged one to the staged abstract interpreter is modular, and does not
@@
 need to rewrite the analysis. We first present the staged lattices and staged
-monads, and then the staged version of primitive operations, especially
+monads, and then discuss the staged version of primitive operations, especially
 @close@, @ap_clo@ and @fix@. At last, we discuss several optimizations.
@@
 In Section \ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
-achieve stage polymorphism. Now we instantiate the type @R@ to @Rep@ and
-still use power-set as an example to describe its staged version.
+achieve the stage polymorphism. Now we instantiate the type @R@ as @Rep@ and
+still use power-set as example to describe its staged version.
 \begin{lstlisting}[escapechar=!]
@@
 
-The trait @RepLattice@ is obtained by specializing @Lattice@ with @Rep@.
+The trait @RepLattice@ is shown by specializing @Lattice@ with type @Rep@.
 The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
-are eventually delegated to operations on a staged set data type, such as
-@subsetOf@, @union@ and @intersect@. We even do not have to change the
-implementation code, except the types -- from @Set[T]@ to @Rep[Set[T]]@, as shown
-in light gray code. The LMS library provides implicit conversions that lift
-current-stage values to next-stage values. In the code generation part, these
-operations emit their corresponding next-stage code. Again, other lattice
-structures such as products and maps are implemented in a similar way.
+are eventually delegated to the operations on a staged set data type, such as
+@subsetOf@, @union@ and @intersect@. We do not need to change the
+implementation code, except the types -- such as from @Set[T]@ to @Rep[Set[T]]@.
+We show the changes in light gray code. The LMS library provides implicit
+conversions that lift the current-stage values to their next-stage values. In
+the code generation part, these operations emit their corresponding next-stage
+code. Again, other lattice structures such as products and maps are implemented
+in a similar way.
 
@@
 
-Now we can implement the staged abstract semantics, where @R@ is instantiated as
-@Rep@, and the types of abstract components are reused from the unstaged version.
+Now we can implement the staged abstract semantics, under that @R@ is instantiated as
+@Rep@. The types of abstract components are reused from the unstaged version.
 
@@
 We use the same monad stack structure from the unstaged abstract interpreter,
-but replacing \textit{all} the transformers to their staged version. The
-following code shows this change. We manually squeeze the three inner
+but replacing \textit{all} the monad transformers to their staged version. The
+following code snippet shows this change. We manually fuse the three inner
 transformers (@IdM@ omitted) into a single monad @RepSetReaderStateM[A, B, C]@,
-where @A@ is the type parameter for the reader effect and @B@ for the state effect.
-In our monad stack scheme, they are both instantiated with type @Cache@. Now the
-grounded result type is a pair of two staged value: @Rep[Set[(Value, Store)]]@ and
-@Rep[Cache]@.
+where @A@ is the type parameter for the reader effects and @B@ for the state effects.
+In our monad stack scheme, they are both instantiated with the type @Cache@. 
+Similar to the unstaged version, the grounded result type is a pair of two
+staged value: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
 \begin{lstlisting}
@@
 
-\paragraph{Primitive Operations} When deriving the staged concrete interpreter
-from the unstaged counterpart, we notice that the recursive calls of @ev@ on the
-lambda body are shifted from @ap_clo@ to @close@. In other words, we need to
-eagerly specialize the interpreter and generate code every time when we see a
-$\lambda$ term, instead of lazily when applications happen.
-With no surprise, it is similar in the staged abstract interpreter;
-additionally, we have to handle nondeterminism occurred by over-approximation
-of runtime behaviors.
+\paragraph{Primitive Operations} 
+When deriving the staged concrete interpreter from its unstaged counterpart, we
+notice that the recursive calls of the evaluator @ev@ on the lambda body are shifted from
+@ap_clo@ to @close@. In other words, we eagerly specialize the interpreter and
+generate code every time when we see a $\lambda$ term, instead of lazily when the
+applications happen.  With no surprise, it is similar here. Additionally,
+in the staged abstract interpreter, we have to handle nondeterminism
+incurred by the over-approximation of runtime behavior.
 In the rest of this part, we keep our roadmap and focus on discussing closure
@@
 
-The @close@ method now is a mixing of the \textit{staged} concrete version and
-unstaged \textit{abstract} version: we build a current-stage function @f@, which
+The @close@ method now is a mix of the \textit{staged} concrete version and
+unstaged \textit{abstract} version. We first build a current-stage function @f@, which
 takes four next-stage values, including the argument and latest store, and
-@in/out@ caches additionally. Inside of @f@, we collapse the @Ans@ monad to
-grounded values of type @Result@ by providing the desired arguments, i.e., the
+@in/out@ caches additionally. Inside of @f@, we collapse the @Ans@ monadic value to
+a grounded value of type @Result@, by providing the desired arguments, i.e., the
 new environment, new store, and caches. The collapsing of monad happens at
-current-stage, so the call of @ev@ is unfolded through staging. Finally, we
+current-stage, so the calls of @ev@ is unfolded within staging. Finally, we
 generate a singleton set containing the compiled next-stage closure
-@emit_compiled_clo(f)@, represented by an IR node in LMS.
+@emit_compiled_clo(f)@, which is represented by an IR node in LMS.
 \begin{lstlisting}
@@
 method @emit_ap_clo@ generates the current-stage representation of the
-future-stage result of application.
+future-stage application result.
 Finally, we put the future @out@ cache, store, and values back into
@@
 \begin{lstlisting}
-  def emit_ap_clo(rator: Rep[AbsValue], rand: Rep[Value], σ: Rep[Store],
+  def emit_ap_clo(fun: Rep[AbsValue], arg: Rep[Value], σ: Rep[Store],
                   in: Rep[Cache], out: Rep[Cache]): Rep[(Set[ValSt], Cache)]
-  def ap_clo(ev: Expr => Ans)(rator: Rep[Value], rand: Rep[Value]): Ans = for {
-    σ   <- get_store;      clo <- lift_nd[AbsValue](rator)
+  def ap_clo(ev: Expr => Ans)(funs: Rep[Value], arg: Rep[Value]): Ans = for {
+    σ   <- get_store;      clo <- lift_nd[AbsValue](funs)
     in  <- ask_in_cache;   out <- get_out_cache
@@
 again relies on the two caches @in@ and @out@, which are both staged maps now.
-Therefore, the query of whether the @out@ cache contains the current
-configuration will produce a next-stage Boolean value, i.e., of type
+Therefore, the query of whether the @in/out@ cache contains the current
+configuration will produce next-stage Boolean values, i.e., of type
 @Rep[Boolean]@. Consequently, the branching condition cannot be determined
-statically -- we need to generate code for @if@. Figure
+statically -- we need to generate code for the whole @if@ expression. Figure
 \ref{fig:staged_coind_cache} shows the staged version of @fix@. The
-monadic-binded variable @res@ is a next-stage result, consisting of a next-stage
-@if@ expression. The true branch simply returns a pair of queried result form
+variable @res@ represents the next-stage result, consisting of a next-stage
+@if@ expression. The true branch simply returns a pair of the queried result form
 the @out@ cache and the @out@ cache itself. The else branch constructs a monadic
 value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@ cache.
-After which, we use the similar technique that collapses the monad @m@ to
-grounded values by providing its desired environment and etc. Finally, we have a
+After which, we use a similar technique that collapses the monadic value @m@ to
+grounded values, by providing its desired environment and etc. Finally, we have a
 current-stage representation of the future-stage values @res@, and put the
@@
       // a next-stage value of type Rep[(Set[ValSt], Cache)] generated by if
-      if (out.contains(cfg)) (repMapToMapOps(out).apply(cfg).toList, out)
+      if (out.contains(cfg)) (out(cfg), out)
       else { val m: Ans = for {
@@
 The code generation for compiled closures and function applications are similar
-to the staged concrete interpreter. We have two the IR node as case classes, they
-take additional caches as arguments. We also elide how to generate next-stage Scala
-code for them.
+to their counterparts in the staged concrete interpreter. We have two IR nodes
+implemented as case classes; they also take additional caches as arguments. We
+elide the code generation part for these IR nodes.
 \begin{lstlisting}
@@
 \subsection{Optimizations} \label{staged_ds}
-
-\iffalse
-Revision: Solving Practical Challenges.
-theoretically, all the things should work nicely. Unfolding the interpreter over the AST.
-But the generated code is blowed up. For example .... This should not affect the correctness,
-but poses burden on the MSP system (ie LMS) and the next stage compiler/runtime (ie, scalac and JVM).
-1) LMS becomes slower since a large IR graph is contructed during the staging.
-2) Scalac becomes slower when reading a such large source code.
-JVM has certain limitation on the size of a single method.
-
-TODO: can we formulate selective caching as a partially-static data law.
-TODO: lambda lifting for if
-\fi
 
@@
     version. For instance, the staged data structures we used here are simply
-    black boxes that wrap the Scala's library data structures (in the next
-    stage).
+    black boxes that wrap the Scala's library data structures.
   \item In our experiments, the staged abstract interpreter produces the same

### 2019-09-06 paper/50_stageabs.tex
typo

@@
 
-Our staging schema works by unfolding the interpreter over the input AST. In
-practice, however, it would suffer from code explosion when analyzing
-(specializing) large programs, which increases the compile time. If we generate
-next stage programs running on JVM, such large generated program would also
-potentially cause runtime GC overhead. In this section, we present several
-optimizations that largely mitigate these issues. For all that, implementing
-these optimizations do not need to change the generic interpreter.
+Our staging schema works by unfolding the interpreter over the input abstract
+syntax tree. In practice, however, it would suffer from code explosion when
+analyzing (specializing) large programs, which increases the compile time. If
+we generate next-stage programs running on JVM, such large generated program
+would also potentially cause runtime GC overhead. In this section, we present
+optimizations that largely mitigate these issues. For all that,
+implementing these optimizations do not need to change the generic interpreter.
 
@@
 
-Although every component except @Expr@s are staged, we treat the data
-structures such as @Map@s as black-boxes, which means the operations on a @Map@
-directly become code in the next stage, but we don't inspect any further inside.
-As we identified when introducing the generic interface, the keys of @Env@s are
-identifiers (i.e., strings) in the program, which are completely known
-statically. This leaves us a chance to further specialize the data structures.
-For example, let us assume that the @Map[K, V]@ is implemented as a hash map. If
-the keys of type @K@ are all known statically, then the indices also can be
-computed statically. Thus the specialized map would be an array of type
-@Array[Rep[V]]@, whose elements are next-stage values; the size of the array is
-also known statically. An access to the map is translated into an access to the
-array with pre-determined index during staging.
+Every components in the interpreter except @Expr@s are staged. The data
+structures representing these components are treated as black-boxes, such as
+@Map@, which means that the operations on a @Map@ directly become nex-stage code,
+but we don't inspect any further inside. As we identified when
+introducing the generic interface, the keys of @Env@s are identifiers (i.e.,
+strings) in the program, which are completely known statically. This
+observation gives us a chance to further specialize the data structures.  For
+example, let us assume that the @Map[K, V]@ is implemented as a hash map. If
+all the keys of type @K@ are all known statically, then the indices also can be
+computed statically. Thus, the specialized map would be an array of type
+@Array[Rep[V]]@, whose elements are next-stage values, and the size of the array is
+known statically. An access to the map is translated into an access to the
+array with the pre-computed index during staging.
 
 Particularly, if we are specializing a monovariant analysis, the address space
-is equivalent to the set of identifiers in the program. Then the accesses to the
-environment can be completely determined statically and generates the addresses
-directly. After which, the stores can be specialized as arrays of @Rep[Value]@
-elements.
+is equivalent to the set of identifiers in the program. Then the accesses to
+the environment can be completely computed statically and we may generate the
+addresses directly during staging. After which, the stores are specialized
+as arrays with element type @Rep[Value]@.
 
-\paragraph{Selective Caching} It is observed that the two-fold co-inductive
-caching is used for every recursive call of our abstract interpreter. But this
-is not necessary and even very expensive when generating code for atomic
+\paragraph{Selective Caching} We observe that the two-fold co-inductive
+caching is used for every recursive calls in our abstract interpreter. But this
+is not necessary and redundant when generating code for atomic
 expressions such as literals or variables, because they always terminate.
 Borrowing the idea from the partition of expressions used in A-Normal Form
-\cite{Flanagan:1993:ECC:155090.155113}, we may use a selective caching
-algorithm that does not generate caching for atomic expressions:
+$\lambda$-calculus \cite{Flanagan:1993:ECC:155090.155113}, we can use a
+selective caching algorithm that does not generate caching code for atomic
+expressions:
 \begin{lstlisting}
@@
 \paragraph{Partially-static Data}
-
-Our treatment to binding-times is coarse-grained: @Expr@s is static, the rest of
-the world are all dynamic. But this is not always true since the static data
-has to be used somewhere with the dynamic operation. Partial-static data is a
+Our treatment of binding-times is coarse-grained: @Expr@s is static, the rest
+of the components are all dynamic. But this is not always true since the static
+data has to be used somewhere with dynamic operations. Partial-static data is a
 way to improve binding-times and optimize the generated code.
 
-For example, to fold a singleton set (often appears in @SetT@), e.g.,
+For example, to fold a static singleton set (often appears in @SetT@), e.g.,
 @Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully apply @foldLeft@ to the set with an anonymous
-function. But we can also utilize the algebraic property of @foldLeft@ to
+generator would faithfully generates code that applies @foldLeft@ with the set and
+function @f@. But we can also utilize the algebraic property of @foldLeft@ to
 generate cheaper code: @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
@@
 operation and function application. We apply several rewritings enabled by
-partial-static data structures, such as @Set@ and @Map@, which greatly reduces
-the size of residual programs.
+partial-static patterns, such as for @Set@ and @Map@. This optimizations
+greatly reduces the size of residual programs.
 
@@
 We have gradually presented the confluence of specialization and abstraction of
-concrete interpreters from an \textit{operational} perspective.
-In this section, we review and summarize our recipe to achieve the staged
-abstract interpreter and discuss correctness issues and different design choices.
+concrete interpreters from an operational perspective.  In this section, we
+review and summarize our recipe to achieve the staged abstract interpreter and
+discuss correctness issues and different design choices.
 
@@
     into monads.
-  \item By using an appropriate monad stack, we can implement a modular, unstaged,
-    abstract interpreter. This step has been explored in the previous literature.
-  \item Replacing the monad stack to a staged monad stack; such monads operate on
-    staged data types, i.e., next-stage values. We also need to reimplement several
-    primitive operations on staged values, in a modular fashion.
+  \item Then we can implement an unstaged abstract interpreter modularly using
+    the appropriate monad stack. This step has been explored in the previous
+    literatures.
+  \item Finally, we replace the monad stack with a staged monad stack, and
+    refactor related primitive operations.  Such staged monads operate on
+    staged data types, i.e., next-stage values.
 \end{itemize}
 
-Monadic interpreters are known to be able to decouple the interpretation process
-and the underlying semantics. The key insight in this paper is that by making the
-monadic interpreter binding-time polymorphic, it can be extended for generating
-efficient code in a staging framework. The underlying semantics and staging are
-two orthogonal dimensions. It is important to note that the computation
-encapsulated by the monads are not staged: only data (such as sets and maps) are
-staged. All the monadic computation, i.e., functions passed to the monadic bind
-operation @flatMap@, are statically known. This is why we can eliminate the
-monadic layer and its associated overhead.
+Monadic interpreters are known to be able to decouple the interpretation
+procedure and the underlying semantics. The key insight in this paper is that
+by making the monadic interpreter binding-time polymorphic, the interpreter can
+be extended for generating efficient code. The underlying semantics and staging
+are two orthogonal dimensions. It is important to note that the computation
+encapsulated by the monads are not staged: only data (such as sets and maps)
+are staged. All the monadic computation, i.e., functions passed to the monadic
+bind operation @flatMap@, are statically known. This is why we can eliminate
+the monadic layer and its associated overhead.
 
@@
 also have several operations on tuples and lists, which are residualized from
-the use of monads internally.
+the internal code fragments of monads.
 
@@
 listed here is based on empirical evidence; a rigorous proof for soundness
-preservation of the staged abstract interpreters remains an open question.
+preservation of the staged abstract interpreter remains an open challenge.
 
 \begin{itemize}
-  \item We usually assume the underlying MSP system achieves cross-stage
-    persistence, i.e., the generated program does not change the observational
+  \item We usually assume the underlying MSP system preserves the meaning during staging,
+    i.e., the generated program does not change the observational
     behaviors of the original program. In the case of this paper, we use the
     LMS framework, which 1) relies on the type checking facilities of Scala, and
-    2) generates code in A-Normal Form \cite{Flanagan:1993:ECC:155090.155113}
+    2) generates code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113}
     that preserves the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
@@
     using unsafe features, such as casts.
-  \item As shown in the previous section, the generic interpreter is untouched
+  \item As shown in the previous sections, the generic interpreter is untouched
     and shared by the four artifacts. We only instantiate binding-time
     annotations and the monadic types. This allows the programmer to easily
-    check the correctness of implementation of staged monads modularly.
+    check the correctness of the implementation of staged monads modularly.
   \item We build the staged monads and staged data structures in a
     correct-by-construct way that directly corresponds to their unstaged
-    version. For instance, the staged data structures we used here are simply
-    black boxes that wrap the Scala's library data structures.
+    versions. For instance, the staged data structures we use here are simply
+    black boxes that wrap the data structures in Scala library.
   \item In our experiments, the staged abstract interpreter produces the same
@@
 
-\paragraph{Monadic-style vs Direct-style} We use a monadic interpreter throughout the paper,
-but it is not necessary for staging. One can inline the monadic
-operations and obtain an abstract interpreter in continuation-passing
+\paragraph{Monadic-style vs Direct-style} We use a monadic interpreter
+throughout the paper, but it is not necessary for staging. One can inline the
+monadic operations and obtain an abstract interpreter in continuation-passing
 style, or even translate back to a direct-style that may use explicit
-side-effects such as mutations. In either case, we can still apply the
-staging idea to the abstract interpreter and remove the interpretation
-overhead. However, monads allow the staged abstract interpreter to be
-implemented in a modular and extensible way.
+side-effects such as mutations. In either case, we can still apply the staging
+idea to the abstract interpreter and remove the interpretation overhead.
+However, monads allow the staged abstract interpreter to be implemented in a
+modular and extensible way.
 
@@
 We implemented a big-step, compositional abstract interpreter in monadic
-style, where \textit{compositional} means that every recursive call of our abstract
-interpreter is applied to proper substructures of the current syntactic
+style, where \textit{compositional} means that every recursive calls of our abstract
+interpreter is applied to the proper substructures of the current syntactic
 parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that
 specialization can be done by unfolding, as well as guarantees the termination
-of the specialization procedure. It is also possible to specialize small-step
+of the specialization procedure. It is also possible to specialize a small-step
 operational abstract semantics through abstract compilation
 \cite{Boucher:1996:ACN:647473.727587} -- as
-\citet{Johnson:2013:OAA:2500365.2500604} implemented it for
+\citet{Johnson:2013:OAA:2500365.2500604} presented for
 optimizing Abstract Abstract Machines. However, the generated abstract

### 2019-09-06 paper/10_intro.tex
typo
revision

@@
 nondeterminism from collecting interpreters.
-%For example, a collecting interpreter that returns all possible runtime values is
-%nondeterministic, which can be modeled by a list monad.
 Although such pure approaches have their own merits, compared with imperative,
@@
 removes interpretative overhead and yields an equivalent executable
-that runs faster. The Futamura projections have been successfully applied to
-other similar performance problems, examples including
-query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \footnote{PyPy:
-\url{http://pypy.org/}}, Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
+that runs faster. 
+In recent years, the idea of Futamura projection has regained more attentions
+and successfully applied to many real-world problems, examples including
+query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
+Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
 The challenge here: can we apply the same idea to a general abstract interpreter

### 2019-09-06 paper/50_stageabs.tex
update

@@
 
-Every components in the interpreter except @Expr@s are staged. The data
+All components in the interpreter except @Expr@s are staged. The data
 structures representing these components are treated as black-boxes, such as
@@
 \paragraph{Selective Caching} We observe that the two-fold co-inductive
-caching is used for every recursive calls in our abstract interpreter. But this
+caching is used for every recursive call in our abstract interpreter. But this
 is not necessary and redundant when generating code for atomic
@@
 We implemented a big-step, compositional abstract interpreter in monadic
-style, where \textit{compositional} means that every recursive calls of our abstract
+style, where \textit{compositional} means that every recursive call of our abstract
 interpreter is applied to the proper substructures of the current syntactic

### 2019-09-06 paper/60_cases.tex
revision
update

@@
 AC was introduced by \citet{Boucher:1996:ACN:647473.727587} as an
-implementation technique for abstract interpretation-based static analysis.
-Similar to the present paper, the idea is inspired by partial evaluation:
+implementation technique for abstract interpretation-based static analyses.
+Similar to the present paper, the idea of AC was inspired by partial evaluation:
 the program is known statically and the interpretive overhead can be
@@
 Shivers:1988:CFA:53990.54007} for continuation-passing style (CPS) programs.
-Since the analyzed program is in CPS form, the analyzer is a big-step
-control-environment abstract interpreter. The abstract compilation transforms
-the analysis as a closure just taking an environment argument. Therefore the
-overhead of traversing the AST of input program has been eliminated.
+The analyzer is a big-step control-environment abstract interpreter. 
+By applying AC, we refactor every function in the analyzer such that it
+return a closure just taking an environment argument.  Therefore the overhead
+of traversing the AST of input program is eliminated.
 
-In this section, we show that the AC (closure generation as example) can be
-understood and implemented as an instance of staging abstract interpreters. We
-first revisit the original implementation of abstract compilation of 0-CFA, and
-then reproduce their result by simply adding stage annotations.
-The program generated by our approach provides approximately the same improvement with regards to
-speed, but without changing a single line of the analyzer program using
-type-based staging annotations. However, AC requires more engineering effort,
-i.e., rewriting the whole analyzer into the closure-generation form. Moreover,
-as shown in Section~\ref{staged_ds}, our approach is able to not only remove the
-interpretive overhead, but also enable several optimizations such as specialized
-data structures and equivalent rewritings.
+In this section, we show AC (closure generation) can be understood and
+implemented as an instance of staging abstract interpreters. We first revisit
+the original implementation of abstract compilation for 0-CFA, and then
+reproduce their result by simply adding stage annotations.
+The program generated by our approach provides approximately the same
+improvement with regards to speed, but without changing a single line of the
+analyzer program using type-based staging annotations. However, AC requires
+more engineering effort, i.e., rewriting the whole analyzer into the
+closure-generation form. Moreover, as shown in Section~\ref{staged_ds}, our
+approach is able to not only remove the interpretive overhead, but also enable
+several optimizations such as specialized data structures and equivalent
+rewritings.
 
@@
 
-We use Scheme as an example, and consider there is a library that contains a
-list map function @map@:
+We use Scheme as an example, and consider a library that contains a
+function @map@:
 \begin{lstlisting}

### 2019-09-07 paper/60_cases.tex
typo & revision

@@
 approach is able to not only remove the interpretive overhead, but also enable
-several optimizations such as specialized data structures and equivalent
-rewritings.
+several optimizations such as specialized data structures.
 
@@
 \paragraph{Original AC}
-The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is 0-CFA
+The analysis presented by \citeauthor{Boucher:1996:ACN:647473.727587} is a 0-CFA
 for a small CPS language consisting of abstractions, applications, recursion, and
@@
 implementation: different syntactic constructs are handled in different functions
-(e.g., @compCall@, @compApp@, etc.).
-Before refactoring into AC, these functions would take the store as a
-regular argument; now every function returns a value of type @CompAnalysis@,
-which is a function (closure) value that takes and returns stores. After the
-first run of analysis, we have traversed the AST and obtained a single closure that
+(e.g., @compCall@ and @compApp@).
+Before refactoring to AC, the original functions take a store as a
+regular argument. Now every function returns a value of type @CompAnalysis@,
+which is a function value (i.e., closure) that takes and returns stores. After the
+first run of the analysis, we have traversed the AST and obtained a single closure that
 takes a store and returns a store. Therefore, the ASTs are static terms, and the
-stores are dynamic terms. Now we can invoke the compiled closure with an initial
+stores are dynamic terms. Now, we can invoke the compiled closure with an initial
 store to complete the analysis.
@@
 
-On the other side, our approach works equivalently through staging: the
-ASTs are static, whereas stores are dynamic; therefore, we annotate the type @Store@
-with @Rep@ (shown in Figure \ref{compare_ac_sai} (right)). In fact, the only
-changes of our approach using the LMS framework are the types of @Store@ to
-@Rep[Store]@, since they will be known at the next stage: no other parts need to change.
-The generated code
-will consist entirely of looks-up and updating the store. The functions at the current stage are
-eliminated in the residual program; therefore, we do not generate higher-order
-functions, which provides additional performance improvement. Nevertheless,
-the support from the LMS framework is necessary, such as staging support for @Map@.
-We consider these efforts reusable and completely orthogonal to the
-implementation of the analysis.
+On the other side, our approach works equivalently through staging: the ASTs
+are static, whereas stores are dynamic. Therefore, we annotate the type @Store@
+with @Rep@ (Figure \ref{compare_ac_sai} (right)). In fact, the only changes of
+our approach using the LMS framework are the types of @Store@ to @Rep[Store]@,
+since they will be known at the next stage. There is no need for other changes.
+The generated programs consist entirely of looks-up and updating operations
+w.r.t the store.  The functions at the current stage are eliminated in the
+residual program; therefore, we do not generate higher-order functions, which
+provides additional performance improvement. Nevertheless, the LMS framework
+provides support for necessary data structures, such as @Map@.  We consider
+these efforts reusable and completely orthogonal to the implementation of the
+analysis.
 
@@
 extend and refactor the staged abstract interpreter we presented in Section
-\ref{sai} to various control-flow analyses (CFA). We first discuss
+\ref{sai} to various control-flow analyses (CFA). We first discuss the
 calling-context sensitive analysis for staged analyzers by tweaking the
-allocation strategy. Then we describe how to implement store-widening with
+allocation strategy. Then, we describe how to implement store-widening with the
 staged monads, which is a common optimization in CFA. Finally, we integrate
-abstract garbage collection with staging. All of them are in fact orthogonal to
+abstract garbage collection with staging. All of them are orthogonal to
 staging.
@@
 analyses \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we demonstrate a
-calling-context-sensitive analysis, i.e., $k$-CFA-like analysis
-\cite{DBLP:journals/jfp/HornM12}, can be implemented our staged abstract
+calling-context sensitive analysis, i.e., $k$-CFA-like analysis
+\cite{DBLP:journals/jfp/HornM12}, can be implemented in our staged abstract
 interpreter.

### 2019-09-07 paper/60_cases.tex
revision

@@
 We then integrate the timestamp into our monad stack by adding a staged
-@StateT@ monad. The result type is accordingly updated to contain a set of
+@StateT@ monad. The result type is accordingly updated to contain the set of
 tuples, where the time is added.
@@
 
-Every time we call the @eval@ function, we refresh the timestamp by
+Every time when we call the @eval@ function, we also refresh the timestamp by
 calling the @tick@ function, which updates the timestamp in the state monad.
@@
 being evaluated (@e@) to the existing calling context, and then taking the
-first $k$ elements from the list.  The type @Config@ used in caching is also
-changed to include timestamps.
+first $k$ elements from the list. The type @Config@ used in the caching
+algorithm is also changed to include timestamps.
 
@@
 
-A commonly-used technique to improve the running time in control-flow analysis
+A commonly-used technique to improve the running time of control-flow analyses
 is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
-DBLP:journals/pacmpl/DaraisLNH17}, store-widening analysis can be easily
-implemented in monadic abstract interpreter by swapping the @StateT@ (for store)
-and the @SetT@ (for nondeterminism). In this section, we show this optimization
-is orthogonal to staging and we just need to similarly swap the same \textit{staged}
-version of monads. After the adjustment, we have the following monad stack:
+DBLP:journals/pacmpl/DaraisLNH17}, a store-widening analysis can be easily
+implemented in the monadic abstract interpreter by swapping the @StateT@ and
+the @SetT@ monad transformer. In this section, we show this optimization is
+orthogonal to staging and we just need to similarly swap the same
+\textit{staged} version of monads. After the adjustment, conceptually, we have
+the following monad stack:
 \begin{lstlisting}
@@
 showed that, for big-step monadic abstract interpreters, we may add a @Reader@
-monad to track a set of root addresses and compute the reachable addresses every
+monad to track the set of root addresses and compute the reachable addresses every
 time when we obtain a value. Here, we adopt the idea with staged monads:
@@
 The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We
-update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@,
-but the caching part is omitted in the presented code. Several auxiliary
+update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@
+(the caching part is omitted in the presented code snippet). Several auxiliary
 functions are used: @ask_root@ retrieves the current root addresses from the
-monad stack; @root@ computes the reachable addresses from a value; and @gc@
-performs garbage collection according to the reachable addresses and returns a new
-store. For the generic interpreter, we also need to compute the root addresses
-for compound expressions, readers may refer to
-\cite{DBLP:journals/pacmpl/DaraisLNH17}.
+monad stack; @root@ computes the reachable addresses given a value; and @gc@
+performs garbage collection according to the given reachable addresses and
+returns a new store. For the generic interpreter, we also need to compute the
+root addresses for compound expressions; for the detail of this part, readers
+may refer to \cite{DBLP:journals/pacmpl/DaraisLNH17}.
 

### 2019-09-07 paper/10_intro.tex
space

@@
 that runs faster. 
-In recent years, the idea of Futamura projection has regained more attentions
+In recent years, the idea of Futamura projections has regained more attentions
 and successfully applied to many real-world problems, examples including

### 2019-09-07 paper/20_prelim.tex
misc
space

@@
 The Scala compiler translates the above @for@-comprehension expression into
-an equivalent one using @flatMap@ and @map@ ~\cite{scala_spec}. The last binding
+an equivalent one using @flatMap@ and @map@. The last binding
 in @for@-comprehensions is translated into a @map@, where the argument of
@@
 Given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$,
-the top-level @run@ method invokes the fixed-point operator directly,
+the top-level @run@ method invokes the fixed-point operator directly.
 \begin{lstlisting}

### 2019-09-07 paper/10_intro.tex
typos

--- 2019-09-06 paper/10_intro.tex
@@
 that runs faster. 
-In recent years, the idea of Futamura projection has regained more attentions
-and successfully applied to many real-world problems, examples including
+In recent years, the idea of Futamura projections has gained more attentions
+and has been successfully applied to many real-world problems, examples including
 query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
@@
 a nonstandard semantics that extensively uses nondeterminism, operates on
-abstract domains, and computes fixed-points iteratively over a cache.
+abstract domains, and computes fixed-points iteratively over caches.
 
@@
 Although the four artifacts may look dissimilar at first glance, they in
-fact are all firmly rooted in the concrete semantics of the language.  This
+fact are all firmly rooted in the concrete semantics.  This
 observation provides a way to abstract over the interpreter and achieve the

### 2019-09-07 paper/50_stageabs.tex
typos

@@
 abstract interpreter in this section uses the same abstract semantics from the
-unstaged version in Section~\ref{unstaged_abs}.  The approach we refactor
+unstaged version in Section~\ref{unstaged_abs}.  The approach we use to refactor
 the unstaged one to the staged abstract interpreter is modular, and does not
@@
 monads, and then discuss the staged version of primitive operations, especially
-@close@, @ap_clo@ and @fix@. At last, we discuss several optimizations.
+@close@, @ap_clo@ and @fix@. In the end, we discuss several optimizations.
 
@@
 We use the same monad stack structure from the unstaged abstract interpreter,
-but replacing \textit{all} the monad transformers to their staged version. The
+but replacing \textit{all} the monad transformers with their staged versions. The
 following code snippet shows this change. We manually fuse the three inner

### 2019-09-07 paper/10_intro.tex
typo

@@
 their implementations.  The soundness of an analysis can also be proven with less
-effort, whether by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
+effort, either by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
 paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.

### 2019-09-07 paper/40_absinterp.tex
typo

@@
 lattice representation with binding-time abstraction, as well as simple
-abstract domains, such as power set lattice.  Then we show the abstract
-components, namely abstract values, abstract environments, and abstract stores.
+abstract domains, such as power sets.  Then, we show the abstract
+components, namely, abstract values, abstract environments, and abstract stores.
 The abstract interpreter we construct in this section is similar to
@@
 After the evaluation, the store @σ@ may have been changed. Therefore, 
-we get the latest store and construct the result @(v, $\sigma$)@; the
+we get the latest store and construct the result @(v, σ)@; the
 result will be used to update the out cache.

### 2019-09-07 paper/50_stageabs.tex
typo

@@
 achieve the stage polymorphism. Now we instantiate the type @R@ as @Rep@ and
-still use power-set as example to describe its staged version.
+still use the power-set as example to describe its staged version.
 \begin{lstlisting}[escapechar=!]
@@
 Similar to the unstaged version, the grounded result type is a pair of two
-staged value: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
+staged values: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
 \begin{lstlisting}
@@
 Our treatment of binding-times is coarse-grained: @Expr@s is static, the rest
-of the components are all dynamic. But this is not always true since the static
-data has to be used somewhere with dynamic operations. Partial-static data is a
+of the components are all dynamic. But this is not always true, because the static
+data have to be used somewhere with the dynamic operations. Partial-static data is a
 way to improve binding-times and optimize the generated code.

### 2019-09-07 paper/80_related.tex
typo

@@
 less engineering efforts to refactor the analyzer (Section \ref{cs_ac}).
-\citet{Johnson:2013:OAA:2500365.2500604} adapted closure generation to optimize
+\citet{Johnson:2013:OAA:2500365.2500604} adapted closure generation to optimize a
 small-step abstract interpreter in state-transition style. The program
-is first compiled to an ``abstract bytecode'' IR. The IR consists of higher-order
+is firstly compiled to an ``abstract bytecode'' IR. The IR consists of higher-order
 functions, which will be executed later on an abstract abstract machine for
-that IR.  \citet{damian1999partial} provided a formal treatment to abstract
+that IR.  \citet{damian1999partial} provided a formal treatment for abstract
 compilation and control-flow analysis, as well as proofs to establish the
@@
 flow-sensitive pointer analysis. The first stage is to analyze the
-programs to obtain a sparse representation, and then the second
-stage conducts the flow-sensitive analysis by refining the first
+programs and obtain a sparse representation, and then the second
+stage conducts a flow-sensitive analysis by refining the first
 one. Another large body of static analyses is implemented by
@@
 Such analyses can be considered having stage distinctions.  One piece of future
-work is to explore whether they can be implemented and optimized from
+work is to explore whether they can be implemented and optimized using
 meta-programming.
@@
 projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
-However, it only specializes Datalog rules to the running engine, instead of a
+However, it only specializes the solving engine to Datalog rules, instead of a
 full pipeline from a program AST to a specialized solver.
@@
 Recent work shows that numerical abstract domains can be
-specialized w.r.t. the static program structure, although not by
-way of staging. For example, decomposing polyhedrons
+specialized to the static program structure, although not by
+ways of staging. For example, decomposing polyhedrons
 \cite{DBLP:conf/popl/SinghPV17, Singh:2017:PCD:3177123.3158143} to

### 2019-09-07 paper/10_intro.tex
intro

@@
 
-Roughly at the same time when abstract interpretation was proposed in the
-1970s, \citet{futamura1971partial} observed a close
+Roughly at the same time when abstract interpretation was proposed,
+\citet{futamura1971partial} observed a close
 connection between interpreters and compilers through a hierarchy of
@@
 that runs faster. 
-In recent years, the idea of Futamura projections has gained more attentions
-and has been successfully applied to many real-world problems, examples including
+In recent years, the idea of Futamura projections has attracted more attentions
+and has been successfully applied to many real-world problems, including
 query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
 Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
-The challenge here: can we apply the same idea to a general abstract interpreter
+The challenge here: can we apply Futamura projections to a general abstract interpreter
 to make it run faster? The solution is not obvious, as we need to specialize

### 2019-09-07 paper/10_intro.tex
intro

@@
 model as possible.
-In short, our approach borrows ideas and techniques from multi-stage
-programming and embedded DSLs, and apply them to abstract interpreters:  1) To
+In short, we borrow ideas and techniques from multi-stage programming and
+embedded DSLs, and apply them to abstract interpreters:  1) To
 remove the overhead from interpretation and effect layers, we specialize the
 abstract interpreter via staging and then generate efficient low-level code
-that does the actual analysis. 2) Inspired by the tagless-final interpreters
+that performs the actual analysis. 2) Inspired by the tagless-final interpreters
 \cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that

### 2019-09-07 paper/30_stageconc.tex
typo

@@
 we also use a @Rep@ prefix on the constructors and types to differentiate them.
-But it is important to note that instances of the case class @ReaderT@/@StateT@
+But it is important to note that the instances of @ReaderT@/@StateT@
 are not staged, and that the monadic computation, such as functions of @R => M[A]@,
@@
 
-The function @f@ passed to @flatMap@ is also a function value known at the current stage.
+The function @f@ passed to @flatMap@ is also a value known at the current stage.
 Therefore all invocations of @flatMap@ can be reduced before generating code.
@@
   case class IRCompiledClo(f: Rep[(ValSt) => ValSt], λ: Lam, ρ: Rep[Env]) extends Def[Value]
-  case class IRApClo(rator: Rep[Value], rand: Rep[Value], σ: Rep[Store]) extends Def[ValSt]
+  case class IRApClo(fun: Rep[Value], arg: Rep[Value], σ: Rep[Store]) extends Def[ValSt]
 \end{lstlisting}
@@
 The IR nodes are manipulated by the LMS passes and finally generated as
-next-stage Scala code. To generate code, LMS provides an @emitNode@ method for
+next-stage Scala programs. To generate code, LMS provides an @emitNode@ method for
 programmers to control over what code is generated for each kind of nodes. Here,
 we match @IRCompiledClo@ and @IRApClo@, and generate definitions via
-@emitValDef@ for them.  For compiled closures, we put the function @f@ and its
-accompanying syntactic term and environment to their next-stage value
-representation @CompiledClo@. For applications, we know the fact that @fun@ is
+@emitValDef@ for them.  For compiled closures, we pass the function @f@ with its
+accompanying syntactic term and environment to the next-stage value
+representation @CompiledClo@. For applications, we know that @fun@ is
 an instance of @CompiledClo@ at the next stage, and it has a callable field
@@
   case IRCompiledClo(f, λ, ρ) => emitValDef(sym, s"CompiledClo($\textdollar${quote(f)}, $\textdollar${quote(λ)}, $\textdollar${quote(ρ)})")
-  case IRApClo(rator, rand, σ) => emitValDef(sym, s"$\textdollar$rator.f($\textdollar${quote(rand)}, $\textdollar${quote(σ)})")
+  case IRApClo(fun, arg, σ) => emitValDef(sym, s"$\textdollar$fun.f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
   ...

### 2019-09-08 paper/50_stageabs.tex
typo

@@
 
-The trait @RepLattice@ is shown by specializing @Lattice@ with type @Rep@.
+The trait @RepLattice@ is shown by instantiating @Lattice@ with type @Rep@.
 The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
@@
 @subsetOf@, @union@ and @intersect@. We do not need to change the
-implementation code, except the types -- such as from @Set[T]@ to @Rep[Set[T]]@.
-We show the changes in light gray code. The LMS library provides implicit
-conversions that lift the current-stage values to their next-stage values. In
-the code generation part, these operations emit their corresponding next-stage
-code. Again, other lattice structures such as products and maps are implemented
-in a similar way.
+implementations, except the types -- such as changing from @Set[T]@ to
+@Rep[Set[T]]@.  We show the changes in light gray code. The LMS library
+provides implicit conversions that lift the current-stage values to their
+next-stage representations. In the code generation part, these operations emit
+their corresponding next-stage code. Again, other lattice structures such as
+products and maps are implemented in a similar way.
 
@@
 @ap_clo@ to @close@. In other words, we eagerly specialize the interpreter and
-generate code every time when we see a $\lambda$ term, instead of lazily when the
-applications happen.  With no surprise, it is similar here. Additionally,
-in the staged abstract interpreter, we have to handle nondeterminism
-incurred by the over-approximation of runtime behavior.
-In the rest of this part, we keep our roadmap and focus on discussing closure
+generate code every time when we see a $\lambda$ term, instead of lazily
+calling @ev@ when the applications happen.  With no surprise, it is similar
+here. Additionally, in the staged abstract interpreter, we have to handle
+nondeterminism incurred by the over-approximation of runtime behavior.  In the
+rest of this part, we keep our roadmap and focus on discussing closure
 representations and function applications for the staged version.
@@
 takes four next-stage values, including the argument and latest store, and
-@in/out@ caches additionally. Inside of @f@, we collapse the @Ans@ monadic value to
-a grounded value of type @Result@, by providing the desired arguments, i.e., the
-new environment, new store, and caches. The collapsing of monad happens at
-current-stage, so the calls of @ev@ is unfolded within staging. Finally, we
-generate a singleton set containing the compiled next-stage closure
-@emit_compiled_clo(f)@, which is represented by an IR node in LMS.
+@in/out@ caches additionally. Inside of @f@, we collapse the monadic value of
+type @Ans@ to grounded values of type @Result@, by providing the desired
+arguments, i.e., the new environment, new store, and caches. The collapsing of
+monads happens at the current-stage, so the calls of @ev@ is unfolded within
+staging. Finally, we generate a singleton set containing the compiled
+next-stage closure @emit_compiled_clo(f)@, which is represented by an IR node
+in LMS.
 \begin{lstlisting}
@@
 future-stage application result.
-Finally, we put the future @out@ cache, store, and values back into
+Finally, we reify the future @out@ cache, store, and values back into
 the current-stage monads.
@@
 again relies on the two caches @in@ and @out@, which are both staged maps now.
-Therefore, the query of whether the @in/out@ cache contains the current
+Because of this, the query of whether the @in/out@ cache contains the current
 configuration will produce next-stage Boolean values, i.e., of type
-@Rep[Boolean]@. Consequently, the branching condition cannot be determined
-statically -- we need to generate code for the whole @if@ expression. Figure
+@Rep[Boolean]@. The branching condition cannot be determined
+statically, consequently we need to generate code for the whole @if@ expression. Figure
 \ref{fig:staged_coind_cache} shows the staged version of @fix@. The
 variable @res@ represents the next-stage result, consisting of a next-stage
-@if@ expression. The true branch simply returns a pair of the queried result form
-the @out@ cache and the @out@ cache itself. The else branch constructs a monadic
-value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@ cache.
-After which, we use a similar technique that collapses the monadic value @m@ to
-grounded values, by providing its desired environment and etc. Finally, we have a
-current-stage representation of the future-stage values @res@, and put the
-content of @res@ back into the monad stack.
+@if@ expression. The true branch simply returns a pair of the queried result
+form the @out@ cache and the @out@ cache itself. The else branch constructs a
+monadic value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@
+cache.  After which, we use a similar technique that collapses the monadic
+value @m@ to grounded values, by providing its desired environment and etc.
+Finally, we have a current-stage representation of the future-stage values
+@res@, and put the content of @res@ back into the monad stack.
 

### 2019-09-08 paper/20_prelim.tex
revision

@@
     \end{lstlisting}
+    \vspace{-0.5em}
     \caption{trait \texttt{Monad}} \label{fig:monad}
@@
     \end{lstlisting}
+    \vspace{-0.5em}
     \caption{trait \texttt{MonadOps}} \label{fig:monadops}
@@
 and an argument value. Note that the @Env@, @Store@, and @Value@ are all
-annotated by @R[_]@ because they can potentially be next-stage values when the
+annotated by @R[_]@, because they are potentially next-stage values when the
 interpreter acts as a compiler.

### 2019-09-08 paper/30_stageconc.tex
revision

@@
 code.
-
 A classic example for introducing multi-stage programming is the power function
@@
 
-If @x@ is a value known at the current stage, we may specialize the power function to
-the value @x@ -- by unrolling the recursive calls. In LMS, this is fulfilled by
+If @x@ is known at the current stage, we may specialize the power function to
+the value @x@ by unrolling the recursive calls. In LMS, this is fulfilled by
 first adding the @Rep@ type annotation to variables known at the next stage. In this case,
-@b : Rep[Int]@ is known later, and @x : Int@ is known currently, as shown in the
-below code snippet.
+@b: Rep[Int]@ is known later, and @x: Int@ is known currently, as shown below.
 The way we use this staged @power@ function is to create a @DslDriver@ and override the
-@snippet@ method that supplies 5 as the currently know value for @x@.
+@snippet@ method that supplies 5 as the currently known value for @x@.
 \begin{lstlisting}

### 2019-09-08 paper/abstract.tex
Proofread abstract

@@
 approach derives a sound staged abstract interpreter based on a
-semantic-agnostic interpreter with type-level binding-time abstraction and
-monadic abstraction. By using different instantiations of these abstractions,
-the generic interpreter can flexibly behave in four modes: unstaged concrete
-interpreter, staged concrete interpreter, unstaged abstract interpreter, or
-staged abstract interpreter. As an example of \emph{abstraction without regret},
-the overhead of these abstraction layers is eliminated in the generated code
-after staging. We show that staging abstract interpreters is practical and
-useful to optimize static analysis, while requiring less engineering effort,
-without compromising soundness. We conduct three case studies, including a
-comparison with Boucher and Feeley's abstract compilation, applications to
-various control-flow analyses, and a demonstration that can be used for modular
-analysis. We also empirically evaluate the effect of staging on the execution
-time. The experiment shows an order of magnitude speedup with staging for
-control-flow analyses.
+semantic-agnostic interpreter with type-level binding-time abstractions and
+monadic abstractions. By using different instantiations of these abstractions,
+the generic interpreter can flexibly behave in one of four modes: as an unstaged
+concrete interpreter, a staged concrete interpreter, an unstaged abstract
+interpreter, or a staged abstract interpreter. As an example of
+\emph{abstraction without regret}, the overhead of these abstraction layers is
+eliminated in the generated code after staging. We show that staging abstract
+interpreters is practical and useful to optimize static analysis, all while
+requiring less engineering effort and without compromising soundness. We conduct
+three case studies, including a comparison with Boucher and Feeley's abstract
+compilation, applications to various control-flow analyses, and a demonstration
+that can be used for modular analysis. We also empirically evaluate the effect
+of staging on the execution time. The experiment shows an order of magnitude
+speedup with staging for control-flow analyses.
 

### 2019-09-08 paper/10_intro.tex
revision

@@
 
-Recent progress on methodologies such as Abstracting Abstract Machine (AAM)
+Recent progress on methodologies such as Abstracting Abstract Machines (AAM)
 \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} uncovers an
@@
 abstract interpreter to be derived from its unstaged counterpart.  As a result,
-the derived staged abstract interpreters has no intrusive changes to the
+the derived staged abstract interpreter has no intrusive changes to the
 underlying abstract semantics, thereby preserving soundness.  In this
@@
   \item We present the unstaged abstract interpreter under the same
-    framework by replacing the environment, store, and values to their
+    framework by replacing the environment, store, and values with their
     abstract counterparts (Section~\ref{unstaged_abs}). We then

### 2019-09-08 paper/20_prelim.tex
revision

@@
 We consider a call-by-value $\lambda$-calculus in direct-style, extended
-with numbers, arithmetic, recursions, and conditionals. Other effects
+with numbers, arithmetic, recursion, and conditionals. Other effects
 such as assignments can also be supported readily.
@@
   case class If0(e1: Expr, e2: Expr, e3: Expr) extends Expr   // conditionals
-  case class Rec(x: String, rhs: Expr, e: Expr) extends Expr  // recursions
+  case class Rec(x: String, rhs: Expr, e: Expr) extends Expr  // recursion
   case class Aop(op: String, e1: Expr, e2: Expr) extends Expr // arithmetic
@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
   \centering
@@
   \end{lstlisting}
+\vspace{-0.5em}
 \caption{The generic interpreter,
   shared by the unstaged/staged + concrete/abstract interpreter.}
+\vspace{-1.5em}
 \label{fig:shared_int}
@@
 A value can be either a tagged number \texttt{IntV}, or a closure
-\texttt{CloV} that contains a lambda term and an environment. The final grounded
-result of the interpreter is a pair of values and stores, where the @Value@ and
-@Store@ can be next-stage objects, as annotated by the binding-time @R@. We
-also define a standard fixed-point combinator to close the open-recursive
-function @ev@.
+\texttt{CloV} that contains a lambda term and an environment. The final
+result of the interpreter is a pair of values and stores. 
+To distinguish from the monadic values of type @Ans@ produced by the
+interpreter, later we use \textit{grounded values} to denote such final values.
+The two elements in the type @Result@ are annotated by the binding-time @R@,
+because they cane be next-stage objects.  We also define a standard fixed-point
+combinator to close the open-recursive function @ev@.
 \begin{lstlisting}
   trait ConcreteComponents extends Semantics {
-    type Addr = Int
-    sealed trait Value
+    type Addr = Int;  sealed trait Value
     case class IntV(i: Int) extends Value

### 2019-09-08 paper/50_stageabs.tex
revision

@@
 
-\begin{figure}[h!]
+\begin{figure}[t]
 \centering
@@
 \end{lstlisting}
+\vspace{-0.5em}
 \caption{The staged co-inductive caching algorithm.}
 \label{fig:staged_coind_cache}
-\vspace{-2em}
+\vspace{-1.5em}
 \end{figure}
@@
 Our treatment of binding-times is coarse-grained: @Expr@s is static, the rest
-of the components are all dynamic. But this is not always true, because the static
-data have to be used somewhere with the dynamic operations. Partial-static data is a
-way to improve binding-times and optimize the generated code.
-
+of the components are all dynamic. But this is not always true, because the
+static data have to be used somewhere with the dynamic operations.
+Partial-static data is a way to improve binding-times and optimize the
+generated code.
 For example, to fold a static singleton set (often appears in @SetT@), e.g.,
 @Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully generates code that applies @foldLeft@ with the set and
-function @f@. But we can also utilize the algebraic property of @foldLeft@ to
-generate cheaper code: @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
+generator would faithfully generates code that applies @foldLeft@ with the set
+and function @f@. But we can also utilize algebraic properties of @foldLeft@
+to generate cheaper code, e.g, @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
 function @f@ is known at the current stage, we completely eliminate the fold
@@
 procedure and the underlying semantics. The key insight in this paper is that
-by making the monadic interpreter binding-time polymorphic, the interpreter can
-be extended for generating efficient code. The underlying semantics and staging
-are two orthogonal dimensions. It is important to note that the computation
-encapsulated by the monads are not staged: only data (such as sets and maps)
-are staged. All the monadic computation, i.e., functions passed to the monadic
-bind operation @flatMap@, are statically known. This is why we can eliminate
-the monadic layer and its associated overhead.
+by making the monadic interpreter stage polymorphic
+\cite{Ofenbeck:2017:SGP:3136040.3136060, Amin:2017:CTI:3177123.3158140}, the
+abstract interpreter can be extended for generating efficient code. The
+underlying semantics and staging are two orthogonal dimensions. It is important
+to note that the computation encapsulated by the monads are not staged: only
+data (such as sets and maps) are staged. All the monadic computation, i.e.,
+functions passed to the monadic bind operation @flatMap@, are statically known.
+This is why we can eliminate the monadic layer and its associated overhead.
 
@@
 
-Soundness is the central concern of static analysis, and as such, is vital
-for prospective users. We now briefly examine how our
-approach preserves the correctness and soundness of the analysis, under the
-assumption that the unstaged one is sound. It is worth noting that the rationale
-listed here is based on empirical evidence; a rigorous proof for soundness
-preservation of the staged abstract interpreter remains an open challenge.
+Soundness is the central concern of static analysis, and as such, is vital for
+prospective users. Our approach does not touch the soundness of the analysis,
+i.e, if the unstaged one does not produces false negative result, the staged
+one also does not. This soundness preservation relies on the correctness
+of the staged implementations and of the underlying MSP system. We now briefly
+examine how this is achieved. Note that the rationale listed
+here is based on empirical evidence; a rigorous proof for soundness
+preservation of our approach remains an open challenge.
 
 \begin{itemize}
-  \item We usually assume the underlying MSP system preserves the meaning during staging,
-    i.e., the generated program does not change the observational
-    behaviors of the original program. In the case of this paper, we use the
-    LMS framework, which 1) relies on the type checking facilities of Scala, and
-    2) generates code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113}
-    that preserves the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
-    However, in general, it is possible to subvert these guarantees by
-    using unsafe features, such as casts.
+  \item We usually assume the underlying MSP system preserves the meaning
+    during staging, i.e., the generated program and original program produce
+    the same result given the same arguments.
+    In this paper, we use the LMS framework, which 1) relies on the
+    type checking facilities of Scala, and 2) generates code in A-Normal form
+    \cite{Flanagan:1993:ECC:155090.155113} that preserves the execution order
+    within a stage \cite{DBLP:conf/birthday/Rompf16}.  However, in general, it
+    is possible to subvert these guarantees by using unsafe features, such as
+    casts.
   \item As shown in the previous sections, the generic interpreter is untouched
-    and shared by the four artifacts. We only instantiate binding-time
-    annotations and the monadic types. This allows the programmer to easily
+    and shared by the four artifacts.  This allows the programmer to
     check the correctness of the implementation of staged monads modularly.
+   % We only instantiate binding-time annotations and the monadic types. 
   \item We build the staged monads and staged data structures in a

### 2019-09-08 paper/80_related.tex
revision

@@
 In this paper, we use mutli-stage programming to specialize programs.
-The Lightweight Modular Staging framework
+The Lightweight Modular Staging (LMS) framework
 \cite{DBLP:conf/gpce/RompfO10} we used in the paper relies on
-type-level stage annotations.  Other notable implementations of MSP
-exist in the ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97} and
-MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03,
-  DBLP:conf/flops/Kiselyov14}.
-The idea of staging an abstract interpreter we presented here
-is also applicable to other MSP systems.
+type-level stage annotations.  Other notable MSP systems using syntactic
+quotations exist in the ML family, e.g., MetaML \cite{DBLP:conf/pepm/TahaS97}
+and MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}. 
+Compared with other approaches using syntactic quotations, LMS provides
+guarantee for preserving evaluation orders within a stage.
+\citet{Ofenbeck:2017:SGP:3136040.3136060} discussed staging for generic
+programming as well as stage polymorphism in LMS.
+%The idea of staging an abstract interpreter we presented here
+%is also applicable to other MSP systems.
 Multi-stage programming has been widely used to improve the

### 2019-09-08 paper/10_intro.tex
Proofreading intro

@@
 Abstract interpretation as a lattice-based approach to sound static analyses was
-proposed by \citet{DBLP:conf/popl/CousotC77, CousotCousot79-1}. Based on the notion of Galois
-connections, the analyzer can soundly approximate the
-concrete program behavior at runtime by computing fixed points on an abstract
-domain. Despite the tremendous theoretical development of abstract
-interpretation over the years, constructing artifacts and analyzers that perform
-sound abstract interpretation for modern and expressive languages was considered
-complicated for a long time.
+proposed by \citet{DBLP:conf/popl/CousotC77, CousotCousot79-1}. Based on the
+notion of Galois connections, an analyzer can soundly approximate concrete
+program behavior at runtime by computing fixed points on an abstract domain.
+Despite the tremendous theoretical development of abstract interpretation over
+the years, constructing artifacts and analyzers that perform sound abstract
+interpretation for modern and expressive languages was considered complicated
+for a long time.
 
 Recent progress on methodologies such as Abstracting Abstract Machines (AAM)
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} uncovers an
-operational approach to constructing abstract interpreters.  By deriving
-semantics artifacts for abstract interpretation from their concrete
-counterparts (for example, abstract machines), soundness can be easily
-established by examining the transformation of semantic artifacts.
-This systematic abstraction approach can be tailored to different
-language features (such as state, first-class control, exceptions, and
-concurrency) and sensitivity analyses \cite{DBLP:conf/icfp/Gilray0M16,
-  DBLP:conf/popl/GilrayL0MH16, Darais:2015:GTM:2814270.2814308}. It
-has also been applied to various small-step abstract machines
-\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
-  Sergey:2013:MAI:2491956.2491979} and big-step definitional
-interpreters \cite{Wei:2018:RAA:3243631.3236800,
-  DBLP:journals/pacmpl/DaraisLNH17, Keidel:2018:CSP:3243631.3236767}.
+\cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10} uncovers an operational
+approach to constructing abstract interpreters.  By deriving semantics artifacts
+for abstract interpretation from their concrete counterparts (for example,
+abstract machines), soundness can be easily established by examining the
+transformation of semantic artifacts. This systematic approach to abstraction
+can be tailored to different language features (such as state, first-class
+control, exceptions, concurrency, etc.) and sensitivity analyses
+\cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16,
+Darais:2015:GTM:2814270.2814308}. It has also been applied to various small-step
+abstract machines \cite{DBLP:journals/jfp/HornM12, DBLP:conf/icfp/HornM10,
+Sergey:2013:MAI:2491956.2491979} and big-step definitional interpreters
+\cite{Wei:2018:RAA:3243631.3236800, DBLP:journals/pacmpl/DaraisLNH17,
+Keidel:2018:CSP:3243631.3236767}.
 
@@
 include monads and monad transformers \cite{DBLP:journals/pacmpl/DaraisLNH17,
-Sergey:2013:MAI:2491956.2491979}, arrows
-\cite{Keidel:2018:CSP:3243631.3236767}, extensible effects
-\cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic}, etc. These pure
-approaches provide certain benefits, for example, the abstract interpretation
-artifacts can be built in a compositional and modular fashion, and referential
-transparency allows programmers to confidently reason about the correctness of
-their implementations.  The soundness of an analysis can also be proven with less
-effort, either by mechanized proofs \cite{Darais:2016:CGC:2951913.2951934} or
-paper-based proofs \cite{Keidel:2018:CSP:3243631.3236767}.
+Sergey:2013:MAI:2491956.2491979}, arrows \cite{Keidel:2018:CSP:3243631.3236767},
+extensible effects \cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic},
+etc. These pure approaches provide certain benefits, including the fact that the
+abstract interpretation artifacts can be built in a compositional and modular
+fashion. Furthermore, referential transparency allows programmers to confidently
+reason about the correctness of their implementations.  The soundness of an
+analysis can also be proven with less effort, whether by mechanized proofs
+\cite{Darais:2016:CGC:2951913.2951934} or paper-based proofs
+\cite{Keidel:2018:CSP:3243631.3236767}.
 
@@
 approaches. First, similar to concrete interpreters, an abstract interpreter
-analyzes a program by traversing its AST. This traversal incurs
-interpretive overhead, such as pattern matching on the AST nodes and recursive
-calls on subexpressions. If it is a single-time traversal, such overhead can be
-negligible. But usually, abstract interpreters iteratively analyze and traverse
-the AST multiple times to reach fixed-points, therefore the accumulated
-overhead is even magnified. 
+analyzes a program by traversing its AST. This traversal incurs interpretive
+overhead, such as pattern matching on the AST nodes and recursive calls on
+subexpressions. If this requires only a single traversal, such overhead can be
+negligible. Usually, however, abstract interpreters iteratively analyze and
+traverse an AST multiple times to reach fixed-points; the accumulated overhead
+is therefore magnified.
 %Repeatedly analyzing library programs also poses significant overhead.
@@
 %significantly if it runs repeatedly (for example, on libraries).
-Second, it is common that abstract interpreters written in pure languages
-extensively use functional encodings of side-effects to implement the semantics
+Second, it is common that abstract interpreters written in pure languages make
+extensive use of functional encodings of side-effects to implement the semantics
 of abstract interpretation.  For example, the list monad can be used to model
-nondeterminism from collecting interpreters.
-Although such pure approaches have their own merits, compared with imperative,
-stateful implementations, they are significantly slower.  In order to
-reconcile these approaches, one must use a programming technique which allows
-for high-level abstractions during implementation, but bypasses the
-incurred overhead at runtime.
+nondeterminism from collecting interpreters. Although such pure approaches have
+their own merits, they are significantly slower when compared with imperative,
+stateful implementations.  In order to reconcile these approaches, one must use
+a programming technique which allows for high-level abstractions during
+implementation, but bypasses the incurred overhead at runtime.
 
-Roughly at the same time when abstract interpretation was proposed,
-\citet{futamura1971partial} observed a close
-connection between interpreters and compilers through a hierarchy of
-specializations, since then known as Futamura Projections. The first Futamura
-projection describes that specializing an interpreter to the input program
-removes interpretative overhead and yields an equivalent executable
-that runs faster. 
-In recent years, the idea of Futamura projections has attracted more attentions
-and has been successfully applied to many real-world problems, including
-query compilers \cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
-Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc.
-The challenge here: can we apply Futamura projections to a general abstract interpreter
-to make it run faster? The solution is not obvious, as we need to specialize
-a nonstandard semantics that extensively uses nondeterminism, operates on
-abstract domains, and computes fixed-points iteratively over caches.
+Roughly at the same time abstract interpretation was proposed,
+\citet{futamura1971partial} observed a close connection between interpreters and
+compilers through a hierarchy of specializations, known since then as the
+Futamura Projections. The first Futamura projection states that specializing an
+interpreter to an input program removes interpretative overhead and yields an
+equivalent executable that runs faster. In recent years, the idea of Futamura
+projections has attracted more attention, and has been successfully applied to
+many real-world problems, including query compilers
+\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
+Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc. The challenge here: can we
+apply Futamura projections to a general abstract interpreter to make it run
+faster? The solution is not obvious, as we need to specialize a nonstandard
+semantics that extensively uses nondeterminism, operates on abstract domains,
+and computes fixed-points iteratively over caches.
 
@@
 interpreters, while keeping the implementation as close to their conceptual
-model as possible.
-In short, we borrow ideas and techniques from multi-stage programming and
-embedded DSLs, and apply them to abstract interpreters:  1) To
+model as possible. In short, we borrow ideas and techniques from multi-stage
+programming and embedded DSLs, and apply them to abstract interpreters:  1) To
 remove the overhead from interpretation and effect layers, we specialize the
-abstract interpreter via staging and then generate efficient low-level code
+abstract interpreter via staging, and then generate efficient low-level code
 that performs the actual analysis. 2) Inspired by the tagless-final interpreters
@@
 the derived staged abstract interpreter has no intrusive changes to the
-underlying abstract semantics, thereby preserving soundness.  In this
-sense, our approach allows for no regret for both performance and engineering
-effort. We elaborate these two main ideas in detail, as follows.
+underlying abstract semantics, thereby preserving soundness.  In this sense, our
+approach allows for a complete absence of regret in terms of both performance
+and engineering effort. We elaborate these two main ideas in detail, as follows.
 
@@
 
-\paragraph{Futamura Projection of Abstract Interpreters}
+\paragraph{Futamura Projections Meet Abstract Interpreters}
 
@@
 instance, assume that @eval : Expr --> Input --> Value@ is an interpreter for some language.
-Given a program $e_0$@ : Expr@, by applying the specialization, we can
+Given a program $e_0$@ : Expr@, by applying the specialization we can
 obtain a specialized program $\texttt{eval}_{\texttt{e0}}$ : @Input --> Value@.
@@
 % By definition of the interpreter, they produce the same result when applied with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
-The first proposed approach to realize Futamura projections is partial
+The first proposed approach to realizing Futamura projections is partial
 evaluation \cite{DBLP:books/daglib/0072559}, which relies on an automatic
@@
 representation can be viewed as a \textit{two-level semantics}
-\cite{NIELSON1989117, NIELSON198859} that distinguishes compile-time
+\cite{NIELSON1989117, NIELSON198859} that distinguishes between compile-time
 computation and run-time computation.
@@
 program. As an alternative approach, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users to explicitly program over
-a two-level syntax \cite{Nielson:1992:TFL:130665}, by manually
-annotating the binding-times of variables. Then, the MSP system
-will check whether these annotations are consistent, and specialize the program
-using that information. The staging annotations can be either syntactic (e.g.,
-quote and quasi-quote in MetaML/MetaOCaml) or type-based (e.g., the Lightweight
-Modular Staging framework \cite{DBLP:conf/gpce/RompfO10} in Scala).
+\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users to explicitly program
+over a two-level syntax \cite{Nielson:1992:TFL:130665} by manually annotating
+the binding-times of variables. The MSP system will then check whether these
+annotations are consistent, and specialize the program using that information.
+The staging annotations can be either syntactic (e.g., quote and quasi-quote in
+MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging framework
+\cite{DBLP:conf/gpce/RompfO10} in Scala).
 
-To apply Futamura projections on abstract interpreters, we consider
-\citet{DBLP:journals/pacmpl/DaraisLNH17}'s big-step monadic abstract
-interpreter as the unstaged baseline.  Similar to the two-level semantics
-\cite{NIELSON1989117}, monads provide a good abstraction to hide the detail of
-abstract interpretation semantics. However, monads do not have a clear
-stage distinction.
-We further introduce binding-times and make the monads binding-time polymorphic.
-Our proposed framework adopts the type-based multi-stage programming from the
-Lightweight Modular Staging framework, and implements the Futamura projection
-of an abstract interpreter for a stateful higher-order
-language. By deriving staged monads that can be used to perform code generation, we
-obtain a staged abstract interpreter that can be used for specialization. Through
-staging, the generated code is specialized to the input program, and all the
-monadic operations are inlined and compiled down to low-level Scala code.
+To apply the Futamura projections to abstract interpreters, we consider
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s big-step monadic abstract interpreter
+as the unstaged baseline.  Similar to two-level semantics \cite{NIELSON1989117},
+monads provide a good abstraction to hide the detail of abstract interpretation
+semantics. However, monads do not have a clear stage distinction. We further
+introduce binding-times and make the monads binding-time polymorphic. Our
+proposed framework adopts the type-based multi-stage programming from the
+Lightweight Modular Staging framework, and implements the Futamura projection of
+an abstract interpreter for a stateful higher-order language. By deriving staged
+monads that can be used to perform code generation, we obtain a staged abstract
+interpreter that can be used for specialization. Through staging, the generated
+code is specialized to the input program, and all the monadic operations are
+inlined and compiled down to low-level Scala code.
 

### 2019-09-08 paper/10_intro.tex
minor

@@
 
-Although the four artifacts may look dissimilar at first glance, they in
-fact are all firmly rooted in the concrete semantics.  This
-observation provides a way to abstract over the interpreter and achieve the
-flexibility of reinterpreting the shared interpreter.  The generic interpreter
-returns a value of monadic type, which can be varied by different semantics.
-The value domains of the interpreter and the effects such as state and nondeterminism
-can all be wrapped into this monadic type.  The binding-time abstraction is
-represented by a higher-kinded type, and injected into the monadic type.  The
-binding-times control whether the interpreter produces values directly or
-generates code.
+Although the four artifacts may look dissimilar at first glance, they are in
+fact all firmly rooted in the concrete semantics.  This observation provides a
+way to abstract over the interpreter and achieve the flexibility of
+reinterpreting the shared interpreter.  The generic interpreter returns a value
+of monadic type, which vary based on semantics. The value domains
+of the interpreter and the effects such as state and nondeterminism can all be
+wrapped into this monadic type.  The binding-time abstraction is represented by
+a higher-kinded type and injected into the monadic type.  The binding-times
+control whether the interpreter produces values directly or generates code.
 

### 2019-09-08 paper/20_prelim.tex
Begone spaces
and -> or

@@
 
-In this section, we first describe the abstract syntax of the language for our interpreters. 
-Then we present the generic interpreter shared among the four different
-semantics. After which, we instantiate the interpreter to the concrete one.
-It is worth noting that we choose to use Scala and monadic style to
-demonstrate the idea, but the approach is not restricted to this
-choice. One can use imperative or direct style in other MSP languages
-(e.g., MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14}
-and Template Haskell \cite{Sheard:2002:TMH:636517.636528}) to construct
-such staged abstract interpreters.
+In this section, we first describe the abstract syntax of the language for our
+interpreters. Then, we present the generic interpreter shared among the four
+different semantics, after which we instantiate the interpreter to the concrete
+one. It is worth noting that we choose to use Scala and a monadic style to
+demonstrate this idea, though the approach is not restricted to this choice.
+Indeed, one can use imperative or direct style in other MSP languages (e.g.,
+MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} or
+Template Haskell \cite{Sheard:2002:TMH:636517.636528}) to construct such staged
+abstract interpreters.
 

### 2019-09-08 paper/20_prelim.tex
typo

@@
 
-We consider a call-by-value $\lambda$-calculus in direct-style, extended
-with numbers, arithmetic, recursion, and conditionals. Other effects
-such as assignments can also be supported readily.
-Since we are mostly interested in analyzing dynamic behavior of
-programs, we elide the static semantics. We also assume that
-input programs are well-typed and all variables are distinct. The abstract
-syntax is as follows:
+We consider a call-by-value $\lambda$-calculus in direct-style, extended with
+numbers, arithmetic, recursion, and conditionals. Other effects such as
+assignments can also be supported readily. Since we are mostly interested in
+analyzing dynamic behavior of programs, we elide the static semantics. We also
+assume that input programs are well-typed and all variables are distinct. The
+abstract syntax is as follows:
 \begin{lstlisting}
@@
 
-The abstract syntax we present can be seen as a deep embedding of the
-language -- we use data-types to represent programs. This design choice allows us
-to use different interpretations over the AST; with the inheritance and
-overriding mechanisms in Scala, we may also add new language constructs and reuse
-existing interpretations.
+The abstract syntax we present can be seen as a deep embedding of the language:
+we use data-types to represent programs. This design choice allows us to use
+different interpretations over the AST; with the inheritance and overriding
+mechanisms in Scala, we may also add new language constructs and reuse existing
+interpretations.
 
@@
 @bind@ unwraps a monadic value and transforms it into a new monadic value.
-Pragmatically in Scala, we define the monad type class using trait @Monad@ (Figure
-\ref{fig:monad}), which it declares the @pure@ \footnote{We elect to use
+Pragmatically in Scala, we define the monad type class using trait @Monad@
+(Figure \ref{fig:monad}), which declares the @pure@ \footnote{We elect to use
 \texttt{pure} as the name, since \texttt{return} is a keyword in Scala and
-\texttt{unit} is a built-in function in LMS.} and @flatMap@ operation. The trait
-itself takes the monad type constructor @M[_]@ as an argument, which, as a
-higher-kinded type, takes a type and returns a type. The method @pure@
-promotes values of type @A@ to values of type @M[A]@. The monadic @bind@
-operation is usually called @flatMap@ in Scala. It takes a
-monad-encapsulated value of type @M[A]@, a function of @A => M[B]@ and returns
-values of type @M[B]@.
+\texttt{unit} is a built-in function in LMS.} and @flatMap@ operations. The
+trait itself takes the monad type constructor @M[_]@ as an argument, which, as a
+higher-kinded type, takes a type and returns a type. The method @pure@ promotes
+values of type @A@ to values of type @M[A]@. The monadic @bind@ operation is
+usually called @flatMap@ in Scala. It takes a monad-encapsulated value of type
+@M[A]@, a function of @A => M[B]@, and returns values of type @M[B]@.
 

### 2019-09-08 paper/20_prelim.tex
space
Typos

@@
   \centering
+  \vspace{-1em}
   \begin{subfigure}[b]{0.55\textwidth}
@@
 
-Note that here, the monadic object @List[_]@ encapsulates the data internally.
-Therefore it only exposes the simplified version of @flatMap@, where the monadic
+Note that here the monadic object @List[_]@ encapsulates the data internally.
+Therefore, it only exposes the simplified version of @flatMap@, where the monadic
 value of @M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
-\ref{fig:monadops}) defines the simplified version of monadic operations that
+\ref{fig:monadops}) defines the simplified versions of monadic operations that
 are necessary for @for@-comprehension. The conversion between @Monad@ and
@@
 In the rest of the paper, we use Scala's @for@-comprehension syntax and a few
-of monads and monad transformers such as @ReaderT@, @StateT@, and @SetT@ to
+monads and monad transformers such as @ReaderT@, @StateT@, and @SetT@ to
 write our interpreters.
 Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
-take a monad as an argument and produces another monad. By using monad
+take a monad as an argument and produce another monad. By using monad
 transformers, we can combine multiple monads into a single one.  The
@@
 In this section, we present the generic interface in the style of a big-step
-definitional interpreter. The key idea is to keep the binding-time type and
-returned monadic type both abstract, so that they can be instantiated
+definitional interpreter. The key idea is to keep both the binding-time type and
+returned monadic type abstract, so that they can be instantiated
 differently; we also need to abstract the primitive operations on those types.
@@
 
-\paragraph{Basic Types}
-We start with some basic type definitions that are used in the interpreter. The
-identifiers in the program are represented by strings. To represent states, two
-required components for the interpreter are environments @Env@ and stores
-@Store@. @Env@ maps identifiers to addresses, and @Store@ maps
-addresses to values, respectively. The @Env@ captures bound variables in the
-scope of current control, and @Store@ models a persistent heap through the
-program run-time.  At this moment, the domains of addresses and values are
-still abstract, and hence, they are declared as abstract types.
+\paragraph{Basic Types} We start with some basic type definitions that are used
+in the interpreter. The identifiers in the program are represented by strings.
+To represent states, two required components for the interpreter are
+environments, represented by type @Env@, and stores, represented by type
+@Store@. An @Env@ maps identifiers to addresses, and a @Store@ maps addresses to
+values. The @Env@ captures bound variables in the scope of current control, and
+@Store@ models a persistent heap through the program run-time.  Currently,
+the domains of addresses and values are still abstract; hence, they are
+declared as abstract types.
 \begin{lstlisting}
@@
   \centering
+\vspace{-1em}
   \begin{subfigure}[b]{0.45\textwidth}

### 2019-09-08 paper/20_prelim.tex
typos

@@
 \paragraph{Monadic Operations} We define the return type of the interpreter
-@Ans@, as a monadic type @AnsM[_]@ wrapping the type @Value@.
-As mentioned in Section ~\ref{monadscala}, to use the @for@-comprehension
-syntax, certain operations have to be added on the type @AnsM@. Here, we use a
-structural type @MonadOps@ to require that type @AnsM@ should at least
+@Ans@ as a monadic type @AnsM[_]@ wrapping the type @Value@.
+As mentioned in Section~\ref{monadscala}, in order to use the @for@-comprehension
+syntax certain operations must be added on the type @AnsM@. Here, we use a
+structural type @MonadOps@ to require that type @AnsM@ must (at least)
 implement @map@ and @flatMap@. It is worth noting that @MonadOps@ takes another
@@
 @ρ@ when evaluating the monadic value @ans@; @set_store@ takes a pair of
-addresses and values (potentially staged) and updates the store accordingly.
+addresses and (potentially staged) values, and updates the store accordingly.
 
@@
 \paragraph{Primitive Operations} Next, we define a few primitive operations.
-First, we delcare two versions of @alloc@. The first one takes a store and an
+First, we declare two versions of @alloc@. The first takes a store and an
 identifier and produces a fresh address of non-monadic type @R[Addr]@. Since
@@
 consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
-of the addresses, and therefore can be used in monadic computations.
-An auxiliary method @get@ retrieves the value of an identifier @x@, through the
+of the addresses, and can therefore be used in monadic computations.
+An auxiliary method @get@ retrieves the value of an identifier @x@ through the
 environment and store.
@@
 and an argument value. Note that the @Env@, @Store@, and @Value@ are all
-annotated by @R[_]@, because they are potentially next-stage values when the
+annotated by @R[_]@, as they are potentially next-stage values when the
 interpreter acts as a compiler.

### 2019-09-08 paper/20_prelim.tex
Typo

@@
 
-\paragraph{The Interpreter} Now we can construct the semantics-agnostic interpreter
+\paragraph{The Interpreter} We can now construct the semantics-agnostic interpreter
 in monadic form, shown in Figure \ref{fig:shared_int}. The basic idea
-of a generic interpretation is to traverse the abstract syntax tree
-while maintaining the effects such as environment and state updates.
+of generic interpretation is to traverse the abstract syntax tree
+while maintaining the effects (such as environment and state updates).
 Note that the interpreter is written in open-recursive style -- it cannot

### 2019-09-08 paper/20_prelim.tex
-- -> :
Typo

@@
 while maintaining the effects (such as environment and state updates).
-Note that the interpreter is written in open-recursive style -- it cannot
+Note that the interpreter is written in open-recursive style: it cannot
 refer to itself directly. Instead, @eval@ takes an additional parameter @ev@ of
@@
 Finally, there is a top-level wrapper @run@. The return type @Result@
-depends on the kind of monads being used; and therefore is also left abstract.
+depends on the kind of monads being used and is therefore also left abstract.
 \begin{lstlisting}

### 2019-09-08 paper/30_stageconc.tex
polish

@@
 presented in Section \ref{prelim}, we show how to stage the concrete
-interpreter by changing the monad type and refactoring primitive
+interpreter by changing the monad type and refactoring the primitive
 operations. We begin by briefly introducing the Lightweight Modular Staging
@@
 
-Lightweight modular staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
+Lightweight Modular Staging (LMS) \cite{DBLP:conf/gpce/RompfO10} is a
 multi-stage programming framework implemented as a Scala library. LMS enables
@@
 MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} which
-use syntactic quotations, LMS distinguishes binding-time based on types. 
-LMS provides a type constructor @Rep[T]@ where @T@ can be
-an arbitrary type, indicating a value of type @T@ will be known at the future stage.
+uses syntactic quotations, LMS distinguishes binding-times based on types. 
+LMS provides a type constructor @Rep[T]@, where @T@ can be an arbitrary type,
+indicating a value of type @T@ will be known at the future stage.
 Operations acting on @Rep[T]@ expressions will be residualized as generated

### 2019-09-08 paper/60_cases.tex
space
revision
polish

@@
 In Section~\ref{sai}, we used a toy language to show that staging an abstract
-interpreter is feasible and can be systematically derived from staged concrete
-interpreters and unstaged abstract interpreters.
+interpreter is feasible and the artifact can be derived by combining the staged
+concrete interpreter and unstaged abstract interpreter.
 In this section, we conduct several case studies to further show that this
@@
 
-In this section, we show AC (closure generation) can be understood and
+In this section, we show AC can be understood and
 implemented as an instance of staging abstract interpreters. We first revisit
@@
   \caption{Comparison of AC (left) and SAI (right). Only core code are shown.}
+  \vspace{-1.5em}
   \label{compare_ac_sai}
@@
 \begin{figure}[h!]
+\vspace{-1em}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\vspace{-0.5em}
 \caption{\texttt{fix\_gc} for abstract garbage collection} \label{fixgc}
+\vspace{-2em}
 \end{figure}

### 2019-09-08 paper/80_related.tex
misc
polish

@@
 The unstaged abstract interpreter in this paper follows the abstracting
-definitional interpreters (ADI) approach from
-\citet{DBLP:journals/pacmpl/DaraisLNH17}.  Similar to ADI, reconstructing
+definitional interpreters (ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17}.  
+Similar to ADI, reconstructing
 big-step abstract interpreters with delimited continuations was shown by
@@
 \citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis, where we borrow the notation of modular analysis.
+modular analysis, from where we borrow the notation of modular analysis.
 However, the staged abstract interpreter does not generate complete summaries for
-modules; instead, it is able to specialize the analysis modularly. Similarly, AC
+modules; instead, it specializes the analysis modularly. Similarly, AC
 is considered as an example of this kind of modular analyses \cite{DBLP:conf/cc/CousotC02}.
@@
 semantics, this paper shows how a staged analyzer can be built and
-used to increase efficiency of static analysis. We augment the monadic
-abstract interpreter from another dimension, using code generation to
-produce efficient low-level code. Hence the presented work can be
-considered as a two-dimensional application of the two-level semantics.
+used to increase efficiency of static analyses. We augment the monadic
+abstract interpreter with binding-time polymorphism, using code generation to
+produce efficient low-level code. The presented work can be considered as
+a two-dimensional application of the two-level semantics.
 
@@
 guarantee for preserving evaluation orders within a stage.
-\citet{Ofenbeck:2017:SGP:3136040.3136060} discussed staging for generic
-programming as well as stage polymorphism in LMS.
+\citet{Ofenbeck:2017:SGP:3136040.3136060} and
+\citet{Amin:2017:CTI:3177123.3158140} discussed staging for generic programming
+as well as stage polymorphism in LMS.
 %The idea of staging an abstract interpreter we presented here

### 2019-09-08 paper/60_cases.tex
polish

@@
 primitive operators. Figure \ref{compare_ac_sai} (left) shows the AC
-implementation: different syntactic constructs are handled in different functions
+implementation: different syntactic constructs are handled by different functions
 (e.g., @compCall@ and @compApp@).
-Before refactoring to AC, the original functions take a store as a
-regular argument. Now every function returns a value of type @CompAnalysis@,
-which is a function value (i.e., closure) that takes and returns stores. After the
-first run of the analysis, we have traversed the AST and obtained a single closure that
-takes a store and returns a store. Therefore, the ASTs are static terms, and the
-stores are dynamic terms. Now, we can invoke the compiled closure with an initial
-store to complete the analysis.
+All original functions take a store as regular argument. After refactoring to
+AC, every function returns a value of type @CompAnalysis@, which is a function
+value that takes and returns stores. After the first run of the
+analysis, we have traversed the AST and obtained a single closure that takes a
+store and returns a store. Therefore, the ASTs are treated as static terms, and
+the stores are dynamic terms. Finally, we may invoke the compiled closure with an
+initial store to complete the analysis.
 

### 2019-09-08 paper/70_eval.tex
space
polish

@@
 To evaluate how performance can be improved by our staged abstract interpreters
-approach, we scale our toy interpreter to a large subset of the Scheme language
-and evaluate the analysis time against the unstaged versions. The result
-shows that the staged versions are, on average, 10+ times as fast as the unstaged
-version, as evidence of our \textit{abstraction without regret} approach.
+approach, we scale our toy interpreter to support a large subset of the Scheme
+language and evaluate the analysis time against the unstaged versions. The
+result shows that the staged versions are, on average, 10+ times as fast as the
+unstaged version, as evidence of our \textit{abstraction without regret}
+approach.
 
+\vspace{-5pt}
 \paragraph{Implementation and Evaluation Environment} We implement the abstract
@@
 
-\vspace{-1em}
-\begin{figure}[h] \footnotesize
+\begin{figure}[t] \footnotesize
 \begin{tabular}{@{}ll|lll|lll@{}}
@@
 \end{tabular}
+\vspace{-0.5em}
 \caption{Evaluation result for monovariant control-flow analysis.} \label{evaluation_result}
+\vspace{-2em}
 \end{figure}

### 2019-09-08 paper/40_absinterp.tex
typo
revision
typos

@@
 
-In this section, we simply instantiate @SPLattice@ with the flat binding-time type
-@R[T] = T@ (i.e., @NoRep@).
+In this section, we simply instantiate @SPLattice@ with the flat binding-time
+type @R[T] = T@ (i.e., @NoRep@).
 An example of lattices we used in the rest of the paper is the power set
@@
 values, indicating that an address may point to an over-approximated set of
-runtime values.  This can be justified by the approximation nature of abstract
-interpretation and nondeterminism in the abstract interpreter.  First, when
+runtime values.  This can be justified by the approximation nature
+and nondeterminism in the abstract interpreter.  First, when
 analyzing a conditional expression, we may not have enough information to
@@
 that type @Value@ is an alias of @Set[AbsValue]@), so we can not directly
-destruct it to syntactic @Lam@ by pattern matching. Instead, we use an
+destruct it to a syntactic @Lam@ by pattern matching. Instead, we use an
 auxiliary function @lift_nd@ that takes a set and lifts the elements in the set
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \paragraph{Caching and Fixpoint Iteration}
@@
 \textit{truncated depth-first evaluation} \cite{Rosendahl:AbsIntPL}. It has
-been used in other abstract interpreter or fixed point computation
+been used in other abstract interpreters or fixed-point computation
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
@@
 input @in@ cache, which indicates that there is no more fact can be discovered.
-Therefore, the iteration should end, and we have reached the least fixed point.
+Therefore, the iteration should end, and we have reached the least fixed-point.
+\begin{lstlisting}
+  def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0)(cache0)(cache0).run
+\end{lstlisting}
 
-\begin{figure}[h!]
+Finally, we override the top-level @run@ method by running the monadic value
+with an empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
+and initial caches @cache@$_0$.
+\begin{figure}[t]
   \centering
@@
 \end{figure}
-
-\begin{lstlisting}
-  def run(e: Expr): Result = fix(eval)(e)(ρ0)(σ0)(cache0)(cache0).run
-\end{lstlisting}
-
-Finally, we override the top-level @run@ method by running the monadic value
-with an empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
-and initial caches @cache@$_0$.

### 2019-09-08 paper/20_prelim.tex
typos

@@
 To distinguish from the monadic values of type @Ans@ produced by the
-interpreter, later we use \textit{grounded values} to denote such final values.
+interpreter, later we use term \textit{grounded values} to denote such final values.
 The two elements in the type @Result@ are annotated by the binding-time @R@,

### 2019-09-09 paper/20_prelim.tex
typo

@@
 types @Env@ and @Store@ are derived automatically. To ensure the freshness of
-address allocations, we use type @Int@ and always return a number greater
+address allocations, we use type @Int@ and always return a number that is greater
 than the size of the current store.
@@
 The two elements in the type @Result@ are annotated by the binding-time @R@,
-because they cane be next-stage objects.  We also define a standard fixed-point
+because they can be next-stage objects.  We also define a standard fixed-point
 combinator to close the open-recursive function @ev@.

### 2019-09-09 paper/20_prelim.tex
polish

@@
 \paragraph{Unstaged Monads}
-For concrete interpretation, the monad needs to model reader and state
-effects, which correspond to the environment and the store, respectively. Following
-the monad transformer approach \cite{DBLP:conf/popl/LiangHJ95}, @ReaderT@ and
-@StateT@ monad transformers are used to compose the monad stack. In other words,
-the type @AnsM@ is instantiated by layering @ReaderT@ and @StateT@
-transformers\footnote{The question mark syntax is a kind projector
-\cite{kindprojector}, thus \texttt{StateT[IdM,Store,?]} is equivalent to
-\newline \texttt{(\{type M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@
-is parameterized by the type @Env@, and @StateT@ is parameterized by the type @Store@,
-and the inner-most monad @IdM@ is merely an identity monad.
+For concrete interpretation, the monad needs to model reader and state effects,
+which correspond to the environment and the store, respectively. We follow the
+monad transformer approach \cite{DBLP:conf/popl/LiangHJ95},
+and use the @ReaderT@ and @StateT@ monad transformers to
+compose the monad stack. In other words, the type @AnsM@ is
+instantiated by layering the @ReaderT@ and @StateT@ transformers\footnote{The
+question mark syntax is a kind projector \cite{kindprojector}, such that
+\texttt{StateT[IdM,Store,?]} is equivalent to \newline \texttt{(\{type
+M[T]=StateT[IdM,Store,T]\})\#M}}, where the @ReaderT@ is parameterized by the
+type @Env@, and the @StateT@ is parameterized by the type @Store@, and the
+inner-most monad @IdM@ is merely an identity monad.
 \begin{lstlisting}
@@
 
-Here we sketch the basic idea of @ReaderT@ and @StateT@. Readers may
+Here, we sketch the basic idea of @ReaderT@ and @StateT@. Readers may
 refer to \cite{DBLP:conf/popl/LiangHJ95, Chiusano:2014:FPS:2688794} for more detail.
 A @ReaderT@ monad transformer encapsulates computation @R => M[A]@, where
-@R@ is the reader type, and @M[_]@ is the inner monadic type.
+@R@ is the environment type, and @M[_]@ is the inner monadic type.
 Given a value of @R@, a @ReaderT@ monad produces a transformed value of type @M[A]@.
@@
 Given a value of @S@, a @StateT@ monad produces a transformed value of type @M[(A, S)]@,
-where the new state (of type @S@) is accompanied with the value (of type @A@).
+where the new state (of type @S@) is accompanied with the result (of type @A@).
 Note that for the moment, the binding-time type @R@ is an identity type; thus,
@@
 top-level of our monad stack. To modify the store, for example, we construct a
-@StateT@ value that transforms the current store $\sigma$ to @σ + αv@, which
-results in a @StateT[IdM, Store, Unit]@, and then lift this @StateT@ value to the
-top-level @ReaderT@ type, i.e., @AnsM[Unit]@.
+@StateT@ value that updates the current store $\sigma$ to @σ + αv@, which
+results in a @StateT[IdM, Store, Unit]@ value. Then, we lift this @StateT@
+value to the top-level @ReaderT@ type, i.e., @AnsM[Unit]@.
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations}
-Other primitive operations over the value domain can be implemented
+Other primitive operations over the value domains can be implemented
 straightforwardly. We elide most of them but describe a bit on how we handle
@@
 part of the control flow. The way we treat them is simple now, but
-will become more involved when staging the abstract interpreter.
-At the moment, the method @close@ takes a lambda term and an environment, and
+will become more involved when we stage the abstract interpreter.
+The method @close@ denotes a lambda term to our representation of functions.
+At the moment, @close@ takes a lambda term and an environment, and
 produces the defunctionalized representation of closures.
+The method @ap_clo@ performs function applications, which takes a function
+value @fun@ and an argument @arg@. It first extracts the lambda term and environment
+enclosed in @fun@, and allocates an addresses for the argument and updates the
+environment and store accordingly.  Finally, body expression of the lambda term
+is evaluated under the new environment and store.
 \begin{lstlisting}
   def close(ev: Expr => Ans)(λ: Lam, ρ: Env): Value = CloV(λ, ρ)
-\end{lstlisting}
-
-When applying a function, @ap_clo@ takes a function value @fun@ and an argument
-@arg@. It extracts the lambda term and environment enclosed in @fun@, and
-allocates the addresses and updates the environment and store accordingly.
-Finally, body expression of the lambda term is evaluated under the new
-environment and store.
-\begin{lstlisting}
   def ap_clo(ev: Expr => Ans)(fun: Value, arg: Value): Ans = fun match {
@@
 
-Given the expression @e@, initial environment $\rho_0$ and store $\sigma_0$,
-the top-level @run@ method invokes the fixed-point operator directly.
+We define the top-level @run@ method as invoking the fixed-point operator with
+the evaluator and input expression @e@, and running the monad stack with the
+initial environment $\rho_0$ and store $\sigma_0$.
 \begin{lstlisting}

### 2019-09-09 paper/30_stageconc.tex
polish

@@
 operations. We begin by briefly introducing the Lightweight Modular Staging
-framework in Scala, and then replay the same steps from the unstaged
+framework in Scala, and then replay the same steps as we did for the unstaged
 counterpart. Finally, we briefly describe the code generation.
@@
 uses syntactic quotations, LMS distinguishes binding-times based on types. 
-LMS provides a type constructor @Rep[T]@, where @T@ can be an arbitrary type,
-indicating a value of type @T@ will be known at the future stage.
-Operations acting on @Rep[T]@ expressions will be residualized as generated
-code.
+To achieve this, LMS provides a type constructor @Rep[T]@, where @T@ can be an
+arbitrary type, indicating that a value of type @T@ will be known at a future
+stage.  Operations acting on @Rep[T]@ expressions will be residualized as
+generated code.
 A classic example for introducing multi-stage programming is the power function
@@
 The LMS framework provides staging support for primitive data types such as
-@Int@ and @Double@, and commonly-used data structures such as lists and maps.
-The framework internally constructs a sea-of-nodes intermediate representation (IR)
-for the next-stage program when executing the current stage \cite{DBLP:conf/birthday/Rompf16}.
+@Int@ and @Double@, and for commonly-used data structures such as lists and
+maps.  The framework internally constructs a sea-of-nodes intermediate
+representation (IR) for the next-stage program when executing the current stage
+program \cite{DBLP:conf/birthday/Rompf16}.
 For convenience, the conversion from expressions of type @Rep[_]@ to their IR
@@
 We use the same monad stack structure as in the unstaged interpreter: a reader
-monad with a state monad. The difference here is that, now the monads operate on
-staged values.  For brevity, we call them \textit{staged monads}. In the code,
-we also use a @Rep@ prefix on the constructors and types to differentiate them.
-But it is important to note that the instances of @ReaderT@/@StateT@
-are not staged, and that the monadic computation, such as functions of @R => M[A]@,
-are also not staged. Instead, the internal data that these monads operate on are
-staged, i.e., @R@ and @A@ in the reader monad, @S@ and @A@ in the state monad.
+monad with a state monad. The difference here is that, now the monads operate
+on staged values.  For brevity, we call them \textit{staged monads}. In the
+code snippet, we also use a @Rep@ prefix on the name of constructors and types
+to differentiate them.
+But it is important to note that the instances of @ReaderT@/@StateT@ are not
+staged, and that the monadic computation, such as functions of @R => M[A]@, are
+also not staged. Instead, the internal data that these monads operate on are
+staged, i.e., values of @R@ and @A@ in the reader monad, and values of @S@ and
+@A@ in the state monad.
 The following code snippet shows the idea. We use \hl{light gray} to highlight
-where @Rep@ types are added:
+where @Rep@ types are added.
 \begin{lstlisting}[escapechar=!]
@@
 
-The function @f@ passed to @flatMap@ is also a value known at the current stage.
-Therefore all invocations of @flatMap@ can be reduced before generating code.
-The fact that we only stage data but not monadic computation or monadic
-values is the reason that we can peel of the monad stack in the generated code.
-Now the type @AnsM[_]@ is instantiated as the same structure as before, but using the
-@Rep@ versions:
+The function @f@ passed to the @flatMap@ is also a value known at the current stage.
+Therefore, all invocations of @flatMap@ can be reduced before generating code.
+The fact that we only stage data but not the monadic computation or monadic
+values is the reason that we can peel of the monad stack in the generated code
+through staging.  Now the type @AnsM[_]@ is instantiated as the same structure
+as before, but using the @Rep@ versions of monad transformers:
 \begin{lstlisting}
@@
 staged, but the whole set is also staged, and we have no knowledge about how
-many elements in the set. In this section, we explicitly distinguish the two
-versions of monadic interfaces. Later, we will instantiate the staged set monad
-by manually fusing a fragment of the monad stack.
+many elements in the set. Therefore, any traversals over the set has to be
+staged.  In this section, we explicitly distinguish the two versions of monadic
+interfaces. Later, we will instantiate the staged set monad by manually fusing
+the fragment of the monad stack inside of @SetT@.
 
 \paragraph{Primitive Operations} Most of the primitive operations can be easily
-translated to their staged versions -- we just need to change the types.
-As we mentioned before, we will take more care on how functions and
-applications are handled.  One problem here is what should we do for
-$\lambda$ terms when staging? We cannot directly create a next-stage
-defunctionalized closure for it, because that means we still need to
-interpret over these closure at the next stage.  The desired goal is to
-eliminate the interpretation overhead by staging, so the right thing is to
-compile the lambda term with its environment to a next-stage Scala function. To
-do this, we need to recursively call the interpreter function at the current stage,
-i.e., the @ev@ argument passed to @close@.  The following code implements the
-idea:
+translated to their staged versions -- we just need to change the types.  As we
+mentioned before, we will illustrate in detail how functions and applications
+are handled.  One problem here is what should we do for $\lambda$ terms when
+staging them? We cannot directly create a next-stage defunctionalized closure
+for it, because that means we still need to interpret over these closures at the
+next stage and the interpretation overhead has not been eliminated. 
+The solution is to compile the $\lambda$ term with its environment by calling
+the evaluator at the current stage. The result is generating a next-stage
+Scala function. The evaluator is passed as @ev@ to the method @close@.
+The following code implements the idea:
 \begin{lstlisting}
@@
 
-In @close@, we first create a current-stage function @f@ that takes a
-next-stage value and store, and returns staged values. Then we use
-@emit_complied_clo@ to denote @f@ to a next-stage Scala function, represented
-by the type @Rep[Value]@. Inside of function @f@, we can access the evaluator via @ev@.
+In @close@, we first create a function @f@ that takes a next-stage value and
+store, and returns staged values and stores. The function @f@ is a
+current-stage and we then use @emit_complied_clo@ to denote it to a next-stage
+Scala function.
+The method @emit_compiled_clo@ produces current-stage representations of
+next-stage function values, i.e., of type @Rep[Value]@.
+Inside of function @f@, we can access the evaluator via @ev@.
 However, @ev@ produces a monadic value of type @AnsM@, which can only exist at
@@
 grounded values, the evaluator @ev@ is supplied with not only the body
-expression @e@ to be specialized, but also the prepared environment and store.
-By doing so, we \textit{collapse} the monadic value of @AnsM@ to grounded
-values within staging; therefore in the future stage, there is no monads.
-
-For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
-and @arg@. But what we can do with these values of type @Rep@ is limited. 
-In fact, @fun@ does not have an intensional application operation we can use
-directly. We only know that it is a staged value produced by @close@, and
-it represents a next-stage Scala function. Therefore, we can only generate a
-next-stage function application, also in Scala. @ap_clo@ produces a monadic
-value of type @Ans@, so that we can still use the monadic operation @get_store@
-to obtain the latest store. Then @emit_ap_clo@ takes the function and argument
-values, and the latest store to generate a current-stage representation of the
-future-stage application result, which is of type @Rep[ValSt]@, a pair of
-values and stores.  After obtaining the result @vs@, we use @put_store@ to
-update the store and reify the result into the current-stage monadic value.
+expression @e@ to be specialized, but also with the prepared environment and
+store.  By doing so, we have \textit{collapsed} the monadic values of @AnsM@ to
+grounded values within staging; therefore in the future stage, there is no
+monads.
 \begin{lstlisting}
@@
 
-Compared with the unstaged version, we can observe a shift of where we use the
-evaluation function @ev@: in the unstaged interpreter, the invocation of @ev@
-happens in @ap_clo@, i.e., at the application time; while in the staged
-interpreter, we eagerly apply @ev@ when denoting the lambda terms to our
-next-stage value domains, i.e., at the value-representation time.
+For function applications @ap_clo@, we have two \textit{next-stage} values @fun@
+and @arg@. But what we can do with these values of type @Rep@ is limited. 
+In fact, @fun@ does not have an intensional application operation we can use
+directly. We only know that it is a current-stage representation of functions produced by @close@,
+whose next-stage form is a Scala function.
+With this fact, we can generate a next-stage function application for it, also in
+Scala. The method @emit_ap_clo@ produces a current-stage representation
+of a next-stage function application, with necessary arguments such as the latest
+store @σ@. Meanwhile, we still want the method @ap_clo@ implemented in monadic
+style, so that it can be smoothly connected with other parts of the
+interpreter.  To achieve this, we lift the result values and stores from
+the future stage into our current-stage monad stack, with a binding @vs@.
+Finally, we use @put_store@ to update the store with @vs._2@ and return the
+value result @vs._1@ into the current-stage monadic value.
+
+Now, compared with the unstaged interpreter, we observe a key difference of where
+we use the evaluation function @ev@.  In the unstaged interpreter, the
+invocation of @ev@ happens in @ap_clo@, i.e., at the application time; while in
+the staged interpreter, we eagerly apply @ev@ when denoting the lambda terms to
+our next-stage value domains, i.e., at the value-representation time.
 

### 2019-09-09 paper/30_stageconc.tex
polish

@@
 To implement the staged concrete interpreter, we replay the steps from the
-instantiation of the unstaged concrete interpreter in Section \ref{unstaged_conc}.
+instantiation of the unstaged counterpart in Section \ref{unstaged_conc}.
 However, now we use the @Rep@ type to annotate the value domains, environments and

### 2019-09-09 paper/40_absinterp.tex
polish

@@
 operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines a type
-class of stage polymorphic lattices. Similar to @MonadOps@ in Section
-\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to
-annotate the binding-times, where @R[_]@ wraps the data type @S@ and @Boolean@
-in these operations.
+class of stage polymorphic lattices. Similar to the @MonadOps@ in Section
+\ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to the
+trait for annotating the binding-times, where @R[_]@ wraps the data type
+@S@ and @Boolean@ in these operations.
 
@@
 In this section, we simply instantiate @SPLattice@ with the flat binding-time
-type @R[T] = T@ (i.e., @NoRep@).
-An example of lattices we used in the rest of the paper is the power set
-abstract domain (shown in Figure \ref{fig:powerset}): the ordering is the subset
-relation, the join is the union operation, and the meet is the intersect
-operation. The bottom element of a power set is empty set, and we do not have a
-representation of top element for power set.  Other lattices used in the paper, such as
-products and maps, can be implemented similarly or by lifting the existing
-lattices element-wise or point-wise. Non-relational numerical abstract domains
-such intervals can also be implemented in a stage polymorphic way.
+type @R[T] = T@ (i.e., @NoRep@); the result is the trait @Lattice[S]@.
+An example of the lattices we used in the rest of the paper is the power set
+abstract domain (shown in Figure \ref{fig:powerset}). For two power sets,
+they are ordered by the subset relation. We use set union to compute their
+join, and set intersection to compute their meet.  The bottom element of a
+power set is empty set, and we do not have a representation of top element for
+power set.  Other lattices used in the paper, such as products and maps, can be
+implemented similarly or by lifting the existing lattices element-wise or
+point-wise.  Non-relational numerical abstract domains such intervals can also
+be implemented in a stage polymorphic way.
 
@@
 
-In this section, we follow the \textit{abstracting definitional interpreters}
-(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our concrete
-interpreter to an abstract interpreter.
+In this section, we follow the abstracting definitional interpreters
+(ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our monadic
+concrete interpreter to an monadic abstract interpreter.
 
@@
 
-We first widen the concrete values to @AbsValue@. There are two variants of
-@AbsValue@: numbers are lifted to a singleton object @IntTop@, which represents
-the set of all numbers; closures remain the same. Then the type @Value@ is
-defined as a set of @AbsValue@.  The address space is constrained to be finite
-to ensure the computability of analysis -- for a simple monovariant analysis,
-we directly use variable names for addresses. After defining the @Value@ and
-@Addr@ type, the types of environments and stores are automatically lifted to
-abstract environments and abstract stores, i.e., @Map[Ident, Addr]@ and
-@Map[Addr, Set[AbsValue]]@, respectively.  
+We first widen the concrete value domain to @AbsValue@. There are two variants of
+@AbsValue@: 1) numbers are lifted to a singleton object @IntTop@, which represents
+the set of all numbers; 2) closures remain the same. Then, the type @Value@ is
+redefined as a set of @AbsValue@ to account for approximation.  The address
+space is constrained to be finite to ensure the computability of the analysis --
+for a simple monovariant analysis, we directly use variable names for
+addresses. After defining the @Value@ and @Addr@ type, the types of
+environments and stores are automatically lifted to their abstract versions,
+i.e., @Map[Ident, Addr]@ and @Map[Addr, Set[AbsValue]]@, respectively.  
 \begin{lstlisting}
@@
     type Value = Set[AbsValue]; case class Addr(x: String)
-    ...
-  }
+    ... }
 \end{lstlisting}
@@
 values, indicating that an address may point to an over-approximated set of
-runtime values.  This can be justified by the approximation nature
-and nondeterminism in the abstract interpreter.  First, when
-analyzing a conditional expression, we may not have enough information to
-decide which branch will be taken, thus a sound treatment is to explore the both
-branches. Also, at some later point, we need to compute the join of two path.
-Another nondeterminism comes from function applications, for instance, @f(a)@,
-the possible callees ($\lambda$ terms) of @f@ is computed approximately.  As a
-result, we may retrieve a set of closures when dereferencing the address of @f@
-and we need to apply all of them.
+runtime values.  This can be justified by the approximation nature and
+nondeterminism during the analysis.  For example, when analyzing a
+conditional expression, we may not have enough information to decide which
+branch will be taken, thus a sound treatment is to explore the both branches.
+Also, at some later point, we need to compute the join of two path.  Another
+nondeterminism comes from function applications, for instance, @f(a)@, the
+possible $\lambda$ terms of @f@ is computed approximately.  As a
+result, we need to store all the possible callees, as well as we will
+retrieve a set of closures when dereferencing an address.
 
@@
 represented by the @Set[M[_], A]@ monad transformer, where @M@ is the inner
-monad type being transformed, and @A@ is the type of elements in the set. We use a
-@ReaderT@ for one cache that is not changed during one fixed-point
+monad type being transformed, and @A@ is the type of elements in the set. 
+We use a @ReaderT@ for one cache that is not changed during one fixed-point
 iteration, and use a @StateT@ for another cache that will be constantly updated
-during the analysis.  \citet{DBLP:journals/pacmpl/DaraisLNH17} discuss
-different permutations of the monad stack for abstract interpretation. The
-following @AnsM@ type shows the monad stack for abstract interpretation (we use
-\hl{light gray} to highlight what have been changed from the monad stack of
-concrete interpretation):
+during the analysis.
+\citet{DBLP:journals/pacmpl/DaraisLNH17} discuss different permutations of the
+monad stack for abstract interpretation. The following @AnsM@ type shows the
+monad stack (we use \hl{light gray} to highlight what have been changed from
+the monad stack of concrete interpretation):
 \begin{lstlisting}[escapechar=!]
@@
   type AnsM[T] = ReaderT[StateT[!\hl{SetT}![!\hl{ReaderT}![!\hl{StateT}![IdM, !\hl{Cache}!, ?], !\hl{Cache}!, ?], ?], Store, ?], Env, T]
-  ...
-}
+  ... }
 \end{lstlisting}
@@
 is of type @M[SetT[A]]@, where @M[_]@ is the inner monad, and @A@ is the element
-type of the set. The @flatMap@ method takes a function that transforms each
-element of type @A@ into a @SetT[M, B]@ value, and then folds
-all the transformed values into a single monadic value of type @Set[M, B]@.
+type of the set. We first apply @flatMap@ on the inner monad to obtain the
+set @s@. Then, we use the function @f@ to transform every element of type @A@
+into a monadic value of type @Set[M, B]@. Finally, we fold all the transformed
+values into a single monadic value of type @Set[M, B]@ by appending all of them.
+The initial value of the fold is an empty @SetT@.
 \begin{lstlisting}
@@
 
-Such monad stack scheme leads to a different type of grounded result. In the concrete
-setting, we only have a reader monad (for environments) and a state monad (for
-stores). The reader effect is not persistent to the final result. Together
-with the value produced by the interpreter, the final result type is a pair of
-@Value@ and @Store@. However, under abstract interpretation, we have a @SetT@
-inside of the environment monad and store monad. The @Set@ type becomes the
-container type of the pairs of values and stores, i.e., @Set[(Value, Store)]@.
-Also, note that the reader and state monad for the caches inhibits internally
-of the nondeterminism monad; therefore, the final result type is paired
-with a cache type from the state monad.
+This monad stack scheme (@AnsM[T]@) leads to a different type of grounded
+result. Recall that in the concrete setting, we only have a reader monad (for
+environments) and a state monad (for stores). The reader effect is not
+persistent to the final result. Hence, together with the value produced by the
+interpreter, the final result type is just a pair of @Value@ and @Store@.
+However, under the new monad stack for abstract interpretation, we have a
+@SetT@ inside of the environment monad and store monad. Therefore, the type
+@Set@ becomes the container type of the pairs of values and stores, i.e.,
+@Set[(Value, Store)]@.  Also, note that the reader and state monad for the
+caches both inhibit internally of the nondeterminism monad; as the result, the
+final result type is paired the set of value-stores with the cache type from
+that state monad, as shown below.
 \begin{lstlisting}
@@
 
-\paragraph{Primitive Operations} The primitive operations are changed according to
-the new monad stack scheme and value domains. One of the notable changes is
-that the store update operator merges the new values with existing values.
-As we mentioned before, sometimes we have to explore the both branches for
+\paragraph{Primitive Operations} The primitive operations are changed according
+to the new monad stack scheme and value domains. One of the notable changes is
+that the store update operator merges the new values with existing values.  As
+we mentioned before, sometimes we have to explore the both branches for
 conditionals: in method @br0@, we combine the results using the @mplus@
@@
 
-The value representation of lambda terms is still a defunctionalized closure
-@CloV@, but we lift it to a singleton set.
+The value representation of lambda terms is still a defunctionalized closure of
+type @CloV@, but we lift it to a singleton set of @AbsValue@ (remember that
+type @Value@ is an alias of @Set[AbsValue]@).
 \begin{lstlisting}
@@
 For function applications, @ap_clo@ looks the same as the concrete interpreter.
-The difference is that now the first argument @funs@ is a set of closures (remember
-that type @Value@ is an alias of @Set[AbsValue]@), so we can not directly
-destruct it to a syntactic @Lam@ by pattern matching. Instead, we use an
-auxiliary function @lift_nd@ that takes a set and lifts the elements in the set
-into the monad stack. Then we can straightforwardly implement the @ap_clo@ in
-monadic style, where the closures come from the @SetT@ and thus the
-nondeterminism can be naturally handled. The light gray code shows what is
-added from the concrete
-@ap_clo@.
+The difference is that now the first argument of @ap_clo@ is a set of closures
+@funs@, so we can not directly destruct it to a syntactic @Lam@ by pattern
+matching. Instead, we use an auxiliary function @lift_nd@ that takes a set and
+lifts the elements in the set into the monad stack. Then we can
+straightforwardly implement the @ap_clo@ in monadic style, where the closures
+come from the monads and thus the nondeterminism can be naturally handled. The
+light gray code shows what is added from the concrete @ap_clo@.
 \begin{lstlisting}[escapechar=!]
@@
 \paragraph{Caching and Fixpoint Iteration}
-
 As we mentioned earlier, the ADI approach uses a two-cache mechanism to
 compute the least fixed-point and prevent non-termination.
-The caching algorithm is also called \textit{co-inductive} caching or
+The caching algorithm is also called a \textit{co-inductive} caching or
 \textit{truncated depth-first evaluation} \cite{Rosendahl:AbsIntPL}. It has
@@
 the depth-first evaluation. The @in@ cache stores the result from the last
-iteration, and @out@ cache is used for updating in the current iteration. In the next
-iteration, the last @out@ cache will be served as @in@ cache, and an empty cache
-is plugged-in to the @out@ slot. Once the @out@ cache does not contain anything
-new than the @in@ cache, the fixed-point is reached.
-
-In our monad stack, the @in@ and @out@ caches are modeled by reader effects and
-state effects, respectively. We first define several methods to manipulate
+iteration, and @out@ cache is used for updating in the current iteration. In
+the next iteration, the last @out@ cache will be served as the @in@ cache, and
+an empty cache is plugged-in to the @out@ slot. Once the @out@ cache does not
+contain any new information compared with the @in@ cache, the fixed-point is
+reached.
+In our monad stack, the @in@ and @out@ caches are modeled by the reader monad and
+state monad, respectively. We first define several methods to manipulate
 the two caches through the monad stack (the implementations are elided).

### 2019-09-09 paper/50_stageabs.tex
polish

@@
     type Result = (Rep[Set[(Value, Store)]], Rep[Cache])
-    ...
-  }
+    ... }
 \end{lstlisting}
-
-\iffalse
-Since our monad stack is uniformly using staged data, the staged @SetT@ now
-stores a staged value of type @Rep[Set[A]]@, inside of another staged monad @M@.
-\begin{lstlisting}
-  case class SetT[M[_]: RepMonad, A](run: M[Set[A]]) {
-    def flatMap[B: Manifest](f: Rep[A] => SetT[M, B]): SetT[M, B] = ...
-  }
-\end{lstlisting}
-\fi
 
@@
         ev(e)(ρ + (unit(x) → α))(σ ⊔ Map(α → arg))(in)(out)
-    }; Set[AbsValue](emit_compiled_clo(f, λ, ρ))
-  }
+    }; Set[AbsValue](emit_compiled_clo(f, λ, ρ)) }
 \end{lstlisting}

### 2019-09-09 paper/40_absinterp.tex
polish

@@
 The instrumentation works as follows. Initially, the two caches are both empty.
-During the iteration, we first check whether @out@ contains the configuration
-@cfg@, representing the current desired computation. If it does, then the result
-is directly returned through the monad stack.
+During the iteration, we first check whether the @out@ cache contains the
+configuration @cfg@, which represents the current desired computation. If @out@ does,
+then the result is directly returned throughout the monad stack.
 Otherwise, we first retrieve the result from @in@ (@⊥@ if @in@ not contains
-@cfg@), and update the result into the @out@ cache in the fashion of join.
-Then, we invoke @ev@ to compute the result for this iteration, where @ev@ takes
-@fix(ev)@ as a self reference.
-After the evaluation, the store @σ@ may have been changed. Therefore, 
-we get the latest store and construct the result @(v, σ)@; the
-result will be used to update the out cache.
+@cfg@), and update this result from @in@ into the @out@ cache in the fashion of
+join. Then, we invoke @ev@ to evaluate the result for this iteration, where
+@ev@ takes @fix(ev)@ as the self reference.  After the evaluation, the store @σ@
+may have been changed, so we obtain the latest store and construct the
+result @(v, σ)@, which will be used to update the @out@ cache.
 
 The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which indicates that there is no more fact can be discovered.
+input @in@ cache, which indicates that there is no more fact has been discovered.
 Therefore, the iteration should end, and we have reached the least fixed-point.
@@
 Finally, we override the top-level @run@ method by running the monadic value
-with an empty abstract environment $\rho_0$, initial abstract store $\sigma_0$,
-and initial caches @cache@$_0$.
+@fix(eval)(e)@ with the initial environment $\rho_0$, initial abstract
+store $\sigma_0$, and initial caches @cache0@; all of which are empty.
 \begin{figure}[t]

### 2019-09-09 paper/50_stageabs.tex
polish

@@
 
-The trait @RepLattice@ is shown by instantiating @Lattice@ with type @Rep@.
-The @RepSetLattice@ is an instance of @RepLattice@, where the lattice operations
-are eventually delegated to the operations on a staged set data type, such as
-@subsetOf@, @union@ and @intersect@. We do not need to change the
-implementations, except the types -- such as changing from @Set[T]@ to
-@Rep[Set[T]]@.  We show the changes in light gray code. The LMS library
-provides implicit conversions that lift the current-stage values to their
-next-stage representations. In the code generation part, these operations emit
-their corresponding next-stage code. Again, other lattice structures such as
-products and maps are implemented in a similar way.
+The trait @RepLattice@ is shown by instantiating @Lattice@ with the type @Rep@.
+The @RepSetLattice@ for power sets is an instance of @RepLattice@, where the
+lattice operations are eventually delegated to the operations on the staged set
+data type, such as @subsetOf@, @union@ and @intersect@. We do not need to
+change the implementations, except the types -- such as changing them from
+@Set[T]@ to @Rep[Set[T]]@.  We show the changes in light gray code. 
+%The LMS library provides implicit conversions that lift the current-stage values to
+%their next-stage representations. 
+In the code generation part, these operations on staged sets emit their
+corresponding next-stage code. Again, other lattice structures such as products
+and maps can be implemented in a similar way.
 
@@
 
-Now we can implement the staged abstract semantics, under that @R@ is instantiated as
-@Rep@. The types of abstract components are reused from the unstaged version.
+Now we can implement the staged abstract semantics, under that the binding-time
+type @R@ is instantiated as @Rep@. The types of abstract components are reused
+from the unstaged version.
 
@@
 but replacing \textit{all} the monad transformers with their staged versions. The
-following code snippet shows this change. We manually fuse the three inner
-transformers (@IdM@ omitted) into a single monad @RepSetReaderStateM[A, B, C]@,
-where @A@ is the type parameter for the reader effects and @B@ for the state effects.
-In our monad stack scheme, they are both instantiated with the type @Cache@. 
-Similar to the unstaged version, the grounded result type is a pair of two
-staged values: @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
+following code snippet shows this change.
+We manually fuse the three inner transformers (@IdM@ is omitted) into a single
+monad @RepSetReaderStateM[R, S, A]@, where @R@ is the type parameter for the
+reader effects, and @S@ is the type for the state.  In our monad stack scheme,
+@R@ and @S@ are both instantiated with the type @Cache@. Similar to the unstaged
+version, the grounded result type is a pair of two staged values:
+@Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
 \begin{lstlisting}
@@
 When deriving the staged concrete interpreter from its unstaged counterpart, we
-notice that the recursive calls of the evaluator @ev@ on the lambda body are shifted from
-@ap_clo@ to @close@. In other words, we eagerly specialize the interpreter and
-generate code every time when we see a $\lambda$ term, instead of lazily
-calling @ev@ when the applications happen.  With no surprise, it is similar
-here. Additionally, in the staged abstract interpreter, we have to handle
-nondeterminism incurred by the over-approximation of runtime behavior.  In the
-rest of this part, we keep our roadmap and focus on discussing closure
-representations and function applications for the staged version.
+notice that the recursive calls of the evaluation function @ev@ on the lambda
+body are shifted from @ap_clo@ to @close@. In other words, we eagerly
+specialize the interpreter and generate code every time when we reach a
+$\lambda$ term, instead of lazily calling @ev@ when the applications happen.
+Similarly, we use the same way to handle staged $\lambda$ terms.  Additionally,
+in the staged abstract interpreter, we have to handle nondeterminism incurred
+by the over-approximation of runtime behavior.  In the rest of this part, we
+keep our roadmap and focus on discussing closure representations and function
+applications for the staged version.
 
@@
 type @Ans@ to grounded values of type @Result@, by providing the desired
-arguments, i.e., the new environment, new store, and caches. The collapsing of
-monads happens at the current-stage, so the calls of @ev@ is unfolded within
-staging. Finally, we generate a singleton set containing the compiled
+arguments, i.e., the new environment, new store, and two caches. The collapsing
+of monadic values happens at the current-stage, so the invocation of @ev@ is unfolded
+within staging. Finally, we generate a singleton set containing the compiled
 next-stage closure @emit_compiled_clo(f)@, which is represented by an IR node
@@
 
-The @ap_clo@ method is also similar: we use the staged version of @lift_nd@ to
-lift the next-stage set of closures into the monad stack, and generate a
-next-stage value for each closure @clo@ in the set, by using @emit_ap_clo@. The
-method @emit_ap_clo@ generates the current-stage representation of the
-future-stage application result.
-Finally, we reify the future @out@ cache, store, and values back into
-the current-stage monads.
+The @ap_clo@ method for function applications is also similarly mixing the two
+previous interpreters.  We use the staged version of @lift_nd@ to lift the
+next-stage set of functions into the monad stack.
+For each closure @clo@ in the set, we a generate next-stage value representing
+the function application @clo(arg)@ by using @emit_ap_clo@.  Again, the method
+@emit_ap_clo@ produces a current-stage representation of the future-stage
+application result.  Finally, we reify the @out@ cache, store, and values,
+which are all from the future-stage, back into the current-stage monadic value.
 \begin{lstlisting}
@@
 
-\paragraph{Staged Caching and Fixpoint Iteration} Our fixed-point iteration
+\paragraph{Staged Caching and Fixpoint Iteration} The fixed-point iteration
 again relies on the two caches @in@ and @out@, which are both staged maps now.
-Because of this, the query of whether the @in/out@ cache contains the current
+Therefore, the query of whether the @in/out@ cache contains the current
 configuration will produce next-stage Boolean values, i.e., of type
-@Rep[Boolean]@. The branching condition cannot be determined
-statically, consequently we need to generate code for the whole @if@ expression. Figure
-\ref{fig:staged_coind_cache} shows the staged version of @fix@. The
+@Rep[Boolean]@, and the branching condition cannot be determined statically.
+We have to generate code for the whole @if@ expression.
+Figure \ref{fig:staged_coind_cache} shows the staged version of @fix@. The
 variable @res@ represents the next-stage result, consisting of a next-stage
-@if@ expression. The true branch simply returns a pair of the queried result
-form the @out@ cache and the @out@ cache itself. The else branch constructs a
+@if@ expression. The true branch simply returns a pair of the query result
+form the @out@ cache, and the @out@ cache itself. The else branch constructs a
 monadic value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@
@@
 Finally, we have a current-stage representation of the future-stage values
-@res@, and put the content of @res@ back into the monad stack.
+@res@, and we reify the content of @res@ back into the current-stage monad stack.
 
@@
 
-Our staging schema works by unfolding the interpreter over the input abstract
-syntax tree. In practice, however, it would suffer from code explosion when
-analyzing (specializing) large programs, which increases the compile time. If
-we generate next-stage programs running on JVM, such large generated program
-would also potentially cause runtime GC overhead. In this section, we present
-optimizations that largely mitigate these issues. For all that,
-implementing these optimizations do not need to change the generic interpreter.
+Our staging schema works by unfolding the interpreter over the abstract syntax
+tree of the input program. In practice, however, it would suffer from code
+explosion when analyzing (specializing) large programs, which increases the
+compile time. If we generate next-stage programs running on JVM, such large
+generated program would also potentially cause runtime GC overhead. In this
+section, we present optimizations that largely mitigate these issues. For all
+that, implementing these optimizations do not need to change the generic
+interpreter.
 
@@
 
-All components in the interpreter except @Expr@s are staged. The data
-structures representing these components are treated as black-boxes, such as
-@Map@, which means that the operations on a @Map@ directly become nex-stage code,
-but we don't inspect any further inside. As we identified when
-introducing the generic interface, the keys of @Env@s are identifiers (i.e.,
-strings) in the program, which are completely known statically. This
-observation gives us a chance to further specialize the data structures.  For
-example, let us assume that the @Map[K, V]@ is implemented as a hash map. If
-all the keys of type @K@ are all known statically, then the indices also can be
-computed statically. Thus, the specialized map would be an array of type
-@Array[Rep[V]]@, whose elements are next-stage values, and the size of the array is
-known statically. An access to the map is translated into an access to the
-array with the pre-computed index during staging.
+In the staged interpreters, all instances of @Env@ and @Store@ are staged. The
+data structures representing these components are treated as black-boxes, i.e.,
+@Rep[Map]@, which means that the operations on a @Rep[Map]@ directly become
+nex-stage code, and we do not inspect any further inside. As we identified when
+introducing the generic interface, the keys of an @Env@ are string-represented
+identifiers in the program, which are completely known statically. This
+observation gives us a chance to further specialize the data structures for environment.
+For example, let us assume that the @Map[K, V]@ is implemented as a hash-map.
+If all the keys of type @K@ are all known statically, then the indices for
+those keys can also be computed statically. Thus, the specialized map would be
+an array of type @Array[Rep[V]]@, whose elements are next-stage values, and the
+size of the array is known statically as the program has finite number of
+identifiers. As result, an access to the map is staged into an access to the array
+with the pre-computed index.
 
 Particularly, if we are specializing a monovariant analysis, the address space
-is equivalent to the set of identifiers in the program. Then the accesses to
-the environment can be completely computed statically and we may generate the
-addresses directly during staging. After which, the stores are specialized
-as arrays with element type @Rep[Value]@.
+is equivalent to the set of all identifiers in the program. Utilizing this fact, the
+result of accesses to the environment can be computed statically and we may
+generate addresses directly during staging. After which, the store can be
+specialized as arrays by using the way mentioned above.
 
@@
 and function @f@. But we can also utilize algebraic properties of @foldLeft@
-to generate cheaper code, e.g, @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
+to generate cheaper code, e.g., @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
 function @f@ is known at the current stage, we completely eliminate the fold
 operation and function application. We apply several rewritings enabled by
-partial-static patterns, such as for @Set@ and @Map@. This optimizations
-greatly reduces the size of residual programs.
+partial-static patterns, such as for @Set@, @Map@, and @Tuple@. This
+optimizations greatly reduces the size of residual programs.
 
@@
 \cite{Ofenbeck:2017:SGP:3136040.3136060, Amin:2017:CTI:3177123.3158140}, the
-abstract interpreter can be extended for generating efficient code. The
+abstract interpreter can be extended to generate efficient code. The
 underlying semantics and staging are two orthogonal dimensions. It is important
@@
 consists of statements and expressions that purely manipulate the environment,
-store, and two caches, whose underlying representations are all @Rep[Map[K,V]]@. We
-also have several operations on tuples and lists, which are residualized from
-the internal code fragments of monads.
+store, and two caches, whose underlying representations are all @Map[K,V]@.
+We also have several operations on tuples and lists, which are residualized from
+the internal code fragments of the monads.
 
@@
 i.e, if the unstaged one does not produces false negative result, the staged
-one also does not. This soundness preservation relies on the correctness
+one also does not. This soundness preservation indeed relies on the correctness
 of the staged implementations and of the underlying MSP system. We now briefly
@@
     the same result given the same arguments.
-    In this paper, we use the LMS framework, which 1) relies on the
-    type checking facilities of Scala, and 2) generates code in A-Normal form
-    \cite{Flanagan:1993:ECC:155090.155113} that preserves the execution order
-    within a stage \cite{DBLP:conf/birthday/Rompf16}.  However, in general, it
-    is possible to subvert these guarantees by using unsafe features, such as
-    casts.
+    In this paper, we use the LMS framework, which 1) checks binding-times by
+    checking types which is enabled by the Scala compiler, and 2) generates
+    code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113} that preserves
+    the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
+    However, in general, it is possible to subvert these guarantees by using
+    unsafe features, such as casts.
   \item As shown in the previous sections, the generic interpreter is untouched
@@
 
-\subsection{Design Choices}
+\subsection{Alternatives}
 

### 2019-09-09 paper/60_cases.tex
polish

@@
     τ <- get_time
-  } yield unit(KCFAAddr(x, τ)) // we use unit to turn a current-stage constant to next-stage value
+  } yield unit(KCFAAddr(x, τ)) // we use unit to turn a current-stage constant to a next-stage value
 \end{lstlisting}
@@
 analyzing libraries. Some static analyzers compute a summary for a function or
-a module, which can be reused later. However, when lacking of the context
+a module, which can be reused later. However, when lacking the context
 information, those analyses can be too conservative or unsound, which both lead
@@
 
-As we have already seen from Section \ref{sai}, the primitive
+As we have already seen in Section \ref{sai}, the primitive
 operation @close@ can denote a $\lambda$ term to a next-stage Scala function.
@@
 We borrow the notation from traditional modular analysis
-\cite{DBLP:conf/cc/CousotC02}: such a generated function is a modularly reusable
-artifact, i.e., independent to its context.
+\cite{DBLP:conf/cc/CousotC02}: such a generated function is a modularly
+reusable artifact, i.e., independent to its calling context.
 The generated functions already eliminate the interpretation overhead and
 monadic overhead, and can therefore be considered as partial summaries. However, we
-still need to run the generated function at next stage to obtain the analysis
+still need to run the generated function at the next stage to obtain the analysis
 result.  At the next stage, we can provide different abstract arguments, stores, and

### 2019-09-09 paper/40_absinterp.tex
polish

@@
 During the iteration, we first check whether the @out@ cache contains the
-configuration @cfg@, which represents the current desired computation. If @out@ does,
-then the result is directly returned throughout the monad stack.
-Otherwise, we first retrieve the result from @in@ (@⊥@ if @in@ not contains
+configuration @cfg@, which represents the current desired computation. If @out@
+does contain @cfg@, then the result is directly returned throughout the monad
+stack.
+Otherwise, we first retrieve the result from @in@ (@⊥@ if @in@ does not contain
 @cfg@), and update this result from @in@ into the @out@ cache in the fashion of

### 2019-09-09 paper/50_stageabs.tex
polish

@@
 The @ap_clo@ method for function applications is also similarly mixing the two
-previous interpreters.  We use the staged version of @lift_nd@ to lift the
+previous versions.  We use the staged version of @lift_nd@ to lift the
 next-stage set of functions into the monad stack.
-For each closure @clo@ in the set, we a generate next-stage value representing
-the function application @clo(arg)@ by using @emit_ap_clo@.  Again, the method
+For each closure @clo@ in the set, we generate a next-stage value, representing
+the function application @clo(arg)@, by using @emit_ap_clo@.  Again, the method
 @emit_ap_clo@ produces a current-stage representation of the future-stage
@@
 generate addresses directly during staging. After which, the store can be
-specialized as arrays by using the way mentioned above.
+specialized as an array by using the way mentioned above.
 

### 2019-09-09 paper/60_cases.tex
polish

@@
 
-In this section, we show AC can be understood and
-implemented as an instance of staging abstract interpreters. We first revisit
-the original implementation of abstract compilation for 0-CFA, and then
-reproduce their result by simply adding stage annotations.
-The program generated by our approach provides approximately the same
-improvement with regards to speed, but without changing a single line of the
-analyzer program using type-based staging annotations. However, AC requires
+We show AC can be understood and implemented as an instance of staging abstract
+interpreters. We first revisit the original implementation of abstract
+compilation for 0-CFA, and then reproduce their result by simply adding stage
+annotations.  The program generated by our approach provides approximately the
+same improvement with regards to speed, but without changing a single line of
+the analyzer program using type-based staging annotations. However, AC requires
 more engineering effort, i.e., rewriting the whole analyzer into the
@@
 
-In this section, to demonstrate that the SAI approach is widely applicable, we
+To demonstrate that the SAI approach is applicable to various analyses, we now
 extend and refactor the staged abstract interpreter we presented in Section
-\ref{sai} to various control-flow analyses (CFA). We first discuss the
-calling-context sensitive analysis for staged analyzers by tweaking the
+\ref{sai} to several variants of control-flow analyses (CFA). We first discuss
+the calling-context sensitive analysis for staged analyzers by tweaking the
 allocation strategy. Then, we describe how to implement store-widening with the
@@
 the address allocation strategy, we can achieve various context-sensitive
-analyses \cite{DBLP:conf/icfp/Gilray0M16}. In this section, we demonstrate a
+analyses \cite{DBLP:conf/icfp/Gilray0M16}. Here, we demonstrate a
 calling-context sensitive analysis, i.e., $k$-CFA-like analysis
@@
 implemented in the monadic abstract interpreter by swapping the @StateT@ and
-the @SetT@ monad transformer. In this section, we show this optimization is
-orthogonal to staging and we just need to similarly swap the same
-\textit{staged} version of monads. After the adjustment, conceptually, we have
-the following monad stack:
+the @SetT@ monad transformer. This optimization is orthogonal to staging.
+Within the monadic interpreter framework, we just need to similarly swap the
+same \textit{staged} version of monads. After this adjustment, conceptually, we
+have the following monad stack:
 \begin{lstlisting}
@@
 
-The idea of abstract garbage collection is shown in Figure \ref{fixgc}. We
-update the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@
+The skeleton of a staged version of abstract garbage collection implementation
+is shown in Figure \ref{fixgc}.
+We refactor the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@
 (the caching part is omitted in the presented code snippet). Several auxiliary
@@
 to imprecision. In this section, we show that meta-programming can provide a
-trade-off between summary-based modular analysis and slow whole-program
-analysis by utilizing an existing whole-program analyzer.
+trade-off between summary-based modular analyses and slow whole-program
+analyses, by utilizing an existing whole-program analyzer.
 
-As we have already seen in Section \ref{sai}, the primitive
-operation @close@ can denote a $\lambda$ term to a next-stage Scala function.
-The specializing of the staged abstract interpreter to a $\lambda$ term is
-independent to other parts of the analyzed program.
+As we have already seen in Section \ref{sai}, the primitive operation @close@
+can denote a $\lambda$ term to a next-stage Scala function. Similar to
+compilers, the specializing of the staged abstract interpreter to a $\lambda$
+term is also independent of other parts of the analyzed program.
 \begin{lstlisting}
@@
 
-We use Scheme as an example, and consider a library that contains a
-function @map@:
+To illustrate the idea concretely, we consider a library that contains a
+function @map@ and use the Scheme language as example:
 \begin{lstlisting}
@@
 \cite{DBLP:conf/cc/CousotC02}: such a generated function is a modularly
-reusable artifact, i.e., independent to its calling context.
+reusable artifact, i.e., independent of its context.
 The generated functions already eliminate the interpretation overhead and
-monadic overhead, and can therefore be considered as partial summaries. However, we
-still need to run the generated function at the next stage to obtain the analysis
-result.  At the next stage, we can provide different abstract arguments, stores, and
-caches, depending on the context where it is called.
+monadic overhead via staging, and can therefore be considered as partial
+summaries. However, we still need to run the generated function at the next
+stage to obtain the analysis result.  At the next stage, we can provide
+different abstract arguments, stores, and caches, depending on the context
+where it is called.
 

### 2019-09-09 paper/70_eval.tex
polish

@@
 
-To evaluate how performance can be improved by our staged abstract interpreters
-approach, we scale our toy interpreter to support a large subset of the Scheme
-language and evaluate the analysis time against the unstaged versions. The
-result shows that the staged versions are, on average, 10+ times as fast as the
-unstaged version, as evidence of our \textit{abstraction without regret}
-approach.
+To evaluate how much performance can be improved by the staging, we scale our
+toy staged abstract interpreter to support a large subset of the Scheme language and evaluate
+the analysis time against the unstaged versions. The result shows that the
+staged versions are, on average, 10+ times as fast as the unstaged version, as
+evidence of our \textit{abstraction without regret} approach.
 

### 2019-09-09 paper/50_stageabs.tex
misc

@@
 \begin{itemize}
-  \item Constructing a generic interpreter that abstracts over binding times,
+  \item First, we construct a generic interpreter that abstracts over binding times,
     value domains, and primitive operations. In this paper, the generic interpreter
@@
     into monads.
-  \item Then we can implement an unstaged abstract interpreter modularly using
-    the appropriate monad stack. This step has been explored in the previous
+  \item Then, we implement an unstaged abstract interpreter modularly using the
+    appropriate monad stack. This step has been explored in the previous
     literatures.

### 2019-09-09 paper/80_related.tex
typos
polish

@@
 Compared with other approaches using syntactic quotations, LMS provides
-guarantee for preserving evaluation orders within a stage.
+guarantees for preserving evaluation orders within a stage.
 \citet{Ofenbeck:2017:SGP:3136040.3136060} and
@@
 monadic interpreters was explored by \citet{danvy1991compiling,
-DBLP:conf/dsl/SheardBP99}. However, they are different from the performance motivation
-here: the generated code is in monadic form.
+DBLP:conf/dsl/SheardBP99}. However, they are different from the performance
+motivation here.

### 2019-09-09 paper/10_intro.tex
typo
misc

@@
 reinterpreting the shared interpreter.  The generic interpreter returns a value
-of monadic type, which vary based on semantics. The value domains
+of monadic types, which vary based on semantics. The value domains
 of the interpreter and the effects such as state and nondeterminism can all be
@@
 annotations, we can achieve the same optimizations. Meanwhile,
-the analyzer does not need to change, except the addition of stage annotations,
+the analyzer does not need to change (modulo the addition of stage annotations),
 thereby requiring significantly less engineering effort.

### 2019-09-09 paper/30_stageconc.tex
typo

@@
 whose next-stage form is a Scala function.
-With this fact, we can generate a next-stage function application for it, also in
-Scala. The method @emit_ap_clo@ produces a current-stage representation
-of a next-stage function application, with necessary arguments such as the latest
-store @σ@. Meanwhile, we still want the method @ap_clo@ implemented in monadic
-style, so that it can be smoothly connected with other parts of the
-interpreter.  To achieve this, we lift the result values and stores from
-the future stage into our current-stage monad stack, with a binding @vs@.
-Finally, we use @put_store@ to update the store with @vs._2@ and return the
-value result @vs._1@ into the current-stage monadic value.
+With this fact, we can generate a next-stage function application for it, also
+in our target language Scala. The method @emit_ap_clo@ produces a current-stage
+representation of a next-stage function application, with necessary arguments
+such as the latest store @σ@. Meanwhile, we still want the method @ap_clo@
+implemented in monadic style, so that it can be smoothly connected with other
+parts of the interpreter.  To achieve this, we lift the result values and
+stores from the future stage into our current-stage monad stack, with a binding
+@vs@.  Finally, we use @put_store@ to update the store with @vs._2@ and return
+the value result @vs._1@ into the current-stage monadic value.
 
@@
 @f@. Hence, we generate a next-stage function application @f(arg, σ)@.  The
-invocations of @quote@ are used to generate proper names for the other
+invocations of @quote@ are used to generate proper names for other
 next-stage values.

### 2019-09-09 paper/40_absinterp.tex
typo

@@
 instantiate it to a commonly-used context-sensitive analysis;
-\citet{Darais:2015:GTM:2814270.2814308} show how to achieve path- and
+\citet{Darais:2015:GTM:2814270.2814308} showed how to achieve path- and
 flow-sensitivity by varying the monads, which is also applicable to our
@@
 (ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our monadic
-concrete interpreter to an monadic abstract interpreter.
+concrete interpreter to a monadic abstract interpreter.
 
@@
 nondeterminism comes from function applications, for instance, @f(a)@, the
-possible $\lambda$ terms of @f@ is computed approximately.  As a
+possible $\lambda$ terms of @f@ are computed approximately.  As the
 result, we need to store all the possible callees, as well as we will

### 2019-09-09 paper/30_stageconc.tex
typo

@@
 However, now we use the @Rep@ type to annotate the value domains, environments and
-stores, and redefine the staged version of monads and primitive operations.
+stores, and define the staged version of monads and primitive operations.
 
@@
 also not staged. Instead, the internal data that these monads operate on are
-staged, i.e., values of @R@ and @A@ in the reader monad, and values of @S@ and
+staged, i.e., values of type @R@ and @A@ in the reader monad, and values of type @S@ and
 @A@ in the state monad.

### 2019-09-09 paper/50_stageabs.tex
misc
typo

@@
 $\lambda$ term, instead of lazily calling @ev@ when the applications happen.
-Similarly, we use the same way to handle staged $\lambda$ terms.  Additionally,
+Similarly, here we use the same way to handle staged $\lambda$ terms.  Additionally,
 in the staged abstract interpreter, we have to handle nondeterminism incurred
@@
 within staging. Finally, we generate a singleton set containing the compiled
-next-stage closure @emit_compiled_clo(f)@, which is represented by an IR node
+next-stage closure (@emit_compiled_clo@), which is represented by an IR node
 in LMS.
@@
 identifiers in the program, which are completely known statically. This
-observation gives us a chance to further specialize the data structures for environment.
+observation gives us a chance to further specialize the data structures for environments.
 For example, let us assume that the @Map[K, V]@ is implemented as a hash-map.
@@
 caching is used for every recursive call in our abstract interpreter. But this
-is not necessary and redundant when generating code for atomic
+is unnecessary and redundant when generating code for atomic
 expressions such as literals or variables, because they always terminate.

### 2019-09-09 paper/60_cases.tex
typo

@@
 By applying AC, we refactor every function in the analyzer such that it
-return a closure just taking an environment argument.  Therefore the overhead
+returns a closure taking an environment argument.  Therefore the overhead
 of traversing the AST of input program is eliminated.
@@
 provides additional performance improvement. Nevertheless, the LMS framework
-provides support for necessary data structures, such as @Map@.  We consider
+provides support for necessary data structures, such as for @Map@.  We consider
 these efforts reusable and completely orthogonal to the implementation of the

### 2019-09-09 paper/20_prelim.tex
polish
typo

@@
 A monad is a type constructor @M[_]: * --> *@ with two operations, often called
-@return@ and @bind@. Informally, @return@ wraps a value into a monad @M@, and
-@bind@ unwraps a monadic value and transforms it into a new monadic value.
+@return@ and @bind@ \cite{Wadler:1992:EFP:143165.143169,
+DBLP:journals/iandc/Moggi91}. Informally, @return@ wraps a value into a monad
+@M@, and @bind@ unwraps a monadic value and transforms it into a new monadic value.
 Pragmatically in Scala, we define the monad type class using trait @Monad@
@@
 transformers, we can combine multiple monads into a single one.  The
-implementation of monads and monad transformers follows the one from Haskell.
+implementation of monads and monad transformers follows the ones from Haskell.
 

### 2019-09-10 paper/paper.tex
space
revision
title and subsection titles

@@
 %\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-\documentclass[acmsmall,review]{acmart}\settopmatter{}
 %% For double-blind review submission, w/ CCS and ACM Reference
@@
 %% For final camera-ready submission, w/ required CCS and ACM Reference
-%\documentclass[acmsmall]{acmart}\settopmatter{}
+\documentclass[acmsmall, screen]{acmart}\settopmatter{}
 
@@
 \setcopyright{rightsretained}
+\acmPrice{}
+\acmDOI{10.1145/3360552}
+\acmYear{2019}
+\copyrightyear{2019}
 \acmJournal{PACMPL}
-\acmYear{2019} 
 \acmVolume{3}
@@
 \acmMonth{10}
-\acmPrice{}\acmDOI{10.1145/3360552}
 
@@
 %\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitle{Fast and Modular Whole-Program Analysis via Meta-Programming}                     %% \subtitle is optional
+\subtitle{Fast and Modular Whole-Program Analysis via Meta-programming}                     %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
 %% comma separated list
-% \keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
+\keywords{multi-stage programming, abstract interpreters, control-flow analysis}  %% \keywords are mandatory in final camera-ready submission
 
@@
 \maketitle
-
-\iffalse
-
-\renewcommand\thefootnotecopyrightpermission{}
-\footnotetextcopyrightpermission{
-Preprint, November 2018.\\ Copyright held by the authors.}
-\fancyhead[RO,LE]{Preprint, November 2018}
-
-\fi
 

### 2019-09-10 paper/60_cases.tex
title and subsection titles
typo
polish

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Control-flow Analysis} \label{cfa}
+\subsection{Control-Flow Analysis} \label{cfa}
 
@@
 
-\subsubsection{Context-sensitivity}
+\subsubsection{Context-Sensitivity}
 
@@
 
-\subsubsection{Store-widened Analysis}
+\subsubsection{Store-Widened Analysis}
 
@@
 is shown in Figure \ref{fixgc}.
-We refactor the @fix@ function with caching (from Section \ref{sai}) to @fix_gc@
+We refactor the @fix@ function with caching (used in Section \ref{sai}) to @fix_gc@
 (the caching part is omitted in the presented code snippet). Several auxiliary
@@
 performs garbage collection according to the given reachable addresses and
-returns a new store. For the generic interpreter, we also need to compute the
-root addresses for compound expressions; for the detail of this part, readers
-may refer to \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+returns a new store. We also need to compute the root addresses for compound
+expressions, which is elided in the code snippet; for the details of this part,
+readers may refer to \cite{DBLP:journals/pacmpl/DaraisLNH17}.
 

### 2019-09-10 paper/20_prelim.tex
polish

@@
 in @for@-comprehensions is translated into a @map@, where the argument of
-@yield@ becomes the body expression inside of that @map@ operation. The
+@yield@ becomes the body expression inside that @map@ operation. The
 bindings before the last one are all translated into calls of @flatMap@:
@@
 returned monadic type abstract, so that they can be instantiated
-differently; we also need to abstract the primitive operations on those types.
+differently. We also need to abstract the primitive operations on those types.
 In later sections, we will instantiate the monadic type to perform concrete
@@
 @Store@. An @Env@ maps identifiers to addresses, and a @Store@ maps addresses to
-values. The @Env@ captures bound variables in the scope of current control, and
+values. An environment captures bound variables in the scope of current control, and
 @Store@ models a persistent heap through the program run-time.  Currently,
@@
 type parameter @R[_]@ as the binding-time, i.e., @MonadOps[R, AnsM, T]@.
-Inside of @MonadOps@, @R[_]@ annotates the data types @A@ and @B@ that are
+Inside @MonadOps@, @R[_]@ annotates the data types @A@ and @B@ that are
 encapsulated by the monad, but not the monad type @M@ itself. When the generic
@@
 lifts lambda terms to closures and method @ap_clo@ that applies functions also
-takes an extra @ev@, because evaluations may happen inside of them.
+takes an extra @ev@, because evaluations may happen inside them.
 To close the open recursion, we use a fixed-point operator @fix@.

### 2019-09-11 paper/30_stageconc.tex
title and subsection titles
typo

@@
 
-\subsection{Multi-Stage Programming with LMS}
+\subsection{Multi-stage Programming with LMS}
 
@@
 interfaces. Later, we will instantiate the staged set monad by manually fusing
-the fragment of the monad stack inside of @SetT@.
+the fragment of the monad stack inside @SetT@.
 
@@
 next-stage function values, i.e., of type @Rep[Value]@.
-Inside of function @f@, we can access the evaluator via @ev@.
+Inside function @f@, we can access the evaluator via @ev@.
 However, @ev@ produces a monadic value of type @AnsM@, which can only exist at

### 2019-09-11 paper/40_absinterp.tex
typo

@@
 However, under the new monad stack for abstract interpretation, we have a
-@SetT@ inside of the environment monad and store monad. Therefore, the type
+@SetT@ monad transformer inside the environment monad and store monad. Therefore, the type
 @Set@ becomes the container type of the pairs of values and stores, i.e.,
@@
 \cite{DBLP:journals/pacmpl/DaraisLNH17, Wei:2018:RAA:3243631.3236800,
-  Rosendahl:AbsIntPL}. The idea is to use a @in@ cache and a @out@ cache during
+  Rosendahl:AbsIntPL}. The idea is to use an @in@ cache and an @out@ cache during
 the depth-first evaluation. The @in@ cache stores the result from the last

### 2019-09-11 paper/60_cases.tex
typo

@@
 Since $k$ is bounded to a fixed number, after the analysis running $k$ steps
-inside of @f@, the calling context will purely depend on the static structure of
+inside @f@, the calling context will purely depend on the static structure of
 our module being analyzed.

### 2019-09-11 paper/20_prelim.tex
polish
typo

@@
 values. An environment captures bound variables in the scope of current control, and
-@Store@ models a persistent heap through the program run-time.  Currently,
+a store models a persistent heap through the program run-time.  Currently,
 the domains of addresses and values are still abstract; hence, they are
@@
 \paragraph{The Interpreter} We can now construct the semantics-agnostic interpreter
-in monadic form, shown in Figure \ref{fig:shared_int}. The basic idea
+in monadic form, as shown in Figure \ref{fig:shared_int}. The basic idea
 of generic interpretation is to traverse the abstract syntax tree

### 2019-09-11 paper/50_stageabs.tex
typo

@@
 compile time. If we generate next-stage programs running on JVM, such large
-generated program would also potentially cause runtime GC overhead. In this
+generated programs would also potentially cause runtime GC overhead. In this
 section, we present optimizations that largely mitigate these issues. For all
@@
 
-We implemented a big-step, compositional abstract interpreter in monadic
+We have implemented a big-step, compositional abstract interpreter in monadic
 style, where \textit{compositional} means that every recursive call of our abstract

### 2019-09-11 paper/50_stageabs.tex
polish

@@
 We have implemented a big-step, compositional abstract interpreter in monadic
-style, where \textit{compositional} means that every recursive call of our abstract
-interpreter is applied to the proper substructures of the current syntactic
+style, where \textit{compositional} means that every recursive call of @eval@ %our abstract
+is applied to the proper substructures of the current syntactic
 parameters \cite{10.1007/3-540-61580-6_11}. This compositionality ensures that

### 2019-09-11 paper/paper.tex
polish

@@
   We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
-  Guanhong Tao, Gr\'egory Essertel, and Fei Wang for their feedback on the early
-  drafts. We thank David Van Horn and David Darais for their encouragement on
-  this work.
+  Guanhong Tao, Gr\'egory Essertel, Fei Wang, Qianchuan Ye, and Yapeng Ye for
+  their feedback on the early drafts. We thank David Van Horn and David Darais
+  for their encouragement on this work.
   This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE

### 2019-09-11 paper/10_intro.tex
typo
misc

@@
 compilers through a hierarchy of specializations, known since then as the
-Futamura Projections. The first Futamura projection states that specializing an
+Futamura projections. The first Futamura projection states that specializing an
 interpreter to an input program removes interpretative overhead and yields an
@@
   that statically computes runtime properties;
-\item With two-level binding-times and abstract domains, it is an optimizing
+\item With two-level binding-times and abstract domains, it is an optimized
   program analyzer, but works in the fashion of compilation.

### 2019-09-11 paper/50_stageabs.tex
typo

@@
 @Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully generates code that applies @foldLeft@ with the set
+generator would faithfully generate code that applies @foldLeft@ with the set
 and function @f@. But we can also utilize algebraic properties of @foldLeft@
@@
 partial-static patterns, such as for @Set@, @Map@, and @Tuple@. This
-optimizations greatly reduces the size of residual programs.
+optimization greatly reduces the size of residual programs.
 

### 2019-09-11 paper/50_stageabs.tex
typo

@@
 
-\paragraph{What has been eliminated?} In the generated code, all the
+\paragraph{What has been eliminated?} In the generated code, all
 primitive operations (such as @eval@, @fix@, @ap_clo@, etc.) and monadic

### 2019-09-11 paper/60_cases.tex
polish

@@
 
-We show AC can be understood and implemented as an instance of staging abstract
+We show that AC can be understood and implemented as an instance of staging abstract
 interpreters. We first revisit the original implementation of abstract
@@
 the analyzer program using type-based staging annotations. However, AC requires
-more engineering effort, i.e., rewriting the whole analyzer into the
+more engineering efforts, i.e., rewriting the whole analyzer into the
 closure-generation form. Moreover, as shown in Section~\ref{staged_ds}, our

### 2019-09-11 paper/70_eval.tex
polish

@@
-\section{Empirical Evaluation} \label{evaluation}
+\section{Performance Evaluation} \label{evaluation}
 
@@
 easy to implement. We implement two common monovariant 0-CFA-like analyses: one
-is equipped with store-widening (0CFA-SW for short), the other is not (0CFA
+is equipped with store-widening (0CFA-SW for short), the other one is not (0CFA
 for short). It is traditionally expected that the store-widened analysis be faster than the
-alternative.  Our prototype implementation currently generates Scala code; the
+alternative.  Our prototype implementation currently generates Scala code. The
 generated Scala code will be compiled and executed on the JVM, though it is possible to
 generate C/C++ code in the future. In the experiments, we implement several
-optimizations mentioned in Section \ref{staged_ds}, specifically, the selective
-caching scheme and IR-level rewriting rules (for pairs, lists, sets, and maps)
-that exploit partially-static data. These optimizations are useful to reduce
-the size of generated code.
+optimizations mentioned in Section \ref{staged_ds}, specifically, including the
+selective caching scheme and IR-level rewriting rules (for pairs, lists, sets,
+and maps) that exploit partially-static data. These optimizations are useful to
+reduce the size of generated code and improve performance.
 
@@
 -XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
-running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All of our
+running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All our
 evaluations were performed on a machine with 4 Intel Xeon Platinum

### 2019-09-11 paper/80_related.tex
polish

@@
 stage conducts a flow-sensitive analysis by refining the first
-one. Another large body of static analyses is implemented by
+one. Many other forms of static analysis are implemented by
 first generating constraints and then using a solver to discharge the

### 2019-09-11 paper/60_cases.tex
typo
polish

--- 2019-09-11 paper/60_cases.tex
@@
 The @tick@ function is implemented as prepending the current control expression
-being evaluated (@e@) to the existing calling context, and then taking the
-first $k$ elements from the list. The type @Config@ used in the caching
-algorithm is also changed to include timestamps.
+@e@ being evaluated to the existing calling context, and then taking the first
+$k$ elements from the list. The type @Config@ used in the caching algorithm is
+also changed to include timestamps.
 

### 2019-09-11 paper/80_related.tex
polish

@@
 one. Many other forms of static analysis are implemented by
-first generating constraints and then using a solver to discharge the
-proof obligations or find the fixed-point. Notable examples include
-SAT- and SMT-based program analysis
+first collecting constraints from programs and then using a solver to discharge
+the proof obligations or find the fixed-point. Notable examples include SAT-
+and SMT-based program analysis
 \cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to

### 2019-09-11 paper/paper.tex
polish

--- 2019-09-10 paper/paper.tex
@@
   We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
-  Guanhong Tao, Gr\'egory Essertel, and Fei Wang for their feedback on the early
-  drafts. We thank David Van Horn and David Darais for their encouragement on
-  this work.
+  Guanhong Tao, Gr\'egory Essertel, Fei Wang, Qianchuan Ye, Yapeng Ye, and Le Yu
+  for their feedback on the early drafts. We thank David Van Horn and David
+  Darais for their encouragement on this work.
   This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE

### 2019-09-11 paper/60_cases.tex
typo

@@
 
-In Section~\ref{sai}, we used a toy language to show that staging an abstract
+In Section~\ref{sai}, we use a toy language to show that staging an abstract
 interpreter is feasible and the artifact can be derived by combining the staged

### 2019-09-11 paper/40_absinterp.tex
misc
polish

@@
 context/path/flow-insensitive by our choice. In Section \ref{cfa}, we will
-instantiate it to a commonly-used context-sensitive analysis;
+instantiate it to a commonly-used context-sensitive analysis.
 \citet{Darais:2015:GTM:2814270.2814308} showed how to achieve path- and
@@
 come from the monads and thus the nondeterminism can be naturally handled. The
-light gray code shows what is added from the concrete @ap_clo@.
+light gray line shows what is added from the concrete @ap_clo@.
 \begin{lstlisting}[escapechar=!]

### 2019-09-11 paper/40_absinterp.tex
polish

@@
 caches both inhibit internally of the nondeterminism monad; as the result, the
-final result type is paired the set of value-stores with the cache type from
+final result type pairs the set of value-stores with the cache type from
 that state monad, as shown below.

### 2019-09-11 paper/80_related.tex
misc
typo

@@
 Splitting an analysis into multiple stages has also been studied for other
-forms of static analyses.
+forms of static analysis.
 \citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
@@
 \citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract
-interpreters is closely related to the two-level semantics: the
+interpreter is closely related to the two-level semantics: the
 use of a generic interface with monads and then reinterpreting it with

### 2019-09-11 paper/70_eval.tex
typo
misc

@@
 evaluations were performed on a machine with 4 Intel Xeon Platinum
-8168 CPU at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores
+8168 CPUs at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores
 (192 threads) in total, the abstract interpreters are single-threaded.
@@
 %booleans.
-For 0CFA without store-widening, we observe an average speedup of ~17x times;
-for 0CFA with store-widening, the average speedup is ~7x times.
+For 0CFA without store-widening, we observe an average speedup of ~17 times;
+for 0CFA with store-widening, the average speedup is ~7 times.
 \texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it

### 2019-09-11 paper/50_stageabs.tex
misc

@@
 takes four next-stage values, including the argument and latest store, and
-@in/out@ caches additionally. Inside of @f@, we collapse the monadic value of
+@in/out@ caches additionally. Inside @f@, we collapse the monadic value of
 type @Ans@ to grounded values of type @Result@, by providing the desired

### 2019-09-11 paper/40_absinterp.tex
misc

@@
 conditional expression, we may not have enough information to decide which
-branch will be taken, thus a sound treatment is to explore the both branches.
+branch will be taken, thus a sound treatment is to explore both of the branches.
 Also, at some later point, we need to compute the join of two path.  Another

### 2019-09-11 paper/20_prelim.tex
polish
lambdas

@@
 in monadic form, as shown in Figure \ref{fig:shared_int}. The basic idea
-of generic interpretation is to traverse the abstract syntax tree
+of generic interpretation is to traverse the AST
 while maintaining the effects (such as environment and state updates).
@@
 type @Expr => Ans@ that refers to itself. Consequently, the method @close@ that
-lifts lambda terms to closures and method @ap_clo@ that applies functions also
+lifts $\lambda$-terms to closures and method @ap_clo@ that applies functions also
 takes an extra @ev@, because evaluations may happen inside them.
@@
 A value can be either a tagged number \texttt{IntV}, or a closure
-\texttt{CloV} that contains a lambda term and an environment. The final
+\texttt{CloV} that contains a $\lambda$-term and an environment. The final
 result of the interpreter is a pair of values and stores.
@@
 straightforwardly. We elide most of them but describe a bit on how we handle
-functions and applications. The reason is that lambda terms are data but are also
+functions and applications. The reason is that $\lambda$-terms are data but are also
 part of the control flow. The way we treat them is simple now, but
 will become more involved when we stage the abstract interpreter.
-The method @close@ denotes a lambda term to our representation of functions.
-At the moment, @close@ takes a lambda term and an environment, and
+The method @close@ denotes a $\lambda$-term to our representation of functions.
+At the moment, @close@ takes a $\lambda$-term and an environment, and
 produces the defunctionalized representation of closures.
 The method @ap_clo@ performs function applications, which takes a function
-value @fun@ and an argument @arg@. It first extracts the lambda term and environment
-enclosed in @fun@, and allocates an addresses for the argument and updates the
-environment and store accordingly.  Finally, body expression of the lambda term
-is evaluated under the new environment and store.
+value @fun@ and an argument @arg@. It first extracts the syntactic
+$\lambda$-term and environment enclosed in @fun@, and allocates an addresses
+for the argument and updates the environment and store accordingly.  Finally,
+body expression of the $\lambda$-term is evaluated under the new environment
+and store.
 \begin{lstlisting}

### 2019-09-11 paper/30_stageconc.tex
space
lambdas

@@
 mentioned before, we will illustrate in detail how functions and applications
-are handled.  One problem here is what should we do for $\lambda$ terms when
+are handled.  One problem here is what should we do for $\lambda$-terms when
 staging them? We cannot directly create a next-stage defunctionalized closure
@@
 next stage and the interpretation overhead has not been eliminated. 
-The solution is to compile the $\lambda$ term with its environment by calling
+The solution is to compile the $\lambda$-term with its environment by calling
 the evaluator at the current stage. The result is generating a next-stage
@@
 invocation of @ev@ happens in @ap_clo@, i.e., at the application time; while in
-the staged interpreter, we eagerly apply @ev@ when denoting the lambda terms to
+the staged interpreter, we eagerly apply @ev@ when denoting the $\lambda$-terms to
 our next-stage value domains, i.e., at the value-representation time.
@@
   case IRApClo(fun, arg, σ) => emitValDef(sym, s"$\textdollar$fun.f($\textdollar${quote(arg)}, $\textdollar${quote(σ)})")
-  ...
-}
+  ... }
 \end{lstlisting}

### 2019-09-11 paper/40_absinterp.tex
lambdas

@@
 nondeterminism comes from function applications, for instance, @f(a)@, the
-possible $\lambda$ terms of @f@ are computed approximately.  As the
+possible $\lambda$-terms of @f@ are computed approximately.  As the
 result, we need to store all the possible callees, as well as we will
@@
 
-The value representation of lambda terms is still a defunctionalized closure of
+The value representation of $\lambda$-terms is still a defunctionalized closure of
 type @CloV@, but we lift it to a singleton set of @AbsValue@ (remember that

### 2019-09-11 paper/50_stageabs.tex
lambdas

@@
 When deriving the staged concrete interpreter from its unstaged counterpart, we
-notice that the recursive calls of the evaluation function @ev@ on the lambda
-body are shifted from @ap_clo@ to @close@. In other words, we eagerly
-specialize the interpreter and generate code every time when we reach a
-$\lambda$ term, instead of lazily calling @ev@ when the applications happen.
-Similarly, here we use the same way to handle staged $\lambda$ terms.  Additionally,
-in the staged abstract interpreter, we have to handle nondeterminism incurred
-by the over-approximation of runtime behavior.  In the rest of this part, we
-keep our roadmap and focus on discussing closure representations and function
+notice that where the evaluation @ev(e)@ is invoked are shifted from @ap_clo@ to
+@close@, where @e@ is the body expression of a $\lambda$-term.
+In other words, when staging, we eagerly specialize the interpreter and
+generate code every time when we reach a $\lambda$-term, instead of lazily
+calling @ev@ when the applications happen.  Similarly, here we use the same way
+to handle staged $\lambda$-terms.  Additionally, in the staged abstract
+interpreter, we have to handle nondeterminism incurred by the
+over-approximation of runtime behavior.  In the rest of this part, we keep our
+roadmap and focus on discussing closure representations and function
 applications for the staged version.

### 2019-09-11 paper/60_cases.tex
lambdas

@@
 As we have already seen in Section \ref{sai}, the primitive operation @close@
-can denote a $\lambda$ term to a next-stage Scala function. Similar to
-compilers, the specializing of the staged abstract interpreter to a $\lambda$
-term is also independent of other parts of the analyzed program.
+can denote a $\lambda$-term to a next-stage Scala function. Similar to
+compilers, the specializing of the staged abstract interpreter to a
+$\lambda$-term is also independent of other parts of the analyzed program.
 \begin{lstlisting}

### 2019-09-11 paper/70_eval.tex
lambdas

@@
 interpreters for control-flow analysis, i.e., a value-flow analysis that
-computes which lambda terms can be called at each call-site.
+computes which $\lambda$-terms can be called at each call-site.
 With other suitable abstract domains, the abstract interpreter
@@
 has a comparably large size of AST, it has a rather simple structure which only
-involves booleans, lambda terms, and applications. We conjecture that the simple
+involves booleans, $\lambda$-terms, and applications. We conjecture that the simple
 domain and structure allow the generated code to be efficiently optimized by the

### 2019-09-11 paper/70_eval.tex
0-cfa

@@
 easy to implement. We implement two common monovariant 0-CFA-like analyses: one
-is equipped with store-widening (0CFA-SW for short), the other one is not (0CFA
+is equipped with store-widening (0-CFA-SW for short), the other one is not (0-CFA
 for short). It is traditionally expected that the store-widened analysis be faster than the
@@
 %booleans.
-For 0CFA without store-widening, we observe an average speedup of ~17 times;
-for 0CFA with store-widening, the average speedup is ~7 times.
+For 0-CFA without store-widening, we observe an average speedup of ~17 times;
+for 0-CFA with store-widening, the average speedup is ~7 times.
 \texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
@@
 domain and structure allow the generated code to be efficiently optimized by the
-JVM. However, on the corresponding 0CFA-SW, we do not observe such a significant
+JVM. However, on the corresponding 0-CFA-SW, we do not observe such a significant
 speedup. Due to the complexity of generated code and next-stage running platform

### 2019-09-11 paper/80_related.tex
0-cfa
misc

@@
 compiler pass, due to its EXPTIME lower bound
-\cite{VanHorn:2008:DKC:1411204.1411243}.  However, several variants of 0CFA
+\cite{VanHorn:2008:DKC:1411204.1411243}.  However, several variants of 0-CFA
 were developed and shown to be useful in practice
@@
 abstract interpretation can be traced to Nielson's two-level semantics
-\cite{NIELSON1989117}; using two-level semantics for code generation
+\cite{NIELSON1989117}. Using two-level semantics for code generation
 was also explored by \citet{NIELSON198859}.

### 2019-09-11 paper/90_conc.tex
conc

@@
 projection to abstract interpreters.
-To realize it, we design a generic interpreter that abstracts over both
-binding times and value domains; by instantiating it with staged monads for
-abstract interpretation, the staged abstract interpreter can be readily derived.
-The meta-programming approach brings new insights and implementation techniques
-to optimize static analyzers. Optimizations can be achieved orthogonally to
-the interpreter and abstract semantics.  In our empirical evaluation, staging
-improves performance by an order of magnitude in terms of running time.
+To realize it, we design a generic interpreter that abstracts over binding
+times and value domains. By instantiating the generic interpreter with the
+staged monads for abstract semantics, the staged abstract interpreter can be
+readily derived.  The meta-programming approach brings new insights and
+implementation techniques to optimize static analyzers via code generation.
+Optimizations can be achieved orthogonally to the interpreter and abstract
+semantics.  In our evaluation, staging improves the performance of running time
+by an order of magnitude.

### 2019-09-11 paper/60_cases.tex
0-cfa
misc

@@
 
-The specialization (0CFA and without store-widening) of @close@ to
+The specialization (0-CFA and without store-widening) of @close@ to
 @map@ generates the following next-stage Scala code, that is, the
@@
 \begin{lstlisting}
-  val x39 = {(x40: List[Set[AbsValue]],          // argument value list
+  val x39 = {(x40: List[Set[AbsValue]],          // list of argument values
               x41: Map[Addr, Set[AbsValue]],     // store
@@
     val x51 = x50.zip(x40)                              // addresses paired with their values
-    val x61 = x51.foldLeft (x41) { case (x52, x53) =>   // prepare a new store that joins
+    val x61 = x51.foldLeft (x41) { case (x52, x53) =>   // preparing a new store that joins
       val x54 = x53._1;  val x55 = x53._2               //   the arguments and their values

### 2019-09-11 paper/10_intro.tex
misc

@@
 projections has attracted more attention, and has been successfully applied to
-many real-world problems, including query compilers
-\cite{DBLP:conf/sigmod/TahboubER18}, PyPy \cite{Bolz:2009:TMP:1565824.1565827},
-Truffle \cite{Marr:2015:TVP:2814270.2814275}, etc. The challenge here: can we
-apply Futamura projections to a general abstract interpreter to make it run
-faster? The solution is not obvious, as we need to specialize a nonstandard
-semantics that extensively uses nondeterminism, operates on abstract domains,
-and computes fixed-points iteratively over caches.
+many real-world problems, including PyPy \cite{Bolz:2009:TMP:1565824.1565827},
+Truffle \cite{Marr:2015:TVP:2814270.2814275}, query compilers
+\cite{DBLP:conf/sigmod/TahboubER18}, etc. The challenge here: can we apply
+Futamura projections to a general abstract interpreter to make it run faster?
+The solution is not obvious, as we need to specialize a nonstandard semantics
+that extensively uses nondeterminism, operates on abstract domains, and
+computes fixed-points iteratively over caches.
 

### 2019-09-11 paper/10_intro.tex
polish

@@
 the years, constructing artifacts and analyzers that perform sound abstract
-interpretation for modern and expressive languages was considered complicated
+interpretation for modern and expressive languages is considered complicated
 for a long time.
@@
 extensive use of functional encodings of side-effects to implement the semantics
-of abstract interpretation.  For example, the list monad can be used to model
-nondeterminism from collecting interpreters. Although such pure approaches have
-their own merits, they are significantly slower when compared with imperative,
-stateful implementations.  In order to reconcile these approaches, one must use
-a programming technique which allows for high-level abstractions during
+of interpreted languages. For example, the list monad can be used to model
+nondeterminism of collecting interpreters. Although such pure approaches have
+their merits, they are significantly slower when compared with imperative,
+stateful implementations.  In order to reconcile these approaches, one need to
+use a programming technique which allows for high-level abstractions during
 implementation, but bypasses the incurred overhead at runtime.
@@
 that extensively uses nondeterminism, operates on abstract domains, and
-computes fixed-points iteratively over caches.
+computes fixed-points iteratively.
 
@@
 abstract interpreter via staging, and then generate efficient low-level code
-that performs the actual analysis. 2) Inspired by the tagless-final interpreters
+that performs the actual analysis. 2) Inspired by tagless-final interpreters
 \cite{DBLP:journals/jfp/CaretteKS09}, we construct a generic interpreter that
-abstracts over binding-times and different semantics, which allows the staged
+abstracts over binding-times \cite{Ofenbeck:2017:SGP:3136040.3136060,
+Amin:2017:CTI:3177123.3158140} and different semantics, which allows the staged
 abstract interpreter to be derived from its unstaged counterpart.  As a result,
@@
 
-The first Futamura projection specifically shows that specializing an
-interpreter w.r.t. the input program yields an equivalent executable. For
+The first Futamura projection shows that specializing an
+interpreter w.r.t. a input program yields an equivalent executable. For
 instance, assume that @eval : Expr --> Input --> Value@ is an interpreter for some language.
@@
 introduce binding-times and make the monads binding-time polymorphic. Our
-proposed framework adopts the type-based multi-stage programming from the
-Lightweight Modular Staging framework, and implements the Futamura projection of
+proposed framework adopts type-based multi-stage programming from the
+Lightweight Modular Staging framework, and implements the Futamura projections of
 an abstract interpreter for a stateful higher-order language. By deriving staged
@@
 fact all firmly rooted in the concrete semantics.  This observation provides a
-way to abstract over the interpreter and achieve the flexibility of
+way to abstract over the semantics and achieve the flexibility of
 reinterpreting the shared interpreter.  The generic interpreter returns a value
@@
 thereby requiring significantly less engineering effort.
-2) We extend the basic staged abstract interpreter to different flow
+2) We extend the basic staged abstract interpreter to cover different flow
 analyses, including a store-widened analysis, a context-sensitive

### 2019-09-11 paper/20_prelim.tex
polish

@@
 
-In this section, we first describe the abstract syntax of the language for our
-interpreters. Then, we present the generic interpreter shared among the four
-different semantics, after which we instantiate the interpreter to the concrete
-one. It is worth noting that we choose to use Scala and a monadic style to
-demonstrate this idea, though the approach is not restricted to this choice.
-Indeed, one can use imperative or direct style in other MSP languages (e.g.,
-MetaOCaml \cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} or
-Template Haskell \cite{Sheard:2002:TMH:636517.636528}) to construct such staged
-abstract interpreters.
+In this section, we first describe the abstract syntax of the language. Then,
+we present the generic interpreter shared among the four different semantics,
+after which we instantiate the interpreter to the concrete one. It is worth
+noting that we choose to use Scala and a monadic style to demonstrate this
+idea, though the approach is not restricted to this choice.  Indeed, one can
+use imperative or direct-style in other MSP languages (e.g., MetaOCaml
+\cite{DBLP:conf/gpce/CalcagnoTHL03, DBLP:conf/flops/Kiselyov14} or Template
+Haskell \cite{Sheard:2002:TMH:636517.636528}) to construct such staged abstract
+interpreters, although the details are different.
 
@@
 numbers, arithmetic, recursion, and conditionals. Other effects such as
-assignments can also be supported readily. Since we are mostly interested in
-analyzing dynamic behavior of programs, we elide the static semantics. We also
+assignments can also be supported readily. Since we are interested in
+analyzing the dynamic behavior of programs, we elide the static semantics. We also
 assume that input programs are well-typed and all variables are distinct. The
@@
 The abstract syntax we present can be seen as a deep embedding of the language:
-we use data-types to represent programs. This design choice allows us to use
-different interpretations over the AST; with the inheritance and overriding
-mechanisms in Scala, we may also add new language constructs and reuse existing
-interpretations.
+we use data-types to represent programs. This is the most natural choice
+for program analysis and allows us to use different interpretations over the
+AST; with the inheritance and overriding mechanisms in Scala, we may also add
+new language constructs and reuse existing interpretations.
 
@@
 an equivalent one using @flatMap@ and @map@. The last binding
-in @for@-comprehensions is translated into a @map@, where the argument of
+in @for@-comprehensions is translated into a @map@ operation, where the argument of
 @yield@ becomes the body expression inside that @map@ operation. The
@@
 
-In this section, we present the generic interface in the style of a big-step
+In this section, we present the generic evaluator interface in the style of a big-step
 definitional interpreter. The key idea is to keep both the binding-time type and
@@
 In later sections, we will instantiate the monadic type to perform concrete
-interpretation \citet{DBLP:conf/popl/LiangHJ95} or abstract interpretation
+interpretation \cite{DBLP:conf/popl/LiangHJ95} or abstract interpretation
 \cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17}.
@@
 \paragraph{Binding-time Abstraction}
-The binding-time type is declared as a higher-kinded type @R[_]@.
+The binding-time type is declared as a higher-kinded type @R[_]@ \cite{Ofenbeck:2017:SGP:3136040.3136060}.
 Now we can use @R@ to annotate other data-types in the interpreter;
 it is also injected into the monadic type @MonadOps@.  If we instantiate
-@R@ as an identity type (i.e., @type R[T] = T@), then the generic interpreter is a
+@R@ as the identity type (i.e., @type R[T] = T@), then the generic interpreter is a
 standard definitional interpreter that will execute the program.  In Section
@@
 Conditionals and arithmetic are handled by @br0@ and @arith@, respectively. The
-method @ap_clo@ is used for applying functions, which takes a function value
+method @ap_clo@ is used for applying functions, by taking a function value
 and an argument value. Note that the @Env@, @Store@, and @Value@ are all
@@
 refer to itself directly. Instead, @eval@ takes an additional parameter @ev@ of
-type @Expr => Ans@ that refers to itself. Consequently, the method @close@ that
+type @Expr => Ans@ that refers to itself. Consequently, method @close@ that
 lifts $\lambda$-terms to closures and method @ap_clo@ that applies functions also
-takes an extra @ev@, because evaluations may happen inside them.
+takes an extra @ev@, because further evaluation may happen inside them.
 To close the open recursion, we use a fixed-point operator @fix@.
@@
 DBLP:conf/ppdp/AgerBDM03}. We first present the concrete components, i.e., the
-value domains, then show the monad stack for concrete interpretation, finally
+value domains, then show the monad stack for concrete interpretation, and finally
 sketch how the primitive operations are implemented.
@@
 type @Env@, and the @StateT@ is parameterized by the type @Store@, and the
-inner-most monad @IdM@ is merely an identity monad.
+inner-most monad @IdM@ is merely the identity monad.
 \begin{lstlisting}
@@
 @S@ is the state type, and @M[_]@ is the inner monad type.
-Given a value of @S@, a @StateT@ monad produces a transformed value of type @M[(A, S)]@,
+Given a value of type @S@, a @StateT@ monad produces a transformed value of type @M[(A, S)]@,
 where the new state (of type @S@) is accompanied with the result (of type @A@).
-Note that for the moment, the binding-time type @R@ is an identity type; thus,
+Note that for the moment, the binding-time type @R@ is the identity type; thus,
 these monads operate on unstaged data. We can also see this from the
@@
 Other primitive operations over the value domains can be implemented
-straightforwardly. We elide most of them but describe a bit on how we handle
+straightforwardly. We elide most of them but describe how we handle
 functions and applications. The reason is that $\lambda$-terms are data but are also
@@
 for the argument and updates the environment and store accordingly.  Finally,
-body expression of the $\lambda$-term is evaluated under the new environment
-and store.
+by recursively calling @ev@, it evaluates the body expression of the
+$\lambda$-term under the new environment and store.
 \begin{lstlisting}

### 2019-09-11 paper/30_stageconc.tex
polish

@@
 We use the same monad stack structure as in the unstaged interpreter: a reader
-monad with a state monad. The difference here is that, now the monads operate
+monad with a state monad. The difference here is that now the monads operate
 on staged values.  For brevity, we call them \textit{staged monads}. In the
@@
 staged, but the whole set is also staged, and we have no knowledge about how
-many elements in the set. Therefore, any traversals over the set has to be
+many elements are in the set. Therefore, any traversals over the set has to be
 staged.  In this section, we explicitly distinguish the two versions of monadic
@@
 Scala function. The evaluator is passed as @ev@ to the method @close@.
-The following code implements the idea:
+The following code implements this idea:
 \begin{lstlisting}
@@
 
-In @close@, we first create a function @f@ that takes a next-stage value and
-store, and returns staged values and stores. The function @f@ is a
-current-stage and we then use @emit_complied_clo@ to denote it to a next-stage
-Scala function.
+In @close@, we first create a function @f@, which input and output are both
+next-stage values and stores.  The function @f@ is a current-stage and we then
+use @emit_complied_clo@ to denote it to a next-stage Scala function.
 The method @emit_compiled_clo@ produces current-stage representations of
@@
 store.  By doing so, we have \textit{collapsed} the monadic values of @AnsM@ to
-grounded values within staging; therefore in the future stage, there is no
-monads.
+grounded values within staging; therefore in the future stage, there will be no
+monadic value.
 \begin{lstlisting}
@@
 whose next-stage form is a Scala function.
-With this fact, we can generate a next-stage function application for it, also
-in our target language Scala. The method @emit_ap_clo@ produces a current-stage
-representation of a next-stage function application, with necessary arguments
+With this knowledge, we can generate a next-stage function application for it. 
+The method @emit_ap_clo@ produces a current-stage
+representation of a next-stage function application, with required arguments
 such as the latest store @σ@. Meanwhile, we still want the method @ap_clo@
@@
 
-Now, compared with the unstaged interpreter, we observe a key difference of where
+Now, compared with the unstaged interpreter, we observe a key difference where
 we use the evaluation function @ev@.  In the unstaged interpreter, the
-invocation of @ev@ happens in @ap_clo@, i.e., at the application time; while in
+invocation of @ev@ happens in @ap_clo@, i.e., at the time of application; while in
 the staged interpreter, we eagerly apply @ev@ when denoting the $\lambda$-terms to
-our next-stage value domains, i.e., at the value-representation time.
+our next-stage value domains, i.e., at the time of value-representation.
 
@@
 In the previous section, we showed the types of @emit_compiled_clo@ and
-@emit_ap_clo@ without concrete implementations. In this section, we briefly
+@emit_ap_clo@ without their concrete implementations. In this section, we briefly
 discuss the IR nodes generated by them and sketch the code generation.

### 2019-09-11 paper/40_absinterp.tex
misc
polish

@@
 
-After seeing the unstaged and staged concrete interpreter, now we turn our
+After seeing the unstaged and staged concrete interpreter, we now turn our
 focus to abstract interpreters under the same framework. We first present a
@@
 
-The abstract domains operated by abstract interpreters usually can be
-formulated by complete lattices.
+Many abstract domains in abstract interpreters can be formulated by complete
+lattices.
 A complete lattice over set $S$ is a tuple $\langle S, \sqsubseteq, \top,
@@
 operator. The trait @SPLattice@ (Figure \ref{fig:splattice}) defines a type
-class of stage polymorphic lattices. Similar to the @MonadOps@ in Section
+class of stage-polymorphic lattices. Similar to the @MonadOps@ in Section
 \ref{generic_if}, we introduce an additional higher-kinded type @R[_]@ to the
@@
 
-In this section, we simply instantiate @SPLattice@ with the flat binding-time
+In this section, we instantiate @SPLattice@ with the flat binding-time
 type @R[T] = T@ (i.e., @NoRep@); the result is the trait @Lattice[S]@.
-An example of the lattices we used in the rest of the paper is the power set
-abstract domain (shown in Figure \ref{fig:powerset}). For two power sets,
-they are ordered by the subset relation. We use set union to compute their
+An example of the lattices we use in the rest of the paper is the power set
+abstract domain (shown in Figure \ref{fig:powerset}). Two power sets are
+ordered by the subset relation. We use set union to compute their
 join, and set intersection to compute their meet.  The bottom element of a
-power set is empty set, and we do not have a representation of top element for
+power set is the empty set, and we do not have a representation of the top element for
 power set.  Other lattices used in the paper, such as products and maps, can be
@@
 point-wise.  Non-relational numerical abstract domains such intervals can also
-be implemented in a stage polymorphic way.
+be implemented in a stage-polymorphic way.
 
@@
 space is constrained to be finite to ensure the computability of the analysis --
-for a simple monovariant analysis, we directly use variable names for
+for a simple monovariant analysis, we directly use variable names as
 addresses. After defining the @Value@ and @Addr@ type, the types of
@@
 branch will be taken, thus a sound treatment is to explore both of the branches.
-Also, at some later point, we need to compute the join of two path.  Another
-nondeterminism comes from function applications, for instance, @f(a)@, the
-possible $\lambda$-terms of @f@ are computed approximately.  As the
-result, we need to store all the possible callees, as well as we will
-retrieve a set of closures when dereferencing an address.
+Also, at some later point, we need to compute the join of two paths.  Another
+source of nondeterminism comes from function applications, for instance,
+consider an application @f(a)@, the possible $\lambda$-terms of @f@ are
+computed approximately.  As the result, we need to store all the possible
+callees, as well as we will retrieve a set of closures when dereferencing an
+address.
 
@@
 monad stack for abstract interpretation. The following @AnsM@ type shows the
-monad stack (we use \hl{light gray} to highlight what have been changed from
+monad stack (we use \hl{light gray} to highlight what has been changed from
 the monad stack of concrete interpretation):
@@
 This monad stack scheme (@AnsM[T]@) leads to a different type of grounded
-result. Recall that in the concrete setting, we only have a reader monad (for
+values. Recall that in the concrete setting, we only have a reader monad (for
 environments) and a state monad (for stores). The reader effect is not
@@
 that the store update operator merges the new values with existing values.  As
-we mentioned before, sometimes we have to explore the both branches for
+we mentioned before, sometimes we have to explore both of the branches for
 conditionals: in method @br0@, we combine the results using the @mplus@
@@
 The difference is that now the first argument of @ap_clo@ is a set of closures
-@funs@, so we can not directly destruct it to a syntactic @Lam@ by pattern
+@funs@, so we cannot directly destruct it with a syntactic @Lam@ by pattern
 matching. Instead, we use an auxiliary function @lift_nd@ that takes a set and
 lifts the elements in the set into the monad stack. Then we can
-straightforwardly implement the @ap_clo@ in monadic style, where the closures
-come from the monads and thus the nondeterminism can be naturally handled. The
-light gray line shows what is added from the concrete @ap_clo@.
+straightforwardly implement the method @ap_clo@ still in monadic style, where the
+closures come from the monads and thus the nondeterminism can be naturally
+handled. The light gray line shows what is added from its concrete counterpart.
 \begin{lstlisting}[escapechar=!]
@@
 the depth-first evaluation. The @in@ cache stores the result from the last
-iteration, and @out@ cache is used for updating in the current iteration. In
-the next iteration, the last @out@ cache will be served as the @in@ cache, and
+iteration, and the @out@ cache is constantly updated during the current iteration. In
+the next iteration, the last @out@ cache is served as the @in@ cache, and
 an empty cache is plugged-in to the @out@ slot. Once the @out@ cache does not

### 2019-09-11 paper/50_stageabs.tex
polish

@@
 their confluence -- a staged abstract interpreter.  Unsurprisingly, the staged
-abstract interpreter in this section uses the same abstract semantics from the
+abstract interpreter in this section uses the same abstract semantics as the
 unstaged version in Section~\ref{unstaged_abs}.  The approach we use to refactor
 the unstaged one to the staged abstract interpreter is modular, and does not
-sacrifice soundness or precision. The designer of the analyzer therefore has no
-need to rewrite the analysis. We first present the staged lattices and staged
+sacrifice soundness or precision. The designer of the analyzer therefore does not
+have to rewrite the analysis. We first present the staged lattices and staged
 monads, and then discuss the staged version of primitive operations, especially
@@
 In Section \ref{stagedpoly_lat}, we exploited the higher-kinded type @R@ to
-achieve the stage polymorphism. Now we instantiate the type @R@ as @Rep@ and
+achieve stage polymorphism. Now we instantiate the type @R@ as @Rep@ and
 still use the power-set as example to describe its staged version.
@@
 The trait @RepLattice@ is shown by instantiating @Lattice@ with the type @Rep@.
-The @RepSetLattice@ for power sets is an instance of @RepLattice@, where the
-lattice operations are eventually delegated to the operations on the staged set
-data type, such as @subsetOf@, @union@ and @intersect@. We do not need to
-change the implementations, except the types -- such as changing them from
-@Set[T]@ to @Rep[Set[T]]@.  We show the changes in light gray code. 
+For all type @T@, method @RepSetLattice@ provides an instance of @RepLattice@
+for @Set[T]@, where the lattice operations are eventually delegated to the
+operations on the staged set data type, such as @subsetOf@, @union@ and
+@intersect@. 
+Compared with the unstaged version, we do not need to change the
+implementations, except the types -- such as changing them from @Set[T]@ to
+@Rep[Set[T]]@.  We show the changes in light gray code. 
 %The LMS library provides implicit conversions that lift the current-stage values to
@@
 \paragraph{Staged Monads for Abstract Interpretation}
-We use the same monad stack structure from the unstaged abstract interpreter,
+We use the same monad stack structure as the unstaged abstract interpreter,
 but replacing \textit{all} the monad transformers with their staged versions. The
@@
 In other words, when staging, we eagerly specialize the interpreter and
-generate code every time when we reach a $\lambda$-term, instead of lazily
+generate code every time we reach a $\lambda$-term, instead of lazily
 calling @ev@ when the applications happen.  Similarly, here we use the same way
@@
 over-approximation of runtime behavior.  In the rest of this part, we keep our
-roadmap and focus on discussing closure representations and function
-applications for the staged version.
+roadmap and focus on discussing the closure representation and function
+application for the staged version.
 
 The @close@ method now is a mix of the \textit{staged} concrete version and
-unstaged \textit{abstract} version. We first build a current-stage function @f@, which
-takes four next-stage values, including the argument and latest store, and
-@in/out@ caches additionally. Inside @f@, we collapse the monadic value of
-type @Ans@ to grounded values of type @Result@, by providing the desired
-arguments, i.e., the new environment, new store, and two caches. The collapsing
-of monadic values happens at the current-stage, so the invocation of @ev@ is unfolded
-within staging. Finally, we generate a singleton set containing the compiled
-next-stage closure (@emit_compiled_clo@), which is represented by an IR node
-in LMS.
+unstaged \textit{abstract} version. We first build a current-stage function
+@f@, which takes four next-stage values, including the argument and latest
+store, and @in/out@ caches additionally. Inside @f@, we collapse the monadic
+value of type @Ans@ to grounded values of type @Result@, by providing the
+desired arguments, i.e., the new environment, new store, and two caches. The
+collapsing of monadic values happens at the current-stage, so the invocation of
+@ev@ is unfolded during staging. Finally, we generate a singleton set
+containing the compiled next-stage closure (@emit_compiled_clo@), which is
+represented by an IR node in LMS.
 \begin{lstlisting}
@@
 monadic value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@
-cache.  After which, we use a similar technique that collapses the monadic
+cache.  After, we use a similar technique that collapses the monadic
 value @m@ to grounded values, by providing its desired environment and etc.
@@
 tree of the input program. In practice, however, it would suffer from code
-explosion when analyzing (specializing) large programs, which increases the
-compile time. If we generate next-stage programs running on JVM, such large
-generated programs would also potentially cause runtime GC overhead. In this
-section, we present optimizations that largely mitigate these issues. For all
-that, implementing these optimizations do not need to change the generic
-interpreter.
+explosion when analyzing (specializing) large programs, which increases
+compile time. If we generate next-stage programs running on the JVM, such large
+generated programs would also incur GC overhead at runtime. In this
+section, we present optimizations that largely mitigate these issues. 
+Implementing all of those optimizations do not need to change the
+generic interpreter.
 
@@
 @Rep[Map]@, which means that the operations on a @Rep[Map]@ directly become
-nex-stage code, and we do not inspect any further inside. As we identified when
+next-stage code, and we do not inspect any further inside. As we identified when
 introducing the generic interface, the keys of an @Env@ are string-represented

### 2019-09-14 paper/10_intro.tex
polish

@@
 can be tailored to different language features (such as state, first-class
-control, exceptions, concurrency, etc.) and sensitivity analyses
+control, exceptions, etc.) and sensitivity analyses
 \cite{DBLP:conf/icfp/Gilray0M16, DBLP:conf/popl/GilrayL0MH16,
@@
 Keidel:2018:CSP:3243631.3236767}.
-
-%Based on the idea of abstracting abstract machines, more
-%pragmatically, several implementation strategies utilizing purely
-%functional programming techniques to build abstract interpreters have
-%emerged.
-
 Based on the idea of AAM, more pragmatically, several implementation strategies
@@
 Sergey:2013:MAI:2491956.2491979}, arrows \cite{Keidel:2018:CSP:3243631.3236767},
-extensible effects \cite{Kiselyov:2015:FMM:2804302.2804319, Githubsemantic},
+extensible effects \cite{Githubsemantic},
 etc. These pure approaches provide certain benefits, including the fact that the
@@
 fashion. Furthermore, referential transparency allows programmers to confidently
-reason about the correctness of their implementations.  The soundness of an
+reason about the correctness. The soundness of an
 analysis can also be proven with less effort, whether by mechanized proofs
@@
 their merits, they are significantly slower when compared with imperative,
-stateful implementations.  In order to reconcile these approaches, one need to
-use a programming technique which allows for high-level abstractions during
-implementation, but bypasses the incurred overhead at runtime.
+stateful implementations. The goal of this paper is to develop an approach for
+implementing abstract interpreters, which allows high-level abstractions during
+implementation for clarity, but passes the incurred overhead at runtime for
+efficiency.
+%In order to reconcile these approaches, one need to
+%use a programming technique which allows for high-level abstractions during
+%implementation, but bypasses the incurred overhead at runtime.
 
@@
 compilers through a hierarchy of specializations, known since then as the
-Futamura projections. The first Futamura projection states that specializing an
-interpreter to an input program removes interpretative overhead and yields an
-equivalent executable that runs faster. In recent years, the idea of Futamura
-projections has attracted more attention, and has been successfully applied to
-many real-world problems, including PyPy \cite{Bolz:2009:TMP:1565824.1565827},
-Truffle \cite{Marr:2015:TVP:2814270.2814275}, query compilers
-\cite{DBLP:conf/sigmod/TahboubER18}, etc. The challenge here: can we apply
-Futamura projections to a general abstract interpreter to make it run faster?
-The solution is not obvious, as we need to specialize a nonstandard semantics
-that extensively uses nondeterminism, operates on abstract domains, and
-computes fixed-points iteratively.
+Futamura projections. 
+The first Futamura projection states that specializing an interpreter to an
+input program removes interpretative overhead and yields an equivalent
+executable that runs faster. Moreover, a procedure that can specialize an
+interpreter to any input program is equivalent to a compiler.
+In recent years, the idea of Futamura projections has attracted more attention,
+and has been successfully applied to many real-world scenarios, such as
+building JIT compilers for dynamic languages
+\cite{Bolz:2009:TMP:1565824.1565827, Marr:2015:TVP:2814270.2814275} and building query
+compilers \cite{DBLP:conf/sigmod/TahboubER18}.  The question here is can we
+apply Futamura projections to an abstract interpreter and obtain a compiler
+for abstract interpretation?
+%apply Futamura projections to a general abstract interpreter to make it run faster? 
+The solution is not obvious, as we need to specialize a nonstandard
+semantics that extensively uses nondeterminism, operates on abstract domains,
+and computes fixed-points iteratively.
 
@@
 
-The first Futamura projection shows that specializing an
-interpreter w.r.t. a input program yields an equivalent executable. For
-instance, assume that @eval : Expr --> Input --> Value@ is an interpreter for some language.
-Given a program $e_0$@ : Expr@, by applying the specialization we can
+The first Futamura projection shows that specializing an interpreter w.r.t. an
+input program yields an equivalent program. For instance, assume that 
+@eval : Expr --> Input --> Value@ is an interpreter for some language.
+Given a program $e_0$@ : Expr@, by applying the specialization we
 obtain a specialized program $\texttt{eval}_{\texttt{e0}}$ : @Input --> Value@.
 %which is the so-called \textit{equivalent executable}.
-% By definition of the interpreter, they produce the same result when applied with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
-The first proposed approach to realizing Futamura projections is partial
-evaluation \cite{DBLP:books/daglib/0072559}, which relies on an automatic
-binding-time analysis (\textit{static} or \textit{dynamic}).  The underlying
-representation can be viewed as a \textit{two-level semantics}
-\cite{NIELSON1989117, NIELSON198859} that distinguishes between compile-time
-computation and run-time computation.
-However, it is hard to precisely analyze binding-times given an arbitrary
-program. As an alternative approach, multi-stage programming (MSP)
-\cite{taha1999multi, DBLP:conf/pepm/TahaS97} allows users to explicitly program
-over a two-level syntax \cite{Nielson:1992:TFL:130665} by manually annotating
-the binding-times of variables. The MSP system will then check whether these
-annotations are consistent, and specialize the program using that information.
-The staging annotations can be either syntactic (e.g., quote and quasi-quote in
-MetaML/MetaOCaml) or type-based (e.g., the Lightweight Modular Staging framework
-\cite{DBLP:conf/gpce/RompfO10} in Scala).
+% By definition of the interpreter, they produce the same result when applied
+% with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
+The earlier approach to realizing Futamura projections is partial
+evaluation (PE) \cite{DBLP:books/daglib/0072559}.
+Offline PE relies on a separate binding-time analysis (\textit{static} or
+\textit{dynamic}), and online PE identifies the binding-times during the
+specialization.
+The underlying representation of binding-times can be viewed as a
+two-level semantics \cite{NIELSON1989117, NIELSON198859,
+Nielson:1992:TFL:130665} that distinguishes between compile-time computation
+and run-time computation.  However, it is hard to precisely analyze
+binding-times given an arbitrary program. As an alternative approach,
+multi-stage programming (MSP) \cite{taha1999multi, DBLP:conf/pepm/TahaS97}
+allows programmers to annotate the binding-times of variables and control which
+part of the program should be specialized.
+At the syntactic level, traditional MSP languages (e.g., MetaML/MetaOCaml) use
+quasi-quotations and escapes as annotations. However, the syntactic
+quasi-quotation is a weak two-level semantics, and does not always preserve the
+evaluation order of staged expressions.
+%The MSP system will then check whether these
+%annotations are consistent, and specialize the program using that information.
+Besides the syntactic approach, staging annotations can also be purely
+type-based.  One of the examples is the Lightweight Modular Staging framework
+\cite{DBLP:conf/gpce/RompfO10} in Scala, which provides a stronger guarantee of
+evaluation order by generating code in administrative normal form.
 
@@
 
-\paragraph{Contributions} Briefly, the contributions of the paper are as follows:
-\begin{itemize}[leftmargin=2em]
-  \item Intellectually, we present a framework that naturally extends the first
-    Futamura projection to definitional abstract interpreters, showing a
-    well-grounded approach to optimizing static analyses via meta-programming.
-  \item Practically, we show that staging an abstract interpreter is useful to
-    improve performance and scalability of analyses by case studies and an
-    empirical evaluation.
-\end{itemize}
+\paragraph{Contributions} 
+1) Intellectually, we present a framework that naturally extends the first
+Futamura projection to abstract interpreters, showing a well-grounded approach
+to optimizing static analyses via meta-programming.  2) Practically, we show
+that staging an abstract interpreter is useful to improve performance and
+scalability of analyses by case studies and an empirical evaluation.
 

### 2019-09-14 paper/20_prelim.tex
polish

@@
 assignments can also be supported readily. Since we are interested in
-analyzing the dynamic behavior of programs, we elide the static semantics. We also
-assume that input programs are well-typed and all variables are distinct. The
+analyzing the dynamic behavior of programs, we elide the precise static semantics. We
+assume that programs are well-typed and variables are distinct. The
 abstract syntax is as follows:
@@
 Monad transformers are type constructors of kind @(* --> *) --> (* --> *)@, which
-take a monad as an argument and produce another monad. By using monad
+take a monad type as argument and produce another monad type. By using monad
 transformers, we can combine multiple monads into a single one.  The
@@
 
-\paragraph{The Interpreter} We can now construct the semantics-agnostic interpreter
+\paragraph{The Interpreter} We can now construct the generic interpreter
 in monadic form, as shown in Figure \ref{fig:shared_int}. The basic idea

### 2019-09-14 paper/50_stageabs.tex
polish

@@
 
-Now we can implement the staged abstract semantics, under that the binding-time
-type @R@ is instantiated as @Rep@. The types of abstract components are reused
+Now we can implement the staged abstract semantics, with the binding-time
+type @R@ instantiated as @Rep@. The types of abstract components are reused
 from the unstaged version.
@@
 interpreter, we have to handle nondeterminism incurred by the
-over-approximation of runtime behavior.  In the rest of this part, we keep our
-roadmap and focus on discussing the closure representation and function
+over-approximation of runtime behavior.  In the rest of this part, we 
+focus on discussing the closure representation and function
 application for the staged version.
@@
 collapsing of monadic values happens at the current-stage, so the invocation of
-@ev@ is unfolded during staging. Finally, we generate a singleton set
+@ev@ is unfolded at staging time. Finally, we generate a singleton set
 containing the compiled next-stage closure (@emit_compiled_clo@), which is
@@
 monadic value @m@ of type @Ans@ first, which evaluates @e@ under the new @out@
-cache.  After, we use a similar technique that collapses the monadic
-value @m@ to grounded values, by providing its desired environment and etc.
-Finally, we have a current-stage representation of the future-stage values
-@res@, and we reify the content of @res@ back into the current-stage monad stack.
+cache.  After this, we use a similar technique that eagerly collapses the
+monadic value @m@ to grounded values, by providing its desired environment and
+other arguments.  Finally, we have a current-stage representation of the
+future-stage values @res@, and we reify the content of @res@ back into the
+current-stage monad stack.
 
@@
 \subsection{A Little Bit of Code Generation, Again}
-The code generation for compiled closures and function applications are similar
+The code generation for compiled closures and function applications is similar
 to their counterparts in the staged concrete interpreter. We have two IR nodes
@@
 Our staging schema works by unfolding the interpreter over the abstract syntax
-tree of the input program. In practice, however, it would suffer from code
-explosion when analyzing (specializing) large programs, which increases
-compile time. If we generate next-stage programs running on the JVM, such large
-generated programs would also incur GC overhead at runtime. In this
+tree of the input program. In practice, however, the staging schema would
+suffer from code explosion when analyzing (specializing) large programs, which
+increases compile time. If we generate next-stage programs running on the JVM,
+such large generated programs would also incur GC overhead at runtime. In this
 section, we present optimizations that largely mitigate these issues.
-Implementing all of those optimizations do not need to change the
-generic interpreter.
+Implementing all of those optimizations do not need to change the generic
+interpreter.
 
@@
 If all the keys of type @K@ are all known statically, then the indices for
-those keys can also be computed statically. Thus, the specialized map would be
-an array of type @Array[Rep[V]]@, whose elements are next-stage values, and the
-size of the array is known statically as the program has finite number of
-identifiers. As result, an access to the map is staged into an access to the array
-with the pre-computed index.
+those keys can also be computed statically. Thus, in this case, the specialized
+map will be an array of type @Array[Rep[V]]@, whose elements are next-stage
+values, and the size of the array is known statically as the program has finite
+number of identifiers. As result, an access to the map is staged into an access
+to the array with a pre-computed index.
 
-Particularly, if we are specializing a monovariant analysis, the address space
+In particular, if we are specializing a monovariant analysis, the address space
 is equivalent to the set of all identifiers in the program. Utilizing this fact, the
 result of accesses to the environment can be computed statically and we may
-generate addresses directly during staging. After which, the store can be
+generate addresses directly during staging. After this, the store can be
 specialized as an array by using the way mentioned above.
@@
 expressions such as literals or variables, because they always terminate.
-Borrowing the idea from the partition of expressions used in A-Normal Form
-$\lambda$-calculus \cite{Flanagan:1993:ECC:155090.155113}, we can use a
-selective caching algorithm that does not generate caching code for atomic
-expressions:
+Borrowing the idea from the partition of expressions used in 
+administrative normal form (ANF) $\lambda$-calculus
+\cite{Flanagan:1993:ECC:155090.155113}, we can use a selective caching
+algorithm that does not generate caching code for atomic expressions:
 \begin{lstlisting}
@@
 \paragraph{Partially-static Data}
-Our treatment of binding-times is coarse-grained: @Expr@s is static, the rest
+Our treatment of binding-times is coarse-grained: @Expr@s are static, the rest
 of the components are all dynamic. But this is not always true, because the
 static data have to be used somewhere with the dynamic operations.
-Partial-static data is a way to improve binding-times and optimize the
+Partially-static data is a way to improve binding-times and optimize the
 generated code.
-For example, to fold a static singleton set (often appears in @SetT@), e.g.,
-@Set(x).foldLeft(init)(f)@ where @x@ and @init@ are staged values, a naive code
-generator would faithfully generate code that applies @foldLeft@ with the set
-and function @f@. But we can also utilize algebraic properties of @foldLeft@
-to generate cheaper code, e.g., @Set(x).foldLeft(init)(f) = f(init, x)@. Since the
+For example, to left-fold a static singleton set (often appears in @SetT@), e.g.,
+@Set(x).fold(init)(f)@ where @x@ and @init@ are staged values, a naive code
+generator would faithfully generate code that applies @fold@ with the set
+and function @f@. But we can also utilize algebraic properties of @fold@
+to generate cheaper code, e.g., @Set(x).fold(init)(f) = f(init, x)@. Since the
 function @f@ is known at the current stage, we completely eliminate the fold
 operation and function application. We apply several rewritings enabled by
-partial-static patterns, such as for @Set@, @Map@, and @Tuple@. This
+partially-static patterns, such as for @Set@, @Map@, and @Tuple@. This
 optimization greatly reduces the size of residual programs.
@@
 review and summarize our recipe to achieve the staged abstract interpreter and
-discuss correctness issues and different design choices.
+discuss correctness and different design choices.
 
@@
     appropriate monad stack. This step has been explored in the previous
-    literatures.
+    literature.
   \item Finally, we replace the monad stack with a staged monad stack, and
@@
 Soundness is the central concern of static analysis, and as such, is vital for
-prospective users. Our approach does not touch the soundness of the analysis,
+prospective users. Our approach does not interact with the soundness of the analysis,
 i.e, if the unstaged one does not produces false negative result, the staged
 one also does not. This soundness preservation indeed relies on the correctness
-of the staged implementations and of the underlying MSP system. We now briefly
+of the staged implementations and the underlying MSP system. We now briefly
 examine how this is achieved. Note that the rationale listed
-here is based on empirical evidence; a rigorous proof for soundness
+here is based on empirical evidence; a rigorous proof of soundness
 preservation of our approach remains an open challenge.
@@
 \begin{itemize}
-  \item We usually assume the underlying MSP system preserves the meaning
-    during staging, i.e., the generated program and original program produce
-    the same result given the same arguments.
+  \item 
+    In general, syntactic MSP systems based on quotations (e.g.,
+    MetaML/MetaOCaml) cannot guarantee that the generated code preserves the
+    meaning after staging. In other words, the generated code may not preserve
+    the evaluation order or may contain undesired duplication of code.
     In this paper, we use the LMS framework, which 1) checks binding-times by
     checking types which is enabled by the Scala compiler, and 2) generates
-    code in A-Normal form \cite{Flanagan:1993:ECC:155090.155113} that preserves
-    the execution order within a stage \cite{DBLP:conf/birthday/Rompf16}.
-    However, in general, it is possible to subvert these guarantees by using
-    unsafe features, such as casts.
+    code in administrative normal form \cite{Flanagan:1993:ECC:155090.155113}
+    that preserves the execution order within a stage
+    \cite{DBLP:conf/birthday/Rompf16}.
+    %However, in general, it is possible to subvert these guarantees by using
+    %unsafe features, such as casts.
   \item As shown in the previous sections, the generic interpreter is untouched
     and shared by the four artifacts.  This allows the programmer to
-    check the correctness of the implementation of staged monads modularly.
+    check the implementation of staged monads and primitive operations,
+    as well as their correctness modularly.
    % We only instantiate binding-time annotations and the monadic types. 
@@
     versions. For instance, the staged data structures we use here are simply
-    black boxes that wrap the data structures in Scala library.
+    black boxes that wrap the data structures in the Scala library.
   \item In our experiments, the staged abstract interpreter produces the same
-    result with the unstaged one on all benchmark programs we tested.
+    result as the unstaged one on all benchmark programs we tested.
 \end{itemize}
@@
 \paragraph{Monadic-style vs Direct-style} We use a monadic interpreter
-throughout the paper, but it is not necessary for staging. One can inline the
-monadic operations and obtain an abstract interpreter in continuation-passing
-style, or even translate back to a direct-style that may use explicit
-side-effects such as mutations. In either case, we can still apply the staging
-idea to the abstract interpreter and remove the interpretation overhead.
-However, monads allow the staged abstract interpreter to be implemented in a
-modular and extensible way.
+throughout the paper, but using monadic-style is not the only choice for
+staging. One can inline the monadic operations and obtain an abstract
+interpreter in continuation-passing style, or even translate back to a
+direct-style that may use explicit side-effects such as mutations. In either
+case, we can still apply the staging idea to the abstract interpreter and
+remove the interpretation overhead.  However, monads allow the staged abstract
+interpreter to be implemented in a modular and extensible way.
 

### 2019-09-14 paper/60_cases.tex
polish

@@
 In Section~\ref{sai}, we use a toy language to show that staging an abstract
-interpreter is feasible and the artifact can be derived by combining the staged
-concrete interpreter and unstaged abstract interpreter.
+interpreter is feasible and that a staged artifact can be derived by combining
+the staged concrete interpreter and unstaged abstract interpreter.
 In this section, we conduct several case studies to further show that this
-methodology is practically useful and widely applicable to diverse analyses.
+methodology is practically useful and widely applicable to a diverse set of
+analyses.
 
@@
 The analyzer is a big-step control-environment abstract interpreter. 
-By applying AC, we refactor every function in the analyzer such that it
-returns a closure taking an environment argument.  Therefore the overhead
-of traversing the AST of input program is eliminated.
+By applying AC, every function in the analyzer is refactored to return a
+closure only taking an environment argument.  Therefore the overhead of
+traversing the AST of input program is eliminated.
 
@@
 same improvement with regards to speed, but without changing a single line of
-the analyzer program using type-based staging annotations. However, AC requires
-more engineering effort, i.e., rewriting the whole analyzer into the
-closure-generation form. Moreover, as shown in Section~\ref{staged_ds}, our
-approach is able to not only remove the interpretive overhead, but also enable
-several optimizations such as specialized data structures.
+the analyzer program relying on type-based staging annotations. With Scala's
+type inference, we only need to add staging annotations for function argument
+types and return types.
+However, AC requires more engineering effort, i.e., rewriting the whole
+analyzer into the closure-generation form. Moreover, as shown in
+Section~\ref{staged_ds}, our approach is able to not only remove the
+interpretive overhead, but also enable several optimizations such as
+specialized data structures.
 
@@
 staged monads, which is a common optimization in CFA. Finally, we integrate
-abstract garbage collection with staging. All of them are orthogonal to
+abstract garbage collection with staging. All of these extensions are orthogonal to
 staging.
@@
 
-The abstract addresses we introduced in Section \ref{sai} are simply identical
+The abstract addresses we introduced in Section \ref{sai} are identical
 to the variable names, which is a context-insensitive schema.  Through tweaking
@@
 
-The timestamp is essentially a finite list of expressions that tracks the $k$-most
+The timestamp is a finite list of expressions that tracks the $k$-most
 recent calling contexts. The definition of abstract addresses @KCFAAddr@ is also
@@
 @StateT@ monad. The result type is accordingly updated to contain the set of
-tuples, where the time is added.
+tuples, where the field of @Time@ is added.
 \begin{lstlisting}
@@
 A commonly-used technique to improve the running time of control-flow analyses
-is store-widening. As shown by \cite{Darais:2015:GTM:2814270.2814308,
+is store-widening. As shown by \citet{Darais:2015:GTM:2814270.2814308,
 DBLP:journals/pacmpl/DaraisLNH17}, a store-widening analysis can be easily
 implemented in the monadic abstract interpreter by swapping the @StateT@ and
-the @SetT@ monad transformer. This optimization is orthogonal to staging.
-Within the monadic interpreter framework, we just need to similarly swap the
-same \textit{staged} version of monads. After this adjustment, conceptually, we
+the @SetT@ monad transformers. This optimization is orthogonal to staging.
+Within the monadic interpreter framework, we just need to swap the
+same \textit{staged} version of monad transformers. After this adjustment, conceptually, we
 have the following monad stack:
@@
 Abstract garbage collection is a technique to reclaim unreachable addresses in
-the store while doing the abstract interpretation
+the store while performing the abstract interpretation
 \cite{Might:2006:IFA:1159803.1159807}. \citet{DBLP:journals/pacmpl/DaraisLNH17}
@@
 expressions, which is elided in the code snippet; for the details of this part,
-readers may refer to \cite{DBLP:journals/pacmpl/DaraisLNH17}.
+readers may refer to \citet{DBLP:journals/pacmpl/DaraisLNH17}.
 
@@
 analyzing libraries. Some static analyzers compute a summary for a function or
-a module, which can be reused later. However, when lacking the context
-information, those analyses can be too conservative or unsound, which both lead
-to imprecision. In this section, we show that meta-programming can provide a
+a module, which can be reused later. However, when lacking context
+information, those analyses can be too conservative or unsound. 
+In this section, we show that meta-programming can provide a
 trade-off between summary-based modular analyses and slow whole-program
@@
 As we have already seen in Section \ref{sai}, the primitive operation @close@
-can denote a $\lambda$-term to a next-stage Scala function. Similar to
-compilers, the specializing of the staged abstract interpreter to a
+can denote a $\lambda$-term to a next-stage Scala function. 
+The specialization of the staged abstract interpreter to a
 $\lambda$-term is also independent of other parts of the analyzed program.

### 2019-09-14 paper/80_related.tex
polish

@@
 small-step abstract interpreter in state-transition style. The program
-is firstly compiled to an ``abstract bytecode'' IR. The IR consists of higher-order
+is first compiled to an ``abstract bytecode'' IR. The IR consists of higher-order
 functions, which will be executed later on an abstract abstract machine for
@@
 \paragraph{Construction of Abstract Interpreters} Abstract interpretation is a
-semantics-based approach to build sound static analysis by approximation
+semantics-based approach of sound static analyses by approximation
 \cite{DBLP:conf/popl/CousotC77, Cousot98-5}.
@@
 big-step abstract interpreters with delimited continuations was shown by
-\citet{Wei:2018:RAA:3243631.3236800}; \citet{Keidel:2018:CSP:3243631.3236767}
+\citet{Wei:2018:RAA:3243631.3236800}.
+\citet{Keidel:2018:CSP:3243631.3236767}
 presented a modular, arrow-based abstract interpreter that allows soundness
-proof to be constructed compositionally.
-\citet{DBLP:conf/cc/CousotC02} proposed an abstract interpretation framework for
-modular analysis, from where we borrow the notation of modular analysis.
-However, the staged abstract interpreter does not generate complete summaries for
-modules; instead, it specializes the analysis modularly. Similarly, AC
-is considered as an example of this kind of modular analyses \cite{DBLP:conf/cc/CousotC02}.
+proofs to be constructed compositionally.  \citet{DBLP:conf/cc/CousotC02}
+proposed an abstract interpretation framework for modular analysis. In Section
+\ref{modular}, we borrow the notation of modular analysis.  However, the staged
+abstract interpreter does not generate complete summaries for modules; instead,
+it specializes the analysis modularly. Similarly, AC is considered as an
+example of this kind of modular analyses \cite{DBLP:conf/cc/CousotC02}.
 
@@
 Partial evaluation is an automatic technique for program
-specialization and was studied comprehensively by
-\citet{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
+specialization \cite{10.1007/3-540-61580-6_11, DBLP:books/daglib/0072559}.
 In this paper, we use mutli-stage programming to specialize programs.
@@
   DBLP:conf/popl/KiselyovBPS17}, query compilation in databases
-\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18},
+\cite{DBLP:conf/osdi/EssertelTDBOR18, DBLP:conf/sigmod/TahboubER18, Rompf:2015:FPS:2784731.2784760},
 etc.
@@
 close relation between interpreters and compilers, which was
-originally proposed by \citet{futamura1971partial}, and later reprinted \cite{Futamura1999}.
-\citeauthor{Amin:2017:CTI:3177123.3158140} considered a tower of concrete
+originally proposed by \citet{futamura1971partial} (reprinted as \cite{Futamura1999}).
+\citet{Amin:2017:CTI:3177123.3158140} considered a tower of concrete
 interpreters and showed how to collapse them by using MSP -- it would be interesting
@@
 LMS, and the memoizing monad is just a combination of a reader monad and a state monad
-operating on staged data.  Similar to the idea in this paper, specializing
-monadic interpreters was explored by \citet{danvy1991compiling,
-DBLP:conf/dsl/SheardBP99}. However, they are different from the performance
+operating on staged data.  Similar to the ideas in this paper, specializing
+monadic interpreters was explored by \citet{danvy1991compiling} and
+\citet{DBLP:conf/dsl/SheardBP99}. However, these work are different from the performance
 motivation here.

### 2019-09-14 paper/paper.tex
conc
subtitle
polish

--- 2019-09-10 paper/paper.tex
@@
 \title{Staged Abstract Interpreters}         %% [Short Title] is optional;
+\subtitle{Fast and Modular Whole-Program Analysis via Meta-programming}                     %% \subtitle is optional
                                         %% when present, will be used in
@@
                                         %% contents suppressed with 'anonymous'
-%\subtitle{Fast and Compositional Whole-Program Analysis for Free}                     %% \subtitle is optional
-\subtitle{Fast and Modular Whole-Program Analysis via Meta-programming}                     %% \subtitle is optional
 %\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
@@
                                         %% contents suppressed with 'anonymous'
-
-\iffalse
-\author
-[Guannan Wei, Yuxuan Chen, Tiark Rompf]
-{
-\vspace{-2ex}
-Guannan Wei, Yuxuan Chen, Tiark Rompf\\
-Purdue University
-\vspace{-0.5ex}
-}
-\fi
 
@@
   We thank the reviewers of PLDI '19 and OOPSLA '19, as well as James Decker,
-  Guanhong Tao, Gr\'egory Essertel, and Fei Wang for their feedback on the early
-  drafts. We thank David Van Horn and David Darais for their encouragement on
-  this work.
+  Guanhong Tao, Gr\'egory Essertel, Fei Wang, Qianchuan Ye, Yapeng Ye, and Le Yu
+  for their feedback on early drafts. We thank David Van Horn and David
+  Darais for their encouragement on this work.
   This work was supported in part by NSF awards 1553471, 1564207, 1918483, DOE

### 2019-09-14 paper/20_prelim.tex
space

@@
     \end{lstlisting}
-    \vspace{-0.5em}
-    \caption{trait \texttt{Monad}} \label{fig:monad}
   \end{subfigure}
@@
     \end{lstlisting}
-    \vspace{-0.5em}
-    \caption{trait \texttt{MonadOps}} \label{fig:monadops}
   \end{subfigure}
+  \vspace{-1em}
+  \caption{trait \texttt{Monad} (left) and trait \texttt{MonadOps} (right)} \label{fig:monad} \label{fig:monadops}
 \end{figure}
@@
 addresses and (potentially staged) values, and updates the store accordingly.
-
-%\vspace{-1em}
-\begin{figure}[h!]
-  \centering
-\vspace{-1em}
-  \begin{subfigure}[b]{0.45\textwidth}
 \begin{lstlisting}
-  type MonadOps[R[_], M[_], A] = {
-    def map[B](f: R[A] => R[B]): M[B]
-    def flatMap[B](f: R[A] => M[B]): M[B]
-  }
-
-  type AnsM[T] <: MonadOps[R, AnsM, T]
-  type Ans = AnsM[Value]
+  type MonadOps[R[_], M[_], A] = {           // Environment operations
+    def map[B](f: R[A] => R[B]): M[B]        def ask_env: AnsM[Env]
+    def flatMap[B](f: R[A] => M[B]): M[B]    def local_env(ans: Ans)(ρ: R[Env]): Ans
+  }                                          // Store operations
+                                             def get_store: AnsM[Store]
+  type AnsM[T] <: MonadOps[R, AnsM, T]       def put_store(σ: R[Store]): AnsM[Unit]
+  type Ans = AnsM[Value]                     def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
 \end{lstlisting}
-  \end{subfigure}
-  ~
-  \begin{subfigure}[b]{0.55\textwidth}
-    \begin{lstlisting}
-// Environment operations
-def ask_env: AnsM[Env]
-def local_env(ans: Ans)(ρ: R[Env]): Ans
-// Store operations
-def get_store: AnsM[Store]
-def put_store(σ: R[Store]): AnsM[Unit]
-def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
-    \end{lstlisting}
-  \end{subfigure}
-\end{figure}
-\vspace{-1em}
 

### 2019-09-14 paper/20_prelim.tex
space

@@
     type Addr = Int;  sealed trait Value
-    case class IntV(i: Int) extends Value
-    case class CloV(λ: Lam, e: Env) extends Value
+    case class IntV(i: Int) extends Value;  case class CloV(λ: Lam, e: Env) extends Value
     type Result = (R[Value], R[Store])
@@
   trait ConcreteSemantics extends ConcreteComponents {
-    type R[T] = T
-    type AnsM[T] = ReaderT[StateT[IdM, Store, ?], Env, T] // the monad stack
+    type R[T] = T;  type AnsM[T] = ReaderT[StateT[IdM, Store, ?], Env, T] // the monad stack
     ... }

### 2019-09-14 paper/50_stageabs.tex
space

@@
   res <- lift_nd[(Set[ValSt], Cache)](Set(
-      // a next-stage value of type Rep[(Set[ValSt], Cache)] generated by if
-      if (out.contains(cfg)) (out(cfg), out)
-      else { val m: Ans = for {
+    if (out.contains(cfg)) (out(cfg), out)  // a next-stage value of type Rep[(Set[ValSt], Cache)]
+    else { val m: Ans = for {               //   generated by the if/else
              _ <- put_out_cache(out + (cfg → in.getOrElse(cfg, ⊥)))
@@
 \end{lstlisting}
-\vspace{-0.5em}
-\caption{The staged co-inductive caching algorithm.}
-\label{fig:staged_coind_cache}
+\vspace{-1em}
+\caption{The staged co-inductive caching algorithm.} \label{fig:staged_coind_cache}
 \vspace{-1.5em}

### 2019-09-15 paper/10_intro.tex
polish

@@
 \cite{Bolz:2009:TMP:1565824.1565827, Marr:2015:TVP:2814270.2814275} and building query
-compilers \cite{DBLP:conf/sigmod/TahboubER18}.  The question here is can we
-apply Futamura projections to an abstract interpreter and obtain a compiler
-for abstract interpretation?
+compilers for databases \cite{DBLP:conf/sigmod/TahboubER18, DBLP:conf/osdi/EssertelTDBOR18}.
+The question here is can we apply Futamura projections to an abstract
+interpreter and obtain a compiler for abstract interpretation?
 %apply Futamura projections to a general abstract interpreter to make it run faster? 
@@
 @eval : Expr --> Input --> Value@ is an interpreter for some language.
-Given a program $e_0$@ : Expr@, by applying the specialization we
-obtain a specialized program $\texttt{eval}_{\texttt{e0}}$ : @Input --> Value@.
+Given a program $e$@ : Expr@, by applying the specialization we
+obtain a specialized program $\texttt{eval}_{\texttt{e}}$ : @Input --> Value@.
 %which is the so-called \textit{equivalent executable}.
 % By definition of the interpreter, they produce the same result when applied
-% with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $.
+% with the argument $\texttt{eval}_{\texttt{e0}}(arg) = [\![ e ]\!] arg $.
 The earlier approach to realizing Futamura projections is partial
@@
 type-based.  One of the examples is the Lightweight Modular Staging framework
-\cite{DBLP:conf/gpce/RompfO10} in Scala, which provides a stronger guarantee of
-evaluation order by generating code in administrative normal form.
+\cite{DBLP:conf/gpce/RompfO10} in Scala, which as well provides a stronger
+guarantee of preserving the evaluation order by generating code in
+administrative normal form.
 
@@
 proposed framework adopts type-based multi-stage programming from the
-Lightweight Modular Staging framework, and implements the Futamura projections of
+Lightweight Modular Staging framework, and implements the first Futamura projection of
 an abstract interpreter for a stateful higher-order language. By deriving staged

### 2019-09-15 paper/20_prelim.tex
polish

@@
 for monadic operations through @for@-comprehension.  For example, an object of
-@List[A]@ is an instance of the @List@ monad, with element type @A@.
+type @List[A]@ is an instance of the list monad, with element type @A@.
 To compute the Cartesian product of two lists of numbers, we can use Scala's
@@
 \begin{lstlisting}
-  val xs = List(1, 2); val ys = List(3, 4)
+  val xs = List(1, 2); val ys = List(3, 4); 
   for { x <- xs; y <- ys } yield (x, y)   // List((1,3), (1,4), (2,3), (2,4))
@@
 an equivalent one using @flatMap@ and @map@. The last binding
-in @for@-comprehensions is translated into a @map@ operation, where the argument of
+of @for@-comprehensions is translated into a @map@ operation, where the argument of
 @yield@ becomes the body expression inside that @map@ operation. The
-bindings before the last one are all translated into calls of @flatMap@:
+bindings before the last one are all translated into @flatMap@:
 \begin{lstlisting}
@@
 
-Note that here the monadic object @List[_]@ encapsulates the data internally.
+Note that here the monadic object of type @List[_]@ encapsulates the data internally.
 Therefore, it only exposes the simplified version of @flatMap@, where the monadic
@@
     type R[_] // Binding-time as a higher-kinded type
-    ... // The definitions in the rest of this section are enclosed in trait Semantics.
-  }
+    ...  }
 \end{lstlisting}
@@
 The binding-time type is declared as a higher-kinded type @R[_]@ \cite{Ofenbeck:2017:SGP:3136040.3136060}.
-Now we can use @R@ to annotate other data-types in the interpreter;
-it is also injected into the monadic type @MonadOps@.  If we instantiate
-@R@ as the identity type (i.e., @type R[T] = T@), then the generic interpreter is a
-standard definitional interpreter that will execute the program.  In Section
-\ref{stagedinterp}, we instantiate @R@ using LMS's built-in next-stage
-type annotation @Rep@, which makes the interpreter act as a compiler.
+Now we can use @R@ to annotate other data types in the interpreter.
+Later, the binding-time information is also injected into the monadic type
+@MonadOps@.  If we instantiate @R@ as the identity type (i.e., @type R[T] = T@), 
+then the generic interpreter is a standard definitional interpreter that
+will execute the program.  In Section \ref{stagedinterp}, we instantiate @R@
+using LMS's built-in next-stage type annotation @Rep@, which makes the
+interpreter act as a compiler.
 
@@
 \texttt{CloV} that contains a $\lambda$-term and an environment. The final
-result of the interpreter is a pair of values and stores.
+result of the interpreter is a @Value@ paired with a @Store@.
 To distinguish from the monadic values of type @Ans@ produced by the
-interpreter, later we use term \textit{grounded values} to denote such final values.
+interpreter, later we use term \textit{grounded values} to denote such final values,
+which can be obtained by running the monads.
 The two elements in the type @Result@ are annotated by the binding-time @R@,

### 2019-09-15 paper/30_stageconc.tex
polish

@@
 runtime code generation in a type-safe manner. Different from the approach of
-MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14, DBLP:conf/gpce/CalcagnoTHL03} which
-uses syntactic quotations, LMS distinguishes binding-times based on types. 
-To achieve this, LMS provides a type constructor @Rep[T]@, where @T@ can be an
-arbitrary type, indicating that a value of type @T@ will be known at a future
-stage.  Operations acting on @Rep[T]@ expressions will be residualized as
-generated code.
+MetaML/MetaOCaml \cite{DBLP:conf/flops/Kiselyov14,
+DBLP:conf/gpce/CalcagnoTHL03} which uses syntactic quotations, LMS
+distinguishes binding-times based on types and preserves the evaluation order
+via an internal ANF transformation of the staged program.
+To achieve typed-based MSP, LMS provides a type constructor @Rep[T]@, where @T@
+can be an arbitrary type, indicating that a value of type @T@ will be known at
+a future stage.  Operations acting on @Rep[T]@ expressions will be residualized
+as generated code.
 A classic example for introducing multi-stage programming is the power function
@@
 
-In @close@, we first create a function @f@, which input and output are both
-next-stage values and stores.  The function @f@ is a current-stage and we then
-use @emit_complied_clo@ to denote it to a next-stage Scala function.
+In @close@, we first create a current-stage function @f@, whose input and
+output are both next-stage values and stores.  We then use @emit_complied_clo@
+to delay @f@ to a next-stage Scala function.
 The method @emit_compiled_clo@ produces current-stage representations of
@@
 expression @e@ to be specialized, but also with the prepared environment and
-store.  By doing so, we have \textit{collapsed} the monadic values of @AnsM@ to
-grounded values within staging; therefore in the future stage, there will be no
-monadic value.
+store.  By doing this, we have \textit{collapsed} the monadic values of type
+@AnsM@ to grounded values (of type @Rep[ValSt]@) within staging; therefore in
+the future stage, there will be no monadic value.
 \begin{lstlisting}

### 2019-09-15 paper/40_absinterp.tex
space
polish

@@
 
-Many abstract domains in abstract interpreters can be formulated by complete
+Many abstract domains in abstract interpreters can be formulated as complete
 lattices.
@@
   \centering
+  \vspace{-1em}
   \begin{subfigure}[b]{0.45\textwidth}
@@
   \end{lstlisting}
-  \caption{trait \texttt{SPLattice} and \texttt{Lattice}} \label{fig:splattice}
   \end{subfigure}
@@
   \end{lstlisting}
-  \caption{The power set instance for lattice} \label{fig:powerset}
   \end{subfigure} 
+  \vspace{-1.5em}
+  \caption{trait \texttt{SPLattice} and \texttt{Lattice} (left),
+  and the power set instance (right) } \label{fig:splattice}
+  \label{fig:powerset}
 \end{figure}
+\vspace{-1em}
 
@@
 Also, at some later point, we need to compute the join of two paths.  Another
-source of nondeterminism comes from function applications, for instance,
-consider an application @f(a)@, the possible $\lambda$-terms of @f@ are
-computed approximately.  As the result, we need to store all the possible
-callees, as well as we will retrieve a set of closures when dereferencing an
-address.
+source of nondeterminism comes from the abstract semantics of function
+application. For instance, consider an expression @f(a)@, if there are multiple
+possible target closures of @f@, then we have to apply all of them
+nondeterminitically.
+%As the result, we need to store all the possible callees, as well as we will
+%retrieve a set of closures when dereferencing an address.
 
@@
 values. Recall that in the concrete setting, we only have a reader monad (for
-environments) and a state monad (for stores). The reader effect is not
-persistent to the final result. Hence, together with the value produced by the
-interpreter, the final result type is just a pair of @Value@ and @Store@.
-However, under the new monad stack for abstract interpretation, we have a
-@SetT@ monad transformer inside the environment monad and store monad. Therefore, the type
-@Set@ becomes the container type of the pairs of values and stores, i.e.,
-@Set[(Value, Store)]@.  Also, note that the reader and state monad for the
-caches both inhibit internally of the nondeterminism monad; as the result, the
-final result type pairs the set of value-stores with the cache type from
-that state monad, as shown below.
+environments) and a state monad (for stores). 
+Different from state effects, reader effects are not visible at the final
+result. Hence, together with the value produced by the interpreter, the final
+result type is just a pair of @Value@ and @Store@.  However, under the new
+monad stack for abstract interpretation, we have a @SetT@ monad transformer
+inside the environment monad and store monad. Therefore, the type @Set@ becomes
+the container type of the pairs of values and stores, i.e., @Set[(Value, Store)]@.
+Also, note that the reader and state monad for the caches both
+inhabit the nondeterminism monad; as the result, the final result
+type pairs the set of value-stores with the cache type from that state monad,
+as shown below.
 \begin{lstlisting}
@@
 conditionals: in method @br0@, we combine the results using the @mplus@
-operation from @MonadPlus@, which requires the value domains are
+operation from @MonadPlus@, which requires the value domains to be
 join-semilattices (in our case, @Set@ and @Map@).
@@
 The difference is that now the first argument of @ap_clo@ is a set of closures
-@funs@, so we cannot directly destruct it with a syntactic @Lam@ by pattern
+@funs@, so we cannot directly match it against with a syntactic @Lam@ by pattern
 matching. Instead, we use an auxiliary function @lift_nd@ that takes a set and
 lifts the elements in the set into the monad stack. Then we can
-straightforwardly implement the method @ap_clo@ still in monadic style, where the
+straightforwardly implement the function @ap_clo@ still in monadic style, where the
 closures come from the monads and thus the nondeterminism can be naturally
@@
 iteration, and the @out@ cache is constantly updated during the current iteration. In
-the next iteration, the last @out@ cache is served as the @in@ cache, and
-an empty cache is plugged-in to the @out@ slot. Once the @out@ cache does not
+the next iteration, the last @out@ cache is used as the @in@ cache, and
+an empty cache is plugged into the @out@ slot. Once the @out@ cache does not
 contain any new information compared with the @in@ cache, the fixed-point is
@@
 The iteration terminates when the resulting @out@ cache is equivalent to the
-input @in@ cache, which indicates that there is no more fact has been discovered.
+input @in@ cache, which indicates that there is no more fact have been discovered.
 Therefore, the iteration should end, and we have reached the least fixed-point.

### 2019-09-16 paper/10_intro.tex
misc

@@
 Sergey:2013:MAI:2491956.2491979}, arrows \cite{Keidel:2018:CSP:3243631.3236767},
-extensible effects \cite{Githubsemantic},
-etc. These pure approaches provide certain benefits, including the fact that the
+and extensible effects \cite{Githubsemantic}.
+These pure approaches provide certain benefits, including the fact that the
 abstract interpretation artifacts can be built in a compositional and modular
@@
 implementing abstract interpreters, which allows high-level abstractions during
-implementation for clarity, but passes the incurred overhead at runtime for
+implementation for clarity, but bypasses the incurred overhead at runtime for
 efficiency.

### 2019-09-16 paper/abstract.tex
misc

@@
 approach derives a sound staged abstract interpreter based on a
-semantic-agnostic interpreter with type-level binding-time abstractions and
+generic interpreter with type-level binding-time abstractions and
 monadic abstractions. By using different instantiations of these abstractions,
@@
 compilation, applications to various control-flow analyses, and a demonstration
-that can be used for modular analysis. We also empirically evaluate the effect
+for modular analysis. We also empirically evaluate the effect
 of staging on the execution time. The experiment shows an order of magnitude

### 2019-09-16 paper/10_intro.tex
misc

@@
 To apply the Futamura projections to abstract interpreters, we consider
-\citet{DBLP:journals/pacmpl/DaraisLNH17}'s big-step monadic abstract interpreter
-as the unstaged baseline.  Similar to two-level semantics \cite{NIELSON1989117},
-monads provide a good abstraction to hide the detail of abstract interpretation
-semantics. However, monads do not have a clear stage distinction. We further
-introduce binding-times and make the monads binding-time polymorphic. Our
-proposed framework adopts type-based multi-stage programming from the
-Lightweight Modular Staging framework, and implements the first Futamura projection of
-an abstract interpreter for a stateful higher-order language. By deriving staged
-monads that can be used to perform code generation, we obtain a staged abstract
-interpreter that can be used for specialization. Through staging, the generated
-code is specialized to the input program, and all the monadic operations are
-inlined and compiled down to low-level Scala code.
+\citet{DBLP:journals/pacmpl/DaraisLNH17}'s big-step monadic abstract
+interpreter as the unstaged baseline.  Our proposed framework adopts type-based
+MSP from the Lightweight Modular Staging framework, and implements the first
+Futamura projection of this monadic abstract interpreter for a stateful
+higher-order language. 
+Similar to two-level semantics \cite{NIELSON1989117}, monads provide a good
+abstraction to hide the detail of abstract interpretation semantics. However,
+monads do not have a clear stage distinction. We further introduce
+binding-times and make the monads binding-time polymorphic. 
+By deriving staged monads that can be used to perform code generation, we
+obtain a staged abstract interpreter that can be used for specialization.
+Through staging, the generated code is specialized to the input program, and
+all the monadic operations are inlined and compiled down to low-level Scala
+code.
 
@@
 3) We show that staging an abstract interpreter enables modular
-compilation of an analysis to programs. Here we borrow the concept of
+compilation of an analysis to programs. Here, we borrow the concept of
 modular analysis, and show that the compiled analysis is reusable.
@@
     (Section~\ref{sai}). We also summarize the approach and discuss
-    soundness properties after showing the four artifacts.
-  \item We summarize our approach and discuss correctness (Section~\ref{discussion}).
-    We conduct three case studies (Section~\ref{cases_study}) and
-    an empirical evaluation on performance improvements (Section~\ref{evaluation}).
+    correctness properties after showing the four artifacts (Section~\ref{discussion}).
+  \item We conduct three case studies (Section~\ref{cases_study}) and an
+    empirical evaluation on performance improvements (Section~\ref{evaluation}).  
     Finally we discuss related work and conclude.

### 2019-09-16 paper/20_prelim.tex
misc

@@
 we use data-types to represent programs. This is the most natural choice
-for program analysis and allows us to use different interpretations over the
+for program analysis and it allows us to use different interpretations over the
 AST; with the inheritance and overriding mechanisms in Scala, we may also add
@@
 Note that here the monadic object of type @List[_]@ encapsulates the data internally.
-Therefore, it only exposes the simplified version of @flatMap@, where the monadic
+Therefore, it exposes the simplified interface of @flatMap@, where the monadic
 value of @M[A]@ is not introduced as a function argument. The trait @MonadOps@ (Figure
-\ref{fig:monadops}) defines the simplified versions of monadic operations that
+\ref{fig:monadops}) defines the simplified signatures of monadic operations that
 are necessary for @for@-comprehension. The conversion between @Monad@ and
@@
 In later sections, we will instantiate the monadic type to perform concrete
-interpretation \cite{DBLP:conf/popl/LiangHJ95} or abstract interpretation
-\cite{Sergey:2013:MAI:2491956.2491979, DBLP:journals/pacmpl/DaraisLNH17}.
+interpretation \cite{DBLP:conf/popl/LiangHJ95, Wadler:1992:EFP:143165.143169}
+or abstract interpretation \cite{Sergey:2013:MAI:2491956.2491979,
+DBLP:journals/pacmpl/DaraisLNH17}.
 

### 2019-09-16 paper/20_prelim.tex
polish

@@
 interpreter act as a compiler.
+\begin{lstlisting}
+  type MonadOps[R[_], M[_], A] = {           // Environment operations
+    def map[B](f: R[A] => R[B]): M[B]        def ask_env: AnsM[Env]
+    def flatMap[B](f: R[A] => M[B]): M[B]    def local_env(ans: Ans)(ρ: R[Env]): Ans
+  }                                          // Store operations
+                                             def get_store: AnsM[Store]
+  type AnsM[T] <: MonadOps[R, AnsM, T]       def put_store(σ: R[Store]): AnsM[Unit]
+  type Ans = AnsM[Value]                     def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
+\end{lstlisting}
 
@@
 addresses and (potentially staged) values, and updates the store accordingly.
-\begin{lstlisting}
-  type MonadOps[R[_], M[_], A] = {           // Environment operations
-    def map[B](f: R[A] => R[B]): M[B]        def ask_env: AnsM[Env]
-    def flatMap[B](f: R[A] => M[B]): M[B]    def local_env(ans: Ans)(ρ: R[Env]): Ans
-  }                                          // Store operations
-                                             def get_store: AnsM[Store]
-  type AnsM[T] <: MonadOps[R, AnsM, T]       def put_store(σ: R[Store]): AnsM[Unit]
-  type Ans = AnsM[Value]                     def set_store(av: (R[Addr], R[Value])): AnsM[Unit]
-\end{lstlisting}
 
-\paragraph{Primitive Operations} Next, we define a few primitive operations.
+\paragraph{Primitive Operations} Next, we declare a few primitive operations.
 First, we declare two versions of @alloc@. The first takes a store and an
@@
 consequently wrapped by @R[_]@. The other @alloc@ is simply the monadic version
-of the addresses, and can therefore be used in monadic computations.
+of the previous one, and can therefore be used in monadic computations.
 An auxiliary method @get@ retrieves the value of an identifier @x@ through the
@@
 Other primitive operations in the interpreter handle the language constructs.  The
-methods @num@ and @close@ deal with primitive values, which lift literal terms
+methods @num@ and @close@ deal with primitive values. For example, @close@ lifts literal terms
 (e.g., lambdas) to our value representation (e.g., closures).

### 2019-09-16 paper/30_stageconc.tex
polish

@@
 as generated code.
-A classic example for introducing multi-stage programming is the power function
+A classic example of multi-stage programming is the power function
 that computes $b^x$. Usually, it is implemented as a recursive function:
@@
 interfaces. Later, we will instantiate the staged set monad by manually fusing
-the fragment of the monad stack inside @SetT@.
+the fragment of the monad stack inside the @SetT@ transformer.
 
@@
 the evaluator at the current stage. The result is generating a next-stage
-Scala function. The evaluator is passed as @ev@ to the method @close@.
+Scala function. The evaluator is already passed as @ev@ to the method @close@.
 The following code implements this idea:

### 2019-09-16 paper/40_absinterp.tex
polish

@@
 power set is the empty set, and we do not have a representation of the top element for
-power set.  Other lattices used in the paper, such as products and maps, can be
+power sets.  Other lattices used in the paper, such as products and maps, can be
 implemented similarly or by lifting the existing lattices element-wise or
@@
 (ADI) approach \cite{DBLP:journals/pacmpl/DaraisLNH17} to refactor our monadic
-concrete interpreter to a monadic abstract interpreter.
+concrete interpreter into a monadic abstract interpreter.
 
@@
 values into a single monadic value of type @Set[M, B]@ by appending all of them.
-The initial value of the fold is an empty @SetT@.
+The initial value of the fold is the empty value of @SetT@.
 \begin{lstlisting}
@@
 \paragraph{Primitive Operations} The primitive operations are changed according
-to the new monad stack scheme and value domains. One of the notable changes is
-that the store update operator merges the new values with existing values.  As
-we mentioned before, sometimes we have to explore both of the branches for
-conditionals: in method @br0@, we combine the results using the @mplus@
-operation from @MonadPlus@, which requires the value domains to be
-join-semilattices (in our case, @Set@ and @Map@).
+to the new monad stack scheme and value domains. One of the notable changes due
+to the nondeterminism is that the store-update operator merges the new values
+with existing values for the same address in the store.  As we mentioned
+before, sometimes we have to explore both of the branches for conditionals: in
+method @br0@, we combine the results using the @mplus@ operation from
+@MonadPlus@, which requires the value domains to be join-semilattices (in our
+case, @Set@ and @Map@).
 \begin{lstlisting}

### 2019-09-29 oopsla19-paper/10_intro_back.tex
reorganize

\section{Introduction} \label{intro}

\iffalse
  Revision: 
  source of overhead: 1) interpreration, dispatch, recursion.
  2) modern ai uses effect systems to construcut abs interp, in a purely functional way.
  which has certain benefits, but introduce additional abstraction layers may lead to 
  performance overhead. Such as MAAM, ADI, Semantic

  also motivate from the perspective that ADI is slow,
  analyze the reason: 1) interpretive overhead, 2) the use of monadic layers,
  3) the intrinsic complexity of abs int of HO language (ie CFA particularly).
  Then introduce the abstraction without regret.
  contribution: derive staged monad transformers.
\fi

Futamura projections \cite{Futamura1999, futamura1971partial} reveal the close
connection between compilers and interpreters. The first Futamura projection
specifically shows that specializing an interpreter with respect to the input
program yields an equivalent executable. Partial evaluation
\cite{DBLP:books/daglib/0072559} was the first proposed approach to realize
Futamura projections: it identifies the binding-time of variables in
the program (i.e., they can be known statically or dynamically),
evaluates the static part, and finally generates a residual program that solely
relies on the dynamic part. However, given an arbitrary program, precisely
analyzing its binding-times is hard in general. As an alternative and pragmatic
approach to specialization and partial evaluation, multi-stage programming (MSP) 
\cite{taha1999multi, DBLP:conf/pepm/TahaS97} requires the staging
annotations (i.e., binding-time annotations) to be given explicitly from the
programmers. The staging annotations can be either syntactic (as in
MetaML/MetaOCaml, through quasi-quote, escape, etc.) or type-based \cite{DBLP:conf/gpce/RompfO10}.
These staging annotations identify which part of the inputs should be
specialized. As a classical example, we use the power function and type-based
annotations to introduce the idea of MSP:

\rev{this power is based on the induction of natural number e}

\begin{lstlisting}
  def power(b: Rep[Int], e: Int): Rep[Int] = if (e == 0) 1 else b * power(b, e - 1)
\end{lstlisting}

If the programmer identifies that @b@ will be known in the future stage (as shown
on its type @Rep[Int]@ -- a representation of @Int@) and @e@ is known at the
current stage, say 5, then we can specialize the function @power@ with
respect to @e = 5@, and generate a specialized function where the overhead of
recursion is eliminated. Conceptually, the generated code looks like the following:

\begin{lstlisting}
  def power5(b: Int): Int = b * b * b * b * b
\end{lstlisting}

Now to concretely illustrate the idea of Futamura projection, consider that we
have an interpreter @eval@ of some language:

\begin{lstlisting}
  def eval(e: Expr)(arg: Input): Value = ...
\end{lstlisting}

where $e$ is a program of type @Expr@ and @arg@ is the input to $e$,
and @eval@ produces a value.
Semantically, the interpreter satisfies $ \texttt{eval}(e)(arg) = [\![ e ]\!] arg$ for all
programs and inputs. Given a program $e_0$, by applying the first Futamura
projection, we can obtain a specialized interpreter
$\texttt{eval}_{\texttt{e0}}$ with respect to $e_0$, which is the so-called \textit{equivalent executable}. 
By definition of the interpreter, 
$\texttt{eval}_{\texttt{e0}}(arg) = [\![ e_0 ]\!] arg $. If we generalize the
argument @arg@ to an environment that maps free variables in $e$ to some
values, the interpreter is a standard environment-passing interpreter,
where the specialization through MSP is also applicable.

Roughly at the same time when Futamura projections were envisioned in the 1970s,
\citet{DBLP:conf/popl/CousotC77} proposed abstract interpretation as a
lattice-based semantic approach to construct sound static analysis, by
approximation of fix-points. However, constructing sound abstract interpreters
was considered abstruse and complicated for a long time.
Recent progress such as Abstracting Abstract Machines (AAM)
\cite{DBLP:conf/icfp/HornM10} uncovers a systematic method to derive sound
abstract interpreters from their concrete counterparts, where the soundness can
be easily established by examining the transformation of semantics. The AAM
approach also has been applied to different variants of definitional
interpreters and abstract machines \cite{DBLP:journals/jfp/HornM12,
DBLP:conf/icfp/HornM10, DBLP:journals/pacmpl/DaraisLNH17}.

\vspace{-1em}
\begin{figure}[h]
  \begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=3em,minimum width=2em]
  {
    \begin{smallmatrix} \text{unstaged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_abs}) \end{smallmatrix} &
    \begin{smallmatrix} \text{staged} \\ \text{abstract} \\ \text{interpreters} \\ (\text{Sec.}~\ref{sai}) \end{smallmatrix} \\
    \begin{smallmatrix} \text{unstaged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{unstaged_conc}) \end{smallmatrix} &
    \begin{smallmatrix} \text{staged} \\ \text{concrete} \\ \text{interpreters} \\ (\text{Sec.}~\ref{stagedinterp}) \end{smallmatrix} \\ };
  \path[-stealth]
    (m-1-1) edge node [above, font=\scriptsize] {$\updownarrows$} (m-1-2)
    (m-2-1) edge node [left, font=\scriptsize] {$\alpha$} (m-1-1)
    (m-2-1) edge node [below, font=\scriptsize] {$\updownarrows$} (m-2-2)
    (m-2-2) edge node [right, font=\scriptsize] {$\alpha$} (m-1-2);
  \end{tikzpicture}
  \caption{The confluence of specialization and approximation.}
  \label{confluence}
\end{figure}
\vspace{-1em}

\paragraph{Intellectual Motivation: Futamura Projection of Abstract Interpreters}

Now that we have seen two orthogonal transformations of interpreters: by
specialization, we can refactor interpreters to code generators; by
approximation, we can refactor interpreters to static analyzers. As an
intellectual quest, we should naturally ask -- starting from an abstract
interpreter, \textit{can we derive a sound abstract interpreter while it
produces optimized code that does the analysis?}
Conceptually, it is similar to the concrete setting: given an
abstract interpreter $\widehat{\texttt{eval}}: \texttt{Expr} \to
\widehat{\texttt{Env}} \to \widehat{\texttt{Value}}$, which takes a program, an abstract
environment and returns abstract values, we would like to specialize it with
respect to some program $e_0$ and produce a compiled analysis
$\widehat{\texttt{eval}}_{\texttt{e0}} : \widehat{\texttt{Env}} \to
\widehat{\texttt{Value}}$, where the abstract environment $\widehat{\rho}$ is kept as input.
The specialized analysis satisfies $\widehat{\texttt{eval}}_{\texttt{e0}}(\widehat{\rho}) =
\widehat{\texttt{eval}}(e_0)(\widehat{\rho})$.

In this paper, we study the confluence of two old ideas --- Futamura projection
and abstract interpretation, but from the perspective of their very recent
realizations --- multi-stage programming and definitional abstract interpreters.
To be specific, we present the application of the first Futamura projection on
definitional abstract interpreters, which intersects as staged abstract
interpreters. We develop a unified approach to fulfill the four different
semantics, as shown in Figure~\ref{confluence}. Starting with a generic but
parametric interpreter shared by the four semantics, by simply instantiating
the interpreter differently, we can automatically derive a concrete interpreter,
or an abstract interpreter, or their staged versions that generate code, respectively.

\iffalse
embedded domain-specific languages \cite{DBLP:conf/snapl/RompfBLSJAOSKDK15,
DBLP:journals/jfp/CaretteKS09, DBLP:conf/icfp/GibbonsW14,
Hofer:2008:PED:1449913.1449935},
The key idea that enables this flexibility is to use abstract type members first
abstract over concrete/abstract components (e.g., concrete values or abstract
values), then as well as the binding-time of them (e.g., static or dynamic). By
staging, the overhead caused by the monadic layers is eliminated in the
generated code.
\fi

\paragraph{Practical Motivation: Abstraction without Regret}
\rev{Abstraction without performance regret; Performance without soundness regret}

To enable such flexibility between different semantics, we combine generic
programming and generative programming, and extensively use abstract type
members over the following abstractions, which can be later instantiated.

\begin{itemize}
\item Env-store-value abstractions, which define the domains that the interpreter will be running with.
  The environments, stores, and values are initially declared as abstract types in the generic interpreter.
  As shown by the AAM approach \cite{DBLP:journals/jfp/HornM12,
  DBLP:conf/icfp/HornM10}, if we widen the values to abstract domains (such as
  intervals) and refactor the store to a mapping from finite addresses to sets of
  values, we can derive a computable and sound abstract interpreter.
\item Binding-time abstractions, which control whether the interpreter produces
  values or generates code. The binding-times differentiate whether data are 
  known statically or dynamically. Unsurprisingly, such binding-time
  information can be encoded into higher-kinded types 
    \cite{DBLP:journals/jfp/CaretteKS09, Ofenbeck:2017:SGP:3136040.3136060}.
\item Monadic abstractions, which encapsulate the above two abstractions.
  The interpreter itself is written in monadic form, and by plugging different monads, 
  the underlying semantics of the interpreter is changed. We also use monad transformers to combine multiple monadic effects.
  %\todo{mention that direct-style also works?}
  %\cite{Steele:1994:BIC:174675.178068, DBLP:conf/popl/LiangHJ95, DBLP:journals/pacmpl/DaraisLNH17, Sergey:2013:MAI:2491956.2491979}
\end{itemize}

We show that these well-known abstractions can be combined smoothly, and more
importantly for free. We implement the four interpreters using Scala and the
Lightweight Modular Staging \cite{DBLP:conf/gpce/RompfO10} framework. Through
staging, the major overhead of abstractions -- from the monadic layers -- is
completely eliminated in the generated code. We observe that the staged version is 
on average 11x times as fast as the unstaged abstract interpreter. 
However, we do not lose any advantages of this high-level programming style,
e.g., modularity and extensibility. The meaning of generated programs also remain
the same, particularly, the soundness of static analysis after staging.

%\paragraph{Performance via Specialization}
Besides the intellectual merit shown in the confluence diagram, this paper
contributes to a practical problem in static analysis: \textit{can we speed up a
static analyzer without any intrusive changes and soundness compromises?} Static
analysis is known as a trade-off game between \textit{precision} and
\textit{efficiency}, we argue that applying meta-programming (multi-stage
programming, in particular) with abstract interpreters is an effective approach
to improve the latter while the former is untouched.
In the literature, specializing analysis with respect to the input program
\cite{damian1999partial, amtoft1999partial, Boucher:1996:ACN:647473.727587,
ashley:practical} is not new, despite the fact that most of the previous works
are ad-hoc on a particular analysis or requiring significant changes to the
analyzer. For example, abstract compilation is one of the techniques
\cite{Boucher:1996:ACN:647473.727587}. But it requires the whole analyzer to be
rewritten to the closure generation form. In this paper, we advocate a systematic
approach of generative programming that achieves the performance goal with less
intrusive changes and engineering efforts.

\paragraph{Applications}

Furthermore, we show various applications of staging an abstract interpreter.
(1) We compare our approach with abstract compilation, and show that we can
achieve the same optimization by utilizing type-based stage annotations,
meanwhile, the analyzer program does not need any changes. (2) We extend the 
basic staged abstract interpreter to diverse control-flow analysis, 
including a store-widened analysis, a context-sensitive analysis and abstract garbage collection.
(3) Staging abstract interpreter enables modular analysis and provides a mechanized
way to create library summaries. We notice that we may apply the staged abstract
interpreter separately on the different part of the analyzed program, and the
generated code is reusable and composable. Therefore, we can turn a
whole-program analyzer into an analyzer that runs modularly.

\iffalse
Our next contribution is to enable modular analysis based on a whole-program
analyzer by meta-programming. Modern software are shipped with large libraries
\cite{toman_et_al:LIPIcs:2017:7121}, and analyzing programs with large libraries
is usually expensive or unnecessarily repeated for whole-program analyzing
algorithms. For example, it has been shown that analyzing a simple ``Hello
World'' program in Java depends on additional 3,000 classes in the library
\cite{DBLP:conf/oopsla/KulkarniMZN16}. A natural idea is to analyze libraries
and generate summaries, later we can plug in the necessary information into the
summaries and finish the analysis; therefore such summaries should be modular
and reusable. Meta-programming gives us a chance to readily tweak a
whole-program analyzing algorithm into an analyzer that works modularly and
produces summaries. Still, the idea is to specialize the abstract interpreter
that was used for whole-program analysis but on a subset of the program, for
examples, a set of generic functions for list manipulation. Such specialization
works on lambda expressions and generates code, which is partial analysis
results and can be completed by running with the necessary arguments (for
example, an environment that closes the lambda term) at next stages. Multiple
partial analysis results also can be reused later and composed together as they
are just normal programs at the next stage. Through staging, we mechanically
obtain a modular analysis, even though the original algorithm is formulated as a
whole-program analysis.
\fi

\paragraph{Contributions} Briefly, the contribution of the paper is twofold:
\begin{enumerate}[leftmargin=2em]
  \item Intellectually, through multi-stage programming, we present a unified construction 
    that naturally extends the first Futamura projection to abstract interpreters.
 \item Pragmatically, by case studies and empirical evaluation, we show that staging 
   abstract interpreters is useful to improve performance while not touching
   the soundness. 
\end{enumerate}
\paragraph{Organization} The paper is organized as follows:
\begin{itemize}[leftmargin=2em]
\item Starting from a generic interface of the interpreter for a small higher-order language
  (Section~\ref{prelim}), we first show the instantiation of concrete
  interpretation (Section~\ref{unstaged_conc}), and then lift it to the staged
  version by replacing the binding-time types (Section~\ref{stagedinterp}).
\item We present the unstaged abstract interpreter under the same framework by
  replacing the environment, store and values (Section~\ref{unstaged_abs}).
  Finally, we show the combination of approximation and specialization, dubbed
  \textit{staged abstract interpreters}, can be readily derived (Section~\ref{sai}).
\item We conduct three case studies on the applicability of our approach
  (Section~\ref{cases_study}).
  \begin{enumerate}
  \item We first revisit the Abstract Compilation (AC)
    \cite{Boucher:1996:ACN:647473.727587} (Section~\ref{cs_ac}), showing that we can
    achieve the same optimization with much less engineering effort.
  \item We extend the staged abstract interpreter to various control-flow
    analysis (Section \ref{cfa}), including context-sensitivity, store-widening, and abstract
    garbage collection, showing that our approach is widely applicable.
  \item We apply our approach to modular analysis (Section \ref{modular}),
    i.e., analyzing the components of the program separately and composing them later,
    showing a novel application of generative programming with program analysis.
  \end{enumerate}
\item We empirically evaluate the performance improvement by staging on
    control-flow analysis (Section~\ref{evaluation}). We compare the both
    context-insensitive and store-widening analysis.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\iffalse
Likewise, specializing static analyses by partial evaluation emerged in the late 90s
, and indeed it is able to effectively remove the interpretive
overhead of repeatedly traversing the abstract syntax tree. However, these
previous works focused mostly on one particular analysis, or required to
completely rewrite the analyzer. Hence, it is worth to investigate the idea from
a modern perspective based on generative programming, especially for a general
abstract interpreter that models direct-style $\lambda$-calculus, imperative
features such as mutation, and different abstract domains. One technical
challenge is the binding-time engineering when non-determinism, fixed-point
iterations, different abstract domains, and staging are introduced at the same
time. In this paper, we present an end-to-end design and implementation of
staging an abstract interpreter; that means not only the interpreter that
traverses the abstract syntax tree, but also the data structure of abstract
domains, abstract environment and heap, and the fixed-point iteration are
staged.

\todo{rewrite this paragraph}
On the other hand, the abstract interpreter, as a semantic artifact, should be
written in a style that is easy for people to communicate the formulation and
abstraction, but also can be implemented efficiently. As the slogan of
multi-stage programming said, "abstraction without regret", we draw a connection
between high-level description and efficient implementation of abstract
interpreters, just like the connection between concrete interpreters and
compilers drawn by the Futamura projections. Particularly, we show an easy but
systematic way of adding stage annotations to the abstract interpreter without
changing the code of the interpreter skeleton, which is shared between four concrete
or abstract, unstaged or staged interpreters. We use the LMS framework for staging,
which allows us only use types to annotate the binding-time. Therefore, the
proposed approach bridges the gap between designing sound static analysis and
implementing efficient program analyzer.

%Of course, all interpreters are metalinguistic abstractions, but some
%interpreters are more "abstract" than others \todo{maybe rephrase}.
%Particularly, we also would like a systematic approach to optimize program
%analyzers and meanwhile minimally modifying the analyzer programs.

When staging a concrete interpreter, the programmers need to distinguish static
and dynamic values --- the given program to be executed by the interpreter is
classified as static because it is known at compile-time, and the inputs to that
program are dynamic. However, when staging an abstract interpreter, this
distinction does not exist anymore. Because the abstract interpreter
instantiates all the inputs as some form of abstract values, which are usually
top elements in their abstract domains and are also statically known. Then what
is the point of staging if there is no such distinction? A surprising by-product
of thinking about this question is to realize that we can apply the staged
abstract interpreter on \textit{open} programs, and the free variables
representing other parts of the program (e.g., libraries) are dynamic inputs,
therefore we obtain a modular analysis through staging, mechanically.
\note{TR: I don't understand this. The program structure is static, the abstract values
are still dynamic, no? They change in every iteration of the fixpoint algorithm}
\note{TR: why is it a big deal. On Closed programs, we obtain a constant factor,
but modular analysis has different asymptotics}

When targeting higher-order programs, either staged concrete interpreters or
staged abstract interpreters are able to compile a closure, i.e., specialize the
call of interpreter with respect to the body expression of the lambda term
without knowing the actual argument value. By generalizing this observation, we
can actually specialize the abstract interpreter with any open programs, which
unexpectedly leads to a modular analysis and improves scalability. An open
program contains free variables, which represent other parts of the program and
will be left as dynamic values. For instance, one challenge in static analysis
of modern software is that the programs are usually shipped with large library
code \cite{toman_et_al:LIPIcs:2017:7121}, for example, it has been shown that
analyzing a simple "Hello World" program in Java depends on additional 3,000
classes in the library \cite{DBLP:conf/oopsla/KulkarniMZN16}. A precise
whole-program analysis formulated as abstract interpreters can be very expensive
due to the scale of the program and the inherent complexity of the algorithm.
However, analyzing these libraries is sometimes unnecessarily repeated. When
applying the staged abstract interpreter on such open programs, we leave the
unknown arguments and calling contexts as dynamic values and generate partial
analyzing result which is represented as a residual program. The partial
analyzing result can be reused and composed with the analyzing result of
application code when available later. Therefore we mechanically obtain a
modular analysis through staging, even though the original algorithm is
formulated as a whole-program analysis.

It has been observed that partially applying context-sensitivity on selected
portions of the program could improve the precision and efficiency
\cite{zipper2018, Kastrinis:2013:HCP:2491956.2462191}. We show that staging
abstract interpreters as an approach to effectively implement hybrid
context-sensitivity\todo{}.
\fi

%\note{Expression problem \cite{DBLP:conf/icfp/GibbonsW14, DBLP:conf/ecoop/KrishnamurthiFF98, expproblem},}

### 2019-09-29 oopsla19-paper/50_stageabs.tex
reorganize

--- 2019-09-14 paper/50_stageabs.tex
@@
 We use the same monad stack structure as the unstaged abstract interpreter,
-but replacing \textit{all} the monad transformers with their staged versions. The
+but replace \textit{all} the monad transformers with their staged versions. The
 following code snippet shows this change.
-We manually fuse the three inner transformers (@IdM@ is omitted) into a single
+We manually fuse the three inner transformers (@IdM@ is omitted) into a single staged
 monad @RepSetReaderStateM[R, S, A]@, where @R@ is the type parameter for the
-reader effects, and @S@ is the type for the state.  In our monad stack scheme,
+reader effects, and @S@ is the type parameter for the state.  In our monad stack scheme,
 @R@ and @S@ are both instantiated with the type @Cache@. Similar to the unstaged
-version, the grounded result type is a pair of two staged values:
+abstract semantics, the grounded result type is a pair of two staged values:
 @Rep[Set[(Value, Store)]]@ and @Rep[Cache]@.
@@
 \paragraph{Primitive Operations} 
-When deriving the staged concrete interpreter from its unstaged counterpart, we
-notice that where the evaluation @ev(e)@ is invoked are shifted from @ap_clo@ to
-@close@, where @e@ is the body expression of a $\lambda$-term.
+When deriving the staged concrete interpreter from its unstaged counterpart, it
+is observed that where we invoke the evaluator @ev(e)@ is shifted
+from @ap_clo@ to @close@, where @e@ is the body expression of a $\lambda$-term.
 In other words, when staging, we eagerly specialize the interpreter and
@@
 section, we present optimizations that largely mitigate these issues.
-Implementing all of those optimizations do not need to change the generic
+Implementing all of those optimizations does not need to change the generic
 interpreter.
@@
 \cite{Flanagan:1993:ECC:155090.155113}, we can use a selective caching
-algorithm that does not generate caching code for atomic expressions:
+algorithm that does not generate caching code for atomic expressions.
+This optimization is implemented by using @fix_select@ to dispatch expressions.
 \begin{lstlisting}
@@
     meaning after staging. In other words, the generated code may not preserve
-    the evaluation order or may contain undesired duplication of code.
+    the evaluation order or may contain undesired duplications of code.
     In this paper, we use the LMS framework, which 1) checks binding-times by
@@
 operational abstract semantics through abstract compilation
-\cite{Boucher:1996:ACN:647473.727587} -- as
-\citet{Johnson:2013:OAA:2500365.2500604} presented for
-optimizing Abstract Abstract Machines. However, the generated abstract
-bytecode still requires another small-step abstract machine to execute, which is
-an additional engineering effort.
+\cite{Boucher:1996:ACN:647473.727587}.
+\citet{Johnson:2013:OAA:2500365.2500604} presented this idea for optimizing
+Abstract Abstract Machines. However, the generated abstract bytecode still
+requires another small-step abstract machine to execute, which is an additional
+engineering effort.

### 2019-09-29 oopsla19-paper/60_cases.tex
reorganize

--- 2019-09-14 paper/60_cases.tex
@@
 loaded.
-
 Specifically, \citeauthor{Boucher:1996:ACN:647473.727587} showed how to compile a
@@
   def fix_gc(ev: (Expr => Ans) => (Expr => Ans))(e: Expr): Ans = for {
-    ψ <- ask_roots ... // omit code for retrieving ρ, σ, in and out
-    val res: Rep[(Set[(Value,Store)], Cache)] = ...  // omit code for checking caches
+    ψ <- ask_roots ... // omit code for retrieving ρ, σ, in and out, and checking caches
+    val res: Rep[(Set[(Value,Store)], Cache)] = ...  // the grounded result
     _ <- put_out_cache(res._2);  vs <- lift_nd(res._1)
@@
 As we have already seen in Section \ref{sai}, the primitive operation @close@
-can denote a $\lambda$-term to a next-stage Scala function. 
-The specialization of the staged abstract interpreter to a
-$\lambda$-term is also independent of other parts of the analyzed program.
+can denote a $\lambda$-term to a next-stage Scala function, which is the
+specialized analyzer to that $\lambda$-term.  The specialization of the staged
+abstract interpreter to a $\lambda$-term is independent of other parts of
+the analyzed program.
 \begin{lstlisting}
@@
 
-The specialization (0-CFA and without store-widening) of @close@ to
+The specialization (0-CFA without store-widening) of @close@ to
 @map@ generates the following next-stage Scala code, that is, the
@@
 different abstract arguments, stores, and caches, depending on the context
-where it is called.
+of where @map@ is called.
 
@@
 transition from a whole-program analyzer to a modular-staged analyzer is
-mechanized and for-free via staging. We do not need to design any
-intermediate representation or data structures of summaries.  In some
+mechanized via staging. We do not need to design any
+intermediate representation or data structure of summaries.  In some
 scenarios, traditional summary-based analyses are able to carry more

### 2019-09-29 oopsla19-paper/70_eval.tex
reorganize

--- 2019-09-11 paper/70_eval.tex
@@
 -XX:MaxMetaspaceSize=2G - XX:ReservedCodeCacheSize=2048M}},
-running on an Ubuntu 16.04 LTS (kernel 4.4.0) machine. All our
+running on an Ubuntu 16.04 LTS machine. All our
 evaluations were performed on a machine with 4 Intel Xeon Platinum
-8168 CPUs at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores
-(192 threads) in total, the abstract interpreters are single-threaded.
-To minimize the effect of warm-up from the HotSpot JIT compiler,
-all the experiments are executed for 20 times and we report the
-statistical median value of the running times. We set a 5 minute
-timeout.
+8168 CPUs at 2.7GHz and 3 TB of RAM. Although the machine has 96 cores in
+total, the abstract interpreters are single-threaded.  To minimize the effect
+of warm-up from the HotSpot JIT compiler, all the experiments are executed 20
+times and we report the statistical median value of the running times. We set a
+5 minute timeout.
 
@@
 
-Figure~\ref{evaluation_result} shows the evaluation result, comparing the
+Figure~\ref{evaluation_result} shows the evaluation result comparing the
 performance improved by staging on the two monovariant CFAs. The
@@
 for 0-CFA with store-widening, the average speedup is ~7 times.
-\texttt{kcfa-worst-64} has the highest speedup among all the tests. Although it
-has a comparably large size of AST, it has a rather simple structure which only
-involves booleans, $\lambda$-terms, and applications. We conjecture that the simple
-domain and structure allow the generated code to be efficiently optimized by the
-JVM. However, on the corresponding 0-CFA-SW, we do not observe such a significant
-speedup. Due to the complexity of generated code and next-stage running platform
-(JVM), we leave the detailed empirical analysis of which static analysis can
-benefit more from the specialization as future work.
+The highest speedup among all the tests is observed on benchmark \texttt{kcfa-worst-64}.
+Although the program size of \texttt{kcfa-worst-64} is large, it has a rather
+simple structure which only involves booleans, $\lambda$-terms, and
+applications. We conjecture that the simple domain and structure allow the
+generated code to be efficiently optimized by the JVM. However, on the
+corresponding 0-CFA-SW, we do not observe a significant speedup. Due to
+the complexity of generated code and next-stage running platform (JVM), we
+leave the detailed empirical analysis of which static analysis can benefit more
+from the specialization as a future work.
 

### 2019-09-29 oopsla19-paper/abstract.tex
reorganize

--- 2019-09-16 paper/abstract.tex
@@
 analysis. What happens when we combine these two ideas, and apply specialization to
-an \emph{abstract} interpreter?
+an abstract interpreter?
 
@@
 interpreter, or a staged abstract interpreter. As an example of
-\emph{abstraction without regret}, the overhead of these abstraction layers is
+abstraction without regret, the overhead of these abstraction layers is
 eliminated in the generated code after staging. We show that staging abstract

### 2019-10-06 draft/paper.tex
misc

--- 2018-01-05 paper.tex
@@
 \acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{CONF} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 %% Title information
-\title[Short Title]{Staged Abstract Interpreter}         %% [Short Title] is optional;
+\title[Short Title]{Full Title}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 \begin{abstract}
-  We present Staged Abstract Interpreter
+Text of abstract \ldots.
 \end{abstract}
@@
 %% comma separated list
-\keywords{generic programming, abstract interpretation, static analysis}  %% \keywords are mandatory in final camera-ready submission
+\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission
 
@@
 
-\textbf{Outline(Proposed)}.
+Text of paper \ldots
 
-1. Revisit abstract compilation (generates textual program and closures)
-
-Show that abstract compilation can be understood as an instance of staged 
-abstract interpreter.
-
-2. From a performance perspective, closure is still not an efficient enough representation.
-
-We implement staged abstract interpreter in a big-step semantics (abstract definitional interpreter) 
-using Scala with LMS framework.  
-Show it generates sound and high-performance low-level analysis code with 
-various precisions and sensitivities, for instances, k-CFA, pushdown or m-CFA.
-
-Store/set operations (join, update) is also expensive, generating efficient store implementation\cite{liang2014fast}.
-
-Extension: can we do it with small-step semantics? or staging a transition system?
-
-3. Modular analysis for free?
-
-Motivation: one of the challenges of modern static analysis is program usually depends on
-large libraries programs\cite{toman_et_al:LIPIcs:2017:7121}. 
-Can we analyze programs and libraries separately without losing precision? So that we can 
-reduce part of the overhead of repeatedly analyzing libraries for different programs.
-Similarly, some static analyzers compute summary for a function or a module, that can be reused
-later(Facebook Infer). But to my knowledge, they are mostly context-insensitive or imprecise.
-
-Application: for example, k-CFA (k > 0) is naturally a kind of whole program analysis, because 
-it need the last k calling contexts as abstraction. But can we analyze programs (libraries) separately 
-which generate the specialized analysis and leave the k calling contexts as dynamic parameters, 
-and then install these contexts when we have the whole program.
-
-Another perspective: programs are data for an abstract interpreter, so if we have $n$ programs, 
-then there can be $n$ stages. 
-Probably we can analyze first $m$ programs, and generate a residual abstract interpreter
-waiting for the rest $(n-m)$ programs.
-
-Not sure how much improvement of performance.
-
-4. Show the evaluation result on benchmark programs.
-
-\section{Background}
-
-\subsection{Abstracting Abstract Machine}
-
-\subsection{Staged Interpreter}
-
-\section{Abstract Compilation, Revisited}
-
-\section{Staged Abstract Interpreter}
-
-\section{Evaluation}
+\cite{Futamura1999, Futamura1971}
 
@@
 
-\textbf{Abstract Compilation}. The idea in this paper was closely inspired by 
-abstract compilation \cite{Boucher:1996:ACN:647473.727587}.
-\citeauthor{Boucher:1996:ACN:647473.727587} presented abstract compilation
-techniques as an efficient implementation of the monovariant flow analysis 
-\textit{0}-CFA for programs in continuation-passing style. The key idea is to 
-reduce the interpretation overhead on traversing the syntax tree by partial evaluation.
-Specifically, they proposed two kinds of abstract compilers to achieve this. 
-The first one is to generate specialized analysis as a textual program, then 
-can be loaded and executed by \texttt{eval} or similar mechanisms. The second 
-one is using closure as a representation of specialized analysis. 
-A closure is a function that remembers the environment where it defined. So the 
-abstract compiler generates higer-order functions on-the-fly rather than textual 
-program, then the functions can be applied immediately in a higher-order host language.
+Along with dynamic taint analysis, forward symbolic execution is widely used in
+software security and bug finding. \citet{DBLP:conf/sp/SchwartzAB10} discussed
+those techniques in a security context. Using a simple low-level intermediate
+language, they presented a formalization of forward symbolic execution in the form of
+small-step operational semantics. Existing challenges of symbolic execution and
+potential mitigations are also discussed.
 
-\citeauthor{Johnson:2013:OAA:2500365.2500604} in their work also adopt the 
-idea closure generation for optimizing small-step abstracting abstract machine
-\cite{Johnson:2013:OAA:2500365.2500604}.
-
-\textbf{Abstracting Abstract Machine}. Abstracting Definitional Interpreter \cite{darais2017abstracting}
-
-Collapsing Towers of Interpreters\cite{Amin:2017:CTI:3177123.3158140}
-
-\section{Conclusion}
+\cite{Siddiqui:2012:SSE:2245276.2231988}
 
@@
 %% Bibliography
-\bibliography{references}
+\bibliography{references,symexe}
 

### 2019-10-14 draft/paper.tex
some symbolic execution

@@
 
-Text of paper \ldots
+"building a symbolic compiler is often the most difficult aspect of
+creating solver-aided tools, especially for general-purpose languages" \cite{Torlak:2014:LSV:2594291.2594340}
 
@@
 
-\cite{Siddiqui:2012:SSE:2245276.2231988}
+Our technique should not be confused with
+\citet{Siddiqui:2012:SSE:2245276.2231988}'s staged symbolic execution, which
+performs symbolic execution in stages. Different from a compilation-based
+staging in this paper, their first stage generates abstract symbolic inputs,
+which then will be shared and used across other methods in the second stage.
 

### 2019-10-14 dev/src/main/scala/sai/lang/SimpIL.scala

@@ -12,11 +12,21 @@ object SimpIL {
   case class Halt() extends Stmt
 
   sealed trait Exp
-  case class Lit(x: Int) extends Exp
-  case class Var(x: String) extends Exp
-  case class Load(e: Exp) extends Exp
-  case class BinOp(op: String, e1: Exp, e2: Exp) extends Exp
-  case class UnaryOp(op: String, e: Exp) extends Exp
+  case class Lit(x: Int) extends Exp {
+    override def toString = x.toString
+  }
+  case class Var(x: String) extends Exp {
+    override def toString = x
+  }
+  case class Load(e: Exp) extends Exp {
+    override def toString = s"(load $e)"
+  }
+  case class BinOp(op: String, e1: Exp, e2: Exp) extends Exp {
+    override def toString = s"($e1 $op $e2)"
+  }
+  case class UnaryOp(op: String, e: Exp) extends Exp {
+    override def toString = s"($op $e)"
+  }
   case class GetInput(src: String) extends Exp
 
   object Values {
@@ -25,7 +35,14 @@ object SimpIL {
   }
 
   object Examples {
-    val ex1 = Prog(List(Assign("x", BinOp("*", Lit(2), GetInput("stdin"))),
-                        Halt()))
+    val ex1 = Prog(List(Assign("x", BinOp("*", Lit(2), GetInput("stdin"))), Halt()))
+
+    val ex2 = Prog(List(
+      Assign("x", BinOp("*", Lit(2), GetInput("stdin"))),
+      Cond(BinOp("==", BinOp("-", Var("x"), Lit(5)), Lit(14)),
+        Lit(2),
+        Lit(3)),
+      Assert(Lit(0)),
+      Halt()))
   }
 }
